### 2019-04-30 warmup.tex
first commit

\documentclass{patmorin}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{pat}
\usepackage{paralist}
%\usepackage{hyperref}

\setlength{\parskip}{1ex}

\title{\MakeUppercase{Notes on Layered $H$-Partitions of $k$-Planar Graphs}%
    \thanks{This work was partly funded by NSERC and MRI.}}

\author{Vida Dujmovi\'c%
        \thanks{School of Computer Science and Electrical Engineering, 
                University of Ottawa},\,\, 
        Pat Morin%
        \thanks{School of Computer Science, Carleton University},\,\, and
        David R. Wood%
        \thanks{School of Mathematical Sciences, Monash University}}

\newcommand{\dual}[1]{{#1}^\star}
\newcommand{\note}[2]{{\color{red}[#1:~#2]}}

\DeclareMathOperator{\dist}{dist}

\begin{document}
\maketitle

\begin{abstract}
  These are some notes on our work of layered $H$-partitions for $k$-planar graphs
\end{abstract}

\section{Introduction}

\section{$1$-Plane Graphs}

Let $G$ be an edge maximal 1-plane graph.  Here, edge-maximal should be taken to mean that any two vertices of $G$ on a common face have an edge joining them.  Assume also that $G$ is crossing minimal so that the embedding of $G$ minimizes the number of crossings over all possible embeddings of the (abstract) graph $G$.  From now on, we treat vertices of $G$ as points and edges as simple curves.

A \emph{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.  It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
The edges $vw$ and $xy$ are called \emph{spars} of $K$.  The cycle $vxwy$ is called the \emph{sail} of $K$.  It follows from crossing minimality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$. Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$. Observe that any spar of $K$ is incident on exactly four \emph{kite faces} of $G$, each of which has three edges and two vertices of $G$ on its boundary.

The 1-plane graph $G$ has a plane triangulation $G'$ as a subgraph that can be obtained by removing one spar from each kite in $G$.  Observe that, for any spar $vw\in E(G)\setminus E(G')$ that crosses $xy\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).  It follows that $\dist_{G'}(x,y)\le 2\dist_G(x,y)$ for any pair of vertices $x,y\in V(G)$.

\begin{lem}\lemlabel{induction} The setup:
  \begin{compactenum}
    \item Let $G$ and $G'$ be defined as above.
    \item Let $T$ be a BFS tree of $G'$ rooted at some vertex $r$.
    \item For each integer $i\ge 0$, let $V_i=\{v\in V(G):\dist_T(r,v)=i\}$. 
    \item Let $C$ be a cycle in $G'$ with $r$ in the exterior of $C$ and such that
    \begin{compactenum} 
      \item No edge of $C$ is crossed by any edge of $G$; and
      \item $V(C)$ can be partitioned into $P_1,\ldots,P_k$, $k\le 3$ such that for each $i\in\{1,\ldots,k\}$,
      \begin{compactenum}
        \item $F[P_i]$ is a path; and
        \item $|P_i\cap V_j| \le 15$ for all integers $j\ge 0$.
      \end{compactenum}
    \end{compactenum}
    \item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $C$ or the interior of $C$.
  \end{compactenum}
  Then $N$ has an $H$-partition $(B_x : x\in V(H))$ such that
  \begin{compactenum}
    \item $H$ is planar;
    \item for all integers $j\ge 0$, and all $x\in V(H)$, $|B_x\cap V_j|\le 15$; 
    \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=B_{x_i}$; and
    \item $H$ has a tree decomposition whose largest bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
  \end{compactenum}
\end{lem}

\begin{proof}
  The proof is by induction on the number of vertices of $N$.
  First note that $N'$ is a near-triangulation.  If $k=3$, set $Q_i=P_i$ for each $i\in\{1,2,3\}$.  Otherwise, as before, split $P_1,\ldots,P_k$ to partition $V(C)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $C[R_i]$ is a path and each $R_i$ contains vertices from at most one of $P_1,\ldots,P_k$.

  For each $i\in\{1,2,3\}$ and $v\in R_i$, assign the colour $i$ to $v$.
  For each $v\in V(N')$, consider the path $P_v$ that consists of the subpath of the $v$ to $r$ path in $T$ that stops at the first vertex $v'\in V(C)$. Assign $v$ to have the same colour as $v'$.

  Sperner's Lemma ensures that $G'$ contains a triangular face $\tau=v_1v_2v_3$ whose three vertices area assigned different colours. For each $i\in\{1,2,3\}$, let $Q_i=P_{v_i}$ be the path in $T$ form $v_i$ to $v_i'$.  Let $Y$ denote the subgraph if $N'$ consisting of vertices and edges $Q_1$, $Q_2$, $Q_3$, and $\tau$.  Let $Y^+$ denote the subgraph of $N$ consisting of the vertices edges of $Y$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $Y$.

  We claim that, for each integer $i\ge 0$, $|V(Y^+)\cap V_i|\le 15$.  To see this, first observe that $Y$ contains at most 3 vertices of $V_i$. If a vertex $x\in V(Y^+)\setminus V(Y)$ is contained in $V_i$, then this is because $Y$ contains an edge $vw$ with $v\in V_{i'}$ and $w\in\{V_{i'+1}$ for some $i'\in\{i-1,i\}$ and $G$ contains an edge $xy$ that crosses $vw$. (Recall that $vx,xw\in E(G')$, so $\dist_{G'}(w,r)-1\le\dist_{G'}(x,r)\le\dist_{G'}(v,r)+1$.)  The graph $Y$ contains at most 6 such edges, each of which accounts for at most 2 additional vertices of $V_i$.  Therefore, in total, $|Y^+\cap V_i|\le 15$.

  Finally, let $S$ and $S^+$ denote the subgraph of $G$ containing the edges and vertices of $Y$, respectively $Y^+$, and the edges and vertices of $C$.  The graph $S^+$ has some number of bounded faces, all contained in the interior of $C$. Some of the bounded faces of $S$ are kite faces. Call the non-kite bounded faces $F_1,\ldots,F_m$ and let $C_1,\ldots,C_m$ denote their boundaries.  We claim that, for each $i\in\{1,\ldots,m\}$, if some portion of $C_i$ is contained in an edge $vw\in E(G)$ then $vw$ is not crossed by any edge of $G$.  To see this, there are three cases to consider:
  \begin{enumerate}
    \item $vw\in E(C)$. By assumption, $vw$ is not crossed by any edge of $G$.
    \item $vw\in E(Y)$. In this case, the kite containing $vw$ is in $Y^+$, so $vw$ is only incident to kite faces.
    \item $vw\in E(Y^+)\setminus E(Y)$. In this case, either $vw$ is a sail edge in which case it is not crossed by definition, or $vw$ is a spar that was added to $Y^+$ because $vw$ crosses some edge $xy\in E(Y)$.  In this latter case, $vw$ is only incident to kite faces.
  \end{enumerate}
  Therefore each $C_i$ is a cycle in $G^+$ consisting entirely of uncrossed edges. The vertices of $C_i$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$ and $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$. Furthermore $C_i[P_j']$ is a path for each $j\in\{1,\ldots,3\}$. Finally, the subgraph $N_i$ of $G$ consisting of the edges and vertices of $G$ contained in $C_i$ or its interior does not contain one of the three vertices of $\tau$. Therefore, we can apply induction using the cycle $C_i$ and the partition $P_1',P_2',P_3'$ of $V(C_i)$ to obtain the desired $H$-partition and tree decomposition of $N_i$.

  The remainder of the proof is as before. We use $V(Y^+)\setminus V(C)$ as an element of our $H$-partition.  The root bag in the tree decomposition contains the vertices obtained by contracting each of $P_1,\ldots,P_k$ and $Y^+$.

  Planarity comes from the fact that if two edges $vw$ and $xy$ cross, then they end up in the same bag of the $H$-partition.  This means that $H$ is actually obtained by contracting connected sets of vertices in the planar graph $G'$.
\end{proof}

\begin{thm}\thmlabel{1-plane}
  Every 1-plane graph $G$ has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
\end{thm}

\begin{proof}
  The same as before except that the layered width 15 from \lemref{induction} becomes 30 because \lemref{induction} uses a layering of $G'$ and distances in $G'$ can be a factor of 2 larger than in $G$.
\end{proof}

\section{Remarks}

\begin{enumerate}
  \item The fact that $G'$ is a triangulation is not really important. We could arbitrarily triangulate $N'$ (compatible with $T$) and still find a Sperner triangle.
  \item The fact that $T$ is plane seems important since it lets us use Sperner's Lemma.  But maybe Sperner's Lemma is overkill. $Y^+$ is just a crappy separator of small layered width that happens to break $N$ into components for which no component is adjacent to more than 2 of $Q_1,\ldots,Q_3$.
  \item The fact that $C_i$ is a cycle of non-crossed edges in $G$ is not necessarily critical.  The non-crossed property isn't so important as long as no edge of $G-V(S^+)$ has endpoints in two different faces of $S^+$.

  More troublesome, again, is the use of Sperner's Lemma since it requires that $P_1,\ldots,P_k$ (and therefore also the boundary of $Y^+$) be not so wild that they interleave.  Again, maybe we can cook up something more robust than Sperner's Lemma.

  \item I think I can push through a proof for 2-plane graphs.
\end{enumerate}

\bibliographystyle{plain}
\bibliography{warmup}

\end{document}

### 2019-05-01 warmup.tex
David's comments.

@@
       \begin{compactenum}
-        \item $F[P_i]$ is a path; and
+        \item $C[P_i]$ is a path; and
         \item $|P_i\cap V_j| \le 15$ for all integers $j\ge 0$.
@@
   First note that $N'$ is a near-triangulation.
-If $k=3$, set $Q_i=P_i$ for each $i\in\{1,2,3\}$.
+If $k=3$, set $R_i=P_i$ for each $i\in\{1,2,3\}$.
 Otherwise, as before, split $P_1,\ldots,P_k$ to partition $V(C)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $C[R_i]$ is a path and each $R_i$ contains vertices from at most one of $P_1,\ldots,P_k$.
@@
 
-  Sperner's Lemma ensures that $G'$ contains a triangular face $\tau=v_1v_2v_3$ whose three vertices area assigned different colours.
+  Sperner's Lemma ensures that $G'$ contains a triangular face $\tau=v_1v_2v_3$ whose three vertices are assigned different colours.
 For each $i\in\{1,2,3\}$, let $Q_i=P_{v_i}$ be the path in $T$ form $v_i$ to $v_i'$.
-Let $Y$ denote the subgraph if $N'$ consisting of vertices and edges $Q_1$, $Q_2$, $Q_3$, and $\tau$.
-Let $Y^+$ denote the subgraph of $N$ consisting of the vertices edges of $Y$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $Y$.
+Let $Y$ denote the subgraph of $N'$ consisting of vertices and edges $Q_1$, $Q_2$, $Q_3$, and $\tau$.
+Let $Y^+$ denote the subgraph of $N$ consisting of the vertices and edges of $Y$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $Y$.
 

### 2019-05-01 warmup.tex
Fuck yeah!

@@
 \usepackage{paralist}
+
+\usepackage{tipa}
+\usepackage{upgreek}
+\usepackage{rotating}
+\newcommand{\V}{\rotatebox[origin=c]{180}{$V$}}
+\newcommand{\Y}{\rotatebox[origin=c]{180}{$Y$}}
 %\usepackage{hyperref}
@@
 \DeclareMathOperator{\dist}{dist}
+\DeclareMathOperator{\depth}{depth}
 
@@
 
+\section{$k$-Planar}
+
+  We make use of the fact that, if $G$ is a $k$-plane graph and $G^+$ is the plane graph obtained by adding a dummy vertex at each crossing in $G$, then $\dist_{G^+}(x,y)\le k\dist_G(x,y)$ for any $x,y\in V(G)$.
+
+\begin{lem}\lemlabel{k-induction} The setup:
+  \begin{compactenum}
+    \item Let $G_0$ be a $k$-plane graph.
+    \item Let $G=(V(G_0), E(G))$ be some subgraph of $G_0$ (so $E(G)\subseteq E(G_0)$).
+    \item Let $G^+$ be the plane graph obtained by adding a vertex at each crossing point in $G_0$.
+    \item Let $T$ be a spanning tree of $G^+$ rooted at some vertex $r$.
+    \item For each integer $i\ge 0$, let $V^+_i=\{v\in V(G^+):\dist_{G^+}(r,v)=i\}$, and let $V_i=V(G)\cap V^+_i$.
+    \item Let $C$ be a cycle in $G^+$ with $r$ in the exterior of $C$ and such that
+    \begin{compactenum} 
+      \item No edge of $C$ is crossed by any edge of $G$; and
+      \item $V(C)$ can be partitioned into $P_1,\ldots,P_k$, $k\le 3$ such that for each $i\in\{1,\ldots,k\}$,
+      \begin{compactenum}
+        \item $C[P_i]$ is a path; and
+        \item $|P_i\cap V_j| \le 24k+4$ for all integers $j\ge 0$.
+      \end{compactenum}
+    \end{compactenum}
+    \item Let $N$ and $N^+$ be the subgraphs of $G$ and $G^+$ consisting only of those edges and vertices contained in $C$ or the interior of $C$.
+  \end{compactenum}
+  Then $N$ has an $H$-partition $(B_x : x\in V(H))$ such that
+  \begin{compactenum}
+    \item for every integer $j\ge 0$, and all $x\in V(H)$, $|B_x\cap V_j|\le 24k+4$; 
+    \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=B_{x_i}$; and
+    \item $H$ has a tree decomposition whose largest bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
+  \end{compactenum}
+\end{lem}
+
+\begin{proof}
+  The proof is by induction on the number of vertices of $N^+$.
+
+  If $k=3$, set $R_i=P_i$ for each $i\in\{1,2,3\}$.
+Otherwise, as before, split $P_1,\ldots,P_k$ to partition $V(C)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $C[R_i]$ is a path and each $R_i$ contains vertices from at most one of $P_1,\ldots,P_k$.
+
+  Next, we 3-colour $V(N^+)$ so that we can apply Sperner's Lemma.
+  For each $i\in\{1,2,3\}$ and $v\in R_i$, assign the colour $i$ to $v$.
+  For each $v\in V(N^+)$, consider the path $P_v$ that consists of the subpath of the $v$ to $r$ path in $T$ that stops at the first vertex $v'\in V(C)$. Assign $v$ to have the same colour as $v'$.
+
+  Add edges to $N^+$ to obtain a near-triangulation $G^\Delta$.  
+  Sperner's Lemma ensures that $G^\Delta$ contains a triangular face $\tau=v_1v_2v_3$ whose three vertices are assigned different colours.
+For each $i\in\{1,2,3\}$, let $Q_i^+=P_{v_i}^+$ be the path in $T$ from $v_i$ to $v_i'$.
+Let $Y^+$ denote the subgraph of $N^+$ consisting of vertices and edges of $Q_1^+$, $Q_2^+$, $Q_3^+$, and $\tau$.
+
+  Define a vertex set $X\subset V(G)$ as follows:  The set $X$ contains $V(Y^+)\cap V(G)\setminus V(C)$.
+Next, for each vertex $v^+\in V(Y^+)\setminus V(G)$, we consider the two edges $vw,xy\in E(G_0)$ whose intersection is $v^+$.
+Next we set
+  \[   X_{vw} = \begin{cases} \{v,w\} & \text{if $vw\in E(G)$} \\
+           \emptyset & \text{if $vw\not\in E(G)$,}
+         \end{cases}
+  \]
+  \[   X_{xy} = \begin{cases} \{x,y\} & \text{if $xy\in E(G)$} \\
+           \emptyset & \text{if $xy\not\in E(G)$} 
+         \end{cases}
+  \]
+  and add $(X_{vw}\cup X_{xy})\setminus V(C)$ to $X$.
+
+  We claim that, for each integer $i\ge 0$, $|V_i\cap X|\le 24k+4$.
+To see why this is so, first note that $|V_i^+\cap V(Y^+)|\le 3$ since $V_i^+$ contains at most one vertex from each of $Q_1^+$, $Q_2^+$, and $Q_3^+$.
+For any vertex $v\in X$, there exists a vertex $v^+\in V(Y^+)$ such that $\dist_{G'}(v,v^+)\le k$.
+Turning this around, we can assign each vertex $v\in X$ to a vertex $v^+\in V(Y^+)$ in such a way that each vertex of $Y^+$ is assigned at most 4 vertices of $X$ (each dummy vertex $v^+$ is assigned the four endpoints of the crossing pair of edges that generated $v^+$).
+Since any vertex $v\in X\cap V_i$ is assigned to one of the at most $6(k+1)$ vertices in $X\cap (V(Y^+)\cap( V_{i-k}^+\cup\cdots\cup V_{i+k}^+))$, we have $|X\cap V_i|\le 24k+4$.
+
+  Let $S^+$ denote the subgraph of $N^+$ consisting of the vertices and edges of $Y^+$ and the vertices and edges of $C$.
+The graph $S^+$ is planar.
+It contains one triangular face $\tau$ and up to three additional bounded faces $F_1$, $F_2$, and $F_3$, where each $F_i$ does not have any vertices in common with $P_i$.
+
+  For each $i\in\{1,2,3\}$, let $Z_i$ be the subset of $V(G)$ that is contained in the interior of $F_i$.
+We claim that $N-X-V(C)$ does not contain any edge $vw$ with $v\in Z_i$ and $w\in Z_j$ for $i\neq j$.
+Indeed, any such edge would contain a vertex of $V(Y^+)\setminus V(G)$ in its interior and therefore each of $v,w\in X\cup V(C)$.
+
+  The set $X$ will become an element $B_x$ in our $H$-partition of $G$.
+All that remains is to apply induction on three subgraphs.
+For each $i\in\{1,\ldots,3\}$, let $N_i^+$ denote the subgraph of $G^+$ containing the vertices and edges of $G^+$ that are contained in the closure of $F_i$.
+  Let $G_i^-$ be the subgraph of $G$ obtained by removing the edges of $G$ that cross the boundary of $F_i$.  
+  Let $N_i$ denote the subgraph of $G_i^-$ containing the vertices and edges of $G_i^-$ that are contained in the closure of $F_i$.
+Note that $G_i^-$ is a subgraph of $G_0$ and that the number of vertices of $N_i^+$ is less than the number of vertices of $N^+$ since $N_i^+$ does not contain one of the vertices of $\tau$.
+Therefore, we can apply induction to obtain the desired $H$-partition of $N_i$ using the original graphs $G_0$ and $G^+$, the subgraph $G_i^-$ in place of $G$, the cycle $F_i$ in place of $C$, and partition of $V(F_i)$ into $V(F_i)\cap X$, $V(F_i)\cap P_{i-1}$, and $V(F_i)\cap P_{i+1}$.
+
+  The remainder of the proof is as before.
+We use $X$ as an element of our $H$-partition.
+The root bag in the tree decomposition contains the vertices obtained by contracting each of $P_1,\ldots,P_k$ and $X$.
+\end{proof}
+
+\begin{thm}\thmlabel{1-plane}
+  Every $k$-plane graph $G$ has a $(3, 24k^2+4k)$-partition.
+\end{thm}
+
+\begin{proof}
+  The same as before except that the layered width $24k+4$ from \lemref{k-induction} becomes $24k^2+4k$ because \lemref{k-induction} uses a layering of $G^+$ and distances in $G^+$ can be a factor of k larger than in $G$.
+\end{proof}
+
+% Let $G$ be a $k$-planar graph with a distinguished vertex $r$.
+For each integer $i\ge 0$, let $V_i=\{v\in V(G):\dist_G(r,v)=i\}$ and let $V_{\le i}=\bigcup_{j=0}^i V_j$.
+We will use distances to $r$ frequently, so for each $v\in V(G)$, we let $d_v=\dist_G(r,v)$ (so that $v\in V_{d_v}$).
+We say that a subgraph $P$ of $G$ has layered width at most $\ell$ if $|V_i\cap V(P)|\le \ell$ for each integer $i\ge 0$.
+% 
+% Let $P_1,\ldots,P_3$ be three vertex-disjoint subgraphs of $G$, each of which has layered width at most $\ell$ and none of which contains $r$.
+Let $B$ be the union of $P_1,\ldots,P_3$.
+Let $N$ be a component of $G-V(B)$ that does not contain $r$.
+Let $G'$ be the subgraph of $G$ consisting of edges and vertices of $N$, edges and vertices of $B$, and all edges $vw\in E(G)$ with $v\in V(N)$ and $w\in V(B)$.
+% 
+% \begin{lem}
+%   There exists a vertex $v\in V(G')$ and three vertex-disjoint paths $Q_1$, $Q_2$, and $Q_3$ where each $Q_i$ begins at $v$, has all of its internal vertices in $V(N)$, ends at a vertex in $V(P_i)$ and has layered width at most 10.
+% \end{lem}
+% 
+% \begin{proof}
+%   For simplicity, assume that every vertex of $P_i$ is adjacent to at least one vertex of $V(N)$ since we can ignore vertices of $P_i$ for which this is not the case.  
+% 
+%   For each integer $r\ge 0$, consider the induced subgraph $G_r'=G'[V(B)\cup V_{\le r}]$.
+Let $r$ be the smallest integer such that $G_{r}'$ has at least one component $A$ that contains vertices from each of $P_1$, $P_2$, and $P_3$.
+Select any vertex $w\in V(A)\setminus V_{<r}$.
+Suppose, for now that $w\not\in V(B)$.
+Then, since no component of $V_{r-1}$ contains vertice from each of $P_1$, $P_2$, and $P_3$, it must be the case that $G_{r-1}'$ has components $A_1$, $A_2$, and $A_3$ where $w$ is adjacent to some vertex in each $A_i$, each $A_i$ contains a vertex of $P_i$, $A_1\not\in\{A_2,A_3\}$ and $A_1$ does not contain a vertex of $P_3$.
+(Note that we may have $A_2=A_3$ and we may have $A_1$ containig vertices of $P_2$.)
+% 
+%   If $A_2\neq A_3$, then we are done.
+We set $v=w$ and, for each $i\in \{1,\ldots,3\}$ we take a path $Q_i$ from $v$ to $V(P_i)$ whose $j$th vertex is in $V_{r-j+1}\cap V(A_i)$.
+These three paths are disjoint since $V(Q_i)\subseteq V(A_i)$ for each $i\in \{1,\ldots,3\}$.
+% 
+%   If $A_2=A_3$, there is still a little more work to do.  In this case we will still use the path $Q_1$ from $w$ to $V(P_1)$.  Now, $w$ is adjacent to a vertex $w_1$ in $A_2=A_3$ so we extend
+% 
+% 
+% 
+% 
+% 
+%    where (without loss of generality) $V(A_1)\cap V(P_1)\neq \emptyset$, $V(A_1)\cap (V(P_2)\cup V(P_3))=\emptyset$, $V(A_2)\cap V(P_2)\neq\emptyset$, $V(A_2)\cap V(P_1)=\emptyset$ and $w$ is adjacent to a vertex $w_i\in A_i$ for each $i\in\{1,2\}$.
+% 
+%   If 
+% 
+% 
+% 
+% 
+% 
+%    contains vertices of $P_1$, $A_2$ contains vertices of $P_2$ and (possibly )
+% 
+% 
+% 
+% 
+%   We claim that $v$ has the desired property.  Indeed
+% 
+% \end{proof}
+% 
+% 
+% 
+% 
+% 
+% \section{A Metric Alternative To Sperner's Lemma}
+% 
+% I'm interested in having something more robust and less magical than Sperner's Lemma. Here's a first attempt.
+% 
+% Let $G$ be a planar triangulation with some distinguished vertex $r$. 
+% 
+% 
+% 
+% 
+% A \emph{$t$-pod} (with respect to $r$) consists of vertex-disjoint paths $Q_1,\ldots,Q_t$, where each $Q_i$ a prefix of some $v_i\leadsto r$ path in $G$ and $v_1,\ldots,v_t$ form a clique in $G$.
+We call $v_1,\ldots,v_t$ the \emph{central vertices} of the tripod and we call $Q_1,\ldots,Q_t$ the \emph{legs} of the tripod.
+The final vertex in each path $Q_i$ is called \emph{foot} if the tripod.
+% 
+% A $t$-pod $\Y$ with $t\le 3$, is called a \emph{tripod}.  A $t$-pod $\V$ with $t\le 2$ is called a \emph{bipod}.  A $t$-pod $I$ with $t\le 1$ is called a \emph{monopod}.
+% 
+% For a $t$-pod $Y$ with central vertices $v_1,\ldots,v_t$, we define the \emph{depth} of $Y$ as 
+% \[
+%     \depth(Y)=\max\{d_G(r,v):v\in V(T)\} 
+%         = \max\{\dist_G(r, v_i): i\in\{1,\ldots,t\}\} \enspace .
+% \]
+% Observe that, since $v_1,\ldots,v_t$ form a clique in $G$, $\depth(Y)-1\le \dist_G(r, v_i) \le \depth(Y)$.
+% 
+% Let $C$ be a cycle in $G$ and $P_1,\ldots,P_a$, $a\le 3$ be a partition of $V(C)$ such that $C[P_i]$ is a path and $C[P_i]$ is also a bipod $\V_i$ in $G$.
+Let $Q_{i,1}$ and (possibly) $Q_{i,2}$ be the legs of $\V_i$ where we use the convention that, for each $i\in\{1,\ldots,a\}$, the last vertex of $Q_{i,2}$ and the first vertex of $Q_{i+1,1}$ are adjacent in $C$.
+Let $v_{i,j}$ be first vertex of $Q_{i,j}$ and let $v_{i,j}'$ be the last vertex of $Q_{i,j}$.
+With these definitions, $v_{i,j}$ is a central vertex of $\V_i$ and $v_{i,j}'$ is a foot of $\V_{i}$.
+% 
+% Note that, for each $i\in\{1,\ldots,a\}$, let $d_i = dist_{G}(r,\{v_{i,2},v_{i+1,1}'\})$.  
+% 
+% 
+% 
+% Let $N$ be the near-triangulation consisting of all edges and vertices of $G$ that are on or in the ineterior of $C$.  
+%
+
 \section{Remarks}

### 2019-05-01 warmup.tex
Small touch-ups

@@
 \begin{abstract}
-  These are some notes on our work of layered $H$-partitions for $k$-planar graphs
+  We prove that every $k$-planar graph has a layered $H$-partition of layered width $k(24k+4)$ in which $H$ has treewidth 3.
+This implies that $k$-planar graphs have a queue number upper-bounded by a polynomial in $k$ and have non-repetitive chromatic number upper-bounded by an exponential function of $k$.
 \end{abstract}
@@
   \begin{compactenum}
-    \item Let $G_0$ be a $k$-plane graph.
+    \item Let $G_0$ be a connected $k$-plane graph.
     \item Let $G=(V(G_0), E(G))$ be some subgraph of $G_0$ (so $E(G)\subseteq E(G_0)$).
-    \item Let $G^+$ be the plane graph obtained by adding a vertex at each crossing point in $G_0$.
-    \item Let $T$ be a spanning tree of $G^+$ rooted at some vertex $r$.
+    \item Let $G^+$ be the triangulation obtained by adding a vertex at each crossing point in $G_0$ and then adding additional edges to obtain a triangulation.
+    \item Let $T$ be a BFS spanning tree of $G^+$ rooted at some vertex $r$.
     \item For each integer $i\ge 0$, let $V^+_i=\{v\in V(G^+):\dist_{G^+}(r,v)=i\}$, and let $V_i=V(G)\cap V^+_i$.
@@
 \begin{proof}
+  We remark that $N^+$ is a near-triangulation.
   The proof is by induction on the number of vertices of $N^+$.
@@
 
-  Add edges to $N^+$ to obtain a near-triangulation $G^\Delta$.  
-  Sperner's Lemma ensures that $G^\Delta$ contains a triangular face $\tau=v_1v_2v_3$ whose three vertices are assigned different colours.
+  Sperner's Lemma ensures that $N^+$ contains a triangular face $\tau=v_1v_2v_3$ whose three vertices are assigned different colours.
 For each $i\in\{1,2,3\}$, let $Q_i^+=P_{v_i}^+$ be the path in $T$ from $v_i$ to $v_i'$.
@@
   For each $i\in\{1,2,3\}$, let $Z_i$ be the subset of $V(G)$ that is contained in the interior of $F_i$.
-We claim that $N-X-V(C)$ does not contain any edge $vw$ with $v\in Z_i$ and $w\in Z_j$ for $i\neq j$.
+We claim that $N-X-V(C)$ does not contain any edge $vw\in E(G)$ with $v\in Z_i$ and $w\in Z_j$ for $i\neq j$.
 Indeed, any such edge would contain a vertex of $V(Y^+)\setminus V(G)$ in its interior and therefore each of $v,w\in X\cup V(C)$.
@@
 For each $i\in\{1,\ldots,3\}$, let $N_i^+$ denote the subgraph of $G^+$ containing the vertices and edges of $G^+$ that are contained in the closure of $F_i$.
-  Let $G_i^-$ be the subgraph of $G$ obtained by removing the edges of $G$ that cross the boundary of $F_i$.  
-  Let $N_i$ denote the subgraph of $G_i^-$ containing the vertices and edges of $G_i^-$ that are contained in the closure of $F_i$.
-Note that $G_i^-$ is a subgraph of $G_0$ and that the number of vertices of $N_i^+$ is less than the number of vertices of $N^+$ since $N_i^+$ does not contain one of the vertices of $\tau$.
-Therefore, we can apply induction to obtain the desired $H$-partition of $N_i$ using the original graphs $G_0$ and $G^+$, the subgraph $G_i^-$ in place of $G$, the cycle $F_i$ in place of $C$, and partition of $V(F_i)$ into $V(F_i)\cap X$, $V(F_i)\cap P_{i-1}$, and $V(F_i)\cap P_{i+1}$.
+  Let $G_i$ be the subgraph of $G$ obtained by removing the edges of $G$ that cross the boundary of $F_i$.  
+  Let $N_i$ denote the subgraph of $G_i$ containing the vertices and edges of $G_i$ that are contained in the closure of $F_i$.
+Note that $G_i$ is a subgraph of $G_0$ and that the number of vertices of $N_i^+$ is less than the number of vertices of $N^+$ since $N_i^+$ does not contain one of the vertices of $\tau$.
+Therefore, we can apply induction to obtain the desired $H$-partition of $N_i$ using the original graphs $G_0$ and $G^+$, the subgraph $G_i$ in place of $G$, the cycle $F_i$ in place of $C$, and partition of $V(F_i)$ into $V(F_i)\cap X$, $V(F_i)\cap P_{i-1}$, and $V(F_i)\cap P_{i+1}$.
 
@@
 % 
-
-\section{Remarks}
-
-\begin{enumerate}
-  \item The fact that $G'$ is a triangulation is not really important. We could arbitrarily triangulate $N'$ (compatible with $T$) and still find a Sperner triangle.
-  \item The fact that $T$ is plane seems important since it lets us use Sperner's Lemma.
+% 
+% \section{Remarks}
+% 
+% \begin{enumerate}
+%   \item The fact that $G'$ is a triangulation is not really important. We could arbitrarily triangulate $N'$ (compatible with $T$) and still find a Sperner triangle.
+%   \item The fact that $T$ is plane seems important since it lets us use Sperner's Lemma.
 But maybe Sperner's Lemma is overkill.
 $Y^+$ is just a crappy separator of small layered width that happens to break $N$ into components for which no component is adjacent to more than 2 of $Q_1,\ldots,Q_3$.
-  \item The fact that $C_i$ is a cycle of non-crossed edges in $G$ is not necessarily critical.
+%   \item The fact that $C_i$ is a cycle of non-crossed edges in $G$ is not necessarily critical.
 The non-crossed property isn't so important as long as no edge of $G-V(S^+)$ has endpoints in two different faces of $S^+$.
-
-  More troublesome, again, is the use of Sperner's Lemma since it requires that $P_1,\ldots,P_k$ (and therefore also the boundary of $Y^+$) be not so wild that they interleave.
+% 
+%   More troublesome, again, is the use of Sperner's Lemma since it requires that $P_1,\ldots,P_k$ (and therefore also the boundary of $Y^+$) be not so wild that they interleave.
 Again, maybe we can cook up something more robust than Sperner's Lemma.
-
-  \item I think I can push through a proof for 2-plane graphs.
-\end{enumerate}
+% 
+%   \item I think I can push through a proof for 2-plane graphs.
+% \end{enumerate}
 

### 2019-05-02 warmup.tex
Tiny change
Fixing....

@@
 
-  We make use of the fact that, if $G$ is a $k$-plane graph and $G^+$ is the plane graph obtained by adding a dummy vertex at each crossing in $G$, then $\dist_{G^+}(x,y)\le k\dist_G(x,y)$ for any $x,y\in V(G)$.
+Let $G$ be a $k$-planar graph.
+We will use of the \emph{plane augmentation} $G^+$ of $G$ obtained by adding a \emph{dummy vertex} at each crossing in $G$, then adding a \emph{star vertex} $v_f$ in each face $f$ of the resulting graph and making $v_f$ adjacent to each vertex of $f$.
+
+We will make use of the fact that $\dist_{G^+}(x,y)\le k\cdot\dist_G(x,y)$ for any $x,y\in V(G)$.
+This follows immediately from the fact that any edge $xy\in E(G)$ has a corresponding path $x-y$ in $G^+$ of length at most $k$.
+We will also make use of the fact that every dummy vertex of $G^+$ has degree 8.
 
@@
   \begin{compactenum}
-    \item Let $G_0$ be a connected $k$-plane graph.
-    \item Let $G=(V(G_0), E(G))$ be some subgraph of $G_0$ (so $E(G)\subseteq E(G_0)$).
-    \item Let $G^+$ be the triangulation obtained by adding a vertex at each crossing point in $G_0$ and then adding additional edges to obtain a triangulation.
+    \item Let $G$ be $k$-plane graph.
+    \item Let $G^+$ be the plane augmentation of $G$.
     \item Let $T$ be a BFS spanning tree of $G^+$ rooted at some vertex $r$.
@@
   Define a vertex set $X\subset V(G)$ as follows:  The set $X$ contains $V(Y^+)\cap V(G)\setminus V(C)$.
-Next, for each vertex $v^+\in V(Y^+)\setminus V(G)$, we consider the two edges $vw,xy\in E(G_0)$ whose intersection is $v^+$.
-Next we set
-  \[   X_{vw} = \begin{cases} \{v,w\} & \text{if $vw\in E(G)$} \\
-           \emptyset & \text{if $vw\not\in E(G)$,}
-         \end{cases}
-  \]
-  \[   X_{xy} = \begin{cases} \{x,y\} & \text{if $xy\in E(G)$} \\
-           \emptyset & \text{if $xy\not\in E(G)$} 
-         \end{cases}
-  \]
-  and add $(X_{vw}\cup X_{xy})\setminus V(C)$ to $X$.
+Next, for each dummy vertex $v^+\in V(Y^+)\setminus V(G)$, we consider the two edges $vw,xy\in E(G_0)$ whose intersection is $v^+$.
+These edges correspond to two paths $v\rightsquigarrow w$ and $x\rightsquigarrow y$.
+We collect the set $A_{x^+}$ of all vertices of $G^+$ that are on any face that contains any dummy vertex in $v\rightsquigarrow w$ or $x\rightsquigarrow y$.
+We add the set $A_{x^+}\cap V(N^+)\setminus V(C)$ to $X$.
+
+  Note that $A_{x^+}$ contains $\{v,w,x,y\}$ and that, for any vertex $z\in A_{x^+}$, $\dist_{G^+}(z,x^+) \le k$.  Careful counting also reveals that the size of $A_{x^+}$ is at most .
 

### 2019-05-02 k-planar.tex
Working version

\documentclass{patmorin}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{pat}
\usepackage{paralist}

\usepackage{tipa}
\usepackage{upgreek}
\usepackage{rotating}
\newcommand{\V}{\rotatebox[origin=c]{180}{$V$}}
\newcommand{\Y}{\rotatebox[origin=c]{180}{$Y$}}
%\usepackage{hyperref}

\setlength{\parskip}{1ex}

\title{\MakeUppercase{Notes on Layered $H$-Partitions of $k$-Planar Graphs}%
    \thanks{This work was partly funded by NSERC and MRI.}}

\author{Vida Dujmovi\'c%
        \thanks{School of Computer Science and Electrical Engineering, 
                University of Ottawa},\,\, 
        Pat Morin%
        \thanks{School of Computer Science, Carleton University},\,\, and
        David R. Wood%
        \thanks{School of Mathematical Sciences, Monash University}}

\newcommand{\dual}[1]{{#1}^\star}
\newcommand{\note}[2]{{\color{red}[#1:~#2]}}

\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\depth}{depth}
\DeclareMathOperator{\ff}{f}
\newcommand{\f}{{\color{red}\ff}}

\newcommand{\fk}{{\color{red}f(k)}}

\begin{document}
\maketitle

\begin{abstract}
  TBD
  % Layered $H$-partitions of graphs (with small layered width, in which $H$ has small treewidth) are a recently-introduced tool that have been used to solve longstanding problems on queue layouts, non-repetitive colouring, and 3-d graph drawing.  Such partitions are known to exist for planar graphs and, more generally, bounded genus graphs.
  % In the current paper, we prove that every $k$-planar graph has a layered $H$-partition of layered width $O(k^3)$ in which $H$ is planar and has treewidth 3. This implies that $k$-planar graphs have a queue number upper-bounded by a polynomial in $k$ and have non-repetitive chromatic number upper-bounded by an exponential function of $k$.
\end{abstract}

\section{Introduction}

% The remainder of the paper is organized as follows:  \secref{1-planar} is a warm-up and gives the proof for 1-plane graphs.  \secref{k-planar} presents the proof for $k$-plane graphs.
% 
% 
% \section{$1$-Plane Graphs}
% \seclabel{1-planar}
% 
% Let $G$ be an edge maximal 1-plane graph.  Here, edge-maximal should be taken to mean that any two vertices of $G$ on a common face have an edge joining them.  Assume also that $G$ is crossing minimal so that the embedding of $G$ minimizes the number of crossings over all possible embeddings of the (abstract) graph $G$.  From now on, we treat vertices of $G$ as points and edges as simple curves.
% 
% A \emph{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.  It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
% The edges $vw$ and $xy$ are called \emph{spars} of $K$.  The cycle $vxwy$ is called the \emph{sail} of $K$.  It follows from crossing minimality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$. Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$. Observe that any spar of $K$ is incident on exactly four \emph{kite faces} of $G$, each of which has three edges and two vertices of $G$ on its boundary.
% 
% The 1-plane graph $G$ has a plane triangulation $G'$ as a subgraph that can be obtained by removing one spar from each kite in $G$.  Observe that, for any spar $vw\in E(G)\setminus E(G')$ that crosses $xy\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).  It follows that $\dist_{G'}(x,y)\le 2\dist_G(x,y)$ for any pair of vertices $x,y\in V(G)$.
% 
% \begin{lem}\lemlabel{induction} The setup:
%   \begin{compactenum}
%     \item Let $G$ and $G'$ be defined as above.
%     \item Let $T$ be a BFS tree of $G'$ rooted at some vertex $r$.
%     \item For each integer $i\ge 0$, let $V_i=\{v\in V(G):\dist_T(r,v)=i\}$. 
%     \item Let $C$ be a cycle in $G'$ with $r$ in the exterior of $C$ and such that
%     \begin{compactenum} 
%       \item No edge of $C$ is crossed by any edge of $G$; and
%       \item $V(C)$ can be partitioned into $P_1,\ldots,P_k$, $k\le 3$ such that for each $i\in\{1,\ldots,k\}$,
%       \begin{compactenum}
%         \item $C[P_i]$ is a path; and
%         \item $|P_i\cap V_j| \le 15$ for all integers $j\ge 0$.
%       \end{compactenum}
%     \end{compactenum}
%     \item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $C$ or the interior of $C$.
%   \end{compactenum}
%   Then $N$ has an $H$-partition $(B_x : x\in V(H))$ such that
%   \begin{compactenum}
%     \item $H$ is planar;
%     \item for all integers $j\ge 0$, and all $x\in V(H)$, $|B_x\cap V_j|\le 15$; 
%     \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=B_{x_i}$; and
%     \item $H$ has a tree decomposition whose largest bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
%   \end{compactenum}
% \end{lem}
% 
% \begin{proof}
%   The proof is by induction on the number of vertices of $N$.
%   First note that $N'$ is a near-triangulation.  If $k=3$, set $R_i=P_i$ for each $i\in\{1,2,3\}$.  Otherwise, as before, split $P_1,\ldots,P_k$ to partition $V(C)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $C[R_i]$ is a path and each $R_i$ contains vertices from at most one of $P_1,\ldots,P_k$. 
% 
%   For each $i\in\{1,2,3\}$ and $v\in R_i$, assign the colour $i$ to $v$.
%   For each $v\in V(N')$, consider the path $P_v$ that consists of the subpath of the $v$ to $r$ path in $T$ that stops at the first vertex $v'\in V(C)$. Assign $v$ to have the same colour as $v'$.
% 
%   Sperner's Lemma ensures that $G'$ contains a triangular face $\tau=v_1v_2v_3$ whose three vertices are assigned different colours. For each $i\in\{1,2,3\}$, let $Q_i=P_{v_i}$ be the path in $T$ form $v_i$ to $v_i'$.  Let $Y$ denote the subgraph of $N'$ consisting of vertices and edges $Q_1$, $Q_2$, $Q_3$, and $\tau$.  Let $Y^+$ denote the subgraph of $N$ consisting of the vertices and edges of $Y$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $Y$.
% 
%   We claim that, for each integer $i\ge 0$, $|V(Y^+)\cap V_i|\le 15$.  To see this, first observe that $Y$ contains at most 3 vertices of $V_i$. If a vertex $x\in V(Y^+)\setminus V(Y)$ is contained in $V_i$, then this is because $Y$ contains an edge $vw$ with $v\in V_{i'}$ and $w\in\{V_{i'+1}$ for some $i'\in\{i-1,i\}$ and $G$ contains an edge $xy$ that crosses $vw$. (Recall that $vx,xw\in E(G')$, so $\dist_{G'}(w,r)-1\le\dist_{G'}(x,r)\le\dist_{G'}(v,r)+1$.)  The graph $Y$ contains at most 6 such edges, each of which accounts for at most 2 additional vertices of $V_i$.  Therefore, in total, $|Y^+\cap V_i|\le 15$.
% 
% 
%   Finally, let $S$ and $S^+$ denote the subgraph of $G$ containing the edges and vertices of $Y$, respectively $Y^+$, and the edges and vertices of $C$.  The graph $S^+$ has some number of bounded faces, all contained in the interior of $C$. Some of the bounded faces of $S$ are kite faces. Call the non-kite bounded faces $F_1,\ldots,F_m$ and let $C_1,\ldots,C_m$ denote their boundaries.  We claim that, for each $i\in\{1,\ldots,m\}$, if some portion of $C_i$ is contained in an edge $vw\in E(G)$ then $vw$ is not crossed by any edge of $G$.  To see this, there are three cases to consider:
%   \begin{enumerate}
%     \item $vw\in E(C)$. By assumption, $vw$ is not crossed by any edge of $G$.
%     \item $vw\in E(Y)$. In this case, the kite containing $vw$ is in $Y^+$, so $vw$ is only incident to kite faces.
%     \item $vw\in E(Y^+)\setminus E(Y)$. In this case, either $vw$ is a sail edge in which case it is not crossed by definition, or $vw$ is a spar that was added to $Y^+$ because $vw$ crosses some edge $xy\in E(Y)$.  In this latter case, $vw$ is only incident to kite faces.
%   \end{enumerate}
%   Therefore each $C_i$ is a cycle in $G^+$ consisting entirely of uncrossed edges. The vertices of $C_i$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$ and $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$. Furthermore $C_i[P_j']$ is a path for each $j\in\{1,\ldots,3\}$. Finally, the subgraph $N_i$ of $G$ consisting of the edges and vertices of $G$ contained in $C_i$ or its interior does not contain one of the three vertices of $\tau$. Therefore, we can apply induction using the cycle $C_i$ and the partition $P_1',P_2',P_3'$ of $V(C_i)$ to obtain the desired $H$-partition and tree decomposition of $N_i$.
% 
%   The remainder of the proof is as before. We use $V(Y^+)\setminus V(C)$ as an element of our $H$-partition.  The root bag in the tree decomposition contains the vertices obtained by contracting each of $P_1,\ldots,P_k$ and $Y^+$.
% 
%   Planarity comes from the fact that if two edges $vw$ and $xy$ cross, then they end up in the same bag of the $H$-partition.  This means that $H$ is actually obtained by contracting connected sets of vertices in the planar graph $G'$.
% \end{proof}
% 
% \begin{thm}\thmlabel{1-plane}
%   Every 1-plane graph $G$ has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
% \end{thm}
% 
% \begin{proof}
%   The same as before except that the layered width 15 from \lemref{induction} becomes 30 because \lemref{induction} uses a layering of $G'$ and distances in $G'$ can be a factor of 2 larger than in $G$.
% \end{proof}

\section{$k$-Planar}
\seclabel{k-planar}

Let $G$ be a $k$-planar graph.  We will use of the \emph{plane augmentation} $G^+$ of $G$ obtained by adding a \emph{dummy vertex} at each crossing in $G$
and then adding edges to the resulting graph to obtain a triangulation.
% 
% triangulating the resulting graph arbitrarily.
% 
% 
%  then adding a \emph{star vertex} $v_f$ in each face $f$ of the resulting graph and making $v_f$ adjacent to each vertex of $f$.

We will make use of the fact that $\dist_{G^+}(x,y)\le k\cdot\dist_G(x,y)$ for any $x,y\in V(G)$.  This follows immediately from the fact that any edge $xy\in E(G)$ has a corresponding path $x\to y$ in $G^+$ of length at most $k$.

The \emph{dummy graph} $G^-$ is the subgraph of $G^+$ induced by the dummy vertices.  Note that the dummy graph has maximum degree 4. This immediately implies the following:

\begin{lem}\lemlabel{small-balls}
  Let $G$ be a $k$-plane graph, let $G^+$ be the plane augmentation of $G$, and let $G^-=G^+-V(G)$.  Then, for any vertex $v\in G^-$, $|\{w\in V(G^-): \dist_{G^-}(v,w)\le k\}|\le 4\cdot 3^{k-1}$.
\end{lem}

Before reading the statement of the following lemma it is worthwhile looking at \figref{M} to understand the role of some of the elements (in particular, $Y^{++}$ and $D^+$).

\begin{lem}\lemlabel{k-induction} The setup:
  \begin{compactenum}
    \item Let $G$ be a $k$-plane graph, let $G^+$ be the plane augmentation of $G$, and let $G^-$ be the dummy graph of $G$.
    \item Let $T$ be a BFS spanning tree of $G^+$ rooted at some vertex $r$.
    \item For each integer $i\ge 0$, let $V^+_i=\{v\in V(G^+):\dist_{G^+}(r,v)=i\}$, and let $V_i=V(G)\cap V^+_i$.
    \item Let $F$ be a cycle in $G^+$ that has $r$ in its exterior.
    \item Let $N^+$ denote the near-triangulation consisting of the edges and vertices of $G^+$ contained in $F$ or the interior of $F$
    \item Let $D^+\subset V(N^+)$ have the following properties:
    \begin{compactenum}
      \item $V(F)\subset D^+$.
      \item For any dummy vertex $x\in V(F)$ and any vertex $y\in V(N^+)\setminus D^+$,  $\dist_{G^-}(x,y) \ge k$.
      \item $D^+$ can be partitioned into $P_1,\ldots,P_k$, $k\le 3$ such that for each $i\in\{1,\ldots,k\}$,
      \begin{compactenum}
        \item $F[P_i]$ is connected; and
        \item $|P_i\cap V_j| \le \fk$ for all integers $j\ge 0$.
      \end{compactenum}
    \end{compactenum}
    \item Let $N$ be the subgraph of $G$ induced by $V(F)\cap V(G)$ plus the vertices of $G$ in the interior of $F$.
  \end{compactenum}
  Then $N$ has an $H$-partition $(B_x : x\in V(H))$ such that
  \begin{compactenum}
    \item for every integer $j\ge 0$, and all $x\in V(H)$, $|B_x\cap V_j|\le \fk$; 
    \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=B_{x_i}$; and
    \item $H$ has a tree decomposition whose largest bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
  \end{compactenum}
\end{lem}

\begin{proof}
  Let $N^+$ be the subgraph of $G^+$ consisting only of those edges and vertices of $G^+$ contained in $F$ or in the interior of $F$.
  We remark that $N^+$ is a near-triangulation.

  The proof is by induction on $|V(N^+)|$.  If $|V(N^+)| = 3$, then the graph $H$ is the complete graph on integers $1,\ldots,k$ and the partition sets $B_{i}=P_i\cap V(G)$ for each $i\in\{1,\ldots,k\}$.

  Since $P_1,\ldots,P_k$ is a partition of $D^+$ and $V(N)\subseteq V(D^+)$, every vertex $v\in V(N)$ is contained in some set $P_i$ and is therefore contained in some bag $B_{i}$ of the $H$-partition.  Every edge $vw\in E(N)$ has $v\in B_{i}$ and $w\in B_{j}$ for some $i,j\in\{1,\ldots,k\}$. Either $i=j$ or the edge $ij\in E(H)$.  Therefore $(B_x : x\in V(H))$ is indeed an $H$-partition of $N$.

  Requirement~1 of the $H$-partition follows from Condition~6.c.ii.  To satisfy Requirements~2 and 3 we take the tree decomposition of $H$ to consist of a single node $u$ whose bag contains $\{1,\ldots,k\}$.  This completes the description of how the base case $|V(N^+)|=3$ is handled. From this point on, we assume $|V(N^+)|> 3$.

  We want to 3-colour $V(N^+)$ so that we can apply Sperner's Lemma, which means we need to partition $V(F)$ into sets $R_1,\ldots,R_3$ such that $F[R_i]$ is a path for each $i\in\{1,\ldots,3\}$.
  If $k=3$, we set $R_i=P_i\cap V(F)$ for each $i\in\{1,2,3\}$.  Condition~6.c.i ensures that $F[R_i]$ is a path, for each $i\in\{1,\ldots,3\}$.  If $k\in\{1,2\}$, then we partition $(P_1\cap V(F))\cup\cdots\cup(P_k\cap V(F))$ into three sets $R_1$, $R_2$, and $R_3$ such that, for each $i\in\{1,\ldots,3\}$, $F[R_i]$ is a path $R_i\subseteq P_j$ for some $j\in\{1,\ldots,k\}$. Again, Condition~6.c.i ensures that this is always possible.

  We now describe the 3-colouring of $V(N^+)$ that allows us to apply Sperner's Lemma. For each $i\in\{1,2,3\}$ and $v\in R_i$, assign the colour $i$ to $v$.
  For each $v\in V(N^+)$, consider the path $P_v$ that consists of the subpath of the $v$ to $r$ path in $T$ that stops at the first vertex $v'\in V(F)$. Assign $v$ to have the same colour as $v'$.

  Sperner's Lemma ensures that $N^+$ contains a triangular face $\tau=v_1v_2v_3$ whose three vertices are assigned different colours. For each $i\in\{1,2,3\}$, let $Q_i^+=P_{v_i}^+$ be the path in $T$ from $v_i$ to $v_i'$.  Let $Y^+$ denote the subgraph of $N^+$ consisting of vertices and edges of $Q_1^+$, $Q_2^+$, $Q_3^+$, and $\tau$.  Before continuing, it is worthwhile examining the graph $M$ consisting of the edges and vertices of $Y^+$ and edges and vertices of $F$ (see \figref{M}.a).  The graph $M$ has an outer face bounded by $F$ an inner face $\tau$ and $f\le 3$ inner faces $F_1,\ldots,F_f$, that are labelled so that each $F_i$ does not contain any vertices of $P_i$ on its boundary.

  \begin{figure}
    \begin{center}
      \begin{tabular}{cc}
        \includegraphics{figs/tripod-faces} &
        \includegraphics{figs/tripod-faces-2} \\
        (a) & (b)
      \end{tabular}
    \end{center}
    \caption{The graph $M=F\cup Y^+$ and the sets $D^+$ and $Y^{++}$.}
    \figlabel{M}
  \end{figure}

  For plane graphs, the tripod $Y^+$ and the esulting graph $M=F\cup Y^+$ are the key ingredients for computing layered $H$-partitions.  The separator $Y^+$ allows us to recurse on the independent subgraphs contained in $F_1$, $F_2$, and $F_3$.  However, for $k$-plane graphs, these are insufficient.  Instead, we need a ``fatter'' version of $M$ that prevents an edge of $G$ from having an endpoint in each of $F_i$ and $F_j$ for $i\neq j$.  The set $D^+$ provides the fat version of $F$ that we need.  Next we describe the fat version of $Y^+$ that we need.

  Consider the vertex set 
  \[  
       X=\bigcup_{x^+\in V(Y^+)\cap V(G^-)}\{z\in V(N^+): \dist_{G^-\cap N^+}(x^+,z)\le k \enspace .
  \]
  By \lemref{small-balls}, every dummy vertex $x^+\in V(Y^+)$ contributes at most $4\cdot3^{k-1}$ elements to $X$.
  Therefore, by the triangle inequality
  \begin{equation}
      \dist{T}(r,x^+) - k \le \dist_{T}(r,z) \le \dist{T}(r,x^+) + k
      \eqlabel{depth-difference}
  \end{equation}

  Next, consider the set 
  \[  
    Y^{++}=(X\cup\{ v: \mbox{$vw\in E(G^+)$ and $w\in X$}\})\cap V(N^+) \setminus V(D^+).  
  \]
  Since every vertex in $X$ has maximum degree $4$, every vertex in $Y+$ contributes at most $16\cdot 3^{k-1}$ elements to $Y^{++}$.

  We claim that, for each integer $i\ge 0$, $|V_i\cap Y^{++}|\le \fk$.  To see why this is so, first note that $|V_i^+\cap V(Y^+)|\le 3$ since $V_i^+$ contains at most one vertex from each of $Q_1^+$, $Q_2^+$, and $Q_3^+$.  By \eqref{depth-difference}, for any vertex $z\in V_i\cap Y^{++}\setminus Y^+$, there is at least one vertex $x^+\in Y^+\cap (V_{i-k}\cup \cdots\cup V_{i+k})$ that contributed $z$.  Since $|Y^+\cap (V_{i-k}\cup \cdots\cup V_{i+k})|\le 3(k+1)$, it follows that $|V_i\cap X\setminus Y^+|\le 3(k+1)16\cdot 3^{k-1}= \fk$.

  For each $i\in\{1,\ldots,f\}$, let $N^+_i$ denote the subgraph of $N^+$ consisting of only the edges and vertices of $N^+_i$ contain in or in the interior of $F_i^+$.  Similarly, let $Y^{++}_i$, $D^+_i$ be the subsets of $Y^{++}$ and $D^+$ contained in the closure of $F^+i$.

  We now make the following crucial claim:
  For any $j\in\{1,\ldots,f\}$ there are two indices $a,b\in\{1,2,3\}$ such that any edge $vw\in E(G)$ with one endpoint $v\in V(N^+_i)\setminus(D^+_i\cup Y^++_i)$ has its other endpoint $w\in V(N^+_i)$.
  Indeed, if this were not the case, then $vw$ would cross the outer face of $N^+_i$ at some dummy vertex.  Let $x^+$ denote the first such crossing point.  If $x^+\in V(F)$, then $vw$ contains $k$ additional dummy vertices of $D^+_i$. Otherwise, $x^+\in V(Y^+)$, in which case $vw$ contains an additional $k$ dummy vertices in $Y^{++}$.  In either case, $vw$ contains at least $k+1$ dummy vertices, contradicting the fact that $vw$ is crossed by at most $k$ edges.

  The set $Y^{++}$ will become an element $B_x$ in our $H$-partition of $G$.
  All that remains is to apply induction on the three subgraphs $N^+_i$ for each $i\in\{1,\ldots,3\}$.  For each $i\in\{1,\ldots,3\}$,
  let $N_i$ denote the subgraph of $G$ containing the vertices and edges of $G$ that are contained in the closure of $F_i$.  Note that the number of vertices of $N_i^+$ is less than the number of vertices of $N^+$ since $N_i^+$ does not contain one of the vertices of $\tau$.  Therefore, we can apply induction to obtain the desired $H$-partition of $N_i$ using the cycle $F_i$ in place of $F$, the set $D' = D^+_i\cap Y^{++}_i$ and the partition of $D'$ into $D'\cap Y^{++}$, $D'\cap P_{i-1}$ and $D'\cap P_{i+1}$.

  The remainder of the proof is as before. We use $Y^{++}$ as an element of our $H$-partition. The root bag in the tree decomposition contains the vertices obtained by contracting each of $P_1,\ldots,P_k$ and $Y^{++}$.  The graph $H$ is planar because $Y^{++}$ is a connected set of vertices in $G^+$. 
\end{proof}

\begin{thm}\thmlabel{1-plane}
  Every $k$-plane graph $G$ has a layered $H$-partition of layered width at most $k\cdot\fk$ in which $H$ is planar and has treewidth at most 3.
\end{thm}

\begin{proof}
  The same as before except that the layered width $\fk$ from \lemref{k-induction} becomes $k\cdot\fk$ because \lemref{k-induction} uses a layering of $G^+$ and distances in $G^+$ can be a factor of $k$ larger than in $G$.
\end{proof}

% Let $G$ be a $k$-planar graph with a distinguished vertex $r$. For each integer $i\ge 0$, let $V_i=\{v\in V(G):\dist_G(r,v)=i\}$ and let $V_{\le i}=\bigcup_{j=0}^i V_j$.  We will use distances to $r$ frequently, so for each $v\in V(G)$, we let $d_v=\dist_G(r,v)$ (so that $v\in V_{d_v}$). We say that a subgraph $P$ of $G$ has layered width at most $\ell$ if $|V_i\cap V(P)|\le \ell$ for each integer $i\ge 0$.
% 
% Let $P_1,\ldots,P_3$ be three vertex-disjoint subgraphs of $G$, each of which has layered width at most $\ell$ and none of which contains $r$. Let $B$ be the union of $P_1,\ldots,P_3$.  Let $N$ be a component of $G-V(B)$ that does not contain $r$.  Let $G'$ be the subgraph of $G$ consisting of edges and vertices of $N$, edges and vertices of $B$, and all edges $vw\in E(G)$ with $v\in V(N)$ and $w\in V(B)$.
% 
% \begin{lem}
%   There exists a vertex $v\in V(G')$ and three vertex-disjoint paths $Q_1$, $Q_2$, and $Q_3$ where each $Q_i$ begins at $v$, has all of its internal vertices in $V(N)$, ends at a vertex in $V(P_i)$ and has layered width at most 10. 
% \end{lem}
% 
% \begin{proof}
%   For simplicity, assume that every vertex of $P_i$ is adjacent to at least one vertex of $V(N)$ since we can ignore vertices of $P_i$ for which this is not the case.  
% 
%   For each integer $r\ge 0$, consider the induced subgraph $G_r'=G'[V(B)\cup V_{\le r}]$.  Let $r$ be the smallest integer such that $G_{r}'$ has at least one component $A$ that contains vertices from each of $P_1$, $P_2$, and $P_3$.  Select any vertex $w\in V(A)\setminus V_{<r}$. Suppose, for now that $w\not\in V(B)$.  Then, since no component of $V_{r-1}$ contains vertice from each of $P_1$, $P_2$, and $P_3$, it must be the case that $G_{r-1}'$ has components $A_1$, $A_2$, and $A_3$ where $w$ is adjacent to some vertex in each $A_i$, each $A_i$ contains a vertex of $P_i$, $A_1\not\in\{A_2,A_3\}$ and $A_1$ does not contain a vertex of $P_3$.  (Note that we may have $A_2=A_3$ and we may have $A_1$ containig vertices of $P_2$.) 
% 
%   If $A_2\neq A_3$, then we are done.  We set $v=w$ and, for each $i\in \{1,\ldots,3\}$ we take a path $Q_i$ from $v$ to $V(P_i)$ whose $j$th vertex is in $V_{r-j+1}\cap V(A_i)$.  These three paths are disjoint since $V(Q_i)\subseteq V(A_i)$ for each $i\in \{1,\ldots,3\}$.
% 
%   If $A_2=A_3$, there is still a little more work to do.  In this case we will still use the path $Q_1$ from $w$ to $V(P_1)$.  Now, $w$ is adjacent to a vertex $w_1$ in $A_2=A_3$ so we extend
% 
% 
% 
% 
% 
%    where (without loss of generality) $V(A_1)\cap V(P_1)\neq \emptyset$, $V(A_1)\cap (V(P_2)\cup V(P_3))=\emptyset$, $V(A_2)\cap V(P_2)\neq\emptyset$, $V(A_2)\cap V(P_1)=\emptyset$ and $w$ is adjacent to a vertex $w_i\in A_i$ for each $i\in\{1,2\}$.
% 
%   If 
% 
% 
% 
% 
% 
%    contains vertices of $P_1$, $A_2$ contains vertices of $P_2$ and (possibly )
% 
% 
% 
% 
%   We claim that $v$ has the desired property.  Indeed
% 
% \end{proof}
% 
% 
% 
% 
% 
% \section{A Metric Alternative To Sperner's Lemma}
% 
% I'm interested in having something more robust and less magical than Sperner's Lemma. Here's a first attempt.
% 
% Let $G$ be a planar triangulation with some distinguished vertex $r$. 
% 
% 
% 
% 
% A \emph{$t$-pod} (with respect to $r$) consists of vertex-disjoint paths $Q_1,\ldots,Q_t$, where each $Q_i$ a prefix of some $v_i\leadsto r$ path in $G$ and $v_1,\ldots,v_t$ form a clique in $G$.  We call $v_1,\ldots,v_t$ the \emph{central vertices} of the tripod and we call $Q_1,\ldots,Q_t$ the \emph{legs} of the tripod.  The final vertex in each path $Q_i$ is called \emph{foot} if the tripod.  
% 
% A $t$-pod $\Y$ with $t\le 3$, is called a \emph{tripod}.  A $t$-pod $\V$ with $t\le 2$ is called a \emph{bipod}.  A $t$-pod $I$ with $t\le 1$ is called a \emph{monopod}.
% 
% For a $t$-pod $Y$ with central vertices $v_1,\ldots,v_t$, we define the \emph{depth} of $Y$ as 
% \[
%     \depth(Y)=\max\{d_G(r,v):v\in V(T)\} 
%         = \max\{\dist_G(r, v_i): i\in\{1,\ldots,t\}\} \enspace .
% \]
% Observe that, since $v_1,\ldots,v_t$ form a clique in $G$, $\depth(Y)-1\le \dist_G(r, v_i) \le \depth(Y)$.
% 
% Let $C$ be a cycle in $G$ and $P_1,\ldots,P_a$, $a\le 3$ be a partition of $V(C)$ such that $C[P_i]$ is a path and $C[P_i]$ is also a bipod $\V_i$ in $G$. Let $Q_{i,1}$ and (possibly) $Q_{i,2}$ be the legs of $\V_i$ where we use the convention that, for each $i\in\{1,\ldots,a\}$, the last vertex of $Q_{i,2}$ and the first vertex of $Q_{i+1,1}$ are adjacent in $C$.  Let $v_{i,j}$ be first vertex of $Q_{i,j}$ and let $v_{i,j}'$ be the last vertex of $Q_{i,j}$.  With these definitions, $v_{i,j}$ is a central vertex of $\V_i$ and $v_{i,j}'$ is a foot of $\V_{i}$.
% 
% Note that, for each $i\in\{1,\ldots,a\}$, let $d_i = dist_{G}(r,\{v_{i,2},v_{i+1,1}'\})$.  
% 
% 
% 
% Let $N$ be the near-triangulation consisting of all edges and vertices of $G$ that are on or in the ineterior of $C$.  
% 
% 
% \section{Remarks}
% 
% \begin{enumerate}
%   \item The fact that $G'$ is a triangulation is not really important. We could arbitrarily triangulate $N'$ (compatible with $T$) and still find a Sperner triangle.
%   \item The fact that $T$ is plane seems important since it lets us use Sperner's Lemma.  But maybe Sperner's Lemma is overkill. $Y^+$ is just a crappy separator of small layered width that happens to break $N$ into components for which no component is adjacent to more than 2 of $Q_1,\ldots,Q_3$.
%   \item The fact that $C_i$ is a cycle of non-crossed edges in $G$ is not necessarily critical.  The non-crossed property isn't so important as long as no edge of $G-V(S^+)$ has endpoints in two different faces of $S^+$.
% 
%   More troublesome, again, is the use of Sperner's Lemma since it requires that $P_1,\ldots,P_k$ (and therefore also the boundary of $Y^+$) be not so wild that they interleave.  Again, maybe we can cook up something more robust than Sperner's Lemma.
% 
%   \item I think I can push through a proof for 2-plane graphs.
% \end{enumerate}

\bibliographystyle{plain}
\bibliography{warmup}

\end{document}

### 2019-05-03 k-planar.tex
This works!

@@
 
-% The remainder of the paper is organized as follows:  \secref{1-planar} is a warm-up and gives the proof for 1-plane graphs.  \secref{k-planar} presents the proof for $k$-plane graphs.
-% 
-% 
-% \section{$1$-Plane Graphs}
-% \seclabel{1-planar}
-% 
-% Let $G$ be an edge maximal 1-plane graph.
-Here, edge-maximal should be taken to mean that any two vertices of $G$ on a common face have an edge joining them.
-Assume also that $G$ is crossing minimal so that the embedding of $G$ minimizes the number of crossings over all possible embeddings of the (abstract) graph $G$.
-From now on, we treat vertices of $G$ as points and edges as simple curves.
-% 
-% A \emph{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.
-It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
-% The edges $vw$ and $xy$ are called \emph{spars} of $K$.
-The cycle $vxwy$ is called the \emph{sail} of $K$.
-It follows from crossing minimality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$.
-Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
-Observe that any spar of $K$ is incident on exactly four \emph{kite faces} of $G$, each of which has three edges and two vertices of $G$ on its boundary.
-% 
-% The 1-plane graph $G$ has a plane triangulation $G'$ as a subgraph that can be obtained by removing one spar from each kite in $G$.
-Observe that, for any spar $vw\in E(G)\setminus E(G')$ that crosses $xy\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
-It follows that $\dist_{G'}(x,y)\le 2\dist_G(x,y)$ for any pair of vertices $x,y\in V(G)$.
-% 
-% \begin{lem}\lemlabel{induction} The setup:
-%   \begin{compactenum}
-%     \item Let $G$ and $G'$ be defined as above.
-%     \item Let $T$ be a BFS tree of $G'$ rooted at some vertex $r$.
-%     \item For each integer $i\ge 0$, let $V_i=\{v\in V(G):\dist_T(r,v)=i\}$. 
-%     \item Let $C$ be a cycle in $G'$ with $r$ in the exterior of $C$ and such that
-%     \begin{compactenum} 
-%       \item No edge of $C$ is crossed by any edge of $G$; and
-%       \item $V(C)$ can be partitioned into $P_1,\ldots,P_k$, $k\le 3$ such that for each $i\in\{1,\ldots,k\}$,
-%       \begin{compactenum}
-%         \item $C[P_i]$ is a path; and
-%         \item $|P_i\cap V_j| \le 15$ for all integers $j\ge 0$.
-%       \end{compactenum}
-%     \end{compactenum}
-%     \item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $C$ or the interior of $C$.
-%   \end{compactenum}
-%   Then $N$ has an $H$-partition $(B_x : x\in V(H))$ such that
-%   \begin{compactenum}
-%     \item $H$ is planar;
-%     \item for all integers $j\ge 0$, and all $x\in V(H)$, $|B_x\cap V_j|\le 15$; 
-%     \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=B_{x_i}$; and
-%     \item $H$ has a tree decomposition whose largest bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
-%   \end{compactenum}
-% \end{lem}
-% 
-% \begin{proof}
-%   The proof is by induction on the number of vertices of $N$.
-%   First note that $N'$ is a near-triangulation.
-If $k=3$, set $R_i=P_i$ for each $i\in\{1,2,3\}$.
-Otherwise, as before, split $P_1,\ldots,P_k$ to partition $V(C)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $C[R_i]$ is a path and each $R_i$ contains vertices from at most one of $P_1,\ldots,P_k$.
-% 
-%   For each $i\in\{1,2,3\}$ and $v\in R_i$, assign the colour $i$ to $v$.
-%   For each $v\in V(N')$, consider the path $P_v$ that consists of the subpath of the $v$ to $r$ path in $T$ that stops at the first vertex $v'\in V(C)$. Assign $v$ to have the same colour as $v'$.
-% 
-%   Sperner's Lemma ensures that $G'$ contains a triangular face $\tau=v_1v_2v_3$ whose three vertices are assigned different colours.
-For each $i\in\{1,2,3\}$, let $Q_i=P_{v_i}$ be the path in $T$ form $v_i$ to $v_i'$.
-Let $Y$ denote the subgraph of $N'$ consisting of vertices and edges $Q_1$, $Q_2$, $Q_3$, and $\tau$.
-Let $Y^+$ denote the subgraph of $N$ consisting of the vertices and edges of $Y$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $Y$.
-% 
-%   We claim that, for each integer $i\ge 0$, $|V(Y^+)\cap V_i|\le 15$.
-To see this, first observe that $Y$ contains at most 3 vertices of $V_i$.
-If a vertex $x\in V(Y^+)\setminus V(Y)$ is contained in $V_i$, then this is because $Y$ contains an edge $vw$ with $v\in V_{i'}$ and $w\in\{V_{i'+1}$ for some $i'\in\{i-1,i\}$ and $G$ contains an edge $xy$ that crosses $vw$.
-(Recall that $vx,xw\in E(G')$, so $\dist_{G'}(w,r)-1\le\dist_{G'}(x,r)\le\dist_{G'}(v,r)+1$.)  The graph $Y$ contains at most 6 such edges, each of which accounts for at most 2 additional vertices of $V_i$.
-Therefore, in total, $|Y^+\cap V_i|\le 15$.
-% 
-% 
-%   Finally, let $S$ and $S^+$ denote the subgraph of $G$ containing the edges and vertices of $Y$, respectively $Y^+$, and the edges and vertices of $C$.
-The graph $S^+$ has some number of bounded faces, all contained in the interior of $C$.
-Some of the bounded faces of $S$ are kite faces.
-Call the non-kite bounded faces $F_1,\ldots,F_m$ and let $C_1,\ldots,C_m$ denote their boundaries.
-We claim that, for each $i\in\{1,\ldots,m\}$, if some portion of $C_i$ is contained in an edge $vw\in E(G)$ then $vw$ is not crossed by any edge of $G$.
-To see this, there are three cases to consider:
-%   \begin{enumerate}
-%     \item $vw\in E(C)$. By assumption, $vw$ is not crossed by any edge of $G$.
-%     \item $vw\in E(Y)$. In this case, the kite containing $vw$ is in $Y^+$, so $vw$ is only incident to kite faces.
-%     \item $vw\in E(Y^+)\setminus E(Y)$.
-In this case, either $vw$ is a sail edge in which case it is not crossed by definition, or $vw$ is a spar that was added to $Y^+$ because $vw$ crosses some edge $xy\in E(Y)$.
-In this latter case, $vw$ is only incident to kite faces.
-%   \end{enumerate}
-%   Therefore each $C_i$ is a cycle in $G^+$ consisting entirely of uncrossed edges.
-The vertices of $C_i$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$ and $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$.
-Furthermore $C_i[P_j']$ is a path for each $j\in\{1,\ldots,3\}$.
-Finally, the subgraph $N_i$ of $G$ consisting of the edges and vertices of $G$ contained in $C_i$ or its interior does not contain one of the three vertices of $\tau$.
-Therefore, we can apply induction using the cycle $C_i$ and the partition $P_1',P_2',P_3'$ of $V(C_i)$ to obtain the desired $H$-partition and tree decomposition of $N_i$.
-% 
-%   The remainder of the proof is as before.
-We use $V(Y^+)\setminus V(C)$ as an element of our $H$-partition.
-The root bag in the tree decomposition contains the vertices obtained by contracting each of $P_1,\ldots,P_k$ and $Y^+$.
-% 
-%   Planarity comes from the fact that if two edges $vw$ and $xy$ cross, then they end up in the same bag of the $H$-partition.
-This means that $H$ is actually obtained by contracting connected sets of vertices in the planar graph $G'$.
-% \end{proof}
-% 
-% \begin{thm}\thmlabel{1-plane}
-%   Every 1-plane graph $G$ has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
-% \end{thm}
-% 
-% \begin{proof}
-%   The same as before except that the layered width 15 from \lemref{induction} becomes 30 because \lemref{induction} uses a layering of $G'$ and distances in $G'$ can be a factor of 2 larger than in $G$.
-% \end{proof}
+\section{Another Way of Looking at It}
 
-\section{$k$-Planar}
-\seclabel{k-planar}
+\subsection{PS-Tree}
 
-Let $G$ be a $k$-planar graph.  We will use of the \emph{plane augmentation} $G^+$ of $G$ obtained by adding a \emph{dummy vertex} at each crossing in $G$
-and then adding edges to the resulting graph to obtain a triangulation.
-% 
-% triangulating the resulting graph arbitrarily.
-% 
-% 
-%  then adding a \emph{star vertex} $v_f$ in each face $f$ of the resulting graph and making $v_f$ adjacent to each vertex of $f$.
+A PS-Partition $(Y_x:x\in V(K))$ for a triangulation $\Delta$ and spanning-tree $T$ of $\Delta$ is a partition of $V(\Delta)$ whose elements are indexed by nodes of a rooted 3-ary tree $K$.
+Each bag $Y_x$ of a PS-partition is a tripod in $\Delta$, with respect to $T$.
+This means that $T[Y_x]$ has at most 3 components, each of which is a vertical path whose lower endpoints form a clique $\tau_x$ in $\Delta$.
+Associated with each node $x\in V(K)$ is a cycle $F_x$ in $\Delta$ and a near-triangulation $N_x$ consisting of the part of $\Delta$ contained in $F_x$.
+Note that $Y_x\subseteq V(N_x)\setminus V(F_x)$ and $V(N_x)\subseteq V(N_a)$ for every ancestor $a$ of $x$.
 
-We will make use of the fact that $\dist_{G^+}(x,y)\le k\cdot\dist_G(x,y)$ for any $x,y\in V(G)$.
-This follows immediately from the fact that any edge $xy\in E(G)$ has a corresponding path $x\to y$ in $G^+$ of length at most $k$.
+It is also useful to have a name for the \emph{Mercedes graph} $M_x$ that consists of the edges and vertices of $F_x$, the clique $\tau_x$, and the paths in $T$ from the vertices of $\tau_x$ to $F_x$.
 
-The \emph{dummy graph} $G^-$ is the subgraph of $G^+$ induced by the dummy vertices.  Note that the dummy graph has maximum degree 4. This immediately implies the following:
+In short, $K$ is the 3-ary tree of tripods we get from our proof.
+It's the tree we use for our tree-decomposition of $H$.
+We've just added some decorations to it.
+Eventually I should write this as a lemma with bullet points for the relevant properties.
 
-\begin{lem}\lemlabel{small-balls}
-  Let $G$ be a $k$-plane graph, let $G^+$ be the plane augmentation of $G$, and let $G^-=G^+-V(G)$.  Then, for any vertex $v\in G^-$, $|\{w\in V(G^-): \dist_{G^-}(v,w)\le k\}|\le 4\cdot 3^{k-1}$.
-\end{lem}
+\subsection{$k$-Planar Graphs}
 
-Before reading the statement of the following lemma it is worthwhile looking at \figref{M} to understand the role of some of the elements (in particular, $Y^{++}$ and $D^+$).
+Let $G$ be a $k$-plane graph and let $G^+$ be the graph obtained by making $G$ planar with dummy vertices and then triangulating the resulting graph.
+Let $T$ be a BFS spanning tree of $G^+$ where the root $r$ of $T$ has degree 3 in $G^+$ (add an additional vertex to $G^+$ if necessary).
+For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{G^+}(r,v)=i\}$.
 
-\begin{lem}\lemlabel{k-induction} The setup:
-  \begin{compactenum}
-    \item Let $G$ be a $k$-plane graph, let $G^+$ be the plane augmentation of $G$, and let $G^-$ be the dummy graph of $G$.
-    \item Let $T$ be a BFS spanning tree of $G^+$ rooted at some vertex $r$.
-    \item For each integer $i\ge 0$, let $V^+_i=\{v\in V(G^+):\dist_{G^+}(r,v)=i\}$, and let $V_i=V(G)\cap V^+_i$.
-    \item Let $F$ be a cycle in $G^+$ that has $r$ in its exterior.
-    \item Let $N^+$ denote the near-triangulation consisting of the edges and vertices of $G^+$ contained in $F$ or the interior of $F$
-    \item Let $D^+\subset V(N^+)$ have the following properties:
-    \begin{compactenum}
-      \item $V(F)\subset D^+$.
-      \item For any dummy vertex $x\in V(F)$ and any vertex $y\in V(N^+)\setminus D^+$,  $\dist_{G^-}(x,y) \ge k$.
-      \item $D^+$ can be partitioned into $P_1,\ldots,P_k$, $k\le 3$ such that for each $i\in\{1,\ldots,k\}$,
-      \begin{compactenum}
-        \item $F[P_i]$ is connected; and
-        \item $|P_i\cap V_j| \le \fk$ for all integers $j\ge 0$.
-      \end{compactenum}
-    \end{compactenum}
-    \item Let $N$ be the subgraph of $G$ induced by $V(F)\cap V(G)$ plus the vertices of $G$ in the interior of $F$.
-  \end{compactenum}
-  Then $N$ has an $H$-partition $(B_x : x\in V(H))$ such that
-  \begin{compactenum}
-    \item for every integer $j\ge 0$, and all $x\in V(H)$, $|B_x\cap V_j|\le \fk$; 
-    \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=B_{x_i}$; and
-    \item $H$ has a tree decomposition whose largest bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
-  \end{compactenum}
-\end{lem}
+Let $K$ be the PS-partition for $G^+$ and $T$.
+We say that a vertex $v\in V(G)$ \emph{appears} at a node $x\in V(K)$ if $u\in Y_x$ or if there exists $vw\in E(G)$ such that $v,w\in V(N_x)\setminus V(F_x)$ and $vw$ crosses $Y_x$ (i.e,.
+$vw$ contains a dummy vertex of $V(Y_x)$ in its interior).
+
+Let $S_x$ (the separator at $x$) denote the set of all vertices $v\in V(G)$ such that $v$ appears at $x$ but does not appear at an ancestor of $x$.
+We claim that $(S_x:x\in V(K))$ is a partition of $V(G)$.
+Indeed, any $v\in V(G)$ can only appear in a node $x$ such that $v\in V(N_x)\setminus V(F_x)$.
+These nodes form a path from the root of $K$ to the unique node $x'$ such that $v\in Y_{x'}$.
+It is worth noting that, if $vw\in E(G)$, and $v,w\in S_x$, then the embedding of $vw$ is contained in the interior of $F_x$ (otherwise $v$ and $w$ would have appeared in $S_a$ for some ancestor $a$ of $x$.)
+
+We use the partition $(S_x: x\in V(K))$ which generates a graph $H$ with vertex set $V(H)=V(K)$ the edge $xy\in E(H)$ if there exists an edge $vw\in E(H)$ with $v\in S_x$ and $w\in S_y$.
+
+First we argue that this partition has small width, with respect to the layering defined by $T$ (note that this is a layering of $G^+$, not a layering of $G$).
+
+\begin{clm}
+  For any $x\in V(H)$, and any integer $i\ge 0$, $|S_x\cap V_i|\le 3(k+1)$.
+\end{clm}
 
 \begin{proof}
-  Let $N^+$ be the subgraph of $G^+$ consisting only of those edges and vertices of $G^+$ contained in $F$ or in the interior of $F$.
-  We remark that $N^+$ is a near-triangulation.
-
-  The proof is by induction on $|V(N^+)|$.
-If $|V(N^+)| = 3$, then the graph $H$ is the complete graph on integers $1,\ldots,k$ and the partition sets $B_{i}=P_i\cap V(G)$ for each $i\in\{1,\ldots,k\}$.
-
-  Since $P_1,\ldots,P_k$ is a partition of $D^+$ and $V(N)\subseteq V(D^+)$, every vertex $v\in V(N)$ is contained in some set $P_i$ and is therefore contained in some bag $B_{i}$ of the $H$-partition.
-Every edge $vw\in E(N)$ has $v\in B_{i}$ and $w\in B_{j}$ for some $i,j\in\{1,\ldots,k\}$.
-Either $i=j$ or the edge $ij\in E(H)$.
-Therefore $(B_x : x\in V(H))$ is indeed an $H$-partition of $N$.
-
-  Requirement~1 of the $H$-partition follows from Condition~6.c.ii.
-To satisfy Requirements~2 and 3 we take the tree decomposition of $H$ to consist of a single node $u$ whose bag contains $\{1,\ldots,k\}$.
-This completes the description of how the base case $|V(N^+)|=3$ is handled.
-From this point on, we assume $|V(N^+)|> 3$.
-
-  We want to 3-colour $V(N^+)$ so that we can apply Sperner's Lemma, which means we need to partition $V(F)$ into sets $R_1,\ldots,R_3$ such that $F[R_i]$ is a path for each $i\in\{1,\ldots,3\}$.
-  If $k=3$, we set $R_i=P_i\cap V(F)$ for each $i\in\{1,2,3\}$.
-Condition~6.c.i ensures that $F[R_i]$ is a path, for each $i\in\{1,\ldots,3\}$.
-If $k\in\{1,2\}$, then we partition $(P_1\cap V(F))\cup\cdots\cup(P_k\cap V(F))$ into three sets $R_1$, $R_2$, and $R_3$ such that, for each $i\in\{1,\ldots,3\}$, $F[R_i]$ is a path $R_i\subseteq P_j$ for some $j\in\{1,\ldots,k\}$.
-Again, Condition~6.c.i ensures that this is always possible.
-
-  We now describe the 3-colouring of $V(N^+)$ that allows us to apply Sperner's Lemma. For each $i\in\{1,2,3\}$ and $v\in R_i$, assign the colour $i$ to $v$.
-  For each $v\in V(N^+)$, consider the path $P_v$ that consists of the subpath of the $v$ to $r$ path in $T$ that stops at the first vertex $v'\in V(F)$. Assign $v$ to have the same colour as $v'$.
-
-  Sperner's Lemma ensures that $N^+$ contains a triangular face $\tau=v_1v_2v_3$ whose three vertices are assigned different colours.
-For each $i\in\{1,2,3\}$, let $Q_i^+=P_{v_i}^+$ be the path in $T$ from $v_i$ to $v_i'$.
-Let $Y^+$ denote the subgraph of $N^+$ consisting of vertices and edges of $Q_1^+$, $Q_2^+$, $Q_3^+$, and $\tau$.
-Before continuing, it is worthwhile examining the graph $M$ consisting of the edges and vertices of $Y^+$ and edges and vertices of $F$ (see \figref{M}.a).
-The graph $M$ has an outer face bounded by $F$ an inner face $\tau$ and $f\le 3$ inner faces $F_1,\ldots,F_f$, that are labelled so that each $F_i$ does not contain any vertices of $P_i$ on its boundary.
-
-  \begin{figure}
-    \begin{center}
-      \begin{tabular}{cc}
-        \includegraphics{figs/tripod-faces} &
-        \includegraphics{figs/tripod-faces-2} \\
-        (a) & (b)
-      \end{tabular}
-    \end{center}
-    \caption{The graph $M=F\cup Y^+$ and the sets $D^+$ and $Y^{++}$.}
-    \figlabel{M}
-  \end{figure}
-
-  For plane graphs, the tripod $Y^+$ and the esulting graph $M=F\cup Y^+$ are the key ingredients for computing layered $H$-partitions.
-The separator $Y^+$ allows us to recurse on the independent subgraphs contained in $F_1$, $F_2$, and $F_3$.
-However, for $k$-plane graphs, these are insufficient.
-Instead, we need a ``fatter'' version of $M$ that prevents an edge of $G$ from having an endpoint in each of $F_i$ and $F_j$ for $i\neq j$.
-The set $D^+$ provides the fat version of $F$ that we need.
-Next we describe the fat version of $Y^+$ that we need.
-
-  Consider the vertex set 
-  \[  
-       X=\bigcup_{x^+\in V(Y^+)\cap V(G^-)}\{z\in V(N^+): \dist_{G^-\cap N^+}(x^+,z)\le k \enspace .
-  \]
-  By \lemref{small-balls}, every dummy vertex $x^+\in V(Y^+)$ contributes at most $4\cdot3^{k-1}$ elements to $X$.
-  Therefore, by the triangle inequality
-  \begin{equation}
-      \dist{T}(r,x^+) - k \le \dist_{T}(r,z) \le \dist{T}(r,x^+) + k
-      \eqlabel{depth-difference}
-  \end{equation}
-
-  Next, consider the set 
-  \[  
-    Y^{++}=(X\cup\{ v: \mbox{$vw\in E(G^+)$ and $w\in X$}\})\cap V(N^+) \setminus V(D^+).  
-  \]
-  Since every vertex in $X$ has maximum degree $4$, every vertex in $Y+$ contributes at most $16\cdot 3^{k-1}$ elements to $Y^{++}$.
-
-  We claim that, for each integer $i\ge 0$, $|V_i\cap Y^{++}|\le \fk$.
-To see why this is so, first note that $|V_i^+\cap V(Y^+)|\le 3$ since $V_i^+$ contains at most one vertex from each of $Q_1^+$, $Q_2^+$, and $Q_3^+$.
-By \eqref{depth-difference}, for any vertex $z\in V_i\cap Y^{++}\setminus Y^+$, there is at least one vertex $x^+\in Y^+\cap (V_{i-k}\cup \cdots\cup V_{i+k})$ that contributed $z$.
-Since $|Y^+\cap (V_{i-k}\cup \cdots\cup V_{i+k})|\le 3(k+1)$, it follows that $|V_i\cap X\setminus Y^+|\le 3(k+1)16\cdot 3^{k-1}= \fk$.
-
-  For each $i\in\{1,\ldots,f\}$, let $N^+_i$ denote the subgraph of $N^+$ consisting of only the edges and vertices of $N^+_i$ contain in or in the interior of $F_i^+$.
-Similarly, let $Y^{++}_i$, $D^+_i$ be the subsets of $Y^{++}$ and $D^+$ contained in the closure of $F^+i$.
-
-  We now make the following crucial claim:
-  For any $j\in\{1,\ldots,f\}$ there are two indices $a,b\in\{1,2,3\}$ such that any edge $vw\in E(G)$ with one endpoint $v\in V(N^+_i)\setminus(D^+_i\cup Y^++_i)$ has its other endpoint $w\in V(N^+_i)$.
-  Indeed, if this were not the case, then $vw$ would cross the outer face of $N^+_i$ at some dummy vertex.
-Let $x^+$ denote the first such crossing point.
-If $x^+\in V(F)$, then $vw$ contains $k$ additional dummy vertices of $D^+_i$.
-Otherwise, $x^+\in V(Y^+)$, in which case $vw$ contains an additional $k$ dummy vertices in $Y^{++}$.
-In either case, $vw$ contains at least $k+1$ dummy vertices, contradicting the fact that $vw$ is crossed by at most $k$ edges.
-
-  The set $Y^{++}$ will become an element $B_x$ in our $H$-partition of $G$.
-  All that remains is to apply induction on the three subgraphs $N^+_i$ for each $i\in\{1,\ldots,3\}$.  For each $i\in\{1,\ldots,3\}$,
-  let $N_i$ denote the subgraph of $G$ containing the vertices and edges of $G$ that are contained in the closure of $F_i$.
-Note that the number of vertices of $N_i^+$ is less than the number of vertices of $N^+$ since $N_i^+$ does not contain one of the vertices of $\tau$.
-Therefore, we can apply induction to obtain the desired $H$-partition of $N_i$ using the cycle $F_i$ in place of $F$, the set $D' = D^+_i\cap Y^{++}_i$ and the partition of $D'$ into $D'\cap Y^{++}$, $D'\cap P_{i-1}$ and $D'\cap P_{i+1}$.
-
-  The remainder of the proof is as before.
-We use $Y^{++}$ as an element of our $H$-partition.
-The root bag in the tree decomposition contains the vertices obtained by contracting each of $P_1,\ldots,P_k$ and $Y^{++}$.
-The graph $H$ is planar because $Y^{++}$ is a connected set of vertices in $G^+$.
+  Done already, several times.
 \end{proof}
 
-\begin{thm}\thmlabel{1-plane}
-  Every $k$-plane graph $G$ has a layered $H$-partition of layered width at most $k\cdot\fk$ in which $H$ is planar and has treewidth at most 3.
+Next we describe a tree decomposition $(B_x:x\in V(K))$ of $H$.
+(Don't be put off here by the fact that $V(H)=V(K)$).
+We do this by definining, for each $x\in V(H)$, the subtree $K[x]=K[\{y:x\in B_y\}]$.
+The subtree $K[x]$ contains $x$.
+Additionally, for every edge $xz\in E(H)$, $K[x]$ includes the path from $x$ to $z$ in $K$.
+This clearly produces a tree decomposition of $H$, what remains is to upper bound the width of this decomposition.
+
+We claim that $K[x]$ is rooted at $x$, i.e., $x$ is the lowest common ancestor of all nodes in $B_x$.
+To see why this is so, suppose $K[x]$ has root $a\neq x$.
+This implies that $a$ has two children $b\neq c$ where $x$ is a descendant of $b$, some node $y$ is a descendant of $c$, and there is an edge $vw\in E(G)$ such that $v\in S_x\subset V(N_x)\subseteq V(N_b)\setminus V(F_a)$
+and $w\in S_y\subset V(N_y)\subseteq V(N_b)\setminus V(F_a)$.
+Therefore, $v$ and $w$ are in the interior of two different faces of the Mercedes graph $M_a$.
+Therefore the curve $vw$ intersects $M_a$.
+If it intersects $M_a$ at some vertex in $Y_a$, then $v$ would be in $S_a$, not $S_x$, a contradiction.
+If it intersects $M_a$ at some vertex in $F_a$, then $v$ would be in $S_{a'}$ for some ancestor $a'$ of $a$, also a contradiction.
+
+Now, fix some node $x\in V(K)$ and consider the contents of $B_x$.
+As we have just argued, $B_x$ contains only ancestors of $x$ (including $x$ itself).
+If $B_x$ contains some ancestor $a$ of $x$, it is because there is an edge $vw\in E(G)$ with $v\in V(N_x)\setminus V(F(x))$ and $w\in S_a$.
+Now, $w\in S_a$ because $w\in Y_a$ or there is an edge $wz\in E(G)$ that contains a point of $Y_a$.
+This implies that there is a curve $C$ from $v$ to $z$ that contains at most $2k$ dummy  vertices of $G^+$.
+
+Now, consider the cubic graph $A$ homeomorphic to the union of $F_{a'}$ for all ancestors $a'$ of $x$.
+It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0$ is the root of $K$.
+We then cut $F_{a_0}$ into two pieces $F^{\bar{x}}_{a_0}$ and $F^x_{a_0}$ where $F^x_{a_0}$ contains $F_x$.
+This process continues by partitioning $F_{a_1}:=F^x_{a_0}$ recursively.
+The process terminates after some number $r$ of iterations, when $F^x_{a_r}=F_x$.
+
+We claim that the curve $C$ begins in $F_x\subseteq F^x_a\subseteq F_a$, remains in the interior of $F_a$, and eventually reaches the boundary $F^{\bar{x}}_a\cap F^x_a$.
+The fact that $C$ remains in $F_a$ is due to the fact that, if it crossed the boundary of $F_a$, then one of $vw$ or $wz$ would interects $Y_{a'}$ for some ancestor $a'$ of $a$.
+But this would contradict the fact that $vw$ (or $wz$) are in $S_a$.
+The fact that $C$ eventually reaches $F^{\bar{x}}_a\cap F^x_a$ follows from the fact that $w\in S_a$ (so either $w\in Y_a$ or $wz$ crosses $Y_a$).
+
+Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of faces $F^{\bar{x}}_a$ of $A$ that can be reached with curves that begin in $F_x$, remain in $F_a$, and terminate in $F^{\bar{x}}_a$ without crossing more than $2k$ edges in $A$.
+To avoid triple subscripts we say that a curve visits $a_i$ if it visits $F^{\bar{x}}_{a_i}$.
+We can further restrict ourselves to studying curves that cross the fewest number of edges of $A$ since if a curve exists that crosses at most $k$ edges then the curve that crosses the fewest number of possible edges also crosses at most $k$ edges.
+We will call such a curve a lightest curve from $F_x$ to $F_a$.
+
+Again, we let $a_0,\ldots,a_r$ denote the vertices of the path, in $K$ from the root $a_0$ to $x=a_r$.
+We claim that, there exists a lightest curve from $a_r$ to $a_0$ that is monotonic in the sense that the curve visits $a_{i_0},\ldots,a_{i_t}$ in this order, where $i_0=r > i_1>\cdots>i_t=0$ the Indeed, this follows from the fact that, for any node $a_i$, $i>0$, the (three) faces of $A$ that share a boundary with $F_{a_i}$ also share a boundary with $F^{\bar{x}}_{a_i}$.
+Therefore it is never advantageous for a path to go from $a_i$ to $a_j$ for some $j>i$ since the path will eventually have to return to one of these three faces.
+
+At this point we are essentially done.
+Any monotonic curve that visits $a_i$ has at most 3 choices for the next face $a_j$ $j<i$ to visit.
+Therefore, the number of faces of $A$ that can be reached from $F_x$ by admissible curves that cross $2k$ edges of $A$ is at most $3^{2k}$.
+Therefore,
+\[
+   |B_x| \le \sum_{j=0}^2k 9^j = (9^{2k+1}-1)/8 \enspace .
+\]
+
+This completes the proof
+
+\begin{thm}
+    Every $k$-planar graph $G$ has a $(3k,k9^{2k+1}/8)$-partition.
 \end{thm}
-
-\begin{proof}
-  The same as before except that the layered width $\fk$ from \lemref{k-induction} becomes $k\cdot\fk$ because \lemref{k-induction} uses a layering of $G^+$ and distances in $G^+$ can be a factor of $k$ larger than in $G$.
-\end{proof}
-
-% Let $G$ be a $k$-planar graph with a distinguished vertex $r$.
-For each integer $i\ge 0$, let $V_i=\{v\in V(G):\dist_G(r,v)=i\}$ and let $V_{\le i}=\bigcup_{j=0}^i V_j$.
-We will use distances to $r$ frequently, so for each $v\in V(G)$, we let $d_v=\dist_G(r,v)$ (so that $v\in V_{d_v}$).
-We say that a subgraph $P$ of $G$ has layered width at most $\ell$ if $|V_i\cap V(P)|\le \ell$ for each integer $i\ge 0$.
-% 
-% Let $P_1,\ldots,P_3$ be three vertex-disjoint subgraphs of $G$, each of which has layered width at most $\ell$ and none of which contains $r$.
-Let $B$ be the union of $P_1,\ldots,P_3$.
-Let $N$ be a component of $G-V(B)$ that does not contain $r$.
-Let $G'$ be the subgraph of $G$ consisting of edges and vertices of $N$, edges and vertices of $B$, and all edges $vw\in E(G)$ with $v\in V(N)$ and $w\in V(B)$.
-% 
-% \begin{lem}
-%   There exists a vertex $v\in V(G')$ and three vertex-disjoint paths $Q_1$, $Q_2$, and $Q_3$ where each $Q_i$ begins at $v$, has all of its internal vertices in $V(N)$, ends at a vertex in $V(P_i)$ and has layered width at most 10.
-% \end{lem}
-% 
-% \begin{proof}
-%   For simplicity, assume that every vertex of $P_i$ is adjacent to at least one vertex of $V(N)$ since we can ignore vertices of $P_i$ for which this is not the case.  
-% 
-%   For each integer $r\ge 0$, consider the induced subgraph $G_r'=G'[V(B)\cup V_{\le r}]$.
-Let $r$ be the smallest integer such that $G_{r}'$ has at least one component $A$ that contains vertices from each of $P_1$, $P_2$, and $P_3$.
-Select any vertex $w\in V(A)\setminus V_{<r}$.
-Suppose, for now that $w\not\in V(B)$.
-Then, since no component of $V_{r-1}$ contains vertice from each of $P_1$, $P_2$, and $P_3$, it must be the case that $G_{r-1}'$ has components $A_1$, $A_2$, and $A_3$ where $w$ is adjacent to some vertex in each $A_i$, each $A_i$ contains a vertex of $P_i$, $A_1\not\in\{A_2,A_3\}$ and $A_1$ does not contain a vertex of $P_3$.
-(Note that we may have $A_2=A_3$ and we may have $A_1$ containig vertices of $P_2$.)
-% 
-%   If $A_2\neq A_3$, then we are done.
-We set $v=w$ and, for each $i\in \{1,\ldots,3\}$ we take a path $Q_i$ from $v$ to $V(P_i)$ whose $j$th vertex is in $V_{r-j+1}\cap V(A_i)$.
-These three paths are disjoint since $V(Q_i)\subseteq V(A_i)$ for each $i\in \{1,\ldots,3\}$.
-% 
-%   If $A_2=A_3$, there is still a little more work to do.  In this case we will still use the path $Q_1$ from $w$ to $V(P_1)$.  Now, $w$ is adjacent to a vertex $w_1$ in $A_2=A_3$ so we extend
-% 
-% 
-% 
-% 
-% 
-%    where (without loss of generality) $V(A_1)\cap V(P_1)\neq \emptyset$, $V(A_1)\cap (V(P_2)\cup V(P_3))=\emptyset$, $V(A_2)\cap V(P_2)\neq\emptyset$, $V(A_2)\cap V(P_1)=\emptyset$ and $w$ is adjacent to a vertex $w_i\in A_i$ for each $i\in\{1,2\}$.
-% 
-%   If 
-% 
-% 
-% 
-% 
-% 
-%    contains vertices of $P_1$, $A_2$ contains vertices of $P_2$ and (possibly )
-% 
-% 
-% 
-% 
-%   We claim that $v$ has the desired property.  Indeed
-% 
-% \end{proof}
-% 
-% 
-% 
-% 
-% 
-% \section{A Metric Alternative To Sperner's Lemma}
-% 
-% I'm interested in having something more robust and less magical than Sperner's Lemma. Here's a first attempt.
-% 
-% Let $G$ be a planar triangulation with some distinguished vertex $r$. 
-% 
-% 
-% 
-% 
-% A \emph{$t$-pod} (with respect to $r$) consists of vertex-disjoint paths $Q_1,\ldots,Q_t$, where each $Q_i$ a prefix of some $v_i\leadsto r$ path in $G$ and $v_1,\ldots,v_t$ form a clique in $G$.
-We call $v_1,\ldots,v_t$ the \emph{central vertices} of the tripod and we call $Q_1,\ldots,Q_t$ the \emph{legs} of the tripod.
-The final vertex in each path $Q_i$ is called \emph{foot} if the tripod.
-% 
-% A $t$-pod $\Y$ with $t\le 3$, is called a \emph{tripod}.  A $t$-pod $\V$ with $t\le 2$ is called a \emph{bipod}.  A $t$-pod $I$ with $t\le 1$ is called a \emph{monopod}.
-% 
-% For a $t$-pod $Y$ with central vertices $v_1,\ldots,v_t$, we define the \emph{depth} of $Y$ as 
-% \[
-%     \depth(Y)=\max\{d_G(r,v):v\in V(T)\} 
-%         = \max\{\dist_G(r, v_i): i\in\{1,\ldots,t\}\} \enspace .
-% \]
-% Observe that, since $v_1,\ldots,v_t$ form a clique in $G$, $\depth(Y)-1\le \dist_G(r, v_i) \le \depth(Y)$.
-% 
-% Let $C$ be a cycle in $G$ and $P_1,\ldots,P_a$, $a\le 3$ be a partition of $V(C)$ such that $C[P_i]$ is a path and $C[P_i]$ is also a bipod $\V_i$ in $G$.
-Let $Q_{i,1}$ and (possibly) $Q_{i,2}$ be the legs of $\V_i$ where we use the convention that, for each $i\in\{1,\ldots,a\}$, the last vertex of $Q_{i,2}$ and the first vertex of $Q_{i+1,1}$ are adjacent in $C$.
-Let $v_{i,j}$ be first vertex of $Q_{i,j}$ and let $v_{i,j}'$ be the last vertex of $Q_{i,j}$.
-With these definitions, $v_{i,j}$ is a central vertex of $\V_i$ and $v_{i,j}'$ is a foot of $\V_{i}$.
-% 
-% Note that, for each $i\in\{1,\ldots,a\}$, let $d_i = dist_{G}(r,\{v_{i,2},v_{i+1,1}'\})$.  
-% 
-% 
-% 
-% Let $N$ be the near-triangulation consisting of all edges and vertices of $G$ that are on or in the ineterior of $C$.  
-% 
-% 
-% \section{Remarks}
-% 
-% \begin{enumerate}
-%   \item The fact that $G'$ is a triangulation is not really important. We could arbitrarily triangulate $N'$ (compatible with $T$) and still find a Sperner triangle.
-%   \item The fact that $T$ is plane seems important since it lets us use Sperner's Lemma.
-But maybe Sperner's Lemma is overkill.
-$Y^+$ is just a crappy separator of small layered width that happens to break $N$ into components for which no component is adjacent to more than 2 of $Q_1,\ldots,Q_3$.
-%   \item The fact that $C_i$ is a cycle of non-crossed edges in $G$ is not necessarily critical.
-The non-crossed property isn't so important as long as no edge of $G-V(S^+)$ has endpoints in two different faces of $S^+$.
-% 
-%   More troublesome, again, is the use of Sperner's Lemma since it requires that $P_1,\ldots,P_k$ (and therefore also the boundary of $Y^+$) be not so wild that they interleave.
-Again, maybe we can cook up something more robust than Sperner's Lemma.
-% 
-%   \item I think I can push through a proof for 2-plane graphs.
-% \end{enumerate}
 

### 2019-05-03 k-planar.tex
Seems to work

@@
 To see why this is so, suppose $K[x]$ has root $a\neq x$.
-This implies that $a$ has two children $b\neq c$ where $x$ is a descendant of $b$, some node $y$ is a descendant of $c$, and there is an edge $vw\in E(G)$ such that $v\in S_x\subset V(N_x)\subseteq V(N_b)\setminus V(F_a)$
+In short: This happens because some $v\in V(N_x)$ is incident on an edge $vw\in E(G)$ that crosses $Y_a$.
+Longer: This implies that $a$ has two children $b\neq c$ where $x$ is a descendant of $b$, some node $y$ is a descendant of $c$, and there is an edge $vw\in E(G)$ such that $v\in V(N_x)\setminus V(F_x)\subseteq V(N_b)\setminus V(F_a)$
 and $w\in S_y\subset V(N_y)\subseteq V(N_b)\setminus V(F_a)$.
-Therefore, $v$ and $w$ are in the interior of two different faces of the Mercedes graph $M_a$.
-Therefore the curve $vw$ intersects $M_a$.
+(Note, we allow $x=b$ and/or $y=c$ here.)  Therefore, $v$ and $w$ are in the interior of two different faces of the Mercedes graph $M_a$.
+Therefore the edge $vw$ intersects $M_a$.
 If it intersects $M_a$ at some vertex in $Y_a$, then $v$ would be in $S_a$, not $S_x$, a contradiction.
@@
 Now, consider the cubic graph $A$ homeomorphic to the union of $F_{a'}$ for all ancestors $a'$ of $x$.
+See \figref{A}.
 It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0$ is the root of $K$.
@@
 The process terminates after some number $r$ of iterations, when $F^x_{a_r}=F_x$.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics{figs/A}
+  \end{center}
+  \caption{The graph $A$.}
+  \figlabel{A}
+\end{figure}
 

### 2019-05-04 k-planar.tex
(t,\ell) instead of (\ell,t)
Fixed boundary issue with where an edge appears

@@
 Let $K$ be the PS-partition for $G^+$ and $T$.
-We say that a vertex $v\in V(G)$ \emph{appears} at a node $x\in V(K)$ if $u\in Y_x$ or if there exists $vw\in E(G)$ such that $v,w\in V(N_x)\setminus V(F_x)$ and $vw$ crosses $Y_x$ (i.e,.
-$vw$ contains a dummy vertex of $V(Y_x)$ in its interior).
+We say that a vertex $v\in V(G)$ \emph{appears} at a node $x\in V(K)$ if $u\in Y_x$ or if there exists $vw\in E(G)$ such that $v,w\in V(N_x)$ and $vw$ contains a vertex of $Y_x$, either as an endpoint or in its interior.
 
@@
 We claim that $(S_x:x\in V(K))$ is a partition of $V(G)$.
-Indeed, any $v\in V(G)$ can only appear in a node $x$ such that $v\in V(N_x)\setminus V(F_x)$.
-These nodes form a path from the root of $K$ to the unique node $x'$ such that $v\in Y_{x'}$.
-It is worth noting that, if $vw\in E(G)$, and $v,w\in S_x$, then the embedding of $vw$ is contained in the interior of $F_x$ (otherwise $v$ and $w$ would have appeared in $S_a$ for some ancestor $a$ of $x$.)
+Indeed, if $v$ appears at nodes $x$ and $y$ where neither $x$ nor $y$ is an ancestor of the other, the $v$ contains a vertex of $Y_a$ for some common ancestor $a$ of $x$ and $y$.
 
@@
 \begin{thm}
-    Every $k$-planar graph $G$ has a $(3k,k9^{2k+1}/8)$-partition.
+    Every $k$-planar graph $G$ has a $(k9^{2k+1}/8,3k)$-partition.
 \end{thm}

### 2019-05-04 k-planar.tex
More correct explanations

@@
 We claim that $(S_x:x\in V(K))$ is a partition of $V(G)$.
-Indeed, if $v$ appears at nodes $x$ and $y$ where neither $x$ nor $y$ is an ancestor of the other, the $v$ contains a vertex of $Y_a$ for some common ancestor $a$ of $x$ and $y$.
+Indeed, if $v$ appears at nodes $x$ and $y$ where neither $x$ nor $y$ is an ancestor of the other, then there exists edges $vw,vz\in E(G)$ (possibly $w=z$) such that $vw$ contains a vertex of $Y_x$ (as an endpoint or in its interior) and $vz$ contains a vertex of $Y_y$ (as endpoint or in its interior).
+At least one of these two edges contains a vertex $Y_a$ for some common ancestor $a$ of $x$ and $y$.
 
@@
 (Don't be put off here by the fact that $V(H)=V(K)$).
-We do this by definining, for each $x\in V(H)$, the subtree $K[x]=K[\{y:x\in B_y\}]$.
+We claim that, for every edge $xz\in V(H)$, $x$ is an ancestor of $z$ (or vice-versa).
+Indeed, this follows from the same argument used to show that $S_x$ is a partition.
+If neither $x$ nor $z$ is an ancestor of the other, then any edge $vw$ with $v\in S_x$ and $w\in S_z$ contains a vertex of $Y_a$ for some common ancestor $a$ of $x$ and $z$.
+
+To describe our tree decomposition we define, for each $x\in V(H)$, the subtree $K[x]=K[\{y:x\in B_y\}]$.
 The subtree $K[x]$ contains $x$.
-Additionally, for every edge $xz\in E(H)$, $K[x]$ includes the path from $x$ to $z$ in $K$.
-This clearly produces a tree decomposition of $H$, what remains is to upper bound the width of this decomposition.
+Additionally, for every edge $xz\in E(H)$, where $x$ is an ancestor of $z$, $K[x]$ includes the path from $x$ to $z$ in $K$.
+That this produces a tree-decomposition is due to the fact that for every edge $xz$ of $H$, $x$ is an ancestor of $z$ (or vice-versa).
+What remains is to upper bound the width of this decomposition.
 
-We claim that $K[x]$ is rooted at $x$, i.e., $x$ is the lowest common ancestor of all nodes in $B_x$.
-To see why this is so, suppose $K[x]$ has root $a\neq x$.
-In short: This happens because some $v\in V(N_x)$ is incident on an edge $vw\in E(G)$ that crosses $Y_a$.
-Longer: This implies that $a$ has two children $b\neq c$ where $x$ is a descendant of $b$, some node $y$ is a descendant of $c$, and there is an edge $vw\in E(G)$ such that $v\in V(N_x)\setminus V(F_x)\subseteq V(N_b)\setminus V(F_a)$
-and $w\in S_y\subset V(N_y)\subseteq V(N_b)\setminus V(F_a)$.
-(Note, we allow $x=b$ and/or $y=c$ here.)  Therefore, $v$ and $w$ are in the interior of two different faces of the Mercedes graph $M_a$.
-Therefore the edge $vw$ intersects $M_a$.
-If it intersects $M_a$ at some vertex in $Y_a$, then $v$ would be in $S_a$, not $S_x$, a contradiction.
-If it intersects $M_a$ at some vertex in $F_a$, then $v$ would be in $S_{a'}$ for some ancestor $a'$ of $a$, also a contradiction.
-
+First observe that the subtree $K[x]$ is rooted at $x$, i.e., $x$ is the lowest common ancestor of all nodes in $B_x$.
 Now, fix some node $x\in V(K)$ and consider the contents of $B_x$.

### 2019-05-04 k-planar.tex
Clearer explanations

@@
 We use the partition $(S_x: x\in V(K))$ which generates a graph $H$ with vertex set $V(H)=V(K)$ the edge $xy\in E(H)$ if there exists an edge $vw\in E(H)$ with $v\in S_x$ and $w\in S_y$.
-
-First we argue that this partition has small width, with respect to the layering defined by $T$ (note that this is a layering of $G^+$, not a layering of $G$).
+First we argue that this partition has small width with respect to the layering defined by $T$ (note that this is a layering of $G^+$, not a layering of $G$).
 
@@
 Next we describe a tree decomposition $(B_x:x\in V(K))$ of $H$.
-(Don't be put off here by the fact that $V(H)=V(K)$).
+(Don't be put off here by the fact that $V(H)=V(K)$, this is actually helpful because the subtree $K[x]$ will be rooted at $x$).
 We claim that, for every edge $xz\in V(H)$, $x$ is an ancestor of $z$ (or vice-versa).
@@
 If neither $x$ nor $z$ is an ancestor of the other, then any edge $vw$ with $v\in S_x$ and $w\in S_z$ contains a vertex of $Y_a$ for some common ancestor $a$ of $x$ and $z$.
+But this implies that neither $v$ nor $w$ appears in $S_x$ nor $S_z$.
 

### 2019-05-04 k-planar.tex
Little bits of work before the house wakes up

@@
 
-\title{\MakeUppercase{Notes on Layered $H$-Partitions of $k$-Planar Graphs}%
+\title{\MakeUppercase{Layered $H$-Partitions of $k$-Planar Graphs}%
     \thanks{This work was partly funded by NSERC and MRI.}}
@@
 \begin{abstract}
-  TBD
-  % Layered $H$-partitions of graphs (with small layered width, in which $H$ has small treewidth) are a recently-introduced tool that have been used to solve longstanding problems on queue layouts, non-repetitive colouring, and 3-d graph drawing.
+  Layered $H$-partitions of graphs (with small layered width, in which $H$ has small treewidth) are a recently-introduced tool that have been used to solve longstanding problems on queue layouts, non-repetitive colouring, and 3-d graph drawing.
 Such partitions are known to exist for planar graphs and, more generally, bounded genus graphs.
-  % In the current paper, we prove that every $k$-planar graph has a layered $H$-partition of layered width $O(k^3)$ in which $H$ is planar and has treewidth 3.
-This implies that $k$-planar graphs have a queue number upper-bounded by a polynomial in $k$ and have non-repetitive chromatic number upper-bounded by an exponential function of $k$.
+In the current paper, we prove that every $k$-planar graph has a layered $H$-partition of layered width $O(k)$ in which $H$ has treewidth $k9^k$.
+This implies that $k$-planar graphs have both queue number and non-repetitive chromatic number upper-bounded by a function of $k$.
 \end{abstract}
@@
 
-\section{Another Way of Looking at It}
+\section{1-Planar Graphs}
+
+Put the 1-Planar stuff back here because the constants are better.
+
+\section{$k$-Planar Graphs}
 
@@
 
-A PS-Partition $(Y_x:x\in V(K))$ for a triangulation $\Delta$ and spanning-tree $T$ of $\Delta$ is a partition of $V(\Delta)$ whose elements are indexed by nodes of a rooted 3-ary tree $K$.
-Each bag $Y_x$ of a PS-partition is a tripod in $\Delta$, with respect to $T$.
-This means that $T[Y_x]$ has at most 3 components, each of which is a vertical path whose lower endpoints form a clique $\tau_x$ in $\Delta$.
-Associated with each node $x\in V(K)$ is a cycle $F_x$ in $\Delta$ and a near-triangulation $N_x$ consisting of the part of $\Delta$ contained in $F_x$.
-Note that $Y_x\subseteq V(N_x)\setminus V(F_x)$ and $V(N_x)\subseteq V(N_a)$ for every ancestor $a$ of $x$.
+Let $\Delta$ be a plane triangulation and let $T$ be a BFS spanning tree of $T$ rooted at some vertex $r$ of degree-3 in $\Delta$.
+Then the \emph{PS-tree} for $(\Delta, T, r)$ is a 3-ary tree $K$ that such that each node $x\in V(K)$ is associated with several objects:
 
-It is also useful to have a name for the \emph{Mercedes graph} $M_x$ that consists of the edges and vertices of $F_x$, the clique $\tau_x$, and the paths in $T$ from the vertices of $\tau_x$ to $F_x$.
+\begin{enumerate}
+  \item  A cycle $F_x$ in $\Delta$.
+There are at most three ancestors $a$, $b$, and $c$, of $x$ in $T$ such that the vertices of $F_x$ are contained in $V(Y_a)\cup V(Y_b)\cup V(Y_c)$.
+(See below for the definition of $Y_a$ $Y_b$, and $Y_c$.)
 
-In short, $K$ is the 3-ary tree of tripods we get from our proof.
-It's the tree we use for our tree-decomposition of $H$.
-We've just added some decorations to it.
-Eventually I should write this as a lemma with bullet points for the relevant properties.
+  \item $A_x=F[Y_a]$, $B_x=F[Y_b]$, $C_x=F[Y_c]$
+
+  \item A near-triangulation $N_x$ consisting of the part of $\Delta$ contained in $F_x$.  Note that $Y_x\subseteq V(N_x)\setminus V(F_x)$ and $V(N_x)\subseteq V(N_a)$ for every ancestor $a$ of $x$.
+
+  \item A \emph{tripod} $Y_x$ in $\Delta$, with respect to $T$.
+This is a clique $\tau_x\in N(x)$ and at most three disjoint vertical paths in $T$ whose lower endpoints are in $\tau_x$.
+Each of these paths is contained in $N_x-V(F_x)$ and the upper endpoint of each path is adjacent, in $T$ to a vertex of $F_x$.
+Furthermore, for each of $A_x$, $B_x$, and $C_x$, at least one upper endpoint is adjacent, in $T$ to it.
+
+  \item A \emph{Mercedes graph} $M_x$ that consists of the edges and vertices of $F_x$, the clique $\tau_x$, and the paths in $T$ from the vertices of $\tau_x$ to $F_x$.
+\end{enumerate}
+
+{\color{red}Expand the preceding points more fully and point out the relationsships better.  Deal with degenernate cases.}
 

### 2019-05-06 k-planar.tex
Cleaning up arguments

@@
 
-\section{1-Planar Graphs}
+\section{$1$-Plane Graphs}
 
-Put the 1-Planar stuff back here because the constants are better.
+Let $G$ be an edge-maximal 1-plane multigraph graph.
+Here, edge-maximal should be taken to mean that, if any two vertices $u$ and $v$ appear on a common face $F$, then there is an edge $uv\in E(G)$ that is contained in the boundary of $F$.
+
+A \emph{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.
+It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
+The edges $vw$ and $xy$ are called \emph{spars} of $K$.
+The cycle $vxwy$ is called the \emph{sail} of $K$.
+It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$.
+Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
+Observe that any spar of $K$ is incident on exactly four \emph{kite faces} of $G$, each of which has three edges and two vertices of $G$ on its boundary.
+
+The 1-plane graph $G$ has a plane triangulation $G'$ as a subgraph that can be obtained by removing one spar from each kite in $G$.
+Observe that, for any spar $vw\in E(G)\setminus E(G')$ that crosses $xy\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
+It follows that $\dist_{G'}(x,y)\le 2\dist_G(x,y)$ for any pair of vertices $x,y\in V(G)$.
+
+\begin{lem}\lemlabel{induction} The setup:
+  \begin{compactenum}
+    \item Let $G$ and $G'$ be defined as above.
+    \item Let $T$ be a BFS tree of $G'$ rooted at some vertex $r$.
+    \item For each integer $i\ge 0$, let $V_i=\{v\in V(G):\dist_T(r,v)=i\}$. 
+    \item Let $C$ be a cycle in $G'$ with $r$ in the exterior of $C$ and such that
+    \begin{compactenum} 
+      \item No edge of $C$ is crossed by any edge of $G$; and
+      \item $V(C)$ can be partitioned into $P_1,\ldots,P_k$, $k\le 3$ such that for each $i\in\{1,\ldots,k\}$,
+      \begin{compactenum}
+        \item $C[P_i]$ is a path; and
+        \item $|P_i\cap V_j| \le 15$ for all integers $j\ge 0$.
+      \end{compactenum}
+    \end{compactenum}
+    \item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $C$ or the interior of $C$.
+  \end{compactenum}
+  Then $N$ has an $H$-partition $(B_x : x\in V(H))$ such that
+  \begin{compactenum}
+    \item $H$ is planar;
+    \item for all integers $j\ge 0$, and all $x\in V(H)$, $|B_x\cap V_j|\le 15$; 
+    \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=B_{x_i}$; and
+    \item $H$ has a tree decomposition whose largest bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
+  \end{compactenum}
+\end{lem}
+
+\begin{proof}
+  The proof is by induction on the number of vertices of $N$.
+  First note that $N'$ is a near-triangulation.
+If $k=3$, set $R_i=P_i$ for each $i\in\{1,2,3\}$.
+Otherwise, as before, split $P_1,\ldots,P_k$ to partition $V(C)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $C[R_i]$ is a path and each $R_i$ contains vertices from at most one of $P_1,\ldots,P_k$.
+
+  For each $i\in\{1,2,3\}$ and $v\in R_i$, assign the colour $i$ to $v$.
+  For each $v\in V(N')$, consider the path $P_v$ that consists of the subpath of the $v$ to $r$ path in $T$ that stops at the first vertex $v'\in V(C)$. Assign $v$ to have the same colour as $v'$.
+
+  Sperner's Lemma ensures that $G'$ contains a triangular face $\tau=v_1v_2v_3$ whose three vertices are assigned different colours.
+For each $i\in\{1,2,3\}$, let $Q_i=P_{v_i}$ be the path in $T$ form $v_i$ to $v_i'$.
+Let $Y$ denote the subgraph of $N'$ consisting of vertices and edges $Q_1$, $Q_2$, $Q_3$, and $\tau$.
+Let $Y^+$ denote the subgraph of $N$ consisting of the vertices and edges of $Y$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $Y$.
+
+  We claim that, for each integer $i\ge 0$, $|V(Y^+)\cap V_i|\le 15$.
+To see this, first observe that $Y$ contains at most 3 vertices of $V_i$.
+If a vertex $x\in V(Y^+)\setminus V(Y)$ is contained in $V_i$, then this is because $Y$ contains an edge $vw$ with $v\in V_{i'}$ and $w\in\{V_{i'+1}$ for some $i'\in\{i-1,i\}$ and $G$ contains an edge $xy$ that crosses $vw$.
+(Recall that $vx,xw\in E(G')$, so $\dist_{G'}(w,r)-1\le\dist_{G'}(x,r)\le\dist_{G'}(v,r)+1$.)  The graph $Y$ contains at most 6 such edges, each of which accounts for at most 2 additional vertices of $V_i$.
+Therefore, in total, $|Y^+\cap V_i|\le 15$.
+
+  Finally, let $S$ and $S^+$ denote the subgraph of $G$ containing the edges and vertices of $Y$, respectively $Y^+$, and the edges and vertices of $C$.
+The graph $S^+$ has some number of bounded faces, all contained in the interior of $C$.
+Some of the bounded faces of $S$ are kite faces.
+Call the non-kite bounded faces $F_1,\ldots,F_m$ and let $C_1,\ldots,C_m$ denote their boundaries.
+We claim that, for each $i\in\{1,\ldots,m\}$, if some portion of $C_i$ is contained in an edge $vw\in E(G)$ then $vw$ is not crossed by any edge of $G$.
+To see this, there are three cases to consider:
+  \begin{enumerate}
+    \item $vw\in E(C)$. By assumption, $vw$ is not crossed by any edge of $G$.
+    \item $vw\in E(Y)$. In this case, the kite containing $vw$ is in $Y^+$, so $vw$ is only incident to kite faces.
+    \item $vw\in E(Y^+)\setminus E(Y)$.
+In this case, either $vw$ is a sail edge in which case it is not crossed by definition, or $vw$ is a spar that was added to $Y^+$ because $vw$ crosses some edge $xy\in E(Y)$.
+In this latter case, $vw$ is only incident to kite faces.
+  \end{enumerate}
+  Therefore each $C_i$ is a cycle in $G^+$ consisting entirely of uncrossed edges.
+The vertices of $C_i$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$ and $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$.
+Furthermore $C_i[P_j']$ is a path for each $j\in\{1,\ldots,3\}$.
+Finally, the subgraph $N_i$ of $G$ consisting of the edges and vertices of $G$ contained in $C_i$ or its interior does not contain one of the three vertices of $\tau$.
+Therefore, we can apply induction using the cycle $C_i$ and the partition $P_1',P_2',P_3'$ of $V(C_i)$ to obtain the desired $H$-partition and tree decomposition of $N_i$.
+
+  The remainder of the proof is as before.
+We use $V(Y^+)\setminus V(C)$ as an element of our $H$-partition.
+The root bag in the tree decomposition contains the vertices obtained by contracting each of $P_1,\ldots,P_k$ and $Y^+$.
+
+  Planarity comes from the fact that if two edges $vw$ and $xy$ cross, then they end up in the same bag of the $H$-partition.
+This means that $H$ is actually obtained by contracting connected sets of vertices in the planar graph $G'$.
+\end{proof}
+
+\begin{thm}\thmlabel{1-plane}
+  Every 1-plane graph $G$ has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
+\end{thm}
+
+\begin{proof}
+  The same as before except that the layered width 15 from \lemref{induction} becomes 30 because \lemref{induction} uses a layering of $G'$ and distances in $G'$ can be a factor of 2 larger than in $G$.
+\end{proof}
 
@@
 
-\subsection{PS-Tree}
+\subsection{The PS-Tree}
 
-Let $\Delta$ be a plane triangulation and let $T$ be a BFS spanning tree of $T$ rooted at some vertex $r$ of degree-3 in $\Delta$.
-Then the \emph{PS-tree} for $(\Delta, T, r)$ is a 3-ary tree $K$ that such that each node $x\in V(K)$ is associated with several objects:
+Let $T$ be a rooted tree rooted at $r\in V(T)$.
+The \emph{depth} of a node $x\in V(T)$ is equal to the length of the path from $x$ to $r$ in $T$.
+A $P$ path in $T$ is \emph{vertical} if, for each integer $i$, there is at most one vertex of depth $i$.
+The deepest vertex in a vertical path $P$ is called  $P$'s \emph{lower endpoint} and the other (shallowest) vertex in $P$ is called $P$'s \emph{upper endpoint}.
+For any node $v\in T$, we call each node on the path, in $T$, from $v$ to $r$ a \emph{$T$-ancestor} of $v$.
+
+A \emph{near-triangulation} $N$ is a plane graph whose outer face is bounded by a simple cycle in $N$ and each of whose inner faces is bounded by a 3-cycle in $N$.
+
+Let $\Delta$ be a plane triangulation and let $T$ be a BFS spanning tree of $T$ rooted at some vertex $r$ of degree 3 that is on the outer face of $\Delta$.
+For any cycle $F$ of $\Delta$ there is an associated near-triangulation $N=\Delta\circledast F$ consisting of all the edges and vertices of $G$ that are contained in $F$ or its interior.
+
+Let $\tau=v_1v_2v_3$ be an inner face of $N$ and consider the three minimal vertical paths $Q_1$, $Q_2$, and $Q_3$ in $T$ such that, for each $i\in\{1,2,3\}$ the lower endpoint of $Q_i$ is $v_i$ and the upper endpoint of $Q_i$ is in $V(F)$.
+If $Q_1$, $Q_2$, and $Q_3$ are vertex disjoint, then the graph $\overline{Y}=\tau\cup Q_1\cup Q_2\cup Q_3$ is called the \emph{closed tripod} in $N$ determined by $\tau$.
+We call $Q_1$, $Q_2$, and $Q_3$ the \emph{legs} of $\overline{Y}$ and we call $\tau$ the \emph{center} of $\overline{Y}$.
+We call $Y=\overline{Y}-F$ the \emph{(open) tripod} of $N$ determined by $\tau$.
+The graph $M=\overline{Y}\cup F$ is called a \emph{Mercedes graph} and the (at most 3) inner faces of $M$ other than $\tau$ are called the \emph{wedges} of $M$.
+
+A \emph{PS-tree}  $K=K(\Delta, T, r)$ is a 3-ary tree, such that each node $x\in V(K)$ is associated with several object (refer to \figref{K-node}):
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{cc}
+      \includegraphics{figs/K-node-1} &
+      \includegraphics{figs/K-node-2}
+    \end{tabular}
+  \end{center}
+  \caption{The elements associated with a node $x\in V(K)$.}
+  \figlabel{K-node}
+\end{figure}
 
 \begin{enumerate}
-  \item  A cycle $F_x$ in $\Delta$.
-There are at most three ancestors $a$, $b$, and $c$, of $x$ in $T$ such that the vertices of $F_x$ are contained in $V(Y_a)\cup V(Y_b)\cup V(Y_c)$.
-(See below for the definition of $Y_a$ $Y_b$, and $Y_c$.)
+  \item A cycle $F_x$ in $\Delta$ and the near-triangulation $N_x=\Delta\circledast F_x$.
 
-  \item $A_x=F[Y_a]$, $B_x=F[Y_b]$, $C_x=F[Y_c]$
+  \item An inner face $\tau_x$ of $N_x$ that determines a closed tripod $\overline{Y}_x$ and open tripod $Y_x$ in $N_x$.
 
-  \item A near-triangulation $N_x$ consisting of the part of $\Delta$ contained in $F_x$.  Note that $Y_x\subseteq V(N_x)\setminus V(F_x)$ and $V(N_x)\subseteq V(N_a)$ for every ancestor $a$ of $x$.
-
-  \item A \emph{tripod} $Y_x$ in $\Delta$, with respect to $T$.
-This is a clique $\tau_x\in N(x)$ and at most three disjoint vertical paths in $T$ whose lower endpoints are in $\tau_x$.
-Each of these paths is contained in $N_x-V(F_x)$ and the upper endpoint of each path is adjacent, in $T$ to a vertex of $F_x$.
-Furthermore, for each of $A_x$, $B_x$, and $C_x$, at least one upper endpoint is adjacent, in $T$ to it.
-
-  \item A \emph{Mercedes graph} $M_x$ that consists of the edges and vertices of $F_x$, the clique $\tau_x$, and the paths in $T$ from the vertices of $\tau_x$ to $F_x$.
+  \item The \emph{Mercedes graph} $M_x=F_x\cup \overline{Y}_x$
 \end{enumerate}
 
-{\color{red}Expand the preceding points more fully and point out the relationsships better.  Deal with degenernate cases.}
+The nodes of the PS-tree $K$ satisfy several conditions:
+\begin{compactenum}[(PR1)]
+  \item For the root $r'$ of $K$, $F_{r'}$ is defined to be the 3-cycle in $\Delta$ formed by the three vertices adjacent to $r$.
+
+  \item \label{children} For each node $x\in V(T)$ such that $M_x$ has $f\le 3$ wedges $F_{x,1},\ldots,F_{x,f}$, the node $x$ has exactly $f$ children $y_1,\ldots,y_f$ in $K$ and $F_{y_i} = F_{x,i}$ for each $i\in\{1,\ldots,f\}$.
+
+  \item \label{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$ \emph{or} $F[P_i]$ is a path in $F_{r'}$.
+
+  \item \label{partition} $\{Y_x : x\in V(K)\}$ is a partition of $V(G^+)\setminus\{r\}$.
+
+  \item Something else?
+\end{compactenum}
+
+Properties~(PR1)--(PR3) are defining properties of a PS-tree, and they imply the remaining properties.
+
+% There are at most three ancestors $a$, $b$, and $c$, of $x$ in $T$ such that the vertices of $F_x$ are contained in $V(Y_a)\cup V(Y_b)\cup V(Y_c)$.
+(See below for the definition of $Y_a$ $Y_b$, and $Y_c$.)
+% 
+%  consisting of the part of $\Delta$ contained in $F_x$.  Note that $Y_x\subseteq V(N_x)\setminus V(F_x)$ and $V(N_x)\subseteq V(N_a)$ for every ancestor $a$ of $x$.
+% 
+% {\color{red}Expand the preceding points more fully and point out the relationsships better.  Deal with degenernate cases.}
 
@@
 
-Let $G$ be a $k$-plane graph and let $G^+$ be the graph obtained by making $G$ planar with dummy vertices and then triangulating the resulting graph.
-Let $T$ be a BFS spanning tree of $G^+$ where the root $r$ of $T$ has degree 3 in $G^+$ (add an additional vertex to $G^+$ if necessary).
+Let $G$ be a $k$-plane graph and let $G^+$ be the graph obtained by making $G$ planar with dummy vertices, then triangulating the resulting graph and then adding a degree-3 vertex $r$ in the outer face of the resulting graph.
+Let $T$ be a BFS spanning tree of $G^+$ rooted at $r$.
 For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{G^+}(r,v)=i\}$.
+Observe that, for any edge $vw\in E(G)$, $\dist_{G^+}(v,w) \le k$ and therefore, for any two vertices $v,w\in V(G)$,
+\begin{equation}
+  \dist_{G^+}(v,w) \le k\cdot \dist_G(v,w) \eqlabel{distance-preserving}
+\end{equation}
 
-Let $K$ be the PS-partition for $G^+$ and $T$.
-We say that a vertex $v\in V(G)$ \emph{appears} at a node $x\in V(K)$ if $u\in Y_x$ or if there exists $vw\in E(G)$ such that $v,w\in V(N_x)$ and $vw$ contains a vertex of $Y_x$, either as an endpoint or in its interior.
+Let $K=K(G^+,T,r)$ be a PS-tree.
+We say that a vertex $v\in V(G)$ \emph{appears} at a node $x\in V(K)$ if $v\in Y_x$ or if there exists $vw\in E(G)$ such that $v,w\in V(N_x)$ and $vw$ contains a vertex of $Y_x$, either as an endpoint or in its interior (see \figref{appearances}).
 
-Let $S_x$ (the separator at $x$) denote the set of all vertices $v\in V(G)$ such that $v$ appears at $x$ but does not appear at an ancestor of $x$.
+\begin{figure}
+  \begin{center}
+    \includegraphics{figs/K-node-3}
+  \end{center}
+  \caption{Examples of vertices (in peach) that appear at a node $x\in V(K)$.}
+  \figlabel{appearances}
+\end{figure}
+
+Let $S_x$ (the separator at $x$) denote the set of all vertices $v\in V(G)$ such that $v$ appears at $x$ but does not appear at any $K$-ancestor of $x$.
 We claim that $(S_x:x\in V(K))$ is a partition of $V(G)$.
-Indeed, if $v$ appears at nodes $x$ and $y$ where neither $x$ nor $y$ is an ancestor of the other, then there exists edges $vw,vz\in E(G)$ (possibly $w=z$) such that $vw$ contains a vertex of $Y_x$ (as an endpoint or in its interior) and $vz$ contains a vertex of $Y_y$ (as endpoint or in its interior).
+By (PR\ref{partition}) every vertex $v\in  V(G)$ is in $V(Y_x)\subseteq V(N_x)$.
+Therefore every vertex $v\in V(G)$ appears at some node $x\in V(K)$, so $\bigcup_{x\in V(K)} S_x \supseteq V(G)$.
+
+Furthermore, if $v$ appears at two nodes $x$ and $y$ then, by definition, $v\in V(N_x)\cap V(N_y)$.
+If neither $x$ nor $y$ is an ancestor of the other then, by (PR\ref{children}), $V(N_x)\cap V(N_y)=V(F_x)\cap V(F_y)$.
+By (PR\ref{ancestor-boundary}), $v\in Y_a$ for some $K$-ancestor $a$ of $x$.
+Similarly, $v\in Y_{b}$ for some ancestor $b$ of $y$.
+By (PR\ref{partition}), $a=b$ is a common $K$-ancestor of $x$ and $y$.
+It follows that there is a unique node $a'$ that is an ancestor of all nodes of $K$ at which $v$ appears, so $v$ appears only in $Y_{a'}$.
+
+% Indeed, if $v$ appears at nodes $x$ and $y$ where neither $x$ nor $y$ is an ancestor of the other, then there exists edges $vw,vz\in E(G)$ (possibly $w=z$) such that $vw$ contains a vertex of $Y_x$ (as an endpoint or in its interior) and $vz$ contains a vertex of $Y_y$ (as endpoint or in its interior).
 At least one of these two edges contains a vertex $Y_a$ for some common ancestor $a$ of $x$ and $y$.
 
-We use the partition $(S_x: x\in V(K))$ which generates a graph $H$ with vertex set $V(H)=V(K)$ the edge $xy\in E(H)$ if there exists an edge $vw\in E(H)$ with $v\in S_x$ and $w\in S_y$.
+We use the partition $(S_x: x\in V(K))$ which generates a graph $H$ with vertex set $V(H)=V(K)$ the edge $xy\in E(H)$ if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
 First we argue that this partition has small width with respect to the layering defined by $T$ (note that this is a layering of $G^+$, not a layering of $G$).

### 2019-05-06 k-planar.tex
Cleaning up

@@
 Next we describe a tree decomposition $(B_x:x\in V(K))$ of $H$.
-(Don't be put off here by the fact that $V(H)=V(K)$, this is actually helpful because the subtree $K[x]$ will be rooted at $x$).
-We claim that, for every edge $xz\in V(H)$, $x$ is an ancestor of $z$ (or vice-versa).
-Indeed, this follows from the same argument used to show that $S_x$ is a partition.
-If neither $x$ nor $z$ is an ancestor of the other, then any edge $vw$ with $v\in S_x$ and $w\in S_z$ contains a vertex of $Y_a$ for some common ancestor $a$ of $x$ and $z$.
-But this implies that neither $v$ nor $w$ appears in $S_x$ nor $S_z$.
+We deliberately use the same node set $V(H)=V(K)$ here, because the subtree $K[x]:=K[\{z:x\in B_z\}]$ will be rooted at $x$.
+We claim that, for every edge $xy\in V(H)$, $x$ is an ancestor of $y$ (or vice-versa).
+Suppose that this is not the case and that $xy\in E(H)$ but neither $x$ nor $y$ is an ancestor of the other.
+Then, since $xy\in E(H)$, there must exist an edge $vw\in E(G)$ such that $v\in S_x$ and $w\in S_y$.
+
+Since $v\in S_x$, $v\in V(N_x)$ and similarly $w\in V(N_y)$.
+Since neither $x$ nor $y$ are ancestors of each other, the interiors of $F_x$ and $F_y$ are disjoint.
+If $v$ is in the interior $V(N_x)\setminus V(F_x)$ of $F_x$ then the edge $vw$ crosses $F_x$ and by (PR\ref{ancestor-boundary}) $v$ appears at some ancestor $a$ of $x$.
+(The node $a$ is the lowest node for which $v,w\in V(N_a)$.)  But this contradicts the assumption that $v\in S_x$, so it must be that $v\in V(F_x)$.
+This is also a contradiction, since (PR\ref{ancestor-boundary}) then implies that $v\in Y_a$ and therefore appears at $a$ for some $K$-ancestor $a$ of $x$, which contradicts the assumption that $v\in V(N_x)$.
 

### 2019-05-06 k-planar.tex
Time to quit for the day

@@
 
-First observe that the subtree $K[x]$ is rooted at $x$, i.e., $x$ is the lowest common ancestor of all nodes in $B_x$.
+First observe that, for every $a\in V(K)$, the subtree $K[a]$ is rooted at $a$, i.e., $a$ is a $K$-ancestor of every node in $x\in B_a$.
 Now, fix some node $x\in V(K)$ and consider the contents of $B_x$.
@@
 If $B_x$ contains some ancestor $a$ of $x$, it is because there is an edge $vw\in E(G)$ with $v\in V(N_x)\setminus V(F(x))$ and $w\in S_a$.
-Now, $w\in S_a$ because $w\in Y_a$ or there is an edge $wz\in E(G)$ that contains a point of $Y_a$.
+{\color{red} CONTINUE here: Explain why $v\in V(N_x)\setminus V(F(x))$.}  Now, $w\in S_a$ because $w\in Y_a$ or there is an edge $wz\in E(G)$ that contains a point of $Y_a$.
 This implies that there is a curve $C$ from $v$ to $z$ that contains at most $2k$ dummy  vertices of $G^+$.

### 2019-05-07 k-planar.tex
Checking

@@
 As we have just argued, $B_x$ contains only ancestors of $x$ (including $x$ itself).
-If $B_x$ contains some ancestor $a$ of $x$, it is because there is an edge $vw\in E(G)$ with $v\in V(N_x)\setminus V(F(x))$ and $w\in S_a$.
-{\color{red} CONTINUE here: Explain why $v\in V(N_x)\setminus V(F(x))$.}  Now, $w\in S_a$ because $w\in Y_a$ or there is an edge $wz\in E(G)$ that contains a point of $Y_a$.
-This implies that there is a curve $C$ from $v$ to $z$ that contains at most $2k$ dummy  vertices of $G^+$.
+If $B_x$ contains some ancestor $a$ of $x$, it is because there is an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_a$.
+Since $v\in S_x$, $v\in V(N_x)$.
+Since, $w\in S_a$, $w\in Y_a$ or there is an edge $wz\in E(G)$ that contains a point of $Y_a$.
+The embedding of the path $vwz$ is a curve that being at $v\in N_x$, ends at
+$z$, contains a point of $V(Y_a)$ and contains at most $2k+1$ vertices of $G^+$.
 
@@
 See \figref{A}.
-It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0$ is the root of $K$.
-We then cut $F_{a_0}$ into two pieces $F^{\bar{x}}_{a_0}$ and $F^x_{a_0}$ where $F^x_{a_0}$ contains $F_x$.
+It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0=r'$ is the root of $K$.
+We then cut $F_{a_0}$ into two pieces $F^{\bar{x}}_{a_0}$ and $F^x_{a_0}$ where $F^x_{a_0}$ is the piece that contains $F_x$.
 This process continues by partitioning $F_{a_1}:=F^x_{a_0}$ recursively.
@@
 We claim that the curve $C$ begins in $F_x\subseteq F^x_a\subseteq F_a$, remains in the interior of $F_a$, and eventually reaches the boundary $F^{\bar{x}}_a\cap F^x_a$.
-The fact that $C$ remains in $F_a$ is due to the fact that, if it crossed the boundary of $F_a$, then one of $vw$ or $wz$ would interects $Y_{a'}$ for some ancestor $a'$ of $a$.
-But this would contradict the fact that $vw$ (or $wz$) are in $S_a$.
-The fact that $C$ eventually reaches $F^{\bar{x}}_a\cap F^x_a$ follows from the fact that $w\in S_a$ (so either $w\in Y_a$ or $wz$ crosses $Y_a$).
+The fact that $C$ remains in $F_a$ is due to the fact that, if it crossed the boundary of $F_a$, then one of $vw$ or $wz$ would intersect $Y_{a'}$ for some ancestor $a'$ of $a$.
+If $vw$ intersects $Y_{a'}$, then $v\not\in S_x$.
+If $wz$ intersects $Y_{a'}$ then $w,z\not\in S_a$.
+The fact that $C$ eventually reaches $F^{\bar{x}}_a\cap F^x_a$ follows from the fact that $w\in S_a$ (so either $w\in Y_a$ or $wz$ crosses $Y_a$) {\color{red} Recheck the reasoning in this last line.}.
 

### 2019-05-07 k-planar.tex
Looking gooder

@@
 We then cut $F_{a_0}$ into two pieces $F^{\bar{x}}_{a_0}$ and $F^x_{a_0}$ where $F^x_{a_0}$ is the piece that contains $F_x$.
-This process continues by partitioning $F_{a_1}:=F^x_{a_0}$ recursively.
+(The path that makes this cut contains two legs $Q_i$ and $Q_j$ of a tripod $Y$ and one edge $v_iv_j$ of its central triangle $\tau$.)  This process continues by partitioning $F_{a_1}:=F^x_{a_0}$ recursively.
 The process terminates after some number $r$ of iterations, when $F^x_{a_r}=F_x$.
@@
 If $wz$ intersects $Y_{a'}$ then $w,z\not\in S_a$.
-The fact that $C$ eventually reaches $F^{\bar{x}}_a\cap F^x_a$ follows from the fact that $w\in S_a$ (so either $w\in Y_a$ or $wz$ crosses $Y_a$) {\color{red} Recheck the reasoning in this last line.}.
+The fact that $C$ eventually reaches follows from the fact that $w\in Y_a$ or $wz$ contains a point of $Y_a$.
 
-Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of faces $F^{\bar{x}}_a$ of $A$ that can be reached with curves that begin in $F_x$, remain in $F_a$, and terminate in $F^{\bar{x}}_a$ without crossing more than $2k$ edges in $A$.
+Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of faces $F^{\bar{x}}_a$ of $A$ that can be reached with curves that begin in $F_x$, remain in the interior of $F_a$, and terminate in $F^{\bar{x}}_a$ without intersecting more than $2k$ edges in $A$.
 To avoid triple subscripts we say that a curve visits $a_i$ if it visits $F^{\bar{x}}_{a_i}$.
 We can further restrict ourselves to studying curves that cross the fewest number of edges of $A$ since if a curve exists that crosses at most $k$ edges then the curve that crosses the fewest number of possible edges also crosses at most $k$ edges.
-We will call such a curve a lightest curve from $F_x$ to $F_a$.
+We will call such a curve a \emph{lightest} curve from $x$ to $a$.
 
 Again, we let $a_0,\ldots,a_r$ denote the vertices of the path, in $K$ from the root $a_0$ to $x=a_r$.
-We claim that, there exists a lightest curve from $a_r$ to $a_0$ that is monotonic in the sense that the curve visits $a_{i_0},\ldots,a_{i_t}$ in this order, where $i_0=r > i_1>\cdots>i_t=0$ the Indeed, this follows from the fact that, for any node $a_i$, $i>0$, the (three) faces of $A$ that share a boundary with $F_{a_i}$ also share a boundary with $F^{\bar{x}}_{a_i}$.
+We claim that, there exists a lightest curve from $a_r$ to $a_0$ that is monotonic in the sense that the curve visits $a_{i_0},\ldots,a_{i_t}$ in this order, where $i_0=r > i_1>\cdots>i_t=0$.
+This follows from (PR\ref{ancestor-boundaries}):  For any node $a_i$, $i>0$, there are at most three faces of $A$ that share a part of their boundary with $F_{a_i}$.
+Each of these faces also necessarily shares a part of its boundary with $F^{\bar{x}}_{a_i}$.
 Therefore it is never advantageous for a path to go from $a_i$ to $a_j$ for some $j>i$ since the path will eventually have to return to one of these three faces.

### 2019-05-07 k-planar.tex
This is working cleanly

@@
 
-  \item \label{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$ \emph{or} $F[P_i]$ is a path in $F_{r'}$.
+  \item \label{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$ \emph{or} $F[P_i]$ is a path or cycle in $F_{r'}$.
 
@@
 
-  \item Something else?
+  % \item Something else?
 \end{compactenum}
 
-Properties~(PR1)--(PR3) are defining properties of a PS-tree, and they imply the remaining properties.
+Properties~(PR1)--(PR3) are defining properties of a PS-tree. Property~(PR4)
 
@@
 The fact that $C$ eventually reaches follows from the fact that $w\in Y_a$ or $wz$ contains a point of $Y_a$.
+We call such a curve an \emph{admissible} curve.
 
@@
 We can further restrict ourselves to studying curves that cross the fewest number of edges of $A$ since if a curve exists that crosses at most $k$ edges then the curve that crosses the fewest number of possible edges also crosses at most $k$ edges.
-We will call such a curve a \emph{lightest} curve from $x$ to $a$.
+We will call such a curve a \emph{lightest} admissible curve from $x$ to $a$.
 
 Again, we let $a_0,\ldots,a_r$ denote the vertices of the path, in $K$ from the root $a_0$ to $x=a_r$.
-We claim that, there exists a lightest curve from $a_r$ to $a_0$ that is monotonic in the sense that the curve visits $a_{i_0},\ldots,a_{i_t}$ in this order, where $i_0=r > i_1>\cdots>i_t=0$.
+We claim that, there exists a lightest admissible curve from $a_r$ to $a_0$ that is monotonic in the sense that the curve visits $a_{i_0},\ldots,a_{i_t}$ in this order, where $i_0=r > i_1>\cdots>i_t=0$.
 This follows from (PR\ref{ancestor-boundaries}):  For any node $a_i$, $i>0$, there are at most three faces of $A$ that share a part of their boundary with $F_{a_i}$.
 Each of these faces also necessarily shares a part of its boundary with $F^{\bar{x}}_{a_i}$.
-Therefore it is never advantageous for a path to go from $a_i$ to $a_j$ for some $j>i$ since the path will eventually have to return to one of these three faces.
+Therefore no lightest admissible curve ever $a_i$ before $a_j$ for some $j>i$ since the curve will eventually have to return to one of these three faces.
 
 At this point we are essentially done.
-Any monotonic curve that visits $a_i$ has at most 3 choices for the next face $a_j$ $j<i$ to visit.
-Therefore, the number of faces of $A$ that can be reached from $F_x$ by admissible curves that cross $2k$ edges of $A$ is at most $3^{2k}$.
+Any lightest admissible curve that visits $a_i$ has at most 3 choices for the next face $a_j$ $j<i$ to visit.
+Therefore, the number of faces of $A$ that can be reached from $F_x$ by admissible curves that intersect at most $\ell$ edges of $A$ is at most $3^{\ell}$.
 Therefore,
 \[
-   |B_x| \le \sum_{j=0}^2k 9^j = (9^{2k+1}-1)/8 \enspace .
+   |B_x| \le \sum_{\ell=0}^{2k+1} 3^\ell = (3^{2k+2}-1)/2 
+       = (9^{k+1}-1)/2 \enspace .
 \]
 
-This completes the proof
+This is all we need to complete the proof of:
 
 \begin{thm}
-    Every $k$-planar graph $G$ has a $(k9^{2k+1}/8,3k)$-partition.
+    Every $k$-planar graph $G$ has a $((9^{k+1}-1)/2-1,3k^2)$-partition.
 \end{thm}
+
+\begin{proof}
+  The $H$-partition $(B_x:x\in V(H))$ of $G$ described above has treewidth $(9^{k+1}-1)/2-1$.
+For each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 3k$.
+\Eqref{distance-preserving} implies setting $V_i' = V_{ki}\cup\cdots\cup V_{k(i+1)-1}$ yields a layering $(V_0',V_1',\ldots)$ of  $G$ and $|B_x\cap V_i'|\le 3k^2$ for all integers $i\ge 0$ and all $x\in V(H)$.
+\end{proof}
 

### 2019-05-07 k-planar.tex
Draft sent to Vida and David

--- 2019-05-07 k-planar.tex
@@
 \DeclareMathOperator{\ff}{f}
-\newcommand{\f}{{\color{red}\ff}}
 
-\newcommand{\fk}{{\color{red}f(k)}}
+\renewcommand{\proplabel}[1]{\label{prop:#1}}
+\renewcommand{\propref}[1]{(PR\ref{prop:#1})}
 
@@
 The \emph{depth} of a node $x\in V(T)$ is equal to the length of the path from $x$ to $r$ in $T$.
-A $P$ path in $T$ is \emph{vertical} if, for each integer $i$, there is at most one vertex of depth $i$.
+A $P$ path in $T$ is \emph{vertical} if, for each integer $i$, $P$ contains at most one vertex of depth $i$.
 The deepest vertex in a vertical path $P$ is called  $P$'s \emph{lower endpoint} and the other (shallowest) vertex in $P$ is called $P$'s \emph{upper endpoint}.
@@
 
-Let $\Delta$ be a plane triangulation and let $T$ be a BFS spanning tree of $T$ rooted at some vertex $r$ of degree 3 that is on the outer face of $\Delta$.
+Let $\Delta$ be a plane triangulation and let $T$ be a BFS spanning tree of $\Delta$ rooted at some vertex $r$ of degree 3 that is on the outer face of $\Delta$.
 For any cycle $F$ of $\Delta$ there is an associated near-triangulation $N=\Delta\circledast F$ consisting of all the edges and vertices of $G$ that are contained in $F$ or its interior.
@@
 
-  \item An inner face $\tau_x$ of $N_x$ that determines a closed tripod $\overline{Y}_x$ and open tripod $Y_x$ in $N_x$.
+  \item An inner face $\tau_x$ of $N_x$ that determines a closed tripod $\overline{Y}_x$ and an open tripod $Y_x$ in $N_x$.
 
@@
 The nodes of the PS-tree $K$ satisfy several conditions:
-\begin{compactenum}[(PR1)]
+\begin{enumerate}[(PR1)]
   \item For the root $r'$ of $K$, $F_{r'}$ is defined to be the 3-cycle in $\Delta$ formed by the three vertices adjacent to $r$.
 
-  \item \label{children} For each node $x\in V(T)$ such that $M_x$ has $f\le 3$ wedges $F_{x,1},\ldots,F_{x,f}$, the node $x$ has exactly $f$ children $y_1,\ldots,y_f$ in $K$ and $F_{y_i} = F_{x,i}$ for each $i\in\{1,\ldots,f\}$.
+  \item \proplabel{children} For each node $x\in V(T)$ such that $M_x$ has $f\le 3$ wedges $F_{x,1},\ldots,F_{x,f}$, the node $x$ has exactly $f$ children $y_1,\ldots,y_f$ in $K$ and $F_{y_i} = F_{x,i}$ for each $i\in\{1,\ldots,f\}$.
 
-  \item \label{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$ \emph{or} $F[P_i]$ is a path in $F_{r'}$.
+  \item \proplabel{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$ \emph{or} $F[P_i]$ is a path or cycle in $F_{r'}$.
 
-  \item \label{partition} $\{Y_x : x\in V(K)\}$ is a partition of $V(G^+)\setminus\{r\}$.
-
-  \item Something else?
-\end{compactenum}
-
-Properties~(PR1)--(PR3) are defining properties of a PS-tree, and they imply the remaining properties.
-
-% There are at most three ancestors $a$, $b$, and $c$, of $x$ in $T$ such that the vertices of $F_x$ are contained in $V(Y_a)\cup V(Y_b)\cup V(Y_c)$.
-(See below for the definition of $Y_a$ $Y_b$, and $Y_c$.)
-% 
-%  consisting of the part of $\Delta$ contained in $F_x$.  Note that $Y_x\subseteq V(N_x)\setminus V(F_x)$ and $V(N_x)\subseteq V(N_a)$ for every ancestor $a$ of $x$.
-% 
-% {\color{red}Expand the preceding points more fully and point out the relationsships better.  Deal with degenernate cases.}
+  \item \proplabel{partition} $\{Y_x : x\in V(K)\}$ is a partition of $V(G^+)\setminus\{r\}\setminus V(F_{r'})$.
+\end{enumerate}
 
@@
 
-Let $G$ be a $k$-plane graph and let $G^+$ be the graph obtained by making $G$ planar with dummy vertices, then triangulating the resulting graph and then adding a degree-3 vertex $r$ in the outer face of the resulting graph.
-Let $T$ be a BFS spanning tree of $G^+$ rooted at $r$.
-For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{G^+}(r,v)=i\}$.
-Observe that, for any edge $vw\in E(G)$, $\dist_{G^+}(v,w) \le k$ and therefore, for any two vertices $v,w\in V(G)$,
+Let $G$ be a $k$-plane graph and let $G^+$ be the graph obtained as follows:
+\begin{compactenum}
+  \item Make $G$ planar by adding dummy vertices at each crossing.
+  \item Add a $K_4$ that contains $G$ in one of its inner faces $F_0$.
+  \item Add edges to complete the resulting graph to a triangulation.
+\end{compactenum}
+Let $r$ be the vertex of the $K_4$ not incident on $F_0$ and let $T$ be a BFS spanning tree of $G^+$ rooted at $r$.  For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{G^+}(r,v)=i\}$.
+
+Observe that, for any edge $vw\in E(G)$, $\dist_{G^+}(v,w) \le k+1$ and therefore, for any two vertices $v,w\in V(G)$,
 \begin{equation} 
-  \dist_{G^+}(v,w) \le k\cdot \dist_G(v,w) \eqlabel{distance-preserving}
+  \dist_{G^+}(v,w) \le (k+1)\cdot \dist_G(v,w) \eqlabel{distance-preserving}
 \end{equation}
@@
 We claim that $(S_x:x\in V(K))$ is a partition of $V(G)$.
-By (PR\ref{partition}) every vertex $v\in  V(G)$ is in $V(Y_x)\subseteq V(N_x)$.
+By \propref{partition} every vertex $v\in  V(G)$ is in $V(Y_x)\subseteq V(N_x)$ for some $x\in V(K)$.
 Therefore every vertex $v\in V(G)$ appears at some node $x\in V(K)$, so $\bigcup_{x\in V(K)} S_x \supseteq V(G)$.
@@
 Furthermore, if $v$ appears at two nodes $x$ and $y$ then, by definition, $v\in V(N_x)\cap V(N_y)$.
-If neither $x$ nor $y$ is an ancestor of the other then, by (PR\ref{children}), $V(N_x)\cap V(N_y)=V(F_x)\cap V(F_y)$.
-By (PR\ref{ancestor-boundary}), $v\in Y_a$ for some $K$-ancestor $a$ of $x$.
-Similarly, $v\in Y_{b}$ for some ancestor $b$ of $y$.
-By (PR\ref{partition}), $a=b$ is a common $K$-ancestor of $x$ and $y$.
-It follows that there is a unique node $a'$ that is an ancestor of all nodes of $K$ at which $v$ appears, so $v$ appears only in $Y_{a'}$.
+If neither $x$ nor $y$ is an ancestor of the other then, by \propref{children}, $V(N_x)\cap V(N_y)=V(F_x)\cap V(F_y)$.
+By \propref{ancestor-boundary}, $v\in V(Y_a)$ for some $K$-ancestor $a$ of $x$.
+Similarly, $v\in V(Y_{b})$ for some ancestor $b$ of $y$.
+By \propref{partition}, $a=b$ is a common $K$-ancestor of $x$ and $y$ and $v$ appears at $a$.
+It follows that there is a unique node $a'$ that is a $K$-ancestor of all nodes of $K$ at which $v$ appears, so $a'$ is the only node of $K$ for which $v\in S_{a'}$.
 
@@
 
-We use the partition $(S_x: x\in V(K))$ which generates a graph $H$ with vertex set $V(H)=V(K)$ the edge $xy\in E(H)$ if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
+We use the partition $(S_x: x\in V(K))$ which generates a graph $H$ with vertex set $V(H)=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
 First we argue that this partition has small width with respect to the layering defined by $T$ (note that this is a layering of $G^+$, not a layering of $G$).
@@
 \begin{clm}
-  For any $x\in V(H)$, and any integer $i\ge 0$, $|S_x\cap V_i|\le 3(k+1)$.
+  For any $x\in V(H)$, and any integer $i\ge 0$, $|S_x\cap V_i|\le 3(4k+3)$.
 \end{clm}
@@
 \begin{proof}
-  Done already, several times.
+  Since the vertices of $Y_x$ come from at most 3 vertical paths in $T$, $|V(Y_x)\cap V_i|\le 3$.
+Any vertex $v\in S_x\cap V_i$ is either in $V(Y_x)$ or is an endpoint of an edge $vw\in E(G)$ that contains a vertex $v'\in V(Y_x)$.
+In the latter case, $|\dist_T(r,v)-\dist_T(r, v')|\le k+1$.
+Therefore $v'\in V(Y_x)\cap V_{j}$ for some $j\in\{i-k-1,\ldots,i+k+1\}$.
+Any such $v'$ is responsible for at most 2 vertices in $S_x\setminus V(Y_x)$.
+Therefore,
+  \[  |S_x\cap V_i| \le |V(Y_x)\cap V_i| + 3\cdot 2\cdot (2k+1) \le 12k+9 \enspace . \qedhere \]
 \end{proof}
@@
 Since neither $x$ nor $y$ are ancestors of each other, the interiors of $F_x$ and $F_y$ are disjoint.
-If $v$ is in the interior $V(N_x)\setminus V(F_x)$ of $F_x$ then the edge $vw$ crosses $F_x$ and by (PR\ref{ancestor-boundary}) $v$ appears at some ancestor $a$ of $x$.
-(The node $a$ is the lowest node for which $v,w\in V(N_a)$.)  But this contradicts the assumption that $v\in S_x$, so it must be that $v\in V(F_x)$.
-This is also a contradiction, since (PR\ref{ancestor-boundary}) then implies that $v\in Y_a$ and therefore appears at $a$ for some $K$-ancestor $a$ of $x$, which contradicts the assumption that $v\in V(N_x)$.
+Therefore, the edge $vw$ contains a vertex of $F_x$ (as an endpoint or in its interior).
+This is also a contradiction, since \propref{ancestor-boundary} then implies that $v\in V(Y_a)$ for some $K$-ancestor $a$ of $x$ and therefore appears at $a$, so $v\not\in S_x$.
 
@@
 Additionally, for every edge $xz\in E(H)$, where $x$ is an ancestor of $z$, $K[x]$ includes the path from $x$ to $z$ in $K$.
-That this produces a tree-decomposition is due to the fact that for every edge $xz$ of $H$, $x$ is an ancestor of $z$ (or vice-versa).
+That this produces a tree-decomposition is due to the fact, established in the preceding paragraph, that for every edge $xz$ of $H$, $x$ is an ancestor of $z$ (or vice-versa).
 What remains is to upper bound the width of this decomposition.
@@
 As we have just argued, $B_x$ contains only ancestors of $x$ (including $x$ itself).
-If $B_x$ contains some ancestor $a$ of $x$, it is because there is an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_a$.
+If $B_x$ contains some $K$-ancestor $a$ of $x$, it is because there is an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_a$.
 Since $v\in S_x$, $v\in V(N_x)$.
-Since, $w\in S_a$, $w\in Y_a$ or there is an edge $wz\in E(G)$ that contains a point of $Y_a$.
-The embedding of the path $vwz$ is a curve that being at $v\in N_x$, ends at
-$z$, contains a point of $V(Y_a)$ and contains at most $2k+1$ vertices of $G^+$.
+Since, $w\in S_a$, either (a)~$w\in V(Y_a)$ or (b)~there is an edge $wz\in E(G)$ with $w,z\in V(N_a)$ that contains a vertex of $Y_a$.
+To avoid having to distinguish between the two cases (a) and (b), we will use $vz$ to denote the embedding of the edge $vw$, in case (a), or the embedding of the path $vwz$, in case~(b).
 
-Now, consider the cubic graph $A$ homeomorphic to the union of $F_{a'}$ for all ancestors $a'$ of $x$.
+Now, consider the graph $A$ that contains $F_{a'}$ for all ancestors $a'$ of $x$.
 See \figref{A}.
@@
   \begin{center}
-    \includegraphics{figs/A}
+      \includegraphics{figs/A-1}
   \end{center}
@@
 
-We claim that the curve $C$ begins in $F_x\subseteq F^x_a\subseteq F_a$, remains in the interior of $F_a$, and eventually reaches the boundary $F^{\bar{x}}_a\cap F^x_a$.
-The fact that $C$ remains in $F_a$ is due to the fact that, if it crossed the boundary of $F_a$, then one of $vw$ or $wz$ would intersect $Y_{a'}$ for some ancestor $a'$ of $a$.
-If $vw$ intersects $Y_{a'}$, then $v\not\in S_x$.
-If $wz$ intersects $Y_{a'}$ then $w,z\not\in S_a$.
-The fact that $C$ eventually reaches follows from the fact that $w\in Y_a$ or $wz$ contains a point of $Y_a$.
+We claim that the curve $vz$ contains a curve $vz'$ that (o)~does not cross any edge of $A$, (i)~begins in $F_x$, (ii)~remains in the interior of $F_a$, (iii)~terminates at $z'$ on the boundary $F^{\bar{x}}_a\cap F^x_a$, and (iv)~contains at most $2k+1$ vertices of $A$.\note{PM}{$2k+1$ should be $2k+3$, but this carries through and I don't have to update right now.}
+\begin{compactenum}[(i)]
+  \item The curve $vz'$ begins at $v\in V(N_x)$;
 
-Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of faces $F^{\bar{x}}_a$ of $A$ that can be reached with curves that begin in $F_x$, remain in the interior of $F_a$, and terminate in $F^{\bar{x}}_a$ without intersecting more than $2k$ edges in $A$.
-To avoid triple subscripts we say that a curve visits $a_i$ if it visits $F^{\bar{x}}_{a_i}$.
-We can further restrict ourselves to studying curves that cross the fewest number of edges of $A$ since if a curve exists that crosses at most $k$ edges then the curve that crosses the fewest number of possible edges also crosses at most $k$ edges.
-We will call such a curve a \emph{lightest} curve from $x$ to $a$.
+  \item The curve $vz'$ remains in the interior of $F_a$ since, if it intersected the boundary of $F_a$, then one of $vw$ or $wz$ would intersect $Y_{a'}$ for some ancestor $a'\neq a$ of $a$.
+If $vw$ contains a vertex of $Y_{a'}$, then $v\not\in S_x$.
+If $wz$ contains a vertex of $Y_{a'}$ then $w,z\not\in S_a$.
 
-Again, we let $a_0,\ldots,a_r$ denote the vertices of the path, in $K$ from the root $a_0$ to $x=a_r$.
-We claim that, there exists a lightest curve from $a_r$ to $a_0$ that is monotonic in the sense that the curve visits $a_{i_0},\ldots,a_{i_t}$ in this order, where $i_0=r > i_1>\cdots>i_t=0$.
-This follows from (PR\ref{ancestor-boundaries}):  For any node $a_i$, $i>0$, there are at most three faces of $A$ that share a part of their boundary with $F_{a_i}$.
-Each of these faces also necessarily shares a part of its boundary with $F^{\bar{x}}_{a_i}$.
-Therefore it is never advantageous for a path to go from $a_i$ to $a_j$ for some $j>i$ since the path will eventually have to return to one of these three faces.
+  \item That $vz'$ eventually reaches $F^{\bar{x}}_a$ follows from the fact that, in either case (a) or (b), $vz$ contains a vertex of $Y_a$.
+The first such vertex encountered walk walking along $vz$ is the endpoint $z'$.
 
-At this point we are essentially done.
-Any monotonic curve that visits $a_i$ has at most 3 choices for the next face $a_j$ $j<i$ to visit.
-Therefore, the number of faces of $A$ that can be reached from $F_x$ by admissible curves that cross $2k$ edges of $A$ is at most $3^{2k}$.
-Therefore,
-\[
-   |B_x| \le \sum_{j=0}^2k 9^j = (9^{2k+1}-1)/8 \enspace .
-\]
+  \item That $vz'$ contains at most $2k+1$ vertices of $A$ follows from the fact that $vz'$ is contained in a path of length at most 2 in $G$, and $G$ is $k$-planar.
+\end{compactenum}
+We call any curve with properties (i)--(iv) an \emph{admissible $xa$-curve} (for $A$).
+Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of faces $F^{\bar{x}}_a$ of $A$ such that there exists an admissible $xa$-curve.
+To avoid triple-subscripts, we will say that an $xa$-curve \emph{visits} $y\in N(K)$ if it contains a vertex of $F^{\bar{x}}_y$.
+It is furthermore sufficient to consider $xa$-curves that are \emph{minimal} in the sense that, if the curve visits $x=x_0,x_1,\ldots,x_\ell=a$ in this order, then $x_i\neq a$ for any $i\in\{0,\ldots,\ell-1\}$.
 
-This completes the proof
+By \propref{ancestor-boundary}, for every $y\in V(K)$, there are $t\le 3$ $K$-ancestors $y_1,\ldots,y_t$ of $y$ such that $F_y$ has a vertex in common with $F^{\bar{x}}_{y_i}$.
+For each $i\in\{1,\ldots,t\}$, colour the pair $(y,y_i)$ with the colour $c(y,y_i)=i$.
+In this way any pair that receives a colour receives one of 3 colours.
+The set $y_1,\ldots,y_t$ behaves like a cutset in the following sense:  If a minimal admissible $xa$-curve visits $y$ (so it contains a point in $F^{\bar{x}}_y$) for some $K$-ancestor $y\neq a$ of $x$, then the curve must subsequently visit one of $y_1,\ldots,y_t$.
+
+\begin{figure}
+  \begin{center}
+      \includegraphics{figs/A-2}
+  \end{center}
+  \caption{The colour sequences defined by two minimal admissible $xa$-curves.}
+  \figlabel{colouring}
+\end{figure}
+
+Consider some minimal admissible $xa$-curve that visits $x=x_0,x_1,\ldots,x_\ell=a$ in this order.
+Refer to \figref{colouring}.
+Let $z_0,\ldots,z_{\ell'}$ be the subsequence of $x_0,\ldots,x_\ell$ consisting only of those $x_i$ that are strict $K$-ancestors of each of $x_0,\ldots,x_{i-1}$.
+This yields a colour sequence $c_1,\ldots,c_{\ell'}$, $\ell'\le \ell$, where $c_i=c(z_{i-1},z_i)$ for each $i\in\{1,\ldots,\ell'\}$.
+There are at most $3^{\ell'}$ such colour sequences and, if an admissible $xa$-curve has the same colour sequence as a minimal admissible $xa'$-curve then $a=a'$.
+Therefore, for some fixed $x\in V(K)$, the $K$-ancestors $a$ of $x$ such that there exists a minimal admissible $xa$-curve is at most
+\begin{equation}
+   |B_x| \le \sum_{\ell=0}^{2k+1} 3^\ell = (3^{2k+2}-1)/2 
+       = (9^{k+1}-1)/2 \enspace .  \eqlabel{bag-size}
+\end{equation}
+
+This is all we need to complete the proof of:
 
 \begin{thm}
-    Every $k$-planar graph $G$ has a $(k9^{2k+1}/8,3k)$-partition.
+    Every $k$-planar graph $G$ has a layered $H$-partition of layered-width at most $(k+1)(12k + 9)$ in which $H$ has treewidth at most $(9^{k+1}-3)/2$
 \end{thm}
+
+\begin{proof}
+  By \eqref{bag-size}, the partition $(S_x:x\in V(K))$ of $G$ described above produces a graph $H$ of treewidth $(9^{k+1}-3)/2$.
+For each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 12k+9$.
+\Eqref{distance-preserving} implies that setting $V_i' = V_{(k+1)i}\cup\cdots\cup V_{(k+1)(i+1)-1}$ yields a layering $(V_0',V_1',\ldots)$ of  $G$ and $|B_x\cap V_i'|\le (k+1)(12k+9)$ for all integers $i\ge 0$ and all $x\in V(H)$.
+\end{proof}
 

### 2019-05-09 k-planar.tex
Intro writing

@@
 \usepackage{paralist}
+
+\usepackage[longnamesfirst,numbers,sort&compress]{natbib}
 
@@
 Such partitions are known to exist for planar graphs and, more generally, bounded genus graphs.
-In the current paper, we prove that every $k$-planar graph has a layered $H$-partition of layered width $O(k)$ in which $H$ has treewidth $k9^k$.
+In the current paper, we prove that every $k$-planar graph has a layered $H$-partition of layered width $O(k)$ in which $H$ has treewidth $O(c^k)$, for some constant $c$.
 This implies that $k$-planar graphs have both queue number and non-repetitive chromatic number upper-bounded by a function of $k$.
@@
 
-\section{$1$-Plane Graphs}
+A \emph{layering} of a graph $G$ is a partition $\mathcal{L}=\{V_0,V_1,\ldots,V_h\}$ of $V(G)$ so that there is no edge $vw\in E(G)$ such that $v\in V_i$, $w\in V_j$ and $|j-i|>1$.
+For any partition $\mathcal{P}=\{S_1,\ldots,S_p$ of $V(G)\}$, the \emph{quotient graph} $H=G/\mathcal{P}$ has $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
+% We call $\mathcal{P}$ an \emph{$H$-partition} of $G$ and we write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.
+A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
+The \emph{layered width} of $(\mathcal{H},\mathcal{P})$ is $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
 
-Let $G$ be an edge-maximal 1-plane multigraph graph.
-Here, edge-maximal should be taken to mean that, if any two vertices $u$ and $v$ appear on a common face $F$, then there is an edge $uv\in E(G)$ that is contained in the boundary of $F$.
+Layered $H$-partition of small layered width in which $H$ has some additional property are a useful tool for a number of graph problems.
+\citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number \cite{X}, track number \cite{X}, etc \cite{X}, etc \cite{X}, etc \cite{X}.
+\citet{dujmovic.esperet.ea:planar} add non-repetitive chromatic number to this list.
+\citet{dujmovic.joret.ea:planar} show that two common families of graphs have such $H$-partitions: planar graphs and (more generally) bounded-genus graphs.
+
+In this paper we show that another family of graphs admits layered $H$-partitions of constant layered-width in which $H$ has constant treewidth: $k$-planar graphs.
+An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve is a continuous function $f:[0,1]\to \R^2$.
+The points $f(0)$ and $f(1)$ are called the \emph{endpoints} of the curve.
+When there is no danger of misunderstanding we treat a curve $f$ as a point set $\{f(t):0\le t\le 1\}$.} with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
+A \emph{crossing} in an embedded graph $G$ is a triple $(p,vw,xy)$ with $p\in\R^2$, $vw,xy\in E(G)$ and such that $p\in (vw\cap xy)\setminus\{v,w,x,y\}$.
+An embedded graph $G$ is \emph{$k$-plane} if each edge of $G$ takes part in at most $k$ crossings.
+A (not necessarily embedded) graph $G'$ is \emph{$k$-planar} if there exists $k$-plane graph $G$ isomorphic to $G'$.
+We prove the following result:
+
+\begin{thm}\thmlabel{k-planar}
+  Every $k$-planar graph has a layered $H$-partition of layered-width at most $(k+1)(12k + 9)$ in which $H$ has treewidth at most $(9^{k+1}-3)/2$.
+\end{thm}
+
+In the special case $k=1$ we obtain better constants and additional property (planarity) of $H$:
+
+\begin{thm}\thmlabel{1-planar}
+  Every 1-planar graph has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
+\end{thm}
+
+These results imply that, for any constant $k$, $k$-planar graphs have all the graph parameters mentioned above upper-bounded by some constant $c_k$.
+
+In \secref{1-planar} we prove \thmref{1-planar}.  In \secref{k-planar} we prove \thmref{k-planar}.
+
+\section{$1$-Planar Graphs}
+\seclabel{1-planar}
+
+Let $G$ be an edge-maximal 1-plane multigraph.
+Here, edge-maximal should be taken to mean that, if any two vertices $u$ and $v$ appear on a common face\footnote{The \emph{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
+We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $uv\in E(G)$ that is contained in the boundary of $F$.
 
@@
 Observe that, for any spar $vw\in E(G)\setminus E(G')$ that crosses $xy\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
-It follows that $\dist_{G'}(x,y)\le 2\dist_G(x,y)$ for any pair of vertices $x,y\in V(G)$.
+It follows that $\dist_{G'}(x,y)\le 2\cdot\dist_G(x,y)$ for any pair of vertices $x,y\in V(G)$.
 
+Our proof of \thmref{1-planar} follows quickly from the following technical lemma, which is an extension of Lemma~X in \citet{dujmovic.joret.ea:planar}.
 \begin{lem}\lemlabel{induction} The setup:
@@
 
-  The remainder of the proof is as before.
-We use $V(Y^+)\setminus V(C)$ as an element of our $H$-partition.
-The root bag in the tree decomposition contains the vertices obtained by contracting each of $P_1,\ldots,P_k$ and $Y^+$.
+  The remainder of the proof finishes in the same way as the proof of Lemma~X in \cite{dujmovic.joret.ea:planar}.
+$P_1,\ldots,P_k$, and $V(Y^+)\setminus V(C)$ become elements of the $H$-partition.
+Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(Y^+)\setminus V(C)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
+The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~X in blah.
 
-  Planarity comes from the fact that if two edges $vw$ and $xy$ cross, then they end up in the same bag of the $H$-partition.
+  The planarity of $H$ comes from the fact that if two edges $vw$ and $xy$ cross, then they end up in the same bag of the $H$-partition.
 This means that $H$ is actually obtained by contracting connected sets of vertices in the planar graph $G'$.
+It is well know that any graph obtained by contracting connected sets of vertices in a planar graph is also planar.
 \end{proof}
 
-\begin{thm}\thmlabel{1-plane}
-  Every 1-plane graph $G$ has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
-\end{thm}
-
-\begin{proof}
+\begin{proof}[Proof of \thmref{1-plane}]
   The same as before except that the layered width 15 from \lemref{induction} becomes 30 because \lemref{induction} uses a layering of $G'$ and distances in $G'$ can be a factor of 2 larger than in $G$.
@@
 
-\begin{thm}
-    Every $k$-planar graph $G$ has a layered $H$-partition of layered-width at most $(k+1)(12k + 9)$ in which $H$ has treewidth at most $(9^{k+1}-3)/2$
-\end{thm}
-
 \begin{proof}
@@
 
-\bibliographystyle{plain}
-\bibliography{warmup}
+\bibliographystyle{plainnat}
+\bibliography{k-planar}
 

### 2019-05-09 k-planar.tex
This branch is going to be a failure

@@
 
-\subsection{The PS-Tree}
+\subsection{PS-Trees}
 
@@
 
+For any triangulation $\Delta$ in which the outer face has a vertex $r$ of degree 3 and for any BFS spanning-tree of $\Delta$ rooted at $r$, there exists a PS-tree $K=K(\Delta,T,r)$.
+Indeed, this tree is constructed (albeit implicitly) by \citet[Proof of Lemma~X]{dujmovic.joret.ea:planar}.
+
+The following property of a PS-Tree $K$ is an immediate consequence of \propref{children} that will be useful for us:
+\begin{compactenum}[(PR5)]
+  \item \proplabel{ancestors} If $v\in V(N_x-F_x)$ and $w\in V(N_y-N_y)$ for some nodes $x,y\in V(K)$, then one of $x$ or $y$ is a $K$-ancestor of the other.
+\end{compactenum}
+
 \subsection{$k$-Planar Graphs}
@@
 \end{compactenum}
-Let $r$ be the vertex of the $K_4$ not incident on $F_0$ and let $T$ be a BFS spanning tree of $G^+$ rooted at $r$.  For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{G^+}(r,v)=i\}$.
-
+Let $r$ be the vertex of the $K_4$ not incident on $F_0$ and let $T$ be a BFS spanning tree of $G^+$ rooted at $r$.
 Observe that, for any edge $vw\in E(G)$, $\dist_{G^+}(v,w) \le k+1$ and therefore, for any two vertices $v,w\in V(G)$,
@@
 \end{equation}
+For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{G^+}(r,v)=i\}$ and observe that $V_0,V_1,\ldots,V_{t}$ is a layering of $G^+$ (but not of $G$).
 
-Let $K=K(G^+,T,r)$ be a PS-tree.
-We say that a vertex $v\in V(G)$ \emph{appears} at a node $x\in V(K)$ if $v\in Y_x$ or if there exists $vw\in E(G)$ such that $v,w\in V(N_x)$ and $vw$ contains a vertex of $Y_x$, either as an endpoint or in its interior (see \figref{appearances}).
+Let $K=K(G^+,T,r)$ be a PS-tree and observe that the root $r'$ of $K$ has $F_{r'}=F_0$.  This implies:
+\begin{compactenum}[(PR6)]
+  \item \proplabel{all-in}  $V(G)\subseteq V(N_{r'}-F_{r'})$.
+\end{compactenum}
 
-\begin{figure}
-  \begin{center}
-    \includegraphics{figs/K-node-3}
-  \end{center}
-  \caption{Examples of vertices (in peach) that appear at a node $x\in V(K)$.}
-  \figlabel{appearances}
-\end{figure}
+For any pair of vertices $v,w\in V(G)$, consider the set $A_{vw}=\{a\in V(K): v,w\in V(N_a-F_a) \}$.
+By \propref{all-in}, $A_{vw}$ contains the root $r'$ of $K$.
+By \propref{children}, any node $a\in A_{vw}$ has at most one child in $A_{vw}$ and any node $a\not\in A_{vw}$ has no child in $A_{vw}$.
+It follows that $K[A_{vw}]$ is a vertical path in $K$ that contains the root.
+We define $a(vw)$ to be the lower endpoint of $K[A_{vw}]$.
 
-Let $S_x$ (the separator at $x$) denote the set of all vertices $v\in V(G)$ such that $v$ appears at $x$ but does not appear at any $K$-ancestor of $x$.
-We claim that $(S_x:x\in V(K))$ is a partition of $V(G)$.
-By \propref{partition} every vertex $v\in  V(G)$ is in $V(Y_x)\subseteq V(N_x)$ for some $x\in V(K)$.
-Therefore every vertex $v\in V(G)$ appears at some node $x\in V(K)$, so $\bigcup_{x\in V(K)} S_x \supseteq V(G)$.
+For any vertex $v\in V(G)$ define $A_v=\bigcap\{vw\in E(G)} A_{vw}$.  $K[A_v]$ is a vertical path in $K$ that contains the root.  We define $a(v)$ to be the lower endpoint of $K[A_v]$.
 
-Furthermore, if $v$ appears at two nodes $x$ and $y$ then, by definition, $v\in V(N_x)\cap V(N_y)$.
-If neither $x$ nor $y$ is an ancestor of the other then, by \propref{children}, $V(N_x)\cap V(N_y)=V(F_x)\cap V(F_y)$.
-By \propref{ancestor-boundary}, $v\in V(Y_a)$ for some $K$-ancestor $a$ of $x$.
-Similarly, $v\in V(Y_{b})$ for some ancestor $b$ of $y$.
-By \propref{partition}, $a=b$ is a common $K$-ancestor of $x$ and $y$ and $v$ appears at $a$.
-It follows that there is a unique node $a'$ that is a $K$-ancestor of all nodes of $K$ at which $v$ appears, so $a'$ is the only node of $K$ for which $v\in S_{a'}$.
+For every node $x\in V(K)$, we define the \emph{separator} of $X$ as
+\[
+   S_x = \{v\in V(G): a(v)=x \} \enspace .
+\]
+Clearly $\mathcal{S}=\{S_x:x\in V(K)\}$ is a partition of $V(G)$.
 
-% Indeed, if $v$ appears at nodes $x$ and $y$ where neither $x$ nor $y$ is an ancestor of the other, then there exists edges $vw,vz\in E(G)$ (possibly $w=z$) such that $vw$ contains a vertex of $Y_x$ (as an endpoint or in its interior) and $vz$ contains a vertex of $Y_y$ (as endpoint or in its interior).
-At least one of these two edges contains a vertex $Y_a$ for some common ancestor $a$ of $x$ and $y$.
-
-We use the partition $(S_x: x\in V(K))$ which generates a graph $H$ with vertex set $V(H)=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
-First we argue that this partition has small width with respect to the layering defined by $T$ (note that this is a layering of $G^+$, not a layering of $G$).
+Consider graph $H=G/\mathcal{S}$ that has vertex set $V(H)=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
+First we argue that this partition has small width with respect to the layering defined by $T$ (recall that this is a layering of $G^+$, not a layering of $G$).
 
@@
 We deliberately use the same node set $V(H)=V(K)$ here, because the subtree $K[x]:=K[\{z:x\in B_z\}]$ will be rooted at $x$.
-We claim that, for every edge $xy\in V(H)$, $x$ is an ancestor of $y$ (or vice-versa).
-Suppose that this is not the case and that $xy\in E(H)$ but neither $x$ nor $y$ is an ancestor of the other.
-Then, since $xy\in E(H)$, there must exist an edge $vw\in E(G)$ such that $v\in S_x$ and $w\in S_y$.
+We claim that, for every edge $xy\in V(H)$, one of $x$ or $y$ is a $K$-ancestor of the other.
+Indeed, since $xy\in E(H)$, there exists an edge $vw\in E(G)$ such that $v\in S_x$ and $w\in S_y$.
+By definition, $x\in A_v\subseteq A_{vw}$ and $y\in A_w\subseteq A_{vw}$.
+Since $K[A_{vw}]$ is a vertical path in $K$, it follows that one of $x$ or $y$ is a $K$-ancestor of the other.
 
-Since $v\in S_x$, $v\in V(N_x)$ and similarly $w\in V(N_y)$.
-Since neither $x$ nor $y$ are ancestors of each other, the interiors of $F_x$ and $F_y$ are disjoint.
-Therefore, the edge $vw$ contains a vertex of $F_x$ (as an endpoint or in its interior).
-This is also a contradiction, since \propref{ancestor-boundary} then implies that $v\in V(Y_a)$ for some $K$-ancestor $a$ of $x$ and therefore appears at $a$, so $v\not\in S_x$.
-
-To describe our tree decomposition we define, for each $x\in V(H)$, the subtree $K[x]=K[\{y:x\in B_y\}]$.
-The subtree $K[x]$ contains $x$.
-Additionally, for every edge $xz\in E(H)$, where $x$ is an ancestor of $z$, $K[x]$ includes the path from $x$ to $z$ in $K$.
-That this produces a tree-decomposition is due to the fact, established in the preceding paragraph, that for every edge $xz$ of $H$, $x$ is an ancestor of $z$ (or vice-versa).
+To describe our tree decomposition we define, for each $y\in V(H)$, the subtree $K[y]=K[\{x:y\in B_x\}]$.
+The subtree $K[y]$ contains $y$.
+Additionally, for every edge $xy\in E(H)$, where $y$ is an ancestor of $x$, $K[y]$ includes the path from $y$ to $x$ in $K$.
+That this produces a tree-decomposition is due to the fact, established in the preceding paragraph, that for every edge $xy$ of $H$, $x$ is a $K$-ancestor of $y$ (or vice-versa).
 What remains is to upper bound the width of this decomposition.
 
-First observe that, for every $a\in V(K)$, the subtree $K[a]$ is rooted at $a$, i.e., $a$ is a $K$-ancestor of every node in $x\in B_a$.
 Now, fix some node $x\in V(K)$ and consider the contents of $B_x$.
-As we have just argued, $B_x$ contains only ancestors of $x$ (including $x$ itself).
-If $B_x$ contains some $K$-ancestor $a$ of $x$, it is because there is an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_a$.
-Since $v\in S_x$, $v\in V(N_x)$.
-Since, $w\in S_a$, either (a)~$w\in V(Y_a)$ or (b)~there is an edge $wz\in E(G)$ with $w,z\in V(N_a)$ that contains a vertex of $Y_a$.
-To avoid having to distinguish between the two cases (a) and (b), we will use $vz$ to denote the embedding of the edge $vw$, in case (a), or the embedding of the path $vwz$, in case~(b).
+As we have just argued, $B_x$ contains only $K$-ancestors of $x$ (including $x$ itself).
+If $B_x$ contains some $K$-ancestor $y$ of $x$, it is because there is an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
+Since $v\in S_x$, $v\in V(N_x-F_x)$.
+Since, $w\in S_y$, $w\in V(N_y-F_y)$.
 
-Now, consider the graph $A$ that contains $F_{a'}$ for all ancestors $a'$ of $x$.
+Now, consider the graph $A$ that contains $F_{a}$ for all $K$-ancestors $a$ of $x$.
 See \figref{A}.
@@
 
-We claim that the curve $vz$ contains a curve $vz'$ that (o)~does not cross any edge of $A$, (i)~begins in $F_x$, (ii)~remains in the interior of $F_a$, (iii)~terminates at $z'$ on the boundary $F^{\bar{x}}_a\cap F^x_a$, and (iv)~contains at most $2k+1$ vertices of $A$.\note{PM}{$2k+1$ should be $2k+3$, but this carries through and I don't have to update right now.}
+The edge $vw$ corresponds to a walk $v=v_0,v_1,\ldots,v_{\ell}$ in $G^+$ that,
+(i)~begins at $v\in V(N_x-F_x)$, (ii)~contains only vertices in $V(N_y-F_y)$, (iii)~ends at $w\in V(N_y-F_y)$, (iv)~has length $\ell\le k+1$.
 \begin{compactenum}[(i)]
-  \item The curve $vz'$ begins at $v\in V(N_x)$;
+  \item $v\in V(N_x)$ since $x\in A_v$ and $v\in V(N_a-F_a)$ for all $a\in A_v$;
 
-  \item The curve $vz'$ remains in the interior of $F_a$ since, if it intersected the boundary of $F_a$, then one of $vw$ or $wz$ would intersect $Y_{a'}$ for some ancestor $a'\neq a$ of $a$.
+  \item $v_0,v_1,\ldots,v_{\ell}\in V(N_a-F_a)$ since, otherwise $v_i\in V(F_y)$.  By \propref{ancestor-boundary}, $v_i\in Y_{a}$ for some $K$-ancestor $a\neq y$ of $y$.
+
+  The curve $vz'$ remains in the interior of $F_a$ since, if it intersected the boundary of $F_a$, then one of $vw$ or $wz$ would intersect $Y_{a'}$ for some ancestor $a'\neq a$ of $a$.
 If $vw$ contains a vertex of $Y_{a'}$, then $v\not\in S_x$.
 If $wz$ contains a vertex of $Y_{a'}$ then $w,z\not\in S_a$.
+
+  \item $w\in V(N_y)$ since $y\in A_w$ and $w\in V(N_a-F_a)$ for all $a\in A_w$;
 

### 2019-05-09 k-planar.tex
Hmmmm

@@
 \begin{compactenum}
-  \item Make $G$ planar by adding dummy vertices at each crossing.
-  \item Add a $K_4$ that contains $G$ in one of its inner faces $F_0$.
-  \item Add edges to complete the resulting graph to a triangulation.
+  \item Make $G$ planar by adding dummy vertices at each crossing to obtain a graph $G_1$.
+  \item Add a $K_4$ that contains $G_1$ in one of its inner faces $F_0$ to obtaina graph $G_2$.
+  \item Add edges to complete $G_2$ to a triangulation $G^+$.
 \end{compactenum}
@@
 \begin{compactenum}[(PR6)]
-  \item \proplabel{all-in}  $V(G)\subseteq V(N_{r'}-F_{r'})$.
+  \item \proplabel{all-in}  $V(G_1)\subseteq V(N_{r'}-F_{r'})$.
 \end{compactenum}
 
-For any pair of vertices $v,w\in V(G)$, consider the set $A_{vw}=\{a\in V(K): v,w\in V(N_a-F_a) \}$.
-By \propref{all-in}, $A_{vw}$ contains the root $r'$ of $K$.
-By \propref{children}, any node $a\in A_{vw}$ has at most one child in $A_{vw}$ and any node $a\not\in A_{vw}$ has no child in $A_{vw}$.
-It follows that $K[A_{vw}]$ is a vertical path in $K$ that contains the root.
-We define $a(vw)$ to be the lower endpoint of $K[A_{vw}]$.
+For each vertex $v\in V(G_1)$, let $P_v=\{a\in V(K): v\in V(N_a-F_a)\}$.  
+By (PR6), $P_{v}$ contains the root $r'$ of $K$.
+By \propref{children}, any node $a\in P_{v}$ has at most one child in $P_{v}$ and any node $a\not\in P_{v}$ has no child in $P_{v}$.
+It follows that $K[P_{v}]$ is a vertical path in $K$ that contains the root.
 
-For any vertex $v\in V(G)$ define $A_v=\bigcap\{vw\in E(G)} A_{vw}$.  $K[A_v]$ is a vertical path in $K$ that contains the root.  We define $a(v)$ to be the lower endpoint of $K[A_v]$.
+For each edge $vw\in E(G)$, consider the walk $W_{vw}$ in $G^+$ that corresponds to $vw$ and define $A_{vw}=\bigcap_{z\in W_{vw}} P_z$.
+Since $W_{vw}\subseteq V(G_1)$, $A_{vw}$ is well-defined.
+For each $z\in W_{vw}$, $P_{z}$ contains $r'$ and is a vertical path in $K$.
+The common intersection of two or more vertical paths is a vertical path.
+Therefore $A_{vw}$ is a non-empty vertical path in $K$.
+For any vertex $v\in V(G)$ define $A_v=\bigcap_{vw\in E(G)} A_{vw}$.
+Again, $K[A_v]$ is a vertical path in $K$ that contains $r'$.
+We define $a(v)$ to be the lower endpoint of $K[A_v]$.
 
-For every node $x\in V(K)$, we define the \emph{separator} of $X$ as
+For every node $x\in V(K)$, define the \emph{separator}
 \[
@@
 
-Next we describe a tree decomposition $(B_x:x\in V(K))$ of $H$.
-We deliberately use the same node set $V(H)=V(K)$ here, because the subtree $K[x]:=K[\{z:x\in B_z\}]$ will be rooted at $x$.
-We claim that, for every edge $xy\in V(H)$, one of $x$ or $y$ is a $K$-ancestor of the other.
-Indeed, since $xy\in E(H)$, there exists an edge $vw\in E(G)$ such that $v\in S_x$ and $w\in S_y$.
-By definition, $x\in A_v\subseteq A_{vw}$ and $y\in A_w\subseteq A_{vw}$.
+Next we describe a tree decomposition $(B_x:x\in V(K))$ of $H$.  We deliberately use the same node set $V(H)=V(K)$ here, because the subtree $K[x]:=K[\{z:x\in B_z\}]$ will be rooted at $x$.
+
+\begin{clm}\clmlabel{xy-ancestor}
+   For every edge $xy\in V(H)$, one of $x$ or $y$ is a $K$-ancestor of the other.
+ \end{clm}
+
+ \begin{proof}
+   Since $xy\in E(H)$, there exists an edge $vw\in E(G)$ such that $v\in S_x$ and $w\in S_y$.
+By definition, $x\in P_v\subseteq A_{vw}$ and $y\in P_w\subseteq A_{vw}$.
 Since $K[A_{vw}]$ is a vertical path in $K$, it follows that one of $x$ or $y$ is a $K$-ancestor of the other.
+\end{proof}
 
-To describe our tree decomposition we define, for each $y\in V(H)$, the subtree $K[y]=K[\{x:y\in B_x\}]$.
-The subtree $K[y]$ contains $y$.
-Additionally, for every edge $xy\in E(H)$, where $y$ is an ancestor of $x$, $K[y]$ includes the path from $y$ to $x$ in $K$.
-That this produces a tree-decomposition is due to the fact, established in the preceding paragraph, that for every edge $xy$ of $H$, $x$ is a $K$-ancestor of $y$ (or vice-versa).
+The tree decomposition $(B_x:x\in V(K))$ of $H$ is defined as follows: For each edge $xy$ where $y$ is a $K$-ancestor of $x$, we add $y$ to be $B_{x'}$ for every node $x'$ on the path from $x$ to $y$ in $K$.
+That this produces a tree-decomposition is due to \clmref{xy-ancestor}.
 What remains is to upper bound the width of this decomposition.
 
-Now, fix some node $x\in V(K)$ and consider the contents of $B_x$.
-As we have just argued, $B_x$ contains only $K$-ancestors of $x$ (including $x$ itself).
+\begin{clm}
+  The tree decomposition $\{B_x: x\in V(K)\}$ of $H$ contains no bag larger than BLAH.
+\end{clm}
+
+\begin{proof}
+  Fix some node $x\in V(K)$ and consider the contents of $B_x$.
+By \clmref{xy-ancestor}, $B_x$ contains only $K$-ancestors of $x$ (including $x$ itself).
 If $B_x$ contains some $K$-ancestor $y$ of $x$, it is because there is an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
-Since $v\in S_x$, $v\in V(N_x-F_x)$.
-Since, $w\in S_y$, $w\in V(N_y-F_y)$.
 
-Now, consider the graph $A$ that contains $F_{a}$ for all $K$-ancestors $a$ of $x$.
+  Consider the graph $A$ that contains $F_{a}$ for all $K$-ancestors $a$ of $x$.
 See \figref{A}.
@@
 
-The edge $vw$ corresponds to a walk $v=v_0,v_1,\ldots,v_{\ell}$ in $G^+$ that,
-(i)~begins at $v\in V(N_x-F_x)$, (ii)~contains only vertices in $V(N_y-F_y)$, (iii)~ends at $w\in V(N_y-F_y)$, (iv)~has length $\ell\le k+1$.
+  The edge $vw$ corresponds to a walk $W_{vw}=v_0,v_1,\ldots,v_{\ell}$ in $G^+$ that
+  (i)~begins at $v_0=v\in V(N_x-F_x)$, (ii)~contains only vertices in $V(N_y-F_y)$, (iii)~ends at $w\in V(N_y)$, \note{PM:notation!} (iv)~has length $\ell\le k+1$.
   \begin{compactenum}[(i)]
-  \item $v\in V(N_x)$ since $x\in A_v$ and $v\in V(N_a-F_a)$ for all $a\in A_v$;
+    \item $v\in V(N_x-F_x)$ since $x\in P_v$ and $v\in V(N_a-F_a)$ for all $a\in P_v$;
 
-  \item $v_0,v_1,\ldots,v_{\ell}\in V(N_a-F_a)$ since, otherwise $v_i\in V(F_y)$.  By \propref{ancestor-boundary}, $v_i\in Y_{a}$ for some $K$-ancestor $a\neq y$ of $y$.
+    \item $v_0,v_1,\ldots,v_{\ell}\in V(N_y-F_y)$ since, otherwise there exists some $i\in\{1,\ldots,v_k\}$ such that $v_i\not\in V(N_y-F_y)$.
+But this implies that $y\not\in P_{v_i}$ which implies that $y\not\in A_{vw}$ which implies that $y\not\in A_w$, a contradiction.
 
-  The curve $vz'$ remains in the interior of $F_a$ since, if it intersected the boundary of $F_a$, then one of $vw$ or $wz$ would intersect $Y_{a'}$ for some ancestor $a'\neq a$ of $a$.
-If $vw$ contains a vertex of $Y_{a'}$, then $v\not\in S_x$.
-If $wz$ contains a vertex of $Y_{a'}$ then $w,z\not\in S_a$.
+    \item That $V_{wv}$ contains a vertex $w'\in F^\bar{x}_{y}$ is due to the fact that $y\in P_w$
 
-  \item $w\in V(N_y)$ since $y\in A_w$ and $w\in V(N_a-F_a)$ for all $a\in A_w$;
+    and $w\in V(N_y)$.
 
-  \item That $vz'$ eventually reaches $F^{\bar{x}}_a$ follows from the fact that, in either case (a) or (b), $vz$ contains a vertex of $Y_a$.
+    That $vz'$ eventually reaches $F^{\bar{x}}_a$ follows from the fact that, in either case (a) or (b), $vz$ contains a vertex of $Y_a$.
 The first such vertex encountered walk walking along $vz$ is the endpoint $z'$.
 
-  \item That $vz'$ contains at most $2k+1$ vertices of $A$ follows from the fact that $vz'$ is contained in a path of length at most 2 in $G$, and $G$ is $k$-planar.
+    \item That $\ell\le k+1$ follows from the fact that $G$ is $k$-planar and that each of $v_1,\ldots,v_{\ell-1}$ corresponds to a crossing involving $vw$.  Therefore $\ell-1\le k$.
   \end{compactenum}
@@
   \end{equation}
-
+\end{proof}
 This is all we need to complete the proof of:

### 2019-05-09 k-planar.tex
Switching from curves to walks

@@
    Since $xy\in E(H)$, there exists an edge $vw\in E(G)$ such that $v\in S_x$ and $w\in S_y$.
-By definition, $x\in P_v\subseteq A_{vw}$ and $y\in P_w\subseteq A_{vw}$.
+By definition, $x\in A_v\subseteq A_{vw}$ and $y\in A_w\subseteq A_{vw}$.
 Since $K[A_{vw}]$ is a vertical path in $K$, it follows that one of $x$ or $y$ is a $K$-ancestor of the other.
@@
 \begin{proof}
-  Fix some node $x\in V(K)$ and consider the contents of $B_x$.
-By \clmref{xy-ancestor}, $B_x$ contains only $K$-ancestors of $x$ (including $x$ itself).
-If $B_x$ contains some $K$-ancestor $y$ of $x$, it is because there is an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
+  Fix some node $x\in V(K)$ and consider the contents of $B_x$.  By \clmref{xy-ancestor}, $B_x$ contains only $K$-ancestors of $x$ (including $x$ itself).
 
@@
 
-  The edge $vw$ corresponds to a walk $W_{vw}=v_0,v_1,\ldots,v_{\ell}$ in $G^+$ that
-  (i)~begins at $v_0=v\in V(N_x-F_x)$, (ii)~contains only vertices in $V(N_y-F_y)$, (iii)~ends at $w\in V(N_y)$, \note{PM:notation!} (iv)~has length $\ell\le k+1$.
+  If $B_x$ contains some $K$-ancestor $a_i$, $i\in\{0,\ldots,r-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_{a_i}$.
+The edge $vw$ corresponds to a walk $W_{vw}$ in $G^+$.
+The walk $W_{vw}$ is contained in $N_x-F_x$, otherwise $x\not\in A_{vw}$ so $x\not\in A_v$.
+
+  However, $w\in S_a$, so there exists an edge $ww'\in E(G)$ such that 
+  the corresponding walk $W_{ww'}$ is contained in $N_{a_i}-F_{a_i}$ but is not
+  contained in $N_{a_{i+1}}-F_{a_{i+1}}$.  Therefore $W_{ww'}$ must contain a first vertex $w^*\in V(F_{a_{i+1}})\setminus V(F_{a_{i}})$.
+
+  Summarizing, we have a walk $w'=v_0,\ldots,v_\ell=w^*$ in $G^+$ such that
   \begin{compactenum}[(i)]
-    \item $v\in V(N_x-F_x)$ since $x\in P_v$ and $v\in V(N_a-F_a)$ for all $a\in P_v$;
+    \item $v_0\in V(N_x-F_x)$;
 
-    \item $v_0,v_1,\ldots,v_{\ell}\in V(N_y-F_y)$ since, otherwise there exists some $i\in\{1,\ldots,v_k\}$ such that $v_i\not\in V(N_y-F_y)$.
-But this implies that $y\not\in P_{v_i}$ which implies that $y\not\in A_{vw}$ which implies that $y\not\in A_w$, a contradiction.
+    \item $v_0,\ldots,v_{\ell-1}\in V(N_{a_{i+1}}-F_{a_{i+1}})$;
 
-    \item That $V_{wv}$ contains a vertex $w'\in F^\bar{x}_{y}$ is due to the fact that $y\in P_w$
+    \item $v_\ell\in V(F_{F_{a_{i+1}}})-V(F_{a_i})$; and
 
-    and $w\in V(N_y)$.
+    \item $\ell\le k+1$.
+  \end{compactenum}
+  We call any walk with properties (i)--(iv) an \emph{$i$-walk}. Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of positive integers $i$ such that there exists an $i$-walk.
 
-    That $vz'$ eventually reaches $F^{\bar{x}}_a$ follows from the fact that, in either case (a) or (b), $vz$ contains a vertex of $Y_a$.
-The first such vertex encountered walk walking along $vz$ is the endpoint $z'$.
+  Define $L_r=V(N_{a_r}-F_{a_r})$ and, for each $j\in\{0,\ldots,r-1\}$, define
+  $L_j = V(N_{a_j}- F_{a_j})\setminus L_{a_{j+1}}$ and observe that $L_i,\ldots,L_r$ is a partition of $V(N_{a_i}-F_{a_i})$.
+Let $v_0,\ldots,v_\ell$ be an $i$-walk and, for each $j\in\{0,\ldots,\ell\}$, let $x_j$ be the unique node $a_f\in\{a_i,\ldots,a_r\}$ such that $v_j\in L_{f}$.
+  Finally, let $x_0',\ldots,x_{\ell'}'$ be the subsequence of $x_0,\ldots,x_{\ell}$ consisting of $x_0$ and only those nodes $x_j$ such that $V(N_{x_j})\supsetneq \bigcup_{q=0}^{j-1} V(N_{x_q})$.
 
-    \item That $\ell\le k+1$ follows from the fact that $G$ is $k$-planar and that each of $v_1,\ldots,v_{\ell-1}$ corresponds to a crossing involving $vw$.  Therefore $\ell-1\le k$.
-  \end{compactenum}
-  We call any curve with properties (i)--(iv) an \emph{admissible $xa$-curve} (for $A$).
-Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of faces $F^{\bar{x}}_a$ of $A$ such that there exists an admissible $xa$-curve.
-To avoid triple-subscripts, we will say that an $xa$-curve \emph{visits} $y\in N(K)$ if it contains a vertex of $F^{\bar{x}}_y$.
-It is furthermore sufficient to consider $xa$-curves that are \emph{minimal} in the sense that, if the curve visits $x=x_0,x_1,\ldots,x_\ell=a$ in this order, then $x_i\neq a$ for any $i\in\{0,\ldots,\ell-1\}$.
-
-  By \propref{ancestor-boundary}, for every $y\in V(K)$, there are $t\le 3$ $K$-ancestors $y_1,\ldots,y_t$ of $y$ such that $F_y$ has a vertex in common with $F^{\bar{x}}_{y_i}$.
-For each $i\in\{1,\ldots,t\}$, colour the pair $(y,y_i)$ with the colour $c(y,y_i)=i$.
+  By \propref{ancestor-boundary}, for every $y\in V(K)$, there are $t\le 3$ $K$-ancestors $y_1,\ldots,y_t$ of $y$ such that $F_y$ has a vertex in common with $F^{\bar{x}}_{y}$.
+For each $j\in\{1,\ldots,t\}$, colour the pair $(y,y_j)$ with the colour $c(y,y_j)=j$.
 In this way any pair that receives a colour receives one of 3 colours.
-The set $y_1,\ldots,y_t$ behaves like a cutset in the following sense:  If a minimal admissible $xa$-curve visits $y$ (so it contains a point in $F^{\bar{x}}_y$) for some $K$-ancestor $y\neq a$ of $x$, then the curve must subsequently visit one of $y_1,\ldots,y_t$.
+The set $y_1,\ldots,y_t$ behaves like a cutset in the following sense: If $x_j'=y$ for $j<\ell'$, then $x_{j+1}'\in\{y_1,\ldots,y_t\}$.
 
@@
 
-  Consider some minimal admissible $xa$-curve that visits $x=x_0,x_1,\ldots,x_\ell=a$ in this order.
-Refer to \figref{colouring}.
-Let $z_0,\ldots,z_{\ell'}$ be the subsequence of $x_0,\ldots,x_\ell$ consisting only of those $x_i$ that are strict $K$-ancestors of each of $x_0,\ldots,x_{i-1}$.
-This yields a colour sequence $c_1,\ldots,c_{\ell'}$, $\ell'\le \ell$, where $c_i=c(z_{i-1},z_i)$ for each $i\in\{1,\ldots,\ell'\}$.
-There are at most $3^{\ell'}$ such colour sequences and, if an admissible $xa$-curve has the same colour sequence as a minimal admissible $xa'$-curve then $a=a'$.
-Therefore, for some fixed $x\in V(K)$, the $K$-ancestors $a$ of $x$ such that there exists a minimal admissible $xa$-curve is at most
+  Therefore $x_0',\ldots,x_{\ell'}'$ defines a color sequence
+  $c_1,\ldots,c_{\ell'}$, $\ell'\le \ell$, where $c_j=c(x_{j-1}',x_j')$ for each $j\in\{1,\ldots,\ell'\}$.
+There are at most $3^{\ell'}$ such colour sequences and, if an $i$-walk has the same colour sequence as an $i'$-walk,
+  then $i=i'$.  Therefore, the number of positive integers $i$ such that there exists and $i$-walk is at most
   \begin{equation}
@@
 \end{proof}
-This is all we need to complete the proof of:
 

### 2019-05-10 k-planar.tex
Almost done, but out of time for today.

@@
 
-\usepackage{tipa}
-\usepackage{upgreek}
-\usepackage{rotating}
-\newcommand{\V}{\rotatebox[origin=c]{180}{$V$}}
-\newcommand{\Y}{\rotatebox[origin=c]{180}{$Y$}}
+% \usepackage{tipa}
+% \usepackage{upgreek}
+% \usepackage{rotating}
+% \newcommand{\V}{\rotatebox[origin=c]{180}{$V$}}
+% \newcommand{\Y}{\rotatebox[origin=c]{180}{$Y$}}
 %\usepackage{hyperref}
@@
 In the current paper, we prove that every $k$-planar graph has a layered $H$-partition of layered width $O(k)$ in which $H$ has treewidth $O(c^k)$, for some constant $c$.
-This implies that $k$-planar graphs have both queue number and non-repetitive chromatic number upper-bounded by a function of $k$.
+This is the first result of this type of a non-minor closed class of graphs and implies that $k$-planar graphs have both queue number and non-repetitive chromatic number upper-bounded by a function of $k$.
+The former result was previously shown using a combination of $H$-partitions for planar graphs with \textit{ad hoc} methods.
+The latter result is new, for all $k\ge 3$.
 \end{abstract}
@@
 A \emph{layering} of a graph $G$ is a partition $\mathcal{L}=\{V_0,V_1,\ldots,V_h\}$ of $V(G)$ so that there is no edge $vw\in E(G)$ such that $v\in V_i$, $w\in V_j$ and $|j-i|>1$.
-For any partition $\mathcal{P}=\{S_1,\ldots,S_p$ of $V(G)\}$, the \emph{quotient graph} $H=G/\mathcal{P}$ has $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
-% We call $\mathcal{P}$ an \emph{$H$-partition} of $G$ and we write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.
-A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
+For any partition $\mathcal{P}=\{S_1,\ldots,S_p$ of $V(G)\}$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
+To highlight the importance of the quotient graph $H$, we can $\mathcal{P}$ an $H$-partition and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.\footnote{An alternative definition of a layering of $G$ is a layered $H$-partition of $G$ where $H$ is a path.}  A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
 The \emph{layered width} of $(\mathcal{H},\mathcal{P})$ is $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
@@
 Layered $H$-partition of small layered width in which $H$ has some additional property are a useful tool for a number of graph problems.
-\citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number \cite{X}, track number \cite{X}, etc \cite{X}, etc \cite{X}, etc \cite{X}.
+\citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number \cite{X}, track number \cite{X}, etc \cite{X}, etc \cite{X}, etc \cite{X}.
 \citet{dujmovic.esperet.ea:planar} add non-repetitive chromatic number to this list.
@@
 
-In this paper we show that another family of graphs admits layered $H$-partitions of constant layered-width in which $H$ has constant treewidth: $k$-planar graphs.
+In this paper we show that another family of graphs admits layered $H$-partitions of constant layered-width in which $H$ has constant treewidth: $k$-planar graphs, which we now define.
 An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve is a continuous function $f:[0,1]\to \R^2$.
@@
 A (not necessarily embedded) graph $G'$ is \emph{$k$-planar} if there exists $k$-plane graph $G$ isomorphic to $G'$.
+
+Under these definitions $0$-planar graphs are exactly planar graphs and $O$-plane graphs are plane graphs.
+Thus, like genus-$g$ graphs, $k$-planar graphs are a generalization of planar graphs (which are genus-$0$ graphs).
+Unlike genus-$g$ graphs, however, the family of $k$-planar graphs is not minor closed.
+A graph $G'$ obtained from a $k$-planar graph $G$ by edge deletions and edge contractions may or may not be $k$-planar.
+
 We prove the following result:
-
 \begin{thm}\thmlabel{k-planar}
-  Every $k$-planar graph has a layered $H$-partition of layered-width at most $(k+1)(12k + 9)$ in which $H$ has treewidth at most $(9^{k+1}-3)/2$.
+  Every $k$-planar graph has a layered $H$-partition of layered-width at most $24k^2 + 36k + 12$ in which $H$ has treewidth at most $(3^{k+2}-3)/2$.
 \end{thm}
 
-In the special case $k=1$ we obtain better constants and additional property (planarity) of $H$:
+In the special case $k=1$ we obtain better constants and an additional property (planarity) of $H$:
 
@@
 
-In \secref{1-planar} we prove \thmref{1-planar}.  In \secref{k-planar} we prove \thmref{k-planar}.
+The remainder of the paper is organized as follows: In \secref{1-planar} we prove \thmref{1-planar}.
+In \secref{k-planar} we prove \thmref{k-planar}.
+In \secref{consequences}, we discuss the implications of Theorems~\ref{thm:k-planar} and \ref{thm:1-planar} for other graph parameters of $k$-planar graphs.
+
+\section{$k$-Planar Graphs}
+\seclabel{k-planar}
+
+The purpose of this section is to prove \thmref{k-planar}.
+In this section, all graphs are finite, simple, an undirected.
+For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$) we use $G[S]$ to denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.
+We use $G-S$ as a shorthand for $G[V(G)\setminus S]$ and for a graph $G'$, we use $G-G'$ as a shorthand for $G-V(G')$.
+
+We begin by introducing a structural tool for planar graphs, called a PS-tree, that is implicit in previous work on $H$-partitions of planar graphs \cite{dujmovic.joret.ea:planar}.
+We then show how the PS-tree of a planarized version of a $k$-planar graph $G$ can be used to find a layered $H$-partition of $G$ of small layered width in which $H$ has small treewidth.
+
+\subsection{PS-Trees}
+
+Let $T$ be a rooted tree rooted at $r\in V(T)$.
+The \emph{depth} of a node $x\in V(T)$ is equal to the length of the path from $x$ to $r$ in $T$.
+A $P$ path in $T$ is \emph{vertical} if, for each integer $i$, $P$ contains at most one vertex of depth $i$.
+The deepest vertex in a vertical path $P$ is called  $P$'s \emph{lower endpoint} and the other (shallowest) vertex in $P$ is called $P$'s \emph{upper endpoint}.
+For any node $v\in T$, we call each node on the path, in $T$, from $v$ to $r$ a \emph{$T$-ancestor} of $v$.
+
+A \emph{near-triangulation} $N$ is a plane graph whose outer face, $F$, is bounded by a simple cycle in $N$ and each of whose inner faces is bounded by a 3-cycle in $N$.
+A \emph{triangulation} is a near-triangulation whose outer face is bounded by a 3-cycle.
+For any cycle $F$ in a near-triangulation $N$ we use $N\circledast F$ to denote the near-triangulation consisting of all the edges and vertices of $G$ that are contained in $F$ or its interior.
+
+For the remainder of this section, $\Delta$ is a triangulation, and $T$ is a BFS spanning tree of $\Delta$ rooted at some degree-3 vertex $r$ on the outer face of $\Delta$.
+
+Fix some cycle $F$ in $\Delta$ and let $N=\Delta\circledast F$.
+Let $\tau=v_1v_2v_3$ be an inner face of $N$ and consider the three minimal vertical paths $Q_1$, $Q_2$, and $Q_3$ in $T$ such that, for each $i\in\{1,2,3\}$ the lower endpoint of $Q_i$ is $v_i$ and the upper endpoint of $Q_i$ is in $V(F)$.
+If $Q_1$, $Q_2$, and $Q_3$ are vertex disjoint, then the graph $\overline{Y}=\tau\cup Q_1\cup Q_2\cup Q_3$ is called the \emph{closed tripod} in $N$ determined by $\tau$.
+We call $Q_1$, $Q_2$, and $Q_3$ the \emph{legs} of $\overline{Y}$ and we call $\tau$ the \emph{center} of $\overline{Y}$.
+We call $Y=\overline{Y}-F$ the \emph{(open) tripod} of $N$ determined by $\tau$.
+The graph $M=\overline{Y}\cup F$ is called a \emph{Mercedes graph} and the (at most 3) inner faces of $M$ other than $\tau$ are called the \emph{wedges} of $M$.
+
+A \emph{PS-tree}  $K=K(\Delta, T, r)$ is a 3-ary tree, such that each node $x\in V(K)$ is associated with several object (refer to \figref{K-node}):
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{cc}
+      \includegraphics{figs/K-node-1} &
+      \includegraphics{figs/K-node-2}
+    \end{tabular}
+  \end{center}
+  \caption{The elements associated with a node $x\in V(K)$.}
+  \figlabel{K-node}
+\end{figure}
+
+\begin{enumerate}
+  \item A cycle $F_x$ in $\Delta$ and the near-triangulation $N_x=\Delta\circledast F_x$.
+
+  \item An inner face $\tau_x$ of $N_x$ that determines a closed tripod $\overline{Y}_x$ and an open tripod $Y_x$ in $N_x$.
+
+  \item The \emph{Mercedes graph} $M_x=F_x\cup \overline{Y}_x$
+\end{enumerate}
+
+The nodes of the PS-tree $K$ satisfy several conditions:
+\begin{enumerate}[(PR1)]
+  \item For the root $r_K$ of $K$, $F_{r_K}$ is the 3-cycle in $\Delta$ formed by the three vertices adjacent to $r$.
+
+  \item \proplabel{children} For each node $x\in V(K)$ such that $M_x$ has $f\le 3$ wedges $F_{x,1},\ldots,F_{x,f}$, the node $x$ has exactly $f$ children $y_1,\ldots,y_f$ in $K$ and $F_{y_i} = F_{x,i}$ for each $i\in\{1,\ldots,f\}$.
+
+  \item \proplabel{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$ \emph{or} $F[P_i]$ is a path or cycle in $F_{r_K}$.
+
+  \item \proplabel{partition} $\{Y_x : x\in V(K)\}$ is a partition of $V(\Delta)\setminus\{r\}\setminus V(F_{r_K})$.
+\end{enumerate}
+
+For any triangulation $\Delta$ in which the outer face has a vertex $r$ of degree 3 and for any BFS spanning-tree of $\Delta$ rooted at $r$, there exists a PS-tree $K=K(\Delta,T,r)$.
+Indeed, this tree is constructed (albeit implicitly) by \citet[Proof of Lemma~14]{dujmovic.joret.ea:planar}.
+
+The following property of a PS-Tree $K$ is an immediate consequence of \propref{children} that will be useful for us: \note{PM}{Do we still use this?}
+\begin{compactenum}[(PR5)]
+  \item \proplabel{ancestors} If $vw\in E(\Delta)$, $v\in V(N_x-F_x)$ and $w\in V(N_y-N_y)$ for some nodes $x,y\in V(K)$, then one of $x$ or $y$ is a $K$-ancestor of the other.
+\end{compactenum}
+
+\subsection{$k$-Planar Graphs}
+
+Let $G$ be a $k$-plane graph and let $G^+$ be the graph obtained as follows:
+\begin{compactenum}
+  \item Make $G$ planar by adding dummy vertices at each crossing to obtain a plane graph and then add edges to this graph to obtain a triangulation $G^+$.
+  \item Add a complete graph $K_4$ that contains $G^{+}$ in one of its inner faces $F_0$ to obtain graph $G^{++}$.
+  \item Add edges to complete $G^{++}$ to a triangulation $G^{+++}$.
+\end{compactenum}
+Let $r$ be the vertex of the $K_4$ not incident on $F_0$ and let $T$ be a BFS spanning tree of $G^+$ rooted at $r$.
+Observe that, for any edge $vw\in E(G)$, $\dist_{G^+}(v,w) \le k+1$ and therefore, for any two vertices $v,w\in V(G)$,
+\begin{equation} 
+  \dist_{G^+}(v,w) \le (k+1)\cdot \dist_G(v,w) \eqlabel{distance-preserving}
+\end{equation}
+For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{T}(r,v)=i\}$ and observe that $\mathcal{L}=V_0,V_1,\ldots,V_{h}$ is a layering of $G^+$ (but not of $G$).
+
+Let $K=K(G^{+++},T,r)$ be a PS-tree and observe that the root $r_K$ of $K$ has $F_{r_K}=F_0$.  This implies:
+\begin{compactenum}[(PR6)]
+  \item \proplabel{all-in}  $V(G^+)\subseteq V(N_{r_K}-F_{r_K})$.
+\end{compactenum}
+
+For each vertex $v\in V(G^+)$, let $P_v=\{a\in V(K): v\in V(N_a-F_a)\}$.  
+By (PR6), $P_{v}$ contains the root $r_K$ of $K$.
+By \propref{children}, $K[P_v]$ is a vertical path (whose lower endpoint is the unique node $x\in V(K)$ such that $v\in Y_x$).
+Therefore $K[P_v]$ is a vertical path in $K$ that contains $r_K$.
+We are about to repeatedly use the fact that the common intersection of two or more vertical paths is a vertical path.
+
+For each edge $vw\in E(G)$, consider the walk $W_{vw}$ in $G^+$ that corresponds to the sequence of vertices in $G^+$ that we encounter while walking along the curve $vw$ from $v$ to $w$.
+We define $A_{vw}=\bigcap_{z\in W_{vw}} P_z$.
+Since $W_{vw}\subseteq V(G^+)$, $A_{vw}$ is well-defined.
+For each $z\in W_{vw}$, $K[P_{z}]$ is a vertical path in $K$ that contains $r_K$.
+Therefore $A_{vw}$ is a non-empty vertical path in $K$.
+For any vertex $v\in V(G)$ define $A_v=\bigcap_{vw\in E(G)} A_{vw}$.
+Again, $K[A_v]$ is a non-empty vertical path in $K$ that contains $r_K$.
+We define $a(v)$ to be the lower endpoint of $K[A_v]$.
+For every node $x\in V(K)$, define the \emph{separator}
+\[
+   S_x = \{v\in V(G): a(v)=x \} \enspace .
+\]
+Clearly $\mathcal{S}=\{S_x:x\in V(K)\}$ is a partition of $V(G)$.
+We will use $\mathcal{S}$ as the partition that defines our quotient graph $H=G/\mathcal{S}$ with vertex set $V(H)=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
+First we argue that this partition has small width with respect to the layering defined by $T$ (recall that this is a layering of $G^+$, not a layering of $G$).
+
+\begin{clm}\clmlabel{chargeback-to-y}
+   For any $x\in V(K)$ and any vertex $v\in S_x$, there exists a vertex $v'\in Y_x$ such that $|\dist_T(r,v')-\dist_T(r,v)|\le k+1$.  
+\end{clm}
+
+\begin{proof}
+  Since $v\in S_x$, the lower endpoint of $K[A_{vw}]$ is $x$ for some edge $vw\in E(G)$.
+This implies that $V(W_{vw})\subset V(N_x-F_x)$ but $V(W_{vw}) \not\subseteq V(N_{x'}-F_{x'})$ for any any child $x'$ of $x$ in $K$.
+This implies that $W_{vw}$ contains a vertex $v'\in Y_x$.
+Since $G$ is $k$-planar and $v'\in V_{vw}$ this implies that $\dist_{G^+}(v,v') \le k+1$.
+  By the triangle inequality,     
+  \[
+    \dist_{G^+}(r,v') \le \dist_{G^+}(r,v) + \dist_{G^+}(v,v') \le \dist_{G^+}(r,v) + k + 1 \enspace .
+  \]
+  Similarly, $\dist_{G^+}(r,v) \le \dist_{G^+}(r, v') + k+1$.  Therefore
+  $|\dist_{G^+}(r,v)-\dist_{G^+}(r,v')| \le k+1$ which establishes the claim, since $T$ is BFS tree, so $\dist_{T}(r,z)=\dist_{G^+}(r,z)$ for all $z\in V(G^+)$.
+\end{proof}
+
+\begin{clm}
+  For any $x\in V(H)$, and any integer $i\ge 0$, $|S_x\cap V_i|\le 24k+12$.
+\end{clm}
+
+\begin{proof}
+  Since the vertices of $Y_x$ come from at most 3 vertical paths in $T$, $|V(Y_x)\cap V_i|\le 3$.
+\clmref{chargeback-to-y} shows that, for any $v\in S_x\cap V_i$, there exists $v'\in Y_x$ with $\dist_T(v,v')\le k+1$.
+Therefore, $v'\in V(Y_x)\cap V_{j}$ for some $j\in\{i-k-1,\ldots,i+k+1\}$.
+If $v'=v$ (so $v'\in V_i$) then $v'$ contributes only one vertex to $S_x$.
+If $v'\neq v$ then $v'\in V(G^+-G)$ is a point where two edges of $G$ cross and $v'$ contributes at most the $4$ endpoints of these edges to $S_x$.
+In either case every $v'\in V(Y_x)\cap \bigcup_{j=i-k-1}^{i+k+1} V_j$ contributes at most $4$ vertices to $S_x\cap V_i$, so $|S_x\cap V_i|\le 3\times 4\times (2k+1) = 24k+12$.
+\end{proof}
+
+Next we describe a tree decomposition $(B_x:x\in V(K))$ of $H$.  We deliberately use the same node set $V(H)=V(K)$ here, because the subtree $K[x]:=K[\{z:x\in B_z\}]$ will be rooted at $x$.
+
+\begin{clm}\clmlabel{xy-ancestor}
+   For every edge $xy\in V(H)$, one of $x$ or $y$ is a $K$-ancestor of the other.
+ \end{clm}
+
+ \begin{proof}
+   Since $xy\in E(H)$, there exists an edge $vw\in E(G)$ such that $v\in S_x$ and $w\in S_y$.
+By definition, $x\in A_v\subseteq A_{vw}$ and $y\in A_w\subseteq A_{vw}$.
+Since $K[A_{vw}]$ is a vertical path in $K$, it follows that one of $x$ or $y$ is a $K$-ancestor of the other.
+\end{proof}
+
+The tree decomposition $(B_x:x\in V(K))$ of $H$ is defined as follows: For each edge $xy\in E(H)$ where $y$ is a $K$-ancestor of $x$, we add $y$ to $B_{x'}$ for every node $x'$ on the path from $x$ to $y$ in $K$.
+That this produces a tree-decomposition is due to \clmref{xy-ancestor}.
+What remains is to upper bound the width of this decomposition.
+
+\begin{clm}
+  The tree decomposition $\{B_x: x\in V(K)\}$ of $H$ contains no bag larger than $(3^{k+2}-3)/2$.
+\end{clm}
+
+\begin{proof}
+  Fix some node $x\in V(K)$ and consider the contents of $B_x$.  By \clmref{xy-ancestor}, $B_x$ contains only $K$-ancestors of $x$ (including $x$ itself).
+
+  Consider the graph $A$ that contains the edges and vertices of $F_{a}$ for all $K$-ancestors $a$ of $x$.
+See \figref{A}.
+It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0=r_K$ is the root of $K$.
+We then cut $F_{a_0}$ into two pieces $F^{\bar{x}}_{a_0}$ and $F^x_{a_0}$ where $F^x_{a_0}$ is the piece that contains $F_x$.
+(The path that makes this cut contains two legs $Q_i$ and $Q_j$ of a tripod $Y$ and one edge $v_iv_j$ of its central triangle $\tau$.)  This process continues by partitioning $F_{a_1}:=F^x_{a_0}$ recursively.
+The process terminates after some number $d$ of iterations, when $F^x_{a_{d-1}}=F_x=F_{d}$.
+
+  \begin{figure}
+    \begin{center}
+        \includegraphics{figs/A-1}
+    \end{center}
+    \caption{The graph $A$.}
+    \figlabel{A}
+  \end{figure}
+
+  If $B_x$ contains some $K$-ancestor $a_i$, $i\in\{0,\ldots,d-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_{a_i}$.
+The edge $vw$ corresponds to a walk $W_{vw}$ in $G^+$.
+The walk $W_{vw}$ is contained in $N_x-F_x$, otherwise $x\not\in A_{vw}$ so $x\not\in A_v$, and $a(v)\neq x$, contradicting the fact that $v\in S_x$.
+
+  However, $w\in S_a$, so there exists an edge $ww'\in E(G)$ such that   $W_{ww'}$ is contained in $V(W_{ww'})\subseteq V(N_{a_i}-F_{a_i})$ but $V(W_{ww'}\not\subseteq V(N_{a_{i+1}}-F_{a_{i+1}})$.
+Therefore $W_{ww'}$ must contain a first vertex $w^*\in V(F_{a_{i+1}}-F_{a_{i}})$.
+
+  Summarizing, we have a walk $w'=v_0,\ldots,v_\ell=w^*$ in $G^+$ such that
+  \begin{compactenum}[(i)]
+    \item $v_0\in V(N_x-F_x)$;
+
+    \item $v_0,\ldots,v_{\ell-1}\in V(N_{a_{i+1}}-F_{a_{i+1}})$;
+
+    \item $v_\ell\in V(F_{a_{i+1}}-F_{a_i})$; and
+
+    \item $\ell\le k+1$.
+  \end{compactenum}
+  We call any walk with properties (i)--(iv) an \emph{$i$-walk}.
+Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of integers $i\in\{0,\ldots,d-1\}$ such that there exists an $i$-walk.
+
+  For each $v'\in V(G^+)$, 
+  \[  
+      d_{v'}=\max\{j\in\{0,\ldots,d\} : v'\in V(N_{a_j}-F_{a_j})
+  \]
+  (Intuitively, $d_{v'}$ is the $K$-depth of the lower endpoint of $K[P_{v'}]$ or the $K$-depth ($d$) of $x$, whichever is shallower.)
+
+  Let $v_0,\ldots,v_\ell$ be an $i$-walk and and consider the subsequence $d_0,\ldots,d_{\ell'}$ of $d_{v_0},\ldots,d_{v_\ell}$ consisting of only $d_{v_0}$ and those (record-breaking) $d_{v_i}$ such that $d_{v_i}<\min\{d_{v_0},\ldots,d_{v_{i-1}}\}$.
+Observe that $d_0=d$ and $d_{\ell'}=i$.
+
+  By \propref{ancestor-boundary}, for every $j\in\{1,\ldots,d\}$, there is a set $X_{j}$, of at most 3 integers $\alpha < j$ such that such that every vertex $q\in V(F_{a_j})$ is in $V(Y_{a_{\alpha}})$ for some $\alpha\in X_j$.
+  Assign a distinct colour $c(j,\alpha)\in\{1,2,3\}$ for each $\alpha\in X_j$ and do this assignment for each $j\in\{1,\ldots,d\}$.
+By definition, for any $j\in\{1,\ldots,d\}$ and any colour $\gamma\in\{1,2,3\}$, there is at most one $\alpha=\alpha(j,\gamma)$ such that $c(j,\alpha)=\gamma$.
+
+  Finally, observe that, for every $j\in\{1,\ldots,\ell'\}$,  $c(d_j,d_{j+1})$ is defined.
+Refer to \figref{colouring}.
+Therefore the \emph{colour sequence} $c_0,\ldots,c_{\ell'-1}$ where $c_j=c(d_j,d_{j+1})$ is well defined.
+Since $d_0=d$ and $d_{\ell'-1}=i$, from the preceding discussion, we have $\alpha(\cdots (\alpha(\alpha(\alpha(d,c_0),c_1),c_2)\cdots),c_{\ell'-1}) = i$.
+
+  \begin{figure}
+    \begin{center}
+        \includegraphics{figs/A-2}
+    \end{center}
+    \caption{The colour sequences defined by two minimal admissible $xa$-curves.}
+    \figlabel{colouring}
+  \end{figure}
+
+  The proof is now almost complete.  Any $i$-walk defines a colour sequence $c_0,\ldots,c_{\ell'}$.  If some $i'$-walk defines the same colour sequence $c_0,\ldots,c_{\ell'}$ then
+  \[
+        i = (\alpha(\alpha(\alpha(d,c_0),c_1),c_2)\cdots),c_{\ell'-1}) = i' \enspace ,
+  \]
+  so $i=i'$.  Since there are at most $3^{\ell'}$ colour sequences of length $\ell'$ and $\ell'\le \ell \le k+1$, the number of $i\in\{0,\ldots,d-1\}$ such that there exists an $i$-walk is at most
+  \[
+     |B_x| \le \sum_{\ell=1}^{k+1} 3^\ell = (3^{k+2}-3)/2 \eqlabel{bag-size} \enspace . \qedhere
+  \]
+\end{proof}
+
+We now have all the parts needed for the proof of \thmref{k-planar}.
+
+\begin{proof}
+  We use the layered $H$-partition $(\mathcal{L}, \mathcal{S})$, where $\mathcal{S}=\{S_x:x\in V(K)\}$ and $\mathcal{L}=\{V_i': i=0,1,2\ldots\}$ where $V_i' = V_{(k+1)i}\cup\cdots\cup V_{(k+1)(i+1)-1}$.
+\Eqref{distance-preserving} shows that $\mathcal{L}$ is indeed a layering of $G$.
+
+  By \clmref{bag-size}, $\mathcal{S}$ produces a graph $H=G/\mathcal{S}$ of treewidth $(3^{k+2}-1)/2$.
+By \clmref{layered-width}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+12$, so $|B_x\cap V_i'|\le (k+1)(24k + 12) = 24k^2+36k+12$.
+\end{proof}
 
@@
 
-\section{$k$-Planar Graphs}
-
-\subsection{PS-Trees}
-
-Let $T$ be a rooted tree rooted at $r\in V(T)$.
-The \emph{depth} of a node $x\in V(T)$ is equal to the length of the path from $x$ to $r$ in $T$.
-A $P$ path in $T$ is \emph{vertical} if, for each integer $i$, $P$ contains at most one vertex of depth $i$.
-The deepest vertex in a vertical path $P$ is called  $P$'s \emph{lower endpoint} and the other (shallowest) vertex in $P$ is called $P$'s \emph{upper endpoint}.
-For any node $v\in T$, we call each node on the path, in $T$, from $v$ to $r$ a \emph{$T$-ancestor} of $v$.
-
-A \emph{near-triangulation} $N$ is a plane graph whose outer face is bounded by a simple cycle in $N$ and each of whose inner faces is bounded by a 3-cycle in $N$.
-
-Let $\Delta$ be a plane triangulation and let $T$ be a BFS spanning tree of $\Delta$ rooted at some vertex $r$ of degree 3 that is on the outer face of $\Delta$.
-For any cycle $F$ of $\Delta$ there is an associated near-triangulation $N=\Delta\circledast F$ consisting of all the edges and vertices of $G$ that are contained in $F$ or its interior.
-
-Let $\tau=v_1v_2v_3$ be an inner face of $N$ and consider the three minimal vertical paths $Q_1$, $Q_2$, and $Q_3$ in $T$ such that, for each $i\in\{1,2,3\}$ the lower endpoint of $Q_i$ is $v_i$ and the upper endpoint of $Q_i$ is in $V(F)$.
-If $Q_1$, $Q_2$, and $Q_3$ are vertex disjoint, then the graph $\overline{Y}=\tau\cup Q_1\cup Q_2\cup Q_3$ is called the \emph{closed tripod} in $N$ determined by $\tau$.
-We call $Q_1$, $Q_2$, and $Q_3$ the \emph{legs} of $\overline{Y}$ and we call $\tau$ the \emph{center} of $\overline{Y}$.
-We call $Y=\overline{Y}-F$ the \emph{(open) tripod} of $N$ determined by $\tau$.
-The graph $M=\overline{Y}\cup F$ is called a \emph{Mercedes graph} and the (at most 3) inner faces of $M$ other than $\tau$ are called the \emph{wedges} of $M$.
-
-A \emph{PS-tree}  $K=K(\Delta, T, r)$ is a 3-ary tree, such that each node $x\in V(K)$ is associated with several object (refer to \figref{K-node}):
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{cc}
-      \includegraphics{figs/K-node-1} &
-      \includegraphics{figs/K-node-2}
-    \end{tabular}
-  \end{center}
-  \caption{The elements associated with a node $x\in V(K)$.}
-  \figlabel{K-node}
-\end{figure}
-
-\begin{enumerate}
-  \item A cycle $F_x$ in $\Delta$ and the near-triangulation $N_x=\Delta\circledast F_x$.
-
-  \item An inner face $\tau_x$ of $N_x$ that determines a closed tripod $\overline{Y}_x$ and an open tripod $Y_x$ in $N_x$.
-
-  \item The \emph{Mercedes graph} $M_x=F_x\cup \overline{Y}_x$
-\end{enumerate}
-
-The nodes of the PS-tree $K$ satisfy several conditions:
-\begin{enumerate}[(PR1)]
-  \item For the root $r'$ of $K$, $F_{r'}$ is defined to be the 3-cycle in $\Delta$ formed by the three vertices adjacent to $r$.
-
-  \item \proplabel{children} For each node $x\in V(T)$ such that $M_x$ has $f\le 3$ wedges $F_{x,1},\ldots,F_{x,f}$, the node $x$ has exactly $f$ children $y_1,\ldots,y_f$ in $K$ and $F_{y_i} = F_{x,i}$ for each $i\in\{1,\ldots,f\}$.
-
-  \item \proplabel{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$ \emph{or} $F[P_i]$ is a path or cycle in $F_{r'}$.
-
-  \item \proplabel{partition} $\{Y_x : x\in V(K)\}$ is a partition of $V(G^+)\setminus\{r\}\setminus V(F_{r'})$.
-\end{enumerate}
-
-For any triangulation $\Delta$ in which the outer face has a vertex $r$ of degree 3 and for any BFS spanning-tree of $\Delta$ rooted at $r$, there exists a PS-tree $K=K(\Delta,T,r)$.
-Indeed, this tree is constructed (albeit implicitly) by \citet[Proof of Lemma~X]{dujmovic.joret.ea:planar}.
-
-The following property of a PS-Tree $K$ is an immediate consequence of \propref{children} that will be useful for us:
-\begin{compactenum}[(PR5)]
-  \item \proplabel{ancestors} If $v\in V(N_x-F_x)$ and $w\in V(N_y-N_y)$ for some nodes $x,y\in V(K)$, then one of $x$ or $y$ is a $K$-ancestor of the other.
-\end{compactenum}
-
-\subsection{$k$-Planar Graphs}
-
-Let $G$ be a $k$-plane graph and let $G^+$ be the graph obtained as follows:
-\begin{compactenum}
-  \item Make $G$ planar by adding dummy vertices at each crossing to obtain a graph $G_1$.
-  \item Add a $K_4$ that contains $G_1$ in one of its inner faces $F_0$ to obtaina graph $G_2$.
-  \item Add edges to complete $G_2$ to a triangulation $G^+$.
-\end{compactenum}
-Let $r$ be the vertex of the $K_4$ not incident on $F_0$ and let $T$ be a BFS spanning tree of $G^+$ rooted at $r$.
-Observe that, for any edge $vw\in E(G)$, $\dist_{G^+}(v,w) \le k+1$ and therefore, for any two vertices $v,w\in V(G)$,
-\begin{equation} 
-  \dist_{G^+}(v,w) \le (k+1)\cdot \dist_G(v,w) \eqlabel{distance-preserving}
-\end{equation}
-For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{G^+}(r,v)=i\}$ and observe that $V_0,V_1,\ldots,V_{t}$ is a layering of $G^+$ (but not of $G$).
-
-Let $K=K(G^+,T,r)$ be a PS-tree and observe that the root $r'$ of $K$ has $F_{r'}=F_0$.  This implies:
-\begin{compactenum}[(PR6)]
-  \item \proplabel{all-in}  $V(G_1)\subseteq V(N_{r'}-F_{r'})$.
-\end{compactenum}
-
-For each vertex $v\in V(G_1)$, let $P_v=\{a\in V(K): v\in V(N_a-F_a)\}$.  
-By (PR6), $P_{v}$ contains the root $r'$ of $K$.
-By \propref{children}, any node $a\in P_{v}$ has at most one child in $P_{v}$ and any node $a\not\in P_{v}$ has no child in $P_{v}$.
-It follows that $K[P_{v}]$ is a vertical path in $K$ that contains the root.
-
-For each edge $vw\in E(G)$, consider the walk $W_{vw}$ in $G^+$ that corresponds to $vw$ and define $A_{vw}=\bigcap_{z\in W_{vw}} P_z$.
-Since $W_{vw}\subseteq V(G_1)$, $A_{vw}$ is well-defined.
-For each $z\in W_{vw}$, $P_{z}$ contains $r'$ and is a vertical path in $K$.
-The common intersection of two or more vertical paths is a vertical path.
-Therefore $A_{vw}$ is a non-empty vertical path in $K$.
-For any vertex $v\in V(G)$ define $A_v=\bigcap_{vw\in E(G)} A_{vw}$.
-Again, $K[A_v]$ is a vertical path in $K$ that contains $r'$.
-We define $a(v)$ to be the lower endpoint of $K[A_v]$.
-
-For every node $x\in V(K)$, define the \emph{separator}
-\[
-   S_x = \{v\in V(G): a(v)=x \} \enspace .
-\]
-Clearly $\mathcal{S}=\{S_x:x\in V(K)\}$ is a partition of $V(G)$.
-
-Consider graph $H=G/\mathcal{S}$ that has vertex set $V(H)=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
-First we argue that this partition has small width with respect to the layering defined by $T$ (recall that this is a layering of $G^+$, not a layering of $G$).
-
-\begin{clm}
-  For any $x\in V(H)$, and any integer $i\ge 0$, $|S_x\cap V_i|\le 3(4k+3)$.
-\end{clm}
-
-\begin{proof}
-  Since the vertices of $Y_x$ come from at most 3 vertical paths in $T$, $|V(Y_x)\cap V_i|\le 3$.
-Any vertex $v\in S_x\cap V_i$ is either in $V(Y_x)$ or is an endpoint of an edge $vw\in E(G)$ that contains a vertex $v'\in V(Y_x)$.
-In the latter case, $|\dist_T(r,v)-\dist_T(r, v')|\le k+1$.
-Therefore $v'\in V(Y_x)\cap V_{j}$ for some $j\in\{i-k-1,\ldots,i+k+1\}$.
-Any such $v'$ is responsible for at most 2 vertices in $S_x\setminus V(Y_x)$.
-Therefore,
-  \[  |S_x\cap V_i| \le |V(Y_x)\cap V_i| + 3\cdot 2\cdot (2k+1) \le 12k+9 \enspace . \qedhere \]
-\end{proof}
-
-Next we describe a tree decomposition $(B_x:x\in V(K))$ of $H$.  We deliberately use the same node set $V(H)=V(K)$ here, because the subtree $K[x]:=K[\{z:x\in B_z\}]$ will be rooted at $x$.
-
-\begin{clm}\clmlabel{xy-ancestor}
-   For every edge $xy\in V(H)$, one of $x$ or $y$ is a $K$-ancestor of the other.
- \end{clm}
-
- \begin{proof}
-   Since $xy\in E(H)$, there exists an edge $vw\in E(G)$ such that $v\in S_x$ and $w\in S_y$.
-By definition, $x\in A_v\subseteq A_{vw}$ and $y\in A_w\subseteq A_{vw}$.
-Since $K[A_{vw}]$ is a vertical path in $K$, it follows that one of $x$ or $y$ is a $K$-ancestor of the other.
-\end{proof}
-
-The tree decomposition $(B_x:x\in V(K))$ of $H$ is defined as follows: For each edge $xy$ where $y$ is a $K$-ancestor of $x$, we add $y$ to be $B_{x'}$ for every node $x'$ on the path from $x$ to $y$ in $K$.
-That this produces a tree-decomposition is due to \clmref{xy-ancestor}.
-What remains is to upper bound the width of this decomposition.
-
-\begin{clm}
-  The tree decomposition $\{B_x: x\in V(K)\}$ of $H$ contains no bag larger than BLAH.
-\end{clm}
-
-\begin{proof}
-  Fix some node $x\in V(K)$ and consider the contents of $B_x$.  By \clmref{xy-ancestor}, $B_x$ contains only $K$-ancestors of $x$ (including $x$ itself).
-
-  Consider the graph $A$ that contains $F_{a}$ for all $K$-ancestors $a$ of $x$.
-See \figref{A}.
-It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0=r'$ is the root of $K$.
-We then cut $F_{a_0}$ into two pieces $F^{\bar{x}}_{a_0}$ and $F^x_{a_0}$ where $F^x_{a_0}$ is the piece that contains $F_x$.
-(The path that makes this cut contains two legs $Q_i$ and $Q_j$ of a tripod $Y$ and one edge $v_iv_j$ of its central triangle $\tau$.)  This process continues by partitioning $F_{a_1}:=F^x_{a_0}$ recursively.
-The process terminates after some number $r$ of iterations, when $F^x_{a_r}=F_x$.
-
-  \begin{figure}
-    \begin{center}
-        \includegraphics{figs/A-1}
-    \end{center}
-    \caption{The graph $A$.}
-    \figlabel{A}
-  \end{figure}
-
-  If $B_x$ contains some $K$-ancestor $a_i$, $i\in\{0,\ldots,r-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_{a_i}$.
-The edge $vw$ corresponds to a walk $W_{vw}$ in $G^+$.
-The walk $W_{vw}$ is contained in $N_x-F_x$, otherwise $x\not\in A_{vw}$ so $x\not\in A_v$.
-
-  However, $w\in S_a$, so there exists an edge $ww'\in E(G)$ such that 
-  the corresponding walk $W_{ww'}$ is contained in $N_{a_i}-F_{a_i}$ but is not
-  contained in $N_{a_{i+1}}-F_{a_{i+1}}$.  Therefore $W_{ww'}$ must contain a first vertex $w^*\in V(F_{a_{i+1}})\setminus V(F_{a_{i}})$.
-
-  Summarizing, we have a walk $w'=v_0,\ldots,v_\ell=w^*$ in $G^+$ such that
-  \begin{compactenum}[(i)]
-    \item $v_0\in V(N_x-F_x)$;
-
-    \item $v_0,\ldots,v_{\ell-1}\in V(N_{a_{i+1}}-F_{a_{i+1}})$;
-
-    \item $v_\ell\in V(F_{F_{a_{i+1}}})-V(F_{a_i})$; and
-
-    \item $\ell\le k+1$.
-  \end{compactenum}
-  We call any walk with properties (i)--(iv) an \emph{$i$-walk}. Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of positive integers $i$ such that there exists an $i$-walk.
-
-  Define $L_r=V(N_{a_r}-F_{a_r})$ and, for each $j\in\{0,\ldots,r-1\}$, define
-  $L_j = V(N_{a_j}- F_{a_j})\setminus L_{a_{j+1}}$ and observe that $L_i,\ldots,L_r$ is a partition of $V(N_{a_i}-F_{a_i})$.
-Let $v_0,\ldots,v_\ell$ be an $i$-walk and, for each $j\in\{0,\ldots,\ell\}$, let $x_j$ be the unique node $a_f\in\{a_i,\ldots,a_r\}$ such that $v_j\in L_{f}$.
-  Finally, let $x_0',\ldots,x_{\ell'}'$ be the subsequence of $x_0,\ldots,x_{\ell}$ consisting of $x_0$ and only those nodes $x_j$ such that $V(N_{x_j})\supsetneq \bigcup_{q=0}^{j-1} V(N_{x_q})$.
-
-  By \propref{ancestor-boundary}, for every $y\in V(K)$, there are $t\le 3$ $K$-ancestors $y_1,\ldots,y_t$ of $y$ such that $F_y$ has a vertex in common with $F^{\bar{x}}_{y}$.
-For each $j\in\{1,\ldots,t\}$, colour the pair $(y,y_j)$ with the colour $c(y,y_j)=j$.
-In this way any pair that receives a colour receives one of 3 colours.
-The set $y_1,\ldots,y_t$ behaves like a cutset in the following sense: If $x_j'=y$ for $j<\ell'$, then $x_{j+1}'\in\{y_1,\ldots,y_t\}$.
-
-  \begin{figure}
-    \begin{center}
-        \includegraphics{figs/A-2}
-    \end{center}
-    \caption{The colour sequences defined by two minimal admissible $xa$-curves.}
-    \figlabel{colouring}
-  \end{figure}
-
-  Therefore $x_0',\ldots,x_{\ell'}'$ defines a color sequence
-  $c_1,\ldots,c_{\ell'}$, $\ell'\le \ell$, where $c_j=c(x_{j-1}',x_j')$ for each $j\in\{1,\ldots,\ell'\}$.
-There are at most $3^{\ell'}$ such colour sequences and, if an $i$-walk has the same colour sequence as an $i'$-walk,
-  then $i=i'$.  Therefore, the number of positive integers $i$ such that there exists and $i$-walk is at most
-  \begin{equation}
-     |B_x| \le \sum_{\ell=0}^{2k+1} 3^\ell = (3^{2k+2}-1)/2 
-         = (9^{k+1}-1)/2 \enspace .  \eqlabel{bag-size}
-  \end{equation}
-\end{proof}
-
-\begin{proof}
-  By \eqref{bag-size}, the partition $(S_x:x\in V(K))$ of $G$ described above produces a graph $H$ of treewidth $(9^{k+1}-3)/2$.
-For each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 12k+9$.
-\Eqref{distance-preserving} implies that setting $V_i' = V_{(k+1)i}\cup\cdots\cup V_{(k+1)(i+1)-1}$ yields a layering $(V_0',V_1',\ldots)$ of  $G$ and $|B_x\cap V_i'|\le (k+1)(12k+9)$ for all integers $i\ge 0$ and all $x\in V(H)$.
-\end{proof}
+\section{Consequences}
+\seclabel{consequences}
 

### 2019-05-10 k-planar.tex
Getting closer....

@@
 
-A \emph{layering} of a graph $G$ is a partition $\mathcal{L}=\{V_0,V_1,\ldots,V_h\}$ of $V(G)$ so that there is no edge $vw\in E(G)$ such that $v\in V_i$, $w\in V_j$ and $|j-i|>1$.
+A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle V_0,V_1,\ldots,V_h\rangle$ such that $\{V_0,V_1,\ldots,V_h\}$ is a partition of $V(G)$ and there is no edge $vw\in E(G)$ such that $v\in V_i$, $w\in V_j$ and $|j-i|>1$.
 For any partition $\mathcal{P}=\{S_1,\ldots,S_p$ of $V(G)\}$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
-To highlight the importance of the quotient graph $H$, we can $\mathcal{P}$ an $H$-partition and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.\footnote{An alternative definition of a layering of $G$ is a layered $H$-partition of $G$ where $H$ is a path.}  A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
+To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an $H$-partition and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.\footnote{An alternative definition of a layering of $G$ is a layered $H$-partition of $G$ where $H$ is a path.}  A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
 The \emph{layered width} of $(\mathcal{H},\mathcal{P})$ is $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
 
-Layered $H$-partition of small layered width in which $H$ has some additional property are a useful tool for a number of graph problems.
+Layered $H$-partitions of small layered width in which $H$ has some additional property are a useful tool for a number of graph problems.
 \citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number \cite{X}, track number \cite{X}, etc \cite{X}, etc \cite{X}, etc \cite{X}.
@@
 
-Under these definitions $0$-planar graphs are exactly planar graphs and $O$-plane graphs are plane graphs.
+Under these definitions $0$-planar graphs are exactly planar graphs and $0$-plane graphs are plane graphs.
 Thus, like genus-$g$ graphs, $k$-planar graphs are a generalization of planar graphs (which are genus-$0$ graphs).
@@
 Let $T$ be a rooted tree rooted at $r\in V(T)$.
-The \emph{depth} of a node $x\in V(T)$ is equal to the length of the path from $x$ to $r$ in $T$.
-A $P$ path in $T$ is \emph{vertical} if, for each integer $i$, $P$ contains at most one vertex of depth $i$.
+The \emph{$T$-depth} of a node $x\in V(T)$ is equal to the length of the path from $x$ to $r$ in $T$.
+A $P$ path in $T$ is \emph{vertical} if, for each integer $i$, $P$ contains at most one vertex of $T$-depth $i$.
 The deepest vertex in a vertical path $P$ is called  $P$'s \emph{lower endpoint} and the other (shallowest) vertex in $P$ is called $P$'s \emph{upper endpoint}.
@@
 
-The following property of a PS-Tree $K$ is an immediate consequence of \propref{children} that will be useful for us: \note{PM}{Do we still use this?}
-\begin{compactenum}[(PR5)]
-  \item \proplabel{ancestors} If $vw\in E(\Delta)$, $v\in V(N_x-F_x)$ and $w\in V(N_y-N_y)$ for some nodes $x,y\in V(K)$, then one of $x$ or $y$ is a $K$-ancestor of the other.
-\end{compactenum}
+% The following property of a PS-Tree $K$ is an immediate consequence of \propref{children} that will be useful for us: \note{PM}{Do we still use this?}
+% \begin{compactenum}[(PR5)]
+%   \item \proplabel{ancestors} If $vw\in E(\Delta)$, $v\in V(N_x-F_x)$ and $w\in V(N_y-N_y)$ for some nodes $x,y\in V(K)$, then one of $x$ or $y$ is a $K$-ancestor of the other.
+% \end{compactenum}
 
@@
 
-Let $G$ be a $k$-plane graph and let $G^+$ be the graph obtained as follows:
+Let $G$ be a $k$-plane graph.
+We will assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$.
+This assumption is not critical since it can be enforced by a slight deformation of the edges of $G$.
+
+Let $G^+$ be the graph obtained as follows:
 \begin{compactenum}
-  \item Make $G$ planar by adding dummy vertices at each crossing to obtain a plane graph and then add edges to this graph to obtain a triangulation $G^+$.
-  \item Add a complete graph $K_4$ that contains $G^{+}$ in one of its inner faces $F_0$ to obtain graph $G^{++}$.
-  \item Add edges to complete $G^{++}$ to a triangulation $G^{+++}$.
+  \item Add a dummy vertex at each crossing in $G$ to obtain a plane graph and then add edges to this graph to obtain a triangulation $G^+$.
+  \item Add a complete graph $K_4$ that contains $G^{+}$ in one of its inner faces, $F_0$, and then add edges to the resulting graph to obtain a triangulation $G^{++}$.
 \end{compactenum}
@@
 \end{equation}
-For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{T}(r,v)=i\}$ and observe that $\mathcal{L}=V_0,V_1,\ldots,V_{h}$ is a layering of $G^+$ (but not of $G$).
+For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{T}(r,v)=i\}$ and observe that $\{V_0,V_1,\ldots,V_{h}\}$ is a layering of $G^+$ (but not of $G$).
 
-Let $K=K(G^{+++},T,r)$ be a PS-tree and observe that the root $r_K$ of $K$ has $F_{r_K}=F_0$.  This implies:
-\begin{compactenum}[(PR6)]
+Let $K=K(G^{++},T,r)$ be a PS-tree and observe that the root $r_K$ of $K$ has $F_{r_K}=F_0$.  This implies:
+\begin{compactenum}[(PR1)]\setcounter{enumi}{4}
   \item \proplabel{all-in}  $V(G^+)\subseteq V(N_{r_K}-F_{r_K})$.
@@
 For each vertex $v\in V(G^+)$, let $P_v=\{a\in V(K): v\in V(N_a-F_a)\}$.  
-By (PR6), $P_{v}$ contains the root $r_K$ of $K$.
+By \propref{all-in}, $P_{v}$ contains the root $r_K$ of $K$.
 By \propref{children}, $K[P_v]$ is a vertical path (whose lower endpoint is the unique node $x\in V(K)$ such that $v\in Y_x$).
@@
 
-\begin{clm}
+\begin{clm}\clmlabel{width-of-g-plus}
   For any $x\in V(H)$, and any integer $i\ge 0$, $|S_x\cap V_i|\le 24k+12$.
@@
 
-\begin{clm}
+\begin{clm}\clmlabel{bag-size}
   The tree decomposition $\{B_x: x\in V(K)\}$ of $H$ contains no bag larger than $(3^{k+2}-3)/2$.
@@
     \end{center}
-    \caption{The colour sequences defined by two minimal admissible $xa$-curves.}
+    \caption{The colour sequences defined by two $i$-curves.}
     \figlabel{colouring}
@@
   By \clmref{bag-size}, $\mathcal{S}$ produces a graph $H=G/\mathcal{S}$ of treewidth $(3^{k+2}-1)/2$.
-By \clmref{layered-width}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+12$, so $|B_x\cap V_i'|\le (k+1)(24k + 12) = 24k^2+36k+12$.
+By \clmref{width-of-g-plus}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+12$, so $|B_x\cap V_i'|\le (k+1)(24k + 12) = 24k^2+36k+12$.
 \end{proof}

### 2019-05-12 k-planar.tex
Polishing...

@@
 Such partitions are known to exist for planar graphs and, more generally, bounded genus graphs.
-In the current paper, we prove that every $k$-planar graph has a layered $H$-partition of layered width $O(k)$ in which $H$ has treewidth $O(c^k)$, for some constant $c$.
-This is the first result of this type of a non-minor closed class of graphs and implies that $k$-planar graphs have both queue number and non-repetitive chromatic number upper-bounded by a function of $k$.
+In the current paper, we prove that every $k$-planar graph has a layered $H$-partition of layered width $O(k^2)$ in which $H$ has treewidth $O(c^k)$, for some constant $c$.
+This is the first result of this type for a non-minor-closed class of graphs and implies that $k$-planar graphs have both queue number and non-repetitive chromatic number upper-bounded by a function of $k$.
 The former result was previously shown using a combination of $H$-partitions for planar graphs with \textit{ad hoc} methods.
-The latter result is new, for all $k\ge 3$.
+The latter result is new, for all $k\ge 1$.
 \end{abstract}
@@
 A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle V_0,V_1,\ldots,V_h\rangle$ such that $\{V_0,V_1,\ldots,V_h\}$ is a partition of $V(G)$ and there is no edge $vw\in E(G)$ such that $v\in V_i$, $w\in V_j$ and $|j-i|>1$.
-For any partition $\mathcal{P}=\{S_1,\ldots,S_p$ of $V(G)\}$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
+For any partition $\mathcal{P}=\{S_1,\ldots,S_p\}$ of $V(G)$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
 To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an $H$-partition and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.\footnote{An alternative definition of a layering of $G$ is a layered $H$-partition of $G$ where $H$ is a path.}  A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
@@
 Layered $H$-partitions of small layered width in which $H$ has some additional property are a useful tool for a number of graph problems.
-\citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number \cite{X}, track number \cite{X}, etc \cite{X}, etc \cite{X}, etc \cite{X}.
+\citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number \cite{X}, track number \cite{X}, layered treewidth \cite{X}, etc \cite{X}, etc \cite{X}, etc \cite{X}.
 \citet{dujmovic.esperet.ea:planar} add non-repetitive chromatic number to this list.
@@
 
-Under these definitions $0$-planar graphs are exactly planar graphs and $0$-plane graphs are plane graphs.
+Under these definitions $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
 Thus, like genus-$g$ graphs, $k$-planar graphs are a generalization of planar graphs (which are genus-$0$ graphs).
@@
 \begin{thm}\thmlabel{k-planar}
-  Every $k$-planar graph has a layered $H$-partition of layered-width at most $24k^2 + 36k + 12$ in which $H$ has treewidth at most $(3^{k+2}-3)/2$.
+  Every $k$-planar graph has a layered $H$-partition of layered-width at most $24k^2 + 36k + 12$ in which $H$ has treewidth at most $(3^{k+2}-5)/2$.
 \end{thm}
@@
 We use $G-S$ as a shorthand for $G[V(G)\setminus S]$ and for a graph $G'$, we use $G-G'$ as a shorthand for $G-V(G')$.
+We use $G'\subseteq G$ to denote subgraph containment, i.e., $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.
 
@@
 
-Let $T$ be a rooted tree rooted at $r\in V(T)$.
+Let $T$ be a rooted tree rooted at some node $r\in V(T)$.
 The \emph{$T$-depth} of a node $x\in V(T)$ is equal to the length of the path from $x$ to $r$ in $T$.
 A $P$ path in $T$ is \emph{vertical} if, for each integer $i$, $P$ contains at most one vertex of $T$-depth $i$.
-The deepest vertex in a vertical path $P$ is called  $P$'s \emph{lower endpoint} and the other (shallowest) vertex in $P$ is called $P$'s \emph{upper endpoint}.
+The deepest vertex in a vertical path $P$ is called  $P$'s \emph{lower endpoint} and the shallowest vertex in $P$ is called $P$'s \emph{upper endpoint}.
 For any node $v\in T$, we call each node on the path, in $T$, from $v$ to $r$ a \emph{$T$-ancestor} of $v$.
@@
 
-For the remainder of this section, $\Delta$ is a triangulation, and $T$ is a BFS spanning tree of $\Delta$ rooted at some degree-3 vertex $r$ on the outer face of $\Delta$.
+For the remainder of this section, $\Delta$ is a triangulation, $r$ is a degree-3 vertex on the outer face of $\Delta$, and $T$ is a BFS spanning tree of $\Delta$ rooted at $r$.
 
@@
 
-  \item An inner face $\tau_x$ of $N_x$ that determines a closed tripod $\overline{Y}_x$ and an open tripod $Y_x$ in $N_x$.
+  \item An inner face $\tau_x$ of $N_x$ that determines a closed tripod $\overline{Y}_x$ and an open tripod $Y_x=\overline{Y}_x-F_x$ in $N_x$.
 
@@
 
-  \item \proplabel{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$ \emph{or} $F[P_i]$ is a path or cycle in $F_{r_K}$.
-
+  \item \proplabel{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, either:
+  \begin{compactenum}
+    \item $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$;
+    \item $F[P_i]$ is a cycle and there exists an edge of $P_i$ whose removal produces a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$; or 
+    \item $F[P_i]$ is a path or cycle in $F_{r_K}$.
+  \end{compactenum}
   \item \proplabel{partition} $\{Y_x : x\in V(K)\}$ is a partition of $V(\Delta)\setminus\{r\}\setminus V(F_{r_K})$.
@@
 This assumption is not critical since it can be enforced by a slight deformation of the edges of $G$.
-
-Let $G^+$ be the graph obtained as follows:
+Consider the following plane graphs:
 \begin{compactenum}
@@
 \end{compactenum}
-Let $r$ be the vertex of the $K_4$ not incident on $F_0$ and let $T$ be a BFS spanning tree of $G^+$ rooted at $r$.
-Observe that, for any edge $vw\in E(G)$, $\dist_{G^+}(v,w) \le k+1$ and therefore, for any two vertices $v,w\in V(G)$,
-\begin{equation} 
-  \dist_{G^+}(v,w) \le (k+1)\cdot \dist_G(v,w) \eqlabel{distance-preserving}
-\end{equation}
+Let $r$ be the vertex of the $K_4$ not incident on $F_0$ and let $T$ be a BFS spanning tree of $G^{++}$ rooted at $r$. Observe that, for any edge $vw\in E(G)$, 
+\[  \dist_{G^+}(v,w) \le k+1 \enspace . \]
+%  and therefore, for any two vertices $v,w\in V(G)$,
+% \begin{equation} 
+%   \dist_{G^+}(v,w) \le (k+1)\cdot \dist_G(v,w) \eqlabel{distance-preserving}
+% \end{equation}
 For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{T}(r,v)=i\}$ and observe that $\{V_0,V_1,\ldots,V_{h}\}$ is a layering of $G^+$ (but not of $G$).
@@
 
-For each vertex $v\in V(G^+)$, let $P_v=\{a\in V(K): v\in V(N_a-F_a)\}$.  
-By \propref{all-in}, $P_{v}$ contains the root $r_K$ of $K$.
-By \propref{children}, $K[P_v]$ is a vertical path (whose lower endpoint is the unique node $x\in V(K)$ such that $v\in Y_x$).
-Therefore $K[P_v]$ is a vertical path in $K$ that contains $r_K$.
+For each vertex $z\in V(G^+)$, let $P_z=\{a\in V(K): z\in V(N_a-F_a)\}$.  
+By \propref{all-in}, $P_{z}$ contains the root $r_K$ of $K$.
+By \propref{children}, $K[P_z]$ is a vertical path (whose lower endpoint is the unique node $x\in V(K)$ such that $z\in Y_x$).
+Therefore $K[P_z]$ is a vertical path in $K$ that contains $r_K$.
 We are about to repeatedly use the fact that the common intersection of two or more vertical paths is a vertical path.
 
-For each edge $vw\in E(G)$, consider the walk $W_{vw}$ in $G^+$ that corresponds to the sequence of vertices in $G^+$ that we encounter while walking along the curve $vw$ from $v$ to $w$.
+For each edge $vw\in E(G)$, consider the walk $W_{vw}$ in $G^+$ that corresponds to the sequence of vertices in $G^+$ encountered while walking along the curve $vw$ from $v$ to $w$.
 We define $A_{vw}=\bigcap_{z\in W_{vw}} P_z$.
-Since $W_{vw}\subseteq V(G^+)$, $A_{vw}$ is well-defined.
-For each $z\in W_{vw}$, $K[P_{z}]$ is a vertical path in $K$ that contains $r_K$.
+Since $V(W_{vw})\subseteq V(G^+)$, $A_{vw}$ is well-defined.
+For each $z\in V(W_{vw})$, $K[P_{z}]$ is a vertical path in $K$ that contains $r_K$.
 Therefore $A_{vw}$ is a non-empty vertical path in $K$.
@@
 We will use $\mathcal{S}$ as the partition that defines our quotient graph $H=G/\mathcal{S}$ with vertex set $V(H)=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
-First we argue that this partition has small width with respect to the layering defined by $T$ (recall that this is a layering of $G^+$, not a layering of $G$).
+First we argue that this partition has small width with respect to the layering $\{V_0,\ldots,V_h\}$ defined by $T$ (recall that this is a layering of $G^+$, not a layering of $G$).
 
@@
   Since $v\in S_x$, the lower endpoint of $K[A_{vw}]$ is $x$ for some edge $vw\in E(G)$.
-This implies that $V(W_{vw})\subset V(N_x-F_x)$ but $V(W_{vw}) \not\subseteq V(N_{x'}-F_{x'})$ for any any child $x'$ of $x$ in $K$.
-This implies that $W_{vw}$ contains a vertex $v'\in Y_x$.
-Since $G$ is $k$-planar and $v'\in V_{vw}$ this implies that $\dist_{G^+}(v,v') \le k+1$.
+This implies that $W_{vw}\subset N_x-F_x$ but $W_{vw} \not\subseteq N_{x'}-F_{x'}$ for any any child $x'$ of $x$ in $K$.
+Therefore, \propref{children} and \propref{ancestor-boundary} imply that $W_{vw}$ contains a vertex $v'\in Y_x$.
+Since $G$ is $k$-planar and $v'\in V(W_{vw})$ this implies that $\dist_{G^+}(v,v') \le k+1$.
   By the triangle inequality,     
@@
   Similarly, $\dist_{G^+}(r,v) \le \dist_{G^+}(r, v') + k+1$.  Therefore
-  $|\dist_{G^+}(r,v)-\dist_{G^+}(r,v')| \le k+1$ which establishes the claim, since $T$ is BFS tree, so $\dist_{T}(r,z)=\dist_{G^+}(r,z)$ for all $z\in V(G^+)$.
+  $|\dist_{G^+}(r,v)-\dist_{G^+}(r,v')| \le k+1$ which establishes the claim, since $T$ is BFS tree rooted at $r$, so $\dist_{T}(r,z)=\dist_{G^+}(r,z)$ for all $z\in V(G^+)$.
 \end{proof}
@@
 
-Next we describe a tree decomposition $(B_x:x\in V(K))$ of $H$.  We deliberately use the same node set $V(H)=V(K)$ here, because the subtree $K[x]:=K[\{z:x\in B_z\}]$ will be rooted at $x$.
+Next we describe a tree decomposition $\{B_x:x\in V(K)\}$ of $H$.  We deliberately use the same node set $V(H)=V(K)$ here, because the subtree $K[x]:=K[\{z:x\in B_z\}]$ will be rooted at $x$.
 
@@
 (The path that makes this cut contains two legs $Q_i$ and $Q_j$ of a tripod $Y$ and one edge $v_iv_j$ of its central triangle $\tau$.)  This process continues by partitioning $F_{a_1}:=F^x_{a_0}$ recursively.
-The process terminates after some number $d$ of iterations, when $F^x_{a_{d-1}}=F_x=F_{d}$.
+The process terminates after some number $d$ of iterations, when $F^x_{a_{d-1}}=F_x=F_{a_d}$.
 
@@
 
-  However, $w\in S_a$, so there exists an edge $ww'\in E(G)$ such that   $W_{ww'}$ is contained in $V(W_{ww'})\subseteq V(N_{a_i}-F_{a_i})$ but $V(W_{ww'}\not\subseteq V(N_{a_{i+1}}-F_{a_{i+1}})$.
-Therefore $W_{ww'}$ must contain a first vertex $w^*\in V(F_{a_{i+1}}-F_{a_{i}})$.
+  However, $w\in S_{_i}$, so there exists an edge $ww'\in E(G)$ such that   $W_{ww'}$ is contained in $W_{ww'}\subseteq N_{a_i}-F_{a_i}$ but $W_{ww'}\not\subseteq N_{a_{i+1}}-F_{a_{i+1}}$.
+Therefore $W_{ww'}$ must contain a first vertex $w^*\in V(F_{a_{i+1}}-F_{a_{i}})\subseteq Y_{a_i}$.
 
-  Summarizing, we have a walk $w'=v_0,\ldots,v_\ell=w^*$ in $G^+$ such that
+  Summarizing, we have a walk $w'=v_0,v_1\ldots,v_{\ell-1},v_\ell=w^*$ in $G^+$ such that
   \begin{compactenum}[(i)]
@@
   \end{compactenum}
-  We call any walk with properties (i)--(iv) an \emph{$i$-walk}.
-Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of integers $i\in\{0,\ldots,d-1\}$ such that there exists an $i$-walk.
+  We call any walk with properties (i)--(iv) an \emph{$i$-walk}. Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $i\in\{0,\ldots,d-1\}$ such that there exists an $i$-walk.
 
-  For each $v'\in V(G^+)$, 
+  For each $v'\in V(G^+)$, define
   \[  
@@
 
-  Let $v_0,\ldots,v_\ell$ be an $i$-walk and and consider the subsequence $d_0,\ldots,d_{\ell'}$ of $d_{v_0},\ldots,d_{v_\ell}$ consisting of only $d_{v_0}$ and those (record-breaking) $d_{v_i}$ such that $d_{v_i}<\min\{d_{v_0},\ldots,d_{v_{i-1}}\}$.
+  Let $v_0,\ldots,v_\ell$ be an $i$-walk and consider the subsequence $d_0,\ldots,d_{\ell'}$ of $d_{v_0},\ldots,d_{v_\ell}$ consisting of only $d_{v_0}$ and those (record-breaking) $d_{v_j}$ such that $d_{v_j}<\min\{d_{v_0},\ldots,d_{v_{j-1}}\}$.
 Observe that $d_0=d$ and $d_{\ell'}=i$.
 
-  By \propref{ancestor-boundary}, for every $j\in\{1,\ldots,d\}$, there is a set $X_{j}$, of at most 3 integers $\alpha < j$ such that such that every vertex $q\in V(F_{a_j})$ is in $V(Y_{a_{\alpha}})$ for some $\alpha\in X_j$.
+  By \propref{ancestor-boundary}, for every $j\in\{1,\ldots,d\}$, there is a set $X_{j}$, of at most 3 integers $\alpha < j$ such that every vertex $q\in V(F_{a_j})$ is in $V(Y_{a_{\alpha}})$ for some $\alpha\in X_j$.
   Assign a distinct colour $c(j,\alpha)\in\{1,2,3\}$ for each $\alpha\in X_j$ and do this assignment for each $j\in\{1,\ldots,d\}$.
@@
   \[
-     |B_x| \le \sum_{\ell=1}^{k+1} 3^\ell = (3^{k+2}-3)/2 \eqlabel{bag-size} \enspace . \qedhere
+     \sum_{\ell=1}^{k+1} 3^\ell = (3^{k+2}-3)/2 \ge |B_x| \eqlabel{bag-size} \enspace . \qedhere
   \]
@@
 
+\note{PM}{I know the analysis in \clmref{bag-width} is not tight---the base $3$ of the exponent can be reduced.  Is it even exponential though?}
 We now have all the parts needed for the proof of \thmref{k-planar}.
@@
 
-  By \clmref{bag-size}, $\mathcal{S}$ produces a graph $H=G/\mathcal{S}$ of treewidth $(3^{k+2}-1)/2$.
+  By \clmref{bag-size}, $\mathcal{S}$ produces a graph $H=G/\mathcal{S}$ of treewidth $(3^{k+2}-5)/2$.
 By \clmref{width-of-g-plus}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+12$, so $|B_x\cap V_i'|\le (k+1)(24k + 12) = 24k^2+36k+12$.

### 2019-05-12 k-planar.tex
Added a reminder about David's suggestions

@@
     \item $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$;
-    \item $F[P_i]$ is a cycle and there exists an edge of $P_i$ whose removal produces a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$; or 
-    \item $F[P_i]$ is a path or cycle in $F_{r_K}$.
+    \item $F[P_i]$ is a cycle and there exists an edge of $F[P_i]$ whose removal produces a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$; or 
+    \item $F[P_i]\subseteq F_{r_K}$.
   \end{compactenum}
@@
 
+\noindent
+\framebox{
+\begin{minipage}{\textwidth}
+  \note{DW}{%
+    I think we need to define layered treewidth, and
+    give the result of Dujmovic-Eppstein-Wood that (g,k)-planar graphs
+    have O(gk)  layered treewidth. This is the best previous structural
+    description of k-planar graphs. State that layered partitions are
+    richer than  layered treewidth, and lead to O(1) bounds on non-rep
+    chromatic number and queue-number, instead of O(log n) bounds via
+    layered treewidth.
+
+    I also think we should give explicit upper bounds on non-rep chromatic
+    number for k-planar graphs, and also for queue-number (especially if
+    they improve the bound in the FOCS / J.ACM paper).
+
+    I think we should think about generalising the new theorem for
+    (g,k)-planar graphs.
+
+    We should add corollaries for d-map graphs, since they are $d^2$-planar.
+    }
+\end{minipage}
+}
 \section{$1$-Planar Graphs}

### 2019-05-13 k-planar.tex
Rewording
Added a root-root vertex to $K$.

@@
 Layered $H$-partitions of small layered width in which $H$ has some additional property are a useful tool for a number of graph problems.
-\citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number \cite{X}, track number \cite{X}, layered treewidth \cite{X}, etc \cite{X}, etc \cite{X}, etc \cite{X}.
+\citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number \cite{X}, track number \cite{X}, treewidth, layered treewidth \cite{X}, etc \cite{X}, etc \cite{X}, etc \cite{X}.
 \citet{dujmovic.esperet.ea:planar} add non-repetitive chromatic number to this list.
@@
 
-A \emph{PS-tree}  $K=K(\Delta, T, r)$ is a 3-ary tree, such that each node $x\in V(K)$ is associated with several object (refer to \figref{K-node}):
+A \emph{PS-tree}  $K=K(\Delta, T, r)$ is a 3-ary tree, such that each non-root node $x\in V(K)$ is associated with several object (refer to \figref{K-node}):
 
@@
 
-The nodes of the PS-tree $K$ satisfy several conditions:
+The root node $r^0_K$ of $K$ is exceptional.
+We define $F_{r^0_K}$ to be the single-vertex graph containing $r$, we define $\overline{Y}_{r^0_K}$ to be the star formed by $r$ and its three neighbours in $\Delta$, and (as usual) $Y_{r^0_K}=\overline{Y}_{r^0_K}-F_{r^0_K}$.
+The root node $r^0_K$ has exactly one child $r_K$, that we call the \emph{sub-root} of $K$, for which $F_{r_K}$ is the 3-cycle formed by the neighbours of $r$ in $\Delta$.
+
+The non-root nodes of the PS-tree $K$ (including the sub-root $r_K$) satisfy several conditions:
 \begin{enumerate}[(PR1)]
-  \item For the root $r_K$ of $K$, $F_{r_K}$ is the 3-cycle in $\Delta$ formed by the three vertices adjacent to $r$.
+  \item \proplabel{children} For each non-root node $x\in V(K)$ such that $M_x$ has $f\le 3$ wedges $F_{x,1},\ldots,F_{x,f}$, the node $x$ has exactly $f$ children $y_1,\ldots,y_f$ in $K$ and $F_{y_i} = F_{x,i}$ for each $i\in\{1,\ldots,f\}$.
 
-  \item \proplabel{children} For each node $x\in V(K)$ such that $M_x$ has $f\le 3$ wedges $F_{x,1},\ldots,F_{x,f}$, the node $x$ has exactly $f$ children $y_1,\ldots,y_f$ in $K$ and $F_{y_i} = F_{x,i}$ for each $i\in\{1,\ldots,f\}$.
+  \item \proplabel{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$.
 
-  \item \proplabel{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, either:
-  \begin{compactenum}
-    \item $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$;
-    \item $F[P_i]$ is a cycle and there exists an edge of $F[P_i]$ whose removal produces a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$; or 
-    \item $F[P_i]\subseteq F_{r_K}$.
-  \end{compactenum}
-  \item \proplabel{partition} $\{Y_x : x\in V(K)\}$ is a partition of $V(\Delta)\setminus\{r\}\setminus V(F_{r_K})$.
+  \item \proplabel{partition} $\{Y_x : x\in V(K)\}$ is a partition of $V(\Delta)\setminus\{r\}$.
 \end{enumerate}
@@
 
-% The following property of a PS-Tree $K$ is an immediate consequence of \propref{children} that will be useful for us: \note{PM}{Do we still use this?}
-% \begin{compactenum}[(PR5)]
-%   \item \proplabel{ancestors} If $vw\in E(\Delta)$, $v\in V(N_x-F_x)$ and $w\in V(N_y-N_y)$ for some nodes $x,y\in V(K)$, then one of $x$ or $y$ is a $K$-ancestor of the other.
-% \end{compactenum}
+\subsubsection{Layered $H$-Partitions from PS-Trees}
+\seclabel{k-to-h}
+
+Before continuing, it is worth pointing out how an $H$-partition of $\Delta-\{r\}$ can be obtained from a PS-tree $K=K(\Delta,T,r)$.
+(Note: There is nothing new in this section.
+This is just an alternative presentation of material from \citet{dujmovic.joret.ea:planar}.) Given $K=K(\Delta,T,r)$, we obtain a layered $H$-partition $(\mathcal{L},\mathcal{P})$ as follows.
+
+We use the layering $\mathcal{L}=\{V_0,V_1,\ldots,V_h\}$ where $V_i=\{ v\in V(\Delta)\setminus\{r\}: \dist_{\Delta}(r,v)=i\}$ and $h=\max\{\dist_\Delta(r,v):v\in V(\Delta)\}$.
+The fact that $T$ is a BFS tree ensures that $\mathcal{L}$ is, indeed, a layering of $\Delta-\{r\}$.
+We use the vertex partition $\mathcal{P}=\{ S_x:=Y_x : x\in V(K)\}$.
+\propref{partition} ensures that $\mathcal{P}$ is, indeed a partition of $V(\Delta)-\{r\}$.
+
+The resulting layered $H$-partition $(\mathcal{L},\mathcal{P})$ has layered width at most 3 because, for each $x\in V(K)$, $S_x$ is made up of vertices from at most 3 vertical paths in $T$.
+These are either the three vertices of $Y_{r_k}$ (if $x=r_K$) or the vertices of the open tripod $Y_x$ whose vertices are included in three vertical paths $Q_1,Q_2,Q_3$ that form the legs of the closed tripod $\overline{Y}_x$.
+Therefore $|S_x\cap V_i|\le 3$.
+
+To show that the quotient graph $H=\Delta/\mathcal{P}$ has small treewidth, we produce a tree-decomposition $\mathcal{K}=\{B_x: x\in V(K)\}$ of $H$ using the tree $K$.
+For each $x\in V(K)$, the subtree $K[x]=K[\{y:x\in B_y\}]$ contains
+$x$ as well as every node $y$ such that $V(Y_x)\cap V(F_y) \neq\emptyset$.
+It is straightforward to verify that $K[x]$ is connected for each $x\in V(K)$.
+The fact that $F_x$ is a cycle in $\Delta$ and \propref{ancestor-boundary} ensures that, if $xy\in E(H)$, then one of $x$ or $y$ is a $K$-ancestor of the other.
+\propref{ancestor-boundary} also implies that, for any node $y\in V(K)$, $|B_y|\le 4$, since $B_y$ contains $y$ and at most 3 $K$-ancestors of $y$.
+
+Therefore $(\mathcal{L},\mathcal{P})$ is a layered $H$-partition of $\Delta-\{r\}$ with layered-width 3 in which $H$ has treewidth 3.
+Furthermore, for each $P\in\mathcal{P}$, $\Delta[P]$ is connected.
+Since $\Delta$ is planar, this implies that $H$ is planar.\footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
+Repeatedly applying this fact implies that $\Delta/\mathcal{P}$ is planar provided that $\Delta[P]$ is connected for every $P\in\mathcal{P}$.}
 
@@
 We will assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$.
-This assumption is not critical since it can be enforced by a slight deformation of the edges of $G$.
+This assumption is not critical since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
 Consider the following plane graphs:
@@
   \item Add a dummy vertex at each crossing in $G$ to obtain a plane graph and then add edges to this graph to obtain a triangulation $G^+$.
+
   \item Add a complete graph $K_4$ that contains $G^{+}$ in one of its inner faces, $F_0$, and then add edges to the resulting graph to obtain a triangulation $G^{++}$.
 \end{compactenum}
-Let $r$ be the vertex of the $K_4$ not incident on $F_0$ and let $T$ be a BFS spanning tree of $G^{++}$ rooted at $r$. Observe that, for any edge $vw\in E(G)$, 
-\[  \dist_{G^+}(v,w) \le k+1 \enspace . \]
-%  and therefore, for any two vertices $v,w\in V(G)$,
-% \begin{equation} 
-%   \dist_{G^+}(v,w) \le (k+1)\cdot \dist_G(v,w) \eqlabel{distance-preserving}
-% \end{equation}
+Observe that, for any edge $vw\in E(G)$, 
+\begin{equation}  
+  \dist_{G^+}(v,w) \le k+1 \enspace , 
+\end{equation}
+since (the curve) $vw$ contains a sequence $W_{vw}$ of at most $k+2$ vertices of $G^+$ that form a walk from $v$ to $w$ in $G^+$.
+
+Let $r$ be the vertex of the $K_4$ not incident on $F_0$ and let $T$ be a BFS spanning tree of $G^{++}$ rooted at $r$. 
 For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{T}(r,v)=i\}$ and observe that $\{V_0,V_1,\ldots,V_{h}\}$ is a layering of $G^+$ (but not of $G$).
 
-Let $K=K(G^{++},T,r)$ be a PS-tree and observe that the root $r_K$ of $K$ has $F_{r_K}=F_0$.  This implies:
+Let $K=K(G^{++},T,r)$ be a PS-tree and observe that the sub-root $r_K$ of $K$ has $F_{r_K}=F_0$.  This implies:
 \begin{compactenum}[(PR1)]\setcounter{enumi}{4}
-  \item \proplabel{all-in}  $V(G^+)\subseteq V(N_{r_K}-F_{r_K})$.
+  \item \proplabel{all-in}  $G^+\subseteq N_{r_K}-F_{r_K}$.
 \end{compactenum}
@@
 For each vertex $z\in V(G^+)$, let $P_z=\{a\in V(K): z\in V(N_a-F_a)\}$.  
-By \propref{all-in}, $P_{z}$ contains the root $r_K$ of $K$.
+By \propref{all-in}, $P_{z}$ contains the sub-root $r_K$ of $K$.
 By \propref{children}, $K[P_z]$ is a vertical path (whose lower endpoint is the unique node $x\in V(K)$ such that $z\in Y_x$).
@@
 
-\note{PM}{I know the analysis in \clmref{bag-width} is not tight---the base $3$ of the exponent can be reduced.  Is it even exponential though?}
-We now have all the parts needed for the proof of \thmref{k-planar}.
+\note{PM}{I know the analysis in \clmref{bag-size} is not tight---the base $3$ of the exponent can be reduced.  Is it even exponential though?}
+We now have all the parts needed to complete the proof of \thmref{k-planar}.
 
@@
 
-Let $G$ be an edge-maximal 1-plane multigraph.
-Here, edge-maximal should be taken to mean that, if any two vertices $u$ and $v$ appear on a common face\footnote{The \emph{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
-We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $uv\in E(G)$ that is contained in the boundary of $F$.
+In this section, $G$ is a 1-plane graph. We will construct a   To construct a layered $H$-decomposition of $G$. First, we augment $G$ as follows:
+\begin{enumerate}
+  \item Add edges to $G$ to obtain a multigraph $G^+$ that is edge-maximal in the following sense:  if any two vertices $v,w\in V(G^+)$ appear on a common face\footnote{The \emph{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
+We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$ of $G^+$, then there is an edge $vw\in E(G^+)$ that is contained in the boundary of $F$.
 
-A \emph{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.
+  \item Add a complete graph $K_4$ that contains $G^{+}$ in one of its inner faces, $F_0$, and then add edges to the resulting graph to obtain a triangulation $G^{++}$.
+\end{enumerate}
+
+Note that, in contrast to \secref{k-planar}, neither $G^+$ nor $G^{++}$ is necessarily a plane graph; crossings that occur in $G$ also occur in $G^+$ and $G^{++}$.
+
+A \emph{kite} in $G^+$ is the subgraph $K=G^+[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G^+)$.
 It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
@@
 The cycle $vxwy$ is called the \emph{sail} of $K$.
-It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$.
+It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G^+$.
 Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
-Observe that any spar of $K$ is incident on exactly four \emph{kite faces} of $G$, each of which has three edges and two vertices of $G$ on its boundary.
+Observe that any spar of $K$ is incident on exactly four \emph{kite faces} of $G^+$, each of which has three edges and two vertices of $G$ on its boundary.
 
-The 1-plane graph $G$ has a plane triangulation $G'$ as a subgraph that can be obtained by removing one spar from each kite in $G$.
-Observe that, for any spar $vw\in E(G)\setminus E(G')$ that crosses $xy\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
-It follows that $\dist_{G'}(x,y)\le 2\cdot\dist_G(x,y)$ for any pair of vertices $x,y\in V(G)$.
+The 1-plane graph $G^+$ has a plane triangulation $G^-$ as a subgraph that can be obtained by removing one spar from each kite in $G^+$.
+Observe that, for any spar $vw\in E(G)\setminus E(G')$ that crosses $xy\in E(G^-)$, $G^-$ contains the path $vxw$ (and $vyw$).
+It follows that $\dist_{G^-}(x,y)\le 2\cdot\dist_{G^+}(x,y)$ for any pair of vertices $x,y\in V(G)$.
 

### 2019-05-13 k-planar.tex
kpfonts in figures

@@
 Layered $H$-partitions of small layered width in which $H$ has some additional property are a useful tool for a number of graph problems.
-\citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number \cite{X}, track number \cite{X}, treewidth, layered treewidth \cite{X}, etc \cite{X}, etc \cite{X}, etc \cite{X}.
+\citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number, track number, layered treewidth, and 3-D grid drawing volume.
 \citet{dujmovic.esperet.ea:planar} add non-repetitive chromatic number to this list.
@@
 
-The remainder of the paper is organized as follows: In \secref{1-planar} we prove \thmref{1-planar}.
-In \secref{k-planar} we prove \thmref{k-planar}.
-In \secref{consequences}, we discuss the implications of Theorems~\ref{thm:k-planar} and \ref{thm:1-planar} for other graph parameters of $k$-planar graphs.
+Prior to this work, the strongest structural characterization of $k$-planar graphs was in terms of layered treewidth, which we now define.
+A \emph{tree-decomposition} $\mathcal{K}$ of a graph $G$ is consists of a tree $K$ and a collection $\mathcal{K}=\{B_x:x\in V(K)\}$ of subsets of $V(G)$ indexed by nodes of $K$ such that (i)~for every $vw\in E(G)$, there exists some $x\in V(K)$ with that $v,w\in B_x$; and (ii)~for every $v\in V(G)$, $K[x] = K[\{y: x\in B_y\}]$ is connected.
+A \emph{layered tree-decomposition} $(\mathcal{L},\mathcal{K})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{K}$ of $G$.
+The layered-width of $(\mathcal{L},\mathcal{K})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{K}\}$.
+The \emph{layered treewidth} of $G$ is the minimum layered-width of any layered tree-decomposition of $G$.
+
+\citet{dujmovic.eppstein.ea:structure} show that $k$-planar graphs have layered treewidth $O(k+1)$.
+This already implies that several of the graph parameters mentioned above, including queue-number and non-repetitive chromatic number are upper-bounded by $O(k\log n)$ for $n$-vertex $k$-planar graphs.
+\thmref{k-planar} is a significant strengthening of this characterization.
+It implies that $k$-planar graphs have layered-treewidth bounded by a function of $k$ and gives upper bounds on queue-number and non-repetitive chromatic number of $n$-vertex $k$-planar graphs that are independent of $n$.
+
+The remainder of the paper is organized as follows: In \secref{1-planar} we prove \thmref{1-planar}.  In \secref{k-planar} we prove \thmref{k-planar}.
 
@@
 
-\subsubsection{Layered $H$-Partitions from PS-Trees}
-\seclabel{k-to-h}
-
-Before continuing, it is worth pointing out how an $H$-partition of $\Delta-\{r\}$ can be obtained from a PS-tree $K=K(\Delta,T,r)$.
+% \subsubsection{Layered $H$-Partitions from PS-Trees}
+% \seclabel{k-to-h}
+% 
+% Before continuing, it is worth pointing out how an $H$-partition of $\Delta-\{r\}$ can be obtained from a PS-tree $K=K(\Delta,T,r)$.
 (Note: There is nothing new in this section.
 This is just an alternative presentation of material from \citet{dujmovic.joret.ea:planar}.) Given $K=K(\Delta,T,r)$, we obtain a layered $H$-partition $(\mathcal{L},\mathcal{P})$ as follows.
-
-We use the layering $\mathcal{L}=\{V_0,V_1,\ldots,V_h\}$ where $V_i=\{ v\in V(\Delta)\setminus\{r\}: \dist_{\Delta}(r,v)=i\}$ and $h=\max\{\dist_\Delta(r,v):v\in V(\Delta)\}$.
+% 
+% We use the layering $\mathcal{L}=\{V_0,V_1,\ldots,V_h\}$ where $V_i=\{ v\in V(\Delta)\setminus\{r\}: \dist_{\Delta}(r,v)=i\}$ and $h=\max\{\dist_\Delta(r,v):v\in V(\Delta)\}$.
 The fact that $T$ is a BFS tree ensures that $\mathcal{L}$ is, indeed, a layering of $\Delta-\{r\}$.
-We use the vertex partition $\mathcal{P}=\{ S_x:=Y_x : x\in V(K)\}$.
+We use the vertex partition $\mathcal{P}=\{ S_x:=V(Y_x) : x\in V(K)\}$.
 \propref{partition} ensures that $\mathcal{P}$ is, indeed a partition of $V(\Delta)-\{r\}$.
-
-The resulting layered $H$-partition $(\mathcal{L},\mathcal{P})$ has layered width at most 3 because, for each $x\in V(K)$, $S_x$ is made up of vertices from at most 3 vertical paths in $T$.
+% 
+% The resulting layered $H$-partition $(\mathcal{L},\mathcal{P})$ has layered width at most 3 because, for each $x\in V(K)$, $S_x$ is made up of vertices from at most 3 vertical paths in $T$.
 These are either the three vertices of $Y_{r_k}$ (if $x=r_K$) or the vertices of the open tripod $Y_x$ whose vertices are included in three vertical paths $Q_1,Q_2,Q_3$ that form the legs of the closed tripod $\overline{Y}_x$.
 Therefore $|S_x\cap V_i|\le 3$.
-
-To show that the quotient graph $H=\Delta/\mathcal{P}$ has small treewidth, we produce a tree-decomposition $\mathcal{K}=\{B_x: x\in V(K)\}$ of $H$ using the tree $K$.
+% 
+% To show that the quotient graph $H=\Delta/\mathcal{P}$ has small treewidth, we produce a tree-decomposition $\mathcal{K}=\{B_x: x\in V(K)\}$ of $H$ using the tree $K$.
 For each $x\in V(K)$, the subtree $K[x]=K[\{y:x\in B_y\}]$ contains
-$x$ as well as every node $y$ such that $V(Y_x)\cap V(F_y) \neq\emptyset$.
-It is straightforward to verify that $K[x]$ is connected for each $x\in V(K)$.
+% $x$ as well as every node $y$ such that $V(Y_x)\cap V(F_y) \neq\emptyset$.
+% It is straightforward to verify that $K[x]$ is connected for each $x\in V(K)$.
 The fact that $F_x$ is a cycle in $\Delta$ and \propref{ancestor-boundary} ensures that, if $xy\in E(H)$, then one of $x$ or $y$ is a $K$-ancestor of the other.
 \propref{ancestor-boundary} also implies that, for any node $y\in V(K)$, $|B_y|\le 4$, since $B_y$ contains $y$ and at most 3 $K$-ancestors of $y$.
-
-Therefore $(\mathcal{L},\mathcal{P})$ is a layered $H$-partition of $\Delta-\{r\}$ with layered-width 3 in which $H$ has treewidth 3.
+% 
+% Therefore $(\mathcal{L},\mathcal{P})$ is a layered $H$-partition of $\Delta-\{r\}$ with layered-width 3 in which $H$ has treewidth 3.
 Furthermore, for each $P\in\mathcal{P}$, $\Delta[P]$ is connected.
@@
   Consider the graph $A$ that contains the edges and vertices of $F_{a}$ for all $K$-ancestors $a$ of $x$.
-See \figref{A}.
+See \figref{A}(a).
 It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0=r_K$ is the root of $K$.
@@
   \begin{figure}
+
     \begin{center}
-        \includegraphics{figs/A-1}
+      \begin{tabular}{cc}
+        \includegraphics{figs/A-1} &
+        \includegraphics{figs/A-3} \\ (a) & (b)
+      \end{tabular}
     \end{center}
-    \caption{The graph $A$.}
+    \caption{(a)~the graph $A$ and (b)~the walks $W_{vw}$ and $W_{ww'}$ that contain the walk $w=v_0,v_1\ldots,v_{\ell-1},v_\ell=w^*$.}
     \figlabel{A}
@@
 
+  Refer to \figref{A}(b)
   If $B_x$ contains some $K$-ancestor $a_i$, $i\in\{0,\ldots,d-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_{a_i}$.
@@
 
-  Summarizing, we have a walk $w'=v_0,v_1\ldots,v_{\ell-1},v_\ell=w^*$ in $G^+$ such that
+  Summarizing, we have a walk $w=v_0,v_1\ldots,v_{\ell-1},v_\ell=w^*$ in $G^+$ such that
   \begin{compactenum}[(i)]
@@
   \note{DW}{%
-    I think we need to define layered treewidth, and
-    give the result of Dujmovic-Eppstein-Wood that (g,k)-planar graphs
-    have O(gk)  layered treewidth. This is the best previous structural
-    description of k-planar graphs. State that layered partitions are
-    richer than  layered treewidth, and lead to O(1) bounds on non-rep
-    chromatic number and queue-number, instead of O(log n) bounds via
-    layered treewidth.
-
+    % I think we need to define layered treewidth, and
+    % give the result of Dujmovic-Eppstein-Wood that (g,k)-planar graphs
+    % have O(gk)  layered treewidth. This is the best previous structural
+    % description of k-planar graphs. State that layered partitions are
+    % richer than  layered treewidth, and lead to O(1) bounds on non-rep
+    % chromatic number and queue-number, instead of O(log n) bounds via
+    % layered treewidth.
     I also think we should give explicit upper bounds on non-rep chromatic
@@
     }
+
+    \note{PM}{
+    A couple of other things to look into: 
+    \begin{compactenum}
+      \item $p$-centered colourings
+      \item low treewidth colourings
+    \end{compactenum}
+    }
 \end{minipage}
@@
 
-In this section, $G$ is a 1-plane graph. We will construct a   To construct a layered $H$-decomposition of $G$. First, we augment $G$ as follows:
-\begin{enumerate}
-  \item Add edges to $G$ to obtain a multigraph $G^+$ that is edge-maximal in the following sense:  if any two vertices $v,w\in V(G^+)$ appear on a common face\footnote{The \emph{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
-We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$ of $G^+$, then there is an edge $vw\in E(G^+)$ that is contained in the boundary of $F$.
+Let $G$ be an edge-maximal 1-plane multigraph.
+Here, edge-maximal should be taken to mean that, if any two vertices $v$ and $w$ appear on a common face\footnote{The \emph{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
+We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.
 
-  \item Add a complete graph $K_4$ that contains $G^{+}$ in one of its inner faces, $F_0$, and then add edges to the resulting graph to obtain a triangulation $G^{++}$.
-\end{enumerate}
-
-Note that, in contrast to \secref{k-planar}, neither $G^+$ nor $G^{++}$ is necessarily a plane graph; crossings that occur in $G$ also occur in $G^+$ and $G^{++}$.
-
-A \emph{kite} in $G^+$ is the subgraph $K=G^+[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G^+)$.
+A \emph{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.
 It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
@@
 The cycle $vxwy$ is called the \emph{sail} of $K$.
-It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G^+$.
+It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$.
 Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
-Observe that any spar of $K$ is incident on exactly four \emph{kite faces} of $G^+$, each of which has three edges and two vertices of $G$ on its boundary.
+Observe that any spar of $K$ is incident on exactly four \emph{kite faces} of $G$, each of which has three edges and two vertices of $G$ on its boundary.
 
-The 1-plane graph $G^+$ has a plane triangulation $G^-$ as a subgraph that can be obtained by removing one spar from each kite in $G^+$.
-Observe that, for any spar $vw\in E(G)\setminus E(G')$ that crosses $xy\in E(G^-)$, $G^-$ contains the path $vxw$ (and $vyw$).
-It follows that $\dist_{G^-}(x,y)\le 2\cdot\dist_{G^+}(x,y)$ for any pair of vertices $x,y\in V(G)$.
+The 1-plane graph $G$ has a plane triangulation $G'$ as a subgraph that can be obtained by removing one spar from each kite in $G$.
+Observe that, for any spar $xy\in E(G-G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
+It follows that
+\begin{equation}
+  \dist_{G'}(v,w)\le 2 \enspace . \eqlabel{spar-distance}
+\end{equation}
 
-Our proof of \thmref{1-planar} follows quickly from the following technical lemma, which is an extension of Lemma~X in \citet{dujmovic.joret.ea:planar}.
+Our proof of \thmref{1-planar} follows quickly from the following technical lemma, which is an extension of the equivalent result for plane graphs \cite{dujmovic.joret.ea:planar}.
 \begin{lem}\lemlabel{induction} The setup:
@@
     \item Let $T$ be a BFS tree of $G'$ rooted at some vertex $r$.
-    \item For each integer $i\ge 0$, let $V_i=\{v\in V(G):\dist_T(r,v)=i\}$. 
-    \item Let $C$ be a cycle in $G'$ with $r$ in the exterior of $C$ and such that
+    \item For each integer $j\ge 0$, let $V_j=\{v\in V(G):\dist_T(r,v)=j\}$. 
+    \item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that
     \begin{compactenum} 
-      \item No edge of $C$ is crossed by any edge of $G$; and
-      \item $V(C)$ can be partitioned into $P_1,\ldots,P_k$, $k\le 3$ such that for each $i\in\{1,\ldots,k\}$,
+      \item No edge of $F$ is crossed by any edge of $G$; and
+      \item $V(F)$ can be partitioned into $P_1,\ldots,P_k$, $k\le 3$ such that for each $i\in\{1,\ldots,k\}$,
       \begin{compactenum}
-        \item $C[P_i]$ is a path; and
-        \item $|P_i\cap V_j| \le 15$ for all integers $j\ge 0$.
+        \item $F[P_i]$ is a path; and
+        \item $|V(P_i)\cap V_j| \le 15$ for all integers $j\ge 0$.
       \end{compactenum}
     \end{compactenum}
-    \item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $C$ or the interior of $C$.
+    \item Let $N^+$ and $N^-$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
   \end{compactenum}
-  Then $N$ has an $H$-partition $(B_x : x\in V(H))$ such that
+  Then $N$ has an $H$-partition $\{S_x : x\in V(H)\}$ such that
   \begin{compactenum}
     \item $H$ is planar;
-    \item for all integers $j\ge 0$, and all $x\in V(H)$, $|B_x\cap V_j|\le 15$; 
-    \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=B_{x_i}$; and
+    \item for all integers $j\ge 0$, and all $x\in V(H)$, $|S_x\cap V_j|\le 15$; 
+    \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; and
     \item $H$ has a tree decomposition whose largest bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
@@
 \begin{proof}
+  This proof very similar to the proof of Lemma~14 in \citet{dujmovic.joret.ea:planar}.
+Rather than duplicate every detail of that proof here, we focus on the differences and refer the reader to the original proof for the remaining details.
+
   The proof is by induction on the number of vertices of $N$.
@@
 If $k=3$, set $R_i=P_i$ for each $i\in\{1,2,3\}$.
-Otherwise, as before, split $P_1,\ldots,P_k$ to partition $V(C)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $C[R_i]$ is a path and each $R_i$ contains vertices from at most one of $P_1,\ldots,P_k$.
+Otherwise, as in \citet{dujmovic.joret.ea:planar}, split $P_1,\ldots,P_k$ to partition $V(C)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a path and each $R_i$ contains vertices from at most one of $P_1,\ldots,P_k$.
 
-  For each $i\in\{1,2,3\}$ and $v\in R_i$, assign the colour $i$ to $v$.
-  For each $v\in V(N')$, consider the path $P_v$ that consists of the subpath of the $v$ to $r$ path in $T$ that stops at the first vertex $v'\in V(C)$. Assign $v$ to have the same colour as $v'$.
+  Next, as in \citet{dujmovic.joret.ea:planar}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $V(N'-F)$.
 
-  Sperner's Lemma ensures that $G'$ contains a triangular face $\tau=v_1v_2v_3$ whose three vertices are assigned different colours.
-For each $i\in\{1,2,3\}$, let $Q_i=P_{v_i}$ be the path in $T$ form $v_i$ to $v_i'$.
-Let $Y$ denote the subgraph of $N'$ consisting of vertices and edges $Q_1$, $Q_2$, $Q_3$, and $\tau$.
-Let $Y^+$ denote the subgraph of $N$ consisting of the vertices and edges of $Y$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $Y$.
+  Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
+Let $\overline{Y}^+$ denote the subgraph of $N$ consisting of the vertices and edges of $\overline{Y}$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $\overline{Y}$.
 
   We claim that, for each integer $i\ge 0$, $|V(Y^+)\cap V_i|\le 15$.
-To see this, first observe that $Y$ contains at most 3 vertices of $V_i$.
-If a vertex $x\in V(Y^+)\setminus V(Y)$ is contained in $V_i$, then this is because $Y$ contains an edge $vw$ with $v\in V_{i'}$ and $w\in\{V_{i'+1}$ for some $i'\in\{i-1,i\}$ and $G$ contains an edge $xy$ that crosses $vw$.
-(Recall that $vx,xw\in E(G')$, so $\dist_{G'}(w,r)-1\le\dist_{G'}(x,r)\le\dist_{G'}(v,r)+1$.)  The graph $Y$ contains at most 6 such edges, each of which accounts for at most 2 additional vertices of $V_i$.
-Therefore, in total, $|Y^+\cap V_i|\le 15$.
+First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most 3 vertices of $V_i$, each incident on at most two edges of $\overline{Y}$.
+By \eqref{spar-distance}, any vertex $x\in V(\overline{Y}^+-\overline{Y})\cap V_i$, is incident on an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $V_i$.
+These at most six edges have at most 12 endpoints.
+Therefore $|V(\overline{Y}^+-\overline{Y})\cap V_i|\le 6\times 2=12$, so $|\overline{Y}^+\cap V_i|\le 12+3=15$.
 
-  Finally, let $S$ and $S^+$ denote the subgraph of $G$ containing the edges and vertices of $Y$, respectively $Y^+$, and the edges and vertices of $C$.
-The graph $S^+$ has some number of bounded faces, all contained in the interior of $C$.
-Some of the bounded faces of $S$ are kite faces.
-Call the non-kite bounded faces $F_1,\ldots,F_m$ and let $C_1,\ldots,C_m$ denote their boundaries.
-We claim that, for each $i\in\{1,\ldots,m\}$, if some portion of $C_i$ is contained in an edge $vw\in E(G)$ then $vw$ is not crossed by any edge of $G$.
-To see this, there are three cases to consider:
+  Finally, let $M$ and $M^+$ denote the subgraph of $G$ containing the edges and vertices of $Y$, respectively $Y^+$, and the edges and vertices of $F$.
+The graph $M^+$ has some number of bounded faces, all contained in the interior of $F$.
+Some of the bounded faces of $M^+$ are kite faces.
+Let $F_1,\ldots,F_m$ be the non-kite bounded faces of $M^+$.
+
+  We claim that, for each $i\in\{1,\ldots,m\}$, the boundary of $F_i$ is a cycle in $G'$.
+Otherwise, some edge $vw$ contributes to the boundary of $F_i$ but is crossed by an edge $xy\in E(G)$.
+Then, $vw\not\in E(F)$ since no edge of $F$ is crossed by any edge of $G$.
+Therefore $vw\in E(\overline{Y}^+)$ so $xy\in E(Y^+)$.
+But then the only faces of $M^+$ incident to $vw$ are kite faces.
+In particular $vw$ can not be incident the non-kite face $F_i$.
+
+  Note that $M$ is a Mercedes graph and that each of the faces $F_1,\ldots,F_m$ is contained in a single wedge of $M$.
+Let $Y^+ = \overline{Y}^+-F$.
+Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.
+
+  Finally, the subgraph $N_i$ of $G$ consisting of the edges and vertices of $G$ contained in $F_i$ or its interior does not contain one of the three vertices of $\tau$.
+Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(C_i)$ to obtain the desired $H$-partition and tree decomposition of $N_i$.
+
+  The remainder of the proof finishes in the same way as the proof of Lemma~14 in \cite{dujmovic.joret.ea:planar}.
+$P_1,\ldots,P_k$, and $V(Y^+-F)$ become elements of the $H$-partition.
+Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
+The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{dujmovic.joret.ea:planar}.
+
+  The planarity of $H$ comes from two properties:
   \begin{enumerate}
-    \item $vw\in E(C)$. By assumption, $vw$ is not crossed by any edge of $G$.
-    \item $vw\in E(Y)$. In this case, the kite containing $vw$ is in $Y^+$, so $vw$ is only incident to kite faces.
-    \item $vw\in E(Y^+)\setminus E(Y)$.
-In this case, either $vw$ is a sail edge in which case it is not crossed by definition, or $vw$ is a spar that was added to $Y^+$ because $vw$ crosses some edge $xy\in E(Y)$.
-In this latter case, $vw$ is only incident to kite faces.
+    \item $G/\mathcal{P}$ and $G'\mathcal{P}$ are isomorphic.
+This follows immediately from the fact that, for each edge $xy\in E(G-G')$, $x,y\in S_a$ for some $a\in V(H)$.
+Indeed, $S_a=Y^+_a$ where $\overline{Y}_a$ contains the edge $vw\in E(G')$ that crosses $xy$.
+
+    \item $G[\overline{Y}^+-F]$ is connected.
+To see why this is so, first observe that $\overline{Y}-F$ is connected, and then observe that every vertex of $\overline{Y^+}$ is either a vertex of $\overline{Y}$ or adjacent to a vertex of $\overline{Y}$.
   \end{enumerate}
-  Therefore each $C_i$ is a cycle in $G^+$ consisting entirely of uncrossed edges.
-The vertices of $C_i$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$ and $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$.
-Furthermore $C_i[P_j']$ is a path for each $j\in\{1,\ldots,3\}$.
-Finally, the subgraph $N_i$ of $G$ consisting of the edges and vertices of $G$ contained in $C_i$ or its interior does not contain one of the three vertices of $\tau$.
-Therefore, we can apply induction using the cycle $C_i$ and the partition $P_1',P_2',P_3'$ of $V(C_i)$ to obtain the desired $H$-partition and tree decomposition of $N_i$.
-
-  The remainder of the proof finishes in the same way as the proof of Lemma~X in \cite{dujmovic.joret.ea:planar}.
-$P_1,\ldots,P_k$, and $V(Y^+)\setminus V(C)$ become elements of the $H$-partition.
-Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(Y^+)\setminus V(C)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
-The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~X in blah.
-
-  The planarity of $H$ comes from the fact that if two edges $vw$ and $xy$ cross, then they end up in the same bag of the $H$-partition.
-This means that $H$ is actually obtained by contracting connected sets of vertices in the planar graph $G'$.
-It is well know that any graph obtained by contracting connected sets of vertices in a planar graph is also planar.
+  Since $G'$ is planar, this implies that $H=G'/\mathcal{P}$ is planar.\footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
+Repeatedly applying this fact implies that $\Delta/\mathcal{P}$ is planar provided that $\Delta[P]$ is connected for every $P\in\mathcal{P}$.}
 \end{proof}
 
-\begin{proof}[Proof of \thmref{1-plane}]
-  The same as before except that the layered width 15 from \lemref{induction} becomes 30 because \lemref{induction} uses a layering of $G'$ and distances in $G'$ can be a factor of 2 larger than in $G$.
+Finally, we can complete the proof of \thmref{1-planar}.
+
+\begin{proof}[Proof of \thmref{1-planar}]
+  Using \lemref{induction}, the proof of \thmref{1-planar} is now straightforward.
+Given a 1-plane graph $G$, add edges to make it edge maximal so that it has an outer face $F=v_1v_2v_3$.
+Next, add a vertex $r$ adjacent to $v_1$, $v_2$, and $v_3$ to obtain an edge-maximal 1-plane graph $\overline{G}$ with one vertex $r$ of degree 3 on its outer face.
+
+  Let $G'$ be the plane graph obtained by removing one spar from each kite of $\overline{G}$ and let $T$ be a BFS tree of $G'$ rooted at $r$.
+Now apply \lemref{induction} with $G=\overline{G}$, $G'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
+This gives an $H$-partition $\{S_x:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ is planar and has treewidth at most 3.
+
+  Use the layering $\mathcal{L}=\{V_0',\ldots,V_h'\}$ where $V_i'=V_{2i}\cup V_{2i+1}$ for each $i\in\{0,\ldots,h\}$.
+That this is a layering of $G$ follows from \eqref{spar-distance}.
+Since $|V_j\cap S_x|\le 15$ for every integer $i\ge 0$, $|V_i'\cap S_x|\le 30$ for every integer $i\ge 0$ and $x\in V(H)$.
 \end{proof}
 
-\section{Consequences}
-\seclabel{consequences}
+% \section{Consequences}
+% \seclabel{consequences}
 

### 2019-05-13 k-planar.tex
Last undefined reference
Small updates

@@
 \begin{equation}  
-  \dist_{G^+}(v,w) \le k+1 \enspace , 
+  \dist_{G^+}(v,w) \le k+1 \enspace ,  \eqlabel{distance-preserving}
 \end{equation}
@@
   \end{compactenum}
-  Then $N$ has an $H$-partition $\{S_x : x\in V(H)\}$ such that
+  Then $N$ has an $H$-partition $\mathcal{P}=\{S_x : x\in V(H)\}$ such that
   \begin{compactenum}
@@
   The remainder of the proof finishes in the same way as the proof of Lemma~14 in \cite{dujmovic.joret.ea:planar}.
-$P_1,\ldots,P_k$, and $V(Y^+-F)$ become elements of the $H$-partition.
+$P_1,\ldots,P_k$, and $V(Y^+)$ become elements of the $H$-partition.
 Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
@@
   \begin{enumerate}
-    \item $G/\mathcal{P}$ and $G'\mathcal{P}$ are isomorphic.
+    \item $G/\mathcal{P}$ and $G'/\mathcal{P}$ are isomorphic.
 This follows immediately from the fact that, for each edge $xy\in E(G-G')$, $x,y\in S_a$ for some $a\in V(H)$.
@@
     \item $G[\overline{Y}^+-F]$ is connected.
-To see why this is so, first observe that $\overline{Y}-F$ is connected, and then observe that every vertex of $\overline{Y^+}$ is either a vertex of $\overline{Y}$ or adjacent to a vertex of $\overline{Y}$.
+To see why this is so, first observe that $\overline{Y}-F$ is connected, and then observe that every vertex of $\overline{Y}^+$ is either a vertex of $\overline{Y}$ or adjacent to a vertex of $\overline{Y}$.
   \end{enumerate}

### 2019-05-13 k-planar.tex
langle and rangle for L

@@
 \begin{proof}
-  We use the layered $H$-partition $(\mathcal{L}, \mathcal{S})$, where $\mathcal{S}=\{S_x:x\in V(K)\}$ and $\mathcal{L}=\{V_i': i=0,1,2\ldots\}$ where $V_i' = V_{(k+1)i}\cup\cdots\cup V_{(k+1)(i+1)-1}$.
+  We use the layered $H$-partition $(\mathcal{L}, \mathcal{S})$, where $\mathcal{S}=\{S_x:x\in V(K)\}$ and $\mathcal{L}=\langle V_i': i=0,1,2\ldots\rangle$ where $V_i' = V_{(k+1)i}\cup\cdots\cup V_{(k+1)(i+1)-1}$.
 \Eqref{distance-preserving} shows that $\mathcal{L}$ is indeed a layering of $G$.
@@
 
-  Use the layering $\mathcal{L}=\{V_0',\ldots,V_h'\}$ where $V_i'=V_{2i}\cup V_{2i+1}$ for each $i\in\{0,\ldots,h\}$.
+  Use the layering $\mathcal{L}=\langle V_0',\ldots,V_h'\rangle$ where $V_i'=V_{2i}\cup V_{2i+1}$ for each $i\in\{0,\ldots,h\}$.
 That this is a layering of $G$ follows from \eqref{spar-distance}.

### 2019-05-13 k-planar.tex
Small update

@@
 For any partition $\mathcal{P}=\{S_1,\ldots,S_p\}$ of $V(G)$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
-To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an $H$-partition and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.\footnote{An alternative definition of a layering of $G$ is a layered $H$-partition of $G$ where $H$ is a path.}  A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
+To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an $H$-partition and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.\footnote{An alternative definition of a layering of $G$ is an $H$-partition of $G$ where $H$ is a path.}  A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
 The \emph{layered width} of $(\mathcal{H},\mathcal{P})$ is $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.

### 2019-05-14 k-planar.tex
First crack at (g,k)-planar graphs

@@
 The latter result is new, for all $k\ge 1$.
+These results extend to $(g,k)$-planar graphs, the natural generalization of $k$-planar graphs to to genus-$g$ surfaces (rather than the genus-$0$ plane).
 \end{abstract}
@@
 In this paper we show that another family of graphs admits layered $H$-partitions of constant layered-width in which $H$ has constant treewidth: $k$-planar graphs, which we now define.
-An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve is a continuous function $f:[0,1]\to \R^2$.
+An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve in a surface $S$ is a continuous function $f:[0,1]\to S$.
 The points $f(0)$ and $f(1)$ are called the \emph{endpoints} of the curve.
-When there is no danger of misunderstanding we treat a curve $f$ as a point set $\{f(t):0\le t\le 1\}$.} with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
+When there is no danger of misunderstanding we treat a curve $f$ as a point set $\{f(t):0\le t\le 1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
 A \emph{crossing} in an embedded graph $G$ is a triple $(p,vw,xy)$ with $p\in\R^2$, $vw,xy\in E(G)$ and such that $p\in (vw\cap xy)\setminus\{v,w,x,y\}$.
@@
 
-The remainder of the paper is organized as follows: In \secref{1-planar} we prove \thmref{1-planar}.  In \secref{k-planar} we prove \thmref{k-planar}.
+There is a natural generalization of $k$-planar to genus-$g$ surfaces, that we now describe.
+A graph $G$ embedded on a surface $S$ is $(S,k)$-plane if no edge of $G$ is involved in more than $k$ crossings.
+A (not necessarily embedded) graph $G$ is $(g,k)$-planar if it is isomorphic to some $(S,k)$-plane graph $G'$ where $S$ is a surface with Euler genus $g$.
+
+\begin{thm}\thmlabel{g-k-planar}
+  Every $(g,k)$-planar graph has a layered $H$-partition of layered-width at most $\max\{16gk^2 + 24gk + 8g, 24k^2 + 36k + 12\}$ in which $H$ has treewidth at most $(3^{k+2}-3)/2$.
+
+  Every $(g,1)$-planar graph has a layered $H$-partition of layered-width at most $\max\{48g, 30\}$ in which $H$ has treewidth at most $4$.  
+\end{thm}
+
+The remainder of the paper is organized as follows: In \secref{k-planar} we prove \thmref{k-planar}.  In \secref{1-planar} we prove \thmref{1-planar}.  
+In \secref{genus-g} we prove \thmref{g-k-planar}.
 
@@
 
+\section{$(g,k)$-Planar Graphs}
+\seclabel{g-k-planar}
+
+\begin{lem} \lemlabel{make-planar}
+Let $G$ be a $(S,0)$-plane graph where $S$ is a surface with Euler genus $g$. 
+%For every BFS layering $(V_0,V_1,\dots)$ of $G$,
+%there is a set $Z\subseteq V(G)$ with at most $2g$ vertices in each layer $V_i$, such that $G-Z$ is planar. 
+%Moreover, there is a connected planar graph $G^+$ containing $G-Z$ as a subgraph, 
+%such that there is a BFS layering $(W_0,W_1,\dots)$ of $G^+$ and $W_{i} \cap V(G-Z) = V_i  \setminus Z$ for all $i\geq 0$. 
+For every BFS spanning tree $T$ of $G$ rooted at some vertex $r\in V(G)$ with corresponding BFS layering $(V_0,V_1,\dots)$, 
+there is a subgraph $Z\subseteq G$ with at most $2g$ vertices in each layer $V_i$, such that $G-V(Z)$ is planar. 
+Moreover, there is a connected planar graph $G^*$ containing $G-V(Z)$ as a subgraph, and 
+there is a BFS spanning tree $T^*$ of $G^*$ rooted at some vertex $r^*$ 
+with corresponding BFS layering $(W_0,W_1,\dots)$ of $G^*$, 
+such that  $W_{i} \cap V(G) \setminus V(Z) = V_i  \setminus V(Z)$ for all $i\geq 0$.
+% , and
+% $P \cap V(G) \setminus V(Z)$ is a vertical path in $T$ for every vertical path $P$ in $T^+$. 
+\end{lem}
+
+\begin{proof}[Proof of \thmref{g-k-planar}]
+  Let $G$ be an $(S,k)$-plane graph where $S$ is a surface of Euler genus $g$.
+Let $G^+$ be the $(S,0)$-plane graph obtained by adding a dummy vertex at each crossing in $G$.
+Let $r$ be any vertex in $V(G)$ and let $T$ be BFS-spanning tree of $G^+$ rooted at $r$ and let $\mathcal{L}=\rangle V_0,ldots,V_h\rangle$ be the resulting layering of $G^+$.
+
+  Apply \lemref{make-planar} to $G^{++}$ to obtain the set $Z\subseteq V(G^{++})$, the planar graph $G^*$, and the BFS spanning tree $T^*$ rooted at $r^*$.
+For each edge $vw\in E(G)$, let $W_{vw}$ be the corresponding walk in $G^+$.
+The set $Z$ contains vertices in $V(G)$ as well as dummy vertices in $V(G^+-G)$.
+Create a new set $Z'\subseteq V(G)$ as
+  \[
+     Z' = \{ v : V(W_{vw})\cap Z\neq\emptyset \} \enspace .
+  \]
+  Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \lemref{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 4\times (2k+1) = 16gk+8g$ for all $i\in\{1,\ldots,h\}$.
+
+  Furthermore, for every edge $vw\in E(G-Z')$, $\dist_{G^+-Z}(v,w)\le k+1$, since $W_{vw}\subseteq G^+-Z$.
+
+  Next, observe that any plane embedding of $G^*\supseteq G^+-Z$ gives a $k$-plane embedding of $G-Z'$.
+Indeed, for every edge $vw\in E(G-Z')$, $G^+-Z$ contains a walk $W_{vw}$ that contains at most $k$ dummy vertices of $G^+-Z$.
+Therefore, replacing edge edge $vw\in E(G-Z')$ with the corresponding walk in $G^*$ gives a $k$-plane embedding of $G-Z'$.
+
+  Apply \thmref{k-planar} or (in the case $k=1$) \thmref{1-planar} to the $k$-plane embedding of $G-Z'$ to obtain a layered $H$-partition $(\mathcal{L}',\mathcal{P})$ of $G-Z'$.
+Though not explicitly in the statement of either theorem, the proof of each theorem shows that it is possible to choose a vertex $r^*\in G^*$ so that each layer $V'_i\in \mathcal{L}'$ is contained in at most $k+1$ consecutive layers of $\mathcal{L}$.
+Therefore, $|Z'\cap V_i'| \le (k+1)(16gk+8g) = 16gk^2 + 24gk + 8g$.
+
+  We turn $\mathcal{P}$, the $H$-partition of $G-Z'$ into a partition $\mathcal{P}'=\mathcal{P}\cup \{Z'\}$.
+Clearly, the treewidth of $H'=G/\mathcal{P'}$ is at most one more than the treewidth of $H=G/\mathcal{P}$.
+Finally
+  \[
+      \max\{ |P\cap L| : P\in \mathcal{P}',\, L\in \mathcal{L}' \} \le
+      \max\{ 16gk^2 + 24gk + 8g, 24k^2 + 36k + 12 \} \enspace .
+  \]
+  \note{PM}{All of the above still needs work...}
+\end{proof}
+
 % \section{Consequences}

### 2019-05-14 k-planar.tex
A couple of notes to myself.

@@
 \]
-Clearly $\mathcal{S}=\{S_x:x\in V(K)\}$ is a partition of $V(G)$.
+\note{PM}{We might be able to save a factor of 2 here on the layered width, because we only need one endpoint of each edge $vw$.}Clearly $\mathcal{S}=\{S_x:x\in V(K)\}$ is a partition of $V(G)$.
 We will use $\mathcal{S}$ as the partition that defines our quotient graph $H=G/\mathcal{S}$ with vertex set $V(H)=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
@@
   \]
+  \note{PM}{We can certainly save a factor of 2 here because we only need one endpoint of each $vw$ that crosses $Z$ to take part in $Z'$.}
   Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \lemref{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 4\times (2k+1) = 16gk+8g$ for all $i\in\{1,\ldots,h\}$.

### 2019-05-15 k-planar.tex
Fixed v in S_x versus v in S_{x'} issue

@@
   Refer to \figref{A}(b)
-  If $B_x$ contains some $K$-ancestor $a_i$, $i\in\{0,\ldots,d-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_{a_i}$.
+  If $B_x$ contains some $K$-ancestor $a_i$, $i\in\{0,\ldots,d-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $w\in S_{a_i}$ and $v\in S_{x'}$ for some $K$-descendant of $x$.
 The edge $vw$ corresponds to a walk $W_{vw}$ in $G^+$.
-The walk $W_{vw}$ is contained in $N_x-F_x$, otherwise $x\not\in A_{vw}$ so $x\not\in A_v$, and $a(v)\neq x$, contradicting the fact that $v\in S_x$.
+The walk $W_{vw}$ is contained in $N_x-F_x$, otherwise $x\not\in A_{vw}$ so $x'\not\in A_{vw}$, so $x'\not\in A_v$, and $a(v)\neq x'$, contradicting the fact that $v\in S_{x'}$.
 
-  However, $w\in S_{_i}$, so there exists an edge $ww'\in E(G)$ such that   $W_{ww'}$ is contained in $W_{ww'}\subseteq N_{a_i}-F_{a_i}$ but $W_{ww'}\not\subseteq N_{a_{i+1}}-F_{a_{i+1}}$.
+  However, $w\in S_{a_i}$, so there exists an edge $ww'\in E(G)$ such that   $W_{ww'}$ is contained in $W_{ww'}\subseteq N_{a_i}-F_{a_i}$ but $W_{ww'}\not\subseteq N_{a_{i+1}}-F_{a_{i+1}}$.
 Therefore $W_{ww'}$ must contain a first vertex $w^*\in V(F_{a_{i+1}}-F_{a_{i}})\subseteq Y_{a_i}$.

### 2019-05-15 k-planar.tex
Getting k^3

@@
 The deepest vertex in a vertical path $P$ is called  $P$'s \emph{lower endpoint} and the shallowest vertex in $P$ is called $P$'s \emph{upper endpoint}.
-For any node $v\in T$, we call each node on the path, in $T$, from $v$ to $r$ a \emph{$T$-ancestor} of $v$.
+For any node $v\in V(T)$, we call each node on the path, in $T$, from $v$ to $r$ (including $v$ itself) a \emph{$T$-ancestor} of $v$.
+If $v'$ is a $K$-ancestor of $v$, then $v$ is a \emph{$K$-descendant} of $v'$.
 
@@
 Fix some cycle $F$ in $\Delta$ and let $N=\Delta\circledast F$.
-Let $\tau=v_1v_2v_3$ be an inner face of $N$ and consider the three minimal vertical paths $Q_1$, $Q_2$, and $Q_3$ in $T$ such that, for each $i\in\{1,2,3\}$ the lower endpoint of $Q_i$ is $v_i$ and the upper endpoint of $Q_i$ is in $V(F)$.
+Let $\tau=v_1v_2v_3$ be an inner face of $N$ and consider the three minimal vertical paths $Q_1$, $Q_2$, and $Q_3$ in $T$ such that, for each $i\in\{1,2,3\}$ the lower endpoint of $Q_i$ is $v_i$ and the upper endpoint $v_i'$ of $Q_i$ is in $V(F)$.
 If $Q_1$, $Q_2$, and $Q_3$ are vertex disjoint, then the graph $\overline{Y}=\tau\cup Q_1\cup Q_2\cup Q_3$ is called the \emph{closed tripod} in $N$ determined by $\tau$.
-We call $Q_1$, $Q_2$, and $Q_3$ the \emph{legs} of $\overline{Y}$ and we call $\tau$ the \emph{center} of $\overline{Y}$.
+We call $Q_1$, $Q_2$, and $Q_3$ the \emph{legs}, $v_1'$, $v_2'$ and $v_3'$ the \emph{feet}, and $\tau$ the \emph{crotch} of $\overline{Y}$.
 We call $Y=\overline{Y}-F$ the \emph{(open) tripod} of $N$ determined by $\tau$.
@@
 
-  \item \proplabel{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$.
+  \item \proplabel{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ non-empty sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$.
+
+  \item \proplabel{tripod-feet} For each $i\in\{1,\ldots,g\}$, at least one vertex in $P_i$ is a foot of the closed tripod $\overline{Y}_x$.
 
@@
 Indeed, this tree is constructed (albeit implicitly) by \citet[Proof of Lemma~14]{dujmovic.joret.ea:planar}.
+
+The following claim is not used until later in the paper. However, it's proof is a helpful exercise in understanding the properties of a PS-Tree.
+
+\begin{clm}\clmlabel{strong-ancestor-boundary}
+  Let $K$ be a PS-tree and
+  let $y\in V(K)$ be a non-root node with parent $p$ such that $V(F_y\cap Y_a)\neq\emptyset$ for some $K$-ancestor $a\neq p$ of $p$. Then $V(F_p\cap Y_a)\neq\emptyset$.
+\end{clm}
+
+\begin{proof}
+  By \propref{children}, $F_y$ is a wedge in the Mercedes graph $M_p$, so $V(F_y)$ can be partitioned into  $P_1'=V(F_y\cap Y_p)$ and $P_{2,3}'=V(F_y\cap F_p)$.
+Let $P_1,\ldots,P_g$ denote the partition of $V(F_p)$ given by \propref{ancestor-boundary}.
+
+  We claim that \propref{tripod-feet} ensures that $P_{2,3}'\subset P_i\cup P_j$ for some $i,j\in \{1,\ldots,g\}$.
+If $g=2$, this claim is immediate, so assume $g=3$.
+Then, $F_y\cap \overline{Y}_p$ is the path formed by two legs $Q_2$ and $Q_3$ of $\overline{Y}_p$ and one edge $v_2v_3$ of $\tau_p$.
+  By \propref{tripod-feet} and the assumption that $q=3$, $Q_2$ has a foot $v_2'\in P_i$ and $Q_3$ has a foot $v_3'\in P_j$, for some $i\neq j$.
+The third foot $v_1'$ of $\overline{Y}_p$ is therefore in $P_\ell$ for the unique $\ell \in\{1,2,3\}\setminus \{i,j\}$.
+It follows that $V(P_\ell\cap F_y)=\emptyset$.
+
+  Setting $P_2'=V(P_i\cup F_y)$ and $P_3'=V(P_j\cup F_y)$ yields a partition $\{P_1',P_2',P_3'\}$ of $V(F_y)$ such that $P_1'\subset V(Y_{p})$,   
+  $P_2'\subset V(Y_{a_2})$, and $P_3'\subset V(Y_{a_3})$  for some $K$-ancestors $a_2$ and $a_3$ of $a$.
+Therefore $a_2$ and $a_3$ are the only $K$-ancestors $a\neq p$ of $p$, such that $V(F_y\cap Y_a)\neq\emptyset$.
+  Now, each foot $v_2',v_3'\in V(F_p)$.
+Furthermore, $v_2'\in P_i\subseteq V(Y_{a_2})$ and $v_3'\in P_j\subseteq V(Y_{a_3})$.
+Therefore, $V(F_p\cap Y_{a_2})\supseteq\{v_2'\}\neq\emptyset$ and $V(F_p\cap Y_{a_3})\supseteq\{v_3'\}\neq\emptyset$, as claimed.
+\end{proof}
+
+Contained in the proof of \clmref{strong-ancestor-boundary} is the following addendum to \propref{ancestor-boundary}:
+
+\begin{enumerate}[(PR1')]\setcounter{enumi}{1}
+  \item \proplabel{ancestor-boundary-ii} Furthermore, if $g=3$, then  $P_i=V(F_x\cap Y_p)$ where $p$ is the parent of $x$ in $K$, for one $i\in\{1,2,3\}$.
+\end{enumerate}
 
@@
 We then cut $F_{a_0}$ into two pieces $F^{\bar{x}}_{a_0}$ and $F^x_{a_0}$ where $F^x_{a_0}$ is the piece that contains $F_x$.
-(The path that makes this cut contains two legs $Q_i$ and $Q_j$ of a tripod $Y$ and one edge $v_iv_j$ of its central triangle $\tau$.)  This process continues by partitioning $F_{a_1}:=F^x_{a_0}$ recursively.
+(The path that makes this cut contains two legs $Q_i$ and $Q_j$ of a tripod $Y$ and one edge $v_iv_j$ of its crotch $\tau=v_1v_2v_3$.)  This process continues by partitioning $F_{a_1}:=F^x_{a_0}$ recursively.
 The process terminates after some number $d$ of iterations, when $F^x_{a_{d-1}}=F_x=F_{a_d}$.
@@
   Refer to \figref{A}(b)
-  If $B_x$ contains some $K$-ancestor $a_i$, $i\in\{0,\ldots,d-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $w\in S_{a_i}$ and $v\in S_{x'}$ for some $K$-descendant of $x$.
+  If $B_x$ contains some $K$-ancestor $a_\delta$, $\delta\in\{0,\ldots,d-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $w\in S_{a_\delta}$ and $v\in S_{x'}$ for some $K$-descendant of $x$.
 The edge $vw$ corresponds to a walk $W_{vw}$ in $G^+$.
@@
 
-  However, $w\in S_{a_i}$, so there exists an edge $ww'\in E(G)$ such that   $W_{ww'}$ is contained in $W_{ww'}\subseteq N_{a_i}-F_{a_i}$ but $W_{ww'}\not\subseteq N_{a_{i+1}}-F_{a_{i+1}}$.
-Therefore $W_{ww'}$ must contain a first vertex $w^*\in V(F_{a_{i+1}}-F_{a_{i}})\subseteq Y_{a_i}$.
+  However, $w\in S_{a_\delta}$, so there exists an edge $ww'\in E(G)$ such that $W_{ww'}\subseteq N_{a_\delta}-F_{a_\delta}$ but $W_{ww'}\not\subseteq N_{a_{\delta+1}}-F_{a_{\delta+1}}$.
+Therefore $W_{ww'}$ must contain a first vertex $w^*\in V(F_{a_{\delta+1}}-F_{a_\delta})\subseteq Y_{a_\delta}$.
 
@@
 
-    \item $v_0,\ldots,v_{\ell-1}\in V(N_{a_{i+1}}-F_{a_{i+1}})$;
+    \item $v_0,\ldots,v_{\ell-1}\in V(N_{a_{\delta+1}}-F_{a_{\delta+1}})$;
 
-    \item $v_\ell\in V(F_{a_{i+1}}-F_{a_i})$; and
+    \item $v_\ell\in V(F_{a_{\delta+1}}-F_{a_\delta})$; and
 
@@
   \end{compactenum}
-  We call any walk with properties (i)--(iv) an \emph{$i$-walk}. Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $i\in\{0,\ldots,d-1\}$ such that there exists an $i$-walk.
+  We call any walk with properties (i)--(iv) a \emph{$\delta$-walk}.
+Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $\delta\in\{0,\ldots,d-1\}$ such that there exists an $\delta$-walk.
 
-  For each $v'\in V(G^+)$, define
+  In the following, when $i$ is an integer, we will use $F_i$, $Y_i$, and so on as shorthands for $F_{a_i}$, $Y_{a_i}$, and so on, respectively.
+
+  To begin, let $d_0=d$, and $d_1=\min\{i: V(F_d\cap Y_i)\neq\emptyset \}$.
+Observe that $F_{d}=F_{d_0}$ is contained in the interior of $F_{d_1}$, i.e., $N_{d_0}\subseteq N_{d_1}-F_{d_1}$.
+We claim that
   \[  
-      d_{v'}=\max\{j\in\{0,\ldots,d\} : v'\in V(N_{a_j}-F_{a_j})
+     |\{\delta\in\{d_1,\ldots,d_0-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+1}{2} \enspace .
   \]
-  (Intuitively, $d_{v'}$ is the $K$-depth of the lower endpoint of $K[P_{v'}]$ or the $K$-depth ($d$) of $x$, whichever is shallower.)
+  To prove this claim, consider the graph $I$ with vertex set $V(I)=\{d_1,\ldots,d_0\}$ and $ij\in E(I)$ if and only $i=j$ or $|F_i\cap Y_j|\neq\emptyset$.
+Any $\delta$-walk $v_0,\ldots,v_\ell$ with $\delta\in V(I)$ corresponds to a walk $\rho_0,\ldots,\rho_{\ell}$ in $I$, where $\rho_i=\max\{j : v_i\in N_{j}-F_j\}$.
 
-  Let $v_0,\ldots,v_\ell$ be an $i$-walk and consider the subsequence $d_0,\ldots,d_{\ell'}$ of $d_{v_0},\ldots,d_{v_\ell}$ consisting of only $d_{v_0}$ and those (record-breaking) $d_{v_j}$ such that $d_{v_j}<\min\{d_{v_0},\ldots,d_{v_{j-1}}\}$.
+  The graph $I$ has the following properties:
+  \begin{enumerate}
+    \item For each $i\in V(I)$, $i$ has at most 2 neighbours among $0,\ldots,i-1$.  If $i$ has two such neighbours, then one of them is $i-1$.
+
+    \item For each $ij\in E(I)$ with $j<i$, and each $i'\in\{j+2,\ldots,i\}$, the vertex $i'$ has at most 2 neighbours and these are contained in $\{i'-1,j\}$.
+  \end{enumerate}
+  These two properties imply that $I$ has a very simple structure.
+We call $j\in\{d_1,\ldots,d_0\}$ a \emph{left endpoint} if there exists an edge $ij\in E(I)$ with $j < i - 1$.
+Let $j_1>j_1>\cdots> j_p$ denote the sorted sequence of left endpoints and, for convenience, let $j_{0}=d$.
+Then
+  \[
+      E(I) \subseteq \{ (i,i+1): i\in\{d_1,\ldots,d_0-1\} \cup 
+         \bigcup_{t=1}^p \{ ij_t : i\in\{ j_t,\ldots,j_{t-1} \}\} \enspace . 
+  \]
+  (See, \figref{i-graph}.)  Observe that shortest path from in $i$ from $d_0$ to $j_t$ has length at least $t$.
+Let $X_t=\{j_t,\ldots,j_{t+1}\}$ and observe that the shortest path from $d_0$ to some value $i\in X_t$ that does not contain any $i' <i$ goes through $j_{t}$.
+Therefore, $X_t$ contains at most $k+2-t$ values of $\delta$ such that there exists a $\delta$-walk.
+Therefore,
+  \[  
+     |\{\delta\in\{d_1,\ldots,d_0-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \sum_{t=0}^{k+1} (k+2-t) = \sum_{t=1}^{k+2} t = \binom{k+2}{2} \enspace .
+  \]
+
+  \begin{figure}
+    \begin{center}
+      \includegraphics{figs/i-graph}
+    \end{center}
+    \caption{The graph $I$ used in the proof of \clmref{bag-size}}
+    \figlabel{i-graph}
+  \end{figure}
+
+  More generally, for $s>1$, let $d_s=\min\{ i: V(F_{d_s}\cap Y_i)\neq\emptyset \}$.
+Consider some $\delta$-walk $v_0,\ldots,v_\ell$ for $\delta\in \{ d_{s},\ldots,d_{s-1}-1 \}$, for some $s>1$.
+Since $v_0\in V(N_{d_0}-F_{d_0})$ and $v_\ell\in F_{\delta+1}-F_\delta$, there exists some maximum value $i$ such that $v_i\in N_{d_{s-1}}-F_{d_{s-1}}$.
+Furthermore, $i\ge s-2$ because $v_1,\ldots,v_i$ contains a vertex from each of $F_{d_0},F_{d_1},\ldots,F_{d_{s-2}}$.
+But this means that we can treat $v_i,\ldots,v_{\ell}$ as a $\delta$-walk of length at most $\ell-s+2 \le k-s+3$ that begins in $N_{d_{s-1}}-F_{d_{s-1}}$.
+By the same argument used above,
+  \[  |\{\delta\in\{d_s,\ldots,d_{s-1}-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k-s+3}{2} \enspace .\]
+
+  We conclude that the maximum number of $\delta\in\{0,\ldots,d-1\}$ such that there exists a $\delta$-walk is at most
+  \[
+      \sum_{s=1}^{k+2} \binom{k-s+3}{2} = O(k^3)
+  \]
+  \note{PM}{Gotta run!}
+
+  % 
+  % 
+  % Next observe that, if some $\delta$-walk
+  % 
+  % 
+  % 
+  % 
+  % For each $v'\in V(G^+)$, define
+  % \[  
+  %     d_{v'}=\max\{i\in\{0,\ldots,d\} : v'\in V(N_{a_i}-F_{a_i})
+  % \]
+  % (Intuitively, $d_{v'}$ is the $K$-depth of the lower endpoint of $K[P_{v'}]$ or the $K$-depth ($d$) of $x$, whichever is shallower.)
+  % 
+  % Consider some $\delta$-walk $v_0,\ldots,v_{\ell}$ and, for each $j\in\{0,\ldots,\ell\}$, let $d_j=d_{v_j}$.
+Then the sequence $d_0,\ldots,d_{\ell}$ has the following properties, corresponding to properties (i)--(iii) above:
+  % \begin{compactenum}[(i)]
+  %   \item $d_0=d$;
+  %   \item $d_0,\ldots,d_{\ell-1} > \delta$
+  %   \item $d_\ell = \delta$;
+  %   \item $\ell\le k+1$.
+  % \end{compactenum}
+  % 
+  % Consider an auxilliary graph $I$ with vertex set $V(I)=\{0,\ldots,d\}$ and edge $ij\in E(I)$ if and only if $V(F_{a_i}\cap Y_j))\neq\emptyset$. Observe that $d_0,\ldots,d_\ell$ is a walk in $I$.
+  % 
+  % We are particularly interested in the \emph{up neighbours} of $i$, defined as
+  % $U(i) = \{q : iq\in E(I),\, q< i-1\}$.  
+  % 
+  % Claim: For each $j\in\{i_1-2,\ldots,d\}$,  $U(j) = U(d)=\{i_1,i_2\}$.
+  % 
+  % Claim: $U(i_1)= \{i_2, i_3\}$ for some $i_3< i_1-1$.
+  % 
+  % Claim: $U(i_1+1)= \{i_1, i_2, i_3\}$.
+  % 
+  %  U(i)\cup\{i-2\}\setminus\{i-1\}$. 
+  % 
+  % Let $i_0=d$, $i_1=\min U(d)$, and $i_2=U(d)$.  
+  % 
+  % 
+  % 
+  % By \clmref{strong-ancestor-boundary}, $I(j)=U(d)=\{i_1,i_2\}$ for all $j\in\{i_1-1,\ldots,d\}$.
+This immediately implies that there can be no $\delta$-walk for any $\delta\in \{ i\in\Z : k+1 < i < i_1 \}$.
+Indeed, such a walk either has to contain $i_1$ or $i_2$ or it has to contain $d,d-1,\ldots,d-k-2$.
+In the former case, the walk violates (ii) or (iii) since it contains values smaller than $\delta$.
+In the latter case, the walk violates (iv) since it has length $k+2>k+1$.
+  % 
+  % Next, consider $\max U(i_1)$ and let $i_3=\max U(i_1)$.  By \propref{strong-ancestor-boundary} 
+  % 
+  % 
+  % 
+  %  since $\delta > i_1 > i_2$ 
+  % 
+  % Claim: For all $j\in\{\min(U(i)),\ldots,i\}$, $U(j) = U(i)$. 
+  % 
+  % Proof: 
+  % The proof is by induction on $i-j$.  The base case $i=j$ is immediate.
+  % 
+  % Now consider some $j\in\{\min(U(i)), i-1\}$.  By the inductive hypothesis $U(j+1)=U(i)$ so $F_{j+1}$ contains vertices of $Y_q$ for each $q\in U(i)$. 
+  % These vertices are not in $Y_j$ since, by the definition of $U(j+1)$, $q < (j+1)-1 = j$ for each $q\in U(j+1)$.  
+  % 
+  % 
+  %  Recall that $F_{j+1}$ is a wedge in the Mercedes graph $M_j$. This wedge is determined by the closed tripod $\overline{Y}_j$
+  % 
+  % First the base case $j=i$:
+  % 
+  % Observe that $F_i$ is one wedge of the Mercedes graph $M_i$, so $F_i \subset F_{i-1}\cup \overline{Y}_{i-1}$.
+Therefore $V(F_i)$ can be partitioned into $A=V(F_i \cap F_{i-1})$ and $B=V(F_i\cap Y_{i-1})$.
+The set $B$ consists entirely of vertices in $N_{i-1}-F_{i-1}-(N_i-F_i)$.
+The set $A$ consists of vertices in $V(F_{i-1})$.
+By \propref{ancestor-boundary},
+  % 
+  % First observe that it is sufficient to prove this for $j=i-1$ since, in this case, we can apply
+  % Consider the vertices of $F_{i}$.  Each of these vertices belongs $Y_{q}$ for some $q\in U(i)$.
+  % 
+  % A critical property of these up edges is that, for any $i' > i$, $U(i') \subseteq \{i'+1,\ldots,i\} \cup U(i)$.  
+  % 
+  % 
+  % , there are at most 3 edges $ij\in E(I)$ with $j<i$.  Furthermore, for each $i'< i$, 
+  % 
+  % 
+  % 
+  % The graph $I$ has some additional structure, which we now describe.  
+  % First, consider the at most three neighbours $i_0>i_1>i_2$ of $d$ in $I$.
+  % In the walk $d_0,\ldots,d_{\ell}$, the first value $d_i\neq d$ is in $\{i_0,i_1,i_2\}$.  Furthermore, 
+  % 
+  % 
+  % 
+  % 
+  % First observe that, for each $i\in\{1,\ldots,d\}$, $(i,(i+1))\in E(I)$.  
+  % 
+  % 
+  % 
+  % $I$ has maximum out-degree $3$ and each directed edge $ij\in E(I)$ has $j < i$.
+  % 
+  % 
+  % 
+  % 
+  % 
+  % 
+  % Let $v_0,\ldots,v_\ell$ be an $i$-walk and consider the subsequence $d_0,\ldots,d_{\ell'}$ of $d_{v_0},\ldots,d_{v_\ell}$ consisting of only $d_{v_0}$ and those (record-breaking) $d_{v_j}$ such that $d_{v_j}<\min\{d_{v_0},\ldots,d_{v_{j-1}}\}$.
 Observe that $d_0=d$ and $d_{\ell'}=i$.
-
-  By \propref{ancestor-boundary}, for every $j\in\{1,\ldots,d\}$, there is a set $X_{j}$, of at most 3 integers $\alpha < j$ such that every vertex $q\in V(F_{a_j})$ is in $V(Y_{a_{\alpha}})$ for some $\alpha\in X_j$.
-  Assign a distinct colour $c(j,\alpha)\in\{1,2,3\}$ for each $\alpha\in X_j$ and do this assignment for each $j\in\{1,\ldots,d\}$.
+  % 
+  % By \propref{ancestor-boundary}, for every $j\in\{1,\ldots,d\}$, there is a set $X_{j}$, of at most 3 integers $\alpha < j$ such that every vertex $q\in V(F_{a_j})$ is in $V(Y_{a_{\alpha}})$ for some $\alpha\in X_j$.
+  % Assign a distinct colour $c(j,\alpha)\in\{1,2,3\}$ for each $\alpha\in X_j$ and do this assignment for each $j\in\{1,\ldots,d\}$.
 By definition, for any $j\in\{1,\ldots,d\}$ and any colour $\gamma\in\{1,2,3\}$, there is at most one $\alpha=\alpha(j,\gamma)$ such that $c(j,\alpha)=\gamma$.
-
-  Finally, observe that, for every $j\in\{1,\ldots,\ell'\}$,  $c(d_j,d_{j+1})$ is defined.
+  % 
+  % Finally, observe that, for every $j\in\{1,\ldots,\ell'\}$,  $c(d_j,d_{j+1})$ is defined.
 Refer to \figref{colouring}.
@@
 Since $d_0=d$ and $d_{\ell'-1}=i$, from the preceding discussion, we have $\alpha(\cdots (\alpha(\alpha(\alpha(d,c_0),c_1),c_2)\cdots),c_{\ell'-1}) = i$.
-
-  \begin{figure}
-    \begin{center}
-        \includegraphics{figs/A-2}
-    \end{center}
-    \caption{The colour sequences defined by two $i$-curves.}
-    \figlabel{colouring}
-  \end{figure}
-
-  The proof is now almost complete.  Any $i$-walk defines a colour sequence $c_0,\ldots,c_{\ell'}$.  If some $i'$-walk defines the same colour sequence $c_0,\ldots,c_{\ell'}$ then
-  \[
-        i = (\alpha(\alpha(\alpha(d,c_0),c_1),c_2)\cdots),c_{\ell'-1}) = i' \enspace ,
-  \]
-  so $i=i'$.  Since there are at most $3^{\ell'}$ colour sequences of length $\ell'$ and $\ell'\le \ell \le k+1$, the number of $i\in\{0,\ldots,d-1\}$ such that there exists an $i$-walk is at most
-  \[
-     \sum_{\ell=1}^{k+1} 3^\ell = (3^{k+2}-3)/2 \ge |B_x| \eqlabel{bag-size} \enspace . \qedhere
-  \]
+  % 
+  % \begin{figure}
+  %   \begin{center}
+  %       \includegraphics{figs/A-2}
+  %   \end{center}
+  %   \caption{The colour sequences defined by two $i$-curves.}
+  %   \figlabel{colouring}
+  % \end{figure}
+  % 
+  % The proof is now almost complete.  Any $i$-walk defines a colour sequence $c_0,\ldots,c_{\ell'}$.  If some $i'$-walk defines the same colour sequence $c_0,\ldots,c_{\ell'}$ then
+  % \[
+  %       i = (\alpha(\alpha(\alpha(d,c_0),c_1),c_2)\cdots),c_{\ell'-1}) = i' \enspace ,
+  % \]
+  % so $i=i'$.  Since there are at most $3^{\ell'}$ colour sequences of length $\ell'$ and $\ell'\le \ell \le k+1$, the number of $i\in\{0,\ldots,d-1\}$ such that there exists an $i$-walk is at most
+  % \[
+  %    \sum_{\ell=1}^{k+1} 3^\ell = (3^{k+2}-3)/2 \ge |B_x| \eqlabel{bag-size} \enspace . \qedhere
+  % \]
 \end{proof}

### 2019-05-16 k-planar.tex
Cubic bound seems to be working

@@
 
-The following claim is not used until later in the paper. However, it's proof is a helpful exercise in understanding the properties of a PS-Tree.
-
-\begin{clm}\clmlabel{strong-ancestor-boundary}
-  Let $K$ be a PS-tree and
-  let $y\in V(K)$ be a non-root node with parent $p$ such that $V(F_y\cap Y_a)\neq\emptyset$ for some $K$-ancestor $a\neq p$ of $p$. Then $V(F_p\cap Y_a)\neq\emptyset$.
-\end{clm}
-
-\begin{proof}
-  By \propref{children}, $F_y$ is a wedge in the Mercedes graph $M_p$, so $V(F_y)$ can be partitioned into  $P_1'=V(F_y\cap Y_p)$ and $P_{2,3}'=V(F_y\cap F_p)$.
+% The following claim is not used until later in the paper. However, it's proof is a helpful exercise in understanding the properties of a PS-Tree.
+% 
+% \begin{clm}\clmlabel{strong-ancestor-boundary}
+%   Let $K$ be a PS-tree and
+%   let $y\in V(K)$ be a non-root node with parent $p$ such that $V(F_y\cap Y_a)\neq\emptyset$ for some $K$-ancestor $a\neq p$ of $p$. Then $V(F_p\cap Y_a)\neq\emptyset$.
+% \end{clm}
+% 
+% \begin{proof}
+%   By \propref{children}, $F_y$ is a wedge in the Mercedes graph $M_p$, so $V(F_y)$ can be partitioned into  $P_1'=V(F_y\cap Y_p)$ and $P_{2,3}'=V(F_y\cap F_p)$.
 Let $P_1,\ldots,P_g$ denote the partition of $V(F_p)$ given by \propref{ancestor-boundary}.
-
-  We claim that \propref{tripod-feet} ensures that $P_{2,3}'\subset P_i\cup P_j$ for some $i,j\in \{1,\ldots,g\}$.
+% 
+%   We claim that \propref{tripod-feet} ensures that $P_{2,3}'\subset P_i\cup P_j$ for some $i,j\in \{1,\ldots,g\}$.
 If $g=2$, this claim is immediate, so assume $g=3$.
 Then, $F_y\cap \overline{Y}_p$ is the path formed by two legs $Q_2$ and $Q_3$ of $\overline{Y}_p$ and one edge $v_2v_3$ of $\tau_p$.
-  By \propref{tripod-feet} and the assumption that $q=3$, $Q_2$ has a foot $v_2'\in P_i$ and $Q_3$ has a foot $v_3'\in P_j$, for some $i\neq j$.
+%   By \propref{tripod-feet} and the assumption that $q=3$, $Q_2$ has a foot $v_2'\in P_i$ and $Q_3$ has a foot $v_3'\in P_j$, for some $i\neq j$.
 The third foot $v_1'$ of $\overline{Y}_p$ is therefore in $P_\ell$ for the unique $\ell \in\{1,2,3\}\setminus \{i,j\}$.
 It follows that $V(P_\ell\cap F_y)=\emptyset$.
-
-  Setting $P_2'=V(P_i\cup F_y)$ and $P_3'=V(P_j\cup F_y)$ yields a partition $\{P_1',P_2',P_3'\}$ of $V(F_y)$ such that $P_1'\subset V(Y_{p})$,   
-  $P_2'\subset V(Y_{a_2})$, and $P_3'\subset V(Y_{a_3})$  for some $K$-ancestors $a_2$ and $a_3$ of $a$.
+% 
+%   Setting $P_2'=V(P_i\cup F_y)$ and $P_3'=V(P_j\cup F_y)$ yields a partition $\{P_1',P_2',P_3'\}$ of $V(F_y)$ such that $P_1'\subset V(Y_{p})$,   
+%   $P_2'\subset V(Y_{a_2})$, and $P_3'\subset V(Y_{a_3})$  for some $K$-ancestors $a_2$ and $a_3$ of $a$.
 Therefore $a_2$ and $a_3$ are the only $K$-ancestors $a\neq p$ of $p$, such that $V(F_y\cap Y_a)\neq\emptyset$.
-  Now, each foot $v_2',v_3'\in V(F_p)$.
+%   Now, each foot $v_2',v_3'\in V(F_p)$.
 Furthermore, $v_2'\in P_i\subseteq V(Y_{a_2})$ and $v_3'\in P_j\subseteq V(Y_{a_3})$.
 Therefore, $V(F_p\cap Y_{a_2})\supseteq\{v_2'\}\neq\emptyset$ and $V(F_p\cap Y_{a_3})\supseteq\{v_3'\}\neq\emptyset$, as claimed.
-\end{proof}
-
-Contained in the proof of \clmref{strong-ancestor-boundary} is the following addendum to \propref{ancestor-boundary}:
-
-\begin{enumerate}[(PR1')]\setcounter{enumi}{1}
-  \item \proplabel{ancestor-boundary-ii} Furthermore, if $g=3$, then  $P_i=V(F_x\cap Y_p)$ where $p$ is the parent of $x$ in $K$, for one $i\in\{1,2,3\}$.
-\end{enumerate}
+% \end{proof}
+% 
+% Contained in the proof of \clmref{strong-ancestor-boundary} is the following addendum to \propref{ancestor-boundary}:
+% 
+% \begin{enumerate}[(PR1')]\setcounter{enumi}{1}
+%   \item \proplabel{ancestor-boundary-ii} Furthermore, if $g=3$, then  $P_i=V(F_x\cap Y_p)$ where $p$ is the parent of $x$ in $K$, for one $i\in\{1,2,3\}$.
+% \end{enumerate}
 
@@
   \[
-    \dist_{G^+}(r,v') \le \dist_{G^+}(r,v) + \dist_{G^+}(v,v') \le \dist_{G^+}(r,v) + k + 1 \enspace .
+    \dist_{G^+}(r,v') \le \dist_{G^+}(r,v) + \dist_{G^+}(v,v') 
+       \le \dist_{G^+}(r,v) + k + 1 \enspace .
   \]
@@
   \begin{figure}
-
     \begin{center}
@@
 Observe that $F_{d}=F_{d_0}$ is contained in the interior of $F_{d_1}$, i.e., $N_{d_0}\subseteq N_{d_1}-F_{d_1}$.
+Refer to \figref{d0d1}.
 We claim that
@@
   \]
+  \begin{figure}
+    \begin{center}
+      \includegraphics{figs/d0d1-5}
+    \end{center}
+    \caption{$F_{d_1},\ldots,F_{d_0}$ and the resulting graph $I$.}
+    \figlabel{d0d1}
+  \end{figure}
   To prove this claim, consider the graph $I$ with vertex set $V(I)=\{d_1,\ldots,d_0\}$ and $ij\in E(I)$ if and only $i=j$ or $|F_i\cap Y_j|\neq\emptyset$.
-Any $\delta$-walk $v_0,\ldots,v_\ell$ with $\delta\in V(I)$ corresponds to a walk $\rho_0,\ldots,\rho_{\ell}$ in $I$, where $\rho_i=\max\{j : v_i\in N_{j}-F_j\}$.
+Any $\delta$-walk $v_0,\ldots,v_\ell$ with $\delta\in V(I)$ corresponds to a walk $\rho_0,\ldots,\rho_{\ell}$ in $I$, where $\rho_i=\max\{j\in\{0,\ldots,d\} : v_i\in N_{j}-F_j\}$.
 
@@
   \begin{enumerate}
-    \item For each $i\in V(I)$, $i$ has at most 2 neighbours among $0,\ldots,i-1$.  If $i$ has two such neighbours, then one of them is $i-1$.
+    \item For each $i\in V(I)$, $i$ has at most 2 neighbours among $d_1,\ldots,i-1$.  If $i$ has two such neighbours, then one of them is $i-1$.
 
@@
 We call $j\in\{d_1,\ldots,d_0\}$ a \emph{left endpoint} if there exists an edge $ij\in E(I)$ with $j < i - 1$.
-Let $j_1>j_1>\cdots> j_p$ denote the sorted sequence of left endpoints and, for convenience, let $j_{0}=d$.
+Let $j_1>j_2>\cdots> j_p$ denote the sorted sequence of left endpoints and, for convenience, let $j_{0}=d$.
 Then
@@
   \]
-  (See, \figref{i-graph}.)  Observe that shortest path from in $i$ from $d_0$ to $j_t$ has length at least $t$.
+  (See the lower half of \figref{d0d1}.)  Observe that shortest path from in $i$ from $d_0$ to $j_t$ has length at least $t$.
 Let $X_t=\{j_t,\ldots,j_{t+1}\}$ and observe that the shortest path from $d_0$ to some value $i\in X_t$ that does not contain any $i' <i$ goes through $j_{t}$.
@@
 
-  \begin{figure}
-    \begin{center}
-      \includegraphics{figs/i-graph}
-    \end{center}
-    \caption{The graph $I$ used in the proof of \clmref{bag-size}}
-    \figlabel{i-graph}
-  \end{figure}
+  % \begin{figure}
+  %   \begin{center}
+  %     \includegraphics{figs/i-graph}
+  %   \end{center}
+  %   \caption{The graph $I$ used in the proof of \clmref{bag-size}}
+  %   \figlabel{i-graph}
+  % \end{figure}
 

### 2019-05-16 k-planar.tex
Almost done with k^3 proof

@@
   We call any walk with properties (i)--(iv) a \emph{$\delta$-walk}.
-Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $\delta\in\{0,\ldots,d-1\}$ such that there exists an $\delta$-walk.
+Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $\delta\in\{0,\ldots,d-1\}$ such that there exists a $\delta$-walk.
 
-  In the following, when $i$ is an integer, we will use $F_i$, $Y_i$, and so on as shorthands for $F_{a_i}$, $Y_{a_i}$, and so on, respectively.
+  In order to avoid excessive levels of subscripts the remainder of this proof will slightly abuse notation in the following way: For $i\in\{0,\ldots,d\}$, we use $F_i$, $N_i$, and $Y_i$, as shorthands for $F_{a_i}$, $Y_{a_i}$, respectively.
 
@@
 Any $\delta$-walk $v_0,\ldots,v_\ell$ with $\delta\in V(I)$ corresponds to a walk $\rho_0,\ldots,\rho_{\ell}$ in $I$, where $\rho_i=\max\{j\in\{0,\ldots,d\} : v_i\in N_{j}-F_j\}$.
+Because of Properties~(i)--(iv) of $\delta$-walks, any walk $\rho_0,\ldots,\rho_\ell$ obtained this way has (i)~$\rho_0=d_0$, (ii--iii)~$\delta=\rho_\ell < \min\{\rho_0,\ldots,\rho_{\ell-1}\}$, and (iv)~$\ell \le k+1$.
 
@@
   \begin{enumerate}
-    \item For each $i\in V(I)$, $i$ has at most 2 neighbours among $d_1,\ldots,i-1$.  If $i$ has two such neighbours, then one of them is $i-1$.
+    \item For each $i\in V(i)$, $id_1\in E(I)$.
+    \item For each $i\in V(I)$, $i$ has at most 2 neighbours in $\{d_1+1,\ldots,i-1\}$. If $i$ has 2 such neighbours, then one of them is $i-1$.
 
-    \item For each $ij\in E(I)$ with $j<i$, and each $i'\in\{j+2,\ldots,i\}$, the vertex $i'$ has at most 2 neighbours and these are contained in $\{i'-1,j\}$.
+    \item For each $ij\in E(I)$ with $j<i-1$, and each $i'\in\{j+1,\ldots,i+1\}$, the neighbours of $i'$ in $\{d_1+1,\ldots,i'-1\}$ are contained in $\{i'-1,j\}$.
   \end{enumerate}

### 2019-05-16 k-planar.tex
No more details hiding

--- 2019-05-16 k-planar.tex
@@
 
+\newcommand{\jlabel}[1]{\label{j:#1}}
+\newcommand{\jref}[1]{(J\ref{j:#1})}
+
 \begin{document}
@@
 Such partitions are known to exist for planar graphs and, more generally, bounded genus graphs.
-In the current paper, we prove that every $k$-planar graph has a layered $H$-partition of layered width $O(k^2)$ in which $H$ has treewidth $O(c^k)$, for some constant $c$.
+In the current paper, we prove that every $k$-planar graph has a layered $H$-partition of layered width $O(k^2)$ in which $H$ has treewidth $O(k^3)$.
 This is the first result of this type for a non-minor-closed class of graphs and implies that $k$-planar graphs have both queue number and non-repetitive chromatic number upper-bounded by a function of $k$.
@@
 \begin{thm}\thmlabel{k-planar}
-  Every $k$-planar graph has a layered $H$-partition of layered-width at most $24k^2 + 36k + 12$ in which $H$ has treewidth at most $(3^{k+2}-5)/2$.
+  Every $k$-planar graph has a layered $H$-partition of layered-width at most $24k^2 + 36k + 12$ in which $H$ has treewidth at most $k^3/6+k^2/+k/3-1$.
 \end{thm}
@@
 \begin{clm}\clmlabel{bag-size}
-  The tree decomposition $\{B_x: x\in V(K)\}$ of $H$ contains no bag larger than $(3^{k+2}-3)/2$.
+  The tree decomposition $\{B_x: x\in V(K)\}$ of $H$ contains no bag larger than $k^3/6 + k^2/2 + k/3$.
 \end{clm}
@@
   We call any walk with properties (i)--(iv) a \emph{$\delta$-walk}.
-Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $\delta\in\{0,\ldots,d-1\}$ such that there exists an $\delta$-walk.
+Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $\delta\in\{0,\ldots,d-1\}$ such that there exists a $\delta$-walk.
+This is our goal for the remainder of this proof.
 
-  In the following, when $i$ is an integer, we will use $F_i$, $Y_i$, and so on as shorthands for $F_{a_i}$, $Y_{a_i}$, and so on, respectively.
+  In order to avoid excessive levels of subscripts, we introduce the following abbreviated notation: For $i\in\{0,\ldots,d\}$, we use $F_i$, $N_i$, and $Y_i$, as shorthands for $F_{a_i}$, $N_{a_i}$, and $Y_{a_i}$, respectively.
 
-  To begin, let $d_0=d$, and $d_1=\min\{i: V(F_d\cap Y_i)\neq\emptyset \}$.
-Observe that $F_{d}=F_{d_0}$ is contained in the interior of $F_{d_1}$, i.e., $N_{d_0}\subseteq N_{d_1}-F_{d_1}$.
+  Let $d_0=d$, and $d_1=\min\{i: V(F_d\cap Y_i)\neq\emptyset \}$.
+By \propref{ancestor-boundary}, $F_d$ is contained in the interior of  $F_{d_1}$, i.e., $N_{d_0}\subseteq N_{d_1}-F_{d_1}$.
 Refer to \figref{d0d1}.
-We claim that
-  \[  
-     |\{\delta\in\{d_1,\ldots,d_0-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+1}{2} \enspace .
-  \]
+We will now prove that
+  \begin{equation}
+     |\{\delta\in\{d_1,\ldots,d_0-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+1}{2} \enspace . \eqlabel{one-round}
+  \end{equation}
   \begin{figure}
     \begin{center}
-      \includegraphics{figs/d0d1-5}
+      \includegraphics{figs/d0d1}
     \end{center}
-    \caption{$F_{d_1},\ldots,F_{d_0}$ and the resulting graph $I$.}
+    \caption{$F_{d_1},\ldots,F_{d_0}$ and the resulting graph $I$.  For readability, the edges $\{id_1:i\in\{d_1-2,\ldots,d_0\}$ are not shown.}
     \figlabel{d0d1}
   \end{figure}
-  To prove this claim, consider the graph $I$ with vertex set $V(I)=\{d_1,\ldots,d_0\}$ and $ij\in E(I)$ if and only $i=j$ or $|F_i\cap Y_j|\neq\emptyset$.
+
+  To prove this claim, consider the graph $I$, illustrated in \figref{d0d1}, with vertex set $V(I)=\{d_1+1,\ldots,d_0\}$ and $ij\in E(I)$ if and only $i=j$ or $|F_i\cap Y_j|\neq\emptyset$.
 Any $\delta$-walk $v_0,\ldots,v_\ell$ with $\delta\in V(I)$ corresponds to a walk $\rho_0,\ldots,\rho_{\ell}$ in $I$, where $\rho_i=\max\{j\in\{0,\ldots,d\} : v_i\in N_{j}-F_j\}$.
+To see why this is true, consider $\rho_{t}= i$ and $\rho_{t+1}=j$ for some $t\in\{0,\ldots,\ell-1\}$:
+  \begin{compactenum}
+    \item If $i=j$, then $ij$ is an edge of $I$ since $I$ contains the self-loop $ii$ for each $i\in V(I)$.
 
-  The graph $I$ has the following properties:
-  \begin{enumerate}
-    \item For each $i\in V(I)$, $i$ has at most 2 neighbours among $d_1,\ldots,i-1$.  If $i$ has two such neighbours, then one of them is $i-1$.
+    \item If $j\neq i$ then assume, without loss of generality, that $j < i$ so that $a_i$ is a $K$-descendant of $a_j$.
+Now $v_t\in V(N_i-F_i)$, but $v_{t+1}\not\in V(N_i-F_i)$ otherwise we would have $j\ge i$.
+Therefore, $v_{t+1}\in V(F_i)$.
+By \propref{ancestor-boundary}, $v_{t+1}$ is in $Y_{j'}$ for some $j'< i$ and the edge $ij'\in E(I)$.
+By \propref{children}, $v_{t+1}\not\in V(N_{j''}-F_{j''})$ for any $K$-descedant $j''>j'$ of $j'$.
+Therefore $j'=\max\{j : v_{t+1}\in N_j-F_j\}$.
+Therefore, by definition $j'=\rho(v_{t+1})=j$, so the edge $ij=ij'\in E(I)$.
+  \end{compactenum}
 
-    \item For each $ij\in E(I)$ with $j<i$, and each $i'\in\{j+2,\ldots,i\}$, the vertex $i'$ has at most 2 neighbours and these are contained in $\{i'-1,j\}$.
+  Properties~(i)--(iv) of $\delta$-walks imply that any walk $\rho_0,\ldots,\rho_\ell$ obtained this way 
+  \begin{compactenum}[(i')]
+    \item begins at $\rho_0=d_0$ 
+    \item is contained in $I[\delta,\ldots,d_0]$, 
+    \item ends at $\rho_\ell=\delta$
+    \item has length $\ell \le k+1$.
+  \end{compactenum}
+  The edges in $E(I)$ can be partitioned into two classes:  A set $B(I)=\{ij\in E(I): |i-j|\le 1\}$ of \emph{boring edges} and a set $J(I)=\{ij\in E(I): |i-j|>1\}$ of \emph{jump edges}.
+For a jump edge $ij$ with $i>j$, $j$ is called a \emph{landing} and $i$ is called a \emph{takeoff}.
+The jump edges of $I$ satisfy the following properties:
+  \begin{enumerate}[(J1)]
+    \item \jlabel{single} Every $i\in\{d_1+1,\ldots,d_0\}$ is a takeoff for at most one jump edge.
+
+    \item \jlabel{nested} For every jump edge $ij\in E(I)$ with takeoff $i$ and landing $j$, and for every $q\in\{j+2,\ldots,i\}$, if $q$ is a takeoff some jump edge, then this edge is $qj$.
   \end{enumerate}
-  These two properties imply that $I$ has a very simple structure.
-We call $j\in\{d_1,\ldots,d_0\}$ a \emph{left endpoint} if there exists an edge $ij\in E(I)$ with $j < i - 1$.
-Let $j_1>j_2>\cdots> j_p$ denote the sorted sequence of left endpoints and, for convenience, let $j_{0}=d$.
-Then
+  Property \jref{nested} can be interpreted informally as saying that jump edges can cross, but barely:  If $i_1j_1$ and $i_2j_2$ are jump edges with $i_1> i_2>j_1>j_2$, then $j_1=i_2-1$.
+
+  We now prove that jump edges satisfy \jref{single}.
+  First We claim that, for each $i\in\{d_1+1,\ldots,d_0\}$, $F_{i}$ contains a vertex of $Y_{d_1}$.
+For $i=d_0$, this is true by definition so $F_{d_0}$ contains some vertex $v\in V(Y_{d_1})$.
+For $i>0$, we have $F_{d_0}-F_{i}\subseteq N_i-F_i$, but $N_i-F_i$ contains no vertex of $Y_{d_1}$.
+It must therefore be the case that $v\in V(F_i)$, otherwise $F_{d_0}-F_i$ would contain $v$.
+
+  Now, consider the parent $a_{i-1}$ of $a_i$ in $K$.
+By \propref{children} $F_i$ is a wedge of the Mercedes graph $M_{i-1}$.
+Let $P_1,\ldots,P_3$ be the partition of $V(F_{i-1})$ guaranteed by \propref{ancestor-boundary}.
+Without loss of generality, we can assume that $P_1=V(Y_{d_1}\cap F_{i-1})$.
+Then, by \propref{tripod-feet}, $V(F_i)$ contains vertices from at most one of $P_2$ or $P_3$---say $P_2$---and suppose $P_2=V(F_{i-1}\cap Y_j)$ for some $j< i-1$.
+This accounts for a single jump edge $ij$ for which $i$ is the takeoff.
+All remaining vertices of $F_i$ are contained in $Y_{i-1}$, which only introduces the boring edge $(i,i-1)$.
+Thus we have established that jump edges satisfy \jref{single}.
+
+  To establish that jump edges satisfy \jref{nested} we can use the same argument used to establish that each $F_i$ contains a vertex of $Y_{d_1}$ to establish that each $F_q$ contains a vertex of $Y_j$ and therefore $qj\in E(I)$.
+Since $q \ge j+2$, $qj$ is a jump edge and, by \propref{single}, it is the only jump edge with takeoff $q$.
+
+  Let $j_0=d$ and let $j_1>j_2>\cdots> j_p$ denote the sequence of landings, in decreasing order.  Then \jref{single} and \jref{nested} imply that
   \[
-      E(I) \subseteq \{ (i,i+1): i\in\{d_1,\ldots,d_0-1\} \cup 
-         \bigcup_{t=1}^p \{ ij_t : i\in\{ j_t,\ldots,j_{t-1} \}\} \enspace . 
+      J(I) \subseteq J^+(I) :=
+         \bigcup_{t=1}^p \{ ij_t : i\in\{ j_t,\ldots,j_{t-1}+1 \}\} \enspace . 
   \]
-  (See the lower half of \figref{d0d1}.)  Observe that shortest path from in $i$ from $d_0$ to $j_t$ has length at least $t$.
-Let $X_t=\{j_t,\ldots,j_{t+1}\}$ and observe that the shortest path from $d_0$ to some value $i\in X_t$ that does not contain any $i' <i$ goes through $j_{t}$.
-Therefore, $X_t$ contains at most $k+2-t$ values of $\delta$ such that there exists a $\delta$-walk.
-Therefore,
+  and clearly $B(I)\subseteq B^+(I) := \{(i,i+1): i\in\{d_1+2,\ldots,d_0\}\}$.
+  Any walk in $I$ is also a walk in the supergraph $I^+$ with $V(I^+)=V(I)$ and $E(I^+)=B^+(I)\cup J^+(I)$, so it we can focus our attention on walks $\rho_0,\ldots,\rho_{\ell}$ in $I^+$ that satisify properties (i')--(iv').
+We will also call these $\delta$-walks.
+
+  We are almost done proving \eqref{one-round}.
+Refer to \figref{one-round}.
+For any landing $j_t$, $t>0$, $I^+[j_t,\ldots,d_0]$ contains the path $d_0,j_1,\ldots,j_t$ of length $t$.
+A few moments consideration will also convince the reader that $d_0,j_1,\ldots,j_t$ is a shortest path, in $I[j_{t+1}-1,\ldots,d_0]$, from $d_0$ to $j_t$.
+Therefore, by (iv') no $\delta$-walk can containd $j_t$ for any $t> k+1$.
+
+  \begin{figure}
+    \begin{center}
+      \includegraphics{figs/one-round}
+    \end{center}
+    \caption{The graph $I^+[\{j_{4}+1,\ldots,d_0\}]$.  The shortest path $d_0,j_1,\ldots,j_3$ is highlighted.}
+    \figlabel{one-round}
+  \end{figure}
+
+  Next, observe that, in $I^+[j_{t+1}+1,\ldots,d_0]$ every path from $d_0$ to some vertex  $j_{t}+i$, $i\ge 0$ contains $j_t$.
+Furthermore, the only edges of $I^+[j_{t+1}+1,\ldots,d_0]$ incident to vertices $j_t+i$, $i>0$ are boring edges.
+Therefore, for any $i\ge 0$, the shortest path from $d_0$ to $j_t+i$ in $I^+[j_{t+1}+1,\ldots,d_0]$ has length $t+i$.
+It follows that there are at most $k+2-t$ values $\delta\in\{j_{t+1},\ldots,j_t\}$ such that there exists a $\delta$-walk.
+
+  Therefore, the total number of $\delta\in\{d_1+1,\ldots,d_0\}$ such that there exists a $\delta$-walk is at most
   \[
-     |\{\delta\in\{d_1,\ldots,d_0-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \sum_{t=0}^{k+1} (k+2-t) = \sum_{t=1}^{k+2} t = \binom{k+2}{2} \enspace .
+     \sum_{t=0}^{k+1} (k+2-t) = \sum_{t=1}^{k+2} t = \binom{k+1}{2}
   \]
+  We are now very close to \eqref{one-round}, which asks for the number of $\delta$-walks for $\delta\in\{d_1,\ldots,d_0-1\}$.
+  The preceding formula counts the $d_0$-walk, but does not count the $d_1$ walk.  These two effects cancel each other, thus establishing \eqref{one-round}.
 
-  % \begin{figure}
-  %   \begin{center}
-  %     \includegraphics{figs/i-graph}
-  %   \end{center}
-  %   \caption{The graph $I$ used in the proof of \clmref{bag-size}}
-  %   \figlabel{i-graph}
-  % \end{figure}
+  Refer to \figref{nested}.
+  More generally, for $s>1$, let $d_s=\min\{i: V(F_{d_s}\cap Y_i)\neq\emptyset \}$.
 
-  More generally, for $s>1$, let $d_s=\min\{ i: V(F_{d_s}\cap Y_i)\neq\emptyset \}$.
-Consider some $\delta$-walk $v_0,\ldots,v_\ell$ for $\delta\in \{ d_{s},\ldots,d_{s-1}-1 \}$, for some $s>1$.
-Since $v_0\in V(N_{d_0}-F_{d_0})$ and $v_\ell\in F_{\delta+1}-F_\delta$, there exists some maximum value $i$ such that $v_i\in N_{d_{s-1}}-F_{d_{s-1}}$.
-Furthermore, $i\ge s-2$ because $v_1,\ldots,v_i$ contains a vertex from each of $F_{d_0},F_{d_1},\ldots,F_{d_{s-2}}$.
-But this means that we can treat $v_i,\ldots,v_{\ell}$ as a $\delta$-walk of length at most $\ell-s+2 \le k-s+3$ that begins in $N_{d_{s-1}}-F_{d_{s-1}}$.
-By the same argument used above,
-  \[  |\{\delta\in\{d_s,\ldots,d_{s-1}-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k-s+3}{2} \enspace .\]
+  \begin{figure}
+    \begin{center}
+      \includegraphics{figs/nested}
+    \end{center}
+    \caption{The nested cycles $F_{d_0}, F_{d_1},F_{d_2}\ldots$.}
+    \figlabel{nested}
+  \end{figure}
+
+  Consider some $\delta$-walk $v_0,\ldots,v_\ell$ for $\delta\in \{ d_{s+1},\ldots,d_{s}-1\}$.
+By (i) and (ii), $v_0\in V(N_{d_0}-F_{d_0})\subseteq V(N_{d_s}-F_{d_s})$ and $v_\ell\in F_{\delta+1}-F_\delta$.
+Therefore, there exists some maximum value $i$ such that $v_i\in V(N_{d_{s}}-F_{d_{s}})$.
+(In words, $i$ is the last time the walk is in the interior of $F_{d_s}$ before it leaves forever.)
+  Observe that $i\ge s-1$ because $v_0,\ldots,v_i$ contains a vertex from each of $F_{d_0},\ldots,F_{d_{s-1}}$.
+
+  Now, consider the subwalk $v_i,\ldots,v_\ell$.  This subwalk obeys Properties~(i)--(iv) with respect to $F_{d_s}$.  Specifically:
+  \begin{compactenum}[(i)]
+    \item $v_i\in V(N_{d_s}-F_{d_s})$;
+
+    \item $v_i,\ldots,v_{\ell-1}\in V(N_{a_{\delta+1}}-F_{a_{\delta+1}})$;
+
+    \item $v_\ell\in V(F_{a_{\delta+1}}-F_{a_\delta})$; and
+
+    \item $\ell\le k+1$, so $\ell-i \le \ell-s \le \le k+1-s$.
+  \end{compactenum}
+  Exactly the same analysis used to establish \eqref{one-round} shows that
+  \begin{equation}
+     |\{\delta\in\{d_{s+1},\ldots,d_s-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+1-s}{2} \enspace . \eqlabel{one-round-s}
+  \end{equation}
 
@@
   \[
-      \sum_{s=1}^{k+2} \binom{k-s+3}{2} = O(k^3)
+      \sum_{s=0}^{k+1} \binom{k+1-s}{2}
+      = \sum_{s=0}^{k+1} \binom{s}{2}
+      = \frac{k^3}{6} + \frac{k^2}{2} + \frac{k}{3} \ge |B_x| \enspace . \qedhere
   \]
-  \note{PM}{Gotta run!}
-
-  % 
-  % 
-  % Next observe that, if some $\delta$-walk
-  % 
-  % 
-  % 
-  % 
-  % For each $v'\in V(G^+)$, define
-  % \[  
-  %     d_{v'}=\max\{i\in\{0,\ldots,d\} : v'\in V(N_{a_i}-F_{a_i})
-  % \]
-  % (Intuitively, $d_{v'}$ is the $K$-depth of the lower endpoint of $K[P_{v'}]$ or the $K$-depth ($d$) of $x$, whichever is shallower.)
-  % 
-  % Consider some $\delta$-walk $v_0,\ldots,v_{\ell}$ and, for each $j\in\{0,\ldots,\ell\}$, let $d_j=d_{v_j}$.
-Then the sequence $d_0,\ldots,d_{\ell}$ has the following properties, corresponding to properties (i)--(iii) above:
-  % \begin{compactenum}[(i)]
-  %   \item $d_0=d$;
-  %   \item $d_0,\ldots,d_{\ell-1} > \delta$
-  %   \item $d_\ell = \delta$;
-  %   \item $\ell\le k+1$.
-  % \end{compactenum}
-  % 
-  % Consider an auxilliary graph $I$ with vertex set $V(I)=\{0,\ldots,d\}$ and edge $ij\in E(I)$ if and only if $V(F_{a_i}\cap Y_j))\neq\emptyset$. Observe that $d_0,\ldots,d_\ell$ is a walk in $I$.
-  % 
-  % We are particularly interested in the \emph{up neighbours} of $i$, defined as
-  % $U(i) = \{q : iq\in E(I),\, q< i-1\}$.  
-  % 
-  % Claim: For each $j\in\{i_1-2,\ldots,d\}$,  $U(j) = U(d)=\{i_1,i_2\}$.
-  % 
-  % Claim: $U(i_1)= \{i_2, i_3\}$ for some $i_3< i_1-1$.
-  % 
-  % Claim: $U(i_1+1)= \{i_1, i_2, i_3\}$.
-  % 
-  %  U(i)\cup\{i-2\}\setminus\{i-1\}$. 
-  % 
-  % Let $i_0=d$, $i_1=\min U(d)$, and $i_2=U(d)$.  
-  % 
-  % 
-  % 
-  % By \clmref{strong-ancestor-boundary}, $I(j)=U(d)=\{i_1,i_2\}$ for all $j\in\{i_1-1,\ldots,d\}$.
-This immediately implies that there can be no $\delta$-walk for any $\delta\in \{ i\in\Z : k+1 < i < i_1 \}$.
-Indeed, such a walk either has to contain $i_1$ or $i_2$ or it has to contain $d,d-1,\ldots,d-k-2$.
-In the former case, the walk violates (ii) or (iii) since it contains values smaller than $\delta$.
-In the latter case, the walk violates (iv) since it has length $k+2>k+1$.
-  % 
-  % Next, consider $\max U(i_1)$ and let $i_3=\max U(i_1)$.  By \propref{strong-ancestor-boundary} 
-  % 
-  % 
-  % 
-  %  since $\delta > i_1 > i_2$ 
-  % 
-  % Claim: For all $j\in\{\min(U(i)),\ldots,i\}$, $U(j) = U(i)$. 
-  % 
-  % Proof: 
-  % The proof is by induction on $i-j$.  The base case $i=j$ is immediate.
-  % 
-  % Now consider some $j\in\{\min(U(i)), i-1\}$.  By the inductive hypothesis $U(j+1)=U(i)$ so $F_{j+1}$ contains vertices of $Y_q$ for each $q\in U(i)$. 
-  % These vertices are not in $Y_j$ since, by the definition of $U(j+1)$, $q < (j+1)-1 = j$ for each $q\in U(j+1)$.  
-  % 
-  % 
-  %  Recall that $F_{j+1}$ is a wedge in the Mercedes graph $M_j$. This wedge is determined by the closed tripod $\overline{Y}_j$
-  % 
-  % First the base case $j=i$:
-  % 
-  % Observe that $F_i$ is one wedge of the Mercedes graph $M_i$, so $F_i \subset F_{i-1}\cup \overline{Y}_{i-1}$.
-Therefore $V(F_i)$ can be partitioned into $A=V(F_i \cap F_{i-1})$ and $B=V(F_i\cap Y_{i-1})$.
-The set $B$ consists entirely of vertices in $N_{i-1}-F_{i-1}-(N_i-F_i)$.
-The set $A$ consists of vertices in $V(F_{i-1})$.
-By \propref{ancestor-boundary},
-  % 
-  % First observe that it is sufficient to prove this for $j=i-1$ since, in this case, we can apply
-  % Consider the vertices of $F_{i}$.  Each of these vertices belongs $Y_{q}$ for some $q\in U(i)$.
-  % 
-  % A critical property of these up edges is that, for any $i' > i$, $U(i') \subseteq \{i'+1,\ldots,i\} \cup U(i)$.  
-  % 
-  % 
-  % , there are at most 3 edges $ij\in E(I)$ with $j<i$.  Furthermore, for each $i'< i$, 
-  % 
-  % 
-  % 
-  % The graph $I$ has some additional structure, which we now describe.  
-  % First, consider the at most three neighbours $i_0>i_1>i_2$ of $d$ in $I$.
-  % In the walk $d_0,\ldots,d_{\ell}$, the first value $d_i\neq d$ is in $\{i_0,i_1,i_2\}$.  Furthermore, 
-  % 
-  % 
-  % 
-  % 
-  % First observe that, for each $i\in\{1,\ldots,d\}$, $(i,(i+1))\in E(I)$.  
-  % 
-  % 
-  % 
-  % $I$ has maximum out-degree $3$ and each directed edge $ij\in E(I)$ has $j < i$.
-  % 
-  % 
-  % 
-  % 
-  % 
-  % 
-  % Let $v_0,\ldots,v_\ell$ be an $i$-walk and consider the subsequence $d_0,\ldots,d_{\ell'}$ of $d_{v_0},\ldots,d_{v_\ell}$ consisting of only $d_{v_0}$ and those (record-breaking) $d_{v_j}$ such that $d_{v_j}<\min\{d_{v_0},\ldots,d_{v_{j-1}}\}$.
-Observe that $d_0=d$ and $d_{\ell'}=i$.
-  % 
-  % By \propref{ancestor-boundary}, for every $j\in\{1,\ldots,d\}$, there is a set $X_{j}$, of at most 3 integers $\alpha < j$ such that every vertex $q\in V(F_{a_j})$ is in $V(Y_{a_{\alpha}})$ for some $\alpha\in X_j$.
-  % Assign a distinct colour $c(j,\alpha)\in\{1,2,3\}$ for each $\alpha\in X_j$ and do this assignment for each $j\in\{1,\ldots,d\}$.
-By definition, for any $j\in\{1,\ldots,d\}$ and any colour $\gamma\in\{1,2,3\}$, there is at most one $\alpha=\alpha(j,\gamma)$ such that $c(j,\alpha)=\gamma$.
-  % 
-  % Finally, observe that, for every $j\in\{1,\ldots,\ell'\}$,  $c(d_j,d_{j+1})$ is defined.
-Refer to \figref{colouring}.
-Therefore the \emph{colour sequence} $c_0,\ldots,c_{\ell'-1}$ where $c_j=c(d_j,d_{j+1})$ is well defined.
-Since $d_0=d$ and $d_{\ell'-1}=i$, from the preceding discussion, we have $\alpha(\cdots (\alpha(\alpha(\alpha(d,c_0),c_1),c_2)\cdots),c_{\ell'-1}) = i$.
-  % 
-  % \begin{figure}
-  %   \begin{center}
-  %       \includegraphics{figs/A-2}
-  %   \end{center}
-  %   \caption{The colour sequences defined by two $i$-curves.}
-  %   \figlabel{colouring}
-  % \end{figure}
-  % 
-  % The proof is now almost complete.  Any $i$-walk defines a colour sequence $c_0,\ldots,c_{\ell'}$.  If some $i'$-walk defines the same colour sequence $c_0,\ldots,c_{\ell'}$ then
-  % \[
-  %       i = (\alpha(\alpha(\alpha(d,c_0),c_1),c_2)\cdots),c_{\ell'-1}) = i' \enspace ,
-  % \]
-  % so $i=i'$.  Since there are at most $3^{\ell'}$ colour sequences of length $\ell'$ and $\ell'\le \ell \le k+1$, the number of $i\in\{0,\ldots,d-1\}$ such that there exists an $i$-walk is at most
-  % \[
-  %    \sum_{\ell=1}^{k+1} 3^\ell = (3^{k+2}-3)/2 \ge |B_x| \eqlabel{bag-size} \enspace . \qedhere
-  % \]
 \end{proof}
 
-\note{PM}{I know the analysis in \clmref{bag-size} is not tight---the base $3$ of the exponent can be reduced.  Is it even exponential though?}
-We now have all the parts needed to complete the proof of \thmref{k-planar}.
-
-\begin{proof}
+\begin{proof}[Proof of \thmref{k-planar}]
   We use the layered $H$-partition $(\mathcal{L}, \mathcal{S})$, where $\mathcal{S}=\{S_x:x\in V(K)\}$ and $\mathcal{L}=\langle V_i': i=0,1,2\ldots\rangle$ where $V_i' = V_{(k+1)i}\cup\cdots\cup V_{(k+1)(i+1)-1}$.
@@
 
-  By \clmref{bag-size}, $\mathcal{S}$ produces a graph $H=G/\mathcal{S}$ of treewidth $(3^{k+2}-5)/2$.
+  By \clmref{bag-size}, $\mathcal{S}$ produces a graph $H=G/\mathcal{S}$ of treewidth at most $k^3/6+k^2/+k/3-1$.
 By \clmref{width-of-g-plus}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+12$, so $|B_x\cap V_i'|\le (k+1)(24k + 12) = 24k^2+36k+12$.

### 2019-05-16 k-planar.tex
Reading time.....

@@
 \begin{thm}\thmlabel{k-planar}
-  Every $k$-planar graph has a layered $H$-partition of layered-width at most $24k^2 + 36k + 12$ in which $H$ has treewidth at most $k^3/6+k^2/+k/3-1$.
+  Every $k$-planar graph has a layered $H$-partition of layered-width at most $24k^2 + 36k + 12$ in which $H$ has treewidth at most $k^3/6+k^2/2+k/3-1$.
 \end{thm}
@@
 \]
-\note{PM}{We might be able to save a factor of 2 here on the layered width, because we only need one endpoint of each edge $vw$.}Clearly $\mathcal{S}=\{S_x:x\in V(K)\}$ is a partition of $V(G)$.
+Clearly $\mathcal{S}=\{S_x:x\in V(K)\}$ is a partition of $V(G)$.
 We will use $\mathcal{S}$ as the partition that defines our quotient graph $H=G/\mathcal{S}$ with vertex set $V(H)=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
@@
 
-  By \clmref{bag-size}, $\mathcal{S}$ produces a graph $H=G/\mathcal{S}$ of treewidth at most $k^3/6+k^2/+k/3-1$.
+  By \clmref{bag-size}, $\mathcal{S}$ produces a graph $H=G/\mathcal{S}$ of treewidth at most $k^3/6+k^2/2+k/3-1$.
 By \clmref{width-of-g-plus}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+12$, so $|B_x\cap V_i'|\le (k+1)(24k + 12) = 24k^2+36k+12$.
@@
   \note{PM}{We can certainly save a factor of 2 here because we only need one endpoint of each $vw$ that crosses $Z$ to take part in $Z'$.}
-  Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \lemref{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 4\times (2k+1) = 16gk+8g$ for all $i\in\{1,\ldots,h\}$.
+  Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \lemlabel{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 4\times (2k+1) = 16gk+8g$ for all $i\in\{1,\ldots,h\}$.
 

### 2019-05-16 k-planar.tex
Pffft

@@
   Refer to \figref{A}(b)
-  If $B_x$ contains some $K$-ancestor $a_\delta$, $\delta\in\{0,\ldots,d-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $w\in S_{a_\delta}$ and $v\in S_{x'}$ for some $K$-descendant of $x$.
+  If $B_x$ contains some $K$-ancestor $a_\delta$, $\delta\in\{0,\ldots,d-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $w\in S_{a_\delta}$ and $v\in S_{x'}$ for some $K$-descendant $x'$ of $x$.
 The edge $vw$ corresponds to a walk $W_{vw}$ in $G^+$.
@@
 
-  In order to avoid excessive levels of subscripts, we introduce the following abbreviated notation: For $i\in\{0,\ldots,d\}$, we use $F_i$, $N_i$, and $Y_i$, as shorthands for $F_{a_i}$, $N_{a_i}$, and $Y_{a_i}$, respectively.
+  In order to avoid excessive levels of subscripts, we introduce the following abbreviated notation: For $i\in\{0,\ldots,d\}$, we use $F_i$, $N_i$, $Y_i$, and $M_i$ as shorthands for $F_{a_i}$, $N_{a_i}$, $Y_{a_i}$, and $M_{a_i}$, respectively.
 
-  Let $d_0=d$, and $d_1=\min\{i: V(F_d\cap Y_i)\neq\emptyset \}$.
-By \propref{ancestor-boundary}, $F_d$ is contained in the interior of  $F_{d_1}$, i.e., $N_{d_0}\subseteq N_{d_1}-F_{d_1}$.
-Refer to \figref{d0d1}.
-We will now prove that
+  Let $d_0=d$, and $d_1=\min\{i: V(F_d\cap Y_i)\neq\emptyset \}$.  By \propref{ancestor-boundary}, any vertex $z\in V(F_d)$ is contained in 
+  $Y_j$ for some $d_1\le j < d$.  Therefore, $F_d$ is contained in the interior of  $F_{d_1}$, i.e., $N_{d_0}\subseteq N_{d_1}-F_{d_1}$.
+
+  Refer to \figref{d0d1}.  We will now prove that 
   \begin{equation}
@@
   \begin{compactenum}
-    \item If $i=j$, then $ij$ is an edge of $I$ since $I$ contains the self-loop $ii$ for each $i\in V(I)$.
+    \item If $i=j$, then $ij\in E(I)$ since $ii\in E(I)$ for each $i\in V(I)$.
 
-    \item If $j\neq i$ then assume, without loss of generality, that $j < i$ so that $a_i$ is a $K$-descendant of $a_j$.
+    \item If $j\neq i$ then assume, without loss of generality, that $j < i$ so that $a_j$ is a $K$-ancestor of $a_i$.
 Now $v_t\in V(N_i-F_i)$, but $v_{t+1}\not\in V(N_i-F_i)$ otherwise we would have $j\ge i$.
 Therefore, $v_{t+1}\in V(F_i)$.
-By \propref{ancestor-boundary}, $v_{t+1}$ is in $Y_{j'}$ for some $j'< i$ and the edge $ij'\in E(I)$.
+By \propref{ancestor-boundary}, $v_{t+1}\in V(Y_{j'})$ for some $j'< i$ and therefore $ij'\in E(I)$.
 By \propref{children}, $v_{t+1}\not\in V(N_{j''}-F_{j''})$ for any $K$-descedant $j''>j'$ of $j'$.
-Therefore $j'=\max\{j : v_{t+1}\in N_j-F_j\}$.
-Therefore, by definition $j'=\rho(v_{t+1})=j$, so the edge $ij=ij'\in E(I)$.
+Therefore $j'=\max\{j : v_{t+1}\in V(N_j-F_j)\}$.
+By definition $j'=\rho(v_{t+1})=j$, so the edge $ij=ij'\in E(I)$.
   \end{compactenum}
@@
     \item begins at $\rho_0=d_0$ 
-    \item is contained in $I[\delta,\ldots,d_0]$, 
+    \item is contained in $I[\{\delta,\ldots,d_0\}]$, 
     \item ends at $\rho_\ell=\delta$
@@
   \end{compactenum}
+  We will call any walk in $I$ that satisfies (i')--(iv') a \emph{$\delta$-walk}.
+
   The edges in $E(I)$ can be partitioned into two classes:  A set $B(I)=\{ij\in E(I): |i-j|\le 1\}$ of \emph{boring edges} and a set $J(I)=\{ij\in E(I): |i-j|>1\}$ of \emph{jump edges}.
 For a jump edge $ij$ with $i>j$, $j$ is called a \emph{landing} and $i$ is called a \emph{takeoff}.
-The jump edges of $I$ satisfy the following properties:
+We will show that the jump edges of $I$ satisfy the following properties:
   \begin{enumerate}[(J1)]
@@
 
-    \item \jlabel{nested} For every jump edge $ij\in E(I)$ with takeoff $i$ and landing $j$, and for every $q\in\{j+2,\ldots,i\}$, if $q$ is a takeoff some jump edge, then this edge is $qj$.
+    \item \jlabel{nested} For every jump edge $ij\in E(I)$ with takeoff $i$ and landing $j$, and for every $q\in\{j+2,\ldots,i\}$, if $q$ is a takeoff for some jump edge, then this edge is $qj$.
   \end{enumerate}
@@
 By \propref{children} $F_i$ is a wedge of the Mercedes graph $M_{i-1}$.
-Let $P_1,\ldots,P_3$ be the partition of $V(F_{i-1})$ guaranteed by \propref{ancestor-boundary}.
-Without loss of generality, we can assume that $P_1=V(Y_{d_1}\cap F_{i-1})$.
-Then, by \propref{tripod-feet}, $V(F_i)$ contains vertices from at most one of $P_2$ or $P_3$---say $P_2$---and suppose $P_2=V(F_{i-1}\cap Y_j)$ for some $j< i-1$.
-This accounts for a single jump edge $ij$ for which $i$ is the takeoff.
-All remaining vertices of $F_i$ are contained in $Y_{i-1}$, which only introduces the boring edge $(i,i-1)$.
-Thus we have established that jump edges satisfy \jref{single}.
+Let $P_1,\ldots,P_g$ be the partition of $V(F_{i-1})$ guaranteed by \propref{ancestor-boundary}.
+  By \propref{tripod-feet}, $V(F_i) = V(F_i)\cap(P_\alpha \cup P_j\cup V(Y_{i-1}))$ for some $\alpha,j\in \{1,\ldots,g\}$.
+Since $F_i$ contains $v\in V(Y_{d_1})$, at least one of $\alpha$ or $j$, say $\alpha$, is equal $d_1$.
+Since $d_1\not\in V(I)$ and the (possible) edge $(i,i+1)$ is a boring edge, the only (possible) jump edge with $i$ as takeoff is $ij$.
+This establishes \jref{single}.
 
   To establish that jump edges satisfy \jref{nested} we can use the same argument used to establish that each $F_i$ contains a vertex of $Y_{d_1}$ to establish that each $F_q$ contains a vertex of $Y_j$ and therefore $qj\in E(I)$.
-Since $q \ge j+2$, $qj$ is a jump edge and, by \propref{single}, it is the only jump edge with takeoff $q$.
+Since $q \ge j+2$, $qj$ is a jump edge and, by \jref{single}, it is the only jump edge with takeoff $q$.
 
-  Let $j_0=d$ and let $j_1>j_2>\cdots> j_p$ denote the sequence of landings, in decreasing order.  Then \jref{single} and \jref{nested} imply that
+  It will be helpful to consider an \emph{edge maximal} version, $I^+$, of $I$ that we now define.
+Let $j_0=d$ and let $j_1>j_2>\cdots> j_p$ denote the sequence of landings, in decreasing order.
+Then \jref{single} and \jref{nested} imply that
   \[
@@
   \]
-  and clearly $B(I)\subseteq B^+(I) := \{(i,i+1): i\in\{d_1+2,\ldots,d_0\}\}$.
-  Any walk in $I$ is also a walk in the supergraph $I^+$ with $V(I^+)=V(I)$ and $E(I^+)=B^+(I)\cup J^+(I)$, so it we can focus our attention on walks $\rho_0,\ldots,\rho_{\ell}$ in $I^+$ that satisify properties (i')--(iv').
-We will also call these $\delta$-walks.
+  and clearly $B(I)\subseteq B^+(I) := \{(i,j): i\in\{d_1+2,\ldots,d_0-1\},\, j\in\{i-1,i,i+1\}\}$.
+  Any walk in $I$ is also a walk in the supergraph $I^+$ with $V(I^+)=V(I)$ and $E(I^+)=B^+(I)\cup J^+(I)$, so we can focus our attention on $\delta$-walks in $I^+$.
 
@@
 Refer to \figref{one-round}.
-For any landing $j_t$, $t>0$, $I^+[j_t,\ldots,d_0]$ contains the path $d_0,j_1,\ldots,j_t$ of length $t$.
-A few moments consideration will also convince the reader that $d_0,j_1,\ldots,j_t$ is a shortest path, in $I[j_{t+1}-1,\ldots,d_0]$, from $d_0$ to $j_t$.
-Therefore, by (iv') no $\delta$-walk can containd $j_t$ for any $t> k+1$.
+For any landing $j_t$, $t>0$, $I^+[\{j_t,\ldots,d_0\}]$ contains the path $d_0,j_1,\ldots,j_t$ of length $t$.
+A few moments consideration will also convince the reader that $d_0,j_1,\ldots,j_t$ is a shortest path, in $I^+[\{j_{t+1}-1,\ldots,d_0\}]$, from $d_0$ to $j_t$.
+Therefore, by (iv') no $\delta$-walk contains $j_t$ for any $t> k+1$.
 
@@
 
-  Next, observe that, in $I^+[j_{t+1}+1,\ldots,d_0]$ every path from $d_0$ to some vertex  $j_{t}+i$, $i\ge 0$ contains $j_t$.
-Furthermore, the only edges of $I^+[j_{t+1}+1,\ldots,d_0]$ incident to vertices $j_t+i$, $i>0$ are boring edges.
-Therefore, for any $i\ge 0$, the shortest path from $d_0$ to $j_t+i$ in $I^+[j_{t+1}+1,\ldots,d_0]$ has length $t+i$.
+  Next, observe that every path in $I^+[\{j_{t+1}+1,\ldots,d_0\}]$ from $d_0$ to some vertex  $j_{t}+i$, $i\ge 0$, contains $j_t$.
+Furthermore, the only edges of $I^+[\{j_{t+1}+1,\ldots,d_0\}]$ incident to vertices $j_t+i$, $i>0$, are boring edges.
+Therefore, for any $i\ge 0$, the shortest path from $d_0$ to $j_t+i$ in $I^+[\{j_{t+1}+1,\ldots,d_0\}]$ has length $t+i$.
 It follows that there are at most $k+2-t$ values $\delta\in\{j_{t+1},\ldots,j_t\}$ such that there exists a $\delta$-walk.
@@
   \]
-  We are now very close to \eqref{one-round}, which asks for the number of $\delta$-walks for $\delta\in\{d_1,\ldots,d_0-1\}$.
-  The preceding formula counts the $d_0$-walk, but does not count the $d_1$ walk.  These two effects cancel each other, thus establishing \eqref{one-round}.
-
-  Refer to \figref{nested}.
-  More generally, for $s>1$, let $d_s=\min\{i: V(F_{d_s}\cap Y_i)\neq\emptyset \}$.
+  We are now very close to \eqref{one-round}, which asks for the number of $\delta\in\{d_1,\ldots,d_0-1\}$ such that there exists a $\delta$-walk.
+  The preceding formula counts $\delta=d_0$, but does not count $\delta=d_1$.  These two effects cancel each other, thus establishing \eqref{one-round}.
 
@@
 
+  Refer to \figref{nested}.
+For $s>1$, let $d_s=\min\{i: V(F_{d_s}\cap Y_i)\neq\emptyset \}$.
 Consider some $\delta$-walk $v_0,\ldots,v_\ell$ for $\delta\in \{ d_{s+1},\ldots,d_{s}-1\}$.
@@
 (In words, $i$ is the last time the walk is in the interior of $F_{d_s}$ before it leaves forever.)
-  Observe that $i\ge s-1$ because $v_0,\ldots,v_i$ contains a vertex from each of $F_{d_0},\ldots,F_{d_{s-1}}$.
+  Observe that $i\ge s$ because $v_0,\ldots,v_i$ contains $v_0\in N_{d_0}-F_{d_0}$ as well as a vertex from each of $F_{d_0},\ldots,F_{d_{s-1}}$.
 
@@
 
-    \item $\ell\le k+1$, so $\ell-i \le \ell-s \le \le k+1-s$.
+    \item $\ell\le k+1$, so $\ell-i \le \ell-s \le k+1-s$.
   \end{compactenum}

### 2019-05-17 k-planar.tex
Fixing an off-by-one error
Fixed off-by-one error added example for k=1

@@
 
+\newcommand{\Bagsize}{\ensuremath{\frac{k^3}{6} + \frac{3k^2}{2} + \frac{13k}{3} + 4}}
+\newcommand{\bagsize}{\ensuremath{k^3/6 + 3k^2/2 + 13k/3 + 4}}
+\newcommand{\treewidth}{\ensuremath{k^3/6 + 3k^2/2 + 13k/3 + 3}}
+
 \begin{document}
@@
 \begin{thm}\thmlabel{k-planar}
-  Every $k$-planar graph has a layered $H$-partition of layered-width at most $24k^2 + 36k + 12$ in which $H$ has treewidth at most $k^3/6+k^2/2+k/3-1$.
+  Every $k$-planar graph has a layered $H$-partition of layered-width at most $24k^2 + 36k + 12$ in which $H$ has treewidth at most $\treewidth$.
 \end{thm}
@@
 \begin{thm}\thmlabel{g-k-planar}
-  Every $(g,k)$-planar graph has a layered $H$-partition of layered-width at most $\max\{16gk^2 + 24gk + 8g, 24k^2 + 36k + 12\}$ in which $H$ has treewidth at most $(3^{k+2}-3)/2$.
+  Every $(g,k)$-planar graph has a layered $H$-partition of layered-width at most $\max\{16gk^2 + 24gk + 8g, 24k^2 + 36k + 12\}$ in which $H$ has treewidth at most $\treewidth$.
 
@@
 The remainder of the paper is organized as follows: In \secref{k-planar} we prove \thmref{k-planar}.  In \secref{1-planar} we prove \thmref{1-planar}.  
-In \secref{genus-g} we prove \thmref{g-k-planar}.
+In \secref{g-k-planar} we prove \thmref{g-k-planar}.
 
@@
 \begin{clm}\clmlabel{bag-size}
-  The tree decomposition $\{B_x: x\in V(K)\}$ of $H$ contains no bag larger than $k^3/6 + k^2/2 + k/3$.
+  The tree decomposition $\{B_x: x\in V(K)\}$ of $H$ contains no bag larger than $\bagsize$.
 \end{clm}
@@
   \begin{equation}
-     |\{\delta\in\{d_1,\ldots,d_0-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+1}{2} \enspace . \eqlabel{one-round}
+     |\{\delta\in\{d_1+1,\ldots,d_0\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+3}{2} \enspace . \eqlabel{one-round}
   \end{equation}
@@
     \end{center}
-    \caption{$F_{d_1},\ldots,F_{d_0}$ and the resulting graph $I$.  For readability, the edges $\{id_1:i\in\{d_1-2,\ldots,d_0\}$ are not shown.}
+    \caption{$F_{d_1},\ldots,F_{d_0}$ and the resulting graph $I$.} 
     \figlabel{d0d1}
@@
   \[
-     \sum_{t=0}^{k+1} (k+2-t) = \sum_{t=1}^{k+2} t = \binom{k+1}{2}
+     \sum_{t=0}^{k+1} (k+2-t) = \sum_{t=1}^{k+2} t = \binom{k+3}{2}
   \]
-  We are now very close to \eqref{one-round}, which asks for the number of $\delta\in\{d_1,\ldots,d_0-1\}$ such that there exists a $\delta$-walk.
-  The preceding formula counts $\delta=d_0$, but does not count $\delta=d_1$.  These two effects cancel each other, thus establishing \eqref{one-round}.
-
+  This establishes \eqref{one-round}.  An illustration of this analysis, showing the case $k=1$ is shown in \figref{one-round-example}.
   \begin{figure}
     \begin{center}
-      \includegraphics{figs/nested}
+      \includegraphics{figs/one-round-example}
     \end{center}
-    \caption{The nested cycles $F_{d_0}, F_{d_1},F_{d_2}\ldots$.}
-    \figlabel{nested}
+    \caption{The set of $6=\binom{k+3}{2}$ vertices in $I^+$ reachable by paths of walks of length at most $k+1$ (for $k=1$).}
+    \figlabel{one-round-example}
   \end{figure}
@@
   Observe that $i\ge s$ because $v_0,\ldots,v_i$ contains $v_0\in N_{d_0}-F_{d_0}$ as well as a vertex from each of $F_{d_0},\ldots,F_{d_{s-1}}$.
+
+  \begin{figure}
+    \begin{center}
+      \includegraphics{figs/nested}
+    \end{center}
+    \caption{The nested cycles $F_{d_0}, F_{d_1},F_{d_2}\ldots$.}
+    \figlabel{nested}
+  \end{figure}
 
@@
   \begin{equation}
-     |\{\delta\in\{d_{s+1},\ldots,d_s-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+1-s}{2} \enspace . \eqlabel{one-round-s}
+     |\{\delta\in\{d_{s+1}+1,\ldots,d_s\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+3-s}{2} \enspace . \eqlabel{one-round-s}
   \end{equation}
 
-  We conclude that the maximum number of $\delta\in\{0,\ldots,d-1\}$ such that there exists a $\delta$-walk is at most
-  \[
-      \sum_{s=0}^{k+1} \binom{k+1-s}{2}
-      = \sum_{s=0}^{k+1} \binom{s}{2}
-      = \frac{k^3}{6} + \frac{k^2}{2} + \frac{k}{3} \ge |B_x| \enspace . \qedhere
-  \]
+  We conclude the proof with 
+  \begin{align*}
+    |B_x| & \le |\{\delta\in\{0,\ldots,d\}:  \mbox{$\exists$ a $\delta$-walk} \}| \\
+      & = \sum_{s=0}^{k+1}|\{\delta\in\{d_{s+1}+1,\ldots,d_s\}:  \mbox{$\exists$ a $\delta$-walk} \}| \\
+      & \le\sum_{s=0}^{k+1} \binom{k+3-s}{2} \\
+      & = \sum_{s=2}^{k+3} \binom{s}{2} \\
+      & = \frac{1}{6}(k+2)(k+3)(k+4) \\
+      & = \frac{k^3}{6} + \frac{3k^2}{2} + \frac{13k}{3} + 4  \enspace . \qedhere
+  \end{align*}
 \end{proof}
+
+We now have all the pieces needed to complete the proof of \thmref{k-planar}.
 
@@
 
-  By \clmref{bag-size}, $\mathcal{S}$ produces a graph $H=G/\mathcal{S}$ of treewidth at most $k^3/6+k^2/2+k/3-1$.
+  By \clmref{bag-size}, $\mathcal{S}$ produces a graph $H=G/\mathcal{S}$ of treewidth at most $\treewidth$.
 By \clmref{width-of-g-plus}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+12$, so $|B_x\cap V_i'|\le (k+1)(24k + 12) = 24k^2+36k+12$.

### 2019-05-17 k-planar.tex
Resolving undefined references

@@
   We call any walk with properties (i)--(iv) a \emph{$\delta$-walk}.
-Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $\delta\in\{0,\ldots,d-1\}$ such that there exists a $\delta$-walk.
-This is our goal for the remainder of this proof.
+Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $\delta\in\{0,\ldots,d\}$ such that there exists a $\delta$-walk.\footnote{To account for the fact that $B_x$ contains $x$ we count $\delta=d$.
+A $d$-walk corresponds, for example, to a 0-length walk that begins and ends at any vertex $v_0\in V(N_x-F_x)$.}  This is our goal for the remainder of this proof.
 
@@
   \note{PM}{We can certainly save a factor of 2 here because we only need one endpoint of each $vw$ that crosses $Z$ to take part in $Z'$.}
-  Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \lemlabel{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 4\times (2k+1) = 16gk+8g$ for all $i\in\{1,\ldots,h\}$.
+  Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \clmref{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 4\times (2k+1) = 16gk+8g$ for all $i\in\{1,\ldots,h\}$.
 

### 2019-05-17 k-planar.tex
Cleaned-up 1-planar proof

@@
 The remainder of the paper is organized as follows: In \secref{k-planar} we prove \thmref{k-planar}.  In \secref{1-planar} we prove \thmref{1-planar}.  
-In \secref{g-k-planar} we prove \thmref{g-k-planar}.
+In \secref{g-k-planar} we prove \thmref{g-k-planar}. In \secref{consequences} we point out some consequences of these results for some of the other graph parameters mentioned above.
 
@@
 
-  Finally, let $M$ and $M^+$ denote the subgraph of $G$ containing the edges and vertices of $Y$, respectively $Y^+$, and the edges and vertices of $F$.
+  Let $M$ and $M^+$ denote the subgraph of $G$ containing the edges and vertices of $\overline{Y}$, respectively $\overline{Y}^+$, and the edges and vertices of $F$.
 The graph $M^+$ has some number of bounded faces, all contained in the interior of $F$.
@@
 
-  We claim that, for each $i\in\{1,\ldots,m\}$, the boundary of $F_i$ is a cycle in $G'$.
+  We claim that, for each $i\in\{1,\ldots,m\}$, the boundary of $F_i$ is a cycle in $G'$ that contains no spars.
 Otherwise, some edge $vw$ contributes to the boundary of $F_i$ but is crossed by an edge $xy\in E(G)$.
@@
   The remainder of the proof finishes in the same way as the proof of Lemma~14 in \cite{dujmovic.joret.ea:planar}.
-$P_1,\ldots,P_k$, and $V(Y^+)$ become elements of the $H$-partition.
+$P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
 Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
-The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{dujmovic.joret.ea:planar}.
+The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{dujmovic.joret.ea:planar}, except that now each node $x$ has a child for each face $F_i$ of $M^+_x$ that contains a vertex of $G$ in its interior.
 
@@
   \begin{enumerate}
-    \item $G/\mathcal{P}$ and $G'/\mathcal{P}$ are isomorphic.
-This follows immediately from the fact that, for each edge $xy\in E(G-G')$, $x,y\in S_a$ for some $a\in V(H)$.
-Indeed, $S_a=Y^+_a$ where $\overline{Y}_a$ contains the edge $vw\in E(G')$ that crosses $xy$.
+    \item $G/\mathcal{P}$ and $G^+/\mathcal{P^+}$ are isomorphic, where $G^+$ is the triangulation obtained by adding dummy vertices at each crossing in $G$ and $\mathcal{P}^+$ is the partition we obtain by adding a dummy vertex $z$ to $\overline{Y}^+$ if $\overline{Y}^+$ contains an edge $vw$ that contains $z$ in its interior.
 
-    \item $G[\overline{Y}^+-F]$ is connected.
+    \item $G^+[\overline{Y}^+-F]$ is connected.
 To see why this is so, first observe that $\overline{Y}-F$ is connected, and then observe that every vertex of $\overline{Y}^+$ is either a vertex of $\overline{Y}$ or adjacent to a vertex of $\overline{Y}$.
   \end{enumerate}
-  Since $G'$ is planar, this implies that $H=G'/\mathcal{P}$ is planar.\footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
-Repeatedly applying this fact implies that $\Delta/\mathcal{P}$ is planar provided that $\Delta[P]$ is connected for every $P\in\mathcal{P}$.}
+  Since $G^+$ is planar, the second point implies that $H=G^+/\mathcal{P}$ is planar.\footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
+Repeatedly applying this fact implies that $\Delta/\mathcal{P}$ is planar provided that $\Delta[P]$ is connected for every $P\in\mathcal{P}$.} The first point then ensures that $G/\mathcal{P}$ is also planar.
 \end{proof}
 
-Finally, we can complete the proof of \thmref{1-planar}.
+Using \lemref{induction}, the proof of \thmref{1-planar} is now straightforward.
 
 \begin{proof}[Proof of \thmref{1-planar}]
-  Using \lemref{induction}, the proof of \thmref{1-planar} is now straightforward.
 Given a 1-plane graph $G$, add edges to make it edge maximal so that it has an outer face $F=v_1v_2v_3$.
@@
 That this is a layering of $G$ follows from \eqref{spar-distance}.
-Since $|V_j\cap S_x|\le 15$ for every integer $i\ge 0$, $|V_i'\cap S_x|\le 30$ for every integer $i\ge 0$ and $x\in V(H)$.
+Since $|V_j\cap S_x|\le 15$ for every integer $i\ge 0$, $|V_i'\cap S_x|\le 30$ for every integer $i\ge 0$ and every $x\in V(H)$.
 \end{proof}
@@
 \seclabel{g-k-planar}
+
+Next we consider a generalization of our results to $(g,k)$-plane graphs.
 
@@
 
-% \section{Consequences}
-% \seclabel{consequences}
+\section{Consequences}
+\seclabel{consequences}
+
+\begin{enumerate}
+  \item $p$-centered colourings;
+  \item non-repetitive chromatic number;
+  \item connection with queue layouts;
+  \item connection with low-treewidth/treedepth colourings
+\end{enumerate}
 

### 2019-05-21 k-planar.tex
Testing
Polishing the (g,k)-planar result

@@
 \DeclareMathOperator{\ff}{f}
+\DeclareMathOperator{\tw}{tw}
 
@@
 
-A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle V_0,V_1,\ldots,V_h\rangle$ such that $\{V_0,V_1,\ldots,V_h\}$ is a partition of $V(G)$ and there is no edge $vw\in E(G)$ such that $v\in V_i$, $w\in V_j$ and $|j-i|>1$.
+A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle V_0,V_1,\ldots\rangle$ such that $\{V_0,V_1,\ldots\}$ is a partition of $V(G)$ and there is no edge $vw\in E(G)$ such that $v\in V_i$, $w\in V_j$ and $|j-i|>1$.
 For any partition $\mathcal{P}=\{S_1,\ldots,S_p\}$ of $V(G)$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
@@
 \begin{thm}\thmlabel{g-k-planar}
-  Every $(g,k)$-planar graph has a layered $H$-partition of layered-width at most $\max\{16gk^2 + 24gk + 8g, 24k^2 + 36k + 12\}$ in which $H$ has treewidth at most $\treewidth$.
+  Every $(g,k)$-planar graph has a layered $H$-partition of layered-width at most $\max\{8gk^2 + 12gk + 4g, 24k^2 + 36k + 12\}$ in which $H$ has treewidth at most $\treewidth$.
 
@@
 This assumption is not critical since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
-Consider the following plane graphs:
+Consider the following plane graphs (see \figref{wrapper}(a)):
 \begin{compactenum}
@@
 since (the curve) $vw$ contains a sequence $W_{vw}$ of at most $k+2$ vertices of $G^+$ that form a walk from $v$ to $w$ in $G^+$.
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c@{\hspace{2cm}}c}
+      \includegraphics{figs/wrapper-1} &
+      %\includegraphics{figs/wrapper-3} &
+      \includegraphics{figs/wrapper-3} \\
+      (a) & (b) 
+    \end{tabular}
+  \end{center}  
+  \caption{Wrapping a $K_4$ around $G^+$: (a)~the graph $G^{++}$
+  %, (b)~the spanning tree $T$ of $G^{++}$, 
+  and (b)~the modified version of $G^{++}$ used in the proof of \corref{k-planar}.}
+  \figlabel{wrapper}
+\end{figure}
 
@@
 We will use $\mathcal{S}$ as the partition that defines our quotient graph $H=G/\mathcal{S}$ with vertex set $V(H)=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
-First we argue that this partition has small width with respect to the layering $\{V_0,\ldots,V_h\}$ defined by $T$ (recall that this is a layering of $G^+$, not a layering of $G$).
+First we argue that this partition has small width with respect to the layering $\{V_0,V_1,\ldots\}$ defined by $T$ (recall that this is a layering of $G^+$, not a layering of $G$).
 
@@
 
-\noindent
-\framebox{
-\begin{minipage}{\textwidth}
-  \note{DW}{%
-    % I think we need to define layered treewidth, and
-    % give the result of Dujmovic-Eppstein-Wood that (g,k)-planar graphs
-    % have O(gk)  layered treewidth. This is the best previous structural
-    % description of k-planar graphs. State that layered partitions are
-    % richer than  layered treewidth, and lead to O(1) bounds on non-rep
-    % chromatic number and queue-number, instead of O(log n) bounds via
-    % layered treewidth.
-    I also think we should give explicit upper bounds on non-rep chromatic
-    number for k-planar graphs, and also for queue-number (especially if
-    they improve the bound in the FOCS / J.ACM paper).
+We conclude this section with a small extension of \thmref{k-planar} that is useful when studying $(g,k)$-planar graphs in \secref{g-k-planar}.
 
-    I think we should think about generalising the new theorem for
-    (g,k)-planar graphs.
+\begin{cor}\corlabel{k-planar}
+  Let $G$ be a $k$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
+Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \thmref{k-planar} and such that $V_i' = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\}$ for each integer $i\ge 0$.
+\end{cor}
 
-    We should add corollaries for d-map graphs, since they are $d^2$-planar.
-    }
+\begin{proof}
+  Without loss of generality, assume that $r_0$ is on the outer face of $G^+$ and let $v_1$ and $v_2$ be the other two vertices of the outer face of $G^+$.
+  The proof of \thmref{k-planar} begins with the graph $G^+$ and then creates a graph $G^{++}$ consisting of a $K_4$ containing $G^+$ in one of its inner faces, $F_0$, and six edges required to complete $G^{++}$ to a triangulation.
+Instead, modify $G^{++}$ by subdividing the edges joining $F_0$ to $v_1$ and $v_2$, as shown in \figref{wrapper}(b).
+In this modified version of $G^{++}$, we have $\dist_{G^++}(r,v)=\dist_{T}(r,v)=\dist_{G^+}(r_0,v)+2$
+  which leads to the layering $\langle V_0,V_1,\ldots\rangle$ of $G^+$ where $V_i=\{v\in V(G^+): \dist_{T}(r,v) = i\}$.
 
-    \note{PM}{
-    A couple of other things to look into: 
-    \begin{compactenum}
-      \item $p$-centered colourings
-      \item low treewidth colourings
-    \end{compactenum}
-    }
-\end{minipage}
-}
+  The rest of proof remains unchanged until the point where we derive the layering $\mathcal{L}=\langle V_0',V_1',\ldots\rangle$ of $G$.  At this point, we set
+  \[  
+    V_i' = \{v\in V(G): \floor{\dist_{T}(r,v)-2}=i\} 
+         = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\} \enspace ,
+  \]
+  for all integers $i\ge 0$.  This layering satisfies the conditions of the corollary.
+\end{proof}
+
 \section{$1$-Planar Graphs}
@@
 
-  Use the layering $\mathcal{L}=\langle V_0',\ldots,V_h'\rangle$ where $V_i'=V_{2i}\cup V_{2i+1}$ for each $i\in\{0,\ldots,h\}$.
+  Use the layering $\mathcal{L}=\langle V_0',V_1'\ldots\rangle$ where $V_i'=V_{2i}\cup V_{2i+1}$ for each integer $i\ge 0$.
 That this is a layering of $G$ follows from \eqref{spar-distance}.
@@
 
+Exactly the same argument used to establish \corref{k-planar} establishes the analagous result for \thmref{1-planar}:
+
+\begin{cor}\corlabel{1-planar}
+  Let $G$ be a $1$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
+Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \thmref{1-planar} and such that $V_i' = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/2}=i\}$ for each integer $i\ge 0$.
+\end{cor}
+
 \section{$(g,k)$-Planar Graphs}
@@
 
-Next we consider a generalization of our results to $(g,k)$-plane graphs.
+Next we prove \thmref{g-k-planar}, which generalizes of our results to $(g,k)$-plane graphs.  We make use of the following lemma due to \citet[Lemma~21]{dujmovic.joret.ea:planar}:
 
 \begin{lem} \lemlabel{make-planar}
-Let $G$ be a $(S,0)$-plane graph where $S$ is a surface with Euler genus $g$. 
-%For every BFS layering $(V_0,V_1,\dots)$ of $G$,
-%there is a set $Z\subseteq V(G)$ with at most $2g$ vertices in each layer $V_i$, such that $G-Z$ is planar. 
-%Moreover, there is a connected planar graph $G^+$ containing $G-Z$ as a subgraph, 
-%such that there is a BFS layering $(W_0,W_1,\dots)$ of $G^+$ and $W_{i} \cap V(G-Z) = V_i  \setminus Z$ for all $i\geq 0$. 
-For every BFS spanning tree $T$ of $G$ rooted at some vertex $r\in V(G)$ with corresponding BFS layering $(V_0,V_1,\dots)$, 
-there is a subgraph $Z\subseteq G$ with at most $2g$ vertices in each layer $V_i$, such that $G-V(Z)$ is planar. 
-Moreover, there is a connected planar graph $G^*$ containing $G-V(Z)$ as a subgraph, and 
-there is a BFS spanning tree $T^*$ of $G^*$ rooted at some vertex $r^*$ 
-with corresponding BFS layering $(W_0,W_1,\dots)$ of $G^*$, 
-such that  $W_{i} \cap V(G) \setminus V(Z) = V_i  \setminus V(Z)$ for all $i\geq 0$.
-% , and
-% $P \cap V(G) \setminus V(Z)$ is a vertical path in $T$ for every vertical path $P$ in $T^+$. 
+Let $G^+$ be a $(S,0)$-plane graph where $S$ is a surface with Euler genus $g$. 
+For each $r\in V(G^+)$ there is a vertex set $Z\subseteq V(G^+)$ such that $G^+-V(Z)$ is planar and there is a connected planar graph $G^*$ containing  $G^+-Z$ as a subgraph with a vertex $r^*\in V(G^*)$ such that, for every integer $i\ge 0$,
+\[  \{v\in V(G^*): \dist_{G^*}(r^*,v) = i\} \supseteq \{v\in V(G^+): \dist_{G^+}(r,v) = i\} \]
+and
+\[ |\{v\in V(G^+): \dist_{G^+}(r,v) = i\}| \le 2g \enspace . \]
+
 \end{lem}
@@
 Let $G^+$ be the $(S,0)$-plane graph obtained by adding a dummy vertex at each crossing in $G$.
-Let $r$ be any vertex in $V(G)$ and let $T$ be BFS-spanning tree of $G^+$ rooted at $r$ and let $\mathcal{L}=\rangle V_0,ldots,V_h\rangle$ be the resulting layering of $G^+$.
+Let $r_0$ be any vertex in $V(G)$, let $T$ be BFS spanning tree of $G^+$ rooted at $r_0$ and let $\langle V_0,V_1,\ldots\rangle$ be the resulting layering of $G^+$.
 
-  Apply \lemref{make-planar} to $G^{++}$ to obtain the set $Z\subseteq V(G^{++})$, the planar graph $G^*$, and the BFS spanning tree $T^*$ rooted at $r^*$.
+  Apply \lemref{make-planar} to $G^{+}$ to obtain the set $Z\subseteq V(G^{+})$ and the planar graph $G^*$ with special vertex $r^*$.
 For each edge $vw\in E(G)$, let $W_{vw}$ be the corresponding walk in $G^+$.
 The set $Z$ contains vertices in $V(G)$ as well as dummy vertices in $V(G^+-G)$.
-Create a new set $Z'\subseteq V(G)$ as
-  \[
-     Z' = \{ v : V(W_{vw})\cap Z\neq\emptyset \} \enspace .
-  \]
-  \note{PM}{We can certainly save a factor of 2 here because we only need one endpoint of each $vw$ that crosses $Z$ to take part in $Z'$.}
-  Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \clmref{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 4\times (2k+1) = 16gk+8g$ for all $i\in\{1,\ldots,h\}$.
+Create a new set $Z'\subseteq V(G)$ by taking, for each $vw\in E(G)$ such that $V(W_{vw})\cap Z\neq\emptyset$, one of the vertices $v$ or $w$.
+In this way, each vertex $z\in Z$ can be charged for at most two vertices in $v,x\in Z'$ and this happens only if $(z,vw,xy)$ is a crossing in $G$.
+  Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \clmref{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 2\times (2k+1) = 4g(2k+1)$ for each integer $i\ge 0$.
 
-  Furthermore, for every edge $vw\in E(G-Z')$, $\dist_{G^+-Z}(v,w)\le k+1$, since $W_{vw}\subseteq G^+-Z$.
+  % For every edge $vw\in E(G)$, $W_{vw}$ has length at most $k+1$.
+Therefore $\dist_{G^+}(v,w)\le k+1$ for each $v\in E(G)$.
+Since $W_{vw}\subseteq G^+-Z$ for each $vw\in E(G-Z')$,  $\dist_{G^+-Z}(v,w)\le k+1$ for each $vw\in E(G-Z')$.
 
@@
 Indeed, for every edge $vw\in E(G-Z')$, $G^+-Z$ contains a walk $W_{vw}$ that contains at most $k$ dummy vertices of $G^+-Z$.
-Therefore, replacing edge edge $vw\in E(G-Z')$ with the corresponding walk in $G^*$ gives a $k$-plane embedding of $G-Z'$.
+Therefore, replacing edge $vw\in E(G-Z')$ with the corresponding walk in $G^*$ gives a $k$-plane embedding of $G-Z'$.
+Call the resulting $k$-plane graph $G^*_0$.
 
-  Apply \thmref{k-planar} or (in the case $k=1$) \thmref{1-planar} to the $k$-plane embedding of $G-Z'$ to obtain a layered $H$-partition $(\mathcal{L}',\mathcal{P})$ of $G-Z'$.
-Though not explicitly in the statement of either theorem, the proof of each theorem shows that it is possible to choose a vertex $r^*\in G^*$ so that each layer $V'_i\in \mathcal{L}'$ is contained in at most $k+1$ consecutive layers of $\mathcal{L}$.
-Therefore, $|Z'\cap V_i'| \le (k+1)(16gk+8g) = 16gk^2 + 24gk + 8g$.
+  Let $G^*_1$ be the supergraph of $G^*_0$ obtained by adding every vertex and edge of $G^*$ that is not contained in some edge of $G^*_0$.
+Finally, let $G^{*+}_1$ be the triangulation obtained by adding edges to $G^*$ in such a way that $\dist_{G^*}(r^*,v)=\dist_{G^{*+}}(r^*,v)$ for every $v\in V(G^*)$.\footnote{The fact that $G^*$ can be completed to a triangulation $G^{*+}$ while preserving distances to $r^*$ follows from the fact that any face $F$ of $G^*$ with 4 or more vertices contains two non-adjacent vertices $x$ and $y$ such that $|\dist_{G^*}(r^*,x)-\dist_{G^*}(r^*,y)|\le 1$.
+Indeed, if $z$ is a vertex of $F$ that minimizes $\dist_{G^*}(r^*,z)$, then two neighbours $x$ and $y$ of $z$ have this property.} Now, observe that $G^*_1$ is a $k$-plane graph and $G^{*+}_1$ is obtained by adding dummy vertices at crossings in $G^*_1$ and triangulating the resulting graph.
 
-  We turn $\mathcal{P}$, the $H$-partition of $G-Z'$ into a partition $\mathcal{P}'=\mathcal{P}\cup \{Z'\}$.
-Clearly, the treewidth of $H'=G/\mathcal{P'}$ is at most one more than the treewidth of $H=G/\mathcal{P}$.
-Finally
+  Apply \corref{k-planar} or (in the case $k=1$) \corref{1-planar} to $G^*_1$ using $G^{*+}_1$ and the distinguished vertex $r^*$ to obtain a layered $H$-partition of $G^*_1$.
+Since $G^*_1\supseteq G^*_0$ and $G^*_0$ is isomorphic to $G-Z'$, this gives a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ of $G-Z'$ that satisfies the conditions of \corref{k-planar} or \corref{1-planar}, as appropriate.
+In particular, for every integer $i\ge 0$,
+  \begin{align*} 
+    V_i'& = \{v\in V(G-Z'):\floor{\dist_{G^*}(r^*,v)/(k+1)} = i\} \\
+        & = \{v\in V(G-Z'):\floor{\dist_{G^+}(r_0,v)/(k+1)} = i\} \enspace .
+  \end{align*}
+  It follows that $|V_i'\cap Z'|\le 4g(2k+1)(k+1) = 8gk^2+12gk+4g$.
+
+  We turn $\mathcal{P}$, the $H$-partition of $G-Z'$ into a partition $\mathcal{P}'=\mathcal{P}\cup \{Z'\}$ of $G$.
+That is, we place all vertices of $Z'$ in a common part.
+In the worst-case, the vertex of $H'=G/ \mathcal{P}'$ corresponding to $V(G)\cap Z'$ is adjacent to every vertex of $H=(G-Z')/\mathcal{P}$, so the treewidth of $H'$ is at most one more than the treewidth of $H$.
+
+  Summarizing, we obtain an $H'$-partition $(\mathcal{L},\mathcal{P}')$ of $G$ where, for each $P\in\mathcal{P}'$,
+  \[ |V_i' \cap P| \le \begin{cases} 8gk^2+12gk+4g & \text{if $P=Z'$} \\
+                                      30 & \text{if $P\neq Z'$ and $k=1$} \\
+                                  24k^2+36k+12 & \text{otherwise.} 
+                    \end{cases}
+  \]
+  and $H'$ has treewidth 
   \[
-      \max\{ |P\cap L| : P\in \mathcal{P}',\, L\in \mathcal{L}' \} \le
-      \max\{ 16gk^2 + 24gk + 8g, 24k^2 + 36k + 12 \} \enspace .
+      \tw(H') 
+        \le \begin{cases}
+              4 & \text{if $k=1$} \\
+              k^3/6 + 3k^2/2 + 13k/3+4 & \text{otherwise.}
+            \end{cases} \qedhere
   \]
-  \note{PM}{All of the above still needs work...}
 \end{proof}
@@
 
+Some possibilities
 \begin{enumerate}
   \item $p$-centered colourings;
+  \item clustered colourings;
   \item non-repetitive chromatic number;
@@
   \item connection with low-treewidth/treedepth colourings
+  \item An important open question is whether $k$-planar graphs have layered $H$-partitions whose layered width is upper bounded by a function of $k$ but in which the treewidth of $H$ is bounded by an absolute constant, independent of $k$.
+  \item Are our current $H$-partitions $K_{3,s}$-free for some $s=s(k)$?  If so, this is good news for clustered colouring.
 \end{enumerate}
+
+\noindent
+\framebox{
+\begin{minipage}{\textwidth}
+  \note{DW}{%
+    % I think we need to define layered treewidth, and
+    % give the result of Dujmovic-Eppstein-Wood that (g,k)-planar graphs
+    % have O(gk)  layered treewidth. This is the best previous structural
+    % description of k-planar graphs. State that layered partitions are
+    % richer than  layered treewidth, and lead to O(1) bounds on non-rep
+    % chromatic number and queue-number, instead of O(log n) bounds via
+    % layered treewidth.
+    I also think we should give explicit upper bounds on non-rep chromatic
+    number for k-planar graphs, and also for queue-number (especially if
+    they improve the bound in the FOCS / J.ACM paper).
+
+    I think we should think about generalising the new theorem for
+    (g,k)-planar graphs.
+
+    We should add corollaries for d-map graphs, since they are $d^2$-planar.
+    }
+
+    \note{PM}{
+    A couple of other things to look into: 
+    \begin{compactenum}
+      \item $p$-centered colourings
+      \item low treewidth colourings
+    \end{compactenum}
+    }
+\end{minipage}
+}
 

### 2019-05-22 k-planar.tex
Implemented (most of) David's comments

@@
 
-\title{\MakeUppercase{Layered $H$-Partitions of $k$-Planar Graphs}%
-    \thanks{This work was partly funded by NSERC and MRI.}}
-
-\author{Vida Dujmovi\'c%
-        \thanks{School of Computer Science and Electrical Engineering,
-                University of Ottawa},\,\,
-        Pat Morin%
-        \thanks{School of Computer Science, Carleton University},\,\, and
-        David R. Wood%
-        \thanks{School of Mathematical Sciences, Monash University}}
-
-\newcommand{\dual}[1]{{#1}^\star}
 \newcommand{\note}[2]{{\color{red}[#1:~#2]}}
@@
 \renewcommand{\propref}[1]{(PR\ref{prop:#1})}
-
 \newcommand{\jlabel}[1]{\label{j:#1}}
@@
 
+\title{\MakeUppercase{Layered Partitions of $k$-Planar Graphs}%
+    \thanks{This work was partly funded by NSERC and MRI.}}
+
+\author{Vida Dujmovi\'c%
+        \thanks{School of Computer Science and Electrical Engineering,
+                University of Ottawa},\,\,
+        Pat Morin%
+        \thanks{School of Computer Science, Carleton University},\,\, and
+        David R. Wood%
+        \thanks{School of Mathematics, Monash University}}
+
+%\renewcommand{\ge}{\geqslant}
+%\renewcommand{\le}{\leqslant}
+
 \begin{document}
@@
 \begin{abstract}
-  Layered $H$-partitions of graphs (with small layered width, in which $H$ has small treewidth) are a recently-introduced tool that have been used to solve longstanding problems on queue layouts, non-repetitive colouring, and 3-d graph drawing.
-Such partitions are known to exist for planar graphs and, more generally, bounded genus graphs.
-In the current paper, we prove that every $k$-planar graph has a layered $H$-partition of layered width $O(k^2)$ in which $H$ has treewidth $O(k^3)$.
+  Layered partitions of graphs (with small layered width, in which the quotient has small treewidth) are a recently-introduced tool that have been used to solve longstanding problems on queue layouts, non-repetitive colouring, and 3-d graph drawing.
+Such partitions are known to exist for planar graphs and, more generally, apex-minor-free graphs.
+In the current paper, we prove that every $k$-planar graph has a layered partition of layered width $O(k^2)$ in  which the quotient graph has treewidth $O(k^3)$.
 This is the first result of this type for a non-minor-closed class of graphs and implies that $k$-planar graphs have both queue number and non-repetitive chromatic number upper-bounded by a function of $k$.
-The former result was previously shown using a combination of $H$-partitions for planar graphs with \textit{ad hoc} methods.
+The former result was previously shown by combining layered partitions of planar graphs with \textit{ad hoc} methods.
 The latter result is new, for all $k\ge 1$.
@@
 A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle V_0,V_1,\ldots\rangle$ such that $\{V_0,V_1,\ldots\}$ is a partition of $V(G)$ and there is no edge $vw\in E(G)$ such that $v\in V_i$, $w\in V_j$ and $|j-i|>1$.
-For any partition $\mathcal{P}=\{S_1,\ldots,S_p\}$ of $V(G)$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
-To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an $H$-partition and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.\footnote{An alternative definition of a layering of $G$ is an $H$-partition of $G$ where $H$ is a path.}  A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
-The \emph{layered width} of $(\mathcal{H},\mathcal{P})$ is $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
+For any partition $\mathcal{P}=\{S_1,\ldots,S_p\}$ of $V(G)$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
+To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an \emph{$H$-partition} and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.\footnote{An alternative definition of a layering of $G$ is an $H$-partition of $G$ where $H$ is a path.}  A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
+The \emph{layered width} of $(\mathcal{L},\mathcal{P})$ is $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
 
 Layered $H$-partitions of small layered width in which $H$ has some additional property are a useful tool for a number of graph problems.
-\citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number, track number, layered treewidth, and 3-D grid drawing volume.
+One such useful property is small \emph{treewidth}, which we now define.
+A \emph{tree-decomposition} $\mathcal{K}$ of a graph $G$ consists of a tree $K$ and a collection $\mathcal{K}=\{B_x:x\in V(K)\}$ of subsets of $V(G)$ indexed by nodes of $K$ such that (i)~for every $vw\in E(G)$, there exists some $x\in V(K)$ with that $v,w\in B_x$; and (ii)~for every $v\in V(G)$, $K[x] = K[\{y: x\in B_y\}]$ is connected.
+The \emph{width} of the tree-decomposition $\mathcal{K}$ is $\max\{|B_x|:x\in V(K)\}-1$.
+The \emph{treewidth} $\tw(G)$ of a graph $G$ is the minimum integer $t$ such that $G$ has a tree-decomposition of width $t$.
+
+ \citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number, track number, and layered treewidth.
 \citet{dujmovic.esperet.ea:planar} add non-repetitive chromatic number to this list.
-\citet{dujmovic.joret.ea:planar} show that two common families of graphs have such $H$-partitions: planar graphs and (more generally) bounded-genus graphs.
+\citet{dujmovic.joret.ea:planar} show that two common families of graphs have such $H$-partitions: planar graphs and (more generally) apex-minor-free graphs.\footnote{A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar.
+A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
+\note{PM}{Help me finish this definition of apex-minor-free.}}
 
-In this paper we show that another family of graphs admits layered $H$-partitions of constant layered-width in which $H$ has constant treewidth: $k$-planar graphs, which we now define.
+In this paper we show that another family of graphs admits layered $H$-partitions of bounded layered-width in which $H$ has bounded treewidth: $k$-planar graphs, which we now define.
 An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve in a surface $S$ is a continuous function $f:[0,1]\to S$.
 The points $f(0)$ and $f(1)$ are called the \emph{endpoints} of the curve.
-When there is no danger of misunderstanding we treat a curve $f$ as a point set $\{f(t):0\le t\le 1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
+When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
 A \emph{crossing} in an embedded graph $G$ is a triple $(p,vw,xy)$ with $p\in\R^2$, $vw,xy\in E(G)$ and such that $p\in (vw\cap xy)\setminus\{v,w,x,y\}$.
@@
 
-Under these definitions $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
-Thus, like genus-$g$ graphs, $k$-planar graphs are a generalization of planar graphs (which are genus-$0$ graphs).
-Unlike genus-$g$ graphs, however, the family of $k$-planar graphs is not minor closed.
-A graph $G'$ obtained from a $k$-planar graph $G$ by edge deletions and edge contractions may or may not be $k$-planar.
+Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
+Thus, like genus-$g$ graphs,\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
+The \textit{Euler genus} of the non-orientable surface with $c$ cross-caps is $c$.
+The \textit{Euler genus} of a graph $G$ is the minimum integer $g$ such that $G$ embeds in a surface of Euler genus $g$.
+Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.} $k$-planar graphs are a generalization of planar graphs (which are genus-$0$ graphs).
+Unlike genus-$g$ graphs, however, the family of $k$-planar graphs is not minor-closed.\footnote{A class $\mathcal{G}$ of graphs is \emph{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$.
+A minor-closed class is \emph{proper} if it is not the class of all graphs.
+For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class.}  A graph $G'$ obtained from a $k$-planar graph $G$ by edge deletions and edge contractions may or may not be $k$-planar.
+Indeed, \citet{dujmovic.eppstein.ea:structure} construct 1-planar graphs that contain arbitrarily large complete graph minors.
 
@@
 \begin{thm}\thmlabel{k-planar}
-  Every $k$-planar graph has a layered $H$-partition of layered-width at most $24k^2 + 36k + 12$ in which $H$ has treewidth at most $\treewidth$.
+  Every $k$-planar graph has a layered $H$-partition of layered-width at most $24k^2 + 60k + 36$ in which $H$ has treewidth at most $\treewidth$.
 \end{thm}
@@
 
-These results imply that, for any constant $k$, $k$-planar graphs have all the graph parameters mentioned above upper-bounded by some constant $c_k$.
+These results imply that $k$-planar graphs have all the graph parameters mentioned above upper-bounded by some constant $c_k$.
 
 Prior to this work, the strongest structural characterization of $k$-planar graphs was in terms of layered treewidth, which we now define.
-A \emph{tree-decomposition} $\mathcal{K}$ of a graph $G$ is consists of a tree $K$ and a collection $\mathcal{K}=\{B_x:x\in V(K)\}$ of subsets of $V(G)$ indexed by nodes of $K$ such that (i)~for every $vw\in E(G)$, there exists some $x\in V(K)$ with that $v,w\in B_x$; and (ii)~for every $v\in V(G)$, $K[x] = K[\{y: x\in B_y\}]$ is connected.
 A \emph{layered tree-decomposition} $(\mathcal{L},\mathcal{K})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{K}$ of $G$.
@@
 \citet{dujmovic.eppstein.ea:structure} show that $k$-planar graphs have layered treewidth $O(k+1)$.
-This already implies that several of the graph parameters mentioned above, including queue-number and non-repetitive chromatic number are upper-bounded by $O(k\log n)$ for $n$-vertex $k$-planar graphs.
+This already implies that several of the graph parameters mentioned above, including queue-number and non-repetitive chromatic number are upper-bounded by $O(k\log n)$ for $n$-vertex $k$-planar graphs \cite{dujmovic.morin.ea:layered}.
 \thmref{k-planar} is a significant strengthening of this characterization.
@@
 The purpose of this section is to prove \thmref{k-planar}.
-In this section, all graphs are finite, simple, an undirected.
+In this section, all graphs are finite, simple, and undirected.
 For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$) we use $G[S]$ to denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.
@@
 We use $G'\subseteq G$ to denote subgraph containment, i.e., $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.
+\note{PM}{TODO: Hunt down occurrences of $G-G'$ and replace with $G-V(G')$.}
 
@@
 
-Let $T$ be a rooted tree rooted at some node $r\in V(T)$.
+Let $T$ be a tree rooted at some node $r\in V(T)$.
 The \emph{$T$-depth} of a node $x\in V(T)$ is equal to the length of the path from $x$ to $r$ in $T$.
-A $P$ path in $T$ is \emph{vertical} if, for each integer $i$, $P$ contains at most one vertex of $T$-depth $i$.
-The deepest vertex in a vertical path $P$ is called  $P$'s \emph{lower endpoint} and the shallowest vertex in $P$ is called $P$'s \emph{upper endpoint}.
-For any node $v\in V(T)$, we call each node on the path, in $T$, from $v$ to $r$ (including $v$ itself) a \emph{$T$-ancestor} of $v$.
+A path $P$ in $T$ is \emph{vertical} if, for each integer $i$, $P$ contains at most one vertex of $T$-depth $i$.
+The deepest vertex in a vertical path $P$ is called  the \emph{lower endpoint} of $P$ and the shallowest vertex in $P$ is called the \emph{upper endpoint} of $P$.
+For each node $v\in V(T)$, we call each node on the path, in $T$, from $v$ to $r$ (including $v$ itself) a \emph{$T$-ancestor} of $v$.
 If $v'$ is a $K$-ancestor of $v$, then $v$ is a \emph{$K$-descendant} of $v'$.
@@
 A \emph{triangulation} is a near-triangulation whose outer face is bounded by a 3-cycle.
-For any cycle $F$ in a near-triangulation $N$ we use $N\circledast F$ to denote the near-triangulation consisting of all the edges and vertices of $G$ that are contained in $F$ or its interior.
+For any cycle $F$ in a near-triangulation $N$ we use $N\circledast F$ to denote the near-triangulation consisting of all the edges and vertices of $N$ that are contained in $F$ or its interior.
+
+For a connected graph $G$ and two vertices $v,w\in V(G)$, we use $\dist_{G}(v,w)$ to denote the length of a shortest path, in $G$, with endpoints $v$ and $w$.
+For a vertex $r\in V(G)$, the \emph{BFS layering} of $G$ from $r$ is the layering $\langle V_0,V_1,\ldots\rangle$ of $G$ where $V_i=\{v\in V(G): \dist_{G}(r,v)=i\}$ for each integer $i\ge 0$.
+A \emph{BFS spanning tree} $T$ of $G$ rooted at $r$ is a spanning tree of $G$ rooted at $r$ where, for every $v\in V(G)\setminus\{r\}$, the parent $v'$ of $v$ in $T$ has $\dist_G(r,v')= \dist_G(r,v)-1$.
 
@@
 We call $Y=\overline{Y}-F$ the \emph{(open) tripod} of $N$ determined by $\tau$.
-The graph $M=\overline{Y}\cup F$ is called a \emph{Mercedes graph} and the (at most 3) inner faces of $M$ other than $\tau$ are called the \emph{wedges} of $M$.
+The graph $M=\overline{Y}\cup F$ is called a \emph{Mercedes graph} and the (at most three) inner faces of $M$ other than $\tau$ are called the \emph{wedges} of $M$.
 
-A \emph{PS-tree}  $K=K(\Delta, T, r)$ is a 3-ary tree, such that each non-root node $x\in V(K)$ is associated with several object (refer to \figref{K-node}):
+Let $r$ be a degree-3 node on the outer face of a triangulation $\Delta$.
+A \emph{PS-tree}  $K=K(\Delta, T, r)$ is a rooted tree in which each non-root node $x\in V(K)$ is associated with several objects (refer to \figref{K-node}):
 
@@
 
-The non-root nodes of the PS-tree $K$ (including the sub-root $r_K$) satisfy several conditions:
+The non-root nodes of the PS-tree $K$ (including the sub-root $r_K$) satisfy the following conditions:
 \begin{enumerate}[(PR1)]
@@
 
-  \item \proplabel{ancestor-boundary} For each non-root node $x\in V(T)$, there exists a partition of $V(F_x)$ into $g\le 3$ non-empty sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$.
+  \item \proplabel{ancestor-boundary} For each non-root node $x\in V(K)$, there exists a partition of $V(F_x)$ into $g\le 3$ non-empty sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$.
 
@@
 
-For any triangulation $\Delta$ in which the outer face has a vertex $r$ of degree 3 and for any BFS spanning-tree of $\Delta$ rooted at $r$, there exists a PS-tree $K=K(\Delta,T,r)$.
-Indeed, this tree is constructed (albeit implicitly) by \citet[Proof of Lemma~14]{dujmovic.joret.ea:planar}.
+For any triangulation $\Delta$ in which the outer face has a vertex $r$ of degree 3 and for any BFS spanning tree of $\Delta$ rooted at $r$, there exists a PS-tree $K=K(\Delta,T,r)$.
+Indeed, this tree is constructed (albeit implicitly) by \citet[Proof of Lemma~14]{dujmovic.joret.ea:planar}.\footnote{\citet{dujmovic.joret.ea:planar} do not require that $T$ be a BFS spanning tree, only that its root, $r$ have degree 3 in $\Delta$ and in $T$.}
 
@@
 Let $r$ be the vertex of the $K_4$ not incident on $F_0$ and let $T$ be a BFS spanning tree of $G^{++}$ rooted at $r$. 
-For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{T}(r,v)=i\}$ and observe that $\{V_0,V_1,\ldots,V_{h}\}$ is a layering of $G^+$ (but not of $G$).
+For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{T}(r,v)=i\}$ and observe that $\{V_2,V_3,\ldots,V_{h}\}$ is a layering of $G^+$ (but not of $G$).
 
@@
 
-For each vertex $z\in V(G^+)$, let $P_z=\{a\in V(K): z\in V(N_a-F_a)\}$.  
 By \propref{all-in}, $P_{z}$ contains the sub-root $r_K$ of $K$.
-By \propref{children}, $K[P_z]$ is a vertical path (whose lower endpoint is the unique node $x\in V(K)$ such that $z\in Y_x$).
-Therefore $K[P_z]$ is a vertical path in $K$ that contains $r_K$.
+By \propref{children}, any node $x\in P_z$ has at most one of its children in $P_z$ and any node $x'\not\in P_z$ has none of its children in $P_z$.
+Therefore $K[P_z]$ is a vertical path that contains $r_K$ (and whose lower endpoint is the unique node $x\in V(K)$ such that $z\in Y_x$).
 We are about to repeatedly use the fact that the common intersection of two or more vertical paths is a vertical path.
@@
 For each $z\in V(W_{vw})$, $K[P_{z}]$ is a vertical path in $K$ that contains $r_K$.
-Therefore $A_{vw}$ is a non-empty vertical path in $K$.
-For any vertex $v\in V(G)$ define $A_v=\bigcap_{vw\in E(G)} A_{vw}$.
-Again, $K[A_v]$ is a non-empty vertical path in $K$ that contains $r_K$.
+Therefore $K[A_{vw}]$ is a vertical path in $K$ that contains $r_K$.
+For each vertex $v\in V(G)$ define $A_v=\bigcap_{vw\in E(G)} A_{vw}$.
+Again, $K[A_v]$ is a vertical path in $K$ that contains $r_K$.
 We define $a(v)$ to be the lower endpoint of $K[A_v]$.
-For every node $x\in V(K)$, define the \emph{separator}
+In words, $a(v)$ is the lowest node in $K$ such that the interior of $F_{a(v)}$ contains all the edges of $G$ incident to $v$.
+
+For each node $x\in V(K)$, define the \emph{separator}
 \[
@@
 We will use $\mathcal{S}$ as the partition that defines our quotient graph $H=G/\mathcal{S}$ with vertex set $V(H)=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
-First we argue that this partition has small width with respect to the layering $\{V_0,V_1,\ldots\}$ defined by $T$ (recall that this is a layering of $G^+$, not a layering of $G$).
+First we argue that this partition has small width with respect to the layering $\{V_2,V_3,\ldots\}$ of $G^+$ defined by $T$ (recall that this is a layering of $G^+$, not a layering of $G$).
 
 \begin{clm}\clmlabel{chargeback-to-y}
-   For any $x\in V(K)$ and any vertex $v\in S_x$, there exists a vertex $v'\in Y_x$ such that $|\dist_T(r,v')-\dist_T(r,v)|\le k+1$.  
+   For any $x\in V(K)$ and any vertex $v\in S_x$, there exists an edge $vw\in E(G)$ and a vertex $v'\in Y_x\cap V(W_{vw})$ such that $|\dist_T(r,v')-\dist_T(r,v)|\le k+1$.
 \end{clm}
@@
   Since $v\in S_x$, the lower endpoint of $K[A_{vw}]$ is $x$ for some edge $vw\in E(G)$.
-This implies that $W_{vw}\subset N_x-F_x$ but $W_{vw} \not\subseteq N_{x'}-F_{x'}$ for any any child $x'$ of $x$ in $K$.
+This implies that $W_{vw}\subseteq N_x-F_x$ but $W_{vw} \not\subseteq N_{x'}-F_{x'}$ for every child $x'$ of $x$ in $K$.
 Therefore, \propref{children} and \propref{ancestor-boundary} imply that $W_{vw}$ contains a vertex $v'\in Y_x$.
-Since $G$ is $k$-planar and $v'\in V(W_{vw})$ this implies that $\dist_{G^+}(v,v') \le k+1$.
+Since $G$ is $k$-planar, the length of $W_{vw}$ is at most $k+1$.
+Therefore, $\dist_{G^{++}}(v,v') \le k+1$.
   By the triangle inequality,     
   \[
-    \dist_{G^+}(r,v') \le \dist_{G^+}(r,v) + \dist_{G^+}(v,v') 
-       \le \dist_{G^+}(r,v) + k + 1 \enspace .
+    \dist_{G^{++}}(r,v') \le \dist_{G^{++}}(r,v) + \dist_{G^{++}}(v,v') 
+       \le \dist_{G^{++}}(r,v) + k + 1 \enspace .
   \]
-  Similarly, $\dist_{G^+}(r,v) \le \dist_{G^+}(r, v') + k+1$.  Therefore
-  $|\dist_{G^+}(r,v)-\dist_{G^+}(r,v')| \le k+1$ which establishes the claim, since $T$ is BFS tree rooted at $r$, so $\dist_{T}(r,z)=\dist_{G^+}(r,z)$ for all $z\in V(G^+)$.
+  Similarly, $\dist_{G^{++}}(r,v) \le \dist_{G^{++}}(r, v') + k+1$.  Therefore
+  $|\dist_{G^+}(r,v)-\dist_{G^+}(r,v')| \le k+1$ which establishes the claim, since $T$ is BFS spanning tree of $G^{++}$ rooted at $r$, so $\dist_{T}(r,z)=\dist_{G^{++}}(r,z)$ for all $z\in V(G^+)$.
 \end{proof}
@@
 \begin{proof}
-  Since the vertices of $Y_x$ come from at most 3 vertical paths in $T$, $|V(Y_x)\cap V_i|\le 3$.
-\clmref{chargeback-to-y} shows that, for any $v\in S_x\cap V_i$, there exists $v'\in Y_x$ with $\dist_T(v,v')\le k+1$.
+  Since the vertices of $Y_x$ come from at most three vertical paths in $T$, $|V(Y_x)\cap V_i|\le 3$.
+\clmref{chargeback-to-y} shows that, for any $v\in S_x\cap V_i$, there exists $vw\in E(G)$ and $v'\in Y_x\cap V(W_{vw})$ with $\dist_T(v,v')\le k+1$.
 Therefore, $v'\in V(Y_x)\cap V_{j}$ for some $j\in\{i-k-1,\ldots,i+k+1\}$.
 If $v'=v$ (so $v'\in V_i$) then $v'$ contributes only one vertex to $S_x$.
-If $v'\neq v$ then $v'\in V(G^+-G)$ is a point where two edges of $G$ cross and $v'$ contributes at most the $4$ endpoints of these edges to $S_x$.
-In either case every $v'\in V(Y_x)\cap \bigcup_{j=i-k-1}^{i+k+1} V_j$ contributes at most $4$ vertices to $S_x\cap V_i$, so $|S_x\cap V_i|\le 3\times 4\times (2k+1) = 24k+12$.
+If $v'\neq v$ then $v'\in V(G^+-G)$ is a point where two edges of $G$ cross and $v'$ contributes at most the four endpoints of these edges to $S_x$.
+In either case each $v'\in V(Y_x)\cap \bigcup_{j=i-k-1}^{i+k+1} V_j$ contributes at most four vertices to $S_x\cap V_i$, so $|S_x\cap V_i|\le 3\times 4\times (2k+3) = 24k+36$.
 \end{proof}
@@
 
-The tree decomposition $(B_x:x\in V(K))$ of $H$ is defined as follows: For each edge $xy\in E(H)$ where $y$ is a $K$-ancestor of $x$, we add $y$ to $B_{x'}$ for every node $x'$ on the path from $x$ to $y$ in $K$.
+The tree decomposition $\{B_x:x\in V(K)\}$ of $H$ is defined as follows: For each edge $xy\in E(H)$ where $y$ is a $K$-ancestor of $x$, we add $y$ to $B_{x'}$ for every node $x'$ on the path from $x$ to $y$ in $K$.
 That this produces a tree-decomposition is due to \clmref{xy-ancestor}.
@@
 
-  Refer to \figref{A}(b)
+  Refer to \figref{A}(b).
 If $B_x$ contains some $K$-ancestor $a_\delta$, $\delta\in\{0,\ldots,d-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $w\in S_{a_\delta}$ and $v\in S_{x'}$ for some $K$-descendant $x'$ of $x$.
@@
   By \clmref{bag-size}, $\mathcal{S}$ produces a graph $H=G/\mathcal{S}$ of treewidth at most $\treewidth$.
-By \clmref{width-of-g-plus}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+12$, so $|B_x\cap V_i'|\le (k+1)(24k + 12) = 24k^2+36k+12$.
+By \clmref{width-of-g-plus}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+36$, so $|B_x\cap V_i'|\le (k+1)(24k + 36) = 24k^2+60k+36$.
 \end{proof}
@@
 
-  Summarizing, we obtain an $H'$-partition $(\mathcal{L},\mathcal{P}')$ of $G$ where, for each $P\in\mathcal{P}'$,
-  \[ |V_i' \cap P| \le \begin{cases} 8gk^2+12gk+4g & \text{if $P=Z'$} \\
-                                      30 & \text{if $P\neq Z'$ and $k=1$} \\
-                                  24k^2+36k+12 & \text{otherwise.} 
-                    \end{cases}
-  \]
-  and $H'$ has treewidth 
+  Summarizing, we obtain an $H'$-partition $(\mathcal{L},\mathcal{P}')$ of $G$ where $H'$ has treewidth 
   \[
@@
               4 & \text{if $k=1$} \\
-              k^3/6 + 3k^2/2 + 13k/3+4 & \text{otherwise.}
+              k^3/6 + 3k^2/2 + 13k/3+4 & \text{otherwise}
+            \end{cases} 
+  \]
+  and, for each $P\in\mathcal{P}'$,
+  \[ 
+    |V_i' \cap P| \le \begin{cases} 8gk^2+12gk+4g & \text{if $P=Z'$} \\
+                                    30 & \text{if $P\neq Z'$ and $k=1$} \\
+                                    24k^2+60k+36 & \text{otherwise.} 
                       \end{cases} \qedhere
@@
   \item $p$-centered colourings;
-  \item clustered colourings;
+  \item clustered colouring;
   \item non-repetitive chromatic number;
   \item connection with queue layouts;
-  \item connection with low-treewidth/treedepth colourings
+  \item $d$-map graphs, which are $d^2$-planar;
+  \item connection with low-treewidth/treedepth colourings;
   \item An important open question is whether $k$-planar graphs have layered $H$-partitions whose layered width is upper bounded by a function of $k$ but in which the treewidth of $H$ is bounded by an absolute constant, independent of $k$.
@@
 
-    I think we should think about generalising the new theorem for
-    (g,k)-planar graphs.
-
     We should add corollaries for d-map graphs, since they are $d^2$-planar.

### 2019-05-22 k-planar.tex
Added consequences for Thue and p-centered colourings

@@
 
+For a graph $G$ and a set $C$ of size $c$, any function $\phi\colon V(G)\to C$  is called a \emph{$c$-colouring} of $G$.
+
+\subsection{Non-Repetitive Colouring}
+
+A $c$-colouring $\phi$ of $G$ is \emph{non-repetitive} if, for every path $v_1,\ldots,v_{2h}$ in $G$, there exists some $i\in\{1,\ldots,k\}$ such that $\phi(v_i)\neq\phi(v_{i+h})$.
+The \emph{non-repetitive chromatic number} $\pi(G)$ of $G$ is the minimum value $c$ such that $G$ has a non-repetitive $c$-colouring.
+
+\citet[Corollaries~9 and 10]{dujmovic.esperet.ea:planar} prove the following result:
+
+\begin{lem}\lemlabel{non-repetitive}
+  If a graph $G$ has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$, then $\pi(G)\le \ell 4^{t+1}$.
+\end{lem}
+
+Combining \lemref{non-repetitive} with Theorems~\ref{thm:1-planar}--\ref{thm:g-k-planar} we immediately obtain the following corollaries:
+
+\begin{cor}
+  For every $k$-planar graph $G$, $\pi(G)\le 12(2k+3)(k+1)\times 4^{(1/6)(k+2)(k+3)(k+4)}$.
+\end{cor}
+
+\begin{cor}
+  For every $1$-planar graph $G$, $\pi(G)\le 30\times 4^4=7680$. 
+\end{cor}
+
+\begin{cor}
+  For every $(g,k)$-planar graph $G$, $\pi(G)\le \max\{4g(3k+1),12(2k+3)(k+1)\}\times 4^{(1/6)(k+2)(k+3)(k+4)}$.
+\end{cor}
+
+\begin{cor}
+  For every $(g,1)$-planar graph $G$, $g\ge 1$, $\pi(G)\le 48g\times 256= 12288g$.
+\end{cor}
+
+\subsection{Centered Colourings}
+
+A $c$-colouring $\phi$ of $G$ is $p$-centered if, for every connected subgraph $X\subseteq G$, $|\{\phi(v):v\in V(X)\}| > p$ or there exists some $v\in V(X)$ such that $\phi(v)\neq \phi(w)$ for every $w\in V(X)\setminus\{v\}$.
+In words, either $X$ receives more than $p$ colours or some vertex of $X$ receives a unique colour.
+Let $\chi_p(G)$ denote the smallest integer $c$ such that $G$ has a $p$-centered $c$-colouring.
+
+We make use of the following lemma due to \cite{pilipczuk.sieberts:polynomial-arxiv,pilipczuk.sieberts:polynomial-soda}
+\begin{lem}\lemlabel{p-centered-treewidth}
+  Any graph $H$ of treewidth at most $t$ has $\chi_p(H)\in O(p^t)$.
+\end{lem}
+
+The following lemma is due to Micek \cite{micek:personal}. We include the proof here for completeness.
+
+\begin{lem}\lemlabel{p-centered}
+  Any graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$, has $\chi_p(G)\le \ell\times (p+1)\times O(p^t)$.
+\end{lem}
+
+\begin{proof}
+  Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=\{B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$ where $H$ as treewidth at most $t$.
+  We use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,O(p^t)\}$.
+For each integer $i\ge 0$ and each $x\in V(H)$, the colour $\phi(v)=(\alpha(v),\beta(v),\gamma(v))$ of a vertex $v\in V_i\cap B_x$ is assigned to satisfy the following conditions:
+  \begin{enumerate}
+    \item $\alpha(v)$ is unique among $\{\phi(w): w\in V_i\cap B_x\}$.
+
+    \item $\beta(v)= i\bmod (p+1)$.
+
+    \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,O(p^t)\}$ is a $p$-centered colouring of $H$.
+  \end{enumerate}
+  Condition 1 is easy to satisfy since $|V_i\cap B_x|\le \ell$.
+Condition~2 is straightforward.
+Condition~3 is possible to satisfy thanks to \lemref{p-centered-treewidth}.
+To see why this is a $p$-centered colouring, consider some $X\subseteq V(G)$ such that $G[X]$ is connected.
+  \begin{enumerate}
+    \item If there exists $v,w\in X$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$ then, since $G[X]$ is connected, $G[X]$ contains a path from $v$ to $w$.
+This path contains at least one vertex from $V_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$.
+Therefore, $|\{\beta(v'):v'\in X\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
+
+    \item Otherwise, $X\subseteq V_{i},\ldots,V_{i+s}$ for some $s<p$.
+Let $H'=H[\{x\in V(H):B_x\cap X\neq\emptyset]$.
+If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in X\}|> p$ so $|\{\phi(v):v\in X\}|> p$ and we are done.
+Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
+    For any $v,w\in B_x$ with $v\neq w$, either $v,w\in V_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in V_{i'}$ and $w\in V_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
+Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.
+  \end{enumerate}
+\end{proof}
+
+Combining \lemref{p-centered} with Theorems~\ref{thm:1-planar}--\ref{thm:g-k-planar} immediately gives the following results, for every $p\ge 2$:
+
+\begin{cor}
+  For every $k$-planar graph $G$, $\chi_p(G)\le p^{O(k^3)}$.
+\end{cor}
+
+\begin{cor}
+  For every $1$-planar graph $G$, $\chi_p(G)\le O(p^4)$. 
+\end{cor}
+
+\begin{cor}
+  For every $(g,k)$-planar graph $G$, $\chi_p(G)\le O(gk^2)\times p^{O(k^3)}$. 
+\end{cor}
+
+\begin{cor}
+  For every $(g,1)$-planar graph $G$, $g\ge 1$, $\chi_p(G)\le O(gp^4)$.
+\end{cor}
+
 Some possibilities

### 2019-05-22 k-planar.tex
Pffft

@@
 
+Prior to the current work, the strongest upper bound on the non-repetitive chromatic number of $n$-vertex planar $k$-planar graphs was $O(k\log n)$ \cite{dujmovic.morin.ea:layered}.
+
 \subsection{Centered Colourings}
@@
 
-We make use of the following lemma due to \cite{pilipczuk.sieberts:polynomial-arxiv,pilipczuk.sieberts:polynomial-soda}
+We make use of the following lemma due to \citet{pilipczuk.siebertz:polynomial-arxiv,pilipczuk.siebertz:polynomial-soda}
 \begin{lem}\lemlabel{p-centered-treewidth}
@@
 Condition~3 is possible to satisfy thanks to \lemref{p-centered-treewidth}.
-To see why this is a $p$-centered colouring, consider some $X\subseteq V(G)$ such that $G[X]$ is connected.
+To see why this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
   \begin{enumerate}
-    \item If there exists $v,w\in X$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$ then, since $G[X]$ is connected, $G[X]$ contains a path from $v$ to $w$.
-This path contains at least one vertex from $V_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$.
+    \item If there exists $v,w\in V(X)$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$ then, since $G[X]$ is connected, $G[X]$ contains a path from $v$ to $w$.
+By the definition of layering, this path contains at least one vertex from $V_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$.
 Therefore, $|\{\beta(v'):v'\in X\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
@@
 Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.
+Therefore, every vertex in $B_x$ receives colour distinct from every other vertex in $X$.
   \end{enumerate}
@@
 
-Some possibilities
+Prior to the current work, the strongest upper bounds on the number of colours required for $p$-centered colourings of $k$-planar graphs used the fact that $k$-planar graphs have bounded expansion.
+\note{PM}{I didn't follow the discussion on the previous best-known bounds for $p$-centered colourings.
+Can someone fill me in?}
+
+\subsection{More Stuff}
+
 \begin{enumerate}
-  \item $p$-centered colourings;
   \item clustered colouring;
-  \item non-repetitive chromatic number;
   \item connection with queue layouts;

### 2019-05-23 k-planar.tex
Added a couple more applications

@@
 \begin{document}
+\begin{titlepage}
 \maketitle
@@
 \end{abstract}
+\end{titlepage}
 
+\tableofcontents
+\newpage
 \section{Introduction}
@@
 
+Here we discuss some of the consequences of Theorems~\ref{thm:1-planar}--\ref{thm:g-k-planar} for other graph parameters that can be bounded using layered partitions and other graph classes that are $k$-planar.
+
+\subsection{Applications to Colouring}
+
 For a graph $G$ and a set $C$ of size $c$, any function $\phi\colon V(G)\to C$  is called a \emph{$c$-colouring} of $G$.
 
-\subsection{Non-Repetitive Colouring}
+\subsubsection{Non-Repetitive Colouring}
 
@@
 
-\begin{cor}
+\begin{cor}\corlabel{first}
   For every $k$-planar graph $G$, $\pi(G)\le 12(2k+3)(k+1)\times 4^{(1/6)(k+2)(k+3)(k+4)}$.
@@
 
-\subsection{Centered Colourings}
+\subsubsection{Centered Colourings}
 
@@
 
-\begin{cor}
+\begin{cor}\corlabel{last}
   For every $(g,1)$-planar graph $G$, $g\ge 1$, $\chi_p(G)\le O(gp^4)$.
@@
 
+\subsection{Other Graph Classes}
+
+We conclude by pointing out two other graph classes, namely $(g,d)$-map graphs and $(g,\delta)$-string graphs that are $(g,k)$-planar for $k=O(d^2)$ and $k=\delta^2$, respectively.
+This implies that Theorems~\ref{thm:1-planar}--\ref{thm:g-k-planar} and Corollaries~\ref{cor:first}--\ref{cor:last} apply to $(g,d)$-map graphs and to $(g,\delta)$-string graphs.
+
+\subsubsection{Map Graphs}
+
+Map graphs are defined as follows.
+Start with a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
+Let $G$ be the graph whose vertices are the nations of $G_0$, where two vertices are adjacent in $G$ if the corresponding faces in $G_0$ share a vertex.
+Then $G$ is called a \emph{$(g,d)$-map graph}.
+% A $(0,d)$-map graph is called a (plane) \emph{$d$-map graph}; such graphs have been extensively studied \citep{FLS-SODA12,DFHT05,CGP02,Chen07,Chen01}.
+The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$ (see \citep{DEW17}).
+So $(g,d)$-map graphs provide a natural generalisation of graphs embedded in a surface.
+
+The following lemma comes from the natural drawing of a map graph obtained by positioning each vertex of $G$ inside the corresponding nation and each edge of $G$ as a curve passing through the corresponding vertex of $G_0$.
+See \citet[Page~4]{dujmovic.eppstein.ea:structure}.
+
+\begin{lem}
+  Every $(g, d)$-map graph is $(g, \floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
+\end{lem}
+
+\subsubsection{String Graphs}
+
+For a surface $S$ and a graph $G$, an \emph{$S$-string representation} $f$ of $G$ is obtained by representing each vertex of $v\in V(G)$ as a curve $f(v)$ in $S$ so that, for any pair of vertices $v,w\in V(G)$, the curves $f(v)$ and $f(w)$ intersect if and only if $vw\in E(G)$.
+If $G$ has maximum degree $\delta$, then an $S$-representation of $G$ is also called an $(S,\delta)$-string representation.
+A graph $G$ is a \emph{$(g,\delta)$-string graph} if it has an $(S,\delta)$-string representation where $S$ is a surface of Euler genus $g$.
+
+\begin{lem}
+  Every $(g,\delta)$-string graph $G$ is a $(g,\delta^2)$-planar graph.
+\end{lem}
+
+\begin{proof}
+  Consider some $(S,\delta)$-string representation, $f$, of $G$.
+We will describe an embedding of $G$ into $S$ that is an $(S,k)$-plane graph.
+Each vertex $v\in v(G)$ is embedded at $f(v)(0)$.
+For each edge $vw\in E(G)$ there is an intersection point $p$ between $f(v)$ and $f(w)$, i.e., $f(v)(t_0)=f(w)(t_1)=p$ for some $t_0,t_1\in[0,1]$.
+Embed the edge $vw$ very close to the curve whose image is $\{f(v)(t): 0\le t\le t_1\}\cup f(w)(t): 0 \le t\le t_2\}$.
+In this embedding, for every vertex $x\in V(G)$, every portion of the curve $f(x)$ between two crossing points has up to $\delta$ edges of $G$ embedded close to it.
+The embedding of any edge $vw\in E(G)$ crosses at most $\delta$ other curves $f(x_1),\ldots,f(x_d)$ and each of these curves $f(x_i)$ is used to embed at most $\delta$ edges incident to $x_i$.
+Therefore the embedding of any edge $vw\in E(G)$ crosses at most $\delta^2$ other edges.
+\end{proof}
+
 \subsection{More Stuff}
@@
 \begin{enumerate}
-  \item clustered colouring;
+  % \item clustered colouring;
   \item connection with queue layouts;
-  \item $d$-map graphs, which are $d^2$-planar;
+  % \item $d$-map graphs, which are $d^2$-planar;
   \item connection with low-treewidth/treedepth colourings;

### 2019-05-23 k-planar.tex
Some rewording

@@
 \begin{lem}
-  Every $(g,\delta)$-string graph $G$ is a $(g,\delta^2)$-planar graph.
+  Every $(g,\delta)$-string graph $G$ is a $(g,2\delta^2)$-planar graph.
 \end{lem}
@@
 Embed the edge $vw$ very close to the curve whose image is $\{f(v)(t): 0\le t\le t_1\}\cup f(w)(t): 0 \le t\le t_2\}$.
-In this embedding, for every vertex $x\in V(G)$, every portion of the curve $f(x)$ between two crossing points has up to $\delta$ edges of $G$ embedded close to it.
-The embedding of any edge $vw\in E(G)$ crosses at most $\delta$ other curves $f(x_1),\ldots,f(x_d)$ and each of these curves $f(x_i)$ is used to embed at most $\delta$ edges incident to $x_i$.
-Therefore the embedding of any edge $vw\in E(G)$ crosses at most $\delta^2$ other edges.
+The embedding of any edge $vw\in E(G)$ is contained in $f(v)$ and $f(w)$.
+Therefore the embedding of $vw$ crosses at most $2\delta$ other curves $f(x_1),\ldots,f(x_d)$.
+Each curve $f(x_i)$ is used to embed at most $\delta$ edges incident to $x_i$.
+Therefore the embedding of any edge $vw\in E(G)$ crosses at most $2\delta^2$ other edges.
 \end{proof}

### 2019-05-23 k-planar.tex
Fixed theorem numbering

@@
 
-Here we discuss some of the consequences of Theorems~\ref{thm:1-planar}--\ref{thm:g-k-planar} for other graph parameters that can be bounded using layered partitions and other graph classes that are $k$-planar.
+Here we discuss some of the consequences of Theorems~\ref{thm:k-planar}--\ref{thm:g-k-planar} for other graph parameters that can be bounded using layered partitions and other graph classes that are $k$-planar.
 
@@
 
-Combining \lemref{non-repetitive} with Theorems~\ref{thm:1-planar}--\ref{thm:g-k-planar} we immediately obtain the following corollaries:
+Combining \lemref{non-repetitive} with Theorems~\ref{thm:k-planar}--\ref{thm:g-k-planar} we immediately obtain the following corollaries:
 
@@
 
-Combining \lemref{p-centered} with Theorems~\ref{thm:1-planar}--\ref{thm:g-k-planar} immediately gives the following results, for every $p\ge 2$:
+Combining \lemref{p-centered} with Theorems~\ref{thm:k-planar}--\ref{thm:g-k-planar} immediately gives the following results, for every $p\ge 2$:
 
@@
 We conclude by pointing out two other graph classes, namely $(g,d)$-map graphs and $(g,\delta)$-string graphs that are $(g,k)$-planar for $k=O(d^2)$ and $k=\delta^2$, respectively.
-This implies that Theorems~\ref{thm:1-planar}--\ref{thm:g-k-planar} and Corollaries~\ref{cor:first}--\ref{cor:last} apply to $(g,d)$-map graphs and to $(g,\delta)$-string graphs.
+This implies that Theorems~\ref{thm:k-planar}--\ref{thm:g-k-planar} and Corollaries~\ref{cor:first}--\ref{cor:last} apply to $(g,d)$-map graphs and to $(g,\delta)$-string graphs.
 

### 2019-05-28 k-planar.tex
Uniform statements of lemmas
Small fixes

@@
 One such useful property is small \emph{treewidth}, which we now define.
-A \emph{tree-decomposition} $\mathcal{K}$ of a graph $G$ consists of a tree $K$ and a collection $\mathcal{K}=\{B_x:x\in V(K)\}$ of subsets of $V(G)$ indexed by nodes of $K$ such that (i)~for every $vw\in E(G)$, there exists some $x\in V(K)$ with that $v,w\in B_x$; and (ii)~for every $v\in V(G)$, $K[x] = K[\{y: x\in B_y\}]$ is connected.
+A \emph{tree-decomposition} $\mathcal{K}$ of a graph $G$ consists of a tree $K$ and a collection $\mathcal{K}=\{B_x:x\in V(K)\}$ of subsets of $V(G)$ indexed by nodes of $K$ such that (i)~for every $vw\in E(G)$, there exists some $x\in V(K)$ with $v,w\in B_x$; and (ii)~for every $v\in V(G)$, $K[x] = K[\{y: x\in B_y\}]$ is connected.
 The \emph{width} of the tree-decomposition $\mathcal{K}$ is $\max\{|B_x|:x\in V(K)\}-1$.
@@
 
- \citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered-width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number, track number, and layered treewidth.
+ \citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number, track number, and layered treewidth.
 \citet{dujmovic.esperet.ea:planar} add non-repetitive chromatic number to this list.
@@
 
-In this paper we show that another family of graphs admits layered $H$-partitions of bounded layered-width in which $H$ has bounded treewidth: $k$-planar graphs, which we now define.
+In this paper we show that another family of graphs admits layered $H$-partitions of bounded layered width in which $H$ has bounded treewidth: $k$-planar graphs, which we now define.
 An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve in a surface $S$ is a continuous function $f:[0,1]\to S$.
@@
 \begin{thm}\thmlabel{k-planar}
-  Every $k$-planar graph has a layered $H$-partition of layered-width at most $24k^2 + 60k + 36$ in which $H$ has treewidth at most $\treewidth$.
+  Every $k$-planar graph has a layered $H$-partition of layered width at most $24k^2 + 60k + 36$ in which $H$ has treewidth at most $\treewidth$.
 \end{thm}
@@
 A \emph{layered tree-decomposition} $(\mathcal{L},\mathcal{K})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{K}$ of $G$.
-The layered-width of $(\mathcal{L},\mathcal{K})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{K}\}$.
-The \emph{layered treewidth} of $G$ is the minimum layered-width of any layered tree-decomposition of $G$.
+The layered width of $(\mathcal{L},\mathcal{K})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{K}\}$.
+The \emph{layered treewidth} of $G$ is the minimum layered width of any layered tree-decomposition of $G$.
 
@@
 \begin{thm}\thmlabel{g-k-planar}
-  Every $(g,k)$-planar graph has a layered $H$-partition of layered-width at most $\max\{8gk^2 + 12gk + 4g, 24k^2 + 36k + 12\}$ in which $H$ has treewidth at most $\treewidth$.
-
-  Every $(g,1)$-planar graph has a layered $H$-partition of layered-width at most $\max\{48g, 30\}$ in which $H$ has treewidth at most $4$.  
+  Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most $\max\{8gk^2 + 12gk + 4g, 24k^2 + 36k + 12\}$ in which $H$ has treewidth at most $\treewidth$.
+
+  Every $(g,1)$-planar graph has a layered $H$-partition of layered width at most $\max\{48g, 30\}$ in which $H$ has treewidth at most $4$.  
 \end{thm}
@@
 % 
-% Therefore $(\mathcal{L},\mathcal{P})$ is a layered $H$-partition of $\Delta-\{r\}$ with layered-width 3 in which $H$ has treewidth 3.
+% Therefore $(\mathcal{L},\mathcal{P})$ is a layered $H$-partition of $\Delta-\{r\}$ with layered width 3 in which $H$ has treewidth 3.
 Furthermore, for each $P\in\mathcal{P}$, $\Delta[P]$ is connected.
@@
 
-A $c$-colouring $\phi$ of $G$ is $p$-centered if, for every connected subgraph $X\subseteq G$, $|\{\phi(v):v\in V(X)\}| > p$ or there exists some $v\in V(X)$ such that $\phi(v)\neq \phi(w)$ for every $w\in V(X)\setminus\{v\}$.
+A $c$-colouring $\phi$ of $G$ is \emph{$p$-centered} if, for every connected subgraph $X\subseteq G$, $|\{\phi(v):v\in V(X)\}| > p$ or there exists some $v\in V(X)$ such that $\phi(v)\neq \phi(w)$ for every $w\in V(X)\setminus\{v\}$.
 In words, either $X$ receives more than $p$ colours or some vertex of $X$ receives a unique colour.
@@
 \begin{lem}
-  Every $(g,\delta)$-string graph $G$ is a $(g,2\delta^2)$-planar graph.
+  Every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.
 \end{lem}

### 2019-05-28 k-planar.tex
Small change

@@
 We will use $\mathcal{S}$ as the partition that defines our quotient graph $H=G/\mathcal{S}$ with vertex set $V(H)=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
-First we argue that this partition has small width with respect to the layering $\{V_2,V_3,\ldots\}$ of $G^+$ defined by $T$ (recall that this is a layering of $G^+$, not a layering of $G$).
+First we argue that this partition has small layered width with respect to the layering $\{V_2,V_3,\ldots\}$ of $G^+$ defined by $T$ (recall that this is a layering of $G^+$, not a layering of $G$).
 

### 2019-06-17 k-planar.tex
Added proof for k-NN graphs

@@
 
-We conclude by pointing out two other graph classes, namely $(g,d)$-map graphs and $(g,\delta)$-string graphs that are $(g,k)$-planar for $k=O(d^2)$ and $k=\delta^2$, respectively.
-This implies that Theorems~\ref{thm:k-planar}--\ref{thm:g-k-planar} and Corollaries~\ref{cor:first}--\ref{cor:last} apply to $(g,d)$-map graphs and to $(g,\delta)$-string graphs.
+We conclude by pointing out three other graph classes, namely $(g,d)$-map graphs, $(g,\delta)$-string graphs, and $k$-nearest-neighbour graphs that are $(g,k)$-planar for $k=O(d^2)$ and $k=\delta^2$, respectively.
+This implies that Theorems~\ref{thm:k-planar}--\ref{thm:g-k-planar} and Corollaries~\ref{cor:first}--\ref{cor:last} apply to $(g,d)$-map graphs, $(g,\delta)$-string graphs, and $k$-nearest-neighbour graphs.
 
@@
 Therefore the embedding of any edge $vw\in E(G)$ crosses at most $2\delta^2$ other edges.
+\end{proof}
+
+\section{$k$-Nearest-Neighbour Graphs}
+
+In this section, we show that $k$-Nearest neighbour graphs of point in the plane are $O(k^2)$-planar.
+For two points $x,y\in\R^2$, let $d_2(x,y)$ denote the Euclidean distance between $x$ and $y$.
+The $k$-nearest-neighbour graph of a point set $P\subset\R^2$ is the geometric graph $G$ with vertex set $V(G)=P$ and edge set
+\[  
+  E= \{vw : v,w\in V(G),\, |\{x\in P\setminus\{v\}: d_2(v,x)\le d_2(v,w)\}|\le k\} \enspace .
+\] 
+(The edges of $G$ are straight-line segments joining their endpoints.)
+
+The following result, which is immediate from \citet[Corollary~4.2.6]{abrego.munroy.ea:on} states that $k$-nearest-neighbour graphs have bounded maximum degree:
+\begin{lem}\lemlabel{k-nn-max-degree}
+  The degree of any vertex in a $k$-nearest-neighbour graph is at most $6k$.
+\end{lem}
+
+We make use of the following well-known observation (see e.g., \citet[Lemma~2]{bose.morin.ea:routing}):
+\begin{obs}\obslabel{convex}
+  If $v_0,\ldots,v_3$ are the vertices of a convex quadrilateral in counterclockwise order then there exists at least one $i\in\{0,\ldots,3\}$ such that $\max\{d_2(v_i,v_{i-1}), d_2(v_i,v_{i+1})\} < d_2(v_{i-1},v_{i+1})$, where subscripts are taken modulo 4.
+\end{obs}
+
+\begin{lem}
+  Every $k$-nearest-neighbour graph is $O(k^2)$ planar.
+\end{lem}
+
+\begin{proof}
+  Let $G$ be a $k$-nearest-neighbour graph and consider any edge $vw\in E(G)$.
+  Let $xy\in E(G)$ be an edge that crosses $vw$.  Note that $vxwy$ are the vertices of a convex quadrilateral in (without loss of generality) counterclockwise order. Then we say that 
+  \begin{compactenum}
+    \item $xy$ is of Type~$v$ if $\max\{d_2(v,x), d_2(v,y)\}< d_2(x,y)$;
+    \item $xy$ is of Type~$w$ if $\max\{d_2(w,x), d_2(w,y)\}< d_2(x,y)$; or
+    \item $xy$ is of Type~C otherwise.
+  \end{compactenum}
+  If $xy$ is of Type~C, then \obsref{convex} implies that (without loss of generality) $\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)$.
+In this case, we call $x$ a Type~C vertex.
+We claim that $V(G)$ contains at most $k-1$ Type~C vertices.
+Indeed, more than $k-1$ Type~C vertices would contradict the fact that $vw\in E(G)$ since every Type~C vertex is closer to both $v$ and $w$ than $d_2(v,w)$.
+
+  Next oberve that, if $xy$ is of Type~$v$, then at least one of $xv$ or $yv$ is in $E(G)$ in which case we call $x$ (respectively $y$) a Type~$v$ vertex.
+By \lemref{k-nn-max-degree}, there are at most $6k$ Type~$v$ vertices.
+Similarly, there are at most $6k$ Type~$w$ vertices.
+
+  Thus, in all, there are at most $13k-1$ Type~$v$, Type~$w$, and Type~C vertices in total.
+By \lemref{k-nn-max-degree}, each of these vertices is incident on at most $6k$ edges that cross $vw$.
+Therefore, there are at most $78k^2-6k$ edges of $G$ that cross $vw$.
+Since this is true for every edge $vw\in E(G)$, $G$ is $(78k^2-6k)$-planar.
 \end{proof}

### 2019-06-17 k-planar.tex
Added a note about tightness.

@@
 
+We note that \lemref{nearest-neighbour} is tight up to the leading constant:  Every $n$-vertex nearest neighbour graph has at least $kn/2$ at at most $kn$ edges.
+For $k\ge 8$, the Crossing Lemma \cite{X} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is at $\Omega(k^2)$.
+
 \subsection{More Stuff}

### 2019-06-17 k-planar.tex
Started writing intro

--- 2019-06-17 k-planar.tex
@@
 
-\title{\MakeUppercase{Layered Partitions of $k$-Planar Graphs}%
+\title{\MakeUppercase{The Structure of $k$-Planar Graphs}%
     \thanks{This work was partly funded by NSERC and MRI.}}
@@
 
+In 1965, \citet{ringel:sechsfarbenproblem} introduced the notion of $1$-planar graphs, a generalization of planar graphs that allows crossings provided that no edge is involved in more than one crossing.
+This simple generalization is a rich source of problems and results.
+Indeed, the annotated bibliography on 1-planar graphs by \cite{kobourov.liotta.ea:annotated} contains more than 140 entries.
+
+The natural generalization of $1$-planar graphs, $k$-planar graphs (that allow crossings so long that no edge is involved in more than $k$ crossings), has been the topic of important recent work on graph drawing.
+Though planar graphs are, by now, quite well-understood, the situation is much less clear for $k$-planar graphs.
+Even the maximum edge-density of $k$-planar graphs is not completely settled.
+For $k\le 4$, the maximum number of edges in an $n$-vertex $k$-planar graph is $(k+3)(n-2)$, which is tight for $k=1,2$ \cite{pach.toth:graphs}.
+A near-tight bound of $11(n-2)/2$ is known for $k=3$ \cite{pach.radoici.ea:improving} but for general $k$, the current best upper bound is $4.108\sqrt{k}n$ \cite{pach.toth:graphs}.
+
+In the current paper we investigate the structure of $k$-planar graphs.
+Since planarity is closed under taking minors, planar graphs can be characterized by a finite set of forbidden minors.
+Indeed, Wagner's Theorem states that this set consists of the complete 5-vertex graph $K_5$ and the complete bipartite graph $K_{3,3}$.
+Unfortunately, $k$-planarity is not closed under taking minors, ruling out this kind of clean structural result.
+Rather, we present a rough structural characterization of $k$-planar graphs that, despite its roughness, provides several new results for $k$-planar and related graph families.
+At the heart of this characterization is layered $H$-partitions, which we now define.
+
 A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle V_0,V_1,\ldots\rangle$ such that $\{V_0,V_1,\ldots\}$ is a partition of $V(G)$ and there is no edge $vw\in E(G)$ such that $v\in V_i$, $w\in V_j$ and $|j-i|>1$.
@@
 
-In this paper we show that another family of graphs admits layered $H$-partitions of bounded layered width in which $H$ has bounded treewidth: $k$-planar graphs, which we now define.
+% In this paper we show that another family of graphs admits layered $H$-partitions of bounded layered width in which $H$ has bounded treewidth:
+In the current paper we show that $k$-planar graphs also admit layered $H$-partitions in which $H$ has bounded treewidth.
+First, we formally define $k$-planar graphs.
 An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve in a surface $S$ is a continuous function $f:[0,1]\to S$.
@@
 \end{thm}
+
+TODO: Write here about applications. $(g,k)$-planar, map graphs, $k$-NN, David's rough characterization.
 

### 2019-06-17 k-planar.tex
Added references for crossing lemma

--- 2019-06-17 k-planar.tex
@@
 
-\begin{lem}
+\begin{lem}\lemlabel{nearest-neighbour}
   Every $k$-nearest-neighbour graph is $O(k^2)$ planar.
@@
 \end{proof}
+
+We note that \lemref{nearest-neighbour} is tight up to the leading constant:  Every $n$-vertex nearest neighbour graph has at least $kn/2$ at at most $kn$ edges.
+For $k\ge 8$, the Crossing Lemma \cite{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is at $\Omega(k^2)$.
 

### 2019-06-18 k-planar.tex
Sending to coauthors

--- 2019-06-17 k-planar.tex
@@
 \usepackage{paralist}
+\usepackage{stmaryrd}
 
 \usepackage[longnamesfirst,numbers,sort&compress]{natbib}
-
-% \usepackage{tipa}
-% \usepackage{upgreek}
-% \usepackage{rotating}
-% \newcommand{\V}{\rotatebox[origin=c]{180}{$V$}}
-% \newcommand{\Y}{\rotatebox[origin=c]{180}{$Y$}}
-%\usepackage{hyperref}
 
@@
 
+In 1965, \citet{ringel:sechsfarbenproblem} introduced the notion of $1$-planar graphs, a generalization of planar graphs that allows edges to cross provided that no edge is involved in more than one crossing.
+This simple generalization has since become a rich source of problems and results.
+Indeed, the annotated bibliography on 1-planar graphs by \cite{kobourov.liotta.ea:annotated} contains more than 140 entries.
+
+The natural generalization of $1$-planar graphs, $k$-planar graphs (that allow edge to be involved in up to $k$ crossings), has been the topic of important recent work on graph drawing.
+Though planar graphs are, by now, quite well-understood, the situation is much less clear for $k$-planar graphs; even the maximum edge-density of $k$-planar graphs is not completely settled.
+For $k\le 4$, the maximum number of edges in an $n$-vertex $k$-planar graph is $(k+3)(n-2)$, which is tight for $k=1,2$ \cite{pach.toth:graphs}.
+A near-tight bound of $11(n-2)/2$ is known for $k=3$ \cite{pach.radoicic.ea:improving} but for $k>4$, the current best upper bound is $4.108\sqrt{k}n$ \cite{pach.toth:graphs}.
+
+% In the current paper we investigate the structure of $k$-planar graphs.
+Since planarity is closed under taking minors, planar graphs can be characterized by a finite set of forbidden minors.
+Indeed, Wagner's Theorem states that this set consists of the complete 5-vertex graph $K_5$ and the complete bipartite graph $K_{3,3}$.
+Unfortunately, $k$-planarity is not closed under taking minors, ruling out this kind of clean structural result.
+The purpose of the current paper is to extend a new structural result for planar graphs to $k$-planar graphs.
+
+The purpose of the current paper is to show that new structural results for planar graphs can be extended to $k$-planar graphs.
+These results can be used to give a rough characterization of $k$-planar graphs of powers of bounded-degree planar graphs
+% Rather, we present a rough structural characterization of $k$-planar graphs that, despite its roughness, provides several new results for $k$-planar and related graph families.  
+All of these results are based on layered $H$-partitions, which we now define.
+{\color{red}TODO: Help.}
+
 A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle V_0,V_1,\ldots\rangle$ such that $\{V_0,V_1,\ldots\}$ is a partition of $V(G)$ and there is no edge $vw\in E(G)$ such that $v\in V_i$, $w\in V_j$ and $|j-i|>1$.
@@
 One such useful property is small \emph{treewidth}, which we now define.
-A \emph{tree-decomposition} $\mathcal{K}$ of a graph $G$ consists of a tree $K$ and a collection $\mathcal{K}=\{B_x:x\in V(K)\}$ of subsets of $V(G)$ indexed by nodes of $K$ such that (i)~for every $vw\in E(G)$, there exists some $x\in V(K)$ with $v,w\in B_x$; and (ii)~for every $v\in V(G)$, $K[x] = K[\{y: x\in B_y\}]$ is connected.
+A \emph{tree-decomposition} $\mathcal{K}$ of a graph $G$ consists of a tree $K$ and a collection $\mathcal{K}=(B_x:x\in V(K))$ of subsets of $V(G)$ indexed by nodes of $K$ such that (i)~for every $vw\in E(G)$, there exists some $x\in V(K)$ with $v,w\in B_x$; and (ii)~for every $v\in V(G)$, the induced subgraph $K[x] := K[\{y: x\in B_y\}]$ is connected.
 The \emph{width} of the tree-decomposition $\mathcal{K}$ is $\max\{|B_x|:x\in V(K)\}-1$.
-The \emph{treewidth} $\tw(G)$ of a graph $G$ is the minimum integer $t$ such that $G$ has a tree-decomposition of width $t$.
+The \emph{treewidth} $\tw(G)$ of a graph $G$ is the minimum width of a tree-decomposition of $G$.
 
@@
 \citet{dujmovic.esperet.ea:planar} add non-repetitive chromatic number to this list.
-\citet{dujmovic.joret.ea:planar} show that two common families of graphs have such $H$-partitions: planar graphs and (more generally) apex-minor-free graphs.\footnote{A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar.
-A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
-\note{PM}{Help me finish this definition of apex-minor-free.}}
-
-In this paper we show that another family of graphs admits layered $H$-partitions of bounded layered width in which $H$ has bounded treewidth: $k$-planar graphs, which we now define.
-An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve in a surface $S$ is a continuous function $f:[0,1]\to S$.
+\citet{dujmovic.joret.ea:planar} show that two common families of graphs have such $H$-partitions: planar graphs, graphs of bounded genus, and (more generally) apex-minor-free graphs.\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
+A class $\mathcal{G}$ of graphs is \emph{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$.
+A minor-closed class is \emph{proper} if it is not the class of all graphs.
+For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class.
+
+ A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar.
+A 1-apex graph is \emph{apex}.
+A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.}
+
+% In this paper we show that another family of graphs admits layered $H$-partitions of bounded layered width in which $H$ has bounded treewidth:
+In the current paper we show that $k$-planar graphs also admit layered $H$-partitions in which $H$ has bounded treewidth.
+First, we formally define $k$-planar graphs.
+An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
 The points $f(0)$ and $f(1)$ are called the \emph{endpoints} of the curve.
@@
 Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.} $k$-planar graphs are a generalization of planar graphs (which are genus-$0$ graphs).
-Unlike genus-$g$ graphs, however, the family of $k$-planar graphs is not minor-closed.\footnote{A class $\mathcal{G}$ of graphs is \emph{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$.
-A minor-closed class is \emph{proper} if it is not the class of all graphs.
-For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class.}  A graph $G'$ obtained from a $k$-planar graph $G$ by edge deletions and edge contractions may or may not be $k$-planar.
+Unlike genus-$g$ graphs, however, the family of $k$-planar graphs is not minor-closed.
+A graph $G'$ obtained from a $k$-planar graph $G$ by edge deletions and edge contractions may or may not be $k$-planar.
 Indeed, \citet{dujmovic.eppstein.ea:structure} construct 1-planar graphs that contain arbitrarily large complete graph minors.
@@
 
+\note{PM}{TODO: Write here about applications. non-repetitive colouring, $p$-centered colouring, $(g,k)$-planar graph, $(g,d)$-map graphs, $k$-NN graphs, David's rough characterization.}
+
 The remainder of the paper is organized as follows: In \secref{k-planar} we prove \thmref{k-planar}.  In \secref{1-planar} we prove \thmref{1-planar}.  
@@
 The purpose of this section is to prove \thmref{k-planar}.
-In this section, all graphs are finite, simple, and undirected.
+In this paper, all graphs are finite and undirected.
+Unless specifically mentioned otherwise, all graphs are also simple.
 For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$) we use $G[S]$ to denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.
 We use $G-S$ as a shorthand for $G[V(G)\setminus S]$ and for a graph $G'$, we use $G-G'$ as a shorthand for $G-V(G')$.
-We use $G'\subseteq G$ to denote subgraph containment, i.e., $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.
-\note{PM}{TODO: Hunt down occurrences of $G-G'$ and replace with $G-V(G')$.}
-
-We begin by introducing a structural tool for planar graphs, called a PS-tree, that is implicit in previous work on $H$-partitions of planar graphs \cite{dujmovic.joret.ea:planar}.
-We then show how the PS-tree of a planarized version of a $k$-planar graph $G$ can be used to find a layered $H$-partition of $G$ of small layered width in which $H$ has small treewidth.
+We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.
+
+We begin by introducing a structural tool for planar graphs, called a tripod decomposition, that is implicit in previous work on $H$-partitions of planar graphs \cite{dujmovic.joret.ea:planar}.
+We then show how the tripod decomposition of a planarized version of a $k$-planar graph $G$ can be used to find a layered $H$-partition of $G$ of small layered width in which $H$ has small treewidth.
 
@@
 For each node $v\in V(T)$, we call each node on the path, in $T$, from $v$ to $r$ (including $v$ itself) a \emph{$T$-ancestor} of $v$.
-If $v'$ is a $K$-ancestor of $v$, then $v$ is a \emph{$K$-descendant} of $v'$.
+If $v'$ is a $T$-ancestor of $v$, then $v$ is a \emph{$T$-descendant} of $v'$.
 
@@
 A \emph{triangulation} is a near-triangulation whose outer face is bounded by a 3-cycle.
-For any cycle $F$ in a near-triangulation $N$ we use $N\circledast F$ to denote the near-triangulation consisting of all the edges and vertices of $N$ that are contained in $F$ or its interior.
+For any cycle $F$ in a near-triangulation $N$ we use $N\llparenthesis F\rrparenthesis$ to denote the near-triangulation consisting of all the edges and vertices of $N$ that are contained in $F$ or its interior.
 
@@
 For a vertex $r\in V(G)$, the \emph{BFS layering} of $G$ from $r$ is the layering $\langle V_0,V_1,\ldots\rangle$ of $G$ where $V_i=\{v\in V(G): \dist_{G}(r,v)=i\}$ for each integer $i\ge 0$.
-A \emph{BFS spanning tree} $T$ of $G$ rooted at $r$ is a spanning tree of $G$ rooted at $r$ where, for every $v\in V(G)\setminus\{r\}$, the parent $v'$ of $v$ in $T$ has $\dist_G(r,v')= \dist_G(r,v)-1$.
+A \emph{BFS spanning tree} $T$ of $G$ rooted at $r$ is a spanning tree of $G$ rooted at $r$ where, for every $v\in V(G)$, $\dist_T(v,r)=\dist_G(v,r)$.
 
@@
 
-Fix some cycle $F$ in $\Delta$ and let $N=\Delta\circledast F$.
+Fix some cycle $F$ in $\Delta$ and let $N=\Delta\llparenthesis F\rrparenthesis$.
 Let $\tau=v_1v_2v_3$ be an inner face of $N$ and consider the three minimal vertical paths $Q_1$, $Q_2$, and $Q_3$ in $T$ such that, for each $i\in\{1,2,3\}$ the lower endpoint of $Q_i$ is $v_i$ and the upper endpoint $v_i'$ of $Q_i$ is in $V(F)$.
-If $Q_1$, $Q_2$, and $Q_3$ are vertex disjoint, then the graph $\overline{Y}=\tau\cup Q_1\cup Q_2\cup Q_3$ is called the \emph{closed tripod} in $N$ determined by $\tau$.
+If $Q_1$, $Q_2$, and $Q_3$ are vertex-disjoint, then the graph $\overline{Y}:=\tau\cup Q_1\cup Q_2\cup Q_3$ is called the \emph{closed tripod} in $N$ determined by $\tau$.
 We call $Q_1$, $Q_2$, and $Q_3$ the \emph{legs}, $v_1'$, $v_2'$ and $v_3'$ the \emph{feet}, and $\tau$ the \emph{crotch} of $\overline{Y}$.
-We call $Y=\overline{Y}-F$ the \emph{(open) tripod} of $N$ determined by $\tau$.
-The graph $M=\overline{Y}\cup F$ is called a \emph{Mercedes graph} and the (at most three) inner faces of $M$ other than $\tau$ are called the \emph{wedges} of $M$.
+We call $Y:=\overline{Y}-F$ the \emph{(open) tripod} of $N$ determined by $\tau$.
+The graph $M:=\overline{Y}\cup F$ is called a \emph{Mercedes graph} and the (at most three) inner faces of $M$ other than $\tau$ are called the \emph{wedges} of $M$.
 
 Let $r$ be a degree-3 node on the outer face of a triangulation $\Delta$.
-A \emph{PS-tree}  $K=K(\Delta, T, r)$ is a rooted tree in which each non-root node $x\in V(K)$ is associated with several objects (refer to \figref{K-node}):
+A \emph{tripod decomposition}  $K=K(\Delta, T, r)$ is a rooted tree in which each non-root node $x\in V(K)$ is associated with several objects (refer to \figref{K-node}):
 
@@
 \begin{enumerate}
-  \item A cycle $F_x$ in $\Delta$ and the near-triangulation $N_x=\Delta\circledast F_x$.
-
-  \item An inner face $\tau_x$ of $N_x$ that determines a closed tripod $\overline{Y}_x$ and an open tripod $Y_x=\overline{Y}_x-F_x$ in $N_x$.
-
-  \item The \emph{Mercedes graph} $M_x=F_x\cup \overline{Y}_x$
+  \item A cycle $F_x$ in $\Delta$ and the near-triangulation $N_x:=\Delta\llparenthesis F_x\rrparenthesis$.
+
+  \item An inner face $\tau_x$ of $N_x$ that determines a closed tripod $\overline{Y}_x$ and an open tripod $Y_x:=\overline{Y}_x-F_x$ in $N_x$, both with crotch $\tau_x$.
+
+  \item The \emph{Mercedes graph} $M_x:=F_x\cup \overline{Y}_x$
 \end{enumerate}
@@
 The root node $r^0_K$ of $K$ is exceptional.
-We define $F_{r^0_K}$ to be the single-vertex graph containing $r$, we define $\overline{Y}_{r^0_K}$ to be the star formed by $r$ and its three neighbours in $\Delta$, and (as usual) $Y_{r^0_K}=\overline{Y}_{r^0_K}-F_{r^0_K}$.
-The root node $r^0_K$ has exactly one child $r_K$, that we call the \emph{sub-root} of $K$, for which $F_{r_K}$ is the 3-cycle formed by the neighbours of $r$ in $\Delta$.
-
-The non-root nodes of the PS-tree $K$ (including the sub-root $r_K$) satisfy the following conditions:
+We define $F_{r^0_K}$ to be the single-vertex graph containing $r$, we define $\overline{Y}_{r^0_K}$ to be the star formed by $r$ and its three neighbours in $\Delta$, and (as usual) $Y_{r^0_K}:=\overline{Y}_{r^0_K}-F_{r^0_K}$.
+The root node $r^0_K$ has exactly one child $r_K$, called the \emph{sub-root} of $K$, for which $F_{r_K}$ is the 3-cycle formed by the neighbours of $r$ in $\Delta$.
+
+The non-root nodes of the tripod decomposition $K$ (including the sub-root $r_K$) satisfy the following conditions:
 \begin{enumerate}[(PR1)]
@@
 
-For any triangulation $\Delta$ in which the outer face has a vertex $r$ of degree 3 and for any BFS spanning tree of $\Delta$ rooted at $r$, there exists a PS-tree $K=K(\Delta,T,r)$.
-Indeed, this tree is constructed (albeit implicitly) by \citet[Proof of Lemma~14]{dujmovic.joret.ea:planar}.\footnote{\citet{dujmovic.joret.ea:planar} do not require that $T$ be a BFS spanning tree, only that its root, $r$ have degree 3 in $\Delta$ and in $T$.}
+For any triangulation $\Delta$ in which the outer face has a vertex $r$ of degree 3 and for any BFS spanning tree of $\Delta$ rooted at $r$, there exists a tripod decomposition $K=K(\Delta,T,r)$.
+Indeed, this tree is constructed (albeit implicitly) by \citet[Proof of Lemma~14]{dujmovic.joret.ea:planar}.\footnote{\citet{dujmovic.joret.ea:planar} do not require that $T$ be a BFS spanning tree, only that its root, $r$ has degree 3 in $\Delta$ and in $T$.}
 
@@
 Let $r$ be the vertex of the $K_4$ not incident on $F_0$ and let $T$ be a BFS spanning tree of $G^{++}$ rooted at $r$. 
-For each integer $i\ge 0$, let $V_i=\{v\in V(G^+): \dist_{T}(r,v)=i\}$ and observe that $\{V_2,V_3,\ldots,V_{h}\}$ is a layering of $G^+$ (but not of $G$).
-
-Let $K=K(G^{++},T,r)$ be a PS-tree and observe that the sub-root $r_K$ of $K$ has $F_{r_K}=F_0$.  This implies:
+For each integer $i\ge 0$, let $V_i:=\{v\in V(G^+): \dist_{T}(r,v)=i\}$ and observe that $\{V_2,V_3,\ldots,V_{h}\}$ is a layering of $G^+$ (but not of $G$).
+
+Let $K=K(G^{++},T,r)$ be a tripod decomposition and observe that the sub-root $r_K$ of $K$ has $F_{r_K}=F_0$.  This implies:
 \begin{compactenum}[(PR1)]\setcounter{enumi}{4}
@@
 
+For each vertex $z\in V(G^+)$, let $P_z=\{a\in V(K): z\in V(N_a-F_a)\}$.
 By \propref{all-in}, $P_{z}$ contains the sub-root $r_K$ of $K$.
 By \propref{children}, any node $x\in P_z$ has at most one of its children in $P_z$ and any node $x'\not\in P_z$ has none of its children in $P_z$.
-Therefore $K[P_z]$ is a vertical path that contains $r_K$ (and whose lower endpoint is the unique node $x\in V(K)$ such that $z\in Y_x$).
+Therefore $K[P_z]$ is a vertical path in $K$ that contains $r_K$ (and whose lower endpoint is the unique node $x\in V(K)$ such that $z\in Y_x$).
 We are about to repeatedly use the fact that the common intersection of two or more vertical paths is a vertical path.
@@
 For each edge $vw\in E(G)$, consider the walk $W_{vw}$ in $G^+$ that corresponds to the sequence of vertices in $G^+$ encountered while walking along the curve $vw$ from $v$ to $w$.
-We define $A_{vw}=\bigcap_{z\in W_{vw}} P_z$.
+We define $A_{vw}:=\bigcap_{z\in W_{vw}} P_z$.
 Since $V(W_{vw})\subseteq V(G^+)$, $A_{vw}$ is well-defined.
@@
 Therefore $K[A_{vw}]$ is a vertical path in $K$ that contains $r_K$.
-For each vertex $v\in V(G)$ define $A_v=\bigcap_{vw\in E(G)} A_{vw}$.
+For each vertex $v\in V(G)$ define $A_v:=\bigcap_{vw\in E(G)} A_{vw}$.
 Again, $K[A_v]$ is a vertical path in $K$ that contains $r_K$.
@@
 \[
-   S_x = \{v\in V(G): a(v)=x \} \enspace .
+   S_x := \{v\in V(G): a(v)=x \} \enspace .
 \]
-Clearly $\mathcal{S}=\{S_x:x\in V(K)\}$ is a partition of $V(G)$.
-We will use $\mathcal{S}$ as the partition that defines our quotient graph $H=G/\mathcal{S}$ with vertex set $V(H)=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
-First we argue that this partition has small layered width with respect to the layering $\{V_2,V_3,\ldots\}$ of $G^+$ defined by $T$ (recall that this is a layering of $G^+$, not a layering of $G$).
+Clearly $\mathcal{S}:=\{S_x:x\in V(K)\}$ is a partition of $V(G)$.
+We will use $\mathcal{S}$ as the partition that defines our quotient graph $H:=G/\mathcal{S}$ with vertex set $V(H):=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
+First we argue that this partition has small layered width with respect to the layering $\{V_2,V_3,\ldots\}$ of $G^+$ defined by $T$.
+ % (recall that this is a layering of $G^+$, not a layering of $G$).
 
@@
   Similarly, $\dist_{G^{++}}(r,v) \le \dist_{G^{++}}(r, v') + k+1$.  Therefore
-  $|\dist_{G^+}(r,v)-\dist_{G^+}(r,v')| \le k+1$ which establishes the claim, since $T$ is BFS spanning tree of $G^{++}$ rooted at $r$, so $\dist_{T}(r,z)=\dist_{G^{++}}(r,z)$ for all $z\in V(G^+)$.
+  $|\dist_{G^{++}}(r,v)-\dist_{G^{++}}(r,v')| \le k+1$ which establishes the claim, since $T$ is BFS spanning tree of $G^{++}$ rooted at $r$, so $\dist_{T}(r,z)=\dist_{G^{++}}(r,z)$ for all $z\in V(G^+)$.
 \end{proof}
@@
 \begin{clm}\clmlabel{width-of-g-plus}
-  For any $x\in V(H)$, and any integer $i\ge 0$, $|S_x\cap V_i|\le 24k+12$.
+  For any $x\in V(H)$, and any integer $i\ge 0$, $|S_x\cap V_i|\le 24k+36$.
 \end{clm}
@@
 If $v'\neq v$ then $v'\in V(G^+-G)$ is a point where two edges of $G$ cross and $v'$ contributes at most the four endpoints of these edges to $S_x$.
-In either case each $v'\in V(Y_x)\cap \bigcup_{j=i-k-1}^{i+k+1} V_j$ contributes at most four vertices to $S_x\cap V_i$, so $|S_x\cap V_i|\le 3\times 4\times (2k+3) = 24k+36$.
+In either case, each $v'\in V(Y_x)\cap \bigcup_{j=i-k-1}^{i+k+1} V_j$ contributes at most four vertices to $S_x\cap V_i$, so $|S_x\cap V_i|\le 3\times 4\times (2k+3) = 24k+36$.
 \end{proof}
 
-Next we describe a tree decomposition $\{B_x:x\in V(K)\}$ of $H$.  We deliberately use the same node set $V(H)=V(K)$ here, because the subtree $K[x]:=K[\{z:x\in B_z\}]$ will be rooted at $x$.
+Next we describe a tree decomposition $(B_x:x\in V(K))$ of $H$.  We deliberately use the same node set $V(H)=V(K)$ here, because the subtree $K[x]:=K[\{z:x\in B_z\}]$ will be rooted at $x$.
 
@@
 
-The tree decomposition $\{B_x:x\in V(K)\}$ of $H$ is defined as follows: For each edge $xy\in E(H)$ where $y$ is a $K$-ancestor of $x$, we add $y$ to $B_{x'}$ for every node $x'$ on the path from $x$ to $y$ in $K$.
+The tree decomposition $(B_x:x\in V(K))$ of $H$ is defined as follows: For each edge $xy\in E(H)$ where $y$ is a $K$-ancestor of $x$, we add $y$ to $B_{x'}$ for every node $x'$ on the path from $x$ to $y$ in $K$.
 That this produces a tree-decomposition is due to \clmref{xy-ancestor}.
@@
 \begin{clm}\clmlabel{bag-size}
-  The tree decomposition $\{B_x: x\in V(K)\}$ of $H$ contains no bag larger than $\bagsize$.
+  The tree decomposition $(B_x: x\in V(K))$ of $H$ contains no bag larger than $\bagsize$.
 \end{clm}
@@
 See \figref{A}(a).
-It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0=r_K$ is the root of $K$.
+It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0:=r_K$ is the root of $K$.
 We then cut $F_{a_0}$ into two pieces $F^{\bar{x}}_{a_0}$ and $F^x_{a_0}$ where $F^x_{a_0}$ is the piece that contains $F_x$.
@@
 The edge $vw$ corresponds to a walk $W_{vw}$ in $G^+$.
-The walk $W_{vw}$ is contained in $N_x-F_x$, otherwise $x\not\in A_{vw}$ so $x'\not\in A_{vw}$, so $x'\not\in A_v$, and $a(v)\neq x'$, contradicting the fact that $v\in S_{x'}$.
+The walk $W_{vw}$ is contained in $N_x-F_x$, otherwise $x\not\in A_{vw}$ which would imply that $x'\not\in A_{vw}$, $x'\not\in A_v$, and $a(v)\neq x'$, contradicting the fact that $v\in S_{x'}$.
 
@@
 Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $\delta\in\{0,\ldots,d\}$ such that there exists a $\delta$-walk.\footnote{To account for the fact that $B_x$ contains $x$ we count $\delta=d$.
-A $d$-walk corresponds, for example, to a 0-length walk that begins and ends at any vertex $v_0\in V(N_x-F_x)$.}  This is our goal for the remainder of this proof.
+A $d$-walk corresponds, for example, to a length-0 walk that begins and ends at any vertex $v_0\in V(N_x-F_x)$.}  This is our goal for the remainder of this proof.
 
@@
 
-  Let $d_0=d$, and $d_1=\min\{i: V(F_d\cap Y_i)\neq\emptyset \}$.  By \propref{ancestor-boundary}, any vertex $z\in V(F_d)$ is contained in 
-  $Y_j$ for some $d_1\le j < d$.  Therefore, $F_d$ is contained in the interior of  $F_{d_1}$, i.e., $N_{d_0}\subseteq N_{d_1}-F_{d_1}$.
+  Let $d_0:=d$, and $d_1:=\min\{i: V(F_d\cap Y_i)\neq\emptyset \}$.  By \propref{ancestor-boundary}, any vertex $z\in V(F_d)$ is contained in 
+  $Y_j$ for some $j\in \{d_1,d_1+1,\ldots,d-1\}$.  Therefore, $F_d$ is contained in the interior of  $F_{d_1}$, that is, $N_{d_0}\subseteq N_{d_1}-F_{d_1}$.
 
@@
 
-  To prove this claim, consider the graph $I$, illustrated in \figref{d0d1}, with vertex set $V(I)=\{d_1+1,\ldots,d_0\}$ and $ij\in E(I)$ if and only $i=j$ or $|F_i\cap Y_j|\neq\emptyset$.
-Any $\delta$-walk $v_0,\ldots,v_\ell$ with $\delta\in V(I)$ corresponds to a walk $\rho_0,\ldots,\rho_{\ell}$ in $I$, where $\rho_i=\max\{j\in\{0,\ldots,d\} : v_i\in N_{j}-F_j\}$.
+  To prove \eqref{one-round}, consider the graph $I$, illustrated in \figref{d0d1}, with vertex set $V(I)=\{d_1+1,\ldots,d_0\}$ and $ij\in E(I)$ if and only $i=j$ or $|F_i\cap Y_j|\neq\emptyset$.
+Any $\delta$-walk $v_0,\ldots,v_\ell$ with $\delta\in V(I)$ corresponds to a walk $\rho_0,\ldots,\rho_{\ell}$ in $I$, where $\rho_i:=\max\{j\in\{0,\ldots,d\} : v_i\in N_{j}-F_j\}$.
 To see why this is true, consider $\rho_{t}= i$ and $\rho_{t+1}=j$ for some $t\in\{0,\ldots,\ell-1\}$:
@@
 
-  The edges in $E(I)$ can be partitioned into two classes:  A set $B(I)=\{ij\in E(I): |i-j|\le 1\}$ of \emph{boring edges} and a set $J(I)=\{ij\in E(I): |i-j|>1\}$ of \emph{jump edges}.
+  The edges in $E(I)$ can be partitioned into two classes:  A set $B(I):=\{ij\in E(I): |i-j|\le 1\}$ of \emph{boring edges} and a set $J(I):=\{ij\in E(I): |i-j|>1\}$ of \emph{jump edges}.
 For a jump edge $ij$ with $i>j$, $j$ is called a \emph{landing} and $i$ is called a \emph{takeoff}.
@@
   We now prove that jump edges satisfy \jref{single}.
-  First We claim that, for each $i\in\{d_1+1,\ldots,d_0\}$, $F_{i}$ contains a vertex of $Y_{d_1}$.
+  First we claim that, for each $i\in\{d_1+1,\ldots,d_0\}$, $F_{i}$ contains a vertex of $Y_{d_1}$.
 For $i=d_0$, this is true by definition so $F_{d_0}$ contains some vertex $v\in V(Y_{d_1})$.
@@
   Now, consider the parent $a_{i-1}$ of $a_i$ in $K$.
-By \propref{children} $F_i$ is a wedge of the Mercedes graph $M_{i-1}$.
+By \propref{children}, $F_i$ is a wedge of the Mercedes graph $M_{i-1}$.
 Let $P_1,\ldots,P_g$ be the partition of $V(F_{i-1})$ guaranteed by \propref{ancestor-boundary}.
@@
 Since $F_i$ contains $v\in V(Y_{d_1})$, at least one of $\alpha$ or $j$, say $\alpha$, is equal $d_1$.
-Since $d_1\not\in V(I)$ and the (possible) edge $(i,i+1)$ is a boring edge, the only (possible) jump edge with $i$ as takeoff is $ij$.
+Since $d_1\not\in V(I)$ and the (possible) edge $(i,i+1)$ is boring, the only (possible) jump edge with $i$ as takeoff is $ij$.
 This establishes \jref{single}.
@@
   This establishes \eqref{one-round}.  An illustration of this analysis, showing the case $k=1$ is shown in \figref{one-round-example}.
-  \begin{figure}
+  \begin{figure}[htbp]
     \begin{center}
@@
   Refer to \figref{nested}.
-For $s>1$, let $d_s=\min\{i: V(F_{d_s}\cap Y_i)\neq\emptyset \}$.
+For $s>1$, let $d_s:=\min\{i: V(F_{d_s}\cap Y_i)\neq\emptyset \}$.
 Consider some $\delta$-walk $v_0,\ldots,v_\ell$ for $\delta\in \{ d_{s+1},\ldots,d_{s}-1\}$.
@@
 
-  By \clmref{bag-size}, $\mathcal{S}$ produces a graph $H=G/\mathcal{S}$ of treewidth at most $\treewidth$.
+  By \clmref{bag-size}, $\mathcal{S}$ produces a quotient graph $H=G/\mathcal{S}$ of treewidth at most $\treewidth$.
 By \clmref{width-of-g-plus}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+36$, so $|B_x\cap V_i'|\le (k+1)(24k + 36) = 24k^2+60k+36$.
@@
   Let $G$ be a $k$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
-Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \thmref{k-planar} and such that $V_i' = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\}$ for each integer $i\ge 0$.
+Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \thmref{k-planar} and such that $V_i' := \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\}$ for each integer $i\ge 0$.
 \end{cor}
@@
   \[  
-    V_i' = \{v\in V(G): \floor{\dist_{T}(r,v)-2}=i\} 
+    V_i' := \{v\in V(G): \floor{\dist_{T}(r,v)-2}=i\} 
          = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\} \enspace ,
@@
 
-Let $G$ be an edge-maximal 1-plane multigraph.
+Let $G$ be an edge-maximal 1-plane multigraph with no two parallel edges on the boundary of a single face.
 Here, edge-maximal should be taken to mean that, if any two vertices $v$ and $w$ appear on a common face\footnote{The \emph{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
 We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.
+We assume that no two edges incident to a common vertex cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossing on a particular edge from $k$ to $2(k-1)$.}
 
@@
 The 1-plane graph $G$ has a plane triangulation $G'$ as a subgraph that can be obtained by removing one spar from each kite in $G$.
-Observe that, for any spar $xy\in E(G-G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
+Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
 It follows that
@@
 
-Our proof of \thmref{1-planar} follows quickly from the following technical lemma, which is an extension of the equivalent result for plane graphs \cite{dujmovic.joret.ea:planar}.
+Our proof of \thmref{1-planar} follows quickly from the following technical lemma, which is an extension of the analagous result for plane graphs \cite{dujmovic.joret.ea:planar}.
 \begin{lem}\lemlabel{induction} The setup:
@@
     \item Let $G$ and $G'$ be defined as above.
-    \item Let $T$ be a BFS tree of $G'$ rooted at some vertex $r$.
+    \item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
     \item For each integer $j\ge 0$, let $V_j=\{v\in V(G):\dist_T(r,v)=j\}$. 
@@
       \item No edge of $F$ is crossed by any edge of $G$; and
-      \item $V(F)$ can be partitioned into $P_1,\ldots,P_k$, $k\le 3$ such that for each $i\in\{1,\ldots,k\}$,
+      \item $V(F)$ can be partitioned into $P_1,\ldots,P_k$, for some $k\in\{1,2,3\}$ such that for each $i\in\{1,\ldots,k\}$,
       \begin{compactenum}
@@
     \end{compactenum}
-    \item Let $N^+$ and $N^-$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
+    \item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
   \end{compactenum}
@@
 \begin{proof}
-  This proof very similar to the proof of Lemma~14 in \citet{dujmovic.joret.ea:planar}.
+  This proof is very similar to the proof of Lemma~14 in \citet{dujmovic.joret.ea:planar}.
 Rather than duplicate every detail of that proof here, we focus on the differences and refer the reader to the original proof for the remaining details.
@@
 If $k=3$, set $R_i=P_i$ for each $i\in\{1,2,3\}$.
-Otherwise, as in \citet{dujmovic.joret.ea:planar}, split $P_1,\ldots,P_k$ to partition $V(C)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a path and each $R_i$ contains vertices from at most one of $P_1,\ldots,P_k$.
-
-  Next, as in \citet{dujmovic.joret.ea:planar}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $V(N'-F)$.
+Otherwise, as in \citet{dujmovic.joret.ea:planar}, split $P_1,\ldots,P_k$ to partition $V(F)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a non-empty path and each $R_i$ contains vertices from exactly one of $P_1,\ldots,P_k$.
+
+  Next, as in \citet{dujmovic.joret.ea:planar}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
 
@@
 
-  We claim that, for each integer $i\ge 0$, $|V(Y^+)\cap V_i|\le 15$.
-First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most 3 vertices of $V_i$, each incident on at most two edges of $\overline{Y}$.
-By \eqref{spar-distance}, any vertex $x\in V(\overline{Y}^+-\overline{Y})\cap V_i$, is incident on an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $V_i$.
+  We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap V_i|\le 15$.
+First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $V_i$, each incident on at most two edges of $\overline{Y}$.
+By \eqref{spar-distance}, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap V_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $V_i$.
 These at most six edges have at most 12 endpoints.
-Therefore $|V(\overline{Y}^+-\overline{Y})\cap V_i|\le 6\times 2=12$, so $|\overline{Y}^+\cap V_i|\le 12+3=15$.
+Therefore $|V(\overline{Y}^+)\setminus V(\overline{Y})\cap V_i|\le 6\times 2=12$, so $|V(\overline{Y}^+)\cap V_i|\le 12+3=15$.
 
@@
 But then the only faces of $M^+$ incident to $vw$ are kite faces.
-In particular $vw$ can not be incident the non-kite face $F_i$.
+In particular $vw$ cannot be incident to the non-kite face $F_i$.
 
   Note that $M$ is a Mercedes graph and that each of the faces $F_1,\ldots,F_m$ is contained in a single wedge of $M$.
-Let $Y^+ = \overline{Y}^+-F$.
+Let $Y^+ := \overline{Y}^+-F$.
 Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.
@@
   The remainder of the proof finishes in the same way as the proof of Lemma~14 in \cite{dujmovic.joret.ea:planar}.
-$P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
+The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
 Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
@@
   \end{enumerate}
-  Since $G^+$ is planar, the second point implies that $H=G^+/\mathcal{P}$ is planar.\footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
+  Since $G^+$ is planar, the second point implies that $H=G^+/\mathcal{P}$ is planar.
+  % \footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
 Repeatedly applying this fact implies that $\Delta/\mathcal{P}$ is planar provided that $\Delta[P]$ is connected for every $P\in\mathcal{P}$.} The first point then ensures that $G/\mathcal{P}$ is also planar.
@@
 
-Exactly the same argument used to establish \corref{k-planar} establishes the analagous result for \thmref{1-planar}:
+Exactly the same argument used to establish \corref{k-planar} as a consequence of \thmref{k-planar} establishes the following result as a consequence of \thmref{1-planar}:
 
@@
 
-Next we prove \thmref{g-k-planar}, which generalizes of our results to $(g,k)$-plane graphs.  We make use of the following lemma due to \citet[Lemma~21]{dujmovic.joret.ea:planar}:
+Next we prove \thmref{g-k-planar}, which generalizes our results to $(g,k)$-plane graphs.  We make use of the following lemma due to \citet[Lemma~21]{dujmovic.joret.ea:planar}:
 
 \begin{lem} \lemlabel{make-planar}
-Let $G^+$ be a $(S,0)$-plane graph where $S$ is a surface with Euler genus $g$. 
+Let $G^+$ be a $(\Sigma,0)$-plane graph where $\Sigma$ is a surface with Euler genus $g$. 
 For each $r\in V(G^+)$ there is a vertex set $Z\subseteq V(G^+)$ such that $G^+-V(Z)$ is planar and there is a connected planar graph $G^*$ containing  $G^+-Z$ as a subgraph with a vertex $r^*\in V(G^*)$ such that, for every integer $i\ge 0$,
@@
 and
-\[ |\{v\in V(G^+): \dist_{G^+}(r,v) = i\}| \le 2g \enspace . \]
+\[ |\{v\in Z: \dist_{G^+}(r,v) = i\}| \le 2g \enspace . \]
 
@@
 \begin{proof}[Proof of \thmref{g-k-planar}]
-  Let $G$ be an $(S,k)$-plane graph where $S$ is a surface of Euler genus $g$.
-Let $G^+$ be the $(S,0)$-plane graph obtained by adding a dummy vertex at each crossing in $G$.
+  Let $G$ be an $(\Sigma,k)$-plane graph where $\Sigma$ is a surface of Euler genus $g$.
+Let $G^+$ be the $(\Sigma,0)$-plane graph obtained by adding a dummy vertex at each crossing in $G$.
 Let $r_0$ be any vertex in $V(G)$, let $T$ be BFS spanning tree of $G^+$ rooted at $r_0$ and let $\langle V_0,V_1,\ldots\rangle$ be the resulting layering of $G^+$.
@@
 For each edge $vw\in E(G)$, let $W_{vw}$ be the corresponding walk in $G^+$.
-The set $Z$ contains vertices in $V(G)$ as well as dummy vertices in $V(G^+-G)$.
-Create a new set $Z'\subseteq V(G)$ by taking, for each $vw\in E(G)$ such that $V(W_{vw})\cap Z\neq\emptyset$, one of the vertices $v$ or $w$.
-In this way, each vertex $z\in Z$ can be charged for at most two vertices in $v,x\in Z'$ and this happens only if $(z,vw,xy)$ is a crossing in $G$.
+The set $Z$ contains vertices in $V(G)$ as well as dummy vertices in $V(G^+)-V(G)$.
+Create a new set $Z'\subseteq V(G)$ from $Z$ by replacing each dummy vertex $z\in Z\cap (V(G^+)\setminus V(G))$ with one endpoint from each of the edges that cross $vw,xy\in E(G)$ that cross at $Z$.
+In this way, each vertex $z\in Z\cap V(G^+)\setminus V(G)$ can be charged for at most two vertices in $v,x\in Z'$ and this happens only if $(z,vw,xy)$ is a crossing in $G$.
   Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \clmref{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 2\times (2k+1) = 4g(2k+1)$ for each integer $i\ge 0$.
@@
   Let $G^*_1$ be the supergraph of $G^*_0$ obtained by adding every vertex and edge of $G^*$ that is not contained in some edge of $G^*_0$.
-Finally, let $G^{*+}_1$ be the triangulation obtained by adding edges to $G^*$ in such a way that $\dist_{G^*}(r^*,v)=\dist_{G^{*+}}(r^*,v)$ for every $v\in V(G^*)$.\footnote{The fact that $G^*$ can be completed to a triangulation $G^{*+}$ while preserving distances to $r^*$ follows from the fact that any face $F$ of $G^*$ with 4 or more vertices contains two non-adjacent vertices $x$ and $y$ such that $|\dist_{G^*}(r^*,x)-\dist_{G^*}(r^*,y)|\le 1$.
-Indeed, if $z$ is a vertex of $F$ that minimizes $\dist_{G^*}(r^*,z)$, then two neighbours $x$ and $y$ of $z$ have this property.} Now, observe that $G^*_1$ is a $k$-plane graph and $G^{*+}_1$ is obtained by adding dummy vertices at crossings in $G^*_1$ and triangulating the resulting graph.
+Finally, let $G^{*+}_1$ be the triangulation obtained by adding edges to $G^*$ so that $\dist_{G^*}(r^*,v)=\dist_{G^{*+}}(r^*,v)$ for every $v\in V(G^*)$.
+The fact that $G^*$ can be completed to a triangulation $G^{*+}$ while preserving distances to $r^*$ follows from the fact that any face $F$ of $G^*$ with four or more vertices contains two non-adjacent vertices $x$ and $y$ such that $|\dist_{G^*}(r^*,x)-\dist_{G^*}(r^*,y)|\le 1$.
+Indeed, if $z$ is a vertex of $F$ that minimizes $\dist_{G^*}(r^*,z)$, then two neighbours $x$ and $y$ of $z$ have this property.
+Now, observe that $G^*_1$ is a $k$-plane graph and $G^{*+}_1$ is obtained by adding dummy vertices at crossings in $G^*_1$ and triangulating the resulting graph.
 
@@
 
-  We turn $\mathcal{P}$, the $H$-partition of $G-Z'$ into a partition $\mathcal{P}'=\mathcal{P}\cup \{Z'\}$ of $G$.
-That is, we place all vertices of $Z'$ in a common part.
-In the worst-case, the vertex of $H'=G/ \mathcal{P}'$ corresponding to $V(G)\cap Z'$ is adjacent to every vertex of $H=(G-Z')/\mathcal{P}$, so the treewidth of $H'$ is at most one more than the treewidth of $H$.
+  We now turn $\mathcal{P}$, the $H$-partition of $G-Z'$, into a partition $\mathcal{P}'=\mathcal{P}\cup \{Z'\}$ of $G$ by placing all vertices of $Z'$ into one new common part.
+Let $H'=G/ \mathcal{P}'$.
+Then the treewidth of $H'$ is at most one more than the treewidth of $H$.
+
+  % In the worst-case, the vertex of $H'=G/ \mathcal{P}'$ corresponding to $V(G)\cap Z'$ is adjacent to every vertex of $H=(G-Z')/\mathcal{P}$, so the treewidth of $H'$ is at most one more than the treewidth of $H$.
 
@@
               4 & \text{if $k=1$} \\
-              k^3/6 + 3k^2/2 + 13k/3+4 & \text{otherwise}
+              k^3/6 + 3k^2/2 + 13k/3+5 & \text{otherwise}
             \end{cases} 
@@
 \begin{cor}\corlabel{first}
-  For every $k$-planar graph $G$, $\pi(G)\le 12(2k+3)(k+1)\times 4^{(1/6)(k+2)(k+3)(k+4)}$.
+  For every $k$-planar graph $G$, $\pi(G)\le 12(2k+3)(k+1)\times 4^{(k+2)(k+3)(k+4)/6}$.
 \end{cor}
@@
 \begin{cor}
-  For every $(g,k)$-planar graph $G$, $\pi(G)\le \max\{4g(3k+1),12(2k+3)(k+1)\}\times 4^{(1/6)(k+2)(k+3)(k+4)}$.
+  For every $(g,k)$-planar graph $G$, 
+  \[ 
+    \pi(G)\le \max\{4g(3k+1),12(2k+3)(k+1)\}\times 4^{(k+2)(k+3)(k+4)/6} 
+    \enspace .
+  \]
 \end{cor}
@@
 \begin{cor}
-  For every $(g,1)$-planar graph $G$, $g\ge 1$, $\pi(G)\le 48g\times 256= 12288g$.
+  For every $g\ge 1$ and every $(g,1)$-planar graph $G$, $\pi(G)\le 48g\times 256= 12288g$.
 \end{cor}
@@
 A $c$-colouring $\phi$ of $G$ is \emph{$p$-centered} if, for every connected subgraph $X\subseteq G$, $|\{\phi(v):v\in V(X)\}| > p$ or there exists some $v\in V(X)$ such that $\phi(v)\neq \phi(w)$ for every $w\in V(X)\setminus\{v\}$.
-In words, either $X$ receives more than $p$ colours or some vertex of $X$ receives a unique colour.
+In words, either $X$ receives more than $p$ colours or some vertex in $X$ receives a unique colour.
 Let $\chi_p(G)$ denote the smallest integer $c$ such that $G$ has a $p$-centered $c$-colouring.
@@
 \begin{proof}
-  Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=\{B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$ where $H$ as treewidth at most $t$.
+  Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$ where $H$ as treewidth at most $t$.
   We use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,O(p^t)\}$.
@@
 
-Prior to the current work, the strongest upper bounds on the number of colours required for $p$-centered colourings of $k$-planar graphs used the fact that $k$-planar graphs have bounded expansion.
-\note{PM}{I didn't follow the discussion on the previous best-known bounds for $p$-centered colourings.
-Can someone fill me in?}
+Prior to the current work, the strongest upper bounds on the number of colours required for $p$-centered colourings of $(g,k)$-planar graphs used the fact that $(g,k)$-planar graphs are have layered treewidth $O(gk)$ \cite{dujmovic.eppstein.ea:structure} and therefore have bounded (indeed, linear) expansion \cite{dujmovic.morin.ea:layered}.
+Using a result of \citet{zhu:colouring} this leads to bounds on $\chi_p(G)$ that are doubly-exponential in $gk$.
+
+\note{PM}{Please check the preceding paragraph.}
+
+% \note{PM}{I didn't follow the discussion on the previous best-known bounds for $p$-centered colourings.  Can someone fill me in?}
 
@@
 
-\subsection{More Stuff}
-
-\begin{enumerate}
-  % \item clustered colouring;
-  \item connection with queue layouts;
-  % \item $d$-map graphs, which are $d^2$-planar;
-  \item connection with low-treewidth/treedepth colourings;
-  \item An important open question is whether $k$-planar graphs have layered $H$-partitions whose layered width is upper bounded by a function of $k$ but in which the treewidth of $H$ is bounded by an absolute constant, independent of $k$.
-  \item Are our current $H$-partitions $K_{3,s}$-free for some $s=s(k)$?  If so, this is good news for clustered colouring.
-\end{enumerate}
-
-\noindent
-\framebox{
-\begin{minipage}{\textwidth}
-  \note{DW}{%
-    % I think we need to define layered treewidth, and
-    % give the result of Dujmovic-Eppstein-Wood that (g,k)-planar graphs
-    % have O(gk)  layered treewidth. This is the best previous structural
-    % description of k-planar graphs. State that layered partitions are
-    % richer than  layered treewidth, and lead to O(1) bounds on non-rep
-    % chromatic number and queue-number, instead of O(log n) bounds via
-    % layered treewidth.
-    I also think we should give explicit upper bounds on non-rep chromatic
-    number for k-planar graphs, and also for queue-number (especially if
-    they improve the bound in the FOCS / J.ACM paper).
-
-    We should add corollaries for d-map graphs, since they are $d^2$-planar.
-    }
-
-    \note{PM}{
-    A couple of other things to look into: 
-    \begin{compactenum}
-      \item $p$-centered colourings
-      \item low treewidth colourings
-    \end{compactenum}
-    }
-\end{minipage}
-}
+% \subsection{More Stuff}
+% 
+% \begin{enumerate}
+%   % \item clustered colouring;
+%   \item connection with queue layouts;
+%   % \item $d$-map graphs, which are $d^2$-planar;
+%   \item connection with low-treewidth/treedepth colourings;
+%   \item An important open question is whether $k$-planar graphs have layered $H$-partitions whose layered width is upper bounded by a function of $k$ but in which the treewidth of $H$ is bounded by an absolute constant, independent of $k$.
+%   \item Are our current $H$-partitions $K_{3,s}$-free for some $s=s(k)$?  If so, this is good news for clustered colouring.
+% \end{enumerate}
+% 
+% \noindent
+% \framebox{
+% \begin{minipage}{\textwidth}
+%   \note{DW}{%
+%     % I think we need to define layered treewidth, and
+%     % give the result of Dujmovic-Eppstein-Wood that (g,k)-planar graphs
+%     % have O(gk)  layered treewidth. This is the best previous structural
+%     % description of k-planar graphs. State that layered partitions are
+%     % richer than  layered treewidth, and lead to O(1) bounds on non-rep
+%     % chromatic number and queue-number, instead of O(log n) bounds via
+%     % layered treewidth.
+%     I also think we should give explicit upper bounds on non-rep chromatic
+%     number for k-planar graphs, and also for queue-number (especially if
+%     they improve the bound in the FOCS / J.ACM paper).
+% 
+%     We should add corollaries for d-map graphs, since they are $d^2$-planar.
+%     }
+% 
+%     \note{PM}{
+%     A couple of other things to look into: 
+%     \begin{compactenum}
+%       \item $p$-centered colourings
+%       \item low treewidth colourings
+%     \end{compactenum}
+%     }
+% \end{minipage}
+% }
+
+\section{Conclusions}
+
+We have shown that $k$-planar graphs and several related graph families have layered $H$-partitions of layered width $O(k^2)$ in which $H$ has treewidth $O(k^3)$.
+We conclude with the following question:  Does there exist a function $f:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph has a layered $H$-partition of layered width at most $f(k)$ in which $H$ has treewidth at most $C$?
+Perhaps $C=3$.
 

### 2019-07-02 k-planar.tex
Working....

@@
   for all integers $i\ge 0$.  This layering satisfies the conditions of the corollary.
+\end{proof}
+
+\section{A Generalization}
+
+We now generalize \thmref{k-planar} to a more abstract setting.
+
+\begin{lem}\lemlabel{nice-decomposition}
+  Let $H$ be a graph of treewidth $t-1$.
+Then there exists a graph $H'$ isomorphic to $H$ that has a tree-decomposition $(B_x:x\in V(T))$ of width $t-1$ where $T$ is rooted a vertex $x_1\in V(T)$ and has following additional properties:
+  \begin{compactenum}
+    \item $V(H')\subseteq V(T)$;
+    \item For each $v'\in V(H')$, the subtree $T[v']:=T[\{x\in V(T):v'\in B_x\}]$ is rooted at $v'$.
+  \end{compactenum}
+\end{lem}
+
+\begin{proof}
+  Begin with a width-$t$ tree-decomposition $(C_x:x\in V(T_0))$ of $H$ using a tree $T_0$ with the property that, for any edge $xy\in E(T_0)$, $|C_x\ominus C_y|=2$, where $\ominus$ denotes the symmetric difference between two sets.
+That is, $C_x$ contains exactly one vertex not in $C_y$ and $C_y$ contains exactly one vertex not in $C_x$.
+It is well-known that such a tree-decomposition exists.
+(Note that this implies that $|C_x|=t$ for every $x\in V(T_0)$.)
+
+  Next, select an arbitrary node $x\in V(T_0)$ and let $C_x:=\{v_1,\ldots,v_t\}$ denote the contents of $C_x$.
+  Create a new tree $T$ from $T_0$ by adding a path of nodes $x_1,x_2,\ldots,x_{t-1},x_t=x$ and assign bags to these new nodes so that $C_{x_i}:=\{v_1,\ldots,v_i\}$.
+Note that $(C_x: x\in V(T))$ is a width-$t$ tree-decomposition of $H$.
+
+  Let $T$ be rooted at $x_1$.
+  For each $v\in V(H)$, let $v'\in V(T)$ be the unique node such that $v\in C_{v'}$ and $v'$ has minimum $T$-depth.
+Finally, for each $x\in V(T)$, let $B_x=\{v' : v\in C_x\}$.
+Observe that $(B_x: x\in V(T))$ is a tree-decomposition satisfying the conditions of the lemma.
+\end{proof}
+
+\begin{lem}
+  Let $H$ be a graph of treewidth $t$ and let $H'$ and $(B_x:x\in V(T))$ be a graph (isomorphic to $H$) and a tree-decomposition of $H'$ satisyfing the conditions of \lemref{nice-decomposition}.
+Then, each $w'\in V(H)$ has at most at most $t$ $T$-ancestors $w'$ such that there exists a $T$-descendant $v'_0$ of $w'$ with $v'v'_0\in E(H)$.
+\end{lem}
+
+\begin{proof}
+  Since $vv_0\in E(H)$, $v'$ is the root of $T[v']$, and $v'_0$ is the root of $T[v'_0]$, the node $v'$ is contained in $B_x$ for every node $x$ on the path, in $T$, from $v'$ to $v'_0$.
+In particular, $w'$ is a node on this path, so $v'\in B_{w'}$.
+The lemma then follows from the fact $|B_w'|\le t$.
+\end{proof}
+
+\begin{thm}
+  Let $t\ge 2$, $d\ge 1$, and $k\ge 1$ be integers, let $G$ be a graph having a layered $H$-partition of width 1 in which $H$ has treewidth at most $t-1$ and let $G'$ be the supergraph of $G$ where, for each distinct pair $v,w\in V(G)$, $vw\in E(G')$ if and only if $G$ contains a path from $v$ to $w$ of length at most $k$ whose internal vertices each have degree at most $d$.
+Then $G'$ has a layered $I$-partition of layered width $O(kd^k)$ where the treewidth of $I$ is at most $(td)^k$.
+\end{thm}
+
+\begin{proof}
+  Apply \lemref{nice-decomposition} to $H$ to obtain a tree-decomposition $(C_x: x\in V(T))$ of $H$.
+(We skip the introduction of the isomorphic graph $H'$ and assume from the outset that $V(H)\subseteq V(T))$.)
+
+  For $v,w\in V(G)$, a \emph{$vw$-shortcut} is any path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.
+That is, $vw\in E(G')$ if and only if $G$ contains a $vw$-shortcut.
+
+  Let $(Q_x: x\in V(H))$ denote the given $H$-partition of $G$.
+For a node $y\in V(T)$, we abuse notation slightly and let $Q_y=\bigcup_{x\in C_y} Q_{x}$.
+For a node $y\in V(T)$, we say that a $vw$-shortcut $P$ \emph{touches} $x$ if $V(P)\cap Q_y\neq\emptyset$.
+If $G$ contains a $vw$-shortcut that touches $x$, then we say that $v$ and $w$ \emph{participate} in $x$.
+
+  For any $v\in V(G)$, the union of all shortcuts in $G$ that have $v$ as an endpoint is a connected subgraph of $G$.
+It follows that the set of nodes $x\in V(T)$ in which $v$ participates is a (connected) subtree $T_v$ of $T$ we let $a(v)$ denote the root of $T_v$.
+For each $x\in V(T)$, we define the \emph{separator}
+  \[  
+       S_x := \{v\in V(G): a(v)= x\} \enspace .
+  \]
+  Observe that $(S_x : x\in V(T))$ is a partition of $V(G)$.  We let $I:=G'/(S_x : x\in V(T))$ denote the resulting quotient graph and we let $V(I)=V(T)$ in the obvious way.
+
+  We claim that, for any edge $xy\in E(I)$, $x$ is a $T$-ancestor of $y$ or $y$ is a $T$-ancestor of $x$.
+Indeed, if $xy\in E(I)$, then there exists an edge $vw\in E(G')$ with $v\in S_x$ and $w\in S_y$.
+Therefore $G$ contains a $vw$-shortcut $P$ where $v\in Q_{x'}$ for some $T$-descendant $x'$ of $x$ and $w\in Q_{y'}$ for some $T$-descendant $y'$ of $y$.
+If neither $x$ nor $y$ is were a $T$-ancestor of the other, then neither $x'$ nor $y'$ would be an ancestor of the the other, but both $v$ and $w$ would participate in a node $z$ that is a $T$-ancestor of both $x$ and $y$.
+This would contradict the fact that $v\in S_x$ and $w\in S_y$.
+
+  We construct a tree-decomposition $\mathcal{T}:=(B_x:x\in V(T))$ of $I$ as follows: For each edge $xy\in E(I)$ we include $x$ in $B_z$ for each node $z$ on the path, in $T$ from $x$ to $y$.
+This is clearly a tree-decomposition of $I$ and it has the property that, for any $x\in V(I)$, $T[x]$ is rooted at $x$.
+All that remains is to bound the width of $\mathcal{T}$.
+
+  We now focus on one particular bag $B_x$ for some $x\in V(T)$.  
+  Let $x_0,\ldots,x_d$ be the path from $x_0:=x$ to the root of $T$.  
+  Let $T_i$ denote the subtree of $T$ rooted at $x_i$ and let
+  \[ 
+      G_i := G[\bigcup_{x'\in V(T_i)}\bigcup Q_{x'}]-Q_{x_{i-1}} \enspace .
+  \]
+  In other words, $G_i$ is the part of $G$ that contains that is separated by the removal of $S$
+
+  Define $N_{x_0}:=G$ and, for each $i\in\{1,\ldots,d\}$, define $N_{x_i}$ to be the subgraph of $G$ induced by the union, over all $x'$ in the subtree of
+
+  containing exactly those components of $N_{x_{i-1}}-Q_{x_i}$ that contain vertices in $
+
+  As discussed above, for any node $x\in V(T)$, the bag $B_x$ contains only $T$-ancestors of $x$ (including $x$ itself).
+If $B_x$ contains some $T$-ancestor $y$ then this is because $xy\in E(I)$ which implies that there exists $vw\in E(G')$ with $v\in S_{x'}$ and $w\in S_y$ for some $T$-descendant $x'$ of $x$.
+This implies that $G$ contains a $vw$-shortcut $P=v_0,\ldots,v_p$ with $v_0=v$ and $v_p=w$.
+  Note that neither $v$ nor $w$ participates any $T$-ancestor of $y$ other than $y$.  Therefore $P$ does not contain any vertices of $Q_{y'}$ for any ancestor $y'$ of $y$.
 \end{proof}

### 2019-07-03 k-planar.tex
Moving....

--- 2019-06-18 k-planar.tex
@@
 \newcommand{\jref}[1]{(J\ref{j:#1})}
+\newcommand{\tref}[1]{(T\ref{#1})}
+\newcommand{\ylabel}[1]{\label{y:#1}}
+\newcommand{\yref}[1]{(Y\ref{y:#1})}
 
@@
   for all integers $i\ge 0$.  This layering satisfies the conditions of the corollary.
+\end{proof}
+
+\newpage
+\section{A Generalization}
+
+We now generalize \thmref{k-planar} to a more abstract setting.
+For two sets $A$ and $B$, we use $A\ominus B := (A\cup B)\setminus(A\cap B)$ to denote the symmetric difference between $A$ and $B$.
+We begin with a lemma about normalized tree decompositions:
+
+\begin{lem}\lemlabel{nice-decomposition}
+  Let $H$ be a graph of treewidth $t-1$.
+Then there exists a graph $H'$ isomorphic to $H$ that has a tree-decomposition $(B_x:x\in V(T))$ of width at most $t-1$ that has following additional properties:
+  \begin{compactenum}[(T1)]
+    \item\label{t:rooted}\label{t:first} $T$ is rooted at some node $x_0\in V(T)$ with $B_{x_0}=\emptyset$.
+    \item\label{t:node-set} $V(H')\subseteq V(T)$.
+    \item\label{t:diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y||\le 1$.
+    \item\label{t:subtree-root} For each $v'\in V(H')$, the subtree $T[v']:=T[\{x\in V(T):v'\in B_x\}]$ is rooted at $v'$.
+    \item\label{t:ancestor-edge}\label{t:last} For each edge $v'w'\in E(H)$, one of $v'$ or $w'$ is a $T$-ancestor of the other.
+  \end{compactenum}
+\end{lem}
+
+\begin{proof}
+  Begin with a width-$(t-1)$ tree-decomposition $(C_x:x\in V(T_0))$ of $H$ using a tree $T_0$ with the property that, for any edge $xy\in E(T_0)$, $|C_x\ominus C_y|\le 1$.
+That is, $C_x$ contains exactly one vertex not in $C_y$ or vice-versa.
+It is well-known that such a tree-decomposition exists.
+
+  Next, select an arbitrary node $x\in V(T_0)$ and let $C_x:=\{v_1,\ldots,v_{t'}\}$ denote the contents of $C_x$.
+  Create a new tree $T$ from $T_0$ by adding a path of nodes $x_0,x_1,x_2,\ldots,x_{t'-1},x_{t'}=x$.
+Let $C_{x_0}:=\emptyset$ and, for each $i\in\{1,\ldots,t'\}$, $C_{x_i}:=\{v_1,\ldots,v_i\}$.
+Note that $(C_x: x\in V(T))$ is a width-$(t-1)$ tree-decomposition of $H$.
+Let $T$ be rooted at $x_0$.
+Note that $(C_x:x\in V(T))$ satisfies Properties~\tref{t:rooted}, and \tref{t:diff}.
+For the remaining properties some renaming is required, which we now describe.
+
+  For each $v\in V(H)$, let $v'\in V(T)$ be the unique node such that $v\in C_{v'}$ and $v'$ has minimum $T$-depth.
+The uniqueness of $v'$ follows from the fact that $T[v]:=T[\{x\in V(T):v\in B_x\}]$ is connected, so $v'$ is the root of $T[v]$.
+Furthermore, for any pair of distinct vertices $v,w\in V(H)$, \tref{t:rooted} and \tref{t:diff} ensure that $v'\neq w'$ since, otherwise, $C_{v'}\ominus C_{x'}\supseteq \{v',w'\}$ where $x'$ is the parent of $v'=w'$.
+Property~\tref{t:rooted} would ensure that $x'$ exists and Property~\tref{t:diff} would ensure that $|C_{x'}\ominus C_{v'}|\le 1$.
+
+  For each $x\in V(T)$, let $B_x=\{v' : v\in C_x\}$.
+Now define the graph $H'$ to have vertex set $V(H')=\{v':v\in V(H)\}$ and edge set $E(H')=\{v'w': vw\in E(H)\}$.
+Obviously, $H'$ isomorphic to $H$ and $V(H')\subseteq V(T)$ and $(B_x: x\in V(T))$ is a width=$(t-1)$ tree-decomposition of $H'$.
+The decomposition $(B_x: x\in V(T))$ inherits Properties~\tref{t:rooted} and \tref{t:diff} from $(C_x:x\in V(T))$ and satisfies Property~\tref{t:node-set} and \tref{t:subtree-root} by definition.
+
+  To see that $(B_:x\in V(T))$ satisifies Property~\tref{t:ancestor-edge}, observe that, if $vw\in E(H)$, then at least one of $v'$ or $w'$ is contained in $B_x$ for every node $x$ on the path from $v'$ to $w'$ in $T$.
+If neither $v'$ nor $w'$ is an ancestor of the other, then some node $x$ on this path has $T$-depth less than that of $v'$ and $w'$.
+If $v'\in B_x$ this contradicts the fact that $v'$ is the root of $T[v']$.
+If $w'\in B_x$ this contradicts the fact that $w'$ is the root of $T[w']$.
+\end{proof}
+
+\newpage
+\subsection{Generalized Tripod Decompositions}
+
+Let $G$ be a graph and let $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{P}:=(Q_x: x\in V(H)))$ be a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{P}$ is at most $t$.
+In this section we define a hierarchical structure on $G$ that can be obtained from $(\mathcal{L},\mathcal{P})$.
+
+\begin{thm}
+  There exists a tree $T$ rooted at a node $x_0$ where each node $x\in V(T)$ is associated with a subset $Y_x\subseteq V(G)$ that defines several related sets $V_x = \bigcup_{y\in V(T_x) Y_y}$, $F_x=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x=V_x\cup F_x$ that satisfy the following properties:
+  \begin{compactenum}[(Y1)]
+    \item\ylabel{thickness} $\mathcal{Y}:=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width (with respect to $\mathcal{L}$) at most $t\ell$.
+    \item\ylabel{separator} There is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$. 
+    \item\ylabel{ancestor-edge} $x$ has $t'\le t$ $T$-ancestors $a_1,\ldots,a_{t'}$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
+  \end{compactenum}
+\end{thm}
+
+\begin{proof}
+  Apply \lemref{nice-decomposition} to $I$ to obtain a tree-decomposition $(B_x: x\in V(T))$ of $I$ where $T$ is rooted at some node $x_0$ and that satisifies Properties~\tref{t:first}--\tref{t:last}.
+(We skip the introduction of the isomorphic graph $H'$ and assume from the outset that $V(I)\subseteq V(T))$.)
+
+  For each $v\in V(G)$, let $q(v)$ denote the unique node $q\in V(I)$ such that $v\in Q_q$ and let $b(v)$ denote the root of the subtree $T[v]:=T[\{x\in V(T):q(v)\in B_x]$.
+For each $x\in V(T)$, define $Y_x=\{v\in V(G):c(v)=x\}$.
+  Clearly, $\mathcal{Y}:=(Y_x:x\in V(T))$ is a partition of $V(G)$.  Furthermore, for any $i\in\in\{1,\ldots,h\}$ 
+  \[  |Y_x| = \sum_{y\in C_x}|Q_y\cap V_i|\le |C_x|\ell \le t\ell ,
+  \] 
+  so $\mathcal{Y}$ has layered width at most $t\ell$ with respect to $\mathcal{L}$.  This establishes \yref{thickness}.
+
+  Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
+
+  To establish Property~\yref{ancestor-edge}, consider some edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.
+By the definition of $V_x$, $v\in Y_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
+  Since $\mathcal{Y}$ is a partition, $w\in Y_{a}$ for some $a\in V(T)$, $a\neq x$.
+Since $v\in Y_{x'}$ and $w\in Y_a$, $q(v)=x'$ and $q(w)=a$, so $x'\in Q_v$ and $w\in Q_a$.
+Therefore $H$ contains the edge $ax'$.
+
+  By \tref{ancestor-edge}, $a$ is an ancestor of $x'$  
+\end{proof}
+
+\newpage
+\subsection{Low-Degree Powers}
+
+For non-negative integers $k$ and $d$, a graph $G$ and distinct vertices $v,w\in V(G)$, a \emph{$(k,d)$ $vw$-shortcut} is a path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.
+The \emph{$(k,d)$th power of $G$}, denoted $G^{k,d}$, is the graph with vertex $V(G^{k,d})=V(G)$ in which, for each distinct pair $v,w\in V(G)$, $vw\in E(G^{d,k})$ if and only if $G$ contains a $(k,d)$ $vw$-shortcut.
+
+\begin{thm}
+  Let $\ell\ge 1$, $t\ge 2$, $d\ge 1$, and $k\ge 1$ be integers, let $G$ be a graph having a layered $I$-partition of layered width $\ell$ in which $I$ has treewidth at most $t-1$.
+Then $G^{k,d}$ has a layered $H$-partition of layered width $O(\ell kd^k)$ where the treewidth of $H$ is at most $\binom{t+k-1}{k}$.
+\end{thm}
+
+\begin{proof}
+  Let $(Q_x: x\in V(I))$ denote the given $I$-partition of $G$ and let $\langle V_1,\ldots,V_h\rangle$ be the given layering of $G$.  For a node $x\in V(T)$, let $R_x:=\bigcup_{y\in C_x} Q_{y}$.
+  Apply \lemref{nice-decomposition} to $I$ to obtain a tree-decomposition $(C_x: x\in V(T))$ of $I$ where $T$ is rooted at some node $x_0$ and that satisifies Properties~\tref{t:first}--\tref{t:last}.
+(We skip the introduction of the isomorphic graph $H'$ and assume from the outset that $V(I)\subseteq V(T))$.)
+
+  For each $v\in V(G)$, let $q(v)$ denote the unique node $q\in V(I)$ such that $v\in Q_q$ and let $c(v)$ denote the root of the subtree $T[v]:=T[\{x\in V(T):q(v)\in C_x]$.
+For each $x\in V(T)$, define the following objects:
+  \begin{enumerate}
+    \item $Y_x=\{v\in V(G):c(v)=x\}$.  
+    \item $V_x=\bigcup_{y\in V(T_x)} V(Y_y)$.  
+    \item $F_x = \{w\in V(G): vw\in E(G), w\in V_x$. 
+    \item $N_x = G[V_x\cup V(F_x)]$.  
+  \end{enumerate}
+
+  These objects have the following properties:
+
+  \begin{enumerate}
+    \item $(Y_x: x\in V(T))$ is a partition of $V(G)$.
+    \item For all $i\in\{1,\ldots,h\}$, $|Y_x\cap L_i| \le t\ell$. (Since $Y_x\subseteq\bigcup_{y\in C_x} Q_y$, $|C_x|\le t$, and $|Q_y\cap L_i|\le\ell$.)
+    \item $N_x\subseteq N_{y}$ for every $T$-ancestor $y$ of $x$.
+    \item $x$ has $t'\le t$ $T$-ancestors $a_1,\ldots,a_{t'}$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
+  \end{enumerate}
+  Property~4 follows from \tref{t:ancestor-edge}:  For each edge $vw\in E(G)$ with $w\in V_x$ and $v\in Y_{a_i}$, the graph $I$ contains the edge $q(v)q(w)=a_ix'$ where (by Property~3) $x'$ is a $T$-descendant of $x$ (possibly $x'=x$).
+By \tref{t:ancestor-edge} one of $x'$ or $a_i$ is a $T$-ancestor of the other.
+For some reason (????) $a_i$ is an ancestor of $x'$.
+Therefore $a_i$ is contained in $C_x$.
+Property~4 then follows from the fact that $|C_x|\le t$.
+
+  =====
+
+  Let $(Q_x: x\in V(I))$ denote the given $I$-partition of $G$.
+For a node $x\in V(T)$, let $R_x:=\bigcup_{y\in C_x} Q_{y}$.
+For a node $x\in V(T)$, we say that a $(k,d)$ $vw$-shortcut $P$ in $G$ \emph{touches} $x$ if $V(P)\cap R_x\neq\emptyset$.
+If $G$ contains a $vw$-shortcut that touches $x$, then we say that $v$ and $w$ \emph{participate} in $x$.
+
+  For any $v\in V(G)$, the union of all $(k,d)$ shortcuts in $G$ that have $v$ as an endpoint is a connected subgraph of $G$.
+It follows that the set of nodes $x\in V(T)$ in which $v$ participates is a non-empty (connected) subtree of $T$ and we let $a(v)$ denote the root of this subtree.
+For each $x\in V(T)$, we define the \emph{separator}
+  \[  
+       S_x := \{v\in V(G): a(v)= x\} \enspace .
+  \]
+  Observe that $(S_x : x\in V(T))$ is a partition of $V(G)$.
+We let $H:=G^{k,d}/(S_x : x\in V(T))$ denote the resulting quotient graph and we let $V(H)=V(T)$ in the obvious way, so $x\in V(H)$ is the vertex obtained by contracting $S_x$ in $G^{k,d}$.
+
+  For a node $x$ in $T$ with ancestors $x_0,x_1,\ldots,x_r=x$ that roots a maximal subtree $T_x$, we define several subgraphs of $G$:  $\overline{G}_x = G[\bigcup_{y\in V(T_x)} R_y$, $G_x=\overline{G}_x\setminus(\bigcup_{i=1}^{r-1} R_{x_i})$, $F_x=G[\{w\in V(G): vw\in E(G), v\in V(G_x)\}]$, $Y_x=G[\{v\in V(G): a(v)=x]$
+
+   $F_x=G[R_x\setminus \bigcup_{i=1}^{r-1} R_{x_i}]$.  We let $T_x$ denote the subtree of $T$ rooted at $x$ and let $G_x := G[\bigcup_{y\in V(T_x)} R_y\setminus R_x]$.
+
+  We claim that, for any edge $xy\in E(H)$, $x$ is a $T$-ancestor of $y$ or vice-versa.  Indeed, if $xy\in E(H)$, then there exists an edge $vw\in E(G^{k,d})$ with $v\in S_x$ and $w\in S_y$.
+
+  Therefore $G$ contains a $vw$-shortcut $P$ where $v\in Q_{x'}$ for some $T$-descendant $x'$ of $x$ and $w\in Q_{y'}$ for some $T$-descendant $y'$ of $y$.
+If neither $x$ nor $y$ is were a $T$-ancestor of the other, then neither $x'$ nor $y'$ would be an ancestor of the the other, but both $v$ and $w$ would participate in a node $z$ that is a $T$-ancestor of both $x$ and $y$.
+This would contradict the fact that $v\in S_x$ and $w\in S_y$.
+
+  We construct a tree-decomposition $\mathcal{T}:=(B_x:x\in V(T))$ of $I$ as follows: For each edge $xy\in E(I)$ we include $x$ in $B_z$ for each node $z$ on the path, in $T$ from $x$ to $y$.
+This is clearly a tree-decomposition of $I$ and it has the property that, for any $x\in V(I)$, $T[x]$ is rooted at $x$.
+All that remains is to bound the width of $\mathcal{T}$.
+
+  We now focus on one particular bag $B_x$ for some $x\in V(T)$.  
+  Let $x_0,\ldots,x_d$ be the path from $x_0:=x$ to the root of $T$.  
+  Let $T_i$ denote the subtree of $T$ rooted at $x_i$ and let
+  \[ 
+      G_i := G[\bigcup_{x'\in V(T_i)}\bigcup Q_{x'}]-Q_{x_{i-1}} \enspace .
+  \]
+  In other words, $G_i$ is the part of $G$ that contains that is separated by the removal of $S$
+
+  Define $N_{x_0}:=G$ and, for each $i\in\{1,\ldots,d\}$, define $N_{x_i}$ to be the subgraph of $G$ induced by the union, over all $x'$ in the subtree of
+
+  % containing exactly those components of $N_{x_{i-1}}-Q_{x_i}$ that contain vertices in $
+
+  As discussed above, for any node $x\in V(T)$, the bag $B_x$ contains only $T$-ancestors of $x$ (including $x$ itself).
+If $B_x$ contains some $T$-ancestor $y$ then this is because $xy\in E(I)$ which implies that there exists $vw\in E(G')$ with $v\in S_{x'}$ and $w\in S_y$ for some $T$-descendant $x'$ of $x$.
+This implies that $G$ contains a $vw$-shortcut $P=v_0,\ldots,v_p$ with $v_0=v$ and $v_p=w$.
+  Note that neither $v$ nor $w$ participates any $T$-ancestor of $y$ other than $y$.  Therefore $P$ does not contain any vertices of $Q_{y'}$ for any ancestor $y'$ of $y$.
 \end{proof}

### 2019-07-03 k-planar.tex
Moving....

@@
     \item\label{t:node-set} $V(H')\subseteq V(T)$.
-    \item\label{t:diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y||\le 1$.
+    \item\label{t:diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
     \item\label{t:subtree-root} For each $v'\in V(H')$, the subtree $T[v']:=T[\{x\in V(T):v'\in B_x\}]$ is rooted at $v'$.
@@
 
-\begin{thm}
-  There exists a tree $T$ rooted at a node $x_0$ where each node $x\in V(T)$ is associated with a subset $Y_x\subseteq V(G)$ that defines several related sets $V_x = \bigcup_{y\in V(T_x) Y_y}$, $F_x=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x=V_x\cup F_x$ that satisfy the following properties:
+\begin{thm}\thmlabel{generalized-tripod}
+  There exists a tree $T$ rooted at a node $x_0$ where each node $x\in V(T)$ is associated with a subset $Y_x\subseteq V(G)$ that defines several related sets $V_x = \bigcup_{y\in V(T_x)} Y_y$, $F_x=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x=V_x\cup F_x$ that satisfy the following properties:
   \begin{compactenum}[(Y1)]
@@
 
+Before proving \thmref{generalized-tripod} we point out another properties that are immediately implied by it:
+
+\begin{compactenum}[(Y1)]\setcounter{enumi}{3}
+  \item\ylabel{containment} $V_x\subseteq V_a$ and $N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
+\end{compactenum}
+
 \begin{proof}
@@
 
-  For each $v\in V(G)$, let $q(v)$ denote the unique node $q\in V(I)$ such that $v\in Q_q$ and let $b(v)$ denote the root of the subtree $T[v]:=T[\{x\in V(T):q(v)\in B_x]$.
+  For each $v\in V(G)$, let $q(v)$ denote the unique node $q\in V(I)$ such that $v\in Q_q$ and let $b(v)$ denote the root of the subtree $T[v]:=T[\{x\in V(T):q(v)\in B_x\}]$.
 For each $x\in V(T)$, define $Y_x=\{v\in V(G):c(v)=x\}$.
-  Clearly, $\mathcal{Y}:=(Y_x:x\in V(T))$ is a partition of $V(G)$.  Furthermore, for any $i\in\in\{1,\ldots,h\}$ 
-  \[  |Y_x| = \sum_{y\in C_x}|Q_y\cap V_i|\le |C_x|\ell \le t\ell ,
+  Clearly, $\mathcal{Y}:=(Y_x:x\in V(T))$ is a partition of $V(G)$.  Furthermore, for any $i\in\{1,\ldots,h\}$ 
+  \[  |Y_x\cap V_i| = \sum_{y\in C_x}|Q_y\cap V_i|\le |C_x|\ell \le t\ell ,
   \] 
@@
 Therefore $H$ contains the edge $ax'$.
-
-  By \tref{ancestor-edge}, $a$ is an ancestor of $x'$  
+By \tref{ancestor-edge}, $a$ is an ancestor of $x'$ or $x'$ is an ancestor of $a$.
+However, \yref{containment} rules out the possibility that $x'$ is a $T$-ancestor of $a$.
+Therefore, $a$ is a $T$-ancestor of $x$ which is a $T$-ancestor of $x'$.
+Let $x'=z_0,\ldots,z_r=a$ be the path, in $T$ from $x'$ to $a$.
+For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $C_{z_i}$.
+However, $b(x')=x'$, so $x'$ is not contained in $C_{x_i}$ for any $i\in\{1,\ldots,r\}$.
+Therefore $a\in C_{x_i}$ for each $i\in\{0,\ldots,r\}$.
+In particular, $a$ is contained in $C_x$.
+  Property~\yref{ancestor-edge} then immediately follows from the fact that $|C_x|\le t$.
 \end{proof}

### 2019-07-04 k-planar.tex
Working....

@@
     \item\ylabel{separator} There is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$. 
-    \item\ylabel{ancestor-edge} $x$ has $t'\le t$ $T$-ancestors $a_1,\ldots,a_{t'}$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
+    \item\ylabel{ancestor-edge} $x$ has $t'\le t$ strict $T$-ancestors $a_1,\ldots,a_{t'}$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
   \end{compactenum}
@@
 
-Before proving \thmref{generalized-tripod} we point out another properties that are immediately implied by it:
+Before proving \thmref{generalized-tripod} we point two another properties that are immediately implied by it:
 
 \begin{compactenum}[(Y1)]\setcounter{enumi}{3}
+  \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$. 
   \item\ylabel{containment} $V_x\subseteq V_a$ and $N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
@@
 
+Let $T$ and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ be the decomposition obtained by applying \thmref{generalized-tripod} ot $G$.
+For a node $x\in V(T)$, we say that a $(k,d)$ $vw$-shortcut $P$ in $G$ \emph{touches} $x$ if $v\in Y_x$ or $V_x$ contains an internal vertex of $P$.
+If $G$ contains a $vw$-shortcut that touches $x$, then we say that $v$ and $w$ \emph{participate} in $x$.
+
+For any $v\in V(G)$, the union of all $(k,d)$ shortcuts in $G$ that have $v$ as an endpoint is a connected subgraph of $G$.
+It follows that the set of nodes $x\in V(T)$ in which $v$ participates is a non-empty (connected) subtree of $T$ and we let $a(v)$ denote the root of this subtree.
+For each $x\in V(T)$, we define the \emph{separator}
+\[  
+     S_x := \{v\in V(G): a(v)= x\} \enspace .
+\]
+Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
+We let $I:=G^{k,d}/(S_x : x\in V(T))$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{k,d}$.
+(Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $I$.)
+
+\begin{clm}
+  For every $x\in V(T)$, $S_x\subseteq N_x$.
+\end{clm}
+
+\begin{proof}
+  For the sake of contradiction, assume otherwise.
+  Then, $G$ contains a $(k,d)$ $vw$-shortcut $P$ that contains an internal vertex $v'\in Y_x$ and such that $w\not\in N_x$.
+By \yref{separator}, $P$ contains an internal vertex $v''\in F_x$.
+By \yref{ancestor-edge}, $v''\in Y_a$ for some strict $T$-ancestor $a$ of $x$, contradicting the assumption that $v\in S_x$ (since $v$ would would be in $Y_{a'}$ for some $T$-ancestor $a'$ of $a$).
+\end{proof}
+
+Next we show that $\mathcal{S}$ has small layered width with respect to $\mathcal{L}$:
+
+\begin{clm}
+  For each $i\in\{1,\ldots,h\}$ and each $x\in V(I)$, $|S_x\cap V_i|\le t\ell(1+(d-1)^{k-1})$.  
+\end{clm}
+
+\begin{proof}
+  Consider some vertex $v\in S_x\cap L_i$.  There are two cases to consider:
+  \begin{enumerate}
+    \item $v\in Y_x$.  By \yref{thickness}, there are at most $t\ell$ such vertices in $S_x\cap L_i$.
+    \item $G$ contains a $(k,d)$ $vw$-shortcut $P$ and $Y_x$ contains an internal vertex $v'$ of $P$.
+In this case, $v'\in L_{i+j}\cap Y_x$ for some $j\in\{-k+1,\ldots,k-1\}$.
+Since $|L_{i+j}\cap Y_x|\le t\ell$, the length of $P$ is at most $k-1$, and the internal vertices of $P$ have degree at most $d$ in $G$, there are at most $t\ell(d-1)^{k-1}$ such vertices in $S_x\cap L_i$.
+\qedhere
+  \end{enumerate}
+\end{proof}
+
+\begin{clm}
+  For each edge $xy\in E(I)$, one of $x$ or $y$ is a $T$-ancestor of the other.
+\end{clm}
+
+\begin{proof}
+  Consider the lowest common $T$-ancestor $a$ of $x$ and $y$ and assume, for the sake of contradiction that $a\not\in\{x,y\}$.
+
+  By \yref{separator}, $G-(F_a\cup Y_a)$ does not contain any path from a vertex $v\in V_x$ to a vertex $w\in V_y$.  
+\end{proof}
+
+\begin{clm}
+  The graph $I$ has a tree-decomposition whose largest bag has size at most $t^k$.
+\end{clm}
+
+\begin{proof}
+  For the tree-decomposition of $I$ we use the same tree $T$ used in the tree-decomposition of $H$.  First we claim that, for each edge
+
+\end{proof}
+
+\newpage
+
 \begin{thm}
@@
 
-  Let $(Q_x: x\in V(I))$ denote the given $I$-partition of $G$.
-For a node $x\in V(T)$, let $R_x:=\bigcup_{y\in C_x} Q_{y}$.
+  Let $(Q_x: x\in V(I))$ denote the given $I$-partition of $G$.  For a node $x\in V(T)$, let $R_x:=\bigcup_{y\in C_x} Q_{y}$.
+
   For a node $x\in V(T)$, we say that a $(k,d)$ $vw$-shortcut $P$ in $G$ \emph{touches} $x$ if $V(P)\cap R_x\neq\emptyset$.

### 2019-07-04 k-planar.tex
It's gonna work

--- 2019-07-03 k-planar.tex
@@
 \newcommand{\jref}[1]{(J\ref{j:#1})}
-\newcommand{\tref}[1]{(T\ref{#1})}
+\newcommand{\tlabel}[1]{\label{t:#1}}
+\newcommand{\tref}[1]{(T\ref{t:#1})}
 \newcommand{\ylabel}[1]{\label{y:#1}}
@@
   \begin{compactenum}[(T1)]
-    \item\label{t:rooted}\label{t:first} $T$ is rooted at some node $x_0\in V(T)$ with $B_{x_0}=\emptyset$.
-    \item\label{t:node-set} $V(H')\subseteq V(T)$.
-    \item\label{t:diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
-    \item\label{t:subtree-root} For each $v'\in V(H')$, the subtree $T[v']:=T[\{x\in V(T):v'\in B_x\}]$ is rooted at $v'$.
-    \item\label{t:ancestor-edge}\label{t:last} For each edge $v'w'\in E(H)$, one of $v'$ or $w'$ is a $T$-ancestor of the other.
+    \item\tlabel{rooted}\tlabel{first} $T$ is rooted at some node $x_0\in V(T)$ with $B_{x_0}=\emptyset$.
+    \item\tlabel{node-set} $V(H')\subseteq V(T)$.
+    \item\tlabel{diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
+    \item\tlabel{subtree-root} For each $v'\in V(H')$, the subtree $T[v']:=T[\{x\in V(T):v'\in B_x\}]$ is rooted at $v'$.
+    \item\tlabel{ancestor-edge}\tlabel{last} For each edge $v'w'\in E(H)$, one of $v'$ or $w'$ is a $T$-ancestor of the other.
   \end{compactenum}
@@
 Let $T$ be rooted at $x_0$.
-Note that $(C_x:x\in V(T))$ satisfies Properties~\tref{t:rooted}, and \tref{t:diff}.
+Note that $(C_x:x\in V(T))$ satisfies Properties~\tref{rooted}, and \tref{diff}.
 For the remaining properties some renaming is required, which we now describe.
@@
 The uniqueness of $v'$ follows from the fact that $T[v]:=T[\{x\in V(T):v\in B_x\}]$ is connected, so $v'$ is the root of $T[v]$.
-Furthermore, for any pair of distinct vertices $v,w\in V(H)$, \tref{t:rooted} and \tref{t:diff} ensure that $v'\neq w'$ since, otherwise, $C_{v'}\ominus C_{x'}\supseteq \{v',w'\}$ where $x'$ is the parent of $v'=w'$.
-Property~\tref{t:rooted} would ensure that $x'$ exists and Property~\tref{t:diff} would ensure that $|C_{x'}\ominus C_{v'}|\le 1$.
+Furthermore, for any pair of distinct vertices $v,w\in V(H)$, \tref{rooted} and \tref{diff} ensure that $v'\neq w'$ since, otherwise, $C_{v'}\ominus C_{x'}\supseteq \{v',w'\}$ where $x'$ is the parent of $v'=w'$.
+Property~\tref{rooted} would ensure that $x'$ exists and Property~\tref{diff} would ensure that $|C_{x'}\ominus C_{v'}|\le 1$.
 
@@
 Obviously, $H'$ isomorphic to $H$ and $V(H')\subseteq V(T)$ and $(B_x: x\in V(T))$ is a width=$(t-1)$ tree-decomposition of $H'$.
-The decomposition $(B_x: x\in V(T))$ inherits Properties~\tref{t:rooted} and \tref{t:diff} from $(C_x:x\in V(T))$ and satisfies Property~\tref{t:node-set} and \tref{t:subtree-root} by definition.
-
-  To see that $(B_:x\in V(T))$ satisifies Property~\tref{t:ancestor-edge}, observe that, if $vw\in E(H)$, then at least one of $v'$ or $w'$ is contained in $B_x$ for every node $x$ on the path from $v'$ to $w'$ in $T$.
+The decomposition $(B_x: x\in V(T))$ inherits Properties~\tref{rooted} and \tref{diff} from $(C_x:x\in V(T))$ and satisfies Property~\tref{node-set} and \tref{subtree-root} by definition.
+
+  To see that $(B_:x\in V(T))$ satisifies Property~\tref{ancestor-edge}, observe that, if $vw\in E(H)$, then at least one of $v'$ or $w'$ is contained in $B_x$ for every node $x$ on the path from $v'$ to $w'$ in $T$.
 If neither $v'$ nor $w'$ is an ancestor of the other, then some node $x$ on this path has $T$-depth less than that of $v'$ and $w'$.
@@
 
-Before proving \thmref{generalized-tripod} we point out another properties that are immediately implied by it:
+Before proving \thmref{generalized-tripod} we point two properties that are immediately implied by it:
 
 \begin{compactenum}[(Y1)]\setcounter{enumi}{3}
-  \item\ylabel{containment} $V_x\subseteq V_a$ and $N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
+  \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$. 
+  \item\ylabel{containment-i} $V_x\subseteq V_a$ for every $T$-ancestor $a$ of $x$. 
+  \item\ylabel{containment-ii}$N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
 \end{compactenum}
 
-\begin{proof}
-  Apply \lemref{nice-decomposition} to $I$ to obtain a tree-decomposition $(B_x: x\in V(T))$ of $I$ where $T$ is rooted at some node $x_0$ and that satisifies Properties~\tref{t:first}--\tref{t:last}.
+Property~\yref{y-subsets} follows from the fact that $V_x$ is the union of several sets, one of which is $Y_x$.
+Property~\yref{containment-i} follows from the definitions of $V_x$ and the fact that the subtree $V(T_x)\subseteq V(T_a)$.
+To show Property~\yref{containment-ii} first note that, by \yref{containment-i} it suffices to consider vertices in $w\in F_x=N_x\setminus V_x$.
+By definition, every vertex $w\in F_x$ is adjacent, in $G$, to a vertex $v\in V_x$.
+By \yref{containment-i}, $v\in V_a$, so $w$ is either in $V_a$ or $w$ satisfies the condition $vw\in E(G)$, $v\in V_a$, and $w\not\in V_a$, so $w\in F_a$.
+In either case $w\in N_a=V_a\cup F_a$.
+Note that none of \yref{y-subsets}--\yref{containment-ii} depends on \yref{ancestor-edge} (which is important, since \yref{containment-i} is used to establish \yref{ancestor-edge} below).
+
+\begin{proof}[Proof of \thmref{generalized-tripod}]
+  Apply \lemref{nice-decomposition} to $I$ to obtain a tree-decomposition $(B_x: x\in V(T))$ of $I$ where $T$ is rooted at some node $x_0$ and that satisifies Properties~\tref{first}--\tref{last}.
 (We skip the introduction of the isomorphic graph $H'$ and assume from the outset that $V(I)\subseteq V(T))$.)
@@
 
-  To establish Property~\yref{ancestor-edge}, consider some edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.
-By the definition of $V_x$, $v\in Y_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
+  To establish Property~\yref{ancestor-edge}, consider some vertex $w\in F_x$.
+Since $w\in F_x$, there exists an edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.
+Since $v\in V_x$, $v\in Y_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
 Since $\mathcal{Y}$ is a partition, $w\in Y_{a}$ for some $a\in V(T)$, $a\neq x$.
@@
 By \tref{ancestor-edge}, $a$ is an ancestor of $x'$ or $x'$ is an ancestor of $a$.
-However, \yref{containment} rules out the possibility that $x'$ is a $T$-ancestor of $a$.
+Since $w\in Y_a\subseteq V_a$ and $w\not\in V_x\supseteq V_{x'}$, \yref{containment-i} rules out the possibility that $x'$ is a $T$-ancestor of $a$.
 Therefore, $a$ is a $T$-ancestor of $x$ which is a $T$-ancestor of $x'$.
@@
 In particular, $a$ is contained in $C_x$.
-  Property~\yref{ancestor-edge} then immediately follows from the fact that $|C_x|\le t$.
+  Property~\yref{ancestor-edge} now follows from the fact that $|C_x|\le t$.
 \end{proof}
@@
 The \emph{$(k,d)$th power of $G$}, denoted $G^{k,d}$, is the graph with vertex $V(G^{k,d})=V(G)$ in which, for each distinct pair $v,w\in V(G)$, $vw\in E(G^{d,k})$ if and only if $G$ contains a $(k,d)$ $vw$-shortcut.
+
+Let $T$ and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ be a decomposition obtained by applying \thmref{generalized-tripod} ot $G$, let $H:=G/\{Y_x:x\in V(T)\}$ denote the resulting quotient graph, and let $(B_x:x\in V(T))$ be a tree-decomposition of $H$ given by \lemref{nice-decomposition}.
+For simplicity, we bypass the graph $H'$ and assume, from the outset, that $V(H)\subseteq V(T)$.
+
+For two nodes $vw\in V(G)$, a $(k,d)$ \emph{$vw$-longcut} is a walk $W=v_0,\ldots,v_{r}$ in $G$ of length $r\le k$ with $v_0=v$, $v_r=w$, in which, for each $i\in\{1,\ldots,r-1\}$, the degree of $v_i$ in $G$ is at most $d$.
+Observe that every $(k,d)$ $vw$-shortcut is also a $(k,d)$ $vw$-longcut and that $G$ contains a $(k,d)$ $vw$-shortcut if and only if it contains a $(k,d)$ $vw$-longcut.
+
+For a node $x\in V(T)$, we say that a $(k,d)$ $vw$-longcut $P$ in $G$ \emph{crosses} $x$ if $Y_x$ contains an internal vertex of $P$.
+We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$ or $G$ contains a $vw$-longcut that crosses $x$.
+
+\begin{clm}
+  For every vertex $v\in V(G)$, there is a (unique) node $a(v)\in V(T)$ such that $v$ participates in $a(v)$ and $a(v)$ is a $T$-ancestor of every node in which $x$ participates. 
+\end{clm}
+
+\begin{proof}
+  Consider the subset $X$ of vertices $v'\in V(G)$ such that $G$ contains a path from $v$ to $v'$ of length at most $k$ in which every vertex other than (possibly) $v$ has degree at most $d$.
+Clearly $G[X]$ is connected and, by definition, $v$ participates in a node $x\in V(T)$ if and only if $X\cap Y_x\neq\emptyset$.
+
+  Let $X_H:=\{x\in V(H): X\cap Y_x\neq\emptyset\}$.
+The connectivity of $G[X]$ implies that $H[X_H]$ is connected.
+Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
+Choose $a(v)\in X_H$ to be any member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.
+Transitivity of the $T$-ancestor relationship, \tref{ancestor-edge}, and connectivity of $H[X_H]$ implies that $a(v)$ is a $T$-ancestor of every node $x\in X_H$, as required.
+\end{proof}
+
+For each $x\in V(T)$, we define the \emph{separator}
+\[  
+     S_x := \{v\in V(G): a(v)= x\} \enspace .
+\]
+Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
+We let $I:=G^{k,d}/(S_x : x\in V(T))$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{k,d}$.
+(Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $I$.)
+
+\begin{clm}\clmlabel{s-subset}
+  For every $x\in V(T)$, $S_x\subseteq V_x$.
+\end{clm}
+
+\begin{proof}
+  For the sake of contradiction, assume otherwise, so there exists some $v\in S_x\setminus V_x$.
+
+  By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.
+Therefore, $G$ contains a path $v=v_0,v_1,\ldots,v_{r}$ such that $r <k$, $v_r\in Y_x$ and the degree of $v_i$ in $G$ is at most $d$ for each $i\in\{1,\ldots,r\}$.
+Since $v\not\in V_x$ and $v_r\in V_x$, \yref{separator}, implies that $v_i\in F_x$ for some $i\in\{0,\ldots,r-1\}$.
+Now \yref{ancestor-edge} implies $v_i\in Y_a$ for some srict $T$-ancestor $a$ of $x$.
+But the existence of the path $v_0,\ldots,v_i$ in $G$ implies that $v=v_0\in S_{a'}$ for some ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
+\end{proof}
+
+Next we show that $\mathcal{S}$ has small layered width with respect to $\mathcal{L}$:
+
+\begin{clm}\clmlabel{general-width}
+  For each $i\in\{1,\ldots,h\}$ and each $x\in V(I)$, $|S_x\cap V_i|\le t\ell(1 + 2(k-1)d^{k-1})$.  
+\end{clm}
+
+\begin{proof}
+  Consider some vertex $v\in S_x\cap L_i$.  There are two cases to consider:
+  \begin{enumerate}
+    \item $v\in Y_x$.  By \yref{thickness}, there are at most $t\ell$ such vertices in $S_x\cap L_i$.
+    \item $G$ contains a path $v=v_0,\ldots,v_r$ with $r<k$ such that $v_r\in Y_x$ and the degree of $v_i$ in $G$ is at most $d$.
+    In this case, $v_r\in L_{i+j}\cap Y_x$ for some $j\in\{-k+1,\ldots,k-1\}$.
+    For each $w\in L_{i+j}\cap Y_x$, the number of such paths originating from $w$ is at most $d^{k-1}$. Therefore, there are at most $t\ell(d-1)^{k-1}$ such vertices in $S_x\cap L_i$. \qedhere
+  \end{enumerate}
+\end{proof}
+
+\begin{clm}\clmlabel{i-ancestor}
+  For each edge $xy\in E(I)$, one of $x$ or $y$ is a $T$-ancestor of the other.
+\end{clm}
+
+\begin{proof}
+  Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.
+Since $xy\in E(I)$, $G^{k,d}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
+Therefore $G$ contains a $(k,d)$ $vw$-shortcut $P$.
+By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.
+If neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
+By \yref{ancestor-edge} $N_x$ and $V_y$ are also disjoint.
+By \yref{separator} $P$ contains an internal vertex $v'\in F_x$.
+By \yref{ancestor-edge}, $v'\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
+But this implies that $v\in S_{a'}$ for some $T$-ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
+\end{proof}
+
+\begin{clm}\clmlabel{general-bag-size}
+  The graph $I$ has a tree-decomposition whose largest bag has size at most $t^k$.
+\end{clm}
+
+\begin{proof}
+  For the tree-decomposition $(B_x:x\in V(T))$ of $I$ we use the same tree $T$ used in the tree-decomposition of $H$.
+For each node $x$ of $T$, $B_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $I$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
+  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(I)$, $a,x'\in B_{x'}$ and the connectivity of $T[a]:=T[\{x\in V(T):a\in B_x]$ is obvious.
+Therefore $(B_x:x\in V(T))$ is indeed a tree-decomposition of $I$.
+What remains is to bound the size of the largest bag.
+
+  Consider an arbitray node $x\in V(T)$ whose ancestors are $x_0,\ldots,x_{r-1},x_r=x$.
+If $x_\delta\in B_x$, it is because $x_\delta x'\in E(I)$ for some $T$-descendant $x'$ of $x$.
+This implies $G^{k,d}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_i}$.
+This implies that $G$ contains a $(k,d)$ $vw$-shortcut.
+Since $w\in S_{x_i}$, $G$ contains a $(k,d)$ $ww'$-longcut $P_{ww'}$ that has an internal vertex in $w''\in Y_{x_i}$.
+
+  Let $w_0,\ldots,w_p$ be the subpath of $P_{ww'}$ that begins at $w_0:=w$ and ends at $w_p:=w''$.
+For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j: \{w_0,\ldots,w_i\}\subseteq V_{x_j}\}\}$.
+Note that $s_0,\ldots,s_p$ is a non-increasing sequence with $s_0=r$ and $s_p=\delta$.
+By \yref{ancestor-edge}, given $s_i$, there are at most $t+1$ possible values of $s_{i+1}$:  Either $s_{i+1}=s_i$ or $s_{i+1}$ is the index of one of the $t$ ancestors described in \yref{ancestor-edge}.
+This immediately implies that $|B_x|\le \sum_{i=0}^{k-1} t^i < t^k$ for $t\ge 2$.
+But we can do better....
+\end{proof}
+
+Together, \clmref{general-width} and \clmref{general-bag-size} imply the following generalization of \thmref{k-planar}:
 
@@
 \end{thm}
-
-\begin{proof}
-  Let $(Q_x: x\in V(I))$ denote the given $I$-partition of $G$ and let $\langle V_1,\ldots,V_h\rangle$ be the given layering of $G$.  For a node $x\in V(T)$, let $R_x:=\bigcup_{y\in C_x} Q_{y}$.
-  Apply \lemref{nice-decomposition} to $I$ to obtain a tree-decomposition $(C_x: x\in V(T))$ of $I$ where $T$ is rooted at some node $x_0$ and that satisifies Properties~\tref{t:first}--\tref{t:last}.
-(We skip the introduction of the isomorphic graph $H'$ and assume from the outset that $V(I)\subseteq V(T))$.)
-
-  For each $v\in V(G)$, let $q(v)$ denote the unique node $q\in V(I)$ such that $v\in Q_q$ and let $c(v)$ denote the root of the subtree $T[v]:=T[\{x\in V(T):q(v)\in C_x]$.
-For each $x\in V(T)$, define the following objects:
-  \begin{enumerate}
-    \item $Y_x=\{v\in V(G):c(v)=x\}$.  
-    \item $V_x=\bigcup_{y\in V(T_x)} V(Y_y)$.  
-    \item $F_x = \{w\in V(G): vw\in E(G), w\in V_x$. 
-    \item $N_x = G[V_x\cup V(F_x)]$.  
-  \end{enumerate}
-
-  These objects have the following properties:
-
-  \begin{enumerate}
-    \item $(Y_x: x\in V(T))$ is a partition of $V(G)$.
-    \item For all $i\in\{1,\ldots,h\}$, $|Y_x\cap L_i| \le t\ell$. (Since $Y_x\subseteq\bigcup_{y\in C_x} Q_y$, $|C_x|\le t$, and $|Q_y\cap L_i|\le\ell$.)
-    \item $N_x\subseteq N_{y}$ for every $T$-ancestor $y$ of $x$.
-    \item $x$ has $t'\le t$ $T$-ancestors $a_1,\ldots,a_{t'}$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
-  \end{enumerate}
-  Property~4 follows from \tref{t:ancestor-edge}:  For each edge $vw\in E(G)$ with $w\in V_x$ and $v\in Y_{a_i}$, the graph $I$ contains the edge $q(v)q(w)=a_ix'$ where (by Property~3) $x'$ is a $T$-descendant of $x$ (possibly $x'=x$).
-By \tref{t:ancestor-edge} one of $x'$ or $a_i$ is a $T$-ancestor of the other.
-For some reason (????) $a_i$ is an ancestor of $x'$.
-Therefore $a_i$ is contained in $C_x$.
-Property~4 then follows from the fact that $|C_x|\le t$.
-
-  =====
-
-  Let $(Q_x: x\in V(I))$ denote the given $I$-partition of $G$.
-For a node $x\in V(T)$, let $R_x:=\bigcup_{y\in C_x} Q_{y}$.
-For a node $x\in V(T)$, we say that a $(k,d)$ $vw$-shortcut $P$ in $G$ \emph{touches} $x$ if $V(P)\cap R_x\neq\emptyset$.
-If $G$ contains a $vw$-shortcut that touches $x$, then we say that $v$ and $w$ \emph{participate} in $x$.
-
-  For any $v\in V(G)$, the union of all $(k,d)$ shortcuts in $G$ that have $v$ as an endpoint is a connected subgraph of $G$.
-It follows that the set of nodes $x\in V(T)$ in which $v$ participates is a non-empty (connected) subtree of $T$ and we let $a(v)$ denote the root of this subtree.
-For each $x\in V(T)$, we define the \emph{separator}
-  \[  
-       S_x := \{v\in V(G): a(v)= x\} \enspace .
-  \]
-  Observe that $(S_x : x\in V(T))$ is a partition of $V(G)$.
-We let $H:=G^{k,d}/(S_x : x\in V(T))$ denote the resulting quotient graph and we let $V(H)=V(T)$ in the obvious way, so $x\in V(H)$ is the vertex obtained by contracting $S_x$ in $G^{k,d}$.
-
-  For a node $x$ in $T$ with ancestors $x_0,x_1,\ldots,x_r=x$ that roots a maximal subtree $T_x$, we define several subgraphs of $G$:  $\overline{G}_x = G[\bigcup_{y\in V(T_x)} R_y$, $G_x=\overline{G}_x\setminus(\bigcup_{i=1}^{r-1} R_{x_i})$, $F_x=G[\{w\in V(G): vw\in E(G), v\in V(G_x)\}]$, $Y_x=G[\{v\in V(G): a(v)=x]$
-
-   $F_x=G[R_x\setminus \bigcup_{i=1}^{r-1} R_{x_i}]$.  We let $T_x$ denote the subtree of $T$ rooted at $x$ and let $G_x := G[\bigcup_{y\in V(T_x)} R_y\setminus R_x]$.
-
-  We claim that, for any edge $xy\in E(H)$, $x$ is a $T$-ancestor of $y$ or vice-versa.  Indeed, if $xy\in E(H)$, then there exists an edge $vw\in E(G^{k,d})$ with $v\in S_x$ and $w\in S_y$.
-
-  Therefore $G$ contains a $vw$-shortcut $P$ where $v\in Q_{x'}$ for some $T$-descendant $x'$ of $x$ and $w\in Q_{y'}$ for some $T$-descendant $y'$ of $y$.
-If neither $x$ nor $y$ is were a $T$-ancestor of the other, then neither $x'$ nor $y'$ would be an ancestor of the the other, but both $v$ and $w$ would participate in a node $z$ that is a $T$-ancestor of both $x$ and $y$.
-This would contradict the fact that $v\in S_x$ and $w\in S_y$.
-
-  We construct a tree-decomposition $\mathcal{T}:=(B_x:x\in V(T))$ of $I$ as follows: For each edge $xy\in E(I)$ we include $x$ in $B_z$ for each node $z$ on the path, in $T$ from $x$ to $y$.
-This is clearly a tree-decomposition of $I$ and it has the property that, for any $x\in V(I)$, $T[x]$ is rooted at $x$.
-All that remains is to bound the width of $\mathcal{T}$.
-
-  We now focus on one particular bag $B_x$ for some $x\in V(T)$.  
-  Let $x_0,\ldots,x_d$ be the path from $x_0:=x$ to the root of $T$.  
-  Let $T_i$ denote the subtree of $T$ rooted at $x_i$ and let
-  \[ 
-      G_i := G[\bigcup_{x'\in V(T_i)}\bigcup Q_{x'}]-Q_{x_{i-1}} \enspace .
-  \]
-  In other words, $G_i$ is the part of $G$ that contains that is separated by the removal of $S$
-
-  Define $N_{x_0}:=G$ and, for each $i\in\{1,\ldots,d\}$, define $N_{x_i}$ to be the subgraph of $G$ induced by the union, over all $x'$ in the subtree of
-
-  % containing exactly those components of $N_{x_{i-1}}-Q_{x_i}$ that contain vertices in $
-
-  As discussed above, for any node $x\in V(T)$, the bag $B_x$ contains only $T$-ancestors of $x$ (including $x$ itself).
-If $B_x$ contains some $T$-ancestor $y$ then this is because $xy\in E(I)$ which implies that there exists $vw\in E(G')$ with $v\in S_{x'}$ and $w\in S_y$ for some $T$-descendant $x'$ of $x$.
-This implies that $G$ contains a $vw$-shortcut $P=v_0,\ldots,v_p$ with $v_0=v$ and $v_p=w$.
-  Note that neither $v$ nor $w$ participates any $T$-ancestor of $y$ other than $y$.  Therefore $P$ does not contain any vertices of $Q_{y'}$ for any ancestor $y'$ of $y$.
-\end{proof}
 

### 2019-07-04 k-planar.tex
Done first draft!

@@
 For two sets $A$ and $B$, we use $A\ominus B := (A\cup B)\setminus(A\cap B)$ to denote the symmetric difference between $A$ and $B$.
-We begin with a lemma about normalized tree decompositions:
+We begin with a fairly standard lemma about normalized tree decompositions:
 
@@
 
-  To see that $(B_:x\in V(T))$ satisifies Property~\tref{ancestor-edge}, observe that, if $vw\in E(H)$, then at least one of $v'$ or $w'$ is contained in $B_x$ for every node $x$ on the path from $v'$ to $w'$ in $T$.
+  To see that $(B_x:x\in V(T))$ satisifies Property~\tref{ancestor-edge}, observe that, if $vw\in E(H)$, then at least one of $v'$ or $w'$ is contained in $B_x$ for every node $x$ on the path from $v'$ to $w'$ in $T$.
 If neither $v'$ nor $w'$ is an ancestor of the other, then some node $x$ on this path has $T$-depth less than that of $v'$ and $w'$.
@@
 
-\newpage
-\subsection{Generalized Tripod Decompositions}
+\subsection{Generalized Tripod Partitions}
 
@@
 
-\begin{thm}\thmlabel{generalized-tripod}
+\begin{lem}\thmlabel{generalized-tripod}
   There exists a tree $T$ rooted at a node $x_0$ where each node $x\in V(T)$ is associated with a subset $Y_x\subseteq V(G)$ that defines several related sets $V_x = \bigcup_{y\in V(T_x)} Y_y$, $F_x=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x=V_x\cup F_x$ that satisfy the following properties:
@@
   \end{compactenum}
-\end{thm}
+\end{lem}
 
@@
 
-\newpage
 \subsection{Low-Degree Powers}
@@
 \begin{clm}\clmlabel{general-bag-size}
-  The graph $I$ has a tree-decomposition whose largest bag has size at most $t^k$.
+  The graph $I$ has a tree-decomposition whose largest bag has size at most $\binom{k+t}{t}$.
 \end{clm}
@@
 \begin{proof}
-  For the tree-decomposition $(B_x:x\in V(T))$ of $I$ we use the same tree $T$ used in the tree-decomposition of $H$.
-For each node $x$ of $T$, $B_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $I$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
-  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(I)$, $a,x'\in B_{x'}$ and the connectivity of $T[a]:=T[\{x\in V(T):a\in B_x]$ is obvious.
-Therefore $(B_x:x\in V(T))$ is indeed a tree-decomposition of $I$.
-What remains is to bound the size of the largest bag.
+  For the tree-decomposition $(C_x:x\in V(T))$ of $I$ we use the same tree $T$ used in the tree-decomposition of $H$.
+For each node $x$ of $T$, $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $I$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
+  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(I)$, $a,x'\in C_{x'}$ and the connectivity of $T[a]:=T[\{x\in V(T):a\in C_x]$ is obvious.
+Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $I$.
+What remains is to bound the size, $\max\{|C_x|:x\in V(T)\}$, of the largest bag.
 
   Consider an arbitray node $x\in V(T)$ whose ancestors are $x_0,\ldots,x_{r-1},x_r=x$.
-If $x_\delta\in B_x$, it is because $x_\delta x'\in E(I)$ for some $T$-descendant $x'$ of $x$.
-This implies $G^{k,d}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_i}$.
-This implies that $G$ contains a $(k,d)$ $vw$-shortcut.
-Since $w\in S_{x_i}$, $G$ contains a $(k,d)$ $ww'$-longcut $P_{ww'}$ that has an internal vertex in $w''\in Y_{x_i}$.
-
-  Let $w_0,\ldots,w_p$ be the subpath of $P_{ww'}$ that begins at $w_0:=w$ and ends at $w_p:=w''$.
-For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j: \{w_0,\ldots,w_i\}\subseteq V_{x_j}\}\}$.
-Note that $s_0,\ldots,s_p$ is a non-increasing sequence with $s_0=r$ and $s_p=\delta$.
-By \yref{ancestor-edge}, given $s_i$, there are at most $t+1$ possible values of $s_{i+1}$:  Either $s_{i+1}=s_i$ or $s_{i+1}$ is the index of one of the $t$ ancestors described in \yref{ancestor-edge}.
-This immediately implies that $|B_x|\le \sum_{i=0}^{k-1} t^i < t^k$ for $t\ge 2$.
-But we can do better....
+To avoid triple-subscripts in what follows, we abuse notation slightly by using $V_i$, $F_i$, and $N_i$,  as shorthands for $V_{x_i}$, $F_{x_i}$ and $N_{x_i}$, respectively.
+
+  If $x_\delta\in C_x$, it is because $x_\delta x'\in E(I)$ for some $T$-descendant $x'$ of $x$.
+This implies $G^{k,d}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
+This implies that $G$ contains a $(k,d)$ $vw$-shortcut $P_{vw}$.
+Let $v'$ be the second-last vertex of $P_{vw}$ (so $v'w\in E(G)$).
+
+  Since $w\in S_{\delta}$, at least one of the following is true:
+  \begin{enumerate}
+    \item $w\in Y_\delta$; or 
+    \item $G$ contains a $(k,d)$ $ww'$-longcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$.
+  \end{enumerate}
+  In the former case we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$.  Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{x_i}$.
+
+  Let $w_0,w_1,\ldots,w_p$ be the path that begins $w_0:=v'$ and then follows the subpath of $P_{ww'}$ that begins at $w_1:=w$ and ends at $w_p:=w''$.
+For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{1,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.
+Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root is non-decreasing.
+
+  By definition, $a_0=x_r$.  We claim that $a_p=x_\delta$, i.e., $s_p=\delta$. 
+  To see this, first observe that, for each $i\in\{1,\ldots,p\}$, $w_p\in V_{\delta}$ since, otherwise, an internal vertex of $P_{ww'}$ belong to $F_\delta$, which would imply (by \yref{ancestor-edge}) that $w\in S_{\delta'}$ for some $\delta' < \delta$, contradicting the assumption that $w\in S_\delta$.
+Therefore $s_p\ge\delta$.
+To see that $s_p>\delta-1$,
+  observe that either $w=w''\in Y_\delta$ or $P_{ww'}$ contains an internal vertex $w''$ in $Y_\delta$.
+By \yref{thickness} and the defnition of $V_x$, $V_{\delta-1}$ does not contain $w''$, so $s_p>\delta-1$.
+
+  Let $H^+$ denote the supergraph of $H$ with vertex set $V(T)$ and in which $xy\in E(H^+)$ if and only there exists some $z\in V(T)$ such that $x,y\in B_z$.
+Note that, by \tref{ancestor-edge}, $xy\in E(H^+)$ if and only if $x\in B_y$ or $y\in B_x$.
+
+  Next we observe that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \emph{lazy walk} in a graph $H$ is a walk in the non-simple graph $H'$ obtained by adding a self loop to each vertex of $H$.} in $H$.
+Indeed, if $a_i\neq a_{i+1}$ for some $i\in\{0,\ldots,p-1\}$ then this is precisely because $w_i\in V_{a_i}$ but $w_{i+1}\not\in V_{a_i}$.
+By definition, $w_i\in Y_{a_i'}$ for some $T$-descendant $a_i'$ of $a_i$.
+  By \yref{separator}, $w_{i+1}\in F_{a_i}$ so by \yref{ancestor-edge} $w_{i+1}\in Y_{a_i''}$ for some strict $T$-ancestor $a_i''$ of $a_i$.
+Since $w_iw_{i+1}\in E(G)$, $a_i'a_i''\in E(H)$.
+By \tref{subtree-root}, $a_i''\in B_{a_i''}$ and $a_i''\in B_{a_i'}$.
+Since $a_i$ is on the path from $a_i'$ to $a_i''$ in $T$ this implies that $a_i''\in B_{a_i}$.
+Therefore $a_ia_i''\in E(H^+)$ as claimed.
+
+  Thus, $a_0,\ldots,a_p$ is a lazy walk in $H^+$ of length $p\le k$ where the distance $s_i$ between $a_i$ and the root $x_0$ of $T$ is non-decreasing.
+By removing repeated vertices this gives a path in the directed graph $\overrightarrow{H}^+$ obtained by directing each edge $xy\in E(H^+)$ from its $T$-descendant $x$ towards its $T$-ancestor $y$.
+  Finally, we are in a position to appeal to \cite[Lemma~24]{pilipczuk.siebertz:polynomial-arxiv} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t-1}{t}$.
 \end{proof}
@@
 
-\begin{thm}
-  Let $\ell\ge 1$, $t\ge 2$, $d\ge 1$, and $k\ge 1$ be integers, let $G$ be a graph having a layered $I$-partition of layered width $\ell$ in which $I$ has treewidth at most $t-1$.
-Then $G^{k,d}$ has a layered $H$-partition of layered width $O(\ell kd^k)$ where the treewidth of $H$ is at most $\binom{t+k-1}{k}$.
+\begin{thm}\thmlabel{modern-major-general}
+  Let $\ell\ge 1$, $t\ge 2$, $d\ge 1$, and $k\ge 1$ be integers, let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t-1$.
+Then $G^{k,d}$ has a layered $I$-partition of layered width at most $O(t\ell k^2d^k)$ where the treewidth of $H$ is at most $\binom{k+t-1}{t}-1$.
 \end{thm}
+
+\begin{proof}
+  The proof is almost immediate, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{k,d}$.
+In particular, $G^{k,d}$ may contain an edge $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.
+To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.
+This increases the thickness given by \clmref{general-width} from $O(t\ell kd^k)$ to $O(t\ell k^2d^k)$.
+\end{proof}
+
+It is natural to ask if \thmref{modern-major-general} is tight.
+We do not answer this question, but we do point out that, even when $t=1$ and $d=3$, an exponential dependence on $k$ is necessary.
+If $G$ is a complete binary tree on $2^{k/2}$ leaves, then $G^{k,3}$ contains a clique $K$ of size  $s=2^{k/2+1}-1$.
+If these vertices appear in $r$ parts of the $I$-partition, then the treewidth of $I$ is at least $r-1$.
+On the other hand, these vertices must all appear on two consecutive layers, so the layered width of the partition is at least $s/(2r)$.
+Therefore any layered $I$ partition of $G^{k,d}$ of layered width $\ell'$ in which the treewidth of $I$ is at most $t'-1$ must satisfy
+\[
+    \ell'\times t' \ge 2^{k/2}-1/2 \enspace .
+\]
+
+\newpage
 

### 2019-07-04 k-planar.tex
Revised page 1

@@
 \begin{lem}\lemlabel{nice-decomposition}
-  Let $H$ be a graph of treewidth $t-1$.
-Then there exists a graph $H'$ isomorphic to $H$ that has a tree-decomposition $(B_x:x\in V(T))$ of width at most $t-1$ that has following additional properties:
+  Let $H$ be a graph of treewidth $t$.
+Then there exists a graph $H'$ isomorphic to $H$ that has a tree-decomposition $(B_x:x\in V(T))$ of width at most $t$ that has following additional properties:
   \begin{compactenum}[(T1)]
@@
 \begin{proof}
-  Begin with a width-$(t-1)$ tree-decomposition $(C_x:x\in V(T_0))$ of $H$ using a tree $T_0$ with the property that, for any edge $xy\in E(T_0)$, $|C_x\ominus C_y|\le 1$.
+  Begin with a width-$t$ tree-decomposition $(C_x:x\in V(T_0))$ of $H$ using a tree $T_0$ with the property that, for any edge $xy\in E(T_0)$, $|C_x\ominus C_y|\le 1$.
 That is, $C_x$ contains exactly one vertex not in $C_y$ or vice-versa.
 It is well-known that such a tree-decomposition exists.
-
-  Next, select an arbitrary node $x\in V(T_0)$ and let $C_x:=\{v_1,\ldots,v_{t'}\}$ denote the contents of $C_x$.
+Indeed, any tree-decomposition of $H$ can be made to have this property adding $|C_x\ominus C_y|-1$ subdivision vertices along each edge $xy\in E(T_0)$ and assigning approriate bags to the newly introduced subdivision vertices.
+
+  Next, select an arbitrary node $x\in V(T_0)$ and let $\{v_1,\ldots,v_{t'}\}:=C_x$ denote the contents of $C_x$.
 Create a new tree $T$ from $T_0$ by adding a path of nodes $x_0,x_1,x_2,\ldots,x_{t'-1},x_{t'}=x$.
 Let $C_{x_0}:=\emptyset$ and, for each $i\in\{1,\ldots,t'\}$, $C_{x_i}:=\{v_1,\ldots,v_i\}$.
-Note that $(C_x: x\in V(T))$ is a width-$(t-1)$ tree-decomposition of $H$.
-Let $T$ be rooted at $x_0$.
-Note that $(C_x:x\in V(T))$ satisfies Properties~\tref{rooted}, and \tref{diff}.
+Root $T_0$ at $x_0$ and note that $(C_x: x\in V(T))$ is a width-$t$ tree-decomposition of $H$ that satisfies Properties~\tref{rooted}, and \tref{diff}.
 For the remaining properties some renaming is required, which we now describe.
@@
 The uniqueness of $v'$ follows from the fact that $T[v]:=T[\{x\in V(T):v\in B_x\}]$ is connected, so $v'$ is the root of $T[v]$.
-Furthermore, for any pair of distinct vertices $v,w\in V(H)$, \tref{rooted} and \tref{diff} ensure that $v'\neq w'$ since, otherwise, $C_{v'}\ominus C_{x'}\supseteq \{v',w'\}$ where $x'$ is the parent of $v'=w'$.
-Property~\tref{rooted} would ensure that $x'$ exists and Property~\tref{diff} would ensure that $|C_{x'}\ominus C_{v'}|\le 1$.
+Furthermore, for any pair of distinct vertices $v,w\in V(H)$, \tref{rooted} and \tref{diff} ensure that $v'\neq w'$: Otherwise, $C_{v'}\ominus C_{x'}\supseteq \{v',w'\}$ where $x'$ is the parent of $v'=w'$.
+Property~\tref{rooted} ensures that $x'$ exists but then Property~\tref{diff} contradicts $C_{v'}\ominus C_{x'}\supseteq \{v',w'\}$.
 
@@
 Now define the graph $H'$ to have vertex set $V(H')=\{v':v\in V(H)\}$ and edge set $E(H')=\{v'w': vw\in E(H)\}$.
-Obviously, $H'$ isomorphic to $H$ and $V(H')\subseteq V(T)$ and $(B_x: x\in V(T))$ is a width=$(t-1)$ tree-decomposition of $H'$.
-The decomposition $(B_x: x\in V(T))$ inherits Properties~\tref{rooted} and \tref{diff} from $(C_x:x\in V(T))$ and satisfies Property~\tref{node-set} and \tref{subtree-root} by definition.
-
-  To see that $(B_x:x\in V(T))$ satisifies Property~\tref{ancestor-edge}, observe that, if $vw\in E(H)$, then at least one of $v'$ or $w'$ is contained in $B_x$ for every node $x$ on the path from $v'$ to $w'$ in $T$.
+Obviously, $H'$ isomorphic to $H$, $V(H')\subseteq V(T)$ and $(B_x: x\in V(T))$ is a width-$t$ tree-decomposition of $H'$.
+The decomposition $(B_x: x\in V(T))$ inherits Properties~\tref{rooted} and \tref{diff} from $(C_x:x\in V(T))$ and satisfies \tref{node-set} and \tref{subtree-root} by definition.
+
+  To see that $(B_x:x\in V(T))$ satisifies \tref{ancestor-edge}, observe that, if $vw\in E(H)$, then at least one of $v'$ or $w'$ is contained in $B_x$ for every node $x$ on the path from $v'$ to $w'$ in $T$.
 If neither $v'$ nor $w'$ is an ancestor of the other, then some node $x$ on this path has $T$-depth less than that of $v'$ and $w'$.
@@
 \]   
-
 \newpage

### 2019-07-04 k-planar.tex
Revised page 2

@@
 For two sets $A$ and $B$, we use $A\ominus B := (A\cup B)\setminus(A\cap B)$ to denote the symmetric difference between $A$ and $B$.
+For a rooted tree $T$ and any node $x\in V(T)$, we define $T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}]$ to be the maximal subtree of $T$ rooted at $x$.
 We begin with a fairly standard lemma about normalized tree decompositions:
@@
 
-Let $G$ be a graph and let $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{P}:=(Q_x: x\in V(H)))$ be a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{P}$ is at most $t$.
-In this section we define a hierarchical structure on $G$ that can be obtained from $(\mathcal{L},\mathcal{P})$.
+For the remainder of this section, $G$ is a graph and $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{P}:=(Q_x: x\in V(H)))$ is a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{P}$ is at most $t-1$.
+The following lemma shows how to use $(\mathcal{L},\mathcal{P})$ to define a hierarchical decomposition of $G$.
 
@@
   \begin{compactenum}[(Y1)]
-    \item\ylabel{thickness} $\mathcal{Y}:=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width (with respect to $\mathcal{L}$) at most $t\ell$.
+    \item\ylabel{thickness} $\mathcal{Y}:=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width (with respect to $\mathcal{L}$) at most $\ell$.
     \item\ylabel{separator} There is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$. 
@@
 
-Before proving \thmref{generalized-tripod} we point two properties that are immediately implied by it:
+Before proving \thmref{generalized-tripod} we point out more properties that are immediately implied by it:
 
@@
 Property~\yref{y-subsets} follows from the fact that $V_x$ is the union of several sets, one of which is $Y_x$.
-Property~\yref{containment-i} follows from the definitions of $V_x$ and the fact that the subtree $V(T_x)\subseteq V(T_a)$.
+Property~\yref{containment-i} follows from the definition of $V_x$ and the fact that $V(T_x)\subseteq V(T_a)$.
 To show Property~\yref{containment-ii} first note that, by \yref{containment-i} it suffices to consider vertices in $w\in F_x=N_x\setminus V_x$.
@@
 In either case $w\in N_a=V_a\cup F_a$.
-Note that none of \yref{y-subsets}--\yref{containment-ii} depends on \yref{ancestor-edge} (which is important, since \yref{containment-i} is used to establish \yref{ancestor-edge} below).
+Note that none of \yref{y-subsets}--\yref{containment-ii} depends on \yref{ancestor-edge} (which is important, since \yref{containment-i} is used to establish \yref{ancestor-edge} in the following proof).
 
 \begin{proof}[Proof of \thmref{generalized-tripod}]
-  Apply \lemref{nice-decomposition} to $I$ to obtain a tree-decomposition $(B_x: x\in V(T))$ of $I$ where $T$ is rooted at some node $x_0$ and that satisifies Properties~\tref{first}--\tref{last}.
-(We skip the introduction of the isomorphic graph $H'$ and assume from the outset that $V(I)\subseteq V(T))$.)
-
-  For each $v\in V(G)$, let $q(v)$ denote the unique node $q\in V(I)$ such that $v\in Q_q$ and let $b(v)$ denote the root of the subtree $T[v]:=T[\{x\in V(T):q(v)\in B_x\}]$.
-For each $x\in V(T)$, define $Y_x=\{v\in V(G):c(v)=x\}$.
-  Clearly, $\mathcal{Y}:=(Y_x:x\in V(T))$ is a partition of $V(G)$.  Furthermore, for any $i\in\{1,\ldots,h\}$ 
-  \[  |Y_x\cap V_i| = \sum_{y\in C_x}|Q_y\cap V_i|\le |C_x|\ell \le t\ell ,
+  Apply \lemref{nice-decomposition} to $H$ to obtain a tree-decomposition $(B_x: x\in V(T))$ of $H$ where $T$ is rooted at some node $x_0$ and that satisifies Properties~\tref{first}--\tref{last}.
+(We skip the introduction of the isomorphic graph $H'$ and assume from the outset that $V(H)\subseteq V(T))$.)
+
+  For each $v\in V(G)$, let $q(v)$ denote the unique node $q\in V(H)$ such that $v\in Q_q$.  For each $x\in V(T)$, define
+  \[
+      Y_x = \begin{cases} Q_x & \text{if $x\in V(H)$} \\
+                          \emptyset & \text{otherwise} .
+            \end{cases}
   \]
+   % 
+   % 
+   % and let $b(v)$ denote the root of the subtree $T[v]:=T[\{x\in V(T):q(v)\in B_x\}]$. For each $x\in V(T)$, define $Y_x=\{v\in V(G):c(v)=x\}$.  
+  Clearly, $\mathcal{Y}:=(Y_x:x\in V(T))$ is a partition of $V(G)$ and, for any $i\in\{1,\ldots,h\}$ $|Y_x\cap V_i|\le \ell$.
   so $\mathcal{Y}$ has layered width at most $t\ell$ with respect to $\mathcal{L}$.  This establishes \yref{thickness}.
@@
 Since $w\in F_x$, there exists an edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.
-Since $v\in V_x$, $v\in Y_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
-Since $\mathcal{Y}$ is a partition, $w\in Y_{a}$ for some $a\in V(T)$, $a\neq x$.
-Since $v\in Y_{x'}$ and $w\in Y_a$, $q(v)=x'$ and $q(w)=a$, so $x'\in Q_v$ and $w\in Q_a$.
-Therefore $H$ contains the edge $ax'$.
-By \tref{ancestor-edge}, $a$ is an ancestor of $x'$ or $x'$ is an ancestor of $a$.
+Since $v\in V_x$, $v\in Y_{x'}=Q_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
+Since $\mathcal{Y}$ is a partition, $w\in Y_{a}=Q_a$ for some $a\not\in V(T_x)$.
+Since $vw\in E(G)$, $x'a\in E(H)$.
+By \tref{ancestor-edge}, one of $a$ or $x'$ is a $T$-ancestor of the other.
 Since $w\in Y_a\subseteq V_a$ and $w\not\in V_x\supseteq V_{x'}$, \yref{containment-i} rules out the possibility that $x'$ is a $T$-ancestor of $a$.
@@
 For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $C_{z_i}$.
-However, $b(x')=x'$, so $x'$ is not contained in $C_{x_i}$ for any $i\in\{1,\ldots,r\}$.
+However, by \tref{subtree-root} $x'$ is not contained in $C_{x_i}$ for any $i\in\{1,\ldots,r\}$.
 Therefore $a\in C_{x_i}$ for each $i\in\{0,\ldots,r\}$.
 In particular, $a$ is contained in $C_x$.
-  Property~\yref{ancestor-edge} now follows from the fact that $|C_x|\le t$.
+  Property~\yref{ancestor-edge} now follows from the fact that $|C_x|\le t+1$.
 \end{proof}

### 2019-07-04 k-planar.tex
Revising...

@@
 
-For the remainder of this section, $G$ is a graph and $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{P}:=(Q_x: x\in V(H)))$ is a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{P}$ is at most $t-1$.
+For the remainder of this section, $G$ is a graph and $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{P}:=(Q_x: x\in V(H)))$ is a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{P}$ is at most $t$.
 The following lemma shows how to use $(\mathcal{L},\mathcal{P})$ to define a hierarchical decomposition of $G$.
@@
 Therefore, $a$ is a $T$-ancestor of $x$ which is a $T$-ancestor of $x'$.
-Let $x'=z_0,\ldots,z_r=a$ be the path, in $T$ from $x'$ to $a$.
-For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $C_{z_i}$.
-However, by \tref{subtree-root} $x'$ is not contained in $C_{x_i}$ for any $i\in\{1,\ldots,r\}$.
-Therefore $a\in C_{x_i}$ for each $i\in\{0,\ldots,r\}$.
-In particular, $a$ is contained in $C_x$.
-  Property~\yref{ancestor-edge} now follows from the fact that $|C_x|\le t+1$.
+Let $z_0,\ldots,z_r$ be the path, in $T$ from $z_0:=x'$ to $z_r:=a$.
+For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $B_{z_i}$.
+However, by \tref{subtree-root} $x'$ is not contained in $B_{x_i}$ for any $i\in\{1,\ldots,r\}$.
+Therefore $a\in B_{x_i}$ for each $i\in\{0,\ldots,r\}$.
+In particular, $a$ is contained in $B_x$.
+  Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $C_x$ contains $x$.
 \end{proof}
@@
 
-For non-negative integers $k$ and $d$, a graph $G$ and distinct vertices $v,w\in V(G)$, a \emph{$(k,d)$ $vw$-shortcut} is a path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.
+For non-negative integers $k$ and $d$, a graph $G$ and distinct vertices $v,w\in V(G)$, a \emph{$(k,d)$ $vw$-shortcut} in $G$ is a path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.
 The \emph{$(k,d)$th power of $G$}, denoted $G^{k,d}$, is the graph with vertex $V(G^{k,d})=V(G)$ in which, for each distinct pair $v,w\in V(G)$, $vw\in E(G^{d,k})$ if and only if $G$ contains a $(k,d)$ $vw$-shortcut.

### 2019-07-04 k-planar.tex
Pffftt

@@
 For non-negative integers $k$ and $d$, a graph $G$ and distinct vertices $v,w\in V(G)$, a \emph{$(k,d)$ $vw$-shortcut} in $G$ is a path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.
-The \emph{$(k,d)$th power of $G$}, denoted $G^{k,d}$, is the graph with vertex $V(G^{k,d})=V(G)$ in which, for each distinct pair $v,w\in V(G)$, $vw\in E(G^{d,k})$ if and only if $G$ contains a $(k,d)$ $vw$-shortcut.
-
-Let $T$ and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ be a decomposition obtained by applying \thmref{generalized-tripod} ot $G$, let $H:=G/\{Y_x:x\in V(T)\}$ denote the resulting quotient graph, and let $(B_x:x\in V(T))$ be a tree-decomposition of $H$ given by \lemref{nice-decomposition}.
+The \emph{$(k,d)^\text{th}$ power of $G$}, denoted $G^{k,d}$, is the graph with vertex $V(G^{k,d})=V(G)$ in which, for each distinct pair $v,w\in V(G)$, $vw\in E(G^{k,d})$ if and only if $G$ contains a $(k,d)$ $vw$-shortcut.
+
+Let $T$ and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ be a decomposition obtained by applying \thmref{generalized-tripod} to $G$, let $H:=G/\{Y_x:x\in V(T)\}$ denote the resulting quotient graph, and let $(B_x:x\in V(T))$ be a tree-decomposition of $H$ given by \lemref{nice-decomposition}.
 For simplicity, we bypass the graph $H'$ and assume, from the outset, that $V(H)\subseteq V(T)$.
 
-For two nodes $vw\in V(G)$, a $(k,d)$ \emph{$vw$-longcut} is a walk $W=v_0,\ldots,v_{r}$ in $G$ of length $r\le k$ with $v_0=v$, $v_r=w$, in which, for each $i\in\{1,\ldots,r-1\}$, the degree of $v_i$ in $G$ is at most $d$.
+For two nodes $v,w\in V(G)$, a $(k,d)$ \emph{$vw$-longcut} is a walk $W=v_0,\ldots,v_{r}$ in $G$ of length $r\le k$ with $v_0=v$, $v_r=w$, in which, for each $i\in\{1,\ldots,r-1\}$, the degree of $v_i$ in $G$ is at most $d$.
 Observe that every $(k,d)$ $vw$-shortcut is also a $(k,d)$ $vw$-longcut and that $G$ contains a $(k,d)$ $vw$-shortcut if and only if it contains a $(k,d)$ $vw$-longcut.
 
-For a node $x\in V(T)$, we say that a $(k,d)$ $vw$-longcut $P$ in $G$ \emph{crosses} $x$ if $Y_x$ contains an internal vertex of $P$.
+For a node $x\in V(T)$, we say that a $(k,d)$ $vw$-longcut $W$ in $G$ \emph{crosses} $x$ if $Y_x$ contains an internal vertex of $W$.
 We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$ or $G$ contains a $vw$-longcut that crosses $x$.
@@
 \begin{clm}
-  For every vertex $v\in V(G)$, there is a (unique) node $a(v)\in V(T)$ such that $v$ participates in $a(v)$ and $a(v)$ is a $T$-ancestor of every node in which $x$ participates. 
+  For every vertex $v\in V(G)$, there is a (unique) node $a(v)\in V(T)$ such that $v$ participates in $a(v)$ and $a(v)$ is a $T$-ancestor of every node $x$ in which $v$ participates. 
 \end{clm}
@@
 
-  Let $X_H:=\{x\in V(H): X\cap Y_x\neq\emptyset\}$.
-The connectivity of $G[X]$ implies that $H[X_H]$ is connected.
-Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
+  Let $X_H:=\{x\in V(H): X\cap Y_x\neq\emptyset\}$.  The connectivity of $G[X]$ implies that $H[X_H]$ is connected.  
+  %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other. 
   Choose $a(v)\in X_H$ to be any member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.
@@
 Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
-We let $I:=G^{k,d}/(S_x : x\in V(T))$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{k,d}$.
+We let $I:=G^{k,d}/\mathcal{S}$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{k,d}$.
 (Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $I$.)
@@
   By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.
-Therefore, $G$ contains a path $v=v_0,v_1,\ldots,v_{r}$ such that $r <k$, $v_r\in Y_x$ and the degree of $v_i$ in $G$ is at most $d$ for each $i\in\{1,\ldots,r\}$.
+Therefore, $G$ contains a path $v_0,v_1,\ldots,v_{r}$ such that $r <k$, $v_0=v$, $v_r\in Y_x$ and the degree of $v_i$ in $G$ is at most $d$ for each $i\in\{1,\ldots,r\}$.
 Since $v\not\in V_x$ and $v_r\in V_x$, \yref{separator}, implies that $v_i\in F_x$ for some $i\in\{0,\ldots,r-1\}$.
-Now \yref{ancestor-edge} implies $v_i\in Y_a$ for some srict $T$-ancestor $a$ of $x$.
+Now \yref{ancestor-edge} implies $v_i\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
 But the existence of the path $v_0,\ldots,v_i$ in $G$ implies that $v=v_0\in S_{a'}$ for some ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
@@
 \begin{clm}\clmlabel{general-width}
-  For each $i\in\{1,\ldots,h\}$ and each $x\in V(I)$, $|S_x\cap V_i|\le t\ell(1 + 2(k-1)d^{k-1})$.  
+  For each $i\in\{1,\ldots,h\}$ and each $x\in V(I)$, $|S_x\cap V_i|\le (2k-1)\ell d^k$.  
 \end{clm}
@@
 \begin{proof}
-  Consider some vertex $v\in S_x\cap L_i$.  There are two cases to consider:
-  \begin{enumerate}
-    \item $v\in Y_x$.  By \yref{thickness}, there are at most $t\ell$ such vertices in $S_x\cap L_i$.
-    \item $G$ contains a path $v=v_0,\ldots,v_r$ with $r<k$ such that $v_r\in Y_x$ and the degree of $v_i$ in $G$ is at most $d$.
-    In this case, $v_r\in L_{i+j}\cap Y_x$ for some $j\in\{-k+1,\ldots,k-1\}$.
-    For each $w\in L_{i+j}\cap Y_x$, the number of such paths originating from $w$ is at most $d^{k-1}$. Therefore, there are at most $t\ell(d-1)^{k-1}$ such vertices in $S_x\cap L_i$. \qedhere
-  \end{enumerate}
+  Consider some vertex $v\in S_x\cap L_i$.
+Since $v\in S_x$, $G$ contains a path $v_0,\ldots,v_r$ such that $r<k$, $v_0\in Y_x$, $v_r=v$, and the degree of $v_i$ in $G$ is at most $d$, for each $i\in\{0,\ldots,r-1\}$.
+Note that $v_0$ is one of the at most $(2k-1)\ell$ vertices in $Z=\bigcup_{j=-k+1}^{k-1}(V_i\cap Y_x)$.
+For each $w\in Z$, the number of such paths originating from $w$ is at most $\sum_{r=0}^{k-1} d^{r}< d^k$ (for $d\ge 2$).
+Therefore, $|S_x\cap L_i|\le |Z|d^k\le (2k-1)\ell d^k$.
+  % \end{enumerate}
 \end{proof}
@@
 By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.
-If neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
+By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
 By \yref{ancestor-edge} $N_x$ and $V_y$ are also disjoint.
@@
 \begin{proof}
-  For the tree-decomposition $(C_x:x\in V(T))$ of $I$ we use the same tree $T$ used in the tree-decomposition of $H$.
+  For the tree-decomposition $(C_x:x\in V(T))$ of $I$ we use the same tree $T$ used in the tree-decomposition $(B_x:x\in V(T))$ of $H$.
 For each node $x$ of $T$, $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $I$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
@@
 
-  Consider an arbitray node $x\in V(T)$ whose ancestors are $x_0,\ldots,x_{r-1},x_r=x$.
+  Consider an arbitrary node $x\in V(T)$ whose ancestors are $x_0,\ldots,x_{r-1},x_r$ where $x_0$ is the root of $T$ and $x_r:=x$.
 To avoid triple-subscripts in what follows, we abuse notation slightly by using $V_i$, $F_i$, and $N_i$,  as shorthands for $V_{x_i}$, $F_{x_i}$ and $N_{x_i}$, respectively.
@@
   \end{enumerate}
-  In the former case we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$.  Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{x_i}$.
+  In the former case we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$.  Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{\delta}$.
 
@@
 For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{1,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.
-Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root is non-decreasing.
+Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root is non-increasing.
 
   By definition, $a_0=x_r$.  We claim that $a_p=x_\delta$, i.e., $s_p=\delta$. 
-  To see this, first observe that, for each $i\in\{1,\ldots,p\}$, $w_p\in V_{\delta}$ since, otherwise, an internal vertex of $P_{ww'}$ belong to $F_\delta$, which would imply (by \yref{ancestor-edge}) that $w\in S_{\delta'}$ for some $\delta' < \delta$, contradicting the assumption that $w\in S_\delta$.
+  To see this, first observe that, for each $i\in\{1,\ldots,p\}$, $w_i\in V_{\delta}$ since, otherwise, an internal vertex of $P_{ww'}$ belong to $F_\delta$, which would imply (by \yref{ancestor-edge}) that $w\in S_{\delta'}$ for some $\delta' < \delta$, contradicting the assumption that $w\in S_\delta$.
 Therefore $s_p\ge\delta$.
-To see that $s_p>\delta-1$,
+To see that $s_p<\delta-1$,
   observe that either $w=w''\in Y_\delta$ or $P_{ww'}$ contains an internal vertex $w''$ in $Y_\delta$.
-By \yref{thickness} and the defnition of $V_x$, $V_{\delta-1}$ does not contain $w''$, so $s_p>\delta-1$.
+By \yref{thickness} and the defnition of $V_x$, $V_{\delta-1}$ does not contain $w''$, so $s_p<\delta-1$.
 
   Let $H^+$ denote the supergraph of $H$ with vertex set $V(T)$ and in which $xy\in E(H^+)$ if and only there exists some $z\in V(T)$ such that $x,y\in B_z$. 
-Note that, by \tref{ancestor-edge}, $xy\in E(H^+)$ if and only if $x\in B_y$ or $y\in B_x$.
-
-  Next we observe that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \emph{lazy walk} in a graph $H$ is a walk in the non-simple graph $H'$ obtained by adding a self loop to each vertex of $H$.} in $H$.
+  % Note that, by \tref{ancestor-edge}, $xy\in E(H^+)$ if and only if $x\in B_y$ or $y\in B_x$.
+  We claim that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \emph{lazy walk} in a graph $H$ is a walk in the non-simple graph $H'$ obtained by adding a self loop to each vertex of $H$.} in $H^+$.
 Indeed, if $a_i\neq a_{i+1}$ for some $i\in\{0,\ldots,p-1\}$ then this is precisely because $w_i\in V_{a_i}$ but $w_{i+1}\not\in V_{a_i}$.
@@
 By removing repeated vertices this gives a path in the directed graph $\overrightarrow{H}^+$ obtained by directing each edge $xy\in E(H^+)$ from its $T$-descendant $x$ towards its $T$-ancestor $y$.
-  Finally, we are in a position to appeal to \cite[Lemma~24]{pilipczuk.siebertz:polynomial-arxiv} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t-1}{t}$.
+  Finally, we are in a position to appeal to \cite[Lemma~24]{pilipczuk.siebertz:polynomial-arxiv} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
 \end{proof}
@@
   Let $\ell\ge 1$, $t\ge 2$, $d\ge 1$, and $k\ge 1$ be integers, let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t-1$.
-Then $G^{k,d}$ has a layered $I$-partition of layered width at most $O(t\ell k^2d^k)$ where the treewidth of $H$ is at most $\binom{k+t-1}{t}-1$.
+Then $G^{k,d}$ has a layered $I$-partition of layered width at most $O(k^2\ell d^k)$ where the treewidth of $H$ is at most $\binom{k+t}{t}-1$.
 \end{thm}
@@
 To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.
-This increases the thickness given by \clmref{general-width} from $O(t\ell kd^k)$ to $O(t\ell k^2d^k)$.
+This increases the thickness given by \clmref{general-width} from $O(k\ell d^k)$ to $O(k^2\ell d^k)$.
 \end{proof}
@@
 We do not answer this question, but we do point out that, even when $t=1$ and $d=3$, an exponential dependence on $k$ is necessary.
-If $G$ is a complete binary tree on $2^{k/2}$ leaves, then $G^{k,3}$ contains a clique $K$ of size  $s=2^{k/2+1}-1$.
-If these vertices appear in $r$ parts of the $I$-partition, then the treewidth of $I$ is at least $r-1$.
-On the other hand, these vertices must all appear on two consecutive layers, so the layered width of the partition is at least $s/(2r)$.
-Therefore any layered $I$ partition of $G^{k,d}$ of layered width $\ell'$ in which the treewidth of $I$ is at most $t'-1$ must satisfy
+If $G$ is a complete binary tree on $2^{k/2}$ leaves, then $G^{k,3}$ is a clique of size  $s=2^{k/2+1}-1$.
+Any partition of $G^{k,3}$ into $r$ parts will give an quotient graph $I$ that is an $r$-clique, which has treewidth $r-1$.
+On the other hand, any layering of $V(G^{k,3})$ has at most two layers, so the layered width of any partition into $r$ parts is at least $s/(2r)$.
+Therefore any layered $I$-partition of $G^{k,d}$ of layered width $\ell'$ in which the treewidth of $I$ is at most $t'-1$ must satisfy
 \[

### 2019-07-05 k-planar.tex
Workgin

@@
 
-\author{Vida Dujmovi\'c%
+\author{Vida Dujmović%
         \thanks{School of Computer Science and Electrical Engineering,
@@
 
-For the remainder of this section, $G$ is a graph and $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{P}:=(Q_x: x\in V(H)))$ is a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{P}$ is at most $t$.
-The following lemma shows how to use $(\mathcal{L},\mathcal{P})$ to define a hierarchical decomposition of $G$.
+For the remainder of this section, $G$ is a graph and $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{Q}:=(Q_x: x\in V(H)))$ is a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{Q}$ is at most $t$.
+The following lemma shows how to use $(\mathcal{L},\mathcal{Q})$ to define a hierarchical decomposition of $G$.
 
@@
 
-\subsection{Low-Degree Powers}
-
-For non-negative integers $k$ and $d$, a graph $G$ and distinct vertices $v,w\in V(G)$, a \emph{$(k,d)$ $vw$-shortcut} in $G$ is a path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.
+\subsection{Shortcut Systems}
+
+% For non-negative integers $k$ and $d$, a graph $G$ and distinct vertices $v,w\in V(G)$, a \emph{$(k,d)$ $vw$-shortcut} in $G$ is a path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.
 The \emph{$(k,d)^\text{th}$ power of $G$}, denoted $G^{k,d}$, is the graph with vertex $V(G^{k,d})=V(G)$ in which, for each distinct pair $v,w\in V(G)$, $vw\in E(G^{k,d})$ if and only if $G$ contains a $(k,d)$ $vw$-shortcut.
 
-Let $T$ and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ be a decomposition obtained by applying \thmref{generalized-tripod} to $G$, let $H:=G/\{Y_x:x\in V(T)\}$ denote the resulting quotient graph, and let $(B_x:x\in V(T))$ be a tree-decomposition of $H$ given by \lemref{nice-decomposition}.
-For simplicity, we bypass the graph $H'$ and assume, from the outset, that $V(H)\subseteq V(T)$.
-
-For two nodes $v,w\in V(G)$, a $(k,d)$ \emph{$vw$-longcut} is a walk $W=v_0,\ldots,v_{r}$ in $G$ of length $r\le k$ with $v_0=v$, $v_r=w$, in which, for each $i\in\{1,\ldots,r-1\}$, the degree of $v_i$ in $G$ is at most $d$.
-Observe that every $(k,d)$ $vw$-shortcut is also a $(k,d)$ $vw$-longcut and that $G$ contains a $(k,d)$ $vw$-shortcut if and only if it contains a $(k,d)$ $vw$-longcut.
-
-For a node $x\in V(T)$, we say that a $(k,d)$ $vw$-longcut $W$ in $G$ \emph{crosses} $x$ if $Y_x$ contains an internal vertex of $W$.
-We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$ or $G$ contains a $vw$-longcut that crosses $x$.
+A collection $\mathcal{P}$ of paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if every path in $\mathcal{P}$ has length at most $k$ and if, for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
+Each path in $P\in\mathcal{P}$ is called a \emph{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \emph{$vw$-shortcut}.
+A $vw$-shortcut is also referred to as \emph{$v$-shortcut} (and a \emph{$w$-shortcut}).
+Given a graph $G$ and a $(k,d)$-shortcut system $\mathcal{P}$ for $G$, we use $G^{\mathcal{P}}$ to denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\mathcal{P}$.
+
+In this section we will show how, given a $(k,d)$-shortcut system $\mathcal{P}$, a layered $H$-partition of $G$ can be used to obtain a layered $I$-partition of $G^{\mathcal{P}}$ where the layered width of the partition does not increase dramatically and the treewidth of $I$ is not much more than the treewidth of $H$.
+
+For convenience, it will be helpful to assume that $\mathcal{P}$ contains a
+length-1 $vw$-shortcut for every edge $vw\in E(G)$.  This assumption has no effect on the resulting graph $G^{\mathcal{P}}$ but avoids special cases in some of our proofs.
+
+As in the rest of this section, $(\mathcal{L},\mathcal{Q})$ denotes a layered $H$-partition of $G$; $T$ is a tree used in a tree-decomposition $(B_x:x\in V(T))$ of $H:=G/\mathcal{Q}$ satisfying the conditions of \lemref{nice-decomposition}; and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ are obtained obtained by applying \thmref{generalized-tripod} to $G$ $(\mathcal{L},\mathcal{Q})$, and $H$.
+
+For a node $x\in V(T)$, we say that a shortcut $P\in\mathcal{P}$ \emph{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
+We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$ or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
+We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
 
 \begin{clm}
-  For every vertex $v\in V(G)$, there is a (unique) node $a(v)\in V(T)$ such that $v$ participates in $a(v)$ and $a(v)$ is a $T$-ancestor of every node $x$ in which $v$ participates. 
+  For any $v\in V(G)$ there exists a (unique) node $a(v)\in X_v$ such that
+  $a(v)$ is a $T$-ancestor of every node in $X_v$.
 \end{clm}
@@
 \begin{proof}
-  Consider the subset $X$ of vertices $v'\in V(G)$ such that $G$ contains a path from $v$ to $v'$ of length at most $k$ in which every vertex other than (possibly) $v$ has degree at most $d$.
-Clearly $G[X]$ is connected and, by definition, $v$ participates in a node $x\in V(T)$ if and only if $X\cap Y_x\neq\emptyset$.
-
-  Let $X_H:=\{x\in V(H): X\cap Y_x\neq\emptyset\}$.  The connectivity of $G[X]$ implies that $H[X_H]$ is connected.  
+  Let 
+  \[
+     Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:\mbox{$(v_0,\ldots,v_r)\in\mathcal{P}$, $v\in \{v_0,\ldots,v_r\}$}\}
+  \]
+  Clearly $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
+
+  We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
+If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
+In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$.
+
+  Let $Z_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.  The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.  
   %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other. 
@@
 Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
-We let $I:=G^{k,d}/\mathcal{S}$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{k,d}$.
+We let $I:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{k,d}$.
 (Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $I$.)
@@
   By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.
-Therefore, $G$ contains a path $v_0,v_1,\ldots,v_{r}$ such that $r <k$, $v_0=v$, $v_r\in Y_x$ and the degree of $v_i$ in $G$ is at most $d$ for each $i\in\{1,\ldots,r\}$.
-Since $v\not\in V_x$ and $v_r\in V_x$, \yref{separator}, implies that $v_i\in F_x$ for some $i\in\{0,\ldots,r-1\}$.
+Therefore, $\mathcal{P}$ contains a path $P$ that crosses $x$.
+The path $P$ contains a subpath $v_0,v_1,\ldots,v_{r}$ such that $v=v_0$ and $v_r\in Y_x$.
+Since $v\not\in V_x$ and $v_r\in Y_x\subseteq V_x$, \yref{separator}, implies that $v_i\in F_x$ for some $i\in\{0,\ldots,r-1\}$.
 Now \yref{ancestor-edge} implies $v_i\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
-But the existence of the path $v_0,\ldots,v_i$ in $G$ implies that $v=v_0\in S_{a'}$ for some ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
+Therefore, either $v\in Y_a$ or $P$ crosses $a$.
+But this implies that $a(v)$ is an ancestor of $a$, contradicting the assumption that $v\in S_x$.
 \end{proof}
@@
 \begin{clm}\clmlabel{general-width}
-  For each $i\in\{1,\ldots,h\}$ and each $x\in V(I)$, $|S_x\cap V_i|\le (2k-1)\ell d^k$.  
+  For each $i\in\{1,\ldots,h\}$ and each $x\in V(I)$, $|S_x\cap V_i|\le d\ell(k^2+k+4)/2$.  
 \end{clm}
@@
 \begin{proof}
-  Consider some vertex $v\in S_x\cap L_i$.
-Since $v\in S_x$, $G$ contains a path $v_0,\ldots,v_r$ such that $r<k$, $v_0\in Y_x$, $v_r=v$, and the degree of $v_i$ in $G$ is at most $d$, for each $i\in\{0,\ldots,r-1\}$.
-Note that $v_0$ is one of the at most $(2k-1)\ell$ vertices in $Z=\bigcup_{j=-k+1}^{k-1}(V_i\cap Y_x)$.
-For each $w\in Z$, the number of such paths originating from $w$ is at most $\sum_{r=0}^{k-1} d^{r}< d^k$ (for $d\ge 2$).
-Therefore, $|S_x\cap L_i|\le |Z|d^k\le (2k-1)\ell d^k$.
-  % \end{enumerate}
+  We will count the number of vertices in $S_x\cap V_i$ by upper-bounding the number of vertices contributed to $S_x\cap V_i$ by each vertex $w\in Y_x$.
+  If $w\in Y_x\cap V_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes only one vertex, itself, to $S_x\cap V_i$.
+
+  Otherwise, consider some path $P\in\mathcal{P}$ that contains $w$ as an internal vertex.
+If $w\in V_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap V_i$.
+If $w\in V_{i-1}\cup V_{i+1}$, then $P$ contributes at most $k$ vertices to $S_x\cap V_i$.
+If $w\in V_{i-j}\cup V_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap V_i$.
+
+  For any $j$, the number of vertices $w\in V_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\mathcal{P}$. Therefore, 
+  \[  |S_x\cap V_i|\le d\ell\cdot\left(k+1 + k + \sum_{j=2}^k (k-j)\right) %= d(2k+1) + \sum_{i=1}^{k-2} i
+      = d\ell(k^2 +k + 4)/2 \enspace . \qedhere
+  \]
 \end{proof}
@@
   Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.
-Since $xy\in E(I)$, $G^{k,d}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
-Therefore $G$ contains a $(k,d)$ $vw$-shortcut $P$.
+Since $xy\in E(I)$, $G^\mathcal{P}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
+Since $xy\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.
 By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.
@@
 \begin{clm}\clmlabel{general-bag-size}
-  The graph $I$ has a tree-decomposition whose largest bag has size at most $\binom{k+t}{t}$.
+  The graph $I$ has a tree-decomposition in which every bag has size at most $\binom{k+t}{t}$.
 \end{clm}
@@
   If $x_\delta\in C_x$, it is because $x_\delta x'\in E(I)$ for some $T$-descendant $x'$ of $x$.
-This implies $G^{k,d}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
-This implies that $G$ contains a $(k,d)$ $vw$-shortcut $P_{vw}$.
+This implies $G^{\mathcal{P}}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
+This implies that $\mathcal{P}$ a $vw$-shortcut $P_{vw}$.
 Let $v'$ be the second-last vertex of $P_{vw}$ (so $v'w\in E(G)$).
@@
   \begin{enumerate}
-    \item $w\in Y_\delta$; or 
-    \item $G$ contains a $(k,d)$ $ww'$-longcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$.
+    \item $\mathcal{P}$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
+    \item $w\in Y_\delta$.  In this case, we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$. 
   \end{enumerate}
-  In the former case we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$.  Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{\delta}$.
+  Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{\delta}$.
 
@@
 
-Together, \clmref{general-width} and \clmref{general-bag-size} imply the following generalization of \thmref{k-planar}:
+% Together, \clmref{general-width} and \clmref{general-bag-size} imply the following generalization of \thmref{k-planar}:
 
 \begin{thm}\thmlabel{modern-major-general}
-  Let $\ell\ge 1$, $t\ge 2$, $d\ge 1$, and $k\ge 1$ be integers, let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t-1$.
-Then $G^{k,d}$ has a layered $I$-partition of layered width at most $O(k^2\ell d^k)$ where the treewidth of $H$ is at most $\binom{k+t}{t}-1$.
+  Let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\mathcal{P}$ be a $(k,d)$-shortcut system for $G$.
+Then $G^{\mathcal{P}}$ has a layered $I$-partition of layered width at most $d\ell(k^3+k^2+4k)/2$ where the treewidth of $H$ is at most $\binom{k+t}{t}-1$.
 \end{thm}

### 2019-07-05 k-planar.tex
Generalization is done

@@
 \begin{proof}
-  The proof is almost immediate, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{k,d}$.
-In particular, $G^{k,d}$ may contain an edge $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.
+  The proof is almost immediate, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
+In particular, $G^{\mathcal{P}}$ may contain an edge $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.
 To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.
-This increases the thickness given by \clmref{general-width} from $O(k\ell d^k)$ to $O(k^2\ell d^k)$.
-\end{proof}
-
-It is natural to ask if \thmref{modern-major-general} is tight.
+This increases the thickness given by \clmref{general-width} from $d\ell(k^2+k+4)/2$ to $d\ell(k^3+k^2+4k)/2$.
+\end{proof}
+
+% It is natural to ask if \thmref{modern-major-general} is tight.
 We do not answer this question, but we do point out that, even when $t=1$ and $d=3$, an exponential dependence on $k$ is necessary.

### 2019-07-05 k-planar.tex
Working....

@@
 \begin{lem}\lemlabel{nice-decomposition}
-  Let $H$ be a graph of treewidth $t$.
-Then there exists a graph $H'$ isomorphic to $H$ that has a tree-decomposition $(B_x:x\in V(T))$ of width at most $t$ that has following additional properties:
+  For any graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:  
   \begin{compactenum}[(T1)]
-    \item\tlabel{rooted}\tlabel{first} $T$ is rooted at some node $x_0\in V(T)$ with $B_{x_0}=\emptyset$.
-    \item\tlabel{node-set} $V(H')\subseteq V(T)$.
-    \item\tlabel{diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
-    \item\tlabel{subtree-root} For each $v'\in V(H')$, the subtree $T[v']:=T[\{x\in V(T):v'\in B_x\}]$ is rooted at $v'$.
-    \item\tlabel{ancestor-edge}\tlabel{last} For each edge $v'w'\in E(H)$, one of $v'$ or $w'$ is a $T$-ancestor of the other.
+    % \item\tlabel{rooted}\tlabel{first} $T$ is rooted at some node $x_0\in V(T)$ with $B_{x_0}=\emptyset$.
+    % \item\tlabel{node-set} $V(T)= V(H)$.
+    % \item\tlabel{diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
+    \item\tlabel{subtree-root} For each $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$.
+    \item\tlabel{ancestor-edge}\tlabel{last} For each edge $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
   \end{compactenum}
@@
 \begin{proof}
+  Begin with any width-$t$ tree decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some rooted tree $T_0$ that is rooted at some node $x_0$ with $C_{B_0}=\emptyset$.
+Let $f:V(H)\to V(T)$ be function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y]$.
+If $f$ is one-to-one, then proceed to the next paragraph.
+Otherwise, as long as there exists some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$ we subdivide the edge between $a$ and its parent by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
+This modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
+
+  Now, consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.
+In this case, $B_{a}\subseteq B_{a'}$ where $a'$ is the parent of $a$ since any $x\in B_a\setminus B_{a'}$ would have $f(x)=a$.
+In this case, contract the edge $aa'$ in $T_0$, eliminating the node $a$.
+Repeating this operation will eventually produce a width-$t$ tree-decomposition of $(B_x:x\in V(T_0))$ where $f$ is a bijection between $V(H)$ and $V(T_0)$.
+Renaming each node $a\in T_0$ as $f^{-1}(a)$ gives a tree-decomposition $(B_x:x\in V(T))$ with $V(T)=V(H)$.
+
+  Now, for each vertex $x\in V(H)$, rename the node $f(x)$ in $T_0$ as $x$. The resulting tree decomposition uses a tree $T_0$ with $V(H)\subseteq V(T_0)$.
+
+     (or introduce a new node)   
   Begin with a width-$t$ tree-decomposition $(C_x:x\in V(T_0))$ of $H$ using a tree $T_0$ with the property that, for any edge $xy\in E(T_0)$, $|C_x\ominus C_y|\le 1$.

### 2019-07-05 k-planar.tex
Last commit on this branch

@@
 We now generalize \thmref{k-planar} to a more abstract setting.
-For two sets $A$ and $B$, we use $A\ominus B := (A\cup B)\setminus(A\cap B)$ to denote the symmetric difference between $A$ and $B$.
 For a rooted tree $T$ and any node $x\in V(T)$, we define $T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}]$ to be the maximal subtree of $T$ rooted at $x$.
@@
 \begin{proof}
-  Begin with any width-$t$ tree decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some rooted tree $T_0$ that is rooted at some node $x_0$ with $C_{B_0}=\emptyset$.
+  Begin with any width-$t$ tree decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some rooted tree $T_0$.
+Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$.
 Let $f:V(H)\to V(T)$ be function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y]$.
-If $f$ is one-to-one, then proceed to the next paragraph.
-Otherwise, as long as there exists some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$ we subdivide the edge between $a$ and its parent by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
+If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.
+Subdivide the edge between $a$ and its parent in $T$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
 This modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
 
-  Now, consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.
+  Next, consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.
 In this case, $B_{a}\subseteq B_{a'}$ where $a'$ is the parent of $a$ since any $x\in B_a\setminus B_{a'}$ would have $f(x)=a$.
@@
 Repeating this operation will eventually produce a width-$t$ tree-decomposition of $(B_x:x\in V(T_0))$ where $f$ is a bijection between $V(H)$ and $V(T_0)$.
-Renaming each node $a\in T_0$ as $f^{-1}(a)$ gives a tree-decomposition $(B_x:x\in V(T))$ with $V(T)=V(H)$.
-
-  Now, for each vertex $x\in V(H)$, rename the node $f(x)$ in $T_0$ as $x$. The resulting tree decomposition uses a tree $T_0$ with $V(H)\subseteq V(T_0)$.
-
-     (or introduce a new node)   
-  Begin with a width-$t$ tree-decomposition $(C_x:x\in V(T_0))$ of $H$ using a tree $T_0$ with the property that, for any edge $xy\in E(T_0)$, $|C_x\ominus C_y|\le 1$.
-That is, $C_x$ contains exactly one vertex not in $C_y$ or vice-versa.
-It is well-known that such a tree-decomposition exists.
-Indeed, any tree-decomposition of $H$ can be made to have this property adding $|C_x\ominus C_y|-1$ subdivision vertices along each edge $xy\in E(T_0)$ and assigning approriate bags to the newly introduced subdivision vertices.
-
-  Next, select an arbitrary node $x\in V(T_0)$ and let $\{v_1,\ldots,v_{t'}\}:=C_x$ denote the contents of $C_x$.
-Create a new tree $T$ from $T_0$ by adding a path of nodes $x_0,x_1,x_2,\ldots,x_{t'-1},x_{t'}=x$.
-Let $C_{x_0}:=\emptyset$ and, for each $i\in\{1,\ldots,t'\}$, $C_{x_i}:=\{v_1,\ldots,v_i\}$.
-Root $T_0$ at $x_0$ and note that $(C_x: x\in V(T))$ is a width-$t$ tree-decomposition of $H$ that satisfies Properties~\tref{rooted}, and \tref{diff}.
-For the remaining properties some renaming is required, which we now describe.
-
-  For each $v\in V(H)$, let $v'\in V(T)$ be the unique node such that $v\in C_{v'}$ and $v'$ has minimum $T$-depth.
-The uniqueness of $v'$ follows from the fact that $T[v]:=T[\{x\in V(T):v\in B_x\}]$ is connected, so $v'$ is the root of $T[v]$.
-Furthermore, for any pair of distinct vertices $v,w\in V(H)$, \tref{rooted} and \tref{diff} ensure that $v'\neq w'$: Otherwise, $C_{v'}\ominus C_{x'}\supseteq \{v',w'\}$ where $x'$ is the parent of $v'=w'$.
-Property~\tref{rooted} ensures that $x'$ exists but then Property~\tref{diff} contradicts $C_{v'}\ominus C_{x'}\supseteq \{v',w'\}$.
-
-  For each $x\in V(T)$, let $B_x=\{v' : v\in C_x\}$.
-Now define the graph $H'$ to have vertex set $V(H')=\{v':v\in V(H)\}$ and edge set $E(H')=\{v'w': vw\in E(H)\}$.
-Obviously, $H'$ isomorphic to $H$, $V(H')\subseteq V(T)$ and $(B_x: x\in V(T))$ is a width-$t$ tree-decomposition of $H'$.
-The decomposition $(B_x: x\in V(T))$ inherits Properties~\tref{rooted} and \tref{diff} from $(C_x:x\in V(T))$ and satisfies \tref{node-set} and \tref{subtree-root} by definition.
-
-  To see that $(B_x:x\in V(T))$ satisifies \tref{ancestor-edge}, observe that, if $vw\in E(H)$, then at least one of $v'$ or $w'$ is contained in $B_x$ for every node $x$ on the path from $v'$ to $w'$ in $T$.
-If neither $v'$ nor $w'$ is an ancestor of the other, then some node $x$ on this path has $T$-depth less than that of $v'$ and $w'$.
-If $v'\in B_x$ this contradicts the fact that $v'$ is the root of $T[v']$.
-If $w'\in B_x$ this contradicts the fact that $w'$ is the root of $T[w']$.
+Renaming each node $a\in V(T_0)$ as $f^{-1}(a)$ gives a tree-decomposition $(B_x:x\in V(T))$ with $V(T)=V(H)$.
+
+  By the definition of $f$, the tree-decomposition $(B_x:x\in V(T))$ satisfies \tref{subtree-root}.
+To see that $(B_x:x\in V(T))$ satisifies \tref{ancestor-edge}, observe that, if $xy\in E(H)$, then at least one of $x$ or $y$ is contained in $B_z$ for every node $z$ on the path from $x$ to $y$ in $T$.
+If neither $x$ nor $y$ is an ancestor of the other, then some node $z$ on this path has $T$-depth less than that of $x$ and $y$.
+If $x\in B_z$ this contradicts the fact that $x$ is the root of $T[x]$.
+If $y\in B_z$ this contradicts the fact that $y$ is the root of $T[y]$.
 \end{proof}
@@
 
-For the remainder of this section, $G$ is a graph and $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{Q}:=(Q_x: x\in V(H)))$ is a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{Q}$ is at most $t$.
-The following lemma shows how to use $(\mathcal{L},\mathcal{Q})$ to define a hierarchical decomposition of $G$.
-
-\begin{lem}\thmlabel{generalized-tripod}
-  There exists a tree $T$ rooted at a node $x_0$ where each node $x\in V(T)$ is associated with a subset $Y_x\subseteq V(G)$ that defines several related sets $V_x = \bigcup_{y\in V(T_x)} Y_y$, $F_x=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x=V_x\cup F_x$ that satisfy the following properties:
+For the remainder of this section, $G$ is a graph and $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{Y}:=(Y_x: x\in V(H)))$ is a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{Y}$ is at most $t$.
+In addition, $(B_x:x\in V(T))$ is a tree-decomposition of $H$ obtained by applying \lemref{nice-decomposition} to $H$.
+The following lemma shows how to interpret these objects as a hierarchical decomposition of $G$.
+
+\begin{clm}\clmlabel{generalized-tripod}
+  For each $x\in V(T)$, let $V_x = \bigcup_{y\in V(T_x)} Y_y$, $F_x=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x=V_x\cup F_x$.  Then, 
   \begin{compactenum}[(Y1)]
     \item\ylabel{thickness} $\mathcal{Y}:=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width (with respect to $\mathcal{L}$) at most $\ell$.
-    \item\ylabel{separator} There is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$. 
-    \item\ylabel{ancestor-edge} $x$ has $t'\le t$ strict $T$-ancestors $a_1,\ldots,a_{t'}$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
+    \item\ylabel{separator} For every $x\in V(T)$, there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$. 
+    \item\ylabel{ancestor-edge} For every $x\in V(T)$, $x$ has $t'\le t$ strict $T$-ancestors $a_1,\ldots,a_{t'}$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
   \end{compactenum}
-\end{lem}
-
-Before proving \thmref{generalized-tripod} we point out more properties that are immediately implied by it:
+\end{clm}
+
+Before proving \clmref{generalized-tripod} we point out more properties that are immediately implied by it:
 
@@
 
-\begin{proof}[Proof of \thmref{generalized-tripod}]
-  Apply \lemref{nice-decomposition} to $H$ to obtain a tree-decomposition $(B_x: x\in V(T))$ of $H$ where $T$ is rooted at some node $x_0$ and that satisifies Properties~\tref{first}--\tref{last}.
-(We skip the introduction of the isomorphic graph $H'$ and assume from the outset that $V(H)\subseteq V(T))$.)
-
-  For each $v\in V(G)$, let $q(v)$ denote the unique node $q\in V(H)$ such that $v\in Q_q$.  For each $x\in V(T)$, define
-  \[
-      Y_x = \begin{cases} Q_x & \text{if $x\in V(H)$} \\
-                          \emptyset & \text{otherwise} .
-            \end{cases}
-  \]
-   % 
-   % 
-   % and let $b(v)$ denote the root of the subtree $T[v]:=T[\{x\in V(T):q(v)\in B_x\}]$. For each $x\in V(T)$, define $Y_x=\{v\in V(G):c(v)=x\}$.  
-  Clearly, $\mathcal{Y}:=(Y_x:x\in V(T))$ is a partition of $V(G)$ and, for any $i\in\{1,\ldots,h\}$ $|Y_x\cap V_i|\le \ell$.
-  so $\mathcal{Y}$ has layered width at most $t\ell$ with respect to $\mathcal{L}$.  This establishes \yref{thickness}.
-
-  Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
+\begin{proof}[Proof of \clmref{generalized-tripod}]
+  % Apply \lemref{nice-decomposition} to $H$ to obtain a tree-decomposition $(B_x: x\in V(T))$ of $H$ satisfying the conditions of \lemref{nice-decomposition}.
+  % 
+  % For each $v\in V(G)$, let $y(v)$ denote the unique node $y\in V(H)$ such that $v\in Y_y$.
+Observe that, by \tref{subtree-root}, $y(v)$ is the root of $T[y(v)]=T[\{x\in V(T):y(v)\in B_x\}]$.
+By definition, $\mathcal{Y}:=(Y_x:x\in V(T))$ is a partition of $V(G)$ and, for any $i\in\{1,\ldots,h\}$ $|Y_x\cap V_i|\le \ell$, so $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
+This establishes \yref{thickness}.
+
+  Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$ and $\mathcal{Y}$ is a partition.
+Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.
+In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
 
@@
 Each path in $P\in\mathcal{P}$ is called a \emph{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \emph{$vw$-shortcut}.
-A $vw$-shortcut is also referred to as \emph{$v$-shortcut} (and a \emph{$w$-shortcut}).
 Given a graph $G$ and a $(k,d)$-shortcut system $\mathcal{P}$ for $G$, we use $G^{\mathcal{P}}$ to denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\mathcal{P}$.
@@
 
-As in the rest of this section, $(\mathcal{L},\mathcal{Q})$ denotes a layered $H$-partition of $G$; $T$ is a tree used in a tree-decomposition $(B_x:x\in V(T))$ of $H:=G/\mathcal{Q}$ satisfying the conditions of \lemref{nice-decomposition}; and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ are obtained obtained by applying \thmref{generalized-tripod} to $G$ $(\mathcal{L},\mathcal{Q})$, and $H$.
+As in the rest of this section, $(\mathcal{L},\mathcal{Y})$ denotes a layered $H$-partition of $G$; $T$ is a tree used in a tree-decomposition $(B_x:x\in V(T))$ of $H:=G/\mathcal{Y}$ satisfying the conditions of \lemref{nice-decomposition}; and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ are defined as in \clmref{generalized-tripod}.
 
@@
 Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
-We let $I:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{k,d}$.
+We let $I:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
 (Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $I$.)
@@
 Therefore any layered $I$-partition of $G^{k,d}$ of layered width $\ell'$ in which the treewidth of $I$ is at most $t'-1$ must satisfy
-\[
-    \ell'\times t' \ge 2^{k/2}-1/2 \enspace .
-\]   
+% \[
+%     \ell'\times t' \ge 2^{k/2}-1/2 \enspace .
+% \]   
 \newpage

### 2019-07-05 k-planar.tex
First commit of David's version

@@
 \usepackage{enumerate}
+\usepackage{amsfonts}
+\usepackage{amsthm,mathtools}
 \usepackage{pat}
@@
 \usepackage[longnamesfirst,numbers,sort&compress]{natbib}
+\usepackage[noabbrev,capitalise]{cleveref}
+\allowdisplaybreaks
+%\sloppy
+%%%%%%%%%%
+\makeatletter
+\def\NAT@spacechar{~}
+\makeatother
+%%%%%%%%%%
+\crefname{lem}{Lemma}{Lemmas}
+\crefname{thm}{Theorem}{Theorems}
+\crefname{cor}{Corollary}{Corollaries}
+\crefname{conj}{Conjecture}{Conjectures}
+\crefname{obs}{Observation}{Observations}
+\crefname{prop}{Proposition}{Propositions}
+\crefname{clm}{Claim}{Claims}
+\crefname{openproblem}{Open Problem}{Open Problems}
+\crefformat{equation}{(#2#1#3)}
+\Crefformat{equation}{Equation #2(#1)#3}
 
@@
 \DeclareMathOperator{\tw}{tw}
-
-\renewcommand{\proplabel}[1]{\label{prop:#1}}
-\renewcommand{\propref}[1]{(PR\ref{prop:#1})}
+\DeclareMathOperator{\pw}{pw}
+\DeclareMathOperator{\qn}{qn}
+\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
+\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
+
+\newcommand{\PRlabel}[1]{\label{PR:#1}}
+\newcommand{\PRref}[1]{(PR\ref{PR:#1})}
 \newcommand{\jlabel}[1]{\label{j:#1}}
 \newcommand{\jref}[1]{(J\ref{j:#1})}
-\newcommand{\tlabel}[1]{\label{t:#1}}
-\newcommand{\tref}[1]{(T\ref{t:#1})}
-\newcommand{\ylabel}[1]{\label{y:#1}}
-\newcommand{\yref}[1]{(Y\ref{y:#1})}
+
+\renewcommand{\ge}{\geqslant}
+\renewcommand{\le}{\leqslant}
+\renewcommand{\geq}{\geqslant}
+\renewcommand{\leq}{\leqslant}
 
@@
 
-\title{\MakeUppercase{The Structure of $k$-Planar Graphs}%
-    \thanks{This work was partly funded by NSERC and MRI.}}
-
-\author{Vida Dujmović%
+\title{\MakeUppercase{The Structure of $k$-Planar Graphs}}
+
+\author{Vida Dujmovi\'c%
         \thanks{School of Computer Science and Electrical Engineering,
-                University of Ottawa},\,\,
+                University of Ottawa, Ottawa, Canada (\texttt{vida.dujmovic@uottawa.ca}).
+                Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\,
         Pat Morin%
-        \thanks{School of Computer Science, Carleton University},\,\, and
-        David R. Wood%
-        \thanks{School of Mathematics, Monash University}}
-
-%\renewcommand{\ge}{\geqslant}
-%\renewcommand{\le}{\leqslant}
+        \thanks{School of Computer Science, Carleton University, Ottawa, Canada (\texttt{morin@scs.carleton.ca}).
+Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\, and
+        David R. Wood\thanks{School of Mathematics, Monash University, Melbourne, Australia (\texttt{david.wood@monash.edu}). Research supported by the Australian Research Council.}
+}
 
@@
 \begin{abstract}
-  Layered partitions of graphs (with small layered width, in which the quotient has small treewidth) are a recently-introduced tool that have been used to solve longstanding problems on queue layouts, non-repetitive colouring, and 3-d graph drawing.
+Dujmovi\'c~et~al.~(FOCS 2019) recently proved that every planar graph is a subgraph of the strong product of a graph of bounded treewidth and a path.
+This tool has been used to solve longstanding problems on queue layouts, non-repetitive colouring, and $p$-centred colouring.
+%, and graph encoding.
+We generalise this result for $k$-planar graphs, where a graph is \emph{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings.
+In particular, we prove that every $k$-planar graph is a subgraph of the strong product of a graph of treewidth $O(k^5)$ and a path.
+This is the first result of this type for a non-minor-closed class of graphs.
+It implies, amongst other results, that $k$-planar graphs have non-repetitive chromatic number upper-bounded by a function of $k$.
+All these results generalise for drawings of graphs on arbitrary surfaces.
+\end{abstract}
+\end{titlepage}
+
+% Layered partitions of graphs (with small layered width, in which the quotient has small treewidth) are a recently-introduced tool that have been used to solve longstanding problems on queue layouts, non-repetitive colouring, and 3-d graph drawing.
 Such partitions are known to exist for planar graphs and, more generally, apex-minor-free graphs.
@@
 These results extend to $(g,k)$-planar graphs, the natural generalization of $k$-planar graphs to to genus-$g$ surfaces (rather than the genus-$0$ plane).
-\end{abstract}
-\end{titlepage}
 
@@
 
-In 1965, \citet{ringel:sechsfarbenproblem} introduced the notion of $1$-planar graphs, a generalization of planar graphs that allows edges to cross provided that no edge is involved in more than one crossing.
+A graph is \emph{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings.
+Such graphs provide a natural generalisation of planar graphs, and are important in graph drawing research; see the recent bibliography on 1-planar graphs and the 140 references therein \citep{kobourov.liotta.ea:annotated}.
+The present paper studies the structure of $k$-planar graphs (and other more general classes).
+Various structural results about planar graphs generalise for $k$-planar graphs.
+For example, \citet{FP08} generalised the Lipton-Tarjan separator theorem to show that every $k$-planar graph with $n$ vertices has a balanced separator of order $O(\sqrt{kn})$.
+
+The main results of this paper generalise the following recent theorem of \citet{dujmovic.joret.ea:planar} to the setting of $k$-planar graphs (and other more general classes).
+
+\begin{thm}[\citep{dujmovic.joret.ea:planar}] 
+\label{PlanarBasic}
+Every planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most $8$ and for some path $P$. 
+\end{thm}
+
+Here $\boxtimes$ is the strong product, and treewidth is an invariant that measures how `tree-like' a given graph is; see \cref{Tools} for formal definitions and see \cref{ProductExample} for an example.
+So, loosely speaking, \cref{PlanarBasic} says that every planar graph is contained in the product of a tree-like graph and a path.
+This enables combinatorial results for graphs of bounded treewidth to be generalised for planar graphs (with different constants).
+\cref{kPlanarBasic} has been the key tool in solving several well-known open problems.
+In particular, \citet{dujmovic.joret.ea:planar} proved that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92}), and \citet{dujmovic.esperet.ea:planar}  proved that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
+
+\begin{figure}[!h]
+\centering
+\includegraphics{ProductExample}
+\caption{Example of a strong product.}
+\label{ProductExample}
+\end{figure}
+
+We generalise \cref{PlanarBasic} as follows.
+
+\begin{thm}
+\label{kPlanarBasic}
+Every $k$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most $(4+o(1))k^5$ and for some path $P$. 
+\end{thm}
+
+This theorem and several related results are proved in \cref{Structure}.
+In particular, in \cref{sec-1-planar} we present a refinement of \cref{kPlanarBasic} in the case $k=1$.
+In \cref{sec-g-k-planar} we generalise \cref{kPlanarBasic} for graphs that have a drawing on an arbitrary surface with a bounded number of crossings per edge.
+
+\cref{kPlanarBasic} (and its generalisations) have applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, and $p$-centred colouring  \citep{micek:personal}.
+%, and graph encoding \citep{BGP}. 
+These applications in the context of $k$-planar graphs are explored in \cref{Applications}.
+For example, we prove that $k$-planar graphs have bounded non-repetitive chromatic number (for fixed $k$).
+Prior to the recent work of
+\citet{dujmovic.esperet.ea:planar}, it was even open whether planar graphs have bounded non-repetitive chromatic number.
+
+\cref{Characterisation} presents a rough characterisation of $k$-planar graphs in terms of an underlying planar graph.
+This is useful for showing that various families of graphs are $k$-planar.
+For example, we conclude that bounded powers of bounded-degree planar graphs are $k$-planar, for a suitable constant $k$.
+Thus
+\cref{kPlanarBasic} and the above applications are applicable to this class. These results also generalise for arbitrary surfaces.
+
+\cref{Examples} presents several examples of $k$-planar graphs, including $k$-map graphs, $k$-string graphs, and $k$-nearest neighbour graphs.
+All of the above results are applicable to these examples.
+
+%\citet{dujmovic.joret.ea:planar} generalised their result for planar graphs for graphs embeddable on any surface.
+In particular, they proved that every graph of Euler genus $g$ is a subgraph of $H\boxtimes P$ for some graph $H$ of treewidth at most $???k$ and for some path $P$.
+We provide a further generalisation allowing crossings.
+Say a graph is $(g,k)$-planar if it has a drawing in a surface of Euler genus at most $g$ in which each edge is involved in at most $k$ crossings.
+%
+%\begin{thm}
+%For all integers $g,k\geq 1$ there is an integer $c_{k,g}$, such that every $(g,k)$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most $c_{k,g}$  and for some path $P$.
+%\end{thm}
+
+%In 1965, \citet{ringel:sechsfarbenproblem} introduced the notion of $1$-planar graphs, a generalization of planar graphs that allows edges to cross provided that no edge is involved in more than one crossing.
 This simple generalization has since become a rich source of problems and results.
 Indeed, the annotated bibliography on 1-planar graphs by \cite{kobourov.liotta.ea:annotated} contains more than 140 entries.
-
-The natural generalization of $1$-planar graphs, $k$-planar graphs (that allow edge to be involved in up to $k$ crossings), has been the topic of important recent work on graph drawing.
+%
+%The natural generalization of $1$-planar graphs, $k$-planar graphs (that allow edge to be involved in up to $k$ crossings), has been the topic of important recent work on graph drawing.
 Though planar graphs are, by now, quite well-understood, the situation is much less clear for $k$-planar graphs; even the maximum edge-density of $k$-planar graphs is not completely settled.
@@
 % In the current paper we investigate the structure of $k$-planar graphs.
-Since planarity is closed under taking minors, planar graphs can be characterized by a finite set of forbidden minors.
+Since planarity is closed under taking minors, planar graphs can be characterised by a finite set of forbidden minors.
 Indeed, Wagner's Theorem states that this set consists of the complete 5-vertex graph $K_5$ and the complete bipartite graph $K_{3,3}$.
@@
 
-The purpose of the current paper is to show that new structural results for planar graphs can be extended to $k$-planar graphs.
+%The purpose of the current paper is to show that new structural results for planar graphs can be extended to $k$-planar graphs.
 These results can be used to give a rough characterization of $k$-planar graphs of powers of bounded-degree planar graphs
 % Rather, we present a rough structural characterization of $k$-planar graphs that, despite its roughness, provides several new results for $k$-planar and related graph families.  
-All of these results are based on layered $H$-partitions, which we now define.
-{\color{red}TODO: Help.}
-
-A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle V_0,V_1,\ldots\rangle$ such that $\{V_0,V_1,\ldots\}$ is a partition of $V(G)$ and there is no edge $vw\in E(G)$ such that $v\in V_i$, $w\in V_j$ and $|j-i|>1$.
-For any partition $\mathcal{P}=\{S_1,\ldots,S_p\}$ of $V(G)$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
-To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an \emph{$H$-partition} and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.\footnote{An alternative definition of a layering of $G$ is an $H$-partition of $G$ where $H$ is a path.}  A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
-The \emph{layered width} of $(\mathcal{L},\mathcal{P})$ is $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
-
-Layered $H$-partitions of small layered width in which $H$ has some additional property are a useful tool for a number of graph problems.
-One such useful property is small \emph{treewidth}, which we now define.
-A \emph{tree-decomposition} $\mathcal{K}$ of a graph $G$ consists of a tree $K$ and a collection $\mathcal{K}=(B_x:x\in V(K))$ of subsets of $V(G)$ indexed by nodes of $K$ such that (i)~for every $vw\in E(G)$, there exists some $x\in V(K)$ with $v,w\in B_x$; and (ii)~for every $v\in V(G)$, the induced subgraph $K[x] := K[\{y: x\in B_y\}]$ is connected.
-The \emph{width} of the tree-decomposition $\mathcal{K}$ is $\max\{|B_x|:x\in V(K)\}-1$.
-The \emph{treewidth} $\tw(G)$ of a graph $G$ is the minimum width of a tree-decomposition of $G$.
-
- \citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number, track number, and layered treewidth.
-\citet{dujmovic.esperet.ea:planar} add non-repetitive chromatic number to this list.
-\citet{dujmovic.joret.ea:planar} show that two common families of graphs have such $H$-partitions: planar graphs, graphs of bounded genus, and (more generally) apex-minor-free graphs.\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
-A class $\mathcal{G}$ of graphs is \emph{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$.
-A minor-closed class is \emph{proper} if it is not the class of all graphs.
-For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class.
-
- A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar.
-A 1-apex graph is \emph{apex}.
-A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.}
-
-% In this paper we show that another family of graphs admits layered $H$-partitions of bounded layered width in which $H$ has bounded treewidth:
-In the current paper we show that $k$-planar graphs also admit layered $H$-partitions in which $H$ has bounded treewidth.
-First, we formally define $k$-planar graphs.
+%All of these results are based on layered $H$-partitions, which we now define.{\color{red}TODO: Help.}
+
+%Thus, like genus-$g$ graphs,$k$-planar graphs are a generalization of planar graphs (which are genus-$0$ graphs).
+Unlike genus-$g$ graphs, however, the family of $k$-planar graphs is not minor-closed.
+A graph $G'$ obtained from a $k$-planar graph $G$ by edge deletions and edge contractions may or may not be $k$-planar.
+Indeed, \citet{dujmovic.eppstein.ea:structure} construct 1-planar graphs that contain arbitrarily large complete graph minors.
+
+%\note{PM}{TODO: Write here about applications. non-repetitive colouring, $p$-centered colouring, $(g,k)$-planar graph, $(g,d)$-map graphs, $k$-NN graphs, David's rough characterization.}
+
+%The remainder of the paper is organised as follows: In \cref{k-planar} we prove \cref{sec-k-planar}.
+In \cref{sec-1-planar} we prove \cref{1-planar}.
+In \cref{g-k-planar} we prove \cref{g-k-planar}.
+In \cref{consequences} we point out some consequences of these results for some of the other graph parameters mentioned above.
+\comment{Re-DO}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Tools}
+\label{Tools}
+
+This section defines concepts from from the literature that will be important for our work.
+
+In this paper, all graphs are finite and undirected.
+Unless specifically mentioned otherwise, all graphs are also simple.
+For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$) we use $G[S]$ to denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.
+We use $G-S$ as a shorthand for $G[V(G)\setminus S]$.
+We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.
+
+We now formally define $k$-planar graphs.
 An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
@@
 A (not necessarily embedded) graph $G'$ is \emph{$k$-planar} if there exists $k$-plane graph $G$ isomorphic to $G'$.
-
 Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
-Thus, like genus-$g$ graphs,\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
+
+We now define two concepts used in the statements of \cref{PlanarBasic,kPlanarBasic}: strong products and treewidth.
+The \emph{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if:
+\begin{compactitem}
+\item  $v=w$ and $xy\in E(B)$, or 
+\item  $x=y$ and $vw\in E(A)$, or  
+\item  $vw\in E(A)$ and $xy\in E(B)$. 
+\end{compactitem}
+A \emph{tree-decomposition} $\mathcal{K}$ of a graph $G$ consists of a tree $K$ and a collection $\mathcal{K}=(B_x:x\in V(K))$ of subsets of $V(G)$ indexed by the nodes of $K$ such that:
+\begin{compactenum}[(i)]
+\item for every $vw\in E(G)$, there exists some $x\in V(K)$ with $v,w\in B_x$; and 
+\item for every $v\in V(G)$, the induced subgraph $K[x] := K[\{y: x\in B_y\}]$ is connected.  
+\end{compactenum}
+The \emph{width} of the tree-decomposition $\mathcal{K}$ is $\max\{|B_x|:x\in V(K)\}-1$.
+The \emph{treewidth} $\tw(G)$ of a graph $G$ is the minimum width of a tree-decomposition of $G$.
+Treewidth is the standard measure of how similar a graph is to a tree.
+Indeed, a connected graph has treewidth 1 if and only if it is a tree.
+Treewidth is of fundamental importance in structural and algorithmic graph theory; see \citep{Reed03,HW17,Bodlaender-TCS98} for surveys.
+
+While strong products enable an easy statement of \cref{PlanarBasic,kPlanarBasic}, to prove such theorems it is helpful to work with layerings and partitions, which we now introduce.
+A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle V_0,V_1,\ldots\rangle$ such that $\{V_0,V_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in V_i$ and $w\in V_j$ then $|j-i|\leq 1$.
+For any partition $\mathcal{P}=\{S_1,\ldots,S_p\}$ of $V(G)$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
+To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an \emph{$H$-partition} and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.\footnote{An alternative definition of a layering of $G$ is an $H$-partition of $G$ where $H$ is a path.}  A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
+The \emph{layered width} of $(\mathcal{L},\mathcal{P})$ is $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
+These definitions relate to strong products as follows.
+
+\begin{lem}[\citep{dujmovic.joret.ea:planar}] 
+\label{PartitionProduct}
+For every graph $H$, a graph $G$ has an $H$-partition of layered width at most $\ell$ if and only if $G$ is a subgraph of 
+$H \boxtimes P \boxtimes K_\ell$ for some path $P$.
+\end{lem}
+
+\citet{dujmovic.joret.ea:planar} also showed it suffices to consider partitions of layered width 1.
+
+\begin{lem}[\citep{dujmovic.joret.ea:planar}] 
+\label{MakeWidth1}
+If a graph $G$ has an $H$-partition of layered width $\ell$ with respect to layering $\langle V_0,V_1,\dots\rangle$, for some graph $H$ of treewidth at most $k$, then $G$ has an $H'$-partition of layered width 1 with respect to the same layering, for some graph $H'$ of treewidth at most $(k+1)\ell-1$.
+That is, if $G\subseteq H\boxtimes P\boxtimes K_\ell$ for some graph $H$ of treewidth at most $k$  and for some path $P$, then $G\subseteq H' \boxtimes P$ for some graph $H'$ of treewidth at most $(k+1)\ell-1$.
+\end{lem}
+
+\citet{dujmovic.joret.ea:planar} proved the following result, which with \cref{PartitionProduct,MakeWidth1}, implies \cref{PlanarBasic}.
+
+\begin{thm}[\citep{dujmovic.joret.ea:planar}]
+\label{PlanarPartition}
+Every planar graph has:
+\begin{compactitem}
+\item an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$, and
+\item an $H$-partition of layered width $3$ for some planar graph $H$ of treewidth at most $3$.
+\end{compactitem}
+\end{thm}
+
+%Layered $H$-partitions of small layered width in which $H$ has some additional property are a useful tool for a number of graph problems.
+One such useful property is small \emph{treewidth}, which we now define.
+
+% \citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number, track number, and layered treewidth.
+\citet{dujmovic.esperet.ea:planar} add non-repetitive chromatic number to this list.
+\citet{dujmovic.joret.ea:planar} show that two common families of graphs have such $H$-partitions: planar graphs, graphs of bounded genus, and (more generally) apex-minor-free graphs.
+
+% In this paper we show that another family of graphs admits layered $H$-partitions of bounded layered width in which $H$ has bounded treewidth:
+
+\citet{dujmovic.joret.ea:planar} proved the following generalisation of \cref{PlanarBasic,PlanarPartition} for graphs embeddable on other surfaces.\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
 The \textit{Euler genus} of the non-orientable surface with $c$ cross-caps is $c$.
 The \textit{Euler genus} of a graph $G$ is the minimum integer $g$ such that $G$ embeds in a surface of Euler genus $g$.
-Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.} $k$-planar graphs are a generalization of planar graphs (which are genus-$0$ graphs).
-Unlike genus-$g$ graphs, however, the family of $k$-planar graphs is not minor-closed.
+Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.}
+
+\begin{thm}[\citep{dujmovic.joret.ea:planar}]
+\label{SurfacePartition}
+Every graph of Euler genus $g$ is a subgraph of:
+\begin{compactenum}[(a)]
+\item  $H \boxtimes P \boxtimes K_{\max\{2g,1\}}$ for some graph $H$ of treewidth at most $9$ and for some path $P$;
+\item  $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $4$ and for some path $P$;
+\item $(K_{2g} + H )  \boxtimes P$ for some graph $H$ of treewidth at most $8$  and some path $P$. 
+\end{compactenum}
+Equivalently, every graph of Euler genus $g$ has:
+\begin{compactenum}[(a)]
+\item an $H$-partition with layered width at most $\max\{2g,1\}$ such that $H$ has treewidth at most $9$;
+\item an $H$-partition with layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most $4$; 
+\item an $H$-partition with layered width at most $1$ such that $H$ has treewidth at most $2g+7$. 
+\end{compactenum}
+%\textcolor{red}{eliminate some of these results?}
+\end{thm}
+
+\citet{dujmovic.joret.ea:planar} generalised \cref{SurfacePartition} further to show that for every apex\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
+A class $\mathcal{G}$ of graphs is \emph{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$.
+A minor-closed class is \emph{proper} if it is not the class of all graphs.
+For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class.
+A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar.
+A 1-apex graph is \emph{apex}.
+A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.}
+ graph $X$, there is a constant $c$ such that every $X$-minor-free graph is a subgraph of $H\boxtimes P$ for some graph $H$ of treewidth at most $c$.
+Apex-minor-free graphs are the largest minor-closed class for which such a theorem holds \citep{dujmovic.joret.ea:planar}.
+However, the family of $k$-planar graphs is not minor-closed.
 A graph $G'$ obtained from a $k$-planar graph $G$ by edge deletions and edge contractions may or may not be $k$-planar.
 Indeed, \citet{dujmovic.eppstein.ea:structure} construct 1-planar graphs that contain arbitrarily large complete graph minors.
-
-We prove the following result:
-\begin{thm}\thmlabel{k-planar}
+Our results for $k$-planar graphs are the first of this type for a non-minor-closed class.
+
+The following is the main result of this paper; it shows that $k$-planar graphs admit $H$-partitions with bounded layered width, in which $H$ has bounded treewidth.
+With \cref{PartitionProduct,MakeWidth1}, this proves \cref{kPlanarBasic}.
+% with $c_k\leq(4+o(1))k^5$.
+
+\begin{thm}
+\label{k-planar}
 Every $k$-planar graph has a layered $H$-partition of layered width at most $24k^2 + 60k + 36$ in which $H$ has treewidth at most $\treewidth$.
@@
 
-\begin{thm}\thmlabel{1-planar}
+\begin{thm}
+\label{1-planar}
 Every 1-planar graph has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
@@
 
-These results imply that $k$-planar graphs have all the graph parameters mentioned above upper-bounded by some constant $c_k$.
-
-Prior to this work, the strongest structural characterization of $k$-planar graphs was in terms of layered treewidth, which we now define.
+%These results imply that $k$-planar graphs have all the graph parameters mentioned above upper-bounded by some constant $c_k$.
+
+Prior to this work, the strongest structural description of $k$-planar graphs was in terms of layered treewidth, which we now define.
 A \emph{layered tree-decomposition} $(\mathcal{L},\mathcal{K})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{K}$ of $G$.
@@
 The \emph{layered treewidth} of $G$ is the minimum layered width of any layered tree-decomposition of $G$.
-
-\citet{dujmovic.eppstein.ea:structure} show that $k$-planar graphs have layered treewidth $O(k+1)$.
-This already implies that several of the graph parameters mentioned above, including queue-number and non-repetitive chromatic number are upper-bounded by $O(k\log n)$ for $n$-vertex $k$-planar graphs \cite{dujmovic.morin.ea:layered}.
-\thmref{k-planar} is a significant strengthening of this characterization.
-It implies that $k$-planar graphs have layered-treewidth bounded by a function of $k$ and gives upper bounds on queue-number and non-repetitive chromatic number of $n$-vertex $k$-planar graphs that are independent of $n$.
-
-There is a natural generalization of $k$-planar to genus-$g$ surfaces, that we now describe.
-A graph $G$ embedded on a surface $S$ is $(S,k)$-plane if no edge of $G$ is involved in more than $k$ crossings.
-A (not necessarily embedded) graph $G$ is $(g,k)$-planar if it is isomorphic to some $(S,k)$-plane graph $G'$ where $S$ is a surface with Euler genus $g$.
-
-\begin{thm}\thmlabel{g-k-planar}
-  Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most $\max\{8gk^2 + 12gk + 4g, 24k^2 + 36k + 12\}$ in which $H$ has treewidth at most $\treewidth$.
-
-  Every $(g,1)$-planar graph has a layered $H$-partition of layered width at most $\max\{48g, 30\}$ in which $H$ has treewidth at most $4$.  
-\end{thm}
-
-\note{PM}{TODO: Write here about applications. non-repetitive colouring, $p$-centered colouring, $(g,k)$-planar graph, $(g,d)$-map graphs, $k$-NN graphs, David's rough characterization.}
-
-The remainder of the paper is organized as follows: In \secref{k-planar} we prove \thmref{k-planar}.  In \secref{1-planar} we prove \thmref{1-planar}.  
-In \secref{g-k-planar} we prove \thmref{g-k-planar}. In \secref{consequences} we point out some consequences of these results for some of the other graph parameters mentioned above.
-
-\section{$k$-Planar Graphs}
-\seclabel{k-planar}
-
-The purpose of this section is to prove \thmref{k-planar}.
-In this paper, all graphs are finite and undirected.
-Unless specifically mentioned otherwise, all graphs are also simple.
-For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$) we use $G[S]$ to denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.
-We use $G-S$ as a shorthand for $G[V(G)\setminus S]$ and for a graph $G'$, we use $G-G'$ as a shorthand for $G-V(G')$.
-We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.
-
-We begin by introducing a structural tool for planar graphs, called a tripod decomposition, that is implicit in previous work on $H$-partitions of planar graphs \cite{dujmovic.joret.ea:planar}.
-We then show how the tripod decomposition of a planarized version of a $k$-planar graph $G$ can be used to find a layered $H$-partition of $G$ of small layered width in which $H$ has small treewidth.
-
-\subsection{PS-Trees}
+\citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph.
+\citet{dujmovic.eppstein.ea:structure} show that $k$-planar graphs have layered treewidth at most $6(k+1)$.
+
+If a graph has a partition with bounded layered width, where the quotient has bounded treewidth, then the graph has bounded layered treewidth.
+In particular, suppose that an $n$-vertex graph $G$ has an $H$-partition with layered width at most $\ell$ such that $H$ has treewidth at most $k$.
+\citet{dujmovic.joret.ea:planar} proved that $G$ has layered treewidth at most $(k+1)\ell$.
+So any property that holds for graphs of bounded layered treewidth also holds for $G$.
+For example, Norin proved that every $n$-vertex graph with layered treewidth at most $t$ has treewidth less than $2\sqrt{tn}$ (see \citep{dujmovic.morin.ea:layered}).
+Thus, $G$ has treewidth at most $2\sqrt{(k+1)\ell n}$.
+This in turn leads to a $O(\sqrt{(k+1)\ell n})$ balanced separator theorem for such graphs \citep{RS-II}.
+So, in the case of $k$-planar graphs, \cref{k-planar} implies an analogue of the separator theorem of \citet{FP08} mentioned in the introduction, albeit with worse dependence on $k$.
+What sets layered partitions apart from layered treewidth is that they lead to constant upper bounds on the queue-number  and non-repetitive chromatic number, whereas for both these parameters, the best known upper bound obtainable via layered treewidth is $O(\log n)$; see \cref{Applications}.
+
+%Indeed the former structure leads to $O(1)$ bounds on the queue-number, instead of $O(\log n)$ bounds obtained via layered treewidth.
+That said, it is open whether graphs of bounded layered treewidth have bounded queue-number.
+It is even possible that graphs of bounded layered treewidth have partitions of bounded layered width, whose quotient has bounded treewidth.
+%
+%This already implies that several of the graph parameters mentioned above, including queue-number and non-repetitive chromatic number are upper-bounded by $O(k\log n)$ for $n$-vertex $k$-planar graphs \cite{dujmovic.morin.ea:layered}.
+\cref{k-planar} is a significant strengthening of this result.
+It implies that $k$-planar graphs have layered treewidth bounded by a function of $k$, and gives upper bounds on the queue-number and non-repetitive chromatic number of $n$-vertex $k$-planar graphs that are independent of $n$.
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Structure}
+\label{Structure}
+
+The purpose of this section is to prove \cref{k-planar} and related structural results. 
+We begin by introducing a tool for studying planar graphs, called a tripod decomposition, that is implicit in previous work on $H$-partitions of planar graphs \cite{dujmovic.joret.ea:planar}.
+We then show how the tripod decomposition of a planarised version of a $k$-planar graph $G$ can be used to find a layered $H$-partition of $G$ of small layered width in which $H$ has small treewidth.
+
+\subsection{Tripod Decompositions}
+\label{TripodDecompositions}
 
@@
 Let $r$ be a degree-3 node on the outer face of a triangulation $\Delta$.
-A \emph{tripod decomposition}  $K=K(\Delta, T, r)$ is a rooted tree in which each non-root node $x\in V(K)$ is associated with several objects (refer to \figref{K-node}):
+A \emph{tripod decomposition}  $K=K(\Delta, T, r)$ is a rooted tree in which each non-root node $x\in V(K)$ is associated with several objects (refer to \cref{K-node}):
 
@@
   \caption{The elements associated with a node $x\in V(K)$.}
-  \figlabel{K-node}
+  \label{K-node}
 \end{figure}
@@
 \begin{enumerate}[(PR1)]
-  \item \proplabel{children} For each non-root node $x\in V(K)$ such that $M_x$ has $f\le 3$ wedges $F_{x,1},\ldots,F_{x,f}$, the node $x$ has exactly $f$ children $y_1,\ldots,y_f$ in $K$ and $F_{y_i} = F_{x,i}$ for each $i\in\{1,\ldots,f\}$.
-
-  \item \proplabel{ancestor-boundary} For each non-root node $x\in V(K)$, there exists a partition of $V(F_x)$ into $g\le 3$ non-empty sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$.
-
-  \item \proplabel{tripod-feet} For each $i\in\{1,\ldots,g\}$, at least one vertex in $P_i$ is a foot of the closed tripod $\overline{Y}_x$.
-
-  \item \proplabel{partition} $\{Y_x : x\in V(K)\}$ is a partition of $V(\Delta)\setminus\{r\}$.
+  \item \PRlabel{children} For each non-root node $x\in V(K)$ such that $M_x$ has $f\le 3$ wedges $F_{x,1},\ldots,F_{x,f}$, the node $x$ has exactly $f$ children $y_1,\ldots,y_f$ in $K$ and $F_{y_i} = F_{x,i}$ for each $i\in\{1,\ldots,f\}$.
+
+  \item \PRlabel{ancestor-boundary} For each non-root node $x\in V(K)$, there exists a partition of $V(F_x)$ into $g\le 3$ non-empty sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$.
+
+  \item \PRlabel{tripod-feet} For each $i\in\{1,\ldots,g\}$, at least one vertex in $P_i$ is a foot of the closed tripod $\overline{Y}_x$.
+
+  \item \PRlabel{partition} $\{Y_x : x\in V(K)\}$ is a partition of $V(\Delta)\setminus\{r\}$.
 \end{enumerate}
@@
 % 
-% \begin{clm}\clmlabel{strong-ancestor-boundary}
+% \begin{clm}\label{strong-ancestor-boundary}
 %   Let $K$ be a PS-tree and
@@
 % 
-%   We claim that \propref{tripod-feet} ensures that $P_{2,3}'\subset P_i\cup P_j$ for some $i,j\in \{1,\ldots,g\}$.
+%   We claim that \PRref{tripod-feet} ensures that $P_{2,3}'\subset P_i\cup P_j$ for some $i,j\in \{1,\ldots,g\}$.
 If $g=2$, this claim is immediate, so assume $g=3$.
 Then, $F_y\cap \overline{Y}_p$ is the path formed by two legs $Q_2$ and $Q_3$ of $\overline{Y}_p$ and one edge $v_2v_3$ of $\tau_p$.
-%   By \propref{tripod-feet} and the assumption that $q=3$, $Q_2$ has a foot $v_2'\in P_i$ and $Q_3$ has a foot $v_3'\in P_j$, for some $i\neq j$.
+%   By \PRref{tripod-feet} and the assumption that $q=3$, $Q_2$ has a foot $v_2'\in P_i$ and $Q_3$ has a foot $v_3'\in P_j$, for some $i\neq j$.
 The third foot $v_1'$ of $\overline{Y}_p$ is therefore in $P_\ell$ for the unique $\ell \in\{1,2,3\}\setminus \{i,j\}$.
@@
 We use the vertex partition $\mathcal{P}=\{ S_x:=V(Y_x) : x\in V(K)\}$.
-\propref{partition} ensures that $\mathcal{P}$ is, indeed a partition of $V(\Delta)-\{r\}$.
+\PRref{partition} ensures that $\mathcal{P}$ is, indeed a partition of $V(\Delta)-\{r\}$.
 % 
@@
 \subsection{$k$-Planar Graphs}
+\label{sec-k-planar}
 
@@
 This assumption is not critical since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
-Consider the following plane graphs (see \figref{wrapper}(a)):
+Consider the following plane graphs (see \cref{wrapper}(a)):
 \begin{compactenum}
@@
   %, (b)~the spanning tree $T$ of $G^{++}$, 
-  and (b)~the modified version of $G^{++}$ used in the proof of \corref{k-planar}.}
-  \figlabel{wrapper}
+  and (b)~the modified version of $G^{++}$ used in the proof of \cref{cor-k-planar}.}
+  \label{wrapper}
 \end{figure}
@@
 \begin{compactenum}[(PR1)]\setcounter{enumi}{4}
-  \item \proplabel{all-in}  $G^+\subseteq N_{r_K}-F_{r_K}$.
+  \item \PRlabel{all-in}  $G^+\subseteq N_{r_K}-F_{r_K}$.
 \end{compactenum}
@@
 For each vertex $z\in V(G^+)$, let $P_z=\{a\in V(K): z\in V(N_a-F_a)\}$.
-By \propref{all-in}, $P_{z}$ contains the sub-root $r_K$ of $K$.
-By \propref{children}, any node $x\in P_z$ has at most one of its children in $P_z$ and any node $x'\not\in P_z$ has none of its children in $P_z$.
+By \PRref{all-in}, $P_{z}$ contains the sub-root $r_K$ of $K$.
+By \PRref{children}, any node $x\in P_z$ has at most one of its children in $P_z$ and any node $x'\not\in P_z$ has none of its children in $P_z$.
 Therefore $K[P_z]$ is a vertical path in $K$ that contains $r_K$ (and whose lower endpoint is the unique node $x\in V(K)$ such that $z\in Y_x$).
@@
 
-\begin{clm}\clmlabel{chargeback-to-y}
+\begin{clm}\label{chargeback-to-y}
    For any $x\in V(K)$ and any vertex $v\in S_x$, there exists an edge $vw\in E(G)$ and a vertex $v'\in Y_x\cap V(W_{vw})$ such that $|\dist_T(r,v')-\dist_T(r,v)|\le k+1$.
@@
 This implies that $W_{vw}\subseteq N_x-F_x$ but $W_{vw} \not\subseteq N_{x'}-F_{x'}$ for every child $x'$ of $x$ in $K$.
-Therefore, \propref{children} and \propref{ancestor-boundary} imply that $W_{vw}$ contains a vertex $v'\in Y_x$.
+Therefore, \PRref{children} and \PRref{ancestor-boundary} imply that $W_{vw}$ contains a vertex $v'\in Y_x$.
 Since $G$ is $k$-planar, the length of $W_{vw}$ is at most $k+1$.
@@
 
-\begin{clm}\clmlabel{width-of-g-plus}
+\begin{clm}
+\label{width-of-g-plus}
   For any $x\in V(H)$, and any integer $i\ge 0$, $|S_x\cap V_i|\le 24k+36$.
@@
   Since the vertices of $Y_x$ come from at most three vertical paths in $T$, $|V(Y_x)\cap V_i|\le 3$.
-\clmref{chargeback-to-y} shows that, for any $v\in S_x\cap V_i$, there exists $vw\in E(G)$ and $v'\in Y_x\cap V(W_{vw})$ with $\dist_T(v,v')\le k+1$.
+\cref{chargeback-to-y} shows that, for any $v\in S_x\cap V_i$, there exists $vw\in E(G)$ and $v'\in Y_x\cap V(W_{vw})$ with $\dist_T(v,v')\le k+1$.
 Therefore, $v'\in V(Y_x)\cap V_{j}$ for some $j\in\{i-k-1,\ldots,i+k+1\}$.
 If $v'=v$ (so $v'\in V_i$) then $v'$ contributes only one vertex to $S_x$.
-If $v'\neq v$ then $v'\in V(G^+-G)$ is a point where two edges of $G$ cross and $v'$ contributes at most the four endpoints of these edges to $S_x$.
+If $v'\neq v$ then $v'\in V(G^+)\setminus V(G)$ is a point where two edges of $G$ cross and $v'$ contributes at most the four endpoints of these edges to $S_x$.
 In either case, each $v'\in V(Y_x)\cap \bigcup_{j=i-k-1}^{i+k+1} V_j$ contributes at most four vertices to $S_x\cap V_i$, so $|S_x\cap V_i|\le 3\times 4\times (2k+3) = 24k+36$.
@@
 
-\begin{clm}\clmlabel{xy-ancestor}
+\begin{clm}\label{xy-ancestor}
    For every edge $xy\in V(H)$, one of $x$ or $y$ is a $K$-ancestor of the other.
@@
 The tree decomposition $(B_x:x\in V(K))$ of $H$ is defined as follows: For each edge $xy\in E(H)$ where $y$ is a $K$-ancestor of $x$, we add $y$ to $B_{x'}$ for every node $x'$ on the path from $x$ to $y$ in $K$.
-That this produces a tree-decomposition is due to \clmref{xy-ancestor}.
+That this produces a tree-decomposition is due to \cref{xy-ancestor}.
 What remains is to upper bound the width of this decomposition.
 
-\begin{clm}\clmlabel{bag-size}
+\begin{clm}\label{bag-size}
   The tree decomposition $(B_x: x\in V(K))$ of $H$ contains no bag larger than $\bagsize$.
@@
 \begin{proof}
-  Fix some node $x\in V(K)$ and consider the contents of $B_x$.  By \clmref{xy-ancestor}, $B_x$ contains only $K$-ancestors of $x$ (including $x$ itself).
+  Fix some node $x\in V(K)$ and consider the contents of $B_x$.  By \cref{xy-ancestor}, $B_x$ contains only $K$-ancestors of $x$ (including $x$ itself).
 
   Consider the graph $A$ that contains the edges and vertices of $F_{a}$ for all $K$-ancestors $a$ of $x$.
-See \figref{A}(a).
+See \cref{A}(a).
 It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0:=r_K$ is the root of $K$.
@@
     \caption{(a)~the graph $A$ and (b)~the walks $W_{vw}$ and $W_{ww'}$ that contain the walk $w=v_0,v_1\ldots,v_{\ell-1},v_\ell=w^*$.}
-    \figlabel{A}
+    \label{A}
   \end{figure}
 
-  Refer to \figref{A}(b).
+  Refer to \cref{A}(b).
 If $B_x$ contains some $K$-ancestor $a_\delta$, $\delta\in\{0,\ldots,d-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $w\in S_{a_\delta}$ and $v\in S_{x'}$ for some $K$-descendant $x'$ of $x$.
@@
 
-  Let $d_0:=d$, and $d_1:=\min\{i: V(F_d\cap Y_i)\neq\emptyset \}$.  By \propref{ancestor-boundary}, any vertex $z\in V(F_d)$ is contained in 
+  Let $d_0:=d$, and $d_1:=\min\{i: V(F_d\cap Y_i)\neq\emptyset \}$.  By \PRref{ancestor-boundary}, any vertex $z\in V(F_d)$ is contained in 
   $Y_j$ for some $j\in \{d_1,d_1+1,\ldots,d-1\}$.  Therefore, $F_d$ is contained in the interior of  $F_{d_1}$, that is, $N_{d_0}\subseteq N_{d_1}-F_{d_1}$.
 
-  Refer to \figref{d0d1}.  We will now prove that 
+  Refer to \cref{d0d1}.  We will now prove that 
   \begin{equation}
-     |\{\delta\in\{d_1+1,\ldots,d_0\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+3}{2} \enspace . \eqlabel{one-round}
+     |\{\delta\in\{d_1+1,\ldots,d_0\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+3}{2} \enspace . 
+     \eqlabel{one-round}
   \end{equation}
+  %
   \begin{figure}
@@
     \caption{$F_{d_1},\ldots,F_{d_0}$ and the resulting graph $I$.} 
-    \figlabel{d0d1}
+    \label{d0d1}
   \end{figure}
 
-  To prove \eqref{one-round}, consider the graph $I$, illustrated in \figref{d0d1}, with vertex set $V(I)=\{d_1+1,\ldots,d_0\}$ and $ij\in E(I)$ if and only $i=j$ or $|F_i\cap Y_j|\neq\emptyset$.
+  To prove \eqref{one-round}, consider the graph $I$, illustrated in \cref{d0d1}, with vertex set $V(I)=\{d_1+1,\ldots,d_0\}$ and $ij\in E(I)$ if and only $i=j$ or $|F_i\cap Y_j|\neq\emptyset$.
 Any $\delta$-walk $v_0,\ldots,v_\ell$ with $\delta\in V(I)$ corresponds to a walk $\rho_0,\ldots,\rho_{\ell}$ in $I$, where $\rho_i:=\max\{j\in\{0,\ldots,d\} : v_i\in N_{j}-F_j\}$.
@@
 Therefore, $v_{t+1}\in V(F_i)$.
-By \propref{ancestor-boundary}, $v_{t+1}\in V(Y_{j'})$ for some $j'< i$ and therefore $ij'\in E(I)$.
-By \propref{children}, $v_{t+1}\not\in V(N_{j''}-F_{j''})$ for any $K$-descedant $j''>j'$ of $j'$.
+By \PRref{ancestor-boundary}, $v_{t+1}\in V(Y_{j'})$ for some $j'< i$ and therefore $ij'\in E(I)$.
+By \PRref{children}, $v_{t+1}\not\in V(N_{j''}-F_{j''})$ for any $K$-descedant $j''>j'$ of $j'$.
 Therefore $j'=\max\{j : v_{t+1}\in V(N_j-F_j)\}$.
@@
   Now, consider the parent $a_{i-1}$ of $a_i$ in $K$.
-By \propref{children}, $F_i$ is a wedge of the Mercedes graph $M_{i-1}$.
-Let $P_1,\ldots,P_g$ be the partition of $V(F_{i-1})$ guaranteed by \propref{ancestor-boundary}.
-  By \propref{tripod-feet}, $V(F_i) = V(F_i)\cap(P_\alpha \cup P_j\cup V(Y_{i-1}))$ for some $\alpha,j\in \{1,\ldots,g\}$.
+By \PRref{children}, $F_i$ is a wedge of the Mercedes graph $M_{i-1}$.
+Let $P_1,\ldots,P_g$ be the partition of $V(F_{i-1})$ guaranteed by \PRref{ancestor-boundary}.
+  By \PRref{tripod-feet}, $V(F_i) = V(F_i)\cap(P_\alpha \cup P_j\cup V(Y_{i-1}))$ for some $\alpha,j\in \{1,\ldots,g\}$.
 Since $F_i$ contains $v\in V(Y_{d_1})$, at least one of $\alpha$ or $j$, say $\alpha$, is equal $d_1$.
@@
   We are almost done proving \eqref{one-round}.
-Refer to \figref{one-round}.
+Refer to \cref{fig-one-round}.
 For any landing $j_t$, $t>0$, $I^+[\{j_t,\ldots,d_0\}]$ contains the path $d_0,j_1,\ldots,j_t$ of length $t$.
@@
     \caption{The graph $I^+[\{j_{4}+1,\ldots,d_0\}]$.  The shortest path $d_0,j_1,\ldots,j_3$ is highlighted.}
-    \figlabel{one-round}
+    \label{fig-one-round}
   \end{figure}
@@
   \]
-  This establishes \eqref{one-round}.  An illustration of this analysis, showing the case $k=1$ is shown in \figref{one-round-example}.
+  This establishes \eqref{one-round}.  An illustration of this analysis, showing the case $k=1$ is shown in \cref{one-round-example}.
   \begin{figure}[htbp]
@@
     \caption{The set of $6=\binom{k+3}{2}$ vertices in $I^+$ reachable by paths of walks of length at most $k+1$ (for $k=1$).}
-    \figlabel{one-round-example}
+    \label{one-round-example}
   \end{figure}
 
-  Refer to \figref{nested}.
+  Refer to \cref{fig-nested}.
 For $s>1$, let $d_s:=\min\{i: V(F_{d_s}\cap Y_i)\neq\emptyset \}$.
@@
     \caption{The nested cycles $F_{d_0}, F_{d_1},F_{d_2}\ldots$.}
-    \figlabel{nested}
+    \label{fig-nested}
   \end{figure}
@@
 
-We now have all the pieces needed to complete the proof of \thmref{k-planar}.
-
-\begin{proof}[Proof of \thmref{k-planar}]
+We now have all the pieces needed to complete the proof of \cref{k-planar}.
+
+\begin{proof}[Proof of \cref{k-planar}]
   We use the layered $H$-partition $(\mathcal{L}, \mathcal{S})$, where $\mathcal{S}=\{S_x:x\in V(K)\}$ and $\mathcal{L}=\langle V_i': i=0,1,2\ldots\rangle$ where $V_i' = V_{(k+1)i}\cup\cdots\cup V_{(k+1)(i+1)-1}$.
@@
 
-  By \clmref{bag-size}, $\mathcal{S}$ produces a quotient graph $H=G/\mathcal{S}$ of treewidth at most $\treewidth$.
-By \clmref{width-of-g-plus}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+36$, so $|B_x\cap V_i'|\le (k+1)(24k + 36) = 24k^2+60k+36$.
+  By \cref{bag-size}, $\mathcal{S}$ produces a quotient graph $H=G/\mathcal{S}$ of treewidth at most $\treewidth$.
+By \cref{width-of-g-plus}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+36$, so $|B_x\cap V_i'|\le (k+1)(24k + 36) = 24k^2+60k+36$.
 \end{proof}
 
-We conclude this section with a small extension of \thmref{k-planar} that is useful when studying $(g,k)$-planar graphs in \secref{g-k-planar}.
-
-\begin{cor}\corlabel{k-planar}
+We conclude this section with a small extension of \cref{k-planar} that will be useful in \cref{sec-g-k-planar}.
+
+\begin{cor}
+\label{cor-k-planar}
 Let $G$ be a $k$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
-Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \thmref{k-planar} and such that $V_i' := \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\}$ for each integer $i\ge 0$.
+Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \cref{k-planar} and such that $V_i' := \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\}$ for each integer $i\ge 0$.
 \end{cor}
@@
   Without loss of generality, assume that $r_0$ is on the outer face of $G^+$ and let $v_1$ and $v_2$ be the other two vertices of the outer face of $G^+$.
-  The proof of \thmref{k-planar} begins with the graph $G^+$ and then creates a graph $G^{++}$ consisting of a $K_4$ containing $G^+$ in one of its inner faces, $F_0$, and six edges required to complete $G^{++}$ to a triangulation.
-Instead, modify $G^{++}$ by subdividing the edges joining $F_0$ to $v_1$ and $v_2$, as shown in \figref{wrapper}(b).
+  The proof of \cref{k-planar} begins with the graph $G^+$ and then creates a graph $G^{++}$ consisting of a $K_4$ containing $G^+$ in one of its inner faces, $F_0$, and six edges required to complete $G^{++}$ to a triangulation.
+Instead, modify $G^{++}$ by subdividing the edges joining $F_0$ to $v_1$ and $v_2$, as shown in \cref{wrapper}(b).
 In this modified version of $G^{++}$, we have $\dist_{G^++}(r,v)=\dist_{T}(r,v)=\dist_{G^+}(r_0,v)+2$
@@
 
-\newpage
-\section{A Generalization}
-
-We now generalize \thmref{k-planar} to a more abstract setting.
-For a rooted tree $T$ and any node $x\in V(T)$, we define $T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}]$ to be the maximal subtree of $T$ rooted at $x$.
-We begin with a fairly standard lemma about normalized tree decompositions:
-
-\begin{lem}\lemlabel{nice-decomposition}
-  For any graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:  
-  \begin{compactenum}[(T1)]
-    % \item\tlabel{rooted}\tlabel{first} $T$ is rooted at some node $x_0\in V(T)$ with $B_{x_0}=\emptyset$.
-    % \item\tlabel{node-set} $V(T)= V(H)$.
-    % \item\tlabel{diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
-    \item\tlabel{subtree-root} For each $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$.
-    \item\tlabel{ancestor-edge}\tlabel{last} For each edge $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
-  \end{compactenum}
-\end{lem}
-
-\begin{proof}
-  Begin with any width-$t$ tree decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some rooted tree $T_0$.
-Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$.
-Let $f:V(H)\to V(T)$ be function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y]$.
-If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.
-Subdivide the edge between $a$ and its parent in $T$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
-This modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
-
-  Next, consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.
-In this case, $B_{a}\subseteq B_{a'}$ where $a'$ is the parent of $a$ since any $x\in B_a\setminus B_{a'}$ would have $f(x)=a$.
-In this case, contract the edge $aa'$ in $T_0$, eliminating the node $a$.
-Repeating this operation will eventually produce a width-$t$ tree-decomposition of $(B_x:x\in V(T_0))$ where $f$ is a bijection between $V(H)$ and $V(T_0)$.
-Renaming each node $a\in V(T_0)$ as $f^{-1}(a)$ gives a tree-decomposition $(B_x:x\in V(T))$ with $V(T)=V(H)$.
-
-  By the definition of $f$, the tree-decomposition $(B_x:x\in V(T))$ satisfies \tref{subtree-root}.
-To see that $(B_x:x\in V(T))$ satisifies \tref{ancestor-edge}, observe that, if $xy\in E(H)$, then at least one of $x$ or $y$ is contained in $B_z$ for every node $z$ on the path from $x$ to $y$ in $T$.
-If neither $x$ nor $y$ is an ancestor of the other, then some node $z$ on this path has $T$-depth less than that of $x$ and $y$.
-If $x\in B_z$ this contradicts the fact that $x$ is the root of $T[x]$.
-If $y\in B_z$ this contradicts the fact that $y$ is the root of $T[y]$.
-\end{proof}
-
-\subsection{Generalized Tripod Partitions}
-
-For the remainder of this section, $G$ is a graph and $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{Y}:=(Y_x: x\in V(H)))$ is a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{Y}$ is at most $t$.
-In addition, $(B_x:x\in V(T))$ is a tree-decomposition of $H$ obtained by applying \lemref{nice-decomposition} to $H$.
-The following lemma shows how to interpret these objects as a hierarchical decomposition of $G$.
-
-\begin{clm}\clmlabel{generalized-tripod}
-  For each $x\in V(T)$, let $V_x = \bigcup_{y\in V(T_x)} Y_y$, $F_x=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x=V_x\cup F_x$.  Then, 
-  \begin{compactenum}[(Y1)]
-    \item\ylabel{thickness} $\mathcal{Y}:=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width (with respect to $\mathcal{L}$) at most $\ell$.
-    \item\ylabel{separator} For every $x\in V(T)$, there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$. 
-    \item\ylabel{ancestor-edge} For every $x\in V(T)$, $x$ has $t'\le t$ strict $T$-ancestors $a_1,\ldots,a_{t'}$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
-  \end{compactenum}
-\end{clm}
-
-Before proving \clmref{generalized-tripod} we point out more properties that are immediately implied by it:
-
-\begin{compactenum}[(Y1)]\setcounter{enumi}{3}
-  \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$. 
-  \item\ylabel{containment-i} $V_x\subseteq V_a$ for every $T$-ancestor $a$ of $x$. 
-  \item\ylabel{containment-ii}$N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
-\end{compactenum}
-
-Property~\yref{y-subsets} follows from the fact that $V_x$ is the union of several sets, one of which is $Y_x$.
-Property~\yref{containment-i} follows from the definition of $V_x$ and the fact that $V(T_x)\subseteq V(T_a)$.
-To show Property~\yref{containment-ii} first note that, by \yref{containment-i} it suffices to consider vertices in $w\in F_x=N_x\setminus V_x$.
-By definition, every vertex $w\in F_x$ is adjacent, in $G$, to a vertex $v\in V_x$.
-By \yref{containment-i}, $v\in V_a$, so $w$ is either in $V_a$ or $w$ satisfies the condition $vw\in E(G)$, $v\in V_a$, and $w\not\in V_a$, so $w\in F_a$.
-In either case $w\in N_a=V_a\cup F_a$.
-Note that none of \yref{y-subsets}--\yref{containment-ii} depends on \yref{ancestor-edge} (which is important, since \yref{containment-i} is used to establish \yref{ancestor-edge} in the following proof).
-
-\begin{proof}[Proof of \clmref{generalized-tripod}]
-  % Apply \lemref{nice-decomposition} to $H$ to obtain a tree-decomposition $(B_x: x\in V(T))$ of $H$ satisfying the conditions of \lemref{nice-decomposition}.
-  % 
-  % For each $v\in V(G)$, let $y(v)$ denote the unique node $y\in V(H)$ such that $v\in Y_y$.
-Observe that, by \tref{subtree-root}, $y(v)$ is the root of $T[y(v)]=T[\{x\in V(T):y(v)\in B_x\}]$.
-By definition, $\mathcal{Y}:=(Y_x:x\in V(T))$ is a partition of $V(G)$ and, for any $i\in\{1,\ldots,h\}$ $|Y_x\cap V_i|\le \ell$, so $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
-This establishes \yref{thickness}.
-
-  Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$ and $\mathcal{Y}$ is a partition.
-Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.
-In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
-
-  To establish Property~\yref{ancestor-edge}, consider some vertex $w\in F_x$.
-Since $w\in F_x$, there exists an edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.
-Since $v\in V_x$, $v\in Y_{x'}=Q_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
-Since $\mathcal{Y}$ is a partition, $w\in Y_{a}=Q_a$ for some $a\not\in V(T_x)$.
-Since $vw\in E(G)$, $x'a\in E(H)$.
-By \tref{ancestor-edge}, one of $a$ or $x'$ is a $T$-ancestor of the other.
-Since $w\in Y_a\subseteq V_a$ and $w\not\in V_x\supseteq V_{x'}$, \yref{containment-i} rules out the possibility that $x'$ is a $T$-ancestor of $a$.
-Therefore, $a$ is a $T$-ancestor of $x$ which is a $T$-ancestor of $x'$.
-Let $z_0,\ldots,z_r$ be the path, in $T$ from $z_0:=x'$ to $z_r:=a$.
-For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $B_{z_i}$.
-However, by \tref{subtree-root} $x'$ is not contained in $B_{x_i}$ for any $i\in\{1,\ldots,r\}$.
-Therefore $a\in B_{x_i}$ for each $i\in\{0,\ldots,r\}$.
-In particular, $a$ is contained in $B_x$.
-  Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $C_x$ contains $x$.
-\end{proof}
-
-\subsection{Shortcut Systems}
-
-% For non-negative integers $k$ and $d$, a graph $G$ and distinct vertices $v,w\in V(G)$, a \emph{$(k,d)$ $vw$-shortcut} in $G$ is a path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.
-The \emph{$(k,d)^\text{th}$ power of $G$}, denoted $G^{k,d}$, is the graph with vertex $V(G^{k,d})=V(G)$ in which, for each distinct pair $v,w\in V(G)$, $vw\in E(G^{k,d})$ if and only if $G$ contains a $(k,d)$ $vw$-shortcut.
-
-A collection $\mathcal{P}$ of paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if every path in $\mathcal{P}$ has length at most $k$ and if, for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
-Each path in $P\in\mathcal{P}$ is called a \emph{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \emph{$vw$-shortcut}.
-Given a graph $G$ and a $(k,d)$-shortcut system $\mathcal{P}$ for $G$, we use $G^{\mathcal{P}}$ to denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\mathcal{P}$.
-
-In this section we will show how, given a $(k,d)$-shortcut system $\mathcal{P}$, a layered $H$-partition of $G$ can be used to obtain a layered $I$-partition of $G^{\mathcal{P}}$ where the layered width of the partition does not increase dramatically and the treewidth of $I$ is not much more than the treewidth of $H$.
-
-For convenience, it will be helpful to assume that $\mathcal{P}$ contains a
-length-1 $vw$-shortcut for every edge $vw\in E(G)$.  This assumption has no effect on the resulting graph $G^{\mathcal{P}}$ but avoids special cases in some of our proofs.
-
-As in the rest of this section, $(\mathcal{L},\mathcal{Y})$ denotes a layered $H$-partition of $G$; $T$ is a tree used in a tree-decomposition $(B_x:x\in V(T))$ of $H:=G/\mathcal{Y}$ satisfying the conditions of \lemref{nice-decomposition}; and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ are defined as in \clmref{generalized-tripod}.
-
-For a node $x\in V(T)$, we say that a shortcut $P\in\mathcal{P}$ \emph{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
-We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$ or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
-We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
-
-\begin{clm}
-  For any $v\in V(G)$ there exists a (unique) node $a(v)\in X_v$ such that
-  $a(v)$ is a $T$-ancestor of every node in $X_v$.
-\end{clm}
-
-\begin{proof}
-  Let 
-  \[
-     Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:\mbox{$(v_0,\ldots,v_r)\in\mathcal{P}$, $v\in \{v_0,\ldots,v_r\}$}\}
-  \]
-  Clearly $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
-
-  We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
-If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
-In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$.
-
-  Let $Z_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.  The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.  
-  %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other. 
-  Choose $a(v)\in X_H$ to be any member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.
-Transitivity of the $T$-ancestor relationship, \tref{ancestor-edge}, and connectivity of $H[X_H]$ implies that $a(v)$ is a $T$-ancestor of every node $x\in X_H$, as required.
-\end{proof}
-
-For each $x\in V(T)$, we define the \emph{separator}
-\[  
-     S_x := \{v\in V(G): a(v)= x\} \enspace .
-\]
-Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
-We let $I:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
-(Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $I$.)
-
-\begin{clm}\clmlabel{s-subset}
-  For every $x\in V(T)$, $S_x\subseteq V_x$.
-\end{clm}
-
-\begin{proof}
-  For the sake of contradiction, assume otherwise, so there exists some $v\in S_x\setminus V_x$.
-
-  By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.
-Therefore, $\mathcal{P}$ contains a path $P$ that crosses $x$.
-The path $P$ contains a subpath $v_0,v_1,\ldots,v_{r}$ such that $v=v_0$ and $v_r\in Y_x$.
-Since $v\not\in V_x$ and $v_r\in Y_x\subseteq V_x$, \yref{separator}, implies that $v_i\in F_x$ for some $i\in\{0,\ldots,r-1\}$.
-Now \yref{ancestor-edge} implies $v_i\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
-Therefore, either $v\in Y_a$ or $P$ crosses $a$.
-But this implies that $a(v)$ is an ancestor of $a$, contradicting the assumption that $v\in S_x$.
-\end{proof}
-
-Next we show that $\mathcal{S}$ has small layered width with respect to $\mathcal{L}$:
-
-\begin{clm}\clmlabel{general-width}
-  For each $i\in\{1,\ldots,h\}$ and each $x\in V(I)$, $|S_x\cap V_i|\le d\ell(k^2+k+4)/2$.  
-\end{clm}
-
-\begin{proof}
-  We will count the number of vertices in $S_x\cap V_i$ by upper-bounding the number of vertices contributed to $S_x\cap V_i$ by each vertex $w\in Y_x$.
-  If $w\in Y_x\cap V_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes only one vertex, itself, to $S_x\cap V_i$.
-
-  Otherwise, consider some path $P\in\mathcal{P}$ that contains $w$ as an internal vertex.
-If $w\in V_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap V_i$.
-If $w\in V_{i-1}\cup V_{i+1}$, then $P$ contributes at most $k$ vertices to $S_x\cap V_i$.
-If $w\in V_{i-j}\cup V_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap V_i$.
-
-  For any $j$, the number of vertices $w\in V_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\mathcal{P}$. Therefore, 
-  \[  |S_x\cap V_i|\le d\ell\cdot\left(k+1 + k + \sum_{j=2}^k (k-j)\right) %= d(2k+1) + \sum_{i=1}^{k-2} i
-      = d\ell(k^2 +k + 4)/2 \enspace . \qedhere
-  \]
-\end{proof}
-
-\begin{clm}\clmlabel{i-ancestor}
-  For each edge $xy\in E(I)$, one of $x$ or $y$ is a $T$-ancestor of the other.
-\end{clm}
-
-\begin{proof}
-  Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.
-Since $xy\in E(I)$, $G^\mathcal{P}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
-Since $xy\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.
-By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.
-By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
-By \yref{ancestor-edge} $N_x$ and $V_y$ are also disjoint.
-By \yref{separator} $P$ contains an internal vertex $v'\in F_x$.
-By \yref{ancestor-edge}, $v'\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
-But this implies that $v\in S_{a'}$ for some $T$-ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
-\end{proof}
-
-\begin{clm}\clmlabel{general-bag-size}
-  The graph $I$ has a tree-decomposition in which every bag has size at most $\binom{k+t}{t}$.
-\end{clm}
-
-\begin{proof}
-  For the tree-decomposition $(C_x:x\in V(T))$ of $I$ we use the same tree $T$ used in the tree-decomposition $(B_x:x\in V(T))$ of $H$.
-For each node $x$ of $T$, $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $I$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
-  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(I)$, $a,x'\in C_{x'}$ and the connectivity of $T[a]:=T[\{x\in V(T):a\in C_x]$ is obvious.
-Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $I$.
-What remains is to bound the size, $\max\{|C_x|:x\in V(T)\}$, of the largest bag.
-
-  Consider an arbitrary node $x\in V(T)$ whose ancestors are $x_0,\ldots,x_{r-1},x_r$ where $x_0$ is the root of $T$ and $x_r:=x$.
-To avoid triple-subscripts in what follows, we abuse notation slightly by using $V_i$, $F_i$, and $N_i$,  as shorthands for $V_{x_i}$, $F_{x_i}$ and $N_{x_i}$, respectively.
-
-  If $x_\delta\in C_x$, it is because $x_\delta x'\in E(I)$ for some $T$-descendant $x'$ of $x$.
-This implies $G^{\mathcal{P}}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
-This implies that $\mathcal{P}$ a $vw$-shortcut $P_{vw}$.
-Let $v'$ be the second-last vertex of $P_{vw}$ (so $v'w\in E(G)$).
-
-  Since $w\in S_{\delta}$, at least one of the following is true:
-  \begin{enumerate}
-    \item $\mathcal{P}$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
-    \item $w\in Y_\delta$.  In this case, we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$. 
-  \end{enumerate}
-  Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{\delta}$.
-
-  Let $w_0,w_1,\ldots,w_p$ be the path that begins $w_0:=v'$ and then follows the subpath of $P_{ww'}$ that begins at $w_1:=w$ and ends at $w_p:=w''$.
-For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{1,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.
-Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root is non-increasing.
-
-  By definition, $a_0=x_r$.  We claim that $a_p=x_\delta$, i.e., $s_p=\delta$. 
-  To see this, first observe that, for each $i\in\{1,\ldots,p\}$, $w_i\in V_{\delta}$ since, otherwise, an internal vertex of $P_{ww'}$ belong to $F_\delta$, which would imply (by \yref{ancestor-edge}) that $w\in S_{\delta'}$ for some $\delta' < \delta$, contradicting the assumption that $w\in S_\delta$.
-Therefore $s_p\ge\delta$.
-To see that $s_p<\delta-1$,
-  observe that either $w=w''\in Y_\delta$ or $P_{ww'}$ contains an internal vertex $w''$ in $Y_\delta$.
-By \yref{thickness} and the defnition of $V_x$, $V_{\delta-1}$ does not contain $w''$, so $s_p<\delta-1$.
-
-  Let $H^+$ denote the supergraph of $H$ with vertex set $V(T)$ and in which $xy\in E(H^+)$ if and only there exists some $z\in V(T)$ such that $x,y\in B_z$. 
-  % Note that, by \tref{ancestor-edge}, $xy\in E(H^+)$ if and only if $x\in B_y$ or $y\in B_x$.
-  We claim that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \emph{lazy walk} in a graph $H$ is a walk in the non-simple graph $H'$ obtained by adding a self loop to each vertex of $H$.} in $H^+$.
-Indeed, if $a_i\neq a_{i+1}$ for some $i\in\{0,\ldots,p-1\}$ then this is precisely because $w_i\in V_{a_i}$ but $w_{i+1}\not\in V_{a_i}$.
-By definition, $w_i\in Y_{a_i'}$ for some $T$-descendant $a_i'$ of $a_i$.
-  By \yref{separator}, $w_{i+1}\in F_{a_i}$ so by \yref{ancestor-edge} $w_{i+1}\in Y_{a_i''}$ for some strict $T$-ancestor $a_i''$ of $a_i$.
-Since $w_iw_{i+1}\in E(G)$, $a_i'a_i''\in E(H)$.
-By \tref{subtree-root}, $a_i''\in B_{a_i''}$ and $a_i''\in B_{a_i'}$.
-Since $a_i$ is on the path from $a_i'$ to $a_i''$ in $T$ this implies that $a_i''\in B_{a_i}$.
-Therefore $a_ia_i''\in E(H^+)$ as claimed.
-
-  Thus, $a_0,\ldots,a_p$ is a lazy walk in $H^+$ of length $p\le k$ where the distance $s_i$ between $a_i$ and the root $x_0$ of $T$ is non-decreasing.
-By removing repeated vertices this gives a path in the directed graph $\overrightarrow{H}^+$ obtained by directing each edge $xy\in E(H^+)$ from its $T$-descendant $x$ towards its $T$-ancestor $y$.
-  Finally, we are in a position to appeal to \cite[Lemma~24]{pilipczuk.siebertz:polynomial-arxiv} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
-\end{proof}
-
-% Together, \clmref{general-width} and \clmref{general-bag-size} imply the following generalization of \thmref{k-planar}:
-
-\begin{thm}\thmlabel{modern-major-general}
-  Let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\mathcal{P}$ be a $(k,d)$-shortcut system for $G$.
-Then $G^{\mathcal{P}}$ has a layered $I$-partition of layered width at most $d\ell(k^3+k^2+4k)/2$ where the treewidth of $H$ is at most $\binom{k+t}{t}-1$.
-\end{thm}
-
-\begin{proof}
-  The proof is almost immediate, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
-In particular, $G^{\mathcal{P}}$ may contain an edge $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.
-To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.
-This increases the thickness given by \clmref{general-width} from $d\ell(k^2+k+4)/2$ to $d\ell(k^3+k^2+4k)/2$.
-\end{proof}
-
-% It is natural to ask if \thmref{modern-major-general} is tight.
-We do not answer this question, but we do point out that, even when $t=1$ and $d=3$, an exponential dependence on $k$ is necessary.
-If $G$ is a complete binary tree on $2^{k/2}$ leaves, then $G^{k,3}$ is a clique of size  $s=2^{k/2+1}-1$.
-Any partition of $G^{k,3}$ into $r$ parts will give an quotient graph $I$ that is an $r$-clique, which has treewidth $r-1$.
-On the other hand, any layering of $V(G^{k,3})$ has at most two layers, so the layered width of any partition into $r$ parts is at least $s/(2r)$.
-Therefore any layered $I$-partition of $G^{k,d}$ of layered width $\ell'$ in which the treewidth of $I$ is at most $t'-1$ must satisfy
-% \[
-%     \ell'\times t' \ge 2^{k/2}-1/2 \enspace .
-% \]   
-\newpage
-
-\section{$1$-Planar Graphs}
-\seclabel{1-planar}
+\subsection{$1$-Planar Graphs}
+\label{sec-1-planar}
 
@@
 
-Our proof of \thmref{1-planar} follows quickly from the following technical lemma, which is an extension of the analagous result for plane graphs \cite{dujmovic.joret.ea:planar}.
-\begin{lem}\lemlabel{induction} The setup:
+Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analagous result for plane graphs \cite{dujmovic.joret.ea:planar}.
+\begin{lem}
+\label{induction} The setup:
   \begin{compactenum}
@@
   \end{compactenum}
-  Then $N$ has an $H$-partition $\mathcal{P}=\{S_x : x\in V(H)\}$ such that
+  Then $N$ has an $H$-partition $\mathcal{P}=\{S_x : x\in V(H)\}$ such that:
   \begin{compactenum}
@@
 
-Using \lemref{induction}, the proof of \thmref{1-planar} is now straightforward.
-
-\begin{proof}[Proof of \thmref{1-planar}]
+Using \cref{induction}, the proof of \cref{1-planar} is now straightforward.
+
+\begin{proof}[Proof of \cref{1-planar}]
 Given a 1-plane graph $G$, add edges to make it edge maximal so that it has an outer face $F=v_1v_2v_3$.
@@
   Let $G'$ be the plane graph obtained by removing one spar from each kite of $\overline{G}$ and let $T$ be a BFS tree of $G'$ rooted at $r$.
-Now apply \lemref{induction} with $G=\overline{G}$, $G'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
+Now apply \cref{induction} with $G=\overline{G}$, $G'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
 This gives an $H$-partition $\{S_x:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ is planar and has treewidth at most 3.
@@
 
-Exactly the same argument used to establish \corref{k-planar} as a consequence of \thmref{k-planar} establishes the following result as a consequence of \thmref{1-planar}:
-
-\begin{cor}\corlabel{1-planar}
+Exactly the same argument used to establish \cref{cor-k-planar} as a consequence of \cref{k-planar} establishes the following result as a consequence of \cref{1-planar}:
+
+\begin{cor}
+\label{cor-1-planar}
 Let $G$ be a $1$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
-Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \thmref{1-planar} and such that $V_i' = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/2}=i\}$ for each integer $i\ge 0$.
+Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \cref{1-planar} and such that $V_i' = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/2}=i\}$ for each integer $i\ge 0$.
 \end{cor}
 
-\section{$(g,k)$-Planar Graphs}
-\seclabel{g-k-planar}
-
-Next we prove \thmref{g-k-planar}, which generalizes our results to $(g,k)$-plane graphs.  We make use of the following lemma due to \citet[Lemma~21]{dujmovic.joret.ea:planar}:
-
-\begin{lem} \lemlabel{make-planar}
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{$(g,k)$-Planar Graphs}
+\label{sec-g-k-planar}
+
+The definition of $k$-planar graphs naturally generalises for other surfaces.
+A graph $G$ drawn on a surface $\Sigma$ is $(\Sigma,k)$-plane if every edge of $G$ is involved in at most $k$ crossings.
+A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
+We prove the following generalisation of \cref{k-planar}.
+
+\begin{thm}
+\label{g-k-planar}
+Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most 
+$\max\{8gk^2 + 12gk + 4g, 24k^2 + 36k + 12\}$ in which $H$ has treewidth at most $\treewidth$.
+\end{thm}
+
+Again, in the case $k=1$ we obtain better constants.
+
+\begin{thm}
+\label{g-1-planar}
+Every $(g,1)$-planar graph has a layered $H$-partition of layered width at most $\max\{48g, 30\}$ in which $H$ has treewidth at most $4$.  
+\end{thm}
+
+The proofs of these results make use of the following lemma due to \citet[Lemma~21]{dujmovic.joret.ea:planar}:
+
+\begin{lem}[\citep{dujmovic.joret.ea:planar}] 
+\label{make-planar}
 Let $G^+$ be a $(\Sigma,0)$-plane graph where $\Sigma$ is a surface with Euler genus $g$. 
@@
 \[ |\{v\in Z: \dist_{G^+}(r,v) = i\}| \le 2g \enspace . \]
-
 \end{lem}
 
-\begin{proof}[Proof of \thmref{g-k-planar}]
+\begin{proof}[Proof of \cref{g-k-planar,g-1-planar}]
   Let $G$ be an $(\Sigma,k)$-plane graph where $\Sigma$ is a surface of Euler genus $g$.
@@
 
-  Apply \lemref{make-planar} to $G^{+}$ to obtain the set $Z\subseteq V(G^{+})$ and the planar graph $G^*$ with special vertex $r^*$.
+  Apply \cref{make-planar} to $G^{+}$ to obtain the set $Z\subseteq V(G^{+})$ and the planar graph $G^*$ with special vertex $r^*$.
 For each edge $vw\in E(G)$, let $W_{vw}$ be the corresponding walk in $G^+$.
@@
 In this way, each vertex $z\in Z\cap V(G^+)\setminus V(G)$ can be charged for at most two vertices in $v,x\in Z'$ and this happens only if $(z,vw,xy)$ is a crossing in $G$.
-  Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \clmref{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 2\times (2k+1) = 4g(2k+1)$ for each integer $i\ge 0$.
+  Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \cref{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 2\times (2k+1) = 4g(2k+1)$ for each integer $i\ge 0$.
 
@@
 The fact that $G^*$ can be completed to a triangulation $G^{*+}$ while preserving distances to $r^*$ follows from the fact that any face $F$ of $G^*$ with four or more vertices contains two non-adjacent vertices $x$ and $y$ such that $|\dist_{G^*}(r^*,x)-\dist_{G^*}(r^*,y)|\le 1$.
-Indeed, if $z$ is a vertex of $F$ that minimizes $\dist_{G^*}(r^*,z)$, then two neighbours $x$ and $y$ of $z$ have this property.
+Indeed, if $z$ is a vertex of $F$ that minimises $\dist_{G^*}(r^*,z)$, then two neighbours $x$ and $y$ of $z$ have this property.
 Now, observe that $G^*_1$ is a $k$-plane graph and $G^{*+}_1$ is obtained by adding dummy vertices at crossings in $G^*_1$ and triangulating the resulting graph.
 
-  Apply \corref{k-planar} or (in the case $k=1$) \corref{1-planar} to $G^*_1$ using $G^{*+}_1$ and the distinguished vertex $r^*$ to obtain a layered $H$-partition of $G^*_1$.
-Since $G^*_1\supseteq G^*_0$ and $G^*_0$ is isomorphic to $G-Z'$, this gives a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ of $G-Z'$ that satisfies the conditions of \corref{k-planar} or \corref{1-planar}, as appropriate.
+  Apply \cref{cor-k-planar} or (in the case $k=1$) \cref{cor-1-planar} to $G^*_1$ using $G^{*+}_1$ and the distinguished vertex $r^*$ to obtain a layered $H$-partition of $G^*_1$.
+Since $G^*_1\supseteq G^*_0$ and $G^*_0$ is isomorphic to $G-Z'$, this gives a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ of $G-Z'$ that satisfies the conditions of \cref{cor-k-planar} or \cref{cor-1-planar}, as appropriate.
 In particular, for every integer $i\ge 0$,
@@
 
-\section{Consequences}
-\seclabel{consequences}
-
-Here we discuss some of the consequences of Theorems~\ref{thm:k-planar}--\ref{thm:g-k-planar} for other graph parameters that can be bounded using layered partitions and other graph classes that are $k$-planar.
-
-\subsection{Applications to Colouring}
-
-For a graph $G$ and a set $C$ of size $c$, any function $\phi\colon V(G)\to C$  is called a \emph{$c$-colouring} of $G$.
-
-\subsubsection{Non-Repetitive Colouring}
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Applications}
+\label{Applications}
+
+Here we discuss some of the consequences of our main theorems for $k$-planar and $(g,k)$-planar graphs.
+
+%%%%%%%%%%%%%%%%%%%
+\subsection{Queue Layouts}
+
+For an integer $k\geq 0$, a \textit{$k$-queue layout} of a graph $G$ consists of a linear ordering $\preceq$ of $V(G)$ and a partition $\{E_1,E_2,\dots,E_k\}$ of $E(G)$, such that for $i\in\{1,2,\dots,k\}$, no two edges in $E_i$ are nested with respect to $\preceq$.
+That is, it is not the case that $v\prec x \prec y \prec w$ for edges $vw,xy\in E_i$.
+The \textit{queue-number} of a graph $G$, denoted by $\qn(G)$, is the minimum integer $k$ such that $G$ has a $k$-queue layout.
+Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
+\citet{dujmovic.joret.ea:planar} recently solved this question using \cref{PlanarPartition} and the following lemma.
+Indeed, resolving this question was the motivation for the development of \cref{PlanarPartition}.
+
+\begin{lem}[\citep{dujmovic.joret.ea:planar}]
+\label{qn}
+If a graph $G$ has a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$, then $\qn(G)\le 3 \ell (2^t-1) + \floor{\tfrac{3}{2}\ell}$.
+\end{lem}
+
+\cref{qn,g-k-planar} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$. 
+Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of 
+$O(g^{k+2})$ using \cref{SurfacePartition} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$. In particular, in the case for $k=1$, \cref{qn,1-planar,g-1-planar} imply:
+
+\begin{prop}
+For $g\geq 1$, every $(g,1)$-planar graph has queue-number at most $3 \times 48g (2^4-1) + \floor{\tfrac{3}{2} \times 48g }= 2232 g $.
+\end{prop}
+
+The previous best bound on the queue-number of $(g,1)$-planar graphs was $2(40g+490)^{3}$ due to \citet{dujmovic.joret.ea:planar}.
+For $g=0$ we can improve further.
+\citet{ABGKP18} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
+Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$.
+\cref{qn,1-planar} then imply:
+
+\begin{prop}
+Every 1-planar graph has queue-number at most $3 \times 30 * 5 + \floor{\tfrac{3}{2} \times 30} = 495$. \end{prop}
+
+%%%%%%%%%%%%%%%%%%%
+\subsection{Non-Repetitive Colouring}
+
+The next three applications are in the field of graph colouring. For our purposes, a \emph{$c$-colouring} of a graph $G$ is any function $\phi\colon V(G)\to C$, where $C$ is a set of size at most $c$.
 
 A $c$-colouring $\phi$ of $G$ is \emph{non-repetitive} if, for every path $v_1,\ldots,v_{2h}$ in $G$, there exists some $i\in\{1,\ldots,k\}$ such that $\phi(v_i)\neq\phi(v_{i+h})$.
-The \emph{non-repetitive chromatic number} $\pi(G)$ of $G$ is the minimum value $c$ such that $G$ has a non-repetitive $c$-colouring.
-
-\citet[Corollaries~9 and 10]{dujmovic.esperet.ea:planar} prove the following result:
-
-\begin{lem}\lemlabel{non-repetitive}
+The \emph{non-repetitive chromatic number} $\pi(G)$ of $G$ is the minimum integer $c$ such that $G$ has a non-repetitive $c$-colouring.
+This concept was introduced by \citet{AGHR-RSA02} and has since been widely studied; see \citep{dujmovic.esperet.ea:planar} for more than 40 references.
+Up until recently the main open problem in the field has been whether planar graphs have bounded non-repetitive chromatic number, first asked by \citet{AGHR-RSA02}.
+\citet{dujmovic.esperet.ea:planar} recently solved this question using \cref{PlanarPartition} and the following lemma.
+
+%\citet[Corollaries~9 and 10]{dujmovic.esperet.ea:planar} prove the following result:
+
+\begin{lem}[\citep{dujmovic.esperet.ea:planar}]
+\label{non-repetitive}
 If a graph $G$ has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$, then $\pi(G)\le \ell 4^{t+1}$.
@@
 
-Combining \lemref{non-repetitive} with Theorems~\ref{thm:k-planar}--\ref{thm:g-k-planar} we immediately obtain the following corollaries:
+\cref{non-repetitive,k-planar,1-planar,g-k-planar} immediately imply the following results:
 
@@
 
-Prior to the current work, the strongest upper bound on the non-repetitive chromatic number of $n$-vertex planar $k$-planar graphs was $O(k\log n)$ \cite{dujmovic.morin.ea:layered}.
-
-\subsubsection{Centered Colourings}
+Prior to the current work, the strongest upper bound on the non-repetitive chromatic number of $n$-vertex  $k$-planar graphs was $O(k\log n)$ \cite{dujmovic.morin.ea:layered}.
+
+%%%%%%%%%%%%%%%%%%%
+\subsection{Centered Colourings}
 
@@
 In words, either $X$ receives more than $p$ colours or some vertex in $X$ receives a unique colour.
-Let $\chi_p(G)$ denote the smallest integer $c$ such that $G$ has a $p$-centered $c$-colouring.
-
-We make use of the following lemma due to \citet{pilipczuk.siebertz:polynomial-arxiv,pilipczuk.siebertz:polynomial-soda}
-\begin{lem}\lemlabel{p-centered-treewidth}
-  Any graph $H$ of treewidth at most $t$ has $\chi_p(H)\in O(p^t)$.
+Let $\chi_p(G)$ be the minimum integer $c$ such that $G$ has a $p$-centered $c$-colouring.
+Centered colourings are important since they characterise classes of bounded expansion, which is a key concept in the sparsity theory of \citet{Sparsity}.
+
+We make use of the following lemma due to \citet{pilipczuk.siebertz:polynomial-soda}.
+
+\begin{lem}[\citep{pilipczuk.siebertz:polynomial-soda}]
+\label{p-centered-treewidth}
+Every graph $H$ of treewidth at most $t$ has $\chi_p(H)\in (t+1)\binom{p+t}{t}$.
 \end{lem}
 
-The following lemma is due to Micek \cite{micek:personal}. We include the proof here for completeness.
-
-\begin{lem}\lemlabel{p-centered}
-  Any graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$, has $\chi_p(G)\le \ell\times (p+1)\times O(p^t)$.
+The following lemma is due to \citet{micek:personal}. We include the proof  for completeness.
+
+\begin{lem}[\citep{micek:personal}]
+\label{p-centered}
+Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
 \end{lem}
@@
   Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$ where $H$ as treewidth at most $t$.
-  We use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,O(p^t)\}$.
-For each integer $i\ge 0$ and each $x\in V(H)$, the colour $\phi(v)=(\alpha(v),\beta(v),\gamma(v))$ of a vertex $v\in V_i\cap B_x$ is assigned to satisfy the following conditions:
-  \begin{enumerate}
-    \item $\alpha(v)$ is unique among $\{\phi(w): w\in V_i\cap B_x\}$.
-
-    \item $\beta(v)= i\bmod (p+1)$.
-
-    \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,O(p^t)\}$ is a $p$-centered colouring of $H$.
-  \end{enumerate}
-  Condition 1 is easy to satisfy since $|V_i\cap B_x|\le \ell$.
-Condition~2 is straightforward.
-Condition~3 is possible to satisfy thanks to \lemref{p-centered-treewidth}.
-To see why this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
-  \begin{enumerate}
-    \item If there exists $v,w\in V(X)$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$ then, since $G[X]$ is connected, $G[X]$ contains a path from $v$ to $w$.
+  We use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,
+  (t+1)\binom{p+t}{t}\}$.  For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in V_i\cap B_x$ such that:
+  \begin{compactenum}
+    \item $\alpha(v)$ is unique among $\{\phi(w): w\in V_i\cap B_x\}$, which is possible 
+    since $|V_i\cap B_x|\le \ell$,
+    \item $\beta(v)= i\bmod (p+1)$, and     
+    \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,(t+1)\binom{p+t}{t}\}$ is a $p$-centered colouring of $H$ from     \cref{p-centered-treewidth}.
+  \end{compactenum}
+To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
+
+First suppose that there exists $v,w\in V(X)$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$.
+Since $G[X]$ is connected, $G[X]$ contains a path from $v$ to $w$.
 By the definition of layering, this path contains at least one vertex from $V_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$.
@@
 
-    \item Otherwise, $X\subseteq V_{i},\ldots,V_{i+s}$ for some $s<p$.
-Let $H'=H[\{x\in V(H):B_x\cap X\neq\emptyset]$.
+Otherwise, $X\subseteq V_{i},\ldots,V_{i+s}$ for some $s<p$.
+Let $H':=H[\{x\in V(H):B_x\cap X\neq\emptyset]$.
 If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in X\}|> p$ so $|\{\phi(v):v\in X\}|> p$ and we are done.
@@
 Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.
-Therefore, every vertex in $B_x$ receives colour distinct from every other vertex in $X$.
-  \end{enumerate}
-\end{proof}
-
-Combining \lemref{p-centered} with Theorems~\ref{thm:k-planar}--\ref{thm:g-k-planar} immediately gives the following results, for every $p\ge 2$:
+Therefore, every vertex in $B_x$ receives colour distinct from every other vertex in $X$.\end{proof}
+
+\cref{p-centered,k-planar,1-planar,g-k-planar} immediately imply the following results, for every $p\ge 2$:
 
 \begin{cor}
+\label{pck}
 For every $k$-planar graph $G$, $\chi_p(G)\le p^{O(k^3)}$.
@@
 \begin{cor}
+\label{pc1}
 For every $1$-planar graph $G$, $\chi_p(G)\le O(p^4)$. 
@@
 \begin{cor}
+\label{pcgk}
 For every $(g,k)$-planar graph $G$, $\chi_p(G)\le O(gk^2)\times p^{O(k^3)}$. 
@@
 
-\begin{cor}\corlabel{last}
+\begin{cor}
+\label{pcg1}
 For every $(g,1)$-planar graph $G$, $g\ge 1$, $\chi_p(G)\le O(gp^4)$.
@@
 
-Prior to the current work, the strongest upper bounds on the number of colours required for $p$-centered colourings of $(g,k)$-planar graphs used the fact that $(g,k)$-planar graphs are have layered treewidth $O(gk)$ \cite{dujmovic.eppstein.ea:structure} and therefore have bounded (indeed, linear) expansion \cite{dujmovic.morin.ea:layered}.
+%Prior to the current work, the strongest upper bounds on the number of colours required for $p$-centered colourings of $(g,k)$-planar graphs used the fact that $(g,k)$-planar graphs have layered treewidth $O(gk)$ \cite{dujmovic.eppstein.ea:structure} and therefore have bounded (indeed, linear) expansion \cite{dujmovic.morin.ea:layered}.
 Using a result of \citet{zhu:colouring} this leads to bounds on $\chi_p(G)$ that are doubly-exponential in $gk$.
 
-\note{PM}{Please check the preceding paragraph.}
-
-% \note{PM}{I didn't follow the discussion on the previous best-known bounds for $p$-centered colourings.  Can someone fill me in?}
-
-\subsection{Other Graph Classes}
-
-We conclude by pointing out three other graph classes, namely $(g,d)$-map graphs, $(g,\delta)$-string graphs, and $k$-nearest-neighbour graphs that are $(g,k)$-planar for $k=O(d^2)$ and $k=\delta^2$, respectively.
-This implies that Theorems~\ref{thm:k-planar}--\ref{thm:g-k-planar} and Corollaries~\ref{cor:first}--\ref{cor:last} apply to $(g,d)$-map graphs, $(g,\delta)$-string graphs, and $k$-nearest-neighbour graphs.
-
-\subsubsection{Map Graphs}
+Prior to the current work, the strongest known upper bounds on the $p$-centered chromatic number of $(g,k)$-planar graphs $G$ were doubly exponential in $p$, as we now explain. 
+\citet{dujmovic.eppstein.ea:structure} proved that $G$ has layered treewidth $(4g+6)(k+1)$. 
+Van den Heuvel and Wood~\citep{vdHW17} showed that this implies that $G$  has $r$-strong colouring number at most $(4g + 6)(k + 1)(2r + 1)$.
+By a result of \citet{zhu:colouring}, $G$ has $r$-weak colouring number at most $( (4g + 6)(k + 1)(2r + 1) )^r$, which by another result of  \citet{zhu:colouring} implies that $G$ has  $p$-centred chromatic number at most $( (4g+6)(k+1)(2^{p-1} + 1) )^{2^{p-2}}$.
+\cref{pc1,pck,pcgk,pcg1} are substantial improvements over these bounds.
+
+%%%%%%%%%%%%%%%%
+\subsection{Low Treewidth Colouring}
+
+\citet{DDOSRSV04} proved the conjecture fo Robin Thomas that every graph in a proper minor-closed class has an edge 2-colouring so that each monochromatic subgraph has bounded treewidth, and more generally, that for fixed $c\geq 2$, every such graph can be edge $c$-coloured such that the union of any $c-1$ colour classes has bounded treewidth.
+They also showed analogous vertex-colouring results.
+\citet{dujmovic.joret.ea:planar} gave simple proofs of these results via layered partitions.
+The following lemma is implicit in \citep{dujmovic.joret.ea:planar}.
+
+\begin{lem}[\citep{dujmovic.joret.ea:planar}]
+\label{LowTreewidth}
+Let $G$ be a graph that has an $H$-partition of layered width $\ell$ such that $H$ has treewidth at most $k$.
+Fix $c\geq 2$.
+Then (a) $G$ has an edge $c$-colouring such that the spanning subgraph of $G$ consisting of the union of any $c-1$ colour classes has treewidth at most $\ell(k+1)(2c-1)$, and (b) $G$ has a vertex $c$-colouring such that the subgraph of $G$ induced by the union of any $c-1$ colour classes has treewidth at most $\ell(k+1)(2c-1)$.
+\end{lem}
+
+One feature of this result compared to that of \citet{DDOSRSV04}  is that it applies for non-minor-closed classes, such as $k$-planar graphs. In particular, \cref{LowTreewidth,g-k-planar} imply:
+
+\begin{prop}[\citep{dujmovic.joret.ea:planar}]
+For every $(g,k)$-planar graph $G$ and integer $c\geq 2$, 
+(a) $G$ has an edge $c$-colouring such that the spanning subgraph of $G$ consisting of the union of any $c-1$ colour classes has treewidth $O(cgk^5 )$, and (b) $G$ has a vertex $c$-colouring such that the subgraph of $G$ induced by the union of any $c-1$ colour classes has treewidth $O(cgk^5 )$.
+\end{prop}
+
+%%%%%%%%%%%%%%%%
+%\subsection{Graph Encodings}
+%
+%\textcolor{red}{Whether we include this section will depend on whether we get a copy of \citep{BGP} prior to SODA submission. They are submitting to SODA too (as is \citep{micek:personal})}
+
+%%%%%%%
+\section{Rough Characterisation}
+\label{Characterisation}
+
+This section presents a rough characterisation of $(g,k)$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $(g,k)$-planar.
+
+\begin{lem} 
+\label{DrawG}
+Fix integers $g\geq 0$ and $k,\Delta\geq 2$. 
+Let $G_0$ be a graph of Euler genus at most $g$. Let $G$ be
+a graph with $V(G) \subseteq V(G_0)$ such that for every edge $vw \in
+E(G)$ there is a $vw$-path $P_{vw}$ in $G_0$ of length at most $k$, such
+that every internal vertex on $P_{vw}$ has degree at most $\Delta$ in
+$G_0$. Then $G$ is $(g, (k+1)\Delta^{2(k-1)} )$-planar.
+\end{lem}
+
+\begin{proof}
+For a vertex $x$ of $G_0$ with degree at most $\Delta$, say a vertex $v$ is \emph{close} to $x$ if there is a $vx$-path $P$ in $G_0$ of length at most $k-1$ such that every internal vertex in $P$  has degree at most $\Delta$ in $G_0$.
+Note that there are at most $\Delta^{k-1}$ vertices close to $x$.
+For each edge $vw$ of $G$, say that $vw$ \emph{passes through} each internal vertex on $P_{vw}$. 
+If $vw$ passes through $x$, then both $v$ and $w$ are close to $x$. 
+Thus at most $\Delta^{2(k-1)}$ edges of $G$ pass through $x$. 
+Draw each edge $vw$ of $G$ alongside $P_{vw}$ in $G_0$, so that every pair of edges cross at most once. 
+Every edge of $G$ that crosses $vw$ passes through a vertex on $P_{vw}$ (including $v$ and/or $w$ if they too have degree at most $\Delta$). 
+Since $P_{vw}$ has at most $k+1$ vertices, and at most $\Delta^{2(k-1)}$ edges of $G$ pass through each vertex on $P_{vw}$, 
+the edge $vw$ is crossed by at most $(k+1)\Delta^{2(k-1)}$ edges in $G$. 
+Hence $G$ is $(g, (k+1)\Delta^{2(k-1)} )$-planar.
+\end{proof}
+
+Note that the following converse of \cref{DrawG}  obviously holds (just add a dummy vertex at each crossing point):
+
+\begin{lem}
+\label{AddDummy}
+For every $(g,k)$-planar graph $G$ there is a graph $G_0$ of Euler
+genus at most $g$, such that $V(G) \subseteq V(G_0)$ and for every
+edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most
+$k+1$, such that every internal vertex in $P$ has degree at most $4$
+in $G_0$.
+\end{lem}
+
+Together \cref{DrawG,AddDummy} provide a rough characterisation of $(g,k)$-planar graphs.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Examples}
+\label{Examples}
+
+This section describes several graph classes that are $k$-planar or $(g,k)$-planar. The first two examples are well known, the second two are new.
+
+%We conclude by pointing out three other graph classes, namely $(g,d)$-map graphs, $(g,\delta)$-string graphs, and $k$-nearest-neighbour graphs that are $(g,k)$-planar for $k=O(d^2)$ and $k=\delta^2$, respectively.
+This implies that Theorems~\ref{k-planar}--\ref{g-k-planar} and Corollaries~\ref{cor:first}--\ref{cor:last} apply to $(g,d)$-map graphs, $(g,\delta)$-string graphs, and $k$-nearest-neighbour graphs.
+
+\subsection{Map Graphs}
 
@@
 Then $G$ is called a \emph{$(g,d)$-map graph}.
-% A $(0,d)$-map graph is called a (plane) \emph{$d$-map graph}; such graphs have been extensively studied \citep{FLS-SODA12,DFHT05,CGP02,Chen07,Chen01}.
-The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$ (see \citep{DEW17}).
-So $(g,d)$-map graphs provide a natural generalisation of graphs embedded in a surface.
-
-The following lemma comes from the natural drawing of a map graph obtained by positioning each vertex of $G$ inside the corresponding nation and each edge of $G$ as a curve passing through the corresponding vertex of $G_0$.
-See \citet[Page~4]{dujmovic.eppstein.ea:structure}.
+A $(0,d)$-map graph is called a (plane) \emph{$d$-map graph}; see \citep{FLS-SODA12,CGP02} for example.
+%such graphs have been extensively studied \citep{FLS-SODA12,DFHT05,CGP02,Chen07,Chen01}. 
+The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}. So $(g,d)$-map graphs generalise graphs embedded in a surface.
+
+There is a natural drawing of a map graph obtained by positioning each vertex of $G$ inside the corresponding nation and each edge of $G$ as a curve passing through the corresponding vertex of $G_0$.
+It is easily seen that each edge is in at most $\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}}$ crossings; see \citet[Page~4]{dujmovic.eppstein.ea:structure}.
+Thus $G$ is a $(g,\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
+(Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar; this implicitly uses the characterisation of map graphs via half-squares of bipartite graphs.)\ % \citep{??}.)\
+
+\cref{g-k-planar} with $k=\frac14 (d-2)^2$ then implies:
+
+\begin{prop}
+\label{MapGraphs}
+Every $(g,d)$-map graph has a layered $H$-partition of layered width at most 
+%$$\max\{g(\tfrac12 d^4 - 4 d^3 + 15 d^2 - 28 d + 24), \, \tfrac{3}{2} d^4 - 12 d^3 + 45 d^2  - 84 d + 72 \}$$ 
+$$\max\{\tfrac12 gd^4 + O(gd^3), \, \tfrac{3}{2}d^4 +O(d^3) \}$$ 
+in which $H$ has treewidth at most 
+%$$\tfrac{1}{384} d^6 - \tfrac{1}{32} d^5  +  \tfrac{1}{4} d^4 - \tfrac{7}{6} d^3 + \tfrac{95}{24} d^2 - \tfrac{47}{6} d + 9.$$
+$\tfrac{1}{384} d^6 + O(d^5) $.
+\end{prop}
+
+%that every $(g,d)$-map graph $G$ has queue-number  $$\qn(G)\le \max\{\tfrac32 gd^4 + O(gd^3), \, \tfrac{9}{2}d^4 +O(d^3) \} \,  2^{d^6/384 + O(d^5)} .$$ previous bounds obtained by \citet{dujmovic.joret.ea:planar} previously proved the bound of  $O(g^{d^2/4+2})$.
+Our result provides a better bound when $g\gg d$.
+For example, with $d=4$, every $(g,4)$-
+
+\cref{MapGraphs,qn} imply an upper bound on the queue-number of $(g,d)$-map graphs that is better than the previous bound obtained by \citet{dujmovic.joret.ea:planar} when $g\gg d$.
+For example, since every $(g,4)$-map graph is $(g,1)$-planar, \cref{qn,1-planar,g-1-planar} imply that every 4-map graph has queue-number at most $495$, and every $(g,4)$-map graph has queue-number at most $2232 g$.
+
+\cref{MapGraphs,non-repetitive} imply that every $(g,d)$-map graph $G$ has non-repetitive chromatic number 
+$$\pi(G)\le\max\{(\tfrac{1}{2}+o(1)) g(d-2)^4, (\tfrac{3}{2}+o(1)) (d-2)^4\} 4^{ (\frac{1}{384}+o(1))(d-2)^6+1}.$$
+This is the first known constant upper bound (for fixed $g,d$).
+
+\cref{p-centered,MapGraphs} imply that every $(g,d)$-map graph 
+satisfies $\chi_p(G)\le \ell (p+1)(t+1)\binom{p+t}{t}$ where $t:=  \tfrac{1}{384} d^6 + O(d^5)$.
+These are the first bounds on $\chi_p(G)$ that are polynomial in $p$ for $(g,d)$-map graphs with $d$ fixed.
+
+%%%%%%%
+\subsection{String Graphs}
+
+A \emph{string graph} is the intersection graph of a set of curves in the plane with no three curves meeting at a single point; see  \cite{PachToth-DCG02,FP10,FP14} for example.
+For an integer $\delta\geq 2$, if each curve is in at most $\delta$ intersections with other curves, then the corresponding string graph is called a \emph{$\delta$-string graph}.
+A \emph{$(g,\delta)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
+
+%For a surface $\Sigma$ and a graph $G$, a \emph{$\Sigma$-string representation} $f$ of $G$ is obtained by representing each vertex $v\in V(G)$ as a curve $f(v)$ in $\Sigma$ so that, for any pair of vertices $v,w\in V(G)$, the curves $f(v)$ and $f(w)$ intersect if and only if $vw\in E(G)$.
+A $\Sigma$-string representation is called a $(\Sigma,\delta)$-string representation if each curve is in at most $\delta$ intersections with other curves.
+A graph $G$ is a \emph{$(g,\delta)$-string graph} if it has a $(\Sigma,\delta)$-string representation where $\Sigma$ is a surface of Euler genus $g$.
 
 \begin{lem}
-  Every $(g, d)$-map graph is $(g, \floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
+\label{StringGraph}
+Every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar. 
+%\textcolor{red}{Eliminate this lemma, since \cref{DrawG} shows the same result via a result of \citet{dujmovic.joret.ea:planar}}
 \end{lem}
 
-\subsubsection{String Graphs}
-
-For a surface $S$ and a graph $G$, an \emph{$S$-string representation} $f$ of $G$ is obtained by representing each vertex of $v\in V(G)$ as a curve $f(v)$ in $S$ so that, for any pair of vertices $v,w\in V(G)$, the curves $f(v)$ and $f(w)$ intersect if and only if $vw\in E(G)$.
-If $G$ has maximum degree $\delta$, then an $S$-representation of $G$ is also called an $(S,\delta)$-string representation.
-A graph $G$ is a \emph{$(g,\delta)$-string graph} if it has an $(S,\delta)$-string representation where $S$ is a surface of Euler genus $g$.
-
-\begin{lem}
-  Every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.
-\end{lem}
-
 \begin{proof}
-  Consider some $(S,\delta)$-string representation, $f$, of $G$.
-We will describe an embedding of $G$ into $S$ that is an $(S,k)$-plane graph.
-Each vertex $v\in v(G)$ is embedded at $f(v)(0)$.
+Consider some $(\Sigma,\delta)$-string representation, $f$, of $G$.
+We will describe an embedding of $G$ into $\Sigma$ that is an $(\Sigma,k)$-plane graph.
+Each vertex $v\in V(G)$ is embedded at $f(v)(0)$.
 For each edge $vw\in E(G)$ there is an intersection point $p$ between $f(v)$ and $f(w)$, i.e., $f(v)(t_0)=f(w)(t_1)=p$ for some $t_0,t_1\in[0,1]$.
-Embed the edge $vw$ very close to the curve whose image is $\{f(v)(t): 0\le t\le t_1\}\cup f(w)(t): 0 \le t\le t_2\}$.
+Embed the edge $vw$ very close to the curve whose image is $\{f(v)(t): 0\le t\le t_1\}\cup \{ f(w)(t): 0 \le t\le t_2\}$.
 The embedding of any edge $vw\in E(G)$ is contained in $f(v)$ and $f(w)$.
@@
 
-\section{$k$-Nearest-Neighbour Graphs}
-
-In this section, we show that $k$-Nearest neighbour graphs of point in the plane are $O(k^2)$-planar.
+\cref{StringGraph,g-k-planar} with $k=2\delta^2$ implies:
+
+\begin{prop}
+Every $(g,\delta)$-string graph has a layered $H$-partition of layered width at most 
+$\max\{(32+o(1))g\delta^4, (96+o(1))\delta^4\}$ in which $H$ has treewidth at most $(\frac{4}{3}+o(1)) \delta^6$.
+\end{prop}
+
+%A \emph{string graph} is the intersection graph of a set of curves in the plane with no three curves meeting at a single point \cite{PachToth-DCG02,SS-JCSS04,SSS-JCSS03,Krat-JCTB91,FP10,FP14}.
+For an integer $k\geq 2$, if each curve is in at most $k$ intersections with other curves, then the corresponding string graph is called a \emph{$k$-string graph}.
+A \emph{$(g,k)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
+
+% from previous paper: For all integers $g\geq 0$ and $k\geq 2$, every $(g,k)$-string graph has queue-number at most $2(40g+490)^{2k+1}$.
+
+%%%%%%%%%%
+\subsection{Powers of Bounded Degree Graphs}
+
+The \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
+\cref{DrawG,g-k-planar,qn,p-centered} immediately imply the following result:
+
+\begin{prop}
+For all integers $k\geq 1$ and $g\geq 0$, if $G$ is a graph with Euler genus $g$ and maximum degree $\Delta$, then:
+\begin{compactitem}
+%%%%%%
+\item $G^k$ is $(g, (k+1)\Delta^{2(k-1)} )$-planar, 
+%%%%%%
+\item $G^k$ has a partition of layered width $(8+o(1))(g+1)k^2\Delta^{4(k-1)}$ such that the quotient has treewidth at most $t:= (\frac{1}{6}+o(1))k^3\Delta^{6(k-1)}$,
+%%%%%%%
+\item $G^k$ has queue-number at most $g 2^{O( k^3\Delta^{6(k-1)} )}$, 
+\note{DW}{the result of  \citet{dujmovic.joret.ea:planar} mentioned previously implies an upper bound of  $O(g^{ (k+1)\Delta^{2(k-1)}+2})$}
+%%%%%%
+\item $G^k$ has  $p$-centred chromatic number at most $(8+o(1))(g+1)k^2\Delta^{4(k-1)}   (p+1) (t+1)\binom{p+t}{t}$, 
+\end{compactitem}
+\end{prop}
+
+\cref{DrawG,non-repetitive,g-k-planar} also give bounds on the non-repetitive chromatic number of $G^k$, 
+%at most $(8+o(1))(g+1)k^2\Delta^{4(k-1)} 4^{(\frac{1}{6}+o(1))k^3\Delta^{6(k-1)}+1}$.\\
+but the bound is weak, since $G^k$ has maximum degree at most $\Delta^k$, implying that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$ by a result of \citet{DJKW16}.
+
+\subsection{$k$-Nearest-Neighbour Graphs}
+
+In this section, we show that $k$-nearest neighbour graphs of point in the plane are $O(k^2)$-planar.
 For two points $x,y\in\R^2$, let $d_2(x,y)$ denote the Euclidean distance between $x$ and $y$.
-The $k$-nearest-neighbour graph of a point set $P\subset\R^2$ is the geometric graph $G$ with vertex set $V(G)=P$ and edge set
-\[  
-  E= \{vw : v,w\in V(G),\, |\{x\in P\setminus\{v\}: d_2(v,x)\le d_2(v,w)\}|\le k\} \enspace .
-\] 
+The $k$-nearest-neighbour graph of a point set $P\subset\R^2$ is the geometric graph $G$ with vertex set $V(G)=P$, where the edge set is defined as follows.
+For each point $v\in P$, let $N_k(v)$ be the set of $k$ points in $P$ closest to $v$.
+Then $vw\in E(G)$ if and only if $w\in N_k(v)$ or $v\in N_k(w)$.
+% and edge set
+%\[  
+%  E= \{vw : v,w\in V(G),\, |\{x\in P\setminus\{v\}: d_2(v,x)\le d_2(v,w)\}|\le k\} \enspace .
+%\] 
 (The edges of $G$ are straight-line segments joining their endpoints.)
@@
 The following result, which is immediate from \citet[Corollary~4.2.6]{abrego.munroy.ea:on} states that $k$-nearest-neighbour graphs have bounded maximum degree:
-\begin{lem}\lemlabel{k-nn-max-degree}
-  The degree of any vertex in a $k$-nearest-neighbour graph is at most $6k$.
+\begin{lem}
+\label{k-nn-max-degree}
+The degree of every vertex in a $k$-nearest-neighbour graph is at most $6k$.
 \end{lem}
@@
 We make use of the following well-known observation (see e.g., \citet[Lemma~2]{bose.morin.ea:routing}):
-\begin{obs}\obslabel{convex}
+\begin{obs}
+\label{convex}
 If $v_0,\ldots,v_3$ are the vertices of a convex quadrilateral in counterclockwise order then there exists at least one $i\in\{0,\ldots,3\}$ such that $\max\{d_2(v_i,v_{i-1}), d_2(v_i,v_{i+1})\} < d_2(v_{i-1},v_{i+1})$, where subscripts are taken modulo 4.
@@
 
-\begin{lem}\lemlabel{nearest-neighbour}
-  Every $k$-nearest-neighbour graph is $O(k^2)$ planar.
+\begin{lem}
+\label{nearest-neighbour}
+  Every $k$-nearest-neighbour graph is $O(k^2)$-planar.
 \end{lem}
@@
   \end{compactenum}
-  If $xy$ is of Type~C, then \obsref{convex} implies that (without loss of generality) $\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)$.
+  If $xy$ is of Type~C, then \cref{convex} implies that (without loss of generality) $\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)$.
 In this case, we call $x$ a Type~C vertex.
@@
   Next oberve that, if $xy$ is of Type~$v$, then at least one of $xv$ or $yv$ is in $E(G)$ in which case we call $x$ (respectively $y$) a Type~$v$ vertex.
-By \lemref{k-nn-max-degree}, there are at most $6k$ Type~$v$ vertices.
+By \cref{k-nn-max-degree}, there are at most $6k$ Type~$v$ vertices.
 Similarly, there are at most $6k$ Type~$w$ vertices.
 
-  Thus, in all, there are at most $13k-1$ Type~$v$, Type~$w$, and Type~C vertices in total.
-By \lemref{k-nn-max-degree}, each of these vertices is incident on at most $6k$ edges that cross $vw$.
+  Thus, in total, there are at most $13k-1$ Type~$v$, Type~$w$, and Type~C vertices.
+By \cref{k-nn-max-degree}, each of these vertices is incident with at most $6k$ edges that cross $vw$.
 Therefore, there are at most $78k^2-6k$ edges of $G$ that cross $vw$.
@@
 
-We note that \lemref{nearest-neighbour} is tight up to the leading constant:  Every $n$-vertex nearest neighbour graph has at least $kn/2$ at at most $kn$ edges.
-For $k\ge 8$, the Crossing Lemma \cite{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is at $\Omega(k^2)$.
+We note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n$ vertices has at least $kn/2$ edges and at most $kn$ edges.
+For $k\ge 8$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is at $\Omega(k^2)$.
+
+\note{DW}{This actually works for $k\geq 7$, since the crossing lemma holds for graphs with at least $(3+\epsilon)n$ edges, for any fixed $\epsilon$.}
 
@@
 Perhaps $C=3$.
+Note that $C\geq 3$ even for planar graphs \citep{dujmovic.joret.ea:planar}.
 
@@
 \bibliography{k-planar}
-
+\newpage
+\appendix
+
+%%%%%%%%%%%%%%
+\section{Future Work}
+
+\note{DW}{Can we strengthen the result of \citep{MTTV97} to prove a structure theorem, not just a separator theorem.
+
+Then show that $k$-nearest neighbour graphs in any fixed dimension have bounded queue-number, etc.
+
+Conjecture: Every $k$-nearest neighbour graph in $d$ dimensions is a subgraph of $P_1 \boxtimes\dots \boxtimes P_{d-1} \boxtimes H$, where $P_1,\dots,P_{d-1}$ are paths and $H$ has
+treewidth at most $f(k,d)$.\\ This  is true for $d=1$ and $d=2$.
+
+It might be true that ``treewidth at most $f(k,d)$'' can be replaced by ``pathwidth at most $f(k,d)$'', which is true for $d=1$ and open for $d=2$.
+The obvious counterexample, is $T^+:=$ the complete binary tree plus one dominant vertex (since if $T^+$ is a subgraph of $H\boxtimes P$ then $H$ has unbounded pathwidth).
+But $T^+$ has unbounded degree, so it is not  a $k$-nearest neighbour graph for any fixed $k$.
+}
+
+%%%%%%%%%%%%%%%%%%
+\subsection{Unit Disc Graphs}
+
+Let $P$ be a set of points in $\mathbb{R}^d$.
+The \emph{unit disc} graph of $P$ is the graph $G$ with $V(G)=P$ where $vw\in E(G)$ if and only if $\dist(v,w)\leq 1$.
+Here $\dist$ is the Euclidean distance in $\mathbb{R}^d$.
+Let $P^{(d)}$ be the strong product of $d$ paths $P\boxtimes \dots \boxtimes P$ (the $d$-dimensional grid graph with crosses).
+
+\begin{lem}
+\label{UnitDiscGraph}
+Every unit disc graph $G$ in $\mathbb{R}^d$ with no $(k+1)$-clique is a subgraph of 
+$P^{(d)} \boxtimes K_{k \ceil{\sqrt{d}}^d}$.
+\end{lem}
+
+\begin{proof}
+Let $t:= k \ceil{\sqrt{d}}^d$. 
+Let $x_i(v)$ be the $i$-th coordinate of each vertex $v\in V(G)$. 
+For $p_1,\dots,p_d\in\mathbb{Z}$, let $V\langle p_1,\dots,p_d\rangle$ be the set of vertices $v\in V(G)$ such that $x_i(v)\in[p_i,p_i+1)$ for each $i\in[d]$.
+Thus the  $V\langle p_1,\dots,p_d\rangle$ partition $V(G)$.
+Each set $V\langle p_1,\dots,p_d\rangle$ consists of the set of vertices in a particular unit cube.
+Note that the unit cube can be partitioned into $\ceil{\sqrt{d}}^d$  sub-cubes, each with side length at most $\frac{1}{\sqrt{d}}$ and thus with diameter at most 1.
+The set of vertices in a sub-cube with diameter at most 1 is a clique in $G$.
+Thus at most $k$ vertices lie in a single sub-cube, and  $|V\langle p_1,\dots,p_d\rangle | \leq t$.
+Injectively label the vertices in $V\langle p_1,\dots,p_d\rangle$ by $1,2,\dots,t$.
+Map each vertex $v$ in 
+$V\langle p_1,\dots,p_d\rangle$ labelled $\ell(v)$ to the vertex  $(p_1,\dots,p_d,\ell(v))$ of 
+$P^{(d)} \boxtimes K_{t}$. 
+Thus the vertices of $G$ are mapped to distinct vertices of $P^{(d)} \boxtimes K_{t}$. 
+For each edge $vw\in E(G)$, if $v\in V\langle p_1,\dots,p_d\rangle$ and 
+$w\in V\langle q_1,\dots,q_d\rangle$, then $|p_i-q_i|\leq 1$ for each $i\in[d]$, and if $p_i=q_i$ for each $i\in[d]$, then $\ell(v)\neq\ell(w)$. Thus $v$ and $w$ are mapped to adjacent vertices in 
+$P^{(d)} \boxtimes K_{t}$. 
+\end{proof}
+
+\note{DW}{By a volume argument, every covering of the unit cube by balls of diameter 1 uses at least $(\frac{d}{18})^{d/2}$ balls.
+So the $k \ceil{\sqrt{d}}^d$ term in the above theorem cannot be drastically improved.
+}
+
+A result of \citet{Wood-Queue-DMTCS05} implies that $\qn(H \boxtimes P ) \leq 3 \qn(H)+ 1$ for every graph $H$ and path $P$.
+It follows that  $\qn( P^{(d)} \boxtimes H ) \leq \frac{1}{6} (  3^{d+1} (2 \qn(H) + 1 ) -3 )$.
+Since $\qn(K_n)=\floor{\frac{n}{2}}$, by \cref{UnitDiscGraph}, every unit disc graph $G$ in $\mathbb{R}^d$ with no $(k+1)$-clique satisfies
+$$\qn(G) \leq \qn( P^{(d)} \boxtimes K_{k \ceil{\sqrt{d}}^d} ) 
+\leq
+\frac{1}{6} (  3^{d+1} (k \ceil{\sqrt{d}}^d + 1 ) -3 ).$$
+
+\note{DW}{Add non-rep and $p$-centred colouring}
+
+\begin{lem}
+\label{TWhighDimProduct}
+Let $G$ be an $n$-vertex subgraph of $P^{(d)} \boxtimes H$ for some graph $H$. 
+Then 
+$$\tw(G) \leq 2  ( \tw(H)+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1.$$
+$$\pw(G) \leq 2  ( \pw(H)+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1.$$
+\end{lem}
+
+\begin{proof}
+Let $\langle{V^i_0,V^i_1,\dots}\rangle$ be the layering of $G$ determined by the $i$-th dimension. 
+Let $$m:= \ceil*{ \left(\frac{dn}{t+1} \right)^{1/(d+1)} } .$$ 
+For $\alpha\in[0,m-1]$, let $V^{i,\alpha} := \bigcup\{ V^i_j : j\equiv \alpha \pmod{m} \}$. 
+Thus $V^{i,0},\dots,V^{i,m-1}$ is a partition of $V(G)$. 
+Hence $|V^{i,\alpha_i}| \leq \frac{n}{m}$ for some $\alpha_i\in[0,m-1]$. 
+Let $X:= \bigcup_{i=1}^d V^{i,\alpha_i}$. Thus $|X| \leq \frac{dn}{m}$. 
+Note that each component of $G-X$ is a subgraph of $Q^{(d)} \boxtimes H$, where $Q$ is the path on $m-1$ vertices.
+Since $\tw(G)$ equals the maximum treewidth of the connected components of $G$, we have
+$\tw(G) \leq \tw( Q^{(d)} \boxtimes H ) + |X| $. 
+To obtain a tree-decomposition of $Q^{(d)} \boxtimes H$ with width $(t+1) (m-1)^d - 1$, start with an optimal tree-decomposition of $H$, and replace each instance of a vertex of $H$ by the corresponding copy of $Q^{(d)}$.
+Thus
+\begin{align*}
+\tw(G) \leq  (t+1) (m-1)^d - 1 + \frac{dn}{m}
+ \leq 2  (t+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1.
+\end{align*}
+\end{proof}
+
+\cref{UnitDiscGraph,TWhighDimProduct} with $H=P\boxtimes K_{k \ceil{\sqrt{d}}^d}$ (which has pathwidth 
+at most $2k \ceil{\sqrt{d}}^d-1$) imply:
+
+\begin{cor}
+For every unit disc graph $G$ in $\mathbb{R}^d$ on $n$ vertices and with no $(k+1)$-clique,
+%$$\pw(G) \leq 2  k^{ 1/(d+1) }  \ceil{\sqrt{d}}^{d/(d+1) } (dn)^{1- 1/(d+1)}   -  1.$$
+$$\pw(G) \leq 2  (2k)^{ 1/d } \ceil{\sqrt{d}} ((d-1)n)^{1- 1/d}   -  1.$$
+\end{cor}
+
+%K_{k \ceil{\sqrt{d}}^d}$.
 \end{document}

### 2019-07-05 k-planar.tex
First major merge done

@@
 \newcommand{\PRref}[1]{(PR\ref{PR:#1})}
+
 \newcommand{\jlabel}[1]{\label{j:#1}}
 \newcommand{\jref}[1]{(J\ref{j:#1})}
+
+\newcommand{\tlabel}[1]{\label{t:#1}}
+\newcommand{\tref}[1]{(T\ref{t:#1})}
+\newcommand{\ylabel}[1]{\label{y:#1}}
+\newcommand{\yref}[1]{(Y\ref{y:#1})}
 
@@
 \newcommand{\Bagsize}{\ensuremath{\frac{k^3}{6} + \frac{3k^2}{2} + \frac{13k}{3} + 4}}
-\newcommand{\bagsize}{\ensuremath{k^3/6 + 3k^2/2 + 13k/3 + 4}}
-\newcommand{\treewidth}{\ensuremath{k^3/6 + 3k^2/2 + 13k/3 + 3}}
+% \newcommand{\bagsize}{\ensuremath{k^3/6 + 3k^2/2 + 13k/3 + 4}}
+\newcommand{\treewidth}{\ensuremath{\binom{k+4}{3}}-1}
 
@@
 \label{k-planar}
-Every $k$-planar graph has a layered $H$-partition of layered width at most $24k^2 + 60k + 36$ in which $H$ has treewidth at most $\treewidth$.
+Every $k$-planar graph has a layered $H$-partition of layered width at most $15k^2 + 30k$ in which $H$ has treewidth at most $\treewidth$.
 \end{thm}
@@
 The purpose of this section is to prove \cref{k-planar} and related structural results.
-We begin by introducing a tool for studying planar graphs, called a tripod decomposition, that is implicit in previous work on $H$-partitions of planar graphs \cite{dujmovic.joret.ea:planar}.
-We then show how the tripod decomposition of a planarised version of a $k$-planar graph $G$ can be used to find a layered $H$-partition of $G$ of small layered width in which $H$ has small treewidth.
-
-\subsection{Tripod Decompositions}
-\label{TripodDecompositions}
-
-Let $T$ be a tree rooted at some node $r\in V(T)$.
-The \emph{$T$-depth} of a node $x\in V(T)$ is equal to the length of the path from $x$ to $r$ in $T$.
-A path $P$ in $T$ is \emph{vertical} if, for each integer $i$, $P$ contains at most one vertex of $T$-depth $i$.
-The deepest vertex in a vertical path $P$ is called  the \emph{lower endpoint} of $P$ and the shallowest vertex in $P$ is called the \emph{upper endpoint} of $P$.
-For each node $v\in V(T)$, we call each node on the path, in $T$, from $v$ to $r$ (including $v$ itself) a \emph{$T$-ancestor} of $v$.
-If $v'$ is a $T$-ancestor of $v$, then $v$ is a \emph{$T$-descendant} of $v'$.
-
-A \emph{near-triangulation} $N$ is a plane graph whose outer face, $F$, is bounded by a simple cycle in $N$ and each of whose inner faces is bounded by a 3-cycle in $N$.
-A \emph{triangulation} is a near-triangulation whose outer face is bounded by a 3-cycle.
-For any cycle $F$ in a near-triangulation $N$ we use $N\llparenthesis F\rrparenthesis$ to denote the near-triangulation consisting of all the edges and vertices of $N$ that are contained in $F$ or its interior.
-
-For a connected graph $G$ and two vertices $v,w\in V(G)$, we use $\dist_{G}(v,w)$ to denote the length of a shortest path, in $G$, with endpoints $v$ and $w$.
-For a vertex $r\in V(G)$, the \emph{BFS layering} of $G$ from $r$ is the layering $\langle V_0,V_1,\ldots\rangle$ of $G$ where $V_i=\{v\in V(G): \dist_{G}(r,v)=i\}$ for each integer $i\ge 0$.
-A \emph{BFS spanning tree} $T$ of $G$ rooted at $r$ is a spanning tree of $G$ rooted at $r$ where, for every $v\in V(G)$, $\dist_T(v,r)=\dist_G(v,r)$.
-
-For the remainder of this section, $\Delta$ is a triangulation, $r$ is a degree-3 vertex on the outer face of $\Delta$, and $T$ is a BFS spanning tree of $\Delta$ rooted at $r$.
-
-Fix some cycle $F$ in $\Delta$ and let $N=\Delta\llparenthesis F\rrparenthesis$.
-Let $\tau=v_1v_2v_3$ be an inner face of $N$ and consider the three minimal vertical paths $Q_1$, $Q_2$, and $Q_3$ in $T$ such that, for each $i\in\{1,2,3\}$ the lower endpoint of $Q_i$ is $v_i$ and the upper endpoint $v_i'$ of $Q_i$ is in $V(F)$.
-If $Q_1$, $Q_2$, and $Q_3$ are vertex-disjoint, then the graph $\overline{Y}:=\tau\cup Q_1\cup Q_2\cup Q_3$ is called the \emph{closed tripod} in $N$ determined by $\tau$.
-We call $Q_1$, $Q_2$, and $Q_3$ the \emph{legs}, $v_1'$, $v_2'$ and $v_3'$ the \emph{feet}, and $\tau$ the \emph{crotch} of $\overline{Y}$.
-We call $Y:=\overline{Y}-F$ the \emph{(open) tripod} of $N$ determined by $\tau$.
-The graph $M:=\overline{Y}\cup F$ is called a \emph{Mercedes graph} and the (at most three) inner faces of $M$ other than $\tau$ are called the \emph{wedges} of $M$.
-
-Let $r$ be a degree-3 node on the outer face of a triangulation $\Delta$.
-A \emph{tripod decomposition}  $K=K(\Delta, T, r)$ is a rooted tree in which each non-root node $x\in V(K)$ is associated with several objects (refer to \cref{K-node}):
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{cc}
-      \includegraphics{figs/K-node-1} &
-      \includegraphics{figs/K-node-2}
-    \end{tabular}
-  \end{center}
-  \caption{The elements associated with a node $x\in V(K)$.}
-  \label{K-node}
-\end{figure}
-
+We begin with a more general abstraction called $(k,d)$-shortcut systems and prove a structural result, \cref{modern-major-general}, for graphs derived from $(k,d)$-shortcut systems.
+The results on $k$-planar (\cref{k-planar}) and $(g,k)$-planar graphs (\cref{g-k-planar}) can then be obtained from \cref{modern-major-general} and existing results for planar and bounded-genus graphs.
+
+\subsection{Shortcut Systems}
+
+For a rooted tree $T$ and any node $x\in V(T)$, we define $T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}]$ to be the maximal subtree of $T$ rooted at $x$.
+We begin with a fairly standard lemma about normalized tree decompositions:
+
+\begin{lem}\lemlabel{nice-decomposition}
+  For any graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:  
+  \begin{compactenum}[(T1)]
+    % \item\tlabel{rooted}\tlabel{first} $T$ is rooted at some node $x_0\in V(T)$ with $B_{x_0}=\emptyset$.
+    % \item\tlabel{node-set} $V(T)= V(H)$.
+    % \item\tlabel{diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
+    \item\tlabel{subtree-root} For each $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$.
+    \item\tlabel{ancestor-edge}\tlabel{last} For each edge $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
+  \end{compactenum}
+\end{lem}
+
+\begin{proof}
+  Begin with any width-$t$ tree decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some rooted tree $T_0$.
+Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$.
+Let $f:V(H)\to V(T)$ be function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y]$.
+If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.
+Subdivide the edge between $a$ and its parent in $T$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
+This modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
+
+  Next, consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.
+In this case, $B_{a}\subseteq B_{a'}$ where $a'$ is the parent of $a$ since any $x\in B_a\setminus B_{a'}$ would have $f(x)=a$.
+In this case, contract the edge $aa'$ in $T_0$, eliminating the node $a$.
+Repeating this operation will eventually produce a width-$t$ tree-decomposition of $(B_x:x\in V(T_0))$ where $f$ is a bijection between $V(H)$ and $V(T_0)$.
+Renaming each node $a\in V(T_0)$ as $f^{-1}(a)$ gives a tree-decomposition $(B_x:x\in V(T))$ with $V(T)=V(H)$.
+
+  By the definition of $f$, the tree-decomposition $(B_x:x\in V(T))$ satisfies \tref{subtree-root}.
+To see that $(B_x:x\in V(T))$ satisifies \tref{ancestor-edge}, observe that, if $xy\in E(H)$, then at least one of $x$ or $y$ is contained in $B_z$ for every node $z$ on the path from $x$ to $y$ in $T$.
+If neither $x$ nor $y$ is an ancestor of the other, then some node $z$ on this path has $T$-depth less than that of $x$ and $y$.
+If $x\in B_z$ this contradicts the fact that $x$ is the root of $T[x]$.
+If $y\in B_z$ this contradicts the fact that $y$ is the root of $T[y]$.
+\end{proof}
+
+\subsubsection{Generalized Tripod Partitions}
+
+For the remainder of this section, $G$ is a graph and $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{Y}:=(Y_x: x\in V(H)))$ is a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{Y}$ is at most $t$.
+In addition, $(B_x:x\in V(T))$ is a tree-decomposition of $H$ obtained by applying \lemref{nice-decomposition} to $H$.
+The following lemma shows how to interpret these objects as a hierarchical decomposition of $G$.
+
+\begin{clm}\clmlabel{generalized-tripod}
+  For each $x\in V(T)$, let $V_x = \bigcup_{y\in V(T_x)} Y_y$, $F_x=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x=V_x\cup F_x$.  Then, 
+  \begin{compactenum}[(Y1)]
+    \item\ylabel{thickness} $\mathcal{Y}:=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width (with respect to $\mathcal{L}$) at most $\ell$.
+    \item\ylabel{separator} For every $x\in V(T)$, there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$. 
+    \item\ylabel{ancestor-edge} For every $x\in V(T)$, $x$ has $t'\le t$ strict $T$-ancestors $a_1,\ldots,a_{t'}$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
+  \end{compactenum}
+\end{clm}
+
+Before proving \clmref{generalized-tripod} we point out more properties that are immediately implied by it:
+
+\begin{compactenum}[(Y1)]\setcounter{enumi}{3}
+  \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$. 
+  \item\ylabel{containment-i} $V_x\subseteq V_a$ for every $T$-ancestor $a$ of $x$. 
+  \item\ylabel{containment-ii}$N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
+\end{compactenum}
+
+Property~\yref{y-subsets} follows from the fact that $V_x$ is the union of several sets, one of which is $Y_x$.
+Property~\yref{containment-i} follows from the definition of $V_x$ and the fact that $V(T_x)\subseteq V(T_a)$.
+To show Property~\yref{containment-ii} first note that, by \yref{containment-i} it suffices to consider vertices in $w\in F_x=N_x\setminus V_x$.
+By definition, every vertex $w\in F_x$ is adjacent, in $G$, to a vertex $v\in V_x$.
+By \yref{containment-i}, $v\in V_a$, so $w$ is either in $V_a$ or $w$ satisfies the condition $vw\in E(G)$, $v\in V_a$, and $w\not\in V_a$, so $w\in F_a$.
+In either case $w\in N_a=V_a\cup F_a$.
+Note that none of \yref{y-subsets}--\yref{containment-ii} depends on \yref{ancestor-edge} (which is important, since \yref{containment-i} is used to establish \yref{ancestor-edge} in the following proof).
+
+\begin{proof}[Proof of \clmref{generalized-tripod}]
+  Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$ and the fact that $(\mathcal{L},\mathcal{Y})$ is a layered partition of layered width $\ell$.
+Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.
+In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
+
+  To establish Property~\yref{ancestor-edge}, consider some vertex $w\in F_x$.
+Since $w\in F_x$, there exists an edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.
+Since $v\in V_x$, $v\in Y_{x'}=Q_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
+Since $\mathcal{Y}$ is a partition, $w\in Y_{a}=Q_a$ for some $a\not\in V(T_x)$.
+Since $vw\in E(G)$, $x'a\in E(H)$.
+By \tref{ancestor-edge}, one of $a$ or $x'$ is a $T$-ancestor of the other.
+Since $w\in Y_a\subseteq V_a$ and $w\not\in V_x\supseteq V_{x'}$, \yref{containment-i} rules out the possibility that $x'$ is a $T$-ancestor of $a$.
+Therefore, $a$ is a $T$-ancestor of $x$ which is a $T$-ancestor of $x'$.
+Let $z_0,\ldots,z_r$ be the path, in $T$ from $z_0:=x'$ to $z_r:=a$.
+For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $B_{z_i}$.
+However, by \tref{subtree-root} $x'$ is not contained in $B_{x_i}$ for any $i\in\{1,\ldots,r\}$.
+Therefore $a\in B_{x_i}$ for each $i\in\{0,\ldots,r\}$.
+In particular, $a$ is contained in $B_x$.
+  Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $C_x$ contains $x$.
+\end{proof}
+
+\subsubsection{Shortcut Systems}
+
+% For non-negative integers $k$ and $d$, a graph $G$ and distinct vertices $v,w\in V(G)$, a \emph{$(k,d)$ $vw$-shortcut} in $G$ is a path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.
+The \emph{$(k,d)^\text{th}$ power of $G$}, denoted $G^{k,d}$, is the graph with vertex $V(G^{k,d})=V(G)$ in which, for each distinct pair $v,w\in V(G)$, $vw\in E(G^{k,d})$ if and only if $G$ contains a $(k,d)$ $vw$-shortcut.
+
+A collection $\mathcal{P}$ of paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if every path in $\mathcal{P}$ has length at most $k$ and if, for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
+Each path in $P\in\mathcal{P}$ is called a \emph{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \emph{$vw$-shortcut}.
+Given a graph $G$ and a $(k,d)$-shortcut system $\mathcal{P}$ for $G$, we use $G^{\mathcal{P}}$ to denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\mathcal{P}$.
+
+In this section we will show how, given a $(k,d)$-shortcut system $\mathcal{P}$, a layered $H$-partition of $G$ can be used to obtain a layered $I$-partition of $G^{\mathcal{P}}$ where the layered width of the partition does not increase dramatically and the treewidth of $I$ is not much more than the treewidth of $H$.
+
+For convenience, it will be helpful to assume that $\mathcal{P}$ contains a
+length-1 $vw$-shortcut for every edge $vw\in E(G)$.  This assumption has no effect on the resulting graph $G^{\mathcal{P}}$ but avoids special cases in some of our proofs.
+
+As in the rest of this section, $(\mathcal{L},\mathcal{Y})$ denotes a layered $H$-partition of $G$; $T$ is a tree used in a tree-decomposition $(B_x:x\in V(T))$ of $H:=G/\mathcal{Y}$ satisfying the conditions of \lemref{nice-decomposition}; and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ are defined as in \clmref{generalized-tripod}.
+
+For a node $x\in V(T)$, we say that a shortcut $P\in\mathcal{P}$ \emph{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
+We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$ or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
+We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
+
+\begin{clm}
+  For any $v\in V(G)$ there exists a (unique) node $a(v)\in X_v$ such that
+  $a(v)$ is a $T$-ancestor of every node in $X_v$.
+\end{clm}
+
+\begin{proof}
+  Let 
+  \[
+     Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:\mbox{$(v_0,\ldots,v_r)\in\mathcal{P}$, $v\in \{v_0,\ldots,v_r\}$}\}
+  \]
+  Clearly $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
+
+  We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
+If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
+In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$.
+
+  Let $Z_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.  The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.  
+  %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other. 
+  Choose $a(v)\in X_H$ to be any member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.
+Transitivity of the $T$-ancestor relationship, \tref{ancestor-edge}, and connectivity of $H[X_H]$ implies that $a(v)$ is a $T$-ancestor of every node $x\in X_H$, as required.
+\end{proof}
+
+For each $x\in V(T)$, we define the \emph{separator}
+\[  
+     S_x := \{v\in V(G): a(v)= x\} \enspace .
+\]
+Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
+We let $I:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
+(Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $I$.)
+
+\begin{clm}\clmlabel{s-subset}
+  For every $x\in V(T)$, $S_x\subseteq V_x$.
+\end{clm}
+
+\begin{proof}
+  For the sake of contradiction, assume otherwise, so there exists some $v\in S_x\setminus V_x$.
+
+  By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.
+Therefore, $\mathcal{P}$ contains a path $P$ that crosses $x$.
+The path $P$ contains a subpath $v_0,v_1,\ldots,v_{r}$ such that $v=v_0$ and $v_r\in Y_x$.
+Since $v\not\in V_x$ and $v_r\in Y_x\subseteq V_x$, \yref{separator}, implies that $v_i\in F_x$ for some $i\in\{0,\ldots,r-1\}$.
+Now \yref{ancestor-edge} implies $v_i\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
+Therefore, either $v\in Y_a$ or $P$ crosses $a$.
+But this implies that $a(v)$ is an ancestor of $a$, contradicting the assumption that $v\in S_x$.
+\end{proof}
+
+Next we show that $\mathcal{S}$ has small layered width with respect to $\mathcal{L}$:
+
+\begin{clm}\clmlabel{general-width}
+  For each $i\in\{1,\ldots,h\}$ and each $x\in V(I)$, $|S_x\cap V_i|\le d\ell(k^2+k+4)$.  
+\end{clm}
+
+\begin{proof}
+  We will count the number of vertices in $S_x\cap V_i$ by upper-bounding the number of vertices contributed to $S_x\cap V_i$ by each vertex $w\in Y_x$.
+  If $w\in Y_x\cap V_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes only one vertex, itself, to $S_x\cap V_i$.
+
+  Otherwise, consider some path $P\in\mathcal{P}$ that contains $w$ as an internal vertex.
+If $w\in V_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap V_i$.
+If $w\in V_{i-1}\cup V_{i+1}$, then $P$ contributes at most $k$ vertices to $S_x\cap V_i$.
+If $w\in V_{i-j}\cup V_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap V_i$.
+
+  For any $j$, the number of vertices $w\in V_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\mathcal{P}$. Therefore, 
+  \[  |S_x\cap V_i|\le d\ell\cdot\left(k+1 + k + \sum_{j=2}^k 2(k-j)\right) %= d(2k+1) + \sum_{i=1}^{k-2} i
+      = d\ell(k^2 +k + 4) \enspace . \qedhere
+  \]
+\end{proof}
+
+\begin{clm}\clmlabel{i-ancestor}
+  For each edge $xy\in E(I)$, one of $x$ or $y$ is a $T$-ancestor of the other.
+\end{clm}
+
+\begin{proof}
+  Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.
+Since $xy\in E(I)$, $G^\mathcal{P}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
+Since $xy\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.
+By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.
+By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
+By \yref{ancestor-edge} $N_x$ and $V_y$ are also disjoint.
+By \yref{separator} $P$ contains an internal vertex $v'\in F_x$.
+By \yref{ancestor-edge}, $v'\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
+But this implies that $v\in S_{a'}$ for some $T$-ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
+\end{proof}
+
+\begin{clm}\clmlabel{general-bag-size}
+  The graph $I$ has a tree-decomposition in which every bag has size at most $\binom{k+t}{t}$.
+\end{clm}
+
+\begin{proof}
+  For the tree-decomposition $(C_x:x\in V(T))$ of $I$ we use the same tree $T$ used in the tree-decomposition $(B_x:x\in V(T))$ of $H$.
+For each node $x$ of $T$, $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $I$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
+  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(I)$, $a,x'\in C_{x'}$ and the connectivity of $T[a]:=T[\{x\in V(T):a\in C_x]$ is obvious.
+Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $I$.
+What remains is to bound the size, $\max\{|C_x|:x\in V(T)\}$, of the largest bag.
+
+  Consider an arbitrary node $x\in V(T)$ whose ancestors are $x_0,\ldots,x_{r-1},x_r$ where $x_0$ is the root of $T$ and $x_r:=x$.
+To avoid triple-subscripts in what follows, we abuse notation slightly by using $V_i$, $F_i$, and $N_i$,  as shorthands for $V_{x_i}$, $F_{x_i}$ and $N_{x_i}$, respectively.
+
+  If $x_\delta\in C_x$, it is because $x_\delta x'\in E(I)$ for some $T$-descendant $x'$ of $x$.
+This implies $G^{\mathcal{P}}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
+This implies that $\mathcal{P}$ a $vw$-shortcut $P_{vw}$.
+Let $v'$ be the second-last vertex of $P_{vw}$ (so $v'w\in E(G)$).
+
+  Since $w\in S_{\delta}$, at least one of the following is true:
   \begin{enumerate}
-  \item A cycle $F_x$ in $\Delta$ and the near-triangulation $N_x:=\Delta\llparenthesis F_x\rrparenthesis$.
-
-  \item An inner face $\tau_x$ of $N_x$ that determines a closed tripod $\overline{Y}_x$ and an open tripod $Y_x:=\overline{Y}_x-F_x$ in $N_x$, both with crotch $\tau_x$.
-
-  \item The \emph{Mercedes graph} $M_x:=F_x\cup \overline{Y}_x$
+    \item $\mathcal{P}$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
+    \item $w\in Y_\delta$.  In this case, we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$. 
   \end{enumerate}
-
-The root node $r^0_K$ of $K$ is exceptional.
-We define $F_{r^0_K}$ to be the single-vertex graph containing $r$, we define $\overline{Y}_{r^0_K}$ to be the star formed by $r$ and its three neighbours in $\Delta$, and (as usual) $Y_{r^0_K}:=\overline{Y}_{r^0_K}-F_{r^0_K}$.
-The root node $r^0_K$ has exactly one child $r_K$, called the \emph{sub-root} of $K$, for which $F_{r_K}$ is the 3-cycle formed by the neighbours of $r$ in $\Delta$.
-
-The non-root nodes of the tripod decomposition $K$ (including the sub-root $r_K$) satisfy the following conditions:
-\begin{enumerate}[(PR1)]
-  \item \PRlabel{children} For each non-root node $x\in V(K)$ such that $M_x$ has $f\le 3$ wedges $F_{x,1},\ldots,F_{x,f}$, the node $x$ has exactly $f$ children $y_1,\ldots,y_f$ in $K$ and $F_{y_i} = F_{x,i}$ for each $i\in\{1,\ldots,f\}$.
-
-  \item \PRlabel{ancestor-boundary} For each non-root node $x\in V(K)$, there exists a partition of $V(F_x)$ into $g\le 3$ non-empty sets $P_1,\ldots,P_g$ such that, for each $i\in\{1,\ldots,g\}$, $F[P_i]$ is a path that is contained in $Y_a$ for some $K$-ancestor $a$ of $x$.
-
-  \item \PRlabel{tripod-feet} For each $i\in\{1,\ldots,g\}$, at least one vertex in $P_i$ is a foot of the closed tripod $\overline{Y}_x$.
-
-  \item \PRlabel{partition} $\{Y_x : x\in V(K)\}$ is a partition of $V(\Delta)\setminus\{r\}$.
-\end{enumerate}
-
-For any triangulation $\Delta$ in which the outer face has a vertex $r$ of degree 3 and for any BFS spanning tree of $\Delta$ rooted at $r$, there exists a tripod decomposition $K=K(\Delta,T,r)$.
-Indeed, this tree is constructed (albeit implicitly) by \citet[Proof of Lemma~14]{dujmovic.joret.ea:planar}.\footnote{\citet{dujmovic.joret.ea:planar} do not require that $T$ be a BFS spanning tree, only that its root, $r$ has degree 3 in $\Delta$ and in $T$.}
-
-% The following claim is not used until later in the paper. However, it's proof is a helpful exercise in understanding the properties of a PS-Tree.
-% 
-% \begin{clm}\label{strong-ancestor-boundary}
-%   Let $K$ be a PS-tree and
-%   let $y\in V(K)$ be a non-root node with parent $p$ such that $V(F_y\cap Y_a)\neq\emptyset$ for some $K$-ancestor $a\neq p$ of $p$. Then $V(F_p\cap Y_a)\neq\emptyset$.
-% \end{clm}
-% 
-% \begin{proof}
-%   By \propref{children}, $F_y$ is a wedge in the Mercedes graph $M_p$, so $V(F_y)$ can be partitioned into  $P_1'=V(F_y\cap Y_p)$ and $P_{2,3}'=V(F_y\cap F_p)$.
-Let $P_1,\ldots,P_g$ denote the partition of $V(F_p)$ given by \propref{ancestor-boundary}.
-% 
-%   We claim that \PRref{tripod-feet} ensures that $P_{2,3}'\subset P_i\cup P_j$ for some $i,j\in \{1,\ldots,g\}$.
-If $g=2$, this claim is immediate, so assume $g=3$.
-Then, $F_y\cap \overline{Y}_p$ is the path formed by two legs $Q_2$ and $Q_3$ of $\overline{Y}_p$ and one edge $v_2v_3$ of $\tau_p$.
-%   By \PRref{tripod-feet} and the assumption that $q=3$, $Q_2$ has a foot $v_2'\in P_i$ and $Q_3$ has a foot $v_3'\in P_j$, for some $i\neq j$.
-The third foot $v_1'$ of $\overline{Y}_p$ is therefore in $P_\ell$ for the unique $\ell \in\{1,2,3\}\setminus \{i,j\}$.
-It follows that $V(P_\ell\cap F_y)=\emptyset$.
-% 
-%   Setting $P_2'=V(P_i\cup F_y)$ and $P_3'=V(P_j\cup F_y)$ yields a partition $\{P_1',P_2',P_3'\}$ of $V(F_y)$ such that $P_1'\subset V(Y_{p})$,   
-%   $P_2'\subset V(Y_{a_2})$, and $P_3'\subset V(Y_{a_3})$  for some $K$-ancestors $a_2$ and $a_3$ of $a$.
-Therefore $a_2$ and $a_3$ are the only $K$-ancestors $a\neq p$ of $p$, such that $V(F_y\cap Y_a)\neq\emptyset$.
-%   Now, each foot $v_2',v_3'\in V(F_p)$.
-Furthermore, $v_2'\in P_i\subseteq V(Y_{a_2})$ and $v_3'\in P_j\subseteq V(Y_{a_3})$.
-Therefore, $V(F_p\cap Y_{a_2})\supseteq\{v_2'\}\neq\emptyset$ and $V(F_p\cap Y_{a_3})\supseteq\{v_3'\}\neq\emptyset$, as claimed.
-% \end{proof}
-% 
-% Contained in the proof of \clmref{strong-ancestor-boundary} is the following addendum to \propref{ancestor-boundary}:
-% 
-% \begin{enumerate}[(PR1')]\setcounter{enumi}{1}
-%   \item \proplabel{ancestor-boundary-ii} Furthermore, if $g=3$, then  $P_i=V(F_x\cap Y_p)$ where $p$ is the parent of $x$ in $K$, for one $i\in\{1,2,3\}$.
-% \end{enumerate}
-
-% \subsubsection{Layered $H$-Partitions from PS-Trees}
-% \seclabel{k-to-h}
-% 
-% Before continuing, it is worth pointing out how an $H$-partition of $\Delta-\{r\}$ can be obtained from a PS-tree $K=K(\Delta,T,r)$.
-(Note: There is nothing new in this section.
-This is just an alternative presentation of material from \citet{dujmovic.joret.ea:planar}.) Given $K=K(\Delta,T,r)$, we obtain a layered $H$-partition $(\mathcal{L},\mathcal{P})$ as follows.
-% 
-% We use the layering $\mathcal{L}=\{V_0,V_1,\ldots,V_h\}$ where $V_i=\{ v\in V(\Delta)\setminus\{r\}: \dist_{\Delta}(r,v)=i\}$ and $h=\max\{\dist_\Delta(r,v):v\in V(\Delta)\}$.
-The fact that $T$ is a BFS tree ensures that $\mathcal{L}$ is, indeed, a layering of $\Delta-\{r\}$.
-We use the vertex partition $\mathcal{P}=\{ S_x:=V(Y_x) : x\in V(K)\}$.
-\PRref{partition} ensures that $\mathcal{P}$ is, indeed a partition of $V(\Delta)-\{r\}$.
-% 
-% The resulting layered $H$-partition $(\mathcal{L},\mathcal{P})$ has layered width at most 3 because, for each $x\in V(K)$, $S_x$ is made up of vertices from at most 3 vertical paths in $T$.
-These are either the three vertices of $Y_{r_k}$ (if $x=r_K$) or the vertices of the open tripod $Y_x$ whose vertices are included in three vertical paths $Q_1,Q_2,Q_3$ that form the legs of the closed tripod $\overline{Y}_x$.
-Therefore $|S_x\cap V_i|\le 3$.
-% 
-% To show that the quotient graph $H=\Delta/\mathcal{P}$ has small treewidth, we produce a tree-decomposition $\mathcal{K}=\{B_x: x\in V(K)\}$ of $H$ using the tree $K$.
-For each $x\in V(K)$, the subtree $K[x]=K[\{y:x\in B_y\}]$ contains
-% $x$ as well as every node $y$ such that $V(Y_x)\cap V(F_y) \neq\emptyset$.
-% It is straightforward to verify that $K[x]$ is connected for each $x\in V(K)$.
-The fact that $F_x$ is a cycle in $\Delta$ and \propref{ancestor-boundary} ensures that, if $xy\in E(H)$, then one of $x$ or $y$ is a $K$-ancestor of the other.
-\propref{ancestor-boundary} also implies that, for any node $y\in V(K)$, $|B_y|\le 4$, since $B_y$ contains $y$ and at most 3 $K$-ancestors of $y$.
-% 
-% Therefore $(\mathcal{L},\mathcal{P})$ is a layered $H$-partition of $\Delta-\{r\}$ with layered width 3 in which $H$ has treewidth 3.
-Furthermore, for each $P\in\mathcal{P}$, $\Delta[P]$ is connected.
-Since $\Delta$ is planar, this implies that $H$ is planar.\footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
-Repeatedly applying this fact implies that $\Delta/\mathcal{P}$ is planar provided that $\Delta[P]$ is connected for every $P\in\mathcal{P}$.}
+  Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{\delta}$.
+
+  Let $w_0,w_1,\ldots,w_p$ be the path that begins $w_0:=v'$ and then follows the subpath of $P_{ww'}$ that begins at $w_1:=w$ and ends at $w_p:=w''$.
+For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{1,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.
+Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root is non-increasing.
+
+  By definition, $a_0=x_r$.  We claim that $a_p=x_\delta$, i.e., $s_p=\delta$. 
+  To see this, first observe that, for each $i\in\{1,\ldots,p\}$, $w_i\in V_{\delta}$ since, otherwise, an internal vertex of $P_{ww'}$ belong to $F_\delta$, which would imply (by \yref{ancestor-edge}) that $w\in S_{\delta'}$ for some $\delta' < \delta$, contradicting the assumption that $w\in S_\delta$.
+Therefore $s_p\ge\delta$.
+To see that $s_p<\delta-1$,
+  observe that either $w=w''\in Y_\delta$ or $P_{ww'}$ contains an internal vertex $w''$ in $Y_\delta$.
+By \yref{thickness} and the defnition of $V_x$, $V_{\delta-1}$ does not contain $w''$, so $s_p<\delta-1$.
+
+  Let $H^+$ denote the supergraph of $H$ with vertex set $V(T)$ and in which $xy\in E(H^+)$ if and only there exists some $z\in V(T)$ such that $x,y\in B_z$. 
+  % Note that, by \tref{ancestor-edge}, $xy\in E(H^+)$ if and only if $x\in B_y$ or $y\in B_x$.
+  We claim that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \emph{lazy walk} in a graph $H$ is a walk in the non-simple graph $H'$ obtained by adding a self loop to each vertex of $H$.} in $H^+$.
+Indeed, if $a_i\neq a_{i+1}$ for some $i\in\{0,\ldots,p-1\}$ then this is precisely because $w_i\in V_{a_i}$ but $w_{i+1}\not\in V_{a_i}$.
+By definition, $w_i\in Y_{a_i'}$ for some $T$-descendant $a_i'$ of $a_i$.
+  By \yref{separator}, $w_{i+1}\in F_{a_i}$ so by \yref{ancestor-edge} $w_{i+1}\in Y_{a_i''}$ for some strict $T$-ancestor $a_i''$ of $a_i$.
+Since $w_iw_{i+1}\in E(G)$, $a_i'a_i''\in E(H)$.
+By \tref{subtree-root}, $a_i''\in B_{a_i''}$ and $a_i''\in B_{a_i'}$.
+Since $a_i$ is on the path from $a_i'$ to $a_i''$ in $T$ this implies that $a_i''\in B_{a_i}$.
+Therefore $a_ia_i''\in E(H^+)$ as claimed.
+
+  Thus, $a_0,\ldots,a_p$ is a lazy walk in $H^+$ of length $p\le k$ where the distance $s_i$ between $a_i$ and the root $x_0$ of $T$ is non-decreasing.
+By removing repeated vertices this gives a path in the directed graph $\overrightarrow{H}^+$ obtained by directing each edge $xy\in E(H^+)$ from its $T$-descendant $x$ towards its $T$-ancestor $y$.
+  Finally, we are in a position to appeal to \cite[Lemma~24]{pilipczuk.siebertz:polynomial-arxiv} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
+\end{proof}
+
+% Together, \clmref{general-width} and \clmref{general-bag-size} imply the following generalization of \thmref{k-planar}:
+
+\begin{thm}\label{modern-major-general}
+  Let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\mathcal{P}$ be a $(k,d)$-shortcut system for $G$.
+Then $G^{\mathcal{P}}$ has a layered $I$-partition of layered width at most $d\ell(k^3+k^2+4k)$ where the treewidth of $H$ is at most $\binom{k+t}{t}-1$.
+\end{thm}
+
+\begin{proof}
+  The proof is almost immediate, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
+In particular, $G^{\mathcal{P}}$ may contain an edge $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.
+To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.
+This increases the thickness given by \clmref{general-width} from $d\ell(k^2+k+4)$ to $d\ell(k^3+k^2+4k)$.
+\end{proof}
 
@@
 This assumption is not critical since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
-Consider the following plane graphs (see \cref{wrapper}(a)):
-\begin{compactenum}
-  \item Add a dummy vertex at each crossing in $G$ to obtain a plane graph and then add edges to this graph to obtain a triangulation $G^+$.
-
-  \item Add a complete graph $K_4$ that contains $G^{+}$ in one of its inner faces, $F_0$, and then add edges to the resulting graph to obtain a triangulation $G^{++}$.
-\end{compactenum}
-Observe that, for any edge $vw\in E(G)$, 
-\begin{equation}  
-  \dist_{G^+}(v,w) \le k+1 \enspace ,  \eqlabel{distance-preserving}
-\end{equation}
-since (the curve) $vw$ contains a sequence $W_{vw}$ of at most $k+2$ vertices of $G^+$ that form a walk from $v$ to $w$ in $G^+$.
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c@{\hspace{2cm}}c}
-      \includegraphics{figs/wrapper-1} &
-      %\includegraphics{figs/wrapper-3} &
-      \includegraphics{figs/wrapper-3} \\
-      (a) & (b) 
-    \end{tabular}
-  \end{center}  
-  \caption{Wrapping a $K_4$ around $G^+$: (a)~the graph $G^{++}$
-  %, (b)~the spanning tree $T$ of $G^{++}$, 
-  and (b)~the modified version of $G^{++}$ used in the proof of \cref{cor-k-planar}.}
-  \label{wrapper}
-\end{figure}
-
-Let $r$ be the vertex of the $K_4$ not incident on $F_0$ and let $T$ be a BFS spanning tree of $G^{++}$ rooted at $r$. 
-For each integer $i\ge 0$, let $V_i:=\{v\in V(G^+): \dist_{T}(r,v)=i\}$ and observe that $\{V_2,V_3,\ldots,V_{h}\}$ is a layering of $G^+$ (but not of $G$).
-
-Let $K=K(G^{++},T,r)$ be a tripod decomposition and observe that the sub-root $r_K$ of $K$ has $F_{r_K}=F_0$.  This implies:
-\begin{compactenum}[(PR1)]\setcounter{enumi}{4}
-  \item \PRlabel{all-in}  $G^+\subseteq N_{r_K}-F_{r_K}$.
-\end{compactenum}
-
-For each vertex $z\in V(G^+)$, let $P_z=\{a\in V(K): z\in V(N_a-F_a)\}$.
-By \PRref{all-in}, $P_{z}$ contains the sub-root $r_K$ of $K$.
-By \PRref{children}, any node $x\in P_z$ has at most one of its children in $P_z$ and any node $x'\not\in P_z$ has none of its children in $P_z$.
-Therefore $K[P_z]$ is a vertical path in $K$ that contains $r_K$ (and whose lower endpoint is the unique node $x\in V(K)$ such that $z\in Y_x$).
-We are about to repeatedly use the fact that the common intersection of two or more vertical paths is a vertical path.
-
-For each edge $vw\in E(G)$, consider the walk $W_{vw}$ in $G^+$ that corresponds to the sequence of vertices in $G^+$ encountered while walking along the curve $vw$ from $v$ to $w$.
-We define $A_{vw}:=\bigcap_{z\in W_{vw}} P_z$.
-Since $V(W_{vw})\subseteq V(G^+)$, $A_{vw}$ is well-defined.
-For each $z\in V(W_{vw})$, $K[P_{z}]$ is a vertical path in $K$ that contains $r_K$.
-Therefore $K[A_{vw}]$ is a vertical path in $K$ that contains $r_K$.
-For each vertex $v\in V(G)$ define $A_v:=\bigcap_{vw\in E(G)} A_{vw}$.
-Again, $K[A_v]$ is a vertical path in $K$ that contains $r_K$.
-We define $a(v)$ to be the lower endpoint of $K[A_v]$.
-In words, $a(v)$ is the lowest node in $K$ such that the interior of $F_{a(v)}$ contains all the edges of $G$ incident to $v$.
-
-For each node $x\in V(K)$, define the \emph{separator}
-\[
-   S_x := \{v\in V(G): a(v)=x \} \enspace .
-\]
-Clearly $\mathcal{S}:=\{S_x:x\in V(K)\}$ is a partition of $V(G)$.
-We will use $\mathcal{S}$ as the partition that defines our quotient graph $H:=G/\mathcal{S}$ with vertex set $V(H):=V(K)$ and the edge $xy\in E(H)$ if and only if there exists an edge $vw\in E(G)$ with $v\in S_x$ and $w\in S_y$.
-First we argue that this partition has small layered width with respect to the layering $\{V_2,V_3,\ldots\}$ of $G^+$ defined by $T$.
- % (recall that this is a layering of $G^+$, not a layering of $G$).
-
-\begin{clm}\label{chargeback-to-y}
-   For any $x\in V(K)$ and any vertex $v\in S_x$, there exists an edge $vw\in E(G)$ and a vertex $v'\in Y_x\cap V(W_{vw})$ such that $|\dist_T(r,v')-\dist_T(r,v)|\le k+1$.
-\end{clm}
-
-\begin{proof}
-  Since $v\in S_x$, the lower endpoint of $K[A_{vw}]$ is $x$ for some edge $vw\in E(G)$.
-This implies that $W_{vw}\subseteq N_x-F_x$ but $W_{vw} \not\subseteq N_{x'}-F_{x'}$ for every child $x'$ of $x$ in $K$.
-Therefore, \PRref{children} and \PRref{ancestor-boundary} imply that $W_{vw}$ contains a vertex $v'\in Y_x$.
-Since $G$ is $k$-planar, the length of $W_{vw}$ is at most $k+1$.
-Therefore, $\dist_{G^{++}}(v,v') \le k+1$.
-  By the triangle inequality,     
-  \[
-    \dist_{G^{++}}(r,v') \le \dist_{G^{++}}(r,v) + \dist_{G^{++}}(v,v') 
-       \le \dist_{G^{++}}(r,v) + k + 1 \enspace .
-  \]
-  Similarly, $\dist_{G^{++}}(r,v) \le \dist_{G^{++}}(r, v') + k+1$.  Therefore
-  $|\dist_{G^{++}}(r,v)-\dist_{G^{++}}(r,v')| \le k+1$ which establishes the claim, since $T$ is BFS spanning tree of $G^{++}$ rooted at $r$, so $\dist_{T}(r,z)=\dist_{G^{++}}(r,z)$ for all $z\in V(G^+)$.
-\end{proof}
-
-\begin{clm}
-\label{width-of-g-plus}
-  For any $x\in V(H)$, and any integer $i\ge 0$, $|S_x\cap V_i|\le 24k+36$.
-\end{clm}
-
-\begin{proof}
-  Since the vertices of $Y_x$ come from at most three vertical paths in $T$, $|V(Y_x)\cap V_i|\le 3$.
-\cref{chargeback-to-y} shows that, for any $v\in S_x\cap V_i$, there exists $vw\in E(G)$ and $v'\in Y_x\cap V(W_{vw})$ with $\dist_T(v,v')\le k+1$.
-Therefore, $v'\in V(Y_x)\cap V_{j}$ for some $j\in\{i-k-1,\ldots,i+k+1\}$.
-If $v'=v$ (so $v'\in V_i$) then $v'$ contributes only one vertex to $S_x$.
-If $v'\neq v$ then $v'\in V(G^+)\setminus V(G)$ is a point where two edges of $G$ cross and $v'$ contributes at most the four endpoints of these edges to $S_x$.
-In either case, each $v'\in V(Y_x)\cap \bigcup_{j=i-k-1}^{i+k+1} V_j$ contributes at most four vertices to $S_x\cap V_i$, so $|S_x\cap V_i|\le 3\times 4\times (2k+3) = 24k+36$.
-\end{proof}
-
-Next we describe a tree decomposition $(B_x:x\in V(K))$ of $H$.  We deliberately use the same node set $V(H)=V(K)$ here, because the subtree $K[x]:=K[\{z:x\in B_z\}]$ will be rooted at $x$.
-
-\begin{clm}\label{xy-ancestor}
-   For every edge $xy\in V(H)$, one of $x$ or $y$ is a $K$-ancestor of the other.
- \end{clm}
-
- \begin{proof}
-   Since $xy\in E(H)$, there exists an edge $vw\in E(G)$ such that $v\in S_x$ and $w\in S_y$.
-By definition, $x\in A_v\subseteq A_{vw}$ and $y\in A_w\subseteq A_{vw}$.
-Since $K[A_{vw}]$ is a vertical path in $K$, it follows that one of $x$ or $y$ is a $K$-ancestor of the other.
-\end{proof}
-
-The tree decomposition $(B_x:x\in V(K))$ of $H$ is defined as follows: For each edge $xy\in E(H)$ where $y$ is a $K$-ancestor of $x$, we add $y$ to $B_{x'}$ for every node $x'$ on the path from $x$ to $y$ in $K$.
-That this produces a tree-decomposition is due to \cref{xy-ancestor}.
-What remains is to upper bound the width of this decomposition.
-
-\begin{clm}\label{bag-size}
-  The tree decomposition $(B_x: x\in V(K))$ of $H$ contains no bag larger than $\bagsize$.
-\end{clm}
-
-\begin{proof}
-  Fix some node $x\in V(K)$ and consider the contents of $B_x$.  By \cref{xy-ancestor}, $B_x$ contains only $K$-ancestors of $x$ (including $x$ itself).
-
-  Consider the graph $A$ that contains the edges and vertices of $F_{a}$ for all $K$-ancestors $a$ of $x$.
-See \cref{A}(a).
-It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0:=r_K$ is the root of $K$.
-We then cut $F_{a_0}$ into two pieces $F^{\bar{x}}_{a_0}$ and $F^x_{a_0}$ where $F^x_{a_0}$ is the piece that contains $F_x$.
-(The path that makes this cut contains two legs $Q_i$ and $Q_j$ of a tripod $Y$ and one edge $v_iv_j$ of its crotch $\tau=v_1v_2v_3$.)  This process continues by partitioning $F_{a_1}:=F^x_{a_0}$ recursively.
-The process terminates after some number $d$ of iterations, when $F^x_{a_{d-1}}=F_x=F_{a_d}$.
-
-  \begin{figure}
-    \begin{center}
-      \begin{tabular}{cc}
-        \includegraphics{figs/A-1} &
-        \includegraphics{figs/A-3} \\ (a) & (b)
-      \end{tabular}
-    \end{center}
-    \caption{(a)~the graph $A$ and (b)~the walks $W_{vw}$ and $W_{ww'}$ that contain the walk $w=v_0,v_1\ldots,v_{\ell-1},v_\ell=w^*$.}
-    \label{A}
-  \end{figure}
-
-  Refer to \cref{A}(b).
-If $B_x$ contains some $K$-ancestor $a_\delta$, $\delta\in\{0,\ldots,d-1\}$, of $x$, it is because there is an edge $vw\in E(G)$ with $w\in S_{a_\delta}$ and $v\in S_{x'}$ for some $K$-descendant $x'$ of $x$.
-The edge $vw$ corresponds to a walk $W_{vw}$ in $G^+$.
-The walk $W_{vw}$ is contained in $N_x-F_x$, otherwise $x\not\in A_{vw}$ which would imply that $x'\not\in A_{vw}$, $x'\not\in A_v$, and $a(v)\neq x'$, contradicting the fact that $v\in S_{x'}$.
-
-  However, $w\in S_{a_\delta}$, so there exists an edge $ww'\in E(G)$ such that $W_{ww'}\subseteq N_{a_\delta}-F_{a_\delta}$ but $W_{ww'}\not\subseteq N_{a_{\delta+1}}-F_{a_{\delta+1}}$.
-Therefore $W_{ww'}$ must contain a first vertex $w^*\in V(F_{a_{\delta+1}}-F_{a_\delta})\subseteq Y_{a_\delta}$.
-
-  Summarizing, we have a walk $w=v_0,v_1\ldots,v_{\ell-1},v_\ell=w^*$ in $G^+$ such that
-  \begin{compactenum}[(i)]
-    \item $v_0\in V(N_x-F_x)$;
-
-    \item $v_0,\ldots,v_{\ell-1}\in V(N_{a_{\delta+1}}-F_{a_{\delta+1}})$;
-
-    \item $v_\ell\in V(F_{a_{\delta+1}}-F_{a_\delta})$; and
-
-    \item $\ell\le k+1$.
-  \end{compactenum}
-  We call any walk with properties (i)--(iv) a \emph{$\delta$-walk}.
-Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $\delta\in\{0,\ldots,d\}$ such that there exists a $\delta$-walk.\footnote{To account for the fact that $B_x$ contains $x$ we count $\delta=d$.
-A $d$-walk corresponds, for example, to a length-0 walk that begins and ends at any vertex $v_0\in V(N_x-F_x)$.}  This is our goal for the remainder of this proof.
-
-  In order to avoid excessive levels of subscripts, we introduce the following abbreviated notation: For $i\in\{0,\ldots,d\}$, we use $F_i$, $N_i$, $Y_i$, and $M_i$ as shorthands for $F_{a_i}$, $N_{a_i}$, $Y_{a_i}$, and $M_{a_i}$, respectively.
-
-  Let $d_0:=d$, and $d_1:=\min\{i: V(F_d\cap Y_i)\neq\emptyset \}$.  By \PRref{ancestor-boundary}, any vertex $z\in V(F_d)$ is contained in 
-  $Y_j$ for some $j\in \{d_1,d_1+1,\ldots,d-1\}$.  Therefore, $F_d$ is contained in the interior of  $F_{d_1}$, that is, $N_{d_0}\subseteq N_{d_1}-F_{d_1}$.
-
-  Refer to \cref{d0d1}.  We will now prove that 
-  \begin{equation}
-     |\{\delta\in\{d_1+1,\ldots,d_0\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+3}{2} \enspace . 
-     \eqlabel{one-round}
-  \end{equation}
-  %
-  \begin{figure}
-    \begin{center}
-      \includegraphics{figs/d0d1}
-    \end{center}
-    \caption{$F_{d_1},\ldots,F_{d_0}$ and the resulting graph $I$.} 
-    \label{d0d1}
-  \end{figure}
-
-  To prove \eqref{one-round}, consider the graph $I$, illustrated in \cref{d0d1}, with vertex set $V(I)=\{d_1+1,\ldots,d_0\}$ and $ij\in E(I)$ if and only $i=j$ or $|F_i\cap Y_j|\neq\emptyset$.
-Any $\delta$-walk $v_0,\ldots,v_\ell$ with $\delta\in V(I)$ corresponds to a walk $\rho_0,\ldots,\rho_{\ell}$ in $I$, where $\rho_i:=\max\{j\in\{0,\ldots,d\} : v_i\in N_{j}-F_j\}$.
-To see why this is true, consider $\rho_{t}= i$ and $\rho_{t+1}=j$ for some $t\in\{0,\ldots,\ell-1\}$:
-  \begin{compactenum}
-    \item If $i=j$, then $ij\in E(I)$ since $ii\in E(I)$ for each $i\in V(I)$.
-
-    \item If $j\neq i$ then assume, without loss of generality, that $j < i$ so that $a_j$ is a $K$-ancestor of $a_i$.
-Now $v_t\in V(N_i-F_i)$, but $v_{t+1}\not\in V(N_i-F_i)$ otherwise we would have $j\ge i$.
-Therefore, $v_{t+1}\in V(F_i)$.
-By \PRref{ancestor-boundary}, $v_{t+1}\in V(Y_{j'})$ for some $j'< i$ and therefore $ij'\in E(I)$.
-By \PRref{children}, $v_{t+1}\not\in V(N_{j''}-F_{j''})$ for any $K$-descedant $j''>j'$ of $j'$.
-Therefore $j'=\max\{j : v_{t+1}\in V(N_j-F_j)\}$.
-By definition $j'=\rho(v_{t+1})=j$, so the edge $ij=ij'\in E(I)$.
-  \end{compactenum}
-
-  Properties~(i)--(iv) of $\delta$-walks imply that any walk $\rho_0,\ldots,\rho_\ell$ obtained this way 
-  \begin{compactenum}[(i')]
-    \item begins at $\rho_0=d_0$ 
-    \item is contained in $I[\{\delta,\ldots,d_0\}]$, 
-    \item ends at $\rho_\ell=\delta$
-    \item has length $\ell \le k+1$.
-  \end{compactenum}
-  We will call any walk in $I$ that satisfies (i')--(iv') a \emph{$\delta$-walk}.
-
-  The edges in $E(I)$ can be partitioned into two classes:  A set $B(I):=\{ij\in E(I): |i-j|\le 1\}$ of \emph{boring edges} and a set $J(I):=\{ij\in E(I): |i-j|>1\}$ of \emph{jump edges}.
-For a jump edge $ij$ with $i>j$, $j$ is called a \emph{landing} and $i$ is called a \emph{takeoff}.
-We will show that the jump edges of $I$ satisfy the following properties:
-  \begin{enumerate}[(J1)]
-    \item \jlabel{single} Every $i\in\{d_1+1,\ldots,d_0\}$ is a takeoff for at most one jump edge.
-
-    \item \jlabel{nested} For every jump edge $ij\in E(I)$ with takeoff $i$ and landing $j$, and for every $q\in\{j+2,\ldots,i\}$, if $q$ is a takeoff for some jump edge, then this edge is $qj$.
+
+Let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$.
+In this way, each edge $vw\in E(G)$ corresponds naturally to a path $P_{vw}$ of length at most $k+1$ in $G_0$.
+Take $\mathcal{P}=\{P_{vw}: vw\in E(G)\}$ and observe that $\mathcal{P}$ is a $(k+1,2)$-shortcut for $G_0$ and that $G_0^{\mathcal{P}}\supseteq G$.
+Specifically, $G_0^{\mathcal{P}}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.
+
+Since $G_0$ is planar, the second part of \cref{PlanarPartition} implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
+Applying \cref{modern-major-general} to $G_0$ and $\mathcal{P}$ immediately implies that $G$ (an arbitrary $k$-planar graph) has a layered $H$-partition of layered width $3(k^3+k^2+4k)$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
+
+We can reduce the layered width of the $H$-partition of $G$ from $O(k^3)$ to $O(k^2)$ by observing that the dummy vertices in $V(G_0)\setminus V(G)$ do not contribute to the layered width of this partition.
+In this setting, the proof of \cref{clm:general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $V_i\cap Y_x$.
+More precisely, a path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$
+\begin{enumerate}
+  \item Contributes at most 2 vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}$;
+  \item Contributes at most 1 vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2},\ldots,k+1\}$; 
+  \item Contributes no vertices to $S_x\cap V_j$ for $j > k+1$.
 \end{enumerate}
-  Property \jref{nested} can be interpreted informally as saying that jump edges can cross, but barely:  If $i_1j_1$ and $i_2j_2$ are jump edges with $i_1> i_2>j_1>j_2$, then $j_1=i_2-1$.
-
-  We now prove that jump edges satisfy \jref{single}.
-  First we claim that, for each $i\in\{d_1+1,\ldots,d_0\}$, $F_{i}$ contains a vertex of $Y_{d_1}$.
-For $i=d_0$, this is true by definition so $F_{d_0}$ contains some vertex $v\in V(Y_{d_1})$.
-For $i>0$, we have $F_{d_0}-F_{i}\subseteq N_i-F_i$, but $N_i-F_i$ contains no vertex of $Y_{d_1}$.
-It must therefore be the case that $v\in V(F_i)$, otherwise $F_{d_0}-F_i$ would contain $v$.
-
-  Now, consider the parent $a_{i-1}$ of $a_i$ in $K$.
-By \PRref{children}, $F_i$ is a wedge of the Mercedes graph $M_{i-1}$.
-Let $P_1,\ldots,P_g$ be the partition of $V(F_{i-1})$ guaranteed by \PRref{ancestor-boundary}.
-  By \PRref{tripod-feet}, $V(F_i) = V(F_i)\cap(P_\alpha \cup P_j\cup V(Y_{i-1}))$ for some $\alpha,j\in \{1,\ldots,g\}$.
-Since $F_i$ contains $v\in V(Y_{d_1})$, at least one of $\alpha$ or $j$, say $\alpha$, is equal $d_1$.
-Since $d_1\not\in V(I)$ and the (possible) edge $(i,i+1)$ is boring, the only (possible) jump edge with $i$ as takeoff is $ij$.
-This establishes \jref{single}.
-
-  To establish that jump edges satisfy \jref{nested} we can use the same argument used to establish that each $F_i$ contains a vertex of $Y_{d_1}$ to establish that each $F_q$ contains a vertex of $Y_j$ and therefore $qj\in E(I)$.
-Since $q \ge j+2$, $qj$ is a jump edge and, by \jref{single}, it is the only jump edge with takeoff $q$.
-
-  It will be helpful to consider an \emph{edge maximal} version, $I^+$, of $I$ that we now define.
-Let $j_0=d$ and let $j_1>j_2>\cdots> j_p$ denote the sequence of landings, in decreasing order.
-Then \jref{single} and \jref{nested} imply that
-  \[
-      J(I) \subseteq J^+(I) :=
-         \bigcup_{t=1}^p \{ ij_t : i\in\{ j_t,\ldots,j_{t-1}+1 \}\} \enspace . 
-  \]
-  and clearly $B(I)\subseteq B^+(I) := \{(i,j): i\in\{d_1+2,\ldots,d_0-1\},\, j\in\{i-1,i,i+1\}\}$.
-  Any walk in $I$ is also a walk in the supergraph $I^+$ with $V(I^+)=V(I)$ and $E(I^+)=B^+(I)\cup J^+(I)$, so we can focus our attention on $\delta$-walks in $I^+$.
-
-  We are almost done proving \eqref{one-round}.
-Refer to \cref{fig-one-round}.
-For any landing $j_t$, $t>0$, $I^+[\{j_t,\ldots,d_0\}]$ contains the path $d_0,j_1,\ldots,j_t$ of length $t$.
-A few moments consideration will also convince the reader that $d_0,j_1,\ldots,j_t$ is a shortest path, in $I^+[\{j_{t+1}-1,\ldots,d_0\}]$, from $d_0$ to $j_t$.
-Therefore, by (iv') no $\delta$-walk contains $j_t$ for any $t> k+1$.
-
-  \begin{figure}
-    \begin{center}
-      \includegraphics{figs/one-round}
-    \end{center}
-    \caption{The graph $I^+[\{j_{4}+1,\ldots,d_0\}]$.  The shortest path $d_0,j_1,\ldots,j_3$ is highlighted.}
-    \label{fig-one-round}
-  \end{figure}
-
-  Next, observe that every path in $I^+[\{j_{t+1}+1,\ldots,d_0\}]$ from $d_0$ to some vertex  $j_{t}+i$, $i\ge 0$, contains $j_t$.
-Furthermore, the only edges of $I^+[\{j_{t+1}+1,\ldots,d_0\}]$ incident to vertices $j_t+i$, $i>0$, are boring edges.
-Therefore, for any $i\ge 0$, the shortest path from $d_0$ to $j_t+i$ in $I^+[\{j_{t+1}+1,\ldots,d_0\}]$ has length $t+i$.
-It follows that there are at most $k+2-t$ values $\delta\in\{j_{t+1},\ldots,j_t\}$ such that there exists a $\delta$-walk.
-
-  Therefore, the total number of $\delta\in\{d_1+1,\ldots,d_0\}$ such that there exists a $\delta$-walk is at most
-  \[
-     \sum_{t=0}^{k+1} (k+2-t) = \sum_{t=1}^{k+2} t = \binom{k+3}{2}
-  \]
-  This establishes \eqref{one-round}.  An illustration of this analysis, showing the case $k=1$ is shown in \cref{one-round-example}.
-  \begin{figure}[htbp]
-    \begin{center}
-      \includegraphics{figs/one-round-example}
-    \end{center}
-    \caption{The set of $6=\binom{k+3}{2}$ vertices in $I^+$ reachable by paths of walks of length at most $k+1$ (for $k=1$).}
-    \label{one-round-example}
-  \end{figure}
-
-  Refer to \cref{fig-nested}.
-For $s>1$, let $d_s:=\min\{i: V(F_{d_s}\cap Y_i)\neq\emptyset \}$.
-Consider some $\delta$-walk $v_0,\ldots,v_\ell$ for $\delta\in \{ d_{s+1},\ldots,d_{s}-1\}$.
-By (i) and (ii), $v_0\in V(N_{d_0}-F_{d_0})\subseteq V(N_{d_s}-F_{d_s})$ and $v_\ell\in F_{\delta+1}-F_\delta$.
-Therefore, there exists some maximum value $i$ such that $v_i\in V(N_{d_{s}}-F_{d_{s}})$.
-(In words, $i$ is the last time the walk is in the interior of $F_{d_s}$ before it leaves forever.)
-  Observe that $i\ge s$ because $v_0,\ldots,v_i$ contains $v_0\in N_{d_0}-F_{d_0}$ as well as a vertex from each of $F_{d_0},\ldots,F_{d_{s-1}}$.
-
-  \begin{figure}
-    \begin{center}
-      \includegraphics{figs/nested}
-    \end{center}
-    \caption{The nested cycles $F_{d_0}, F_{d_1},F_{d_2}\ldots$.}
-    \label{fig-nested}
-  \end{figure}
-
-  Now, consider the subwalk $v_i,\ldots,v_\ell$.  This subwalk obeys Properties~(i)--(iv) with respect to $F_{d_s}$.  Specifically:
-  \begin{compactenum}[(i)]
-    \item $v_i\in V(N_{d_s}-F_{d_s})$;
-
-    \item $v_i,\ldots,v_{\ell-1}\in V(N_{a_{\delta+1}}-F_{a_{\delta+1}})$;
-
-    \item $v_\ell\in V(F_{a_{\delta+1}}-F_{a_\delta})$; and
-
-    \item $\ell\le k+1$, so $\ell-i \le \ell-s \le k+1-s$.
-  \end{compactenum}
-  Exactly the same analysis used to establish \eqref{one-round} shows that
-  \begin{equation}
-     |\{\delta\in\{d_{s+1}+1,\ldots,d_s\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+3-s}{2} \enspace . \eqlabel{one-round-s}
-  \end{equation}
-
-  We conclude the proof with 
-  \begin{align*}
-    |B_x| & \le |\{\delta\in\{0,\ldots,d\}:  \mbox{$\exists$ a $\delta$-walk} \}| \\
-      & = \sum_{s=0}^{k+1}|\{\delta\in\{d_{s+1}+1,\ldots,d_s\}:  \mbox{$\exists$ a $\delta$-walk} \}| \\
-      & \le\sum_{s=0}^{k+1} \binom{k+3-s}{2} \\
-      & = \sum_{s=2}^{k+3} \binom{s}{2} \\
-      & = \frac{1}{6}(k+2)(k+3)(k+4) \\
-      & = \frac{k^3}{6} + \frac{3k^2}{2} + \frac{13k}{3} + 4  \enspace . \qedhere
-  \end{align*}
-\end{proof}
-
-We now have all the pieces needed to complete the proof of \cref{k-planar}.
-
-\begin{proof}[Proof of \cref{k-planar}]
-  We use the layered $H$-partition $(\mathcal{L}, \mathcal{S})$, where $\mathcal{S}=\{S_x:x\in V(K)\}$ and $\mathcal{L}=\langle V_i': i=0,1,2\ldots\rangle$ where $V_i' = V_{(k+1)i}\cup\cdots\cup V_{(k+1)(i+1)-1}$.
-\Eqref{distance-preserving} shows that $\mathcal{L}$ is indeed a layering of $G$.
-
-  By \cref{bag-size}, $\mathcal{S}$ produces a quotient graph $H=G/\mathcal{S}$ of treewidth at most $\treewidth$.
-By \cref{width-of-g-plus}, for each $x\in V(H)$ and each integer $i$, $|B_x\cap V_i|\le 24k+36$, so $|B_x\cap V_i'|\le (k+1)(24k + 36) = 24k^2+60k+36$.
-\end{proof}
-
-We conclude this section with a small extension of \cref{k-planar} that will be useful in \cref{sec-g-k-planar}.
-
-\begin{cor}
-\label{cor-k-planar}
-Let $G$ be a $k$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
-Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \cref{k-planar} and such that $V_i' := \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\}$ for each integer $i\ge 0$.
-\end{cor}
-
-\begin{proof}
-  Without loss of generality, assume that $r_0$ is on the outer face of $G^+$ and let $v_1$ and $v_2$ be the other two vertices of the outer face of $G^+$.
-  The proof of \cref{k-planar} begins with the graph $G^+$ and then creates a graph $G^{++}$ consisting of a $K_4$ containing $G^+$ in one of its inner faces, $F_0$, and six edges required to complete $G^{++}$ to a triangulation.
-Instead, modify $G^{++}$ by subdividing the edges joining $F_0$ to $v_1$ and $v_2$, as shown in \cref{wrapper}(b).
-In this modified version of $G^{++}$, we have $\dist_{G^++}(r,v)=\dist_{T}(r,v)=\dist_{G^+}(r_0,v)+2$
-  which leads to the layering $\langle V_0,V_1,\ldots\rangle$ of $G^+$ where $V_i=\{v\in V(G^+): \dist_{T}(r,v) = i\}$.
-
-  The rest of proof remains unchanged until the point where we derive the layering $\mathcal{L}=\langle V_0',V_1',\ldots\rangle$ of $G$.  At this point, we set
-  \[  
-    V_i' := \{v\in V(G): \floor{\dist_{T}(r,v)-2}=i\} 
-         = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\} \enspace ,
-  \]
-  for all integers $i\ge 0$.  This layering satisfies the conditions of the corollary.
-\end{proof}
+Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields $|S_x\cap Y_i|\le d\ell(5k/2+5)=15k+30$.
+With this change, the layered width of the partition given by \cref{modern-major-general} becomes $15k^2+30k$.
+This establishes \cref{k-planar}.
 
@@
 
-Exactly the same argument used to establish \cref{cor-k-planar} as a consequence of \cref{k-planar} establishes the following result as a consequence of \cref{1-planar}:
-
-\begin{cor}
-\label{cor-1-planar}
-Let $G$ be a $1$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
+% Exactly the same argument used to establish \cref{cor-k-planar} as a consequence of \cref{k-planar} establishes the following result as a consequence of \cref{1-planar}:
+% 
+% \begin{cor}
+% \label{cor-1-planar}
+% Let $G$ be a $1$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
 Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \cref{1-planar} and such that $V_i' = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/2}=i\}$ for each integer $i\ge 0$.
-\end{cor}
+% \end{cor}
 
@@
 A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
-We prove the following generalisation of \cref{k-planar}.
+
+\citet[Theorem~22]{dujmovic.joret.ea:planar} show that every genus-$g$ graph has a layered $H$-partition of layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most 4.
+Using this result and proceeding exactly as in \cref{sec-k-planar} establishes the following generalisation of \cref{k-planar}.
 
@@
 Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most 
-$\max\{8gk^2 + 12gk + 4g, 24k^2 + 36k + 12\}$ in which $H$ has treewidth at most $\treewidth$.
+$\max\{2g,3\}(5k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}$.
 \end{thm}
 
-Again, in the case $k=1$ we obtain better constants.
-
+\note{PM}{I remove the $(g,1)$-specialization. It requires cutting surfaces.  For reference, the statement was:
 \begin{thm}
@@
 \end{thm}
-
-The proofs of these results make use of the following lemma due to \citet[Lemma~21]{dujmovic.joret.ea:planar}:
-
-\begin{lem}[\citep{dujmovic.joret.ea:planar}] 
-\label{make-planar}
-Let $G^+$ be a $(\Sigma,0)$-plane graph where $\Sigma$ is a surface with Euler genus $g$. 
-For each $r\in V(G^+)$ there is a vertex set $Z\subseteq V(G^+)$ such that $G^+-V(Z)$ is planar and there is a connected planar graph $G^*$ containing  $G^+-Z$ as a subgraph with a vertex $r^*\in V(G^*)$ such that, for every integer $i\ge 0$,
-\[  \{v\in V(G^*): \dist_{G^*}(r^*,v) = i\} \supseteq \{v\in V(G^+): \dist_{G^+}(r,v) = i\} \]
-and
-\[ |\{v\in Z: \dist_{G^+}(r,v) = i\}| \le 2g \enspace . \]
-\end{lem}
-
-\begin{proof}[Proof of \cref{g-k-planar,g-1-planar}]
-  Let $G$ be an $(\Sigma,k)$-plane graph where $\Sigma$ is a surface of Euler genus $g$.
+}
+% Again, in the case $k=1$ we obtain better constants. 
+% 
+% \begin{thm}
+% \label{g-1-planar}
+% Every $(g,1)$-planar graph has a layered $H$-partition of layered width at most $\max\{48g, 30\}$ in which $H$ has treewidth at most $4$.  
+% \end{thm}
+% 
+% The proofs of these results make use of the following lemma due to \citet[Lemma~21]{dujmovic.joret.ea:planar}:
+% 
+% \begin{lem}[\citep{dujmovic.joret.ea:planar}] 
+% \label{make-planar}
+% Let $G^+$ be a $(\Sigma,0)$-plane graph where $\Sigma$ is a surface with Euler genus $g$. 
+% For each $r\in V(G^+)$ there is a vertex set $Z\subseteq V(G^+)$ such that $G^+-V(Z)$ is planar and there is a connected planar graph $G^*$ containing  $G^+-Z$ as a subgraph with a vertex $r^*\in V(G^*)$ such that, for every integer $i\ge 0$,
+% \[  \{v\in V(G^*): \dist_{G^*}(r^*,v) = i\} \supseteq \{v\in V(G^+): \dist_{G^+}(r,v) = i\} \]
+% and
+% \[ |\{v\in Z: \dist_{G^+}(r,v) = i\}| \le 2g \enspace . \]
+% \end{lem}
+% 
+% \begin{proof}[Proof of \cref{g-k-planar,g-1-planar}]
+%   Let $G$ be an $(\Sigma,k)$-plane graph where $\Sigma$ is a surface of Euler genus $g$.
 Let $G^+$ be the $(\Sigma,0)$-plane graph obtained by adding a dummy vertex at each crossing in $G$.
 Let $r_0$ be any vertex in $V(G)$, let $T$ be BFS spanning tree of $G^+$ rooted at $r_0$ and let $\langle V_0,V_1,\ldots\rangle$ be the resulting layering of $G^+$.
-
-  Apply \cref{make-planar} to $G^{+}$ to obtain the set $Z\subseteq V(G^{+})$ and the planar graph $G^*$ with special vertex $r^*$.
+% 
+%   Apply \cref{make-planar} to $G^{+}$ to obtain the set $Z\subseteq V(G^{+})$ and the planar graph $G^*$ with special vertex $r^*$.
 For each edge $vw\in E(G)$, let $W_{vw}$ be the corresponding walk in $G^+$.
@@
 In this way, each vertex $z\in Z\cap V(G^+)\setminus V(G)$ can be charged for at most two vertices in $v,x\in Z'$ and this happens only if $(z,vw,xy)$ is a crossing in $G$.
-  Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \cref{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 2\times (2k+1) = 4g(2k+1)$ for each integer $i\ge 0$.
-
-  % For every edge $vw\in E(G)$, $W_{vw}$ has length at most $k+1$.
+%   Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \cref{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 2\times (2k+1) = 4g(2k+1)$ for each integer $i\ge 0$.
+% 
+%   % For every edge $vw\in E(G)$, $W_{vw}$ has length at most $k+1$.
 Therefore $\dist_{G^+}(v,w)\le k+1$ for each $v\in E(G)$.
 Since $W_{vw}\subseteq G^+-Z$ for each $vw\in E(G-Z')$,  $\dist_{G^+-Z}(v,w)\le k+1$ for each $vw\in E(G-Z')$.
-
-  Next, observe that any plane embedding of $G^*\supseteq G^+-Z$ gives a $k$-plane embedding of $G-Z'$.
+% 
+%   Next, observe that any plane embedding of $G^*\supseteq G^+-Z$ gives a $k$-plane embedding of $G-Z'$.
 Indeed, for every edge $vw\in E(G-Z')$, $G^+-Z$ contains a walk $W_{vw}$ that contains at most $k$ dummy vertices of $G^+-Z$.
@@
 Call the resulting $k$-plane graph $G^*_0$.
-
-  Let $G^*_1$ be the supergraph of $G^*_0$ obtained by adding every vertex and edge of $G^*$ that is not contained in some edge of $G^*_0$.
+% 
+%   Let $G^*_1$ be the supergraph of $G^*_0$ obtained by adding every vertex and edge of $G^*$ that is not contained in some edge of $G^*_0$.
 Finally, let $G^{*+}_1$ be the triangulation obtained by adding edges to $G^*$ so that $\dist_{G^*}(r^*,v)=\dist_{G^{*+}}(r^*,v)$ for every $v\in V(G^*)$.
@@
 Now, observe that $G^*_1$ is a $k$-plane graph and $G^{*+}_1$ is obtained by adding dummy vertices at crossings in $G^*_1$ and triangulating the resulting graph.
-
-  Apply \cref{cor-k-planar} or (in the case $k=1$) \cref{cor-1-planar} to $G^*_1$ using $G^{*+}_1$ and the distinguished vertex $r^*$ to obtain a layered $H$-partition of $G^*_1$.
+% 
+%   Apply \cref{cor-k-planar} or (in the case $k=1$) \cref{cor-1-planar} to $G^*_1$ using $G^{*+}_1$ and the distinguished vertex $r^*$ to obtain a layered $H$-partition of $G^*_1$.
 Since $G^*_1\supseteq G^*_0$ and $G^*_0$ is isomorphic to $G-Z'$, this gives a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ of $G-Z'$ that satisfies the conditions of \cref{cor-k-planar} or \cref{cor-1-planar}, as appropriate.
 In particular, for every integer $i\ge 0$,
-  \begin{align*} 
-    V_i'& = \{v\in V(G-Z'):\floor{\dist_{G^*}(r^*,v)/(k+1)} = i\} \\
-        & = \{v\in V(G-Z'):\floor{\dist_{G^+}(r_0,v)/(k+1)} = i\} \enspace .
-  \end{align*}
-  It follows that $|V_i'\cap Z'|\le 4g(2k+1)(k+1) = 8gk^2+12gk+4g$.
-
-  We now turn $\mathcal{P}$, the $H$-partition of $G-Z'$, into a partition $\mathcal{P}'=\mathcal{P}\cup \{Z'\}$ of $G$ by placing all vertices of $Z'$ into one new common part.
+%   \begin{align*} 
+%     V_i'& = \{v\in V(G-Z'):\floor{\dist_{G^*}(r^*,v)/(k+1)} = i\} \\
+%         & = \{v\in V(G-Z'):\floor{\dist_{G^+}(r_0,v)/(k+1)} = i\} \enspace .
+%   \end{align*}
+%   It follows that $|V_i'\cap Z'|\le 4g(2k+1)(k+1) = 8gk^2+12gk+4g$.  
+% 
+%   We now turn $\mathcal{P}$, the $H$-partition of $G-Z'$, into a partition $\mathcal{P}'=\mathcal{P}\cup \{Z'\}$ of $G$ by placing all vertices of $Z'$ into one new common part.
 Let $H'=G/ \mathcal{P}'$.
 Then the treewidth of $H'$ is at most one more than the treewidth of $H$.
-
-  % In the worst-case, the vertex of $H'=G/ \mathcal{P}'$ corresponding to $V(G)\cap Z'$ is adjacent to every vertex of $H=(G-Z')/\mathcal{P}$, so the treewidth of $H'$ is at most one more than the treewidth of $H$.
-
-  Summarizing, we obtain an $H'$-partition $(\mathcal{L},\mathcal{P}')$ of $G$ where $H'$ has treewidth 
-  \[
-      \tw(H') 
-        \le \begin{cases}
-              4 & \text{if $k=1$} \\
-              k^3/6 + 3k^2/2 + 13k/3+5 & \text{otherwise}
-            \end{cases} 
-  \]
-  and, for each $P\in\mathcal{P}'$,
-  \[ 
-    |V_i' \cap P| \le \begin{cases} 8gk^2+12gk+4g & \text{if $P=Z'$} \\
-                                    30 & \text{if $P\neq Z'$ and $k=1$} \\
-                                    24k^2+60k+36 & \text{otherwise.} 
-                      \end{cases} \qedhere
-  \]
-\end{proof}
+% 
+%   % In the worst-case, the vertex of $H'=G/ \mathcal{P}'$ corresponding to $V(G)\cap Z'$ is adjacent to every vertex of $H=(G-Z')/\mathcal{P}$, so the treewidth of $H'$ is at most one more than the treewidth of $H$.
+% 
+%   Summarizing, we obtain an $H'$-partition $(\mathcal{L},\mathcal{P}')$ of $G$ where $H'$ has treewidth 
+%   \[
+%       \tw(H') 
+%         \le \begin{cases}
+%               4 & \text{if $k=1$} \\
+%               k^3/6 + 3k^2/2 + 13k/3+5 & \text{otherwise}
+%             \end{cases} 
+%   \]
+%   and, for each $P\in\mathcal{P}'$,
+%   \[ 
+%     |V_i' \cap P| \le \begin{cases} 8gk^2+12gk+4g & \text{if $P=Z'$} \\
+%                                     30 & \text{if $P\neq Z'$ and $k=1$} \\
+%                                     24k^2+60k+36 & \text{otherwise.} 
+%                       \end{cases} \qedhere
+%   \]
+% \end{proof}
 
@@
 \cref{non-repetitive,k-planar,1-planar,g-k-planar} immediately imply the following results:
+
+\noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}
 
@@
 
+\noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}
+
 \begin{cor}
@@
 
-\citet{DDOSRSV04} proved the conjecture fo Robin Thomas that every graph in a proper minor-closed class has an edge 2-colouring so that each monochromatic subgraph has bounded treewidth, and more generally, that for fixed $c\geq 2$, every such graph can be edge $c$-coloured such that the union of any $c-1$ colour classes has bounded treewidth.
+\citet{DDOSRSV04} proved the conjecture of Robin Thomas that every graph in a proper minor-closed class has an edge 2-colouring so that each monochromatic subgraph has bounded treewidth, and more generally, that for fixed $c\geq 2$, every such graph can be edge $c$-coloured such that the union of any $c-1$ colour classes has bounded treewidth.
 They also showed analogous vertex-colouring results.
@@
 
+\noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}
+
 \begin{prop}
@@
 \cref{DrawG,g-k-planar,qn,p-centered} immediately imply the following result:
+
+\noindent\note{PM}{We can probably do better now by using a $(k,(\Delta-1)^{k-1})$-shortcut system and \cref{modern-major-general}.}
 

### 2019-07-05 k-planar.tex
Careful check of calculations

@@
 \label{k-planar}
-Every $k$-planar graph has a layered $H$-partition of layered width at most $15k^2 + 30k$ in which $H$ has treewidth at most $\treewidth$.
+Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\treewidth$.
 \end{thm}
@@
 \begin{clm}\clmlabel{generalized-tripod}
-  For each $x\in V(T)$, let $V_x = \bigcup_{y\in V(T_x)} Y_y$, $F_x=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x=V_x\cup F_x$.  Then, 
+  For each $x\in V(T)$, let $V_x := \bigcup_{y\in V(T_x)} Y_y$, $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x:=V_x\cup F_x$.  Then, 
   \begin{compactenum}[(Y1)]
     \item\ylabel{thickness} $\mathcal{Y}:=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width (with respect to $\mathcal{L}$) at most $\ell$.
-    \item\ylabel{separator} For every $x\in V(T)$, there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$. 
-    \item\ylabel{ancestor-edge} For every $x\in V(T)$, $x$ has $t'\le t$ strict $T$-ancestors $a_1,\ldots,a_{t'}$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
+    \item\ylabel{separator} For each $x\in V(T)$, there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$. 
+    \item\ylabel{ancestor-edge} For each $x\in V(T)$, there is a set $\{a_1,\ldots,a_{t'}\}$ of $t'\le t$ strict $T$-ancestors of $x$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
   \end{compactenum}
@@
 Since $w\in F_x$, there exists an edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.
-Since $v\in V_x$, $v\in Y_{x'}=Q_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
-Since $\mathcal{Y}$ is a partition, $w\in Y_{a}=Q_a$ for some $a\not\in V(T_x)$.
-Since $vw\in E(G)$, $x'a\in E(H)$.
+Since $v\in V_x$, $v\in Y_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
+Since $\mathcal{Y}$ is a partition, $w\in Y_{a}$ for some $a\not\in V(T_x)$.
+Since $vw\in E(G)$, we have $x'a\in E(H)$.
 By \tref{ancestor-edge}, one of $a$ or $x'$ is a $T$-ancestor of the other.
@@
 Therefore, $a$ is a $T$-ancestor of $x$ which is a $T$-ancestor of $x'$.
-Let $z_0,\ldots,z_r$ be the path, in $T$ from $z_0:=x'$ to $z_r:=a$.
+Let $z_0,\ldots,z_r$ be the path in $T$ from $z_0:=x'$ to $z_r:=a$.
 For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $B_{z_i}$.
@@
 In particular, $a$ is contained in $B_x$.
-  Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $C_x$ contains $x$.
+  Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $B_x$ contains $x$.
 \end{proof}
@@
   % Note that, by \tref{ancestor-edge}, $xy\in E(H^+)$ if and only if $x\in B_y$ or $y\in B_x$.
-  We claim that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \emph{lazy walk} in a graph $H$ is a walk in the non-simple graph $H'$ obtained by adding a self loop to each vertex of $H$.} in $H^+$.
+  We claim that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \emph{lazy walk} in a graph $H$ is a walk in the pseudograph $H'$ obtained by adding a self loop to each vertex of $H$.} in $H^+$.
 Indeed, if $a_i\neq a_{i+1}$ for some $i\in\{0,\ldots,p-1\}$ then this is precisely because $w_i\in V_{a_i}$ but $w_{i+1}\not\in V_{a_i}$.
@@
 \end{enumerate}
-Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields $|S_x\cap Y_i|\le d\ell(5k/2+5)=15k+30$.
-With this change, the layered width of the partition given by \cref{modern-major-general} becomes $15k^2+30k$.
-This establishes \cref{k-planar}.
+Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields 
+\begin{align*}
+    |S_x\cap Y_i| &\le d\ell\left( 
+       2 
+       + 4\left\lfloor\frac{k+1}{2}\right\rfloor 
+       + 2\left(k+1-\left\lfloor\frac{k+1}{2}\right\rfloor\right)
+       \right)\\
+       & =
+       d\ell\left(
+         2k + 2\left\lfloor\frac{k+1}{2}\right\rfloor + 4
+       \right) \\
+       & \le
+       d\ell(3k+5) = 18k+30 \enspace .
+\end{align*}    
+With this change, the layered width of the partition given by \cref{modern-major-general} becomes $18k^2+30k$.  This establishes \cref{k-planar}.
 
@@
 Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most 
-$\max\{2g,3\}(5k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}$.
+$\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}$.
 \end{thm}
 
-\note{PM}{I remove the $(g,1)$-specialization. It requires cutting surfaces.  For reference, the statement was:
+\note{PM}{I removed (commented out) the $(g,1)$-specialization. It requires cutting surfaces.  For reference, the statement was:
 \begin{thm}
@@
 This section presents a rough characterisation of $(g,k)$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $(g,k)$-planar.
+
+\note{PM}{By now this rough characterization is closely related to $(k+1,2)$-shortcut systems.  Let's think about how to present this.}
 

### 2019-07-08 k-planar.tex
David's changes

@@
 \usepackage[utf8]{inputenc}
-\usepackage{amsmath}
+\usepackage{amsmath,wrapfig}
 \usepackage{amsfonts}
@@
 
+\newcommand{\PP}{\mathcal{P}}
+
 \renewcommand{\ge}{\geqslant}
@@
 Here $\boxtimes$ is the strong product, and treewidth is an invariant that measures how `tree-like' a given graph is; see \cref{Tools} for formal definitions and see \cref{ProductExample} for an example.
-So, loosely speaking, \cref{PlanarBasic} says that every planar graph is contained in the product of a tree-like graph and a path.
-This enables combinatorial results for graphs of bounded treewidth to be generalised for planar graphs (with different constants).
-\cref{kPlanarBasic} has been the key tool in solving several well-known open problems.
-In particular, \citet{dujmovic.joret.ea:planar} proved that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92}), and \citet{dujmovic.esperet.ea:planar}  proved that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
-
-\begin{figure}[!h]
+So, loosely speaking, \cref{PlanarBasic} says that every planar graph is contained in the product of a tree-like graph and a path. This enables combinatorial results for graphs of bounded 
+\begin{wrapfigure}{r}{0.5\textwidth} 
 \centering
@@
 \label{ProductExample}
-\end{figure}
+\end{wrapfigure}
+treewidth to be generalised for planar graphs (with different constants).
+\cref{kPlanarBasic} has been the key tool in solving several well-known open problems.
+In particular, \citet{dujmovic.joret.ea:planar} proved that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92}), and \citet{dujmovic.esperet.ea:planar}  proved that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
 
@@
 \label{kPlanarBasic}
-Every $k$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most $(4+o(1))k^5$ and for some path $P$. 
+Every $k$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth $O(k^5)$ and for some path $P$. 
 \end{thm}
 
-This theorem and several related results are proved in \cref{Structure}.
+We actually prove a substantially stronger result than \cref{kPlanarBasic} using the following definition.
+A collection $\mathcal{P}$ of paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if:
+\begin{compactitem}
+\item every path in $\mathcal{P}$ has length at most $k$, and
+\item for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
+\end{compactitem} 
+Each path $P\in\mathcal{P}$ is called a \emph{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \emph{$vw$-shortcut}.
+Given a graph $G$ and a $(k,d)$-shortcut system $\mathcal{P}$ for $G$, let $G^{\mathcal{P}}$ denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\mathcal{P}$.
+
+This definition is related to $k$-planarity because of the following observation
+
+\begin{obs}
+\label{AddDummy}
+Every $k$-planar graph is isomorphic to $G^\PP$ for some planar graph $G$ and some $(k+1,2)$-shortcut system $\PP$ for $G$. 
+\end{obs}
+
+The proof of \cref{AddDummy} is trivial: Given a $k$-planar graph $G'$, create a planar graph $G$ by adding a dummy vertex at each crossing point.
+For each edge $vw\in E(G')\setminus E(G)$ there is a path $P$ in $G$ between $v$ and $w$ of length at most $k+1$.
+Let $\PP$ be the set of such paths $P$.
+For each vertex $v$ of $G$, at most two paths in $\PP$ use $v$ as an internal vertex (since no original vertex of $G'$ is an internal vertex of a path in $\PP$).
+Thus $\PP$ is a $(k+1,2)$-shortcut system for $G$, such that $G'=G^\PP$.
+
+We prove the following theorem.
+
+\begin{thm}
+\label{ShortcutProduct}
+Let $G$ be a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most $t$ and for some path $P$. 
+Let $\PP$ be a $(k,d)$-shortcut system for $G$. Then $G^\PP$ is a subgraph of $J\boxtimes P$ for some graph $J$ of treewidth at most $d(k^3+k^2+4k)\binom{k+t}{t}-1$ and some path $P$. 
+\end{thm}
+
+\cref{ShortcutProduct,AddDummy} immediately imply \cref{kPlanarBasic} with $O(k^{11})$ instead of $O(k^5)$.
+Some further observations reduce the bound to $O(k^5)$; see \cref{Tools}.
+All of these results  generalise for graphs that have a drawing on an arbitrary surface with a bounded number of crossings per edge.
+
+\cref{ShortcutProduct} leads to several other results of interest.
+Here is one example.
+The \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
+If $G$ has maximum degree $\Delta$, then $G^k = G^\PP$ for some $(k,2k\Delta^{k})$-shortcut system $\PP$; see \cref{PowerShortcut}.
+\cref{kPlanarBasic,ShortcutProduct} then imply:
+
+\begin{thm}
+\label{kPowerBasic}
+For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most $O( k^7 \Delta^{k})$ and for some path $P$.
+\note{DW}{Improve to $k^6\Delta^{k}$?}
+\end{thm}
+
+%This theorem and several related results are proved in \cref{Structure}.
 In particular, in \cref{sec-1-planar} we present a refinement of \cref{kPlanarBasic} in the case $k=1$.
-In \cref{sec-g-k-planar} we generalise \cref{kPlanarBasic} for graphs that have a drawing on an arbitrary surface with a bounded number of crossings per edge.
-
-\cref{kPlanarBasic} (and its generalisations) have applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, and $p$-centred colouring  \citep{micek:personal}.
+In \cref{sec-g-k-planar} we generalise \cref{kPlanarBasic} for
+
+These theorems have applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, and $p$-centred colouring  \citep{micek:personal},
 %, and graph encoding \citep{BGP}. 
-These applications in the context of $k$-planar graphs are explored in \cref{Applications}.
+which we explore in \cref{Applications}.
 For example, we prove that $k$-planar graphs have bounded non-repetitive chromatic number (for fixed $k$).
-Prior to the recent work of
-\citet{dujmovic.esperet.ea:planar}, it was even open whether planar graphs have bounded non-repetitive chromatic number.
-
-\cref{Characterisation} presents a rough characterisation of $k$-planar graphs in terms of an underlying planar graph.
+Prior to the recent work of \citet{dujmovic.esperet.ea:planar}, it was even open whether planar graphs have bounded non-repetitive chromatic number.
+
+%\cref{Characterisation} presents a rough characterisation of $k$-planar graphs in terms of an underlying planar graph.
 This is useful for showing that various families of graphs are $k$-planar.
 For example, we conclude that bounded powers of bounded-degree planar graphs are $k$-planar, for a suitable constant $k$.
-Thus
-\cref{kPlanarBasic} and the above applications are applicable to this class. These results also generalise for arbitrary surfaces.
-
-\cref{Examples} presents several examples of $k$-planar graphs, including $k$-map graphs, $k$-string graphs, and $k$-nearest neighbour graphs.
+Thus  \cref{kPlanarBasic} and the above applications are applicable to this class.
+These results also generalise for arbitrary surfaces.
+
+\cref{Examples} presents several examples of $k$-planar graphs, including graph powers, map graphs, string graphs, and nearest neighbour graphs.
 All of the above results are applicable to these examples.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Tools}
+\section{Layerings, Decompositions and Partitions}
 \label{Tools}
@@
 The \emph{layered width} of $(\mathcal{L},\mathcal{P})$ is $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
+
+\citet{dujmovic.joret.ea:planar} introduced the study of partitions with bounded layered width such that the quotient has bounded treewidth.
+They defined a class $\mathcal{G}$ of graphs to \emph{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition with layered width at most $\ell$ such that $H$ has treewidth at most $t$.
+
 These definitions relate to strong products as follows.
@@
 
-\citet{dujmovic.joret.ea:planar} proved the following result, which with \cref{PartitionProduct,MakeWidth1}, implies \cref{PlanarBasic}.
+\citet{dujmovic.joret.ea:planar} proved the following result, which with \cref{PartitionProduct}, implies \cref{PlanarBasic}.
 
@@
 \end{compactenum}
-%\textcolor{red}{eliminate some of these results?}
+\textcolor{red}{eliminate some of these results? just keep (b)? }
 \end{thm}
@@
  graph $X$, there is a constant $c$ such that every $X$-minor-free graph is a subgraph of $H\boxtimes P$ for some graph $H$ of treewidth at most $c$.
-Apex-minor-free graphs are the largest minor-closed class for which such a theorem holds \citep{dujmovic.joret.ea:planar}.
+Apex-minor-free graphs are the largest minor-closed class that admit bounded layered partitions~\citep{dujmovic.joret.ea:planar}.
 However, the family of $k$-planar graphs is not minor-closed.
@@
 
-The following is the main result of this paper; it shows that $k$-planar graphs admit $H$-partitions with bounded layered width, in which $H$ has bounded treewidth.
-With \cref{PartitionProduct,MakeWidth1}, this proves \cref{kPlanarBasic}.
-% with $c_k\leq(4+o(1))k^5$.
+\section{Our Results}
+
+The following result is the main theorem of the paper.
+Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so to does $G^\PP$ for every $(k,d)$-shortcut system $\PP$.
+
+\begin{thm}\label{modern-major-general}
+  Let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\PP$ be a $(k,d)$-shortcut system for $G$.
+Then $G^\PP$ has a layered $J$-partition of layered width at most $d\ell(k^3+k^2+4k)$ for some graph $J$ with treewidth at most $\binom{k+t}{t}-1$.
+\end{thm}
+
+\cref{modern-major-general,PartitionProduct,MakeWidth1} immediately imply \cref{ShortcutProduct} in the introduction.
+
+The following result shows that $k$-planar graphs admit bounded layered partitions. With \cref{PartitionProduct,MakeWidth1}, this proves \cref{kPlanarBasic}. % with $c_k\leq(4+o(1))k^5$.
 
@@
 \label{k-planar}
-Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\treewidth$.
+Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
 \end{thm}
 
-In the special case $k=1$ we obtain better constants and an additional property (planarity) of $H$:
+\cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with a weaker bound on the layered width.
+We prove the stronger bound in \cref{sec-k-planar}.
+Also note that \cref{k-planar,PartitionProduct} imply \cref{kPlanarBasic} with the $O(k^5)$ bound on the treewidth of $H$, as promised.
+
+In the special case $k=1$ we obtain better constants and an additional property (planarity) of $H$ (see \cref{sec-1-planar} for the proof):
 
@@
 
-Prior to this work, the strongest structural description of $k$-planar graphs was in terms of layered treewidth, which we now define.
+The definition of $k$-planar graphs naturally generalises for other surfaces.
+A graph $G$ drawn on a surface $\Sigma$ is $(\Sigma,k)$-plane if every edge of $G$ is involved in at most $k$ crossings.
+A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
+ %\citet[Theorem~22]{dujmovic.joret.ea:planar} show that every genus-$g$ graph has a layered $H$-partition of layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most 4. 
+\cref{AddDummy} immediately generalises as follows:
+
+\begin{obs}
+\label{gAddDummy}
+Every $(g,k)$-planar graph is isomorphic to $G^\PP$ for some graph $G$ of Euler genus $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G$. 
+\end{obs}
+
+We prove that $(g,k)$-planar graphs admits bounded layered partitions.
+
+\begin{thm}
+\label{g-k-planar}
+Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most 
+$\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
+\end{thm}
+
+Note that \cref{SurfacePartition}(b) and \cref{modern-major-general} imply \cref{k-planar} with a weaker bound on the layered width. We prove the stronger bound in \cref{sec-k-planar}.
+
+Finally, we state the following corollary of \cref{modern-major-general,MakeWidth1,PartitionProduct} and the above-mentioned theorem for apex-minor-free graphs.
+This is the strongest result that follows from our work and that of \citet{dujmovic.joret.ea:planar}.
+
+\begin{thm}
+For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\PP$ for $G$,  $G^\PP$ has an $H$-partition of layered width 1 such that $H$ has treewidth at most $c$; that is $G^\PP \subseteq H \boxtimes P$ for some path $P$.
+\end{thm}
+
+%%%%%%%%%%%%%%%%
+\subsection{Prior Work}
+\label{PriorWork}
+
+\note{DW}{should this subsection be moved to the end of \cref{Tools}?}
+
+Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs was in terms of layered treewidth, which we now define.
 A \emph{layered tree-decomposition} $(\mathcal{L},\mathcal{K})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{K}$ of $G$.
@@
 \citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph.
-\citet{dujmovic.eppstein.ea:structure} show that $k$-planar graphs have layered treewidth at most $6(k+1)$.
-
-If a graph has a partition with bounded layered width, where the quotient has bounded treewidth, then the graph has bounded layered treewidth.
+\citet{dujmovic.eppstein.ea:structure} show that every $k$-planar graph has layered treewidth at most $6(k+1)$, and more generally that
+every $(g,k)$-planar graph has layered treewidth at most $(4g+6)(k+1)$.
+
+If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
 In particular, suppose that an $n$-vertex graph $G$ has an $H$-partition with layered width at most $\ell$ such that $H$ has treewidth at most $k$.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Structure}
+\section{Shortcut Systems}
 \label{Structure}
 
-The purpose of this section is to prove \cref{k-planar} and related structural results.
-We begin with a more general abstraction called $(k,d)$-shortcut systems and prove a structural result, \cref{modern-major-general}, for graphs derived from $(k,d)$-shortcut systems.
-The results on $k$-planar (\cref{k-planar}) and $(g,k)$-planar graphs (\cref{g-k-planar}) can then be obtained from \cref{modern-major-general} and existing results for planar and bounded-genus graphs.
-
-\subsection{Shortcut Systems}
-
-For a rooted tree $T$ and any node $x\in V(T)$, we define $T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}]$ to be the maximal subtree of $T$ rooted at $x$.
-We begin with a fairly standard lemma about normalized tree decompositions:
+The purpose of this section is to prove our main result, \cref{modern-major-general}.
+This theorem shows how, given a $(k,d)$-shortcut system $\mathcal{P}$ of a graph $G$, a layered $H$-partition of $G$ can be used to obtain a layered $J$-partition of $G^{\mathcal{P}}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
+Our main results for $k$-planar graphs (\cref{k-planar}) and $(g,k)$-planar graphs (\cref{g-k-planar}) follow; see \cref{sec-k-planar}.
+
+For convenience, it will be helpful to assume that $\mathcal{P}$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.
+This assumption has no effect on the resulting graph $G^{\mathcal{P}}$ but avoids special cases in some of our proofs.
+
+We begin with a fairly standard lemma about normalized tree decompositions. 
+For a rooted tree $T$ and any node $x\in V(T)$, let  
+$$T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}]$$ 
+be the maximal subtree of $T$ rooted at $x$.
 
@@
 
-\subsubsection{Generalized Tripod Partitions}
+%\subsection{Generalized Tripod Partitions}
 
@@
 
-\subsubsection{Shortcut Systems}
-
 % For non-negative integers $k$ and $d$, a graph $G$ and distinct vertices $v,w\in V(G)$, a \emph{$(k,d)$ $vw$-shortcut} in $G$ is a path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.
@@
 
-A collection $\mathcal{P}$ of paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if every path in $\mathcal{P}$ has length at most $k$ and if, for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
+%A collection $\mathcal{P}$ of paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if every path in $\mathcal{P}$ has length at most $k$ and if, for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
 Each path in $P\in\mathcal{P}$ is called a \emph{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \emph{$vw$-shortcut}.
 Given a graph $G$ and a $(k,d)$-shortcut system $\mathcal{P}$ for $G$, we use $G^{\mathcal{P}}$ to denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\mathcal{P}$.
-
-In this section we will show how, given a $(k,d)$-shortcut system $\mathcal{P}$, a layered $H$-partition of $G$ can be used to obtain a layered $I$-partition of $G^{\mathcal{P}}$ where the layered width of the partition does not increase dramatically and the treewidth of $I$ is not much more than the treewidth of $H$.
-
-For convenience, it will be helpful to assume that $\mathcal{P}$ contains a
-length-1 $vw$-shortcut for every edge $vw\in E(G)$.  This assumption has no effect on the resulting graph $G^{\mathcal{P}}$ but avoids special cases in some of our proofs.
 
@@
 % Together, \clmref{general-width} and \clmref{general-bag-size} imply the following generalization of \thmref{k-planar}:
-
-\begin{thm}\label{modern-major-general}
-  Let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\mathcal{P}$ be a $(k,d)$-shortcut system for $G$.
-Then $G^{\mathcal{P}}$ has a layered $I$-partition of layered width at most $d\ell(k^3+k^2+4k)$ where the treewidth of $H$ is at most $\binom{k+t}{t}-1$.
-\end{thm}
-
-\begin{proof}
+We are now ready to prove the main result of the paper.
+
+\begin{proof}[Proof of \cref{modern-major-general}] 
   The proof is almost immediate, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
@@
 To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.
-This increases the thickness given by \clmref{general-width} from $d\ell(k^2+k+4)$ to $d\ell(k^3+k^2+4k)$.
+This increases the layered width given by \clmref{general-width} from $d\ell(k^2+k+4)$ to $d\ell(k^3+k^2+4k)$.
 \end{proof}
 
-\subsection{$k$-Planar Graphs}
+%%%%%%%%%%%%%%%%%%%
+\section{\boldmath $k$-Planar Graphs}
 \label{sec-k-planar}
@@
 In this way, each edge $vw\in E(G)$ corresponds naturally to a path $P_{vw}$ of length at most $k+1$ in $G_0$.
-Take $\mathcal{P}=\{P_{vw}: vw\in E(G)\}$ and observe that $\mathcal{P}$ is a $(k+1,2)$-shortcut for $G_0$ and that $G_0^{\mathcal{P}}\supseteq G$.
+Let $\PP := \{P_{vw}: vw\in E(G)\}$.
+Observe that $\PP$ is a $(k+1,2)$-shortcut for $G_0$ and that $G_0^{\mathcal{P}}\supseteq G$.
 Specifically, $G_0^{\mathcal{P}}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.
@@
 More precisely, a path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$
-\begin{enumerate}
-  \item Contributes at most 2 vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}$;
-  \item Contributes at most 1 vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2},\ldots,k+1\}$; 
+\begin{compactenum}
+  \item Contributes at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}$;
+  \item Contributes at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2},\ldots,k+1\}$; 
   \item Contributes no vertices to $S_x\cap V_j$ for $j > k+1$.
-\end{enumerate}
+\end{compactenum}
 Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields 
@@
 
-\subsection{$1$-Planar Graphs}
+Exactly the same approach using \cref{SurfacePartition}(b) instead of \cref{PlanarPartition} establishes 
+\cref{g-k-planar} for $(g,k)$-planar graphs.
+
+\subsection{\boldmath  $1$-Planar Graphs}
 \label{sec-1-planar}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{$(g,k)$-Planar Graphs}
+\subsection{\boldmath  $(g,k)$-Planar Graphs}
 \label{sec-g-k-planar}
-
-The definition of $k$-planar graphs naturally generalises for other surfaces.
-A graph $G$ drawn on a surface $\Sigma$ is $(\Sigma,k)$-plane if every edge of $G$ is involved in at most $k$ crossings.
-A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
-
-\citet[Theorem~22]{dujmovic.joret.ea:planar} show that every genus-$g$ graph has a layered $H$-partition of layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most 4.
-Using this result and proceeding exactly as in \cref{sec-k-planar} establishes the following generalisation of \cref{k-planar}.
-
-\begin{thm}
-\label{g-k-planar}
-Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most 
-$\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}$.
-\end{thm}
 
@@
 \end{thm}
+}
+\note{DW}{I can understand you doing this, but for some our applications we only get the best known bounds in the case when $g\gg k$, in which case this is best illustrated by the $k=1$ case.
+I have commented out these results which rely on  \cref{g-1-planar} for now, but I would advocate including  \cref{g-1-planar} and its nice corollaries.
 }
@@
 Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of 
-$O(g^{k+2})$ using \cref{SurfacePartition} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$. In particular, in the case for $k=1$, \cref{qn,1-planar,g-1-planar} imply:
-
-\begin{prop}
-For $g\geq 1$, every $(g,1)$-planar graph has queue-number at most $3 \times 48g (2^4-1) + \floor{\tfrac{3}{2} \times 48g }= 2232 g $.
-\end{prop}
-
-The previous best bound on the queue-number of $(g,1)$-planar graphs was $2(40g+490)^{3}$ due to \citet{dujmovic.joret.ea:planar}.
+$O(g^{k+2})$ using \cref{SurfacePartition} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$. 
+%In particular, in the case for $k=1$, \cref{qn,1-planar,g-1-planar} imply:
+%
+%\begin{prop}
+%For $g\geq 1$, every $(g,1)$-planar graph has queue-number at most $3 \times 48g (2^4-1) + \floor{\tfrac{3}{2} \times 48g }= 2232 g $.
+%\end{prop}
+%
+%The previous best bound on the queue-number of $(g,1)$-planar graphs was $2(40g+490)^{3}$ due to \citet{dujmovic.joret.ea:planar}.
 For $g=0$ we can improve further.
+\citet{ABGKP18} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
+Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$.
+\cref{qn,1-planar} then imply:
+In the case of 1-planar graphs we can improve further.
 \citet{ABGKP18} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
@@
 \label{non-repetitive}
-If a graph $G$ has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$, then $\pi(G)\le \ell 4^{t+1}$.
+If a graph $G$ has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$, then $\pi(G)\le \ell\, 4^{t+1}$.
 \end{lem}
@@
 
-\noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}
+\noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}\note{DW}{done}
 
 \begin{cor}\corlabel{first}
-  For every $k$-planar graph $G$, $\pi(G)\le 12(2k+3)(k+1)\times 4^{(k+2)(k+3)(k+4)/6}$.
+  For every $k$-planar graph $G$, 
+  %$\pi(G)\le 12(2k+3)(k+1)\times 4^{(k+2)(k+3)(k+4)/6}$.
+  $\pi(G)\le (18k^2 + 30k) 4^{\binom{k+4}{3}}$.
+%  Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}}-1$.
 \end{cor}
@@
   For every $1$-planar graph $G$, $\pi(G)\le 30\times 4^4=7680$. 
+%  Every 1-planar graph has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
 \end{cor}
@@
   \[ 
-    \pi(G)\le \max\{4g(3k+1),12(2k+3)(k+1)\}\times 4^{(k+2)(k+3)(k+4)/6} 
-    \enspace .
+%    \pi(G)\le \max\{4g(3k+1),12(2k+3)(k+1)\}\times 4^{(k+2)(k+3)(k+4)/6} 
+    \pi(G)\le \max\{2g,3\}\cdot(6k^2 + 10k) 4^{\binom{k+5}{4}+1}. 
      \]
+%  \label{g-k-planar} Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most  
+% $\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}$.
 \end{cor}
 
-\begin{cor}
-  For every $g\ge 1$ and every $(g,1)$-planar graph $G$, $\pi(G)\le 48g\times 256= 12288g$.
-\end{cor}
+% DW: I have commented out the result for $(g,1)$-planar graph. 
+%\begin{cor}
+%  For every $g\ge 1$ and every $(g,1)$-planar graph $G$, $\pi(G)\le 48g\times 256= 12288g$.
+%\end{cor}
 
@@
 
-The following lemma is due to \citet{micek:personal}. We include the proof  for completeness.
+The following lemma is due to \citet{micek:personal}. We include the proof  for completeness. \note{DW}{can we replace by a result of \citet{micek:personal}?}
 
@@
 
-\noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}
+\noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.} \note{DW}{done}
 
@@
 \label{pck}
-For every $k$-planar graph $G$, $\chi_p(G)\le p^{O(k^3)}$.
+For every $k$-planar graph $G$, for some integer $t\leq \binom{k+4}{3}-1$, 
+$$\chi_p(G)\le (p+1) (t+1)\tbinom{p+t}{t}.$$
+%$\chi_p(G)\le p^{O(k^3)}$.
+%Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
+%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
 \end{cor}
@@
 \label{pc1}
-For every $1$-planar graph $G$, $\chi_p(G)\le O(p^4)$. 
+For every $1$-planar graph $G$, 
+$$\chi_p(G)\le 20 (p+3)(p+2)(p+1)^2\le O(p^4).$$
+%120 (p+1) \binom{p+3}{3} 
+%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
+%Every 1-planar graph has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
 \end{cor}
@@
 \label{pcgk}
-For every $(g,k)$-planar graph $G$, $\chi_p(G)\le O(gk^2)\times p^{O(k^3)}$. 
+For every $(g,k)$-planar graph $G$,  for some integer $t\leq \binom{k+5}{4}-1$, 
+$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) (t+1)\tbinom{p+t}{t}.$$
+%O(gk^2)\times p^{O(k^3)}$. 
+%%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
+%Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most 
+%$\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
 \end{cor}
 
-\begin{cor}
-\label{pcg1}
-For every $(g,1)$-planar graph $G$, $g\ge 1$, $\chi_p(G)\le O(gp^4)$.
-\end{cor}
+%\begin{cor}
+%\label{pcg1}
+%For every $(g,1)$-planar graph $G$, $g\ge 1$, $\chi_p(G)\le O(gp^4)$.
+%\end{cor}
 
@@
 By a result of \citet{zhu:colouring}, $G$ has $r$-weak colouring number at most $( (4g + 6)(k + 1)(2r + 1) )^r$, which by another result of  \citet{zhu:colouring} implies that $G$ has  $p$-centred chromatic number at most $( (4g+6)(k+1)(2^{p-1} + 1) )^{2^{p-2}}$.
-\cref{pc1,pck,pcgk,pcg1} are substantial improvements over these bounds.
+\cref{pc1,pck,pcgk} are substantial improvements over these bounds. 
+%\cref{pc1,pck,pcgk,pcg1} are substantial improvements over these bounds.
 
@@
 \subsection{Low Treewidth Colouring}
+
+\note{DW}{The results in this section are trivial because $(g,k)$-planar graphs have bounded layered treewidth. Probably delete this section.}
 
@@
 
-%%%%%%%
+%%%%%%%%%
 \section{Rough Characterisation}
@@
 This section presents a rough characterisation of $(g,k)$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $(g,k)$-planar.
-
-\note{PM}{By now this rough characterization is closely related to $(k+1,2)$-shortcut systems.  Let's think about how to present this.}
+First note the following easy result (just add a degree-4 vertex at each crossing point of $G$).
+
+\begin{lem}
+\label{gAddDummyDegree}
+For every $(g,k)$-planar graph $G$ there is a graph $G_0$ of Euler
+genus at most $g$, such that $V(G) \subseteq V(G_0)$ and for every
+edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most
+$k+1$, such that every internal vertex in $P$ has degree at most $4$
+in $G_0$.
+\end{lem}
+
+%Recall that \cref{gAddDummy} says that every $(g,k)$-planar graph is isomorphic to $G^\PP$ for some graph $G$ of Euler genus $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G$.
+Here $G$ is the graph obtained by adding a degree-4 dummy vertex at each crossing point.
+Thus for each path $P\in \PP$ each internal vertex in $P$ has degree 4.
+
+We now show that following converse of \cref{gAddDummyDegree}  holds.
 
@@
 
-Note that the following converse of \cref{DrawG}  obviously holds (just add a dummy vertex at each crossing point):
-
-\begin{lem}
-\label{AddDummy}
-For every $(g,k)$-planar graph $G$ there is a graph $G_0$ of Euler
-genus at most $g$, such that $V(G) \subseteq V(G_0)$ and for every
-edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most
-$k+1$, such that every internal vertex in $P$ has degree at most $4$
-in $G_0$.
-\end{lem}
-
 Together \cref{DrawG,AddDummy} provide a rough characterisation of $(g,k)$-planar graphs.
@@
 
-The \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
-\cref{DrawG,g-k-planar,qn,p-centered} immediately imply the following result:
-
-\noindent\note{PM}{We can probably do better now by using a $(k,(\Delta-1)^{k-1})$-shortcut system and \cref{modern-major-general}.}
-
-\begin{prop}
+Recall that the \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
+If $G$ is planar with maximum degree $\Delta$, then $G^k$ is $(k+1)\Delta^{2(k-1)} $-planar by \cref{DrawG}.
+Thus we can immediately conclude that bounded powers of planar graphs of bounded degree admit bounded layered partitions.
+However, the bounds we obtain are improved by the following lemma that constructs a shortcut system directly.
+
+\begin{lem}
+\label{PowerShortcut}
+If a graph $G$ has maximum degree $\Delta$, then $G^k = G^\PP$ for some $(k,2k \Delta^{k})$-shortcut system $\PP$.
+\end{lem}
+
+\begin{proof}
+For each pair of vertices $x$ and $y$ in $G$ with $\dist_G(x,y)\in\{1,\dots,k\}$, fix an $xy$-path $P_{xy}$ of length 
+$\dist_G(x,y)$  in $G$.
+Let $\PP:=\{P_{xy}: \dist_G(x,y)\in\{1,\dots,k\} \}$.
+Say $P_{xy}$ uses some vertex $v$ as an internal vertex.
+If $\dist_G(v,x)=i$ and $\dist_G(v,y)=j$ then $i\in\{1,\dots,k\}$ and $j\in\{1,\dots,k-i\}$.
+The number of vertices at distance $i$ from $v$ is at most $\Delta^i$.
+Thus the number of paths in $\PP$ that use $v$ as an internal vertex is at most
+$$\sum_{i=1}^k \sum_{j=1}^{k-i} \Delta^i\Delta^j 
+= \sum_{i=1}^k \Delta^i \sum_{j=1}^{k-i} \Delta^j 
+< \sum_{i=1}^k \Delta^i ( 2 \Delta^{k-i} )
+= 2k \Delta^k.$$
+Hence $\PP$ is a $(k, 2k \Delta^k)$-shortcut system. 
+\end{proof}
+
+\cref{SurfacePartition}(b) and \cref{modern-major-general,PowerShortcut} imply:
+
+\begin{thm}
+\label{PowerPartition}
+Let $G$ be a graph of Euler genus $g$ and maximum degree $\Delta$. Let $k$ be a positive integer. 
+Then $G^k$ has a layered $H$-partition of layered width at most $\max\{2g,3\} (2k^4+2k^3+8k^2) \Delta^{k} $ for some graph $J$ with treewidth at most $\binom{k+4}{4}-1$.
+\end{thm}
+
+\cref{DrawG,qn,p-centered,PowerPartition} immediately imply the following result:
+
+\noindent\note{PM}{We can probably do better now by using a $(k,(\Delta-1)^{k-1})$-shortcut system and \cref{modern-major-general}.} \note{DW}{An exponent of $k-1$ here is false for $k=2$.}
+
+\begin{cor}
 For all integers $k\geq 1$ and $g\geq 0$, if $G$ is a graph with Euler genus $g$ and maximum degree $\Delta$, then:
@@
 %%%%%%
-\item $G^k$ has a partition of layered width $(8+o(1))(g+1)k^2\Delta^{4(k-1)}$ such that the quotient has treewidth at most $t:= (\frac{1}{6}+o(1))k^3\Delta^{6(k-1)}$,
-%%%%%%%
-\item $G^k$ has queue-number at most $g 2^{O( k^3\Delta^{6(k-1)} )}$, 
-\note{DW}{the result of  \citet{dujmovic.joret.ea:planar} mentioned previously implies an upper bound of  $O(g^{ (k+1)\Delta^{2(k-1)}+2})$}
+\item $G^k$ has queue-number at most 
+$$\max\{2g,3\} (6k^4+6k^3+24k^6) \,\Delta^{k}\, 2^{\binom{k+4}{4}-1},$$
+%If a graph $G$ has a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$, then $\qn(G)\le 3 \ell (2^t-1) + \floor{\tfrac{3}{2}\ell} \le 3 \ell (2^t-\frac12 )$.
 %%%%%%
-\item $G^k$ has  $p$-centred chromatic number at most $(8+o(1))(g+1)k^2\Delta^{4(k-1)}   (p+1) (t+1)\binom{p+t}{t}$, 
+\item $G^k$ has  $p$-centred chromatic number at most 
+$$\max\{2g,3\} (2k^4+2k^3+8k^2) \Delta^{k}  (p+1) (t+1)\tbinom{p+t}{t}$$
+for some integer $t \leq \binom{k+4}{4}-1$. 
+%\label{p-centered} Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
 \end{compactitem}
-\end{prop}
+\end{cor}
 
@@
 but the bound is weak, since $G^k$ has maximum degree at most $\Delta^k$, implying that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$ by a result of \citet{DJKW16}.
+
+Also note that the above results in conjunction with the comments about separators in \cref{PriorWork} imply that for fixed integers $g,k,\Delta$, the $k$-th power of a graph with Euler genus $g$ and maximum degree $\Delta$ has $O(\sqrt{n})$ treewidth  and thus admits balanced separators of order $O(\sqrt{n})$.
 
@@
 
-\note{DW}{This actually works for $k\geq 7$, since the crossing lemma holds for graphs with at least $(3+\epsilon)n$ edges, for any fixed $\epsilon$.}
+\note{DW}{This actually works for $k\geq 7$, since the crossing lemma holds for graphs with at least $(3+\epsilon)n$ edges, for any fixed $\epsilon$. Of course, delete this comment.}
+
+\cref{nearest-neighbour,k-planar,qn,p-centered} imply:
+
+\begin{cor}
+\label{k-nn}
+Every $k$-nearest-neighbour graph has:
+\begin{compactitem}
+\item a layered $H$-partition of layered width $O(k^4)$ in which $H$ has treewidth $O(k^6)$,
+\item queue-number at most $2^{O(k^6)}$,
+\item $p$-centred chromatic number at most $O(k^4) (p+1) (t+1)\binom{p+t}{t}$, for some integer $t\leq O(k^6)$.
+\end{compactitem}
+\end{cor}
+
+\cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
+However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (6+o(1))k$ by a result of \citet{DJKW16}.
+
+\section{Conclusions}
+
+We have shown that $k$-planar graphs and several related graph families have layered $H$-partitions of layered width $O(k^2)$ in which $H$ has treewidth $O(k^3)$.
+We conclude with the following question:  Does there exist a function $f:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph has a layered $H$-partition of layered width at most $f(k)$ in which $H$ has treewidth at most $C$?
+Perhaps $C=3$.
+Note that $C\geq 3$ even for planar graphs \citep{dujmovic.joret.ea:planar}.
+
+\bibliographystyle{plainnat}
+\bibliography{k-planar}
+\newpage
+\appendix
+
+%%%%%%%%%%%%%%
+\section{Future Work}
 
@@
 % }
-
-\section{Conclusions}
-
-We have shown that $k$-planar graphs and several related graph families have layered $H$-partitions of layered width $O(k^2)$ in which $H$ has treewidth $O(k^3)$.
-We conclude with the following question:  Does there exist a function $f:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph has a layered $H$-partition of layered width at most $f(k)$ in which $H$ has treewidth at most $C$?
-Perhaps $C=3$.
-Note that $C\geq 3$ even for planar graphs \citep{dujmovic.joret.ea:planar}.
-
-\bibliographystyle{plainnat}
-\bibliography{k-planar}
-\newpage
-\appendix
-
-%%%%%%%%%%%%%%
-\section{Future Work}
+%
 

### 2019-07-08 k-planar.tex
Small changes, include K -> T

@@
 \cref{kPlanarBasic} has been the key tool in solving several well-known open problems.
-In particular, \citet{dujmovic.joret.ea:planar} proved that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92}), and \citet{dujmovic.esperet.ea:planar}  proved that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
+In particular, \citet{dujmovic.joret.ea:planar} proved that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992), and \citet{dujmovic.esperet.ea:planar}  proved that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002).
 
@@
 \label{AddDummy}
-Every $k$-planar graph is isomorphic to $G^\PP$ for some planar graph $G$ and some $(k+1,2)$-shortcut system $\PP$ for $G$. 
+Every $k$-planar graph is isomorphic to a subgraph of $G^\PP$ for some planar graph $G$ and some $(k+1,2)$-shortcut system $\PP$ for $G$. 
 \end{obs}
@@
 \end{thm}
+\note{PM}{TODO: Work out the treewidth more carefully. The use of big-O is here difficult to decipher (and probably incorrect).}
 
@@
 \end{compactitem}
-A \emph{tree-decomposition} $\mathcal{K}$ of a graph $G$ consists of a tree $K$ and a collection $\mathcal{K}=(B_x:x\in V(K))$ of subsets of $V(G)$ indexed by the nodes of $K$ such that:
+A \emph{tree-decomposition} $\mathcal{T}$ of a graph $G$ consists of a tree $T$ and a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $K$ such that:
 \begin{compactenum}[(i)]
-\item for every $vw\in E(G)$, there exists some $x\in V(K)$ with $v,w\in B_x$; and 
-\item for every $v\in V(G)$, the induced subgraph $K[x] := K[\{y: x\in B_y\}]$ is connected.  
+\item for every $vw\in E(G)$, there exists some node $x\in V(T)$ with $v,w\in B_x$; and 
+\item for every $v\in V(G)$, the induced subgraph $T[v] := T[\{x: v\in B_x\}]$ is connected.  
 \end{compactenum}
-The \emph{width} of the tree-decomposition $\mathcal{K}$ is $\max\{|B_x|:x\in V(K)\}-1$.
+The \emph{width} of the tree-decomposition $\mathcal{T}$ is $\max\{|B_x|:x\in V(T)\}-1$.
 The \emph{treewidth} $\tw(G)$ of a graph $G$ is the minimum width of a tree-decomposition of $G$.
@@
 Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs was in terms of layered treewidth, which we now define.
-A \emph{layered tree-decomposition} $(\mathcal{L},\mathcal{K})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{K}$ of $G$.
-The layered width of $(\mathcal{L},\mathcal{K})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{K}\}$.
+A \emph{layered tree-decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{T}$ of $G$.
+The layered width of $(\mathcal{L},\mathcal{T})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{T}\}$.
 The \emph{layered treewidth} of $G$ is the minimum layered width of any layered tree-decomposition of $G$.

### 2019-07-08 k-planar.tex
Changed width from dl(k^3+k^2+4k) to dl(k^3+3k)

@@
 
-\newcommand{\Bagsize}{\ensuremath{\frac{k^3}{6} + \frac{3k^2}{2} + \frac{13k}{3} + 4}}
+%\newcommand{\Bagsize}{\ensuremath{\frac{k^3}{6} + \frac{3k^2}{2} + \frac{13k}{3} + 4}}
 % \newcommand{\bagsize}{\ensuremath{k^3/6 + 3k^2/2 + 13k/3 + 4}}
@@
 
-\citet{dujmovic.joret.ea:planar} proved the following result, which with \cref{PartitionProduct}, implies \cref{PlanarBasic}.
+\citet{dujmovic.joret.ea:planar} proved the following result, which with \cref{MakeWidth1}, implies \cref{PlanarBasic}.
 
@@
 \end{thm}
+Their proof is constructive and gives a simple quadratic-time algorithm for finding these partitions.
 
@@
 \end{compactenum}
-\textcolor{red}{eliminate some of these results? just keep (b)? }
+%PM: Keep 'em all: \textcolor{red}{eliminate some of these results? just keep (b)? }
 \end{thm}
@@
   Let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\PP$ be a $(k,d)$-shortcut system for $G$.
-Then $G^\PP$ has a layered $J$-partition of layered width at most $d\ell(k^3+k^2+4k)$ for some graph $J$ with treewidth at most $\binom{k+t}{t}-1$.
+Then $G^\PP$ has a layered $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ with treewidth at most $\binom{k+t}{t}-1$.
 \end{thm}
@@
 To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.
-This increases the layered width given by \clmref{general-width} from $d\ell(k^2+k+4)$ to $d\ell(k^3+k^2+4k)$.
+This increases the layered width given by \clmref{general-width} from $d\ell(k^2+3)$ to $d\ell(k^3+3k)$.
 \end{proof}
@@
 Since $G_0$ is planar, the second part of \cref{PlanarPartition} implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
-Applying \cref{modern-major-general} to $G_0$ and $\mathcal{P}$ immediately implies that $G$ (an arbitrary $k$-planar graph) has a layered $H$-partition of layered width $3(k^3+k^2+4k)$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
+Applying \cref{modern-major-general} to $G_0$ and $\mathcal{P}$ immediately implies that $G$ (an arbitrary $k$-planar graph) has a layered $H$-partition of layered width $3(k^3+3k)$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
 
@@
 \end{thm}
+\note{PM}{TODO: Deal with the change caused by the change to layered-width in \cref{modern-major-general}}.
 
@@
 \noindent\note{PM}{We can probably do better now by using a $(k,(\Delta-1)^{k-1})$-shortcut system and \cref{modern-major-general}.} \note{DW}{An exponent of $k-1$ here is false for $k=2$.}
+
+\note{PM}{TODO: Deal with the change caused by the change to layered-width in \cref{modern-major-general}}.
 
@@
 We note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n$ vertices has at least $kn/2$ edges and at most $kn$ edges.
-For $k\ge 8$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is at $\Omega(k^2)$.
-
-\note{DW}{This actually works for $k\geq 7$, since the crossing lemma holds for graphs with at least $(3+\epsilon)n$ edges, for any fixed $\epsilon$. Of course, delete this comment.}
+For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is at $\Omega(k^2)$.
 

### 2019-07-08 k-planar.tex
Changed notation from k to t in MakeWidth1

@@
 Let $G$ be a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most $t$ and for some path $P$. 
-Let $\PP$ be a $(k,d)$-shortcut system for $G$. Then $G^\PP$ is a subgraph of $J\boxtimes P$ for some graph $J$ of treewidth at most $d(k^3+k^2+4k)\binom{k+t}{t}-1$ and some path $P$. 
+Let $\PP$ be a $(k,d)$-shortcut system for $G$. Then $G^\PP$ is a subgraph of $J\boxtimes P$ for some graph $J$ of treewidth at most $d(k^3+3k)\binom{k+t}{t}-1$ and some path $P$. 
 \end{thm}
@@
 \label{MakeWidth1}
-If a graph $G$ has an $H$-partition of layered width $\ell$ with respect to layering $\langle V_0,V_1,\dots\rangle$, for some graph $H$ of treewidth at most $k$, then $G$ has an $H'$-partition of layered width 1 with respect to the same layering, for some graph $H'$ of treewidth at most $(k+1)\ell-1$.
-That is, if $G\subseteq H\boxtimes P\boxtimes K_\ell$ for some graph $H$ of treewidth at most $k$  and for some path $P$, then $G\subseteq H' \boxtimes P$ for some graph $H'$ of treewidth at most $(k+1)\ell-1$.
+If a graph $G$ has an $H$-partition of layered width $\ell$ with respect to layering $\langle V_0,V_1,\dots\rangle$, for some graph $H$ of treewidth at most $t$, then $G$ has an $H'$-partition of layered width 1 with respect to the same layering, for some graph $H'$ of treewidth at most $(t+1)\ell-1$.
+That is, if $G\subseteq H\boxtimes P\boxtimes K_\ell$ for some graph $H$ of treewidth at most $t$  and for some path $P$, then $G\subseteq H' \boxtimes P$ for some graph $H'$ of treewidth at most $(t+1)\ell-1$.
 \end{lem}

### 2019-07-08 k-planar.tex
Quitting for the day

@@
 Various structural results about planar graphs generalise for $k$-planar graphs.
-For example, \citet{FP08} generalised the Lipton-Tarjan separator theorem to show that every $k$-planar graph with $n$ vertices has a balanced separator of order $O(\sqrt{kn})$.
+For example, \citet{FP08} generalised the Lipton-Tarjan separator theorem to show that every $k$-planar graph with $n$ vertices has a balanced separator of order $O(\sqrt{kn})$.\note{PM}{I must be missing something.
+The planarized graph $G'$ has $O(kn)$ vertices.
+Let $S$ be the set of non-dummy vertices and take a separation $(A,B)$ of $G'$ with $|A\cap B|=O(\sqrt{kn})$, $|S\cap (A\setminus B)|\le 2|S|/3$ and $|S\cap (B\setminus A)|\le 2|S|/3$.
+Now take $X=A\cap B$ plus the endpoints of every edge of $G$ that contains a dummy vertex in $A\cap B$.
+$|X|\le 5|A\cap B|=O(\sqrt{kn})$ and $X$ is a separator of $G$.}
 
@@
 
-\cref{modern-major-general,PartitionProduct,MakeWidth1} immediately imply \cref{ShortcutProduct} in the introduction.
+\cref{modern-major-general,MakeWidth1} immediately imply \cref{ShortcutProduct} in the introduction.
 
@@
 
-\cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with a weaker bound on the layered width.
+\cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with the weaker bound of $(2\times 3)((k+1)^3+3(k+1))=6k^3 + 18k^2 + 36k+24$ on the layered width.
 We prove the stronger bound in \cref{sec-k-planar}.
-Also note that \cref{k-planar,PartitionProduct} imply \cref{kPlanarBasic} with the $O(k^5)$ bound on the treewidth of $H$, as promised.
-
-In the special case $k=1$ we obtain better constants and an additional property (planarity) of $H$ (see \cref{sec-1-planar} for the proof):
+%Also note that \cref{k-planar,MakeWidth1} imply \cref{kPlanarBasic} with the $O(k^5)$ bound on the treewidth of $H$, as promised. 
+% PM: I removed the last line because it's already stated in the line preceding the theorem
+
+In the important special case of $k=1$ we obtain better constants and an additional property (planarity) of $H$ (see \cref{sec-1-planar} for the proof):
 
@@
 
-We prove that $(g,k)$-planar graphs admits bounded layered partitions.
+We prove that $(g,k)$-planar graphs admit bounded layered partitions.
 
@@
 
-\note{DW}{should this subsection be moved to the end of \cref{Tools}?}
+\note{DW}{should this subsection be moved to the end of \cref{Tools}?}\note{PM}{No, I don't think so.
+The information given in Section 2 isn't detailed enough to compare with layered treewidth versus layered partitions.}
 
@@
 For convenience, it will be helpful to assume that $\mathcal{P}$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.
-This assumption has no effect on the resulting graph $G^{\mathcal{P}}$ but avoids special cases in some of our proofs.
+Since $G^\PP$ is defined to be a supergraph of $G$, this assumption has no effect on $G^{\mathcal{P}}$ but avoids special cases in some of our proofs.
+
+For a rooted tree $T$ rooted at some node $x_0\in V(T)$, we we say that a node $a\in V(T)$ is a \emph{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \emph{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
+Note that each node $x\in v(T)$ is a $T$-ancestor and $T$-descendant of itself.
+We say that a $T$-ancestor $a\in V(T)$ of $x\in V(T)$ is a \emph{strict} $T$-ancestor of $x$ if $a\neq x$.
+For a rooted tree $T$ and any node $x\in V(T)$, let
+\[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
+be the maximal subtree of $T$ rooted at $x$.
 
 We begin with a fairly standard lemma about normalized tree decompositions.
-For a rooted tree $T$ and any node $x\in V(T)$, let  
-$$T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}]$$ 
-be the maximal subtree of $T$ rooted at $x$.
 
@@
     % \item\tlabel{diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
-    \item\tlabel{subtree-root} For each $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$.
+    \item\tlabel{subtree-root} For each node $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$.
     \item\tlabel{ancestor-edge}\tlabel{last} For each edge $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
@@
 \begin{proof}
-  Begin with any width-$t$ tree decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some rooted tree $T_0$.
+  Begin with any width-$t$ tree decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some tree $T_0$.
 Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$.
-Let $f:V(H)\to V(T)$ be function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y]$.
+Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y]$.
 If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.
@@
 For the remainder of this section, $G$ is a graph and $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{Y}:=(Y_x: x\in V(H)))$ is a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{Y}$ is at most $t$.
-In addition, $(B_x:x\in V(T))$ is a tree-decomposition of $H$ obtained by applying \lemref{nice-decomposition} to $H$.
-The following lemma shows how to interpret these objects as a hierarchical decomposition of $G$.
+In addition, $\mathcal{T}:=(B_x:x\in V(T))$ is a tree-decomposition of $H$ obtained by applying \lemref{nice-decomposition} to $H$.
+The following lemma shows how to interpret the pair $(\mathcal{Y},\mathcal{T})$ as a hierarchical decomposition of $G$.
+Refer to \cref{fig:generalized-tripod}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics{figs/tripoddo}
+  \end{center}
+  \caption{The sets $Y_x$, $F_x$, and $N_x$ associated with $x\in V(T)$
+  and the ancestors $a_1,\ldots,a_{t'}$ of $X$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.}
+  \label{fig:generalized-tripod}
+\end{figure}
 
@@
 
-As in the rest of this section, $(\mathcal{L},\mathcal{Y})$ denotes a layered $H$-partition of $G$; $T$ is a tree used in a tree-decomposition $(B_x:x\in V(T))$ of $H:=G/\mathcal{Y}$ satisfying the conditions of \lemref{nice-decomposition}; and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ are defined as in \clmref{generalized-tripod}.
+In the rest of this section, 
+%$(\mathcal{L},\mathcal{Y})$ denotes a layered $H$-partition of $G$;
+$\mathcal{T}=(B_x:x\in V(T))$ is a tree-decomposition of $H:=G/\mathcal{Y}$ satisfying the conditions of \lemref{nice-decomposition} and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ are defined as in \clmref{generalized-tripod}.
 
@@
 Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
-We let $I:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
-(Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $I$.)
+We let $J:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph and we let $V(J)\subseteq V(T)$ in the obvious way, so that each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
+(Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $J$.)
 
@@
   For the sake of contradiction, assume otherwise, so there exists some $v\in S_x\setminus V_x$.
-
 By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.
@@
 Therefore, either $v\in Y_a$ or $P$ crosses $a$.
-But this implies that $a(v)$ is an ancestor of $a$, contradicting the assumption that $v\in S_x$.
+But this implies that $a(v)$ is a $T$-ancestor of $a$, which is a strict $T$-ancestor of $x$, contradicting the assumption that $v\in S_x$.
 \end{proof}
@@
 \begin{clm}\clmlabel{general-width}
-  For each $i\in\{1,\ldots,h\}$ and each $x\in V(I)$, $|S_x\cap V_i|\le d\ell(k^2+k+4)$.  
+  For each $i\in\{1,\ldots,h\}$ and each $x\in V(J)$, $|S_x\cap V_i|\le d\ell(k^2+3)$.  
 \end{clm}
@@
 \begin{proof}
-  We will count the number of vertices in $S_x\cap V_i$ by upper-bounding the number of vertices contributed to $S_x\cap V_i$ by each vertex $w\in Y_x$.
+  We will count the number of vertices in $S_x\cap V_i$ by upper-bounding the number of vertices contributed to $S_x\cap V_i$ by each vertex $w\in Y_x$. Refer to \cref{contribute}.
   If $w\in Y_x\cap V_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes only one vertex, itself, to $S_x\cap V_i$.
+  \begin{figure}
+    \begin{center}
+      \includegraphics{figs/contribute}
+    \end{center}
+    \caption{A path $P$ containing an internal vertex $w\in Y_x\cap V_{i-j}$.}
+    \label{contribute}
+  \end{figure}
 
@@
   For any $j$, the number of vertices $w\in V_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\mathcal{P}$. Therefore, 
-  \[  |S_x\cap V_i|\le d\ell\cdot\left(k+1 + k + \sum_{j=2}^k 2(k-j)\right) %= d(2k+1) + \sum_{i=1}^{k-2} i
-      = d\ell(k^2 +k + 4) \enspace . \qedhere
+  \[  |S_x\cap V_i|\le d\ell\cdot\left(k+1 + 2k + \sum_{j=2}^k 2(k-j)\right) %= d(2k+1) + \sum_{i=1}^{k-2} i
+      = d\ell(k^2 +3) \enspace . \qedhere
   \]
@@
 \begin{clm}\clmlabel{i-ancestor}
-  For each edge $xy\in E(I)$, one of $x$ or $y$ is a $T$-ancestor of the other.
+  For each edge $xy\in E(J)$, one of $x$ or $y$ is a $T$-ancestor of the other.
 \end{clm}
@@
   Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.
-Since $xy\in E(I)$, $G^\mathcal{P}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
+Since $xy\in E(J)$, $G^\mathcal{P}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
 Since $xy\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.
@@
 \begin{clm}\clmlabel{general-bag-size}
-  The graph $I$ has a tree-decomposition in which every bag has size at most $\binom{k+t}{t}$.
+  The graph $J$ has a tree-decomposition in which every bag has size at most $\binom{k+t}{t}$.
 \end{clm}
@@
 \begin{proof}
-  For the tree-decomposition $(C_x:x\in V(T))$ of $I$ we use the same tree $T$ used in the tree-decomposition $(B_x:x\in V(T))$ of $H$.
-For each node $x$ of $T$, $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $I$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
-  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(I)$, $a,x'\in C_{x'}$ and the connectivity of $T[a]:=T[\{x\in V(T):a\in C_x]$ is obvious.
-Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $I$.
+  For the tree-decomposition $(C_x:x\in V(T))$ of $J$ we use the same tree $T$ used in the tree-decomposition $(B_x:x\in V(T))$ of $H$.
+For each node $x$ of $T$, $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
+  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, $a,x'\in C_{x'}$ and the connectivity of $T[a]:=T[\{x\in V(T):a\in C_x]$ is obvious.
+Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $J$.
 What remains is to bound the size, $\max\{|C_x|:x\in V(T)\}$, of the largest bag.
@@
 
-  If $x_\delta\in C_x$, it is because $x_\delta x'\in E(I)$ for some $T$-descendant $x'$ of $x$.
+  If $x_\delta\in C_x$, it is because $x_\delta x'\in E(J)$ for some $T$-descendant $x'$ of $x$.
 This implies $G^{\mathcal{P}}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
@@
 
+In this section we show that the constants in \cref{modern-major-general} can be improved for the important special case of $1$-planar graphs.
 Let $G$ be a $k$-plane graph.
@@
 Let $\PP := \{P_{vw}: vw\in E(G)\}$.
-Observe that $\PP$ is a $(k+1,2)$-shortcut for $G_0$ and that $G_0^{\mathcal{P}}\supseteq G$.
+Observe that $\PP$ is a $(k+1,2)$-shortcut system for $G_0$ and that $G_0^{\mathcal{P}}\supseteq G$.
 Specifically, $G_0^{\mathcal{P}}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.
@@
 Since $G_0$ is planar, the second part of \cref{PlanarPartition} implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
-Applying \cref{modern-major-general} to $G_0$ and $\mathcal{P}$ immediately implies that $G$ (an arbitrary $k$-planar graph) has a layered $H$-partition of layered width $3(k^3+3k)$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
+Applying \cref{modern-major-general} to $G_0$ and $\mathcal{P}$ immediately implies that $G$ (an arbitrary $k$-planar graph) has a layered $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
 
@@
 \begin{compactenum}
-  \item Contributes at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}$;
+  \item Contributes at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$;
   \item Contributes at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2},\ldots,k+1\}$; 
@@
 }
+
 \note{DW}{I can understand you doing this, but for some our applications we only get the best known bounds in the case when $g\gg k$, in which case this is best illustrated by the $k=1$ case.
@@
 }
+
+\note{PM}{I propose we leave it out of the SODA submission, but I'm happy to include it in the DCG/ArXiV submission (which we can hopefully finish in the next few days.)}
+
 % Again, in the case $k=1$ we obtain better constants. 
@@
 Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
-\citet{dujmovic.joret.ea:planar} recently solved this question using \cref{PlanarPartition} and the following lemma.
+\citet{dujmovic.joret.ea:planar} recently proved this conjecture using \cref{PlanarPartition} and the following lemma.
 Indeed, resolving this question was the motivation for the development of \cref{PlanarPartition}.
@@
 \label{qn}
-If a graph $G$ has a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$, then $\qn(G)\le 3 \ell (2^t-1) + \floor{\tfrac{3}{2}\ell}$.
+If a graph $G$ has a layered $H$-partition of layered width $\ell$ in which $H$ has queue number at most $q$, then $\qn(G)\le 3 \ell q + \floor{\tfrac{3}{2}\ell}$.
 \end{lem}
 
-\cref{qn,g-k-planar} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$. 
+Several authors have shown that queue-number is bounded by a function of treewidth. The best current result of this type is due to \cite{wiechert:on}
+
+\begin{thm}\label{qn-tw}
+    For every graph $G$, $\qn(G)\le 2^{\tw(G)}-1$.
+\end{thm}
+
+\cref{qn,qn-tw,g-k-planar} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$. 
 Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of 
@@
 \begin{prop}
-Every 1-planar graph has queue-number at most $3 \times 30 * 5 + \floor{\tfrac{3}{2} \times 30} = 495$. \end{prop}
+Every 1-planar graph has queue-number at most $3 \times 30 \times 5 + \floor{\tfrac{3}{2} \times 30} = 495$. \end{prop}
 
@@
 
-A $c$-colouring $\phi$ of $G$ is \emph{non-repetitive} if, for every path $v_1,\ldots,v_{2h}$ in $G$, there exists some $i\in\{1,\ldots,k\}$ such that $\phi(v_i)\neq\phi(v_{i+h})$.
+A $c$-colouring $\phi$ of $G$ is \emph{non-repetitive} if, for every path $v_1,\ldots,v_{2h}$ in $G$, there exists some $i\in\{1,\ldots,h\}$ such that $\phi(v_i)\neq\phi(v_{i+h})$.
 The \emph{non-repetitive chromatic number} $\pi(G)$ of $G$ is the minimum integer $c$ such that $G$ has a non-repetitive $c$-colouring.
@@
 
-\noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}\note{DW}{done}
+% \noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}\note{DW}{done}
 
@@
 
-The following lemma is due to \citet{micek:personal}. We include the proof  for completeness. \note{DW}{can we replace by a result of \citet{micek:personal}?}
+The following lemma is due to \citet{micek:personal}. We include the proof  for completeness. 
+% \note{DW}{can we replace by a result of \citet{micek:personal}?}
+% \note{PM}{I didn't see a statement of this exact result in the last draft Piotr sent.  Though their Theorem~7 is this proof for the special case of planar graphs.}
 
@@
 
-\noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.} \note{DW}{done}
+% \noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.} \note{DW}{done}
 
@@
 For every $k$-planar graph $G$, for some integer $t\leq \binom{k+4}{3}-1$, 
-$$\chi_p(G)\le (p+1) (t+1)\tbinom{p+t}{t}.$$
+\note{PM}{This used to say, \color{red} $$\chi_p(G)\le (p+1) (t+1)\tbinom{p+t}{t},$$ but I think we forgot the factor $\ell$.}
+$$\chi_p(G)\le (18k^2+30k)(p+1) (t+1)\tbinom{p+t}{t}.$$
 %$\chi_p(G)\le p^{O(k^3)}$.
@@
 %%%%%%%%%%%%%%%%
-\subsection{Low Treewidth Colouring}
-
-\note{DW}{The results in this section are trivial because $(g,k)$-planar graphs have bounded layered treewidth. Probably delete this section.}
-
-\citet{DDOSRSV04} proved the conjecture of Robin Thomas that every graph in a proper minor-closed class has an edge 2-colouring so that each monochromatic subgraph has bounded treewidth, and more generally, that for fixed $c\geq 2$, every such graph can be edge $c$-coloured such that the union of any $c-1$ colour classes has bounded treewidth.
+% \subsection{Low Treewidth Colouring}
+% 
+% \note{DW}{The results in this section are trivial because $(g,k)$-planar graphs have bounded layered treewidth. Probably delete this section.}
+% 
+% \citet{DDOSRSV04} proved the conjecture of Robin Thomas that every graph in a proper minor-closed class has an edge 2-colouring so that each monochromatic subgraph has bounded treewidth, and more generally, that for fixed $c\geq 2$, every such graph can be edge $c$-coloured such that the union of any $c-1$ colour classes has bounded treewidth.
 They also showed analogous vertex-colouring results.
@@
 The following lemma is implicit in \citep{dujmovic.joret.ea:planar}.
-
-\begin{lem}[\citep{dujmovic.joret.ea:planar}]
-\label{LowTreewidth}
-Let $G$ be a graph that has an $H$-partition of layered width $\ell$ such that $H$ has treewidth at most $k$.
+% 
+% \begin{lem}[\citep{dujmovic.joret.ea:planar}]
+% \label{LowTreewidth}
+% Let $G$ be a graph that has an $H$-partition of layered width $\ell$ such that $H$ has treewidth at most $k$.
 Fix $c\geq 2$.
 Then (a) $G$ has an edge $c$-colouring such that the spanning subgraph of $G$ consisting of the union of any $c-1$ colour classes has treewidth at most $\ell(k+1)(2c-1)$, and (b) $G$ has a vertex $c$-colouring such that the subgraph of $G$ induced by the union of any $c-1$ colour classes has treewidth at most $\ell(k+1)(2c-1)$.
-\end{lem}
-
-One feature of this result compared to that of \citet{DDOSRSV04}  is that it applies for non-minor-closed classes, such as $k$-planar graphs. In particular, \cref{LowTreewidth,g-k-planar} imply:
-
-\begin{prop}[\citep{dujmovic.joret.ea:planar}]
-For every $(g,k)$-planar graph $G$ and integer $c\geq 2$, 
-(a) $G$ has an edge $c$-colouring such that the spanning subgraph of $G$ consisting of the union of any $c-1$ colour classes has treewidth $O(cgk^5 )$, and (b) $G$ has a vertex $c$-colouring such that the subgraph of $G$ induced by the union of any $c-1$ colour classes has treewidth $O(cgk^5 )$.
-\end{prop}
+% \end{lem}
+% 
+% One feature of this result compared to that of \citet{DDOSRSV04}  is that it applies for non-minor-closed classes, such as $k$-planar graphs. In particular, \cref{LowTreewidth,g-k-planar} imply:
+% 
+% \begin{prop}[\citep{dujmovic.joret.ea:planar}]
+% For every $(g,k)$-planar graph $G$ and integer $c\geq 2$, 
+% (a) $G$ has an edge $c$-colouring such that the spanning subgraph of $G$ consisting of the union of any $c-1$ colour classes has treewidth $O(cgk^5 )$, and (b) $G$ has a vertex $c$-colouring such that the subgraph of $G$ induced by the union of any $c-1$ colour classes has treewidth $O(cgk^5 )$.
+% \end{prop}
 
@@
 It is easily seen that each edge is in at most $\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}}$ crossings; see \citet[Page~4]{dujmovic.eppstein.ea:structure}.
-Thus $G$ is a $(g,\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
+Thus $G$ is $(g,\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
 (Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar; this implicitly uses the characterisation of map graphs via half-squares of bipartite graphs.)\ % \citep{??}.)\
@@
 
-\noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}
+% \noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}
 
@@
 %$$\max\{g(\tfrac12 d^4 - 4 d^3 + 15 d^2 - 28 d + 24), \, \tfrac{3}{2} d^4 - 12 d^3 + 45 d^2  - 84 d + 72 \}$$ 
-$$\max\{\tfrac12 gd^4 + O(gd^3), \, \tfrac{3}{2}d^4 +O(d^3) \}$$ 
+$$\max\{\tfrac34 gd^4 + O(gd^3), \, \tfrac{9}{8}d^4 +O(d^3) \}$$ 
 in which $H$ has treewidth at most 
@@
 \cref{MapGraphs,qn} imply an upper bound on the queue-number of $(g,d)$-map graphs that is better than the previous bound obtained by \citet{dujmovic.joret.ea:planar} when $g\gg d$.
-For example, since every $(g,4)$-map graph is $(g,1)$-planar, \cref{qn,1-planar,g-1-planar} imply that every 4-map graph has queue-number at most $495$, and every $(g,4)$-map graph has queue-number at most $2232 g$.
+For example, since every $(g,4)$-map graph is $(g,1)$-planar, \cref{qn,1-planar,g-1-planar} imply that every 4-map graph has queue-number at most $495$,
+\note{PM}{This part is no longer true/tight: and every $(g,4)$-map graph has queue-number at most $2232 g$.}
 
 \cref{MapGraphs,non-repetitive} imply that every $(g,d)$-map graph $G$ has non-repetitive chromatic number 
-$$\pi(G)\le\max\{(\tfrac{1}{2}+o(1)) g(d-2)^4, (\tfrac{3}{2}+o(1)) (d-2)^4\} 4^{ (\frac{1}{384}+o(1))(d-2)^6+1}.$$
+$$\pi(G)\le\max\{(\tfrac{3}{4}+o(1)) g(d-2)^4, (\tfrac{9}{8}+o(1)) (d-2)^4\} 4^{ (\frac{1}{384}+o(1))(d-2)^6+1}.$$
 This is the first known constant upper bound (for fixed $g,d$).
 
-\cref{p-centered,MapGraphs} imply that every $(g,d)$-map graph 
-satisfies $\chi_p(G)\le \ell (p+1)(t+1)\binom{p+t}{t}$ where $t:=  \tfrac{1}{384} d^6 + O(d^5)$.
+\cref{p-centered,MapGraphs} imply that every $(g,d)$-map graph satisfies $\chi_p(G)\le \ell (p+1)(t+1)\binom{p+t}{t}$ where $\ell:=\max\{(\tfrac{3}{4}+o(1)) g(d-2)^4, (\tfrac{9}{8}+o(1)) (d-2)^4\}$ and $t:=  \tfrac{1}{384} d^6 + O(d^5)$.
 These are the first bounds on $\chi_p(G)$ that are polynomial in $p$ for $(g,d)$-map graphs with $d$ fixed.
@@
 Every $(g,\delta)$-string graph has a layered $H$-partition of layered width at most 
-$\max\{(32+o(1))g\delta^4, (96+o(1))\delta^4\}$ in which $H$ has treewidth at most $(\frac{4}{3}+o(1)) \delta^6$.
+% $\max\{(48+o(1))g\delta^4, (72+o(1))\delta^4\}$ in which $H$ has treewidth at most $(\frac{4}{3}+o(1)) \delta^6$.
+$\max\{(48+o(1))g\delta^4, (72+o(1))\delta^4\}$ in which $H$ has treewidth at most $(\frac{2}{3}+o(1)) \delta^8$.
+% tw = 2\delta+5 choose 4 - 1  , for g > 1.
 \end{prop}
@@
 Let $G$ be a graph of Euler genus $g$ and maximum degree $\Delta$. Let $k$ be a positive integer. 
-Then $G^k$ has a layered $H$-partition of layered width at most $\max\{2g,3\} (2k^4+2k^3+8k^2) \Delta^{k} $ for some graph $J$ with treewidth at most $\binom{k+4}{4}-1$.
+Then $G^k$ has a layered $H$-partition of layered width at most $\max\{2g,3\} (2k^4+6k^2) \Delta^{k}$ for some graph $J$ with treewidth at most $\binom{k+4}{4}-1$.
 \end{thm}
-\note{PM}{TODO: Deal with the change caused by the change to layered-width in \cref{modern-major-general}}.
 
@@
 
-\noindent\note{PM}{We can probably do better now by using a $(k,(\Delta-1)^{k-1})$-shortcut system and \cref{modern-major-general}.} \note{DW}{An exponent of $k-1$ here is false for $k=2$.}
-
-\note{PM}{TODO: Deal with the change caused by the change to layered-width in \cref{modern-major-general}}.
+% \noindent\note{PM}{We can probably do better now by using a $(k,(\Delta-1)^{k-1})$-shortcut system and \cref{modern-major-general}.} \note{DW}{An exponent of $k-1$ here is false for $k=2$.}
+
+% \note{PM}{TODO: Deal with the change caused by the change to layered-width in \cref{modern-major-general}}.
 
@@
 \item $G^k$ has queue-number at most 
-$$\max\{2g,3\} (6k^4+6k^3+24k^6) \,\Delta^{k}\, 2^{\binom{k+4}{4}-1},$$
+$$3\,\max\{2g,3\} (2k^4+6k^2) \,\Delta^{k}\, 2^{\binom{k+4}{4}-1},$$
 %If a graph $G$ has a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$, then $\qn(G)\le 3 \ell (2^t-1) + \floor{\tfrac{3}{2}\ell} \le 3 \ell (2^t-\frac12 )$.
@@
 \item $G^k$ has  $p$-centred chromatic number at most 
-$$\max\{2g,3\} (2k^4+2k^3+8k^2) \Delta^{k}  (p+1) (t+1)\tbinom{p+t}{t}$$
+$$\max\{2g,3\}(2k^4+6k^2) \Delta^{k}  (p+1) (t+1)\tbinom{p+t}{t}$$
 for some integer $t \leq \binom{k+4}{4}-1$. 
@@
 Also note that the above results in conjunction with the comments about separators in \cref{PriorWork} imply that for fixed integers $g,k,\Delta$, the $k$-th power of a graph with Euler genus $g$ and maximum degree $\Delta$ has $O(\sqrt{n})$ treewidth  and thus admits balanced separators of order $O(\sqrt{n})$.
+\note{PM}{How interesting is this last note?
+A $\sqrt{n}$-separator of the genus-$g$ graph $G$ has a $k$-neighbourhood of size at most $\sqrt{n}\Delta^k$ and is this neighbourhood is a separator of the $G^k$.}
 

### 2019-07-09 k-planar.tex
David's work

@@
 All these results generalise for drawings of graphs on arbitrary surfaces.
+In fact, we work in a much more general setting based on so-called shortcut systems that are of independent interest.
 \end{abstract}
@@
 \section{Introduction}
+\label{Introduction}
 
@@
 Now take $X=A\cap B$ plus the endpoints of every edge of $G$ that contains a dummy vertex in $A\cap B$.
-$|X|\le 5|A\cap B|=O(\sqrt{kn})$ and $X$ is a separator of $G$.}
+$|X|\le 5|A\cap B|=O(\sqrt{kn})$ and $X$ is a separator of $G$.} \note{DW}{I think you have just found a new proof of this fact.
+Note that you are actually using that $G'$ has treewidth $O(\sqrt{kn})$ and then applying the Robertson-Seymour separator lemma for graphs of given treewidth (so that $|S\cap (A\setminus B)|\le 2|S|/3$ and $|S\cap (B\setminus A)|\le 2|S|/3$).
+}
 
@@
 \label{kPlanarBasic}
-Every $k$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth $O(k^5)$ and for some path $P$. 
+Every $k$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth $(18k^3+30k)\binom{k+4}{3}-1$ and for some path $P$. 
 \end{thm}
@@
 
-We prove the following theorem.
+We prove the following theorem that shows if a graph $G$ is a subgraph of $H\boxtimes P$ and $\PP$ is a shortcut system for $G$, then $G^{\mathcal{P}}$ is a subgraph of $J\boxtimes P$, where the treewidth of $J$ is bounded by a function of the treewidth of $H$.
 
@@
 
-\cref{ShortcutProduct,AddDummy} immediately imply \cref{kPlanarBasic} with $O(k^{11})$ instead of $O(k^5)$.
-Some further observations reduce the bound to $O(k^5)$; see \cref{Tools}.
-All of these results  generalise for graphs that have a drawing on an arbitrary surface with a bounded number of crossings per edge.
+\cref{ShortcutProduct,AddDummy} immediately imply \cref{kPlanarBasic} with $O(k^{11})$ instead of $O(k^5)$. Some further observations reduce the bound to $O(k^5)$; see \cref{Tools}.
 
@@
 \label{kPowerBasic}
-For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most $O( k^7 \Delta^{k})$ and for some path $P$.
-\note{DW}{Improve to $k^6\Delta^{k}$?}
+For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most 
+$\Delta^{k} (2k^4+6k^2)\binom{k+8}{8}-1$. 
 \end{thm}
-\note{PM}{TODO: Work out the treewidth more carefully. The use of big-O is here difficult to decipher (and probably incorrect).}
+
+Note that the dependence on $k$ in the bound on the treewidth of $H$ in \cref{kPowerBasic} is improved from $O(k^{12})$ to $O(k^7)$ in \cref{Powers}.
 
@@
 
-\cref{Examples} presents several examples of $k$-planar graphs, including graph powers, map graphs, string graphs, and nearest neighbour graphs.
+\cref{Examples} presents several examples of graph classes that can be obtained from a shortcut system applied to a planar graphs, including graph powers, map graphs, string graphs, and nearest neighbour graphs.
 All of the above results are applicable to these examples.
+
+All of the above results apply, where instead of planar graphs, we consider graphs of bounded Euler genus.
 
@@
 Unless specifically mentioned otherwise, all graphs are also simple.
-For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$) we use $G[S]$ to denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.
+For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$), let $G[S]$  denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.
 We use $G-S$ as a shorthand for $G[V(G)\setminus S]$.
@@
 
-We now define two concepts used in the statements of \cref{PlanarBasic,kPlanarBasic}: strong products and treewidth.
+We now define two concepts used in the theorems in \cref{Introduction}: strong products and treewidth.
 The \emph{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if:
@@
 
-While strong products enable an easy statement of \cref{PlanarBasic,kPlanarBasic}, to prove such theorems it is helpful to work with layerings and partitions, which we now introduce.
+While strong products enable an easy statement of the theorems in \cref{Introduction}, to prove such results it is helpful to work with layerings and partitions, which we now introduce.
 A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle V_0,V_1,\ldots\rangle$ such that $\{V_0,V_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in V_i$ and $w\in V_j$ then $|j-i|\leq 1$.
@@
 The \emph{layered width} of $(\mathcal{L},\mathcal{P})$ is $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
+
+\note{DW}{I  prefer to not define ``layered $H$-partition''.
+Just define the ``layered width'' of an $H$-partition, which is what we did in \citep{dujmovic.joret.ea:planar}.
+There are many places where we write ``$H$-partition with layered width'' without writing ``layered $H$-partition''.
+I prefer it this way.}
 
@@
 Every planar graph has:
-\begin{compactitem}
+\begin{compactenum}[(a)]
 \item an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$, and
 \item an $H$-partition of layered width $3$ for some planar graph $H$ of treewidth at most $3$.
-\end{compactitem}
+\end{compactenum}
 \end{thm}
@@
 
-\citet{dujmovic.joret.ea:planar} generalised \cref{SurfacePartition} further to show that for every apex\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
+\citet{dujmovic.joret.ea:planar} generalised \cref{SurfacePartition} further as follows.\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
 A class $\mathcal{G}$ of graphs is \emph{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$.
@@
 A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.}
- graph $X$, there is a constant $c$ such that every $X$-minor-free graph is a subgraph of $H\boxtimes P$ for some graph $H$ of treewidth at most $c$.
+
+\begin{thm}[\citep{dujmovic.joret.ea:planar}] 
+\label{ApexMinorFree}
+For every apex graph $X$, there exists $c\in\mathbb{N}$ such that every $X$-minor-free graph $G$ has
+an $H$-partition with layered width 1 such that $H$ has treewidth at most $c$. Equivalently, $G\subseteq H\boxtimes P$ for some path $P$. 
+\end{thm}
+
 Apex-minor-free graphs are the largest minor-closed class that admit bounded layered partitions~\citep{dujmovic.joret.ea:planar}.
@@
 \cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with the weaker bound of $(2\times 3)((k+1)^3+3(k+1))=6k^3 + 18k^2 + 36k+24$ on the layered width. 
+\note{DW}{I would just write, ``\cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with a weaker bound on the layered width.}
 We prove the stronger bound in \cref{sec-k-planar}. 
@@
 \label{gAddDummy}
-Every $(g,k)$-planar graph is isomorphic to $G^\PP$ for some graph $G$ of Euler genus $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G$. 
+Every $(g,k)$-planar graph is isomorphic to $G^\PP$ for some graph $G$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G$. 
 \end{obs}
@@
 
-Finally, we state the following corollary of \cref{modern-major-general,MakeWidth1,PartitionProduct} and the above-mentioned theorem for apex-minor-free graphs.
+Finally, we state the following corollary of \cref{ApexMinorFree,modern-major-general,MakeWidth1,PartitionProduct}.
 This is the strongest result that follows from our work and that of \citet{dujmovic.joret.ea:planar}.
@@
 
-\note{DW}{should this subsection be moved to the end of \cref{Tools}?}\note{PM}{No, I don't think so.
+%\note{DW}{should this subsection be moved to the end of \cref{Tools}?}\note{PM}{No, I don't think so.
 The information given in Section 2 isn't detailed enough to compare with layered treewidth versus layered partitions.}
@@
 
-For a rooted tree $T$ rooted at some node $x_0\in V(T)$, we we say that a node $a\in V(T)$ is a \emph{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \emph{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
+For a tree $T$ rooted at some node $x_0\in V(T)$, we we say that a node $a\in V(T)$ is a \emph{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \emph{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
 Note that each node $x\in v(T)$ is a $T$-ancestor and $T$-descendant of itself.
@@
 \begin{lem}\lemlabel{nice-decomposition}
-  For any graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:  
+  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:  
   \begin{compactenum}[(T1)]
@@
 In addition, $\mathcal{T}:=(B_x:x\in V(T))$ is a tree-decomposition of $H$ obtained by applying \lemref{nice-decomposition} to $H$.
-The following lemma shows how to interpret the pair $(\mathcal{Y},\mathcal{T})$ as a hierarchical decomposition of $G$.
-Refer to \cref{fig:generalized-tripod}
+The following lemma shows how to interpret the pair $(\mathcal{Y},\mathcal{T})$ as a hierarchical decomposition of $G$; refer to \cref{fig:generalized-tripod}.
 
@@
 \end{figure}
+
+\note{W}{I think \clmref{generalized-tripod} should be presented as a stand-alone lemma.
+Then we repeat the statement of \cref{modern-major-general} followed by $\\begin \{proof\} ....
+\\end\{proof\}$.}
+
+\note{DW}{Is some sort of converse to \clmref{generalized-tripod}  also true?
+\clmref{generalized-tripod}  shows that if a graph $G$ has a layered partition with all the nice properties, then $G$ has this generalised tripod partition.
+Isn't also true that if $G$ has a generalised tripod partition, then $G$ has a layered partition with all the nice properties.
+The proof is basically in \citep{dujmovic.joret.ea:planar}, right?}
 
@@
 Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $J$.
-What remains is to bound the size, $\max\{|C_x|:x\in V(T)\}$, of the largest bag.
+It remains to bound the size of each bag $C_x$.
 
@@
 \begin{proof}[Proof of \cref{modern-major-general}] 
-  The proof is almost immediate, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
+  The proof is almost immediate from \clmref{general-width} and \clmref{general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
 In particular, $G^{\mathcal{P}}$ may contain an edge $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.
@@
 
-In this section we show that the constants in \cref{modern-major-general} can be improved for the important special case of $1$-planar graphs.
+This section shows that the constants in \cref{modern-major-general} can be improved in the case of $k$-planar graphs.
 Let $G$ be a $k$-plane graph.
@@
 
-Let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$.
+As in the proof of \cref{AddDummy}, let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$.
 In this way, each edge $vw\in E(G)$ corresponds naturally to a path $P_{vw}$ of length at most $k+1$ in $G_0$.
@@
 
-Since $G_0$ is planar, the second part of \cref{PlanarPartition} implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
+Since $G_0$ is planar,  \cref{PlanarPartition}(b) implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
 Applying \cref{modern-major-general} to $G_0$ and $\mathcal{P}$ immediately implies that $G$ (an arbitrary $k$-planar graph) has a layered $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
@@
 In this setting, the proof of \cref{clm:general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $V_i\cap Y_x$.
-More precisely, a path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$
-\begin{compactenum}
-  \item Contributes at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$;
-  \item Contributes at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2},\ldots,k+1\}$; 
-  \item Contributes no vertices to $S_x\cap V_j$ for $j > k+1$.
-\end{compactenum}
+More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$ contributes:
+\begin{compactitem}
+  \item  at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$;
+  \item at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2},\ldots,k+1\}$; 
+  \item no vertices to $S_x\cap V_j$ for $j > k+1$.
+\end{compactitem}
+\note{DW}{do we need the bulletted list here?}
 Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields 
@@
 
-Exactly the same approach using \cref{SurfacePartition}(b) instead of \cref{PlanarPartition} establishes 
+Exactly the same approach using \cref{SurfacePartition}(b) instead of \cref{PlanarPartition}(b) establishes 
 \cref{g-k-planar} for $(g,k)$-planar graphs.
@@
 
+This section shows that the constants in \cref{modern-major-general} can be further improved in the important case of $1$-planar graphs.
 Let $G$ be an edge-maximal 1-plane multigraph with no two parallel edges on the boundary of a single face.
@@
 \begin{proof}
-  This proof is very similar to the proof of Lemma~14 in \citet{dujmovic.joret.ea:planar}.
+  This proof is very similar to the proof of Lemma~14 by \citet{dujmovic.joret.ea:planar}.
 Rather than duplicate every detail of that proof here, we focus on the differences and refer the reader to the original proof for the remaining details.
@@
   First note that $N'$ is a near-triangulation.
-If $k=3$, set $R_i=P_i$ for each $i\in\{1,2,3\}$.
-Otherwise, as in \citet{dujmovic.joret.ea:planar}, split $P_1,\ldots,P_k$ to partition $V(F)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a non-empty path and each $R_i$ contains vertices from exactly one of $P_1,\ldots,P_k$.
-
-  Next, as in \citet{dujmovic.joret.ea:planar}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
+If $k=3$, set $R_i := P_i$ for each $i\in\{1,2,3\}$.
+Otherwise, as in \citep{dujmovic.joret.ea:planar}, split $P_1,\ldots,P_k$ to partition $V(F)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a non-empty path and each $R_i$ contains vertices from exactly one of $P_1,\ldots,P_k$.
+
+  Next, as in \citep{dujmovic.joret.ea:planar}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
 
@@
 
-  Note that $M$ is a Mercedes graph and that each of the faces $F_1,\ldots,F_m$ is contained in a single wedge of $M$.
+  Note that $M$ is a Mercedes graph 
+  \note{DW}{``Mercedes'' not defined} and that each of the faces $F_1,\ldots,F_m$ is contained in a single wedge of $M$.
 Let $Y^+ := \overline{Y}^+-F$.
@@
 
-  The remainder of the proof finishes in the same way as the proof of Lemma~14 in \cite{dujmovic.joret.ea:planar}.
-The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
+  The proof finishes in the same way as the proof in \cite{dujmovic.joret.ea:planar}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition. 
+  \note{DW}{Do we need $S$ here?}
    Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
@@
 That this is a layering of $G$ follows from \eqref{spar-distance}.
-Since $|V_j\cap S_x|\le 15$ for every integer $i\ge 0$, $|V_i'\cap S_x|\le 30$ for every integer $i\ge 0$ and every $x\in V(H)$.
+Since $|V_j\cap S_x|\le 15$ for every integer $i\ge 0$, \note{DW}{replace $i$ by $j$?} $|V_i'\cap S_x|\le 30$ for every integer $i\ge 0$ and every $x\in V(H)$.
 \end{proof}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\boldmath  $(g,k)$-Planar Graphs}
-\label{sec-g-k-planar}
-
-\note{PM}{I removed (commented out) the $(g,1)$-specialization. It requires cutting surfaces.  For reference, the statement was:
-\begin{thm}
-\label{g-1-planar}
-Every $(g,1)$-planar graph has a layered $H$-partition of layered width at most $\max\{48g, 30\}$ in which $H$ has treewidth at most $4$.  
-\end{thm}
-}
-
-\note{DW}{I can understand you doing this, but for some our applications we only get the best known bounds in the case when $g\gg k$, in which case this is best illustrated by the $k=1$ case.
+%\subsection{\boldmath  $(g,k)$-Planar Graphs}
+%\label{sec-g-k-planar}
+%
+%
+%\note{PM}{I removed (commented out) the $(g,1)$-specialization. It requires cutting surfaces.  For reference, the statement was:
+%\begin{thm}
+%\label{g-1-planar}
+%Every $(g,1)$-planar graph has a layered $H$-partition of layered width at most $\max\{48g, 30\}$ in which $H$ has treewidth at most $4$.  
+%\end{thm}
+%}
+%
+%\note{DW}{I can understand you doing this, but for some our applications we only get the best known bounds in the case when $g\gg k$, in which case this is best illustrated by the $k=1$ case.
 I have commented out these results which rely on  \cref{g-1-planar} for now, but I would advocate including  \cref{g-1-planar} and its nice corollaries.
 }
-
-\note{PM}{I propose we leave it out of the SODA submission, but I'm happy to include it in the DCG/ArXiV submission (which we can hopefully finish in the next few days.)}
+%
+%\note{PM}{I propose we leave it out of the SODA submission, but I'm happy to include it in the DCG/ArXiV submission (which we can hopefully finish in the next few days.)}
+%
 
@@
 \citet{dujmovic.joret.ea:planar} recently proved this conjecture using \cref{PlanarPartition} and the following lemma.
-Indeed, resolving this question was the motivation for the development of \cref{PlanarPartition}.
+Indeed, resolving this question was the motivation for the development of \cref{PlanarBasic,PlanarPartition}.
 
@@
 \label{qn}
-If a graph $G$ has a layered $H$-partition of layered width $\ell$ in which $H$ has queue number at most $q$, then $\qn(G)\le 3 \ell q + \floor{\tfrac{3}{2}\ell}$.
+If a graph $G$ has a layered $H$-partition of layered width $\ell$, then 
+$$\qn(G)\le 3 \ell \, \qn(H) + \floor{\tfrac{3}{2}\ell}\enspace.$$
 \end{lem}
 
-Several authors have shown that queue-number is bounded by a function of treewidth. The best current result of this type is due to \cite{wiechert:on}
-
-\begin{thm}\label{qn-tw}
+\citet{dujmovic.morin.ea:layered} proved that queue-number is bounded by a function of treewidth. 
+The best known bound is due to \citet{wiechert:on}:
+
+\begin{thm}[\citep{wiechert:on}]
+\label{qn-tw}
     For every graph $G$, $\qn(G)\le 2^{\tw(G)}-1$.
@@
 
-\cref{qn,qn-tw,g-k-planar} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$. 
+\cref{qn,qn-tw,g-k-planar} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^4)}$. 
 Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of 
-$O(g^{k+2})$ using \cref{SurfacePartition} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$. 
+$O(g^{k+2})$ using \cref{SurfacePartition} and an ad-hoc method. Our result provides a better bound when $g>2^{k^3}$. 
 %In particular, in the case for $k=1$, \cref{qn,1-planar,g-1-planar} imply:
@@
 
-The next three applications are in the field of graph colouring. For our purposes, a \emph{$c$-colouring} of a graph $G$ is any function $\phi\colon V(G)\to C$, where $C$ is a set of size at most $c$.
+The next two applications are in the field of graph colouring. For our purposes, a \emph{$c$-colouring} of a graph $G$ is any function $\phi\colon V(G)\to C$, where $C$ is a set of size at most $c$.
 
@@
 %    \pi(G)\le \max\{4g(3k+1),12(2k+3)(k+1)\}\times 4^{(k+2)(k+3)(k+4)/6} 
-    \pi(G)\le \max\{2g,3\}\cdot(6k^2 + 10k) 4^{\binom{k+5}{4}+1}. 
+    \pi(G)\le \max\{2g,3\}\cdot(6k^2 + 10k) 4^{\binom{k+5}{4}}. 
      \]
@@
 \label{pck}
-For every $k$-planar graph $G$, for some integer $t\leq \binom{k+4}{3}-1$, 
-\note{PM}{This used to say, \color{red} $$\chi_p(G)\le (p+1) (t+1)\tbinom{p+t}{t},$$ but I think we forgot the factor $\ell$.}
-$$\chi_p(G)\le (18k^2+30k)(p+1) (t+1)\tbinom{p+t}{t}.$$
-%$\chi_p(G)\le p^{O(k^3)}$.
+For every $k$-planar graph $G$, 
+%for some integer $t\leq \binom{k+4}{3}-1$, 
+%$$\chi_p(G)\le (18k^2+30k)(p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
+$$\chi_p(G)\le (18k^2+30k)(p+1) \binom{k+4}{3} \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}\enspace.$$
+%\note{DW}{I previously wrote ``for some integer $t\leq \binom{k+4}{3}-1$'' thinking that $\binom{p+t}{t}$ might not be monotone with $t$, but it is because of the $+t$.}
 %Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
@@
 For every $1$-planar graph $G$, 
-$$\chi_p(G)\le 20 (p+3)(p+2)(p+1)^2\le O(p^4).$$
+$$\chi_p(G)\le 20 (p+3)(p+2)(p+1)^2\enspace.$$
 %120 (p+1) \binom{p+3}{3} 
@@
 \label{pcgk}
-For every $(g,k)$-planar graph $G$,  for some integer $t\leq \binom{k+5}{4}-1$, 
-$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) (t+1)\tbinom{p+t}{t}.$$
+For every $(g,k)$-planar graph $G$,  
+%for some integer $t\leq \binom{k+5}{4}-1$, 
+%$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
+$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) \binom{k+5}{4}\binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}\enspace.$$
 %O(gk^2)\times p^{O(k^3)}$. 
@@
 By a result of \citet{zhu:colouring}, $G$ has $r$-weak colouring number at most $( (4g + 6)(k + 1)(2r + 1) )^r$, which by another result of  \citet{zhu:colouring} implies that $G$ has  $p$-centred chromatic number at most $( (4g+6)(k+1)(2^{p-1} + 1) )^{2^{p-2}}$.
-\cref{pc1,pck,pcgk} are substantial improvements over these bounds. 
+\cref{pc1,pck,pcgk} are substantial improvements over these results, providing bounds on $\chi_p(G)$ that are polynomial in $p$ for fixed $g$ and $k$. 
 %\cref{pc1,pck,pcgk,pcg1} are substantial improvements over these bounds.
@@
 This section presents a rough characterisation of $(g,k)$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $(g,k)$-planar.
-First note the following easy result (just add a degree-4 vertex at each crossing point of $G$).
+First note the following easy result (just add a degree-4 vertex at each crossing point of $G$, as in the proof of \cref{AddDummy,gAddDummy}).
 
@@
 
-We now show that following converse of \cref{gAddDummyDegree}  holds.
+The following converse of \cref{gAddDummyDegree} provides a sufficient condition for a graph to be $(g,k)$-planar that is slightly more restrictive than being obtained from a $(k,d)$-shortcut system of a $(g,k)$-planar graph.
+\note{DW}{Is this sentence clear enough?}
 
@@
 that every internal vertex on $P_{vw}$ has degree at most $\Delta$ in
-$G_0$. Then $G$ is $(g, (k+1)\Delta^{2(k-1)} )$-planar.
+$G_0$. Then $G$ is $(g, 2k(k+1)\Delta^{k} )$-planar.
 \end{lem}
@@
 \begin{proof}
-For a vertex $x$ of $G_0$ with degree at most $\Delta$, say a vertex $v$ is \emph{close} to $x$ if there is a $vx$-path $P$ in $G_0$ of length at most $k-1$ such that every internal vertex in $P$  has degree at most $\Delta$ in $G_0$.
-Note that there are at most $\Delta^{k-1}$ vertices close to $x$.
+For a vertex $x$ of $G_0$ with degree at most $\Delta$, and for $i\in\{1,\dots,k-1\}$, say a vertex $v$ is \emph{$i$-close} to $x$ if there is a $vx$-path $P$ in $G_0$ of length at most $i$ such that every internal vertex in $P$  has degree at most $\Delta$ in $G_0$.
 For each edge $vw$ of $G$, say that $vw$ \emph{passes through} each internal vertex on $P_{vw}$. 
-If $vw$ passes through $x$, then both $v$ and $w$ are close to $x$. 
-Thus at most $\Delta^{2(k-1)}$ edges of $G$ pass through $x$. 
+Say $vw$ passes through $x$. Then $v$ is $i$-close to $x$ and $w$ is $j$-close to $x$ for some $i,j\in\{1,\dots,k-1\}$ with $i+j\leq k$. At most $\Delta^{i}$ vertices are $i$-close to $x$. 
+Thus, the number of edges of $G$ that pass through $x$ is at most 
+$$
+\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i \Delta^j 
+= \sum_{i=1}^{k-1} \Delta^i  \sum_{j=1}^{k-i} \Delta^j 
+< \sum_{i=1}^{k-1} \Delta^i  2 \Delta^{k-i}
+= \sum_{i=1}^{k-1} 2\Delta^k 
+< 2k \Delta^k \enspace.
+$$
 Draw each edge $vw$ of $G$ alongside $P_{vw}$ in $G_0$, so that every pair of edges cross at most once. 
 Every edge of $G$ that crosses $vw$ passes through a vertex on $P_{vw}$ (including $v$ and/or $w$ if they too have degree at most $\Delta$).
-Since $P_{vw}$ has at most $k+1$ vertices, and at most $\Delta^{2(k-1)}$ edges of $G$ pass through each vertex on $P_{vw}$, 
-the edge $vw$ is crossed by at most $(k+1)\Delta^{2(k-1)}$ edges in $G$. 
-Hence $G$ is $(g, (k+1)\Delta^{2(k-1)} )$-planar.
+Since $P_{vw}$ has at most $k+1$ vertices, and less than $2k\Delta^{k}$ edges of $G$ pass through each vertex on $P_{vw}$, the edge $vw$ is crossed by less than $2k(k+1)\Delta^{k}$ edges in $G$.
+Hence $G$ is $(g, 2k(k+1)\Delta^{k} )$-planar.
 \end{proof}
 
-Together \cref{DrawG,AddDummy} provide a rough characterisation of $(g,k)$-planar graphs.
+Together \cref{DrawG,gAddDummyDegree} provide a rough characterisation of $(g,k)$-planar graphs.
 
@@
 
-This section describes several graph classes that are $k$-planar or $(g,k)$-planar. The first two examples are well known, the second two are new.
+This section describes several examples of graph classes that can be obtained from a shortcut system applied to graphs of bounded Euler genus.
+
+%This section describes several graph classes that are $k$-planar or $(g,k)$-planar. The first two examples are well known, the second two are new.
 
@@
 %such graphs have been extensively studied \citep{FLS-SODA12,DFHT05,CGP02,Chen07,Chen01}. 
-The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}. So $(g,d)$-map graphs generalise graphs embedded in a surface.
+The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.
+So $(g,d)$-map graphs generalise graphs embedded in a surface, and we now assume that $d\geq 4$ for the remainder of this section.
 
 There is a natural drawing of a map graph obtained by positioning each vertex of $G$ inside the corresponding nation and each edge of $G$ as a curve passing through the corresponding vertex of $G_0$.
-It is easily seen that each edge is in at most $\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}}$ crossings; see \citet[Page~4]{dujmovic.eppstein.ea:structure}.
+It is easily seen that each edge is in at most $\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}}$ crossings; see \citep{dujmovic.eppstein.ea:structure}.
 Thus $G$ is $(g,\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
-(Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar; this implicitly uses the characterisation of map graphs via half-squares of bipartite graphs.)\ % \citep{??}.)\
-
-\cref{g-k-planar} with $k=\frac14 (d-2)^2$ then implies:
-
-% \noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}
-
-\begin{prop}
-\label{MapGraphs}
-Every $(g,d)$-map graph has a layered $H$-partition of layered width at most 
-%$$\max\{g(\tfrac12 d^4 - 4 d^3 + 15 d^2 - 28 d + 24), \, \tfrac{3}{2} d^4 - 12 d^3 + 45 d^2  - 84 d + 72 \}$$ 
-$$\max\{\tfrac34 gd^4 + O(gd^3), \, \tfrac{9}{8}d^4 +O(d^3) \}$$ 
-in which $H$ has treewidth at most 
-%$$\tfrac{1}{384} d^6 - \tfrac{1}{32} d^5  +  \tfrac{1}{4} d^4 - \tfrac{7}{6} d^3 + \tfrac{95}{24} d^2 - \tfrac{47}{6} d + 9.$$
-$\tfrac{1}{384} d^6 + O(d^5) $.
-\end{prop}
-
-%that every $(g,d)$-map graph $G$ has queue-number  $$\qn(G)\le \max\{\tfrac32 gd^4 + O(gd^3), \, \tfrac{9}{2}d^4 +O(d^3) \} \,  2^{d^6/384 + O(d^5)} .$$ previous bounds obtained by \citet{dujmovic.joret.ea:planar} previously proved the bound of  $O(g^{d^2/4+2})$.
+Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar.
+This shows that map graphs admit bounded layered partitions, but we get much better bounds by constructing a shortcut system directly.
+The following lemma is reminiscent of the characterisation of $(g,d)$-map graphs in terms of the half-square of bipartite graphs \citep{CGP02,dujmovic.eppstein.ea:structure}.
+
+%Consider a bipartite graph $H$ with bipartition $\{A,B\}$.
+Define the \emph{half-square} graph $H^2[A]$ with vertex set $A$, where two vertices in $A$ are adjacent if and only if they have a common neighbour in $B$.
+\citet{qqq} proved that a graph $G$ is a $(g,d)$-map graph if and only if $G$ is isomorphic to $H^2[A]$ for some bipartite graph $H$ with Euler genus at most $g$ and bipartition $\{A,B\}$, where vertices in $B$ have maximum degree at most $d$.
+This was proved earlier in the case $g=0$ by \citep{CGP02}.
+
+\begin{lem}
+\label{MapShortcut}
+Every $(g,d)$-map graph $G$ is a subgraph of $G_1^\PP$ for some graph $G_1$ with Euler genus at most $g$ and some $(2,\tfrac12 d(d-3) )$-shortcut system $\PP$ for $G_1$. 
+\end{lem}
+
+\begin{proof}
+Let $G$ be a $(g,d)$-map graph.
+So there is a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
+Let $N$ be the set of nations.
+Then $V(G)=N$ where two vertices are adjacent in $G$ if the corresponding nation faces of $G_0$ share a vertex.
+Let $G_1$ be the graph with $V(G_1):=V(G_0) \cup N$, where distinct vertices $v,w\in N$ are adjacent in $G_1$ if the boundaries of the corresponding nations have an edge of $G_0$ in common, and $v\in V(G_0)$ and $w\in N$ are adjacent in $G_1$ if $v$ is on the boundary of the nation corresponding to $w$.
+Observe that $G_1$ embeds in the same surface as $G_0$ with no crossings, and that each vertex in $V(G_0)$ has degree at most $d$ in $G_1$.
+Consider an edge $vw\in E(G)$.
+If the nations corresponding to $v$ and $w$ share an edge of $G_0$, then $vw$ is an edge of $G_1$.
+Otherwise,  $v$ and $w$ have a common neighbour $x$ in $V(G_0)$.
+In the latter case, let $P_{vw}$ be the path $(v,x,w)$.
+Let $\PP$ be the set of all such paths $P_{vw}$.
+Each vertex $x\in V(G_0)$ is the middle vertex on at most $\tfrac12 d(d-3)$  paths in $\PP$.
+Thus $\PP$ is a $(2,\tfrac12 d(d-3))$-shortcut system for $G_1$, and by construction, $G$ is a subgraph of $G_1^\PP$.
+\end{proof}
+
+\cref{PlanarPartition}(b)  and \cref{SurfacePartition}(b) and \cref{modern-major-general,MapShortcut} imply:
+
+\begin{thm}
+\label{MapPartition}
+Every $(g,d)$-map graph has an $H$-partition with layered width at most 
+$7d(d-3)\, \max\{2g,3\}$ for some graph $H$ with treewidth at most $\binom{2+4}{4}-1=14$ 
+or treewidth at most $\binom{2+3}{3}-1=9$ if $g=0$.
+%
+% an $H$-partition with layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most $4$; 
+%
+%  Let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\PP$ be a $(k,d)$-shortcut system for $G$.
+Then $G^\PP$ has a layered $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ with treewidth at most $\binom{k+t}{t}-1$.
+%
+\end{thm}
+
+\cref{MapPartition,non-repetitive} imply the first known constant upper bound on the
+non-repetitive chromatic number of map graphs:
+
+\begin{cor}
+Every $(g,d)$-map graph $G$ has non-repetitive chromatic number 
+$\pi(G) \le  7 \cdot 4^{15} \, d(d-3)\, \max\{2g,3\}$, and if $g=0$ then 
+$\pi(G) \le  21 \cdot 4^{10} \, d(d-3)$.
+%\label{non-repetitive} If a graph $G$ has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$, then $\pi(G)\le \ell\, 4^{t+1}$.
+\end{cor}
+
+\cref{p-centered,MapPartition} imply that the first known bounds on the $p$-centred chromatic number of map graphs that are polynomial in $p$.
+
+\begin{cor}
+Every $(g,d)$-map graph has $p$-centred chromatic number  
+$$\chi_p(G)\le 105\, d(d-3)\, \max\{2g,3\}\, (p+1) \tbinom{p+14}{14}\enspace,$$
+and every plane $d$-map graph has $p$-centred chromatic number  
+$$\chi_p(G)\le 210\,d(d-3)\, (p+1) \tbinom{p+9}{9}\enspace.$$
+\end{cor}
+
+\cref{MapPartition,qn,qn-tw} imply the best known bounds on the queue-number of map graphs:
+
+\begin{cor}
+Every $(g,d)$-map graph has queue-number at most $ 21\cdot 2^{14} \, d(d-3)\, \max\{2g,3\} $, 
+and every plane $d$-map graph  has queue-number at most 
+$ 63 \cdot 2^9\, d(d-3)$. 
+%If a graph $G$ has a layered $H$-partition of layered width $\ell$, then 
+%$$\qn(G)\le 3 \ell ( 2^t- 1/2) \enspace.$$
+\end{cor}
+
+%an upper bound on the queue-number of $(g,d)$-map graphs that is better than the previous bound obtained by \citet{dujmovic.joret.ea:planar} when $g\gg d$.
+For example, since every $(g,4)$-map graph is $(g,1)$-planar, \cref{qn,1-planar,g-1-planar} imply that every 4-map graph has queue-number at most $495$,  \note{PM}{This part is no longer true/tight: and every $(g,4)$-map graph has queue-number at most $2232 g$.}
+
+%that every $(g,d)$-map graph $G$ has queue-number  $$\qn(G)\le \max\{\tfrac32 gd^4 + O(gd^3), \, \tfrac{9}{2}d^4 +O(d^3) \} \,  2^{d^6/384 + O(d^5)} \enspace.$$ previous bounds obtained by \citet{dujmovic.joret.ea:planar} previously proved the bound of  $O(g^{d^2/4+2})$.
 Our result provides a better bound when $g\gg d$.
 For example, with $d=4$, every $(g,4)$-
-
-\cref{MapGraphs,qn} imply an upper bound on the queue-number of $(g,d)$-map graphs that is better than the previous bound obtained by \citet{dujmovic.joret.ea:planar} when $g\gg d$.
-For example, since every $(g,4)$-map graph is $(g,1)$-planar, \cref{qn,1-planar,g-1-planar} imply that every 4-map graph has queue-number at most $495$,
-\note{PM}{This part is no longer true/tight: and every $(g,4)$-map graph has queue-number at most $2232 g$.}
-
-\cref{MapGraphs,non-repetitive} imply that every $(g,d)$-map graph $G$ has non-repetitive chromatic number 
-$$\pi(G)\le\max\{(\tfrac{3}{4}+o(1)) g(d-2)^4, (\tfrac{9}{8}+o(1)) (d-2)^4\} 4^{ (\frac{1}{384}+o(1))(d-2)^6+1}.$$
-This is the first known constant upper bound (for fixed $g,d$).
-
-\cref{p-centered,MapGraphs} imply that every $(g,d)$-map graph satisfies $\chi_p(G)\le \ell (p+1)(t+1)\binom{p+t}{t}$ where $\ell:=\max\{(\tfrac{3}{4}+o(1)) g(d-2)^4, (\tfrac{9}{8}+o(1)) (d-2)^4\}$ and $t:=  \tfrac{1}{384} d^6 + O(d^5)$.
-These are the first bounds on $\chi_p(G)$ that are polynomial in $p$ for $(g,d)$-map graphs with $d$ fixed.
 
@@
 \begin{lem}
-\label{StringGraph}
-Every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar. 
-%\textcolor{red}{Eliminate this lemma, since \cref{DrawG} shows the same result via a result of \citet{dujmovic.joret.ea:planar}}
+\label{StringShortcut}
+Every $(g,\delta)$-string graph $G$ is a subgraph of $G_0^\PP$ for some graph $G_0$ with Euler genus at most $g$ and some $(\delta+1,\delta+1 )$-shortcut system $\PP$ for $G_0$. 
+%Moreover, every internal vertex on the paths in $\PP$ has degree 4 in $G_0$. 
 \end{lem}
@@
 \begin{proof}
-Consider some $(\Sigma,\delta)$-string representation, $f$, of $G$.
+Let $\mathcal{C}=\{C_v:v\in V(G)\}$ be a set of curves in a surface of Euler genus at most $g$ whose intersection graph is $G$.
+Let $G_0$ be the graph obtained by adding a vertex at the intersection point of every pair of curves in $\mathcal{C}$ that intersect,  where two such consecutive vertices on a curve $C_v$ are adjacent in $G_0$.
+For each vertex $v\in V(G)$, introduce a new vertex called $v$ on $C_v$ between the
+$\floor{\frac{\delta}{2}}$-th vertex already on $C_v$ and the $(\floor{\frac{\delta}{2}}+1)$-th such vertex.
+For each edge $vw$ of $G$, there is a path $P_{vw}$ of length at most $2\ceil{\frac{\delta}{2}}\leq \delta+1$ in $G_0$ between $v$ and $w$.
+Let $\PP$ be the set of all such paths $P_{vw}$.
+Consider a vertex $x$ in $G_0$ that is an internal vertex on some path in $\PP$.
+Then $x$ is at the intersection of $C_v$ and $C_w$ for some edge $vw\in E(G)$.
+If some path $P\in \PP$ passes through $x$, then $P=P_{vu}$ for some edge $vu$ incident to $v$, or $P=P_{wu}$ for some edge $wu$ incident to $w$.
+At most $\ceil{\frac{\delta}{2}}$ paths in $\PP$ corresponding to edges incident to $v$ pass through $x$, and similarly for edges incident to $w$.
+Thus at most $2\ceil{\frac{\delta}{2}}\leq\delta+1$ paths in $\PP$ use $x$ as an internal vertex.
+Thus $\PP$ is a $(\delta+1,\delta+1)$-shortcut system for $G_0$, and by construction, $G \subseteq G_0^\PP$.
+\end{proof}
+
+A similar proof to that of \cref{StringShortcut} shows that every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.
+
+%\begin{lem}
+%\label{StringGraph}
+%Every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar. 
+%\end{lem}
+%
+%\begin{proof}
+%Consider some $(\Sigma,\delta)$-string representation, $f$, of $G$.
 We will describe an embedding of $G$ into $\Sigma$ that is an $(\Sigma,k)$-plane graph.
@@
 Therefore the embedding of any edge $vw\in E(G)$ crosses at most $2\delta^2$ other edges.
-\end{proof}
-
-\cref{StringGraph,g-k-planar} with $k=2\delta^2$ implies:
-
-% \noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}
-
-\begin{prop}
-Every $(g,\delta)$-string graph has a layered $H$-partition of layered width at most 
-% $\max\{(48+o(1))g\delta^4, (72+o(1))\delta^4\}$ in which $H$ has treewidth at most $(\frac{4}{3}+o(1)) \delta^6$.
-$\max\{(48+o(1))g\delta^4, (72+o(1))\delta^4\}$ in which $H$ has treewidth at most $(\frac{2}{3}+o(1)) \delta^8$.
-% tw = 2\delta+5 choose 4 - 1  , for g > 1.
-\end{prop}
+%\end{proof}
+
+\cref{StringShortcut,modern-major-general} and \cref{PlanarPartition}(b) and \cref{SurfacePartition}(b) imply:
+
+\begin{thm}
+\label{StringPartition}
+Every $(g,\delta)$-string graph has a layered $H$-partition of layered 
+width at most $\max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$ 
+for some graph $H$ with treewidth at most $\binom{ \delta+5}{4}-1$, 
+and treewidth at most $\binom{ \delta+4}{3}-1$ if $g=0$. 
+%
+%Every $(g,\delta)$-string graph $G$ is a subgraph of $G_0^\PP$ for some graph $G_0$ with Euler genus at most $g$ and some $(\delta+1,\delta+1 )$-shortcut system $\PP$ for $G_0$. 
+%
+ % Let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\PP$ be a $(k,d)$-shortcut system for $G$.
+Then $G^\PP$ has a layered $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ with treewidth at most $\binom{k+t}{t}-1$.
+%
+% an $H$-partition with layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most $4$; 
+\end{thm}
+
+\cref{p-centered,StringPartition} imply that the first known bounds on the $p$-centred chromatic number of map graphs that are polynomial in $p$.
+
+\begin{cor}
+Every $(g,\delta)$-string graph has $p$-centred chromatic number  
+$$\chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1) 
+\tbinom{ \delta+5}{4} \binom{p+\binom{ \delta+5}{4}-1}{\binom{ \delta+5}{4}-1}\enspace,$$
+and every plane $(0,\delta)$-string graph has $p$-centred chromatic number  
+$$\chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1) 
+\tbinom{ \delta+4}{3}\binom{p+\binom{ \delta+4}{3}-1}{\binom{ \delta+4}{3}-1}\enspace.$$
+\end{cor}
+%\label{p-centered}
+%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
+
+\cref{StringPartition,qn,qn-tw} imply the following bounds on the queue-number of map graphs:
+
+\begin{cor}
+Every $(g,\delta)$-string graph has queue-number at most 
+$$3 \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)  2^{\binom{ \delta+5}{4}-1} \enspace.$$
+and every plane $(0,\delta)$-map graph  has queue-number at most 
+$$ 9 (\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)  2^{\binom{ \delta+4}{3}-1} \enspace.$$
+%If a graph $G$ has a layered $H$-partition of layered width $\ell$, then 
+%$$\qn(G)\le 3 \ell ( 2^t- 1/2) \enspace.$$
+\end{cor}
 
@@
 
+\cref{non-repetitive,StringPartition} also give bounds on the non-repetitive chromatic number of $(g,\delta)$-string graphs, but the bound is weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ by a result of \citet{DJKW16}.
+
 %%%%%%%%%%
 \subsection{Powers of Bounded Degree Graphs}
+\label{Powers}
 
 Recall that the \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
-If $G$ is planar with maximum degree $\Delta$, then $G^k$ is $(k+1)\Delta^{2(k-1)} $-planar by \cref{DrawG}.
+If $G$ is planar with maximum degree $\Delta$, then $G^k$ is $2k(k+1)\Delta^{k}$-planar by \cref{DrawG}.
 Thus we can immediately conclude that bounded powers of planar graphs of bounded degree admit bounded layered partitions.
@@
 Say $P_{xy}$ uses some vertex $v$ as an internal vertex.
-If $\dist_G(v,x)=i$ and $\dist_G(v,y)=j$ then $i\in\{1,\dots,k\}$ and $j\in\{1,\dots,k-i\}$.
+If $\dist_G(v,x)=i$ and $\dist_G(v,y)=j$, then $i,j\in\{1,\dots,k-1\}$ and $i+j\leq k$.
 The number of vertices at distance $i$ from $v$ is at most $\Delta^i$.
 Thus the number of paths in $\PP$ that use $v$ as an internal vertex is at most
-$$\sum_{i=1}^k \sum_{j=1}^{k-i} \Delta^i\Delta^j 
-= \sum_{i=1}^k \Delta^i \sum_{j=1}^{k-i} \Delta^j 
-< \sum_{i=1}^k \Delta^i ( 2 \Delta^{k-i} )
-= 2k \Delta^k.$$
+$$\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i\Delta^j 
+= \sum_{i=1}^{k-1} \Delta^i \sum_{j=1}^{k-i} \Delta^j 
+< \sum_{i=1}^{k-1} \Delta^i ( 2 \Delta^{k-i} )
+< 2k \Delta^k\enspace.$$
 Hence $\PP$ is a $(k, 2k \Delta^k)$-shortcut system. 
@@
 
-\cref{SurfacePartition}(b) and \cref{modern-major-general,PowerShortcut} imply:
+\cref{PlanarPartition}(b) and \cref{SurfacePartition}(b) and \cref{modern-major-general,PowerShortcut} imply the following result, which with \cref{PartitionProduct} implies \cref{kPowerBasic} in \cref{Introduction}.
 
@@
 \label{PowerPartition}
-Let $G$ be a graph of Euler genus $g$ and maximum degree $\Delta$. Let $k$ be a positive integer. 
-Then $G^k$ has a layered $H$-partition of layered width at most $\max\{2g,3\} (2k^4+6k^2) \Delta^{k}$ for some graph $J$ with treewidth at most $\binom{k+4}{4}-1$.
+Fix $k\geq 1$. Let $G$ be a graph of Euler genus $g$ and maximum degree $\Delta$. 
+Then $G^k$ has a layered $H$-partition of layered width at most $\max\{2g,3\} (2k^4+6k^2) \Delta^{k}$ 
+for some graph $H$ with treewidth at most $\binom{k+4}{4}-1$, and treewidth 
+at most $\binom{k+3}{3}-1$ if $g=0$. 
 \end{thm}
@@
 %%%%%%
-\item $G^k$ is $(g, (k+1)\Delta^{2(k-1)} )$-planar, 
+\item $G^k$ is $(g, 2k(k+1)\Delta^{k} )$-planar, 
 %%%%%%
@@
 \item $G^k$ has  $p$-centred chromatic number at most 
-$$\max\{2g,3\}(2k^4+6k^2) \Delta^{k}  (p+1) (t+1)\tbinom{p+t}{t}$$
-for some integer $t \leq \binom{k+4}{4}-1$. 
+$$\max\{2g,3\}(2k^4+6k^2) \Delta^{k}  (p+1) \tbinom{k+4}{4} 
+\binom{p+\binom{k+4}{4}-1}{\binom{k+4}{4}-1}\enspace.$$
 %\label{p-centered} Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
@@
 
-Also note that the above results in conjunction with the comments about separators in \cref{PriorWork} imply that for fixed integers $g,k,\Delta$, the $k$-th power of a graph with Euler genus $g$ and maximum degree $\Delta$ has $O(\sqrt{n})$ treewidth  and thus admits balanced separators of order $O(\sqrt{n})$.
+%Also note that the above results in conjunction with the comments about separators in \cref{PriorWork} imply that for fixed integers $g,k,\Delta$, the $k$-th power of a graph with Euler genus $g$ and maximum degree $\Delta$ has $O(\sqrt{n})$ treewidth  and thus admits balanced separators of order $O(\sqrt{n})$.
 \note{PM}{How interesting is this last note?
@@
 
-In this section, we show that $k$-nearest neighbour graphs of point in the plane are $O(k^2)$-planar.
+In this section, we show that $k$-nearest neighbour graphs of point sets in the plane are $O(k^2)$-planar.
 For two points $x,y\in\R^2$, let $d_2(x,y)$ denote the Euclidean distance between $x$ and $y$.
@@
 
-We make use of the following well-known observation (see e.g., \citet[Lemma~2]{bose.morin.ea:routing}):
+We make use of the following well-known observation (see for example, \citet[Lemma~2]{bose.morin.ea:routing}):
 \begin{obs}
@@
   \end{compactenum}
-  If $xy$ is of Type~C, then \cref{convex} implies that (without loss of generality) $\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)$.
+  If $xy$ is of Type~C, then \cref{convex} implies that (without loss of generality) 
+  $$\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)\enspace.$$ 
    In this case, we call $x$ a Type~C vertex.
@@
 \cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
-However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (6+o(1))k$ by a result of \citet{DJKW16}.
-
-\section{Conclusions}
-
-We have shown that $k$-planar graphs and several related graph families have layered $H$-partitions of layered width $O(k^2)$ in which $H$ has treewidth $O(k^3)$.
+However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (36+o(1))k^2$ by a result of \citet{DJKW16}.
+
+\section{An Open Problem}
+
+\cref{k-planar} shows that $k$-planar graphs have layered $H$-partitions of layered width $O(k^2)$ in which $H$ has treewidth $O(k^3)$.
 We conclude with the following question:  Does there exist a function $f:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph has a layered $H$-partition of layered width at most $f(k)$ in which $H$ has treewidth at most $C$?
@@
 \bibliography{k-planar}
+\end{document}
+
 \newpage
@@
 \leq
-\frac{1}{6} (  3^{d+1} (k \ceil{\sqrt{d}}^d + 1 ) -3 ).$$
+\frac{1}{6} (  3^{d+1} (k \ceil{\sqrt{d}}^d + 1 ) -3 )\enspace.$$
 
@@
 Then 
-$$\tw(G) \leq 2  ( \tw(H)+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1.$$
-$$\pw(G) \leq 2  ( \pw(H)+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1.$$
+$$\tw(G) \leq 2  ( \tw(H)+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1\enspace.$$
+$$\pw(G) \leq 2  ( \pw(H)+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1\enspace.$$
 \end{lem}
@@
 Let $\langle{V^i_0,V^i_1,\dots}\rangle$ be the layering of $G$ determined by the $i$-th dimension. 
-Let $$m:= \ceil*{ \left(\frac{dn}{t+1} \right)^{1/(d+1)} } .$$ 
+Let $$m:= \ceil*{ \left(\frac{dn}{t+1} \right)^{1/(d+1)} } \enspace.$$ 
 For $\alpha\in[0,m-1]$, let $V^{i,\alpha} := \bigcup\{ V^i_j : j\equiv \alpha \pmod{m} \}$. 
@@
 For every unit disc graph $G$ in $\mathbb{R}^d$ on $n$ vertices and with no $(k+1)$-clique,
-%$$\pw(G) \leq 2  k^{ 1/(d+1) }  \ceil{\sqrt{d}}^{d/(d+1) } (dn)^{1- 1/(d+1)}   -  1.$$
-$$\pw(G) \leq 2  (2k)^{ 1/d } \ceil{\sqrt{d}} ((d-1)n)^{1- 1/d}   -  1.$$
+%$$\pw(G) \leq 2  k^{ 1/(d+1) }  \ceil{\sqrt{d}}^{d/(d+1) } (dn)^{1- 1/(d+1)}   -  1\enspace.$$
+$$\pw(G) \leq 2  (2k)^{ 1/d } \ceil{\sqrt{d}} ((d-1)n)^{1- 1/d}   -  1\enspace.$$
 \end{cor}

### 2019-07-09 k-planar.tex
Added David's figure

@@
 We note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n$ vertices has at least $kn/2$ edges and at most $kn$ edges.
-For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is at $\Omega(k^2)$.
+For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is $\Omega(k^2)$.
 

### 2019-07-09 k-planar.tex
Vida's changes and other stuff

@@
 Such graphs provide a natural generalisation of planar graphs, and are important in graph drawing research; see the recent bibliography on 1-planar graphs and the 140 references therein \citep{kobourov.liotta.ea:annotated}.
-The present paper studies the structure of $k$-planar graphs (and other more general classes).
-Various structural results about planar graphs generalise for $k$-planar graphs.
+The present paper studies the structure of $k$-planar graphs and other more general classes of graphs.
+% Various structural results about planar graphs generalise for $k$-planar graphs.
 For example, \citet{FP08} generalised the Lipton-Tarjan separator theorem to show that every $k$-planar graph with $n$ vertices has a balanced separator of order $O(\sqrt{kn})$.\note{PM}{I must be missing something.
@@
 }
-
-The main results of this paper generalise the following recent theorem of \citet{dujmovic.joret.ea:planar} to the setting of $k$-planar graphs (and other more general classes).
+The main results of this paper generalise the following recent theorem of \citet{dujmovic.joret.ea:planar} to this setting, and are the first results of this type for non-proper-minor-closed graph classes.
 
@@
 Here $\boxtimes$ is the strong product, and treewidth is an invariant that measures how `tree-like' a given graph is; see \cref{Tools} for formal definitions and see \cref{ProductExample} for an example.
-So, loosely speaking, \cref{PlanarBasic} says that every planar graph is contained in the product of a tree-like graph and a path. This enables combinatorial results for graphs of bounded 
+Loosely speaking, \cref{PlanarBasic} says that every planar graph is contained in the product of a tree-like graph and a path. This enables combinatorial results for graphs of bounded 
 \begin{wrapfigure}{r}{0.5\textwidth} 
@@
 treewidth to be generalised for planar graphs (with different constants).
-\cref{kPlanarBasic} has been the key tool in solving several well-known open problems.
+\cref{PlanarBasic} has been the key tool in solving several well-known open problems.
 In particular, \citet{dujmovic.joret.ea:planar} proved that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992), and \citet{dujmovic.esperet.ea:planar}  proved that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002).
@@
 \begin{compactitem}
-\item every path in $\mathcal{P}$ has length at most $k$, and
+\item every path in $\mathcal{P}$ has length at most $k$,\footnote{A path of length $k$ consists of $k$ edges and $k+1$ vertices.} and
 \item for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
@@
 
-\cref{ShortcutProduct,AddDummy} immediately imply \cref{kPlanarBasic} with $O(k^{11})$ instead of $O(k^5)$. Some further observations reduce the bound to $O(k^5)$; see \cref{Tools}.
+\cref{PlanarBasic,ShortcutProduct,AddDummy} immediately imply \cref{kPlanarBasic} with $O(k^{11})$ instead of $O(k^5)$. Some further observations reduce the bound to $O(k^5)$; see \cref{Tools}.
 
@@
 Here is one example.
-The \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
-If $G$ has maximum degree $\Delta$, then $G^k = G^\PP$ for some $(k,2k\Delta^{k})$-shortcut system $\PP$; see \cref{PowerShortcut}.
+The \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.\footnote{For a graph $G$ and two vertices $v,w\in V(G)$, we use $\dist_G(v,w)$ to denote the length of a shortest path with endpoints $v$ and $w$.
+If $v$ and $w$ are in different components of $G$ then we use the convention that $\dist_G(v,w)=\infty$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\PP$ for some $(k,2k\Delta^{k})$-shortcut system $\PP$; see \cref{PowerShortcut}.
 \cref{kPlanarBasic,ShortcutProduct} then imply:
@@
 
-Note that the dependence on $k$ in the bound on the treewidth of $H$ in \cref{kPowerBasic} is improved from $O(k^{12})$ to $O(k^7)$ in \cref{Powers}.
+% Note that the dependence on $k$ in the bound on the treewidth of $H$ in \cref{kPowerBasic} is improved from $O(k^{12})$ to $O(k^7)$ in \cref{Powers}. 
+% PM: I removed this because the bound is exponential in $k$ anyway. You only see the difference when you separate treewidth and layered width.
 
@@
 
-\cref{Examples} presents several examples of graph classes that can be obtained from a shortcut system applied to a planar graphs, including graph powers, map graphs, string graphs, and nearest neighbour graphs.
-All of the above results are applicable to these examples.
-
-All of the above results apply, where instead of planar graphs, we consider graphs of bounded Euler genus.
+\cref{Examples} presents several examples of graph classes that can be obtained from a shortcut system applied to a planar graph, including graph powers, map graphs, string graphs, and nearest neighbour graphs.
+All of these results also apply, where instead of planar graphs, we consider graphs of bounded Euler genus.
+All of the applications in \cref{Applications}  are applicable to these examples.
 
@@
 
-This section defines concepts from from the literature that will be important for our work.
+This section defines concepts and results from the literature that will be important for our work.
 
@@
 
+\subsection{Previous Work on Partitions of Minor-Closed Classes}
 \citet{dujmovic.joret.ea:planar} introduced the study of partitions with bounded layered width such that the quotient has bounded treewidth.
@@
 
-Apex-minor-free graphs are the largest minor-closed class that admit bounded layered partitions~\citep{dujmovic.joret.ea:planar}.
+\section{New Results}
+
+Apex-minor-free graphs, addressed by \cref{ApexMinorFree}, are the largest minor-closed class that admit bounded layered partitions~\citep{dujmovic.joret.ea:planar}.
 However, the family of $k$-planar graphs is not minor-closed.
@@
 
-\section{Our Results}
-
 The following result is the main theorem of the paper.
-Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so to does $G^\PP$ for every $(k,d)$-shortcut system $\PP$.
+Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so to does $G^\PP$ for every $(k,d)$-shortcut system $\PP$ of $G$.
 
@@
 
-The following result shows that $k$-planar graphs admit bounded layered partitions. With \cref{PartitionProduct,MakeWidth1}, this proves \cref{kPlanarBasic}. % with $c_k\leq(4+o(1))k^5$.
+The relationship between $k$-planar graphs and $(k+1,2)$-shortcut systems of planar graphs shows that \cref{modern-major-general} has immediate applications to $k$-planar graphs.
+The following result shows that $k$-planar graphs admit bounded layered partitions.
+With \cref{PartitionProduct,MakeWidth1}, this proves \cref{kPlanarBasic}.
+% with $c_k\leq(4+o(1))k^5$.
 
@@
 %%%%%%%%%%%%%%%%
-\subsection{Prior Work}
+\subsection{Previous Work on the Structure of $(g,k)$-Planar Graphs}
 \label{PriorWork}
@@
 \citet{dujmovic.joret.ea:planar} proved that $G$ has layered treewidth at most $(k+1)\ell$.
-So any property that holds for graphs of bounded layered treewidth also holds for $G$.
+% So any property that holds for graphs of bounded layered treewidth also holds for $G$.
 For example, Norin proved that every $n$-vertex graph with layered treewidth at most $t$ has treewidth less than $2\sqrt{tn}$ (see \citep{dujmovic.morin.ea:layered}).

### 2019-07-09 k-planar.tex
Final preparations before submission

@@
 \usepackage{stmaryrd}
+\usepackage{thm-restate}
 
@@
 \usepackage[noabbrev,capitalise]{cleveref}
+
+\usepackage{doi} % To make plainnat handle doi's properly
+
 \allowdisplaybreaks
@@
 
-\newcommand{\note}[2]{{\color{red}[#1:~#2]}}
+% \newcommand{\note}[2]{{\color{red}[#1:~#2]}}
+% \newcommand{\notex}[2]{{\color{blue}[#1:~#2]}}
+\newcommand{\notex}[2]{}
 
@@
 \renewcommand{\leq}{\leqslant}
+
+% \newcommand*{\doi}[1]{\href{https://dx.doi.org/#1}{\verb+doi: #1+}}
 
@@
 Dujmovi\'c~et~al.~(FOCS 2019) recently proved that every planar graph is a subgraph of the strong product of a graph of bounded treewidth and a path.
-This tool has been used to solve longstanding problems on queue layouts, non-repetitive colouring, and $p$-centred colouring.
-%, and graph encoding.
+This tool has been used to solve longstanding problems on queue layouts, non-repetitive colouring, $p$-centred colouring, and implicit graph encoding.
 We generalise this result for $k$-planar graphs, where a graph is \emph{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings.
@@
 
-\tableofcontents
-\newpage
+% \tableofcontents
+% \newpage
+
+%\setcounter{\page}{1}
 \section{Introduction}
@@
 The present paper studies the structure of $k$-planar graphs and other more general classes of graphs.
-% Various structural results about planar graphs generalise for $k$-planar graphs.
-For example, \citet{FP08} generalised the Lipton-Tarjan separator theorem to show that every $k$-planar graph with $n$ vertices has a balanced separator of order $O(\sqrt{kn})$.\note{PM}{I must be missing something.
+\notex{PM}{I removed the following sentences because neither Vida nor could find anything about $k$-planar graphs in \cite{FP08} and, in any case, the proof is easy.
+This could easily upset a referee.
+% \note{PM}{I must be missing something.
 The planarized graph $G'$ has $O(kn)$ vertices.
@@
 }
-The main results of this paper generalise the following recent theorem of \citet{dujmovic.joret.ea:planar} to this setting, and are the first results of this type for non-proper-minor-closed graph classes.
+Various structural results about planar graphs generalise for $k$-planar graphs.
+For example, \citet{FP08} generalised the Lipton-Tarjan separator theorem to show that every $k$-planar graph with $n$ vertices has a balanced separator of order $O(\sqrt{kn})$.}
+The main results of this paper generalise the following recent theorem of \citet{dujmovic.joret.ea:planar} to a setting that includes $k$-planar graphs.
+ % as well as some other interesting non-minor-closed graph families.
 
@@
 Here $\boxtimes$ is the strong product, and treewidth is an invariant that measures how `tree-like' a given graph is; see \cref{Tools} for formal definitions and see \cref{ProductExample} for an example.
-Loosely speaking, \cref{PlanarBasic} says that every planar graph is contained in the product of a tree-like graph and a path. This enables combinatorial results for graphs of bounded 
+Loosely speaking, \cref{PlanarBasic} says that every planar graph is contained in the product of a tree-like graph and a path.
+This enables combinatorial results for graphs of bounded
 \begin{wrapfigure}{r}{0.5\textwidth} 
@@
 \cref{PlanarBasic} has been the key tool in solving several well-known open problems.
-In particular, \citet{dujmovic.joret.ea:planar} proved that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992), and \citet{dujmovic.esperet.ea:planar}  proved that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002).
+In particular, \citet{dujmovic.joret.ea:planar} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992), \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002), \citet{bonamy.gavoille.ea:shorter} use it to find shorter implicit representations of planar graphs.
 
@@
 
-We actually prove a substantially stronger result than \cref{kPlanarBasic} using the following definition.
+Although $k$-planar graphs are the most high-profile target for a generalization \cref{PlanarBasic}, we actually prove a substantially stronger result than \cref{kPlanarBasic} using the following definition.
 A collection $\mathcal{P}$ of paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if:
@@
 
-The proof of \cref{AddDummy} is trivial: Given a $k$-planar graph $G'$, create a planar graph $G$ by adding a dummy vertex at each crossing point.
-For each edge $vw\in E(G')\setminus E(G)$ there is a path $P$ in $G$ between $v$ and $w$ of length at most $k+1$.
+The proof of \cref{AddDummy} is trivial: Given a $k$-plane embedding of a graph $G'$, create a planar graph $G$ by adding a dummy vertex at each crossing point.
+For each edge $vw\in E(G')$ there is a path $P$ in $G$ between $v$ and $w$ of length at most $k+1$.
 Let $\PP$ be the set of such paths $P$.
 For each vertex $v$ of $G$, at most two paths in $\PP$ use $v$ as an internal vertex (since no original vertex of $G'$ is an internal vertex of a path in $\PP$).
-Thus $\PP$ is a $(k+1,2)$-shortcut system for $G$, such that $G'=G^\PP$.
+Thus $\PP$ is a $(k+1,2)$-shortcut system for $G$, such that $G'\subseteq G^\PP$.
 
@@
 Here is one example.
-The \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.\footnote{For a graph $G$ and two vertices $v,w\in V(G)$, we use $\dist_G(v,w)$ to denote the length of a shortest path with endpoints $v$ and $w$.
-If $v$ and $w$ are in different components of $G$ then we use the convention that $\dist_G(v,w)=\infty$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\PP$ for some $(k,2k\Delta^{k})$-shortcut system $\PP$; see \cref{PowerShortcut}.
+The \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.\footnote{For a graph $G$ and two vertices $v,w\in V(G)$, $\dist_G(v,w)$  denotes the length of a shortest path, in $G$, with endpoints $v$ and $w$.
+We define $\dist_G(v,w):=\infty$ if $v$ and $w$ are in different connected components of $G$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\PP$ for some $(k,2k\Delta^{k})$-shortcut system $\PP$; see \cref{PowerShortcut}.
 \cref{kPlanarBasic,ShortcutProduct} then imply:
@@
 
-These theorems have applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, and $p$-centred colouring  \citep{micek:personal},
+These theorems have applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, $p$-centred colouring  \citep{micek:personal}, and succinct graph representations \citep{bonamy.gavoille.ea:shorter}, some of which
 %, and graph encoding \citep{BGP}. 
@@
 All of these results also apply, where instead of planar graphs, we consider graphs of bounded Euler genus.
-All of the applications in \cref{Applications}  are applicable to these examples.
+All of the applications discussed in \cref{Applications} work on these graph classes.
 
@@
 
-While strong products enable an easy statement of the theorems in \cref{Introduction}, to prove such results it is helpful to work with layerings and partitions, which we now introduce.
+While strong products enable concise statements of the theorems in \cref{Introduction}, to prove such results it is helpful to work with layerings and partitions, which we now introduce.
+
 A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle V_0,V_1,\ldots\rangle$ such that $\{V_0,V_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in V_i$ and $w\in V_j$ then $|j-i|\leq 1$.
 For any partition $\mathcal{P}=\{S_1,\ldots,S_p\}$ of $V(G)$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
-To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an \emph{$H$-partition} and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.\footnote{An alternative definition of a layering of $G$ is an $H$-partition of $G$ where $H$ is a path.}  A layered $H$-partition $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
+To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an \emph{$H$-partition} and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.
+
+For any partition $\mathcal{P}$ of $V(G)$ and any layering $\mathcal{L}$ of $G$ we define the \emph{layered width} of $\mathcal{P}$ with respect to $\mathcal{L}$ as $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
+For any partition $\mathcal{P}$ of $V(G)$, we define the \emph{layered width} of $\mathcal{P}$ as the minimum, over all layerings $\mathcal{L}$ of $G$, of the layered width of $\mathcal{P}$ with respect to $\mathcal{L}$.
+
+% \footnote{An alternative definition of a layering of $G$ is an $H$-partition of $G$ where $H$ is a path.}
+
+% A \emph{layered $H$-partition} $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
 The \emph{layered width} of $(\mathcal{L},\mathcal{P})$ is $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
 
-\note{DW}{I  prefer to not define ``layered $H$-partition''.
+% \note{DW}{I  prefer to not define ``layered $H$-partition''.
 Just define the ``layered width'' of an $H$-partition, which is what we did in \citep{dujmovic.joret.ea:planar}.
@@
 \subsection{Previous Work on Partitions of Minor-Closed Classes}
-\citet{dujmovic.joret.ea:planar} introduced the study of partitions with bounded layered width such that the quotient has bounded treewidth.
-They defined a class $\mathcal{G}$ of graphs to \emph{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition with layered width at most $\ell$ such that $H$ has treewidth at most $t$.
+
+\citet{dujmovic.joret.ea:planar} introduced the study of partitions with bounded layered width such that the quotient has some additional desirable property, like small treewidth. 
+% Say that a graph $G$ admits a $(t,\ell)$-partition if $V(G)$ has a partition $\mathcal{P}$ of layered width at most $\ell$ and the treewidth of $H:=G/\mathcal{P}$ is at most $t$. 
+They defined a class $\mathcal{G}$ of graphs to \emph{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
 
@@
 \end{thm}
-Their proof is constructive and gives a simple quadratic-time algorithm for finding these partitions.
-
+Their proof is constructive and gives a simple quadratic-time algorithm for finding these partitions and corresponding layerings.
 %Layered $H$-partitions of small layered width in which $H$ has some additional property are a useful tool for a number of graph problems.
 One such useful property is small \emph{treewidth}, which we now define.
-
+%
 % \citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number, track number, and layered treewidth.
@@
 \citet{dujmovic.joret.ea:planar} show that two common families of graphs have such $H$-partitions: planar graphs, graphs of bounded genus, and (more generally) apex-minor-free graphs.
-
+%
 % In this paper we show that another family of graphs admits layered $H$-partitions of bounded layered width in which $H$ has bounded treewidth:
-
-\citet{dujmovic.joret.ea:planar} proved the following generalisation of \cref{PlanarBasic,PlanarPartition} for graphs embeddable on other surfaces.\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
+The same authors proved the following generalisation of \cref{PlanarBasic,PlanarPartition} for graphs embeddable on other surfaces.\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
 The \textit{Euler genus} of the non-orientable surface with $c$ cross-caps is $c$.
@@
 
-\begin{thm}\label{modern-major-general}
-  Let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\PP$ be a $(k,d)$-shortcut system for $G$.
-Then $G^\PP$ has a layered $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ with treewidth at most $\binom{k+t}{t}-1$.
-\end{thm}
+\begin{restatable}{thm}{mmg}
+  %\begin{thm}
+  \label{modern-major-general}
+    Let $G$ be a graph having an $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\PP$ be a $(k,d)$-shortcut system for $G$.
+Then $G^\PP$ has a $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ with treewidth at most $\binom{k+t}{t}-1$.
+  %\end{thm}
+\end{restatable}
 
@@
 
-The relationship between $k$-planar graphs and $(k+1,2)$-shortcut systems of planar graphs shows that \cref{modern-major-general} has immediate applications to $k$-planar graphs.
+Using relationship between $k$-planar graphs and $(k+1,2)$-shortcut systems along with a direct application of \cref{modern-major-general} we can obtain a slightly weaker version of \cref{k-planar}, below.
+The minor modifications needed to obtain the stronger bound are described in \cref{sec-k-planar}.
+
+ % of planar graphs shows that \cref{modern-major-general} has immediate applications to $k$-planar graphs.
 The following result shows that $k$-planar graphs admit bounded layered partitions.
@@
 \label{k-planar}
-Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
+Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
 \end{thm}
 
-\cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with the weaker bound of $(2\times 3)((k+1)^3+3(k+1))=6k^3 + 18k^2 + 36k+24$ on the layered width. 
-\note{DW}{I would just write, ``\cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with a weaker bound on the layered width.}
-We prove the stronger bound in \cref{sec-k-planar}. 
-%Also note that \cref{k-planar,MakeWidth1} imply \cref{kPlanarBasic} with the $O(k^5)$ bound on the treewidth of $H$, as promised. 
-% PM: I removed the last line because it's already stated in the line preceding the theorem
+% \cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with the weaker bound of $O(k^3)$
+%  % $(2\times 3)((k+1)^3+3(k+1))=6k^3 + 18k^2 + 36k+24$
+% on the layered width. 
+% % \note{DW}{I would just write, ``\cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with a weaker bound on the layered width.}
+% We prove the stronger bound in \cref{sec-k-planar}. 
+% %Also note that \cref{k-planar,MakeWidth1} imply \cref{kPlanarBasic} with the $O(k^5)$ bound on the treewidth of $H$, as promised. 
+% % PM: I removed the last line because it's already stated in the line preceding the theorem
 
@@
 \label{1-planar}
-Every 1-planar graph has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
+Every 1-planar graph has an $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
 \end{thm}
@@
 A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
+\citet{grigoriev.bodlaender:algorithms} study approximation algorithms for $(g,k)$-plane graphs.
  %\citet[Theorem~22]{dujmovic.joret.ea:planar} show that every genus-$g$ graph has a layered $H$-partition of layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most 4. 
@@
 \label{gAddDummy}
-Every $(g,k)$-planar graph is isomorphic to $G^\PP$ for some graph $G$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G$. 
+Every $(g,k)$-planar graph is isomorphic to a subgraph of $G^\PP$ for some graph $G$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G$. 
 \end{obs}
@@
 \label{g-k-planar}
-Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most 
+Every $(g,k)$-planar graph has an $H$-partition of layered width at most 
 $\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
@@
 
-Note that \cref{SurfacePartition}(b) and \cref{modern-major-general} imply \cref{k-planar} with a weaker bound on the layered width. We prove the stronger bound in \cref{sec-k-planar}.
-
-Finally, we state the following corollary of \cref{ApexMinorFree,modern-major-general,MakeWidth1,PartitionProduct}.
-This is the strongest result that follows from our work and that of \citet{dujmovic.joret.ea:planar}.
+Again, a direct application of \cref{modern-major-general} using \cref{SurfacePartition}(b) implies \cref{k-planar} with a weaker bound on the layered width.
+We prove the stronger bound in \cref{sec-k-planar}.
+
+Finally, we state the following corollary of \cref{ApexMinorFree,modern-major-general,MakeWidth1}.
+This is the most general result that follows from \cref{modern-major-general} and that of \citet{dujmovic.joret.ea:planar}.
 
@@
 The purpose of this section is to prove our main result, \cref{modern-major-general}.
-This theorem shows how, given a $(k,d)$-shortcut system $\mathcal{P}$ of a graph $G$, a layered $H$-partition of $G$ can be used to obtain a layered $J$-partition of $G^{\mathcal{P}}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
+This theorem shows how, given a $(k,d)$-shortcut system $\mathcal{P}$ of a graph $G$, a $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\mathcal{P}}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
 Our main results for $k$-planar graphs (\cref{k-planar}) and $(g,k)$-planar graphs (\cref{g-k-planar}) follow; see \cref{sec-k-planar}.
@@
 
-For the remainder of this section, $G$ is a graph and $(\mathcal{L}:=\langle V_1,\ldots,V_h\rangle,\mathcal{Y}:=(Y_x: x\in V(H)))$ is a layered partition of $G$ of layered width $\ell$ where the treewidth of the quotient graph $H=G/\mathcal{Y}$ is at most $t$.
-In addition, $\mathcal{T}:=(B_x:x\in V(T))$ is a tree-decomposition of $H$ obtained by applying \lemref{nice-decomposition} to $H$.
-The following lemma shows how to interpret the pair $(\mathcal{Y},\mathcal{T})$ as a hierarchical decomposition of $G$; refer to \cref{fig:generalized-tripod}.
-
 \begin{figure}
@@
 
-\note{W}{I think \clmref{generalized-tripod} should be presented as a stand-alone lemma.
+% \notex{W}{I think \clmref{generalized-tripod} should be presented as a stand-alone lemma.
 Then we repeat the statement of \cref{modern-major-general} followed by $\\begin \{proof\} ....
@@
 
-\note{DW}{Is some sort of converse to \clmref{generalized-tripod}  also true?
-\clmref{generalized-tripod}  shows that if a graph $G$ has a layered partition with all the nice properties, then $G$ has this generalised tripod partition.
+\notex{DW}{Is some sort of converse to \cref{generalized-tripod}  also true?
+\cref{generalized-tripod}  shows that if a graph $G$ has a layered partition with all the nice properties, then $G$ has this generalised tripod partition.
 Isn't also true that if $G$ has a generalised tripod partition, then $G$ has a layered partition with all the nice properties.
 The proof is basically in \citep{dujmovic.joret.ea:planar}, right?}
-
-\begin{clm}\clmlabel{generalized-tripod}
-  For each $x\in V(T)$, let $V_x := \bigcup_{y\in V(T_x)} Y_y$, $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x:=V_x\cup F_x$.  Then, 
+\notex{PM}{Yes, if by generalized tripod partition you mean $T$ and $\mathcal{Y}=(Y_x:x\in V(T))$ that satisify \yref{thickness}, \yref{separator}, and \yref{ancestor-edge}.
+From \yref{separator} and \yref{ancestor-edge} you can construct a width-$t$ $T$-decomposition of $H=G/\mathcal{Y}$.
+By \yref{thickness}, $\mathcal{Y}$, has layered width at most $\ell$.}
+
+The following lemma shows how to interpret an $H$-partition of $G$ and a tree-partition of $H$ as a hierarchical decomposition of $G$; refer to \cref{fig:generalized-tripod}.
+
+\begin{lem}\label{generalized-tripod}
+  Let $G$ be a graph; let $\mathcal{L}:=\langle V_1,\ldots,V_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \lemref{nice-decomposition}.
+For each $x\in V(T)$, let $V_x := \bigcup_{y\in V(T_x)} Y_y$, $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x:=V_x\cup F_x$.
+Then,
   \begin{compactenum}[(Y1)]
@@
   \end{compactenum}
-\end{clm}
-
-Before proving \clmref{generalized-tripod} we point out more properties that are immediately implied by it:
+\end{lem}
+
+Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:
 
@@
 
-\begin{proof}[Proof of \clmref{generalized-tripod}]
-  Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$ and the fact that $(\mathcal{L},\mathcal{Y})$ is a layered partition of layered width $\ell$.
+\begin{proof}[Proof of \cref{generalized-tripod}]
+  Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$
+  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
 Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.
@@
 
-In the rest of this section, 
-%$(\mathcal{L},\mathcal{Y})$ denotes a layered $H$-partition of $G$;
-$\mathcal{T}=(B_x:x\in V(T))$ is a tree-decomposition of $H:=G/\mathcal{Y}$ satisfying the conditions of \lemref{nice-decomposition} and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ are defined as in \clmref{generalized-tripod}.
+% In the rest of this section, 
+% %$(\mathcal{L},\mathcal{Y})$ denotes a layered $H$-partition of $G$;
+% $\mathcal{T}=(B_x:x\in V(T))$ is a tree-decomposition of $H:=G/\mathcal{Y}$ satisfying the conditions of \lemref{nice-decomposition} and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ are defined as in \clmref{generalized-tripod}.
+
+We are now ready to prove our main result, which we now restate for convenience:
+
+\mmg*
+
+\begin{proof}
+Apply \cref{generalized-tripod} to $G$ and let 
+let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $Y_x$, $V_x$, $F_x$, and $N_x$ be defined as in \cref{generalized-tripod}.
 
@@
 
-\begin{clm}
+\begin{clm}\label{x-v-ancestor}
   For any $v\in V(G)$ there exists a (unique) node $a(v)\in X_v$ such that
@@
 % Together, \clmref{general-width} and \clmref{general-bag-size} imply the following generalization of \thmref{k-planar}:
-We are now ready to prove the main result of the paper.
-
-\begin{proof}[Proof of \cref{modern-major-general}] 
-  The proof is almost immediate from \clmref{general-width} and \clmref{general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
+
+  At this point, the proof of \cref{modern-major-general} is almost immediate from \clmref{general-width} and \clmref{general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
 In particular, $G^{\mathcal{P}}$ may contain an edge $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.
@@
 This increases the layered width given by \clmref{general-width} from $d\ell(k^2+3)$ to $d\ell(k^3+3k)$.
+Therefore $G$ has an $H$-partition of layered width at most $d\ell(k^3+3k)$ in which $H$ has treewidth at most $\binom{k+t}{t}$, completing the proof of \cref{modern-major-general}.
 \end{proof}
@@
 Since $G_0$ is planar,  \cref{PlanarPartition}(b) implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
-Applying \cref{modern-major-general} to $G_0$ and $\mathcal{P}$ immediately implies that $G$ (an arbitrary $k$-planar graph) has a layered $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
+Applying \cref{modern-major-general} to $G_0$ and $\mathcal{P}$ immediately implies that $G$ (an arbitrary $k$-planar graph) has an $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
 
@@
 In this setting, the proof of \cref{clm:general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $V_i\cap Y_x$.
-More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$ contributes:
-\begin{compactitem}
-  \item  at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$;
-  \item at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2},\ldots,k+1\}$; 
-  \item no vertices to $S_x\cap V_j$ for $j > k+1$.
-\end{compactitem}
-\note{DW}{do we need the bulletted list here?}
+More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2},\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap V_j$ for $j > k+1$.
+% \end{compactitem}
+% \note{DW}{do we need the bulletted list here?}
 Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields 
@@
 Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
-It follows that
-\begin{equation}
-  \dist_{G'}(v,w)\le 2 \enspace . \eqlabel{spar-distance}
-\end{equation}
+It follows that $\dist_{G'}(v,w)\le 2$.
 
@@
 First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $V_i$, each incident on at most two edges of $\overline{Y}$.
-By \eqref{spar-distance}, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap V_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $V_i$.
+Since $\dist_{G'}(v,w)\le 2$ for each $vw\in E(G)$, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap V_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $V_i$.
 These at most six edges have at most 12 endpoints.
@@
 
-  Note that $M$ is a Mercedes graph 
-  \note{DW}{``Mercedes'' not defined} and that each of the faces $F_1,\ldots,F_m$ is contained in a single wedge of $M$.
+  % Note that $M$ is a Mercedes graph 
+  % \note{DW}{``Mercedes'' not defined} and that each of the faces
+
+  Observe that each of the faces $F_1,\ldots,F_m$ is contained in a single internal face of $M$.
 Let $Y^+ := \overline{Y}^+-F$.
@@
   The proof finishes in the same way as the proof in \cite{dujmovic.joret.ea:planar}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition. 
-  \note{DW}{Do we need $S$ here?}
+  % \note{DW}{Do we need $S$ here?}
    Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
@@
   Use the layering $\mathcal{L}=\langle V_0',V_1'\ldots\rangle$ where $V_i'=V_{2i}\cup V_{2i+1}$ for each integer $i\ge 0$.
-That this is a layering of $G$ follows from \eqref{spar-distance}.
-Since $|V_j\cap S_x|\le 15$ for every integer $i\ge 0$, \note{DW}{replace $i$ by $j$?} $|V_i'\cap S_x|\le 30$ for every integer $i\ge 0$ and every $x\in V(H)$.
+That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le 2$ for every edge $vw\in E(G)$.
+Since $|V_j\cap S_x|\le 15$ for every integer $j\ge 0$, $|V_i'\cap S_x|\le 30$ for every integer $i\ge 0$ and every $x\in V(H)$.
 \end{proof}
@@
 \label{qn}
-If a graph $G$ has a layered $H$-partition of layered width $\ell$, then 
+If a graph $G$ has an $H$-partition of layered width $\ell$, then 
 $$\qn(G)\le 3 \ell \, \qn(H) + \floor{\tfrac{3}{2}\ell}\enspace.$$
@@
 \label{non-repetitive}
-If a graph $G$ has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$, then $\pi(G)\le \ell\, 4^{t+1}$.
+If a graph $G$ has an $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$, then $\pi(G)\le \ell\, 4^{t+1}$.
 \end{lem}
@@
 \label{p-centered}
-Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
+Every graph $G$ that has an $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
 \end{lem}
@@
 
-Prior to the current work, the strongest known upper bounds on the $p$-centered chromatic number of $(g,k)$-planar graphs $G$ were doubly exponential in $p$, as we now explain. 
+Prior to the current work, the strongest known upper bounds on the $p$-centered chromatic number of $(g,k)$-planar graphs $G$ were doubly-exponential in $p$, as we now explain. 
 \citet{dujmovic.eppstein.ea:structure} proved that $G$ has layered treewidth $(4g+6)(k+1)$. 
@@
 The following converse of \cref{gAddDummyDegree} provides a sufficient condition for a graph to be $(g,k)$-planar that is slightly more restrictive than being obtained from a $(k,d)$-shortcut system of a $(g,k)$-planar graph.
-\note{DW}{Is this sentence clear enough?}
+\notex{DW}{Is this sentence clear enough?}\notex{PM}{Sure.}
 
@@
 \label{StringPartition}
-Every $(g,\delta)$-string graph has a layered $H$-partition of layered 
+Every $(g,\delta)$-string graph has an $H$-partition of layered 
 width at most $\max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$ 
@@
 Fix $k\geq 1$. Let $G$ be a graph of Euler genus $g$ and maximum degree $\Delta$. 
-Then $G^k$ has a layered $H$-partition of layered width at most $\max\{2g,3\} (2k^4+6k^2) \Delta^{k}$ 
+Then $G^k$ has an $H$-partition of layered width at most $\max\{2g,3\} (2k^4+6k^2) \Delta^{k}$ 
 for some graph $H$ with treewidth at most $\binom{k+4}{4}-1$, and treewidth 
@@
 \begin{compactitem}
-\item a layered $H$-partition of layered width $O(k^4)$ in which $H$ has treewidth $O(k^6)$,
+\item an $H$-partition of layered width $O(k^4)$ in which $H$ has treewidth $O(k^6)$,
 \item queue-number at most $2^{O(k^6)}$,
@@
 
-\cref{k-planar} shows that $k$-planar graphs have layered $H$-partitions of layered width $O(k^2)$ in which $H$ has treewidth $O(k^3)$.
-We conclude with the following question:  Does there exist a function $f:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph has a layered $H$-partition of layered width at most $f(k)$ in which $H$ has treewidth at most $C$?
+\cref{k-planar} shows that $k$-planar graphs have $H$-partitions of layered width $O(k^2)$ in which $H$ has treewidth $O(k^3)$.
+We conclude with the following question:  Does there exist a function $f:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph has an $H$-partition of layered width at most $f(k)$ in which $H$ has treewidth at most $C$?
 Perhaps $C=3$.

### 2019-07-09 k-planar.tex
History of implicit representations

@@
 This enables combinatorial results for graphs of bounded
-\begin{wrapfigure}{r}{0.5\textwidth} 
-\centering
+treewidth to be generalised for planar graphs (with different constants).
+\cref{PlanarBasic} has been the key tool in solving several well-known open problems.
+In particular, \citet{dujmovic.joret.ea:planar} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992), \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002), and \citet{bonamy.gavoille.ea:shorter} use it to find shorter implicit representations of planar graphs (making progress on a sequence of results going back to 1988 \citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}).
+
+% \begin{wrapfigure}{r}{0.5\textwidth} 
+\begin{figure}[hbtp]
+\begin{center}
 \includegraphics{ProductExample}
+\end{center}
 \caption{Example of a strong product.}
 \label{ProductExample}
-\end{wrapfigure}
-treewidth to be generalised for planar graphs (with different constants).
-\cref{PlanarBasic} has been the key tool in solving several well-known open problems.
-In particular, \citet{dujmovic.joret.ea:planar} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992), \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002), \citet{bonamy.gavoille.ea:shorter} use it to find shorter implicit representations of planar graphs.
+\end{figure}
+% \end{wrapfigure}
 
@@
 
-\begin{figure}
+\begin{figure}[htbp]
   \begin{center}
@@
   If $w\in Y_x\cap V_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes only one vertex, itself, to $S_x\cap V_i$.
-  \begin{figure}
+
+  \begin{figure}[htbp]
     \begin{center}
@@
 
-The following lemma is due to \citet{micek:personal}. We include the proof  for completeness. 
+The following lemma is due to \citet{micek:personal}. 
+% We include the proof  for completeness. 
 % \note{DW}{can we replace by a result of \citet{micek:personal}?}
@@
 \label{p-centered}
-Every graph $G$ that has an $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
+Every graph $G$ that has an $H$-partition of layered width at most $\ell$ has $\chi_p(G)\le \ell (p+1)\chi_p(H)$. 
 \end{lem}
 
-\begin{proof}
-  Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$ where $H$ as treewidth at most $t$.
-  We use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,
-  (t+1)\binom{p+t}{t}\}$.  For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in V_i\cap B_x$ such that:
-  \begin{compactenum}
-    \item $\alpha(v)$ is unique among $\{\phi(w): w\in V_i\cap B_x\}$, which is possible 
-    since $|V_i\cap B_x|\le \ell$,
-    \item $\beta(v)= i\bmod (p+1)$, and     
-    \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,(t+1)\binom{p+t}{t}\}$ is a $p$-centered colouring of $H$ from \cref{p-centered-treewidth}.
-  \end{compactenum}
-To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
-
-First suppose that there exists $v,w\in V(X)$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$.
+% \begin{proof}
+%   Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$.
+%   Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,
+%   (t+1)\binom{p+t}{t}\}$.  For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in V_i\cap B_x$ such that:
+%   \begin{compactenum}
+%     \item $\alpha(v)$ is unique among $\{\phi(w): w\in V_i\cap B_x\}$, which is possible 
+%     since $|V_i\cap B_x|\le \ell$,
+%     \item $\beta(v)= i\bmod (p+1)$, and     
+%     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
+%   \end{compactenum}
+% To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
+% 
+% First suppose that there exists $v,w\in V(X)$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$.
 Since $G[X]$ is connected, $G[X]$ contains a path from $v$ to $w$.
@@
 Therefore, $|\{\beta(v'):v'\in X\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
-
-Otherwise, $X\subseteq V_{i},\ldots,V_{i+s}$ for some $s<p$.
+% 
+% Otherwise, $X\subseteq V_{i},\ldots,V_{i+s}$ for some $s<p$.
 Let $H':=H[\{x\in V(H):B_x\cap X\neq\emptyset]$.
@@
 Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
-    For any $v,w\in B_x$ with $v\neq w$, either $v,w\in V_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in V_{i'}$ and $w\in V_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
+% For any $v,w\in B_x$ with $v\neq w$, either $v,w\in V_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in V_{i'}$ and $w\in V_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
 Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.
-Therefore, every vertex in $B_x$ receives colour distinct from every other vertex in $X$.\end{proof}
-
-\cref{p-centered,k-planar,1-planar,g-k-planar} immediately imply the following results, for every $p\ge 2$:
+Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
+% \end{proof}
+
+\cref{p-centered,p-centered-treewidth,k-planar,1-planar,g-k-planar} immediately imply the following results, for every $p\ge 2$:
 

### 2019-07-09 k-planar.tex
Fixed serious typo ---- in Theorem 2

@@
 \label{kPlanarBasic}
-Every $k$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth $(18k^3+30k)\binom{k+4}{3}-1$ and for some path $P$. 
+Every $k$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth $(18k^2+30k)\binom{k+4}{3}-1$ and for some path $P$. 
 \end{thm}

### 2019-07-09 k-planar.tex
Last changes for SODA submission

@@
 \usepackage{doi} % To make plainnat handle doi's properly
+
+\usepackage[mathlines]{lineno}
+% \setlength{\linenumbersep}{2em}
+\linenumbers
+% \rightlinenumbers
+% \linenumbers
+\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+ \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+ \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+ \renewenvironment{#1}%
+    {\linenomath\csname old#1\endcsname}%
+    {\csname oldend#1\endcsname\endlinenomath}}% 
+\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+ \patchAmsMathEnvironmentForLineno{#1}%
+ \patchAmsMathEnvironmentForLineno{#1*}}%
+\AtBeginDocument{%
+\patchBothAmsMathEnvironmentsForLineno{equation}%
+\patchBothAmsMathEnvironmentsForLineno{align}%
+\patchBothAmsMathEnvironmentsForLineno{flalign}%
+\patchBothAmsMathEnvironmentsForLineno{alignat}%
+\patchBothAmsMathEnvironmentsForLineno{gather}%
+\patchBothAmsMathEnvironmentsForLineno{multline}%
+}
 
@@
 
-These theorems have applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, $p$-centred colouring  \citep{micek:personal}, and succinct graph representations \citep{bonamy.gavoille.ea:shorter}, some of which
+These theorems have applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, and $p$-centred colouring  \citep{micek:personal},
+% and succinct graph representations \citep{bonamy.gavoille.ea:shorter}, some of which
 %, and graph encoding \citep{BGP}. 
@@
 \end{compactitem}
-A \emph{tree-decomposition} $\mathcal{T}$ of a graph $G$ consists of a tree $T$ and a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $K$ such that:
+A \emph{tree-decomposition} $\mathcal{T}$ of a graph $G$ consists of a tree $T$ and a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that:
 \begin{compactenum}[(i)]
@@
     Let $G$ be a graph having an $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\PP$ be a $(k,d)$-shortcut system for $G$.
-Then $G^\PP$ has a $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ with treewidth at most $\binom{k+t}{t}-1$.
+Then $G^\PP$ has a $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$.
   %\end{thm}
@@
 
-Again, a direct application of \cref{modern-major-general} using \cref{SurfacePartition}(b) implies \cref{k-planar} with a weaker bound on the layered width.
+Again, a direct application of \cref{modern-major-general} using \cref{SurfacePartition}(b) implies \cref{g-k-planar} with a weaker bound on the layered width.
 We prove the stronger bound in \cref{sec-k-planar}.
@@
 For convenience, it will be helpful to assume that $\mathcal{P}$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.
-Since $G^\PP$ is defined to be a supergraph of $G$, this assumption has no effect on $G^{\mathcal{P}}$ but avoids special cases in some of our proofs.
+Since $G^\PP$ is defined to be a supergraph of $G$, this assumption has no effect on $G^{\mathcal{P}}$ but eliminates special cases in some of our proofs.
 
@@
 We say that a $T$-ancestor $a\in V(T)$ of $x\in V(T)$ is a \emph{strict} $T$-ancestor of $x$ if $a\neq x$.
-For a rooted tree $T$ and any node $x\in V(T)$, let
+The \emph{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.  For any node $x\in V(T)$, define
 \[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
-be the maximal subtree of $T$ rooted at $x$.
+to be the maximal subtree of $T$ rooted at $x$.
 
@@
   \end{center}
-  \caption{The sets $Y_x$, $F_x$, and $N_x$ associated with $x\in V(T)$
+  \caption{The sets $Y_x$, $F_x$, and $V_x$ associated with $x\in V(T)$
   and the ancestors $a_1,\ldots,a_{t'}$ of $X$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.}
@@
   \begin{compactenum}[(Y1)]
-    \item\ylabel{thickness} $\mathcal{Y}:=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width (with respect to $\mathcal{L}$) at most $\ell$.
+    \item\ylabel{thickness} $\mathcal{Y}:=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
     \item\ylabel{separator} For each $x\in V(T)$, there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$. 
@@
 Property~\yref{containment-i} follows from the definition of $V_x$ and the fact that $V(T_x)\subseteq V(T_a)$.
-To show Property~\yref{containment-ii} first note that, by \yref{containment-i} it suffices to consider vertices in $w\in F_x=N_x\setminus V_x$.
+To show Property~\yref{containment-ii} first note that, by \yref{containment-i} it suffices to consider vertices $w\in F_x=N_x\setminus V_x$.
 By definition, every vertex $w\in F_x$ is adjacent, in $G$, to a vertex $v\in V_x$.
@@
 
-We are now ready to prove our main result, which we now restate for convenience:
+We are now ready to prove our main result, which we restate here convenience:
 
@@
 \begin{proof}
-Apply \cref{generalized-tripod} to $G$ and let 
-let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $Y_x$, $V_x$, $F_x$, and $N_x$ be defined as in \cref{generalized-tripod}.
+Apply \cref{generalized-tripod} to $G$ and let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $Y_x$, $V_x$, $F_x$, and $N_x$ be defined as in \cref{generalized-tripod}.
 
@@
 If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
-In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$.
+In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$, $(v_0,\ldots,v_r)\in\PP$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$.
 
@@
 By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.
-Therefore, $\mathcal{P}$ contains a path $P$ that crosses $x$.
+Therefore, $\mathcal{P}$ contains a path $P$, with $v\in V(P)$, that crosses $x$.
 The path $P$ contains a subpath $v_0,v_1,\ldots,v_{r}$ such that $v=v_0$ and $v_r\in Y_x$.
@@
 \begin{proof}
-  We will count the number of vertices in $S_x\cap V_i$ by upper-bounding the number of vertices contributed to $S_x\cap V_i$ by each vertex $w\in Y_x$. Refer to \cref{contribute}.
+  We count the number of vertices in $S_x\cap V_i$ by upper-bounding the number of vertices contributed to $S_x\cap V_i$ by each vertex $w\in Y_x$. Refer to \cref{contribute}.
   If $w\in Y_x\cap V_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes only one vertex, itself, to $S_x\cap V_i$.
@@
 Since $xy\in E(J)$, $G^\mathcal{P}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
-Since $xy\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.
+Since $vw\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.
 By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.
@@
 By \yref{ancestor-edge}, $v'\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
-But this implies that $v\in S_{a'}$ for some $T$-ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
+But this implies that $a(v)=a'$ so $v\in S_{a'}$ for some $T$-ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
 \end{proof}
@@
 
-  Consider an arbitrary node $x\in V(T)$ whose ancestors are $x_0,\ldots,x_{r-1},x_r$ where $x_0$ is the root of $T$ and $x_r:=x$.
+  Consider an arbitrary node $x\in V(T)$ where $x_0,\ldots,x_r$ is the path from the root $x_0$ of $T$ to $x_r:=x$.
 To avoid triple-subscripts in what follows, we abuse notation slightly by using $V_i$, $F_i$, and $N_i$,  as shorthands for $V_{x_i}$, $F_{x_i}$ and $N_{x_i}$, respectively.
@@
 This implies $G^{\mathcal{P}}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
-This implies that $\mathcal{P}$ a $vw$-shortcut $P_{vw}$.
+This implies that $\mathcal{P}$ contains a $vw$-shortcut $P_{vw}$.
 Let $v'$ be the second-last vertex of $P_{vw}$ (so $v'w\in E(G)$).
@@
 For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{1,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.
-Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root is non-increasing.
+Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root, $x_0$, of $T$ is non-increasing.
 
@@
   At this point, the proof of \cref{modern-major-general} is almost immediate from \clmref{general-width} and \clmref{general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
-In particular, $G^{\mathcal{P}}$ may contain an edge $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.
+In particular, $G^{\mathcal{P}}$ may contain edges $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.
 To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.
@@
 If a graph $G$ has an $H$-partition of layered width $\ell$, then 
-$$\qn(G)\le 3 \ell \, \qn(H) + \floor{\tfrac{3}{2}\ell}\enspace.$$
+\[ \qn(G)\le 3 \ell \, \qn(H) + \floor{\tfrac{3}{2}\ell}\enspace.\]
 \end{lem}
@@
 %$$\chi_p(G)\le (18k^2+30k)(p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
-$$\chi_p(G)\le (18k^2+30k)(p+1) \binom{k+4}{3} \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}\enspace.$$
+\[ \chi_p(G)\le (18k^2+30k)(p+1) \binom{k+4}{3} \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}\enspace. \]
 %\note{DW}{I previously wrote ``for some integer $t\leq \binom{k+4}{3}-1$'' thinking that $\binom{p+t}{t}$ might not be monotone with $t$, but it is because of the $+t$.}
@@
 For every $1$-planar graph $G$, 
-$$\chi_p(G)\le 20 (p+3)(p+2)(p+1)^2\enspace.$$
+\[ \chi_p(G)\le 20 (p+3)(p+2)(p+1)^2\enspace. \]
 %120 (p+1) \binom{p+3}{3} 
@@
 %$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
-$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) \binom{k+5}{4}\binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}\enspace.$$
+\[ \chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) \binom{k+5}{4}\binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}\enspace. \]
 %O(gk^2)\times p^{O(k^3)}$. 
@@
 Thus, the number of edges of $G$ that pass through $x$ is at most 
-$$
+\[
 \sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i \Delta^j 
@@
 < 2k \Delta^k \enspace.
-$$
+\]
 Draw each edge $vw$ of $G$ alongside $P_{vw}$ in $G_0$, so that every pair of edges cross at most once. 
@@
 Every $(g,d)$-map graph has $p$-centred chromatic number  
-$$\chi_p(G)\le 105\, d(d-3)\, \max\{2g,3\}\, (p+1) \tbinom{p+14}{14}\enspace,$$
+\[ \chi_p(G)\le 105\, d(d-3)\, \max\{2g,3\}\, (p+1) \tbinom{p+14}{14}\enspace, \]
 and every plane $d$-map graph has $p$-centred chromatic number  
-$$\chi_p(G)\le 210\,d(d-3)\, (p+1) \tbinom{p+9}{9}\enspace.$$
+\[ \chi_p(G)\le 210\,d(d-3)\, (p+1) \tbinom{p+9}{9}\enspace. \]
 \end{cor}
@@
 Every $(g,\delta)$-string graph has $p$-centred chromatic number  
-$$\chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1) 
-\tbinom{ \delta+5}{4} \binom{p+\binom{ \delta+5}{4}-1}{\binom{ \delta+5}{4}-1}\enspace,$$
+\[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1) 
+\tbinom{ \delta+5}{4} \binom{p+\binom{ \delta+5}{4}-1}{\binom{ \delta+5}{4}-1}\enspace, \]
 and every plane $(0,\delta)$-string graph has $p$-centred chromatic number  
-$$\chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1) 
-\tbinom{ \delta+4}{3}\binom{p+\binom{ \delta+4}{3}-1}{\binom{ \delta+4}{3}-1}\enspace.$$
+\[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1) 
+\tbinom{ \delta+4}{3}\binom{p+\binom{ \delta+4}{3}-1}{\binom{ \delta+4}{3}-1}\enspace.\]
 \end{cor}
@@
 Every $(g,\delta)$-string graph has queue-number at most 
-$$3 \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)  2^{\binom{ \delta+5}{4}-1} \enspace.$$
+\[3 \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)  2^{\binom{ \delta+5}{4}-1} \enspace.\]
 and every plane $(0,\delta)$-map graph  has queue-number at most 
-$$ 9 (\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)  2^{\binom{ \delta+4}{3}-1} \enspace.$$
+\[ 9 (\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)  2^{\binom{ \delta+4}{3}-1} \enspace.\]
 %If a graph $G$ has a layered $H$-partition of layered width $\ell$, then 
@@
 Thus the number of paths in $\PP$ that use $v$ as an internal vertex is at most
-$$\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i\Delta^j 
+\[\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i\Delta^j 
 = \sum_{i=1}^{k-1} \Delta^i \sum_{j=1}^{k-i} \Delta^j 
 < \sum_{i=1}^{k-1} \Delta^i ( 2 \Delta^{k-i} )
-< 2k \Delta^k\enspace.$$
+< 2k \Delta^k\enspace.\]
 Hence $\PP$ is a $(k, 2k \Delta^k)$-shortcut system. 
@@
 \item $G^k$ has queue-number at most 
-$$3\,\max\{2g,3\} (2k^4+6k^2) \,\Delta^{k}\, 2^{\binom{k+4}{4}-1},$$
+\[3\,\max\{2g,3\} (2k^4+6k^2) \,\Delta^{k}\, 2^{\binom{k+4}{4}-1},\]
 %If a graph $G$ has a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$, then $\qn(G)\le 3 \ell (2^t-1) + \floor{\tfrac{3}{2}\ell} \le 3 \ell (2^t-\frac12 )$.
@@
 \item $G^k$ has  $p$-centred chromatic number at most 
-$$\max\{2g,3\}(2k^4+6k^2) \Delta^{k}  (p+1) \tbinom{k+4}{4} 
-\binom{p+\binom{k+4}{4}-1}{\binom{k+4}{4}-1}\enspace.$$
+\[\max\{2g,3\}(2k^4+6k^2) \Delta^{k}  (p+1) \tbinom{k+4}{4} 
+\binom{p+\binom{k+4}{4}-1}{\binom{k+4}{4}-1}\enspace.\]
 %\label{p-centered} Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
@@
   If $xy$ is of Type~C, then \cref{convex} implies that (without loss of generality) 
-  $$\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)\enspace.$$ 
+  \[\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)\enspace.\] 
    In this case, we call $x$ a Type~C vertex.
@@
 Since $\qn(K_n)=\floor{\frac{n}{2}}$, by \cref{UnitDiscGraph}, every unit disc graph $G$ in $\mathbb{R}^d$ with no $(k+1)$-clique satisfies
-$$\qn(G) \leq \qn( P^{(d)} \boxtimes K_{k \ceil{\sqrt{d}}^d} ) 
+\[\qn(G) \leq \qn( P^{(d)} \boxtimes K_{k \ceil{\sqrt{d}}^d} ) 
 \leq
-\frac{1}{6} (  3^{d+1} (k \ceil{\sqrt{d}}^d + 1 ) -3 )\enspace.$$
+\frac{1}{6} (  3^{d+1} (k \ceil{\sqrt{d}}^d + 1 ) -3 )\enspace.\]
 
@@
 Then 
-$$\tw(G) \leq 2  ( \tw(H)+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1\enspace.$$
-$$\pw(G) \leq 2  ( \pw(H)+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1\enspace.$$
+\[\tw(G) \leq 2  ( \tw(H)+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1\enspace.\]
+\[\pw(G) \leq 2  ( \pw(H)+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1\enspace.\]
 \end{lem}
@@
 Let $\langle{V^i_0,V^i_1,\dots}\rangle$ be the layering of $G$ determined by the $i$-th dimension. 
-Let $$m:= \ceil*{ \left(\frac{dn}{t+1} \right)^{1/(d+1)} } \enspace.$$ 
+Let \[m:= \ceil*{ \left(\frac{dn}{t+1} \right)^{1/(d+1)} } \enspace.\]
 For $\alpha\in[0,m-1]$, let $V^{i,\alpha} := \bigcup\{ V^i_j : j\equiv \alpha \pmod{m} \}$. 
@@
 %$$\pw(G) \leq 2  k^{ 1/(d+1) }  \ceil{\sqrt{d}}^{d/(d+1) } (dn)^{1- 1/(d+1)}   -  1\enspace.$$
-$$\pw(G) \leq 2  (2k)^{ 1/d } \ceil{\sqrt{d}} ((d-1)n)^{1- 1/d}   -  1\enspace.$$
+\[\pw(G) \leq 2  (2k)^{ 1/d } \ceil{\sqrt{d}} ((d-1)n)^{1- 1/d}   -  1\enspace.\]
 \end{cor}

### 2019-07-11 k-planar.tex
Added a note about David's conjecture

@@
 
+\note{PM}{
+I'm not sure what the role of $P_1,\ldots,P_{d-1}$ is here.
+If the result is true, then for some graph $H$ with $\tw(H)\le f(k,d)$, $G\subseteq H\boxtimes P_1 \boxtimes\dots \boxtimes P_{d-1} \subseteq H\boxtimes P\boxtimes K_{d-1}$  so, $G\subgraph H'\boxtimes P$ where $tw(H')\le f'(k,d):=(f(k,d)+1))(d-1)$ by \cref{MakeWidth1}.
+On the other hand, if $G\subgraph H'\boxtimes P$ then $G\subgraph H'\boxtimes P_1\boxtimes\cdots\boxtimes P_{d-1}$ so the two statements are equivalent up to a factor of $d-1$ in $f(k,d)$..
+}
+
 %%%%%%%%%%%%%%%%%%

### 2019-07-11 k-planar.tex
Pfffft

@@
 If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
-In particular, suppose that an $n$-vertex graph $G$ has an $H$-partition with layered width at most $\ell$ such that $H$ has treewidth at most $k$.
-\citet{dujmovic.joret.ea:planar} proved that $G$ has layered treewidth at most $(k+1)\ell$.
+In particular, suppose that an $n$-vertex graph $G$ has an $H$-partition with layered width at most $\ell$ such that $H$ has treewidth at most $t$.
+\citet{dujmovic.joret.ea:planar} proved that $G$ has layered treewidth at most $(t+1)\ell$.
 % So any property that holds for graphs of bounded layered treewidth also holds for $G$.
@@
 
+Finally, we show that these results cannot be generalized to $k$-nearest-neighbour graphs of points in $\R^3$, even for $k=4$.
+\begin{prop}\label{3d-4nn}
+  The class of $4$-nearest-neighbour graphs of point sets in $\R^3$ does not admit bounded layered partitions.
+\end{prop}
+
+\begin{proof}
+  We actually prove something stronger, that 4-nearest-neighbour graphs of point sets in $\R^3$ do not have bounded layered treewidth.
+As discussed in \cref{PriorWork}, if $G$ is a graph with an $H$-partition of layered width $\ell$ in which $\tw(H)\le t$, then the layered treewidth of $G$ is at most $(t+1)\ell$.
+Next we give an examples of arbitrarily large $n$-vertex $4$-nearest-neighbour graphs in $\R^3$ having layered treewidth $\Omega(\sqrt{n}/\log n)$, thus establishing \cref{3d-4nn}.
+
+  Let $r:=2^h$ for any integer $h>0$. For $i\in\{0,\ldots,h\}$, let $q_i:=\sum_{j=1}^i2^{i-2}$, let
+  \[
+     Z_i:=\left\{ q_i+x2^i, q_i+y2^i, 2^i)   : x,y\in\{0,r/2^{i}-1\} \right\}
+  \]
+  and define the quadtree-like point set $P = \bigcup_{i=0}^{h} Z_i$.
+Let $G_P$ denote the $4$-nearest-neighbour graph of $P$.
+For any $p\in Z_0$, the distance to any point $q\in P\setminus Z_0$ is at least
+  \[ \sqrt{(1/2)^2 + (1/2)^2 + (2^1-2^0)} > 1 \enspace .\]
+  Therefore $G_P[Z_0]$ contains an $r\times r$ grid.
+Therefore $G_P$ has treewidth at least $r-1$ and any tree decomposition $\mathcal{T}=(B_x:x\in V(T))$ of $G_P$ contains a bag $B_x$ of size at least $r$.
+
+  We can readily verify that $G_P$ contains a complete $4$-ary spanning tree $T_4$ of height $h$.
+Indeed, $T_4$ is rooted at the unique point in $Z_h$ and, for each $i\in\{0,\ldots,h\}$, the points in $Z_i$ have $T_4$-depth $h-i$.
+Therefore $G_P$ has diameter at most $2h$.
+Therefore, any layering of $G_P$ has at most $2h$ non-empty layers.
+Therefore, in any layering $\mathcal{L}:=\langle V_1,\ldots,V_{2h}\rangle$ of $G_P$, there is a layer $L_i\in\mathcal{L}$ such that $|L_i\cap B_x| > r/(2h) = r/(2\log_2 r)$.
+Therefore the layered treewidth of $G_P$ is least $r/(2\log_2 r)$.
+Letting $n:=|V(G)| = \sum_{i=0}^h r^2/4^i < 2r^2$ we see that $G_P$ has layered treewidth $\Omega(\sqrt{n}/\log n)$.
+\end{proof}
+
 \section{An Open Problem}
@@
 If the result is true, then for some graph $H$ with $\tw(H)\le f(k,d)$, $G\subseteq H\boxtimes P_1 \boxtimes\dots \boxtimes P_{d-1} \subseteq H\boxtimes P\boxtimes K_{d-1}$  so, $G\subgraph H'\boxtimes P$ where $tw(H')\le f'(k,d):=(f(k,d)+1))(d-1)$ by \cref{MakeWidth1}.
-On the other hand, if $G\subgraph H'\boxtimes P$ then $G\subgraph H'\boxtimes P_1\boxtimes\cdots\boxtimes P_{d-1}$ so the two statements are equivalent up to a factor of $d-1$ in $f(k,d)$..
+On the other hand, if $G\subgraph H'\boxtimes P$ then $G\subgraph H'\boxtimes P_1\boxtimes\cdots\boxtimes P_{d-1}$ so the two statements are equivalent up to a factor of $d-1$ in $f(k,d)$.
+It would be quite cool, though, if the statement were true with $f(k)$.
 }

### 2019-07-15 k-planar.tex
Arxiv v1

--- 2019-07-09 k-planar.tex
@@
 \usepackage{stmaryrd}
+%\usepackage{iwona}
 \usepackage{thm-restate}
@@
 
-\usepackage[mathlines]{lineno}
-% \setlength{\linenumbersep}{2em}
-\linenumbers
-% \rightlinenumbers
-% \linenumbers
-\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
- \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
- \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
- \renewenvironment{#1}%
-    {\linenomath\csname old#1\endcsname}%
-    {\csname oldend#1\endcsname\endlinenomath}}% 
-\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
- \patchAmsMathEnvironmentForLineno{#1}%
- \patchAmsMathEnvironmentForLineno{#1*}}%
-\AtBeginDocument{%
-\patchBothAmsMathEnvironmentsForLineno{equation}%
-\patchBothAmsMathEnvironmentsForLineno{align}%
-\patchBothAmsMathEnvironmentsForLineno{flalign}%
-\patchBothAmsMathEnvironmentsForLineno{alignat}%
-\patchBothAmsMathEnvironmentsForLineno{gather}%
-\patchBothAmsMathEnvironmentsForLineno{multline}%
-}
+%\usepackage[mathlines]{lineno}
+%% \setlength{\linenumbersep}{2em}
+%\linenumbers
+%% \rightlinenumbers
+%% \linenumbers
+%\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+% \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+% \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+% \renewenvironment{#1}%
+%    {\linenomath\csname old#1\endcsname}%
+%    {\csname oldend#1\endcsname\endlinenomath}}% 
+%\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+% \patchAmsMathEnvironmentForLineno{#1}%
+% \patchAmsMathEnvironmentForLineno{#1*}}%
+%\AtBeginDocument{%
+%\patchBothAmsMathEnvironmentsForLineno{equation}%
+%\patchBothAmsMathEnvironmentsForLineno{align}%
+%\patchBothAmsMathEnvironmentsForLineno{flalign}%
+%\patchBothAmsMathEnvironmentsForLineno{alignat}%
+%\patchBothAmsMathEnvironmentsForLineno{gather}%
+%\patchBothAmsMathEnvironmentsForLineno{multline}%
+%}
+%
+%
+%
 
@@
 
-\setlength{\parskip}{1ex}
-
-% \newcommand{\note}[2]{{\color{red}[#1:~#2]}}
+\setlength{\parskip}{0.85ex}
+
+ \newcommand{\note}[2]{{\color{red}[#1:~#2]}}
 % \newcommand{\notex}[2]{{\color{blue}[#1:~#2]}}
@@
 Dujmovi\'c~et~al.~(FOCS 2019) recently proved that every planar graph is a subgraph of the strong product of a graph of bounded treewidth and a path.
-This tool has been used to solve longstanding problems on queue layouts, non-repetitive colouring, $p$-centred colouring, and implicit graph encoding.
+This tool has been used to solve longstanding problems on queue layouts, non-repetitive colouring, $p$-centered colouring, and implicit graph encoding.
 We generalise this result for $k$-planar graphs, where a graph is \emph{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings.
@@
 \end{titlepage}
-
-% Layered partitions of graphs (with small layered width, in which the quotient has small treewidth) are a recently-introduced tool that have been used to solve longstanding problems on queue layouts, non-repetitive colouring, and 3-d graph drawing.
-Such partitions are known to exist for planar graphs and, more generally, apex-minor-free graphs.
-In the current paper, we prove that every $k$-planar graph has a layered partition of layered width $O(k^2)$ in  which the quotient graph has treewidth $O(k^3)$.
-This is the first result of this type for a non-minor-closed class of graphs and implies that $k$-planar graphs have both queue number and non-repetitive chromatic number upper-bounded by a function of $k$.
-The former result was previously shown by combining layered partitions of planar graphs with \textit{ad hoc} methods.
-The latter result is new, for all $k\ge 1$.
-These results extend to $(g,k)$-planar graphs, the natural generalization of $k$-planar graphs to to genus-$g$ surfaces (rather than the genus-$0$ plane).
 
@@
 \cref{PlanarBasic} has been the key tool in solving several well-known open problems.
-In particular, \citet{dujmovic.joret.ea:planar} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992), \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002), and \citet{bonamy.gavoille.ea:shorter} use it to find shorter implicit representations of planar graphs (making progress on a sequence of results going back to 1988 \citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}).
+In particular, \citet{dujmovic.joret.ea:planar} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992);  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002); and \citet{bonamy.gavoille.ea:shorter} use it to find shorter implicit representations of planar graphs (making progress on a sequence of results going back to 1988 \citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}).
 
@@
 
-Although $k$-planar graphs are the most high-profile target for a generalization \cref{PlanarBasic}, we actually prove a substantially stronger result than \cref{kPlanarBasic} using the following definition.
+Although $k$-planar graphs are the most high-profile target for a generalization of \cref{PlanarBasic}, we actually prove a substantially stronger result than \cref{kPlanarBasic} using the following definition.
 A collection $\mathcal{P}$ of paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if:
@@
 
-This definition is related to $k$-planarity because of the following observation
+This definition is related to $k$-planarity because of the following observation:
 
@@
 \label{AddDummy}
-Every $k$-planar graph is isomorphic to a subgraph of $G^\PP$ for some planar graph $G$ and some $(k+1,2)$-shortcut system $\PP$ for $G$. 
+Every $k$-planar graph is a subgraph of $G^\PP$ for some planar graph $G$ and some $(k+1,2)$-shortcut system $\PP$ for $G$. 
 \end{obs}
@@
 
-These theorems have applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, and $p$-centred colouring  \citep{micek:personal},
+These theorems have applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, and $p$-centered colouring  \citep{micek:personal},
 % and succinct graph representations \citep{bonamy.gavoille.ea:shorter}, some of which
@@
 
-%\note{PM}{TODO: Write here about applications. non-repetitive colouring, $p$-centered colouring, $(g,k)$-planar graph, $(g,d)$-map graphs, $k$-NN graphs, David's rough characterization.}
+%\notex{PM}{TODO: Write here about applications. non-repetitive colouring, $p$-centered colouring, $(g,k)$-planar graph, $(g,d)$-map graphs, $k$-NN graphs, David's rough characterization.}
 
@@
 An embedded graph $G$ is \emph{$k$-plane} if each edge of $G$ takes part in at most $k$ crossings.
-A (not necessarily embedded) graph $G'$ is \emph{$k$-planar} if there exists $k$-plane graph $G$ isomorphic to $G'$.
+A (not necessarily embedded) graph $G'$ is \emph{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
 Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
@@
 
-% \note{DW}{I  prefer to not define ``layered $H$-partition''.
+% \notex{DW}{I  prefer to not define ``layered $H$-partition''.
 Just define the ``layered width'' of an $H$-partition, which is what we did in \citep{dujmovic.joret.ea:planar}.
@@
 \label{MakeWidth1}
-If a graph $G$ has an $H$-partition of layered width $\ell$ with respect to layering $\langle V_0,V_1,\dots\rangle$, for some graph $H$ of treewidth at most $t$, then $G$ has an $H'$-partition of layered width 1 with respect to the same layering, for some graph $H'$ of treewidth at most $(t+1)\ell-1$.
+If a graph $G$ has an $H$-partition of layered width $\ell$  
+%with respect to layering $\langle V_0,V_1,\dots\rangle$, 
+for some graph $H$ of treewidth at most $t$, then $G$ has an $H'$-partition of layered width 1   
+%with respect to the same layering, 
+for some graph $H'$ of treewidth at most $(t+1)\ell-1$.
 That is, if $G\subseteq H\boxtimes P\boxtimes K_\ell$ for some graph $H$ of treewidth at most $t$  and for some path $P$, then $G\subseteq H' \boxtimes P$ for some graph $H'$ of treewidth at most $(t+1)\ell-1$.
@@
 
-Using relationship between $k$-planar graphs and $(k+1,2)$-shortcut systems along with a direct application of \cref{modern-major-general} we can obtain a slightly weaker version of \cref{k-planar}, below.
+Using the relationship between $k$-planar graphs and $(k+1,2)$-shortcut systems along with a direct application of \cref{modern-major-general} we can obtain a slightly weaker version of \cref{k-planar}, below.
 The minor modifications needed to obtain the stronger bound are described in \cref{sec-k-planar}.
-
- % of planar graphs shows that \cref{modern-major-general} has immediate applications to $k$-planar graphs.
-The following result shows that $k$-planar graphs admit bounded layered partitions.
-With \cref{PartitionProduct,MakeWidth1}, this proves \cref{kPlanarBasic}.
-% with $c_k\leq(4+o(1))k^5$.
 
@@
 % on the layered width. 
-% % \note{DW}{I would just write, ``\cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with a weaker bound on the layered width.}
+% % \notex{DW}{I would just write, ``\cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with a weaker bound on the layered width.}
 % We prove the stronger bound in \cref{sec-k-planar}. 
@@
 \label{gAddDummy}
-Every $(g,k)$-planar graph is isomorphic to a subgraph of $G^\PP$ for some graph $G$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G$. 
+Every $(g,k)$-planar graph is a subgraph of $G^\PP$ for some graph $G$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G$. 
 \end{obs}
@@
 Finally, we state the following corollary of \cref{ApexMinorFree,modern-major-general,MakeWidth1}.
-This is the most general result that follows from \cref{modern-major-general} and that of \citet{dujmovic.joret.ea:planar}.
+This is the most general result that follows from \cref{modern-major-general} and the work of \citet{dujmovic.joret.ea:planar}.
 
@@
 
-%\note{DW}{should this subsection be moved to the end of \cref{Tools}?}\note{PM}{No, I don't think so.
+%\notex{DW}{should this subsection be moved to the end of \cref{Tools}?}\notex{PM}{No, I don't think so.
 The information given in Section 2 isn't detailed enough to compare with layered treewidth versus layered partitions.}
@@
 
-If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
-In particular, suppose that an $n$-vertex graph $G$ has an $H$-partition with layered width at most $\ell$ such that $H$ has treewidth at most $k$.
-\citet{dujmovic.joret.ea:planar} proved that $G$ has layered treewidth at most $(k+1)\ell$.
+If a graph class admits bounded layered partitions, then it also has bounded layered treewidth. In particular, 
+\citet{dujmovic.joret.ea:planar} proved that if a graph $G$ has an $H$-partition with layered width at most $\ell$ such that $H$ has treewidth at most $k$, then $G$ has layered treewidth at most $(k+1)\ell$.
 % So any property that holds for graphs of bounded layered treewidth also holds for $G$.
@@
 We say that a $T$-ancestor $a\in V(T)$ of $x\in V(T)$ is a \emph{strict} $T$-ancestor of $x$ if $a\neq x$.
-The \emph{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.  For any node $x\in V(T)$, define
+The \emph{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.  For each node $x\in V(T)$, define
 \[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
@@
 
-We begin with a fairly standard lemma about normalized tree decompositions.
+We begin with a fairly standard lemma about normalised tree decompositions.
 
 \begin{lem}\lemlabel{nice-decomposition}
-  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:  
+  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a $T$-decomposition $(B_x:x\in V(T))$ of $H$ with width $t$ that has following additional properties:  
   \begin{compactenum}[(T1)]
@@
 
-We are now ready to prove our main result, which we restate here convenience:
+We are now ready to prove our main result, which we restate here for convenience:
 
@@
 By removing repeated vertices this gives a path in the directed graph $\overrightarrow{H}^+$ obtained by directing each edge $xy\in E(H^+)$ from its $T$-descendant $x$ towards its $T$-ancestor $y$.
-  Finally, we are in a position to appeal to \cite[Lemma~24]{pilipczuk.siebertz:polynomial-arxiv} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
+  Finally, we are in a position to appeal to \cite[Lemma~24]{pilipczuk.siebertz:polynomial-soda} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
 \end{proof}
@@
 We will assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$.
-This assumption is not critical since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
+This assumption is justified since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
 
@@
 More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2},\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap V_j$ for $j > k+1$.
-% \end{compactitem}
-% \note{DW}{do we need the bulletted list here?}
 Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields 
 \begin{align*}
-    |S_x\cap Y_i| &\le d\ell\left( 
+    |S_x\cap Y_i| \le d\ell\left( 
        2 
@@
        + 2\left(k+1-\left\lfloor\frac{k+1}{2}\right\rfloor\right)
-       \right)\\
+       \right)
        & =
@@
        & \le
-       d\ell(3k+5) = 18k+30 \enspace .
+       d\ell(3k+5) 
+      = 18k+30 \enspace .
 \end{align*}    
@@
     \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; and
-    \item $H$ has a tree decomposition whose largest bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
+    \item $H$ has a tree decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
   \end{compactenum}
@@
   % Note that $M$ is a Mercedes graph
-  % \note{DW}{``Mercedes'' not defined} and that each of the faces
 
@@
   The proof finishes in the same way as the proof in \cite{dujmovic.joret.ea:planar}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition. 
-  % \note{DW}{Do we need $S$ here?}
    Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
@@
 \begin{proof}[Proof of \cref{1-planar}]
-Given a 1-plane graph $G$, add edges to make it edge maximal so that it has an outer face $F=v_1v_2v_3$.
+Given a 1-plane graph $G$, add edges to make it edge-maximal so that it has an outer face $F=v_1v_2v_3$.
 Next, add a vertex $r$ adjacent to $v_1$, $v_2$, and $v_3$ to obtain an edge-maximal 1-plane graph $\overline{G}$ with one vertex $r$ of degree 3 on its outer face.
@@
 %
-%
-%\note{PM}{I removed (commented out) the $(g,1)$-specialization. It requires cutting surfaces.  For reference, the statement was:
-%\begin{thm}
-%\label{g-1-planar}
-%Every $(g,1)$-planar graph has a layered $H$-partition of layered width at most $\max\{48g, 30\}$ in which $H$ has treewidth at most $4$.  
-%\end{thm}
-%}
-%
-%\note{DW}{I can understand you doing this, but for some our applications we only get the best known bounds in the case when $g\gg k$, in which case this is best illustrated by the $k=1$ case.
-I have commented out these results which rely on  \cref{g-1-planar} for now, but I would advocate including  \cref{g-1-planar} and its nice corollaries.
-}
-%
-%\note{PM}{I propose we leave it out of the SODA submission, but I'm happy to include it in the DCG/ArXiV submission (which we can hopefully finish in the next few days.)}
-%
-
+%We conclude this section with a small extension of \cref{k-planar} that will be useful in \cref{sec-g-k-planar}.
+%
+%\begin{cor}
+%\label{cor-k-planar}
+%Let $G$ be a $k$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
+Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \cref{k-planar} and such that $V_i' := \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\}$ for each integer $i\ge 0$.
+%\end{cor}
+%
+%\begin{proof}
+%  Without loss of generality, assume that $r_0$ is on the outer face of $G^+$ and let $v_1$ and $v_2$ be the other two vertices of the outer face of $G^+$.
+%  The proof of \cref{k-planar} begins with the graph $G^+$ and then creates a graph $G^{++}$ consisting of a $K_4$ containing $G^+$ in one of its inner faces, $F_0$, and six edges required to complete $G^{++}$ to a triangulation.
+Instead, modify $G^{++}$ by subdividing the edges joining $F_0$ to $v_1$ and $v_2$, as shown in \cref{wrapper}(b).
+In this modified version of $G^{++}$, we have $\dist_{G^++}(r,v)=\dist_{T}(r,v)=\dist_{G^+}(r_0,v)+2$
+%  which leads to the layering $\langle V_0,V_1,\ldots\rangle$ of $G^+$ where $V_i=\{v\in V(G^+): \dist_{T}(r,v) = i\}$.
+%  
+%  The rest of proof remains unchanged until the point where we derive the layering $\mathcal{L}=\langle V_0',V_1',\ldots\rangle$ of $G$.  At this point, we set
+%  \[  
+%    V_i' := \{v\in V(G): \floor{\dist_{T}(r,v)-2}=i\} 
+%         = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\} \enspace ,
+%  \]
+%  for all integers $i\ge 0$.  This layering satisfies the conditions of the corollary.
+%\end{proof}
+%  
+%
+%
+%
 % Again, in the case $k=1$ we obtain better constants. 
@@
 Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of 
-$O(g^{k+2})$ using \cref{SurfacePartition} and an ad-hoc method. Our result provides a better bound when $g>2^{k^3}$. 
-%In particular, in the case for $k=1$, \cref{qn,1-planar,g-1-planar} imply:
-%
-%\begin{prop}
-%For $g\geq 1$, every $(g,1)$-planar graph has queue-number at most $3 \times 48g (2^4-1) + \floor{\tfrac{3}{2} \times 48g }= 2232 g $.
-%\end{prop}
-%
-%The previous best bound on the queue-number of $(g,1)$-planar graphs was $2(40g+490)^{3}$ due to \citet{dujmovic.joret.ea:planar}.
-For $g=0$ we can improve further.
-\citet{ABGKP18} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
-Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$.
-\cref{qn,1-planar} then imply:
+$O(g^{k+2})$ using \cref{SurfacePartition} and an ad-hoc method.
+Our result provides a better bound when $g>2^{k^3}$.
 In the case of 1-planar graphs we can improve further.
@@
 \cref{non-repetitive,k-planar,1-planar,g-k-planar} immediately imply the following results:
-
-% \noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.}\note{DW}{done}
 
@@
 \label{p-centered-treewidth}
-Every graph $H$ of treewidth at most $t$ has $\chi_p(H)\in (t+1)\binom{p+t}{t}$.
+Every graph $H$ of treewidth at most $t$ has $\chi_p(H)\leq  (t+1)\binom{p+t}{t}$.
 \end{lem}
 
-The following lemma is due to \citet{micek:personal}. 
-% We include the proof  for completeness. 
+The following lemma is implicitly due to \citet{micek:personal}. We include the proof for completeness.
+
 % \note{DW}{can we replace by a result of \citet{micek:personal}?}
@@
 \label{p-centered}
-Every graph $G$ that has an $H$-partition of layered width at most $\ell$ has $\chi_p(G)\le \ell (p+1)\chi_p(H)$. 
+Every graph $G$ that has an $H$-partition of layered width at most $\ell$ has $\chi_p(G)\le \ell (p+1)\, \chi_p(H)$. 
 \end{lem}
 
-% \begin{proof}
-%   Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$.
-%   Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,
-%   (t+1)\binom{p+t}{t}\}$.  For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in V_i\cap B_x$ such that:
-%   \begin{compactenum}
-%     \item $\alpha(v)$ is unique among $\{\phi(w): w\in V_i\cap B_x\}$, which is possible 
-%     since $|V_i\cap B_x|\le \ell$,
-%     \item $\beta(v)= i\bmod (p+1)$, and     
-%     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
-%   \end{compactenum}
-% To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
-% 
-% First suppose that there exists $v,w\in V(X)$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$.
+\begin{proof}
+   Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$.
+   Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.
+For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in V_i\cap B_x$ such that:
+   \begin{compactenum}
+     \item $\alpha(v)$ is unique among $\{\phi(w): w\in V_i\cap B_x\}$, which is possible 
+     since $|V_i\cap B_x|\le \ell$,
+     \item $\beta(v)= i\bmod (p+1)$, and     
+     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
+   \end{compactenum}
+ To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
+
+ First suppose that there exists $v,w\in V(X)$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$.
 Since $G[X]$ is connected, $G[X]$ contains a path from $v$ to $w$.
@@
 Therefore, $|\{\beta(v'):v'\in X\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
-% 
-% Otherwise, $X\subseteq V_{i},\ldots,V_{i+s}$ for some $s<p$.
+
+ Otherwise, $X\subseteq V_{i},\ldots,V_{i+s}$ for some $s<p$.
 Let $H':=H[\{x\in V(H):B_x\cap X\neq\emptyset]$.
@@
 Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
-% For any $v,w\in B_x$ with $v\neq w$, either $v,w\in V_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in V_{i'}$ and $w\in V_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
+ For any $v,w\in B_x$ with $v\neq w$, either $v,w\in V_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in V_{i'}$ and $w\in V_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
 Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.
 Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
-% \end{proof}
+ \end{proof}
 
 \cref{p-centered,p-centered-treewidth,k-planar,1-planar,g-k-planar} immediately imply the following results, for every $p\ge 2$:
-
-% \noindent\note{PM}{Double-check the calculation for \cref{k-planar} and \cref{g-k-planar} and then fix the constants here.} \note{DW}{done}
 
@@
 %For every $(g,1)$-planar graph $G$, $g\ge 1$, $\chi_p(G)\le O(gp^4)$.
+%\note{DW}{update}
 %\end{cor}
@@
 Van den Heuvel and Wood~\citep{vdHW17} showed that this implies that $G$  has $r$-strong colouring number at most $(4g + 6)(k + 1)(2r + 1)$.
-By a result of \citet{zhu:colouring}, $G$ has $r$-weak colouring number at most $( (4g + 6)(k + 1)(2r + 1) )^r$, which by another result of  \citet{zhu:colouring} implies that $G$ has  $p$-centred chromatic number at most $( (4g+6)(k+1)(2^{p-1} + 1) )^{2^{p-2}}$.
+By a result of \citet{zhu:colouring}, $G$ has $r$-weak colouring number at most $( (4g + 6)(k + 1)(2r + 1) )^r$, which by another result of  \citet{zhu:colouring} implies that $G$ has  $p$-centered chromatic number at most $( (4g+6)(k+1)(2^{p-1} + 1) )^{2^{p-2}}$.
 \cref{pc1,pck,pcgk} are substantial improvements over these results, providing bounds on $\chi_p(G)$ that are polynomial in $p$ for fixed $g$ and $k$. 
@@
 \end{lem}
-
-%Recall that \cref{gAddDummy} says that every $(g,k)$-planar graph is isomorphic to $G^\PP$ for some graph $G$ of Euler genus $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G$.
-Here $G$ is the graph obtained by adding a degree-4 dummy vertex at each crossing point.
-Thus for each path $P\in \PP$ each internal vertex in $P$ has degree 4.
 
@@
 A $(0,d)$-map graph is called a (plane) \emph{$d$-map graph}; see \citep{FLS-SODA12,CGP02} for example.
-%such graphs have been extensively studied \citep{FLS-SODA12,DFHT05,CGP02,Chen07,Chen01}. 
 The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.
@@
 Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar.
-This shows that map graphs admit bounded layered partitions, but we get much better bounds by constructing a shortcut system directly.
+\cref{g-k-planar} then shows that map graphs admit bounded layered partitions, but we get much better bounds by constructing a shortcut system directly.
 The following lemma is reminiscent of the characterisation of $(g,d)$-map graphs in terms of the half-square of bipartite graphs \citep{CGP02,dujmovic.eppstein.ea:structure}.
@@
 Each vertex $x\in V(G_0)$ is the middle vertex on at most $\tfrac12 d(d-3)$  paths in $\PP$.
-Thus $\PP$ is a $(2,\tfrac12 d(d-3))$-shortcut system for $G_1$, and by construction, $G$ is a subgraph of $G_1^\PP$.
+Thus $\PP$ is a $(2,\tfrac12 d(d-3))$-shortcut system for $G_1$, and by construction, $G \subseteq G_1^\PP$.
 \end{proof}
 
-\cref{PlanarPartition}(b)  and \cref{SurfacePartition}(b) and \cref{modern-major-general,MapShortcut} imply:
+\cref{PlanarPartition}(b), \cref{SurfacePartition}(b) and \cref{modern-major-general,MapShortcut} imply:
 
@@
 
-\cref{p-centered,MapPartition} imply that the first known bounds on the $p$-centred chromatic number of map graphs that are polynomial in $p$.
+\cref{p-centered,p-centered-treewidth,MapPartition} imply that the first known bounds on the $p$-centered chromatic number of map graphs that are polynomial in $p$.
 
 \begin{cor}
-Every $(g,d)$-map graph has $p$-centred chromatic number  
+Every $(g,d)$-map graph $G$ has $p$-centered chromatic number  
 \[ \chi_p(G)\le 105\, d(d-3)\, \max\{2g,3\}\, (p+1) \tbinom{p+14}{14}\enspace, \]
-and every plane $d$-map graph has $p$-centred chromatic number  
+and every plane $d$-map graph $G$ has $p$-centered chromatic number  
 \[ \chi_p(G)\le 210\,d(d-3)\, (p+1) \tbinom{p+9}{9}\enspace. \]
@@
 Let $G_0$ be the graph obtained by adding a vertex at the intersection point of every pair of curves in $\mathcal{C}$ that intersect,  where two such consecutive vertices on a curve $C_v$ are adjacent in $G_0$.
-For each vertex $v\in V(G)$, introduce a new vertex called $v$ on $C_v$ between the
-$\floor{\frac{\delta}{2}}$-th vertex already on $C_v$ and the $(\floor{\frac{\delta}{2}}+1)$-th such vertex.
+For each vertex $v\in V(G)$, if $C_v$ intersects $k\leq\delta$ other curves, then introduce a new vertex called $v$ on $C_v$ between the
+$\floor{\frac{k}{2}}$-th vertex already on $C_v$ and the $\floor{\frac{k}{2}+1}$-th such vertex.
 For each edge $vw$ of $G$, there is a path $P_{vw}$ of length at most $2\ceil{\frac{\delta}{2}}\leq \delta+1$ in $G_0$ between $v$ and $w$.
@@
 
-\cref{StringShortcut,modern-major-general} and \cref{PlanarPartition}(b) and \cref{SurfacePartition}(b) imply:
+\cref{StringShortcut,modern-major-general}, \cref{PlanarPartition}(b) and \cref{SurfacePartition}(b) imply:
 
@@
 width at most $\max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$ 
-for some graph $H$ with treewidth at most $\binom{ \delta+5}{4}-1$, 
-and treewidth at most $\binom{ \delta+4}{3}-1$ if $g=0$. 
+for some graph $H$ with treewidth at most $\binom{ \delta+5}{4}-1$ or treewidth at most $\binom{ \delta+4}{3}-1$ if $g=0$. 
 %
@@
 
-\cref{p-centered,StringPartition} imply that the first known bounds on the $p$-centred chromatic number of map graphs that are polynomial in $p$.
+\cref{p-centered,p-centered-treewidth,StringPartition} imply the first known bounds on the $p$-centered chromatic number of map graphs that are polynomial in $p$.
 
 \begin{cor}
-Every $(g,\delta)$-string graph has $p$-centred chromatic number  
+Every $(g,\delta)$-string graph has $p$-centered chromatic number  
 \[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1) 
 \tbinom{ \delta+5}{4} \binom{p+\binom{ \delta+5}{4}-1}{\binom{ \delta+5}{4}-1}\enspace, \]
-and every plane $(0,\delta)$-string graph has $p$-centred chromatic number  
+and every plane $(0,\delta)$-string graph has $p$-centered chromatic number  
 \[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1) 
@@
 
-\cref{PlanarPartition}(b) and \cref{SurfacePartition}(b) and \cref{modern-major-general,PowerShortcut} imply the following result, which with \cref{PartitionProduct} implies \cref{kPowerBasic} in \cref{Introduction}.
+\cref{PlanarPartition}(b), \cref{SurfacePartition}(b), \cref{modern-major-general,PowerShortcut} imply the following result, which with \cref{PartitionProduct} implies \cref{kPowerBasic} in \cref{Introduction}.
 
@@
 
-\cref{DrawG,qn,p-centered,PowerPartition} immediately imply the following result:
+\cref{DrawG,qn,p-centered,p-centered-treewidth,PowerPartition,non-repetitive} immediately imply the following result:
 
@@
 \begin{cor}
+\label{PowerCor}
 For all integers $k\geq 1$ and $g\geq 0$, if $G$ is a graph with Euler genus $g$ and maximum degree $\Delta$, then:
@@
 %%%%%%
-\item $G^k$ has queue-number at most 
-\[3\,\max\{2g,3\} (2k^4+6k^2) \,\Delta^{k}\, 2^{\binom{k+4}{4}-1},\]
+\item $G^k$ has queue-number 
+\[\qn(G^k) \leq 3\,\max\{2g,3\} (2k^4+6k^2) \,\Delta^{k}\, 2^{\binom{k+4}{4}-1},\]
 %If a graph $G$ has a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$, then $\qn(G)\le 3 \ell (2^t-1) + \floor{\tfrac{3}{2}\ell} \le 3 \ell (2^t-\frac12 )$.
 %%%%%%
-\item $G^k$ has  $p$-centred chromatic number at most 
-\[\max\{2g,3\}(2k^4+6k^2) \Delta^{k}  (p+1) \tbinom{k+4}{4} 
-\binom{p+\binom{k+4}{4}-1}{\binom{k+4}{4}-1}\enspace.\]
+\item $G^k$ has  $p$-centered chromatic number 
+\[\chi_p(G^k) \leq \max\{2g,3\}(2k^4+6k^2) \Delta^{k}  (p+1) \tbinom{k+4}{4} 
+\binom{p+\binom{k+4}{4}-1}{\binom{k+4}{4}-1}\enspace,\]
 %\label{p-centered} Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
+\item $G^k$ has  non-repetitive chromatic number
+\[ \pi(G^k) \leq  \max\{2g,3\} (2k^4+6k^2) \Delta^{k} \, 4^{\binom{k+4}{4}}\enspace.\]
+%If a graph $G$ has an $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$,  then $\pi(G)\le \ell\, 4^{t+1}$.
 \end{compactitem}
@@
 
-\cref{DrawG,non-repetitive,g-k-planar} also give bounds on the non-repetitive chromatic number of $G^k$, 
-%at most $(8+o(1))(g+1)k^2\Delta^{4(k-1)} 4^{(\frac{1}{6}+o(1))k^3\Delta^{6(k-1)}+1}$.\\
-but the bound is weak, since $G^k$ has maximum degree at most $\Delta^k$, implying that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$ by a result of \citet{DJKW16}.
+This result is the first constant upper bound on the queue-number and $p$-centered chromatic number of bounded powers of graphs with bounded degree and bounded Euler genus.
+For every graph $G$, since $G^k$ has maximum degree at most $\Delta^k$, a result of \citet{DJKW16} implies that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$.
+\cref{PowerCor} improves upon this bound when $k\ll\Delta$ and $G$ has Euler genus $g$.
 
@@
 %\] 
-(The edges of $G$ are straight-line segments joining their endpoints.)
+(The edges of $G$ are straight-line segments joining their endpoints.) See \citep{ProximityGraphs} for a survey of results on $k$-nearest neighbour graphs and other related proximity graphs.
 
@@
 
-\cref{nearest-neighbour,k-planar,qn,p-centered} imply:
+\cref{nearest-neighbour,k-planar,qn,p-centered,p-centered-treewidth} imply:
 
@@
 \item queue-number at most $2^{O(k^6)}$,
-\item $p$-centred chromatic number at most $O(k^4) (p+1) (t+1)\binom{p+t}{t}$, for some integer $t\leq O(k^6)$.
+\item $p$-centered chromatic number at most $O(k^4) (p+1) (t+1)\binom{p+t}{t}$ for some integer $t\leq O(k^6)$.
 \end{compactitem}
@@
 
-\bibliographystyle{plainnat}
-\bibliography{k-planar}
+%%%  Squashing the bibliography 
+  \let\oldthebibliography=\thebibliography
+  \let\endoldthebibliography=\endthebibliography
+  \renewenvironment{thebibliography}[1]{%
+    \begin{oldthebibliography}{#1}%
+      \setlength{\parskip}{0.25ex}%
+      \setlength{\itemsep}{0.25ex}%
+  }{\end{oldthebibliography}}
+
+%\bibliographystyle{plainnat}
+%\bibliography{k-planar}
+
+% NOTE THAT MANUAL CHANGES HAVE BEEN MADE BELOW.
+
+\begin{thebibliography}{33}
+\providecommand{\natexlab}[1]{#1}
+\providecommand{\url}[1]{\texttt{#1}}
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{doi: #1}\else
+  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
+
+\bibitem[{\'{A}}brego et~al.(2011){\'{A}}brego, Monroy,
+  Fern{\'{a}}ndez{-}Merchant, Flores{-}Pe{\~{n}}aloza, Hurtado,
+  Sacrist{\'{a}}n, and Saumell]{abrego.munroy.ea:on}
+Bernardo~M. {\'{A}}brego, Ruy~Fabila Monroy, Silvia Fern{\'{a}}ndez{-}Merchant,
+  David Flores{-}Pe{\~{n}}aloza, Ferran Hurtado, Vera Sacrist{\'{a}}n, and
+  Maria Saumell.
+\newblock On crossing numbers of geometric proximity graphs.
+\newblock \emph{Comput. Geom.}, 44\penalty0 (4):\penalty0 216--233, 2011.
+\newblock \doi{10.1016/j.comgeo.2010.11.003}.
+
+\bibitem[Ajtai et~al.(1982)Ajtai, Chv{\'a}tal, Newborn, and
+  Szemer{\'e}di]{ajtai.chvatal.ea:crossing-free}
+Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, and Endre
+  Szemer{\'e}di.
+\newblock Crossing-free subgraphs.
+\newblock \emph{Ann. Discrete Math.}, 12:\penalty0 9--12, 1982.
+\newblock \doi{10.1016/S0304-0208(08)73484-4}.
+
+\bibitem[Alam et~al.(2018)Alam, Bekos, Gronemann, Kaufmann, and
+  Pupyrev]{ABGKP18}
+Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael Kaufmann, and
+  Sergey Pupyrev.
+\newblock Queue layouts of planar 3-trees.
+\newblock In Therese~C. Biedl and Andreas Kerren, editors, \emph{Proc. 26th
+  International Symposium on Graph Drawing and Network Visualization} ({GD}
+  2018), vol.~11282 of \emph{Lecture Notes in Computer Science}, pp.~213--226. Springer, 2018.
+\newblock \doi{10.1007/978-3-030-04414-5\_15}.
+
+\bibitem[Alon et~al.(2002)Alon, Grytczuk, Ha{\l}uszczak, and
+  Riordan]{AGHR-RSA02}
+Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, and Oliver Riordan.
+\newblock Nonrepetitive colorings of graphs.
+\newblock \emph{Random Structures Algorithms}, 21\penalty0 (3-4):\penalty0
+  336--346, 2002.
+\newblock \doi{10.1002/rsa.10057}.
+
+\bibitem[Bodlaender(1998)]{Bodlaender-TCS98}
+Hans~L. Bodlaender.
+\newblock A partial $k$-arboretum of graphs with bounded treewidth.
+\newblock \emph{Theoret. Comput. Sci.}, 209\penalty0 (1-2):\penalty0 1--45,
+  1998.
+\newblock \doi{10.1016/S0304-3975(97)00228-4}.
+
+\bibitem[Bonamy et~al.(2019)Bonamy, Gavoille, and
+  Pilipczuk]{bonamy.gavoille.ea:shorter}
+Marthe Bonamy, Cyril Gavoille, and Michał Pilipczuk.
+\newblock Shorter implicit representation for planar graphs.
+\newblock In S.~Cabello and B.~Mohar, editors, \emph{9th Slovenian
+  International Conference on Graph Theory} (SICGT), page~63, June 2019.
+
+\bibitem[Bose et~al.(2001)Bose, Morin, Stojmenović, and
+  Urrutia]{bose.morin.ea:routing}
+Prosenjit Bose, Pat Morin, Ivan Stojmenovi\'c, and Jorge Urrutia.
+\newblock Routing with guaranteed delivery in ad hoc wireless networks.
+\newblock \emph{Wireless Networks}, 7\penalty0 (6):\penalty0 609--616, 2001.
+\newblock \doi{10.1023/A:1012319418150}.
+
+\bibitem[Bose et~al.(2012)Bose, Dujmovi\'c, Hurtado, Iacono, Langerman, Meijer,
+  Sacrist\'an, Saumell, and Wood]{ProximityGraphs}
+Prosenjit Bose, Vida Dujmovi\'c, Ferran Hurtado, John Iacono, Stefan Langerman,
+  Henk Meijer, Vera Sacrist\'an, Maria Saumell, and David~R. Wood.
+\newblock Proximity graphs: $e$, $\delta$, $\delta$, $\delta$, $\chi$ and
+  $\omega$.
+\newblock \emph{Internat. J. Comput. Geom.  Appl.},
+  22\penalty0 (5):\penalty0 439--469, 2012.
+
+\bibitem[Chen et~al.(2002)Chen, Grigni, and Papadimitriou]{CGP02}
+Zhi-Zhong Chen, Michelangelo Grigni, and Christos~H. Papadimitriou.
+\newblock Map graphs.
+\newblock \emph{J. ACM}, 49\penalty0 (2):\penalty0 127--138, 2002.
+\newblock \doi{10.1145/506147.506148}.
+
+\bibitem[D{\k e}bski et~al.(2019)D{\k e}bski, Felsner, Micek, and Schr{\"
+  o}der]{micek:personal}
+  Micha{\l} D{\k e}bski, Stefan Felsner, Piotr Micek, and Felix Schr{\" o}der. 
+\newblock Improved bounds for centered colorings, 2019.
+\newblock \href{https://arxiv.org/abs/1907.04586}{arXiv:1907.04586}.
+
+\bibitem[Dujmovi\'{c} et~al.(2016)Dujmovi\'{c}, Joret, Kozik, and Wood]{DJKW16}
+Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, and David~R. Wood.
+\newblock Nonrepetitive colouring via entropy compression.
+\newblock \emph{Combinatorica}, 36\penalty0 (6):\penalty0 661--686, 2016.
+\newblock \doi{10.1007/s00493-015-3070-6}.
+
+\bibitem[Dujmovi\'c et~al.(2017{\natexlab{a}})Dujmovi\'c, Eppstein, and
+  Wood]{dujmovic.eppstein.ea:structure}
+Vida Dujmovi\'c, David Eppstein, and David~R. Wood.
+\newblock Structure of graphs with locally restricted crossings.
+\newblock \emph{{SIAM} J. Discrete Math.}, 31\penalty0 (2):\penalty0 805--824,
+  2017{\natexlab{a}}.
+\newblock \doi{10.1137/16M1062879}.
+
+\bibitem[Dujmovi\'c et~al.(2017{\natexlab{b}})Dujmovi\'c, Morin, and
+  Wood]{dujmovic.morin.ea:layered}
+Vida Dujmovi\'c, Pat Morin, and David~R. Wood.
+\newblock Layered separators in minor-closed graph classes with applications.
+\newblock \emph{J. Comb. Theory, Ser. {B}}, 127:\penalty0 111--147,
+  2017{\natexlab{b}}.
+\newblock \doi{10.1016/j.jctb.2017.05.006}.
+
+\bibitem[Dujmovi\'c et~al.(2019{\natexlab{a}})Dujmovi\'c, Esperet, Joret,
+  Walczak, and Wood]{dujmovic.esperet.ea:planar}
+Vida Dujmovi\'c, Louis Esperet, Gwena{\"{e}}l Joret, Bartosz Walczak,   David~R. Wood.
+\newblock Planar graphs have bounded nonrepetitive chromatic number.
+\newblock  \href{http://arxiv.org/abs/1904.05269}{arXiv:1904.05269}, 2019{\natexlab{a}}.
+
+\bibitem[Dujmovi\'c et~al.(2019{\natexlab{b}})Dujmovi\'c, Joret, Micek, Morin,
+  Ueckerdt, and Wood]{dujmovic.joret.ea:planar}
+Vida Dujmovi\'c, Gwena{\"{e}}l Joret, Piotr Micek, Pat Morin, Torsten Ueckerdt,
+  and David~R. Wood.
+\newblock Planar graphs have bounded queue-number.
+\newblock \emph{Proc. 60th Annual Symp. on Foundations of Computer 
+  Science} (FOCS '19), to appear. 
+  \newblock  \href{http://arxiv.org/abs/1904.04791}{arXiv:1904.04791}, 2019{\natexlab{b}}.
+
+\bibitem[Fomin et~al.(2012)Fomin, Lokshtanov, and Saurabh]{FLS-SODA12}
+Fedor~V. Fomin, Daniel Lokshtanov, and Saket Saurabh.
+\newblock Bidimensionality and geometric graphs.
+\newblock In \emph{Proc. 23rd {A}nnual {ACM}-{SIAM} {S}ymposium on {D}iscrete
+  {A}lgorithms} (SODA 2012), pp.~1563--1575, 2012.
+\newblock \doi{10.1137/1.9781611973099.124}.
+
+\bibitem[Fox and Pach(2010)]{FP10}
+Jacob Fox and J{\'a}nos Pach.
+\newblock A separator theorem for string graphs and its applications.
+\newblock \emph{Combin. Probab. Comput.}, 19\penalty0 (3):\penalty0 371--390,
+  2010.
+\newblock \doi{10.1017/S0963548309990459}.
+
+\bibitem[Fox and Pach(2014)]{FP14}
+Jacob Fox and J{\'a}nos Pach.
+\newblock Applications of a new separator theorem for string graphs.
+\newblock \emph{Combin. Probab. Comput.}, 23\penalty0 (1):\penalty0 66--74,
+  2014.
+\newblock \doi{10.1017/S0963548313000412}.
+
+\bibitem[Grigoriev and Bodlaender(2007)]{grigoriev.bodlaender:algorithms}
+Alexander Grigoriev and Hans~L. Bodlaender.
+\newblock Algorithms for graphs embeddable with few crossings per edge.
+\newblock \emph{Algorithmica}, 49\penalty0 (1):\penalty0 1--11, 2007.
+\newblock \doi{10.1007/s00453-007-0010-x}.
+
+\bibitem[Harvey and Wood(2017)]{HW17}
+Daniel~J. Harvey and David~R. Wood.
+\newblock Parameters tied to treewidth.
+\newblock \emph{J. Graph Theory}, 84\penalty0 (4):\penalty0 364--385, 2017.
+\newblock \doi{10.1002/jgt.22030}.
+
+\bibitem[Heath et~al.(1992)Heath, Leighton, and Rosenberg]{HLR92}
+Lenwood~S. Heath, F.~Thomson Leighton, and Arnold~L. Rosenberg.
+\newblock Comparing queues and stacks as mechanisms for laying out graphs.
+\newblock \emph{SIAM J. Discrete Math.}, 5\penalty0 (3):\penalty0 398--412,
+  1992.
+\newblock \doi{10.1137/0405031}.
+
+\bibitem[van~den Heuvel and Wood(2017)]{vdHW17}
+Jan van~den Heuvel and David~R. Wood.
+\newblock Improper colourings inspired by {H}adwiger's conjecture.
+\newblock \emph{J. London Math. Soc.}, 98(1):129--148, 2018. 
+\newblock  \doi{10.1112/jlms.12127}. See   \href{https://arxiv.org/abs/1704.06536}{arXiv:1704.06536} for the cited results.
+
+\bibitem[Kannan et~al.(1988)Kannan, Naor, and
+  Rudich]{kannan.naor.ea:implicit-stoc}
+Sampath Kannan, Moni Naor, and Steven Rudich.
+\newblock Implicit representation of graphs.
+\newblock In Janos Simon, editor, \emph{Proc.\ 20th Annual {ACM} Symposium on
+  Theory of Computing} (STOC '88), pp.~334--343. {ACM}, 1988.
+\newblock \doi{10.1145/62212.62244}.
+
+\bibitem[Kannan et~al.(1992)Kannan, Naor, and Rudich]{kannan.naor.ea:implicit}
+Sampath Kannan, Moni Naor, and Steven Rudich.
+\newblock Implicit representation of graphs.
+\newblock \emph{{SIAM} J. Discrete Math.}, 5\penalty0 (4):\penalty0 596--603,
+  1992.
+\newblock \doi{10.1137/0405049}.
+
+\bibitem[Kobourov et~al.(2017)Kobourov, Liotta, and
+  Montecchiani]{kobourov.liotta.ea:annotated}
+Stephen~G. Kobourov, Giuseppe Liotta, and Fabrizio Montecchiani.
+\newblock An annotated bibliography on 1-planarity.
+\newblock \emph{Comput. Sci. Rev.}, 25:\penalty0 49--67, 2017.
+\newblock \doi{10.1016/j.cosrev.2017.06.002}.
+
+\bibitem[Leighton(1983)]{leighton:complexity}
+F.~Thomson Leighton.
+\newblock \emph{Complexity issues in {VLSI}}.
+\newblock MIT Press, Cambridge, 1983.
+
+\bibitem[Mohar and Thomassen(2001)]{mohar.thomassen:graphs}
+Bojan Mohar and Carsten Thomassen.
+\newblock \emph{Graphs on surfaces}.
+\newblock Johns Hopkins Univ.\ Press, 2001.
+
+\bibitem[Ne{\v{s}}et{\v{r}}il and Ossona~de Mendez(2012)]{Sparsity}
+Jaroslav Ne{\v{s}}et{\v{r}}il and Patrice Ossona~de Mendez.
+\newblock \emph{Sparsity}, vol.~28 of \emph{Algorithms and Combinatorics}.
+\newblock Springer, 2012.
+\newblock \doi{10.1007/978-3-642-27875-4}.
+
+\bibitem[Pach and T{\'o}th(2002)]{PachToth-DCG02}
+J{\'a}nos Pach and G{\'e}za T{\'o}th.
+\newblock Recognizing string graphs is decidable.
+\newblock \emph{Discrete Comput. Geom.}, 28\penalty0 (4):\penalty0 593--606,
+  2002.
+\newblock \doi{10.1007/s00454-002-2891-4}.
+
+\bibitem[Pilipczuk and Siebertz(2019)]{pilipczuk.siebertz:polynomial-soda}
+Micha{\l} Pilipczuk and Sebastian Siebertz.
+\newblock Polynomial bounds for centered colorings on proper minor-closed graph
+  classes.
+\newblock In Timothy~M. Chan, editor, \emph{Proc.\ 30th Annual {ACM-SIAM}
+  Symposium on Discrete Algorithms} ({SODA} 2019), pp.~1501--1520. {SIAM},
+  2019.
+\newblock \doi{10.1137/1.9781611975482.91}.
+
+\bibitem[Reed(2003)]{Reed03}
+Bruce~A. Reed.
+\newblock Algorithmic aspects of tree width.
+\newblock In \emph{Recent Advances in Algorithms and Combinatorics}, vol.~11,
+  pp.~85--107. Springer, 2003, 
+\newblock \doi{10.1007/0-387-22444-0_4}.
+
+\bibitem[Wiechert(2017)]{wiechert:on}
+Veit Wiechert.
+\newblock On the queue-number of graphs with bounded tree-width.
+\newblock \emph{Electr. J. Comb.}, 24\penalty0 (1):\penalty0 P1.65, 2017.
+\newblock  \url{http://www.combinatorics.org/v24i1p65}.
+
+\bibitem[Zhu(2009)]{zhu:colouring}
+Xuding Zhu.
+\newblock Colouring graphs with bounded generalized colouring number.
+\newblock \emph{Discrete Mathematics}, 309\penalty0 (18):\penalty0 5562--5568,
+  2009.
+\newblock \doi{10.1016/j.disc.2008.03.024}.
+
+\end{thebibliography}
+
 \end{document}
@@
 
-\note{DW}{Add non-rep and $p$-centred colouring}
+\note{DW}{Add non-rep and $p$-centered colouring}
 
@@
 %K_{k \ceil{\sqrt{d}}^d}$.
-\end{document}

### 2019-07-15 k-planar.tex
Version submitted to DCG

@@
 \usepackage{stmaryrd}
-%\usepackage{iwona}
 \usepackage{thm-restate}
@@
 
-%\usepackage[mathlines]{lineno}
-%% \setlength{\linenumbersep}{2em}
-%\linenumbers
-%% \rightlinenumbers
-%% \linenumbers
-%\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
-% \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
-% \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
-% \renewenvironment{#1}%
-%    {\linenomath\csname old#1\endcsname}%
-%    {\csname oldend#1\endcsname\endlinenomath}}% 
-%\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
-% \patchAmsMathEnvironmentForLineno{#1}%
-% \patchAmsMathEnvironmentForLineno{#1*}}%
-%\AtBeginDocument{%
-%\patchBothAmsMathEnvironmentsForLineno{equation}%
-%\patchBothAmsMathEnvironmentsForLineno{align}%
-%\patchBothAmsMathEnvironmentsForLineno{flalign}%
-%\patchBothAmsMathEnvironmentsForLineno{alignat}%
-%\patchBothAmsMathEnvironmentsForLineno{gather}%
-%\patchBothAmsMathEnvironmentsForLineno{multline}%
-%}
-%
-%
-%
+\usepackage[mathlines]{lineno}
+% \setlength{\linenumbersep}{2em}
+\linenumbers
+% \rightlinenumbers
+% \linenumbers
+\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+ \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+ \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+ \renewenvironment{#1}%
+    {\linenomath\csname old#1\endcsname}%
+    {\csname oldend#1\endcsname\endlinenomath}}% 
+\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+ \patchAmsMathEnvironmentForLineno{#1}%
+ \patchAmsMathEnvironmentForLineno{#1*}}%
+\AtBeginDocument{%
+\patchBothAmsMathEnvironmentsForLineno{equation}%
+\patchBothAmsMathEnvironmentsForLineno{align}%
+\patchBothAmsMathEnvironmentsForLineno{flalign}%
+\patchBothAmsMathEnvironmentsForLineno{alignat}%
+\patchBothAmsMathEnvironmentsForLineno{gather}%
+\patchBothAmsMathEnvironmentsForLineno{multline}%
+}
 
@@
 
-\setlength{\parskip}{0.85ex}
+\setlength{\parskip}{0.82ex}
 
@@
 For example, \citet{FP08} generalised the Lipton-Tarjan separator theorem to show that every $k$-planar graph with $n$ vertices has a balanced separator of order $O(\sqrt{kn})$.}
-The main results of this paper generalise the following recent theorem of \citet{dujmovic.joret.ea:planar} to a setting that includes $k$-planar graphs.
+
+Our main results generalise the following recent theorem of \citet{dujmovic.joret.ea:planar} to a setting that includes $k$-planar graphs.
  % as well as some other interesting non-minor-closed graph families.
@@
 \item  $H \boxtimes P \boxtimes K_{\max\{2g,1\}}$ for some graph $H$ of treewidth at most $9$ and for some path $P$;
-\item  $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $4$ and for some path $P$;
-\item $(K_{2g} + H )  \boxtimes P$ for some graph $H$ of treewidth at most $8$  and some path $P$. 
+\item  $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $4$ and for some path $P$.
+%\item $(K_{2g} + H )  \boxtimes P$ for some graph $H$ of treewidth at most $8$  and some path $P$. 
 \end{compactenum}
@@
 \item an $H$-partition with layered width at most $\max\{2g,1\}$ such that $H$ has treewidth at most $9$;
-\item an $H$-partition with layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most $4$; 
-\item an $H$-partition with layered width at most $1$ such that $H$ has treewidth at most $2g+7$. 
+\item an $H$-partition with layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most $4$.
+%\item an $H$-partition with layered width at most $1$ such that $H$ has treewidth at most $2g+8$. 
 \end{compactenum}
@@
 A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
-\citet{grigoriev.bodlaender:algorithms} study approximation algorithms for $(g,k)$-plane graphs.
  %\citet[Theorem~22]{dujmovic.joret.ea:planar} show that every genus-$g$ graph has a layered $H$-partition of layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most 4. 
@@
 \citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph.
-\citet{dujmovic.eppstein.ea:structure} show that every $k$-planar graph has layered treewidth at most $6(k+1)$, and more generally that
-every $(g,k)$-planar graph has layered treewidth at most $(4g+6)(k+1)$.
+\citet{dujmovic.eppstein.ea:structure} show that every $k$-planar graph has layered treewidth at most $6(k+1)$, and more generally that every $(g,k)$-planar graph has layered treewidth at most $(4g+6)(k+1)$.
+It follows from this result that $(g,k)$-planar graphs have treewidth $O(\sqrt{(g+1)(k+1)n})$ and thus have balanced separators of the same order, which can also be concluded from the work of
+\citet{FP08}.
+In related work, \citet{grigoriev.bodlaender:algorithms} used structural results to obtain approximation algorithms for $(g,k)$-planar graphs, and \citet{PachToth97} determined the maximum number of edges in a $k$-planar graph (up to a constant factor).
 
@@
 \citet{dujmovic.joret.ea:planar} proved that if a graph $G$ has an $H$-partition with layered width at most $\ell$ such that $H$ has treewidth at most $k$, then $G$ has layered treewidth at most $(k+1)\ell$.
-% So any property that holds for graphs of bounded layered treewidth also holds for $G$.
-For example, Norin proved that every $n$-vertex graph with layered treewidth at most $t$ has treewidth less than $2\sqrt{tn}$ (see \citep{dujmovic.morin.ea:layered}).
+So any property that holds for graphs of bounded layered treewidth also holds for $G$.
+%For example, Norin proved that every $n$-vertex graph with layered treewidth at most $t$ has treewidth less than $2\sqrt{tn}$ (see \citep{dujmovic.morin.ea:layered}).
 Thus, $G$ has treewidth at most $2\sqrt{(k+1)\ell n}$.
@@
 
-The following lemma shows how to interpret an $H$-partition of $G$ and a tree-partition of $H$ as a hierarchical decomposition of $G$; refer to \cref{fig:generalized-tripod}.
+The following lemma shows how to interpret an $H$-partition of $G$ and a tree-decomposition of $H$ as a hierarchical decomposition of $G$; refer to \cref{fig:generalized-tripod}.
 
@@
     \begin{oldthebibliography}{#1}%
-      \setlength{\parskip}{0.25ex}%
-      \setlength{\itemsep}{0.25ex}%
+      \setlength{\parskip}{0.4ex}%
+      \setlength{\itemsep}{0.4ex}%
   }{\end{oldthebibliography}}
@@
 % NOTE THAT MANUAL CHANGES HAVE BEEN MADE BELOW.
+
+\fontsize{11}{11.75}\selectfont
 
@@
   Szemer{\'e}di]{ajtai.chvatal.ea:crossing-free}
-Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, and Endre
+Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, Endre
   Szemer{\'e}di.
@@
   Pupyrev]{ABGKP18}
-Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael Kaufmann, and
-  Sergey Pupyrev.
+Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael Kaufmann, Sergey Pupyrev.
 \newblock Queue layouts of planar 3-trees.
 \newblock In Therese~C. Biedl and Andreas Kerren, editors, \emph{Proc. 26th
-  International Symposium on Graph Drawing and Network Visualization} ({GD}
-  2018), vol.~11282 of \emph{Lecture Notes in Computer Science}, pp.~213--226. Springer, 2018.
+  International Symposium on Graph Drawing and Network Visualization} (GD '18), vol.~11282 of \emph{Lecture Notes in Computer Science}, pp.~213--226. Springer, 2018.
 \newblock \doi{10.1007/978-3-030-04414-5\_15}.
@@
   Riordan]{AGHR-RSA02}
-Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, and Oliver Riordan.
+Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, Oliver Riordan.
 \newblock Nonrepetitive colorings of graphs.
@@
   Pilipczuk]{bonamy.gavoille.ea:shorter}
-Marthe Bonamy, Cyril Gavoille, and Michał Pilipczuk.
+Marthe Bonamy, Cyril Gavoille, and Micha{\l} Pilipczuk.
 \newblock Shorter implicit representation for planar graphs.
 \newblock In S.~Cabello and B.~Mohar, editors, \emph{9th Slovenian
-  International Conference on Graph Theory} (SICGT), page~63, June 2019.
-
-\bibitem[Bose et~al.(2001)Bose, Morin, Stojmenović, and
+  International Conference on Graph Theory} (SICGT), p.~63, June 2019.
+
+\bibitem[Bose et~al.(2001)Bose, Morin, Stojmenovi\'c, and
   Urrutia]{bose.morin.ea:routing}
-Prosenjit Bose, Pat Morin, Ivan Stojmenovi\'c, and Jorge Urrutia.
+Prosenjit Bose, Pat Morin, Ivan Stojmenovi\'c, Jorge Urrutia.
 \newblock Routing with guaranteed delivery in ad hoc wireless networks.
@@
 Prosenjit Bose, Vida Dujmovi\'c, Ferran Hurtado, John Iacono, Stefan Langerman,
-  Henk Meijer, Vera Sacrist\'an, Maria Saumell, and David~R. Wood.
+  Henk Meijer, Vera Sacrist\'an, Maria Saumell, David~R. Wood.
 \newblock Proximity graphs: $e$, $\delta$, $\delta$, $\delta$, $\chi$ and
@@
 \bibitem[Chen et~al.(2002)Chen, Grigni, and Papadimitriou]{CGP02}
-Zhi-Zhong Chen, Michelangelo Grigni, and Christos~H. Papadimitriou.
+Zhi-Zhong Chen, Michelangelo Grigni, Christos~H. Papadimitriou.
 \newblock Map graphs.
@@
 
-\bibitem[D{\k e}bski et~al.(2019)D{\k e}bski, Felsner, Micek, and Schr{\"
-  o}der]{micek:personal}
-  Micha{\l} D{\k e}bski, Stefan Felsner, Piotr Micek, and Felix Schr{\" o}der. 
+\bibitem[D{\k e}bski et~al.(2019)D{\k e}bski, Felsner, Micek, and Schr{\"o}der]{micek:personal}
+  Micha{\l} D{\k e}bski, Stefan Felsner, Piotr Micek, Felix Schr{\" o}der. 
 \newblock Improved bounds for centered colorings, 2019.
@@
 \bibitem[Dujmovi\'{c} et~al.(2016)Dujmovi\'{c}, Joret, Kozik, and Wood]{DJKW16}
-Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, and David~R. Wood.
+Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, David~R. Wood.
 \newblock Nonrepetitive colouring via entropy compression.
@@
   Wood]{dujmovic.eppstein.ea:structure}
-Vida Dujmovi\'c, David Eppstein, and David~R. Wood.
+Vida Dujmovi\'c, David Eppstein, David~R. Wood.
 \newblock Structure of graphs with locally restricted crossings.
@@
   Wood]{dujmovic.morin.ea:layered}
-Vida Dujmovi\'c, Pat Morin, and David~R. Wood.
+Vida Dujmovi\'c, Pat Morin, David~R. Wood.
 \newblock Layered separators in minor-closed graph classes with applications.
@@
   Ueckerdt, and Wood]{dujmovic.joret.ea:planar}
-Vida Dujmovi\'c, Gwena{\"{e}}l Joret, Piotr Micek, Pat Morin, Torsten Ueckerdt,
-  and David~R. Wood.
+Vida Dujmovi\'c, Gwena{\"{e}}l Joret, Piotr Micek, Pat Morin, Torsten Ueckerdt, David~R. Wood.
 \newblock Planar graphs have bounded queue-number.
@@
 \bibitem[Fomin et~al.(2012)Fomin, Lokshtanov, and Saurabh]{FLS-SODA12}
-Fedor~V. Fomin, Daniel Lokshtanov, and Saket Saurabh.
+Fedor~V. Fomin, Daniel Lokshtanov, Saket Saurabh.
 \newblock Bidimensionality and geometric graphs.
@@
 
+\bibitem[Fox and Pach(2008)]{FP08}
+Jacob Fox, J{\'a}nos Pach.
+\newblock Separator theorems and {T}ur\'an-type results for planar intersection graphs.
+\newblock \emph{Adv. Math.}, 219(3):1070--1080, 2008.
+\newblock \doi{10.1016/j.aim.2008.06.002}.
+
 \bibitem[Fox and Pach(2010)]{FP10}
-Jacob Fox and J{\'a}nos Pach.
+Jacob Fox, J{\'a}nos Pach.
 \newblock A separator theorem for string graphs and its applications.
@@
 \bibitem[Fox and Pach(2014)]{FP14}
-Jacob Fox and J{\'a}nos Pach.
+Jacob Fox, J{\'a}nos Pach.
 \newblock Applications of a new separator theorem for string graphs.
@@
 \bibitem[Grigoriev and Bodlaender(2007)]{grigoriev.bodlaender:algorithms}
-Alexander Grigoriev and Hans~L. Bodlaender.
+Alexander Grigoriev, Hans~L. Bodlaender.
 \newblock Algorithms for graphs embeddable with few crossings per edge.
@@
 \bibitem[Harvey and Wood(2017)]{HW17}
-Daniel~J. Harvey and David~R. Wood.
+Daniel~J. Harvey, David~R. Wood.
 \newblock Parameters tied to treewidth.
@@
 \bibitem[Heath et~al.(1992)Heath, Leighton, and Rosenberg]{HLR92}
-Lenwood~S. Heath, F.~Thomson Leighton, and Arnold~L. Rosenberg.
+Lenwood~S. Heath, F.~Thomson Leighton, Arnold~L. Rosenberg.
 \newblock Comparing queues and stacks as mechanisms for laying out graphs.
@@
 \bibitem[van~den Heuvel and Wood(2017)]{vdHW17}
-Jan van~den Heuvel and David~R. Wood.
+Jan van~den Heuvel, David~R. Wood.
 \newblock Improper colourings inspired by {H}adwiger's conjecture.
@@
   Rudich]{kannan.naor.ea:implicit-stoc}
-Sampath Kannan, Moni Naor, and Steven Rudich.
+Sampath Kannan, Moni Naor, Steven Rudich.
 \newblock Implicit representation of graphs.
@@
 \bibitem[Kannan et~al.(1992)Kannan, Naor, and Rudich]{kannan.naor.ea:implicit}
-Sampath Kannan, Moni Naor, and Steven Rudich.
+Sampath Kannan, Moni Naor, Steven Rudich.
 \newblock Implicit representation of graphs.
@@
   Montecchiani]{kobourov.liotta.ea:annotated}
-Stephen~G. Kobourov, Giuseppe Liotta, and Fabrizio Montecchiani.
+Stephen~G. Kobourov, Giuseppe Liotta, Fabrizio Montecchiani.
 \newblock An annotated bibliography on 1-planarity.
@@
 \bibitem[Mohar and Thomassen(2001)]{mohar.thomassen:graphs}
-Bojan Mohar and Carsten Thomassen.
+Bojan Mohar, Carsten Thomassen.
 \newblock \emph{Graphs on surfaces}.
-\newblock Johns Hopkins Univ.\ Press, 2001.
+\newblock Johns Hopkins Press, 2001.
 
 \bibitem[Ne{\v{s}}et{\v{r}}il and Ossona~de Mendez(2012)]{Sparsity}
-Jaroslav Ne{\v{s}}et{\v{r}}il and Patrice Ossona~de Mendez.
+Jaroslav Ne{\v{s}}et{\v{r}}il, Patrice Ossona~de Mendez.
 \newblock \emph{Sparsity}, vol.~28 of \emph{Algorithms and Combinatorics}.
@@
 
+\bibitem[Pach and T{\'o}th(1997)]{PachToth97}
+J{\'a}nos Pach, G{\'e}za T{\'o}th.
+\newblock Graphs drawn with few crossings per edge. 
+\newblock \emph{Combinatorica}, 17(3):427--439, 1997
+  2002.
+\newblock \doi{10.1007/BF01215922}.
+
 \bibitem[Pach and T{\'o}th(2002)]{PachToth-DCG02}
-J{\'a}nos Pach and G{\'e}za T{\'o}th.
+J{\'a}nos Pach, G{\'e}za T{\'o}th.
 \newblock Recognizing string graphs is decidable.
@@
 \bibitem[Pilipczuk and Siebertz(2019)]{pilipczuk.siebertz:polynomial-soda}
-Micha{\l} Pilipczuk and Sebastian Siebertz.
+Micha{\l} Pilipczuk, Sebastian Siebertz.
 \newblock Polynomial bounds for centered colorings on proper minor-closed graph
@@
 \newblock Algorithmic aspects of tree width.
-\newblock In \emph{Recent Advances in Algorithms and Combinatorics}, vol.~11,
+\newblock In \emph{Recent Advances in Algorithms and Combinat.}, vol.~11,
   pp.~85--107. Springer, 2003, 
@@
 \newblock Colouring graphs with bounded generalized colouring number.
-\newblock \emph{Discrete Mathematics}, 309\penalty0 (18):\penalty0 5562--5568,
+\newblock \emph{Discrete Math.}, 309\penalty0 (18):\penalty0 5562--5568,
   2009.

### 2019-12-04 k-planar.tex
Shortcut systems must be non-empty

@@
 This tool has been used to solve longstanding problems on queue layouts, non-repetitive colouring, $p$-centered colouring, and implicit graph encoding.
-We generalise this result for $k$-planar graphs, where a graph is \emph{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings.
+We extend this result for $k$-planar graphs, where a graph is \emph{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings.
 In particular, we prove that every $k$-planar graph is a subgraph of the strong product of a graph of treewidth $O(k^5)$ and a path.
@@
 
-Our main results generalise the following recent theorem of \citet{dujmovic.joret.ea:planar} to a setting that includes $k$-planar graphs.
+Our main results extends the following recent theorem of \citet{dujmovic.joret.ea:planar} to a setting that includes $k$-planar graphs.
  % as well as some other interesting non-minor-closed graph families.
@@
 
-We generalise \cref{PlanarBasic} as follows.
+We extend \cref{PlanarBasic} as follows.
 
@@
 Although $k$-planar graphs are the most high-profile target for a generalization of \cref{PlanarBasic}, we actually prove a substantially stronger result than \cref{kPlanarBasic} using the following definition.
-A collection $\mathcal{P}$ of paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if:
+A non-empty set $\mathcal{P}$ of non-trivial paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if:
 \begin{compactitem}
-\item every path in $\mathcal{P}$ has length at most $k$,\footnote{A path of length $k$ consists of $k$ edges and $k+1$ vertices.} and
+\item every path in $\mathcal{P}$ has length at most $k$,\footnote{A path of length $k$ consists of $k$ edges and $k+1$ vertices.
+A path is \emph{trivial} if it has length 0 and \emph{non-trivial} otherwise.} and
 \item for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
@@
 In this setting, the proof of \cref{clm:general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $V_i\cap Y_x$.
-More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2},\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap V_j$ for $j > k+1$.
+More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap V_j$ for $j > k+1$.
 Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields 
@@
        + 4\left\lfloor\frac{k+1}{2}\right\rfloor 
-       + 2\left(k+1-\left\lfloor\frac{k+1}{2}\right\rfloor\right)
+       + 2\left\lceil\frac{k+1}{2}\right\rceil
        \right)
@@
        d\ell\left(
-         2k + 2\left\lfloor\frac{k+1}{2}\right\rfloor + 4
+         2 + 2(k+1) + 2\left\lfloor\frac{k+1}{2}\right\rfloor
        \right) \\

### 2019-12-04 k-planar.tex
Fixed mistake caused by k+1 versus k

@@
 \label{kPlanarBasic}
-Every $k$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth $(18k^2+30k)\binom{k+4}{3}-1$ and for some path $P$. 
+Every $k$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth $(18k^2+48k+30)\binom{k+4}{3}-1$ and for some path $P$. 
 \end{thm}
@@
 \label{k-planar}
-Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
+Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 48k+30$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
 \end{thm}
@@
 Every $(g,k)$-planar graph has an $H$-partition of layered width at most 
-$\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
+$\max\{2g,3\}\cdot(6k^2+16k+10)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
 \end{thm}
@@
 \end{align*}    
-With this change, the layered width of the partition given by \cref{modern-major-general} becomes $18k^2+30k$.  This establishes \cref{k-planar}.
+With this change, the layered width of the partition given by \cref{modern-major-general} becomes $(18k+30)(k+1)=18k^2+48k+30$.  This establishes \cref{k-planar}.
+% Note to self-and-others. In our original SODA submission we had 18k^2+30k here because we forgot that the layer compression step has compress k+1 consecutive layers, not just k consecutive layers.
 
@@
   %$\pi(G)\le 12(2k+3)(k+1)\times 4^{(k+2)(k+3)(k+4)/6}$.
-  $\pi(G)\le (18k^2 + 30k) 4^{\binom{k+4}{3}}$.
+  $\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
 %  Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}}-1$.
@@
 %    \pi(G)\le \max\{4g(3k+1),12(2k+3)(k+1)\}\times 4^{(k+2)(k+3)(k+4)/6} 
-    \pi(G)\le \max\{2g,3\}\cdot(6k^2 + 10k) 4^{\binom{k+5}{4}}. 
+    \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+5}{4}}. 
      \]
@@
 %$$\chi_p(G)\le (18k^2+30k)(p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
-\[ \chi_p(G)\le (18k^2+30k)(p+1) \binom{k+4}{3} \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}\enspace. \]
+\[ \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{k+4}{3} \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}\enspace. \]
 %\note{DW}{I previously wrote ``for some integer $t\leq \binom{k+4}{3}-1$'' thinking that $\binom{p+t}{t}$ might not be monotone with $t$, but it is because of the $+t$.}
@@
 %$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
-\[ \chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) \binom{k+5}{4}\binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}\enspace. \]
+\[ \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{k+5}{4}\binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}\enspace. \]
 %O(gk^2)\times p^{O(k^3)}$. 

### 2019-12-04 k-planar.tex
Beat this up into SoCG format

@@
-\documentclass{patmorin}
+\documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
+
+% \documentclass{socg-lipics-v2019}
+
 \usepackage[T1]{fontenc}
@@
 
-\usepackage[mathlines]{lineno}
-% \setlength{\linenumbersep}{2em}
-\linenumbers
-% \rightlinenumbers
+% \usepackage[mathlines]{lineno}
+% % \setlength{\linenumbersep}{2em}
 % \linenumbers
-\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
- \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
- \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
- \renewenvironment{#1}%
-    {\linenomath\csname old#1\endcsname}%
-    {\csname oldend#1\endcsname\endlinenomath}}% 
-\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
- \patchAmsMathEnvironmentForLineno{#1}%
- \patchAmsMathEnvironmentForLineno{#1*}}%
-\AtBeginDocument{%
-\patchBothAmsMathEnvironmentsForLineno{equation}%
-\patchBothAmsMathEnvironmentsForLineno{align}%
-\patchBothAmsMathEnvironmentsForLineno{flalign}%
-\patchBothAmsMathEnvironmentsForLineno{alignat}%
-\patchBothAmsMathEnvironmentsForLineno{gather}%
-\patchBothAmsMathEnvironmentsForLineno{multline}%
-}
+% % \rightlinenumbers
+% % \linenumbers
+% \newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+%  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+%  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+%  \renewenvironment{#1}%
+%     {\linenomath\csname old#1\endcsname}%
+%     {\csname oldend#1\endcsname\endlinenomath}}% 
+% \newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+%  \patchAmsMathEnvironmentForLineno{#1}%
+%  \patchAmsMathEnvironmentForLineno{#1*}}%
+% \AtBeginDocument{%
+% \patchBothAmsMathEnvironmentsForLineno{equation}%
+% \patchBothAmsMathEnvironmentsForLineno{align}%
+% \patchBothAmsMathEnvironmentsForLineno{flalign}%
+% \patchBothAmsMathEnvironmentsForLineno{alignat}%
+% \patchBothAmsMathEnvironmentsForLineno{gather}%
+% \patchBothAmsMathEnvironmentsForLineno{multline}%
+% }
 
@@
 
- \newcommand{\note}[2]{{\color{red}[#1:~#2]}}
+ % \newcommand{\note}[2]{{\color{red}[#1:~#2]}}
 % \newcommand{\notex}[2]{{\color{blue}[#1:~#2]}}
-\newcommand{\notex}[2]{}
+% \newcommand{\notex}[2]{}
 
@@
 
-\title{\MakeUppercase{The Structure of $k$-Planar Graphs}}
-
-\author{Vida Dujmovi\'c%
-        \thanks{School of Computer Science and Electrical Engineering,
-                University of Ottawa, Ottawa, Canada (\texttt{vida.dujmovic@uottawa.ca}).
-                Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\,
-        Pat Morin%
-        \thanks{School of Computer Science, Carleton University, Ottawa, Canada (\texttt{morin@scs.carleton.ca}).
-Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\, and
-        David R. Wood\thanks{School of Mathematics, Monash University, Melbourne, Australia (\texttt{david.wood@monash.edu}). Research supported by the Australian Research Council.}
-}
+% \title{\MakeUppercase{The Structure of $k$-Planar Graphs}}
+\title{The Structure of $k$-Planar Graphs}
+\titlerunning{The Structure of $k$-Planar Graphs}
+
+\author{Vida Dujmović}{School of Computer Science and Electrical Engineering, University of Ottawa}{vida.dujmovic@uottawa.ca}{[0000-0003-0471-4118]}{Research  supported by NSERC and the Ontario Ministry of Research and Innovation.}
+% , Ottawa, Canada (\texttt{vida.dujmovic@uottawa.ca}).
+%                 Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\,
+\author{Pat Morin}{School of Computer Science, Carleton University}{morin@scs.carleton.ca}{[0000-0003-0471-4118]}{Research supported by NSERC.}
+% , Ottawa, Canada (\texttt{morin@scs.carleton.ca}).                 Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\, and
+\author{David R. Wood}{School of Mathematics, Monash University}{david.wood@monash.edu}{[0000-0003-0471-4118]}{Research supported by the Australian Research Council.}
+% , Melbourne, Australia (\texttt{david.wood@monash.edu}). Research supported by the Australian Research Council.}
+% }
+
+\authorrunning{V. Dujmović, P. Morin, and D. R. Wood}
+
+\keywords{$k$-planar graphs, layered partitions, queue-number, nonrepetitive colouring, centered colouring, $k$-nearest-neighbour graphs}
+
+\Copyright{Vida Dujmović, Pat Morin, and David R. Wood}
 
 \begin{document}
-\begin{titlepage}
+% \begin{titlepage}
 \maketitle
@@
 \end{abstract}
-\end{titlepage}
+% \end{titlepage}
 
@@
 The present paper studies the structure of $k$-planar graphs and other more general classes of graphs.
-\notex{PM}{I removed the following sentences because neither Vida nor could find anything about $k$-planar graphs in \cite{FP08} and, in any case, the proof is easy.
+% \notex{PM}{I removed the following sentences because neither Vida nor could find anything about $k$-planar graphs in \cite{FP08} and, in any case, the proof is easy.
 This could easily upset a referee.
-% \note{PM}{I must be missing something.
+% % \note{PM}{I must be missing something.
 The planarized graph $G'$ has $O(kn)$ vertices.
@@
 }
-Various structural results about planar graphs generalise for $k$-planar graphs.
+% Various structural results about planar graphs generalise for $k$-planar graphs.
 For example, \citet{FP08} generalised the Lipton-Tarjan separator theorem to show that every $k$-planar graph with $n$ vertices has a balanced separator of order $O(\sqrt{kn})$.}
@@
 
-\notex{DW}{Is some sort of converse to \cref{generalized-tripod}  also true?
+% \notex{DW}{Is some sort of converse to \cref{generalized-tripod}  also true?
 \cref{generalized-tripod}  shows that if a graph $G$ has a layered partition with all the nice properties, then $G$ has this generalised tripod partition.
@@
 The proof is basically in \citep{dujmovic.joret.ea:planar}, right?}
-\notex{PM}{Yes, if by generalized tripod partition you mean $T$ and $\mathcal{Y}=(Y_x:x\in V(T))$ that satisify \yref{thickness}, \yref{separator}, and \yref{ancestor-edge}.
+% \notex{PM}{Yes, if by generalized tripod partition you mean $T$ and $\mathcal{Y}=(Y_x:x\in V(T))$ that satisify \yref{thickness}, \yref{separator}, and \yref{ancestor-edge}.
 From \yref{separator} and \yref{ancestor-edge} you can construct a width-$t$ $T$-decomposition of $H=G/\mathcal{Y}$.
@@
 The following converse of \cref{gAddDummyDegree} provides a sufficient condition for a graph to be $(g,k)$-planar that is slightly more restrictive than being obtained from a $(k,d)$-shortcut system of a $(g,k)$-planar graph.
-\notex{DW}{Is this sentence clear enough?}\notex{PM}{Sure.}
 

### 2019-12-05 k-planar.tex
Nearly done

@@
-\documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
+ \documentclass[a4paper,UKenglish,cleveref,autoref]{lipics-v2019}
 
@@
 \usepackage[utf8]{inputenc}
-\usepackage{amsmath,wrapfig}
+\usepackage{amsmath}
 \usepackage{amsfonts}
@@
 \usepackage{paralist}
-\usepackage{stmaryrd}
+% \usepackage{stmaryrd}
 \usepackage{thm-restate}
@@
 
-\allowdisplaybreaks
-%\sloppy
-%%%%%%%%%%
-\makeatletter
-\def\NAT@spacechar{~}
-\makeatother
+% \allowdisplaybreaks
+% %\sloppy
+% %%%%%%%%%%
+% \makeatletter
+% \def\NAT@spacechar{~}
+% \makeatother
 %%%%%%%%%%
@@
 
-\author{Vida Dujmović}{School of Computer Science and Electrical Engineering, University of Ottawa}{vida.dujmovic@uottawa.ca}{[0000-0003-0471-4118]}{Research  supported by NSERC and the Ontario Ministry of Research and Innovation.}
+\author{Vida Dujmović}{School of Computer Science and Electrical Engineering, University of Ottawa}{vida.dujmovic@uottawa.ca}{}{Research  supported by NSERC.}
 % , Ottawa, Canada (\texttt{vida.dujmovic@uottawa.ca}).
@@
 % , Ottawa, Canada (\texttt{morin@scs.carleton.ca}).                 Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\, and
-\author{David R. Wood}{School of Mathematics, Monash University}{david.wood@monash.edu}{[0000-0003-0471-4118]}{Research supported by the Australian Research Council.}
+\author{David R. Wood}{School of Mathematics, Monash University}{david.wood@monash.edu}{}{Research supported by the Australian Research Council.}
 % , Melbourne, Australia (\texttt{david.wood@monash.edu}). Research supported by the Australian Research Council.}
@@
 \Copyright{Vida Dujmović, Pat Morin, and David R. Wood}
+\ccsdesc{Mathematics of computing~Graph theory}
+\ccsdesc{Mathematics of computing~Extremal graph theory}
+\ccsdesc{Human-centered computing~Graph drawings}
+\relatedversion{This paper has been submitted to Discrete \&
+Computational Geometry. The current status, as of December 4th, 2019, is ``under review''.  A full version of this paper is available at
+                 \url{https://arxiv.org/abs/1907.05168}.}
+\funding{This work was partly funded by NSERC and MRI.}
 
 \begin{document}
-% \begin{titlepage}
+
 \maketitle
-
 \begin{abstract}
@@
 A path is \emph{trivial} if it has length 0 and \emph{non-trivial} otherwise.} and
-\item for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
+\item for every $v\in V(G)$, the number of paths in $\mathcal{P}$ with $v$ as an internal vertex is at most $d$.
 \end{compactitem} 
@@
 
-We prove the following theorem that shows if a graph $G$ is a subgraph of $H\boxtimes P$ and $\PP$ is a shortcut system for $G$, then $G^{\mathcal{P}}$ is a subgraph of $J\boxtimes P$, where the treewidth of $J$ is bounded by a function of the treewidth of $H$.
+We prove the following theorem that shows if a graph $G$ is a subgraph of $H\boxtimes P$ and $\PP$ is a shortcut system for $G$, then $G^{\mathcal{P}}$ is a subgraph of $J\boxtimes P$, where the treewidth of $J$ is bounded by a function of $k$, $d$, and the treewidth of $H$.
 
@@
 
-\cref{ShortcutProduct} leads to several other results of interest.
+\cref{ShortcutProduct} leads to several other results.
 Here is one example.
@@
 
-This section defines concepts and results from the literature that will be important for our work.
+This section defines concepts and results from the literature that are used in our work.
 
@@
 They defined a class $\mathcal{G}$ of graphs to \emph{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
-
 These definitions relate to strong products as follows.
@@
 \begin{compactenum}[(a)]
-\item an $H$-partition with layered width at most $\max\{2g,1\}$ such that $H$ has treewidth at most $9$;
-\item an $H$-partition with layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most $4$.
+\item an $H$-partition with layered width at most $\max\{2g,1\}$ where $H$ has treewidth at most $9$;
+\item an $H$-partition with layered width at most $\max\{2g,3\}$ where $H$ has treewidth at most $4$.
 %\item an $H$-partition with layered width at most $1$ such that $H$ has treewidth at most $2g+8$. 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Shortcut Systems}
+\section{Shortcut Systems: Proof of \cref{modern-major-general}}
 \label{Structure}
@@
 
-For a tree $T$ rooted at some node $x_0\in V(T)$, we we say that a node $a\in V(T)$ is a \emph{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \emph{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
-Note that each node $x\in v(T)$ is a $T$-ancestor and $T$-descendant of itself.
+For a tree $T$ rooted at some node $x_0\in V(T)$, we say that a node $a\in V(T)$ is a \emph{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \emph{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
+Note that each node $x\in V(T)$ is a $T$-ancestor and $T$-descendant of itself.
 We say that a $T$-ancestor $a\in V(T)$ of $x\in V(T)$ is a \emph{strict} $T$-ancestor of $x$ if $a\neq x$.
-The \emph{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.  For each node $x\in V(T)$, define
-\[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
-to be the maximal subtree of $T$ rooted at $x$.
-
-We begin with a fairly standard lemma about normalised tree decompositions.
+The \emph{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.
+For each node $x\in V(T)$, define $T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}]$ to be the maximal subtree of $T$ rooted at $x$.
+
+We begin with a fairly standard trick of working with a normalised tree decomposition.
 
 \begin{lem}\lemlabel{nice-decomposition}
-  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a $T$-decomposition $(B_x:x\in V(T))$ of $H$ with width $t$ that has following additional properties:  
+  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a $T$-decomposition $(B_x:x\in V(T))$ of $H$ of width at most $t$ that has following additional properties:  
   \begin{compactenum}[(T1)]
@@
 
-Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:
+We first point out more properties that are immediately implied by \cref{generalized-tripod}:
 
@@
 \begin{proof}[Proof of \cref{generalized-tripod}]
-  Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$
-  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
+  Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$ (a condition ensured by \lemref{nice-decomposition}) and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
 Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.
@@
 \begin{proof}
-Apply \cref{generalized-tripod} to $G$ and let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $Y_x$, $V_x$, $F_x$, and $N_x$ be defined as in \cref{generalized-tripod}.
+Apply \cref{generalized-tripod} to $G$ and let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $Y_x$, $V_x$, $F_x$, and $N_x$ be defined as in \cref{generalized-tripod} (so that $(\mathcal{L},\mathcal{Y})$ is an $H$-decomposition of $G$ of layered width $\ell$ where $H$ has treewidth $t$.)
 
@@
 We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$ or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
-We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
+Let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
 
@@
 \begin{proof}
-  Let 
-  \[
-     Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:\mbox{$(v_0,\ldots,v_r)\in\mathcal{P}$, $v\in \{v_0,\ldots,v_r\}$}\}
-  \]
+  Let  $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\mathcal{P}$, $v\in \{v_0,\ldots,v_r\}\}$
   Clearly $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
@@
 
-  Let $Z_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.  The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.  
+  Let $X_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.  The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.  
   %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other. 
@@
 
-For each $x\in V(T)$, we define the \emph{separator}
-\[  
-     S_x := \{v\in V(G): a(v)= x\} \enspace .
-\]
+For each $x\in V(T)$, we define the \emph{separator} $S_x := \{v\in V(G): a(v)= x\}$.
 Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
-We let $J:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph and we let $V(J)\subseteq V(T)$ in the obvious way, so that each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
+Let $J:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph and let $V(J)\subseteq V(T)$ in the obvious way, so that each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
 (Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $J$.)
@@
 %%%%%%%%%%%%%%%%%%%
-\section{\boldmath $k$-Planar Graphs}
+\section{\boldmath $k$-Planar Graphs: Proof of \cref{k-planar}}
 \label{sec-k-planar}
@@
 More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap V_j$ for $j > k+1$.
-Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields 
-\begin{align*}
-    |S_x\cap Y_i| \le d\ell\left( 
+Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields $ |S_x\cap Y_i| \le d\ell\left( 
        2 
@@
        \right)
-       & =
+        =
        d\ell\left(
@@
        \right) \\
-       & \le
+        \le
        d\ell(3k+5) 
-      = 18k+30 \enspace .
-\end{align*}    
+      = 18k+30$.
+
 With this change, the layered width of the partition given by \cref{modern-major-general} becomes $(18k+30)(k+1)=18k^2+48k+30$.  This establishes \cref{k-planar}.
 % Note to self-and-others. In our original SODA submission we had 18k^2+30k here because we forgot that the layer compression step has compress k+1 consecutive layers, not just k consecutive layers.
-
  Exactly the same approach using \cref{SurfacePartition}(b) instead of \cref{PlanarPartition}(b) establishes 
@@
 
-\subsection{\boldmath  $1$-Planar Graphs}
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Applications}
+\label{Applications}
+Our main theorems have a number of applications including queue layouts, non-repetitive colourings and clustered colourings  of $k$-planar and $(g,k)$-planar graphs.
+We present here queue layout results and defer non-repetitive colourings and clustered colourings to \cref{applmore}.
+
+%%%%%%%%%%%%%%%%%%%
+\subsection{Queue Layouts}
+
+For an integer $k\geq 0$, a \textit{$k$-queue layout} of a graph $G$ consists of a linear ordering $\preceq$ of $V(G)$ and a partition $\{E_1,E_2,\dots,E_k\}$ of $E(G)$, such that for $i\in\{1,2,\dots,k\}$, no two edges in $E_i$ are nested with respect to $\preceq$.
+That is, it is not the case that $v\prec x \prec y \prec w$ for edges $vw,xy\in E_i$.
+The \textit{queue-number} of a graph $G$, denoted by $\qn(G)$, is the minimum integer $k$ such that $G$ has a $k$-queue layout.
+Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
+\citet{dujmovic.joret.ea:planar} recently proved this conjecture using \cref{PlanarPartition} and the following lemma.
+Indeed, resolving this question was the motivation for the development of \cref{PlanarBasic,PlanarPartition}.
+
+\begin{lem}[\citep{dujmovic.joret.ea:planar}]
+\label{qn}
+If a graph $G$ has an $H$-partition of layered width $\ell$, then  $\qn(G)\le 3 \ell \, \qn(H) + \floor{\tfrac{3}{2}\ell}$.
+\end{lem}
+
+\citet{dujmovic.morin.ea:layered} proved that queue-number is bounded by a function of treewidth. 
+The best known bound is due to \citet{wiechert:on}:
+
+\begin{thm}[\citep{wiechert:on}]
+\label{qn-tw}
+    For every graph $G$, $\qn(G)\le 2^{\tw(G)}-1$.
+\end{thm}
+
+\cref{qn,qn-tw,g-k-planar} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^4)}$.
+Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of  $O(g^{k+2})$ using \cref{SurfacePartition} and an ad-hoc method.
+Our result provides a better bound when $g>2^{k^3}$.
+In the case of 1-planar graphs we can improve further.
+\citet{ABGKP18} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
+Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$.
+\cref{qn,1-planar} then imply:
+
+\begin{prop}
+Every 1-planar graph has queue-number at most $3 \times 30 \times 5 + \floor{\tfrac{3}{2} \times 30} = 495$. \end{prop}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Graph Classes: Examples}
+\label{Examples}
+There are several examples of graph classes that can be obtained from a shortcut system applied to graphs of bounded Euler genus.
+These include map graphs, string graphs, powers of bounded degree graphs and $k$-nearest-neighbour graphs.
+We present here  $k$-nearest-neighbour graphs results and defer the others to \cref{exmore}.
+
+\subsection{$k$-Nearest-Neighbour Graphs}
+
+In this section, we show that $k$-nearest neighbour graphs of point sets in the plane are $O(k^2)$-planar.
+For two points $x,y\in\R^2$, let $d_2(x,y)$ denote the Euclidean distance between $x$ and $y$.
+The $k$-nearest-neighbour graph of a point set $P\subset\R^2$ is the geometric graph $G$ with vertex set $V(G)=P$, where the edge set is defined as follows.
+For each point $v\in P$, let $N_k(v)$ be the set of $k$ points in $P$ closest to $v$.
+Then $vw\in E(G)$ if and only if $w\in N_k(v)$ or $v\in N_k(w)$.
+% and edge set
+%\[  
+%  E= \{vw : v,w\in V(G),\, |\{x\in P\setminus\{v\}: d_2(v,x)\le d_2(v,w)\}|\le k\} \enspace .
+%\] 
+(The edges of $G$ are straight-line segments joining their endpoints.) See \citep{ProximityGraphs} for a survey of results on $k$-nearest neighbour graphs and other related proximity graphs.
+
+The following result, which is immediate from \citet[Corollary~4.2.6]{abrego.munroy.ea:on} states that $k$-nearest-neighbour graphs have bounded maximum degree:
+\begin{lem}
+\label{k-nn-max-degree}
+The degree of every vertex in a $k$-nearest-neighbour graph is at most $6k$.
+\end{lem}
+
+We make use of the following well-known observation (see for example, \citet[Lemma~2]{bose.morin.ea:routing}):
+\begin{obs}
+\label{convex}
+If $v_0,\ldots,v_3$ are the vertices of a convex quadrilateral in counterclockwise order then there exists at least one $i\in\{0,\ldots,3\}$ such that $\max\{d_2(v_i,v_{i-1}), d_2(v_i,v_{i+1})\} < d_2(v_{i-1},v_{i+1})$, where subscripts are taken modulo 4.
+\end{obs}
+
+\begin{lem}
+\label{nearest-neighbour}
+  Every $k$-nearest-neighbour graph is $O(k^2)$-planar.
+\end{lem}
+
+\begin{proof}
+  Let $G$ be a $k$-nearest-neighbour graph and consider any edge $vw\in E(G)$.
+  Let $xy\in E(G)$ be an edge that crosses $vw$.  Note that $vxwy$ are the vertices of a convex quadrilateral in (without loss of generality) counterclockwise order. Then we say that 
+  \begin{compactenum}
+    \item $xy$ is of Type~$v$ if $\max\{d_2(v,x), d_2(v,y)\}< d_2(x,y)$;
+    \item $xy$ is of Type~$w$ if $\max\{d_2(w,x), d_2(w,y)\}< d_2(x,y)$; or
+    \item $xy$ is of Type~C otherwise.
+  \end{compactenum}
+  If $xy$ is of Type~C, then \cref{convex} implies that (without loss of generality) 
+  \[\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)\enspace.\] 
+   In this case, we call $x$ a Type~C vertex.
+We claim that $V(G)$ contains at most $k-1$ Type~C vertices.
+Indeed, more than $k-1$ Type~C vertices would contradict the fact that $vw\in E(G)$ since every Type~C vertex is closer to both $v$ and $w$ than $d_2(v,w)$.
+
+  Next oberve that, if $xy$ is of Type~$v$, then at least one of $xv$ or $yv$ is in $E(G)$ in which case we call $x$ (respectively $y$) a Type~$v$ vertex.
+By \cref{k-nn-max-degree}, there are at most $6k$ Type~$v$ vertices.
+Similarly, there are at most $6k$ Type~$w$ vertices.
+
+  Thus, in total, there are at most $13k-1$ Type~$v$, Type~$w$, and Type~C vertices.
+By \cref{k-nn-max-degree}, each of these vertices is incident with at most $6k$ edges that cross $vw$.
+Therefore, there are at most $78k^2-6k$ edges of $G$ that cross $vw$.
+Since this is true for every edge $vw\in E(G)$, $G$ is $(78k^2-6k)$-planar.
+\end{proof}
+
+We note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n$ vertices has at least $kn/2$ edges and at most $kn$ edges.
+For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is $\Omega(k^2)$.
+
+\cref{nearest-neighbour,k-planar,qn,p-centered,p-centered-treewidth} imply:
+
+\begin{cor}
+\label{k-nn}
+Every $k$-nearest-neighbour graph has:
+\begin{compactitem}
+\item an $H$-partition of layered width $O(k^4)$ in which $H$ has treewidth $O(k^6)$,
+\item queue-number at most $2^{O(k^6)}$,
+\item $p$-centered chromatic number at most $O(k^4) (p+1) (t+1)\binom{p+t}{t}$ for some integer $t\leq O(k^6)$.
+\end{compactitem}
+\end{cor}
+
+\cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
+However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (36+o(1))k^2$ by a result of \citet{DJKW16}.
+
+%%%  Squashing the bibliography 
+  \let\oldthebibliography=\thebibliography
+  \let\endoldthebibliography=\endthebibliography
+  \renewenvironment{thebibliography}[1]{%
+    \begin{oldthebibliography}{#1}%
+      \setlength{\parskip}{0.4ex}%
+      \setlength{\itemsep}{0.4ex}%
+  }{\end{oldthebibliography}}
+
+\bibliographystyle{plainnat}
+\bibliography{k-planar}
+
+% % NOTE THAT MANUAL CHANGES HAVE BEEN MADE BELOW.
+% 
+% \fontsize{11}{11.75}\selectfont 
+% 
+% \begin{thebibliography}{33}
+% \providecommand{\natexlab}[1]{#1}
+% \providecommand{\url}[1]{\texttt{#1}}
+% \expandafter\ifx\csname urlstyle\endcsname\relax
+%   \providecommand{\doi}[1]{doi: #1}\else
+%   \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
+% 
+% \bibitem[{\'{A}}brego et~al.(2011){\'{A}}brego, Monroy,
+%   Fern{\'{a}}ndez{-}Merchant, Flores{-}Pe{\~{n}}aloza, Hurtado,
+%   Sacrist{\'{a}}n, and Saumell]{abrego.munroy.ea:on}
+% Bernardo~M. {\'{A}}brego, Ruy~Fabila Monroy, Silvia Fern{\'{a}}ndez{-}Merchant,
+%   David Flores{-}Pe{\~{n}}aloza, Ferran Hurtado, Vera Sacrist{\'{a}}n, and
+%   Maria Saumell.
+% \newblock On crossing numbers of geometric proximity graphs.
+% \newblock \emph{Comput. Geom.}, 44\penalty0 (4):\penalty0 216--233, 2011.
+% \newblock \doi{10.1016/j.comgeo.2010.11.003}.
+% 
+% \bibitem[Ajtai et~al.(1982)Ajtai, Chv{\'a}tal, Newborn, and
+%   Szemer{\'e}di]{ajtai.chvatal.ea:crossing-free}
+% Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, Endre
+%   Szemer{\'e}di.
+% \newblock Crossing-free subgraphs.
+% \newblock \emph{Ann. Discrete Math.}, 12:\penalty0 9--12, 1982.
+% \newblock \doi{10.1016/S0304-0208(08)73484-4}.
+% 
+% \bibitem[Alam et~al.(2018)Alam, Bekos, Gronemann, Kaufmann, and
+%   Pupyrev]{ABGKP18}
+% Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael Kaufmann, Sergey Pupyrev.
+% \newblock Queue layouts of planar 3-trees.
+% \newblock In Therese~C. Biedl and Andreas Kerren, editors, \emph{Proc. 26th
+%   International Symposium on Graph Drawing and Network Visualization} (GD '18), vol.~11282 of \emph{Lecture Notes in Computer Science}, pp.~213--226. Springer, 2018.
+% \newblock \doi{10.1007/978-3-030-04414-5\_15}.
+% 
+% \bibitem[Alon et~al.(2002)Alon, Grytczuk, Ha{\l}uszczak, and
+%   Riordan]{AGHR-RSA02}
+% Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, Oliver Riordan.
+% \newblock Nonrepetitive colorings of graphs.
+% \newblock \emph{Random Structures Algorithms}, 21\penalty0 (3-4):\penalty0
+%   336--346, 2002.
+% \newblock \doi{10.1002/rsa.10057}.
+% 
+% \bibitem[Bodlaender(1998)]{Bodlaender-TCS98}
+% Hans~L. Bodlaender.
+% \newblock A partial $k$-arboretum of graphs with bounded treewidth.
+% \newblock \emph{Theoret. Comput. Sci.}, 209\penalty0 (1-2):\penalty0 1--45,
+%   1998.
+% \newblock \doi{10.1016/S0304-3975(97)00228-4}.
+% 
+% \bibitem[Bonamy et~al.(2019)Bonamy, Gavoille, and
+%   Pilipczuk]{bonamy.gavoille.ea:shorter}
+% Marthe Bonamy, Cyril Gavoille, and Micha{\l} Pilipczuk.
+% \newblock Shorter implicit representation for planar graphs.
+% \newblock In S.~Cabello and B.~Mohar, editors, \emph{9th Slovenian
+%   International Conference on Graph Theory} (SICGT), p.~63, June 2019.
+% 
+% \bibitem[Bose et~al.(2001)Bose, Morin, Stojmenovi\'c, and
+%   Urrutia]{bose.morin.ea:routing}
+% Prosenjit Bose, Pat Morin, Ivan Stojmenovi\'c, Jorge Urrutia.
+% \newblock Routing with guaranteed delivery in ad hoc wireless networks.
+% \newblock \emph{Wireless Networks}, 7\penalty0 (6):\penalty0 609--616, 2001.
+% \newblock \doi{10.1023/A:1012319418150}.
+% 
+% \bibitem[Bose et~al.(2012)Bose, Dujmovi\'c, Hurtado, Iacono, Langerman, Meijer,
+%   Sacrist\'an, Saumell, and Wood]{ProximityGraphs}
+% Prosenjit Bose, Vida Dujmovi\'c, Ferran Hurtado, John Iacono, Stefan Langerman,
+%   Henk Meijer, Vera Sacrist\'an, Maria Saumell, David~R. Wood.
+% \newblock Proximity graphs: $e$, $\delta$, $\delta$, $\delta$, $\chi$ and
+%   $\omega$.
+% \newblock \emph{Internat. J. Comput. Geom.  Appl.},
+%   22\penalty0 (5):\penalty0 439--469, 2012.
+% 
+% \bibitem[Chen et~al.(2002)Chen, Grigni, and Papadimitriou]{CGP02}
+% Zhi-Zhong Chen, Michelangelo Grigni, Christos~H. Papadimitriou.
+% \newblock Map graphs.
+% \newblock \emph{J. ACM}, 49\penalty0 (2):\penalty0 127--138, 2002.
+% \newblock \doi{10.1145/506147.506148}.
+% 
+% \bibitem[D{\k e}bski et~al.(2019)D{\k e}bski, Felsner, Micek, and Schr{\"o}der]{micek:personal}
+%   Micha{\l} D{\k e}bski, Stefan Felsner, Piotr Micek, Felix Schr{\" o}der. 
+% \newblock Improved bounds for centered colorings, 2019.
+% \newblock \href{https://arxiv.org/abs/1907.04586}{arXiv:1907.04586}.
+% 
+% \bibitem[Dujmovi\'{c} et~al.(2016)Dujmovi\'{c}, Joret, Kozik, and Wood]{DJKW16}
+% Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, David~R. Wood.
+% \newblock Nonrepetitive colouring via entropy compression.
+% \newblock \emph{Combinatorica}, 36\penalty0 (6):\penalty0 661--686, 2016.
+% \newblock \doi{10.1007/s00493-015-3070-6}.
+% 
+% \bibitem[Dujmovi\'c et~al.(2017{\natexlab{a}})Dujmovi\'c, Eppstein, and
+%   Wood]{dujmovic.eppstein.ea:structure}
+% Vida Dujmovi\'c, David Eppstein, David~R. Wood.
+% \newblock Structure of graphs with locally restricted crossings.
+% \newblock \emph{{SIAM} J. Discrete Math.}, 31\penalty0 (2):\penalty0 805--824,
+%   2017{\natexlab{a}}.
+% \newblock \doi{10.1137/16M1062879}.
+% 
+% \bibitem[Dujmovi\'c et~al.(2017{\natexlab{b}})Dujmovi\'c, Morin, and
+%   Wood]{dujmovic.morin.ea:layered}
+% Vida Dujmovi\'c, Pat Morin, David~R. Wood.
+% \newblock Layered separators in minor-closed graph classes with applications.
+% \newblock \emph{J. Comb. Theory, Ser. {B}}, 127:\penalty0 111--147,
+%   2017{\natexlab{b}}.
+% \newblock \doi{10.1016/j.jctb.2017.05.006}.
+% 
+% \bibitem[Dujmovi\'c et~al.(2019{\natexlab{a}})Dujmovi\'c, Esperet, Joret,
+%   Walczak, and Wood]{dujmovic.esperet.ea:planar}
+% Vida Dujmovi\'c, Louis Esperet, Gwena{\"{e}}l Joret, Bartosz Walczak,   David~R. Wood.
+% \newblock Planar graphs have bounded nonrepetitive chromatic number.
+% \newblock  \href{http://arxiv.org/abs/1904.05269}{arXiv:1904.05269}, 2019{\natexlab{a}}.
+% 
+% \bibitem[Dujmovi\'c et~al.(2019{\natexlab{b}})Dujmovi\'c, Joret, Micek, Morin,
+%   Ueckerdt, and Wood]{dujmovic.joret.ea:planar}
+% Vida Dujmovi\'c, Gwena{\"{e}}l Joret, Piotr Micek, Pat Morin, Torsten Ueckerdt, David~R. Wood.
+% \newblock Planar graphs have bounded queue-number.
+% \newblock \emph{Proc. 60th Annual Symp. on Foundations of Computer 
+%   Science} (FOCS '19), to appear. 
+%   \newblock  \href{http://arxiv.org/abs/1904.04791}{arXiv:1904.04791}, 2019{\natexlab{b}}.
+% 
+% \bibitem[Fomin et~al.(2012)Fomin, Lokshtanov, and Saurabh]{FLS-SODA12}
+% Fedor~V. Fomin, Daniel Lokshtanov, Saket Saurabh.
+% \newblock Bidimensionality and geometric graphs.
+% \newblock In \emph{Proc. 23rd {A}nnual {ACM}-{SIAM} {S}ymposium on {D}iscrete
+%   {A}lgorithms} (SODA 2012), pp.~1563--1575, 2012.
+% \newblock \doi{10.1137/1.9781611973099.124}.
+% 
+% \bibitem[Fox and Pach(2008)]{FP08}
+% Jacob Fox, J{\'a}nos Pach.
+% \newblock Separator theorems and {T}ur\'an-type results for planar intersection graphs.
+% \newblock \emph{Adv. Math.}, 219(3):1070--1080, 2008.
+% \newblock \doi{10.1016/j.aim.2008.06.002}.
+% 
+% \bibitem[Fox and Pach(2010)]{FP10}
+% Jacob Fox, J{\'a}nos Pach.
+% \newblock A separator theorem for string graphs and its applications.
+% \newblock \emph{Combin. Probab. Comput.}, 19\penalty0 (3):\penalty0 371--390,
+%   2010.
+% \newblock \doi{10.1017/S0963548309990459}.
+% 
+% \bibitem[Fox and Pach(2014)]{FP14}
+% Jacob Fox, J{\'a}nos Pach.
+% \newblock Applications of a new separator theorem for string graphs.
+% \newblock \emph{Combin. Probab. Comput.}, 23\penalty0 (1):\penalty0 66--74,
+%   2014.
+% \newblock \doi{10.1017/S0963548313000412}.
+% 
+% \bibitem[Grigoriev and Bodlaender(2007)]{grigoriev.bodlaender:algorithms}
+% Alexander Grigoriev, Hans~L. Bodlaender.
+% \newblock Algorithms for graphs embeddable with few crossings per edge.
+% \newblock \emph{Algorithmica}, 49\penalty0 (1):\penalty0 1--11, 2007.
+% \newblock \doi{10.1007/s00453-007-0010-x}.
+% 
+% \bibitem[Harvey and Wood(2017)]{HW17}
+% Daniel~J. Harvey, David~R. Wood.
+% \newblock Parameters tied to treewidth.
+% \newblock \emph{J. Graph Theory}, 84\penalty0 (4):\penalty0 364--385, 2017.
+% \newblock \doi{10.1002/jgt.22030}.
+% 
+% \bibitem[Heath et~al.(1992)Heath, Leighton, and Rosenberg]{HLR92}
+% Lenwood~S. Heath, F.~Thomson Leighton, Arnold~L. Rosenberg.
+% \newblock Comparing queues and stacks as mechanisms for laying out graphs.
+% \newblock \emph{SIAM J. Discrete Math.}, 5\penalty0 (3):\penalty0 398--412,
+%   1992.
+% \newblock \doi{10.1137/0405031}.
+% 
+% \bibitem[van~den Heuvel and Wood(2017)]{vdHW17}
+% Jan van~den Heuvel, David~R. Wood.
+% \newblock Improper colourings inspired by {H}adwiger's conjecture.
+% \newblock \emph{J. London Math. Soc.}, 98(1):129--148, 2018. 
+% \newblock  \doi{10.1112/jlms.12127}. See   \href{https://arxiv.org/abs/1704.06536}{arXiv:1704.06536} for the cited results. 
+% 
+% \bibitem[Kannan et~al.(1988)Kannan, Naor, and
+%   Rudich]{kannan.naor.ea:implicit-stoc}
+% Sampath Kannan, Moni Naor, Steven Rudich.
+% \newblock Implicit representation of graphs.
+% \newblock In Janos Simon, editor, \emph{Proc.\ 20th Annual {ACM} Symposium on
+%   Theory of Computing} (STOC '88), pp.~334--343. {ACM}, 1988.
+% \newblock \doi{10.1145/62212.62244}.
+% 
+% \bibitem[Kannan et~al.(1992)Kannan, Naor, and Rudich]{kannan.naor.ea:implicit}
+% Sampath Kannan, Moni Naor, Steven Rudich.
+% \newblock Implicit representation of graphs.
+% \newblock \emph{{SIAM} J. Discrete Math.}, 5\penalty0 (4):\penalty0 596--603,
+%   1992.
+% \newblock \doi{10.1137/0405049}.
+% 
+% \bibitem[Kobourov et~al.(2017)Kobourov, Liotta, and
+%   Montecchiani]{kobourov.liotta.ea:annotated}
+% Stephen~G. Kobourov, Giuseppe Liotta, Fabrizio Montecchiani.
+% \newblock An annotated bibliography on 1-planarity.
+% \newblock \emph{Comput. Sci. Rev.}, 25:\penalty0 49--67, 2017.
+% \newblock \doi{10.1016/j.cosrev.2017.06.002}.
+% 
+% \bibitem[Leighton(1983)]{leighton:complexity}
+% F.~Thomson Leighton.
+% \newblock \emph{Complexity issues in {VLSI}}.
+% \newblock MIT Press, Cambridge, 1983.
+% 
+% \bibitem[Mohar and Thomassen(2001)]{mohar.thomassen:graphs}
+% Bojan Mohar, Carsten Thomassen.
+% \newblock \emph{Graphs on surfaces}.
+% \newblock Johns Hopkins Press, 2001.
+% 
+% \bibitem[Ne{\v{s}}et{\v{r}}il and Ossona~de Mendez(2012)]{Sparsity}
+% Jaroslav Ne{\v{s}}et{\v{r}}il, Patrice Ossona~de Mendez.
+% \newblock \emph{Sparsity}, vol.~28 of \emph{Algorithms and Combinatorics}.
+% \newblock Springer, 2012.
+% \newblock \doi{10.1007/978-3-642-27875-4}.
+% 
+% \bibitem[Pach and T{\'o}th(1997)]{PachToth97}
+% J{\'a}nos Pach, G{\'e}za T{\'o}th.
+% \newblock Graphs drawn with few crossings per edge. 
+% \newblock \emph{Combinatorica}, 17(3):427--439, 1997
+%   2002.
+% \newblock \doi{10.1007/BF01215922}.
+% 
+% \bibitem[Pach and T{\'o}th(2002)]{PachToth-DCG02}
+% J{\'a}nos Pach, G{\'e}za T{\'o}th.
+% \newblock Recognizing string graphs is decidable.
+% \newblock \emph{Discrete Comput. Geom.}, 28\penalty0 (4):\penalty0 593--606,
+%   2002.
+% \newblock \doi{10.1007/s00454-002-2891-4}.
+% 
+% \bibitem[Pilipczuk and Siebertz(2019)]{pilipczuk.siebertz:polynomial-soda}
+% Micha{\l} Pilipczuk, Sebastian Siebertz.
+% \newblock Polynomial bounds for centered colorings on proper minor-closed graph
+%   classes.
+% \newblock In Timothy~M. Chan, editor, \emph{Proc.\ 30th Annual {ACM-SIAM}
+%   Symposium on Discrete Algorithms} ({SODA} 2019), pp.~1501--1520. {SIAM},
+%   2019.
+% \newblock \doi{10.1137/1.9781611975482.91}.
+% 
+% \bibitem[Reed(2003)]{Reed03}
+% Bruce~A. Reed.
+% \newblock Algorithmic aspects of tree width.
+% \newblock In \emph{Recent Advances in Algorithms and Combinat.}, vol.~11,
+%   pp.~85--107. Springer, 2003, 
+% \newblock \doi{10.1007/0-387-22444-0_4}.
+% 
+% \bibitem[Wiechert(2017)]{wiechert:on}
+% Veit Wiechert.
+% \newblock On the queue-number of graphs with bounded tree-width.
+% \newblock \emph{Electr. J. Comb.}, 24\penalty0 (1):\penalty0 P1.65, 2017.
+% \newblock  \url{http://www.combinatorics.org/v24i1p65}.
+% 
+% \bibitem[Zhu(2009)]{zhu:colouring}
+% Xuding Zhu.
+% \newblock Colouring graphs with bounded generalized colouring number.
+% \newblock \emph{Discrete Math.}, 309\penalty0 (18):\penalty0 5562--5568,
+%   2009.
+% \newblock \doi{10.1016/j.disc.2008.03.024}.
+% 
+% \end{thebibliography}
+
+\appendix
+
+\section{\boldmath  $1$-Planar Graphs: proof of \cref{1-planar}}
 \label{sec-1-planar}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Applications}
-\label{Applications}
-
-Here we discuss some of the consequences of our main theorems for $k$-planar and $(g,k)$-planar graphs.
-
-%%%%%%%%%%%%%%%%%%%
-\subsection{Queue Layouts}
-
-For an integer $k\geq 0$, a \textit{$k$-queue layout} of a graph $G$ consists of a linear ordering $\preceq$ of $V(G)$ and a partition $\{E_1,E_2,\dots,E_k\}$ of $E(G)$, such that for $i\in\{1,2,\dots,k\}$, no two edges in $E_i$ are nested with respect to $\preceq$.
-That is, it is not the case that $v\prec x \prec y \prec w$ for edges $vw,xy\in E_i$.
-The \textit{queue-number} of a graph $G$, denoted by $\qn(G)$, is the minimum integer $k$ such that $G$ has a $k$-queue layout.
-Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
-\citet{dujmovic.joret.ea:planar} recently proved this conjecture using \cref{PlanarPartition} and the following lemma.
-Indeed, resolving this question was the motivation for the development of \cref{PlanarBasic,PlanarPartition}.
-
-\begin{lem}[\citep{dujmovic.joret.ea:planar}]
-\label{qn}
-If a graph $G$ has an $H$-partition of layered width $\ell$, then 
-\[ \qn(G)\le 3 \ell \, \qn(H) + \floor{\tfrac{3}{2}\ell}\enspace.\]
-\end{lem}
-
-\citet{dujmovic.morin.ea:layered} proved that queue-number is bounded by a function of treewidth. 
-The best known bound is due to \citet{wiechert:on}:
-
-\begin{thm}[\citep{wiechert:on}]
-\label{qn-tw}
-    For every graph $G$, $\qn(G)\le 2^{\tw(G)}-1$.
-\end{thm}
-
-\cref{qn,qn-tw,g-k-planar} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^4)}$. 
-Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of 
-$O(g^{k+2})$ using \cref{SurfacePartition} and an ad-hoc method.
-Our result provides a better bound when $g>2^{k^3}$.
-In the case of 1-planar graphs we can improve further.
-\citet{ABGKP18} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
-Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$.
-\cref{qn,1-planar} then imply:
-
-\begin{prop}
-Every 1-planar graph has queue-number at most $3 \times 30 \times 5 + \floor{\tfrac{3}{2} \times 30} = 495$. \end{prop}
+\label{applmore}
 
@@
 \begin{cor}
-  For every $(g,k)$-planar graph $G$, 
-  \[ 
-%    \pi(G)\le \max\{4g(3k+1),12(2k+3)(k+1)\}\times 4^{(k+2)(k+3)(k+4)/6} 
-    \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+5}{4}}. 
-     \]
+  For every $(g,k)$-planar graph $G$,  $\pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+5}{4}}$.
+
 %  \label{g-k-planar} Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most  
@@
 
-The following lemma is implicitly due to \citet{micek:personal}. We include the proof for completeness.
+The following lemma is implicitly due to \citet{micek:personal}. %We include the proof for completeness.
 
@@
 
-\begin{proof}
-   Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$.
-   Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.
+% \begin{proof}
+%    Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$.
+%    Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.
 For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in V_i\cap B_x$ such that:
-   \begin{compactenum}
-     \item $\alpha(v)$ is unique among $\{\phi(w): w\in V_i\cap B_x\}$, which is possible 
-     since $|V_i\cap B_x|\le \ell$,
-     \item $\beta(v)= i\bmod (p+1)$, and     
-     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
-   \end{compactenum}
- To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
-
- First suppose that there exists $v,w\in V(X)$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$.
+%    \begin{compactenum}
+%      \item $\alpha(v)$ is unique among $\{\phi(w): w\in V_i\cap B_x\}$, which is possible 
+%      since $|V_i\cap B_x|\le \ell$,
+%      \item $\beta(v)= i\bmod (p+1)$, and     
+%      \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
+%    \end{compactenum}
+%  To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
+
+%  First suppose that there exists $v,w\in V(X)$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$.
 Since $G[X]$ is connected, $G[X]$ contains a path from $v$ to $w$.
@@
 
- Otherwise, $X\subseteq V_{i},\ldots,V_{i+s}$ for some $s<p$.
+%  Otherwise, $X\subseteq V_{i},\ldots,V_{i+s}$ for some $s<p$.
 Let $H':=H[\{x\in V(H):B_x\cap X\neq\emptyset]$.
@@
 Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
- For any $v,w\in B_x$ with $v\neq w$, either $v,w\in V_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in V_{i'}$ and $w\in V_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
+%  For any $v,w\in B_x$ with $v\neq w$, either $v,w\in V_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in V_{i'}$ and $w\in V_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
 Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.
 Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
- \end{proof}
+%  \end{proof}
 
@@
 
-%%%%%%%%%
-\section{Rough Characterisation}
-\label{Characterisation}
-
-This section presents a rough characterisation of $(g,k)$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $(g,k)$-planar.
-First note the following easy result (just add a degree-4 vertex at each crossing point of $G$, as in the proof of \cref{AddDummy,gAddDummy}).
-
-\begin{lem}
-\label{gAddDummyDegree}
-For every $(g,k)$-planar graph $G$ there is a graph $G_0$ of Euler
-genus at most $g$, such that $V(G) \subseteq V(G_0)$ and for every
-edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most
-$k+1$, such that every internal vertex in $P$ has degree at most $4$
-in $G_0$.
-\end{lem}
-
-The following converse of \cref{gAddDummyDegree} provides a sufficient condition for a graph to be $(g,k)$-planar that is slightly more restrictive than being obtained from a $(k,d)$-shortcut system of a $(g,k)$-planar graph.
-
-\begin{lem} 
-\label{DrawG}
-Fix integers $g\geq 0$ and $k,\Delta\geq 2$. 
-Let $G_0$ be a graph of Euler genus at most $g$. Let $G$ be
-a graph with $V(G) \subseteq V(G_0)$ such that for every edge $vw \in
-E(G)$ there is a $vw$-path $P_{vw}$ in $G_0$ of length at most $k$, such
-that every internal vertex on $P_{vw}$ has degree at most $\Delta$ in
-$G_0$. Then $G$ is $(g, 2k(k+1)\Delta^{k} )$-planar.
-\end{lem}
-
-\begin{proof}
-For a vertex $x$ of $G_0$ with degree at most $\Delta$, and for $i\in\{1,\dots,k-1\}$, say a vertex $v$ is \emph{$i$-close} to $x$ if there is a $vx$-path $P$ in $G_0$ of length at most $i$ such that every internal vertex in $P$  has degree at most $\Delta$ in $G_0$.
-For each edge $vw$ of $G$, say that $vw$ \emph{passes through} each internal vertex on $P_{vw}$. 
-Say $vw$ passes through $x$. Then $v$ is $i$-close to $x$ and $w$ is $j$-close to $x$ for some $i,j\in\{1,\dots,k-1\}$ with $i+j\leq k$. At most $\Delta^{i}$ vertices are $i$-close to $x$. 
-Thus, the number of edges of $G$ that pass through $x$ is at most 
-\[
-\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i \Delta^j 
-= \sum_{i=1}^{k-1} \Delta^i  \sum_{j=1}^{k-i} \Delta^j 
-< \sum_{i=1}^{k-1} \Delta^i  2 \Delta^{k-i}
-= \sum_{i=1}^{k-1} 2\Delta^k 
-< 2k \Delta^k \enspace.
-\]
-Draw each edge $vw$ of $G$ alongside $P_{vw}$ in $G_0$, so that every pair of edges cross at most once. 
-Every edge of $G$ that crosses $vw$ passes through a vertex on $P_{vw}$ (including $v$ and/or $w$ if they too have degree at most $\Delta$).
-Since $P_{vw}$ has at most $k+1$ vertices, and less than $2k\Delta^{k}$ edges of $G$ pass through each vertex on $P_{vw}$, the edge $vw$ is crossed by less than $2k(k+1)\Delta^{k}$ edges in $G$.
-Hence $G$ is $(g, 2k(k+1)\Delta^{k} )$-planar.
-\end{proof}
-
-Together \cref{DrawG,gAddDummyDegree} provide a rough characterisation of $(g,k)$-planar graphs.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Examples}
-\label{Examples}
-
-This section describes several examples of graph classes that can be obtained from a shortcut system applied to graphs of bounded Euler genus.
-
-%This section describes several graph classes that are $k$-planar or $(g,k)$-planar. The first two examples are well known, the second two are new.
-
-%We conclude by pointing out three other graph classes, namely $(g,d)$-map graphs, $(g,\delta)$-string graphs, and $k$-nearest-neighbour graphs that are $(g,k)$-planar for $k=O(d^2)$ and $k=\delta^2$, respectively.
-This implies that Theorems~\ref{k-planar}--\ref{g-k-planar} and Corollaries~\ref{cor:first}--\ref{cor:last} apply to $(g,d)$-map graphs, $(g,\delta)$-string graphs, and $k$-nearest-neighbour graphs.
+\section{Graph Classes: examples}
+\label{exmore}
 
@@
 Thus $G$ is $(g,\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
-Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar.
+% Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar.
  \cref{g-k-planar} then shows that map graphs admit bounded layered partitions, but we get much better bounds by constructing a shortcut system directly.
@@
 Recall that the \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
-If $G$ is planar with maximum degree $\Delta$, then $G^k$ is $2k(k+1)\Delta^{k}$-planar by \cref{DrawG}.
+If $G$ is planar with maximum degree $\Delta$, then $G^k$ is $2k(k+1)\Delta^{k}$-planar.
 Thus we can immediately conclude that bounded powers of planar graphs of bounded degree admit bounded layered partitions.
@@
 The number of vertices at distance $i$ from $v$ is at most $\Delta^i$.
-Thus the number of paths in $\PP$ that use $v$ as an internal vertex is at most
+Thus the number of paths in $\PP$ with $v$ as an internal vertex is at most
 \[\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i\Delta^j 
@@
 
-\cref{DrawG,qn,p-centered,p-centered-treewidth,PowerPartition,non-repetitive} immediately imply the following result:
+\cref{qn,p-centered,p-centered-treewidth,PowerPartition,non-repetitive} immediately imply the following result:
 
@@
 
-\subsection{$k$-Nearest-Neighbour Graphs}
-
-In this section, we show that $k$-nearest neighbour graphs of point sets in the plane are $O(k^2)$-planar.
-For two points $x,y\in\R^2$, let $d_2(x,y)$ denote the Euclidean distance between $x$ and $y$.
-The $k$-nearest-neighbour graph of a point set $P\subset\R^2$ is the geometric graph $G$ with vertex set $V(G)=P$, where the edge set is defined as follows.
-For each point $v\in P$, let $N_k(v)$ be the set of $k$ points in $P$ closest to $v$.
-Then $vw\in E(G)$ if and only if $w\in N_k(v)$ or $v\in N_k(w)$.
-% and edge set
-%\[  
-%  E= \{vw : v,w\in V(G),\, |\{x\in P\setminus\{v\}: d_2(v,x)\le d_2(v,w)\}|\le k\} \enspace .
-%\] 
-(The edges of $G$ are straight-line segments joining their endpoints.) See \citep{ProximityGraphs} for a survey of results on $k$-nearest neighbour graphs and other related proximity graphs.
-
-The following result, which is immediate from \citet[Corollary~4.2.6]{abrego.munroy.ea:on} states that $k$-nearest-neighbour graphs have bounded maximum degree:
-\begin{lem}
-\label{k-nn-max-degree}
-The degree of every vertex in a $k$-nearest-neighbour graph is at most $6k$.
-\end{lem}
-
-We make use of the following well-known observation (see for example, \citet[Lemma~2]{bose.morin.ea:routing}):
-\begin{obs}
-\label{convex}
-If $v_0,\ldots,v_3$ are the vertices of a convex quadrilateral in counterclockwise order then there exists at least one $i\in\{0,\ldots,3\}$ such that $\max\{d_2(v_i,v_{i-1}), d_2(v_i,v_{i+1})\} < d_2(v_{i-1},v_{i+1})$, where subscripts are taken modulo 4.
-\end{obs}
-
-\begin{lem}
-\label{nearest-neighbour}
-  Every $k$-nearest-neighbour graph is $O(k^2)$-planar.
-\end{lem}
-
-\begin{proof}
-  Let $G$ be a $k$-nearest-neighbour graph and consider any edge $vw\in E(G)$.
-  Let $xy\in E(G)$ be an edge that crosses $vw$.  Note that $vxwy$ are the vertices of a convex quadrilateral in (without loss of generality) counterclockwise order. Then we say that 
-  \begin{compactenum}
-    \item $xy$ is of Type~$v$ if $\max\{d_2(v,x), d_2(v,y)\}< d_2(x,y)$;
-    \item $xy$ is of Type~$w$ if $\max\{d_2(w,x), d_2(w,y)\}< d_2(x,y)$; or
-    \item $xy$ is of Type~C otherwise.
-  \end{compactenum}
-  If $xy$ is of Type~C, then \cref{convex} implies that (without loss of generality) 
-  \[\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)\enspace.\] 
-   In this case, we call $x$ a Type~C vertex.
-We claim that $V(G)$ contains at most $k-1$ Type~C vertices.
-Indeed, more than $k-1$ Type~C vertices would contradict the fact that $vw\in E(G)$ since every Type~C vertex is closer to both $v$ and $w$ than $d_2(v,w)$.
-
-  Next oberve that, if $xy$ is of Type~$v$, then at least one of $xv$ or $yv$ is in $E(G)$ in which case we call $x$ (respectively $y$) a Type~$v$ vertex.
-By \cref{k-nn-max-degree}, there are at most $6k$ Type~$v$ vertices.
-Similarly, there are at most $6k$ Type~$w$ vertices.
-
-  Thus, in total, there are at most $13k-1$ Type~$v$, Type~$w$, and Type~C vertices.
-By \cref{k-nn-max-degree}, each of these vertices is incident with at most $6k$ edges that cross $vw$.
-Therefore, there are at most $78k^2-6k$ edges of $G$ that cross $vw$.
-Since this is true for every edge $vw\in E(G)$, $G$ is $(78k^2-6k)$-planar.
-\end{proof}
-
-We note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n$ vertices has at least $kn/2$ edges and at most $kn$ edges.
-For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is $\Omega(k^2)$.
-
-\cref{nearest-neighbour,k-planar,qn,p-centered,p-centered-treewidth} imply:
-
-\begin{cor}
-\label{k-nn}
-Every $k$-nearest-neighbour graph has:
-\begin{compactitem}
-\item an $H$-partition of layered width $O(k^4)$ in which $H$ has treewidth $O(k^6)$,
-\item queue-number at most $2^{O(k^6)}$,
-\item $p$-centered chromatic number at most $O(k^4) (p+1) (t+1)\binom{p+t}{t}$ for some integer $t\leq O(k^6)$.
-\end{compactitem}
-\end{cor}
-
-\cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
-However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (36+o(1))k^2$ by a result of \citet{DJKW16}.
-
-\section{An Open Problem}
-
-\cref{k-planar} shows that $k$-planar graphs have $H$-partitions of layered width $O(k^2)$ in which $H$ has treewidth $O(k^3)$.
-We conclude with the following question:  Does there exist a function $f:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph has an $H$-partition of layered width at most $f(k)$ in which $H$ has treewidth at most $C$?
-Perhaps $C=3$.
-Note that $C\geq 3$ even for planar graphs \citep{dujmovic.joret.ea:planar}.
-
-%%%  Squashing the bibliography 
-  \let\oldthebibliography=\thebibliography
-  \let\endoldthebibliography=\endthebibliography
-  \renewenvironment{thebibliography}[1]{%
-    \begin{oldthebibliography}{#1}%
-      \setlength{\parskip}{0.4ex}%
-      \setlength{\itemsep}{0.4ex}%
-  }{\end{oldthebibliography}}
-
-%\bibliographystyle{plainnat}
-%\bibliography{k-planar}
-
-% NOTE THAT MANUAL CHANGES HAVE BEEN MADE BELOW.
-
-\fontsize{11}{11.75}\selectfont
-
-\begin{thebibliography}{33}
-\providecommand{\natexlab}[1]{#1}
-\providecommand{\url}[1]{\texttt{#1}}
-\expandafter\ifx\csname urlstyle\endcsname\relax
-  \providecommand{\doi}[1]{doi: #1}\else
-  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
-
-\bibitem[{\'{A}}brego et~al.(2011){\'{A}}brego, Monroy,
-  Fern{\'{a}}ndez{-}Merchant, Flores{-}Pe{\~{n}}aloza, Hurtado,
-  Sacrist{\'{a}}n, and Saumell]{abrego.munroy.ea:on}
-Bernardo~M. {\'{A}}brego, Ruy~Fabila Monroy, Silvia Fern{\'{a}}ndez{-}Merchant,
-  David Flores{-}Pe{\~{n}}aloza, Ferran Hurtado, Vera Sacrist{\'{a}}n, and
-  Maria Saumell.
-\newblock On crossing numbers of geometric proximity graphs.
-\newblock \emph{Comput. Geom.}, 44\penalty0 (4):\penalty0 216--233, 2011.
-\newblock \doi{10.1016/j.comgeo.2010.11.003}.
-
-\bibitem[Ajtai et~al.(1982)Ajtai, Chv{\'a}tal, Newborn, and
-  Szemer{\'e}di]{ajtai.chvatal.ea:crossing-free}
-Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, Endre
-  Szemer{\'e}di.
-\newblock Crossing-free subgraphs.
-\newblock \emph{Ann. Discrete Math.}, 12:\penalty0 9--12, 1982.
-\newblock \doi{10.1016/S0304-0208(08)73484-4}.
-
-\bibitem[Alam et~al.(2018)Alam, Bekos, Gronemann, Kaufmann, and
-  Pupyrev]{ABGKP18}
-Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael Kaufmann, Sergey Pupyrev.
-\newblock Queue layouts of planar 3-trees.
-\newblock In Therese~C. Biedl and Andreas Kerren, editors, \emph{Proc. 26th
-  International Symposium on Graph Drawing and Network Visualization} (GD '18), vol.~11282 of \emph{Lecture Notes in Computer Science}, pp.~213--226. Springer, 2018.
-\newblock \doi{10.1007/978-3-030-04414-5\_15}.
-
-\bibitem[Alon et~al.(2002)Alon, Grytczuk, Ha{\l}uszczak, and
-  Riordan]{AGHR-RSA02}
-Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, Oliver Riordan.
-\newblock Nonrepetitive colorings of graphs.
-\newblock \emph{Random Structures Algorithms}, 21\penalty0 (3-4):\penalty0
-  336--346, 2002.
-\newblock \doi{10.1002/rsa.10057}.
-
-\bibitem[Bodlaender(1998)]{Bodlaender-TCS98}
-Hans~L. Bodlaender.
-\newblock A partial $k$-arboretum of graphs with bounded treewidth.
-\newblock \emph{Theoret. Comput. Sci.}, 209\penalty0 (1-2):\penalty0 1--45,
-  1998.
-\newblock \doi{10.1016/S0304-3975(97)00228-4}.
-
-\bibitem[Bonamy et~al.(2019)Bonamy, Gavoille, and
-  Pilipczuk]{bonamy.gavoille.ea:shorter}
-Marthe Bonamy, Cyril Gavoille, and Micha{\l} Pilipczuk.
-\newblock Shorter implicit representation for planar graphs.
-\newblock In S.~Cabello and B.~Mohar, editors, \emph{9th Slovenian
-  International Conference on Graph Theory} (SICGT), p.~63, June 2019.
-
-\bibitem[Bose et~al.(2001)Bose, Morin, Stojmenovi\'c, and
-  Urrutia]{bose.morin.ea:routing}
-Prosenjit Bose, Pat Morin, Ivan Stojmenovi\'c, Jorge Urrutia.
-\newblock Routing with guaranteed delivery in ad hoc wireless networks.
-\newblock \emph{Wireless Networks}, 7\penalty0 (6):\penalty0 609--616, 2001.
-\newblock \doi{10.1023/A:1012319418150}.
-
-\bibitem[Bose et~al.(2012)Bose, Dujmovi\'c, Hurtado, Iacono, Langerman, Meijer,
-  Sacrist\'an, Saumell, and Wood]{ProximityGraphs}
-Prosenjit Bose, Vida Dujmovi\'c, Ferran Hurtado, John Iacono, Stefan Langerman,
-  Henk Meijer, Vera Sacrist\'an, Maria Saumell, David~R. Wood.
-\newblock Proximity graphs: $e$, $\delta$, $\delta$, $\delta$, $\chi$ and
-  $\omega$.
-\newblock \emph{Internat. J. Comput. Geom.  Appl.},
-  22\penalty0 (5):\penalty0 439--469, 2012.
-
-\bibitem[Chen et~al.(2002)Chen, Grigni, and Papadimitriou]{CGP02}
-Zhi-Zhong Chen, Michelangelo Grigni, Christos~H. Papadimitriou.
-\newblock Map graphs.
-\newblock \emph{J. ACM}, 49\penalty0 (2):\penalty0 127--138, 2002.
-\newblock \doi{10.1145/506147.506148}.
-
-\bibitem[D{\k e}bski et~al.(2019)D{\k e}bski, Felsner, Micek, and Schr{\"o}der]{micek:personal}
-  Micha{\l} D{\k e}bski, Stefan Felsner, Piotr Micek, Felix Schr{\" o}der. 
-\newblock Improved bounds for centered colorings, 2019.
-\newblock \href{https://arxiv.org/abs/1907.04586}{arXiv:1907.04586}.
-
-\bibitem[Dujmovi\'{c} et~al.(2016)Dujmovi\'{c}, Joret, Kozik, and Wood]{DJKW16}
-Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, David~R. Wood.
-\newblock Nonrepetitive colouring via entropy compression.
-\newblock \emph{Combinatorica}, 36\penalty0 (6):\penalty0 661--686, 2016.
-\newblock \doi{10.1007/s00493-015-3070-6}.
-
-\bibitem[Dujmovi\'c et~al.(2017{\natexlab{a}})Dujmovi\'c, Eppstein, and
-  Wood]{dujmovic.eppstein.ea:structure}
-Vida Dujmovi\'c, David Eppstein, David~R. Wood.
-\newblock Structure of graphs with locally restricted crossings.
-\newblock \emph{{SIAM} J. Discrete Math.}, 31\penalty0 (2):\penalty0 805--824,
-  2017{\natexlab{a}}.
-\newblock \doi{10.1137/16M1062879}.
-
-\bibitem[Dujmovi\'c et~al.(2017{\natexlab{b}})Dujmovi\'c, Morin, and
-  Wood]{dujmovic.morin.ea:layered}
-Vida Dujmovi\'c, Pat Morin, David~R. Wood.
-\newblock Layered separators in minor-closed graph classes with applications.
-\newblock \emph{J. Comb. Theory, Ser. {B}}, 127:\penalty0 111--147,
-  2017{\natexlab{b}}.
-\newblock \doi{10.1016/j.jctb.2017.05.006}.
-
-\bibitem[Dujmovi\'c et~al.(2019{\natexlab{a}})Dujmovi\'c, Esperet, Joret,
-  Walczak, and Wood]{dujmovic.esperet.ea:planar}
-Vida Dujmovi\'c, Louis Esperet, Gwena{\"{e}}l Joret, Bartosz Walczak,   David~R. Wood.
-\newblock Planar graphs have bounded nonrepetitive chromatic number.
-\newblock  \href{http://arxiv.org/abs/1904.05269}{arXiv:1904.05269}, 2019{\natexlab{a}}.
-
-\bibitem[Dujmovi\'c et~al.(2019{\natexlab{b}})Dujmovi\'c, Joret, Micek, Morin,
-  Ueckerdt, and Wood]{dujmovic.joret.ea:planar}
-Vida Dujmovi\'c, Gwena{\"{e}}l Joret, Piotr Micek, Pat Morin, Torsten Ueckerdt, David~R. Wood.
-\newblock Planar graphs have bounded queue-number.
-\newblock \emph{Proc. 60th Annual Symp. on Foundations of Computer 
-  Science} (FOCS '19), to appear. 
-  \newblock  \href{http://arxiv.org/abs/1904.04791}{arXiv:1904.04791}, 2019{\natexlab{b}}.
-
-\bibitem[Fomin et~al.(2012)Fomin, Lokshtanov, and Saurabh]{FLS-SODA12}
-Fedor~V. Fomin, Daniel Lokshtanov, Saket Saurabh.
-\newblock Bidimensionality and geometric graphs.
-\newblock In \emph{Proc. 23rd {A}nnual {ACM}-{SIAM} {S}ymposium on {D}iscrete
-  {A}lgorithms} (SODA 2012), pp.~1563--1575, 2012.
-\newblock \doi{10.1137/1.9781611973099.124}.
-
-\bibitem[Fox and Pach(2008)]{FP08}
-Jacob Fox, J{\'a}nos Pach.
-\newblock Separator theorems and {T}ur\'an-type results for planar intersection graphs.
-\newblock \emph{Adv. Math.}, 219(3):1070--1080, 2008.
-\newblock \doi{10.1016/j.aim.2008.06.002}.
-
-\bibitem[Fox and Pach(2010)]{FP10}
-Jacob Fox, J{\'a}nos Pach.
-\newblock A separator theorem for string graphs and its applications.
-\newblock \emph{Combin. Probab. Comput.}, 19\penalty0 (3):\penalty0 371--390,
-  2010.
-\newblock \doi{10.1017/S0963548309990459}.
-
-\bibitem[Fox and Pach(2014)]{FP14}
-Jacob Fox, J{\'a}nos Pach.
-\newblock Applications of a new separator theorem for string graphs.
-\newblock \emph{Combin. Probab. Comput.}, 23\penalty0 (1):\penalty0 66--74,
-  2014.
-\newblock \doi{10.1017/S0963548313000412}.
-
-\bibitem[Grigoriev and Bodlaender(2007)]{grigoriev.bodlaender:algorithms}
-Alexander Grigoriev, Hans~L. Bodlaender.
-\newblock Algorithms for graphs embeddable with few crossings per edge.
-\newblock \emph{Algorithmica}, 49\penalty0 (1):\penalty0 1--11, 2007.
-\newblock \doi{10.1007/s00453-007-0010-x}.
-
-\bibitem[Harvey and Wood(2017)]{HW17}
-Daniel~J. Harvey, David~R. Wood.
-\newblock Parameters tied to treewidth.
-\newblock \emph{J. Graph Theory}, 84\penalty0 (4):\penalty0 364--385, 2017.
-\newblock \doi{10.1002/jgt.22030}.
-
-\bibitem[Heath et~al.(1992)Heath, Leighton, and Rosenberg]{HLR92}
-Lenwood~S. Heath, F.~Thomson Leighton, Arnold~L. Rosenberg.
-\newblock Comparing queues and stacks as mechanisms for laying out graphs.
-\newblock \emph{SIAM J. Discrete Math.}, 5\penalty0 (3):\penalty0 398--412,
-  1992.
-\newblock \doi{10.1137/0405031}.
-
-\bibitem[van~den Heuvel and Wood(2017)]{vdHW17}
-Jan van~den Heuvel, David~R. Wood.
-\newblock Improper colourings inspired by {H}adwiger's conjecture.
-\newblock \emph{J. London Math. Soc.}, 98(1):129--148, 2018. 
-\newblock  \doi{10.1112/jlms.12127}. See   \href{https://arxiv.org/abs/1704.06536}{arXiv:1704.06536} for the cited results.
-
-\bibitem[Kannan et~al.(1988)Kannan, Naor, and
-  Rudich]{kannan.naor.ea:implicit-stoc}
-Sampath Kannan, Moni Naor, Steven Rudich.
-\newblock Implicit representation of graphs.
-\newblock In Janos Simon, editor, \emph{Proc.\ 20th Annual {ACM} Symposium on
-  Theory of Computing} (STOC '88), pp.~334--343. {ACM}, 1988.
-\newblock \doi{10.1145/62212.62244}.
-
-\bibitem[Kannan et~al.(1992)Kannan, Naor, and Rudich]{kannan.naor.ea:implicit}
-Sampath Kannan, Moni Naor, Steven Rudich.
-\newblock Implicit representation of graphs.
-\newblock \emph{{SIAM} J. Discrete Math.}, 5\penalty0 (4):\penalty0 596--603,
-  1992.
-\newblock \doi{10.1137/0405049}.
-
-\bibitem[Kobourov et~al.(2017)Kobourov, Liotta, and
-  Montecchiani]{kobourov.liotta.ea:annotated}
-Stephen~G. Kobourov, Giuseppe Liotta, Fabrizio Montecchiani.
-\newblock An annotated bibliography on 1-planarity.
-\newblock \emph{Comput. Sci. Rev.}, 25:\penalty0 49--67, 2017.
-\newblock \doi{10.1016/j.cosrev.2017.06.002}.
-
-\bibitem[Leighton(1983)]{leighton:complexity}
-F.~Thomson Leighton.
-\newblock \emph{Complexity issues in {VLSI}}.
-\newblock MIT Press, Cambridge, 1983.
-
-\bibitem[Mohar and Thomassen(2001)]{mohar.thomassen:graphs}
-Bojan Mohar, Carsten Thomassen.
-\newblock \emph{Graphs on surfaces}.
-\newblock Johns Hopkins Press, 2001.
-
-\bibitem[Ne{\v{s}}et{\v{r}}il and Ossona~de Mendez(2012)]{Sparsity}
-Jaroslav Ne{\v{s}}et{\v{r}}il, Patrice Ossona~de Mendez.
-\newblock \emph{Sparsity}, vol.~28 of \emph{Algorithms and Combinatorics}.
-\newblock Springer, 2012.
-\newblock \doi{10.1007/978-3-642-27875-4}.
-
-\bibitem[Pach and T{\'o}th(1997)]{PachToth97}
-J{\'a}nos Pach, G{\'e}za T{\'o}th.
-\newblock Graphs drawn with few crossings per edge. 
-\newblock \emph{Combinatorica}, 17(3):427--439, 1997
-  2002.
-\newblock \doi{10.1007/BF01215922}.
-
-\bibitem[Pach and T{\'o}th(2002)]{PachToth-DCG02}
-J{\'a}nos Pach, G{\'e}za T{\'o}th.
-\newblock Recognizing string graphs is decidable.
-\newblock \emph{Discrete Comput. Geom.}, 28\penalty0 (4):\penalty0 593--606,
-  2002.
-\newblock \doi{10.1007/s00454-002-2891-4}.
-
-\bibitem[Pilipczuk and Siebertz(2019)]{pilipczuk.siebertz:polynomial-soda}
-Micha{\l} Pilipczuk, Sebastian Siebertz.
-\newblock Polynomial bounds for centered colorings on proper minor-closed graph
-  classes.
-\newblock In Timothy~M. Chan, editor, \emph{Proc.\ 30th Annual {ACM-SIAM}
-  Symposium on Discrete Algorithms} ({SODA} 2019), pp.~1501--1520. {SIAM},
-  2019.
-\newblock \doi{10.1137/1.9781611975482.91}.
-
-\bibitem[Reed(2003)]{Reed03}
-Bruce~A. Reed.
-\newblock Algorithmic aspects of tree width.
-\newblock In \emph{Recent Advances in Algorithms and Combinat.}, vol.~11,
-  pp.~85--107. Springer, 2003, 
-\newblock \doi{10.1007/0-387-22444-0_4}.
-
-\bibitem[Wiechert(2017)]{wiechert:on}
-Veit Wiechert.
-\newblock On the queue-number of graphs with bounded tree-width.
-\newblock \emph{Electr. J. Comb.}, 24\penalty0 (1):\penalty0 P1.65, 2017.
-\newblock  \url{http://www.combinatorics.org/v24i1p65}.
-
-\bibitem[Zhu(2009)]{zhu:colouring}
-Xuding Zhu.
-\newblock Colouring graphs with bounded generalized colouring number.
-\newblock \emph{Discrete Math.}, 309\penalty0 (18):\penalty0 5562--5568,
-  2009.
-\newblock \doi{10.1016/j.disc.2008.03.024}.
-
-\end{thebibliography}
-
 \end{document}

### 2019-12-05 k-planar.tex
This is what I'm submitting

@@
 \begin{document}
-
+\nolinenumbers
 \maketitle
@@
 %\setcounter{\page}{1}
+\linenumbers
 \section{Introduction}
@@
 Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$.
-Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y]$.
+Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y\}]$.
 If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.
@@
 Let $z_0,\ldots,z_r$ be the path in $T$ from $z_0:=x'$ to $z_r:=a$.
-For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $B_{z_i}$.
+For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $B_{z_i}$ since $x'a\in E(H)$.
 However, by \tref{subtree-root} $x'$ is not contained in $B_{x_i}$ for any $i\in\{1,\ldots,r\}$.
@@
 For a node $x\in V(T)$, we say that a shortcut $P\in\mathcal{P}$ \emph{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
-We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$ or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
+We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$, or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
 Let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
@@
 \begin{clm}\label{x-v-ancestor}
-  For any $v\in V(G)$ there exists a (unique) node $a(v)\in X_v$ such that
+  For each $v\in V(G)$ there exists a (unique) node $a(v)\in X_v$ such that
   $a(v)$ is a $T$-ancestor of every node in $X_v$.
@@
 
-For each $x\in V(T)$, we define the \emph{separator} $S_x := \{v\in V(G): a(v)= x\}$.
+For each $x\in V(T)$, we define $S_x := \{v\in V(G): a(v)= x\}$.
 Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
@@
 Therefore, $\mathcal{P}$ contains a path $P$, with $v\in V(P)$, that crosses $x$.
-The path $P$ contains a subpath $v_0,v_1,\ldots,v_{r}$ such that $v=v_0$ and $v_r\in Y_x$.
+The path $P$ contains a proper subpath $v_0,v_1,\ldots,v_{r}$ such that $v=v_0$ and $v_r\in Y_x$.
 Since $v\not\in V_x$ and $v_r\in Y_x\subseteq V_x$, \yref{separator}, implies that $v_i\in F_x$ for some $i\in\{0,\ldots,r-1\}$.
@@
 
-Next we show that $\mathcal{S}$ has small layered width with respect to $\mathcal{L}$:
+Next we show that $\mathcal{S}$ has small layered width with respect to $\mathcal{L}=(V_1,\ldots,V_h)$:
 
@@
 Since $xy\in E(J)$, $G^\mathcal{P}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
-Since $vw\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.
-By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.
+Since $vw\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.\footnote{Recall that we assume $\mathcal{P}$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.}  By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.
 By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
 By \yref{ancestor-edge} $N_x$ and $V_y$ are also disjoint.
-By \yref{separator} $P$ contains an internal vertex $v'\in F_x$.
+By \yref{separator}, $P$ contains an internal vertex $v'\in F_x$.
 By \yref{ancestor-edge}, $v'\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
@@
   For the tree-decomposition $(C_x:x\in V(T))$ of $J$ we use the same tree $T$ used in the tree-decomposition $(B_x:x\in V(T))$ of $H$.
-For each node $x$ of $T$, $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
-  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, $a,x'\in C_{x'}$ and the connectivity of $T[a]:=T[\{x\in V(T):a\in C_x]$ is obvious.
+For each node $x$ of $T$, we define $C_x$ as follows: $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
+  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, $a,x'\in C_{x'}$ and the connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ is obvious.
 Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $J$.
@@
   \begin{enumerate}
-    \item $\mathcal{P}$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
+    \item there exists $w'\in V(G)$ such that $\mathcal{P}$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
     \item $w\in Y_\delta$.  In this case, we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$. 
@@
   Let $w_0,w_1,\ldots,w_p$ be the path that begins $w_0:=v'$ and then follows the subpath of $P_{ww'}$ that begins at $w_1:=w$ and ends at $w_p:=w''$.
-For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{1,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.
+For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{0,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.
 Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root, $x_0$, of $T$ is non-increasing.
@@
 Therefore $s_p\ge\delta$.
-To see that $s_p<\delta-1$,
+To see that $s_p<\delta+1$,
   observe that either $w=w''\in Y_\delta$ or $P_{ww'}$ contains an internal vertex $w''$ in $Y_\delta$.
-By \yref{thickness} and the defnition of $V_x$, $V_{\delta-1}$ does not contain $w''$, so $s_p<\delta-1$.
+By \yref{thickness} and the defnition of $V_x$, $V_{\delta+1}$ does not contain $w''$, so $s_p<\delta+1$.
 
@@
 Finally, we are in a position to appeal to \cite[Lemma~24]{pilipczuk.siebertz:polynomial-soda} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
+Therefore, the number of $T$-ancestors $a$ of $x$ contained in $C_x$ is at most $\binom{k+t}{t}$, so $|C_x|\le \binom{k+t}{t}$.
 \end{proof}
@@
 %%%  Squashing the bibliography 
-  \let\oldthebibliography=\thebibliography
-  \let\endoldthebibliography=\endthebibliography
-  \renewenvironment{thebibliography}[1]{%
-    \begin{oldthebibliography}{#1}%
-      \setlength{\parskip}{0.4ex}%
-      \setlength{\itemsep}{0.4ex}%
-  }{\end{oldthebibliography}}
-
+  % \let\oldthebibliography=\thebibliography
+  % \let\endoldthebibliography=\endthebibliography
+  % \renewenvironment{thebibliography}[1]{%
+  %   \begin{oldthebibliography}{#1}%
+  %     \setlength{\parskip}{0.4ex}%
+  %     \setlength{\itemsep}{0.4ex}%
+  % }{\end{oldthebibliography}}
+\nolinenumbers
 \bibliographystyle{plainnat}
@@
 
+\linenumbers
 \section{\boldmath  $1$-Planar Graphs: proof of \cref{1-planar}}

### 2019-12-05 k-planar.tex
Using bibtex again

--- 2019-12-05 k-planar.tex
@@
- \documentclass[a4paper,UKenglish,cleveref,autoref]{lipics-v2019}
-
-% \documentclass{socg-lipics-v2019}
-
+\documentclass{patmorin}
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
-\usepackage{amsmath}
+\usepackage{amsmath,wrapfig}
 \usepackage{amsfonts}
@@
 \usepackage{paralist}
-% \usepackage{stmaryrd}
+\usepackage{stmaryrd}
 \usepackage{thm-restate}
@@
 
-% \usepackage[mathlines]{lineno}
-% % \setlength{\linenumbersep}{2em}
+\usepackage[mathlines]{lineno}
+% \setlength{\linenumbersep}{2em}
+\linenumbers
+% \rightlinenumbers
 % \linenumbers
-% % \rightlinenumbers
-% % \linenumbers
-% \newcommand*\patchAmsMathEnvironmentForLineno[1]{%
-%  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
-%  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
-%  \renewenvironment{#1}%
-%     {\linenomath\csname old#1\endcsname}%
-%     {\csname oldend#1\endcsname\endlinenomath}}% 
-% \newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
-%  \patchAmsMathEnvironmentForLineno{#1}%
-%  \patchAmsMathEnvironmentForLineno{#1*}}%
-% \AtBeginDocument{%
-% \patchBothAmsMathEnvironmentsForLineno{equation}%
-% \patchBothAmsMathEnvironmentsForLineno{align}%
-% \patchBothAmsMathEnvironmentsForLineno{flalign}%
-% \patchBothAmsMathEnvironmentsForLineno{alignat}%
-% \patchBothAmsMathEnvironmentsForLineno{gather}%
-% \patchBothAmsMathEnvironmentsForLineno{multline}%
-% }
-
-% \allowdisplaybreaks
-% %\sloppy
-% %%%%%%%%%%
-% \makeatletter
-% \def\NAT@spacechar{~}
-% \makeatother
+\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+ \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+ \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+ \renewenvironment{#1}%
+    {\linenomath\csname old#1\endcsname}%
+    {\csname oldend#1\endcsname\endlinenomath}}% 
+\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+ \patchAmsMathEnvironmentForLineno{#1}%
+ \patchAmsMathEnvironmentForLineno{#1*}}%
+\AtBeginDocument{%
+\patchBothAmsMathEnvironmentsForLineno{equation}%
+\patchBothAmsMathEnvironmentsForLineno{align}%
+\patchBothAmsMathEnvironmentsForLineno{flalign}%
+\patchBothAmsMathEnvironmentsForLineno{alignat}%
+\patchBothAmsMathEnvironmentsForLineno{gather}%
+\patchBothAmsMathEnvironmentsForLineno{multline}%
+}
+
+\allowdisplaybreaks
+%\sloppy
+%%%%%%%%%%
+\makeatletter
+\def\NAT@spacechar{~}
+\makeatother
 %%%%%%%%%%
@@
 
- % \newcommand{\note}[2]{{\color{red}[#1:~#2]}}
+ \newcommand{\note}[2]{{\color{red}[#1:~#2]}}
 % \newcommand{\notex}[2]{{\color{blue}[#1:~#2]}}
-% \newcommand{\notex}[2]{}
+\newcommand{\notex}[2]{}
 
@@
 
-% \title{\MakeUppercase{The Structure of $k$-Planar Graphs}}
-\title{The Structure of $k$-Planar Graphs}
-\titlerunning{The Structure of $k$-Planar Graphs}
-
-\author{Vida Dujmović}{School of Computer Science and Electrical Engineering, University of Ottawa}{vida.dujmovic@uottawa.ca}{}{Research  supported by NSERC.}
-% , Ottawa, Canada (\texttt{vida.dujmovic@uottawa.ca}).
-%                 Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\,
-\author{Pat Morin}{School of Computer Science, Carleton University}{morin@scs.carleton.ca}{[0000-0003-0471-4118]}{Research supported by NSERC.}
-% , Ottawa, Canada (\texttt{morin@scs.carleton.ca}).                 Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\, and
-\author{David R. Wood}{School of Mathematics, Monash University}{david.wood@monash.edu}{}{Research supported by the Australian Research Council.}
-% , Melbourne, Australia (\texttt{david.wood@monash.edu}). Research supported by the Australian Research Council.}
-% }
-
-\authorrunning{V. Dujmović, P. Morin, and D. R. Wood}
-
-\keywords{$k$-planar graphs, layered partitions, queue-number, nonrepetitive colouring, centered colouring, $k$-nearest-neighbour graphs}
-
-\Copyright{Vida Dujmović, Pat Morin, and David R. Wood}
-\ccsdesc{Mathematics of computing~Graph theory}
-\ccsdesc{Mathematics of computing~Extremal graph theory}
-\ccsdesc{Human-centered computing~Graph drawings}
-\relatedversion{This paper has been submitted to Discrete \&
-Computational Geometry. The current status, as of December 4th, 2019, is ``under review''.  A full version of this paper is available at
-                 \url{https://arxiv.org/abs/1907.05168}.}
-\funding{This work was partly funded by NSERC and MRI.}
+\title{\MakeUppercase{The Structure of $k$-Planar Graphs}}
+
+\author{Vida Dujmovi\'c%
+        \thanks{School of Computer Science and Electrical Engineering,
+                University of Ottawa, Ottawa, Canada (\texttt{vida.dujmovic@uottawa.ca}).
+                Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\,
+        Pat Morin%
+        \thanks{School of Computer Science, Carleton University, Ottawa, Canada (\texttt{morin@scs.carleton.ca}).
+Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\, and
+        David R. Wood\thanks{School of Mathematics, Monash University, Melbourne, Australia (\texttt{david.wood@monash.edu}). Research supported by the Australian Research Council.}
+}
 
 \begin{document}
-
+\begin{titlepage}
 \maketitle
+
 \begin{abstract}
@@
 \end{abstract}
-% \end{titlepage}
+\end{titlepage}
 
@@
 The present paper studies the structure of $k$-planar graphs and other more general classes of graphs.
-% \notex{PM}{I removed the following sentences because neither Vida nor could find anything about $k$-planar graphs in \cite{FP08} and, in any case, the proof is easy.
+\notex{PM}{I removed the following sentences because neither Vida nor could find anything about $k$-planar graphs in \cite{FP08} and, in any case, the proof is easy.
 This could easily upset a referee.
-% % \note{PM}{I must be missing something.
+% \note{PM}{I must be missing something.
 The planarized graph $G'$ has $O(kn)$ vertices.
@@
 }
-% Various structural results about planar graphs generalise for $k$-planar graphs.
+Various structural results about planar graphs generalise for $k$-planar graphs.
 For example, \citet{FP08} generalised the Lipton-Tarjan separator theorem to show that every $k$-planar graph with $n$ vertices has a balanced separator of order $O(\sqrt{kn})$.}
@@
 A path is \emph{trivial} if it has length 0 and \emph{non-trivial} otherwise.} and
-\item for every $v\in V(G)$, the number of paths in $\mathcal{P}$ with $v$ as an internal vertex is at most $d$.
+\item for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
 \end{compactitem} 
@@
 
-We prove the following theorem that shows if a graph $G$ is a subgraph of $H\boxtimes P$ and $\PP$ is a shortcut system for $G$, then $G^{\mathcal{P}}$ is a subgraph of $J\boxtimes P$, where the treewidth of $J$ is bounded by a function of $k$, $d$, and the treewidth of $H$.
+We prove the following theorem that shows if a graph $G$ is a subgraph of $H\boxtimes P$ and $\PP$ is a shortcut system for $G$, then $G^{\mathcal{P}}$ is a subgraph of $J\boxtimes P$, where the treewidth of $J$ is bounded by a function of the treewidth of $H$.
 
@@
 
-\cref{ShortcutProduct} leads to several other results.
+\cref{ShortcutProduct} leads to several other results of interest.
 Here is one example.
@@
 
-This section defines concepts and results from the literature that are used in our work.
+This section defines concepts and results from the literature that will be important for our work.
 
@@
 They defined a class $\mathcal{G}$ of graphs to \emph{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
+
 These definitions relate to strong products as follows.
@@
 \begin{compactenum}[(a)]
-\item an $H$-partition with layered width at most $\max\{2g,1\}$ where $H$ has treewidth at most $9$;
-\item an $H$-partition with layered width at most $\max\{2g,3\}$ where $H$ has treewidth at most $4$.
+\item an $H$-partition with layered width at most $\max\{2g,1\}$ such that $H$ has treewidth at most $9$;
+\item an $H$-partition with layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most $4$.
 %\item an $H$-partition with layered width at most $1$ such that $H$ has treewidth at most $2g+8$. 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Shortcut Systems: Proof of \cref{modern-major-general}}
+\section{Shortcut Systems}
 \label{Structure}
@@
 
-For a tree $T$ rooted at some node $x_0\in V(T)$, we say that a node $a\in V(T)$ is a \emph{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \emph{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
-Note that each node $x\in V(T)$ is a $T$-ancestor and $T$-descendant of itself.
+For a tree $T$ rooted at some node $x_0\in V(T)$, we we say that a node $a\in V(T)$ is a \emph{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \emph{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
+Note that each node $x\in v(T)$ is a $T$-ancestor and $T$-descendant of itself.
 We say that a $T$-ancestor $a\in V(T)$ of $x\in V(T)$ is a \emph{strict} $T$-ancestor of $x$ if $a\neq x$.
-The \emph{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.
-For each node $x\in V(T)$, define $T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}]$ to be the maximal subtree of $T$ rooted at $x$.
-
-We begin with a fairly standard trick of working with a normalised tree decomposition.
+The \emph{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.  For each node $x\in V(T)$, define
+\[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
+to be the maximal subtree of $T$ rooted at $x$.
+
+We begin with a fairly standard lemma about normalised tree decompositions.
 
 \begin{lem}\lemlabel{nice-decomposition}
-  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a $T$-decomposition $(B_x:x\in V(T))$ of $H$ of width at most $t$ that has following additional properties:  
+  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a $T$-decomposition $(B_x:x\in V(T))$ of $H$ with width $t$ that has following additional properties:  
   \begin{compactenum}[(T1)]
@@
 
-% \notex{DW}{Is some sort of converse to \cref{generalized-tripod}  also true?
+\notex{DW}{Is some sort of converse to \cref{generalized-tripod}  also true?
 \cref{generalized-tripod}  shows that if a graph $G$ has a layered partition with all the nice properties, then $G$ has this generalised tripod partition.
@@
 The proof is basically in \citep{dujmovic.joret.ea:planar}, right?}
-% \notex{PM}{Yes, if by generalized tripod partition you mean $T$ and $\mathcal{Y}=(Y_x:x\in V(T))$ that satisify \yref{thickness}, \yref{separator}, and \yref{ancestor-edge}.
+\notex{PM}{Yes, if by generalized tripod partition you mean $T$ and $\mathcal{Y}=(Y_x:x\in V(T))$ that satisify \yref{thickness}, \yref{separator}, and \yref{ancestor-edge}.
 From \yref{separator} and \yref{ancestor-edge} you can construct a width-$t$ $T$-decomposition of $H=G/\mathcal{Y}$.
@@
 
-We first point out more properties that are immediately implied by \cref{generalized-tripod}:
+Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:
 
@@
 \begin{proof}[Proof of \cref{generalized-tripod}]
-  Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$ (a condition ensured by \lemref{nice-decomposition}) and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
+  Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$
+  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
 Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.
@@
 \begin{proof}
-Apply \cref{generalized-tripod} to $G$ and let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $Y_x$, $V_x$, $F_x$, and $N_x$ be defined as in \cref{generalized-tripod} (so that $(\mathcal{L},\mathcal{Y})$ is an $H$-decomposition of $G$ of layered width $\ell$ where $H$ has treewidth $t$.)
+Apply \cref{generalized-tripod} to $G$ and let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $Y_x$, $V_x$, $F_x$, and $N_x$ be defined as in \cref{generalized-tripod}.
 
@@
 We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$ or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
-Let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
+We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
 
@@
 \begin{proof}
-  Let  $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\mathcal{P}$, $v\in \{v_0,\ldots,v_r\}\}$
+  Let 
+  \[
+     Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:\mbox{$(v_0,\ldots,v_r)\in\mathcal{P}$, $v\in \{v_0,\ldots,v_r\}$}\}
+  \]
   Clearly $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
@@
 
-  Let $X_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.  The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.  
+  Let $Z_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.  The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.  
   %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other. 
@@
 
-For each $x\in V(T)$, we define the \emph{separator} $S_x := \{v\in V(G): a(v)= x\}$.
+For each $x\in V(T)$, we define the \emph{separator}
+\[  
+     S_x := \{v\in V(G): a(v)= x\} \enspace .
+\]
 Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
-Let $J:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph and let $V(J)\subseteq V(T)$ in the obvious way, so that each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
+We let $J:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph and we let $V(J)\subseteq V(T)$ in the obvious way, so that each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
 (Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $J$.)
@@
 %%%%%%%%%%%%%%%%%%%
-\section{\boldmath $k$-Planar Graphs: Proof of \cref{k-planar}}
+\section{\boldmath $k$-Planar Graphs}
 \label{sec-k-planar}
@@
 More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap V_j$ for $j > k+1$.
-Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields $ |S_x\cap Y_i| \le d\ell\left( 
+Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields 
+\begin{align*}
+    |S_x\cap Y_i| \le d\ell\left( 
        2 
@@
        \right)
-        =
+       & =
        d\ell\left(
@@
        \right) \\
-        \le
+       & \le
        d\ell(3k+5) 
-      = 18k+30$.
-
+      = 18k+30 \enspace .
+\end{align*}    
 With this change, the layered width of the partition given by \cref{modern-major-general} becomes $(18k+30)(k+1)=18k^2+48k+30$.  This establishes \cref{k-planar}.
 % Note to self-and-others. In our original SODA submission we had 18k^2+30k here because we forgot that the layer compression step has compress k+1 consecutive layers, not just k consecutive layers.
+
 Exactly the same approach using \cref{SurfacePartition}(b) instead of \cref{PlanarPartition}(b) establishes 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Applications}
-\label{Applications}
-Our main theorems have a number of applications including queue layouts, non-repetitive colourings and clustered colourings  of $k$-planar and $(g,k)$-planar graphs.
-We present here queue layout results and defer non-repetitive colourings and clustered colourings to \cref{applmore}.
-
-%%%%%%%%%%%%%%%%%%%
-\subsection{Queue Layouts}
-
-For an integer $k\geq 0$, a \textit{$k$-queue layout} of a graph $G$ consists of a linear ordering $\preceq$ of $V(G)$ and a partition $\{E_1,E_2,\dots,E_k\}$ of $E(G)$, such that for $i\in\{1,2,\dots,k\}$, no two edges in $E_i$ are nested with respect to $\preceq$.
-That is, it is not the case that $v\prec x \prec y \prec w$ for edges $vw,xy\in E_i$.
-The \textit{queue-number} of a graph $G$, denoted by $\qn(G)$, is the minimum integer $k$ such that $G$ has a $k$-queue layout.
-Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
-\citet{dujmovic.joret.ea:planar} recently proved this conjecture using \cref{PlanarPartition} and the following lemma.
-Indeed, resolving this question was the motivation for the development of \cref{PlanarBasic,PlanarPartition}.
-
-\begin{lem}[\citep{dujmovic.joret.ea:planar}]
-\label{qn}
-If a graph $G$ has an $H$-partition of layered width $\ell$, then  $\qn(G)\le 3 \ell \, \qn(H) + \floor{\tfrac{3}{2}\ell}$.
-\end{lem}
-
-\citet{dujmovic.morin.ea:layered} proved that queue-number is bounded by a function of treewidth. 
-The best known bound is due to \citet{wiechert:on}:
-
-\begin{thm}[\citep{wiechert:on}]
-\label{qn-tw}
-    For every graph $G$, $\qn(G)\le 2^{\tw(G)}-1$.
-\end{thm}
-
-\cref{qn,qn-tw,g-k-planar} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^4)}$.
-Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of  $O(g^{k+2})$ using \cref{SurfacePartition} and an ad-hoc method.
-Our result provides a better bound when $g>2^{k^3}$.
-In the case of 1-planar graphs we can improve further.
-\citet{ABGKP18} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
-Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$.
-\cref{qn,1-planar} then imply:
-
-\begin{prop}
-Every 1-planar graph has queue-number at most $3 \times 30 \times 5 + \floor{\tfrac{3}{2} \times 30} = 495$. \end{prop}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Graph Classes: Examples}
-\label{Examples}
-There are several examples of graph classes that can be obtained from a shortcut system applied to graphs of bounded Euler genus.
-These include map graphs, string graphs, powers of bounded degree graphs and $k$-nearest-neighbour graphs.
-We present here  $k$-nearest-neighbour graphs results and defer the others to \cref{exmore}.
-
-\subsection{$k$-Nearest-Neighbour Graphs}
-
-In this section, we show that $k$-nearest neighbour graphs of point sets in the plane are $O(k^2)$-planar.
-For two points $x,y\in\R^2$, let $d_2(x,y)$ denote the Euclidean distance between $x$ and $y$.
-The $k$-nearest-neighbour graph of a point set $P\subset\R^2$ is the geometric graph $G$ with vertex set $V(G)=P$, where the edge set is defined as follows.
-For each point $v\in P$, let $N_k(v)$ be the set of $k$ points in $P$ closest to $v$.
-Then $vw\in E(G)$ if and only if $w\in N_k(v)$ or $v\in N_k(w)$.
-% and edge set
-%\[  
-%  E= \{vw : v,w\in V(G),\, |\{x\in P\setminus\{v\}: d_2(v,x)\le d_2(v,w)\}|\le k\} \enspace .
-%\] 
-(The edges of $G$ are straight-line segments joining their endpoints.) See \citep{ProximityGraphs} for a survey of results on $k$-nearest neighbour graphs and other related proximity graphs.
-
-The following result, which is immediate from \citet[Corollary~4.2.6]{abrego.munroy.ea:on} states that $k$-nearest-neighbour graphs have bounded maximum degree:
-\begin{lem}
-\label{k-nn-max-degree}
-The degree of every vertex in a $k$-nearest-neighbour graph is at most $6k$.
-\end{lem}
-
-We make use of the following well-known observation (see for example, \citet[Lemma~2]{bose.morin.ea:routing}):
-\begin{obs}
-\label{convex}
-If $v_0,\ldots,v_3$ are the vertices of a convex quadrilateral in counterclockwise order then there exists at least one $i\in\{0,\ldots,3\}$ such that $\max\{d_2(v_i,v_{i-1}), d_2(v_i,v_{i+1})\} < d_2(v_{i-1},v_{i+1})$, where subscripts are taken modulo 4.
-\end{obs}
-
-\begin{lem}
-\label{nearest-neighbour}
-  Every $k$-nearest-neighbour graph is $O(k^2)$-planar.
-\end{lem}
-
-\begin{proof}
-  Let $G$ be a $k$-nearest-neighbour graph and consider any edge $vw\in E(G)$.
-  Let $xy\in E(G)$ be an edge that crosses $vw$.  Note that $vxwy$ are the vertices of a convex quadrilateral in (without loss of generality) counterclockwise order. Then we say that 
-  \begin{compactenum}
-    \item $xy$ is of Type~$v$ if $\max\{d_2(v,x), d_2(v,y)\}< d_2(x,y)$;
-    \item $xy$ is of Type~$w$ if $\max\{d_2(w,x), d_2(w,y)\}< d_2(x,y)$; or
-    \item $xy$ is of Type~C otherwise.
-  \end{compactenum}
-  If $xy$ is of Type~C, then \cref{convex} implies that (without loss of generality) 
-  \[\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)\enspace.\] 
-   In this case, we call $x$ a Type~C vertex.
-We claim that $V(G)$ contains at most $k-1$ Type~C vertices.
-Indeed, more than $k-1$ Type~C vertices would contradict the fact that $vw\in E(G)$ since every Type~C vertex is closer to both $v$ and $w$ than $d_2(v,w)$.
-
-  Next oberve that, if $xy$ is of Type~$v$, then at least one of $xv$ or $yv$ is in $E(G)$ in which case we call $x$ (respectively $y$) a Type~$v$ vertex.
-By \cref{k-nn-max-degree}, there are at most $6k$ Type~$v$ vertices.
-Similarly, there are at most $6k$ Type~$w$ vertices.
-
-  Thus, in total, there are at most $13k-1$ Type~$v$, Type~$w$, and Type~C vertices.
-By \cref{k-nn-max-degree}, each of these vertices is incident with at most $6k$ edges that cross $vw$.
-Therefore, there are at most $78k^2-6k$ edges of $G$ that cross $vw$.
-Since this is true for every edge $vw\in E(G)$, $G$ is $(78k^2-6k)$-planar.
-\end{proof}
-
-We note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n$ vertices has at least $kn/2$ edges and at most $kn$ edges.
-For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is $\Omega(k^2)$.
-
-\cref{nearest-neighbour,k-planar,qn,p-centered,p-centered-treewidth} imply:
-
-\begin{cor}
-\label{k-nn}
-Every $k$-nearest-neighbour graph has:
-\begin{compactitem}
-\item an $H$-partition of layered width $O(k^4)$ in which $H$ has treewidth $O(k^6)$,
-\item queue-number at most $2^{O(k^6)}$,
-\item $p$-centered chromatic number at most $O(k^4) (p+1) (t+1)\binom{p+t}{t}$ for some integer $t\leq O(k^6)$.
-\end{compactitem}
-\end{cor}
-
-\cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
-However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (36+o(1))k^2$ by a result of \citet{DJKW16}.
-
-%%%  Squashing the bibliography 
-  \let\oldthebibliography=\thebibliography
-  \let\endoldthebibliography=\endthebibliography
-  \renewenvironment{thebibliography}[1]{%
-    \begin{oldthebibliography}{#1}%
-      \setlength{\parskip}{0.4ex}%
-      \setlength{\itemsep}{0.4ex}%
-  }{\end{oldthebibliography}}
-
-\bibliographystyle{plainnat}
-\bibliography{k-planar}
-
-% % NOTE THAT MANUAL CHANGES HAVE BEEN MADE BELOW.
-% 
-% \fontsize{11}{11.75}\selectfont 
-% 
-% \begin{thebibliography}{33}
-% \providecommand{\natexlab}[1]{#1}
-% \providecommand{\url}[1]{\texttt{#1}}
-% \expandafter\ifx\csname urlstyle\endcsname\relax
-%   \providecommand{\doi}[1]{doi: #1}\else
-%   \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
-% 
-% \bibitem[{\'{A}}brego et~al.(2011){\'{A}}brego, Monroy,
-%   Fern{\'{a}}ndez{-}Merchant, Flores{-}Pe{\~{n}}aloza, Hurtado,
-%   Sacrist{\'{a}}n, and Saumell]{abrego.munroy.ea:on}
-% Bernardo~M. {\'{A}}brego, Ruy~Fabila Monroy, Silvia Fern{\'{a}}ndez{-}Merchant,
-%   David Flores{-}Pe{\~{n}}aloza, Ferran Hurtado, Vera Sacrist{\'{a}}n, and
-%   Maria Saumell.
-% \newblock On crossing numbers of geometric proximity graphs.
-% \newblock \emph{Comput. Geom.}, 44\penalty0 (4):\penalty0 216--233, 2011.
-% \newblock \doi{10.1016/j.comgeo.2010.11.003}.
-% 
-% \bibitem[Ajtai et~al.(1982)Ajtai, Chv{\'a}tal, Newborn, and
-%   Szemer{\'e}di]{ajtai.chvatal.ea:crossing-free}
-% Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, Endre
-%   Szemer{\'e}di.
-% \newblock Crossing-free subgraphs.
-% \newblock \emph{Ann. Discrete Math.}, 12:\penalty0 9--12, 1982.
-% \newblock \doi{10.1016/S0304-0208(08)73484-4}.
-% 
-% \bibitem[Alam et~al.(2018)Alam, Bekos, Gronemann, Kaufmann, and
-%   Pupyrev]{ABGKP18}
-% Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael Kaufmann, Sergey Pupyrev.
-% \newblock Queue layouts of planar 3-trees.
-% \newblock In Therese~C. Biedl and Andreas Kerren, editors, \emph{Proc. 26th
-%   International Symposium on Graph Drawing and Network Visualization} (GD '18), vol.~11282 of \emph{Lecture Notes in Computer Science}, pp.~213--226. Springer, 2018.
-% \newblock \doi{10.1007/978-3-030-04414-5\_15}.
-% 
-% \bibitem[Alon et~al.(2002)Alon, Grytczuk, Ha{\l}uszczak, and
-%   Riordan]{AGHR-RSA02}
-% Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, Oliver Riordan.
-% \newblock Nonrepetitive colorings of graphs.
-% \newblock \emph{Random Structures Algorithms}, 21\penalty0 (3-4):\penalty0
-%   336--346, 2002.
-% \newblock \doi{10.1002/rsa.10057}.
-% 
-% \bibitem[Bodlaender(1998)]{Bodlaender-TCS98}
-% Hans~L. Bodlaender.
-% \newblock A partial $k$-arboretum of graphs with bounded treewidth.
-% \newblock \emph{Theoret. Comput. Sci.}, 209\penalty0 (1-2):\penalty0 1--45,
-%   1998.
-% \newblock \doi{10.1016/S0304-3975(97)00228-4}.
-% 
-% \bibitem[Bonamy et~al.(2019)Bonamy, Gavoille, and
-%   Pilipczuk]{bonamy.gavoille.ea:shorter}
-% Marthe Bonamy, Cyril Gavoille, and Micha{\l} Pilipczuk.
-% \newblock Shorter implicit representation for planar graphs.
-% \newblock In S.~Cabello and B.~Mohar, editors, \emph{9th Slovenian
-%   International Conference on Graph Theory} (SICGT), p.~63, June 2019.
-% 
-% \bibitem[Bose et~al.(2001)Bose, Morin, Stojmenovi\'c, and
-%   Urrutia]{bose.morin.ea:routing}
-% Prosenjit Bose, Pat Morin, Ivan Stojmenovi\'c, Jorge Urrutia.
-% \newblock Routing with guaranteed delivery in ad hoc wireless networks.
-% \newblock \emph{Wireless Networks}, 7\penalty0 (6):\penalty0 609--616, 2001.
-% \newblock \doi{10.1023/A:1012319418150}.
-% 
-% \bibitem[Bose et~al.(2012)Bose, Dujmovi\'c, Hurtado, Iacono, Langerman, Meijer,
-%   Sacrist\'an, Saumell, and Wood]{ProximityGraphs}
-% Prosenjit Bose, Vida Dujmovi\'c, Ferran Hurtado, John Iacono, Stefan Langerman,
-%   Henk Meijer, Vera Sacrist\'an, Maria Saumell, David~R. Wood.
-% \newblock Proximity graphs: $e$, $\delta$, $\delta$, $\delta$, $\chi$ and
-%   $\omega$.
-% \newblock \emph{Internat. J. Comput. Geom.  Appl.},
-%   22\penalty0 (5):\penalty0 439--469, 2012.
-% 
-% \bibitem[Chen et~al.(2002)Chen, Grigni, and Papadimitriou]{CGP02}
-% Zhi-Zhong Chen, Michelangelo Grigni, Christos~H. Papadimitriou.
-% \newblock Map graphs.
-% \newblock \emph{J. ACM}, 49\penalty0 (2):\penalty0 127--138, 2002.
-% \newblock \doi{10.1145/506147.506148}.
-% 
-% \bibitem[D{\k e}bski et~al.(2019)D{\k e}bski, Felsner, Micek, and Schr{\"o}der]{micek:personal}
-%   Micha{\l} D{\k e}bski, Stefan Felsner, Piotr Micek, Felix Schr{\" o}der. 
-% \newblock Improved bounds for centered colorings, 2019.
-% \newblock \href{https://arxiv.org/abs/1907.04586}{arXiv:1907.04586}.
-% 
-% \bibitem[Dujmovi\'{c} et~al.(2016)Dujmovi\'{c}, Joret, Kozik, and Wood]{DJKW16}
-% Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, David~R. Wood.
-% \newblock Nonrepetitive colouring via entropy compression.
-% \newblock \emph{Combinatorica}, 36\penalty0 (6):\penalty0 661--686, 2016.
-% \newblock \doi{10.1007/s00493-015-3070-6}.
-% 
-% \bibitem[Dujmovi\'c et~al.(2017{\natexlab{a}})Dujmovi\'c, Eppstein, and
-%   Wood]{dujmovic.eppstein.ea:structure}
-% Vida Dujmovi\'c, David Eppstein, David~R. Wood.
-% \newblock Structure of graphs with locally restricted crossings.
-% \newblock \emph{{SIAM} J. Discrete Math.}, 31\penalty0 (2):\penalty0 805--824,
-%   2017{\natexlab{a}}.
-% \newblock \doi{10.1137/16M1062879}.
-% 
-% \bibitem[Dujmovi\'c et~al.(2017{\natexlab{b}})Dujmovi\'c, Morin, and
-%   Wood]{dujmovic.morin.ea:layered}
-% Vida Dujmovi\'c, Pat Morin, David~R. Wood.
-% \newblock Layered separators in minor-closed graph classes with applications.
-% \newblock \emph{J. Comb. Theory, Ser. {B}}, 127:\penalty0 111--147,
-%   2017{\natexlab{b}}.
-% \newblock \doi{10.1016/j.jctb.2017.05.006}.
-% 
-% \bibitem[Dujmovi\'c et~al.(2019{\natexlab{a}})Dujmovi\'c, Esperet, Joret,
-%   Walczak, and Wood]{dujmovic.esperet.ea:planar}
-% Vida Dujmovi\'c, Louis Esperet, Gwena{\"{e}}l Joret, Bartosz Walczak,   David~R. Wood.
-% \newblock Planar graphs have bounded nonrepetitive chromatic number.
-% \newblock  \href{http://arxiv.org/abs/1904.05269}{arXiv:1904.05269}, 2019{\natexlab{a}}.
-% 
-% \bibitem[Dujmovi\'c et~al.(2019{\natexlab{b}})Dujmovi\'c, Joret, Micek, Morin,
-%   Ueckerdt, and Wood]{dujmovic.joret.ea:planar}
-% Vida Dujmovi\'c, Gwena{\"{e}}l Joret, Piotr Micek, Pat Morin, Torsten Ueckerdt, David~R. Wood.
-% \newblock Planar graphs have bounded queue-number.
-% \newblock \emph{Proc. 60th Annual Symp. on Foundations of Computer 
-%   Science} (FOCS '19), to appear. 
-%   \newblock  \href{http://arxiv.org/abs/1904.04791}{arXiv:1904.04791}, 2019{\natexlab{b}}.
-% 
-% \bibitem[Fomin et~al.(2012)Fomin, Lokshtanov, and Saurabh]{FLS-SODA12}
-% Fedor~V. Fomin, Daniel Lokshtanov, Saket Saurabh.
-% \newblock Bidimensionality and geometric graphs.
-% \newblock In \emph{Proc. 23rd {A}nnual {ACM}-{SIAM} {S}ymposium on {D}iscrete
-%   {A}lgorithms} (SODA 2012), pp.~1563--1575, 2012.
-% \newblock \doi{10.1137/1.9781611973099.124}.
-% 
-% \bibitem[Fox and Pach(2008)]{FP08}
-% Jacob Fox, J{\'a}nos Pach.
-% \newblock Separator theorems and {T}ur\'an-type results for planar intersection graphs.
-% \newblock \emph{Adv. Math.}, 219(3):1070--1080, 2008.
-% \newblock \doi{10.1016/j.aim.2008.06.002}.
-% 
-% \bibitem[Fox and Pach(2010)]{FP10}
-% Jacob Fox, J{\'a}nos Pach.
-% \newblock A separator theorem for string graphs and its applications.
-% \newblock \emph{Combin. Probab. Comput.}, 19\penalty0 (3):\penalty0 371--390,
-%   2010.
-% \newblock \doi{10.1017/S0963548309990459}.
-% 
-% \bibitem[Fox and Pach(2014)]{FP14}
-% Jacob Fox, J{\'a}nos Pach.
-% \newblock Applications of a new separator theorem for string graphs.
-% \newblock \emph{Combin. Probab. Comput.}, 23\penalty0 (1):\penalty0 66--74,
-%   2014.
-% \newblock \doi{10.1017/S0963548313000412}.
-% 
-% \bibitem[Grigoriev and Bodlaender(2007)]{grigoriev.bodlaender:algorithms}
-% Alexander Grigoriev, Hans~L. Bodlaender.
-% \newblock Algorithms for graphs embeddable with few crossings per edge.
-% \newblock \emph{Algorithmica}, 49\penalty0 (1):\penalty0 1--11, 2007.
-% \newblock \doi{10.1007/s00453-007-0010-x}.
-% 
-% \bibitem[Harvey and Wood(2017)]{HW17}
-% Daniel~J. Harvey, David~R. Wood.
-% \newblock Parameters tied to treewidth.
-% \newblock \emph{J. Graph Theory}, 84\penalty0 (4):\penalty0 364--385, 2017.
-% \newblock \doi{10.1002/jgt.22030}.
-% 
-% \bibitem[Heath et~al.(1992)Heath, Leighton, and Rosenberg]{HLR92}
-% Lenwood~S. Heath, F.~Thomson Leighton, Arnold~L. Rosenberg.
-% \newblock Comparing queues and stacks as mechanisms for laying out graphs.
-% \newblock \emph{SIAM J. Discrete Math.}, 5\penalty0 (3):\penalty0 398--412,
-%   1992.
-% \newblock \doi{10.1137/0405031}.
-% 
-% \bibitem[van~den Heuvel and Wood(2017)]{vdHW17}
-% Jan van~den Heuvel, David~R. Wood.
-% \newblock Improper colourings inspired by {H}adwiger's conjecture.
-% \newblock \emph{J. London Math. Soc.}, 98(1):129--148, 2018. 
-% \newblock  \doi{10.1112/jlms.12127}. See   \href{https://arxiv.org/abs/1704.06536}{arXiv:1704.06536} for the cited results. 
-% 
-% \bibitem[Kannan et~al.(1988)Kannan, Naor, and
-%   Rudich]{kannan.naor.ea:implicit-stoc}
-% Sampath Kannan, Moni Naor, Steven Rudich.
-% \newblock Implicit representation of graphs.
-% \newblock In Janos Simon, editor, \emph{Proc.\ 20th Annual {ACM} Symposium on
-%   Theory of Computing} (STOC '88), pp.~334--343. {ACM}, 1988.
-% \newblock \doi{10.1145/62212.62244}.
-% 
-% \bibitem[Kannan et~al.(1992)Kannan, Naor, and Rudich]{kannan.naor.ea:implicit}
-% Sampath Kannan, Moni Naor, Steven Rudich.
-% \newblock Implicit representation of graphs.
-% \newblock \emph{{SIAM} J. Discrete Math.}, 5\penalty0 (4):\penalty0 596--603,
-%   1992.
-% \newblock \doi{10.1137/0405049}.
-% 
-% \bibitem[Kobourov et~al.(2017)Kobourov, Liotta, and
-%   Montecchiani]{kobourov.liotta.ea:annotated}
-% Stephen~G. Kobourov, Giuseppe Liotta, Fabrizio Montecchiani.
-% \newblock An annotated bibliography on 1-planarity.
-% \newblock \emph{Comput. Sci. Rev.}, 25:\penalty0 49--67, 2017.
-% \newblock \doi{10.1016/j.cosrev.2017.06.002}.
-% 
-% \bibitem[Leighton(1983)]{leighton:complexity}
-% F.~Thomson Leighton.
-% \newblock \emph{Complexity issues in {VLSI}}.
-% \newblock MIT Press, Cambridge, 1983.
-% 
-% \bibitem[Mohar and Thomassen(2001)]{mohar.thomassen:graphs}
-% Bojan Mohar, Carsten Thomassen.
-% \newblock \emph{Graphs on surfaces}.
-% \newblock Johns Hopkins Press, 2001.
-% 
-% \bibitem[Ne{\v{s}}et{\v{r}}il and Ossona~de Mendez(2012)]{Sparsity}
-% Jaroslav Ne{\v{s}}et{\v{r}}il, Patrice Ossona~de Mendez.
-% \newblock \emph{Sparsity}, vol.~28 of \emph{Algorithms and Combinatorics}.
-% \newblock Springer, 2012.
-% \newblock \doi{10.1007/978-3-642-27875-4}.
-% 
-% \bibitem[Pach and T{\'o}th(1997)]{PachToth97}
-% J{\'a}nos Pach, G{\'e}za T{\'o}th.
-% \newblock Graphs drawn with few crossings per edge. 
-% \newblock \emph{Combinatorica}, 17(3):427--439, 1997
-%   2002.
-% \newblock \doi{10.1007/BF01215922}.
-% 
-% \bibitem[Pach and T{\'o}th(2002)]{PachToth-DCG02}
-% J{\'a}nos Pach, G{\'e}za T{\'o}th.
-% \newblock Recognizing string graphs is decidable.
-% \newblock \emph{Discrete Comput. Geom.}, 28\penalty0 (4):\penalty0 593--606,
-%   2002.
-% \newblock \doi{10.1007/s00454-002-2891-4}.
-% 
-% \bibitem[Pilipczuk and Siebertz(2019)]{pilipczuk.siebertz:polynomial-soda}
-% Micha{\l} Pilipczuk, Sebastian Siebertz.
-% \newblock Polynomial bounds for centered colorings on proper minor-closed graph
-%   classes.
-% \newblock In Timothy~M. Chan, editor, \emph{Proc.\ 30th Annual {ACM-SIAM}
-%   Symposium on Discrete Algorithms} ({SODA} 2019), pp.~1501--1520. {SIAM},
-%   2019.
-% \newblock \doi{10.1137/1.9781611975482.91}.
-% 
-% \bibitem[Reed(2003)]{Reed03}
-% Bruce~A. Reed.
-% \newblock Algorithmic aspects of tree width.
-% \newblock In \emph{Recent Advances in Algorithms and Combinat.}, vol.~11,
-%   pp.~85--107. Springer, 2003, 
-% \newblock \doi{10.1007/0-387-22444-0_4}.
-% 
-% \bibitem[Wiechert(2017)]{wiechert:on}
-% Veit Wiechert.
-% \newblock On the queue-number of graphs with bounded tree-width.
-% \newblock \emph{Electr. J. Comb.}, 24\penalty0 (1):\penalty0 P1.65, 2017.
-% \newblock  \url{http://www.combinatorics.org/v24i1p65}.
-% 
-% \bibitem[Zhu(2009)]{zhu:colouring}
-% Xuding Zhu.
-% \newblock Colouring graphs with bounded generalized colouring number.
-% \newblock \emph{Discrete Math.}, 309\penalty0 (18):\penalty0 5562--5568,
-%   2009.
-% \newblock \doi{10.1016/j.disc.2008.03.024}.
-% 
-% \end{thebibliography}
-
-\appendix
-
-\section{\boldmath  $1$-Planar Graphs: proof of \cref{1-planar}}
+\subsection{\boldmath  $1$-Planar Graphs}
 \label{sec-1-planar}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Applications}
-\label{applmore}
+\label{Applications}
+
+Here we discuss some of the consequences of our main theorems for $k$-planar and $(g,k)$-planar graphs.
+
+%%%%%%%%%%%%%%%%%%%
+\subsection{Queue Layouts}
+
+For an integer $k\geq 0$, a \textit{$k$-queue layout} of a graph $G$ consists of a linear ordering $\preceq$ of $V(G)$ and a partition $\{E_1,E_2,\dots,E_k\}$ of $E(G)$, such that for $i\in\{1,2,\dots,k\}$, no two edges in $E_i$ are nested with respect to $\preceq$.
+That is, it is not the case that $v\prec x \prec y \prec w$ for edges $vw,xy\in E_i$.
+The \textit{queue-number} of a graph $G$, denoted by $\qn(G)$, is the minimum integer $k$ such that $G$ has a $k$-queue layout.
+Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
+\citet{dujmovic.joret.ea:planar} recently proved this conjecture using \cref{PlanarPartition} and the following lemma.
+Indeed, resolving this question was the motivation for the development of \cref{PlanarBasic,PlanarPartition}.
+
+\begin{lem}[\citep{dujmovic.joret.ea:planar}]
+\label{qn}
+If a graph $G$ has an $H$-partition of layered width $\ell$, then 
+\[ \qn(G)\le 3 \ell \, \qn(H) + \floor{\tfrac{3}{2}\ell}\enspace.\]
+\end{lem}
+
+\citet{dujmovic.morin.ea:layered} proved that queue-number is bounded by a function of treewidth. 
+The best known bound is due to \citet{wiechert:on}:
+
+\begin{thm}[\citep{wiechert:on}]
+\label{qn-tw}
+    For every graph $G$, $\qn(G)\le 2^{\tw(G)}-1$.
+\end{thm}
+
+\cref{qn,qn-tw,g-k-planar} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^4)}$. 
+Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of 
+$O(g^{k+2})$ using \cref{SurfacePartition} and an ad-hoc method.
+Our result provides a better bound when $g>2^{k^3}$.
+In the case of 1-planar graphs we can improve further.
+\citet{ABGKP18} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
+Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$.
+\cref{qn,1-planar} then imply:
+
+\begin{prop}
+Every 1-planar graph has queue-number at most $3 \times 30 \times 5 + \floor{\tfrac{3}{2} \times 30} = 495$. \end{prop}
 
@@
 \begin{cor}
-  For every $(g,k)$-planar graph $G$,  $\pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+5}{4}}$.
-
+  For every $(g,k)$-planar graph $G$, 
+  \[ 
+%    \pi(G)\le \max\{4g(3k+1),12(2k+3)(k+1)\}\times 4^{(k+2)(k+3)(k+4)/6} 
+    \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+5}{4}}. 
+     \]
 %  \label{g-k-planar} Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most  
@@
 
-The following lemma is implicitly due to \citet{micek:personal}. %We include the proof for completeness.
+The following lemma is implicitly due to \citet{micek:personal}. We include the proof for completeness.
 
@@
 
-% \begin{proof}
-%    Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$.
-%    Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.
+\begin{proof}
+   Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$.
+   Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.
 For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in V_i\cap B_x$ such that:
-%    \begin{compactenum}
-%      \item $\alpha(v)$ is unique among $\{\phi(w): w\in V_i\cap B_x\}$, which is possible 
-%      since $|V_i\cap B_x|\le \ell$,
-%      \item $\beta(v)= i\bmod (p+1)$, and     
-%      \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
-%    \end{compactenum}
-%  To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
-
-%  First suppose that there exists $v,w\in V(X)$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$.
+   \begin{compactenum}
+     \item $\alpha(v)$ is unique among $\{\phi(w): w\in V_i\cap B_x\}$, which is possible 
+     since $|V_i\cap B_x|\le \ell$,
+     \item $\beta(v)= i\bmod (p+1)$, and     
+     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
+   \end{compactenum}
+ To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
+
+ First suppose that there exists $v,w\in V(X)$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$.
 Since $G[X]$ is connected, $G[X]$ contains a path from $v$ to $w$.
@@
 
-%  Otherwise, $X\subseteq V_{i},\ldots,V_{i+s}$ for some $s<p$.
+ Otherwise, $X\subseteq V_{i},\ldots,V_{i+s}$ for some $s<p$.
 Let $H':=H[\{x\in V(H):B_x\cap X\neq\emptyset]$.
@@
 Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
-%  For any $v,w\in B_x$ with $v\neq w$, either $v,w\in V_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in V_{i'}$ and $w\in V_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
+ For any $v,w\in B_x$ with $v\neq w$, either $v,w\in V_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in V_{i'}$ and $w\in V_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
 Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.
 Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
-%  \end{proof}
+ \end{proof}
 
@@
 
-\section{Graph Classes: examples}
-\label{exmore}
+%%%%%%%%%
+\section{Rough Characterisation}
+\label{Characterisation}
+
+This section presents a rough characterisation of $(g,k)$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $(g,k)$-planar.
+First note the following easy result (just add a degree-4 vertex at each crossing point of $G$, as in the proof of \cref{AddDummy,gAddDummy}).
+
+\begin{lem}
+\label{gAddDummyDegree}
+For every $(g,k)$-planar graph $G$ there is a graph $G_0$ of Euler
+genus at most $g$, such that $V(G) \subseteq V(G_0)$ and for every
+edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most
+$k+1$, such that every internal vertex in $P$ has degree at most $4$
+in $G_0$.
+\end{lem}
+
+The following converse of \cref{gAddDummyDegree} provides a sufficient condition for a graph to be $(g,k)$-planar that is slightly more restrictive than being obtained from a $(k,d)$-shortcut system of a $(g,k)$-planar graph.
+\notex{DW}{Is this sentence clear enough?}\notex{PM}{Sure.}
+
+\begin{lem} 
+\label{DrawG}
+Fix integers $g\geq 0$ and $k,\Delta\geq 2$. 
+Let $G_0$ be a graph of Euler genus at most $g$. Let $G$ be
+a graph with $V(G) \subseteq V(G_0)$ such that for every edge $vw \in
+E(G)$ there is a $vw$-path $P_{vw}$ in $G_0$ of length at most $k$, such
+that every internal vertex on $P_{vw}$ has degree at most $\Delta$ in
+$G_0$. Then $G$ is $(g, 2k(k+1)\Delta^{k} )$-planar.
+\end{lem}
+
+\begin{proof}
+For a vertex $x$ of $G_0$ with degree at most $\Delta$, and for $i\in\{1,\dots,k-1\}$, say a vertex $v$ is \emph{$i$-close} to $x$ if there is a $vx$-path $P$ in $G_0$ of length at most $i$ such that every internal vertex in $P$  has degree at most $\Delta$ in $G_0$.
+For each edge $vw$ of $G$, say that $vw$ \emph{passes through} each internal vertex on $P_{vw}$. 
+Say $vw$ passes through $x$. Then $v$ is $i$-close to $x$ and $w$ is $j$-close to $x$ for some $i,j\in\{1,\dots,k-1\}$ with $i+j\leq k$. At most $\Delta^{i}$ vertices are $i$-close to $x$. 
+Thus, the number of edges of $G$ that pass through $x$ is at most 
+\[
+\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i \Delta^j 
+= \sum_{i=1}^{k-1} \Delta^i  \sum_{j=1}^{k-i} \Delta^j 
+< \sum_{i=1}^{k-1} \Delta^i  2 \Delta^{k-i}
+= \sum_{i=1}^{k-1} 2\Delta^k 
+< 2k \Delta^k \enspace.
+\]
+Draw each edge $vw$ of $G$ alongside $P_{vw}$ in $G_0$, so that every pair of edges cross at most once. 
+Every edge of $G$ that crosses $vw$ passes through a vertex on $P_{vw}$ (including $v$ and/or $w$ if they too have degree at most $\Delta$).
+Since $P_{vw}$ has at most $k+1$ vertices, and less than $2k\Delta^{k}$ edges of $G$ pass through each vertex on $P_{vw}$, the edge $vw$ is crossed by less than $2k(k+1)\Delta^{k}$ edges in $G$.
+Hence $G$ is $(g, 2k(k+1)\Delta^{k} )$-planar.
+\end{proof}
+
+Together \cref{DrawG,gAddDummyDegree} provide a rough characterisation of $(g,k)$-planar graphs.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Examples}
+\label{Examples}
+
+This section describes several examples of graph classes that can be obtained from a shortcut system applied to graphs of bounded Euler genus.
+
+%This section describes several graph classes that are $k$-planar or $(g,k)$-planar. The first two examples are well known, the second two are new.
+
+%We conclude by pointing out three other graph classes, namely $(g,d)$-map graphs, $(g,\delta)$-string graphs, and $k$-nearest-neighbour graphs that are $(g,k)$-planar for $k=O(d^2)$ and $k=\delta^2$, respectively.
+This implies that Theorems~\ref{k-planar}--\ref{g-k-planar} and Corollaries~\ref{cor:first}--\ref{cor:last} apply to $(g,d)$-map graphs, $(g,\delta)$-string graphs, and $k$-nearest-neighbour graphs.
 
@@
 Thus $G$ is $(g,\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
-% Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar.
+Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar.
 \cref{g-k-planar} then shows that map graphs admit bounded layered partitions, but we get much better bounds by constructing a shortcut system directly.
@@
 Recall that the \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
-If $G$ is planar with maximum degree $\Delta$, then $G^k$ is $2k(k+1)\Delta^{k}$-planar.
+If $G$ is planar with maximum degree $\Delta$, then $G^k$ is $2k(k+1)\Delta^{k}$-planar by \cref{DrawG}.
 Thus we can immediately conclude that bounded powers of planar graphs of bounded degree admit bounded layered partitions.
@@
 The number of vertices at distance $i$ from $v$ is at most $\Delta^i$.
-Thus the number of paths in $\PP$ with $v$ as an internal vertex is at most
+Thus the number of paths in $\PP$ that use $v$ as an internal vertex is at most
 \[\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i\Delta^j 
@@
 
-\cref{qn,p-centered,p-centered-treewidth,PowerPartition,non-repetitive} immediately imply the following result:
+\cref{DrawG,qn,p-centered,p-centered-treewidth,PowerPartition,non-repetitive} immediately imply the following result:
 
@@
 
+\subsection{$k$-Nearest-Neighbour Graphs}
+
+In this section, we show that $k$-nearest neighbour graphs of point sets in the plane are $O(k^2)$-planar.
+For two points $x,y\in\R^2$, let $d_2(x,y)$ denote the Euclidean distance between $x$ and $y$.
+The $k$-nearest-neighbour graph of a point set $P\subset\R^2$ is the geometric graph $G$ with vertex set $V(G)=P$, where the edge set is defined as follows.
+For each point $v\in P$, let $N_k(v)$ be the set of $k$ points in $P$ closest to $v$.
+Then $vw\in E(G)$ if and only if $w\in N_k(v)$ or $v\in N_k(w)$.
+% and edge set
+%\[  
+%  E= \{vw : v,w\in V(G),\, |\{x\in P\setminus\{v\}: d_2(v,x)\le d_2(v,w)\}|\le k\} \enspace .
+%\] 
+(The edges of $G$ are straight-line segments joining their endpoints.) See \citep{ProximityGraphs} for a survey of results on $k$-nearest neighbour graphs and other related proximity graphs.
+
+The following result, which is immediate from \citet[Corollary~4.2.6]{abrego.munroy.ea:on} states that $k$-nearest-neighbour graphs have bounded maximum degree:
+\begin{lem}
+\label{k-nn-max-degree}
+The degree of every vertex in a $k$-nearest-neighbour graph is at most $6k$.
+\end{lem}
+
+We make use of the following well-known observation (see for example, \citet[Lemma~2]{bose.morin.ea:routing}):
+\begin{obs}
+\label{convex}
+If $v_0,\ldots,v_3$ are the vertices of a convex quadrilateral in counterclockwise order then there exists at least one $i\in\{0,\ldots,3\}$ such that $\max\{d_2(v_i,v_{i-1}), d_2(v_i,v_{i+1})\} < d_2(v_{i-1},v_{i+1})$, where subscripts are taken modulo 4.
+\end{obs}
+
+\begin{lem}
+\label{nearest-neighbour}
+  Every $k$-nearest-neighbour graph is $O(k^2)$-planar.
+\end{lem}
+
+\begin{proof}
+  Let $G$ be a $k$-nearest-neighbour graph and consider any edge $vw\in E(G)$.
+  Let $xy\in E(G)$ be an edge that crosses $vw$.  Note that $vxwy$ are the vertices of a convex quadrilateral in (without loss of generality) counterclockwise order. Then we say that 
+  \begin{compactenum}
+    \item $xy$ is of Type~$v$ if $\max\{d_2(v,x), d_2(v,y)\}< d_2(x,y)$;
+    \item $xy$ is of Type~$w$ if $\max\{d_2(w,x), d_2(w,y)\}< d_2(x,y)$; or
+    \item $xy$ is of Type~C otherwise.
+  \end{compactenum}
+  If $xy$ is of Type~C, then \cref{convex} implies that (without loss of generality) 
+  \[\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)\enspace.\] 
+   In this case, we call $x$ a Type~C vertex.
+We claim that $V(G)$ contains at most $k-1$ Type~C vertices.
+Indeed, more than $k-1$ Type~C vertices would contradict the fact that $vw\in E(G)$ since every Type~C vertex is closer to both $v$ and $w$ than $d_2(v,w)$.
+
+  Next oberve that, if $xy$ is of Type~$v$, then at least one of $xv$ or $yv$ is in $E(G)$ in which case we call $x$ (respectively $y$) a Type~$v$ vertex.
+By \cref{k-nn-max-degree}, there are at most $6k$ Type~$v$ vertices.
+Similarly, there are at most $6k$ Type~$w$ vertices.
+
+  Thus, in total, there are at most $13k-1$ Type~$v$, Type~$w$, and Type~C vertices.
+By \cref{k-nn-max-degree}, each of these vertices is incident with at most $6k$ edges that cross $vw$.
+Therefore, there are at most $78k^2-6k$ edges of $G$ that cross $vw$.
+Since this is true for every edge $vw\in E(G)$, $G$ is $(78k^2-6k)$-planar.
+\end{proof}
+
+We note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n$ vertices has at least $kn/2$ edges and at most $kn$ edges.
+For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is $\Omega(k^2)$.
+
+\cref{nearest-neighbour,k-planar,qn,p-centered,p-centered-treewidth} imply:
+
+\begin{cor}
+\label{k-nn}
+Every $k$-nearest-neighbour graph has:
+\begin{compactitem}
+\item an $H$-partition of layered width $O(k^4)$ in which $H$ has treewidth $O(k^6)$,
+\item queue-number at most $2^{O(k^6)}$,
+\item $p$-centered chromatic number at most $O(k^4) (p+1) (t+1)\binom{p+t}{t}$ for some integer $t\leq O(k^6)$.
+\end{compactitem}
+\end{cor}
+
+\cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
+However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (36+o(1))k^2$ by a result of \citet{DJKW16}.
+
+\section{An Open Problem}
+
+\cref{k-planar} shows that $k$-planar graphs have $H$-partitions of layered width $O(k^2)$ in which $H$ has treewidth $O(k^3)$.
+We conclude with the following question:  Does there exist a function $f:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph has an $H$-partition of layered width at most $f(k)$ in which $H$ has treewidth at most $C$?
+Perhaps $C=3$.
+Note that $C\geq 3$ even for planar graphs \citep{dujmovic.joret.ea:planar}.
+
+% %%%  Squashing the bibliography 
+%   \let\oldthebibliography=\thebibliography
+%   \let\endoldthebibliography=\endthebibliography
+%   \renewenvironment{thebibliography}[1]{%
+%     \begin{oldthebibliography}{#1}%
+%       \setlength{\parskip}{0.4ex}%
+%       \setlength{\itemsep}{0.4ex}%
+%   }{\end{oldthebibliography}}
+
+\bibliographystyle{plainnat}
+\bibliography{k-planar}
+
+% % NOTE THAT MANUAL CHANGES HAVE BEEN MADE BELOW.
+% 
+% \fontsize{11}{11.75}\selectfont 
+% 
+% \begin{thebibliography}{33}
+% \providecommand{\natexlab}[1]{#1}
+% \providecommand{\url}[1]{\texttt{#1}}
+% \expandafter\ifx\csname urlstyle\endcsname\relax
+%   \providecommand{\doi}[1]{doi: #1}\else
+%   \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
+% 
+% \bibitem[{\'{A}}brego et~al.(2011){\'{A}}brego, Monroy,
+%   Fern{\'{a}}ndez{-}Merchant, Flores{-}Pe{\~{n}}aloza, Hurtado,
+%   Sacrist{\'{a}}n, and Saumell]{abrego.munroy.ea:on}
+% Bernardo~M. {\'{A}}brego, Ruy~Fabila Monroy, Silvia Fern{\'{a}}ndez{-}Merchant,
+%   David Flores{-}Pe{\~{n}}aloza, Ferran Hurtado, Vera Sacrist{\'{a}}n, and
+%   Maria Saumell.
+% \newblock On crossing numbers of geometric proximity graphs.
+% \newblock \emph{Comput. Geom.}, 44\penalty0 (4):\penalty0 216--233, 2011.
+% \newblock \doi{10.1016/j.comgeo.2010.11.003}.
+% 
+% \bibitem[Ajtai et~al.(1982)Ajtai, Chv{\'a}tal, Newborn, and
+%   Szemer{\'e}di]{ajtai.chvatal.ea:crossing-free}
+% Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, Endre
+%   Szemer{\'e}di.
+% \newblock Crossing-free subgraphs.
+% \newblock \emph{Ann. Discrete Math.}, 12:\penalty0 9--12, 1982.
+% \newblock \doi{10.1016/S0304-0208(08)73484-4}.
+% 
+% \bibitem[Alam et~al.(2018)Alam, Bekos, Gronemann, Kaufmann, and
+%   Pupyrev]{ABGKP18}
+% Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael Kaufmann, Sergey Pupyrev.
+% \newblock Queue layouts of planar 3-trees.
+% \newblock In Therese~C. Biedl and Andreas Kerren, editors, \emph{Proc. 26th
+%   International Symposium on Graph Drawing and Network Visualization} (GD '18), vol.~11282 of \emph{Lecture Notes in Computer Science}, pp.~213--226. Springer, 2018.
+% \newblock \doi{10.1007/978-3-030-04414-5\_15}.
+% 
+% \bibitem[Alon et~al.(2002)Alon, Grytczuk, Ha{\l}uszczak, and
+%   Riordan]{AGHR-RSA02}
+% Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, Oliver Riordan.
+% \newblock Nonrepetitive colorings of graphs.
+% \newblock \emph{Random Structures Algorithms}, 21\penalty0 (3-4):\penalty0
+%   336--346, 2002.
+% \newblock \doi{10.1002/rsa.10057}.
+% 
+% \bibitem[Bodlaender(1998)]{Bodlaender-TCS98}
+% Hans~L. Bodlaender.
+% \newblock A partial $k$-arboretum of graphs with bounded treewidth.
+% \newblock \emph{Theoret. Comput. Sci.}, 209\penalty0 (1-2):\penalty0 1--45,
+%   1998.
+% \newblock \doi{10.1016/S0304-3975(97)00228-4}.
+% 
+% \bibitem[Bonamy et~al.(2019)Bonamy, Gavoille, and
+%   Pilipczuk]{bonamy.gavoille.ea:shorter}
+% Marthe Bonamy, Cyril Gavoille, and Micha{\l} Pilipczuk.
+% \newblock Shorter implicit representation for planar graphs.
+% \newblock In S.~Cabello and B.~Mohar, editors, \emph{9th Slovenian
+%   International Conference on Graph Theory} (SICGT), p.~63, June 2019.
+% 
+% \bibitem[Bose et~al.(2001)Bose, Morin, Stojmenovi\'c, and
+%   Urrutia]{bose.morin.ea:routing}
+% Prosenjit Bose, Pat Morin, Ivan Stojmenovi\'c, Jorge Urrutia.
+% \newblock Routing with guaranteed delivery in ad hoc wireless networks.
+% \newblock \emph{Wireless Networks}, 7\penalty0 (6):\penalty0 609--616, 2001.
+% \newblock \doi{10.1023/A:1012319418150}.
+% 
+% \bibitem[Bose et~al.(2012)Bose, Dujmovi\'c, Hurtado, Iacono, Langerman, Meijer,
+%   Sacrist\'an, Saumell, and Wood]{ProximityGraphs}
+% Prosenjit Bose, Vida Dujmovi\'c, Ferran Hurtado, John Iacono, Stefan Langerman,
+%   Henk Meijer, Vera Sacrist\'an, Maria Saumell, David~R. Wood.
+% \newblock Proximity graphs: $e$, $\delta$, $\delta$, $\delta$, $\chi$ and
+%   $\omega$.
+% \newblock \emph{Internat. J. Comput. Geom.  Appl.},
+%   22\penalty0 (5):\penalty0 439--469, 2012.
+% 
+% \bibitem[Chen et~al.(2002)Chen, Grigni, and Papadimitriou]{CGP02}
+% Zhi-Zhong Chen, Michelangelo Grigni, Christos~H. Papadimitriou.
+% \newblock Map graphs.
+% \newblock \emph{J. ACM}, 49\penalty0 (2):\penalty0 127--138, 2002.
+% \newblock \doi{10.1145/506147.506148}.
+% 
+% \bibitem[D{\k e}bski et~al.(2019)D{\k e}bski, Felsner, Micek, and Schr{\"o}der]{micek:personal}
+%   Micha{\l} D{\k e}bski, Stefan Felsner, Piotr Micek, Felix Schr{\" o}der. 
+% \newblock Improved bounds for centered colorings, 2019.
+% \newblock \href{https://arxiv.org/abs/1907.04586}{arXiv:1907.04586}.
+% 
+% \bibitem[Dujmovi\'{c} et~al.(2016)Dujmovi\'{c}, Joret, Kozik, and Wood]{DJKW16}
+% Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, David~R. Wood.
+% \newblock Nonrepetitive colouring via entropy compression.
+% \newblock \emph{Combinatorica}, 36\penalty0 (6):\penalty0 661--686, 2016.
+% \newblock \doi{10.1007/s00493-015-3070-6}.
+% 
+% \bibitem[Dujmovi\'c et~al.(2017{\natexlab{a}})Dujmovi\'c, Eppstein, and
+%   Wood]{dujmovic.eppstein.ea:structure}
+% Vida Dujmovi\'c, David Eppstein, David~R. Wood.
+% \newblock Structure of graphs with locally restricted crossings.
+% \newblock \emph{{SIAM} J. Discrete Math.}, 31\penalty0 (2):\penalty0 805--824,
+%   2017{\natexlab{a}}.
+% \newblock \doi{10.1137/16M1062879}.
+% 
+% \bibitem[Dujmovi\'c et~al.(2017{\natexlab{b}})Dujmovi\'c, Morin, and
+%   Wood]{dujmovic.morin.ea:layered}
+% Vida Dujmovi\'c, Pat Morin, David~R. Wood.
+% \newblock Layered separators in minor-closed graph classes with applications.
+% \newblock \emph{J. Comb. Theory, Ser. {B}}, 127:\penalty0 111--147,
+%   2017{\natexlab{b}}.
+% \newblock \doi{10.1016/j.jctb.2017.05.006}.
+% 
+% \bibitem[Dujmovi\'c et~al.(2019{\natexlab{a}})Dujmovi\'c, Esperet, Joret,
+%   Walczak, and Wood]{dujmovic.esperet.ea:planar}
+% Vida Dujmovi\'c, Louis Esperet, Gwena{\"{e}}l Joret, Bartosz Walczak,   David~R. Wood.
+% \newblock Planar graphs have bounded nonrepetitive chromatic number.
+% \newblock  \href{http://arxiv.org/abs/1904.05269}{arXiv:1904.05269}, 2019{\natexlab{a}}.
+% 
+% \bibitem[Dujmovi\'c et~al.(2019{\natexlab{b}})Dujmovi\'c, Joret, Micek, Morin,
+%   Ueckerdt, and Wood]{dujmovic.joret.ea:planar}
+% Vida Dujmovi\'c, Gwena{\"{e}}l Joret, Piotr Micek, Pat Morin, Torsten Ueckerdt, David~R. Wood.
+% \newblock Planar graphs have bounded queue-number.
+% \newblock \emph{Proc. 60th Annual Symp. on Foundations of Computer 
+%   Science} (FOCS '19), to appear. 
+%   \newblock  \href{http://arxiv.org/abs/1904.04791}{arXiv:1904.04791}, 2019{\natexlab{b}}.
+% 
+% \bibitem[Fomin et~al.(2012)Fomin, Lokshtanov, and Saurabh]{FLS-SODA12}
+% Fedor~V. Fomin, Daniel Lokshtanov, Saket Saurabh.
+% \newblock Bidimensionality and geometric graphs.
+% \newblock In \emph{Proc. 23rd {A}nnual {ACM}-{SIAM} {S}ymposium on {D}iscrete
+%   {A}lgorithms} (SODA 2012), pp.~1563--1575, 2012.
+% \newblock \doi{10.1137/1.9781611973099.124}.
+% 
+% \bibitem[Fox and Pach(2008)]{FP08}
+% Jacob Fox, J{\'a}nos Pach.
+% \newblock Separator theorems and {T}ur\'an-type results for planar intersection graphs.
+% \newblock \emph{Adv. Math.}, 219(3):1070--1080, 2008.
+% \newblock \doi{10.1016/j.aim.2008.06.002}.
+% 
+% \bibitem[Fox and Pach(2010)]{FP10}
+% Jacob Fox, J{\'a}nos Pach.
+% \newblock A separator theorem for string graphs and its applications.
+% \newblock \emph{Combin. Probab. Comput.}, 19\penalty0 (3):\penalty0 371--390,
+%   2010.
+% \newblock \doi{10.1017/S0963548309990459}.
+% 
+% \bibitem[Fox and Pach(2014)]{FP14}
+% Jacob Fox, J{\'a}nos Pach.
+% \newblock Applications of a new separator theorem for string graphs.
+% \newblock \emph{Combin. Probab. Comput.}, 23\penalty0 (1):\penalty0 66--74,
+%   2014.
+% \newblock \doi{10.1017/S0963548313000412}.
+% 
+% \bibitem[Grigoriev and Bodlaender(2007)]{grigoriev.bodlaender:algorithms}
+% Alexander Grigoriev, Hans~L. Bodlaender.
+% \newblock Algorithms for graphs embeddable with few crossings per edge.
+% \newblock \emph{Algorithmica}, 49\penalty0 (1):\penalty0 1--11, 2007.
+% \newblock \doi{10.1007/s00453-007-0010-x}.
+% 
+% \bibitem[Harvey and Wood(2017)]{HW17}
+% Daniel~J. Harvey, David~R. Wood.
+% \newblock Parameters tied to treewidth.
+% \newblock \emph{J. Graph Theory}, 84\penalty0 (4):\penalty0 364--385, 2017.
+% \newblock \doi{10.1002/jgt.22030}.
+% 
+% \bibitem[Heath et~al.(1992)Heath, Leighton, and Rosenberg]{HLR92}
+% Lenwood~S. Heath, F.~Thomson Leighton, Arnold~L. Rosenberg.
+% \newblock Comparing queues and stacks as mechanisms for laying out graphs.
+% \newblock \emph{SIAM J. Discrete Math.}, 5\penalty0 (3):\penalty0 398--412,
+%   1992.
+% \newblock \doi{10.1137/0405031}.
+% 
+% \bibitem[van~den Heuvel and Wood(2017)]{vdHW17}
+% Jan van~den Heuvel, David~R. Wood.
+% \newblock Improper colourings inspired by {H}adwiger's conjecture.
+% \newblock \emph{J. London Math. Soc.}, 98(1):129--148, 2018. 
+% \newblock  \doi{10.1112/jlms.12127}. See   \href{https://arxiv.org/abs/1704.06536}{arXiv:1704.06536} for the cited results. 
+% 
+% \bibitem[Kannan et~al.(1988)Kannan, Naor, and
+%   Rudich]{kannan.naor.ea:implicit-stoc}
+% Sampath Kannan, Moni Naor, Steven Rudich.
+% \newblock Implicit representation of graphs.
+% \newblock In Janos Simon, editor, \emph{Proc.\ 20th Annual {ACM} Symposium on
+%   Theory of Computing} (STOC '88), pp.~334--343. {ACM}, 1988.
+% \newblock \doi{10.1145/62212.62244}.
+% 
+% \bibitem[Kannan et~al.(1992)Kannan, Naor, and Rudich]{kannan.naor.ea:implicit}
+% Sampath Kannan, Moni Naor, Steven Rudich.
+% \newblock Implicit representation of graphs.
+% \newblock \emph{{SIAM} J. Discrete Math.}, 5\penalty0 (4):\penalty0 596--603,
+%   1992.
+% \newblock \doi{10.1137/0405049}.
+% 
+% \bibitem[Kobourov et~al.(2017)Kobourov, Liotta, and
+%   Montecchiani]{kobourov.liotta.ea:annotated}
+% Stephen~G. Kobourov, Giuseppe Liotta, Fabrizio Montecchiani.
+% \newblock An annotated bibliography on 1-planarity.
+% \newblock \emph{Comput. Sci. Rev.}, 25:\penalty0 49--67, 2017.
+% \newblock \doi{10.1016/j.cosrev.2017.06.002}.
+% 
+% \bibitem[Leighton(1983)]{leighton:complexity}
+% F.~Thomson Leighton.
+% \newblock \emph{Complexity issues in {VLSI}}.
+% \newblock MIT Press, Cambridge, 1983.
+% 
+% \bibitem[Mohar and Thomassen(2001)]{mohar.thomassen:graphs}
+% Bojan Mohar, Carsten Thomassen.
+% \newblock \emph{Graphs on surfaces}.
+% \newblock Johns Hopkins Press, 2001.
+% 
+% \bibitem[Ne{\v{s}}et{\v{r}}il and Ossona~de Mendez(2012)]{Sparsity}
+% Jaroslav Ne{\v{s}}et{\v{r}}il, Patrice Ossona~de Mendez.
+% \newblock \emph{Sparsity}, vol.~28 of \emph{Algorithms and Combinatorics}.
+% \newblock Springer, 2012.
+% \newblock \doi{10.1007/978-3-642-27875-4}.
+% 
+% \bibitem[Pach and T{\'o}th(1997)]{PachToth97}
+% J{\'a}nos Pach, G{\'e}za T{\'o}th.
+% \newblock Graphs drawn with few crossings per edge. 
+% \newblock \emph{Combinatorica}, 17(3):427--439, 1997
+%   2002.
+% \newblock \doi{10.1007/BF01215922}.
+% 
+% \bibitem[Pach and T{\'o}th(2002)]{PachToth-DCG02}
+% J{\'a}nos Pach, G{\'e}za T{\'o}th.
+% \newblock Recognizing string graphs is decidable.
+% \newblock \emph{Discrete Comput. Geom.}, 28\penalty0 (4):\penalty0 593--606,
+%   2002.
+% \newblock \doi{10.1007/s00454-002-2891-4}.
+% 
+% \bibitem[Pilipczuk and Siebertz(2019)]{pilipczuk.siebertz:polynomial-soda}
+% Micha{\l} Pilipczuk, Sebastian Siebertz.
+% \newblock Polynomial bounds for centered colorings on proper minor-closed graph
+%   classes.
+% \newblock In Timothy~M. Chan, editor, \emph{Proc.\ 30th Annual {ACM-SIAM}
+%   Symposium on Discrete Algorithms} ({SODA} 2019), pp.~1501--1520. {SIAM},
+%   2019.
+% \newblock \doi{10.1137/1.9781611975482.91}.
+% 
+% \bibitem[Reed(2003)]{Reed03}
+% Bruce~A. Reed.
+% \newblock Algorithmic aspects of tree width.
+% \newblock In \emph{Recent Advances in Algorithms and Combinat.}, vol.~11,
+%   pp.~85--107. Springer, 2003, 
+% \newblock \doi{10.1007/0-387-22444-0_4}.
+% 
+% \bibitem[Wiechert(2017)]{wiechert:on}
+% Veit Wiechert.
+% \newblock On the queue-number of graphs with bounded tree-width.
+% \newblock \emph{Electr. J. Comb.}, 24\penalty0 (1):\penalty0 P1.65, 2017.
+% \newblock  \url{http://www.combinatorics.org/v24i1p65}.
+% 
+% \bibitem[Zhu(2009)]{zhu:colouring}
+% Xuding Zhu.
+% \newblock Colouring graphs with bounded generalized colouring number.
+% \newblock \emph{Discrete Math.}, 309\penalty0 (18):\penalty0 5562--5568,
+%   2009.
+% \newblock \doi{10.1016/j.disc.2008.03.024}.
+% 
+% \end{thebibliography}
+
 \end{document}

### 2019-12-05 k-planar.tex
Updated master branch with all the referee's comments

@@
 \label{MakeWidth1}
-If a graph $G$ has an $H$-partition of layered width $\ell$  
-%with respect to layering $\langle V_0,V_1,\dots\rangle$, 
-for some graph $H$ of treewidth at most $t$, then $G$ has an $H'$-partition of layered width 1   
-%with respect to the same layering, 
-for some graph $H'$ of treewidth at most $(t+1)\ell-1$.
-That is, if $G\subseteq H\boxtimes P\boxtimes K_\ell$ for some graph $H$ of treewidth at most $t$  and for some path $P$, then $G\subseteq H' \boxtimes P$ for some graph $H'$ of treewidth at most $(t+1)\ell-1$.
+If a graph $G$ has an $H$-partition of layered width $\ell$ for some graph $H$ of treewidth at most $t$, then $G$ has an $H'$-partition of layered width 1 for some graph $H'$ of treewidth at most $(t+1)\ell-1$.
+
+Equivalently (by \cref{PartitionProduct}), if $G\subseteq H\boxtimes P\boxtimes K_\ell$ for some graph $H$ of treewidth at most $t$  and for some path $P$, then $G\subseteq H' \boxtimes P$ for some graph $H'$ of treewidth at most $(t+1)\ell-1$.
 \end{lem}
 
-\citet{dujmovic.joret.ea:planar} proved the following result, which with \cref{MakeWidth1}, implies \cref{PlanarBasic}.
+\citet{dujmovic.joret.ea:planar} proved the following result, which with \cref{PartitionProduct}, implies \cref{PlanarBasic}.
 
@@
 
-If a graph class admits bounded layered partitions, then it also has bounded layered treewidth. In particular, 
-\citet{dujmovic.joret.ea:planar} proved that if a graph $G$ has an $H$-partition with layered width at most $\ell$ such that $H$ has treewidth at most $k$, then $G$ has layered treewidth at most $(k+1)\ell$.
+If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
+In particular, if $\mathcal{P}=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree-decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree-decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{dujmovic.joret.ea:planar}.
+
+Therefore, any property that holds for graphs of bounded layered treewidth also holds for $G$.
+What sets layered partitions apart from layered treewidth is that they lead to constant upper bounds on the queue-number  and non-repetitive chromatic number, whereas for both these parameters, the best known upper bound obtainable via layered treewidth is $O(\log n)$; see \cref{Applications}.
+
+% $(\mathcal{P},\mathcal{L})$ is an $H$-partition of 
+% \citet{dujmovic.joret.ea:planar} proved that if a graph $G$ has an $H$-partition with layered width at most $\ell$ such that $H$ has treewidth at most $k$, then $G$ has layered treewidth at most $(k+1)\ell$.
 So any property that holds for graphs of bounded layered treewidth also holds for $G$.
@@
 So, in the case of $k$-planar graphs, \cref{k-planar} implies an analogue of the separator theorem of \citet{FP08} mentioned in the introduction, albeit with worse dependence on $k$.
-What sets layered partitions apart from layered treewidth is that they lead to constant upper bounds on the queue-number  and non-repetitive chromatic number, whereas for both these parameters, the best known upper bound obtainable via layered treewidth is $O(\log n)$; see \cref{Applications}.
 
@@
 
-For a tree $T$ rooted at some node $x_0\in V(T)$, we we say that a node $a\in V(T)$ is a \emph{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \emph{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
-Note that each node $x\in v(T)$ is a $T$-ancestor and $T$-descendant of itself.
+For a tree $T$ rooted at some node $x_0\in V(T)$, we say that a node $a\in V(T)$ is a \emph{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \emph{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
+Note that each node $x\in V(T)$ is a $T$-ancestor and $T$-descendant of itself.
 We say that a $T$-ancestor $a\in V(T)$ of $x\in V(T)$ is a \emph{strict} $T$-ancestor of $x$ if $a\neq x$.
@@
 
-We begin with a fairly standard lemma about normalised tree decompositions.
+We begin with a standard technique that allows us to work with a normalised tree decomposition:
 
 \begin{lem}\lemlabel{nice-decomposition}
-  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a $T$-decomposition $(B_x:x\in V(T))$ of $H$ with width $t$ that has following additional properties:  
+  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a width-$t$ $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:  
   \begin{compactenum}[(T1)]
@@
 Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$.
-Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y]$.
+(The purpose of $x_0$ is to ensure that every node $x$ for which $B_x$ is non-empty has a parent.)  Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y\}]$.
 If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.
@@
   \begin{compactenum}[(Y1)]
-    \item\ylabel{thickness} $\mathcal{Y}:=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
+    % \item\ylabel{thickness} $\mathcal{Y}=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
     \item\ylabel{separator} For each $x\in V(T)$, there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$. 
@@
 \begin{proof}[Proof of \cref{generalized-tripod}]
-  Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$
-  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
-Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.
-In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
+  % Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.  
+  Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
 
@@
 Let $z_0,\ldots,z_r$ be the path in $T$ from $z_0:=x'$ to $z_r:=a$.
-For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $B_{z_i}$.
+For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $B_{z_i}$, since $x'a\in E(H)$.
 However, by \tref{subtree-root} $x'$ is not contained in $B_{x_i}$ for any $i\in\{1,\ldots,r\}$.
@@
 \begin{proof}
-Apply \cref{generalized-tripod} to $G$ and let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $Y_x$, $V_x$, $F_x$, and $N_x$ be defined as in \cref{generalized-tripod}.
+Apply \cref{generalized-tripod} to $G$ and let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $Y_x$, $V_x$, $F_x$, and $N_x$ be defined as in \cref{generalized-tripod}, where the partition $\mathcal{Y}$ has width $\ell$ with respect to the layering $\mathcal{L}$.
 
 For a node $x\in V(T)$, we say that a shortcut $P\in\mathcal{P}$ \emph{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
-We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$ or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
+We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$, or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
 We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
@@
   \[
-     Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:\mbox{$(v_0,\ldots,v_r)\in\mathcal{P}$, $v\in \{v_0,\ldots,v_r\}$}\}
+     Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:\mbox{$(v_0,\ldots,v_r)\in\mathcal{P}$, $v\in \{v_0,\ldots,v_r\}$}\} \enspace .
   \]
@@
 
-For each $x\in V(T)$, we define the \emph{separator}
+For each $x\in V(T)$, define 
 \[  
@@
 
-Next we show that $\mathcal{S}$ has small layered width with respect to $\mathcal{L}$:
+Next we show that $\mathcal{S}$ has small layered width with respect to the layering $\mathcal{L}=\langle V_1,\ldots,V_h\rangle$:
 
@@
 \begin{proof}
-  We count the number of vertices in $S_x\cap V_i$ by upper-bounding the number of vertices contributed to $S_x\cap V_i$ by each vertex $w\in Y_x$. Refer to \cref{contribute}.
+  We count the number of vertices in $S_x\cap V_i$ by upper-bounding the number of vertices contributed to $S_x\cap V_i$ by each vertex $w\in Y_x$.
+Refer to \cref{contribute}.
 If $w\in Y_x\cap V_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes only one vertex, itself, to $S_x\cap V_i$.
@@
 Since $xy\in E(J)$, $G^\mathcal{P}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
-Since $vw\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.
-By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.
+Since $vw\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.\footnote{Recall that we have made the assumption that $\mathcal{P}$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.
 By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
-By \yref{ancestor-edge} $N_x$ and $V_y$ are also disjoint.
+By \yref{ancestor-edge}, $N_x$ and $V_y$ are also disjoint.
 By \yref{separator} $P$ contains an internal vertex $v'\in F_x$.
@@
   For the tree-decomposition $(C_x:x\in V(T))$ of $J$ we use the same tree $T$ used in the tree-decomposition $(B_x:x\in V(T))$ of $H$.
-For each node $x$ of $T$, $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
-  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, $a,x'\in C_{x'}$ and the connectivity of $T[a]:=T[\{x\in V(T):a\in C_x]$ is obvious.
+For each node $x$ of $T$, we define $C_x$ as follows: $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
+  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, $a,x'\in C_{x'}$.
+The connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ follows from the fact that, for every node $x'\in T[a]$, every node $x$ on the path in $T$ from $x'$ to  $a$ is also a node of $T[a]$.
 Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $J$.
@@
 
-  Since $w\in S_{\delta}$, at least one of the following is true:
+  Since $w\in S_{\delta}$, $a(w)=x_\delta$, so $w$ participates in $x_\delta$, so at least one of the following is true:
   \begin{enumerate}
-    \item $\mathcal{P}$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
+    \item There exists $w'\in V(G)$ such that $\mathcal{P}$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
     \item $w\in Y_\delta$.  In this case, we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$. 
@@
   Let $w_0,w_1,\ldots,w_p$ be the path that begins $w_0:=v'$ and then follows the subpath of $P_{ww'}$ that begins at $w_1:=w$ and ends at $w_p:=w''$.
-For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{1,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.
+For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{0,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.
 Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root, $x_0$, of $T$ is non-increasing.
 
-  By definition, $a_0=x_r$.  We claim that $a_p=x_\delta$, i.e., $s_p=\delta$. 
+  We claim that $a_0=x_r$.
+Since $v\in S_{x'}$, $a(v)=x'$, so $V(P_{vw})\subseteq V_{x'}$, otherwise $v$ participates in $a$ for some node $a$ not in the subtree $T_{x'}$ rooted at $x'$, but this contradicts \clmref{i-ancestor} since $a(v)=x'$ is a $T$-ancestor of every node in which $v$ participates.
+Since $v'=w_0\in V(P_{vw})$, $\{w_0\}\subseteq V_{x'}\subseteq V_{r}$, so $s_0=r$ and $a_0=x_r$.
+
+  We claim that $a_p=x_\delta$---that is, $s_p=\delta$. 
   To see this, first observe that, for each $i\in\{1,\ldots,p\}$, $w_i\in V_{\delta}$ since, otherwise, an internal vertex of $P_{ww'}$ belong to $F_\delta$, which would imply (by \yref{ancestor-edge}) that $w\in S_{\delta'}$ for some $\delta' < \delta$, contradicting the assumption that $w\in S_\delta$.
 Therefore $s_p\ge\delta$.
-To see that $s_p<\delta-1$,
-  observe that either $w=w''\in Y_\delta$ or $P_{ww'}$ contains an internal vertex $w''$ in $Y_\delta$.
-By \yref{thickness} and the defnition of $V_x$, $V_{\delta-1}$ does not contain $w''$, so $s_p<\delta-1$.
+To see that $s_p<\delta+1$,
+  observe that either $w=w''\in Y_\delta$ or $P_{ww'}$ contains an internal vertex $w''$ in $Y_\delta$.  By the definition of $V_x$, $V_{\delta-1}$ does not contain $w''$, so $s_p<\delta+1$.
 
@@
   Finally, we are in a position to appeal to \cite[Lemma~24]{pilipczuk.siebertz:polynomial-soda} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
+
+  At this point, the proof is complete, but let us summarize.
+For each node $x\in V(T)$, $C_x$ contains only $T$-ancestors of $x$ (including $x$ itself).
+For each ancestor $x_\delta$ of $x$ contained in $C_x$, there is path from $x$ to $x_\delta$ in $\overrightarrow{H}^+$ of length at most $k$.
+The number of ancestors of $x$ that can be reached by paths of length at most $k$ in $\overrightarrow{H}^+$ is at most $\binom{k+t}{t}$.
+Therefore $|C_x|\le \binom{k+t}{t}$, as required.
 \end{proof}

### 2020-03-25 k-planar.tex
Added table of contents
Added \listfiles
Small changes + V_i => L_i

@@
 \documentclass{patmorin}
+\listfiles
 \usepackage[T1]{fontenc}
@@
 
-\usepackage[mathlines]{lineno}
+% \usepackage[mathlines]{lineno}
 % \setlength{\linenumbersep}{2em}
-\linenumbers
+% \linenumbers
 % \rightlinenumbers
 % \linenumbers
-\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
- \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
- \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
- \renewenvironment{#1}%
-    {\linenomath\csname old#1\endcsname}%
-    {\csname oldend#1\endcsname\endlinenomath}}% 
-\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
- \patchAmsMathEnvironmentForLineno{#1}%
- \patchAmsMathEnvironmentForLineno{#1*}}%
-\AtBeginDocument{%
-\patchBothAmsMathEnvironmentsForLineno{equation}%
-\patchBothAmsMathEnvironmentsForLineno{align}%
-\patchBothAmsMathEnvironmentsForLineno{flalign}%
-\patchBothAmsMathEnvironmentsForLineno{alignat}%
-\patchBothAmsMathEnvironmentsForLineno{gather}%
-\patchBothAmsMathEnvironmentsForLineno{multline}%
-}
-
-\allowdisplaybreaks
+% \newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+%  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+%  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+%  \renewenvironment{#1}%
+%     {\linenomath\csname old#1\endcsname}%
+%     {\csname oldend#1\endcsname\endlinenomath}}%
+% \newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+%  \patchAmsMathEnvironmentForLineno{#1}%
+%  \patchAmsMathEnvironmentForLineno{#1*}}%
+% \AtBeginDocument{%
+% \patchBothAmsMathEnvironmentsForLineno{equation}%
+% \patchBothAmsMathEnvironmentsForLineno{align}%
+% \patchBothAmsMathEnvironmentsForLineno{flalign}%
+% \patchBothAmsMathEnvironmentsForLineno{alignat}%
+% \patchBothAmsMathEnvironmentsForLineno{gather}%
+% \patchBothAmsMathEnvironmentsForLineno{multline}%
+% }
+%
+%
+% \allowdisplaybreaks
 %\sloppy
 %%%%%%%%%%
-\makeatletter
-\def\NAT@spacechar{~}
-\makeatother
-%%%%%%%%%%
+% \makeatletter
+% \def\NAT@spacechar{~}
+% \makeatother
+
+\newcommand{\snote}[1]{{\color{red}#1}}
+
 \crefname{lem}{Lemma}{Lemmas}
@@
 \end{titlepage}
-
-% \tableofcontents
-% \newpage
-
-%\setcounter{\page}{1}
+\pagenumbering{roman}
+\tableofcontents
+\newpage
+
+\pagenumbering{arabic}
+% \setcounter{\page}{1}
 \section{Introduction}
@@
 
-A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle V_0,V_1,\ldots\rangle$ such that $\{V_0,V_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in V_i$ and $w\in V_j$ then $|j-i|\leq 1$.
+A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle$ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
 For any partition $\mathcal{P}=\{S_1,\ldots,S_p\}$ of $V(G)$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
@@
 A graph $G'$ obtained from a $k$-planar graph $G$ by edge deletions and edge contractions may or may not be $k$-planar.
-Indeed, \citet{dujmovic.eppstein.ea:structure} construct 1-planar graphs that contain arbitrarily large complete graph minors.
+\snote{DCG Reviewer \#1 writes: \emph{Isn't this trivial?
+A grid with all
+diagonals will do.}  Are they correct?
+If so, what to do?} Indeed, \citet{dujmovic.eppstein.ea:structure} construct 1-planar graphs that contain arbitrarily large complete graph minors.
 Our results for $k$-planar graphs are the first of this type for a non-minor-closed class.
@@
 \begin{lem}\label{generalized-tripod}
-  Let $G$ be a graph; let $\mathcal{L}:=\langle V_1,\ldots,V_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \lemref{nice-decomposition}.
+  Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \lemref{nice-decomposition}.
 For each $x\in V(T)$, let $V_x := \bigcup_{y\in V(T_x)} Y_y$, $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x:=V_x\cup F_x$.
@@
 
-Next we show that $\mathcal{S}$ has small layered width with respect to the layering $\mathcal{L}=\langle V_1,\ldots,V_h\rangle$:
+Next we show that $\mathcal{S}$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
 
 \begin{clm}\clmlabel{general-width}
-  For each $i\in\{1,\ldots,h\}$ and each $x\in V(J)$, $|S_x\cap V_i|\le d\ell(k^2+3)$.  
+  For each $i\in\{1,\ldots,h\}$ and each $x\in V(J)$, $|S_x\cap L_i|\le d\ell(k^2+3)$.
 \end{clm}
@@
 \begin{proof}
-  We count the number of vertices in $S_x\cap V_i$ by upper-bounding the number of vertices contributed to $S_x\cap V_i$ by each vertex $w\in Y_x$.
+  We count the number of vertices in $S_x\cap L_i$ by upper-bounding the number of vertices contributed to $S_x\cap L_i$ by each vertex $w\in Y_x$.
 Refer to \cref{contribute}.
-If $w\in Y_x\cap V_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes only one vertex, itself, to $S_x\cap V_i$.
+If $w\in Y_x\cap L_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes only one vertex, itself, to $S_x\cap L_i$.
 
@@
     \end{center}
-    \caption{A path $P$ containing an internal vertex $w\in Y_x\cap V_{i-j}$.}
+    \caption{A path $P$ containing an internal vertex $w\in Y_x\cap L_{i-j}$.}
     \label{contribute}
@@
   Otherwise, consider some path $P\in\mathcal{P}$ that contains $w$ as an internal vertex.
-If $w\in V_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap V_i$.
-If $w\in V_{i-1}\cup V_{i+1}$, then $P$ contributes at most $k$ vertices to $S_x\cap V_i$.
-If $w\in V_{i-j}\cup V_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap V_i$.
-
-  For any $j$, the number of vertices $w\in V_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\mathcal{P}$. Therefore, 
-  \[  |S_x\cap V_i|\le d\ell\cdot\left(k+1 + 2k + \sum_{j=2}^k 2(k-j)\right) %= d(2k+1) + \sum_{i=1}^{k-2} i
+If $w\in L_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap L_i$.
+If $w\in L_{i-1}\cup L_{i+1}$, then $P$ contributes at most $k$ vertices to $S_x\cap L_i$.
+If $w\in L_{i-j}\cup L_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap L_i$.
+
+  For any $j$, the number of vertices $w\in L_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\mathcal{P}$. Therefore,
+  \[  |S_x\cap L_i|\le d\ell\cdot\left(k+1 + 2k + \sum_{j=2}^k 2(k-j)\right) %= d(2k+1) + \sum_{i=1}^{k-2} i
       = d\ell(k^2 +3) \enspace . \qedhere
@@
   At this point, the proof of \cref{modern-major-general} is almost immediate from \clmref{general-width} and \clmref{general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
-In particular, $G^{\mathcal{P}}$ may contain edges $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.
-To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.
+In particular, $G^{\mathcal{P}}$ may contain edges $vw$ with $v\in L_i$ and $w\in L_{i+j}$ for any $j\in\{0,\ldots,k\}$.
+To resolve this, we use a new layering $\mathcal{L}':=\langle L_0',\ldots,L_h'\rangle$ in which $L_i'=\bigcup_{j=ki}^{ki+k-1} L_i$.
 This increases the layered width given by \clmref{general-width} from $d\ell(k^2+3)$ to $d\ell(k^3+3k)$.
@@
 We can reduce the layered width of the $H$-partition of $G$ from $O(k^3)$ to $O(k^2)$ by observing that the dummy vertices in $V(G_0)\setminus V(G)$ do not contribute to the layered width of this partition.
-In this setting, the proof of \cref{clm:general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $V_i\cap Y_x$.
-More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap V_j$ for $j > k+1$.
+In this setting, the proof of \cref{clm:general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $L_i\cap Y_x$.
+More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (L_{i-j}\cup L_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap L_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap L_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap L_j$ for $j > k+1$.
 Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields
@@
     \item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
-    \item For each integer $j\ge 0$, let $V_j=\{v\in V(G):\dist_T(r,v)=j\}$. 
+    \item For each integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
     \item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that
@@
         \item $F[P_i]$ is a path; and
-        \item $|V(P_i)\cap V_j| \le 15$ for all integers $j\ge 0$.
+        \item $|V(P_i)\cap L_j| \le 15$ for all integers $j\ge 0$.
       \end{compactenum}
@@
     \item $H$ is planar;
-    \item for all integers $j\ge 0$, and all $x\in V(H)$, $|S_x\cap V_j|\le 15$; 
+    \item for all integers $j\ge 0$, and all $x\in V(H)$, $|S_x\cap L_j|\le 15$;
     \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; and
@@
 
-  We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap V_i|\le 15$.
-First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $V_i$, each incident on at most two edges of $\overline{Y}$.
-Since $\dist_{G'}(v,w)\le 2$ for each $vw\in E(G)$, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap V_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $V_i$.
+  We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
+First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $L_i$, each incident on at most two edges of $\overline{Y}$.
+Since $\dist_{G'}(v,w)\le 2$ for each $vw\in E(G)$, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $L_i$.
 These at most six edges have at most 12 endpoints.
-Therefore $|V(\overline{Y}^+)\setminus V(\overline{Y})\cap V_i|\le 6\times 2=12$, so $|V(\overline{Y}^+)\cap V_i|\le 12+3=15$.
+Therefore $|V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i|\le 6\times 2=12$, so $|V(\overline{Y}^+)\cap L_i|\le 12+3=15$.
 
@@
 
-  Use the layering $\mathcal{L}=\langle V_0',V_1'\ldots\rangle$ where $V_i'=V_{2i}\cup V_{2i+1}$ for each integer $i\ge 0$.
+  Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{2i}\cup L_{2i+1}$ for each integer $i\ge 0$.
 That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le 2$ for every edge $vw\in E(G)$.
-Since $|V_j\cap S_x|\le 15$ for every integer $j\ge 0$, $|V_i'\cap S_x|\le 30$ for every integer $i\ge 0$ and every $x\in V(H)$.
+Since $|L_j\cap S_x|\le 15$ for every integer $j\ge 0$, $|L_i'\cap S_x|\le 30$ for every integer $i\ge 0$ and every $x\in V(H)$.
 \end{proof}
@@
 % Let $G$ be a $1$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
-Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \cref{1-planar} and such that $V_i' = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/2}=i\}$ for each integer $i\ge 0$.
+Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle L_0',L_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \cref{1-planar} and such that $L_i' = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/2}=i\}$ for each integer $i\ge 0$.
 % \end{cor}
@@
 %Let $G$ be a $k$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
-Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \cref{k-planar} and such that $V_i' := \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\}$ for each integer $i\ge 0$.
+Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle L_0',L_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \cref{k-planar} and such that $L_i' := \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\}$ for each integer $i\ge 0$.
 %\end{cor}
@@
 In this modified version of $G^{++}$, we have $\dist_{G^++}(r,v)=\dist_{T}(r,v)=\dist_{G^+}(r_0,v)+2$
-%  which leads to the layering $\langle V_0,V_1,\ldots\rangle$ of $G^+$ where $V_i=\{v\in V(G^+): \dist_{T}(r,v) = i\}$.
-%  
-%  The rest of proof remains unchanged until the point where we derive the layering $\mathcal{L}=\langle V_0',V_1',\ldots\rangle$ of $G$.  At this point, we set
+%  which leads to the layering $\langle L_0,L_1,\ldots\rangle$ of $G^+$ where $L_i=\{v\in V(G^+): \dist_{T}(r,v) = i\}$.
+%
+%  The rest of proof remains unchanged until the point where we derive the layering $\mathcal{L}=\langle L_0',L_1',\ldots\rangle$ of $G$.  At this point, we set
 %  \[
-%    V_i' := \{v\in V(G): \floor{\dist_{T}(r,v)-2}=i\} 
+%    L_i' := \{v\in V(G): \floor{\dist_{T}(r,v)-2}=i\}
 %         = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\} \enspace ,
@@
 Let $G^+$ be the $(\Sigma,0)$-plane graph obtained by adding a dummy vertex at each crossing in $G$.
-Let $r_0$ be any vertex in $V(G)$, let $T$ be BFS spanning tree of $G^+$ rooted at $r_0$ and let $\langle V_0,V_1,\ldots\rangle$ be the resulting layering of $G^+$.
+Let $r_0$ be any vertex in $V(G)$, let $T$ be BFS spanning tree of $G^+$ rooted at $r_0$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting layering of $G^+$.
 %
@@
 In this way, each vertex $z\in Z\cap V(G^+)\setminus V(G)$ can be charged for at most two vertices in $v,x\in Z'$ and this happens only if $(z,vw,xy)$ is a crossing in $G$.
-%   Since $|Z\cap V_i| \le 2g$, the same argument used in the proof of \cref{width-of-g-plus} shows that $|Z'\cap V_i|\le 2g\times 2\times (2k+1) = 4g(2k+1)$ for each integer $i\ge 0$.
+%   Since $|Z\cap L_i| \le 2g$, the same argument used in the proof of \cref{width-of-g-plus} shows that $|Z'\cap L_i|\le 2g\times 2\times (2k+1) = 4g(2k+1)$ for each integer $i\ge 0$.
 %
@@
 %   Apply \cref{cor-k-planar} or (in the case $k=1$) \cref{cor-1-planar} to $G^*_1$ using $G^{*+}_1$ and the distinguished vertex $r^*$ to obtain a layered $H$-partition of $G^*_1$.
-Since $G^*_1\supseteq G^*_0$ and $G^*_0$ is isomorphic to $G-Z'$, this gives a layered $H$-partition $(\mathcal{L}=\langle V_0',V_1',\ldots\rangle,\mathcal{P})$ of $G-Z'$ that satisfies the conditions of \cref{cor-k-planar} or \cref{cor-1-planar}, as appropriate.
+Since $G^*_1\supseteq G^*_0$ and $G^*_0$ is isomorphic to $G-Z'$, this gives a layered $H$-partition $(\mathcal{L}=\langle L_0',L_1',\ldots\rangle,\mathcal{P})$ of $G-Z'$ that satisfies the conditions of \cref{cor-k-planar} or \cref{cor-1-planar}, as appropriate.
 In particular, for every integer $i\ge 0$,
 %   \begin{align*}
-%     V_i'& = \{v\in V(G-Z'):\floor{\dist_{G^*}(r^*,v)/(k+1)} = i\} \\
+%     L_i'& = \{v\in V(G-Z'):\floor{\dist_{G^*}(r^*,v)/(k+1)} = i\} \\
 %         & = \{v\in V(G-Z'):\floor{\dist_{G^+}(r_0,v)/(k+1)} = i\} \enspace .
 %   \end{align*}
-%   It follows that $|V_i'\cap Z'|\le 4g(2k+1)(k+1) = 8gk^2+12gk+4g$.  
+%   It follows that $|L_i'\cap Z'|\le 4g(2k+1)(k+1) = 8gk^2+12gk+4g$.
 %
@@
 %   \[
-%     |V_i' \cap P| \le \begin{cases} 8gk^2+12gk+4g & \text{if $P=Z'$} \\
+%     |L_i' \cap P| \le \begin{cases} 8gk^2+12gk+4g & \text{if $P=Z'$} \\
 %                                     30 & \text{if $P\neq Z'$ and $k=1$} \\
@@
 \begin{proof}
-   Let $(\mathcal{L}=\langle V_0,V_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$.
+   Let $(\mathcal{L}=\langle L_0,L_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$.
    Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.
-For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in V_i\cap B_x$ such that:
+For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in L_i\cap B_x$ such that:
    \begin{compactenum}
-     \item $\alpha(v)$ is unique among $\{\phi(w): w\in V_i\cap B_x\}$, which is possible 
-     since $|V_i\cap B_x|\le \ell$,
+     \item $\alpha(v)$ is unique among $\{\phi(w): w\in L_i\cap B_x\}$, which is possible
+     since $|L_i\cap B_x|\le \ell$,
      \item $\beta(v)= i\bmod (p+1)$, and
@@
 
- First suppose that there exists $v,w\in V(X)$ with $v\in V_i$ and $w\in V_j$ with $j-i\ge p$.
+ First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$.
 Since $G[X]$ is connected, $G[X]$ contains a path from $v$ to $w$.
-By the definition of layering, this path contains at least one vertex from $V_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$.
+By the definition of layering, this path contains at least one vertex from $L_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$.
 Therefore, $|\{\beta(v'):v'\in X\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
 
- Otherwise, $X\subseteq V_{i},\ldots,V_{i+s}$ for some $s<p$.
+ Otherwise, $X\subseteq L_{i},\ldots,L_{i+s}$ for some $s<p$.
 Let $H':=H[\{x\in V(H):B_x\cap X\neq\emptyset]$.
@@
 Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
- For any $v,w\in B_x$ with $v\neq w$, either $v,w\in V_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in V_{i'}$ and $w\in V_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
+ For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
 Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.

### 2020-03-25 k-planar.tex
Relating our results to  Pilipczuk and Siebertz

@@
 
-These theorems have applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, and $p$-centered colouring  \citep{micek:personal},
+These theorems have applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, and $p$-centered colouring  \citep{debski.felsner.ea:improved},
 % and succinct graph representations \citep{bonamy.gavoille.ea:shorter}, some of which
@@
 % Say that a graph $G$ admits a $(t,\ell)$-partition if $V(G)$ has a partition $\mathcal{P}$ of layered width at most $\ell$ and the treewidth of $H:=G/\mathcal{P}$ is at most $t$.
-They defined a class $\mathcal{G}$ of graphs to \emph{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
+Dujmović \etal\ define a class $\mathcal{G}$ of graphs to \emph{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
 
@@
 \begin{compactenum}[(a)]
-\item an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$, and
-\item an $H$-partition of layered width $3$ for some planar graph $H$ of treewidth at most $3$.
+\item an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$, and \label{subthm:eight-one}
+\item an $H$-partition of layered width $3$ for some planar graph $H$ of treewidth at most $3$.\label{subthm:three-tree}
 \end{compactenum}
@@
 
+At the core of their work is an elegant proof of the following result of  \citet{pilipczuk.siebertz:polynomial-soda} that was introduced in the context of centered colourings (see \cref{sec:centered-colourings}):
+\begin{thm}[\citep{pilipczuk.siebertz:polynomial-soda}]\thmlabel{ps}
+  Every planar triangulation $G$ has an $H$-partition $\mathcal{P}$
+  such that $H$ has treewidth at most 8 and, for each $P\in\mathcal{P}$, $G[P]$ is a shortest path in $G$.
+\end{thm}
+
+Indeed, \cref{PlanarPartition}.\ref{subthm:eight-one} is a slight strengthening of \cref{thm:ps} in which $G$ has a distinguished vertex $r$ and, for each $P\in\mathcal{P}$ no two vertices of $P$ have the same distance to $r$.
+
 \section{New Results}
@@
 \begin{proof}
-  We count the number of vertices in $S_x\cap L_i$ by upper-bounding the number of vertices contributed to $S_x\cap L_i$ by each vertex $w\in Y_x$.
+  Recall that $S_x$ is defined by vertices that participate in $x$, and these are vertices that are either in $Y_x$ or in shortcuts that cross $Y_x$.
+We say that a vertex $w\in Y_x$ \emph{contributes} a vertex $v\in S_x$ if $v$ participates in $x$.
+  We upper bound the number of vertices in $S_x\cap L_i$ by upper-bounding the number of vertices contributed to $S_x\cap L_i$ by each $w\in Y_x$.
 Refer to \cref{contribute}.
-If $w\in Y_x\cap L_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes only one vertex, itself, to $S_x\cap L_i$.
+If $w\in Y_x\cap L_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
 
@@
 \subsection{Centered Colourings}
+\seclabel{centered-colourings}
 
@@
 
+\citet{pilipczuk.siebertz:polynomial-soda} and \citet{debski.felsner.ea:improved} use \cref{thm:ps} and \cref{PlanarPartition}.\ref{subthm:three-tree}
+
 We make use of the following lemma due to \citet{pilipczuk.siebertz:polynomial-soda}.
@@
 
-The following lemma is implicitly due to \citet{micek:personal}. We include the proof for completeness.
+The following lemma is implicitly in the work of \citet[Proof of Theorem 2.1]{debski.felsner.ea:improved}. We include the proof for completeness.
 
@@
 
-\begin{lem}[\citep{micek:personal}]
+\begin{lem}[\citep{debski.felsner.ea:improved}]
 \label{p-centered}

### 2020-03-25 k-planar.tex
Addressed X versus V(X) issue

@@
 A grid with all
-diagonals will do.}  Are they correct?
+diagonals will do.}  Is the reviewer correct?
 If so, what to do?} Indeed, \citet{dujmovic.eppstein.ea:structure} construct 1-planar graphs that contain arbitrarily large complete graph minors.
@@
 
-\citet{pilipczuk.siebertz:polynomial-soda} and \citet{debski.felsner.ea:improved} use \cref{thm:ps} and \cref{PlanarPartition}.\ref{subthm:three-tree}
-
-We make use of the following lemma due to \citet{pilipczuk.siebertz:polynomial-soda}.
+\citet{pilipczuk.siebertz:polynomial-soda} and \citet{debski.felsner.ea:improved} use \cref{thm:ps} and \cref{PlanarPartition}.\ref{subthm:three-tree}, respectively, to show that, when $G$ is planar, or of bounded genus, $\chi_p(G)$ is polynomial in $p$.
+We make use of the following lemma due to \citet[Lemma~15]{pilipczuk.siebertz:polynomial-arxiv}.
 
@@
 \label{p-centered-treewidth}
-Every graph $H$ of treewidth at most $t$ has $\chi_p(H)\leq  (t+1)\binom{p+t}{t}$.
+  Every graph $H$ of treewidth at most $t$ has $\chi_p(H)\leq \binom{p+t}{t}$.
 \end{lem}
@@
  First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$.
-Since $G[X]$ is connected, $G[X]$ contains a path from $v$ to $w$.
+Since $X$ is connected, $X$ contains a path from $v$ to $w$.
 By the definition of layering, this path contains at least one vertex from $L_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$.
-Therefore, $|\{\beta(v'):v'\in X\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
-
- Otherwise, $X\subseteq L_{i},\ldots,L_{i+s}$ for some $s<p$.
-Let $H':=H[\{x\in V(H):B_x\cap X\neq\emptyset]$.
-If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in X\}|> p$ so $|\{\phi(v):v\in X\}|> p$ and we are done.
+Therefore, $|\{\beta(v'):v'\in V(X)\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
+
+ Otherwise, $V(X)\subseteq L_{i},\ldots,L_{i+s}$ for some $s<p$.
+Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.
+If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in V(X)\}|> p$ so $|\{\phi(v):v\in V(X)\}|> p$ and we are done.
 Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
@@
 %$$\chi_p(G)\le (18k^2+30k)(p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
-\[ \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{k+4}{3} \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}\enspace. \]
+\[ \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}\enspace. \]
 %\note{DW}{I previously wrote ``for some integer $t\leq \binom{k+4}{3}-1$'' thinking that $\binom{p+t}{t}$ might not be monotone with $t$, but it is because of the $+t$.}
@@
 For every $1$-planar graph $G$,
-\[ \chi_p(G)\le 20 (p+3)(p+2)(p+1)^2\enspace. \]
+\[ \chi_p(G)\le 5 (p+3)(p+2)(p+1)^2\enspace. \]
 %120 (p+1) \binom{p+3}{3}
@@
 %$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
-\[ \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{k+5}{4}\binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}\enspace. \]
+\[ \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}\enspace. \]
 %O(gk^2)\times p^{O(k^3)}$.

### 2020-03-25 k-planar.tex
Pffft

@@
 
- Otherwise, $V(X)\subseteq L_{i},\ldots,L_{i+s}$ for some $s<p$.
+ Otherwise, $V(X)\subseteq L_{i}\cup\cdots\cup L_{i+s}$ for some $s<p$.
 Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.
@@
 
-\cref{p-centered,p-centered-treewidth,StringPartition} imply the first known bounds on the $p$-centered chromatic number of map graphs that are polynomial in $p$.
-
-\begin{cor}
-Every $(g,\delta)$-string graph has $p$-centered chromatic number
-\[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1)
-\tbinom{ \delta+5}{4} \binom{p+\binom{ \delta+5}{4}-1}{\binom{ \delta+5}{4}-1}\enspace, \]
-and every plane $(0,\delta)$-string graph has $p$-centered chromatic number
-\[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1)
-\tbinom{ \delta+4}{3}\binom{p+\binom{ \delta+4}{3}-1}{\binom{ \delta+4}{3}-1}\enspace.\]
-\end{cor}
+% \snote{Deleted results on p-centered colouring since the bounded-degree results by \citet{debski.felsner.ea:improved} give better dependence on $\delta$.}
+%
+%
+% \cref{p-centered,p-centered-treewidth,StringPartition} imply the first known bounds on the $p$-centered chromatic number of map graphs that are polynomial in $p$.
+%
+% \begin{cor}
+% Every $(g,\delta)$-string graph has $p$-centered chromatic number
+% \[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1)
+% \tbinom{ \delta+5}{4} \binom{p+\binom{ \delta+5}{4}-1}{\binom{ \delta+5}{4}-1}\enspace, \]
+% and every plane $(0,\delta)$-string graph has $p$-centered chromatic number
+% \[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1)
+% \tbinom{ \delta+4}{3}\binom{p+\binom{ \delta+4}{3}-1}{\binom{ \delta+4}{3}-1}\enspace.\]
+% \end{cor}
 %\label{p-centered}
@@
 
-\cref{non-repetitive,StringPartition} also give bounds on the non-repetitive chromatic number of $(g,\delta)$-string graphs, but the bound is weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ by a result of \citet{DJKW16}.
+\cref{non-repetitive,StringPartition} also give bounds on the non-repetitive chromatic number and the $p$-centered colouring number of $(g,\delta)$-string graphs, but the bounds are weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ and $\chi_p(G)\le p(64\delta)^2$ by results of \citet{DJKW16} and \citet{debski.felsner.ea:improved}, respectively.
 
@@
 
-\cref{PlanarPartition}(b), \cref{SurfacePartition}(b), \cref{modern-major-general,PowerShortcut} imply the following result, which with \cref{PartitionProduct} implies \cref{kPowerBasic} in \cref{Introduction}.
+\cref{PlanarPartition}(b), \cref{SurfacePartition}(b), \cref{modern-major-general,PowerShortcut} imply the following result, which with \cref{PartitionProduct} implies \cref{kPowerBasic} in the introduction.
 
@@
 \item $G^k$ has  $p$-centered chromatic number
-\[\chi_p(G^k) \leq \max\{2g,3\}(2k^4+6k^2) \Delta^{k}  (p+1) \tbinom{k+4}{4}
+\[\chi_p(G^k) \leq \max\{2g,3\}(2k^4+6k^2) \Delta^{k}  (p+1)
 \binom{p+\binom{k+4}{4}-1}{\binom{k+4}{4}-1}\enspace,\]
@@
 
-This result is the first constant upper bound on the queue-number and $p$-centered chromatic number of bounded powers of graphs with bounded degree and bounded Euler genus.
+This result is the first constant upper bound on the queue-number of bounded powers of graphs with bounded degree and bounded Euler genus.
 For every graph $G$, since $G^k$ has maximum degree at most $\Delta^k$, a result of \citet{DJKW16} implies that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$.
 \cref{PowerCor} improves upon this bound when $k\ll\Delta$ and $G$ has Euler genus $g$.
+Similarly, a result of \citet{debski.felsner.ea:improved} implies that $\chi_p(G^k)\le 1024p\Delta^{2k}$ and \cref{PowerCor} improves upon this bound when $p,k\ll\Delta$ and $G$ has Euler genus $g$.
 
@@
 
-We note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n$ vertices has at least $kn/2$ edges and at most $kn$ edges.
+We note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n\ge k+1$ vertices has at least $kn/2$ edges and at most $kn$ edges.
 For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is $\Omega(k^2)$.

### 2020-03-25 k-planar.tex
Did comments by referee 1

@@
 
-We prove the following theorem that shows if a graph $G$ is a subgraph of $H\boxtimes P$ and $\PP$ is a shortcut system for $G$, then $G^{\mathcal{P}}$ is a subgraph of $J\boxtimes P$, where the treewidth of $J$ is bounded by a function of $k$, $d$, and the treewidth of $H$.
+We prove the following theorem that shows if a graph $G$ is a subgraph of $H\boxtimes P$ and $\PP$ is a $(k,d)$-shortcut system for $G$, then $G^{\mathcal{P}}$ is a subgraph of $J\boxtimes P$, where the treewidth of $J$ is bounded by a function of $k$, $d$, and the treewidth of $H$.
 
@@
     % \item\tlabel{diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
-    \item\tlabel{subtree-root} For each node $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$.
-    \item\tlabel{ancestor-edge}\tlabel{last} For each edge $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
+    \item\tlabel{subtree-root} for each node $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$; and consequently
+    \item\tlabel{ancestor-edge}\tlabel{last} for each edge $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
   \end{compactenum}

### 2020-03-26 k-planar.tex
Small changes

@@
 A grid with all
-diagonals will do.}  Is the reviewer correct?
-If so, what to do?} Indeed, \citet{dujmovic.eppstein.ea:structure} construct 1-planar graphs that contain arbitrarily large complete graph minors.
+diagonals will do.}  The other reviewer write: \emph{Well,
+such a construction is trivial: just take grids with both diagonals in
+every cell.}  In response, I changed the following sentence from ``construct'' to ``observe''.
+Ok?} Indeed, \citet{dujmovic.eppstein.ea:structure} observe that even 1-planar graphs may contain arbitrarily large complete graph minors.
 Our results for $k$-planar graphs are the first of this type for a non-minor-closed class.
@@
 The following result is the main theorem of the paper.
-Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so to does $G^\PP$ for every $(k,d)$-shortcut system $\PP$ of $G$.
+Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so too does $G^\PP$ for every $(k,d)$-shortcut system $\PP$ of $G$.
 

### 2020-03-26 k-planar.tex
Addressed the triviality of the ancestor lemma

@@
 \end{compactitem}
-A \emph{tree-decomposition} $\mathcal{T}$ of a graph $G$ consists of a tree $T$ and a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that:
+A \emph{tree decomposition} $\mathcal{T}$ of a graph $G$ consists of a tree $T$ and a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that:
 \begin{compactenum}[(i)]
@@
 \end{compactenum}
-The \emph{width} of the tree-decomposition $\mathcal{T}$ is $\max\{|B_x|:x\in V(T)\}-1$.
-The \emph{treewidth} $\tw(G)$ of a graph $G$ is the minimum width of a tree-decomposition of $G$.
+The \emph{width} of the tree decomposition $\mathcal{T}$ is $\max\{|B_x|:x\in V(T)\}-1$.
+The \emph{treewidth} $\tw(G)$ of a graph $G$ is the minimum width of a tree decomposition of $G$.
 Treewidth is the standard measure of how similar a graph is to a tree.
@@
 A grid with all
-diagonals will do.}  The other reviewer write: \emph{Well,
+diagonals will do.}  The other reviewer writes: \emph{Well,
 such a construction is trivial: just take grids with both diagonals in
@@
 Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs was in terms of layered treewidth, which we now define.
-A \emph{layered tree-decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{T}$ of $G$.
+A \emph{layered tree decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree decomposition $\mathcal{T}$ of $G$.
 The layered width of $(\mathcal{L},\mathcal{T})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{T}\}$.
-The \emph{layered treewidth} of $G$ is the minimum layered width of any layered tree-decomposition of $G$.
+The \emph{layered treewidth} of $G$ is the minimum layered width of any layered tree decomposition of $G$.
 \citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph.
@@
 If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
-In particular, if $\mathcal{P}=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree-decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree-decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{dujmovic.joret.ea:planar}.
+In particular, if $\mathcal{P}=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{dujmovic.joret.ea:planar}.
 
@@
 \begin{proof}
+  That \tref{subtree-root} implies \tref{ancestor-edge} is a standard observation: If two subtrees intersect, then one contains the root of the other.
+Thus, it suffices to construct width-$t$ tree decomposition that satisfies \tref{ancestor-edge}.
+
   Begin with any width-$t$ tree decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some tree $T_0$.
@@
 Subdivide the edge between $a$ and its parent in $T$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
-This modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
+This modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
 
@@
 In this case, contract the edge $aa'$ in $T_0$, eliminating the node $a$.
-Repeating this operation will eventually produce a width-$t$ tree-decomposition of $(B_x:x\in V(T_0))$ where $f$ is a bijection between $V(H)$ and $V(T_0)$.
-Renaming each node $a\in V(T_0)$ as $f^{-1}(a)$ gives a tree-decomposition $(B_x:x\in V(T))$ with $V(T)=V(H)$.
-
-  By the definition of $f$, the tree-decomposition $(B_x:x\in V(T))$ satisfies \tref{subtree-root}.
-To see that $(B_x:x\in V(T))$ satisifies \tref{ancestor-edge}, observe that, if $xy\in E(H)$, then at least one of $x$ or $y$ is contained in $B_z$ for every node $z$ on the path from $x$ to $y$ in $T$.
+Repeating this operation will eventually produce a width-$t$ tree decomposition of $(B_x:x\in V(T_0))$ where $f$ is a bijection between $V(H)$ and $V(T_0)$.
+Renaming each node $a\in V(T_0)$ as $f^{-1}(a)$ gives a tree decomposition $(B_x:x\in V(T))$ with $V(T)=V(H)$.
+By the definition of $f$, the tree decomposition $(B_x:x\in V(T))$ satisfies \tref{subtree-root}.
+  % To see that $(B_x:x\in V(T))$ satisifies \tref{ancestor-edge}, observe that,
+  % if $xy\in E(H)$, then at least one of $x$ or $y$ is contained in $B_z$ for every node $z$ on the path from $x$ to $y$ in $T$.
 If neither $x$ nor $y$ is an ancestor of the other, then some node $z$ on this path has $T$-depth less than that of $x$ and $y$.
@@
 
-The following lemma shows how to interpret an $H$-partition of $G$ and a tree-decomposition of $H$ as a hierarchical decomposition of $G$; refer to \cref{fig:generalized-tripod}.
+The following lemma shows how to interpret an $H$-partition of $G$ and a tree decomposition of $H$ as a hierarchical decomposition of $G$; refer to \cref{fig:generalized-tripod}.
 
 \begin{lem}\label{generalized-tripod}
-  Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \lemref{nice-decomposition}.
+  Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree decomposition of $H$ satisfying the conditions of \lemref{nice-decomposition}.
 For each $x\in V(T)$, let $V_x := \bigcup_{y\in V(T_x)} Y_y$, $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x:=V_x\cup F_x$.
@@
 % %$(\mathcal{L},\mathcal{Y})$ denotes a layered $H$-partition of $G$;
-% $\mathcal{T}=(B_x:x\in V(T))$ is a tree-decomposition of $H:=G/\mathcal{Y}$ satisfying the conditions of \lemref{nice-decomposition} and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ are defined as in \clmref{generalized-tripod}.
+% $\mathcal{T}=(B_x:x\in V(T))$ is a tree decomposition of $H:=G/\mathcal{Y}$ satisfying the conditions of \lemref{nice-decomposition} and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ are defined as in \clmref{generalized-tripod}.
 
@@
 \begin{clm}\clmlabel{general-bag-size}
-  The graph $J$ has a tree-decomposition in which every bag has size at most $\binom{k+t}{t}$.
+  The graph $J$ has a tree decomposition in which every bag has size at most $\binom{k+t}{t}$.
 \end{clm}
@@
 \begin{proof}
-  For the tree-decomposition $(C_x:x\in V(T))$ of $J$ we use the same tree $T$ used in the tree-decomposition $(B_x:x\in V(T))$ of $H$.
+  For the tree decomposition $(C_x:x\in V(T))$ of $J$ we use the same tree $T$ used in the tree decomposition $(B_x:x\in V(T))$ of $H$.
 For each node $x$ of $T$, we define $C_x$ as follows: $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
@@
 The connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ follows from the fact that, for every node $x'\in T[a]$, every node $x$ on the path in $T$ from $x'$ to  $a$ is also a node of $T[a]$.
-Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $J$.
+Therefore $(C_x:x\in V(T))$ is indeed a tree decomposition of $J$.
 It remains to bound the size of each bag $C_x$.
@@
 % Veit Wiechert.
-% \newblock On the queue-number of graphs with bounded tree-width.
+% \newblock On the queue-number of graphs with bounded treewidth.
 % \newblock \emph{Electr. J. Comb.}, 24\penalty0 (1):\penalty0 P1.65, 2017.
@@
 $\tw(G) \leq \tw( Q^{(d)} \boxtimes H ) + |X| $.
-To obtain a tree-decomposition of $Q^{(d)} \boxtimes H$ with width $(t+1) (m-1)^d - 1$, start with an optimal tree-decomposition of $H$, and replace each instance of a vertex of $H$ by the corresponding copy of $Q^{(d)}$.
+To obtain a tree decomposition of $Q^{(d)} \boxtimes H$ with width $(t+1) (m-1)^d - 1$, start with an optimal tree decomposition of $H$, and replace each instance of a vertex of $H$ by the corresponding copy of $Q^{(d)}$.
 Thus

### 2020-03-26 k-planar.tex
Done with reviewers comments

@@
   %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
-  Choose $a(v)\in X_H$ to be any member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.
-Transitivity of the $T$-ancestor relationship, \tref{ancestor-edge}, and connectivity of $H[X_H]$ implies that $a(v)$ is a $T$-ancestor of every node $x\in X_H$, as required.
+  Choose $a(v)\in X_H$ to be the member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.
+Transitivity of the $T$-ancestor relationship, \tref{ancestor-edge}, and connectivity of $H[X_H]$ implies that such an $a(v)$ exists and is a $T$-ancestor of every node $x\in X_H$, as required.
 \end{proof}
@@
 (Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $J$.)
+
+From this point onward, the plan is to show that
+\begin{enumerate}[(i)]
+  \item $\mathcal{S}$ has small layered width with respect to the layering $\mathcal{L}$ and that
+  \item $J$ has small treewidth.
+\end{enumerate}
+
+First we need to understand the relationship between sets $S_x\in\mathcal{S}$ and the related sets objects $Y_x$, $V_x$, $F_x$, and $N_x$.
 
@@
 
-Next we show that $\mathcal{S}$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
+Next we complete Step~(i) and show that $\mathcal{S}$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
 
@@
 \end{proof}
+
+% From this point onward, the plan is to show that (i)~$\mathcal{S}$ has small layered width with respect to a $\mathcal{L}$ and that (ii)~$J$ has small treewidth.
+
+\snote{Reviewer \#2 writes: \emph{Lemma 3: the property (T2) is sth peculiar; EVERY tree-decomposition of H has the property that if x and y are adjacent then the root of x and the root of y are comparable in the tree; I suggest to get rid of this one.
+The same story applies to Claim 4 where the authors give an elaborate proof of that fact again.} The reviewer is right about Lemma~3 (the other reviewer noticed it also) but not Claim~4.
+In Claim~4, $T$ is the tree in a tree decomposition of $H$.
+At this point we're setting up to show that $T$ can also be used to get a tree decomposition of $J$ without blowing up the width too much.
+
+To address this, I've added the map breaking things down into Step~(i) and Step~(ii), and the following text.}
+
+We now proceed with Step~(ii), showing that $J$ has small treewidth.
+To accomplish this, we construct a small width tree-decomposition $\mathcal{C}:=(C_x:x\in V(T))$ of $J$ using the same tree $T$ used in the tree decomposition $\mathcal{T}:=(B_x:x\in V(T))$ of $H$.
+The following claim will be useful in showing that the resulting decomposition has small width.
 

### 2021-11-09 k-planar.tex
Changed the title
Major cleanup done by David

@@
 \usepackage{thm-restate}
+
+\newcommand{\arXiv}[1]{arXiv:\,\href{https://arxiv.org/abs/#1}{#1}}
+%\newcommand{\doi}[1]{\href{https://doi.org/#1}{\tt https://doi.org/#1}}
 
@@
 %%%%%%%%%%
-% \makeatletter
-% \def\NAT@spacechar{~}
-% \makeatother
+ \makeatletter
+ \def\NAT@spacechar{~}
+ \makeatother
 
@@
 
-\setlength{\parskip}{0.82ex}
-
  \newcommand{\note}[2]{{\color{red}[#1:~#2]}}
@@
 
-% \newcommand*{\doi}[1]{\href{https://dx.doi.org/#1}{\verb+doi: #1+}}
-
-%\newcommand{\Bagsize}{\ensuremath{\frac{k^3}{6} + \frac{3k^2}{2} + \frac{13k}{3} + 4}}
-% \newcommand{\bagsize}{\ensuremath{k^3/6 + 3k^2/2 + 13k/3 + 4}}
-\newcommand{\treewidth}{\ensuremath{\binom{k+4}{3}}-1}
-
-\title{\MakeUppercase{The Structure of $k$-Planar Graphs}}
+% \title{\MakeUppercase{The Structure of $k$-Planar Graphs}}
+\title{\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}}
 
@@
 Dujmovi\'c~et~al.~(FOCS 2019) recently proved that every planar graph is a subgraph of the strong product of a graph of bounded treewidth and a path.
-This tool has been used to solve longstanding problems on queue layouts, non-repetitive colouring, $p$-centered colouring, and implicit graph encoding.
-We extend this result for $k$-planar graphs, where a graph is \emph{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings.
-In particular, we prove that every $k$-planar graph is a subgraph of the strong product of a graph of treewidth $O(k^5)$ and a path.
+Analogous results were obtained for graphs of bounded Euler genus or apex-minor-free graphs.
+These tools have been used to solve longstanding problems on queue layouts, non-repetitive colouring, $p$-centered colouring, and adjacency labelling.
+This paper proves analogous product structure theorems for various non-minor-closed classes.
+One noteable example is $k$-planar graphs (those with a drawing in the plane in which each edge is involved in at most $k$ crossings).
+We prove that every $k$-planar graph is a subgraph of the strong product of a graph of treewidth $O(k^5)$ and a path.
 This is the first result of this type for a non-minor-closed class of graphs.
@@
 In fact, we work in a much more general setting based on so-called shortcut systems that are of independent interest.
+This leads to analogous results for map graphs, string graphs, graph powers, and nearest neighbour graphs.
 \end{abstract}
@@
 
+The starting point for this work is the following `product structure theorem' for planar graphs\footnote{In this paper, all graphs are finite and undirected.
+Unless specifically mentioned otherwise, all graphs are also simple.
+For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$), let $G[S]$  denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.
+We use $G-S$ as a shorthand for $G[V(G)\setminus S]$.
+We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.} by \citet{dujmovic.joret.ea:planar}.
+
+\begin{thm}[\citep{dujmovic.joret.ea:planar}]
+	\label{PlanarProduct}
+	Every planar graph is a subgraph of:
+	\begin{compactenum}[(a)]
+		\item $H\boxtimes P$ for some graph $H$ of treewidth at most $8$ and for some path $P$,
+		\item $H\boxtimes P \boxtimes K_3$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
+	\end{compactenum}
+\end{thm}
+
+Here $\boxtimes$ is the strong product\footnote{The \emph{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
+	%:\begin{compactitem}
+	%	\item  
+	$v=w$ and $xy\in E(B)$, or
+	%	\item  
+	$x=y$ and $vw\in E(A)$, or
+	%	\item  
+	$vw\in E(A)$ and $xy\in E(B)$.
+	%\end{compactitem}
+}, and treewidth\footnote{A \emph{tree decomposition} $\mathcal{T}$ of a graph $G$ consists of a tree $T$ and a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that
+%:
+%\begin{compactenum}[(i)]
+%\item 
+(i) for every $vw\in E(G)$, there exists some node $x\in V(T)$ with $v,w\in B_x$; and
+%\item 
+(ii) for every $v\in V(G)$, the induced subgraph $T[v] := T[\{x: v\in B_x\}]$ is connected.
+%\end{compactenum}
+The \emph{width} of the tree decomposition $\mathcal{T}$ is $\max\{|B_x|:x\in V(T)\}-1$.
+The \emph{treewidth} $\tw(G)$ of a graph $G$ is the minimum width of a tree decomposition of $G$.
+Treewidth is the standard measure of how similar a graph is to a tree.
+Indeed, a connected graph has treewidth 1 if and only if it is a tree.
+Treewidth is of fundamental importance in structural and algorithmic graph theory; see \citep{Reed03,HW17,Bodlaender-TCS98} for surveys.} is an invariant that measures how `tree-like' a given graph is; see
+%\cref{Tools} for formal definitions and see 
+\cref{ProductExample} for an example.
+Loosely speaking, \cref{PlanarProduct} says that every planar graph is contained in the product of a tree-like graph and a path.
+This enables combinatorial results for graphs of bounded treewidth to be generalised for planar graphs (with different constants).
+
+% \begin{wrapfigure}{r}{0.5\textwidth}
+\begin{figure}[!h]
+\centering%	\begin{center}
+	\includegraphics{ProductExample}
+	%\end{center}
+	\caption{Example of a strong product.
+	\label{ProductExample}}
+\end{figure}
+
+\noindent\cref{PlanarProduct} has been the key tool in solving the following well-known open problems:
+\begin{compactitem}
+\item \citet{dujmovic.joret.ea:planar} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992).
+\item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002).
+\item \citet{debski.felsner.ea:improved} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs. 
+\item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results going back to 1988~\citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}), and \citet{DEJGMM} have since used it to find asymptotically optimal
+adjacency labellings of planar graphs.
+\item The result of \citet{DEJGMM} implies that, for every integer
+$n>0$, there is a \emph{universal} graph $U_n$ with $n^{1+o(1)}$ 
+vertices such that every $n$-vertex planar graph is an induced
+subgraph of $U_n$.
+\end{compactitem}
+
+All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is a subgraph of $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this definition.}.
+These applications motivate finding product structure theorems for other graph classes.
+\citet{dujmovic.joret.ea:planar} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
+The \textit{Euler genus} of the non-orientable surface with $c$ cross-caps is $c$.
+The \textit{Euler genus} of a graph $G$ is the minimum integer $g$ such that $G$ embeds in a surface of Euler genus $g$.
+Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.} and for apex-minor-free graphs\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
+A class $\mathcal{G}$ of graphs is \emph{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$.
+A minor-closed class is \emph{proper} if it is not the class of all graphs.
+For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class.
+A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar.
+A 1-apex graph is \emph{apex}.
+A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.}, and \citet{DEMWW} do so for graphs in any minor-closed class and with bounded maximum degree.
+See \cref{Generalisations} for more precise statements and see \citep{DHGLW} for a survey on this topic.
+
+The purpose of this paper is to prove product structure theorems for several non-minor-closed classes of interest. Our results are the first of this type for non-minor-closed classes.
+
+\subsection{$k$-Planar Graphs}
+
+We start with the example of $k$-planar graphs.
 A graph is \emph{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings.
 Such graphs provide a natural generalisation of planar graphs, and are important in graph drawing research; see the recent bibliography on 1-planar graphs and the 140 references therein \citep{kobourov.liotta.ea:annotated}.
-The present paper studies the structure of $k$-planar graphs and other more general classes of graphs.
-\notex{PM}{I removed the following sentences because neither Vida nor could find anything about $k$-planar graphs in \cite{FP08} and, in any case, the proof is easy.
-This could easily upset a referee.
-% \note{PM}{I must be missing something.
-The planarized graph $G'$ has $O(kn)$ vertices.
-Let $S$ be the set of non-dummy vertices and take a separation $(A,B)$ of $G'$ with $|A\cap B|=O(\sqrt{kn})$, $|S\cap (A\setminus B)|\le 2|S|/3$ and $|S\cap (B\setminus A)|\le 2|S|/3$.
-Now take $X=A\cap B$ plus the endpoints of every edge of $G$ that contains a dummy vertex in $A\cap B$.
-$|X|\le 5|A\cap B|=O(\sqrt{kn})$ and $X$ is a separator of $G$.} \note{DW}{I think you have just found a new proof of this fact.
-Note that you are actually using that $G'$ has treewidth $O(\sqrt{kn})$ and then applying the Robertson-Seymour separator lemma for graphs of given treewidth (so that $|S\cap (A\setminus B)|\le 2|S|/3$ and $|S\cap (B\setminus A)|\le 2|S|/3$).
-}
-Various structural results about planar graphs generalise for $k$-planar graphs.
-For example, \citet{FP08} generalised the Lipton-Tarjan separator theorem to show that every $k$-planar graph with $n$ vertices has a balanced separator of order $O(\sqrt{kn})$.}
-
-Our main results extends the following recent theorem of \citet{dujmovic.joret.ea:planar} to a setting that includes $k$-planar graphs.
- % as well as some other interesting non-minor-closed graph families.
-
-\begin{thm}[\citep{dujmovic.joret.ea:planar}]
-\label{PlanarBasic}
-Every planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most $8$ and for some path $P$.
-\end{thm}
-
-Here $\boxtimes$ is the strong product, and treewidth is an invariant that measures how `tree-like' a given graph is; see \cref{Tools} for formal definitions and see \cref{ProductExample} for an example.
-Loosely speaking, \cref{PlanarBasic} says that every planar graph is contained in the product of a tree-like graph and a path.
-This enables combinatorial results for graphs of bounded
-treewidth to be generalised for planar graphs (with different constants).
-\cref{PlanarBasic} has been the key tool in solving several well-known open problems.
-In particular, \citet{dujmovic.joret.ea:planar} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992);  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002); and \citet{bonamy.gavoille.ea:shorter} use it to find shorter implicit representations of planar graphs (making progress on a sequence of results going back to 1988 \citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}).
-
-% \begin{wrapfigure}{r}{0.5\textwidth}
-\begin{figure}[hbtp]
-\begin{center}
-\includegraphics{ProductExample}
-\end{center}
-\caption{Example of a strong product.}
-\label{ProductExample}
-\end{figure}
-% \end{wrapfigure}
-
-We extend \cref{PlanarBasic} as follows.
+It is well-known that the family of $k$-planar graphs is not minor-closed.
+Indeed, 1-planar graphs may contain arbitrarily large complete graph minors~\citep{dujmovic.eppstein.ea:structure}.
+Hence the above results are not applicable for  $k$-planar graphs.
+We extend \cref{PlanarProduct} as follows.
 
 \begin{thm}
-\label{kPlanarBasic}
-Every $k$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth $(18k^2+48k+30)\binom{k+4}{3}-1$ and for some path $P$.
-\end{thm}
-
-Although $k$-planar graphs are the most high-profile target for a generalization of \cref{PlanarBasic}, we actually prove a substantially stronger result than \cref{kPlanarBasic} using the following definition.
+\label{kPlanarProduct}
+Every $k$-planar graph is a subgraph of $H\boxtimes P\boxtimes K_{18k^2+48k+30}$, for some graph $H$ of treewidth $\binom{k+4}{3}-1$ and for some path $P$.
+\end{thm}
+
+This theorem has applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, $p$-centered colouring  \citep{debski.felsner.ea:improved}, and adjacency labelling \citep{DEJGMM}, which we explore in \cref{Applications}.
+For example, we prove that $k$-planar graphs have bounded non-repetitive chromatic number (for fixed $k$).
+Prior to the recent work of \citet{dujmovic.esperet.ea:planar}, it was even open whether planar graphs have bounded non-repetitive chromatic number.
+
+\subsection{Shortcut Systems}
+
+Although $k$-planar graphs are the most high-profile target for a generalization of \cref{PlanarProduct}, we actually prove a substantially stronger result than \cref{kPlanarProduct} using the following definition.
 A non-empty set $\mathcal{P}$ of non-trivial paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if:
@@
 Thus $\PP$ is a $(k+1,2)$-shortcut system for $G$, such that $G'\subseteq G^\PP$.
-
-We prove the following theorem that shows if a graph $G$ is a subgraph of $H\boxtimes P$ and $\PP$ is a $(k,d)$-shortcut system for $G$, then $G^{\mathcal{P}}$ is a subgraph of $J\boxtimes P$, where the treewidth of $J$ is bounded by a function of $k$, $d$, and the treewidth of $H$.
+This idea can be pushed further to obtain a rough characterisation of $k$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $k$-planar (see \cref{Characterisation}).
+
+The following theorem is the main contribution of the paper.
+It says that if a graph class $\mathcal{G}$ has a product structure theorem, then the class of graphs obtained by applying a shortcut system to graphs in $\mathcal{G}$ also has a product structure theorem.
 
@@
 \label{ShortcutProduct}
-Let $G$ be a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
-Let $\PP$ be a $(k,d)$-shortcut system for $G$. Then $G^\PP$ is a subgraph of $J\boxtimes P$ for some graph $J$ of treewidth at most $d(k^3+3k)\binom{k+t}{t}-1$ and some path $P$.
-\end{thm}
-
-\cref{PlanarBasic,ShortcutProduct,AddDummy} immediately imply \cref{kPlanarBasic} with $O(k^{11})$ instead of $O(k^5)$. Some further observations reduce the bound to $O(k^5)$; see \cref{Tools}.
-
-\cref{ShortcutProduct} leads to several other results.
+Let $G$ be a subgraph of $H\boxtimes P \boxtimes K_\ell$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
+Let $\PP$ be a $(k,d)$-shortcut system for $G$. Then $G^\PP$ is a subgraph of $J\boxtimes P\boxtimes K_{d\ell(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
+\end{thm}
+
+Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$.
+Some further observations presented in \cref{sec-k-planar} lead to the improved result.
+
+\cref{ShortcutProduct} is applicable for many graph classes in addition to $k$-planar graphs.
 Here is one example.
@@
 We define $\dist_G(v,w):=\infty$ if $v$ and $w$ are in different connected components of $G$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\PP$ for some $(k,2k\Delta^{k})$-shortcut system $\PP$; see \cref{PowerShortcut}.
-\cref{kPlanarBasic,ShortcutProduct} then imply:
+Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} then imply:
 
@@
 \label{kPowerBasic}
-For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most
-$\Delta^{k} (2k^4+6k^2)\binom{k+8}{8}-1$.
-\end{thm}
-
-% Note that the dependence on $k$ in the bound on the treewidth of $H$ in \cref{kPowerBasic} is improved from $O(k^{12})$ to $O(k^7)$ in \cref{Powers}.
-% PM: I removed this because the bound is exponential in $k$ anyway. You only see the difference when you separate treewidth and layered width.
-
-%This theorem and several related results are proved in \cref{Structure}.
-In particular, in \cref{sec-1-planar} we present a refinement of \cref{kPlanarBasic} in the case $k=1$.
-In \cref{sec-g-k-planar} we generalise \cref{kPlanarBasic} for
-
-These theorems have applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, and $p$-centered colouring  \citep{debski.felsner.ea:improved},
-% and succinct graph representations \citep{bonamy.gavoille.ea:shorter}, some of which
-%, and graph encoding \citep{BGP}.
-which we explore in \cref{Applications}.
-For example, we prove that $k$-planar graphs have bounded non-repetitive chromatic number (for fixed $k$).
-Prior to the recent work of \citet{dujmovic.esperet.ea:planar}, it was even open whether planar graphs have bounded non-repetitive chromatic number.
-
-%\cref{Characterisation} presents a rough characterisation of $k$-planar graphs in terms of an underlying planar graph.
-This is useful for showing that various families of graphs are $k$-planar.
-For example, we conclude that bounded powers of bounded-degree planar graphs are $k$-planar, for a suitable constant $k$.
-Thus  \cref{kPlanarBasic} and the above applications are applicable to this class.
-These results also generalise for arbitrary surfaces.
-
-\cref{Examples} presents several examples of graph classes that can be obtained from a shortcut system applied to a planar graph, including graph powers, map graphs, string graphs, and nearest neighbour graphs.
-All of these results also apply, where instead of planar graphs, we consider graphs of bounded Euler genus.
-All of the applications discussed in \cref{Applications} work on these graph classes.
-
-%\citet{dujmovic.joret.ea:planar} generalised their result for planar graphs for graphs embeddable on any surface.
-In particular, they proved that every graph of Euler genus $g$ is a subgraph of $H\boxtimes P$ for some graph $H$ of treewidth at most $???k$ and for some path $P$.
-We provide a further generalisation allowing crossings.
-Say a graph is $(g,k)$-planar if it has a drawing in a surface of Euler genus at most $g$ in which each edge is involved in at most $k$ crossings.
+For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is a subgraph of $H\boxtimes P\boxtimes K_{6k^2(k^2+3)\Delta^{k}}$ for some graph $H$ of treewidth at most $\binom{k+3}{3}-1$ and some path $P$.
+\end{thm}
+
+\cref{Examples} presents further examples of graph classes that can be constructed using shortcut systems, including map graphs, string graphs, and $k$-nearest neighbour graphs.
+\cref{ShortcutProduct} implies product structure theorems for each of these classes.
+All of the above-mentioned applications also hold for these examples.
+
+\subsection{Generalisations}
+\label{Generalisations}
+
+As mentioned above, product structure theorems have been established for several minor-closed classes in addition to planar graphs.
+The first generalises \cref{PlanarProduct} for graphs of bounded Euler genus.
+
+\begin{thm}[\citep{dujmovic.joret.ea:planar}]
+	\label{GenusProduct}
+	Every graph of Euler genus $g$ is a subgraph of:
+	\begin{compactenum}[(a)]
+		\item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+8$  and some path $P$.
+		\item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $4$ and for some path $P$.\label{temp}
+	\end{compactenum}
+\end{thm}
+
+\citet{dujmovic.joret.ea:planar} generalised \cref{GenusProduct} for apex-minor-free graphs as follows.
+
+\begin{thm}[\citep{dujmovic.joret.ea:planar}]
+	\label{ApexMinorFree}
+	For every apex graph $X$, there exists $c\in\mathbb{N}$ such that every $X$-minor-free graph is a subgraph of $H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and some path $P$.
+\end{thm}
+
+The assumption that $X$ is apex is needed in \cref{ApexMinorFree}, since if the class of $X$-minor-free graphs has a product structure theorem analogous to \cref{PlanarProduct}, then $X$ is apex \citep{dujmovic.joret.ea:planar}.
+On the other hand, \citet{DEMWW} proved a product structure theorem for bounded degree graphs in any minor-closed class.
+
+\begin{thm}[\citep{DEMWW}]
+	\label{MinorFreeDegree}
+	For every graph $X$ there exists $c\in\mathbb{N}$ such that for every $\Delta\in\mathbb{N}$, every $X$-minor-free graph $G$ with maximum degree at most $\Delta$ is a subgraph of $H\boxtimes P$ for some graph $H$ with $\tw(H) \leq c\Delta$ and for some path $P$.
+\end{thm}
+
+%The next three results are immediate corollaries of \cref{GenusProduct,ApexMinorFree,ShortcutProduct,MinorFreeDegree}.
 %
 %\begin{thm}
-%For all integers $g,k\geq 1$ there is an integer $c_{k,g}$, such that every $(g,k)$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth at most $c_{k,g}$  and for some path $P$.
+%Let $\PP$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$.
+Then $G^\PP$ is a subgraph of $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
 %\end{thm}
-
-%In 1965, \citet{ringel:sechsfarbenproblem} introduced the notion of $1$-planar graphs, a generalization of planar graphs that allows edges to cross provided that no edge is involved in more than one crossing.
-This simple generalization has since become a rich source of problems and results.
-Indeed, the annotated bibliography on 1-planar graphs by \cite{kobourov.liotta.ea:annotated} contains more than 140 entries.
 %
-%The natural generalization of $1$-planar graphs, $k$-planar graphs (that allow edge to be involved in up to $k$ crossings), has been the topic of important recent work on graph drawing.
-Though planar graphs are, by now, quite well-understood, the situation is much less clear for $k$-planar graphs; even the maximum edge-density of $k$-planar graphs is not completely settled.
-For $k\le 4$, the maximum number of edges in an $n$-vertex $k$-planar graph is $(k+3)(n-2)$, which is tight for $k=1,2$ \cite{pach.toth:graphs}.
-A near-tight bound of $11(n-2)/2$ is known for $k=3$ \cite{pach.radoicic.ea:improving} but for $k>4$, the current best upper bound is $4.108\sqrt{k}n$ \cite{pach.toth:graphs}.
-
-% In the current paper we investigate the structure of $k$-planar graphs.
-Since planarity is closed under taking minors, planar graphs can be characterised by a finite set of forbidden minors.
-Indeed, Wagner's Theorem states that this set consists of the complete 5-vertex graph $K_5$ and the complete bipartite graph $K_{3,3}$.
-Unfortunately, $k$-planarity is not closed under taking minors, ruling out this kind of clean structural result.
-The purpose of the current paper is to extend a new structural result for planar graphs to $k$-planar graphs.
-
-%The purpose of the current paper is to show that new structural results for planar graphs can be extended to $k$-planar graphs.
-These results can be used to give a rough characterization of $k$-planar graphs of powers of bounded-degree planar graphs
-% Rather, we present a rough structural characterization of $k$-planar graphs that, despite its roughness, provides several new results for $k$-planar and related graph families.
-%All of these results are based on layered $H$-partitions, which we now define.{\color{red}TODO: Help.}
-
-%Thus, like genus-$g$ graphs,$k$-planar graphs are a generalization of planar graphs (which are genus-$0$ graphs).
-Unlike genus-$g$ graphs, however, the family of $k$-planar graphs is not minor-closed.
-A graph $G'$ obtained from a $k$-planar graph $G$ by edge deletions and edge contractions may or may not be $k$-planar.
-Indeed, \citet{dujmovic.eppstein.ea:structure} construct 1-planar graphs that contain arbitrarily large complete graph minors.
-
-%\notex{PM}{TODO: Write here about applications. non-repetitive colouring, $p$-centered colouring, $(g,k)$-planar graph, $(g,d)$-map graphs, $k$-NN graphs, David's rough characterization.}
-
-%The remainder of the paper is organised as follows: In \cref{k-planar} we prove \cref{sec-k-planar}.
-In \cref{sec-1-planar} we prove \cref{1-planar}.
-In \cref{g-k-planar} we prove \cref{g-k-planar}.
-In \cref{consequences} we point out some consequences of these results for some of the other graph parameters mentioned above.
-\comment{Re-DO}
+%\begin{thm}
+%For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\PP$ for $G$, $G^\PP\subseteq H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
+%\end{thm}
+%
+%\begin{thm}
+%For every graph $X$ and for all integers $k,\Delta\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ with maximum degree $\Delta$ and for every $(k,\Delta)$-shortcut system $\PP$ for $G$, $G^\PP \subseteq H \boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
+%\end{thm}
+%
+%\note{DW}{delete the above three theorems, they are never used}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Layerings, Decompositions and Partitions}
-\label{Tools}
-
-This section defines concepts and results from the literature that will be important for our work.
-
-In this paper, all graphs are finite and undirected.
-Unless specifically mentioned otherwise, all graphs are also simple.
-For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$), let $G[S]$  denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.
-We use $G-S$ as a shorthand for $G[V(G)\setminus S]$.
-We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.
-
-We now formally define $k$-planar graphs.
-An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
-The points $f(0)$ and $f(1)$ are called the \emph{endpoints} of the curve.
-When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
-A \emph{crossing} in an embedded graph $G$ is a triple $(p,vw,xy)$ with $p\in\R^2$, $vw,xy\in E(G)$ and such that $p\in (vw\cap xy)\setminus\{v,w,x,y\}$.
-An embedded graph $G$ is \emph{$k$-plane} if each edge of $G$ takes part in at most $k$ crossings.
-A (not necessarily embedded) graph $G'$ is \emph{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
-Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
-
-We now define two concepts used in the theorems in \cref{Introduction}: strong products and treewidth.
-The \emph{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if:
-\begin{compactitem}
-\item  $v=w$ and $xy\in E(B)$, or
-\item  $x=y$ and $vw\in E(A)$, or
-\item  $vw\in E(A)$ and $xy\in E(B)$.
-\end{compactitem}
-A \emph{tree decomposition} $\mathcal{T}$ of a graph $G$ consists of a tree $T$ and a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that:
-\begin{compactenum}[(i)]
-\item for every $vw\in E(G)$, there exists some node $x\in V(T)$ with $v,w\in B_x$; and
-\item for every $v\in V(G)$, the induced subgraph $T[v] := T[\{x: v\in B_x\}]$ is connected.
-\end{compactenum}
-The \emph{width} of the tree decomposition $\mathcal{T}$ is $\max\{|B_x|:x\in V(T)\}-1$.
-The \emph{treewidth} $\tw(G)$ of a graph $G$ is the minimum width of a tree decomposition of $G$.
-Treewidth is the standard measure of how similar a graph is to a tree.
-Indeed, a connected graph has treewidth 1 if and only if it is a tree.
-Treewidth is of fundamental importance in structural and algorithmic graph theory; see \citep{Reed03,HW17,Bodlaender-TCS98} for surveys.
+\subsection{Layered Partitions}
+\label{LayeredPartitions}
 
@@
 
-% \footnote{An alternative definition of a layering of $G$ is an $H$-partition of $G$ where $H$ is a path.}
-
-% A \emph{layered $H$-partition} $(\mathcal{L},\mathcal{P})$ of a graph $G$ consists of a layering $\mathcal{L}$ of $G$ and an $H$-partition $\mathcal{P}$ of $G$.
-The \emph{layered width} of $(\mathcal{L},\mathcal{P})$ is $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
-
-% \notex{DW}{I  prefer to not define ``layered $H$-partition''.
-Just define the ``layered width'' of an $H$-partition, which is what we did in \citep{dujmovic.joret.ea:planar}.
-There are many places where we write ``$H$-partition with layered width'' without writing ``layered $H$-partition''.
-I prefer it this way.}
-
-\subsection{Previous Work on Partitions of Minor-Closed Classes}
-
 \citet{dujmovic.joret.ea:planar} introduced the study of partitions with bounded layered width such that the quotient has some additional desirable property, like small treewidth.
-% Say that a graph $G$ admits a $(t,\ell)$-partition if $V(G)$ has a partition $\mathcal{P}$ of layered width at most $\ell$ and the treewidth of $H:=G/\mathcal{P}$ is at most $t$.
-Dujmović \etal\ define a class $\mathcal{G}$ of graphs to \emph{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
+Dujmovi\'c~et~al.\ define a class $\mathcal{G}$ of graphs to \emph{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
 
@@
 	\label{PartitionProduct}
-For every graph $H$, a graph $G$ has an $H$-partition of layered width at most $\ell$ if and only if $G$ is a subgraph of
-$H \boxtimes P \boxtimes K_\ell$ for some path $P$.
+	For every graph $H$, a graph $G$ has an $H$-partition of layered width at most $\ell$ if and only if $G$ is a subgraph of $H \boxtimes P \boxtimes K_\ell$ for some path $P$.
 \end{lem}
 
-\citet{dujmovic.joret.ea:planar} also showed it suffices to consider partitions of layered width 1.
-
-\begin{lem}[\citep{dujmovic.joret.ea:planar}]
-\label{MakeWidth1}
-If a graph $G$ has an $H$-partition of layered width $\ell$ for some graph $H$ of treewidth at most $t$, then $G$ has an $H'$-partition of layered width 1 for some graph $H'$ of treewidth at most $(t+1)\ell-1$.
-
-Equivalently (by \cref{PartitionProduct}), if $G\subseteq H\boxtimes P\boxtimes K_\ell$ for some graph $H$ of treewidth at most $t$  and for some path $P$, then $G\subseteq H' \boxtimes P$ for some graph $H'$ of treewidth at most $(t+1)\ell-1$.
-\end{lem}
-
-\citet{dujmovic.joret.ea:planar} proved the following result, which with \cref{PartitionProduct}, implies \cref{PlanarBasic}.
-
-\begin{thm}[\citep{dujmovic.joret.ea:planar}]
-\label{PlanarPartition}
-Every planar graph has:
-\begin{compactenum}[(a)]
-\item an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$, and \label{subthm:eight-one}
-\item an $H$-partition of layered width $3$ for some planar graph $H$ of treewidth at most $3$.\label{subthm:three-tree}
-\end{compactenum}
-\end{thm}
-Their proof is constructive and gives a simple quadratic-time algorithm for finding these partitions and corresponding layerings.
-%Layered $H$-partitions of small layered width in which $H$ has some additional property are a useful tool for a number of graph problems.
-One such useful property is small \emph{treewidth}, which we now define.
-%
-% \citet{dujmovic.joret.ea:planar} show that, if each graph $G$ in some family $\mathcal{G}$ of graphs has a layered $H$-partition for which the layered width of the partition and the treewidth of $H$ are each upper-bounded by some constant $c_\mathcal{G}$, then each of the following quantities are upper bounded by a constant, for every $G\in\mathcal{G}$: queue number, track number, and layered treewidth.
-\citet{dujmovic.esperet.ea:planar} add non-repetitive chromatic number to this list.
-\citet{dujmovic.joret.ea:planar} show that two common families of graphs have such $H$-partitions: planar graphs, graphs of bounded genus, and (more generally) apex-minor-free graphs.
-%
-% In this paper we show that another family of graphs admits layered $H$-partitions of bounded layered width in which $H$ has bounded treewidth:
-The same authors proved the following generalisation of \cref{PlanarBasic,PlanarPartition} for graphs embeddable on other surfaces.\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
-The \textit{Euler genus} of the non-orientable surface with $c$ cross-caps is $c$.
-The \textit{Euler genus} of a graph $G$ is the minimum integer $g$ such that $G$ embeds in a surface of Euler genus $g$.
-Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.}
-
-\begin{thm}[\citep{dujmovic.joret.ea:planar}]
-\label{SurfacePartition}
-Every graph of Euler genus $g$ is a subgraph of:
-\begin{compactenum}[(a)]
-\item  $H \boxtimes P \boxtimes K_{\max\{2g,1\}}$ for some graph $H$ of treewidth at most $9$ and for some path $P$;
-\item  $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $4$ and for some path $P$.
-%\item $(K_{2g} + H )  \boxtimes P$ for some graph $H$ of treewidth at most $8$  and some path $P$.
-\end{compactenum}
-Equivalently, every graph of Euler genus $g$ has:
-\begin{compactenum}[(a)]
-\item an $H$-partition with layered width at most $\max\{2g,1\}$ such that $H$ has treewidth at most $9$;
-\item an $H$-partition with layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most $4$.
-%\item an $H$-partition with layered width at most $1$ such that $H$ has treewidth at most $2g+8$.
-\end{compactenum}
-%PM: Keep 'em all: \textcolor{red}{eliminate some of these results? just keep (b)? }
-\end{thm}
-
-\citet{dujmovic.joret.ea:planar} generalised \cref{SurfacePartition} further as follows.\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
-A class $\mathcal{G}$ of graphs is \emph{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$.
-A minor-closed class is \emph{proper} if it is not the class of all graphs.
-For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class.
-A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar.
-A 1-apex graph is \emph{apex}.
-A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.}
-
-\begin{thm}[\citep{dujmovic.joret.ea:planar}]
-\label{ApexMinorFree}
-For every apex graph $X$, there exists $c\in\mathbb{N}$ such that every $X$-minor-free graph $G$ has
-an $H$-partition with layered width 1 such that $H$ has treewidth at most $c$. Equivalently, $G\subseteq H\boxtimes P$ for some path $P$.
-\end{thm}
-
-At the core of their work is an elegant proof of the following result of  \citet{pilipczuk.siebertz:polynomial-soda} that was introduced in the context of centered colourings (see \cref{sec:centered-colourings}):
-\begin{thm}[\citep{pilipczuk.siebertz:polynomial-soda}]\thmlabel{ps}
+As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a), 
+\citet{dujmovic.joret.ea:planar} showed that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$.
+The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.
+At the core of their work is the elegant proof by \citet{pilipczuk.siebertz:polynomial-soda} of the following result:
+
+\begin{thm}[\citep{pilipczuk.siebertz:polynomial-soda}]
+\label{ps}
 Every planar triangulation $G$ has an $H$-partition $\mathcal{P}$
-  such that $H$ has treewidth at most 8 and, for each $P\in\mathcal{P}$, $G[P]$ is a shortest path in $G$.
-\end{thm}
-
-Indeed, \cref{PlanarPartition}.\ref{subthm:eight-one} is a slight strengthening of \cref{thm:ps} in which $G$ has a distinguished vertex $r$ and, for each $P\in\mathcal{P}$ no two vertices of $P$ have the same distance to $r$.
-
-\section{New Results}
-
-Apex-minor-free graphs, addressed by \cref{ApexMinorFree}, are the largest minor-closed class that admit bounded layered partitions~\citep{dujmovic.joret.ea:planar}.
-However, the family of $k$-planar graphs is not minor-closed.
-A graph $G'$ obtained from a $k$-planar graph $G$ by edge deletions and edge contractions may or may not be $k$-planar.
-\snote{DCG Reviewer \#1 writes: \emph{Isn't this trivial?
-A grid with all
-diagonals will do.}  The other reviewer writes: \emph{Well,
-such a construction is trivial: just take grids with both diagonals in
-every cell.}  In response, I changed the following sentence from ``construct'' to ``observe''.
-Ok?} Indeed, \citet{dujmovic.eppstein.ea:structure} observe that even 1-planar graphs may contain arbitrarily large complete graph minors.
-Our results for $k$-planar graphs are the first of this type for a non-minor-closed class.
-
-The following result is the main theorem of the paper.
-Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so too does $G^\PP$ for every $(k,d)$-shortcut system $\PP$ of $G$.
+such that $\tw(H)\leq 8$ and $G[P]$ is a shortest path in $G$ for each $P\in\mathcal{P}$.
+\end{thm}
+
+Indeed, the above-mentioned result of \citet{dujmovic.joret.ea:planar} is a slight strengthening of \cref{ps}, where for each $P\in\mathcal{P}$ no two vertices of $P$ have the same distance to some fixed root vertex $r$.
+
+The following result is the main technical contribution of the paper.
+Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so too does $G^\PP$ for every shortcut system $\PP$ of $G$.
 
 \begin{restatable}{thm}{mmg}
-  %\begin{thm}
-  \label{modern-major-general}
+	\label{ShortcutPartition}
 	Let $G$ be a graph having an $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\PP$ be a $(k,d)$-shortcut system for $G$.
 Then $G^\PP$ has a $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$.
-  %\end{thm}
 \end{restatable}
 
-\cref{modern-major-general,MakeWidth1} immediately imply \cref{ShortcutProduct} in the introduction.
-
-Using the relationship between $k$-planar graphs and $(k+1,2)$-shortcut systems along with a direct application of \cref{modern-major-general} we can obtain a slightly weaker version of \cref{k-planar}, below.
-The minor modifications needed to obtain the stronger bound are described in \cref{sec-k-planar}.
-
-\begin{thm}
-\label{k-planar}
-Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 48k+30$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
-\end{thm}
-
-% \cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with the weaker bound of $O(k^3)$
-%  % $(2\times 3)((k+1)^3+3(k+1))=6k^3 + 18k^2 + 36k+24$
-% on the layered width.
-% % \notex{DW}{I would just write, ``\cref{PlanarPartition,modern-major-general} imply \cref{k-planar} with a weaker bound on the layered width.}
-% We prove the stronger bound in \cref{sec-k-planar}.
-% %Also note that \cref{k-planar,MakeWidth1} imply \cref{kPlanarBasic} with the $O(k^5)$ bound on the treewidth of $H$, as promised.
-% % PM: I removed the last line because it's already stated in the line preceding the theorem
-
-In the important special case of $k=1$ we obtain better constants and an additional property (planarity) of $H$ (see \cref{sec-1-planar} for the proof):
-
-\begin{thm}
-\label{1-planar}
-Every 1-planar graph has an $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
-\end{thm}
-
-%These results imply that $k$-planar graphs have all the graph parameters mentioned above upper-bounded by some constant $c_k$.
-
-The definition of $k$-planar graphs naturally generalises for other surfaces.
-A graph $G$ drawn on a surface $\Sigma$ is $(\Sigma,k)$-plane if every edge of $G$ is involved in at most $k$ crossings.
-A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
- %\citet[Theorem~22]{dujmovic.joret.ea:planar} show that every genus-$g$ graph has a layered $H$-partition of layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most 4.
-\cref{AddDummy} immediately generalises as follows:
-
-\begin{obs}
-\label{gAddDummy}
-Every $(g,k)$-planar graph is a subgraph of $G^\PP$ for some graph $G$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G$.
-\end{obs}
-
-We prove that $(g,k)$-planar graphs admit bounded layered partitions.
-
-\begin{thm}
-\label{g-k-planar}
-Every $(g,k)$-planar graph has an $H$-partition of layered width at most
-$\max\{2g,3\}\cdot(6k^2+16k+10)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
-\end{thm}
-
-Again, a direct application of \cref{modern-major-general} using \cref{SurfacePartition}(b) implies \cref{g-k-planar} with a weaker bound on the layered width.
-We prove the stronger bound in \cref{sec-k-planar}.
-
-Finally, we state the following corollary of \cref{ApexMinorFree,modern-major-general,MakeWidth1}.
-This is the most general result that follows from \cref{modern-major-general} and the work of \citet{dujmovic.joret.ea:planar}.
-
-\begin{thm}
-For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\PP$ for $G$,  $G^\PP$ has an $H$-partition of layered width 1 such that $H$ has treewidth at most $c$; that is $G^\PP \subseteq H \boxtimes P$ for some path $P$.
-\end{thm}
-
-%%%%%%%%%%%%%%%%
-\subsection{Previous Work on the Structure of $(g,k)$-Planar Graphs}
-\label{PriorWork}
-
-%\notex{DW}{should this subsection be moved to the end of \cref{Tools}?}\notex{PM}{No, I don't think so.
-The information given in Section 2 isn't detailed enough to compare with layered treewidth versus layered partitions.}
-
-Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs was in terms of layered treewidth, which we now define.
-A \emph{layered tree decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree decomposition $\mathcal{T}$ of $G$.
-The layered width of $(\mathcal{L},\mathcal{T})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{T}\}$.
-The \emph{layered treewidth} of $G$ is the minimum layered width of any layered tree decomposition of $G$.
-\citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph.
-\citet{dujmovic.eppstein.ea:structure} show that every $k$-planar graph has layered treewidth at most $6(k+1)$, and more generally that every $(g,k)$-planar graph has layered treewidth at most $(4g+6)(k+1)$.
-It follows from this result that $(g,k)$-planar graphs have treewidth $O(\sqrt{(g+1)(k+1)n})$ and thus have balanced separators of the same order, which can also be concluded from the work of
-\citet{FP08}.
-In related work, \citet{grigoriev.bodlaender:algorithms} used structural results to obtain approximation algorithms for $(g,k)$-planar graphs, and \citet{PachToth97} determined the maximum number of edges in a $k$-planar graph (up to a constant factor).
-
-If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
-In particular, if $\mathcal{P}=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{dujmovic.joret.ea:planar}.
-
-Therefore, any property that holds for graphs of bounded layered treewidth also holds for $G$.
-What sets layered partitions apart from layered treewidth is that they lead to constant upper bounds on the queue-number  and non-repetitive chromatic number, whereas for both these parameters, the best known upper bound obtainable via layered treewidth is $O(\log n)$; see \cref{Applications}.
-
-% $(\mathcal{P},\mathcal{L})$ is an $H$-partition of
-% \citet{dujmovic.joret.ea:planar} proved that if a graph $G$ has an $H$-partition with layered width at most $\ell$ such that $H$ has treewidth at most $k$, then $G$ has layered treewidth at most $(k+1)\ell$.
-So any property that holds for graphs of bounded layered treewidth also holds for $G$.
-%For example, Norin proved that every $n$-vertex graph with layered treewidth at most $t$ has treewidth less than $2\sqrt{tn}$ (see \citep{dujmovic.morin.ea:layered}).
-Thus, $G$ has treewidth at most $2\sqrt{(k+1)\ell n}$.
-This in turn leads to a $O(\sqrt{(k+1)\ell n})$ balanced separator theorem for such graphs \citep{RS-II}.
-So, in the case of $k$-planar graphs, \cref{k-planar} implies an analogue of the separator theorem of \citet{FP08} mentioned in the introduction, albeit with worse dependence on $k$.
-
-%Indeed the former structure leads to $O(1)$ bounds on the queue-number, instead of $O(\log n)$ bounds obtained via layered treewidth.
-That said, it is open whether graphs of bounded layered treewidth have bounded queue-number.
-It is even possible that graphs of bounded layered treewidth have partitions of bounded layered width, whose quotient has bounded treewidth.
-%
-%This already implies that several of the graph parameters mentioned above, including queue-number and non-repetitive chromatic number are upper-bounded by $O(k\log n)$ for $n$-vertex $k$-planar graphs \cite{dujmovic.morin.ea:layered}.
-\cref{k-planar} is a significant strengthening of this result.
-It implies that $k$-planar graphs have layered treewidth bounded by a function of $k$, and gives upper bounds on the queue-number and non-repetitive chromatic number of $n$-vertex $k$-planar graphs that are independent of $n$.
-%
+Note that \cref{ShortcutPartition} is equivalent to \cref{ShortcutProduct} by \cref{PartitionProduct}.
 
@@
 
-The purpose of this section is to prove our main result, \cref{modern-major-general}.
+The purpose of this section is to prove our main technical result, \cref{ShortcutPartition}.
 This theorem shows how, given a $(k,d)$-shortcut system $\mathcal{P}$ of a graph $G$, a $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\mathcal{P}}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
-Our main results for $k$-planar graphs (\cref{k-planar}) and $(g,k)$-planar graphs (\cref{g-k-planar}) follow; see \cref{sec-k-planar}.
 
@@
 
-% \notex{W}{I think \clmref{generalized-tripod} should be presented as a stand-alone lemma.
-Then we repeat the statement of \cref{modern-major-general} followed by $\\begin \{proof\} ....
-\\end\{proof\}$.}
-
-% \notex{DW}{Is some sort of converse to \cref{generalized-tripod}  also true?
-\cref{generalized-tripod}  shows that if a graph $G$ has a layered partition with all the nice properties, then $G$ has this generalised tripod partition.
-Isn't also true that if $G$ has a generalised tripod partition, then $G$ has a layered partition with all the nice properties.
-The proof is basically in \citep{dujmovic.joret.ea:planar}, right?}
-% \notex{PM}{Yes, if by generalized tripod partition you mean $T$ and $\mathcal{Y}=(Y_x:x\in V(T))$ that satisify \yref{thickness}, \yref{separator}, and \yref{ancestor-edge}.
-From \yref{separator} and \yref{ancestor-edge} you can construct a width-$t$ $T$-decomposition of $H=G/\mathcal{Y}$.
-By \yref{thickness}, $\mathcal{Y}$, has layered width at most $\ell$.}
-
 The following lemma shows how to interpret an $H$-partition of $G$ and a tree decomposition of $H$ as a hierarchical decomposition of $G$; refer to \cref{fig:generalized-tripod}.
@@
 
-\begin{compactenum}[(Y1)]\setcounter{enumi}{3}
+\begin{compactenum}[(Y1)]\setcounter{enumi}{2}
   \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$.
@@
 
-% For non-negative integers $k$ and $d$, a graph $G$ and distinct vertices $v,w\in V(G)$, a \emph{$(k,d)$ $vw$-shortcut} in $G$ is a path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.
-The \emph{$(k,d)^\text{th}$ power of $G$}, denoted $G^{k,d}$, is the graph with vertex $V(G^{k,d})=V(G)$ in which, for each distinct pair $v,w\in V(G)$, $vw\in E(G^{k,d})$ if and only if $G$ contains a $(k,d)$ $vw$-shortcut.
-
-%A collection $\mathcal{P}$ of paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if every path in $\mathcal{P}$ has length at most $k$ and if, for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
-Each path in $P\in\mathcal{P}$ is called a \emph{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \emph{$vw$-shortcut}.
-Given a graph $G$ and a $(k,d)$-shortcut system $\mathcal{P}$ for $G$, we use $G^{\mathcal{P}}$ to denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\mathcal{P}$.
-
-% In the rest of this section,
-% %$(\mathcal{L},\mathcal{Y})$ denotes a layered $H$-partition of $G$;
-% $\mathcal{T}=(B_x:x\in V(T))$ is a tree decomposition of $H:=G/\mathcal{Y}$ satisfying the conditions of \lemref{nice-decomposition} and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ are defined as in \clmref{generalized-tripod}.
-
 We are now ready to prove our main result, which we restate here for convenience:
@@
 \begin{proof}
-  Let
-  \[
-     Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:\mbox{$(v_0,\ldots,v_r)\in\mathcal{P}$, $v\in \{v_0,\ldots,v_r\}$}\} \enspace .
-  \]
-  Clearly $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
+  Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\mathcal{P}, v\in \{v_0,\ldots,v_r\}\}$.
+Then $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
 
@@
 
-For each $x\in V(T)$, define
-\[
-     S_x := \{v\in V(G): a(v)= x\} \enspace .
-\]
+For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
 Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
@@
 
-From this point onward, the plan is to show that
-\begin{enumerate}[(i)]
+From this point onward, the plan is to show that:
+\begin{compactenum}[(i)]
   \item $\mathcal{S}$ has small layered width with respect to the layering $\mathcal{L}$ and that
   \item $J$ has small treewidth.
-\end{enumerate}
-
-First we need to understand the relationship between sets $S_x\in\mathcal{S}$ and the related sets objects $Y_x$, $V_x$, $F_x$, and $N_x$.
-
-\begin{clm}\clmlabel{s-subset}
+\end{compactenum}
+
+First we need to understand the relationship between sets $S_x\in\mathcal{S}$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
+
+\begin{clm}
+	\label{s-subset}
   For every $x\in V(T)$, $S_x\subseteq V_x$.
@@
 
-\begin{clm}\clmlabel{general-width}
+\begin{clm}
+	\label{general-width}
   For each $i\in\{1,\ldots,h\}$ and each $x\in V(J)$, $|S_x\cap L_i|\le d\ell(k^2+3)$.
@@
   For any $j$, the number of vertices $w\in L_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\mathcal{P}$. Therefore,
-  \[  |S_x\cap L_i|\le d\ell\cdot\left(k+1 + 2k + \sum_{j=2}^k 2(k-j)\right) %= d(2k+1) + \sum_{i=1}^{k-2} i
+  \[  |S_x\cap L_i|\le d\ell  \, \Big(k+1 + 2k + \sum_{j=2}^k 2(k-j)\Big) %= d(2k+1) + \sum_{i=1}^{k-2} i
       = d\ell(k^2 +3) \enspace . \qedhere
@@
 
-% From this point onward, the plan is to show that (i)~$\mathcal{S}$ has small layered width with respect to a $\mathcal{L}$ and that (ii)~$J$ has small treewidth.
-
-\snote{Reviewer \#2 writes: \emph{Lemma 3: the property (T2) is sth peculiar; EVERY tree-decomposition of H has the property that if x and y are adjacent then the root of x and the root of y are comparable in the tree; I suggest to get rid of this one.
-The same story applies to Claim 4 where the authors give an elaborate proof of that fact again.} The reviewer is right about Lemma~3 (the other reviewer noticed it also) but not Claim~4.
-In Claim~4, $T$ is the tree in a tree decomposition of $H$.
-At this point we're setting up to show that $T$ can also be used to get a tree decomposition of $J$ without blowing up the width too much.
-
-To address this, I've added the map breaking things down into Step~(i) and Step~(ii), and the following text.}
-
 We now proceed with Step~(ii), showing that $J$ has small treewidth.
@@
 
-\begin{clm}\clmlabel{i-ancestor}
+\begin{clm}\label{i-ancestor}
   For each edge $xy\in E(J)$, one of $x$ or $y$ is a $T$-ancestor of the other.
@@
 Since $xy\in E(J)$, $G^\mathcal{P}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
-Since $vw\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.\footnote{Recall that we have made the assumption that $\mathcal{P}$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.
+Since $vw\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.\footnote{Recall that we have made the assumption that $\mathcal{P}$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  By \cref{s-subset}, $v\in V_x$ and $w\in V_y$.
 By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
@@
 
-\begin{clm}\clmlabel{general-bag-size}
+\begin{clm}
+	\label{general-bag-size}
   The graph $J$ has a tree decomposition in which every bag has size at most $\binom{k+t}{t}$.
@@
 For each node $x$ of $T$, we define $C_x$ as follows: $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
-  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, $a,x'\in C_{x'}$.
+  \cref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, $a,x'\in C_{x'}$.
 The connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ follows from the fact that, for every node $x'\in T[a]$, every node $x$ on the path in $T$ from $x'$ to  $a$ is also a node of $T[a]$.
@@
   Since $w\in S_{\delta}$, $a(w)=x_\delta$, so $w$ participates in $x_\delta$, so at least one of the following is true:
-  \begin{enumerate}
+  \begin{compactenum}
     \item There exists $w'\in V(G)$ such that $\mathcal{P}$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
     \item $w\in Y_\delta$.  In this case, we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$.
-  \end{enumerate}
+  \end{compactenum}
   Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{\delta}$.
@@
   We claim that $a_0=x_r$.
-Since $v\in S_{x'}$, $a(v)=x'$, so $V(P_{vw})\subseteq V_{x'}$, otherwise $v$ participates in $a$ for some node $a$ not in the subtree $T_{x'}$ rooted at $x'$, but this contradicts \clmref{i-ancestor} since $a(v)=x'$ is a $T$-ancestor of every node in which $v$ participates.
+Since $v\in S_{x'}$, $a(v)=x'$, so $V(P_{vw})\subseteq V_{x'}$, otherwise $v$ participates in $a$ for some node $a$ not in the subtree $T_{x'}$ rooted at $x'$, but this contradicts \cref{i-ancestor} since $a(v)=x'$ is a $T$-ancestor of every node in which $v$ participates.
 Since $v'=w_0\in V(P_{vw})$, $\{w_0\}\subseteq V_{x'}\subseteq V_{r}$, so $s_0=r$ and $a_0=x_r$.
@@
   % Note that, by \tref{ancestor-edge}, $xy\in E(H^+)$ if and only if $x\in B_y$ or $y\in B_x$.
-  We claim that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \emph{lazy walk} in a graph $H$ is a walk in the pseudograph $H'$ obtained by adding a self loop to each vertex of $H$.} in $H^+$.
+  We claim that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \emph{lazy walk} in a graph $H$ is a walk in the pseudograph $H'$ obtained by adding a loop at each vertex of $H$.} in $H^+$.
 Indeed, if $a_i\neq a_{i+1}$ for some $i\in\{0,\ldots,p-1\}$ then this is precisely because $w_i\in V_{a_i}$ but $w_{i+1}\not\in V_{a_i}$.
@@
 
-% Together, \clmref{general-width} and \clmref{general-bag-size} imply the following generalization of \thmref{k-planar}:
-
-  At this point, the proof of \cref{modern-major-general} is almost immediate from \clmref{general-width} and \clmref{general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
+At this point, the proof of \cref{ShortcutPartition} is almost immediate from \cref{general-width,general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
 In particular, $G^{\mathcal{P}}$ may contain edges $vw$ with $v\in L_i$ and $w\in L_{i+j}$ for any $j\in\{0,\ldots,k\}$.
 To resolve this, we use a new layering $\mathcal{L}':=\langle L_0',\ldots,L_h'\rangle$ in which $L_i'=\bigcup_{j=ki}^{ki+k-1} L_i$.
-This increases the layered width given by \clmref{general-width} from $d\ell(k^2+3)$ to $d\ell(k^3+3k)$.
-Therefore $G$ has an $H$-partition of layered width at most $d\ell(k^3+3k)$ in which $H$ has treewidth at most $\binom{k+t}{t}$, completing the proof of \cref{modern-major-general}.
+This increases the layered width given by \cref{general-width} from $d\ell(k^2+3)$ to $d\ell(k^3+3k)$.
+Therefore $G$ has an $H$-partition of layered width at most $d\ell(k^3+3k)$ in which $H$ has treewidth at most $\binom{k+t}{t}$, completing the proof of \cref{ShortcutPartition}.
 \end{proof}
@@
 %%%%%%%%%%%%%%%%%%%
-\section{\boldmath $k$-Planar Graphs}
+\section{Allowing Crossings}
 \label{sec-k-planar}
 
-This section shows that the constants in \cref{modern-major-general} can be improved in the case of $k$-planar graphs.
+This section applies our main results for shortcut systems to prove graph product structure theorem for graphs drawn with a bounded number of crossings per edge.
+Then we show how to improve the bounds in this case.
+
+\subsection{$k$-Planar Graphs}
+
+We first formally define $k$-planar graphs.
+An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
+The points $f(0)$ and $f(1)$ are called the \emph{endpoints} of the curve.
+When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
+A \emph{crossing} in an embedded graph $G$ is a triple $(p,vw,xy)$ with $p\in\R^2$, $vw,xy\in E(G)$ and such that $p\in (vw\cap xy)\setminus\{v,w,x,y\}$.
+An embedded graph $G$ is \emph{$k$-plane} if each edge of $G$ takes part in at most $k$ crossings.
+A (not necessarily embedded) graph $G'$ is \emph{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
+Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
+
+As mentioned in \cref{Introduction}, \cref{PlanarProduct,ShortcutProduct} imply a product structure theorem for $k$-planar graphs. We get improved bounds as follows.
+
+%\begin{thm}
+%\label{k-planar}
+%Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 48k+30$ in which $\tw(H)\leq \binom{k+4}{3}-1$. 
+%\end{thm}
+
+\begin{proof}[Proof of \cref{kPlanarProduct}]
 	Let $G$ be a $k$-plane graph.
@@
 Let $\PP := \{P_{vw}: vw\in E(G)\}$.
-Observe that $\PP$ is a $(k+1,2)$-shortcut system for $G_0$ and that $G_0^{\mathcal{P}}\supseteq G$.
-Specifically, $G_0^{\mathcal{P}}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.
-
-Since $G_0$ is planar,  \cref{PlanarPartition}(b) implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
-Applying \cref{modern-major-general} to $G_0$ and $\mathcal{P}$ immediately implies that $G$ (an arbitrary $k$-planar graph) has an $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
+Observe that $\PP$ is a $(k+1,2)$-shortcut system for $G_0$ and that $G_0^{\PP}\supseteq G$.
+Specifically, $G_0^{\PP}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.
+
+	Since $G_0$ is planar,  \cref{PlanarProduct}(b) and \cref{PartitionProduct} implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
+Applying \cref{ShortcutPartition} to $G_0$ and $\mathcal{P}$ immediately implies that $G$ (an arbitrary $k$-planar graph) has an $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
 
 	We can reduce the layered width of the $H$-partition of $G$ from $O(k^3)$ to $O(k^2)$ by observing that the dummy vertices in $V(G_0)\setminus V(G)$ do not contribute to the layered width of this partition.
-In this setting, the proof of \cref{clm:general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $L_i\cap Y_x$.
+In this setting, the proof of \cref{general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $L_i\cap Y_x$.
 More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (L_{i-j}\cup L_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap L_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap L_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap L_j$ for $j > k+1$.
-Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields
+	Redoing the calculation at the end of the proof of \cref{general-width} then yields
 	\begin{align*}
@@
 	2
-       + 4\left\lfloor\frac{k+1}{2}\right\rfloor
-       + 2\left\lceil\frac{k+1}{2}\right\rceil
+	+ 4\left\lfloor\tfrac{k+1}{2}\right\rfloor
+	+ 2\left\lceil\tfrac{k+1}{2}\right\rceil
 	\right)
-       & =
+	 =
 	d\ell\left(
-         2 + 2(k+1) + 2\left\lfloor\frac{k+1}{2}\right\rfloor
-       \right) \\
-       & \le
+	2 + 2(k+1) + 2\left\lfloor\tfrac{k+1}{2}\right\rfloor
+	\right) 
+	 \le
 	d\ell(3k+5)
@@
 	\end{align*}
-With this change, the layered width of the partition given by \cref{modern-major-general} becomes $(18k+30)(k+1)=18k^2+48k+30$.  This establishes \cref{k-planar}.
+	With this change, the layered width of the partition given by \cref{ShortcutPartition} becomes $(18k+30)(k+1)=18k^2+48k+30$.  
+	The result follows from \cref{PartitionProduct}.
+	%This establishes \cref{k-planar}.
 	% Note to self-and-others. In our original SODA submission we had 18k^2+30k here because we forgot that the layer compression step has compress k+1 consecutive layers, not just k consecutive layers.
-
-Exactly the same approach using \cref{SurfacePartition}(b) instead of \cref{PlanarPartition}(b) establishes
-\cref{g-k-planar} for $(g,k)$-planar graphs.
+\end{proof}
 
@@
 
-This section shows that the constants in \cref{modern-major-general} can be further improved in the important case of $1$-planar graphs.
+In the important special case of 1-planar graphs we obtain better constants and an additional property (planarity) of $H$.
+
+\begin{thm}
+\label{1-planar}
+Every 1-planar graph is a subgraph of $H\boxtimes P\boxtimes K_{30}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+\end{thm}
+
 Let $G$ be an edge-maximal 1-plane multigraph with no two parallel edges on the boundary of a single face.
@@
 We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.
-We assume that no two edges incident to a common vertex cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossing on a particular edge from $k$ to $2(k-1)$.}
+We assume that no two edges incident to a common vertex cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}
 
@@
 
-  % Note that $M$ is a Mercedes graph
-
 	Observe that each of the faces $F_1,\ldots,F_m$ is contained in a single internal face of $M$.
@@
 	The planarity of $H$ comes from two properties:
-  \begin{enumerate}
+	\begin{compactenum}
 		\item $G/\mathcal{P}$ and $G^+/\mathcal{P^+}$ are isomorphic, where $G^+$ is the triangulation obtained by adding dummy vertices at each crossing in $G$ and $\mathcal{P}^+$ is the partition we obtain by adding a dummy vertex $z$ to $\overline{Y}^+$ if $\overline{Y}^+$ contains an edge $vw$ that contains $z$ in its interior.
@@
 		\item $G^+[\overline{Y}^+-F]$ is connected.
-To see why this is so, first observe that $\overline{Y}-F$ is connected, and then observe that every vertex of $\overline{Y}^+$ is either a vertex of $\overline{Y}$ or adjacent to a vertex of $\overline{Y}$.
-  \end{enumerate}
+To see this, first observe that $\overline{Y}-F$ is connected, and then observe that every vertex of $\overline{Y}^+$ is either a vertex of $\overline{Y}$ or adjacent to a vertex of $\overline{Y}$.
+	\end{compactenum}
 	Since $G^+$ is planar, the second point implies that $H=G^+/\mathcal{P}$ is planar.
@@
 Since $|L_j\cap S_x|\le 15$ for every integer $j\ge 0$, $|L_i'\cap S_x|\le 30$ for every integer $i\ge 0$ and every $x\in V(H)$.
+The result follows from \cref{PartitionProduct}.
 \end{proof}
 
-% Exactly the same argument used to establish \cref{cor-k-planar} as a consequence of \cref{k-planar} establishes the following result as a consequence of \cref{1-planar}:
-%
-% \begin{cor}
-% \label{cor-1-planar}
-% Let $G$ be a $1$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
-Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle L_0',L_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \cref{1-planar} and such that $L_i' = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/2}=i\}$ for each integer $i\ge 0$.
-% \end{cor}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{\boldmath  $(g,k)$-Planar Graphs}
-%\label{sec-g-k-planar}
-%
-%We conclude this section with a small extension of \cref{k-planar} that will be useful in \cref{sec-g-k-planar}.
-%
-%\begin{cor}
-%\label{cor-k-planar}
-%Let $G$ be a $k$-plane graph, let $r_0$ be an arbitrary vertex of $G$, let $G^+$ be a graph obtained by adding a dummy vertex at each crossing in $G$ and then arbitrarily adding non-crossing edges to obtain a triangulation.
-Then $G$ has a layered $H$-partition $(\mathcal{L}=\langle L_0',L_1',\ldots\rangle,\mathcal{P})$ satisfying the conditions of \cref{k-planar} and such that $L_i' := \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\}$ for each integer $i\ge 0$.
-%\end{cor}
-%
-%\begin{proof}
-%  Without loss of generality, assume that $r_0$ is on the outer face of $G^+$ and let $v_1$ and $v_2$ be the other two vertices of the outer face of $G^+$.
-%  The proof of \cref{k-planar} begins with the graph $G^+$ and then creates a graph $G^{++}$ consisting of a $K_4$ containing $G^+$ in one of its inner faces, $F_0$, and six edges required to complete $G^{++}$ to a triangulation.
-Instead, modify $G^{++}$ by subdividing the edges joining $F_0$ to $v_1$ and $v_2$, as shown in \cref{wrapper}(b).
-In this modified version of $G^{++}$, we have $\dist_{G^++}(r,v)=\dist_{T}(r,v)=\dist_{G^+}(r_0,v)+2$
-%  which leads to the layering $\langle L_0,L_1,\ldots\rangle$ of $G^+$ where $L_i=\{v\in V(G^+): \dist_{T}(r,v) = i\}$.
-%
-%  The rest of proof remains unchanged until the point where we derive the layering $\mathcal{L}=\langle L_0',L_1',\ldots\rangle$ of $G$.  At this point, we set
-%  \[
-%    L_i' := \{v\in V(G): \floor{\dist_{T}(r,v)-2}=i\}
-%         = \{v\in V(G): \floor{\dist_{G^+}(r_0,v)/(k+1)}=i\} \enspace ,
-%  \]
-%  for all integers $i\ge 0$.  This layering satisfies the conditions of the corollary.
-%\end{proof}
-%
-%
-%
-%
-% Again, in the case $k=1$ we obtain better constants.
-%
-% \begin{thm}
-% \label{g-1-planar}
-% Every $(g,1)$-planar graph has a layered $H$-partition of layered width at most $\max\{48g, 30\}$ in which $H$ has treewidth at most $4$.
-% \end{thm}
-%
-% The proofs of these results make use of the following lemma due to \citet[Lemma~21]{dujmovic.joret.ea:planar}:
-%
-% \begin{lem}[\citep{dujmovic.joret.ea:planar}]
-% \label{make-planar}
-% Let $G^+$ be a $(\Sigma,0)$-plane graph where $\Sigma$ is a surface with Euler genus $g$.
-% For each $r\in V(G^+)$ there is a vertex set $Z\subseteq V(G^+)$ such that $G^+-V(Z)$ is planar and there is a connected planar graph $G^*$ containing  $G^+-Z$ as a subgraph with a vertex $r^*\in V(G^*)$ such that, for every integer $i\ge 0$,
-% \[  \{v\in V(G^*): \dist_{G^*}(r^*,v) = i\} \supseteq \{v\in V(G^+): \dist_{G^+}(r,v) = i\} \]
-% and
-% \[ |\{v\in Z: \dist_{G^+}(r,v) = i\}| \le 2g \enspace . \]
-% \end{lem}
-%
-% \begin{proof}[Proof of \cref{g-k-planar,g-1-planar}]
-%   Let $G$ be an $(\Sigma,k)$-plane graph where $\Sigma$ is a surface of Euler genus $g$.
-Let $G^+$ be the $(\Sigma,0)$-plane graph obtained by adding a dummy vertex at each crossing in $G$.
-Let $r_0$ be any vertex in $V(G)$, let $T$ be BFS spanning tree of $G^+$ rooted at $r_0$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting layering of $G^+$.
-%
-%   Apply \cref{make-planar} to $G^{+}$ to obtain the set $Z\subseteq V(G^{+})$ and the planar graph $G^*$ with special vertex $r^*$.
-For each edge $vw\in E(G)$, let $W_{vw}$ be the corresponding walk in $G^+$.
-The set $Z$ contains vertices in $V(G)$ as well as dummy vertices in $V(G^+)-V(G)$.
-Create a new set $Z'\subseteq V(G)$ from $Z$ by replacing each dummy vertex $z\in Z\cap (V(G^+)\setminus V(G))$ with one endpoint from each of the edges that cross $vw,xy\in E(G)$ that cross at $Z$.
-In this way, each vertex $z\in Z\cap V(G^+)\setminus V(G)$ can be charged for at most two vertices in $v,x\in Z'$ and this happens only if $(z,vw,xy)$ is a crossing in $G$.
-%   Since $|Z\cap L_i| \le 2g$, the same argument used in the proof of \cref{width-of-g-plus} shows that $|Z'\cap L_i|\le 2g\times 2\times (2k+1) = 4g(2k+1)$ for each integer $i\ge 0$.
-%
-%   % For every edge $vw\in E(G)$, $W_{vw}$ has length at most $k+1$.
-Therefore $\dist_{G^+}(v,w)\le k+1$ for each $v\in E(G)$.
-Since $W_{vw}\subseteq G^+-Z$ for each $vw\in E(G-Z')$,  $\dist_{G^+-Z}(v,w)\le k+1$ for each $vw\in E(G-Z')$.
-%
-%   Next, observe that any plane embedding of $G^*\supseteq G^+-Z$ gives a $k$-plane embedding of $G-Z'$.
-Indeed, for every edge $vw\in E(G-Z')$, $G^+-Z$ contains a walk $W_{vw}$ that contains at most $k$ dummy vertices of $G^+-Z$.
-Therefore, replacing edge $vw\in E(G-Z')$ with the corresponding walk in $G^*$ gives a $k$-plane embedding of $G-Z'$.
-Call the resulting $k$-plane graph $G^*_0$.
-%
-%   Let $G^*_1$ be the supergraph of $G^*_0$ obtained by adding every vertex and edge of $G^*$ that is not contained in some edge of $G^*_0$.
-Finally, let $G^{*+}_1$ be the triangulation obtained by adding edges to $G^*$ so that $\dist_{G^*}(r^*,v)=\dist_{G^{*+}}(r^*,v)$ for every $v\in V(G^*)$.
-The fact that $G^*$ can be completed to a triangulation $G^{*+}$ while preserving distances to $r^*$ follows from the fact that any face $F$ of $G^*$ with four or more vertices contains two non-adjacent vertices $x$ and $y$ such that $|\dist_{G^*}(r^*,x)-\dist_{G^*}(r^*,y)|\le 1$.
-Indeed, if $z$ is a vertex of $F$ that minimises $\dist_{G^*}(r^*,z)$, then two neighbours $x$ and $y$ of $z$ have this property.
-Now, observe that $G^*_1$ is a $k$-plane graph and $G^{*+}_1$ is obtained by adding dummy vertices at crossings in $G^*_1$ and triangulating the resulting graph.
-%
-%   Apply \cref{cor-k-planar} or (in the case $k=1$) \cref{cor-1-planar} to $G^*_1$ using $G^{*+}_1$ and the distinguished vertex $r^*$ to obtain a layered $H$-partition of $G^*_1$.
-Since $G^*_1\supseteq G^*_0$ and $G^*_0$ is isomorphic to $G-Z'$, this gives a layered $H$-partition $(\mathcal{L}=\langle L_0',L_1',\ldots\rangle,\mathcal{P})$ of $G-Z'$ that satisfies the conditions of \cref{cor-k-planar} or \cref{cor-1-planar}, as appropriate.
-In particular, for every integer $i\ge 0$,
-%   \begin{align*}
-%     L_i'& = \{v\in V(G-Z'):\floor{\dist_{G^*}(r^*,v)/(k+1)} = i\} \\
-%         & = \{v\in V(G-Z'):\floor{\dist_{G^+}(r_0,v)/(k+1)} = i\} \enspace .
-%   \end{align*}
-%   It follows that $|L_i'\cap Z'|\le 4g(2k+1)(k+1) = 8gk^2+12gk+4g$.
-%
-%   We now turn $\mathcal{P}$, the $H$-partition of $G-Z'$, into a partition $\mathcal{P}'=\mathcal{P}\cup \{Z'\}$ of $G$ by placing all vertices of $Z'$ into one new common part.
-Let $H'=G/ \mathcal{P}'$.
-Then the treewidth of $H'$ is at most one more than the treewidth of $H$.
-%
-%   % In the worst-case, the vertex of $H'=G/ \mathcal{P}'$ corresponding to $V(G)\cap Z'$ is adjacent to every vertex of $H=(G-Z')/\mathcal{P}$, so the treewidth of $H'$ is at most one more than the treewidth of $H$.
-%
-%   Summarizing, we obtain an $H'$-partition $(\mathcal{L},\mathcal{P}')$ of $G$ where $H'$ has treewidth
-%   \[
-%       \tw(H')
-%         \le \begin{cases}
-%               4 & \text{if $k=1$} \\
-%               k^3/6 + 3k^2/2 + 13k/3+5 & \text{otherwise}
-%             \end{cases}
-%   \]
-%   and, for each $P\in\mathcal{P}'$,
-%   \[
-%     |L_i' \cap P| \le \begin{cases} 8gk^2+12gk+4g & \text{if $P=Z'$} \\
-%                                     30 & \text{if $P\neq Z'$ and $k=1$} \\
-%                                     24k^2+60k+36 & \text{otherwise.}
-%                       \end{cases} \qedhere
-%   \]
-% \end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Applications}
-\label{Applications}
-
-Here we discuss some of the consequences of our main theorems for $k$-planar and $(g,k)$-planar graphs.
-
-%%%%%%%%%%%%%%%%%%%
-\subsection{Queue Layouts}
-
-For an integer $k\geq 0$, a \textit{$k$-queue layout} of a graph $G$ consists of a linear ordering $\preceq$ of $V(G)$ and a partition $\{E_1,E_2,\dots,E_k\}$ of $E(G)$, such that for $i\in\{1,2,\dots,k\}$, no two edges in $E_i$ are nested with respect to $\preceq$.
-That is, it is not the case that $v\prec x \prec y \prec w$ for edges $vw,xy\in E_i$.
-The \textit{queue-number} of a graph $G$, denoted by $\qn(G)$, is the minimum integer $k$ such that $G$ has a $k$-queue layout.
-Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
-\citet{dujmovic.joret.ea:planar} recently proved this conjecture using \cref{PlanarPartition} and the following lemma.
-Indeed, resolving this question was the motivation for the development of \cref{PlanarBasic,PlanarPartition}.
-
-\begin{lem}[\citep{dujmovic.joret.ea:planar}]
-\label{qn}
-If a graph $G$ has an $H$-partition of layered width $\ell$, then
-\[ \qn(G)\le 3 \ell \, \qn(H) + \floor{\tfrac{3}{2}\ell}\enspace.\]
-\end{lem}
-
-\citet{dujmovic.morin.ea:layered} proved that queue-number is bounded by a function of treewidth.
-The best known bound is due to \citet{wiechert:on}:
-
-\begin{thm}[\citep{wiechert:on}]
-\label{qn-tw}
-    For every graph $G$, $\qn(G)\le 2^{\tw(G)}-1$.
-\end{thm}
-
-\cref{qn,qn-tw,g-k-planar} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^4)}$.
-Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of
-$O(g^{k+2})$ using \cref{SurfacePartition} and an ad-hoc method.
-Our result provides a better bound when $g>2^{k^3}$.
-In the case of 1-planar graphs we can improve further.
-\citet{ABGKP18} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
-Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$.
-\cref{qn,1-planar} then imply:
-
-\begin{prop}
-Every 1-planar graph has queue-number at most $3 \times 30 \times 5 + \floor{\tfrac{3}{2} \times 30} = 495$. \end{prop}
-
-%%%%%%%%%%%%%%%%%%%
-\subsection{Non-Repetitive Colouring}
-
-The next two applications are in the field of graph colouring. For our purposes, a \emph{$c$-colouring} of a graph $G$ is any function $\phi\colon V(G)\to C$, where $C$ is a set of size at most $c$.
-
-A $c$-colouring $\phi$ of $G$ is \emph{non-repetitive} if, for every path $v_1,\ldots,v_{2h}$ in $G$, there exists some $i\in\{1,\ldots,h\}$ such that $\phi(v_i)\neq\phi(v_{i+h})$.
-The \emph{non-repetitive chromatic number} $\pi(G)$ of $G$ is the minimum integer $c$ such that $G$ has a non-repetitive $c$-colouring.
-This concept was introduced by \citet{AGHR-RSA02} and has since been widely studied; see \citep{dujmovic.esperet.ea:planar} for more than 40 references.
-Up until recently the main open problem in the field has been whether planar graphs have bounded non-repetitive chromatic number, first asked by \citet{AGHR-RSA02}.
-\citet{dujmovic.esperet.ea:planar} recently solved this question using \cref{PlanarPartition} and the following lemma.
-
-%\citet[Corollaries~9 and 10]{dujmovic.esperet.ea:planar} prove the following result:
-
-\begin{lem}[\citep{dujmovic.esperet.ea:planar}]
-\label{non-repetitive}
-If a graph $G$ has an $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$, then $\pi(G)\le \ell\, 4^{t+1}$.
-\end{lem}
-
-\cref{non-repetitive,k-planar,1-planar,g-k-planar} immediately imply the following results:
-
-\begin{cor}\corlabel{first}
-  For every $k$-planar graph $G$,
-  %$\pi(G)\le 12(2k+3)(k+1)\times 4^{(k+2)(k+3)(k+4)/6}$.
-  $\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
-%  Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}}-1$.
-\end{cor}
-
-\begin{cor}
-  For every $1$-planar graph $G$, $\pi(G)\le 30\times 4^4=7680$.
-%  Every 1-planar graph has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
-\end{cor}
-
-\begin{cor}
-  For every $(g,k)$-planar graph $G$,
-  \[
-%    \pi(G)\le \max\{4g(3k+1),12(2k+3)(k+1)\}\times 4^{(k+2)(k+3)(k+4)/6}
-    \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+5}{4}}.
-     \]
-%  \label{g-k-planar} Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most
-% $\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}$.
-\end{cor}
-
-% DW: I have commented out the result for $(g,1)$-planar graph.
-%\begin{cor}
-%  For every $g\ge 1$ and every $(g,1)$-planar graph $G$, $\pi(G)\le 48g\times 256= 12288g$.
-%\end{cor}
-
-Prior to the current work, the strongest upper bound on the non-repetitive chromatic number of $n$-vertex  $k$-planar graphs was $O(k\log n)$ \cite{dujmovic.morin.ea:layered}.
-
-%%%%%%%%%%%%%%%%%%%
-\subsection{Centered Colourings}
-\seclabel{centered-colourings}
-
-A $c$-colouring $\phi$ of $G$ is \emph{$p$-centered} if, for every connected subgraph $X\subseteq G$, $|\{\phi(v):v\in V(X)\}| > p$ or there exists some $v\in V(X)$ such that $\phi(v)\neq \phi(w)$ for every $w\in V(X)\setminus\{v\}$.
-In words, either $X$ receives more than $p$ colours or some vertex in $X$ receives a unique colour.
-Let $\chi_p(G)$ be the minimum integer $c$ such that $G$ has a $p$-centered $c$-colouring.
-Centered colourings are important since they characterise classes of bounded expansion, which is a key concept in the sparsity theory of \citet{Sparsity}.
-
-\citet{pilipczuk.siebertz:polynomial-soda} and \citet{debski.felsner.ea:improved} use \cref{thm:ps} and \cref{PlanarPartition}.\ref{subthm:three-tree}, respectively, to show that, when $G$ is planar, or of bounded genus, $\chi_p(G)$ is polynomial in $p$.
-We make use of the following lemma due to \citet[Lemma~15]{pilipczuk.siebertz:polynomial-arxiv}.
-
-\begin{lem}[\citep{pilipczuk.siebertz:polynomial-soda}]
-\label{p-centered-treewidth}
-  Every graph $H$ of treewidth at most $t$ has $\chi_p(H)\leq \binom{p+t}{t}$.
-\end{lem}
-
-The following lemma is implicitly in the work of \citet[Proof of Theorem 2.1]{debski.felsner.ea:improved}. We include the proof for completeness.
-
-% \note{DW}{can we replace by a result of \citet{micek:personal}?}
-% \note{PM}{I didn't see a statement of this exact result in the last draft Piotr sent.  Though their Theorem~7 is this proof for the special case of planar graphs.}
-
-\begin{lem}[\citep{debski.felsner.ea:improved}]
-\label{p-centered}
-Every graph $G$ that has an $H$-partition of layered width at most $\ell$ has $\chi_p(G)\le \ell (p+1)\, \chi_p(H)$.
-\end{lem}
-
-\begin{proof}
-   Let $(\mathcal{L}=\langle L_0,L_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ be an $H$-partition of $G$ having layered width at most $\ell$.
-   Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.
-For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in L_i\cap B_x$ such that:
-   \begin{compactenum}
-     \item $\alpha(v)$ is unique among $\{\phi(w): w\in L_i\cap B_x\}$, which is possible
-     since $|L_i\cap B_x|\le \ell$,
-     \item $\beta(v)= i\bmod (p+1)$, and
-     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
-   \end{compactenum}
- To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
-
- First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$.
-Since $X$ is connected, $X$ contains a path from $v$ to $w$.
-By the definition of layering, this path contains at least one vertex from $L_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$.
-Therefore, $|\{\beta(v'):v'\in V(X)\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
-
- Otherwise, $V(X)\subseteq L_{i}\cup\cdots\cup L_{i+s}$ for some $s<p$.
-Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.
-If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in V(X)\}|> p$ so $|\{\phi(v):v\in V(X)\}|> p$ and we are done.
-Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
- For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
-Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.
-Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
- \end{proof}
-
-\cref{p-centered,p-centered-treewidth,k-planar,1-planar,g-k-planar} immediately imply the following results, for every $p\ge 2$:
-
-\begin{cor}
-\label{pck}
-For every $k$-planar graph $G$,
-%for some integer $t\leq \binom{k+4}{3}-1$,
-%$$\chi_p(G)\le (18k^2+30k)(p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
-\[ \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}\enspace. \]
-%\note{DW}{I previously wrote ``for some integer $t\leq \binom{k+4}{3}-1$'' thinking that $\binom{p+t}{t}$ might not be monotone with $t$, but it is because of the $+t$.}
-%Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
-%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
-\end{cor}
-
-\begin{cor}
-\label{pc1}
-For every $1$-planar graph $G$,
-\[ \chi_p(G)\le 5 (p+3)(p+2)(p+1)^2\enspace. \]
-%120 (p+1) \binom{p+3}{3}
-%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
-%Every 1-planar graph has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
-\end{cor}
-
-\begin{cor}
-\label{pcgk}
-For every $(g,k)$-planar graph $G$,
-%for some integer $t\leq \binom{k+5}{4}-1$,
-%$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
-\[ \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}\enspace. \]
-%O(gk^2)\times p^{O(k^3)}$.
-%%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
-%Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most
-%$\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
-\end{cor}
-
-%\begin{cor}
-%\label{pcg1}
-%For every $(g,1)$-planar graph $G$, $g\ge 1$, $\chi_p(G)\le O(gp^4)$.
-%\note{DW}{update}
-%\end{cor}
-
-%Prior to the current work, the strongest upper bounds on the number of colours required for $p$-centered colourings of $(g,k)$-planar graphs used the fact that $(g,k)$-planar graphs have layered treewidth $O(gk)$ \cite{dujmovic.eppstein.ea:structure} and therefore have bounded (indeed, linear) expansion \cite{dujmovic.morin.ea:layered}.
-Using a result of \citet{zhu:colouring} this leads to bounds on $\chi_p(G)$ that are doubly-exponential in $gk$.
-
-Prior to the current work, the strongest known upper bounds on the $p$-centered chromatic number of $(g,k)$-planar graphs $G$ were doubly-exponential in $p$, as we now explain.
-\citet{dujmovic.eppstein.ea:structure} proved that $G$ has layered treewidth $(4g+6)(k+1)$.
-Van den Heuvel and Wood~\citep{vdHW17} showed that this implies that $G$  has $r$-strong colouring number at most $(4g + 6)(k + 1)(2r + 1)$.
-By a result of \citet{zhu:colouring}, $G$ has $r$-weak colouring number at most $( (4g + 6)(k + 1)(2r + 1) )^r$, which by another result of  \citet{zhu:colouring} implies that $G$ has  $p$-centered chromatic number at most $( (4g+6)(k+1)(2^{p-1} + 1) )^{2^{p-2}}$.
-\cref{pc1,pck,pcgk} are substantial improvements over these results, providing bounds on $\chi_p(G)$ that are polynomial in $p$ for fixed $g$ and $k$.
-%\cref{pc1,pck,pcgk,pcg1} are substantial improvements over these bounds.
-
-%%%%%%%%%%%%%%%%
-% \subsection{Low Treewidth Colouring}
-%
-% \note{DW}{The results in this section are trivial because $(g,k)$-planar graphs have bounded layered treewidth. Probably delete this section.}
-%
-% \citet{DDOSRSV04} proved the conjecture of Robin Thomas that every graph in a proper minor-closed class has an edge 2-colouring so that each monochromatic subgraph has bounded treewidth, and more generally, that for fixed $c\geq 2$, every such graph can be edge $c$-coloured such that the union of any $c-1$ colour classes has bounded treewidth.
-They also showed analogous vertex-colouring results.
-\citet{dujmovic.joret.ea:planar} gave simple proofs of these results via layered partitions.
-The following lemma is implicit in \citep{dujmovic.joret.ea:planar}.
-%
-% \begin{lem}[\citep{dujmovic.joret.ea:planar}]
-% \label{LowTreewidth}
-% Let $G$ be a graph that has an $H$-partition of layered width $\ell$ such that $H$ has treewidth at most $k$.
-Fix $c\geq 2$.
-Then (a) $G$ has an edge $c$-colouring such that the spanning subgraph of $G$ consisting of the union of any $c-1$ colour classes has treewidth at most $\ell(k+1)(2c-1)$, and (b) $G$ has a vertex $c$-colouring such that the subgraph of $G$ induced by the union of any $c-1$ colour classes has treewidth at most $\ell(k+1)(2c-1)$.
-% \end{lem}
-%
-% One feature of this result compared to that of \citet{DDOSRSV04}  is that it applies for non-minor-closed classes, such as $k$-planar graphs. In particular, \cref{LowTreewidth,g-k-planar} imply:
-%
-% \begin{prop}[\citep{dujmovic.joret.ea:planar}]
-% For every $(g,k)$-planar graph $G$ and integer $c\geq 2$,
-% (a) $G$ has an edge $c$-colouring such that the spanning subgraph of $G$ consisting of the union of any $c-1$ colour classes has treewidth $O(cgk^5 )$, and (b) $G$ has a vertex $c$-colouring such that the subgraph of $G$ induced by the union of any $c-1$ colour classes has treewidth $O(cgk^5 )$.
-% \end{prop}
-
-%%%%%%%%%%%%%%%%
-%\subsection{Graph Encodings}
-%
-%\textcolor{red}{Whether we include this section will depend on whether we get a copy of \citep{BGP} prior to SODA submission. They are submitting to SODA too (as is \citep{micek:personal})}
-
-%%%%%%%%%
-\section{Rough Characterisation}
+\subsection{\boldmath $(g,k)$-Planar Graphs}
+
+The definition of $k$-planar graphs naturally generalises for other surfaces.
+A graph $G$ drawn on a surface $\Sigma$ is $(\Sigma,k)$-plane if every edge of $G$ is involved in at most $k$ crossings.
+A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
+\cref{AddDummy} immediately generalises as follows:
+
+\begin{obs}
+\label{gAddDummy}
+Every $(g,k)$-planar graph $G$ is a subgraph of $G_0^\PP$ for some graph $G_0$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G_0$.
+Moreover, $V(G) \subseteq V(G_0)$ and for every edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most $k+1$, such that every internal vertex in $P$ has degree at most $4$ in $G_0$.
+\end{obs}
+
+Theorems~\ref{ShortcutProduct}  and \ref{GenusProduct}(b) imply a product structure theorem for $(g,k)$-planar graphs.
+The obtained bounds are improved by the following result, which is proved using exactly the same approach used in the proof of \cref{kPlanarProduct} (applying \cref{GenusProduct}(b) instead of \cref{PlanarProduct}(b)).
+We omit repeating the details.
+
+\begin{thm}
+\label{gkPlanarProduct}
+Every $(g,k)$-planar graph is a subgraph of $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with $\tw(H) \leq \binom{k+5}{4}-1$, where $\ell:=\max\{2g,3\}\cdot(6k^2+16k+10)$.
+\end{thm}
+
+Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs (or any of the other classes presented in \cref{Examples}) was in terms of layered treewidth, which we now define.
+A \emph{layered tree decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree decomposition $\mathcal{T}$ of $G$.
+The layered width of $(\mathcal{L},\mathcal{T})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{T}\}$.
+The \emph{layered treewidth} of $G$ is the minimum layered width of any layered tree decomposition of $G$.
+\citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph.
+\citet{dujmovic.eppstein.ea:structure} show that every $k$-planar graph has layered treewidth at most $6(k+1)$, and more generally that every $(g,k)$-planar graph has layered treewidth at most $(4g+6)(k+1)$.
+It follows from this result that $(g,k)$-planar graphs have treewidth $O(\sqrt{(g+1)(k+1)n})$ and thus have balanced separators of the same order, which can also be concluded from the work of \citet{FP08}.
+In related work, \citet{grigoriev.bodlaender:algorithms} used structural results to obtain approximation algorithms for $(g,k)$-planar graphs, and \citet{PachToth97} determined the maximum number of edges in a $k$-planar graph (up to a constant factor).
+
+If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
+In particular, if $\mathcal{P}=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{dujmovic.joret.ea:planar}.
+Therefore, any property that holds for graphs of bounded layered treewidth also holds for $G$.
+What sets layered partitions apart from layered treewidth is that they lead to constant upper bounds on the queue-number  and non-repetitive chromatic number, whereas for both these parameters, the best known upper bound obtainable via layered treewidth is $O(\log n)$; see \cref{Applications}.
+
+\subsection{Rough Characterisation}
 \label{Characterisation}
 
-This section presents a rough characterisation of $(g,k)$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $(g,k)$-planar.
-First note the following easy result (just add a degree-4 vertex at each crossing point of $G$, as in the proof of \cref{AddDummy,gAddDummy}).
-
-\begin{lem}
-\label{gAddDummyDegree}
-For every $(g,k)$-planar graph $G$ there is a graph $G_0$ of Euler
-genus at most $g$, such that $V(G) \subseteq V(G_0)$ and for every
-edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most
-$k+1$, such that every internal vertex in $P$ has degree at most $4$
-in $G_0$.
-\end{lem}
-
-The following converse of \cref{gAddDummyDegree} provides a sufficient condition for a graph to be $(g,k)$-planar that is slightly more restrictive than being obtained from a $(k,d)$-shortcut system of a $(g,k)$-planar graph.
-\notex{DW}{Is this sentence clear enough?}\notex{PM}{Sure.}
+\cref{gAddDummy} shows that $(g,k)$-planar graphs can be obtained by a shortcut system applied to a graph of Euler genus $g$, where internal vertices on the paths have bounded degree.
+This observation and the following converse result together provide a rough characterisation of $(g,k)$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $(g,k)$-planar.
 
@@
 
-Together \cref{DrawG,gAddDummyDegree} provide a rough characterisation of $(g,k)$-planar graphs.
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Applications}
+\label{Applications}
+
+Here we discuss some of the consequences of the above theorems for $k$-planar and $(g,k)$-planar graphs.
+
+%%%%%%%%%%%%%%%%%%%
+\subsection{Queue Layouts}
+
+For an integer $k\geq 0$, a \textit{$k$-queue layout} of a graph $G$ consists of a linear ordering $\preceq$ of $V(G)$ and a partition $\{E_1,E_2,\dots,E_k\}$ of $E(G)$, such that for $i\in\{1,2,\dots,k\}$, no two edges in $E_i$ are nested with respect to $\preceq$.
+That is, it is not the case that $v\prec x \prec y \prec w$ for edges $vw,xy\in E_i$.
+The \textit{queue-number} of a graph $G$, denoted by $\qn(G)$, is the minimum integer $k$ such that $G$ has a $k$-queue layout.
+Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
+\citet{dujmovic.joret.ea:planar} recently proved this conjecture using \cref{PlanarProduct} and the following lemma.
+Indeed, resolving this question was the motivation for the development of \cref{PlanarProduct}.
+
+\begin{lem}[\citep{dujmovic.joret.ea:planar}]
+\label{qn}
+If $G\subseteq H \boxtimes P \boxtimes K_\ell$ then 
+$\qn(G) \leq  3 \ell \, \qn(H) + \floor{\tfrac{3}{2}\ell}
+\leq 3 \ell \, 2^{\tw(H)}  - \ceil{\tfrac{3}{2}\ell}$. 
+\end{lem}
+
+\cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^4)}$.
+Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of
+$O(g^{k+2})$ using \cref{GenusProduct} and an ad-hoc method.
+Our result provides a better bound when $g>2^{k^3}$.
+In the case of 1-planar graphs we can improve further.
+\citet{ABGKP18} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
+Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$.
+\cref{qn,1-planar} then imply:
+
+\begin{cor}
+Every 1-planar graph has queue-number at most $3 \times 30 \times 5 + \floor{\tfrac{3}{2} \times 30} = 495$. 
+\end{cor}
+
+%%%%%%%%%%%%%%%%%%%
+\subsection{Non-Repetitive Colouring}
+
+The next two applications are in the field of graph colouring. For our purposes, a \emph{$c$-colouring} of a graph $G$ is any function $\phi\colon V(G)\to C$, where $C$ is a set of size at most $c$.
+A $c$-colouring $\phi$ of $G$ is \emph{non-repetitive} if, for every path $v_1,\ldots,v_{2h}$ in $G$, there exists $i\in\{1,\ldots,h\}$ such that $\phi(v_i)\neq\phi(v_{i+h})$.
+The \emph{non-repetitive chromatic number} $\pi(G)$ of $G$ is the minimum integer $c$ such that $G$ has a non-repetitive $c$-colouring.
+This concept, introduced by \citet{AGHR-RSA02}, has since been widely studied; see \citep{dujmovic.esperet.ea:planar} for more than 40 references.
+Up until recently the main open problem in the field has been whether planar graphs have bounded non-repetitive chromatic number, first asked by \citet{AGHR-RSA02}.
+\citet{dujmovic.esperet.ea:planar} recently solved this question using \cref{PlanarProduct} and the following lemma.
+
+\begin{lem}[\citep{dujmovic.esperet.ea:planar}]
+\label{non-repetitive}
+If $G\subseteq H\boxtimes P \boxtimes K_\ell$ then $\pi(G)\le \ell\, 4^{\tw(H)+1}$.
+\end{lem}
+
+\cref{non-repetitive,kPlanarProduct,1-planar,gkPlanarProduct} imply the following result:
+
+\pagebreak
+
+\begin{cor}\quad
+\begin{compactitem}
+\item	For every $1$-planar graph $G$, $\pi(G)\le 30\times 4^4=7680$.
+\item For every $k$-planar graph $G$,
+	$\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
+\item	For every $(g,k)$-planar graph $G$,
+	$  \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+5}{4}}.$
+\end{compactitem}
+\end{cor}
+%\end{samepage}
+
+Prior to the current work, the strongest upper bound on the non-repetitive chromatic number of $n$-vertex  $k$-planar graphs was $O(k\log n)$ \cite{dujmovic.morin.ea:layered}.
+
+%%%%%%%%%%%%%%%%%%%
+\subsection{Centered Colourings}
+\seclabel{centered-colourings}
+
+A $c$-colouring $\phi$ of $G$ is \emph{$p$-centered} if, for every connected subgraph $X\subseteq G$, $|\{\phi(v):v\in V(X)\}| > p$ or there exists some $v\in V(X)$ such that $\phi(v)\neq \phi(w)$ for every $w\in V(X)\setminus\{v\}$.
+In words, either $X$ receives more than $p$ colours or some vertex in $X$ receives a unique colour.
+Let $\chi_p(G)$ be the minimum integer $c$ such that $G$ has a $p$-centered $c$-colouring.
+Centered colourings are important since they characterise classes of bounded expansion, which is a key concept in the sparsity theory of \citet{Sparsity}.
+
+\citet{pilipczuk.siebertz:polynomial-soda} and \citet{debski.felsner.ea:improved} use \cref{ps} and \cref{PlanarProduct}(b), respectively, to show that $\chi_p(G)$ is polynomial in $p$ when $G$ is planar or of bounded Euler genus.
+We use the following lemma due to \citet[Lemma~15]{pilipczuk.siebertz:polynomial-arxiv}.
+
+\begin{lem}[\citep{pilipczuk.siebertz:polynomial-arxiv}]
+\label{p-centered-treewidth}
+  Every graph $H$ of treewidth at most $t$ has $\chi_p(H)\leq \binom{p+t}{t}$.
+\end{lem}
+
+The following lemma is implicitly in the work of \citet[Proof of Theorem~2.1]{debski.felsner.ea:improved}. We include the proof for completeness.
+
+\begin{lem}[\citep{debski.felsner.ea:improved}]
+\label{p-centered}
+If $G\subseteq H\boxtimes P \boxtimes K_\ell$ then $\chi_p(G)\le \ell (p+1)\, \chi_p(H)$.
+\end{lem}
+
+\begin{proof}
+	By \cref{PartitionProduct}, $G$ has an $H$-partition $(\mathcal{L}=\langle L_0,L_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ with layered width at most $\ell$.
+Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.
+For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in L_i\cap B_x$ such that:
+   \begin{compactenum}
+     \item $\alpha(v)$ is unique among $\{\phi(w): w\in L_i\cap B_x\}$, which is possible
+     since $|L_i\cap B_x|\le \ell$,
+     \item $\beta(v)= i\bmod (p+1)$, and
+     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
+   \end{compactenum}
+ To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
+
+ First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$.
+Since $X$ is connected, $X$ contains a path from $v$ to $w$.
+By the definition of layering, this path contains at least one vertex from $L_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$.
+Therefore, $|\{\beta(v'):v'\in V(X)\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
+
+ Otherwise, $V(X)\subseteq L_{i}\cup\cdots\cup L_{i+s}$ for some $s<p$.
+Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.
+If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in V(X)\}|> p$ so $|\{\phi(v):v\in V(X)\}|> p$ and we are done.
+Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
+ For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
+Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.
+Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
+ \end{proof}
+
+\cref{p-centered,p-centered-treewidth,kPlanarProduct,1-planar,gkPlanarProduct} immediately imply the following results, for every $p\ge 2$:
+
+\begin{cor}\quad
+\begin{compactitem}
+\item For every $1$-planar graph $G$,\; $ \chi_p(G)\le 5 (p+3)(p+2)(p+1)^2$. 
+\item For every $k$-planar graph $G$,\; $\displaystyle \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}$. 
+\item For every $(g,k)$-planar graph $G$,\;
+$\displaystyle \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}$. 
+ \end{compactitem}
+\end{cor}
+
+Prior to the current work, the strongest known upper bounds on the $p$-centered chromatic number of $(g,k)$-planar graphs $G$ were doubly-exponential in $p$, as we now explain.
+\citet{dujmovic.eppstein.ea:structure} proved that $G$ has layered treewidth $(4g+6)(k+1)$.
+Van den Heuvel and Wood~\citep{vdHW17} showed that this implies that $G$  has $r$-strong colouring number at most $(4g + 6)(k + 1)(2r + 1)$.
+By a result of \citet{zhu:colouring}, $G$ has $r$-weak colouring number at most $( (4g + 6)(k + 1)(2r + 1) )^r$, which by another result of  \citet{zhu:colouring} implies that $G$ has  $p$-centered chromatic number at most $( (4g+6)(k+1)(2^{p-1} + 1) )^{2^{p-2}}$.
+The above results are substantial improvements, providing bounds on $\chi_p(G)$ that are polynomial in $p$ for fixed $g$ and $k$.
 
@@
 
-This section describes several examples of graph classes that can be obtained from a shortcut system applied to graphs of bounded Euler genus.
-
-%This section describes several graph classes that are $k$-planar or $(g,k)$-planar. The first two examples are well known, the second two are new.
-
-%We conclude by pointing out three other graph classes, namely $(g,d)$-map graphs, $(g,\delta)$-string graphs, and $k$-nearest-neighbour graphs that are $(g,k)$-planar for $k=O(d^2)$ and $k=\delta^2$, respectively.
-This implies that Theorems~\ref{k-planar}--\ref{g-k-planar} and Corollaries~\ref{cor:first}--\ref{cor:last} apply to $(g,d)$-map graphs, $(g,\delta)$-string graphs, and $k$-nearest-neighbour graphs.
+This section describes several examples of graph classes that can be obtained from a shortcut system typically applied to graphs of bounded Euler genus.
 
@@
 Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar.
-\cref{g-k-planar} then shows that map graphs admit bounded layered partitions, but we get much better bounds by constructing a shortcut system directly.
+\cref{gkPlanarProduct} then establishes a product structure theorem for map graphs, but we get much better bounds by constructing a shortcut system directly.
 The following lemma is reminiscent of the characterisation of $(g,d)$-map graphs in terms of the half-square of bipartite graphs \citep{CGP02,dujmovic.eppstein.ea:structure}.
-
-%Consider a bipartite graph $H$ with bipartition $\{A,B\}$.
-Define the \emph{half-square} graph $H^2[A]$ with vertex set $A$, where two vertices in $A$ are adjacent if and only if they have a common neighbour in $B$.
-\citet{qqq} proved that a graph $G$ is a $(g,d)$-map graph if and only if $G$ is isomorphic to $H^2[A]$ for some bipartite graph $H$ with Euler genus at most $g$ and bipartition $\{A,B\}$, where vertices in $B$ have maximum degree at most $d$.
-This was proved earlier in the case $g=0$ by \citep{CGP02}.
 
@@
 
-\cref{PlanarPartition}(b), \cref{SurfacePartition}(b) and \cref{modern-major-general,MapShortcut} imply:
+Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} 
+and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,p-centered-treewidth,non-repetitive} imply the following results.
+
+\begin{thm}
+\label{PlaneMapPartition}
+Every $d$-map graph $G$:
+	\begin{compactitem}
+		\item is a subgraph of $H \boxtimes P \boxtimes K_{21d(d-3)}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
+		\item has queue-number $\qn(G) < 32225\, d(d-3)$. 
+		\item has $p$-centered chromatic number $\chi_p(G) \leq 21d(d-3) (p+1)  \binom{p+9}{9}$, 
+		\item has non-repetitive chromatic number $ \pi(G) \leq 21 \cdot 4^{10} d(d-3)$.
+	\end{compactitem}
+\end{thm}
 
@@
 \label{MapPartition}
-Every $(g,d)$-map graph has an $H$-partition with layered width at most
-$7d(d-3)\, \max\{2g,3\}$ for some graph $H$ with treewidth at most $\binom{2+4}{4}-1=14$
-or treewidth at most $\binom{2+3}{3}-1=9$ if $g=0$.
-%
-% an $H$-partition with layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most $4$;
-%
-%  Let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\PP$ be a $(k,d)$-shortcut system for $G$.
-Then $G^\PP$ has a layered $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ with treewidth at most $\binom{k+t}{t}-1$.
-%
-\end{thm}
-
-\cref{MapPartition,non-repetitive} imply the first known constant upper bound on the
-non-repetitive chromatic number of map graphs:
-
-\begin{cor}
-Every $(g,d)$-map graph $G$ has non-repetitive chromatic number
-$\pi(G) \le  7 \cdot 4^{15} \, d(d-3)\, \max\{2g,3\}$, and if $g=0$ then
-$\pi(G) \le  21 \cdot 4^{10} \, d(d-3)$.
-%\label{non-repetitive} If a graph $G$ has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$, then $\pi(G)\le \ell\, 4^{t+1}$.
-\end{cor}
-
-\cref{p-centered,p-centered-treewidth,MapPartition} imply that the first known bounds on the $p$-centered chromatic number of map graphs that are polynomial in $p$.
-
-\begin{cor}
-Every $(g,d)$-map graph $G$ has $p$-centered chromatic number
-\[ \chi_p(G)\le 105\, d(d-3)\, \max\{2g,3\}\, (p+1) \tbinom{p+14}{14}\enspace, \]
-and every plane $d$-map graph $G$ has $p$-centered chromatic number
-\[ \chi_p(G)\le 210\,d(d-3)\, (p+1) \tbinom{p+9}{9}\enspace. \]
-\end{cor}
-
-\cref{MapPartition,qn,qn-tw} imply the best known bounds on the queue-number of map graphs:
-
-\begin{cor}
-Every $(g,d)$-map graph has queue-number at most $ 21\cdot 2^{14} \, d(d-3)\, \max\{2g,3\} $,
-and every plane $d$-map graph  has queue-number at most
-$ 63 \cdot 2^9\, d(d-3)$.
-%If a graph $G$ has a layered $H$-partition of layered width $\ell$, then
-%$$\qn(G)\le 3 \ell ( 2^t- 1/2) \enspace.$$
-\end{cor}
-
-%an upper bound on the queue-number of $(g,d)$-map graphs that is better than the previous bound obtained by \citet{dujmovic.joret.ea:planar} when $g\gg d$.
-For example, since every $(g,4)$-map graph is $(g,1)$-planar, \cref{qn,1-planar,g-1-planar} imply that every 4-map graph has queue-number at most $495$,  \note{PM}{This part is no longer true/tight: and every $(g,4)$-map graph has queue-number at most $2232 g$.}
-
-%that every $(g,d)$-map graph $G$ has queue-number  $$\qn(G)\le \max\{\tfrac32 gd^4 + O(gd^3), \, \tfrac{9}{2}d^4 +O(d^3) \} \,  2^{d^6/384 + O(d^5)} \enspace.$$ previous bounds obtained by \citet{dujmovic.joret.ea:planar} previously proved the bound of  $O(g^{d^2/4+2})$.
-Our result provides a better bound when $g\gg d$.
-For example, with $d=4$, every $(g,4)$-
+For integers $g\geq 0$ and $d\geq 4$, if $\ell:=  7d(d-3)\, \max\{2g,3\}$ then every $(g,d)$-map graph $G$:
+\begin{compactitem}
+\item is a subgraph of $H \boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 14$,
+\item has queue-number $\qn(G) <  49151\, \ell $. 
+\item has $p$-centered chromatic number $\chi_p(G) \leq \ell (p+1)\,  \binom{p+14}{14}$, 
+\item has non-repetitive chromatic number $ \pi(G) \leq 4^{15}\,\ell $.	
+\end{compactitem}
+\end{thm}
+
+These results give 
+the first constant upper bound on the non-repetitive chromatic number of map graphs,  
+the first polynomial bounds on the $p$-centered chromatic number of map graphs, and
+the best known bounds on the queue-number of map graphs.
 
@@
 A \emph{$(g,\delta)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
-
-%For a surface $\Sigma$ and a graph $G$, a \emph{$\Sigma$-string representation} $f$ of $G$ is obtained by representing each vertex $v\in V(G)$ as a curve $f(v)$ in $\Sigma$ so that, for any pair of vertices $v,w\in V(G)$, the curves $f(v)$ and $f(w)$ intersect if and only if $vw\in E(G)$.
-A $\Sigma$-string representation is called a $(\Sigma,\delta)$-string representation if each curve is in at most $\delta$ intersections with other curves.
-A graph $G$ is a \emph{$(g,\delta)$-string graph} if it has a $(\Sigma,\delta)$-string representation where $\Sigma$ is a surface of Euler genus $g$.
 
@@
 A similar proof to that of \cref{StringShortcut} shows that every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.
-
-%\begin{lem}
-%\label{StringGraph}
-%Every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.
-%\end{lem}
-%
-%\begin{proof}
-%Consider some $(\Sigma,\delta)$-string representation, $f$, of $G$.
-We will describe an embedding of $G$ into $\Sigma$ that is an $(\Sigma,k)$-plane graph.
-Each vertex $v\in V(G)$ is embedded at $f(v)(0)$.
-For each edge $vw\in E(G)$ there is an intersection point $p$ between $f(v)$ and $f(w)$, i.e., $f(v)(t_0)=f(w)(t_1)=p$ for some $t_0,t_1\in[0,1]$.
-Embed the edge $vw$ very close to the curve whose image is $\{f(v)(t): 0\le t\le t_1\}\cup \{ f(w)(t): 0 \le t\le t_2\}$.
-The embedding of any edge $vw\in E(G)$ is contained in $f(v)$ and $f(w)$.
-Therefore the embedding of $vw$ crosses at most $2\delta$ other curves $f(x_1),\ldots,f(x_d)$.
-Each curve $f(x_i)$ is used to embed at most $\delta$ edges incident to $x_i$.
-Therefore the embedding of any edge $vw\in E(G)$ crosses at most $2\delta^2$ other edges.
-%\end{proof}
-
-\cref{StringShortcut,modern-major-general}, \cref{PlanarPartition}(b) and \cref{SurfacePartition}(b) imply:
+Theorems~\ref{PlanarProduct}(b), \ref{ShortcutProduct}  and \ref{GenusProduct}(b) and \cref{StringShortcut,qn} imply:
 
@@
 \label{StringPartition}
-Every $(g,\delta)$-string graph has an $H$-partition of layered
-width at most $\max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$
-for some graph $H$ with treewidth at most $\binom{ \delta+5}{4}-1$ or treewidth at most $\binom{ \delta+4}{3}-1$ if $g=0$.
-%
-%Every $(g,\delta)$-string graph $G$ is a subgraph of $G_0^\PP$ for some graph $G_0$ with Euler genus at most $g$ and some $(\delta+1,\delta+1 )$-shortcut system $\PP$ for $G_0$.
-%
- % Let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\PP$ be a $(k,d)$-shortcut system for $G$.
-Then $G^\PP$ has a layered $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ with treewidth at most $\binom{k+t}{t}-1$.
-%
-% an $H$-partition with layered width at most $\max\{2g,3\}$ such that $H$ has treewidth at most $4$;
-\end{thm}
-
-% \snote{Deleted results on p-centered colouring since the bounded-degree results by \citet{debski.felsner.ea:improved} give better dependence on $\delta$.}
-%
-%
-% \cref{p-centered,p-centered-treewidth,StringPartition} imply the first known bounds on the $p$-centered chromatic number of map graphs that are polynomial in $p$.
-%
-% \begin{cor}
-% Every $(g,\delta)$-string graph has $p$-centered chromatic number
-% \[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1)
-% \tbinom{ \delta+5}{4} \binom{p+\binom{ \delta+5}{4}-1}{\binom{ \delta+5}{4}-1}\enspace, \]
-% and every plane $(0,\delta)$-string graph has $p$-centered chromatic number
-% \[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1)
-% \tbinom{ \delta+4}{3}\binom{p+\binom{ \delta+4}{3}-1}{\binom{ \delta+4}{3}-1}\enspace.\]
-% \end{cor}
-%\label{p-centered}
-%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
-
-\cref{StringPartition,qn,qn-tw} imply the following bounds on the queue-number of map graphs:
-
-\begin{cor}
-Every $(g,\delta)$-string graph has queue-number at most
-\[3 \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)  2^{\binom{ \delta+5}{4}-1} \enspace.\]
-and every plane $(0,\delta)$-map graph  has queue-number at most
-\[ 9 (\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)  2^{\binom{ \delta+4}{3}-1} \enspace.\]
-%If a graph $G$ has a layered $H$-partition of layered width $\ell$, then
-%$$\qn(G)\le 3 \ell ( 2^t- 1/2) \enspace.$$
-\end{cor}
-
-%A \emph{string graph} is the intersection graph of a set of curves in the plane with no three curves meeting at a single point \cite{PachToth-DCG02,SS-JCSS04,SSS-JCSS03,Krat-JCTB91,FP10,FP14}.
-For an integer $k\geq 2$, if each curve is in at most $k$ intersections with other curves, then the corresponding string graph is called a \emph{$k$-string graph}.
-A \emph{$(g,k)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
-
-% from previous paper: For all integers $g\geq 0$ and $k\geq 2$, every $(g,k)$-string graph has queue-number at most $2(40g+490)^{2k+1}$.
-
-\cref{non-repetitive,StringPartition} also give bounds on the non-repetitive chromatic number and the $p$-centered colouring number of $(g,\delta)$-string graphs, but the bounds are weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ and $\chi_p(G)\le p(64\delta)^2$ by results of \citet{DJKW16} and \citet{debski.felsner.ea:improved}, respectively.
+For integers $g\geq 0$ and $\delta\geq 2$, let $\ell:= \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$, 
+and $t:= \binom{ \delta+5}{4}-1$ if $g\geq 1$ and $t:= \binom{ \delta+4}{3}-1$ if $g=0$. 
+Then  every $(g,\delta)$-string graph: 
+\begin{compactitem}
+	\item is a subgraph of $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$,
+	\item has queue-number $\qn(G^k) \leq 3 \ell \, 2^t - \ceil{\tfrac{3}{2}\ell}$,
+\end{compactitem}
+\end{thm}
+
+Our results also give bounds on the non-repetitive chromatic number and the $p$-centered chromatic number of $(g,\delta)$-string graphs, but the bounds are weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ and $\chi_p(G)\le p(64\delta)^2$ by results of \citet{DJKW16} and \citet{debski.felsner.ea:improved}, respectively.
 
@@
 
-\cref{PlanarPartition}(b), \cref{SurfacePartition}(b), \cref{modern-major-general,PowerShortcut} imply the following result, which with \cref{PartitionProduct} implies \cref{kPowerBasic} in the introduction.
+\cref{ShortcutProduct,PowerShortcut} imply:
 
 \begin{thm}
-\label{PowerPartition}
-Fix $k\geq 1$. Let $G$ be a graph of Euler genus $g$ and maximum degree $\Delta$.
-Then $G^k$ has an $H$-partition of layered width at most $\max\{2g,3\} (2k^4+6k^2) \Delta^{k}$
-for some graph $H$ with treewidth at most $\binom{k+4}{4}-1$, and treewidth
-at most $\binom{k+3}{3}-1$ if $g=0$.
-\end{thm}
-
-\cref{DrawG,qn,p-centered,p-centered-treewidth,PowerPartition,non-repetitive} immediately imply the following result:
-
-% \noindent\note{PM}{We can probably do better now by using a $(k,(\Delta-1)^{k-1})$-shortcut system and \cref{modern-major-general}.} \note{DW}{An exponent of $k-1$ here is false for $k=2$.}
-
-% \note{PM}{TODO: Deal with the change caused by the change to layered-width in \cref{modern-major-general}}.
-
-\begin{cor}
-\label{PowerCor}
-For all integers $k\geq 1$ and $g\geq 0$, if $G$ is a graph with Euler genus $g$ and maximum degree $\Delta$, then:
+\label{PowerProduct}
+Let $G$ be a subgraph of $H\boxtimes P\boxtimes K_\ell$ with maximum degree $\Delta$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
+Then for every integer $k\geq 1$, the $k$-th power $G^k$ is a subgraph of $J\boxtimes P\boxtimes K_{2k \ell \Delta^{k}(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
+\end{thm}
+
+Theorems~\ref{PlanarProduct}(b), \ref{GenusProduct}(b) and \cref{ShortcutProduct} and  \cref{DrawG,qn,p-centered,p-centered-treewidth,non-repetitive,PowerShortcut} imply the following result, which with $g=0$ implies  \cref{kPowerBasic} in the introduction.
+
+\begin{thm}
+\label{PowerGenus}
+For integers $g\geq 0$ and $k,\Delta\geq 1$, let $\ell:= \max\{2g,3\} (2k^4+6k^2) \Delta^{k}$, and $t:= \binom{k+4}{4}-1$ if $g\geq 1$ and $t:= \binom{k+3}{3}-1$ if $g=0$.
+Then for every graph $G$ of Euler genus $g$ and maximum degree $\Delta$,
 \begin{compactitem}
-%%%%%%
+\item $G^k$ is a subgraph of $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$ 
 \item $G^k$ is $(g, 2k(k+1)\Delta^{k} )$-planar,
-%%%%%%
-\item $G^k$ has queue-number
-\[\qn(G^k) \leq 3\,\max\{2g,3\} (2k^4+6k^2) \,\Delta^{k}\, 2^{\binom{k+4}{4}-1},\]
-%If a graph $G$ has a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$, then $\qn(G)\le 3 \ell (2^t-1) + \floor{\tfrac{3}{2}\ell} \le 3 \ell (2^t-\frac12 )$.
-%%%%%%
-\item $G^k$ has  $p$-centered chromatic number
-\[\chi_p(G^k) \leq \max\{2g,3\}(2k^4+6k^2) \Delta^{k}  (p+1)
-\binom{p+\binom{k+4}{4}-1}{\binom{k+4}{4}-1}\enspace,\]
-%\label{p-centered} Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
-\item $G^k$ has  non-repetitive chromatic number
-\[ \pi(G^k) \leq  \max\{2g,3\} (2k^4+6k^2) \Delta^{k} \, 4^{\binom{k+4}{4}}\enspace.\]
-%If a graph $G$ has an $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$,  then $\pi(G)\le \ell\, 4^{t+1}$.
+\item $G^k$ has queue-number $\qn(G^k) \leq 3 \ell \cdot 2^t - \ceil{\tfrac{3}{2}\ell}$. 
+\item $G^k$ has  $p$-centered chromatic number $\chi_p(G^k) \leq \ell (p+1)\,  \binom{p+t}{t}$, 
+\item $G^k$ has  non-repetitive chromatic number $ \pi(G^k) \leq \ell \, 4^{t+1}$.
 \end{compactitem}
-\end{cor}
+\end{thm}
 
@@
 For every graph $G$, since $G^k$ has maximum degree at most $\Delta^k$, a result of \citet{DJKW16} implies that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$.
-\cref{PowerCor} improves upon this bound when $k\ll\Delta$ and $G$ has Euler genus $g$.
-Similarly, a result of \citet{debski.felsner.ea:improved} implies that $\chi_p(G^k)\le 1024p\Delta^{2k}$ and \cref{PowerCor} improves upon this bound when $p,k\ll\Delta$ and $G$ has Euler genus $g$.
-
-%Also note that the above results in conjunction with the comments about separators in \cref{PriorWork} imply that for fixed integers $g,k,\Delta$, the $k$-th power of a graph with Euler genus $g$ and maximum degree $\Delta$ has $O(\sqrt{n})$ treewidth  and thus admits balanced separators of order $O(\sqrt{n})$.
-\note{PM}{How interesting is this last note?
-A $\sqrt{n}$-separator of the genus-$g$ graph $G$ has a $k$-neighbourhood of size at most $\sqrt{n}\Delta^k$ and is this neighbourhood is a separator of the $G^k$.}
+\cref{PowerGenus} improves upon this bound when $k,g\ll\Delta$.
+Similarly, a result of \citet{debski.felsner.ea:improved} implies that $\chi_p(G^k)\le 1024p\Delta^{2k}$ and \cref{PowerGenus} improves upon this bound when $p,k,g\ll\Delta$.
+
+\cref{qn,p-centered,p-centered-treewidth,non-repetitive,MinorFreeDegree,PowerProduct} imply the following analogous result for powers of graphs in any minor-closed class with bounded maximum degree.
+
+\begin{thm}
+\label{PowerMinor}
+For every graph $X$ there exists an integer $c$ such that for all integers $k,\Delta\geq 1$, if $t:= 2k\Delta^{k}(k^3+3k)\binom{k+c\Delta}{c\Delta}-1$ and $G$ is an $X$-minor-free graph with maximum degree $\Delta$, then:
+\begin{compactitem}
+	\item $G^k$ is a subgraph of $H\boxtimes P$ for some graph $H$ with treewidth $t$ and for some path $P$.
+	\item $G^k$ has queue-number at most $3\cdot 2^t-2$, 
+	\item $G^k$ has $p$-centered chromatic number at most $(p+1)\binom{p+t}{t}$. 
+\end{compactitem}
+\end{thm}
 
@@
 Then $vw\in E(G)$ if and only if $w\in N_k(v)$ or $v\in N_k(w)$.
-% and edge set
-%\[
-%  E= \{vw : v,w\in V(G),\, |\{x\in P\setminus\{v\}: d_2(v,x)\le d_2(v,w)\}|\le k\} \enspace .
-%\]
 (The edges of $G$ are straight-line segments joining their endpoints.) See \citep{ProximityGraphs} for a survey of results on $k$-nearest neighbour graphs and other related proximity graphs.
@@
   \end{compactenum}
-  If $xy$ is of Type~C, then \cref{convex} implies that (without loss of generality)
-  \[\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)\enspace.\]
+  If $xy$ is of Type~C, then \cref{convex} implies that $\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)$ without loss of generality. 
    In this case, we call $x$ a Type~C vertex.
@@
 
-We note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n\ge k+1$ vertices has at least $kn/2$ edges and at most $kn$ edges.
+Note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n\ge k+1$ vertices has at least $kn/2$ edges and at most $kn$ edges.
 For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is $\Omega(k^2)$.
 
-\cref{nearest-neighbour,k-planar,qn,p-centered,p-centered-treewidth} imply:
+\cref{nearest-neighbour,kPlanarProduct,qn,p-centered,p-centered-treewidth} imply:
 
@@
 \label{k-nn}
-Every $k$-nearest-neighbour graph has:
+For every integer $k\geq 1$ there exists integers $t\leq O(k^6)$ and $\ell\leq O(k^4)$ such that every $k$-nearest-neighbour graph:
 \begin{compactitem}
-\item an $H$-partition of layered width $O(k^4)$ in which $H$ has treewidth $O(k^6)$,
-\item queue-number at most $2^{O(k^6)}$,
-\item $p$-centered chromatic number at most $O(k^4) (p+1) (t+1)\binom{p+t}{t}$ for some integer $t\leq O(k^6)$.
+\item is a subgraph of $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $t$ and some path $P$, 
+\item has queue-number at most $2^{O(k^6)}$, and
+\item has $p$-centered chromatic number at most $\ell (p+1)\binom{p+t}{t}$.
 \end{compactitem}
@@
 
-\section{An Open Problem}
-
-\cref{k-planar} shows that $k$-planar graphs have $H$-partitions of layered width $O(k^2)$ in which $H$ has treewidth $O(k^3)$.
-We conclude with the following question:  Does there exist a function $f:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph has an $H$-partition of layered width at most $f(k)$ in which $H$ has treewidth at most $C$?
+\section{Wrapping Up}
+
+As mentioned in \cref{Introduction}, \citet{DEJGMM} prove results about adjacency labelling schemes for planar graphs.
+Stated in terms of universal graphs, their main theorem is interpreted as follows:
+
+\begin{thm}[\citep{DEJGMM}]
+	\label{Universal}
+	For every fixed integer $t$ and every integer $n>0$ there exists a
+	graph $U_n$ with $n^{1+o(1)}$ vertices such that for every graph $H$ of
+	treewidth at most $t$ and path $P$, every $n$-vertex subgraph of $H\boxtimes P$ is isomorphic to an induced subgraph of $U_n$.
+\end{thm}
+
+Combining this with \cref{gkPlanarProduct,k-nn,PowerMinor,StringPartition,MapPartition} yields the following:
+
+\begin{thm}
+	\label{UniversalUniversal}
+	For every fixed graph $X$ and all fixed integers $d,\delta,\Delta,g,k>0$, and every integer
+	$n>0$, there exists a graph $U_n$ with $n^{1+o(1)}$ vertices such that
+	$U_n$ contains all of the following graphs as induced subgraphs:
+	\begin{compactitem}
+		\item every $n$-vertex $(g,k)$-planar graph;
+		\item every $n$-vertex $(g,d)$-map graph;
+		\item every $n$-vertex $(g,\delta)$-string graph;
+		\item every $n$-vertex graph $G^k$ where $G$ is $X$-minor-free and has maximum degree at most $\Delta$;
+		\item every $k$-nearest neighbour graph of $n$ points in $\R^2$.
+	\end{compactitem}
+\end{thm}
+
+We finish with an open problem.
+\cref{kPlanarProduct} shows that every $k$-planar graph is a subgraph of $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $O(k^3)$ where $\ell\leq O(k^2)$.
+Does there exist a function $\ell:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph is a subgraph of $H\boxtimes P \boxtimes K_{\ell(k)}$ for some graph $H$ with treewidth at most $C$?
 Perhaps $C=3$.
@@
 % %%%  Squashing the bibliography
-%   \let\oldthebibliography=\thebibliography
-%   \let\endoldthebibliography=\endthebibliography
-%   \renewenvironment{thebibliography}[1]{%
-%     \begin{oldthebibliography}{#1}%
-%       \setlength{\parskip}{0.4ex}%
-%       \setlength{\itemsep}{0.4ex}%
-%   }{\end{oldthebibliography}}
-
-\bibliographystyle{plainnat}
-\bibliography{k-planar}
+   \let\oldthebibliography=\thebibliography
+   \let\endoldthebibliography=\endthebibliography
+   \renewenvironment{thebibliography}[1]{%
+     \begin{oldthebibliography}{#1}%
+       \setlength{\parskip}{0ex}%
+       \setlength{\itemsep}{0ex}%
+   }{\end{oldthebibliography}}
+
+%\bibliographystyle{plainnat}
+%\bibliography{k-planar}
 
@@
 %
-% \fontsize{11}{11.75}\selectfont
-%
-% \begin{thebibliography}{33}
-% \providecommand{\natexlab}[1]{#1}
-% \providecommand{\url}[1]{\texttt{#1}}
-% \expandafter\ifx\csname urlstyle\endcsname\relax
-%   \providecommand{\doi}[1]{doi: #1}\else
-%   \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
-%
-% \bibitem[{\'{A}}brego et~al.(2011){\'{A}}brego, Monroy,
-%   Fern{\'{a}}ndez{-}Merchant, Flores{-}Pe{\~{n}}aloza, Hurtado,
-%   Sacrist{\'{a}}n, and Saumell]{abrego.munroy.ea:on}
-% Bernardo~M. {\'{A}}brego, Ruy~Fabila Monroy, Silvia Fern{\'{a}}ndez{-}Merchant,
-%   David Flores{-}Pe{\~{n}}aloza, Ferran Hurtado, Vera Sacrist{\'{a}}n, and
-%   Maria Saumell.
-% \newblock On crossing numbers of geometric proximity graphs.
-% \newblock \emph{Comput. Geom.}, 44\penalty0 (4):\penalty0 216--233, 2011.
-% \newblock \doi{10.1016/j.comgeo.2010.11.003}.
-%
-% \bibitem[Ajtai et~al.(1982)Ajtai, Chv{\'a}tal, Newborn, and
-%   Szemer{\'e}di]{ajtai.chvatal.ea:crossing-free}
-% Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, Endre
-%   Szemer{\'e}di.
-% \newblock Crossing-free subgraphs.
-% \newblock \emph{Ann. Discrete Math.}, 12:\penalty0 9--12, 1982.
-% \newblock \doi{10.1016/S0304-0208(08)73484-4}.
-%
-% \bibitem[Alam et~al.(2018)Alam, Bekos, Gronemann, Kaufmann, and
-%   Pupyrev]{ABGKP18}
-% Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael Kaufmann, Sergey Pupyrev.
-% \newblock Queue layouts of planar 3-trees.
-% \newblock In Therese~C. Biedl and Andreas Kerren, editors, \emph{Proc. 26th
-%   International Symposium on Graph Drawing and Network Visualization} (GD '18), vol.~11282 of \emph{Lecture Notes in Computer Science}, pp.~213--226. Springer, 2018.
-% \newblock \doi{10.1007/978-3-030-04414-5\_15}.
-%
-% \bibitem[Alon et~al.(2002)Alon, Grytczuk, Ha{\l}uszczak, and
-%   Riordan]{AGHR-RSA02}
-% Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, Oliver Riordan.
-% \newblock Nonrepetitive colorings of graphs.
-% \newblock \emph{Random Structures Algorithms}, 21\penalty0 (3-4):\penalty0
-%   336--346, 2002.
-% \newblock \doi{10.1002/rsa.10057}.
-%
-% \bibitem[Bodlaender(1998)]{Bodlaender-TCS98}
-% Hans~L. Bodlaender.
-% \newblock A partial $k$-arboretum of graphs with bounded treewidth.
-% \newblock \emph{Theoret. Comput. Sci.}, 209\penalty0 (1-2):\penalty0 1--45,
-%   1998.
-% \newblock \doi{10.1016/S0304-3975(97)00228-4}.
-%
-% \bibitem[Bonamy et~al.(2019)Bonamy, Gavoille, and
-%   Pilipczuk]{bonamy.gavoille.ea:shorter}
-% Marthe Bonamy, Cyril Gavoille, and Micha{\l} Pilipczuk.
-% \newblock Shorter implicit representation for planar graphs.
-% \newblock In S.~Cabello and B.~Mohar, editors, \emph{9th Slovenian
-%   International Conference on Graph Theory} (SICGT), p.~63, June 2019.
-%
-% \bibitem[Bose et~al.(2001)Bose, Morin, Stojmenovi\'c, and
-%   Urrutia]{bose.morin.ea:routing}
-% Prosenjit Bose, Pat Morin, Ivan Stojmenovi\'c, Jorge Urrutia.
-% \newblock Routing with guaranteed delivery in ad hoc wireless networks.
-% \newblock \emph{Wireless Networks}, 7\penalty0 (6):\penalty0 609--616, 2001.
-% \newblock \doi{10.1023/A:1012319418150}.
-%
-% \bibitem[Bose et~al.(2012)Bose, Dujmovi\'c, Hurtado, Iacono, Langerman, Meijer,
-%   Sacrist\'an, Saumell, and Wood]{ProximityGraphs}
-% Prosenjit Bose, Vida Dujmovi\'c, Ferran Hurtado, John Iacono, Stefan Langerman,
-%   Henk Meijer, Vera Sacrist\'an, Maria Saumell, David~R. Wood.
-% \newblock Proximity graphs: $e$, $\delta$, $\delta$, $\delta$, $\chi$ and
-%   $\omega$.
-% \newblock \emph{Internat. J. Comput. Geom.  Appl.},
-%   22\penalty0 (5):\penalty0 439--469, 2012.
-%
-% \bibitem[Chen et~al.(2002)Chen, Grigni, and Papadimitriou]{CGP02}
-% Zhi-Zhong Chen, Michelangelo Grigni, Christos~H. Papadimitriou.
-% \newblock Map graphs.
-% \newblock \emph{J. ACM}, 49\penalty0 (2):\penalty0 127--138, 2002.
-% \newblock \doi{10.1145/506147.506148}.
-%
-% \bibitem[D{\k e}bski et~al.(2019)D{\k e}bski, Felsner, Micek, and Schr{\"o}der]{micek:personal}
-%   Micha{\l} D{\k e}bski, Stefan Felsner, Piotr Micek, Felix Schr{\" o}der.
-% \newblock Improved bounds for centered colorings, 2019.
-% \newblock \href{https://arxiv.org/abs/1907.04586}{arXiv:1907.04586}.
-%
-% \bibitem[Dujmovi\'{c} et~al.(2016)Dujmovi\'{c}, Joret, Kozik, and Wood]{DJKW16}
-% Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, David~R. Wood.
-% \newblock Nonrepetitive colouring via entropy compression.
-% \newblock \emph{Combinatorica}, 36\penalty0 (6):\penalty0 661--686, 2016.
-% \newblock \doi{10.1007/s00493-015-3070-6}.
-%
-% \bibitem[Dujmovi\'c et~al.(2017{\natexlab{a}})Dujmovi\'c, Eppstein, and
-%   Wood]{dujmovic.eppstein.ea:structure}
-% Vida Dujmovi\'c, David Eppstein, David~R. Wood.
-% \newblock Structure of graphs with locally restricted crossings.
-% \newblock \emph{{SIAM} J. Discrete Math.}, 31\penalty0 (2):\penalty0 805--824,
-%   2017{\natexlab{a}}.
-% \newblock \doi{10.1137/16M1062879}.
-%
-% \bibitem[Dujmovi\'c et~al.(2017{\natexlab{b}})Dujmovi\'c, Morin, and
-%   Wood]{dujmovic.morin.ea:layered}
-% Vida Dujmovi\'c, Pat Morin, David~R. Wood.
-% \newblock Layered separators in minor-closed graph classes with applications.
-% \newblock \emph{J. Comb. Theory, Ser. {B}}, 127:\penalty0 111--147,
-%   2017{\natexlab{b}}.
-% \newblock \doi{10.1016/j.jctb.2017.05.006}.
-%
-% \bibitem[Dujmovi\'c et~al.(2019{\natexlab{a}})Dujmovi\'c, Esperet, Joret,
-%   Walczak, and Wood]{dujmovic.esperet.ea:planar}
-% Vida Dujmovi\'c, Louis Esperet, Gwena{\"{e}}l Joret, Bartosz Walczak,   David~R. Wood.
-% \newblock Planar graphs have bounded nonrepetitive chromatic number.
-% \newblock  \href{http://arxiv.org/abs/1904.05269}{arXiv:1904.05269}, 2019{\natexlab{a}}.
-%
-% \bibitem[Dujmovi\'c et~al.(2019{\natexlab{b}})Dujmovi\'c, Joret, Micek, Morin,
-%   Ueckerdt, and Wood]{dujmovic.joret.ea:planar}
-% Vida Dujmovi\'c, Gwena{\"{e}}l Joret, Piotr Micek, Pat Morin, Torsten Ueckerdt, David~R. Wood.
-% \newblock Planar graphs have bounded queue-number.
-% \newblock \emph{Proc. 60th Annual Symp. on Foundations of Computer
-%   Science} (FOCS '19), to appear.
-%   \newblock  \href{http://arxiv.org/abs/1904.04791}{arXiv:1904.04791}, 2019{\natexlab{b}}.
-%
-% \bibitem[Fomin et~al.(2012)Fomin, Lokshtanov, and Saurabh]{FLS-SODA12}
-% Fedor~V. Fomin, Daniel Lokshtanov, Saket Saurabh.
-% \newblock Bidimensionality and geometric graphs.
-% \newblock In \emph{Proc. 23rd {A}nnual {ACM}-{SIAM} {S}ymposium on {D}iscrete
-%   {A}lgorithms} (SODA 2012), pp.~1563--1575, 2012.
-% \newblock \doi{10.1137/1.9781611973099.124}.
-%
-% \bibitem[Fox and Pach(2008)]{FP08}
-% Jacob Fox, J{\'a}nos Pach.
-% \newblock Separator theorems and {T}ur\'an-type results for planar intersection graphs.
-% \newblock \emph{Adv. Math.}, 219(3):1070--1080, 2008.
-% \newblock \doi{10.1016/j.aim.2008.06.002}.
-%
-% \bibitem[Fox and Pach(2010)]{FP10}
-% Jacob Fox, J{\'a}nos Pach.
-% \newblock A separator theorem for string graphs and its applications.
-% \newblock \emph{Combin. Probab. Comput.}, 19\penalty0 (3):\penalty0 371--390,
-%   2010.
-% \newblock \doi{10.1017/S0963548309990459}.
-%
-% \bibitem[Fox and Pach(2014)]{FP14}
-% Jacob Fox, J{\'a}nos Pach.
-% \newblock Applications of a new separator theorem for string graphs.
-% \newblock \emph{Combin. Probab. Comput.}, 23\penalty0 (1):\penalty0 66--74,
-%   2014.
-% \newblock \doi{10.1017/S0963548313000412}.
-%
-% \bibitem[Grigoriev and Bodlaender(2007)]{grigoriev.bodlaender:algorithms}
-% Alexander Grigoriev, Hans~L. Bodlaender.
-% \newblock Algorithms for graphs embeddable with few crossings per edge.
-% \newblock \emph{Algorithmica}, 49\penalty0 (1):\penalty0 1--11, 2007.
-% \newblock \doi{10.1007/s00453-007-0010-x}.
-%
-% \bibitem[Harvey and Wood(2017)]{HW17}
-% Daniel~J. Harvey, David~R. Wood.
-% \newblock Parameters tied to treewidth.
-% \newblock \emph{J. Graph Theory}, 84\penalty0 (4):\penalty0 364--385, 2017.
-% \newblock \doi{10.1002/jgt.22030}.
-%
-% \bibitem[Heath et~al.(1992)Heath, Leighton, and Rosenberg]{HLR92}
-% Lenwood~S. Heath, F.~Thomson Leighton, Arnold~L. Rosenberg.
-% \newblock Comparing queues and stacks as mechanisms for laying out graphs.
-% \newblock \emph{SIAM J. Discrete Math.}, 5\penalty0 (3):\penalty0 398--412,
-%   1992.
-% \newblock \doi{10.1137/0405031}.
-%
-% \bibitem[van~den Heuvel and Wood(2017)]{vdHW17}
-% Jan van~den Heuvel, David~R. Wood.
-% \newblock Improper colourings inspired by {H}adwiger's conjecture.
-% \newblock \emph{J. London Math. Soc.}, 98(1):129--148, 2018.
-% \newblock  \doi{10.1112/jlms.12127}. See   \href{https://arxiv.org/abs/1704.06536}{arXiv:1704.06536} for the cited results.
-%
-% \bibitem[Kannan et~al.(1988)Kannan, Naor, and
-%   Rudich]{kannan.naor.ea:implicit-stoc}
-% Sampath Kannan, Moni Naor, Steven Rudich.
-% \newblock Implicit representation of graphs.
-% \newblock In Janos Simon, editor, \emph{Proc.\ 20th Annual {ACM} Symposium on
-%   Theory of Computing} (STOC '88), pp.~334--343. {ACM}, 1988.
-% \newblock \doi{10.1145/62212.62244}.
-%
-% \bibitem[Kannan et~al.(1992)Kannan, Naor, and Rudich]{kannan.naor.ea:implicit}
-% Sampath Kannan, Moni Naor, Steven Rudich.
-% \newblock Implicit representation of graphs.
-% \newblock \emph{{SIAM} J. Discrete Math.}, 5\penalty0 (4):\penalty0 596--603,
-%   1992.
-% \newblock \doi{10.1137/0405049}.
-%
-% \bibitem[Kobourov et~al.(2017)Kobourov, Liotta, and
-%   Montecchiani]{kobourov.liotta.ea:annotated}
-% Stephen~G. Kobourov, Giuseppe Liotta, Fabrizio Montecchiani.
-% \newblock An annotated bibliography on 1-planarity.
-% \newblock \emph{Comput. Sci. Rev.}, 25:\penalty0 49--67, 2017.
-% \newblock \doi{10.1016/j.cosrev.2017.06.002}.
-%
-% \bibitem[Leighton(1983)]{leighton:complexity}
-% F.~Thomson Leighton.
-% \newblock \emph{Complexity issues in {VLSI}}.
-% \newblock MIT Press, Cambridge, 1983.
-%
-% \bibitem[Mohar and Thomassen(2001)]{mohar.thomassen:graphs}
-% Bojan Mohar, Carsten Thomassen.
-% \newblock \emph{Graphs on surfaces}.
-% \newblock Johns Hopkins Press, 2001.
-%
-% \bibitem[Ne{\v{s}}et{\v{r}}il and Ossona~de Mendez(2012)]{Sparsity}
-% Jaroslav Ne{\v{s}}et{\v{r}}il, Patrice Ossona~de Mendez.
-% \newblock \emph{Sparsity}, vol.~28 of \emph{Algorithms and Combinatorics}.
-% \newblock Springer, 2012.
-% \newblock \doi{10.1007/978-3-642-27875-4}.
-%
-% \bibitem[Pach and T{\'o}th(1997)]{PachToth97}
-% J{\'a}nos Pach, G{\'e}za T{\'o}th.
-% \newblock Graphs drawn with few crossings per edge.
-% \newblock \emph{Combinatorica}, 17(3):427--439, 1997
-%   2002.
-% \newblock \doi{10.1007/BF01215922}.
-%
-% \bibitem[Pach and T{\'o}th(2002)]{PachToth-DCG02}
-% J{\'a}nos Pach, G{\'e}za T{\'o}th.
-% \newblock Recognizing string graphs is decidable.
-% \newblock \emph{Discrete Comput. Geom.}, 28\penalty0 (4):\penalty0 593--606,
-%   2002.
-% \newblock \doi{10.1007/s00454-002-2891-4}.
-%
-% \bibitem[Pilipczuk and Siebertz(2019)]{pilipczuk.siebertz:polynomial-soda}
-% Micha{\l} Pilipczuk, Sebastian Siebertz.
-% \newblock Polynomial bounds for centered colorings on proper minor-closed graph
-%   classes.
-% \newblock In Timothy~M. Chan, editor, \emph{Proc.\ 30th Annual {ACM-SIAM}
-%   Symposium on Discrete Algorithms} ({SODA} 2019), pp.~1501--1520. {SIAM},
-%   2019.
-% \newblock \doi{10.1137/1.9781611975482.91}.
-%
-% \bibitem[Reed(2003)]{Reed03}
-% Bruce~A. Reed.
-% \newblock Algorithmic aspects of tree width.
-% \newblock In \emph{Recent Advances in Algorithms and Combinat.}, vol.~11,
-%   pp.~85--107. Springer, 2003,
-% \newblock \doi{10.1007/0-387-22444-0_4}.
-%
+\begin{thebibliography}{37}
+	\providecommand{\natexlab}[1]{#1}
+	\providecommand{\url}[1]{\texttt{#1}}
+	\expandafter\ifx\csname urlstyle\endcsname\relax
+	\providecommand{\doi}[1]{doi: #1}\else
+	\providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
+
+	\bibitem[{\'{A}}brego et~al.(2011){\'{A}}brego, Monroy,
+	Fern{\'{a}}ndez{-}Merchant, Flores{-}Pe{\~{n}}aloza, Hurtado,
+	Sacrist{\'{a}}n, and Saumell]{abrego.munroy.ea:on}
+	Bernardo~M. {\'{A}}brego, Ruy~Fabila Monroy, Silvia Fern{\'{a}}ndez{-}Merchant,
+	David Flores{-}Pe{\~{n}}aloza, Ferran Hurtado, Vera Sacrist{\'{a}}n, and
+	Maria Saumell.
+	\newblock On crossing numbers of geometric proximity graphs.
+	\newblock \emph{Comput. Geom.}, 44\penalty0 (4):\penalty0 216--233, 2011.
+	\newblock \doi{10.1016/j.comgeo.2010.11.003}.
+
+	\bibitem[Ajtai et~al.(1982)Ajtai, Chv{\'a}tal, Newborn, and
+	Szemer{\'e}di]{ajtai.chvatal.ea:crossing-free}
+	Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, and Endre
+	Szemer{\'e}di.
+	\newblock Crossing-free subgraphs.
+	\newblock \emph{Ann. Discrete Math.}, 12:\penalty0 9--12, 1982.
+	\newblock \doi{10.1016/S0304-0208(08)73484-4}.
+
+	\bibitem[Alam et~al.(2018)Alam, Bekos, Gronemann, Kaufmann, and
+	Pupyrev]{ABGKP18}
+	Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael Kaufmann, and
+	Sergey Pupyrev.
+	\newblock Queue layouts of planar 3-trees.
+	\newblock In Therese~C.
+Biedl and Andreas Kerren, editors, \emph{Proc.
+26th International Symposium on Graph Drawing and Network Visualization} (GD 2018), volume 11282 of \emph{Lecture Notes in Computer Science}, pages
+	213--226. Springer, 2018.
+	\newblock \doi{10.1007/978-3-030-04414-5\_15}.
+
+	\bibitem[Alon et~al.(2002)Alon, Grytczuk, Ha{\l}uszczak, and
+	Riordan]{AGHR-RSA02}
+	Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, and Oliver Riordan.
+	\newblock Nonrepetitive colorings of graphs.
+	\newblock \emph{Random Structures Algorithms}, 21\penalty0 (3-4):\penalty0
+	336--346, 2002.
+	\newblock \doi{10.1002/rsa.10057}.
+
+	\bibitem[Bodlaender(1998)]{Bodlaender-TCS98}
+	Hans~L. Bodlaender.
+	\newblock A partial $k$-arboretum of graphs with bounded treewidth.
+	\newblock \emph{Theoret. Comput. Sci.}, 209\penalty0 (1-2):\penalty0 1--45,
+	1998.
+	\newblock \doi{10.1016/S0304-3975(97)00228-4}.
+
+	\bibitem[Bonamy et~al.(2020)Bonamy, Gavoille, and
+	Pilipczuk]{bonamy.gavoille.ea:shorter}
+	Marthe Bonamy, Cyril Gavoille, and Micha{\l} Pilipczuk.
+	\newblock Shorter labeling schemes for planar graphs.
+	\newblock In Shuchi Chawla, editor, \emph{Proc. ACM-SIAM Symposium on Discrete Algorithms} (SODA 2020), pages 446--462. {SIAM}, 2020.
+	\newblock \doi{10.1137/1.9781611975994.27}.
+
+	\bibitem[Bose et~al.(2001)Bose, Morin, Stojmenović, and
+	Urrutia]{bose.morin.ea:routing}
+	Prosenjit Bose, Pat Morin, Ivan Stojmenović, and Jorge Urrutia.
+	\newblock Routing with guaranteed delivery in ad hoc wireless networks.
+	\newblock \emph{Wireless Networks}, 7\penalty0 (6):\penalty0 609--616, 2001.
+	\newblock \doi{10.1023/A:1012319418150}.
+
+	\bibitem[Bose et~al.(2012)Bose, Dujmovic, Hurtado, Iacono, Langerman, Meijer,
+	Adinolfi, Saumell, and Wood]{ProximityGraphs}
+	Prosenjit Bose, Vida Dujmovic, Ferran Hurtado, John Iacono, Stefan Langerman,
+	Henk Meijer, Vera~Sacrist{\'{a}}n Adinolfi, Maria Saumell, and David~R. Wood.
+	\newblock Proximity graphs: E, {\(\delta\)}, {\(\Delta\)}, {\(\chi\)} and
+	{\(\omega\)}.
+	\newblock \emph{Int. J. Comput. Geometry Appl.}, 22\penalty0 (5):\penalty0
+	439--470, 2012.
+	\newblock \doi{10.1142/S0218195912500112}.
+
+	\bibitem[Chen et~al.(2002)Chen, Grigni, and Papadimitriou]{CGP02}
+	Zhi-Zhong Chen, Michelangelo Grigni, and Christos~H. Papadimitriou.
+	\newblock Map graphs.
+	\newblock \emph{J. ACM}, 49\penalty0 (2):\penalty0 127--138, 2002.
+	\newblock \doi{10.1145/506147.506148}.
+
+	\bibitem[Debski et~al.(2020)Debski, Felsner, Micek, and
+	Schr{\"{o}}der]{debski.felsner.ea:improved}
+	Michal Debski, Stefan Felsner, Piotr Micek, and Felix Schr{\"{o}}der.
+	\newblock Improved bounds for centered colorings.
+	\newblock In Shuchi Chawla, editor, \emph{Proc. {ACM-SIAM} Symposium on
+		Discrete Algorithms} ({SODA} 2020), pages 2212--2226. {SIAM}, 2020.
+	\newblock \doi{10.1137/1.9781611975994.136}.
+
+	\bibitem[{Dvo{\v{r}}{\'a}k et~al.(2020)Dvo{\v{r}}{\'a}k, Huynh, Joret, and Liu}]{DHGLW}
+	Zden{\v{e}}k Dvo{\v{r}}{\'a}k, Tony Huynh, Gwena\"el Joret, and Chun-Hung Liu.
+	\newblock Notes on graph product structure theory.
+	\newblock 2020.
+	\newblock \arXiv{2001.08860}.
+
+	\bibitem[Dujmovi\'{c} et~al.(2016)Dujmovi\'{c}, Joret, Kozik, and Wood]{DJKW16}
+	Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, and David~R. Wood.
+	\newblock Nonrepetitive colouring via entropy compression.
+	\newblock \emph{Combinatorica}, 36\penalty0 (6):\penalty0 661--686, 2016.
+	\newblock \doi{10.1007/s00493-015-3070-6}.
+
+	\bibitem[Dujmovi\'c et~al.(2017{\natexlab{a}})Dujmovi\'c, Eppstein, and
+	Wood]{dujmovic.eppstein.ea:structure}
+	Vida Dujmovi\'c, David Eppstein, and David~R. Wood.
+	\newblock Structure of graphs with locally restricted crossings.
+	\newblock \emph{{SIAM} J. Discrete Math.}, 31\penalty0 (2):\penalty0 805--824,
+	2017{\natexlab{a}}.
+	\newblock \doi{10.1137/16M1062879}.
+
+	\bibitem[Dujmovi\'c et~al.(2017{\natexlab{b}})Dujmovi\'c, Morin, and
+	Wood]{dujmovic.morin.ea:layered}
+	Vida Dujmovi\'c, Pat Morin, and David~R. Wood.
+	\newblock Layered separators in minor-closed graph classes with applications.
+	\newblock \emph{J. Comb. Theory, Ser. {B}}, 127:\penalty0 111--147,
+	2017{\natexlab{b}}.
+	\newblock \doi{10.1016/j.jctb.2017.05.006}.
+
+	\bibitem[Dujmovi{\'c} et~al.(2019)Dujmovi{\'c}, Joret, Micek, Morin, Ueckerdt,
+	and Wood]{dujmovic.joret.ea:planar}
+	Vida Dujmovi{\'c}, Gwena\"{e}l Joret, Piotr Micek, Pat Morin, Torsten Ueckerdt,
+	and David~R. Wood.
+	\newblock Planar graphs have bounded queue-number.
+	\newblock In David Zuckerman, editor, \emph{Proc. 60th {IEEE} Annual Symp.
+		Foundations Comput. Sci.} (FOCS 2019), pages 862--875. IEEE, 2019.
+	\newblock \doi{10.1109/FOCS.2019.00056}.
+	\newblock To appear in \emph{J. ACM}, \arXiv{1904.04791}.
+
+	\bibitem[Dujmovi\'c et~al.(2020{\natexlab{a}})Dujmovi\'c, Esperet, Joret,
+	Gavoille, Micek, and Morin]{DEJGMM}
+	Vida Dujmovi\'c, Louis Esperet, Gwena\"el Joret, Cyril Gavoille, Piotr Micek,
+	and Pat Morin.
+	\newblock Adjacency labelling for planar graphs (and beyond).
+	\newblock 2020{\natexlab{a}}.
+	\newblock \arXiv{2003.04280}.
+
+	\bibitem[Dujmovi\'c et~al.(2020{\natexlab{b}})Dujmovi\'c, Esperet, Joret,
+	Walczak, and Wood]{dujmovic.esperet.ea:planar}
+	Vida Dujmovi\'c, Louis Esperet, Gwena{\"{e}}l Joret, Bartosz Walczak, and
+	David~R. Wood.
+	\newblock Planar graphs have bounded nonrepetitive chromatic number.
+	\newblock \emph{Advances in Combinatorics}, 5, 2020{\natexlab{b}}.
+	\newblock \doi{10.19086/aic.12100}.
+
+	\bibitem[{Dujmovi{\'c} et~al.(2020)Dujmovi{\'c}, Esperet, Morin, Walczak, and Wood}]{DEMWW}
+	Vida Dujmovi{\'c}, Louis Esperet, Pat Morin, Bartosz Walczak, and
+ David~R. Wood.
+	\newblock Clustered 3-colouring graphs of bounded degree.
+	\newblock 2020.
+	\newblock \arXiv{2002.11721}.
+
+	\bibitem[Fomin et~al.(2012)Fomin, Lokshtanov, and Saurabh]{FLS-SODA12}
+	Fedor~V. Fomin, Daniel Lokshtanov, and Saket Saurabh.
+	\newblock Bidimensionality and geometric graphs.
+	\newblock In \emph{Proc. 23rd {A}nnual {ACM}-{SIAM} {S}ymposium on {D}iscrete
+		{A}lgorithms}, pages 1563--1575, 2012.
+	\newblock \doi{10.1137/1.9781611973099.124}.
+
+	\bibitem[Fox and Pach(2008)]{FP08}
+	Jacob Fox and J{\'a}nos Pach.
+	\newblock Separator theorems and {T}ur\'an-type results for planar intersection
+	graphs.
+	\newblock \emph{Adv. Math.}, 219\penalty0 (3):\penalty0 1070--1080, 2008.
+	\newblock \doi{10.1016/j.aim.2008.06.002}.
+
+	\bibitem[Fox and Pach(2010)]{FP10}
+	Jacob Fox and J{\'a}nos Pach.
+	\newblock A separator theorem for string graphs and its applications.
+	\newblock \emph{Combin. Probab. Comput.}, 19\penalty0 (3):\penalty0 371--390,
+	2010.
+	\newblock \doi{10.1017/S0963548309990459}.
+
+	\bibitem[Fox and Pach(2014)]{FP14}
+	Jacob Fox and J{\'a}nos Pach.
+	\newblock Applications of a new separator theorem for string graphs.
+	\newblock \emph{Combin. Probab. Comput.}, 23\penalty0 (1):\penalty0 66--74,
+	2014.
+	\newblock \doi{10.1017/S0963548313000412}.
+
+	\bibitem[Grigoriev and Bodlaender(2007)]{grigoriev.bodlaender:algorithms}
+	Alexander Grigoriev and Hans~L. Bodlaender.
+	\newblock Algorithms for graphs embeddable with few crossings per edge.
+	\newblock \emph{Algorithmica}, 49\penalty0 (1):\penalty0 1--11, 2007.
+	\newblock \doi{10.1007/s00453-007-0010-x}.
+
+	\bibitem[Harvey and Wood(2017)]{HW17}
+	Daniel~J. Harvey and David~R. Wood.
+	\newblock Parameters tied to treewidth.
+	\newblock \emph{J. Graph Theory}, 84\penalty0 (4):\penalty0 364--385, 2017.
+	\newblock \doi{10.1002/jgt.22030}.
+
+	\bibitem[Heath et~al.(1992)Heath, Leighton, and Rosenberg]{HLR92}
+	Lenwood~S. Heath, F.~Thomson Leighton, and Arnold~L. Rosenberg.
+	\newblock Comparing queues and stacks as mechanisms for laying out graphs.
+	\newblock \emph{SIAM J. Discrete Math.}, 5\penalty0 (3):\penalty0 398--412,
+	1992.
+	\newblock \doi{10.1137/0405031}.
+
+	\bibitem[van~den Heuvel and Wood(2017)]{vdHW17}
+	Jan van~den Heuvel and David~R. Wood.
+	\newblock Improper colourings inspired by {H}adwiger's conjecture.
+	\newblock \arXiv{1704.06536}, 2017.
+
+	\bibitem[Kannan et~al.(1988)Kannan, Naor, and
+	Rudich]{kannan.naor.ea:implicit-stoc}
+	Sampath Kannan, Moni Naor, and Steven Rudich.
+	\newblock Implicit representation of graphs.
+	\newblock In Janos Simon, editor, \emph{Proc. 20th Annual {ACM} Symposium on Theory of Computing} (STOC 1988), pages 334--343. {ACM}, 1988.
+	\newblock \doi{10.1145/62212.62244}.
+
+	\bibitem[Kannan et~al.(1992)Kannan, Naor, and Rudich]{kannan.naor.ea:implicit}
+	Sampath Kannan, Moni Naor, and Steven Rudich.
+	\newblock Implicit representation of graphs.
+	\newblock \emph{{SIAM} J. Discrete Math.}, 5\penalty0 (4):\penalty0 596--603,
+	1992.
+	\newblock \doi{10.1137/0405049}.
+
+	\bibitem[Kobourov et~al.(2017)Kobourov, Liotta, and
+	Montecchiani]{kobourov.liotta.ea:annotated}
+	Stephen~G. Kobourov, Giuseppe Liotta, and Fabrizio Montecchiani.
+	\newblock An annotated bibliography on 1-planarity.
+	\newblock \emph{Comput. Sci. Rev.}, 25:\penalty0 49--67, 2017.
+	\newblock \doi{10.1016/j.cosrev.2017.06.002}.
+
+	\bibitem[Leighton(1983)]{leighton:complexity}
+	F.~Thomson Leighton.
+	\newblock \emph{Complexity issues in {VLSI}}.
+	\newblock MIT Press, Cambridge, MA, 1983.
+
+	\bibitem[Mohar and Thomassen(2001)]{mohar.thomassen:graphs}
+	Bojan Mohar and Carsten Thomassen.
+	\newblock \emph{Graphs on surfaces}.
+	\newblock Johns Hopkins University Press, 2001.
+
+	\bibitem[Ne{\v{s}}et{\v{r}}il and Ossona~de Mendez(2012)]{Sparsity}
+	Jaroslav Ne{\v{s}}et{\v{r}}il and Patrice Ossona~de Mendez.
+	\newblock \emph{Sparsity}, volume~28 of \emph{Algorithms and Combinatorics}.
+	\newblock Springer, 2012.
+	\newblock \doi{10.1007/978-3-642-27875-4}.
+
+	\bibitem[Pach and T{\'{o}}th(1997)]{PachToth97}
+	J{\'{a}}nos Pach and G{\'{e}}za T{\'{o}}th.
+	\newblock Graphs drawn with few crossings per edge.
+	\newblock \emph{Combinatorica}, 17\penalty0 (3):\penalty0 427--439, 1997.
+	\newblock \doi{10.1007/BF01215922}.
+
+	\bibitem[Pach and T{\'o}th(2002)]{PachToth-DCG02}
+	J{\'a}nos Pach and G{\'e}za T{\'o}th.
+	\newblock Recognizing string graphs is decidable.
+	\newblock \emph{Discrete Comput. Geom.}, 28\penalty0 (4):\penalty0 593--606,
+	2002.
+	\newblock \doi{10.1007/s00454-002-2891-4}.
+
+	\bibitem[Pilipczuk and Siebertz(2019)]{pilipczuk.siebertz:polynomial-soda}
+	Micha{\l} Pilipczuk and Sebastian Siebertz.
+	\newblock Polynomial bounds for centered colorings on proper minor-closed graph classes.
+	\newblock In Timothy~M. Chan, editor, \emph{Proc. 13th Annual {ACM-SIAM}
+		Symposium on Discrete Algorithms} (SODA 2019), pages 1501--1520. {SIAM},	2019.
+	\newblock \doi{10.1137/1.9781611975482.91}.
+
+	\bibitem[Pilipczuk and Siebertz(2018)]{pilipczuk.siebertz:polynomial-arxiv}
+	Michał Pilipczuk and Sebastian Siebertz.
+	\newblock Polynomial bounds for centered colorings on proper minor-closed graph
+	classes, 2018.
+	\newblock \arXiv{1807.03683v2}.
+
+	\bibitem[Reed(2003)]{Reed03}
+	Bruce~A. Reed.
+	\newblock Algorithmic aspects of tree width.
+	\newblock In \emph{Recent Advances in Algorithms and Combinatorics}, pages 85--107. Springer, 2003.
+	\newblock \doi{10.1007/0-387-22444-0_4}.
+
 %	\bibitem[Wiechert(2017)]{wiechert:on}
 %	Veit Wiechert.
-% \newblock On the queue-number of graphs with bounded treewidth.
+%	\newblock On the queue-number of graphs with bounded tree-width.
 %	\newblock \emph{Electr. J. Comb.}, 24\penalty0 (1):\penalty0 P1.65, 2017.
-% \newblock  \url{http://www.combinatorics.org/v24i1p65}.
-%
-% \bibitem[Zhu(2009)]{zhu:colouring}
-% Xuding Zhu.
-% \newblock Colouring graphs with bounded generalized colouring number.
-% \newblock \emph{Discrete Math.}, 309\penalty0 (18):\penalty0 5562--5568,
-%   2009.
-% \newblock \doi{10.1016/j.disc.2008.03.024}.
-%
-% \end{thebibliography}
+%	\newblock \doi{10.37236/6429}.
+
+	\bibitem[Zhu(2009)]{zhu:colouring}
+	Xuding Zhu.
+	\newblock Colouring graphs with bounded generalized colouring number.
+	\newblock \emph{Discrete Mathematics}, 309\penalty0 (18):\penalty0 5562--5568,
+	2009.
+	\newblock \doi{10.1016/j.disc.2008.03.024}.
+
+\end{thebibliography}
 
 \end{document}
-
-\newpage
-\appendix
-
-%%%%%%%%%%%%%%
-\section{Future Work}
-
-% \subsection{More Stuff}
-%
-% \begin{enumerate}
-%   % \item clustered colouring;
-%   \item connection with queue layouts;
-%   % \item $d$-map graphs, which are $d^2$-planar;
-%   \item connection with low-treewidth/treedepth colourings;
-%   \item An important open question is whether $k$-planar graphs have layered $H$-partitions whose layered width is upper bounded by a function of $k$ but in which the treewidth of $H$ is bounded by an absolute constant, independent of $k$.
-%   \item Are our current $H$-partitions $K_{3,s}$-free for some $s=s(k)$?  If so, this is good news for clustered colouring.
-% \end{enumerate}
-%
-% \noindent
-% \framebox{
-% \begin{minipage}{\textwidth}
-%   \note{DW}{%
-%     % I think we need to define layered treewidth, and
-%     % give the result of Dujmovic-Eppstein-Wood that (g,k)-planar graphs
-%     % have O(gk)  layered treewidth. This is the best previous structural
-%     % description of k-planar graphs. State that layered partitions are
-%     % richer than  layered treewidth, and lead to O(1) bounds on non-rep
-%     % chromatic number and queue-number, instead of O(log n) bounds via
-%     % layered treewidth.
-%     I also think we should give explicit upper bounds on non-rep chromatic
-%     number for k-planar graphs, and also for queue-number (especially if
-%     they improve the bound in the FOCS / J.ACM paper).
-%
-%     We should add corollaries for d-map graphs, since they are $d^2$-planar.
-%     }
-%
-%     \note{PM}{
-%     A couple of other things to look into:
-%     \begin{compactenum}
-%       \item $p$-centered colourings
-%       \item low treewidth colourings
-%     \end{compactenum}
-%     }
-% \end{minipage}
-% }
-%
-
-\note{DW}{Can we strengthen the result of \citep{MTTV97} to prove a structure theorem, not just a separator theorem.
-
-Then show that $k$-nearest neighbour graphs in any fixed dimension have bounded queue-number, etc.
-
-Conjecture: Every $k$-nearest neighbour graph in $d$ dimensions is a subgraph of $P_1 \boxtimes\dots \boxtimes P_{d-1} \boxtimes H$, where $P_1,\dots,P_{d-1}$ are paths and $H$ has
-treewidth at most $f(k,d)$.\\ This  is true for $d=1$ and $d=2$.
-
-It might be true that ``treewidth at most $f(k,d)$'' can be replaced by ``pathwidth at most $f(k,d)$'', which is true for $d=1$ and open for $d=2$.
-The obvious counterexample, is $T^+:=$ the complete binary tree plus one dominant vertex (since if $T^+$ is a subgraph of $H\boxtimes P$ then $H$ has unbounded pathwidth).
-But $T^+$ has unbounded degree, so it is not  a $k$-nearest neighbour graph for any fixed $k$.
-}
-
-%%%%%%%%%%%%%%%%%%
-\subsection{Unit Disc Graphs}
-
-Let $P$ be a set of points in $\mathbb{R}^d$.
-The \emph{unit disc} graph of $P$ is the graph $G$ with $V(G)=P$ where $vw\in E(G)$ if and only if $\dist(v,w)\leq 1$.
-Here $\dist$ is the Euclidean distance in $\mathbb{R}^d$.
-Let $P^{(d)}$ be the strong product of $d$ paths $P\boxtimes \dots \boxtimes P$ (the $d$-dimensional grid graph with crosses).
-
-\begin{lem}
-\label{UnitDiscGraph}
-Every unit disc graph $G$ in $\mathbb{R}^d$ with no $(k+1)$-clique is a subgraph of
-$P^{(d)} \boxtimes K_{k \ceil{\sqrt{d}}^d}$.
-\end{lem}
-
-\begin{proof}
-Let $t:= k \ceil{\sqrt{d}}^d$.
-Let $x_i(v)$ be the $i$-th coordinate of each vertex $v\in V(G)$.
-For $p_1,\dots,p_d\in\mathbb{Z}$, let $V\langle p_1,\dots,p_d\rangle$ be the set of vertices $v\in V(G)$ such that $x_i(v)\in[p_i,p_i+1)$ for each $i\in[d]$.
-Thus the  $V\langle p_1,\dots,p_d\rangle$ partition $V(G)$.
-Each set $V\langle p_1,\dots,p_d\rangle$ consists of the set of vertices in a particular unit cube.
-Note that the unit cube can be partitioned into $\ceil{\sqrt{d}}^d$  sub-cubes, each with side length at most $\frac{1}{\sqrt{d}}$ and thus with diameter at most 1.
-The set of vertices in a sub-cube with diameter at most 1 is a clique in $G$.
-Thus at most $k$ vertices lie in a single sub-cube, and  $|V\langle p_1,\dots,p_d\rangle | \leq t$.
-Injectively label the vertices in $V\langle p_1,\dots,p_d\rangle$ by $1,2,\dots,t$.
-Map each vertex $v$ in
-$V\langle p_1,\dots,p_d\rangle$ labelled $\ell(v)$ to the vertex  $(p_1,\dots,p_d,\ell(v))$ of
-$P^{(d)} \boxtimes K_{t}$.
-Thus the vertices of $G$ are mapped to distinct vertices of $P^{(d)} \boxtimes K_{t}$.
-For each edge $vw\in E(G)$, if $v\in V\langle p_1,\dots,p_d\rangle$ and
-$w\in V\langle q_1,\dots,q_d\rangle$, then $|p_i-q_i|\leq 1$ for each $i\in[d]$, and if $p_i=q_i$ for each $i\in[d]$, then $\ell(v)\neq\ell(w)$. Thus $v$ and $w$ are mapped to adjacent vertices in
-$P^{(d)} \boxtimes K_{t}$.
-\end{proof}
-
-\note{DW}{By a volume argument, every covering of the unit cube by balls of diameter 1 uses at least $(\frac{d}{18})^{d/2}$ balls.
-So the $k \ceil{\sqrt{d}}^d$ term in the above theorem cannot be drastically improved.
-}
-
-A result of \citet{Wood-Queue-DMTCS05} implies that $\qn(H \boxtimes P ) \leq 3 \qn(H)+ 1$ for every graph $H$ and path $P$.
-It follows that  $\qn( P^{(d)} \boxtimes H ) \leq \frac{1}{6} (  3^{d+1} (2 \qn(H) + 1 ) -3 )$.
-Since $\qn(K_n)=\floor{\frac{n}{2}}$, by \cref{UnitDiscGraph}, every unit disc graph $G$ in $\mathbb{R}^d$ with no $(k+1)$-clique satisfies
-\[\qn(G) \leq \qn( P^{(d)} \boxtimes K_{k \ceil{\sqrt{d}}^d} )
-\leq
-\frac{1}{6} (  3^{d+1} (k \ceil{\sqrt{d}}^d + 1 ) -3 )\enspace.\]
-
-\note{DW}{Add non-rep and $p$-centered colouring}
-
-\begin{lem}
-\label{TWhighDimProduct}
-Let $G$ be an $n$-vertex subgraph of $P^{(d)} \boxtimes H$ for some graph $H$.
-Then
-\[\tw(G) \leq 2  ( \tw(H)+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1\enspace.\]
-\[\pw(G) \leq 2  ( \pw(H)+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1\enspace.\]
-\end{lem}
-
-\begin{proof}
-Let $\langle{V^i_0,V^i_1,\dots}\rangle$ be the layering of $G$ determined by the $i$-th dimension.
-Let \[m:= \ceil*{ \left(\frac{dn}{t+1} \right)^{1/(d+1)} } \enspace.\]
-For $\alpha\in[0,m-1]$, let $V^{i,\alpha} := \bigcup\{ V^i_j : j\equiv \alpha \pmod{m} \}$.
-Thus $V^{i,0},\dots,V^{i,m-1}$ is a partition of $V(G)$.
-Hence $|V^{i,\alpha_i}| \leq \frac{n}{m}$ for some $\alpha_i\in[0,m-1]$.
-Let $X:= \bigcup_{i=1}^d V^{i,\alpha_i}$. Thus $|X| \leq \frac{dn}{m}$.
-Note that each component of $G-X$ is a subgraph of $Q^{(d)} \boxtimes H$, where $Q$ is the path on $m-1$ vertices.
-Since $\tw(G)$ equals the maximum treewidth of the connected components of $G$, we have
-$\tw(G) \leq \tw( Q^{(d)} \boxtimes H ) + |X| $.
-To obtain a tree decomposition of $Q^{(d)} \boxtimes H$ with width $(t+1) (m-1)^d - 1$, start with an optimal tree decomposition of $H$, and replace each instance of a vertex of $H$ by the corresponding copy of $Q^{(d)}$.
-Thus
-\begin{align*}
-\tw(G) \leq  (t+1) (m-1)^d - 1 + \frac{dn}{m}
- \leq 2  (t+1)^{ 1/(d+1) } (dn)^{1- 1/(d+1)}   -  1.
-\end{align*}
-\end{proof}
-
-\cref{UnitDiscGraph,TWhighDimProduct} with $H=P\boxtimes K_{k \ceil{\sqrt{d}}^d}$ (which has pathwidth
-at most $2k \ceil{\sqrt{d}}^d-1$) imply:
-
-\begin{cor}
-For every unit disc graph $G$ in $\mathbb{R}^d$ on $n$ vertices and with no $(k+1)$-clique,
-%$$\pw(G) \leq 2  k^{ 1/(d+1) }  \ceil{\sqrt{d}}^{d/(d+1) } (dn)^{1- 1/(d+1)}   -  1\enspace.$$
-\[\pw(G) \leq 2  (2k)^{ 1/d } \ceil{\sqrt{d}} ((d-1)n)^{1- 1/d}   -  1\enspace.\]
-\end{cor}
-
-%K_{k \ceil{\sqrt{d}}^d}$.

### 2021-11-18 slides/slides.tex
Making slides

\documentclass[xcolor=dvipsnames]{beamer}
\usepackage{ods}
%\usepackage{ods-figs}
\usepackage[cm]{sfmath}
\usepackage[utf8]{inputenc}
\usepackage{array}
%\usepackage{enumitem}
%\usepackage{enumitem}
%\setitemize{itemsep=1.5ex}
%\setlength{\leftmargini}{0pt}
\usepackage{array}

\newcommand{\caressed}{$\kappa$aressed}
\newcommand{\caresses}{$\kappa$aresses}
\newcommand{\R}{\mathbb{R}}
\newcommand{\dual}[1]{#1^\star}
\newcommand{\Fary}{F\'ary}

\DeclareMathOperator{\tw}{tw}

\title{Graph Product Structure for Non-Minor-Closed Classes}
\author{Vida Dujmović \and Pat Morin \and David R. Wood}
\titlegraphic{\includegraphics[height=1em]{by}}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
    \frametitle{Outline}

    \begin{itemize}
        \item $k$-planar graphs
        \item $(k,d)$-shortcut systems
        \item statement of main theorem
        \item applications
        \item proof of main theorem
        \item open problem
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Motivation: $k$-Planar Graphs}
    \framesubtitle{$k$-planar: locally non-planar}

    \begin{itemize}
        \item $G$ is \emph{$k$-planar} if it has a drawing with $\le\!\! k$ crossings per edge
        \begin{center}
            \multiinclude[<+>][format=pdf,start=1]{figs/k-planar}%
        \end{center}
        \item<4-> Is $G\subseteq H\boxtimes P$ where $\tw(H)\le f(k)$?
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Abstraction: Shortcut Systems}

    \begin{itemize}
        \item \emph{$(k,d)$-shortcut system} for $G$:
        \begin{itemize}
            \item<2-> a set $\mathcal{P}$ of length-$\le\!\! k$ paths in $G$
            \item<2-> each $v\in V(G)$ appears interior to at most $d$ paths in $\mathcal{P}$
        \end{itemize}
        \begin{center}
            \multiinclude[<+>][format=pdf,start=1]{figs/shortcut-system}%
        \end{center}
        \item<3-> $G^{\mathcal{P}}:= G\cup\{vw:\mbox{$v$ and $w$ are endpoints of a path in $\mathcal{P}$}\}$
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Observation}
    \framesubtitle{Crossed edges are just shortcuts in a planar graph}

    If $G'$ is $k$-planar then $G'\subseteq G^{\mathcal{P}}$ where
    \begin{itemize}
        \item $G$ is planar
        \item $\mathcal{P}$ is a $(k+1,2)$-shortcut system
    \end{itemize}
    \begin{center}
        \multiinclude[<+>][format=pdf,start=1]{figs/kgp}%
    \end{center}
\end{frame}

\begin{frame}
    \frametitle{Main Theorem}
    \framesubtitle{If $G$ has product structure then so does $G^{\mathcal{P}}$}
    \textbf{Main Theorem:} If
    \begin{itemize}
        \item $G\subseteq H\boxtimes P\boxtimes K$
        \begin{itemize}
            \item $\tw(H)=t$
            \item $K$ is a clique of order $\ell$
        \end{itemize}
        \item $\mathcal{P}$ is a $(k,d)$-shortcut system
    \end{itemize}
    then $G^\mathcal{P}\subseteq H'\boxtimes P\boxtimes K'$
    \begin{itemize}
        \item $\tw(H') \le \binom{k+t}{t}-1$ and
        \item $K'$ is a clique of order $d\ell(k^3+3k)$
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Applications}

    \textbf{($\mathcal{G}$) Graphs:}
    \begin{itemize}
        \item Any subgraph of $G^\mathcal{P}$ where $G$ has \emph{product structure}
        \begin{itemize}
            \item $(g,k)$-planar graphs
            \item $(g,\delta)$-string graphs
            \item $k$-nearest neighbour graphs of points in $2d$
            \item bounded-degree graphs from proper minor-closed families
            \item any $G^k$ where $G$ has bounded-degree and product structure
        \end{itemize}
    \end{itemize}

    \textbf{($\mathcal{A}$) Applications:}
    \begin{itemize}
        \item queue layouts
        \item non-repetitive colouring
        \item $p$-centered colouring and $\ell$-vertex ranking
        \item universal graphs
    \end{itemize}

    Anything in $\mathcal{G}\times \mathcal{A}$
\end{frame}

\end{document}

\begin{frame}
  \frametitle{Proper Good Curves}

  \begin{itemize}
     \item A Jordan curve $C$ is
     \begin{itemize}
        \item \emph{good} if it contains a point \emph{in} the outer face of $G$
        \item \emph{proper} if the intersection of $C$ with each edge of $G$ is
         \begin{itemize}
            \item empty; or
            \item a single point (maybe a vertex); or
            \item the entire edge.
         \end{itemize}
     \end{itemize}
  \end{itemize}
  \begin{center}
    \includegraphics[width=.22\textwidth]{figs/pgc-2}
    \includegraphics[width=.22\textwidth]{figs/pgc-3}
    \includegraphics[width=.22\textwidth]{figs/pgc-4}
    \includegraphics[width=.22\textwidth]{figs/pgc-5} \\
    \includegraphics[width=.22\textwidth]{figs/pgc-6}
    \includegraphics[width=.22\textwidth]{figs/pgc-7}
    \includegraphics[width=.22\textwidth]{figs/pgc-8}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Free Sets}

   \begin{itemize}[<+->]
     \item A set $F\subseteq V(G)$ is \emph{free} if for every
     $X\subset\R^2$ with $|X|=|F|$, $G$ has a \emph{non-crossing
     straight-line} drawing in which the vertices of $F$ are drawn on
     the points of $X$.

     \item Free sets have applications to
     \begin{itemize}
        \item untangling,
        \item column planarity,
        \item universal point subsets,
        \item partial simultaneous geometric drawings.
        \item \ldots
     \end{itemize}
   \end{itemize}
\end{frame}

\begin{frame}
   \frametitle{The Free-Set Theorem}

   \begin{itemize}
       \item[]\textbf{Theorem (Dujmovi\'c--Frati--Gon\c{c}alves--M.--Rote 2019):} A set $F\subseteq V(G)$ is a \emph{free set} if and only if there is a proper good curve that contains $F$.
   \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Proper Good Curves and Dual Cycles}

  \begin{itemize}
     \item Every proper-good curve containing $k$ vertices gives a dual
       cycle of length at least $k$.
       \begin{center}
         \only<1>{\includegraphics{figs/curve-cycle-1}}%
         \only<2>{\includegraphics{figs/curve-cycle-2}}%
         \only<3>{\includegraphics{figs/curve-cycle-3}}%
         \only<4->{\includegraphics{figs/curve-cycle-4}}
       \end{center}
     \item<5-> What about the other direction?
     \item<6-> Can we get a proper good curve containing many vertices from long dual cycle?
  \end{itemize}
\end{frame}

\begin{frame}
   \frametitle{Pause}

   \begin{itemize}
      \item From this point on $G$ is (wlog) a triangulation.
      \item $G$ has maximum degree $\Delta$.
      \item $\dual{G}$ is the dual of $G$, a 3-connected cubic planar
      graph whose largest face has size at most $\Delta$
   \end{itemize}
\end{frame}

\begin{frame}
   \frametitle{Dual Cycles---Circumference}

   \begin{itemize}[<+->]
      \item The longest cycle in $\dual{G}$ is called its \emph{circumference}
      \item Circumference of $n$-vertex 3-connected cubic (planar) $\dual{G}$
      \begin{itemize}
        \item Tait's Conjecture (1884): $\forall \dual{G}: c(\dual{G}) = n$
        \item Disproved by Tutte (1946): $\exists \dual{G}: c(\dual{G})< n$
        \item Gr\"unbaum Walther (1973): $\exists \dual{G}: c(\dual{G}) = O(n^{0.9859})$.
        \item Barnette (XXXX): $\forall \dual{G}: c(\dual{G})=\Omega(\log n)$.
        \item Bondy and Simonovits (XXXX): $\forall \dual{G}: c(\dual{G})=e^{\Omega(\sqrt{\log n})}$
        \item Jackson (XXXX): $\forall \dual{G}: c(\dual{G}) = \Omega(n^{0.6942})$
        \item Billinski et al. (2011): $\forall \dual{G}: c(\dual{G}) = \Omega(n^{0.7532})$
        \item Liu, Yu, Zhang (2019): $\forall \dual{G}: c(\dual{G}) = \Omega(n^{0.8})$
      \end{itemize}
   \end{itemize}
\end{frame}

\begin{frame}
   \frametitle{Dual Cycles and Proper Good Curves}

   \begin{itemize}[<+->]
      \item Every dual cycle defines a proper good curve
      \item But the curve contains no vertices!
      \item Can we ``bend'' the curve to pick up some vertices?
      \item Sometimes yes, sometimes no
      \begin{center}
        \only<1-4>{\includegraphics{figs/cycle-to-curve-1}}%
        \only<5>{\includegraphics{figs/cycle-to-curve-2}}%
        \only<6>{\includegraphics{figs/cycle-to-curve-3}}%
        \only<7>{\includegraphics{figs/cycle-to-curve-4}}%
        \only<8->{\includegraphics{figs/cycle-to-curve-5}}
      \end{center}
   \end{itemize}

\end{frame}

\begin{frame}
   \frametitle{Dual Cycles and Proper Good Curves}

   \begin{itemize}[<+->]
      \item A cycle $C$ in $\dual{G}$
      \begin{itemize}
         \item \emph{touches} $f$ if $f\cap C\neq\emptyset$
         \item \emph{caresses} $f$ if $f\cap C$ is a path
         \item \emph{pinches} $f$ if $f\cap C$ is a cycle or has more
               than one connected component.
      \end{itemize}
      \item $\tau$ouched, $\rho$inched, $\kappa$aressed
      \item $\tau = \rho + \kappa$
   \end{itemize}

   \begin{center}
      \includegraphics[height=.5\textheight]{figs/t0t1-2}
   \end{center}
\end{frame}

\begin{frame}
  \frametitle{Dual Cycles and Proper Good Curves}

  \begin{itemize}[<+->]
    \item[]\textbf{Theorem:} If $C$ \caresses\ $\kappa$ faces of $\dual{G}$ then there is a proper good curve $\tilde{C}$ that contains at least $\kappa/4$ vertices of $G$.
    \item[]\textit{Proof:} Take an independent set $\dual{S}$ of \caressed\ faces of $\dual{G}$ and ``bend'' $C$ so that it passes through each vertex of $S$.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Summary So Far}

  \begin{center}
    \uncover<4->{{\color{blue}cycle in $G$ of length $\ell=c\kappa\Delta^4$ \\ $\Downarrow$} \\}
    \uncover<1->{cycle in $\dual{G}$ that \caresses\ $\kappa$ faces}\\
    \uncover<2->{$\Downarrow$ \\
    proper good curve that contains $\kappa/4$ vertices of $G$}\\
    \uncover<3->{$\Downarrow$ \\
    free set of size $\kappa/4$}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{The Main Theorem}

  \begin{itemize}
     \item[]\textbf{Theorem:} If $G$ has maximum degree $\Delta$ and $\dual{G}$ has a cycle of length $\ell$, then $\dual{G}$ has a cycle $C'$ that \caresses\ $\kappa=\Omega(\ell/\Delta^4)$ faces.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Proof Ideas}

  \begin{itemize}[<+->]
     \item Want to show that $C$ $\kappa$aresses many faces
     \item $C$ has length $\ell$
     \item each face of $\dual{G}$ has most $\Delta$ edges
     \begin{itemize}[<+->]
       \item[$\therefore$] $C$ $\tau$ouches at least $\ell/\Delta$ faces
          ($\tau = \Omega(\ell/\Delta)$)
     \end{itemize}
     \item Recall: $\tau = \kappa + \rho$
     \item If $\kappa =\Omega(\ell/\Delta)$ we are done
     \item Otherwise $C$ $\rho$inches at least $\rho = \Omega(\ell/\Delta)$ faces
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{A Bad Example}

  \begin{center}
     \includegraphics[width=.9\textwidth]{figs/two-caressed}
  \end{center}
  \begin{itemize}
    \item $\tau = \kappa + \rho$
    \item $\tau = 2k+3$, $\kappa = 4$, $\rho = 2k-1$
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Proof Idea}

  \begin{center}
     \only<1,2>{\includegraphics[height=.2\textheight]{figs/two-caressed-1}}%
     \only<3-8>{\includegraphics[height=.2\textheight]{figs/two-caressed-2}}%
     \only<9->{\includegraphics[height=.2\textheight]{figs/two-caressed-3}}
  \end{center}
  \begin{itemize}
    \item<2-> \alert<11>{Define two trees $T_0$ (inside $C$) and $T_1$ (outside $C$)}
    \item<4-> Show that:
    \begin{itemize}
      \item<5-> Each leaf of $T_i$ contains a \caressed\ face
      \item<6-> Each $T_i$ has $\Omega(\ell/\Delta)$ nodes
    \end{itemize}
    \item<7-> We are done if either $T_i$ has many leaves, so each $T_i$ must have few leaves
    \begin{itemize}
      \item<8->[$\therefore$] each $T_i$ has many \emph{bad nodes} \\ (degree-2 and containing no caressed faces)
    \end{itemize}
    \item<10-> With enough bad nodes, we can perform a \emph{surgery} that increases the number of leaves in $T_0$.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{The Trees $T_0$ and $T_1$}

  \begin{center}
   \begin{tabular}{cc}
      \only<1>{\includegraphics[width=.45\textwidth]{figs/t0t1-1}}%
      \only<2->{\includegraphics[width=.45\textwidth]{figs/t0t1-2}}%
        &
      \uncover<3->{\includegraphics[width=.45\textwidth]{figs/t0t1-3}} \\
      \uncover<4->{\includegraphics[width=.45\textwidth]{figs/t0t1-4}} &
      \uncover<5->{\includegraphics[width=.45\textwidth]{figs/t0t1-5}} \\
   \end{tabular}
  \end{center}
\end{frame}

%      \item For a node $u$ of $T_i$,
%        \[ \rho_u \le 2(\kappa_u + \delta_u) \Leftrightarrow
%           \tau_u \le 3\kappa_u + 2\delta_u \]
%    \end{itemize}
%    \item So if $\tau_u$ is big then either
%    \begin{itemize}
%       \item $\kappa_u$ is big; or
%       \item $\delta_u$ is big
%    \end{itemize}
%  \end{itemize}
%\end{frame}
%

\begin{frame}
  \frametitle{Proof Idea}

  \begin{itemize}
    \only<1>{\item Define two trees $T_0$ (inside $C$) and $T_1$ (outside $C$)}
    \only<2->{\item[$\checkmark$] Define two trees $T_0$ (inside $C$) and $T_1$ (outside $C$)}
    \item Show that:
    \begin{itemize}
      \item \alert<3->{Each leaf of $T_i$ contains a \caressed\ face}
      \item Each $T_i$ has $\Omega(\ell/\Delta)$ nodes
    \end{itemize}
    \item We are done if either $T_i$ has many leaves, so each $T_i$ must have few leaves
    \begin{itemize}
      \item[$\therefore$] each $T_i$ has many \emph{bad nodes} \\ (degree-2 and containing no caressed faces)
    \end{itemize}
    \item With enough bad nodes, we can perform a \emph{surgery} that increases the number of leaves in $T_0$.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Every Leaf $\kappa$aresses a Face}

   \noindent\textbf{Lemma:}
   Let $P$ be a \emph{chord path} for $C$ and let $L$ and $R$ be the two faces
   of $P\cup C$ that each contain $P$ in their boundary. Then $R$
   contains at least one face of $\dual{T}$ that is caressed by $C$.

  \uncover<2->{\emph{Proof:} by induction on number of faces in $R$}
  \begin{center}
      \only<1,2>{\includegraphics[width=.45\textwidth]{figs/one-caressed-1}}%
      \only<3,4>{\includegraphics[width=.45\textwidth]{figs/one-caressed-2}}
      \only<5>{\includegraphics[width=.45\textwidth]{figs/one-caressed-3}}
  \end{center}

  \uncover<4->{
   \noindent\textbf{Corollary:} Every leaf of $T_i$ contains at least one
   caressed face.
  }

\end{frame}

\begin{frame}
  \frametitle{Proof Idea}

  \begin{itemize}
    \item[$\checkmark$] Define two trees $T_0$ (inside $C$) and $T_1$ (outside $C$)
    \item Show that:
    \begin{itemize}
      \only<1>{\item Each leaf of $T_i$ contains a \caressed\ face}
      \only<2->{\item[$\checkmark$] Each leaf of $T_i$ contains a \caressed\ face}
      \item Each $T_i$ has $\Omega(\ell/\Delta)$ nodes
    \end{itemize}
    \item We are done if either $T_i$ has many leaves, so each $T_i$ must have few leaves
    \begin{itemize}
      \item[$\therefore$] each $T_i$ has many \emph{bad nodes} \\ (degree-2 and containing no caressed faces)
    \end{itemize}
    \item \alert<3->{With enough bad nodes, we can perform a \emph{surgery} that increases the number of leaves in $T_0$.}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Bad Nodes}

  \begin{itemize}[<+->]
    \item Recall: A node of $T_i$ is \emph{bad} if its has degree 2 and contains no caressed face.
    \item[]\textbf{Lemma:} Every bad node contains exactly one face of $\dual{G}$
    \item Recall: We have lots of bad nodes and very few leaves
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Surgery}

  \begin{center}
    \begin{tabular}{m{.3\textwidth}m{2em}m{.3\textwidth}}
    \only<1>{\includegraphics[height=.7\textheight]{figs/cases-1}}%
    \only<2->{\includegraphics[height=.7\textheight]{figs/cases-3}} &
    \uncover<3->{$\Rightarrow$} &
    \uncover<3->{\includegraphics[height=.7\textheight]{figs/cases-4}}
    \end{tabular}
  \end{center}
  \begin{itemize}
     \item<4-> increases number of leaves in $T_1$
        (see $a_0$)
     \item<5-> decreases length of $C$ by at most $\Delta^2$
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Surgery II}

  \begin{center}
    \begin{tabular}{m{.3\textwidth}m{2em}m{.3\textwidth}}
    \only<1->{\includegraphics[height=.7\textheight]{figs/surgery-1}} &
    \uncover<2->{$\Rightarrow$} &
    \uncover<2->{\includegraphics[height=.7\textheight]{figs/surgery-2}}
    \end{tabular}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Proof Idea}

  \begin{itemize}
    \item[$\checkmark$] Define two trees $T_0$ (inside $C$) and $T_1$ (outside $C$)
    \item Show that:
    \begin{itemize}
      \item[$\checkmark$] Each leaf of $T_i$ contains a \caressed\ face
      \item \alert<3>{Each $T_i$ has $\Omega(\ell/\Delta)$ nodes}
    \end{itemize}
    \item We are done if either $T_i$ has many leaves, so each $T_i$ must have few leaves
    \begin{itemize}
      \item[$\therefore$] each $T_i$ has many \emph{bad nodes} \\ (degree-2 and containing no caressed faces)
    \end{itemize}
    \only<1>{\item With enough bad nodes, we can perform a \emph{surgery} that increases the number of leaves in $T_0$.}
    \only<2->{\item[$\checkmark$] With enough bad nodes, we can perform a \emph{surgery} that increases the number of leaves in $T_0$.}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{$T_i$ Has Many Nodes}

  Situation so far:
  \begin{itemize}[<+->]
    \item Need to show: $T_i$ has many nodes
    \item Danger: $T_i$ contains few nodes, each containing many $\tau$ouched faces
    \item Will show that for any node $u$,
      \[   \rho_u \le 2(\kappa_u + \delta_u)  \]
     ($\rho_u$inched faces (inside $u$), $\kappa_u$aressed faces (inside $u$),
        $\delta$egree of $u$)
    \item Recall: $\tau_u=\kappa_u+\rho_u$
    \item If $\tau_u$ is large then either:
       \begin{enumerate}[<+->]
          \item $\kappa_u$ is large (good, more $\kappa$aressed faces!); and/or
          \item $\delta_u$ is large (good, more leaves!).
       \end{enumerate}
  \end{itemize}
\end{frame}

\begin{frame}

\textbf{Lemma:}
Let $u$ be a node of $T_i$ {\color{gray}and let $\rho_u$, $\kappa_u$, and $\delta_u$ denote the number of pinched faces of $\dual{T}$ in $u$, the number of caressed faces of $\dual{T}$ in $u$, and the degree of $u$ in $T_i$, respectively}.  Then $\rho_u \le 2(\kappa_u+\delta_u)$.

\uncover<2->{\textit{Proof (by discharging):}}
\begin{itemize}
  \item<3-> Assign a charge of 2 to each $\rho$inched face
  \item<4-> Move charge around so that
  \begin{itemize}
    \item<5->No charge is left on any pinched face
    \item<6->Each caressed faces has a charge of at most 2
    \item<7->Each outgoing edge has a charge of at most 2
  \end{itemize}
\end{itemize}
\uncover<2->{
\begin{center}
                \begin{tabular}{cccc}
                        \includegraphics[height=.22\textheight]{figs/discharge-2} &
                        \includegraphics[height=.22\textheight]{figs/discharge-3} &
                        \includegraphics[height=.22\textheight]{figs/discharge-4} &
                        \includegraphics[height=.22\textheight]{figs/discharge-5}
                \end{tabular}
                \end{center}
}
\end{frame}

\begin{frame}
  \frametitle{Proof Idea}

  \begin{itemize}
    \item[$\checkmark$] Define two trees $T_0$ (inside $C$) and $T_1$ (outside $C$)
    \item Show that:
    \begin{itemize}
      \item[$\checkmark$] Each leaf of $T_i$ contains a \caressed\ face
      \only<1>{\item Each $T_i$ has $\Omega(\ell/\Delta)$ nodes}
      \only<2->{\item[$\checkmark$] Each $T_i$ has $\Omega(\ell/\Delta)$ nodes}
    \end{itemize}
    \item We are done if either $T_i$ has many leaves, so each $T_i$ must have few leaves
    \begin{itemize}
      \item[$\therefore$] each $T_i$ has many \emph{bad nodes} \\ (degree-2 and containing no caressed faces)
    \end{itemize}
    \item[$\checkmark$] With enough bad nodes, we can perform a \emph{surgery} that increases the number of leaves in $T_0$.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Conclusion}

  \textbf{Theorem:} If $\dual{G}$ has a cycle of lenth $\ell$, then $\dual{G}$ has a cycle $C'$ that caresses $\Omega(\ell/\Delta^4)$ faces.

  \textbf{Consequence:} Every $n$-vertex planar graph of maximum degree $\Delta$ has a free set of size $\Omega(n^{0.8}/\Delta^4)$.

  \textbf{Open Problem:} Eliminate the dependence on $\Delta$.
\end{frame}

%\closing

\end{document}

### 2021-11-18 slides/slides.tex
Getting there.....

@@
   \framesubtitle{If $G$ has product structure then so does $G^{\mathcal{P}}$}
-    \textbf{Main Theorem:} If
+  \textbf{Main Theorem:} \newline
+  If $G\subseteq H\boxtimes P\boxtimes K_\ell$
   \begin{itemize}
-        \item $G\subseteq H\boxtimes P\boxtimes K$
+    \item  $\tw(H)\le t$
+    \item $\mathcal{P}$ is a $(k,d)$-shortcut system for $G$
+  \end{itemize}
+  then $G^\mathcal{P}\subseteq H'\boxtimes P\boxtimes K_{\ell'}$
   \begin{itemize}
-            \item $\tw(H)=t$
-            \item $K$ is a clique of order $\ell$
+    \item $\tw(H') \le t'=\binom{k+t}{t}-1$ and
+    \item $\ell'=d\ell(k^3+3k)$
+    % \item $K'$ is a clique of order $d\ell(k^3+3k)$
   \end{itemize}
-        \item $\mathcal{P}$ is a $(k,d)$-shortcut system
+  \vspace{1cm}
+  \uncover<2->{\textbf{Example:} $G'$ is ${\color{red} k}$-planar
+    \begin{itemize}
+      \item $t=3$, $\ell=3$, $d=2$, $k={\color{red}k}+1$
+      \item $t'=O(k^3)$, $\ell'=O(k^3)$ 
     \end{itemize}
-    then $G^\mathcal{P}\subseteq H'\boxtimes P\boxtimes K'$
-    \begin{itemize}
-        \item $\tw(H') \le \binom{k+t}{t}-1$ and
-        \item $K'$ is a clique of order $d\ell(k^3+3k)$
-    \end{itemize}
+  }
 \end{frame}
@@
     \frametitle{Applications}
+    \framesubtitle{(Besides $k$-planar graphs)}
 
-    \textbf{($\mathcal{G}$) Graphs:}
-    \begin{itemize}
-        \item Any subgraph of $G^\mathcal{P}$ where $G$ has \emph{product structure}
+    \textbf{($\mathcal{G}$) Graphs:} any subgraph of $G^\mathcal{P}$ where $G$ has \emph{product structure}
     \begin{itemize}
@@
         \item $(g,\delta)$-string graphs
-            \item $k$-nearest neighbour graphs of points in $2d$
+        \item $k$-nearest neighbour graphs of $2d$ point sets
         \item bounded-degree graphs from proper minor-closed families
         \item any $G^k$ where $G$ has bounded-degree and product structure
-        \end{itemize}
     \end{itemize}
@@
 
-\end{document}
+\section{Proof of Main Theorem}
 
 \begin{frame}
-  \frametitle{Proper Good Curves}
+  \frametitle{Proof: The Picture}
+  \begin{itemize}
+    \item $G= H\boxtimes P\boxtimes K_{\ell}$
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Shortcuts over multiple layers}
 
   \begin{itemize}
-     \item A Jordan curve $C$ is
+    \item Shortcuts in $\mathcal{P}$ may jump $k$ layers
+    \item Easy solution: Compress $k$ layers into one
     \begin{itemize}
-        \item \emph{good} if it contains a point \emph{in} the outer face of $G$
-        \item \emph{proper} if the intersection of $C$ with each edge of $G$ is
+      \item Length $k-1$ path $P'$ in $P$ compresses down to vertex
+      \item $P'\boxtimes K_{\ell}\subseteq K_{\ell k}$
+    \end{itemize}
+    \item Increases $\ell$ by a factor of $k$
+    \item $\mathcal{P}$ is a $(k,d)$-shortcut system over $H\boxtimes P\boxtimes K_{\ell k}$ with all shortcuts crossing $\le\!\! 1$ layers
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Projecting shortcuts onto $H\boxtimes P$}
+
   \begin{itemize}
-            \item empty; or
-            \item a single point (maybe a vertex); or
-            \item the entire edge.
+    \item Contract each copy of $K_{k\ell}$, so we are left with $H\boxtimes P$
+    \item Each shortcut in $\mathcal{P}$ projects onto a walk in $H\boxtimes P$
+    \item Each vertex of $H\boxtimes P$ is used in at most $dk\ell$ shortcuts
+    \item Studying a $(k,dk\ell)$-shortcut system $\mathcal{P'}$ over $G'=H\boxtimes P$
   \end{itemize}
-     \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Nice Tree Decompositions}
+
+  \begin{itemize}
+    \item A \emph{nice} rooted tree decomposition $\mathcal{T}=(B_x:x\in V(T))$ of $H$ has 
+    \begin{itemize}
+      \item $V(T) = V(H)$
+      \item $x$ is lowest common ancestor of $\{x'\in V(T):x\in B_{x'}\}$
     \end{itemize}
     \begin{center}
-    \includegraphics[width=.22\textwidth]{figs/pgc-2}
-    \includegraphics[width=.22\textwidth]{figs/pgc-3}
-    \includegraphics[width=.22\textwidth]{figs/pgc-4}
-    \includegraphics[width=.22\textwidth]{figs/pgc-5} \\
-    \includegraphics[width=.22\textwidth]{figs/pgc-6}
-    \includegraphics[width=.22\textwidth]{figs/pgc-7}
-    \includegraphics[width=.22\textwidth]{figs/pgc-8}
+      \includegraphics{figs/nice-td}
+    \end{center}
+    \item If $xy\in E(H)$ then $x\prec_T y$ or $y\prec_T x$
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Alternative View}
+
+  \begin{itemize}
+    \item<2-> Start with $B_x\gets\{x\}$ for each $x\in V(T)$
+    \item<3-> If $xy\in E(H)$ and $x\prec_T y$ then drag $x$ down into $B_y$
+  \end{itemize}
+  \begin{center}
+    \multiinclude[<+>][format=pdf,start=1]{figs/alternative-view}%
   \end{center}
@@
 \begin{frame}
-  \frametitle{Free Sets}
+  \frametitle{A Useful Lemma}
 
-   \begin{itemize}[<+->]
-     \item A set $F\subseteq V(G)$ is \emph{free} if for every
-     $X\subset\R^2$ with $|X|=|F|$, $G$ has a \emph{non-crossing
-     straight-line} drawing in which the vertices of $F$ are drawn on
-     the points of $X$.
+  \textbf{Lemma (Pilipczuk--Siebertz):}\footnote{Yes, the same paper!} For any $z\in V(H)$ the number of $x\in V(H)$ such that $x\prec_T z\prec_T z'$ and $\mathop{dist}_H(x,z')\le k$ is at most $\binom{k+t}{t}$.
+  \begin{center}
+    \includegraphics{figs/pz}%
+  \end{center}
+\end{frame}
 
-     \item Free sets have applications to
+\begin{frame}
+  \frametitle{Projecting shortcuts onto $T$}
+
   \begin{itemize}
-        \item untangling,
-        \item column planarity,
-        \item universal point subsets,
-        \item partial simultaneous geometric drawings.
-        \item \ldots
+    \item A shortcut in $I\in\mathcal{P'}$
+    \begin{itemize}
+      \item<2-> is a path $I$ in $G'=H\boxtimes P$
+      \item<3-> $I$ projects onto a walk $W$ in $H$
+      \item<4-> $W$ is a sequence $S_I$ of nodes in $T$
     \end{itemize}
+  \end{itemize}
+  \begin{center}
+    \multiinclude[<+>][format=pdf,start=2,end=5]{figs/nice-td}%
+  \end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Lifting in $T$}
+  \begin{itemize}
+    \item For $v$ in $G$, $a(v)$ is the LCA of $\cup\{S_I:I\in\mathcal{P'},\, v\in V(I)\}$
+    \item For $x\in V(T)$, $S_x:=\{v\in V(G): a(v)=x\}$
+    \begin{center}
+      \includegraphics{figs/nice-td-6}%
+    \end{center}
+    \item Note: $\mathcal{S}:=(S_x:x\in V(T))$ is a partition of $V(G)$
 \end{itemize}
@@
 \begin{frame}
-   \frametitle{The Free-Set Theorem}
+  \frametitle{Now the Crazy Part}
 
+    Let $H':=G/\mathcal{S}$
+
+    \textbf{Claim:} $\tw(H')\le\binom{k+t}{t}-1$
+
+    \textit{Proof:}
     \begin{itemize}
-       \item[]\textbf{Theorem (Dujmovi\'c--Frati--Gon\c{c}alves--M.--Rote 2019):} A set $F\subseteq V(G)$ is a \emph{free set} if and only if there is a proper good curve that contains $F$.
+      \item Each $x\in V(H')$ is obtained by contracting $S_x$
+      \item Use tree nice tree decomposition with the same tree $T$
+      \item If $xz\in E(H')$ and $x\prec_T z$ then drag $x$ down into $B_z$
+      \item Point: If $x\in B_{z'}$ then $x\prec_T z'\prec_T z$ and $\mathop{dist}_{H}(x,z)\le k$.  
+      \item By Pilipczuk-Siebertz Lemma, $|B_{z'}|\le \binom{k+t}{t}$.
     \end{itemize}
@@
 \begin{frame}
-  \frametitle{Proper Good Curves and Dual Cycles}
-
   \begin{itemize}
-     \item Every proper-good curve containing $k$ vertices gives a dual
-       cycle of length at least $k$.
-       \begin{center}
-         \only<1>{\includegraphics{figs/curve-cycle-1}}%
-         \only<2>{\includegraphics{figs/curve-cycle-2}}%
-         \only<3>{\includegraphics{figs/curve-cycle-3}}%
-         \only<4->{\includegraphics{figs/curve-cycle-4}}
-       \end{center}
-     \item<5-> What about the other direction?
-     \item<6-> Can we get a proper good curve containing many vertices from long dual cycle?
+    \item each $S_x$ contains $O(1)$ vertices from each row of $H\boxtimes P$.
+(Easy, each vertex in $S_x$ from row $i$ comes from a shortcut through a vertex of $Y_x$ in row $i-1$, $i$, or $i+1$, so $3dk\ell(k+1)=3d\ell(k^2+k)$)
   \end{itemize}
@@
 
-\begin{frame}
-   \frametitle{Pause}
-
-   \begin{itemize}
-      \item From this point on $G$ is (wlog) a triangulation.
-      \item $G$ has maximum degree $\Delta$.
-      \item $\dual{G}$ is the dual of $G$, a 3-connected cubic planar
-      graph whose largest face has size at most $\Delta$
-   \end{itemize}
-\end{frame}
-
-\begin{frame}
-   \frametitle{Dual Cycles---Circumference}
-
-   \begin{itemize}[<+->]
-      \item The longest cycle in $\dual{G}$ is called its \emph{circumference}
-      \item Circumference of $n$-vertex 3-connected cubic (planar) $\dual{G}$
-      \begin{itemize}
-        \item Tait's Conjecture (1884): $\forall \dual{G}: c(\dual{G}) = n$
-        \item Disproved by Tutte (1946): $\exists \dual{G}: c(\dual{G})< n$
-        \item Gr\"unbaum Walther (1973): $\exists \dual{G}: c(\dual{G}) = O(n^{0.9859})$.
-        \item Barnette (XXXX): $\forall \dual{G}: c(\dual{G})=\Omega(\log n)$.
-        \item Bondy and Simonovits (XXXX): $\forall \dual{G}: c(\dual{G})=e^{\Omega(\sqrt{\log n})}$
-        \item Jackson (XXXX): $\forall \dual{G}: c(\dual{G}) = \Omega(n^{0.6942})$
-        \item Billinski et al. (2011): $\forall \dual{G}: c(\dual{G}) = \Omega(n^{0.7532})$
-        \item Liu, Yu, Zhang (2019): $\forall \dual{G}: c(\dual{G}) = \Omega(n^{0.8})$
-      \end{itemize}
-   \end{itemize}
-\end{frame}
-
-\begin{frame}
-   \frametitle{Dual Cycles and Proper Good Curves}
-
-   \begin{itemize}[<+->]
-      \item Every dual cycle defines a proper good curve
-      \item But the curve contains no vertices!
-      \item Can we ``bend'' the curve to pick up some vertices?
-      \item Sometimes yes, sometimes no
-      \begin{center}
-        \only<1-4>{\includegraphics{figs/cycle-to-curve-1}}%
-        \only<5>{\includegraphics{figs/cycle-to-curve-2}}%
-        \only<6>{\includegraphics{figs/cycle-to-curve-3}}%
-        \only<7>{\includegraphics{figs/cycle-to-curve-4}}%
-        \only<8->{\includegraphics{figs/cycle-to-curve-5}}
-      \end{center}
-   \end{itemize}
-
-\end{frame}
-
-\begin{frame}
-   \frametitle{Dual Cycles and Proper Good Curves}
-
-   \begin{itemize}[<+->]
-      \item A cycle $C$ in $\dual{G}$
-      \begin{itemize}
-         \item \emph{touches} $f$ if $f\cap C\neq\emptyset$
-         \item \emph{caresses} $f$ if $f\cap C$ is a path
-         \item \emph{pinches} $f$ if $f\cap C$ is a cycle or has more
-               than one connected component.
-      \end{itemize}
-      \item $\tau$ouched, $\rho$inched, $\kappa$aressed
-      \item $\tau = \rho + \kappa$
-   \end{itemize}
-
-   \begin{center}
-      \includegraphics[height=.5\textheight]{figs/t0t1-2}
-   \end{center}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Dual Cycles and Proper Good Curves}
-
-  \begin{itemize}[<+->]
-    \item[]\textbf{Theorem:} If $C$ \caresses\ $\kappa$ faces of $\dual{G}$ then there is a proper good curve $\tilde{C}$ that contains at least $\kappa/4$ vertices of $G$.
-    \item[]\textit{Proof:} Take an independent set $\dual{S}$ of \caressed\ faces of $\dual{G}$ and ``bend'' $C$ so that it passes through each vertex of $S$.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Summary So Far}
-
-  \begin{center}
-    \uncover<4->{{\color{blue}cycle in $G$ of length $\ell=c\kappa\Delta^4$ \\ $\Downarrow$} \\}
-    \uncover<1->{cycle in $\dual{G}$ that \caresses\ $\kappa$ faces}\\
-    \uncover<2->{$\Downarrow$ \\
-    proper good curve that contains $\kappa/4$ vertices of $G$}\\
-    \uncover<3->{$\Downarrow$ \\
-    free set of size $\kappa/4$}
-  \end{center}
-\end{frame}
-
-\begin{frame}
-  \frametitle{The Main Theorem}
-
-  \begin{itemize}
-     \item[]\textbf{Theorem:} If $G$ has maximum degree $\Delta$ and $\dual{G}$ has a cycle of length $\ell$, then $\dual{G}$ has a cycle $C'$ that \caresses\ $\kappa=\Omega(\ell/\Delta^4)$ faces.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Proof Ideas}
-
-  \begin{itemize}[<+->]
-     \item Want to show that $C$ $\kappa$aresses many faces
-     \item $C$ has length $\ell$
-     \item each face of $\dual{G}$ has most $\Delta$ edges
-     \begin{itemize}[<+->]
-       \item[$\therefore$] $C$ $\tau$ouches at least $\ell/\Delta$ faces
-          ($\tau = \Omega(\ell/\Delta)$)
-     \end{itemize}
-     \item Recall: $\tau = \kappa + \rho$
-     \item If $\kappa =\Omega(\ell/\Delta)$ we are done
-     \item Otherwise $C$ $\rho$inches at least $\rho = \Omega(\ell/\Delta)$ faces
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{A Bad Example}
-
-  \begin{center}
-     \includegraphics[width=.9\textwidth]{figs/two-caressed}
-  \end{center}
-  \begin{itemize}
-    \item $\tau = \kappa + \rho$
-    \item $\tau = 2k+3$, $\kappa = 4$, $\rho = 2k-1$
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Proof Idea}
-
-  \begin{center}
-     \only<1,2>{\includegraphics[height=.2\textheight]{figs/two-caressed-1}}%
-     \only<3-8>{\includegraphics[height=.2\textheight]{figs/two-caressed-2}}%
-     \only<9->{\includegraphics[height=.2\textheight]{figs/two-caressed-3}}
-  \end{center}
-  \begin{itemize}
-    \item<2-> \alert<11>{Define two trees $T_0$ (inside $C$) and $T_1$ (outside $C$)}
-    \item<4-> Show that:
-    \begin{itemize}
-      \item<5-> Each leaf of $T_i$ contains a \caressed\ face
-      \item<6-> Each $T_i$ has $\Omega(\ell/\Delta)$ nodes
-    \end{itemize}
-    \item<7-> We are done if either $T_i$ has many leaves, so each $T_i$ must have few leaves
-    \begin{itemize}
-      \item<8->[$\therefore$] each $T_i$ has many \emph{bad nodes} \\ (degree-2 and containing no caressed faces)
-    \end{itemize}
-    \item<10-> With enough bad nodes, we can perform a \emph{surgery} that increases the number of leaves in $T_0$.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{The Trees $T_0$ and $T_1$}
-
-  \begin{center}
-   \begin{tabular}{cc}
-      \only<1>{\includegraphics[width=.45\textwidth]{figs/t0t1-1}}%
-      \only<2->{\includegraphics[width=.45\textwidth]{figs/t0t1-2}}%
-        &
-      \uncover<3->{\includegraphics[width=.45\textwidth]{figs/t0t1-3}} \\
-      \uncover<4->{\includegraphics[width=.45\textwidth]{figs/t0t1-4}} &
-      \uncover<5->{\includegraphics[width=.45\textwidth]{figs/t0t1-5}} \\
-   \end{tabular}
-  \end{center}
-\end{frame}
-
-%      \item For a node $u$ of $T_i$,
-%        \[ \rho_u \le 2(\kappa_u + \delta_u) \Leftrightarrow
-%           \tau_u \le 3\kappa_u + 2\delta_u \]
-%    \end{itemize}
-%    \item So if $\tau_u$ is big then either
-%    \begin{itemize}
-%       \item $\kappa_u$ is big; or
-%       \item $\delta_u$ is big
-%    \end{itemize}
-%  \end{itemize}
-%\end{frame}
-%
-
-\begin{frame}
-  \frametitle{Proof Idea}
-
-  \begin{itemize}
-    \only<1>{\item Define two trees $T_0$ (inside $C$) and $T_1$ (outside $C$)}
-    \only<2->{\item[$\checkmark$] Define two trees $T_0$ (inside $C$) and $T_1$ (outside $C$)}
-    \item Show that:
-    \begin{itemize}
-      \item \alert<3->{Each leaf of $T_i$ contains a \caressed\ face}
-      \item Each $T_i$ has $\Omega(\ell/\Delta)$ nodes
-    \end{itemize}
-    \item We are done if either $T_i$ has many leaves, so each $T_i$ must have few leaves
-    \begin{itemize}
-      \item[$\therefore$] each $T_i$ has many \emph{bad nodes} \\ (degree-2 and containing no caressed faces)
-    \end{itemize}
-    \item With enough bad nodes, we can perform a \emph{surgery} that increases the number of leaves in $T_0$.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Every Leaf $\kappa$aresses a Face}
-
-   \noindent\textbf{Lemma:}
-   Let $P$ be a \emph{chord path} for $C$ and let $L$ and $R$ be the two faces
-   of $P\cup C$ that each contain $P$ in their boundary. Then $R$
-   contains at least one face of $\dual{T}$ that is caressed by $C$.
-
-  \uncover<2->{\emph{Proof:} by induction on number of faces in $R$}
-  \begin{center}
-      \only<1,2>{\includegraphics[width=.45\textwidth]{figs/one-caressed-1}}%
-      \only<3,4>{\includegraphics[width=.45\textwidth]{figs/one-caressed-2}}
-      \only<5>{\includegraphics[width=.45\textwidth]{figs/one-caressed-3}}
-  \end{center}
-
-  \uncover<4->{
-   \noindent\textbf{Corollary:} Every leaf of $T_i$ contains at least one
-   caressed face.
-  }
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Proof Idea}
-
-  \begin{itemize}
-    \item[$\checkmark$] Define two trees $T_0$ (inside $C$) and $T_1$ (outside $C$)
-    \item Show that:
-    \begin{itemize}
-      \only<1>{\item Each leaf of $T_i$ contains a \caressed\ face}
-      \only<2->{\item[$\checkmark$] Each leaf of $T_i$ contains a \caressed\ face}
-      \item Each $T_i$ has $\Omega(\ell/\Delta)$ nodes
-    \end{itemize}
-    \item We are done if either $T_i$ has many leaves, so each $T_i$ must have few leaves
-    \begin{itemize}
-      \item[$\therefore$] each $T_i$ has many \emph{bad nodes} \\ (degree-2 and containing no caressed faces)
-    \end{itemize}
-    \item \alert<3->{With enough bad nodes, we can perform a \emph{surgery} that increases the number of leaves in $T_0$.}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Bad Nodes}
-
-  \begin{itemize}[<+->]
-    \item Recall: A node of $T_i$ is \emph{bad} if its has degree 2 and contains no caressed face.
-    \item[]\textbf{Lemma:} Every bad node contains exactly one face of $\dual{G}$
-    \item Recall: We have lots of bad nodes and very few leaves
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Surgery}
-
-  \begin{center}
-    \begin{tabular}{m{.3\textwidth}m{2em}m{.3\textwidth}}
-    \only<1>{\includegraphics[height=.7\textheight]{figs/cases-1}}%
-    \only<2->{\includegraphics[height=.7\textheight]{figs/cases-3}} &
-    \uncover<3->{$\Rightarrow$} &
-    \uncover<3->{\includegraphics[height=.7\textheight]{figs/cases-4}}
-    \end{tabular}
-  \end{center}
-  \begin{itemize}
-     \item<4-> increases number of leaves in $T_1$
-        (see $a_0$)
-     \item<5-> decreases length of $C$ by at most $\Delta^2$
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Surgery II}
-
-  \begin{center}
-    \begin{tabular}{m{.3\textwidth}m{2em}m{.3\textwidth}}
-    \only<1->{\includegraphics[height=.7\textheight]{figs/surgery-1}} &
-    \uncover<2->{$\Rightarrow$} &
-    \uncover<2->{\includegraphics[height=.7\textheight]{figs/surgery-2}}
-    \end{tabular}
-  \end{center}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Proof Idea}
-
-  \begin{itemize}
-    \item[$\checkmark$] Define two trees $T_0$ (inside $C$) and $T_1$ (outside $C$)
-    \item Show that:
-    \begin{itemize}
-      \item[$\checkmark$] Each leaf of $T_i$ contains a \caressed\ face
-      \item \alert<3>{Each $T_i$ has $\Omega(\ell/\Delta)$ nodes}
-    \end{itemize}
-    \item We are done if either $T_i$ has many leaves, so each $T_i$ must have few leaves
-    \begin{itemize}
-      \item[$\therefore$] each $T_i$ has many \emph{bad nodes} \\ (degree-2 and containing no caressed faces)
-    \end{itemize}
-    \only<1>{\item With enough bad nodes, we can perform a \emph{surgery} that increases the number of leaves in $T_0$.}
-    \only<2->{\item[$\checkmark$] With enough bad nodes, we can perform a \emph{surgery} that increases the number of leaves in $T_0$.}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{$T_i$ Has Many Nodes}
-
-  Situation so far:
-  \begin{itemize}[<+->]
-    \item Need to show: $T_i$ has many nodes
-    \item Danger: $T_i$ contains few nodes, each containing many $\tau$ouched faces
-    \item Will show that for any node $u$,
-      \[   \rho_u \le 2(\kappa_u + \delta_u)  \]
-     ($\rho_u$inched faces (inside $u$), $\kappa_u$aressed faces (inside $u$),
-        $\delta$egree of $u$)
-    \item Recall: $\tau_u=\kappa_u+\rho_u$
-    \item If $\tau_u$ is large then either:
-       \begin{enumerate}[<+->]
-          \item $\kappa_u$ is large (good, more $\kappa$aressed faces!); and/or
-          \item $\delta_u$ is large (good, more leaves!).
-       \end{enumerate}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-
-\textbf{Lemma:}
-Let $u$ be a node of $T_i$ {\color{gray}and let $\rho_u$, $\kappa_u$, and $\delta_u$ denote the number of pinched faces of $\dual{T}$ in $u$, the number of caressed faces of $\dual{T}$ in $u$, and the degree of $u$ in $T_i$, respectively}.
-Then $\rho_u \le 2(\kappa_u+\delta_u)$.
-
-\uncover<2->{\textit{Proof (by discharging):}}
-\begin{itemize}
-  \item<3-> Assign a charge of 2 to each $\rho$inched face
-  \item<4-> Move charge around so that
-  \begin{itemize}
-    \item<5->No charge is left on any pinched face
-    \item<6->Each caressed faces has a charge of at most 2
-    \item<7->Each outgoing edge has a charge of at most 2
-  \end{itemize}
-\end{itemize}
-\uncover<2->{
-\begin{center}
-                \begin{tabular}{cccc}
-                        \includegraphics[height=.22\textheight]{figs/discharge-2} &
-                        \includegraphics[height=.22\textheight]{figs/discharge-3} &
-                        \includegraphics[height=.22\textheight]{figs/discharge-4} &
-                        \includegraphics[height=.22\textheight]{figs/discharge-5}
-                \end{tabular}
-                \end{center}
-}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Proof Idea}
-
-  \begin{itemize}
-    \item[$\checkmark$] Define two trees $T_0$ (inside $C$) and $T_1$ (outside $C$)
-    \item Show that:
-    \begin{itemize}
-      \item[$\checkmark$] Each leaf of $T_i$ contains a \caressed\ face
-      \only<1>{\item Each $T_i$ has $\Omega(\ell/\Delta)$ nodes}
-      \only<2->{\item[$\checkmark$] Each $T_i$ has $\Omega(\ell/\Delta)$ nodes}
-    \end{itemize}
-    \item We are done if either $T_i$ has many leaves, so each $T_i$ must have few leaves
-    \begin{itemize}
-      \item[$\therefore$] each $T_i$ has many \emph{bad nodes} \\ (degree-2 and containing no caressed faces)
-    \end{itemize}
-    \item[$\checkmark$] With enough bad nodes, we can perform a \emph{surgery} that increases the number of leaves in $T_0$.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Conclusion}
-
-  \textbf{Theorem:} If $\dual{G}$ has a cycle of lenth $\ell$, then $\dual{G}$ has a cycle $C'$ that caresses $\Omega(\ell/\Delta^4)$ faces.
-
-  \textbf{Consequence:} Every $n$-vertex planar graph of maximum degree $\Delta$ has a free set of size $\Omega(n^{0.8}/\Delta^4)$.
-
-  \textbf{Open Problem:} Eliminate the dependence on $\Delta$.
-\end{frame}
-
-%\closing
-
 \end{document}

### 2021-11-19 slides/slides.tex
Small stuff

@@
     \textbf{Claim:} $\tw(H')\le\binom{k+t}{t}-1$
-
+    \vspace{1cm}
     \textit{Proof:}
@@
 \begin{frame}
+  \frametitle{Thickness}
+
   \begin{itemize}
-    \item each $S_x$ contains $O(1)$ vertices from each row of $H\boxtimes P$.
-(Easy, each vertex in $S_x$ from row $i$ comes from a shortcut through a vertex of $Y_x$ in row $i-1$, $i$, or $i+1$, so $3dk\ell(k+1)=3d\ell(k^2+k)$)
+    \item Recall that $\mathcal{P'}$ is a $(k,dk\ell)$-shortcut system over $H\boxtimes P$.
+    \item We can show that the subset of $H\boxtimes P\boxtimes K_{\ell}$ corresponding to $S_x$ contains at most $d\ell(k^2+3)$ vertices from each layer of $H\boxtimes P\boxtimes K_{\ell}$.
+    \item Implies $G^{\mathcal{P}}\subseteq H'\boxtimes P\boxtimes K_{d\ell(k^2+3)}$
   \end{itemize}

### 2021-11-19 slides/slides.tex
Pffft

@@
 \begin{frame}
-    \frametitle{Abstraction: Shortcut Systems}
+    \frametitle{$(k,d)$-Shortcut Systems}
+    \framesubtitle{An abstraction of $k$-planarity}
 
@@
         \end{itemize}
+    \end{itemize}
     \begin{center}
         \multiinclude[<+>][format=pdf,start=1]{figs/shortcut-system}%
+        \uncover<2->{\\ $(3,2)$-shortcut system}
     \end{center}
+    \begin{itemize}
         \item<3-> $G^{\mathcal{P}}:= G\cup\{vw:\mbox{$v$ and $w$ are endpoints of a path in $\mathcal{P}$}\}$
@@
 
-    If $G'$ is $k$-planar then $G'\subseteq G^{\mathcal{P}}$ where
+    If $G$ is $k$-planar then $G\subseteq G_0^{\mathcal{P}}$ where
     \begin{itemize}
-        \item $G$ is planar
-        \item $\mathcal{P}$ is a $(k+1,2)$-shortcut system
+        \item<2-> $G_0$ is planar
+        \item<3-> $\mathcal{P}$ is a $(k+1,2)$-shortcut system
     \end{itemize}
@@
 \begin{frame}
-  \frametitle{Proof: The Picture}
+  \frametitle{Mental Map of $H\boxtimes P\boxtimes K_{\ell}$}
   \begin{itemize}
-    \item $G= H\boxtimes P\boxtimes K_{\ell}$
+    \item $H\uncover<2->{\boxtimes P\uncover<3->{\boxtimes K_{\ell}}}$
   \end{itemize}
+  \begin{center}
+    \multiinclude[<+>][format=pdf,start=1]{figs/product-x}%
+  \end{center}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Shortcuts over multiple layers}
+  \frametitle{Shortcuts over multiple rows}
 
   \begin{itemize}
-    \item Shortcuts in $\mathcal{P}$ may jump $k$ layers
-    \item Easy solution: Compress $k$ layers into one
+    \item Shortcuts in $\mathcal{P}$ may jump $k$ rows
+    \item<2-> Easy solution: Compress $k$ rows into one
+  \end{itemize}
+    \begin{center}
+    \only<1>{\includegraphics{figs/layer-compression-1}}%
+    \only<2->{\includegraphics{figs/layer-compression-2}}%
+    \end{center}
   \begin{itemize}
-      \item Length $k-1$ path $P'$ in $P$ compresses down to vertex
-      \item $P'\boxtimes K_{\ell}\subseteq K_{\ell k}$
-    \end{itemize}
-    \item Increases $\ell$ by a factor of $k$
-    \item $\mathcal{P}$ is a $(k,d)$-shortcut system over $H\boxtimes P\boxtimes K_{\ell k}$ with all shortcuts crossing $\le\!\! 1$ layers
+    \item<3-> Increases $\ell$ by a factor of $k$
+    \item<4-> $\mathcal{P}$ is a $(k,d)$-shortcut system over $H\boxtimes P\boxtimes K_{\ell k}$ with all shortcuts crossing $\le\!\! 1$ layers
   \end{itemize}
@@
   \begin{itemize}
-    \item Contract each copy of $K_{k\ell}$, so we are left with $H\boxtimes P$
-    \item Each shortcut in $\mathcal{P}$ projects onto a walk in $H\boxtimes P$
-    \item Each vertex of $H\boxtimes P$ is used in at most $dk\ell$ shortcuts
-    \item Studying a $(k,dk\ell)$-shortcut system $\mathcal{P'}$ over $G'=H\boxtimes P$
+    \item<2-> Contract each copy of $K_{k\ell}$, so we are left with $H\boxtimes P$
+  \end{itemize}
+    \begin{center}
+    \only<1>{\includegraphics{figs/layer-compression-2}}%
+    \only<2->{\includegraphics{figs/layer-compression-3}}%
+    \end{center}
+  \begin{itemize}
+    \item<3-> Each vertex of $H\boxtimes P$ is used in at most $dk\ell$ shortcuts
+    \item<4-> $(k,dk\ell)$-shortcut system $\mathcal{P'}$ over $G'=H\boxtimes P$
+    \item<4-> shortcuts span $\le\!\!1$ rows
   \end{itemize}
@@
   \frametitle{Nice Tree Decompositions}
+  \framesubtitle{Nodes of $T$ are same as vertices of $H$}
 
@@
     \begin{center}
-      \includegraphics{figs/nice-td}
+      \includegraphics[scale=.8]{figs/nice-td}
     \end{center}
@@
   \begin{itemize}
-    \item<2-> Start with $B_x\gets\{x\}$ for each $x\in V(T)$
+    \item<1-> Start with $T$ such that $x\mathbin{\diamond}_T y$ for each $xy\in E(H)$
+    \item<2-> $B_x\gets\{x\}$ for each $x\in V(T)$
     \item<3-> If $xy\in E(H)$ and $x\prec_T y$ then drag $x$ down into $B_y$
@@
   \begin{center}
-    \includegraphics{figs/pz}%
+    \includegraphics[scale=.8]{figs/pz}%
   \end{center}
+  \begin{itemize}
+      \item<2->Problem: $vw$-shortcut in $H\boxtimes P$ projects onto $xz'$ shortcut in $H$, but $x\not\preceq_T z'$ and $z'\not\preceq_T x$
+  \end{itemize}
 \end{frame}
@@
   \begin{itemize}
-    \item For $v$ in $G$, $a(v)$ is the LCA of $\cup\{S_I:I\in\mathcal{P'},\, v\in V(I)\}$
+    \item For $v$ in $H\boxtimes P$, $a(v)$ is the LCA of $\cup\{S_I:I\in\mathcal{P'},\, v\in V(I)\}$
     \item For $x\in V(T)$, $S_x:=\{v\in V(G): a(v)=x\}$
     \begin{center}
-      \includegraphics{figs/nice-td-6}%
+      \includegraphics[scale=.8]{figs/nice-td-6}%
     \end{center}
-    \item Note: $\mathcal{S}:=(S_x:x\in V(T))$ is a partition of $V(G)$
+    \item<2-> $\mathcal{S}:=(S_x:x\in V(T))$ is a partition of $V(G)$
+    \item<3->[$(\star)$] For any $vw$-shortcut in $H\boxtimes P$, $a(v)\preceq_T a(w)$ or $a(w)\preceq_T a(v)$
 \end{itemize}
@@
 
+  \begin{itemize}
+      \item[$(\star)$] For any $vw$-shortcut in $H\boxtimes P$, $a(v)\preceq_T a(w)$ or $a(w)\preceq_T a(v)$
+  \end{itemize}
+
+    \vspace{.5cm}
     Let $H':=G/\mathcal{S}$
+    \vspace{.5cm}
 
-    \textbf{Claim:} $\tw(H')\le\binom{k+t}{t}-1$
-    \vspace{1cm}
-    \textit{Proof:}
+    \textbf{Claim:} $\tw(H')\le\binom{k+t}{t}-1$\\
+    \vspace{.5cm}
+    \uncover<2->{\textit{Proof:}}
     \begin{itemize}
-      \item Each $x\in V(H')$ is obtained by contracting $S_x$
-      \item Use tree nice tree decomposition with the same tree $T$
-      \item If $xz\in E(H')$ and $x\prec_T z$ then drag $x$ down into $B_z$
-      \item Point: If $x\in B_{z'}$ then $x\prec_T z'\prec_T z$ and $\mathop{dist}_{H}(x,z)\le k$.  
-      \item By Pilipczuk-Siebertz Lemma, $|B_{z'}|\le \binom{k+t}{t}$.
+      \item<2-> Each $x\in V(H')$ is obtained by contracting $S_x$
+      \item<3-> Start with $T$ from nice tree decomposition of $H$ (not $H'$)
+      \item<4-> For each $xz\in E(H')$ where $x\prec_T z$, drag $x$ down into $B_z$
+      \item<5-> Correctness follows from $(\star)$
+      \item<6-> Point: If $x\in B_{z'}$ then
+      \begin{itemize}
+          \item $x\prec_T z'\prec_T z$ and
+          \item $\mathop{dist}_{H}(x,z)\le k$ (because shortcuts have length $\le\!\!k$).
+      \end{itemize}
+      \item<7-> By Pilipczuk-Siebertz Lemma, $|B_{z'}|\le \binom{k+t}{t}$.
+      \hfill{\qed}
     \end{itemize}
@@
   \frametitle{Thickness}
+  \framesubtitle{Detail checking}
 
@@
     \item We can show that the subset of $H\boxtimes P\boxtimes K_{\ell}$ corresponding to $S_x$ contains at most $d\ell(k^2+3)$ vertices from each layer of $H\boxtimes P\boxtimes K_{\ell}$.
-    \item Implies $G^{\mathcal{P}}\subseteq H'\boxtimes P\boxtimes K_{d\ell(k^2+3)}$
+    \item Implies $G^{\mathcal{P}}\subseteq H'\boxtimes P\boxtimes K_{d\ell(k^3+3k)}$
   \end{itemize}
@@
 
+\begin{frame}
+
+\end{frame}
+
+% \begin{frame}
+%
+%
+% \end{frame}
+%
+
 \end{document}

### 2022-01-18 response.tex
started  revisions from ref reports

\documentclass[12pt]{article}
\listfiles
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,color}
\usepackage[margin=20mm]{geometry}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{amsfonts}
\usepackage{amsthm,mathtools}
\usepackage{paralist}
\usepackage{stmaryrd}

\setlength{\parindent}{0ex}
\setlength{\parskip}{1ex}

\renewcommand{\ge}{\geqslant}
\renewcommand{\le}{\leqslant}
\renewcommand{\geq}{\geqslant}
\renewcommand{\leq}{\leqslant}

\begin{document}

\subsection*{Graph Product Structure for Non-Minor-Closed Classes\\
Vida Dujmovi\'c, Pat Morin, David R. Wood\\
Journal of Combinatorial Theory Series B (JCTB9758)\\
Response to Referees}

\subsection*{Reviewer \#2}

 The Graph Product Structure Theorem of Dujmović, Joret,
Micek, Morin, Ueckerdt, and Wood (FOCS '19) asserts that every planar
graph is a subgraph of the strong product of a graph of treewidth at
most 8 and a path.  This has been the key tool in resolving several
longstanding open problems on planar graphs (and graphs in
minor-closed classes), including problems on queue layouts,
non-repetitive colouring, p-centered colouring, and universal graphs.
This paper extends the Graph Product Structure Theorem beyond
minor-closed classes.  The authors prove product structure theorems
for k-planar graphs, map graphs, string graphs, graph powers, and
nearest neighbour graphs.   The authors also present several
applications of their results.  These are essentially the same
applications as for the original Graph Product Structure Theorem; the
point being that the applications typically follow quite easily once a
product structure theorem is established.

This paper is a significant step in answering the (ambitious) question
of which graph classes admit product structure theorems. It is
extremely well-written (with very few typos) and was a pleasure to
read.  It is certainly well within the scope of JCTB and I strongly
recommend it for acceptance.

Below are some typos and suggestions for the authors.

Page 3.  The sentence 'For each $vw \in E(G')$ there is a path $P$ in $G$
between v and we of length at most k+1'' is correct but should be made
more precise.  For example, there could be a short path between v and
w which does not involve any new dummy vertices.  You mean to say that
the 'subdivided path' between v and w has length at most k+1.

Page 3.  You should add the assumption that no three edges cross at
the same point (which you can assume by perturbing the embedding).
Otherwise, you will not get a (k+1, 2)-shortcut system.

Page 3. P is used twice in the statement of Theorem 3.  It is probably
safer (and necessary?) to use a different variable for the second
occurrence.

Page 6. Regarding (T1), technically a subtree of a rooted tree can be
rooted at any vertex, so it is a bit imprecise to say that T[x] is
rooted at x.  I suggest adding that a subtree of a tree rooted at r,
is always considered to be rooted at the vertex closest to r.

Page 6. Replace 'that satisfies (T2)' by 'that satisfies (T1)'.

Page 6. Replace 'parent in $T'$ by 'parent in $T_0$'.

Page 6. The term 'hierarchical decomposition' is not defined.

Page 7. Although it is clear for me, 'separation' has not been
defined.  Perhaps a blanket note saying all undefined terms are in
Diestel's textbook should be added.

Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
(twice).  Moreover, the way the proof is written, it seems to suggest
that we only get that $a \in B_{z_i}$ for each $i \in \{1, \dots, r\}$.
	The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
		true though.  In a normalized tree decomposition, if $xy$ is an edge and
		$x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
		and $T_y$ intersect).

		Page 8.  Can you add a brief explanation why $\mathcal{S}$ is a
		partition of V(G)?  It is clear that the sets in $\mathcal{S}$ are
		disjoint, but why do they cover $V(G)$?

		Page 11. Replace 'belong to $F_\delta$' by 'belongs to $F_\delta$'.

		Page 11. In footnote 11, what the authors call a 'closed curve' I
		would call just a 'curve'.  To me, a closed curve satisfies the
		additional condition that f(0)=f(1).

		Page 13. Could it be that the subgraph induced by the vertices of a
		kite is a $K_4$ with some parallel edges?  This seems relevant later in
		the proof.

		Page 13. Perhaps I am misunderstanding something, but I do not see why
		'none of the edges vx, xw, wy, or yv are crossed by any other edges of
		G.'  For example, see the attached PDF for a picture where vw, xw, wy,
		and yv are all crossed by other edges of G.

		Page 13. The attached PDF also shows that the definition of 'kite
		face' may not be well-defined.  For example, I do not see why there
		cannot be some edges of G 'inside' a kite face.  Even if this is not
		the case, I think it is more precise to say that a kite face has 'two
		and a half edges' and two vertices of G on its boundary' rather than
		'three edges and two vertices of G on its boundary'.

		Page 13.  Regarding the proof of Lemma 4, I think it is better to
		include all the details from reference [14].  As far as I know, JCTB
		does not have a page limit, so I do not see an issue with just
		reproducing the entire proof and telling the reader that they can skip
		all the details if they wish.  This is just my personal opinion
		though, so the authors can ignore this request if they choose.

		Page 15. Regarding the proof of Theorem 11, if it really is the same
		proof as the proof of Theorem 2, why not just prove Theorem 11 (from
		which Theorem 2 follows as a special case)?

		Page 16.  At the end of Page 16, I do not see why 'every pair of edges
		cross at most once.'  Without further assumptions, it seems as if two
		paths in the path system could intersect several times.  Therefore, by
		drawing 'each edge vw of G alongside $P_vw$ in $G_0$', there may be two
		edges which cross several times.  It seems this can be fixed by
		choosing the path system carefully by 'uncrossing' paths which
		intersect at more than two internal vertices.

		Page 18. In point 1 in the proof of Lemma 9, I think '$\phi(w)$' should
		be '$\alpha(w)$'.

		Page 18.  Say $H'$ is connected since $X$ is connected.

		Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.

		Page 20.  Where does the bound $d(d-3)/2$ come from?  The naive bound I
		compute is $d(d-1)/2$.

		Page 21.  Replace the comma at the end of the statement of Theorem 14
		with a period.

		Page 23. Replace 'oberve' by 'observe'.

		\textcolor{blue}{Done}

		Page 24. Petr Hliněný (together with his student) has announced that
		the answer to the open question is yes, with $C=3$ and $\ell=O(k^2)$.

\end{document}

### 2022-01-18 k-planar.tex
started  revisions from ref reports

@@
 
-% \title{\MakeUppercase{The Structure of $k$-Planar Graphs}}
 \title{\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}}
@@
   Let $G$ be a $k$-nearest-neighbour graph and consider any edge $vw\in E(G)$.
-  Let $xy\in E(G)$ be an edge that crosses $vw$.  Note that $vxwy$ are the vertices of a convex quadrilateral in (without loss of generality) counterclockwise order. Then we say that
+  Let $xy\in E(G)$ be an edge that crosses $vw$.  Note that $vxwy$ are the vertices of a convex quadrilateral in (without loss of generality) counter-clockwise order. Then we say that
   \begin{compactenum}
@@
 
-  Next oberve that, if $xy$ is of Type~$v$, then at least one of $xv$ or $yv$ is in $E(G)$ in which case we call $x$ (respectively $y$) a Type~$v$ vertex.
+  Next observe that, if $xy$ is of Type~$v$, then at least one of $xv$ or $yv$ is in $E(G)$ in which case we call $x$ (respectively $y$) a Type~$v$ vertex.
 By \cref{k-nn-max-degree}, there are at most $6k$ Type~$v$ vertices.

### 2022-01-18 response.tex
added to responses

@@
 \setlength{\parskip}{1ex}
-
-\renewcommand{\ge}{\geqslant}
-\renewcommand{\le}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\renewcommand{\leq}{\leqslant}
 
@@
 
+\subsection*{Referee \#1}
+
+This paper proves that if a class of graphs F admits a product structure
+theorem, then the class of all graphs which have a 1-subdivision in F admits
+a product structure theorem as well. The main theorem, Theorem 3, while
+equivalent to the above statement using standard techniques, is stated in
+a different way which is considerably more useful for applications and for
+obtaining good bounds. As applications of Theorem 3, the paper shows that
+the following classes admit a product structure theorem:
+
+\begin{itemize}
+\item graphs drawn on a fixed surface with each edge in $\leq k$ crossings,
+\item map graphs of degree $\leq k$ graphs embedded in a fixed surface,
+\item string graphs with each curve in $\leq k$ intersections,
+\item $k$-nearest neighbour graphs in the plane, and
+\item powers of classes of graphs of degree $\leq k$ which admit a product structure theorem.
+\end{itemize}
+
+All but the last bullet point are shown to follow from the first.
+The main results in the paper are correct, strong, and of high interest,
+and the paper should definitely be accepted to JCTB after some
+(fairly significant) revisions. In particular, the proof of Theorem 3 is
+difficult to read due to the notation, and the paper could be substantially
+streamlined without losing the above results. Detailed suggestions follow.
+
+Major comments:
+
+1.
+Lemma 3: This lemma is really about “normalised” tree decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$.
+Call this property (T3) and remove (Y1)  through (Y5), which should now be obvious.
+
+2. Proof of Theorem 9: There is too much notation throughout. As
+one example, it is hard to remember which of $T_x$, $Y_x$, $V_x$, $F_x$, $N_x$,
+$S_x$, $X_v$, $B_x$, $C_x$ is which. If $\phi(S)$ is written for the subset of $V(G)$
+corresponding to a set $S \subseteq V(H)$ and appropriate shorthand and
+standard notation for neighbourhoods are used, then $Y_x$, $V_x$, $F_x$, and
+$N_x$ can be replaced by $\phi(x)$, $\phi(T_x)$, $N(\phi(T_x))$, and $V(G) \setminus N[\phi(T_x)]$.
+The set $X_v$ is only used in the statement of Claim~1, so it does not
+need a name; say what these vertices are explicitly. It may be more
+indicative to write $\phi'(x)$ for $S_x$ and $B'_x$ for $C_x$.
+
+3. Theorem 10: Unfortunately I have to suggest removing it. I don’t see
+the motivation for making H planar or for improving the constants.
+Are the improved constants at all close to tight? Is there another
+reason for including it?
+
+4. Sections 4 and 5: I think as they are currently organized they distract from the purpose of the paper, as stated in the introduction, to
+“prove product structure theorems for several non-minor-closed classes
+of interest.” Which of Corollaries 1–4 and Theorems 12–17 are most
+important? I would remove any mention of the applications from the
+examples section, put that section first, and state in that section a single main theorem with all of the best bounds on the product structure.
+The applications section is mostly a survey and I think should be
+focused on the most exciting new corollaries. Theorems 18 and 19
+should be in the applications section.
+
+Other comments:
+
+1.
+abstract: “This leads to analogous results for map graphs, string graphs, graph powers, and nearest neighbour graphs.” $\longrightarrow$ Add “certain types of” or be more specific.
+Same comment for immediately after Theorem 4.
+
+2. page 1: “A tree-decomposition T consists of a tree T and a collection
+T ” $\longrightarrow$ Delete the first T and define a T-decomposition.
+
+3. pages 2 – 4: The paragraph following Theorem 2 and Section 1.3
+(except for the fact that the apex assumption is necessary) is repetition
+of things discussed earlier on in the introduction.
+
+4. page 6, line 14: “Thus, it suffices to construct width-t tree decomposition that satisfies (T2).” $\longrightarrow$ should be (T1)
+
+5. page 6, line 16: “Select any node $x \in V (T_0)$” $\longrightarrow$ “Select any nonempty node $x \in V (T_0)$”
+
+6. page 6, last paragraph of Lemma 2: This isn’t quite right since you
+have to be careful with x0.
+
+7. page 8, item (i): “S has small layered width with respect to the layering
+L” $\longrightarrow$ “S has small layered width with respect to the layering L of
+G”. Perhaps mention something about how it is easy to find the new
+layering.
+
+8. general comment: Claims 1, 2, and 4 are especially belaboured by
+notation as they are quite straightforward.
+
+9. page 8, line 10: “We say that a vertex $w \in Y_x$ contributes a vertex
+$v \in S_x$ if $v$ participates in $x$.” This needs to be fixed.
+
+10. page 10, Claim 5: Indices could mostly be removed by continuing
+to talk about T-ancestors instead. Move the definition of H+ and its
+directed version up, along with the explanation of what will be proven,
+so that it isn’t necessary to define the si
+.
+11. page 10, Claim 5: The name Pww0 is strange since in case 2. there is
+no w
+0 defined.
+
+12. page 11, first paragraph of Section 3: “This section applies our main
+results for shortcut systems to prove graph product structure theorem...” $\longrightarrow$ “to prove a graph product structure theorem...”
+
 \end{document}

### 2022-01-18 response.tex
added responses

@@
 Journal of Combinatorial Theory Series B (JCTB9758)\\
-Response to Referees}
-
-\subsection*{Reviewer \#2}
-
- The Graph Product Structure Theorem of Dujmović, Joret,
-Micek, Morin, Ueckerdt, and Wood (FOCS '19) asserts that every planar
-graph is a subgraph of the strong product of a graph of treewidth at
-most 8 and a path.  This has been the key tool in resolving several
-longstanding open problems on planar graphs (and graphs in
-minor-closed classes), including problems on queue layouts,
-non-repetitive colouring, p-centered colouring, and universal graphs.
-This paper extends the Graph Product Structure Theorem beyond
-minor-closed classes.  The authors prove product structure theorems
-for k-planar graphs, map graphs, string graphs, graph powers, and
-nearest neighbour graphs.   The authors also present several
-applications of their results.  These are essentially the same
-applications as for the original Graph Product Structure Theorem; the
-point being that the applications typically follow quite easily once a
-product structure theorem is established.
-
-This paper is a significant step in answering the (ambitious) question
-of which graph classes admit product structure theorems. It is
-extremely well-written (with very few typos) and was a pleasure to
-read.  It is certainly well within the scope of JCTB and I strongly
-recommend it for acceptance.
-
-Below are some typos and suggestions for the authors.
-
-Page 3.  The sentence 'For each $vw \in E(G')$ there is a path $P$ in $G$
-between v and we of length at most k+1'' is correct but should be made
-more precise.  For example, there could be a short path between v and
-w which does not involve any new dummy vertices.  You mean to say that
-the 'subdivided path' between v and w has length at most k+1.
-
-Page 3.  You should add the assumption that no three edges cross at
-the same point (which you can assume by perturbing the embedding).
-Otherwise, you will not get a (k+1, 2)-shortcut system.
-
-Page 3. P is used twice in the statement of Theorem 3.  It is probably
-safer (and necessary?) to use a different variable for the second
-occurrence.
-
-Page 6. Regarding (T1), technically a subtree of a rooted tree can be
-rooted at any vertex, so it is a bit imprecise to say that T[x] is
-rooted at x.  I suggest adding that a subtree of a tree rooted at r,
-is always considered to be rooted at the vertex closest to r.
-
-Page 6. Replace 'that satisfies (T2)' by 'that satisfies (T1)'.
-
-Page 6. Replace 'parent in $T'$ by 'parent in $T_0$'.
-
-Page 6. The term 'hierarchical decomposition' is not defined.
-
-Page 7. Although it is clear for me, 'separation' has not been
-defined.  Perhaps a blanket note saying all undefined terms are in
-Diestel's textbook should be added.
-
-Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
-(twice).  Moreover, the way the proof is written, it seems to suggest
-that we only get that $a \in B_{z_i}$ for each $i \in \{1, \dots, r\}$.
-	The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
-		true though.  In a normalized tree decomposition, if $xy$ is an edge and
-		$x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
-		and $T_y$ intersect).
-
-		Page 8.  Can you add a brief explanation why $\mathcal{S}$ is a
-		partition of V(G)?  It is clear that the sets in $\mathcal{S}$ are
-		disjoint, but why do they cover $V(G)$?
-
-		Page 11. Replace 'belong to $F_\delta$' by 'belongs to $F_\delta$'.
-
-		Page 11. In footnote 11, what the authors call a 'closed curve' I
-		would call just a 'curve'.  To me, a closed curve satisfies the
-		additional condition that f(0)=f(1).
-
-		Page 13. Could it be that the subgraph induced by the vertices of a
-		kite is a $K_4$ with some parallel edges?  This seems relevant later in
-		the proof.
-
-		Page 13. Perhaps I am misunderstanding something, but I do not see why
-		'none of the edges vx, xw, wy, or yv are crossed by any other edges of
-		G.'  For example, see the attached PDF for a picture where vw, xw, wy,
-		and yv are all crossed by other edges of G.
-
-		Page 13. The attached PDF also shows that the definition of 'kite
-		face' may not be well-defined.  For example, I do not see why there
-		cannot be some edges of G 'inside' a kite face.  Even if this is not
-		the case, I think it is more precise to say that a kite face has 'two
-		and a half edges' and two vertices of G on its boundary' rather than
-		'three edges and two vertices of G on its boundary'.
-
-		Page 13.  Regarding the proof of Lemma 4, I think it is better to
-		include all the details from reference [14].  As far as I know, JCTB
-		does not have a page limit, so I do not see an issue with just
-		reproducing the entire proof and telling the reader that they can skip
-		all the details if they wish.  This is just my personal opinion
-		though, so the authors can ignore this request if they choose.
-
-		Page 15. Regarding the proof of Theorem 11, if it really is the same
-		proof as the proof of Theorem 2, why not just prove Theorem 11 (from
-		which Theorem 2 follows as a special case)?
-
-		Page 16.  At the end of Page 16, I do not see why 'every pair of edges
-		cross at most once.'  Without further assumptions, it seems as if two
-		paths in the path system could intersect several times.  Therefore, by
-		drawing 'each edge vw of G alongside $P_vw$ in $G_0$', there may be two
-		edges which cross several times.  It seems this can be fixed by
-		choosing the path system carefully by 'uncrossing' paths which
-		intersect at more than two internal vertices.
-
-		Page 18. In point 1 in the proof of Lemma 9, I think '$\phi(w)$' should
-		be '$\alpha(w)$'.
-
-		Page 18.  Say $H'$ is connected since $X$ is connected.
-
-		Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.
-
-		Page 20.  Where does the bound $d(d-3)/2$ come from?  The naive bound I
-		compute is $d(d-1)/2$.
-
-		Page 21.  Replace the comma at the end of the statement of Theorem 14
-		with a period.
-
-		Page 23. Replace 'oberve' by 'observe'.
-
-		\textcolor{blue}{Done}
-
-		Page 24. Petr Hliněný (together with his student) has announced that
-		the answer to the open question is yes, with $C=3$ and $\ell=O(k^2)$.
+Response to Referee Reports}
 
@@
 11. page 10, Claim 5: The name Pww0 is strange since in case 2. there is
-no w
-0 defined.
+no w 0 defined.
 
-12. page 11, first paragraph of Section 3: “This section applies our main
-results for shortcut systems to prove graph product structure theorem...” $\longrightarrow$ “to prove a graph product structure theorem...”
+12.
+page 11, first paragraph of Section 3: “This section applies our main results for shortcut systems to prove graph product structure theorem...” $\longrightarrow$ “to prove a graph product structure theorem...”
+
+\textcolor{blue}{done}
+
+\subsection*{Referee \#2}
+
+The Graph Product Structure Theorem of Dujmović, Joret,
+Micek, Morin, Ueckerdt, and Wood (FOCS '19) asserts that every planar
+graph is a subgraph of the strong product of a graph of treewidth at
+most 8 and a path.  This has been the key tool in resolving several
+longstanding open problems on planar graphs (and graphs in
+minor-closed classes), including problems on queue layouts,
+non-repetitive colouring, p-centered colouring, and universal graphs.
+This paper extends the Graph Product Structure Theorem beyond
+minor-closed classes.  The authors prove product structure theorems
+for k-planar graphs, map graphs, string graphs, graph powers, and
+nearest neighbour graphs.   The authors also present several
+applications of their results.  These are essentially the same
+applications as for the original Graph Product Structure Theorem; the
+point being that the applications typically follow quite easily once a
+product structure theorem is established.
+
+This paper is a significant step in answering the (ambitious) question
+of which graph classes admit product structure theorems. It is
+extremely well-written (with very few typos) and was a pleasure to
+read.  It is certainly well within the scope of JCTB and I strongly
+recommend it for acceptance.
+
+Below are some typos and suggestions for the authors.
+
+Page 3.  The sentence 'For each $vw \in E(G')$ there is a path $P$ in $G$
+between v and we of length at most $k+1$'' is correct but should be made
+more precise.  For example, there could be a short path between $v$ and
+$w$ which does not involve any new dummy vertices.  You mean to say that
+the 'subdivided path' between $v$ and $w$ has length at most $k+1$.
+
+Page 3.  You should add the assumption that no three edges cross at
+the same point (which you can assume by perturbing the embedding).
+Otherwise, you will not get a $(k+1, 2)$-shortcut system.
+
+Page 3. $P$ is used twice in the statement of Theorem 3.  It is probably
+safer (and necessary?) to use a different variable for the second
+occurrence.
+
+Page 6. Regarding (T1), technically a subtree of a rooted tree can be
+rooted at any vertex, so it is a bit imprecise to say that $T[x]$ is
+rooted at $x$.  I suggest adding that a subtree of a tree rooted at $r$,
+is always considered to be rooted at the vertex closest to $r$.
+
+Page 6. Replace 'that satisfies (T2)' by 'that satisfies (T1)'.
+
+Page 6. Replace 'parent in $T'$ by 'parent in $T_0$'.
+
+Page 6. The term 'hierarchical decomposition' is not defined.
+
+Page 7. Although it is clear for me, 'separation' has not been
+defined.  Perhaps a blanket note saying all undefined terms are in
+Diestel's textbook should be added.
+
+Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
+(twice).  Moreover, the way the proof is written, it seems to suggest
+that we only get that $a \in B_{z_i}$ for each $i \in \{1, \dots, r\}$.
+The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
+true though.  In a normalized tree decomposition, if $xy$ is an edge and
+$x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
+and $T_y$ intersect).
+
+Page 8.  Can you add a brief explanation why $\mathcal{S}$ is a
+partition of V(G)?  It is clear that the sets in $\mathcal{S}$ are
+disjoint, but why do they cover $V(G)$?
+
+Page 11. Replace 'belong to $F_\delta'$ by 'belongs to $F_\delta$'.
+
+Page 11. In footnote 11, what the authors call a 'closed curve' I
+would call just a 'curve'.  To me, a closed curve satisfies the
+additional condition that $f(0)=f(1)$.
+
+Page 13. Could it be that the subgraph induced by the vertices of a
+kite is a $K_4$ with some parallel edges?  This seems relevant later in
+the proof.
+
+Page 13. Perhaps I am misunderstanding something, but I do not see why
+'none of the edges vx, xw, wy, or yv are crossed by any other edges of
+G.'  For example, see the attached PDF for a picture where vw, xw, wy,
+and yv are all crossed by other edges of G.
+
+Page 13. The attached PDF also shows that the definition of 'kite
+face' may not be well-defined.  For example, I do not see why there
+cannot be some edges of G 'inside' a kite face.  Even if this is not
+the case, I think it is more precise to say that a kite face has 'two
+and a half edges' and two vertices of G on its boundary' rather than
+'three edges and two vertices of G on its boundary'.
+
+Page 13.  Regarding the proof of Lemma 4, I think it is better to
+include all the details from reference [14].  As far as I know, JCTB
+does not have a page limit, so I do not see an issue with just
+reproducing the entire proof and telling the reader that they can skip
+all the details if they wish.  This is just my personal opinion
+though, so the authors can ignore this request if they choose.
+
+Page 15. Regarding the proof of Theorem 11, if it really is the same
+proof as the proof of Theorem 2, why not just prove Theorem 11 (from
+which Theorem 2 follows as a special case)?
+
+Page 16.  At the end of Page 16, I do not see why 'every pair of edges
+cross at most once.'  Without further assumptions, it seems as if two
+paths in the path system could intersect several times.  Therefore, by
+drawing 'each edge vw of G alongside $P_vw$ in $G_0$', there may be two
+edges which cross several times.  It seems this can be fixed by
+choosing the path system carefully by 'uncrossing' paths which
+intersect at more than two internal vertices.
+
+Page 18. In point 1 in the proof of Lemma 9, I think '$\phi(w)$' should
+be '$\alpha(w)$'.
+
+Page 18.  Say $H'$ is connected since $X$ is connected.
+
+Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.
+
+Page 20.  Where does the bound $d(d-3)/2$ come from?  The naive bound I
+compute is $d(d-1)/2$.
+
+Page 21.  Replace the comma at the end of the statement of Theorem 14
+with a period.
+
+\textcolor{blue}{done}
+
+Page 23. Replace 'oberve' by 'observe'.
+
+\textcolor{blue}{done}
+
+Page 24. Petr Hliněný (together with his student) has announced that
+the answer to the open question is yes, with $C=3$ and $\ell=O(k^2)$.
+
+\textcolor{blue}{We look forward to reading the paper.}
 

### 2022-01-18 k-planar.tex
added responses

@@
 
-This section applies our main results for shortcut systems to prove graph product structure theorem for graphs drawn with a bounded number of crossings per edge.
+This section applies our main results for shortcut systems to prove a product structure theorem for graphs drawn with a bounded number of crossings per edge.
 Then we show how to improve the bounds in this case.
@@
 	\item is a subgraph of $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$,
-	\item has queue-number $\qn(G^k) \leq 3 \ell \, 2^t - \ceil{\tfrac{3}{2}\ell}$,
+	\item has queue-number $\qn(G^k) \leq 3 \ell \, 2^t - \ceil{\tfrac{3}{2}\ell}$.
 \end{compactitem}
@@
 
-% % NOTE THAT MANUAL CHANGES HAVE BEEN MADE BELOW.
-%
+% NOTE THAT MANUAL CHANGES HAVE BEEN MADE BELOW.
+
 \begin{thebibliography}{37}
@@
 	\newblock Planar graphs have bounded queue-number.
-	\newblock In David Zuckerman, editor, \emph{Proc. 60th {IEEE} Annual Symp.
-		Foundations Comput. Sci.} (FOCS 2019), pages 862--875. IEEE, 2019.
-	\newblock \doi{10.1109/FOCS.2019.00056}.
-	\newblock To appear in \emph{J. ACM}, \arXiv{1904.04791}.
+	\newblock \emph{J. ACM} 67(4):22, 2020.
+	\newblock \doi{10.1145/3385731}.
 
@@
 	\newblock Adjacency labelling for planar graphs (and beyond).
-	\newblock 2020{\natexlab{a}}.
-	\newblock \arXiv{2003.04280}.
+	\newblock \emph{J. ACM} 68(6):42, 2021. 
+	\newblock \doi{10.1145/3477542}.
 
@@
 	\newblock Clustered 3-colouring graphs of bounded degree.
-	\newblock 2020.
-	\newblock \arXiv{2002.11721}.
+\newblock \emph{Combin. Probab. Comput.} 31(1):123--135, 2022.
+\newblock \doi{10.1017/S0963548321000213}.
 

### 2022-03-03 response.tex
DW: added all the referee comments to the paper in pretty blue boxes

@@
 \documentclass[12pt]{article}
-\listfiles
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
-\usepackage{amsmath,color}
 \usepackage[margin=20mm]{geometry}
-\usepackage{amsfonts}
-\usepackage{amsthm}
-\usepackage{graphicx}
-\usepackage{enumerate}
-\usepackage{amsfonts}
-\usepackage{amsthm,mathtools}
-\usepackage{paralist}
-\usepackage{stmaryrd}
+\usepackage{amsmath,color,amsthm,enumerate,amsfonts,mathtools}
 
@@
 1.
-Lemma 3: This lemma is really about “normalised” tree decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$.
+Lemma 3: This lemma is really about ``normalised'' tree decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$.
 Call this property (T3) and remove (Y1)  through (Y5), which should now be obvious.
@@
 3. Theorem 10: Unfortunately I have to suggest removing it. I don’t see
-the motivation for making H planar or for improving the constants.
+the motivation for making $H$ planar or for improving the constants.
 Are the improved constants at all close to tight? Is there another
@@
 
-4. Sections 4 and 5: I think as they are currently organized they distract from the purpose of the paper, as stated in the introduction, to
-“prove product structure theorems for several non-minor-closed classes
-of interest.” Which of Corollaries 1–4 and Theorems 12–17 are most
-important? I would remove any mention of the applications from the
-examples section, put that section first, and state in that section a single main theorem with all of the best bounds on the product structure.
-The applications section is mostly a survey and I think should be
-focused on the most exciting new corollaries. Theorems 18 and 19
-should be in the applications section.
+4.
+Sections 4 and 5: I think as they are currently organized they distract from the purpose of the paper, as stated in the introduction, to ``prove product structure theorems for several non-minor-closed classes of interest.'' Which of Corollaries 1--4 and Theorems 12--17 are most important?
+I would remove any mention of the applications from the examples section, put that section first, and state in that section a single main theorem with all of the best bounds on the product structure.
+The applications section is mostly a survey and I think should be focused on the most exciting new corollaries.
+Theorems 18 and 19 should be in the applications section.
 
@@
 
-1.
-abstract: “This leads to analogous results for map graphs, string graphs, graph powers, and nearest neighbour graphs.” $\longrightarrow$ Add “certain types of” or be more specific.
-Same comment for immediately after Theorem 4.
+1. abstract: ``This leads to analogous results for map graphs, string graphs, graph powers, and nearest neighbour graphs.'' $\longrightarrow$ 
+Add ``certain types of'' or be more specific. Same comment for immediately after Theorem 4.
 
-2. page 1: “A tree-decomposition T consists of a tree T and a collection
-T ” $\longrightarrow$ Delete the first T and define a T-decomposition.
+\textcolor{blue}{done}
+
+2. page 1: ``A tree-decomposition $T$ consists of a tree $T$ and a collection
+$T$ '' $\longrightarrow$ Delete the first $T$ and define a $T$-decomposition.
 
@@
 
-4. page 6, line 14: “Thus, it suffices to construct width-t tree decomposition that satisfies (T2).” $\longrightarrow$ should be (T1)
+4. page 6, line 14: ``Thus, it suffices to construct width-$t$ tree decomposition that satisfies (T2).'' $\longrightarrow$ should be (T1)
 
-5. page 6, line 16: “Select any node $x \in V (T_0)$” $\longrightarrow$ “Select any nonempty node $x \in V (T_0)$”
+5. page 6, line 16: ``Select any node $x \in V (T_0)$'' $\longrightarrow$ ``Select any nonempty node $x \in V (T_0)$''
 
 6. page 6, last paragraph of Lemma 2: This isn’t quite right since you
-have to be careful with x0.
+have to be careful with $x_0$.
 
-7. page 8, item (i): “S has small layered width with respect to the layering
-L” $\longrightarrow$ “S has small layered width with respect to the layering L of
-G”. Perhaps mention something about how it is easy to find the new
-layering.
+7. page 8, item (i): ``$S$ has small layered width with respect to the layering
+L'' $\longrightarrow$ ``$S$ has small layered width with respect to the layering $L$ of
+$G$''. Perhaps mention something about how it is easy to find the new layering.
 
@@
 
-9. page 8, line 10: “We say that a vertex $w \in Y_x$ contributes a vertex
-$v \in S_x$ if $v$ participates in $x$.” This needs to be fixed.
+9. page 8, line 10: ``We say that a vertex $w \in Y_x$ contributes a vertex
+$v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.
 
 10. page 10, Claim 5: Indices could mostly be removed by continuing
-to talk about T-ancestors instead. Move the definition of H+ and its
+to talk about $T$-ancestors instead. Move the definition of $H+$ and its
 directed version up, along with the explanation of what will be proven,
-so that it isn’t necessary to define the si
-.
-11. page 10, Claim 5: The name Pww0 is strange since in case 2. there is
-no w 0 defined.
+so that it isn’t necessary to define the $s_i$.
+
+11. page 10, Claim 5: The name $Pww'$ is strange since in case 2. there is
+no $w'$ defined.
 
 12.
-page 11, first paragraph of Section 3: “This section applies our main results for shortcut systems to prove graph product structure theorem...” $\longrightarrow$ “to prove a graph product structure theorem...”
+page 11, first paragraph of Section 3: ``This section applies our main results for shortcut systems to prove graph product structure theorem...'' $\longrightarrow$ ``to prove a graph product structure theorem...''
 
@@
 minor-closed classes), including problems on queue layouts,
-non-repetitive colouring, p-centered colouring, and universal graphs.
+non-repetitive colouring, $p$-centered colouring, and universal graphs.
 This paper extends the Graph Product Structure Theorem beyond
@@
 Page 3.  The sentence 'For each $vw \in E(G')$ there is a path $P$ in $G$
-between v and we of length at most $k+1$'' is correct but should be made
+between $v$ and we of length at most $k+1$'' is correct but should be made
 more precise.  For example, there could be a short path between $v$ and
@@
 Page 8.  Can you add a brief explanation why $\mathcal{S}$ is a
-partition of V(G)?  It is clear that the sets in $\mathcal{S}$ are
+partition of $V(G)$?  It is clear that the sets in $\mathcal{S}$ are
 disjoint, but why do they cover $V(G)$?
@@
 Page 11. Replace 'belong to $F_\delta'$ by 'belongs to $F_\delta$'.
+
+\textcolor{blue}{done}
 
@@
 Page 13. Perhaps I am misunderstanding something, but I do not see why
-'none of the edges vx, xw, wy, or yv are crossed by any other edges of
-G.'  For example, see the attached PDF for a picture where vw, xw, wy,
-and yv are all crossed by other edges of G.
+'none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of
+$G$.'  For example, see the attached PDF for a picture where $vw$, $xw$, $wy$,
+and $yv$ are all crossed by other edges of $G$.
+
+\includegraphics[width=60mm]{kite}
 
@@
 face' may not be well-defined.  For example, I do not see why there
-cannot be some edges of G 'inside' a kite face.  Even if this is not
+cannot be some edges of $G$ 'inside' a kite face.  Even if this is not
 the case, I think it is more precise to say that a kite face has 'two
-and a half edges' and two vertices of G on its boundary' rather than
-'three edges and two vertices of G on its boundary'.
+and a half edges' and two vertices of $G$ on its boundary' rather than
+'three edges and two vertices of $G$ on its boundary'.
 
@@
 
-\textcolor{blue}{We look forward to reading the paper.}
+\textcolor{blue}{We have been told there is an error in their proof.}
 

### 2022-03-03 k-planar.tex
DW: added all the referee comments to the paper in pretty blue boxes

@@
 \newcommand{\notex}[2]{}
+\newcommand{\referee}[2]{\noindent\textcolor{blue}{\framebox{\begin{minipage}{\textwidth} Ref \#{#1}: #2\end{minipage}}}}
 
@@
 In fact, we work in a much more general setting based on so-called shortcut systems that are of independent interest.
-This leads to analogous results for map graphs, string graphs, graph powers, and nearest neighbour graphs.
+This leads to analogous results for certain types of map graphs, string graphs, graph powers, and nearest neighbour graphs.
 \end{abstract}
@@
 \end{thm}
+
+\referee{1}{2. page 1: ``A tree-decomposition $T$ consists of a tree $T$ and a collection
+$T$ '' $\longrightarrow$ Delete the first $T$ and define a $T$-decomposition.}
 
@@
 
+\referee{1}{Page 3.  You should add the assumption that no three edges cross at
+the same point (which you can assume by perturbing the embedding).
+Otherwise, you will not get a $(k+1, 2)$-shortcut system.}
+
 We start with the example of $k$-planar graphs.
@@
 Prior to the recent work of \citet{dujmovic.esperet.ea:planar}, it was even open whether planar graphs have bounded non-repetitive chromatic number.
+
+\referee{1}{3. pages 2 – 4: The paragraph following Theorem 2 and Section 1.3
+(except for the fact that the apex assumption is necessary) is repetition
+of things discussed earlier on in the introduction.}
 
@@
 
+\referee{2}{Page 3.  The sentence 'For each $vw \in E(G')$ there is a path $P$ in $G$
+between $v$ and we of length at most $k+1$'' is correct but should be made
+more precise.  For example, there could be a short path between $v$ and
+$w$ which does not involve any new dummy vertices.  You mean to say that
+the 'subdivided path' between $v$ and $w$ has length at most $k+1$.}
+
 The following theorem is the main contribution of the paper.
@@
 
+\referee{2}{Page 3. $P$ is used twice in the statement of Theorem 3.  It is probably
+safer (and necessary?) to use a different variable for the second
+occurrence.}
+
 Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$.
@@
 
-\cref{Examples} presents further examples of graph classes that can be constructed using shortcut systems, including map graphs, string graphs, and $k$-nearest neighbour graphs.
+\cref{Examples} presents further examples of graph classes that can be constructed using shortcut systems, including certain types of map graphs, string graphs, and $k$-nearest neighbour graphs.
 \cref{ShortcutProduct} implies product structure theorems for each of these classes.
@@
 
+\referee{2}{Page 6. Regarding (T1), technically a subtree of a rooted tree can be
+rooted at any vertex, so it is a bit imprecise to say that $T[x]$ is
+rooted at $x$.  I suggest adding that a subtree of a tree rooted at $r$,
+is always considered to be rooted at the vertex closest to $r$.}
+
 \begin{proof}
@@
 Thus, it suffices to construct width-$t$ tree decomposition that satisfies \tref{ancestor-edge}.
+
+\referee{2}{Page 6. Replace 'that satisfies (T2)' by 'that satisfies (T1)'.}
 
@@
 This modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
+
+\referee{2}{Page 6. Replace 'parent in $T$' by 'parent in $T_0$'.}
 
@@
 
+\referee{1}{6. page 6, last paragraph of Lemma 2: This isn’t quite right since you
+have to be careful with $x_0$.}
+
 %\subsection{Generalized Tripod Partitions}
@@
 The following lemma shows how to interpret an $H$-partition of $G$ and a tree decomposition of $H$ as a hierarchical decomposition of $G$; refer to \cref{fig:generalized-tripod}.
+
+\referee{2}{Page 6. The term 'hierarchical decomposition' is not defined.}
 
@@
 
+\referee{1}{1.
+Lemma 3: This lemma is really about ``normalised'' tree decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$.
+Call this property (T3) and remove (Y1) through (Y5), which should now be obvious.}
+
 Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:
@@
   Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
+
+\referee{2}{Page 7. Although it is clear for me, 'separation' has not been
+defined.  Perhaps a blanket note saying all undefined terms are in
+Diestel's textbook should be added.}
 
@@
 
+\referee{2}{Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
+(twice).  Moreover, the way the proof is written, it seems to suggest
+that we only get that $a \in B_{z_i}$ for each $i \in \{1, \dots, r\}$.
+The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
+true though.  In a normalized tree decomposition, if $xy$ is an edge and
+$x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
+and $T_y$ intersect).}
+
 We are now ready to prove our main result, which we restate here for convenience:
@@
 \mmg*
+
+\referee{1}{2. Proof of Theorem 9: There is too much notation throughout. As
+	one example, it is hard to remember which of $T_x$, $Y_x$, $V_x$, $F_x$, $N_x$,
+	$S_x$, $X_v$, $B_x$, $C_x$ is which. If $\phi(S)$ is written for the subset of $V(G)$
+	corresponding to a set $S \subseteq V(H)$ and appropriate shorthand and
+	standard notation for neighbourhoods are used, then $Y_x$, $V_x$, $F_x$, and
+	$N_x$ can be replaced by $\phi(x)$, $\phi(T_x)$, $N(\phi(T_x))$, and $V(G) \setminus N[\phi(T_x)]$.
+	The set $X_v$ is only used in the statement of Claim~1, so it does not
+	need a name; say what these vertices are explicitly. It may be more
+	indicative to write $\phi'(x)$ for $S_x$ and $B'_x$ for $C_x$.}
 
@@
 
+\referee{2}{Page 8.  Can you add a brief explanation why $\mathcal{S}$ is a
+partition of $V(G)$?  It is clear that the sets in $\mathcal{S}$ are
+disjoint, but why do they cover $V(G)$?}
+
 For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
@@
 
+\referee{1}{7. page 8, item (i): ``$S$ has small layered width with respect to the layering
+L'' $\longrightarrow$ ``$S$ has small layered width with respect to the layering $L$ of
+$G$''. Perhaps mention something about how it is easy to find the new layering.}
+
 First we need to understand the relationship between sets $S_x\in\mathcal{S}$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
+
+\referee{1}{8. general comment: Claims 1, 2, and 4 are especially belaboured by
+notation as they are quite straightforward.}
 
@@
 
+\referee{1}{9. page 8, line 10: ``We say that a vertex $w \in Y_x$ contributes a vertex
+  $v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.}
+
   \begin{figure}[htbp]
@@
 
+\referee{1}{10. page 10, Claim 5: Indices could mostly be removed by continuing
+to talk about $T$-ancestors instead. Move the definition of $H+$ and its
+directed version up, along with the explanation of what will be proven,
+so that it isn’t necessary to define the $s_i$.}
+
 \begin{proof}
@@
 
+\referee{1}{11. page 10, Claim 5: The name $Pww'$ is strange since in case 2. there is
+no $w'$ defined.}
+
   Let $w_0,w_1,\ldots,w_p$ be the path that begins $w_0:=v'$ and then follows the subpath of $P_{ww'}$ that begins at $w_1:=w$ and ends at $w_p:=w''$.
@@
   We claim that $a_p=x_\delta$---that is, $s_p=\delta$.
-  To see this, first observe that, for each $i\in\{1,\ldots,p\}$, $w_i\in V_{\delta}$ since, otherwise, an internal vertex of $P_{ww'}$ belong to $F_\delta$, which would imply (by \yref{ancestor-edge}) that $w\in S_{\delta'}$ for some $\delta' < \delta$, contradicting the assumption that $w\in S_\delta$.
+  To see this, first observe that, for each $i\in\{1,\ldots,p\}$, $w_i\in V_{\delta}$ since, otherwise, an internal vertex of $P_{ww'}$ belongs to $F_\delta$, which would imply (by \yref{ancestor-edge}) that $w\in S_{\delta'}$ for some $\delta' < \delta$, contradicting the assumption that $w\in S_\delta$.
 Therefore $s_p\ge\delta$.
@@
 
+\referee{2}{Page 18.  Say $H'$ is connected since $X$ is connected.}
+
   Thus, $a_0,\ldots,a_p$ is a lazy walk in $H^+$ of length $p\le k$ where the distance $s_i$ between $a_i$ and the root $x_0$ of $T$ is non-decreasing.
@@
 \subsection{$k$-Planar Graphs}
+
+\referee{2}{Page 11. In footnote 11, what the authors call a 'closed curve' I
+would call just a 'curve'.  To me, a closed curve satisfies the
+additional condition that $f(0)=f(1)$.}
 
@@
 
+\referee{1}{3. Theorem 10: Unfortunately I have to suggest removing it. I don’t see
+	the motivation for making $H$ planar or for improving the constants.
+	Are the improved constants at all close to tight? Is there another
+	reason for including it?}
+
 Let $G$ be an edge-maximal 1-plane multigraph with no two parallel edges on the boundary of a single face.
@@
 Observe that any spar of $K$ is incident on exactly four \emph{kite faces} of $G$, each of which has three edges and two vertices of $G$ on its boundary.
+
+\referee{2}{Page 13. Could it be that the subgraph induced by the vertices of a
+	kite is a $K_4$ with some parallel edges?  This seems relevant later in
+	the proof.}
+
+\referee{2}{Page 13. Perhaps I am misunderstanding something, but I do not see why
+'none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of
+$G$.'  For example, see the attached PDF for a picture where $vw$, $xw$, $wy$,
+and $yv$ are all crossed by other edges of $G$.\\
+\includegraphics[width=60mm]{kite}\\
+Page 13. The attached PDF also shows that the definition of 'kite
+face' may not be well-defined.  For example, I do not see why there
+cannot be some edges of $G$ 'inside' a kite face.  Even if this is not
+the case, I think it is more precise to say that a kite face has 'two
+and a half edges' and two vertices of $G$ on its boundary' rather than
+'three edges and two vertices of $G$ on its boundary'.
+}
 
@@
 
+\referee{2}{Page 13.  Regarding the proof of Lemma 4, I think it is better to
+include all the details from reference [14].  As far as I know, JCTB
+does not have a page limit, so I do not see an issue with just
+reproducing the entire proof and telling the reader that they can skip
+all the details if they wish.  This is just my personal opinion
+though, so the authors can ignore this request if they choose.}
+
 \begin{proof}
@@
 \end{thm}
+
+\referee{2}{Page 15. Regarding the proof of Theorem 11, if it really is the same
+	proof as the proof of Theorem 2, why not just prove Theorem 11 (from
+	which Theorem 2 follows as a special case)?}
 
@@
 
+\referee{2}{Page 16.  At the end of Page 16, I do not see why 'every pair of edges
+cross at most once.'  Without further assumptions, it seems as if two
+paths in the path system could intersect several times.  Therefore, by
+drawing 'each edge vw of G alongside $P_vw$ in $G_0$', there may be two
+edges which cross several times.  It seems this can be fixed by
+choosing the path system carefully by 'uncrossing' paths which
+intersect at more than two internal vertices.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 Here we discuss some of the consequences of the above theorems for $k$-planar and $(g,k)$-planar graphs.
+
+\referee{1}{4.
+Sections 4 and 5: I think as they are currently organized they distract from the purpose of the paper, as stated in the introduction, to ``prove product structure theorems for several non-minor-closed classes of interest.'' Which of Corollaries 1--4 and Theorems 12--17 are most important?
+I would remove any mention of the applications from the examples section, put that section first, and state in that section a single main theorem with all of the best bounds on the product structure.
+The applications section is mostly a survey and I think should be focused on the most exciting new corollaries.
+Theorems 18 and 19 should be in the applications section.}
 
@@
 
+\referee{2}{Page 18. In point 1 in the proof of Lemma 9, I think '$\phi(w)$' should
+be '$\alpha(w)$'.}
+
  First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$.
@@
 
+\referee{2}{Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.}
+
 \cref{p-centered,p-centered-treewidth,kPlanarProduct,1-planar,gkPlanarProduct} immediately imply the following results, for every $p\ge 2$:
@@
 \end{lem}
+
+\referee{2}{Page 20.  Where does the bound $d(d-3)/2$ come from?  The naive bound I
+compute is $d(d-1)/2$.}
 

### 2022-03-04 k-planar.tex
revised according to referee comments

@@
 \usepackage{thm-restate}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 
@@
 
+\newcommand{\defin}[1]{\textcolor{Maroon}{\emph{#1}}}
+
 \crefname{lem}{Lemma}{Lemmas}
@@
 
- \newcommand{\note}[2]{{\color{red}[#1:~#2]}}
+ \newcommand{\note}[2]{\noindent{\color{red}[#1:~#2]}}
 % \newcommand{\notex}[2]{{\color{blue}[#1:~#2]}}
@@
 \newcommand{\PP}{\mathcal{P}}
+\renewcommand{\SS}{\mathcal{S}}
 
@@
 \begin{abstract}
-Dujmovi\'c~et~al.~(FOCS 2019) recently proved that every planar graph is a subgraph of the strong product of a graph of bounded treewidth and a path.
+Dujmovi\'c~et~al.~[\emph{J.~ACM}~'20] recently proved that every planar graph is a subgraph of the strong product of a graph of bounded treewidth and a path.
 Analogous results were obtained for graphs of bounded Euler genus or apex-minor-free graphs.
@@
 We use $G-S$ as a shorthand for $G[V(G)\setminus S]$.
-We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.} by \citet{dujmovic.joret.ea:planar}.
-
-\begin{thm}[\citep{dujmovic.joret.ea:planar}]
+We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.
+Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20}.
+
+\begin{thm}[\citep{DJMMUW20}]
 	\label{PlanarProduct}
@@
 
-\referee{1}{2. page 1: ``A tree-decomposition $T$ consists of a tree $T$ and a collection
-$T$ '' $\longrightarrow$ Delete the first $T$ and define a $T$-decomposition.}
-
-Here $\boxtimes$ is the strong product\footnote{The \emph{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
+\note{DW}{In 	\cref{PlanarProduct} and elsewhere, we really should write, ``Every planar graph is isomorphic to a subgraph of''.
+One alternative is to define that $H$ is \defin{contained} in $G$ if $H$ is isomorphic to a subgraph of $G$.
+I suggest the latter.}
+
+Here $\boxtimes$ is the strong product\footnote{The \defin{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
 	%:\begin{compactitem}
@@
 	%\end{compactitem}
-}, and treewidth\footnote{A \emph{tree decomposition} $\mathcal{T}$ of a graph $G$ consists of a tree $T$ and a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that
-%:
-%\begin{compactenum}[(i)]
-%\item 
+}, and treewidth\footnote{For a tree $T$, a \defin{$T$-decomposition} of a graph $G$ is a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that
 (i) for every $vw\in E(G)$, there exists some node $x\in V(T)$ with $v,w\in B_x$; and
-%\item 
 (ii) for every $v\in V(G)$, the induced subgraph $T[v] := T[\{x: v\in B_x\}]$ is connected.
-%\end{compactenum}
-The \emph{width} of the tree decomposition $\mathcal{T}$ is $\max\{|B_x|:x\in V(T)\}-1$.
-The \emph{treewidth} $\tw(G)$ of a graph $G$ is the minimum width of a tree decomposition of $G$.
+The \defin{width} of $\mathcal{T}$ is $\max\{|B_x|:x\in V(T)\}-1$.
+A \defin{tree-decomposition} is a $T$-decomposition for any tree $T$.
+The \defin{treewidth} $\tw(G)$ of a graph $G$ is the minimum width of a tree-decomposition of $G$.
 Treewidth is the standard measure of how similar a graph is to a tree.
 Indeed, a connected graph has treewidth 1 if and only if it is a tree.
-Treewidth is of fundamental importance in structural and algorithmic graph theory; see \citep{Reed03,HW17,Bodlaender-TCS98} for surveys.} is an invariant that measures how `tree-like' a given graph is; see
-%\cref{Tools} for formal definitions and see 
-\cref{ProductExample} for an example.
+Treewidth is of fundamental importance in structural and algorithmic graph theory; see \citep{Reed03,HW17,Bodlaender-TCS98} for surveys.} is an invariant that measures how `tree-like' a given graph is; see \cref{ProductExample} for an example.
 Loosely speaking, \cref{PlanarProduct} says that every planar graph is contained in the product of a tree-like graph and a path.
@@
 \begin{compactitem}
-\item \citet{dujmovic.joret.ea:planar} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992).
+\item \citet{DJMMUW20} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992).
 \item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002).
-\item \citet{debski.felsner.ea:improved} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs. 
-\item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results going back to 1988~\citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}), and \citet{DEJGMM} have since used it to find asymptotically optimal
-adjacency labellings of planar graphs.
-\item The result of \citet{DEJGMM} implies that, for every integer
-$n>0$, there is a \emph{universal} graph $U_n$ with $n^{1+o(1)}$ 
+\item \citet{DFMS21} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs. 
+\item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results going back to 1988~\citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}), and \citet{DEJGMM21} have since used it to find asymptotically optimal adjacency labellings of planar graphs.
+\item The result of \citet{DEJGMM21} implies that, for every integer
+$n>0$, there is a `universal' graph $U_n$ with $n^{1+o(1)}$ 
 vertices such that every $n$-vertex planar graph is an induced
@@
 These applications motivate finding product structure theorems for other graph classes.
-\citet{dujmovic.joret.ea:planar} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
+\citet{DJMMUW20} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
 The \textit{Euler genus} of the non-orientable surface with $c$ cross-caps is $c$.
@@
 Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.} and for apex-minor-free graphs\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
-A class $\mathcal{G}$ of graphs is \emph{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$.
-A minor-closed class is \emph{proper} if it is not the class of all graphs.
+A class $\mathcal{G}$ of graphs is \defin{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$.
+A minor-closed class is \defin{proper} if it is not the class of all graphs.
 For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class.
 A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar.
-A 1-apex graph is \emph{apex}.
-A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.}, and \citet{DEMWW} do so for graphs in any minor-closed class and with bounded maximum degree.
-See \cref{Generalisations} for more precise statements and see \citep{DHGLW} for a survey on this topic.
+A 1-apex graph is \defin{apex}.
+A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.}, and \citet{DEMWW22} do so for graphs in any minor-closed class and with bounded maximum degree.
+See \cref{Generalisations} for more precise statements and see \citep{DHJLW21} for a survey on this topic.
 
@@
 
-\referee{1}{Page 3.  You should add the assumption that no three edges cross at
-the same point (which you can assume by perturbing the embedding).
-Otherwise, you will not get a $(k+1, 2)$-shortcut system.}
-
 We start with the example of $k$-planar graphs.
-A graph is \emph{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings.
+A graph is \defin{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings, where no three edges cross at a single point (see \cref{sec-k-planar} for a formal definition).
 Such graphs provide a natural generalisation of planar graphs, and are important in graph drawing research; see the recent bibliography on 1-planar graphs and the 140 references therein \citep{kobourov.liotta.ea:annotated}.
@@
 
-This theorem has applications in diverse areas, including queue layouts  \citep{dujmovic.joret.ea:planar}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, $p$-centered colouring  \citep{debski.felsner.ea:improved}, and adjacency labelling \citep{DEJGMM}, which we explore in \cref{Applications}.
+This theorem has applications in diverse areas, including queue layouts  \citep{DJMMUW20}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, $p$-centered colouring  \citep{DFMS21}, and adjacency labelling \citep{DEJGMM21}, which we explore in \cref{Applications}.
 For example, we prove that $k$-planar graphs have bounded non-repetitive chromatic number (for fixed $k$).
@@
 Although $k$-planar graphs are the most high-profile target for a generalization of \cref{PlanarProduct}, we actually prove a substantially stronger result than \cref{kPlanarProduct} using the following definition.
-A non-empty set $\mathcal{P}$ of non-trivial paths in a graph $G$ is a \emph{$(k,d)$-shortcut system} (for $G$) if:
+A non-empty set $\mathcal{P}$ of non-trivial paths in a graph $G$ is a \defin{$(k,d)$-shortcut system} (for $G$) if:
 \begin{compactitem}
 \item every path in $\mathcal{P}$ has length at most $k$,\footnote{A path of length $k$ consists of $k$ edges and $k+1$ vertices.
-A path is \emph{trivial} if it has length 0 and \emph{non-trivial} otherwise.} and
+A path is \defin{trivial} if it has length 0 and \defin{non-trivial} otherwise.} and
 \item for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
 \end{compactitem}
-Each path $P\in\mathcal{P}$ is called a \emph{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \emph{$vw$-shortcut}.
+Each path $P\in\mathcal{P}$ is called a \defin{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \defin{$vw$-shortcut}.
 Given a graph $G$ and a $(k,d)$-shortcut system $\mathcal{P}$ for $G$, let $G^{\mathcal{P}}$ denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\mathcal{P}$.
@@
 The proof of \cref{AddDummy} is trivial: Given a $k$-plane embedding of a graph $G'$, create a planar graph $G$ by adding a dummy vertex at each crossing point.
-For each edge $vw\in E(G')$ there is a path $P$ in $G$ between $v$ and $w$ of length at most $k+1$.
+For each edge $vw\in E(G')$ there is a path $P$ in $G$ between $v$ and $w$ of length at most $k+1$ in which every internal vertex is a dummy vertex.
 Let $\PP$ be the set of such paths $P$.
@@
 
-\referee{2}{Page 3.  The sentence 'For each $vw \in E(G')$ there is a path $P$ in $G$
-between $v$ and we of length at most $k+1$'' is correct but should be made
-more precise.  For example, there could be a short path between $v$ and
-$w$ which does not involve any new dummy vertices.  You mean to say that
-the 'subdivided path' between $v$ and $w$ has length at most $k+1$.}
-
 The following theorem is the main contribution of the paper.
@@
 
+\note{DW}{Use $\SS$ for shortcut system, okay?}
+
 Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$.
@@
 Here is one example.
-The \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.\footnote{For a graph $G$ and two vertices $v,w\in V(G)$, $\dist_G(v,w)$  denotes the length of a shortest path, in $G$, with endpoints $v$ and $w$.
+The \defin{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.\footnote{For a graph $G$ and two vertices $v,w\in V(G)$, $\dist_G(v,w)$  denotes the length of a shortest path, in $G$, with endpoints $v$ and $w$.
 We define $\dist_G(v,w):=\infty$ if $v$ and $w$ are in different connected components of $G$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\PP$ for some $(k,2k\Delta^{k})$-shortcut system $\PP$; see \cref{PowerShortcut}.
@@
 
-\begin{thm}[\citep{dujmovic.joret.ea:planar}]
+\begin{thm}[\citep{DJMMUW20}]
 	\label{GenusProduct}
@@
 
-\citet{dujmovic.joret.ea:planar} generalised \cref{GenusProduct} for apex-minor-free graphs as follows.
-
-\begin{thm}[\citep{dujmovic.joret.ea:planar}]
+\citet{DJMMUW20} generalised \cref{GenusProduct} for apex-minor-free graphs as follows.
+
+\begin{thm}[\citep{DJMMUW20}]
 	\label{ApexMinorFree}
@@
 
-The assumption that $X$ is apex is needed in \cref{ApexMinorFree}, since if the class of $X$-minor-free graphs has a product structure theorem analogous to \cref{PlanarProduct}, then $X$ is apex \citep{dujmovic.joret.ea:planar}.
-On the other hand, \citet{DEMWW} proved a product structure theorem for bounded degree graphs in any minor-closed class.
-
-\begin{thm}[\citep{DEMWW}]
+The assumption that $X$ is apex is needed in \cref{ApexMinorFree}, since if the class of $X$-minor-free graphs has a product structure theorem analogous to \cref{PlanarProduct}, then $X$ is apex \citep{DJMMUW20}.
+On the other hand, \citet{DEMWW22} proved a product structure theorem for bounded degree graphs in any minor-closed class.
+
+\begin{thm}[\citep{DEMWW22}]
 	\label{MinorFreeDegree}
@@
 
-A \emph{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle$ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
-For any partition $\mathcal{P}=\{S_1,\ldots,S_p\}$ of $V(G)$, a \emph{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
-To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an \emph{$H$-partition} and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.
-
-For any partition $\mathcal{P}$ of $V(G)$ and any layering $\mathcal{L}$ of $G$ we define the \emph{layered width} of $\mathcal{P}$ with respect to $\mathcal{L}$ as $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
-For any partition $\mathcal{P}$ of $V(G)$, we define the \emph{layered width} of $\mathcal{P}$ as the minimum, over all layerings $\mathcal{L}$ of $G$, of the layered width of $\mathcal{P}$ with respect to $\mathcal{L}$.
-
-\citet{dujmovic.joret.ea:planar} introduced the study of partitions with bounded layered width such that the quotient has some additional desirable property, like small treewidth.
-Dujmovi\'c~et~al.\ define a class $\mathcal{G}$ of graphs to \emph{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
+A \defin{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle$ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
+For any partition $\mathcal{P}=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
+To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an \defin{$H$-partition} and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.
+
+For any partition $\mathcal{P}$ of $V(G)$ and any layering $\mathcal{L}$ of $G$ we define the \defin{layered width} of $\mathcal{P}$ with respect to $\mathcal{L}$ as $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
+For any partition $\mathcal{P}$ of $V(G)$, we define the \defin{layered width} of $\mathcal{P}$ as the minimum, over all layerings $\mathcal{L}$ of $G$, of the layered width of $\mathcal{P}$ with respect to $\mathcal{L}$.
+
+\citet{DJMMUW20} introduced the study of partitions with bounded layered width such that the quotient has some additional desirable property, like small treewidth.
+Dujmovi\'c~et~al.\ define a class $\mathcal{G}$ of graphs to \defin{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
 
@@
 
-\begin{lem}[\citep{dujmovic.joret.ea:planar}]
+\begin{lem}[\citep{DJMMUW20}]
 \label{PartitionProduct}
@@
 As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a), 
-\citet{dujmovic.joret.ea:planar} showed that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$.
+\citet{DJMMUW20} showed that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$.
 The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.
-At the core of their work is the elegant proof by \citet{pilipczuk.siebertz:polynomial-soda} of the following result:
-
-\begin{thm}[\citep{pilipczuk.siebertz:polynomial-soda}]
+At the core of their work is the elegant proof by \citet{PS21} of the following result:
+
+\begin{thm}[\citep{PS21}]
 \label{ps}
@@
 
-Indeed, the above-mentioned result of \citet{dujmovic.joret.ea:planar} is a slight strengthening of \cref{ps}, where for each $P\in\mathcal{P}$ no two vertices of $P$ have the same distance to some fixed root vertex $r$.
+Indeed, the above-mentioned result of \citet{DJMMUW20} is a slight strengthening of \cref{ps}, where for each $P\in\mathcal{P}$ no two vertices of $P$ have the same distance to some fixed root vertex $r$.
 
@@
 
-For a tree $T$ rooted at some node $x_0\in V(T)$, we say that a node $a\in V(T)$ is a \emph{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \emph{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
+Consider a tree $T$ rooted at some node $x_0\in V(T)$.
+Each subtree $T'$ of $T$ is considered to be rooted at the node in $T'$ closest to $x_0$.
+A node $a\in V(T)$ is a \defin{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \defin{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
 Note that each node $x\in V(T)$ is a $T$-ancestor and $T$-descendant of itself.
-We say that a $T$-ancestor $a\in V(T)$ of $x\in V(T)$ is a \emph{strict} $T$-ancestor of $x$ if $a\neq x$.
-The \emph{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.  For each node $x\in V(T)$, define
+We say that a $T$-ancestor $a\in V(T)$ of $x\in V(T)$ is a \defin{strict} $T$-ancestor of $x$ if $a\neq x$.
+The \defin{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.  For each node $x\in V(T)$, define
 \[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
@@
 
-We begin with a standard technique that allows us to work with a normalised tree decomposition:
+We begin with a standard technique that allows us to work with a normalised tree-decomposition:
 
@@
 
-\referee{2}{Page 6. Regarding (T1), technically a subtree of a rooted tree can be
-rooted at any vertex, so it is a bit imprecise to say that $T[x]$ is
-rooted at $x$.  I suggest adding that a subtree of a tree rooted at $r$,
-is always considered to be rooted at the vertex closest to $r$.}
-
 \begin{proof}
   That \tref{subtree-root} implies \tref{ancestor-edge} is a standard observation: If two subtrees intersect, then one contains the root of the other.
-Thus, it suffices to construct width-$t$ tree decomposition that satisfies \tref{ancestor-edge}.
-
-\referee{2}{Page 6. Replace 'that satisfies (T2)' by 'that satisfies (T1)'.}
-
-  Begin with any width-$t$ tree decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some tree $T_0$.
+Thus, it suffices to construct a width-$t$ tree-decomposition that satisfies \tref{subtree-root}.
+
+  Begin with any width-$t$ tree-decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some tree $T_0$.
 Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$.
@@
 If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.
-Subdivide the edge between $a$ and its parent in $T$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
-This modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
-
-\referee{2}{Page 6. Replace 'parent in $T$' by 'parent in $T_0$'.}
+Subdivide the edge between $a$ and its parent in $T_0$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
+This modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
 
@@
 In this case, contract the edge $aa'$ in $T_0$, eliminating the node $a$.
-Repeating this operation will eventually produce a width-$t$ tree decomposition of $(B_x:x\in V(T_0))$ where $f$ is a bijection between $V(H)$ and $V(T_0)$.
-Renaming each node $a\in V(T_0)$ as $f^{-1}(a)$ gives a tree decomposition $(B_x:x\in V(T))$ with $V(T)=V(H)$.
-By the definition of $f$, the tree decomposition $(B_x:x\in V(T))$ satisfies \tref{subtree-root}.
+Repeating this operation will eventually produce a width-$t$ tree-decomposition of $(B_x:x\in V(T_0))$ where $f$ is a bijection between $V(H)$ and $V(T_0)$.
+Renaming each node $a\in V(T_0)$ as $f^{-1}(a)$ gives a tree-decomposition $(B_x:x\in V(T))$ with $V(T)=V(H)$.
+By the definition of $f$, the tree-decomposition $(B_x:x\in V(T))$ satisfies \tref{subtree-root}.
   % To see that $(B_x:x\in V(T))$ satisifies \tref{ancestor-edge}, observe that,
@@
 
-The following lemma shows how to interpret an $H$-partition of $G$ and a tree decomposition of $H$ as a hierarchical decomposition of $G$; refer to \cref{fig:generalized-tripod}.
-
-\referee{2}{Page 6. The term 'hierarchical decomposition' is not defined.}
+The following lemma shows how to interpret an $H$-partition of $G$ and a tree-decomposition of $H$ as a `hierarchical' decomposition of $G$; refer to \cref{fig:generalized-tripod}.
 
 \begin{lem}\label{generalized-tripod}
-  Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree decomposition of $H$ satisfying the conditions of \lemref{nice-decomposition}.
+  Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \lemref{nice-decomposition}.
 For each $x\in V(T)$, let $V_x := \bigcup_{y\in V(T_x)} Y_y$, $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x:=V_x\cup F_x$.
@@
 \referee{1}{1.
-Lemma 3: This lemma is really about ``normalised'' tree decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$.
+Lemma 3: This lemma is really about ``normalised'' tree-decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$.
 Call this property (T3) and remove (Y1) through (Y5), which should now be obvious.}
@@
   Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
-
-\referee{2}{Page 7. Although it is clear for me, 'separation' has not been
-defined.  Perhaps a blanket note saying all undefined terms are in
-Diestel's textbook should be added.}
 
@@
 The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
-true though.  In a normalized tree decomposition, if $xy$ is an edge and
+true though.  In a normalized tree-decomposition, if $xy$ is an edge and
 $x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
@@
 
-For a node $x\in V(T)$, we say that a shortcut $P\in\mathcal{P}$ \emph{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
-We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$, or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
+For a node $x\in V(T)$, we say that a shortcut $P\in\mathcal{P}$ \defin{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
+We say that a vertex $v\in V(G)$ \defin{participates} in $x$ if $v\in Y_x$, or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
 We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
@@
   Recall that $S_x$ is defined by vertices that participate in $x$, and these are vertices that are either in $Y_x$ or in shortcuts that cross $Y_x$.
-We say that a vertex $w\in Y_x$ \emph{contributes} a vertex $v\in S_x$ if $v$ participates in $x$.
+We say that a vertex $w\in Y_x$ \defin{contributes} a vertex $v\in S_x$ if $v$ participates in $x$.
   We upper bound the number of vertices in $S_x\cap L_i$ by upper-bounding the number of vertices contributed to $S_x\cap L_i$ by each $w\in Y_x$.
@@
   $v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.}
+
+\note{DW}{What is wrong here?}
 
@@
 We now proceed with Step~(ii), showing that $J$ has small treewidth.
-To accomplish this, we construct a small width tree-decomposition $\mathcal{C}:=(C_x:x\in V(T))$ of $J$ using the same tree $T$ used in the tree decomposition $\mathcal{T}:=(B_x:x\in V(T))$ of $H$.
+To accomplish this, we construct a small width tree-decomposition $\mathcal{C}:=(C_x:x\in V(T))$ of $J$ using the same tree $T$ used in $\mathcal{T}$.
 The following claim will be useful in showing that the resulting decomposition has small width.
@@
 	\label{general-bag-size}
-  The graph $J$ has a tree decomposition in which every bag has size at most $\binom{k+t}{t}$.
+  The graph $J$ has a tree-decomposition in which every bag has size at most $\binom{k+t}{t}$.
 \end{clm}
@@
 \begin{proof}
-  For the tree decomposition $(C_x:x\in V(T))$ of $J$ we use the same tree $T$ used in the tree decomposition $(B_x:x\in V(T))$ of $H$.
+  For the tree-decomposition $(C_x:x\in V(T))$ of $J$ we use the same tree $T$ used in the tree-decomposition $(B_x:x\in V(T))$ of $H$.
 For each node $x$ of $T$, we define $C_x$ as follows: $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
@@
 The connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ follows from the fact that, for every node $x'\in T[a]$, every node $x$ on the path in $T$ from $x'$ to  $a$ is also a node of $T[a]$.
-Therefore $(C_x:x\in V(T))$ is indeed a tree decomposition of $J$.
+Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $J$.
 It remains to bound the size of each bag $C_x$.
@@
     \item There exists $w'\in V(G)$ such that $\mathcal{P}$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
-    \item $w\in Y_\delta$.  In this case, we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$.
+    \item $w\in Y_\delta$; in this case, define $w':=w$ and let $P_{ww'}$ be the 1-vertex path $w$.
   \end{compactenum}
   Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{\delta}$.
-
-\referee{1}{11. page 10, Claim 5: The name $Pww'$ is strange since in case 2. there is
-no $w'$ defined.}
 
@@
   % Note that, by \tref{ancestor-edge}, $xy\in E(H^+)$ if and only if $x\in B_y$ or $y\in B_x$.
-  We claim that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \emph{lazy walk} in a graph $H$ is a walk in the pseudograph $H'$ obtained by adding a loop at each vertex of $H$.} in $H^+$.
+  We claim that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \defin{lazy walk} in a graph $H$ is a walk in the pseudograph $H'$ obtained by adding a loop at each vertex of $H$.} in $H^+$.
 Indeed, if $a_i\neq a_{i+1}$ for some $i\in\{0,\ldots,p-1\}$ then this is precisely because $w_i\in V_{a_i}$ but $w_{i+1}\not\in V_{a_i}$.
@@
 By removing repeated vertices this gives a path in the directed graph $\overrightarrow{H}^+$ obtained by directing each edge $xy\in E(H^+)$ from its $T$-descendant $x$ towards its $T$-ancestor $y$.
-  Finally, we are in a position to appeal to \cite[Lemma~24]{pilipczuk.siebertz:polynomial-soda} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
+  Finally, we are in a position to appeal to \cite[Lemma~24]{PS21} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
 
@@
 
-\referee{2}{Page 11. In footnote 11, what the authors call a 'closed curve' I
-would call just a 'curve'.  To me, a closed curve satisfies the
-additional condition that $f(0)=f(1)$.}
-
 We first formally define $k$-planar graphs.
-An \emph{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a closed curve\footnote{A closed curve in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
-The points $f(0)$ and $f(1)$ are called the \emph{endpoints} of the curve.
+An \defin{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a curve\footnote{A \defin{curve} in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
+The points $f(0)$ and $f(1)$ are called the \defin{endpoints} of the curve.
 When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
-A \emph{crossing} in an embedded graph $G$ is a triple $(p,vw,xy)$ with $p\in\R^2$, $vw,xy\in E(G)$ and such that $p\in (vw\cap xy)\setminus\{v,w,x,y\}$.
-An embedded graph $G$ is \emph{$k$-plane} if each edge of $G$ takes part in at most $k$ crossings.
-A (not necessarily embedded) graph $G'$ is \emph{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
+A \defin{crossing} in an embedded graph $G$ is a triple $(p,vw,xy)$ with $p\in\R^2$, $vw,xy\in E(G)$ and such that $p\in (vw\cap xy)\setminus\{v,w,x,y\}$.
+An embedded graph $G$ is \defin{$k$-plane} if each edge of $G$ takes part in at most $k$ crossings.
+A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
 Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
@@
 
-\referee{1}{3. Theorem 10: Unfortunately I have to suggest removing it. I don’t see
-	the motivation for making $H$ planar or for improving the constants.
-	Are the improved constants at all close to tight? Is there another
-	reason for including it?}
+\referee{1}{3.
+Theorem 10: Unfortunately I have to suggest removing it.
+I don’t see the motivation for making $H$ planar or for improving the constants.
+Are the improved constants at all close to tight?
+Is there another reason for including it?}
+
+\note{DW}{We can motivate \cref{1-planar} since treewidth 3 is best possible, and the planarity of $H$ is essential in the proof of \cref{1PlanarQueue}.
+Shall we argue this in the response?
+Or do we remove \cref{sec-1-planar}?
+We could keep the statement of \cref{1-planar} in the paper and refer to the arXiv paper for the proof.
+What do you think?
+All this depends on whether we can fix the errors below.
+}
 
 Let $G$ be an edge-maximal 1-plane multigraph with no two parallel edges on the boundary of a single face.
-Here, edge-maximal should be taken to mean that, if any two vertices $v$ and $w$ appear on a common face\footnote{The \emph{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
+Here, edge-maximal should be taken to mean that, if any two vertices $v$ and $w$ appear on a common face\footnote{The \defin{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
 We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.
@@
 
-A \emph{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.
+A \defin{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.
 It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
-The edges $vw$ and $xy$ are called \emph{spars} of $K$.
-The cycle $vxwy$ is called the \emph{sail} of $K$.
+The edges $vw$ and $xy$ are called \defin{spars} of $K$.
+The cycle $vxwy$ is called the \defin{sail} of $K$.
 It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$.
-Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
-Observe that any spar of $K$ is incident on exactly four \emph{kite faces} of $G$, each of which has three edges and two vertices of $G$ on its boundary.
-
-\referee{2}{Page 13. Could it be that the subgraph induced by the vertices of a
-	kite is a $K_4$ with some parallel edges?  This seems relevant later in
-	the proof.}
 
@@
 and $yv$ are all crossed by other edges of $G$.\\
-\includegraphics[width=60mm]{kite}\\
-Page 13. The attached PDF also shows that the definition of 'kite
+\includegraphics[width=60mm]{kite}}
+\note{DW}{I agree with the referee.}
+Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
+\note{DW}{This is wrong I think} 
+Observe that any spar of $K$ is incident on exactly four \defin{kite faces} of $G$, each of which has three edges and two vertices of $G$ on its boundary.
+
+\referee{2}{Page 13. The attached PDF also shows that the definition of 'kite
 face' may not be well-defined.  For example, I do not see why there
@@
 and a half edges' and two vertices of $G$ on its boundary' rather than
-'three edges and two vertices of $G$ on its boundary'.
-}
+'three edges and two vertices of $G$ on its boundary'.}
+\note{DW}{I agree with the referee.}
+
+\referee{2}{Page 13. Could it be that the subgraph induced by the vertices of a
+	kite is a $K_4$ with some parallel edges?  This seems relevant later in
+	the proof.}
+
+\note{DW}{I agree with the referee.}
 
@@
 
-Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analagous result for plane graphs \cite{dujmovic.joret.ea:planar}.
+Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analagous result for plane graphs \cite{DJMMUW20}.
 \begin{lem}
@@
 		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; and
-		\item $H$ has a tree decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
+		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
 	\end{compactenum}
@@
 \begin{proof}
-	This proof is very similar to the proof of Lemma~14 by \citet{dujmovic.joret.ea:planar}.
+	This proof is very similar to the proof of Lemma~14 by \citet{DJMMUW20}.
 Rather than duplicate every detail of that proof here, we focus on the differences and refer the reader to the original proof for the remaining details.
@@
 If $k=3$, set $R_i := P_i$ for each $i\in\{1,2,3\}$.
-Otherwise, as in \citep{dujmovic.joret.ea:planar}, split $P_1,\ldots,P_k$ to partition $V(F)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a non-empty path and each $R_i$ contains vertices from exactly one of $P_1,\ldots,P_k$.
-
-	Next, as in \citep{dujmovic.joret.ea:planar}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
+Otherwise, as in \citep{DJMMUW20}, split $P_1,\ldots,P_k$ to partition $V(F)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a non-empty path and each $R_i$ contains vertices from exactly one of $P_1,\ldots,P_k$.
+
+	Next, as in \citep{DJMMUW20}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
 
@@
 	Finally, the subgraph $N_i$ of $G$ consisting of the edges and vertices of $G$ contained in $F_i$ or its interior does not contain one of the three vertices of $\tau$.
-Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(C_i)$ to obtain the desired $H$-partition and tree decomposition of $N_i$.
-
-	The proof finishes in the same way as the proof in \cite{dujmovic.joret.ea:planar}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
+Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(C_i)$ to obtain the desired $H$-partition and tree-decomposition of $N_i$.
+
+	The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
 	Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
-The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{dujmovic.joret.ea:planar}, except that now each node $x$ has a child for each face $F_i$ of $M^+_x$ that contains a vertex of $G$ in its interior.
+The desired tree-decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{DJMMUW20}, except that now each node $x$ has a child for each face $F_i$ of $M^+_x$ that contains a vertex of $G$ in its interior.
 
@@
 Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs (or any of the other classes presented in \cref{Examples}) was in terms of layered treewidth, which we now define.
-A \emph{layered tree decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree decomposition $\mathcal{T}$ of $G$.
+A \defin{layered tree-decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{T}$ of $G$.
 The layered width of $(\mathcal{L},\mathcal{T})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{T}\}$.
-The \emph{layered treewidth} of $G$ is the minimum layered width of any layered tree decomposition of $G$.
+The \defin{layered treewidth} of $G$ is the minimum layered width of any layered tree-decomposition of $G$.
 \citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph.
@@
 If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
-In particular, if $\mathcal{P}=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{dujmovic.joret.ea:planar}.
+In particular, if $\mathcal{P}=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree-decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree-decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{DJMMUW20}.
 Therefore, any property that holds for graphs of bounded layered treewidth also holds for $G$.
@@
 \begin{proof}
-	For a vertex $x$ of $G_0$ with degree at most $\Delta$, and for $i\in\{1,\dots,k-1\}$, say a vertex $v$ is \emph{$i$-close} to $x$ if there is a $vx$-path $P$ in $G_0$ of length at most $i$ such that every internal vertex in $P$  has degree at most $\Delta$ in $G_0$.
-	For each edge $vw$ of $G$, say that $vw$ \emph{passes through} each internal vertex on $P_{vw}$.
+	For a vertex $x$ of $G_0$ with degree at most $\Delta$, and for $i\in\{1,\dots,k-1\}$, say a vertex $v$ is \defin{$i$-close} to $x$ if there is a $vx$-path $P$ in $G_0$ of length at most $i$ such that every internal vertex in $P$  has degree at most $\Delta$ in $G_0$.
+	For each edge $vw$ of $G$, say that $vw$ \defin{passes through} each internal vertex on $P_{vw}$.
 	Say $vw$ passes through $x$. Then $v$ is $i$-close to $x$ and $w$ is $j$-close to $x$ for some $i,j\in\{1,\dots,k-1\}$ with $i+j\leq k$. At most $\Delta^{i}$ vertices are $i$-close to $x$.
@@
 Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
-\citet{dujmovic.joret.ea:planar} recently proved this conjecture using \cref{PlanarProduct} and the following lemma.
+\citet{DJMMUW20} recently proved this conjecture using \cref{PlanarProduct} and the following lemma.
 Indeed, resolving this question was the motivation for the development of \cref{PlanarProduct}.
 
-\begin{lem}[\citep{dujmovic.joret.ea:planar}]
+\begin{lem}[\citep{DJMMUW20}]
 \label{qn}
@@
 \cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^4)}$.
-Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of
+Note that \citet{DJMMUW20} previously proved the bound of
 $O(g^{k+2})$ using \cref{GenusProduct} and an ad-hoc method.
@@
 \begin{cor}
+	\label{1PlanarQueue}
 Every 1-planar graph has queue-number at most $3 \times 30 \times 5 + \floor{\tfrac{3}{2} \times 30} = 495$. 
@@
 
-The next two applications are in the field of graph colouring. For our purposes, a \emph{$c$-colouring} of a graph $G$ is any function $\phi\colon V(G)\to C$, where $C$ is a set of size at most $c$.
-A $c$-colouring $\phi$ of $G$ is \emph{non-repetitive} if, for every path $v_1,\ldots,v_{2h}$ in $G$, there exists $i\in\{1,\ldots,h\}$ such that $\phi(v_i)\neq\phi(v_{i+h})$.
-The \emph{non-repetitive chromatic number} $\pi(G)$ of $G$ is the minimum integer $c$ such that $G$ has a non-repetitive $c$-colouring.
+The next two applications are in the field of graph colouring. For our purposes, a \defin{$c$-colouring} of a graph $G$ is any function $\phi\colon V(G)\to C$, where $C$ is a set of size at most $c$.
+A $c$-colouring $\phi$ of $G$ is \defin{non-repetitive} if, for every path $v_1,\ldots,v_{2h}$ in $G$, there exists $i\in\{1,\ldots,h\}$ such that $\phi(v_i)\neq\phi(v_{i+h})$.
+The \defin{non-repetitive chromatic number} $\pi(G)$ of $G$ is the minimum integer $c$ such that $G$ has a non-repetitive $c$-colouring.
 This concept, introduced by \citet{AGHR-RSA02}, has since been widely studied; see \citep{dujmovic.esperet.ea:planar} for more than 40 references.
@@
 
-A $c$-colouring $\phi$ of $G$ is \emph{$p$-centered} if, for every connected subgraph $X\subseteq G$, $|\{\phi(v):v\in V(X)\}| > p$ or there exists some $v\in V(X)$ such that $\phi(v)\neq \phi(w)$ for every $w\in V(X)\setminus\{v\}$.
+A $c$-colouring $\phi$ of $G$ is \defin{$p$-centered} if, for every connected subgraph $X\subseteq G$, $|\{\phi(v):v\in V(X)\}| > p$ or there exists some $v\in V(X)$ such that $\phi(v)\neq \phi(w)$ for every $w\in V(X)\setminus\{v\}$.
 In words, either $X$ receives more than $p$ colours or some vertex in $X$ receives a unique colour.
@@
 
-\citet{pilipczuk.siebertz:polynomial-soda} and \citet{debski.felsner.ea:improved} use \cref{ps} and \cref{PlanarProduct}(b), respectively, to show that $\chi_p(G)$ is polynomial in $p$ when $G$ is planar or of bounded Euler genus.
+\citet{PS21} and \citet{DFMS21} use \cref{ps} and \cref{PlanarProduct}(b), respectively, to show that $\chi_p(G)$ is polynomial in $p$ when $G$ is planar or of bounded Euler genus.
 We use the following lemma due to \citet[Lemma~15]{pilipczuk.siebertz:polynomial-arxiv}.
+
+\note{DW}{Is \cref{p-centered-treewidth} in the journal version \citep{PS21}?}
 
@@
 
-The following lemma is implicitly in the work of \citet[Proof of Theorem~2.1]{debski.felsner.ea:improved}. We include the proof for completeness.
-
-\begin{lem}[\citep{debski.felsner.ea:improved}]
+\note{DW}{In the journal version \citet{DFMS21} (under the insistence of the referee :-) ) proved that $\chi_p(H_1\boxtimes H_2) \leq \chi_p(H_1) \cdot \chi(H_2^p)$ where $H_2^p$ is the $p$-th power of $H_2$.
+Apply this with $H_2:= P \boxtimes K_\ell$.
+Then $H_2^p= P^p\boxtimes K_\ell$ and $\chi(H_2^p)=(p+1)\ell$.
+So
+	$\chi_p(H\boxtimes P \boxtimes K_\ell) \leq \chi_p(H_1) \cdot (p+1)\ell$ }
+
+\note{DW}{I suggest we replace the next paragraph by 	``\citet{DFMS21} proved that 
+$\chi_p(H \boxtimes J) \leq \chi_p(H) \cdot \chi(J^p)$ for all graphs $H$ and $J$, which implies the following lemma'' and delete the proof. Okay?}
+
+The following lemma is implicitly in the work of \citet[Proof of Theorem~2.1]{DFMS21}. We include the proof for completeness.
+
+\begin{lem}[\citep{DFMS21}]
 \label{p-centered}
@@
 Let $G$ be the graph whose vertices are the nations of $G_0$, where two vertices are adjacent in $G$ if the corresponding faces in $G_0$ share a vertex.
-Then $G$ is called a \emph{$(g,d)$-map graph}.
-A $(0,d)$-map graph is called a (plane) \emph{$d$-map graph}; see \citep{FLS-SODA12,CGP02} for example.
+Then $G$ is called a \defin{$(g,d)$-map graph}.
+A $(0,d)$-map graph is called a (plane) \defin{$d$-map graph}; see \citep{FLS-SODA12,CGP02} for example.
 The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.
@@
 
-A \emph{string graph} is the intersection graph of a set of curves in the plane with no three curves meeting at a single point; see  \cite{PachToth-DCG02,FP10,FP14} for example.
-For an integer $\delta\geq 2$, if each curve is in at most $\delta$ intersections with other curves, then the corresponding string graph is called a \emph{$\delta$-string graph}.
-A \emph{$(g,\delta)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
+A \defin{string graph} is the intersection graph of a set of curves in the plane with no three curves meeting at a single point; see  \cite{PachToth-DCG02,FP10,FP14} for example.
+For an integer $\delta\geq 2$, if each curve is in at most $\delta$ intersections with other curves, then the corresponding string graph is called a \defin{$\delta$-string graph}.
+A \defin{$(g,\delta)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
 
@@
 
-Our results also give bounds on the non-repetitive chromatic number and the $p$-centered chromatic number of $(g,\delta)$-string graphs, but the bounds are weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ and $\chi_p(G)\le p(64\delta)^2$ by results of \citet{DJKW16} and \citet{debski.felsner.ea:improved}, respectively.
+Our results also give bounds on the non-repetitive chromatic number and the $p$-centered chromatic number of $(g,\delta)$-string graphs, but the bounds are weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ and $\chi_p(G)\le p(64\delta)^2$ by results of \citet{DJKW16} and \citet{DFMS21}, respectively.
 
@@
 
-Recall that the \emph{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
+Recall that the \defin{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
 If $G$ is planar with maximum degree $\Delta$, then $G^k$ is $2k(k+1)\Delta^{k}$-planar by \cref{DrawG}.
@@
 \cref{PowerGenus} improves upon this bound when $k,g\ll\Delta$.
-Similarly, a result of \citet{debski.felsner.ea:improved} implies that $\chi_p(G^k)\le 1024p\Delta^{2k}$ and \cref{PowerGenus} improves upon this bound when $p,k,g\ll\Delta$.
+Similarly, a result of \citet{DFMS21} implies that $\chi_p(G^k)\le 1024p\Delta^{2k}$ and \cref{PowerGenus} improves upon this bound when $p,k,g\ll\Delta$.
 
@@
 
-As mentioned in \cref{Introduction}, \citet{DEJGMM} prove results about adjacency labelling schemes for planar graphs.
+As mentioned in \cref{Introduction}, \citet{DEJGMM21} prove results about adjacency labelling schemes for planar graphs.
 Stated in terms of universal graphs, their main theorem is interpreted as follows:
 
-\begin{thm}[\citep{DEJGMM}]
+\begin{thm}[\citep{DEJGMM21}]
 	\label{Universal}
@@
 Perhaps $C=3$.
-Note that $C\geq 3$ even for planar graphs \citep{dujmovic.joret.ea:planar}.
+Note that $C\geq 3$ even for planar graphs \citep{DJMMUW20}.
+
+\subsection*{Note Added in Proof} Subsequent to the initial release of this paper, the treewidth 8 bound in \cref{PlanarProduct} was improved to 6 by \citet{UWY}, which implies that the treewidth $2g+8$ bound in \cref{GenusProduct}(a) can be improved to $2g+6$.
+Simmilarly, the treewidth 4 bound in \cref{GenusProduct}(b) was improved to 3 by \citet{DHHW}.
+\note{DW}{Say more about the consequence for the bounds in the rest of the paper}
 
@@
 %\bibliographystyle{plainnat}
-%\bibliography{k-planar}
-
-% NOTE THAT MANUAL CHANGES HAVE BEEN MADE BELOW.
-
-\begin{thebibliography}{37}
-	\providecommand{\natexlab}[1]{#1}
-	\providecommand{\url}[1]{\texttt{#1}}
-	\expandafter\ifx\csname urlstyle\endcsname\relax
-	\providecommand{\doi}[1]{doi: #1}\else
-	\providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
-
-	\bibitem[{\'{A}}brego et~al.(2011){\'{A}}brego, Monroy,
-	Fern{\'{a}}ndez{-}Merchant, Flores{-}Pe{\~{n}}aloza, Hurtado,
-	Sacrist{\'{a}}n, and Saumell]{abrego.munroy.ea:on}
-	Bernardo~M. {\'{A}}brego, Ruy~Fabila Monroy, Silvia Fern{\'{a}}ndez{-}Merchant,
-	David Flores{-}Pe{\~{n}}aloza, Ferran Hurtado, Vera Sacrist{\'{a}}n, and
-	Maria Saumell.
-	\newblock On crossing numbers of geometric proximity graphs.
-	\newblock \emph{Comput. Geom.}, 44\penalty0 (4):\penalty0 216--233, 2011.
-	\newblock \doi{10.1016/j.comgeo.2010.11.003}.
-
-	\bibitem[Ajtai et~al.(1982)Ajtai, Chv{\'a}tal, Newborn, and
-	Szemer{\'e}di]{ajtai.chvatal.ea:crossing-free}
-	Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, and Endre
-	Szemer{\'e}di.
-	\newblock Crossing-free subgraphs.
-	\newblock \emph{Ann. Discrete Math.}, 12:\penalty0 9--12, 1982.
-	\newblock \doi{10.1016/S0304-0208(08)73484-4}.
-
-	\bibitem[Alam et~al.(2018)Alam, Bekos, Gronemann, Kaufmann, and
-	Pupyrev]{ABGKP18}
-	Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael Kaufmann, and
-	Sergey Pupyrev.
-	\newblock Queue layouts of planar 3-trees.
-	\newblock In Therese~C.
-Biedl and Andreas Kerren, editors, \emph{Proc.
-26th International Symposium on Graph Drawing and Network Visualization} (GD 2018), volume 11282 of \emph{Lecture Notes in Computer Science}, pages
-	213--226. Springer, 2018.
-	\newblock \doi{10.1007/978-3-030-04414-5\_15}.
-
-	\bibitem[Alon et~al.(2002)Alon, Grytczuk, Ha{\l}uszczak, and
-	Riordan]{AGHR-RSA02}
-	Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, and Oliver Riordan.
-	\newblock Nonrepetitive colorings of graphs.
-	\newblock \emph{Random Structures Algorithms}, 21\penalty0 (3-4):\penalty0
-	336--346, 2002.
-	\newblock \doi{10.1002/rsa.10057}.
-
-	\bibitem[Bodlaender(1998)]{Bodlaender-TCS98}
-	Hans~L. Bodlaender.
-	\newblock A partial $k$-arboretum of graphs with bounded treewidth.
-	\newblock \emph{Theoret. Comput. Sci.}, 209\penalty0 (1-2):\penalty0 1--45,
-	1998.
-	\newblock \doi{10.1016/S0304-3975(97)00228-4}.
-
-	\bibitem[Bonamy et~al.(2020)Bonamy, Gavoille, and
-	Pilipczuk]{bonamy.gavoille.ea:shorter}
-	Marthe Bonamy, Cyril Gavoille, and Micha{\l} Pilipczuk.
-	\newblock Shorter labeling schemes for planar graphs.
-	\newblock In Shuchi Chawla, editor, \emph{Proc. ACM-SIAM Symposium on Discrete Algorithms} (SODA 2020), pages 446--462. {SIAM}, 2020.
-	\newblock \doi{10.1137/1.9781611975994.27}.
-
-	\bibitem[Bose et~al.(2001)Bose, Morin, Stojmenović, and
-	Urrutia]{bose.morin.ea:routing}
-	Prosenjit Bose, Pat Morin, Ivan Stojmenović, and Jorge Urrutia.
-	\newblock Routing with guaranteed delivery in ad hoc wireless networks.
-	\newblock \emph{Wireless Networks}, 7\penalty0 (6):\penalty0 609--616, 2001.
-	\newblock \doi{10.1023/A:1012319418150}.
-
-	\bibitem[Bose et~al.(2012)Bose, Dujmovic, Hurtado, Iacono, Langerman, Meijer,
-	Adinolfi, Saumell, and Wood]{ProximityGraphs}
-	Prosenjit Bose, Vida Dujmovic, Ferran Hurtado, John Iacono, Stefan Langerman,
-	Henk Meijer, Vera~Sacrist{\'{a}}n Adinolfi, Maria Saumell, and David~R. Wood.
-	\newblock Proximity graphs: E, {\(\delta\)}, {\(\Delta\)}, {\(\chi\)} and
-	{\(\omega\)}.
-	\newblock \emph{Int. J. Comput. Geometry Appl.}, 22\penalty0 (5):\penalty0
-	439--470, 2012.
-	\newblock \doi{10.1142/S0218195912500112}.
-
-	\bibitem[Chen et~al.(2002)Chen, Grigni, and Papadimitriou]{CGP02}
-	Zhi-Zhong Chen, Michelangelo Grigni, and Christos~H. Papadimitriou.
-	\newblock Map graphs.
-	\newblock \emph{J. ACM}, 49\penalty0 (2):\penalty0 127--138, 2002.
-	\newblock \doi{10.1145/506147.506148}.
-
-	\bibitem[Debski et~al.(2020)Debski, Felsner, Micek, and
-	Schr{\"{o}}der]{debski.felsner.ea:improved}
-	Michal Debski, Stefan Felsner, Piotr Micek, and Felix Schr{\"{o}}der.
-	\newblock Improved bounds for centered colorings.
-	\newblock In Shuchi Chawla, editor, \emph{Proc. {ACM-SIAM} Symposium on
-		Discrete Algorithms} ({SODA} 2020), pages 2212--2226. {SIAM}, 2020.
-	\newblock \doi{10.1137/1.9781611975994.136}.
-
-	\bibitem[{Dvo{\v{r}}{\'a}k et~al.(2020)Dvo{\v{r}}{\'a}k, Huynh, Joret, and Liu}]{DHGLW}
-	Zden{\v{e}}k Dvo{\v{r}}{\'a}k, Tony Huynh, Gwena\"el Joret, and Chun-Hung Liu.
-	\newblock Notes on graph product structure theory.
-	\newblock 2020.
-	\newblock \arXiv{2001.08860}.
-
-	\bibitem[Dujmovi\'{c} et~al.(2016)Dujmovi\'{c}, Joret, Kozik, and Wood]{DJKW16}
-	Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, and David~R. Wood.
-	\newblock Nonrepetitive colouring via entropy compression.
-	\newblock \emph{Combinatorica}, 36\penalty0 (6):\penalty0 661--686, 2016.
-	\newblock \doi{10.1007/s00493-015-3070-6}.
-
-	\bibitem[Dujmovi\'c et~al.(2017{\natexlab{a}})Dujmovi\'c, Eppstein, and
-	Wood]{dujmovic.eppstein.ea:structure}
-	Vida Dujmovi\'c, David Eppstein, and David~R. Wood.
-	\newblock Structure of graphs with locally restricted crossings.
-	\newblock \emph{{SIAM} J. Discrete Math.}, 31\penalty0 (2):\penalty0 805--824,
-	2017{\natexlab{a}}.
-	\newblock \doi{10.1137/16M1062879}.
-
-	\bibitem[Dujmovi\'c et~al.(2017{\natexlab{b}})Dujmovi\'c, Morin, and
-	Wood]{dujmovic.morin.ea:layered}
-	Vida Dujmovi\'c, Pat Morin, and David~R. Wood.
-	\newblock Layered separators in minor-closed graph classes with applications.
-	\newblock \emph{J. Comb. Theory, Ser. {B}}, 127:\penalty0 111--147,
-	2017{\natexlab{b}}.
-	\newblock \doi{10.1016/j.jctb.2017.05.006}.
-
-	\bibitem[Dujmovi{\'c} et~al.(2019)Dujmovi{\'c}, Joret, Micek, Morin, Ueckerdt,
-	and Wood]{dujmovic.joret.ea:planar}
-	Vida Dujmovi{\'c}, Gwena\"{e}l Joret, Piotr Micek, Pat Morin, Torsten Ueckerdt,
-	and David~R. Wood.
-	\newblock Planar graphs have bounded queue-number.
-	\newblock \emph{J. ACM} 67(4):22, 2020.
-	\newblock \doi{10.1145/3385731}.
-
-	\bibitem[Dujmovi\'c et~al.(2020{\natexlab{a}})Dujmovi\'c, Esperet, Joret,
-	Gavoille, Micek, and Morin]{DEJGMM}
-	Vida Dujmovi\'c, Louis Esperet, Gwena\"el Joret, Cyril Gavoille, Piotr Micek,
-	and Pat Morin.
-	\newblock Adjacency labelling for planar graphs (and beyond).
-	\newblock \emph{J. ACM} 68(6):42, 2021. 
-	\newblock \doi{10.1145/3477542}.
-
-	\bibitem[Dujmovi\'c et~al.(2020{\natexlab{b}})Dujmovi\'c, Esperet, Joret,
-	Walczak, and Wood]{dujmovic.esperet.ea:planar}
-	Vida Dujmovi\'c, Louis Esperet, Gwena{\"{e}}l Joret, Bartosz Walczak, and
-	David~R. Wood.
-	\newblock Planar graphs have bounded nonrepetitive chromatic number.
-	\newblock \emph{Advances in Combinatorics}, 5, 2020{\natexlab{b}}.
-	\newblock \doi{10.19086/aic.12100}.
-
-	\bibitem[{Dujmovi{\'c} et~al.(2020)Dujmovi{\'c}, Esperet, Morin, Walczak, and Wood}]{DEMWW}
-	Vida Dujmovi{\'c}, Louis Esperet, Pat Morin, Bartosz Walczak, and
- David~R. Wood.
-	\newblock Clustered 3-colouring graphs of bounded degree.
-\newblock \emph{Combin. Probab. Comput.} 31(1):123--135, 2022.
-\newblock \doi{10.1017/S0963548321000213}.
-
-	\bibitem[Fomin et~al.(2012)Fomin, Lokshtanov, and Saurabh]{FLS-SODA12}
-	Fedor~V. Fomin, Daniel Lokshtanov, and Saket Saurabh.
-	\newblock Bidimensionality and geometric graphs.
-	\newblock In \emph{Proc. 23rd {A}nnual {ACM}-{SIAM} {S}ymposium on {D}iscrete
-		{A}lgorithms}, pages 1563--1575, 2012.
-	\newblock \doi{10.1137/1.9781611973099.124}.
-
-	\bibitem[Fox and Pach(2008)]{FP08}
-	Jacob Fox and J{\'a}nos Pach.
-	\newblock Separator theorems and {T}ur\'an-type results for planar intersection
-	graphs.
-	\newblock \emph{Adv. Math.}, 219\penalty0 (3):\penalty0 1070--1080, 2008.
-	\newblock \doi{10.1016/j.aim.2008.06.002}.
-
-	\bibitem[Fox and Pach(2010)]{FP10}
-	Jacob Fox and J{\'a}nos Pach.
-	\newblock A separator theorem for string graphs and its applications.
-	\newblock \emph{Combin. Probab. Comput.}, 19\penalty0 (3):\penalty0 371--390,
-	2010.
-	\newblock \doi{10.1017/S0963548309990459}.
-
-	\bibitem[Fox and Pach(2014)]{FP14}
-	Jacob Fox and J{\'a}nos Pach.
-	\newblock Applications of a new separator theorem for string graphs.
-	\newblock \emph{Combin. Probab. Comput.}, 23\penalty0 (1):\penalty0 66--74,
-	2014.
-	\newblock \doi{10.1017/S0963548313000412}.
-
-	\bibitem[Grigoriev and Bodlaender(2007)]{grigoriev.bodlaender:algorithms}
-	Alexander Grigoriev and Hans~L. Bodlaender.
-	\newblock Algorithms for graphs embeddable with few crossings per edge.
-	\newblock \emph{Algorithmica}, 49\penalty0 (1):\penalty0 1--11, 2007.
-	\newblock \doi{10.1007/s00453-007-0010-x}.
-
-	\bibitem[Harvey and Wood(2017)]{HW17}
-	Daniel~J. Harvey and David~R. Wood.
-	\newblock Parameters tied to treewidth.
-	\newblock \emph{J. Graph Theory}, 84\penalty0 (4):\penalty0 364--385, 2017.
-	\newblock \doi{10.1002/jgt.22030}.
-
-	\bibitem[Heath et~al.(1992)Heath, Leighton, and Rosenberg]{HLR92}
-	Lenwood~S. Heath, F.~Thomson Leighton, and Arnold~L. Rosenberg.
-	\newblock Comparing queues and stacks as mechanisms for laying out graphs.
-	\newblock \emph{SIAM J. Discrete Math.}, 5\penalty0 (3):\penalty0 398--412,
-	1992.
-	\newblock \doi{10.1137/0405031}.
-
-	\bibitem[van~den Heuvel and Wood(2017)]{vdHW17}
-	Jan van~den Heuvel and David~R. Wood.
-	\newblock Improper colourings inspired by {H}adwiger's conjecture.
-	\newblock \arXiv{1704.06536}, 2017.
-
-	\bibitem[Kannan et~al.(1988)Kannan, Naor, and
-	Rudich]{kannan.naor.ea:implicit-stoc}
-	Sampath Kannan, Moni Naor, and Steven Rudich.
-	\newblock Implicit representation of graphs.
-	\newblock In Janos Simon, editor, \emph{Proc. 20th Annual {ACM} Symposium on Theory of Computing} (STOC 1988), pages 334--343. {ACM}, 1988.
-	\newblock \doi{10.1145/62212.62244}.
-
-	\bibitem[Kannan et~al.(1992)Kannan, Naor, and Rudich]{kannan.naor.ea:implicit}
-	Sampath Kannan, Moni Naor, and Steven Rudich.
-	\newblock Implicit representation of graphs.
-	\newblock \emph{{SIAM} J. Discrete Math.}, 5\penalty0 (4):\penalty0 596--603,
-	1992.
-	\newblock \doi{10.1137/0405049}.
-
-	\bibitem[Kobourov et~al.(2017)Kobourov, Liotta, and
-	Montecchiani]{kobourov.liotta.ea:annotated}
-	Stephen~G. Kobourov, Giuseppe Liotta, and Fabrizio Montecchiani.
-	\newblock An annotated bibliography on 1-planarity.
-	\newblock \emph{Comput. Sci. Rev.}, 25:\penalty0 49--67, 2017.
-	\newblock \doi{10.1016/j.cosrev.2017.06.002}.
-
-	\bibitem[Leighton(1983)]{leighton:complexity}
-	F.~Thomson Leighton.
-	\newblock \emph{Complexity issues in {VLSI}}.
-	\newblock MIT Press, Cambridge, MA, 1983.
-
-	\bibitem[Mohar and Thomassen(2001)]{mohar.thomassen:graphs}
-	Bojan Mohar and Carsten Thomassen.
-	\newblock \emph{Graphs on surfaces}.
-	\newblock Johns Hopkins University Press, 2001.
-
-	\bibitem[Ne{\v{s}}et{\v{r}}il and Ossona~de Mendez(2012)]{Sparsity}
-	Jaroslav Ne{\v{s}}et{\v{r}}il and Patrice Ossona~de Mendez.
-	\newblock \emph{Sparsity}, volume~28 of \emph{Algorithms and Combinatorics}.
-	\newblock Springer, 2012.
-	\newblock \doi{10.1007/978-3-642-27875-4}.
-
-	\bibitem[Pach and T{\'{o}}th(1997)]{PachToth97}
-	J{\'{a}}nos Pach and G{\'{e}}za T{\'{o}}th.
-	\newblock Graphs drawn with few crossings per edge.
-	\newblock \emph{Combinatorica}, 17\penalty0 (3):\penalty0 427--439, 1997.
-	\newblock \doi{10.1007/BF01215922}.
-
-	\bibitem[Pach and T{\'o}th(2002)]{PachToth-DCG02}
-	J{\'a}nos Pach and G{\'e}za T{\'o}th.
-	\newblock Recognizing string graphs is decidable.
-	\newblock \emph{Discrete Comput. Geom.}, 28\penalty0 (4):\penalty0 593--606,
-	2002.
-	\newblock \doi{10.1007/s00454-002-2891-4}.
-
-	\bibitem[Pilipczuk and Siebertz(2019)]{pilipczuk.siebertz:polynomial-soda}
-	Micha{\l} Pilipczuk and Sebastian Siebertz.
-	\newblock Polynomial bounds for centered colorings on proper minor-closed graph classes.
-	\newblock In Timothy~M. Chan, editor, \emph{Proc. 13th Annual {ACM-SIAM}
-		Symposium on Discrete Algorithms} (SODA 2019), pages 1501--1520. {SIAM},	2019.
-	\newblock \doi{10.1137/1.9781611975482.91}.
-
-	\bibitem[Pilipczuk and Siebertz(2018)]{pilipczuk.siebertz:polynomial-arxiv}
-	Michał Pilipczuk and Sebastian Siebertz.
-	\newblock Polynomial bounds for centered colorings on proper minor-closed graph
-	classes, 2018.
-	\newblock \arXiv{1807.03683v2}.
-
-	\bibitem[Reed(2003)]{Reed03}
-	Bruce~A. Reed.
-	\newblock Algorithmic aspects of tree width.
-	\newblock In \emph{Recent Advances in Algorithms and Combinatorics}, pages 85--107. Springer, 2003.
-	\newblock \doi{10.1007/0-387-22444-0_4}.
-
-%	\bibitem[Wiechert(2017)]{wiechert:on}
-%	Veit Wiechert.
-%	\newblock On the queue-number of graphs with bounded tree-width.
-%	\newblock \emph{Electr. J. Comb.}, 24\penalty0 (1):\penalty0 P1.65, 2017.
-%	\newblock \doi{10.37236/6429}.
-
-	\bibitem[Zhu(2009)]{zhu:colouring}
-	Xuding Zhu.
-	\newblock Colouring graphs with bounded generalized colouring number.
-	\newblock \emph{Discrete Mathematics}, 309\penalty0 (18):\penalty0 5562--5568,
-	2009.
-	\newblock \doi{10.1016/j.disc.2008.03.024}.
-
-\end{thebibliography}
-
+\bibliographystyle{DavidNatbibStyle}
+\bibliography{k-planar}
 \end{document}

### 2022-03-04 k-planar.tex

@@
 \newcommand{\note}[2]{\noindent{\color{red}[#1:~#2]}}
-% \newcommand{\notex}[2]{{\color{blue}[#1:~#2]}}
 \newcommand{\notex}[2]{}
@@
 
-Here $\boxtimes$ is the strong product\footnote{The \defin{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
-	%:\begin{compactitem}
-	%	\item  
+Here $\boxtimes$ is the strong product\footnote{The \emph{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
 	$v=w$ and $xy\in E(B)$, or
-	%	\item  
 	$x=y$ and $vw\in E(A)$, or
-	%	\item  
-	$vw\in E(A)$ and $xy\in E(B)$.
-	%\end{compactitem}
-}, and treewidth\footnote{For a tree $T$, a \defin{$T$-decomposition} of a graph $G$ is a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that
+	$vw\in E(A)$ and $xy\in E(B)$.}, 
+and treewidth\footnote{For a tree $T$, a \defin{$T$-decomposition} of a graph $G$ is a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that
 (i) for every $vw\in E(G)$, there exists some node $x\in V(T)$ with $v,w\in B_x$; and
@@
 
-% \begin{wrapfigure}{r}{0.5\textwidth}
 \begin{figure}[!h]
@@
 \item The result of \citet{DEJGMM21} implies that, for every integer
-$n>0$, there is a `universal' graph $U_n$ with $n^{1+o(1)}$ 
-vertices such that every $n$-vertex planar graph is an induced
+$n>0$, there is a `universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
 subgraph of $U_n$.
@@
 A non-empty set $\mathcal{P}$ of non-trivial paths in a graph $G$ is a \defin{$(k,d)$-shortcut system} (for $G$) if:
+
 \begin{compactitem}
@@
 
-\referee{1}{7. page 8, item (i): ``$S$ has small layered width with respect to the layering
-L'' $\longrightarrow$ ``$S$ has small layered width with respect to the layering $L$ of
-$G$''. Perhaps mention something about how it is easy to find the new layering.}
+\referee{1}{7.
+page 8, item (i): ``$S$ has small layered width with respect to the layering $L$'' $\longrightarrow$ ``$S$ has small layered width with respect to the layering $L$ of $G$''.
+Perhaps mention something about how it is easy to find the new layering.}
 
@@
 	Finally, the subgraph $N_i$ of $G$ consisting of the edges and vertices of $G$ contained in $F_i$ or its interior does not contain one of the three vertices of $\tau$.
-Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(C_i)$ to obtain the desired $H$-partition and tree-decomposition of $N_i$.
-
-	The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
+Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(C_i)$ to obtain the desired $H$-partition and tree decomposition of $N_i$.
+
+	The proof finishes in the same way as the proof in \cite{dujmovic.joret.ea:planar}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
 	Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
-The desired tree-decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{DJMMUW20}, except that now each node $x$ has a child for each face $F_i$ of $M^+_x$ that contains a vertex of $G$ in its interior.
+The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{dujmovic.joret.ea:planar}, except that now each node $x$ has a child for each face $F_i$ of $M^+_x$ that contains a vertex of $G$ in its interior.
 
@@
 	For a vertex $x$ of $G_0$ with degree at most $\Delta$, and for $i\in\{1,\dots,k-1\}$, say a vertex $v$ is \defin{$i$-close} to $x$ if there is a $vx$-path $P$ in $G_0$ of length at most $i$ such that every internal vertex in $P$  has degree at most $\Delta$ in $G_0$.
-	For each edge $vw$ of $G$, say that $vw$ \defin{passes through} each internal vertex on $P_{vw}$.
-	Say $vw$ passes through $x$. Then $v$ is $i$-close to $x$ and $w$ is $j$-close to $x$ for some $i,j\in\{1,\dots,k-1\}$ with $i+j\leq k$. At most $\Delta^{i}$ vertices are $i$-close to $x$.
+	For each edge $vw$ of $G$, say that $vw$ \defin{passes through} each internal vertex on $P_{vw}$.	Say $vw$ passes through $x$.
+Then $v$ is $i$-close to $x$ and $w$ is $j$-close to $x$ for some $i,j\in\{1,\dots,k-1\}$ with $i+j\leq k$.
+At most $\Delta^{i}$ vertices are $i$-close to $x$.
 	Thus, the number of edges of $G$ that pass through $x$ is at most
@@
 
+\note{DW}{We do not want to uncross.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 \cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^4)}$.
+
 Note that \citet{DJMMUW20} previously proved the bound of
@@
 \cref{non-repetitive,kPlanarProduct,1-planar,gkPlanarProduct} imply the following result:
-
-\pagebreak
 
@@
 \end{cor}
-%\end{samepage}
 
@@
 \note{DW}{Is \cref{p-centered-treewidth} in the journal version \citep{PS21}?}
+
+\citet{pilipczuk.siebertz:polynomial-soda} and \citet{debski.felsner.ea:improved} use \cref{ps} and \cref{PlanarProduct}(b), respectively, to show that $\chi_p(G)$ is polynomial in $p$ when $G$ is planar or of bounded Euler genus.
+We use the following lemma due to \citet[Lemma~15]{pilipczuk.siebertz:polynomial-arxiv}.
 
@@
 
-%\bibliographystyle{plainnat}
 \bibliographystyle{DavidNatbibStyle}
 \bibliography{k-planar}
+
 \end{document}

### 2022-03-04 slides/slides.tex

@@
 \newcommand{\Fary}{F\'ary}
+\newcommand{\N}{\mathbb{N}}
 
@@
   \begin{itemize}
-    \item Shortcuts in $\mathcal{P}$ may jump $k$ rows
+    \item Shortcuts in $\mathcal{P}$ may jump up to $k$ rows
+    \item $G^\mathcal{P}$ will have edges between non-consecutive rows
     \item<2-> Easy solution: Compress $k$ rows into one
@@
     \begin{center}
-    \only<1>{\includegraphics{figs/layer-compression-2}}%
-    \only<2->{\includegraphics{figs/layer-compression-3}}%
+    \only<1-2>{\includegraphics{figs/layer-compression-2}}%
+    \only<3->{\includegraphics{figs/layer-compression-3}}%
     \end{center}
@@
     \item<3-> Each vertex of $H\boxtimes P$ is used in at most $dk\ell$ shortcuts
-    \item<4-> $(k,dk\ell)$-shortcut system $\mathcal{P'}$ over $G'=H\boxtimes P$
-    \item<4-> shortcuts span $\le\!\!1$ rows
+    \item<4-> $(k,dk\ell)$-shortcut system $\mathcal{P'}$ over $H\boxtimes P$
+    \item<4-> shortcuts still span $\le\!\!1$ rows
   \end{itemize}
@@
       \item $V(T) = V(H)$
-      \item $x$ is lowest common ancestor of $\{x'\in V(T):x\in B_{x'}\}$
+      \item $x$ is root of $T_x:=T[\{x'\in V(T):x\in B_{x'}\}]$
     \end{itemize}
@@
     \end{center}
-    \item If $xy\in E(H)$ then $x\prec_T y$ or $y\prec_T x$
+    \item<2-> Tree decomp implies: If $xy\in E(H)$ then $x\prec_T y$ or $y\prec_T x$
   \end{itemize}
@@
   \frametitle{Alternative View}
+  \framesubtitle{Constructing the bags from the tree}
 
   \begin{itemize}
-    \item<1-> Start with $T$ such that $x\mathbin{\diamond}_T y$ for each $xy\in E(H)$
+    \item<1-> Start with $T$ such that $x\prec_T y$ or $y\prec_T x$ for each $xy\in E(H)$
     \item<2-> $B_x\gets\{x\}$ for each $x\in V(T)$
@@
   \begin{center}
-    \includegraphics[scale=.8]{figs/pz}%
+    \multiinclude[<+>][format=pdf,start=1]{figs/pz}%
   \end{center}
@@
       \item<3-> Start with $T$ from nice tree decomposition of $H$ (not $H'$)
-      \item<4-> For each $xz\in E(H')$ where $x\prec_T z$, drag $x$ down into $B_z$
-      \item<5-> Correctness follows from $(\star)$
-      \item<6-> Point: If $x\in B_{z'}$ then
+      \item<4-> For each $xz'\in E(H')$ where $x\prec_T z$, drag $x$ down into $B_z'$
+      \item<5-> Correctness follows from $(\star)$\footnote{Technicality: Assume $\mathcal{P}$ contains a length-$1$ shortcut for each edge of $H\boxtimes P$}
+      \item<6-> Point: If $x\in B_{z}$ then
       \begin{itemize}
-          \item $x\prec_T z'\prec_T z$ and
-          \item $\mathop{dist}_{H}(x,z)\le k$ (because shortcuts have length $\le\!\!k$).
+          \item $x\prec_T z\prec_T z'$ for some $z'$ and
+          \item $\mathop{dist}_{H}(x,z')\le k$ (because shortcuts have length $\le\!\!k$).
       \end{itemize}
@@
 \begin{frame}
-  \frametitle{Thickness}
-  \framesubtitle{Detail checking}
+  \frametitle{Finishing Up}
+  \framesubtitle{Verifying the layered width}
 
   \begin{itemize}
-    \item Recall that $\mathcal{P'}$ is a $(k,dk\ell)$-shortcut system over $H\boxtimes P$.
-    \item We can show that the subset of $H\boxtimes P\boxtimes K_{\ell}$ corresponding to $S_x$ contains at most $d\ell(k^2+3)$ vertices from each layer of $H\boxtimes P\boxtimes K_{\ell}$.
+    \item $\mathcal{P'}$: a $(k,dk\ell)$-shortcut system over $H\boxtimes P$.
+    \item Obtained from $\mathcal{P}$: a $(k,d)$-shortcut system over $H\boxtimes P\boxtimes K_{k\ell}$
+    \item We can show that the subset of $H\boxtimes P\boxtimes K_{\ell}$ corresponding to $S_x$ contains at most $d\ell(k^3+3k)$ vertices from each row of $H\boxtimes P\boxtimes K_{\ell}$.
     \item Implies $G^{\mathcal{P}}\subseteq H'\boxtimes P\boxtimes K_{d\ell(k^3+3k)}$
   \end{itemize}
+  \vspace{1cm}
+  \uncover<2->{
+  \textbf{Main Theorem:} \newline
+  If $G\subseteq H\boxtimes P\boxtimes K_\ell$
+    \begin{itemize}
+      \item  $\tw(H)\le t$
+      \item $\mathcal{P}$ is a $(k,d)$-shortcut system for $G$
+    \end{itemize}
+    then $G^\mathcal{P}\subseteq H'\boxtimes P\boxtimes K_{\ell'}$
+    \begin{itemize}
+      \item $\tw(H') \le t'=\binom{k+t}{t}-1$ and
+      \item $\ell'=d\ell(k^3+3k)$
+      % \item $K'$ is a clique of order $d\ell(k^3+3k)$
+    \end{itemize}
+  }
 \end{frame}
@@
 \begin{frame}
+  \frametitle{Discussion}
+
+  \begin{itemize}
+    \item<1-> Optimizing for $k$-planar $G$ shows $G\subseteq H\boxtimes P\boxtimes K_\ell$ for $\tw(H)\le \binom{k+4}{3}-1$ and $\ell\le 18k^{\color{red}2}+48k+30$.
+
+    \item<2-> Optimizing for $1$-planar $G$ shows $G\subseteq H\boxtimes P\boxtimes K_{30}$ with $\tw(H)\le 3$
+
+    \item<3-> \textbf{Open Problem:} Does there exist $C\in\N$ and $f:\N\to\N$ such that for $k$-planar $G$,  $G\subseteq H\boxtimes P\boxtimes K_{f(k)}$ where $\tw(H)\le C$?
+\uncover<4->{(Maybe $C=3$)}\vspace{1.5cm}
+
+    \item<5->[] \centerline{\Large\bfseries Thank You!}
+  \end{itemize}
 

### 2022-03-04 response.tex
revised according to referee comments
merged response

@@
 \usepackage{amsmath,color,amsthm,enumerate,amsfonts,mathtools}
+\usepackage{graphicx}
+\usepackage{enumerate}
+\usepackage{amsfonts}
+\usepackage{paralist}
+\usepackage{stmaryrd}
+
+\newcommand{\done}{\textcolor{blue}{done}}
+\newcommand{\tba}{\textcolor{red}{to be addresssed}}
 
@@
 
+\tba
+
 2. Proof of Theorem 9: There is too much notation throughout. As
@@
 
+\tba
+
 3. Theorem 10: Unfortunately I have to suggest removing it. I don’t see
@@
 reason for including it?
+
+\tba
 
@@
 
-Other comments:
+\tba
 
@@
 
-\textcolor{blue}{done}
+\done
 
@@
 $T$ '' $\longrightarrow$ Delete the first $T$ and define a $T$-decomposition.
+
+\done
+
+Other comments:
+
+1.
+abstract: “This leads to analogous results for map graphs, string graphs, graph powers, and nearest neighbour graphs.” $\longrightarrow$ Add “certain types of” or be more specific.
+Same comment for immediately after Theorem 4.
+
+2. page 1: “A tree-decomposition T consists of a tree T and a collection
+T ” $\longrightarrow$ Delete the first T and define a T-decomposition.
+>>>>>>> f2dea7a16f3c0af4014722c98524e5a9481eb01b
 
@@
 
+\tba
+
 4. page 6, line 14: ``Thus, it suffices to construct width-$t$ tree decomposition that satisfies (T2).'' $\longrightarrow$ should be (T1)
+
+\done
 
@@
 
+\tba
+
 6. page 6, last paragraph of Lemma 2: This isn’t quite right since you
 have to be careful with $x_0$.
+
+\tba
 
@@
 
+\tba
+
 8. general comment: Claims 1, 2, and 4 are especially belaboured by
@@
 
+\tba
+
 9. page 8, line 10: ``We say that a vertex $w \in Y_x$ contributes a vertex
 $v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.
+
+\tba
 
@@
 
+\tba
+
 11. page 10, Claim 5: The name $Pww'$ is strange since in case 2. there is
 no $w'$ defined.
+
+\done
 
@@
 
+\done
+
 Page 3.  You should add the assumption that no three edges cross at
@@
 
-Page 3. $P$ is used twice in the statement of Theorem 3.  It is probably
-safer (and necessary?) to use a different variable for the second
-occurrence.
+\done
 
@@
 
+\done
+
 Page 6. Replace 'that satisfies (T2)' by 'that satisfies (T1)'.
+
+\done
 
@@
 
+\done
+
 Page 6. The term 'hierarchical decomposition' is not defined.
+
+\textcolor{blue}{We have put `hierarchical' in quotes to indicate it is not a defined word.}
 
@@
 
+\done
+
+=======
 Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
@@
 
+\tba
+
 Page 8.  Can you add a brief explanation why $\mathcal{S}$ is a
@@
 
+\tba
+
 Page 11. Replace 'belong to $F_\delta'$ by 'belongs to $F_\delta$'.
 
-\textcolor{blue}{done}
+\done
 
@@
 
+\done
+
 Page 13. Could it be that the subgraph induced by the vertices of a
@@
 the proof.
+
+\tba
 
@@
 
+\tba
+
 Page 13.  Regarding the proof of Lemma 4, I think it is better to
@@
 
+\tba
+
 Page 15. Regarding the proof of Theorem 11, if it really is the same
@@
 which Theorem 2 follows as a special case)?
+
+\tba
 
@@
 
+\tba
+
 Page 18. In point 1 in the proof of Lemma 9, I think '$\phi(w)$' should
@@
 
+\tba
+
 Page 18.  Say $H'$ is connected since $X$ is connected.
 
+\tba
+
 Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.
+
+\tba
 
@@
 
+\tba
+
 Page 21.  Replace the comma at the end of the statement of Theorem 14

### 2022-03-05 k-planar.tex
added comments about how to fix the 1-planar proof;  it is all fixed if we  define kite in the right way

@@
 It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
+\note{DW}{We need to define a kite to be the subgraph with vertex-set $\{v,w,x,y\}$ and edge-set $\{vx,xy,vx,vy,wy,wx\}$ where $vx,vy,wy,wx$ are the edges on the `outside' of the kite-faces.
+Then it really is a simple $K_4$.
+I think this is what we intended.
+We should add a figure showing a kite and kite-faces.
+}
 The edges $vw$ and $xy$ are called \defin{spars} of $K$.
@@
 \includegraphics[width=60mm]{kite}}
-\note{DW}{I agree with the referee.}
+\note{DW}{We need to explain more about how edge-maximality fixes this problem. In this case we can draw other $xv$, $vy$, $yw$ $wx$ edges close to the central crossing edges.}
+\note{DW}{The graph drawing literature defines a 1-planar graph to be \defin{optimal} if it is obtained from a plane quadrangulation by adding the two crossing edges across each face.
+Optimal 1-planar graphs have the maximum number of edges out of all $n$-vertex 1-planar graphs.	But I am pretty sure there are edge-maximal 1-planar graphs that are not optimal.
+We should check that our proof is okay on these examples.
+}
 Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
-\note{DW}{This is wrong I think} 
-Observe that any spar of $K$ is incident on exactly four \defin{kite faces} of $G$, each of which has three edges and two vertices of $G$ on its boundary.
 
@@
 'three edges and two vertices of $G$ on its boundary'.}
-\note{DW}{I agree with the referee.}
+\note{DW}{This should be fixed if we define kite more precisely. A kite-face is bounded by one edge and two half-edges right?}
 
@@
 
-\note{DW}{I agree with the referee.}
+\note{DW}{If we define kite as above, this is not a problem}
 
@@
 
-Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analagous result for plane graphs \cite{DJMMUW20}.
+Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
 \begin{lem}
@@
 
-	The proof finishes in the same way as the proof in \cite{dujmovic.joret.ea:planar}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
+	The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
 	Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
-The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{dujmovic.joret.ea:planar}, except that now each node $x$ has a child for each face $F_i$ of $M^+_x$ that contains a vertex of $G$ in its interior.
+The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{DJMMUW20}, except that now each node $x$ has a child for each face $F_i$ of $M^+_x$ that contains a vertex of $G$ in its interior.
+
+	\note{DW}{Check whether Lemma~14 is the right lemma}
 

### 2022-03-05 k-planar.tex
added more commens on 1-planar section

@@
 \note{DW}{We can motivate \cref{1-planar} since treewidth 3 is best possible, and the planarity of $H$ is essential in the proof of \cref{1PlanarQueue}.
+Futher motivation comes from the open problem at the end of the paper.
 Shall we argue this in the response?
@@
 All this depends on whether we can fix the errors below.
+I vote to fix the errors, add more motivation for \cref{1-planar}, keep the theorem and the proof.
 }
@@
 We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.
+\note{DW}{Each face of $G$ is one of two types, either bounded by three edges (none of which are crossed), or bounded by one edge that is not crossed plus portions of two crossing edges.
+Right?
+If this is true, we should say it here, which would help to clear up the confusion below.
+Add a figure.}
 We assume that no two edges incident to a common vertex cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}
+
+\note{DW}{I suggest we say that it follows from Euler's formula that $G$ is finite.
+(To see this, say $G$ has $n$ vertices, $m$ non-crossed edges, and $k$ crossings.
+Let $G'$ be the plane graph obtained from $G$ by adding a dummy vertex at each crossing.
+So $G'$ is a plane multigraph in which every face is bounded by a 3-cycle.
+It follows from Euler's formula that $m+4k = |E(G')| = 3( |V(G')|-2) = 3 ( n+k-2)$, implying	$m+k = 3 (n-2)$ and $|E(G)| = m+2k \leq 6(n-2)$.) }
+
+\note{DW}{It seems to be that $G$ can be obtained from a plane multigraph with faces of size 3 or 4 such that no edge is in two 3-faces, by adding a pair of crossing edges across each 4-face.
+I think this characterises edge-maximal 1-planar multigraphs with no 2-faces.
+Right?
+Does this viewpoint help?
+Is it known?}
+
+\note{DW}{Say $G$ is obtained from a plane multigraph by making each face a clique. Is $G\subseteq H \boxtimes P \boxtimes K_{f(\omega(G)}$? for some treewidth 3 graph $H$.}
 
@@
 \includegraphics[width=60mm]{kite}}
-\note{DW}{We need to explain more about how edge-maximality fixes this problem. In this case we can draw other $xv$, $vy$, $yw$ $wx$ edges close to the central crossing edges.}
-\note{DW}{The graph drawing literature defines a 1-planar graph to be \defin{optimal} if it is obtained from a plane quadrangulation by adding the two crossing edges across each face.
-Optimal 1-planar graphs have the maximum number of edges out of all $n$-vertex 1-planar graphs.	But I am pretty sure there are edge-maximal 1-planar graphs that are not optimal.
-We should check that our proof is okay on these examples.
-}
-Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
+\note{DW}{We need to explain more about how edge-maximality fixes this problem.
+In this case we can draw other $xv$, $vy$, $yw$ $wx$ edges close to the central crossing edges.} Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
 

### 2022-03-05 k-planar.tex
added comments

@@
 An embedded graph $G$ is \defin{$k$-plane} if each edge of $G$ takes part in at most $k$ crossings.
-A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
+\note{DW}{We need at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}  A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
 Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
@@
 Futher motivation comes from the open problem at the end of the paper.
-Shall we argue this in the response?
-Or do we remove \cref{sec-1-planar}?
+We need to argue this in the response and in the paper.
+The alternative is to remove \cref{sec-1-planar}?
 We could keep the statement of \cref{1-planar} in the paper and refer to the arXiv paper for the proof.
@@
 So $G'$ is a plane multigraph in which every face is bounded by a 3-cycle.
-It follows from Euler's formula that $m+4k = |E(G')| = 3( |V(G')|-2) = 3 ( n+k-2)$, implying	$m+k = 3 (n-2)$ and $|E(G)| = m+2k \leq 6(n-2)$.) }
+It follows from Euler's formula that $m+4k = |E(G')| = 3( |V(G')|-2) = 3 ( n+k-2)$, implying	$m+k = 3 (n-2)$ and $|E(G)| = m+2k \leq 6(n-2)$, which can also be shown by deleting one edge from each crossing point.
+) }
 
@@
 		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
-		\item For each integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
+		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
 		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that
@@
 				\item $F[P_i]$ is a path; and
-				\item $|V(P_i)\cap L_j| \le 15$ for all integers $j\ge 0$.
+				\item $|V(P_i)\cap L_j| \le 15$ for every integer $j\ge 0$.
 			\end{compactenum}
 		\end{compactenum}
+	\note{DW}{Should the vertices already grabbed by $P_i$ be mentioned in the setup?
+The new paths $Q_1,Q_2,Q_3$ need to avoid the vertices already grabbed by $P_1,P_2,P_3$.
+Basically, I am concerned about what happens when distinct vertical paths in $T$ want to grab the same vertex.}
 		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
@@
 		\item $H$ is planar;
-		\item for all integers $j\ge 0$, and all $x\in V(H)$, $|S_x\cap L_j|\le 15$;
-		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; and
+		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 15$;
+		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$
+		\note{DW}{Should this be $P_i\subseteq S_{x_i}$? (since the path grabs other vertices into its part)}; and
 		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
@@
 
-\note{DW}{We do not want to uncross.}
+\note{DW}{But we do not want to uncross.
+Our claim that ``every pair of edges cross at most once'' is not true.
+I am convinced the result is still correct.
+Just need to think  of a better way to write it.
+Each crossing is in the vicinity of a vertex that has degree $\leq\Delta$.
+}
 

### 2022-03-05 k-planar.tex
more updates

@@
 
+\note{DW}{Use $\mathcal{S}$ for the given shortcut system and $\mathcal{P}$ for the partition?}
+
+For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
+Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
+Let $J:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph.
+We consider $V(J)\subseteq V(T)$, where each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
+(Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $J$.)
+
 \referee{2}{Page 8.  Can you add a brief explanation why $\mathcal{S}$ is a
@@
 	disjoint, but why do they cover $V(G)$?}
-
-For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
-Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
-We let $J:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph and we let $V(J)\subseteq V(T)$ in the obvious way, so that each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
-(Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $J$.)
 
@@
 		\end{compactenum}
-	\note{DW}{Should the vertices already grabbed by $P_i$ be mentioned in the setup?
+	\note{DW}{Should the vertices already grabbed by $P_1,P_2,P_3$ be mentioned in the setup?
 The new paths $Q_1,Q_2,Q_3$ need to avoid the vertices already grabbed by $P_1,P_2,P_3$.
@@
 Repeatedly applying this fact implies that $\Delta/\mathcal{P}$ is planar provided that $\Delta[P]$ is connected for every $P\in\mathcal{P}$.} The first point then ensures that $G/\mathcal{P}$ is also planar.
+	\note{DW}{We need a sentence about why $H$ Has the claimed tree-decomposition.}
 \end{proof}
@@
 In the case of 1-planar graphs we can improve further.
-\citet{ABGKP18} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
+\citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
 Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$.
@@
 \citet{PS21} and \citet{DFMS21} use \cref{ps} and \cref{PlanarProduct}(b), respectively, to show that $\chi_p(G)$ is polynomial in $p$ when $G$ is planar or of bounded Euler genus.
-We use the following lemma due to \citet[Lemma~15]{pilipczuk.siebertz:polynomial-arxiv}.
-
-\note{DW}{Is \cref{p-centered-treewidth} in the journal version \citep{PS21}?}
-
-\citet{pilipczuk.siebertz:polynomial-soda} and \citet{debski.felsner.ea:improved} use \cref{ps} and \cref{PlanarProduct}(b), respectively, to show that $\chi_p(G)$ is polynomial in $p$ when $G$ is planar or of bounded Euler genus.
-We use the following lemma due to \citet[Lemma~15]{pilipczuk.siebertz:polynomial-arxiv}.
-
-\begin{lem}[\citep{pilipczuk.siebertz:polynomial-arxiv}]
+We use the following lemma due to \citet[Lemma~15]{PS21}.
+
+\begin{lem}[\citep{PS21}]
 \label{p-centered-treewidth}
@@
 So
-	$\chi_p(H\boxtimes P \boxtimes K_\ell) \leq \chi_p(H_1) \cdot (p+1)\ell$ }
-
-\note{DW}{I suggest we replace the next paragraph by 	``\citet{DFMS21} proved that 
+	$\chi_p(H\boxtimes P \boxtimes K_\ell) \leq \chi_p(H_1) \cdot (p+1)\ell$. So I suggest we replace the next paragraph by 	``\citet{DFMS21} proved that 
 $\chi_p(H \boxtimes J) \leq \chi_p(H) \cdot \chi(J^p)$ for all graphs $H$ and $J$, which implies the following lemma'' and delete the proof. Okay?}
@@
 \end{lem}
+
+\note{DW}{For consistency with \cref{qn}, I suggest we combine \cref{p-centered,p-centered-treewidth}, and just write, ``For every graph $H$ of treewidth at most $t$ and for every path $P$, if $G\subseteq H\boxtimes P \boxtimes K_\ell$ then
+	\[\chi_p(G)\le \ell (p+1)\, \chi_p(H) \leq (p+1)\ell\, \tbinom{p+t}{t}.\]}
 
@@
    \begin{compactenum}
-     \item $\alpha(v)$ is unique among $\{\phi(w): w\in L_i\cap B_x\}$, which is possible
+     \item $\alpha(v)$ is unique among $\{\alpha(w): w\in L_i\cap B_x\}$, which is possible
      since $|L_i\cap B_x|\le \ell$,
@@
  To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
-
-\referee{2}{Page 18. In point 1 in the proof of Lemma 9, I think '$\phi(w)$' should
-be '$\alpha(w)$'.}
 
@@
 compute is $d(d-1)/2$.}
+\note{DW}{consecutive neighbours don't have to be counted. I will revise}
 

### 2022-03-05 k-planar.tex
more updates

@@
 
+\note{DW}{Delete `from 1992' and  `from 2007'? replace ``going back to 1988~\citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}'' by 
+	``starting with the work of \citet{kannan.naor.ea:implicit}''}
+
 All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is a subgraph of $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this definition.}.
@@
 
-The purpose of this paper is to prove product structure theorems for several non-minor-closed classes of interest. Our results are the first of this type for non-minor-closed classes.
+The main purpose of this paper is to prove product structure theorems for several non-minor-closed classes of interest. Our results are the first of this type for non-minor-closed classes.
 
@@
 
-\note{DW}{Use $\mathcal{S}$ for the given shortcut system and $\mathcal{P}$ for the partition?}
+\note{DW}{Use $\SS$ for the given shortcut system and $\PP$ for the partition?}
 
@@
 \end{clm}
+
+\note{DW}{I think $\binom{k+t}{t}$ in \cref{general-bag-size} and $\binom{p+t}{t}$ in \cref{p-centered-treewidth} are no coincidence. The proofs are very similar. Worth exploring further?}
 
@@
 An embedded graph $G$ is \defin{$k$-plane} if each edge of $G$ takes part in at most $k$ crossings.
-\note{DW}{We need at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}  A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
+\note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}  A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
 Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
@@
 
-\note{DW}{We can motivate \cref{1-planar} since treewidth 3 is best possible, and the planarity of $H$ is essential in the proof of \cref{1PlanarQueue}.
+\note{DW}{We can motivate \cref{1-planar} since the graph drawing community particularly cares about 1-planar graphs, since treewidth 3 is best possible, and since the planarity of $H$ is essential in the proof of \cref{1PlanarQueue}.
 Futher motivation comes from the open problem at the end of the paper.
@@
 
+\note{DW}{I agree we need more details. I am having trouble verifying this proof. }
+
 \begin{proof}
@@
 \end{thm}
-
-\referee{2}{Page 15. Regarding the proof of Theorem 11, if it really is the same
-	proof as the proof of Theorem 2, why not just prove Theorem 11 (from
-	which Theorem 2 follows as a special case)?}
 
@@
 
+\note{DW}{I don't want to do a complete restructuring. In our response, let's just say why we structure it the way we do.}
+
 %%%%%%%%%%%%%%%%%%%

### 2022-03-06 response.tex
more updates

@@
 
-Other comments:
-
-1.
-abstract: “This leads to analogous results for map graphs, string graphs, graph powers, and nearest neighbour graphs.” $\longrightarrow$ Add “certain types of” or be more specific.
-Same comment for immediately after Theorem 4.
-
-2. page 1: “A tree-decomposition T consists of a tree T and a collection
-T ” $\longrightarrow$ Delete the first T and define a T-decomposition.
->>>>>>> f2dea7a16f3c0af4014722c98524e5a9481eb01b
-
 3. pages 2 – 4: The paragraph following Theorem 2 and Section 1.3
@@
 
-11. page 10, Claim 5: The name $Pww'$ is strange since in case 2. there is
+11. page 10, Claim 5: The name $P_{ww'}$ is strange since in case 2. there is
 no $w'$ defined.
 
-\done
+\textcolor{blue}{Rearranged the sentence so that it is more clear we are defining $w'=w$. }
 
@@
 
-=======
 Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
@@
 
-Page 11. Replace 'belong to $F_\delta'$ by 'belongs to $F_\delta$'.
+Page 11. Replace 'belong to $F_\delta$' by 'belongs to $F_\delta$'.
 
@@
 
-\tba
+\textcolor{blue}{We have written the entire paper starting with simpler cases before moving onto the more general.
+This approach may use a few more words, but we think the paper becomes more accessible for more readers this way.}
 
@@
 
-\tba
+\done
 
@@
 
-\tba
+\textcolor{blue}{By construction, if the path $vxw$ is in $\mathcal{P}$, then $xv$ and $xw$ are non-consecutive edges incident to $x$.
+The bound of $d(d-1)/2$ arises if one ignores that $xv$ and $xw$ are non-consecutive.}
 

### 2022-03-06 k-planar.tex
more updates

@@
 
-\usepackage{doi} % To make plainnat handle doi's properly
+%\usepackage{doi} % To make plainnat handle doi's properly
 
@@
 
+\note{DW}{I suggest we replace $8$ by $6$ here, and replace ``The starting point for this work is the following `product structure theorem' for planar graphs by \citet{DJMMUW20} (improved by \citet{UWY})''.
+Okay?}
+
 Here $\boxtimes$ is the strong product\footnote{The \emph{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
@@
 \begin{figure}[!h]
-\centering%	\begin{center}
+\centering
 \includegraphics{ProductExample}
-	%\end{center}
 \caption{Example of a strong product.
@@
 	``starting with the work of \citet{kannan.naor.ea:implicit}''}
+
+\note{DW}{Add \citep{EJM} to the story.}
 
@@
 		\item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+8$  and some path $P$.
-		\item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $4$ and for some path $P$.\label{temp}
+		\item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $4$ and for some path $P$.
 	\end{compactenum}
 \end{thm}
+
+\note{DW}{I suggest we replace $2g+8$ by $2g+6$ here and replace 4 by 3, and cite \citep{DHHW,UWY}.
+This will improve the bounds and (more importantly) simplify the presentation of various results (such as \cref{StringPartition,PowerGenus}) which currently distinguish the $g=0$ case.
+}
 
@@
 \end{clm}
-
-\note{DW}{I think $\binom{k+t}{t}$ in \cref{general-bag-size} and $\binom{p+t}{t}$ in \cref{p-centered-treewidth} are no coincidence. The proofs are very similar. Worth exploring further?}
 
@@
 
-\referee{2}{Page 20.  Where does the bound $d(d-3)/2$ come from?  The naive bound I
-compute is $d(d-1)/2$.}
-\note{DW}{consecutive neighbours don't have to be counted. I will revise}
-
 \begin{proof}
@@
 
-Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} 
-and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,p-centered-treewidth,non-repetitive} imply the following results.
+Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,p-centered-treewidth,non-repetitive} imply the following results.
 
@@
 
-\subsection*{Note Added in Proof} Subsequent to the initial release of this paper, the treewidth 8 bound in \cref{PlanarProduct} was improved to 6 by \citet{UWY}, which implies that the treewidth $2g+8$ bound in \cref{GenusProduct}(a) can be improved to $2g+6$.
+%\subsection*{Note Added in Proof} Subsequent to the initial release of this paper, the treewidth 8 bound in \cref{PlanarProduct} was improved to 6 by \citet{UWY}, which implies that the treewidth $2g+8$ bound in \cref{GenusProduct}(a) can be improved to $2g+6$.
 Simmilarly, the treewidth 4 bound in \cref{GenusProduct}(b) was improved to 3 by \citet{DHHW}.
@@
 
-\end{document}
+\end{document}s

### 2022-03-07 k-planar.tex
Pfffft

@@
 \begin{abstract}
-Dujmovi\'c~et~al.~[\emph{J.~ACM}~'20] recently proved that every planar graph is a subgraph of the strong product of a graph of bounded treewidth and a path.
+Dujmovi\'c~et~al.~[\emph{J.~ACM}~'20] recently proved that every planar graph is isomorphic to a subgraph of the strong product of a graph of bounded treewidth and a path.
 Analogous results were obtained for graphs of bounded Euler genus or apex-minor-free graphs.
@@
 One noteable example is $k$-planar graphs (those with a drawing in the plane in which each edge is involved in at most $k$ crossings).
-We prove that every $k$-planar graph is a subgraph of the strong product of a graph of treewidth $O(k^5)$ and a path.
+We prove that every $k$-planar graph is isomorphic to a subgraph of the strong product of a graph of treewidth $O(k^5)$ and a path.
 This is the first result of this type for a non-minor-closed class of graphs.
@@
 We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.
-Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20}.
-
-\begin{thm}[\citep{DJMMUW20}]
+We say that a graph $G$ is \emph{contained in} a graph $X$ if $G$ is isomorphic to a subgraph of $X$.
+Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (improved by \citet{UWY}).
+
+\begin{thm}[\citep{DJMMUW20,UWY}]
 	\label{PlanarProduct}
@@
 
-\note{DW}{In 	\cref{PlanarProduct} and elsewhere, we really should write, ``Every planar graph is isomorphic to a subgraph of''.
+% \note{DW}{In 	\cref{PlanarProduct} and elsewhere, we really should write, ``Every planar graph is isomorphic to a subgraph of''.
 One alternative is to define that $H$ is \defin{contained} in $G$ if $H$ is isomorphic to a subgraph of $G$.
@@
 
-\note{DW}{I suggest we replace $8$ by $6$ here, and replace ``The starting point for this work is the following `product structure theorem' for planar graphs by \citet{DJMMUW20} (improved by \citet{UWY})''.
+% \note{DW}{I suggest we replace $8$ by $6$ here, and replace ``The starting point for this work is the following `product structure theorem' for planar graphs by \citet{DJMMUW20} (improved by \citet{UWY})''.
 Okay?}
 
-Here $\boxtimes$ is the strong product\footnote{The \emph{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
+Here $\boxtimes$ is the strong product,\!\footnote{The \emph{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
 	$v=w$ and $xy\in E(B)$, or
 	$x=y$ and $vw\in E(A)$, or
-	$vw\in E(A)$ and $xy\in E(B)$.}, 
+	$vw\in E(A)$ and $xy\in E(B)$.}
 and treewidth\footnote{For a tree $T$, a \defin{$T$-decomposition} of a graph $G$ is a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that
@@
 \item The result of \citet{DEJGMM21} implies that, for every integer
-$n>0$, there is a `universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
+$n>0$, there is a `induced-universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
 subgraph of $U_n$.
+This has recently been extended by \citet{EJM}, who show the existence of a `subgraph-universal' graph with $(1+\epsilon)n$ vertices that contains every $n$-vertex planar graph as a subgraph as well as the existence of an `induced universal' graph with $n^{1+o(1)}$ vertices \emph{and} edges.
+The former result is the first progress on subgraph-universal graphs for planar graphs since the $O(n^{3/2})$-vertex construction of \citet{babai.chung.ea:on}.
+% aking the first progress on this problem since the work of \citet{babai.xx} in 1984.
 \end{compactitem}
@@
 	``starting with the work of \citet{kannan.naor.ea:implicit}''}
-
-\note{DW}{Add \citep{EJM} to the story.}
+\note{PM}{Was this a referee request, or just because we don't need to sell it anymore?  I'm fine with the change either way, but didn't implement it.}
+
+% \note{DW}{Add \citep{EJM} to the story.}
 
@@
 
+\note{PM}{Kind of, but I suggest we keep that paragraph and section anyway.
+Justification: They're both short and they describe the exact bounds for bounded genus graphs, which you don't get from the one-paragraph description in Section~1.0.
+A}
+
 \subsection{Shortcut Systems}
@@
 \note{DW}{Use $\SS$ for shortcut system, okay?}
+\note{PM}{Ok.}
 
@@
 
-\begin{thm}[\citep{DJMMUW20}]
+\begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
 	\label{GenusProduct}
@@
 	\begin{compactenum}[(a)]
-		\item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+8$  and some path $P$.
-		\item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $4$ and for some path $P$.
+		\item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
+		\item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
 	\end{compactenum}
@@
 }
+\note{PM}{I agree with this suggestion.  I've only added the other two references here, and haven't changed anything elsewhere yet.}
 
@@
 
-  Next, consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.
+  Next, remove the node $x_0$ from $T_0$, and consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.
 In this case, $B_{a}\subseteq B_{a'}$ where $a'$ is the parent of $a$ since any $x\in B_a\setminus B_{a'}$ would have $f(x)=a$.
@@
 
-\referee{1}{6. page 6, last paragraph of Lemma 2: This isn’t quite right since you
-have to be careful with $x_0$.}
+% \referee{1}{6. page 6, last paragraph of Lemma 2: This isn’t quite right since you have to be careful with $x_0$.}
+% \note{PM}{I addressed this by removing $x_0$ from $T_0$ at the beginning of the second paragraph.}
 
@@
 For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
-Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
-Let $J:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph.
+Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.\footnote{The sets in $\mathcal{S}$ are disjoint because $a:V(G)\to V(T)$ is a function.
+The sets in $\mathcal{S}$ cover $V(G)$ since, for each $v\in V(G)$, $v\in S_{a(v)}$.} Let $J:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph.
 We consider $V(J)\subseteq V(T)$, where each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
@@
 	disjoint, but why do they cover $V(G)$?}
+\note{PM}{I added an explanation in a footnote since this is obvious.}
 

### 2022-03-07 k-planar.tex
A big change of notation

@@
 
-\newcommand{\PP}{\mathcal{P}}
+% \newcommand{\PP}{\mathcal{P}}
 \renewcommand{\SS}{\mathcal{S}}
@@
 \begin{compactitem}
-\item \citet{DJMMUW20} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992).
-\item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002).
+\item \citet{DJMMUW20} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92}).
+\item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
 \item \citet{DFMS21} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs.
-\item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results going back to 1988~\citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}), and \citet{DEJGMM21} have since used it to find asymptotically optimal adjacency labellings of planar graphs.
+\item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results starting with the work of \citet{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}).
+\citet{DEJGMM21} have since used it to find asymptotically optimal adjacency labellings of planar graphs.
 \item The result of \citet{DEJGMM21} implies that, for every integer
@@
 
-\note{DW}{Delete `from 1992' and  `from 2007'? replace ``going back to 1988~\citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}'' by
-	``starting with the work of \citet{kannan.naor.ea:implicit}''}
-\note{PM}{Was this a referee request, or just because we don't need to sell it anymore?  I'm fine with the change either way, but didn't implement it.}
+% \note{DW}{Delete `from 1992' and  `from 2007'? replace ``going back to 1988~\citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}'' by
+	% ``starting with the work of \citet{kannan.naor.ea:implicit}''}
+% \note{PM}{Was this a referee request, or just because we don't need to sell it anymore?  I'm fine with the change either way, but didn't implement it.}
 
@@
 
-All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is a subgraph of $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this definition.}.
-These applications motivate finding product structure theorems for other graph classes.
+All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this definition.} These applications motivate finding product structure theorems for other graph classes.
 \citet{DJMMUW20} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
@@
 \note{PM}{Kind of, but I suggest we keep that paragraph and section anyway.
-Justification: They're both short and they describe the exact bounds for bounded genus graphs, which you don't get from the one-paragraph description in Section~1.0.
-A}
+Justification: They're both short and they describe the exact bounds for bounded genus graphs, which you don't get from the one-paragraph description in Section~1.0.}
 
@@
 Although $k$-planar graphs are the most high-profile target for a generalization of \cref{PlanarProduct}, we actually prove a substantially stronger result than \cref{kPlanarProduct} using the following definition.
-A non-empty set $\mathcal{P}$ of non-trivial paths in a graph $G$ is a \defin{$(k,d)$-shortcut system} (for $G$) if:
+A non-empty set $\SS$ of non-trivial paths in a graph $G$ is a \defin{$(k,d)$-shortcut system} (for $G$) if:
 
 \begin{compactitem}
-\item every path in $\mathcal{P}$ has length at most $k$,\footnote{A path of length $k$ consists of $k$ edges and $k+1$ vertices.
+\item every path in $\SS$ has length at most $k$,\footnote{A path of length $k$ consists of $k$ edges and $k+1$ vertices.
 A path is \defin{trivial} if it has length 0 and \defin{non-trivial} otherwise.} and
-\item for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
+\item for every $v\in V(G)$, the number of paths in $\SS$ that use $v$ as an internal vertex is at most $d$.
 \end{compactitem}
-Each path $P\in\mathcal{P}$ is called a \defin{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \defin{$vw$-shortcut}.
-Given a graph $G$ and a $(k,d)$-shortcut system $\mathcal{P}$ for $G$, let $G^{\mathcal{P}}$ denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\mathcal{P}$.
+Each path $P\in\SS$ is called a \defin{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \defin{$vw$-shortcut}.
+Given a graph $G$ and a $(k,d)$-shortcut system $\SS$ for $G$, let $G^{\SS}$ denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\SS$.
 
@@
 \label{AddDummy}
-Every $k$-planar graph is a subgraph of $G^\PP$ for some planar graph $G$ and some $(k+1,2)$-shortcut system $\PP$ for $G$.
+Every $k$-planar graph is a subgraph of $G^\SS$ for some planar graph $G$ and some $(k+1,2)$-shortcut system $\SS$ for $G$.
 \end{obs}
@@
 For each edge $vw\in E(G')$ there is a path $P$ in $G$ between $v$ and $w$ of length at most $k+1$ in which every internal vertex is a dummy vertex.
-Let $\PP$ be the set of such paths $P$.
-For each vertex $v$ of $G$, at most two paths in $\PP$ use $v$ as an internal vertex (since no original vertex of $G'$ is an internal vertex of a path in $\PP$).
-Thus $\PP$ is a $(k+1,2)$-shortcut system for $G$, such that $G'\subseteq G^\PP$.
+Let $\SS$ be the set of such paths $P$.
+For each vertex $v$ of $G$, at most two paths in $\SS$ use $v$ as an internal vertex (since no original vertex of $G'$ is an internal vertex of a path in $\SS$).
+Thus $\SS$ is a $(k+1,2)$-shortcut system for $G$, such that $G'\subseteq G^\SS$.
 This idea can be pushed further to obtain a rough characterisation of $k$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $k$-planar (see \cref{Characterisation}).
@@
 Let $G$ be a subgraph of $H\boxtimes P \boxtimes K_\ell$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
-Let $\PP$ be a $(k,d)$-shortcut system for $G$. Then $G^\PP$ is a subgraph of $J\boxtimes P\boxtimes K_{d\ell(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
+Let $\SS$ be a $(k,d)$-shortcut system for $G$. Then $G^\SS$ is a subgraph of $J\boxtimes P\boxtimes K_{d\ell(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
 \end{thm}
@@
 \note{DW}{Use $\SS$ for shortcut system, okay?}
-\note{PM}{Ok.}
+\note{PM}{Ok.  I've implemented this change.  I was careful, but will still reread.}
 
@@
 The \defin{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.\footnote{For a graph $G$ and two vertices $v,w\in V(G)$, $\dist_G(v,w)$  denotes the length of a shortest path, in $G$, with endpoints $v$ and $w$.
-We define $\dist_G(v,w):=\infty$ if $v$ and $w$ are in different connected components of $G$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\PP$ for some $(k,2k\Delta^{k})$-shortcut system $\PP$; see \cref{PowerShortcut}.
+We define $\dist_G(v,w):=\infty$ if $v$ and $w$ are in different connected components of $G$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\SS$ for some $(k,2k\Delta^{k})$-shortcut system $\SS$; see \cref{PowerShortcut}.
 Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} then imply:
@@
 %\begin{thm}
-%Let $\PP$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$.
-Then $G^\PP$ is a subgraph of $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
+%Let $\SS$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$.
+Then $G^\SS$ is a subgraph of $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
 %\end{thm}
@@
 %\begin{thm}
-%For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\PP$ for $G$, $G^\PP\subseteq H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
+%For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\SS$ for $G$, $G^\SS\subseteq H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
 %\end{thm}
@@
 %\begin{thm}
-%For every graph $X$ and for all integers $k,\Delta\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ with maximum degree $\Delta$ and for every $(k,\Delta)$-shortcut system $\PP$ for $G$, $G^\PP \subseteq H \boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
+%For every graph $X$ and for all integers $k,\Delta\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ with maximum degree $\Delta$ and for every $(k,\Delta)$-shortcut system $\SS$ for $G$, $G^\SS \subseteq H \boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
 %\end{thm}
@@
 The following result is the main technical contribution of the paper.
-Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so too does $G^\PP$ for every shortcut system $\PP$ of $G$.
+Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so too does $G^\SS$ for every shortcut system $\SS$ of $G$.
 
@@
 	\label{ShortcutPartition}
-	Let $G$ be a graph having an $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\PP$ be a $(k,d)$-shortcut system for $G$.
-Then $G^\PP$ has a $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$.
+	Let $G$ be a graph having an $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\SS$ be a $(k,d)$-shortcut system for $G$.
+Then $G^\SS$ has a $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$.
 \end{restatable}
@@
 The purpose of this section is to prove our main technical result, \cref{ShortcutPartition}.
-This theorem shows how, given a $(k,d)$-shortcut system $\mathcal{P}$ of a graph $G$, a $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\mathcal{P}}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
-
-For convenience, it will be helpful to assume that $\mathcal{P}$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.
-Since $G^\PP$ is defined to be a supergraph of $G$, this assumption has no effect on $G^{\mathcal{P}}$ but eliminates special cases in some of our proofs.
+This theorem shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, a $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
+
+For convenience, it will be helpful to assume that $\SS$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.
+Since $G^\SS$ is defined to be a supergraph of $G$, this assumption has no effect on $G^{\SS}$ but eliminates special cases in some of our proofs.
 
@@
 
-For a node $x\in V(T)$, we say that a shortcut $P\in\mathcal{P}$ \defin{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
-We say that a vertex $v\in V(G)$ \defin{participates} in $x$ if $v\in Y_x$, or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
+For a node $x\in V(T)$, we say that a shortcut $P\in\SS$ \defin{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
+We say that a vertex $v\in V(G)$ \defin{participates} in $x$ if $v\in Y_x$, or $\SS$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
 We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
@@
 \begin{proof}
-  Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\mathcal{P}, v\in \{v_0,\ldots,v_r\}\}$.
+  Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\SS, v\in \{v_0,\ldots,v_r\}\}$.
 Then $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
@@
   We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
-If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
-In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$, $(v_0,\ldots,v_r)\in\PP$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$.
+If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\SS$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
+In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$, $(v_0,\ldots,v_r)\in\SS$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\SS$ that crosses $x$.
 
@@
 
-\note{DW}{Use $\SS$ for the given shortcut system and $\PP$ for the partition?}
+\note{DW}{Use $\SS$ for the given shortcut system and $\SS$ for the partition?}
 
 For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
-Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.\footnote{The sets in $\mathcal{S}$ are disjoint because $a:V(G)\to V(T)$ is a function.
-The sets in $\mathcal{S}$ cover $V(G)$ since, for each $v\in V(G)$, $v\in S_{a(v)}$.} Let $J:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph.
-We consider $V(J)\subseteq V(T)$, where each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
-(Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $J$.)
-
-\referee{2}{Page 8.  Can you add a brief explanation why $\mathcal{S}$ is a
-	partition of $V(G)$?  It is clear that the sets in $\mathcal{S}$ are
+Observe that $\mathcal{P}:=(S_x : x\in V(T))$ is a partition of $V(G)$.\footnote{The sets in $\mathcal{P}$ are disjoint because $a:V(G)\to V(T)$ is a function.
+The sets in $\mathcal{P}$ cover $V(G)$ since, for each $v\in V(G)$, $v\in S_{a(v)}$.} Let $J:=G^\SS/\mathcal{P}$ denote the resulting quotient graph.
+We consider $V(J)\subseteq V(T)$, where each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\SS}$.
+(Any node $x\in V(T)$ with $S_x=\emptyset$ does not contribute a vertex to $J$.)
+
+\referee{2}{Page 8.  Can you add a brief explanation why $\mathcal{P}$ is a
+	partition of $V(G)$?  It is clear that the sets in $\mathcal{P}$ are
 	disjoint, but why do they cover $V(G)$?}
-\note{PM}{I added an explanation in a footnote since this is obvious.}
+\note{PM}{I added an explanation in a footnote since this is obvious.  In particular, for any function $f$, $\{f^{-1}(x):x\in\mathop{codomain}(f)\}$ is a partition of the $\mathop{domain}(f)$.}
 
@@
 \begin{compactenum}[(i)]
-  \item $\mathcal{S}$ has small layered width with respect to the layering $\mathcal{L}$ and that
+  \item $\mathcal{P}$ has small layered width with respect to the layering $\mathcal{L}$ and that
   \item $J$ has small treewidth.
@@
 
-First we need to understand the relationship between sets $S_x\in\mathcal{S}$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
+First we need to understand the relationship between sets $S_x\in\mathcal{P}$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
 
@@
 By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.
-Therefore, $\mathcal{P}$ contains a path $P$, with $v\in V(P)$, that crosses $x$.
+Therefore, $\SS$ contains a path $P$, with $v\in V(P)$, that crosses $x$.
 The path $P$ contains a proper subpath $v_0,v_1,\ldots,v_{r}$ such that $v=v_0$ and $v_r\in Y_x$.
@@
 
-Next we complete Step~(i) and show that $\mathcal{S}$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
+Next we complete Step~(i) and show that $\mathcal{P}$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
 
@@
 Refer to \cref{contribute}.
-If $w\in Y_x\cap L_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
+If $w\in Y_x\cap L_i$ and no path in $\SS$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
 
@@
 
-  Otherwise, consider some path $P\in\mathcal{P}$ that contains $w$ as an internal vertex.
+  Otherwise, consider some path $P\in\SS$ that contains $w$ as an internal vertex.
 If $w\in L_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap L_i$.
@@
 
-  For any $j$, the number of vertices $w\in L_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\mathcal{P}$. Therefore,
+  For any $j$, the number of vertices $w\in L_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\SS$. Therefore,
   \[  |S_x\cap L_i|\le d\ell  \, \Big(k+1 + 2k + \sum_{j=2}^k 2(k-j)\Big) %= d(2k+1) + \sum_{i=1}^{k-2} i
@@
   Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.
-Since $xy\in E(J)$, $G^\mathcal{P}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
-Since $vw\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.\footnote{Recall that we have made the assumption that $\mathcal{P}$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  By \cref{s-subset}, $v\in V_x$ and $w\in V_y$.
+Since $xy\in E(J)$, $G^\SS$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
+Since $vw\in E(G^{\SS})$,  $\SS$ contains a $vw$-shortcut $P$.\footnote{Recall that we have made the assumption that $\SS$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  By \cref{s-subset}, $v\in V_x$ and $w\in V_y$.
 By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
@@
   If $x_\delta\in C_x$, it is because $x_\delta x'\in E(J)$ for some $T$-descendant $x'$ of $x$.
-This implies $G^{\mathcal{P}}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
-This implies that $\mathcal{P}$ contains a $vw$-shortcut $P_{vw}$.
+This implies $G^{\SS}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
+This implies that $\SS$ contains a $vw$-shortcut $P_{vw}$.
 Let $v'$ be the second-last vertex of $P_{vw}$ (so $v'w\in E(G)$).
@@
   \begin{compactenum}
-    \item There exists $w'\in V(G)$ such that $\mathcal{P}$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
+    \item There exists $w'\in V(G)$ such that $\SS$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
     \item $w\in Y_\delta$; in this case, define $w':=w$ and let $P_{ww'}$ be the 1-vertex path $w$.
@@
 
-At this point, the proof of \cref{ShortcutPartition} is almost immediate from \cref{general-width,general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
-In particular, $G^{\mathcal{P}}$ may contain edges $vw$ with $v\in L_i$ and $w\in L_{i+j}$ for any $j\in\{0,\ldots,k\}$.
+At this point, the proof of \cref{ShortcutPartition} is almost immediate from \cref{general-width,general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\SS}$.
+In particular, $G^{\SS}$ may contain edges $vw$ with $v\in L_i$ and $w\in L_{i+j}$ for any $j\in\{0,\ldots,k\}$.
 To resolve this, we use a new layering $\mathcal{L}':=\langle L_0',\ldots,L_h'\rangle$ in which $L_i'=\bigcup_{j=ki}^{ki+k-1} L_i$.
@@
 An embedded graph $G$ is \defin{$k$-plane} if each edge of $G$ takes part in at most $k$ crossings.
-\note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}  A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
+\note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}\note{PM}{No, under the current definition, $r$ edges crossing at a single point $p$ generate $\binom{r}{2}$ crossings, with each edge taking part in $r-1$ crossings, just like they would if we perturbed the edges so that they all crossed at individual points.
+I guess we could mention this.
+(See the first paragraph in the proof of Theorem~2 below.)}  A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
 Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
@@
 In this way, each edge $vw\in E(G)$ corresponds naturally to a path $P_{vw}$ of length at most $k+1$ in $G_0$.
-Let $\PP := \{P_{vw}: vw\in E(G)\}$.
-Observe that $\PP$ is a $(k+1,2)$-shortcut system for $G_0$ and that $G_0^{\PP}\supseteq G$.
-Specifically, $G_0^{\PP}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.
+Let $\SS := \{P_{vw}: vw\in E(G)\}$.
+Observe that $\SS$ is a $(k+1,2)$-shortcut system for $G_0$ and that $G_0^{\SS}\supseteq G$.
+Specifically, $G_0^{\SS}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.
 
 	Since $G_0$ is planar,  \cref{PlanarProduct}(b) and \cref{PartitionProduct} implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
-Applying \cref{ShortcutPartition} to $G_0$ and $\mathcal{P}$ immediately implies that $G$ (an arbitrary $k$-planar graph) has an $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
+Applying \cref{ShortcutPartition} to $G_0$ and $\SS$ immediately implies that $G$ (an arbitrary $k$-planar graph) has an $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
 
@@
 In this setting, the proof of \cref{general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $L_i\cap Y_x$.
-More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (L_{i-j}\cup L_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap L_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap L_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap L_j$ for $j > k+1$.
+More precisely, each path $P\in\SS$ containing an internal (dummy) vertex $w\in Y_x\cap (L_{i-j}\cup L_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap L_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap L_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap L_j$ for $j > k+1$.
 	Redoing the calculation at the end of the proof of \cref{general-width} then yields
@@
 \label{gAddDummy}
-Every $(g,k)$-planar graph $G$ is a subgraph of $G_0^\PP$ for some graph $G_0$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G_0$.
+Every $(g,k)$-planar graph $G$ is a subgraph of $G_0^\SS$ for some graph $G_0$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\SS$ for $G_0$.
 Moreover, $V(G) \subseteq V(G_0)$ and for every edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most $k+1$, such that every internal vertex in $P$ has degree at most $4$ in $G_0$.
@@
 \label{MapShortcut}
-Every $(g,d)$-map graph $G$ is a subgraph of $G_1^\PP$ for some graph $G_1$ with Euler genus at most $g$ and some $(2,\tfrac12 d(d-3) )$-shortcut system $\PP$ for $G_1$.
+Every $(g,d)$-map graph $G$ is a subgraph of $G_1^\SS$ for some graph $G_1$ with Euler genus at most $g$ and some $(2,\tfrac12 d(d-3) )$-shortcut system $\SS$ for $G_1$.
 \end{lem}
@@
 In the latter case, let $P_{vw}$ be the path $(v,x,w)$.
-Let $\PP$ be the set of all such paths $P_{vw}$.
-Each vertex $x\in V(G_0)$ is the middle vertex on at most $\tfrac12 d(d-3)$  paths in $\PP$.
-Thus $\PP$ is a $(2,\tfrac12 d(d-3))$-shortcut system for $G_1$, and by construction, $G \subseteq G_1^\PP$.
+Let $\SS$ be the set of all such paths $P_{vw}$.
+Each vertex $x\in V(G_0)$ is the middle vertex on at most $\tfrac12 d(d-3)$  paths in $\SS$.
+Thus $\SS$ is a $(2,\tfrac12 d(d-3))$-shortcut system for $G_1$, and by construction, $G \subseteq G_1^\SS$.
 \end{proof}
@@
 \label{StringShortcut}
-Every $(g,\delta)$-string graph $G$ is a subgraph of $G_0^\PP$ for some graph $G_0$ with Euler genus at most $g$ and some $(\delta+1,\delta+1 )$-shortcut system $\PP$ for $G_0$.
-%Moreover, every internal vertex on the paths in $\PP$ has degree 4 in $G_0$.
+Every $(g,\delta)$-string graph $G$ is a subgraph of $G_0^\SS$ for some graph $G_0$ with Euler genus at most $g$ and some $(\delta+1,\delta+1 )$-shortcut system $\SS$ for $G_0$.
+%Moreover, every internal vertex on the paths in $\SS$ has degree 4 in $G_0$.
 \end{lem}
@@
 For each edge $vw$ of $G$, there is a path $P_{vw}$ of length at most $2\ceil{\frac{\delta}{2}}\leq \delta+1$ in $G_0$ between $v$ and $w$.
-Let $\PP$ be the set of all such paths $P_{vw}$.
-Consider a vertex $x$ in $G_0$ that is an internal vertex on some path in $\PP$.
+Let $\SS$ be the set of all such paths $P_{vw}$.
+Consider a vertex $x$ in $G_0$ that is an internal vertex on some path in $\SS$.
 Then $x$ is at the intersection of $C_v$ and $C_w$ for some edge $vw\in E(G)$.
-If some path $P\in \PP$ passes through $x$, then $P=P_{vu}$ for some edge $vu$ incident to $v$, or $P=P_{wu}$ for some edge $wu$ incident to $w$.
-At most $\ceil{\frac{\delta}{2}}$ paths in $\PP$ corresponding to edges incident to $v$ pass through $x$, and similarly for edges incident to $w$.
-Thus at most $2\ceil{\frac{\delta}{2}}\leq\delta+1$ paths in $\PP$ use $x$ as an internal vertex.
-Thus $\PP$ is a $(\delta+1,\delta+1)$-shortcut system for $G_0$, and by construction, $G \subseteq G_0^\PP$.
+If some path $P\in \SS$ passes through $x$, then $P=P_{vu}$ for some edge $vu$ incident to $v$, or $P=P_{wu}$ for some edge $wu$ incident to $w$.
+At most $\ceil{\frac{\delta}{2}}$ paths in $\SS$ corresponding to edges incident to $v$ pass through $x$, and similarly for edges incident to $w$.
+Thus at most $2\ceil{\frac{\delta}{2}}\leq\delta+1$ paths in $\SS$ use $x$ as an internal vertex.
+Thus $\SS$ is a $(\delta+1,\delta+1)$-shortcut system for $G_0$, and by construction, $G \subseteq G_0^\SS$.
 \end{proof}
@@
 \label{PowerShortcut}
-If a graph $G$ has maximum degree $\Delta$, then $G^k = G^\PP$ for some $(k,2k \Delta^{k})$-shortcut system $\PP$.
+If a graph $G$ has maximum degree $\Delta$, then $G^k = G^\SS$ for some $(k,2k \Delta^{k})$-shortcut system $\SS$.
 \end{lem}
@@
 $\dist_G(x,y)$  in $G$.
-Let $\PP:=\{P_{xy}: \dist_G(x,y)\in\{1,\dots,k\} \}$.
+Let $\SS:=\{P_{xy}: \dist_G(x,y)\in\{1,\dots,k\} \}$.
 Say $P_{xy}$ uses some vertex $v$ as an internal vertex.
@@
 The number of vertices at distance $i$ from $v$ is at most $\Delta^i$.
-Thus the number of paths in $\PP$ that use $v$ as an internal vertex is at most
+Thus the number of paths in $\SS$ that use $v$ as an internal vertex is at most
 \[\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i\Delta^j
@@
 < 2k \Delta^k\enspace.\]
-Hence $\PP$ is a $(k, 2k \Delta^k)$-shortcut system.
+Hence $\SS$ is a $(k, 2k \Delta^k)$-shortcut system.
 \end{proof}

### 2022-03-07 k-planar.tex
Pfffft

@@
 Subdivide the edge between $a$ and its parent in $T_0$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
-This modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
+Now $f(y)=a'$ and $f(x)=a$, so this modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
 
@@
 \note{DW}{Use $\SS$ for the given shortcut system and $\SS$ for the partition?}
+\note{PM}{Yes, I've done this.  Again, I was carefuly, but need to reread.}
 
@@
 
-\referee{2}{Page 8.  Can you add a brief explanation why $\mathcal{P}$ is a
-	partition of $V(G)$?  It is clear that the sets in $\mathcal{P}$ are
-	disjoint, but why do they cover $V(G)$?}
-\note{PM}{I added an explanation in a footnote since this is obvious.  In particular, for any function $f$, $\{f^{-1}(x):x\in\mathop{codomain}(f)\}$ is a partition of the $\mathop{domain}(f)$.}
+% \referee{2}{Page 8.  Can you add a brief explanation why $\mathcal{P}$ is a
+% 	partition of $V(G)$?  It is clear that the sets in $\mathcal{P}$ are
+% 	disjoint, but why do they cover $V(G)$?}
+% \note{PM}{I added an explanation in a footnote since this is obvious.  In particular, for any function $f$, $\{f^{-1}(x):x\in\operatorname{range}(f)\}$ is a partition of $\operatorname{domain}(f)$.}
 
@@
 \begin{compactenum}[(i)]
-  \item $\mathcal{P}$ has small layered width with respect to the layering $\mathcal{L}$ and that
+  \item $\mathcal{P}$ has small layered width with respect to the layering $\mathcal{L}$ of $G$ and that
   \item $J$ has small treewidth.
 \end{compactenum}
-
-\referee{1}{7.
-page 8, item (i): ``$S$ has small layered width with respect to the layering $L$'' $\longrightarrow$ ``$S$ has small layered width with respect to the layering $L$ of $G$''.
-Perhaps mention something about how it is easy to find the new layering.}
+Once we have established (i) and (ii), the result follows easily since a layering of $G^\SS$ is easily obtained from $\mathcal{L}$ by `compressing' groups of $k$ consecutive layers.
+% \referee{1}{7. page 8, item (i): ``$P$ has small layered width with respect to the layering $L$'' $\longrightarrow$ ``$P$ has small layered width with respect to the layering $L$ of $G$''.}
+% \referee{1}{Perhaps mention something about how it is easy to find the new layering.}
 
@@
   Recall that $S_x$ is defined by vertices that participate in $x$, and these are vertices that are either in $Y_x$ or in shortcuts that cross $Y_x$.
-We say that a vertex $w\in Y_x$ \defin{contributes} a vertex $v\in S_x$ if $v$ participates in $x$.
+We say that a vertex $w\in Y_x$ \defin{contributes} a vertex $v\in S_x$ if $v=w$ or if some path in $\SS$ that contains $v$ has $w$ as an internal vertex.
   We upper bound the number of vertices in $S_x\cap L_i$ by upper-bounding the number of vertices contributed to $S_x\cap L_i$ by each $w\in Y_x$.
-Refer to \cref{contribute}.
-If $w\in Y_x\cap L_i$ and no path in $\SS$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
-
-\referee{1}{9. page 8, line 10: ``We say that a vertex $w \in Y_x$ contributes a vertex
-  $v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.}
-
-\note{DW}{What is wrong here?}
+
+  Refer to \cref{contribute}.  If $w\in Y_x\cap L_i$ and no path in $\SS$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
+  Otherwise, consider some path $P\in\SS$ that contains $w$ as an internal vertex.
+If $w\in L_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap L_i$.
+If $w\in L_{i-1}\cup L_{i+1}$, then $P$ contributes at most $k$ vertices to $S_x\cap L_i$.
+If $w\in L_{i-j}\cup L_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap L_i$.
+% 
+% \referee{1}{9. page 8, line 10: ``We say that a vertex $w \in Y_x$ contributes a vertex
+%   $v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.}
+% 
+% \note{DW}{What is wrong here?}
 
@@
   \end{figure}
-
-  Otherwise, consider some path $P\in\SS$ that contains $w$ as an internal vertex.
-If $w\in L_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap L_i$.
-If $w\in L_{i-1}\cup L_{i+1}$, then $P$ contributes at most $k$ vertices to $S_x\cap L_i$.
-If $w\in L_{i-j}\cup L_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap L_i$.
 

### 2022-03-07 k-planar.tex
Pfffffft

@@
 An embedded graph $G$ is \defin{$k$-plane} if each edge of $G$ takes part in at most $k$ crossings.
-\note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}\note{PM}{No, under the current definition, $r$ edges crossing at a single point $p$ generate $\binom{r}{2}$ crossings, with each edge taking part in $r-1$ crossings, just like they would if we perturbed the edges so that they all crossed at individual points.
+\note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}\note{PM}{No, under the current definition, $r$ edges crossing at a single point $p$ generate $\binom{r}{2}$ crossings, with each edge taking part in $r-1$ crossings, just like they would if we perturbed the edges so that each pair crosses at a distinct point.
 I guess we could mention this.
@@
 }
+
+\note{PM}{I agree that we should fix it and keep it.
+Another reason: For vertex ranking, the bound is $O(\log n/\log^{(t)} n)$ where $t:=\tw(H)$, so this is an example where $\tw(H)$ affects more than just the constant.}
 

### 2022-03-07 k-planar.tex
Cleaning up the 1-planar proof

@@
 
-Let $G$ be an edge-maximal 1-plane multigraph with no two parallel edges on the boundary of a single face.
+Let $G$ be an edge-maximal $1$-plane multigraph with no two parallel edges on the boundary of a single face.
 Here, edge-maximal should be taken to mean that, if any two vertices $v$ and $w$ appear on a common face\footnote{The \defin{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
 We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.
-\note{DW}{Each face of $G$ is one of two types, either bounded by three edges (none of which are crossed), or bounded by one edge that is not crossed plus portions of two crossing edges.
+Refer to \cref{edge_maximal}.
+We assume that no two edges incident to a common vertex of $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c@{\hspace{1cm}}c}
+      \includegraphics{figs/one_planar_example-1} &
+      \includegraphics{figs/one_planar_example-2} \\
+      (a) & (b)
+    \end{tabular}
+  \end{center}
+  \caption{Examples of (a)~a $1$-plane graph that is not edge maximal; and (b)~an edge-maximal $1$-plane graph.}
+\end{figure}
+
+To understand the structure of an edge maximal $1$-plane multigraph $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of crossing edges $vw$ and $xy$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \emph{dummy vertex} $p$.
+Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_r$ be the facial walk around $F$.
+By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
+
+Furthermore, for any edge $w_iw_{i+1}$ of $W$, at least one of $w_i$ or $w_{i+1}$ is not a dummy vertex since, otherwise this would imply that the edge of $G$ that contains $w_iw_{i+1}$ is involved in at least two crossings.
+Therefore each dummy vertex $w_i$ of $W$ is surrounded by two vertices $w_{i-1},w_{i+1}\in V(G)$.
+Since $w_i$ has degree $4$ in $G_0$, $w_{i-1}\neq w_i$.
+Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.
+Therefore, if $F$ has a dummy vertex in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.
+
+Therefore, in such a graph each face of $G$ is one of two types:
+\begin{inparaenum}[(i)] 
+  \item bounded by three edges (none of which are crossed) or 
+  \item bounded by one edge that is not crossed plus portions of two edges that cross each other.
+\end{inparaenum}
+Consider some crossing $(p,vw,xy)$ in $G$.
+The point $p$ at which this crossing occurs is a vertex of $G_0$ that has degree $4$.
+There are four faces $F_1,\ldots,F_4$ of $G_0$ that contain $p$.
+Each of these four faces of is of Type~(ii), so the boundary of $F_1\cup\cdots\cup F_4$ includes a $4$-cycle $C:=vxwy$ in $G$ whose edges are not involved in any crossings.
+We call the subgraph $K:=C\cup\{vw,xy\}$ a \defin{kite} of $G$.
+The edges $vw$ and $xy$ are called \defin{spars} of $K$.
+The cycle $vxwy$ is called the \defin{sail} of $K$.
+Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
+
+\note{PM}{I've commented out a lot of red and blue here to make this readable.}
+% 
+% It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$.
+
+% 
+% \note{DW}{Each face of $G$ is one of two types, either bounded by three edges (none of which are crossed), or bounded by one edge that is not crossed plus portions of two crossing edges.
 Right?
@@
 Add a figure.}
-We assume that no two edges incident to a common vertex cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}
-
-\note{DW}{I suggest we say that it follows from Euler's formula that $G$ is finite.
+% 
+% \note{DW}{I suggest we say that it follows from Euler's formula that $G$ is finite.
 (To see this, say $G$ has $n$ vertices, $m$ non-crossed edges, and $k$ crossings.
@@
 ) }
-
-\note{DW}{It seems to be that $G$ can be obtained from a plane multigraph with faces of size 3 or 4 such that no edge is in two 3-faces, by adding a pair of crossing edges across each 4-face.
+% 
+% \note{DW}{It seems to be that $G$ can be obtained from a plane multigraph with faces of size 3 or 4 such that no edge is in two 3-faces, by adding a pair of crossing edges across each 4-face.
 I think this characterises edge-maximal 1-planar multigraphs with no 2-faces.
@@
 Is it known?}
-
-\note{DW}{Say $G$ is obtained from a plane multigraph by making each face a clique. Is $G\subseteq H \boxtimes P \boxtimes K_{f(\omega(G)}$? for some treewidth 3 graph $H$.}
-
-A \defin{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.
+% 
+% \note{DW}{Say $G$ is obtained from a plane multigraph by making each face a clique. Is $G\subseteq H \boxtimes P \boxtimes K_{f(\omega(G)}$? for some treewidth 3 graph $H$.}
+% 
+% % A \defin{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.
 It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
@@
 }
-The edges $vw$ and $xy$ are called \defin{spars} of $K$.
-The cycle $vxwy$ is called the \defin{sail} of $K$.
-It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$.
-
-\referee{2}{Page 13. Perhaps I am misunderstanding something, but I do not see why
-'none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of
-$G$.'  For example, see the attached PDF for a picture where $vw$, $xw$, $wy$,
-and $yv$ are all crossed by other edges of $G$.\\
-\includegraphics[width=60mm]{kite}}
-\note{DW}{We need to explain more about how edge-maximality fixes this problem.
-In this case we can draw other $xv$, $vy$, $yw$ $wx$ edges close to the central crossing edges.} Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
-
-\referee{2}{Page 13. The attached PDF also shows that the definition of 'kite
-face' may not be well-defined.  For example, I do not see why there
-cannot be some edges of $G$ 'inside' a kite face.  Even if this is not
-the case, I think it is more precise to say that a kite face has 'two
-and a half edges' and two vertices of $G$ on its boundary' rather than
-'three edges and two vertices of $G$ on its boundary'.}
-\note{DW}{This should be fixed if we define kite more precisely. A kite-face is bounded by one edge and two half-edges right?}
-
-\referee{2}{Page 13. Could it be that the subgraph induced by the vertices of a
-	kite is a $K_4$ with some parallel edges?  This seems relevant later in
-	the proof.}
-
-\note{DW}{If we define kite as above, this is not a problem}
-
-The 1-plane graph $G$ has a plane triangulation $G'$ as a subgraph that can be obtained by removing one spar from each kite in $G$.
+% 
+% \referee{2}{Page 13. Perhaps I am misunderstanding something, but I do not see why
+% 'none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of
+% $G$.'  For example, see the attached PDF for a picture where $vw$, $xw$, $wy$,
+% and $yv$ are all crossed by other edges of $G$.\\
+% \includegraphics[width=60mm]{kite}}
+% \note{DW}{We need to explain more about how edge-maximality fixes this problem.
+In this case we can draw other $xv$, $vy$, $yw$ $wx$ edges close to the central crossing edges.} \note{PM}{In the example above, $v$ and $y$ appear on a common (inner) face $F$, so by definition of edge-maximality there is an edge $xy$  on the boundary of $F$.
+}
+% 
+% \referee{2}{Page 13. The attached PDF also shows that the definition of 'kite
+% face' may not be well-defined.  For example, I do not see why there
+% cannot be some edges of $G$ 'inside' a kite face.  Even if this is not
+% the case, I think it is more precise to say that a kite face has 'two
+% and a half edges' and two vertices of $G$ on its boundary' rather than
+% 'three edges and two vertices of $G$ on its boundary'.}
+% \note{DW}{This should be fixed if we define kite more precisely. A kite-face is bounded by one edge and two half-edges right?}
+% 
+% \referee{2}{Page 13. Could it be that the subgraph induced by the vertices of a
+% 	kite is a $K_4$ with some parallel edges?  This seems relevant later in
+% 	the proof.}
+% 
+% \note{DW}{If we define kite as above, this is not a problem}
+
+By removing one spar from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three edges.
 Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).

### 2022-03-07 k-planar.tex
Now we have good definitions

@@
 We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.
-Refer to \cref{edge_maximal}.
-We assume that no two edges incident to a common vertex of $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}
+(Refer to \cref{one_planar_example}(a,b).) We assume that no two edges incident to a common vertex of $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}
 
@@
       \includegraphics{figs/one_planar_example-2} \\
-      (a) & (b)
+      (a) & (b) \\[1em]
+      \includegraphics{figs/one_planar_example-3} &
+      \includegraphics{figs/one_planar_example-4} \\
+      (c) & (d) \\[1em]
     \end{tabular}
   \end{center}
-  \caption{Examples of (a)~a $1$-plane graph that is not edge maximal; and (b)~an edge-maximal $1$-plane graph.}
+  \caption{Examples of (a)~a $1$-plane graph that is not edge maximal; (b)~an edge-maximal $1$-plane graph $G$; (c)~the kites in $G$; (d)~a plane multigraph $G'$ obtained by removing one spar from each kite of $G$.}
+  \label{one_planar_example} 
+  \note{PM}{I suggest we only keep parts (c) and (d) of this figure.}
 \end{figure}
 
-To understand the structure of an edge maximal $1$-plane multigraph $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of crossing edges $vw$ and $xy$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \emph{dummy vertex} $p$.
+To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \emph{dummy vertex} $p$.
 Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_r$ be the facial walk around $F$.
@@
 
-Furthermore, for any edge $w_iw_{i+1}$ of $W$, at least one of $w_i$ or $w_{i+1}$ is not a dummy vertex since, otherwise this would imply that the edge of $G$ that contains $w_iw_{i+1}$ is involved in at least two crossings.
+If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) is involved in at least two crossings.
 Therefore each dummy vertex $w_i$ of $W$ is surrounded by two vertices $w_{i-1},w_{i+1}\in V(G)$.
@@
 Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.
-Therefore, if $F$ has a dummy vertex in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.
-
-Therefore, in such a graph each face of $G$ is one of two types:
+Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.
+
+Therefore, each face of $G$ is one of two types:
 \begin{inparaenum}[(i)] 
-  \item bounded by three edges (none of which are crossed) or 
-  \item bounded by one edge that is not crossed plus portions of two edges that cross each other.
+  \item bounded by three edges of $G$ (none of which are crossed) or 
+  \item bounded by one edge of $G$ that is not crossed plus portions of two edges of $G$ that cross each other.
 \end{inparaenum}
@@
 We call the subgraph $K:=C\cup\{vw,xy\}$ a \defin{kite} of $G$.
-The edges $vw$ and $xy$ are called \defin{spars} of $K$.
+(See \cref{one_planar_example}(c).)  The edges $vw$ and $xy$ are called \defin{spars} of $K$.
 The cycle $vxwy$ is called the \defin{sail} of $K$.
@@
 
-\note{PM}{I've commented out a lot of red and blue here to make this readable.}
+\note{PM}{I've commented out a lot of red and blue here to make this readable. Those notes are still in the \LaTeX\ source.}
 % 
@@
 
+Refer to \cref{one_planar_example}(d).
 By removing one spar from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three edges.

### 2022-03-07 k-planar.tex
updates

--- 2022-03-07 k-planar.tex
@@
 
-% \newcommand{\PP}{\mathcal{P}}
+\newcommand{\PP}{\mathcal{P}}
 \renewcommand{\SS}{\mathcal{S}}
@@
 \begin{abstract}
-Dujmovi\'c~et~al.~[\emph{J.~ACM}~'20] recently proved that every planar graph is isomorphic to a subgraph of the strong product of a graph of bounded treewidth and a path.
+Dujmovi\'c~et~al.~[\emph{J.~ACM}~'20] recently proved that every planar graph is a subgraph of the strong product of a graph of bounded treewidth and a path.
 Analogous results were obtained for graphs of bounded Euler genus or apex-minor-free graphs.
@@
 One noteable example is $k$-planar graphs (those with a drawing in the plane in which each edge is involved in at most $k$ crossings).
-We prove that every $k$-planar graph is isomorphic to a subgraph of the strong product of a graph of treewidth $O(k^5)$ and a path.
+We prove that every $k$-planar graph is a subgraph of the strong product of a graph of treewidth $O(k^5)$ and a path.
 This is the first result of this type for a non-minor-closed class of graphs.
@@
 We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.
-We say that a graph $G$ is \emph{contained in} a graph $X$ if $G$ is isomorphic to a subgraph of $X$.
-Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (improved by \citet{UWY}).
-
-\begin{thm}[\citep{DJMMUW20,UWY}]
+Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20}.
+
+\begin{thm}[\citep{DJMMUW20}]
 	\label{PlanarProduct}
@@
 
-% \note{DW}{In 	\cref{PlanarProduct} and elsewhere, we really should write, ``Every planar graph is isomorphic to a subgraph of''.
+\note{DW}{In 	\cref{PlanarProduct} and elsewhere, we really should write, ``Every planar graph is isomorphic to a subgraph of''.
 One alternative is to define that $H$ is \defin{contained} in $G$ if $H$ is isomorphic to a subgraph of $G$.
@@
 
-% \note{DW}{I suggest we replace $8$ by $6$ here, and replace ``The starting point for this work is the following `product structure theorem' for planar graphs by \citet{DJMMUW20} (improved by \citet{UWY})''.
+\note{DW}{I suggest we replace $8$ by $6$ here, and replace ``The starting point for this work is the following `product structure theorem' for planar graphs by \citet{DJMMUW20} (improved by \citet{UWY})''.
 Okay?}
 
-Here $\boxtimes$ is the strong product,\!\footnote{The \emph{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
+Here $\boxtimes$ is the strong product\footnote{The \emph{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
 	$v=w$ and $xy\in E(B)$, or
 	$x=y$ and $vw\in E(A)$, or
-	$vw\in E(A)$ and $xy\in E(B)$.}
+	$vw\in E(A)$ and $xy\in E(B)$.}, 
 and treewidth\footnote{For a tree $T$, a \defin{$T$-decomposition} of a graph $G$ is a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that
@@
 \begin{compactitem}
-\item \citet{DJMMUW20} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92}).
-\item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
+\item \citet{DJMMUW20} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92} from 1992).
+\item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02} from 2002).
 \item \citet{DFMS21} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs. 
-\item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results starting with the work of \citet{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}).
-\citet{DEJGMM21} have since used it to find asymptotically optimal adjacency labellings of planar graphs.
+\item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results going back to 1988~\citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}), and \citet{DEJGMM21} have since used it to find asymptotically optimal adjacency labellings of planar graphs.
 \item The result of \citet{DEJGMM21} implies that, for every integer
-$n>0$, there is a `induced-universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
+$n>0$, there is a `universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
 subgraph of $U_n$.
-This has recently been extended by \citet{EJM}, who show the existence of a `subgraph-universal' graph with $(1+\epsilon)n$ vertices that contains every $n$-vertex planar graph as a subgraph as well as the existence of an `induced universal' graph with $n^{1+o(1)}$ vertices \emph{and} edges.
-The former result is the first progress on subgraph-universal graphs for planar graphs since the $O(n^{3/2})$-vertex construction of \citet{babai.chung.ea:on}.
-% aking the first progress on this problem since the work of \citet{babai.xx} in 1984.
 \end{compactitem}
 
-% \note{DW}{Delete `from 1992' and  `from 2007'? replace ``going back to 1988~\citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}'' by
-	% ``starting with the work of \citet{kannan.naor.ea:implicit}''}
-% \note{PM}{Was this a referee request, or just because we don't need to sell it anymore?  I'm fine with the change either way, but didn't implement it.}
-
-% \note{DW}{Add \citep{EJM} to the story.}
-
-All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this definition.} These applications motivate finding product structure theorems for other graph classes.
+\note{DW}{Delete `from 1992' and  `from 2007'? replace ``going back to 1988~\citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}'' by 
+	``starting with the work of \citet{kannan.naor.ea:implicit}''}
+
+\note{DW}{Add \citep{EJM} to the story.}
+
+All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is a subgraph of $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this definition.}.
+These applications motivate finding product structure theorems for other graph classes.
 \citet{DJMMUW20} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
@@
 
-\note{PM}{Kind of, but I suggest we keep that paragraph and section anyway.
-Justification: They're both short and they describe the exact bounds for bounded genus graphs, which you don't get from the one-paragraph description in Section~1.0.}
-
 \subsection{Shortcut Systems}
@@
 Although $k$-planar graphs are the most high-profile target for a generalization of \cref{PlanarProduct}, we actually prove a substantially stronger result than \cref{kPlanarProduct} using the following definition.
-A non-empty set $\SS$ of non-trivial paths in a graph $G$ is a \defin{$(k,d)$-shortcut system} (for $G$) if:
+A non-empty set $\mathcal{P}$ of non-trivial paths in a graph $G$ is a \defin{$(k,d)$-shortcut system} (for $G$) if:
 
 \begin{compactitem}
-\item every path in $\SS$ has length at most $k$,\footnote{A path of length $k$ consists of $k$ edges and $k+1$ vertices.
+\item every path in $\mathcal{P}$ has length at most $k$,\footnote{A path of length $k$ consists of $k$ edges and $k+1$ vertices.
 A path is \defin{trivial} if it has length 0 and \defin{non-trivial} otherwise.} and
-\item for every $v\in V(G)$, the number of paths in $\SS$ that use $v$ as an internal vertex is at most $d$.
+\item for every $v\in V(G)$, the number of paths in $\mathcal{P}$ that use $v$ as an internal vertex is at most $d$.
 \end{compactitem}
-Each path $P\in\SS$ is called a \defin{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \defin{$vw$-shortcut}.
-Given a graph $G$ and a $(k,d)$-shortcut system $\SS$ for $G$, let $G^{\SS}$ denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\SS$.
+Each path $P\in\mathcal{P}$ is called a \defin{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \defin{$vw$-shortcut}.
+Given a graph $G$ and a $(k,d)$-shortcut system $\mathcal{P}$ for $G$, let $G^{\mathcal{P}}$ denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\mathcal{P}$.
 
@@
 \label{AddDummy}
-Every $k$-planar graph is a subgraph of $G^\SS$ for some planar graph $G$ and some $(k+1,2)$-shortcut system $\SS$ for $G$.
+Every $k$-planar graph is a subgraph of $G^\PP$ for some planar graph $G$ and some $(k+1,2)$-shortcut system $\PP$ for $G$.
 \end{obs}
@@
 For each edge $vw\in E(G')$ there is a path $P$ in $G$ between $v$ and $w$ of length at most $k+1$ in which every internal vertex is a dummy vertex.
-Let $\SS$ be the set of such paths $P$.
-For each vertex $v$ of $G$, at most two paths in $\SS$ use $v$ as an internal vertex (since no original vertex of $G'$ is an internal vertex of a path in $\SS$).
-Thus $\SS$ is a $(k+1,2)$-shortcut system for $G$, such that $G'\subseteq G^\SS$.
+Let $\PP$ be the set of such paths $P$.
+For each vertex $v$ of $G$, at most two paths in $\PP$ use $v$ as an internal vertex (since no original vertex of $G'$ is an internal vertex of a path in $\PP$).
+Thus $\PP$ is a $(k+1,2)$-shortcut system for $G$, such that $G'\subseteq G^\PP$.
 This idea can be pushed further to obtain a rough characterisation of $k$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $k$-planar (see \cref{Characterisation}).
@@
 Let $G$ be a subgraph of $H\boxtimes P \boxtimes K_\ell$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
-Let $\SS$ be a $(k,d)$-shortcut system for $G$. Then $G^\SS$ is a subgraph of $J\boxtimes P\boxtimes K_{d\ell(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
+Let $\PP$ be a $(k,d)$-shortcut system for $G$. Then $G^\PP$ is a subgraph of $J\boxtimes P\boxtimes K_{d\ell(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
 \end{thm}
@@
 \note{DW}{Use $\SS$ for shortcut system, okay?}
-\note{PM}{Ok.  I've implemented this change.  I was careful, but will still reread.}
 
@@
 The \defin{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.\footnote{For a graph $G$ and two vertices $v,w\in V(G)$, $\dist_G(v,w)$  denotes the length of a shortest path, in $G$, with endpoints $v$ and $w$.
-We define $\dist_G(v,w):=\infty$ if $v$ and $w$ are in different connected components of $G$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\SS$ for some $(k,2k\Delta^{k})$-shortcut system $\SS$; see \cref{PowerShortcut}.
+We define $\dist_G(v,w):=\infty$ if $v$ and $w$ are in different connected components of $G$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\PP$ for some $(k,2k\Delta^{k})$-shortcut system $\PP$; see \cref{PowerShortcut}.
 Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} then imply:
@@
 
-\begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
+\begin{thm}[\citep{DJMMUW20}]
 	\label{GenusProduct}
@@
 	\begin{compactenum}[(a)]
-		\item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
-		\item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
+		\item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+8$  and some path $P$.
+		\item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $4$ and for some path $P$.
 	\end{compactenum}
@@
 }
-\note{PM}{I agree with this suggestion.  I've only added the other two references here, and haven't changed anything elsewhere yet.}
 
@@
 %\begin{thm}
-%Let $\SS$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$.
-Then $G^\SS$ is a subgraph of $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
+%Let $\PP$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$.
+Then $G^\PP$ is a subgraph of $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
 %\end{thm}
@@
 %\begin{thm}
-%For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\SS$ for $G$, $G^\SS\subseteq H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
+%For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\PP$ for $G$, $G^\PP\subseteq H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
 %\end{thm}
@@
 %\begin{thm}
-%For every graph $X$ and for all integers $k,\Delta\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ with maximum degree $\Delta$ and for every $(k,\Delta)$-shortcut system $\SS$ for $G$, $G^\SS \subseteq H \boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
+%For every graph $X$ and for all integers $k,\Delta\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ with maximum degree $\Delta$ and for every $(k,\Delta)$-shortcut system $\PP$ for $G$, $G^\PP \subseteq H \boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
 %\end{thm}
@@
 The following result is the main technical contribution of the paper.
-Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so too does $G^\SS$ for every shortcut system $\SS$ of $G$.
+Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so too does $G^\PP$ for every shortcut system $\PP$ of $G$.
 
@@
 	\label{ShortcutPartition}
-	Let $G$ be a graph having an $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\SS$ be a $(k,d)$-shortcut system for $G$.
-Then $G^\SS$ has a $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$.
+	Let $G$ be a graph having an $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\PP$ be a $(k,d)$-shortcut system for $G$.
+Then $G^\PP$ has a $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$.
 \end{restatable}
@@
 The purpose of this section is to prove our main technical result, \cref{ShortcutPartition}.
-This theorem shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, a $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
-
-For convenience, it will be helpful to assume that $\SS$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.
-Since $G^\SS$ is defined to be a supergraph of $G$, this assumption has no effect on $G^{\SS}$ but eliminates special cases in some of our proofs.
+This theorem shows how, given a $(k,d)$-shortcut system $\mathcal{P}$ of a graph $G$, a $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\mathcal{P}}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
+
+For convenience, it will be helpful to assume that $\mathcal{P}$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.
+Since $G^\PP$ is defined to be a supergraph of $G$, this assumption has no effect on $G^{\mathcal{P}}$ but eliminates special cases in some of our proofs.
 
@@
 Subdivide the edge between $a$ and its parent in $T_0$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
-Now $f(y)=a'$ and $f(x)=a$, so this modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
-
-  Next, remove the node $x_0$ from $T_0$, and consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.
+This modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
+
+  Next, consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.
 In this case, $B_{a}\subseteq B_{a'}$ where $a'$ is the parent of $a$ since any $x\in B_a\setminus B_{a'}$ would have $f(x)=a$.
@@
 
-% \referee{1}{6. page 6, last paragraph of Lemma 2: This isn’t quite right since you have to be careful with $x_0$.}
-% \note{PM}{I addressed this by removing $x_0$ from $T_0$ at the beginning of the second paragraph.}
+\referee{1}{6. page 6, last paragraph of Lemma 2: This isn’t quite right since you
+have to be careful with $x_0$.}
 
@@
 
-For a node $x\in V(T)$, we say that a shortcut $P\in\SS$ \defin{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
-We say that a vertex $v\in V(G)$ \defin{participates} in $x$ if $v\in Y_x$, or $\SS$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
+For a node $x\in V(T)$, we say that a shortcut $P\in\mathcal{P}$ \defin{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
+We say that a vertex $v\in V(G)$ \defin{participates} in $x$ if $v\in Y_x$, or $\mathcal{P}$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
 We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
@@
 \begin{proof}
-  Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\SS, v\in \{v_0,\ldots,v_r\}\}$.
+  Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\mathcal{P}, v\in \{v_0,\ldots,v_r\}\}$.
 Then $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
@@
   We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
-If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\SS$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
-In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$, $(v_0,\ldots,v_r)\in\SS$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\SS$ that crosses $x$.
+If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
+In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$, $(v_0,\ldots,v_r)\in\PP$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\mathcal{P}$ that crosses $x$.
 
@@
 
-\note{DW}{Use $\SS$ for the given shortcut system and $\SS$ for the partition?}
-\note{PM}{Yes, I've done this.  Again, I was carefuly, but need to reread.}
+\note{DW}{Use $\SS$ for the given shortcut system and $\PP$ for the partition?}
 
 For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
-Observe that $\mathcal{P}:=(S_x : x\in V(T))$ is a partition of $V(G)$.\footnote{The sets in $\mathcal{P}$ are disjoint because $a:V(G)\to V(T)$ is a function.
-The sets in $\mathcal{P}$ cover $V(G)$ since, for each $v\in V(G)$, $v\in S_{a(v)}$.} Let $J:=G^\SS/\mathcal{P}$ denote the resulting quotient graph.
-We consider $V(J)\subseteq V(T)$, where each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\SS}$.
-(Any node $x\in V(T)$ with $S_x=\emptyset$ does not contribute a vertex to $J$.)
-
-% \referee{2}{Page 8.  Can you add a brief explanation why $\mathcal{P}$ is a
-% 	partition of $V(G)$?  It is clear that the sets in $\mathcal{P}$ are
-% 	disjoint, but why do they cover $V(G)$?}
-% \note{PM}{I added an explanation in a footnote since this is obvious.  In particular, for any function $f$, $\{f^{-1}(x):x\in\operatorname{range}(f)\}$ is a partition of $\operatorname{domain}(f)$.}
+Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.
+Let $J:=G^\mathcal{P}/\mathcal{S}$ denote the resulting quotient graph.
+We consider $V(J)\subseteq V(T)$, where each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\mathcal{P}}$.
+(Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $J$.)
+
+\referee{2}{Page 8.  Can you add a brief explanation why $\mathcal{S}$ is a
+	partition of $V(G)$?  It is clear that the sets in $\mathcal{S}$ are
+	disjoint, but why do they cover $V(G)$?}
 
@@
 \begin{compactenum}[(i)]
-  \item $\mathcal{P}$ has small layered width with respect to the layering $\mathcal{L}$ of $G$ and that
+  \item $\mathcal{S}$ has small layered width with respect to the layering $\mathcal{L}$ and that
   \item $J$ has small treewidth.
 \end{compactenum}
-Once we have established (i) and (ii), the result follows easily since a layering of $G^\SS$ is easily obtained from $\mathcal{L}$ by `compressing' groups of $k$ consecutive layers.
-% \referee{1}{7. page 8, item (i): ``$P$ has small layered width with respect to the layering $L$'' $\longrightarrow$ ``$P$ has small layered width with respect to the layering $L$ of $G$''.}
-% \referee{1}{Perhaps mention something about how it is easy to find the new layering.}
-
-First we need to understand the relationship between sets $S_x\in\mathcal{P}$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
+
+\referee{1}{7.
+page 8, item (i): ``$S$ has small layered width with respect to the layering $L$'' $\longrightarrow$ ``$S$ has small layered width with respect to the layering $L$ of $G$''.
+Perhaps mention something about how it is easy to find the new layering.}
+
+First we need to understand the relationship between sets $S_x\in\mathcal{S}$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
 
@@
 By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.
-Therefore, $\SS$ contains a path $P$, with $v\in V(P)$, that crosses $x$.
+Therefore, $\mathcal{P}$ contains a path $P$, with $v\in V(P)$, that crosses $x$.
 The path $P$ contains a proper subpath $v_0,v_1,\ldots,v_{r}$ such that $v=v_0$ and $v_r\in Y_x$.
@@
 
-Next we complete Step~(i) and show that $\mathcal{P}$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
+Next we complete Step~(i) and show that $\mathcal{S}$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
 
@@
   Recall that $S_x$ is defined by vertices that participate in $x$, and these are vertices that are either in $Y_x$ or in shortcuts that cross $Y_x$.
-We say that a vertex $w\in Y_x$ \defin{contributes} a vertex $v\in S_x$ if $v=w$ or if some path in $\SS$ that contains $v$ has $w$ as an internal vertex.
+We say that a vertex $w\in Y_x$ \defin{contributes} a vertex $v\in S_x$ if $v$ participates in $x$.
   We upper bound the number of vertices in $S_x\cap L_i$ by upper-bounding the number of vertices contributed to $S_x\cap L_i$ by each $w\in Y_x$.
-
-  Refer to \cref{contribute}.  If $w\in Y_x\cap L_i$ and no path in $\SS$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
-  Otherwise, consider some path $P\in\SS$ that contains $w$ as an internal vertex.
-If $w\in L_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap L_i$.
-If $w\in L_{i-1}\cup L_{i+1}$, then $P$ contributes at most $k$ vertices to $S_x\cap L_i$.
-If $w\in L_{i-j}\cup L_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap L_i$.
-% 
-% \referee{1}{9. page 8, line 10: ``We say that a vertex $w \in Y_x$ contributes a vertex
-%   $v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.}
-% 
-% \note{DW}{What is wrong here?}
+Refer to \cref{contribute}.
+If $w\in Y_x\cap L_i$ and no path in $\mathcal{P}$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
+
+\referee{1}{9. page 8, line 10: ``We say that a vertex $w \in Y_x$ contributes a vertex
+  $v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.}
+
+\note{DW}{What is wrong here?}
 
@@
 
-  For any $j$, the number of vertices $w\in L_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\SS$. Therefore,
+  Otherwise, consider some path $P\in\mathcal{P}$ that contains $w$ as an internal vertex.
+If $w\in L_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap L_i$.
+If $w\in L_{i-1}\cup L_{i+1}$, then $P$ contributes at most $k$ vertices to $S_x\cap L_i$.
+If $w\in L_{i-j}\cup L_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap L_i$.
+
+  For any $j$, the number of vertices $w\in L_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\mathcal{P}$. Therefore,
   \[  |S_x\cap L_i|\le d\ell  \, \Big(k+1 + 2k + \sum_{j=2}^k 2(k-j)\Big) %= d(2k+1) + \sum_{i=1}^{k-2} i
@@
   Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.
-Since $xy\in E(J)$, $G^\SS$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
-Since $vw\in E(G^{\SS})$,  $\SS$ contains a $vw$-shortcut $P$.\footnote{Recall that we have made the assumption that $\SS$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  By \cref{s-subset}, $v\in V_x$ and $w\in V_y$.
+Since $xy\in E(J)$, $G^\mathcal{P}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
+Since $vw\in E(G^{\mathcal{P}})$,  $\mathcal{P}$ contains a $vw$-shortcut $P$.\footnote{Recall that we have made the assumption that $\mathcal{P}$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  By \cref{s-subset}, $v\in V_x$ and $w\in V_y$.
 By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
@@
   If $x_\delta\in C_x$, it is because $x_\delta x'\in E(J)$ for some $T$-descendant $x'$ of $x$.
-This implies $G^{\SS}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
-This implies that $\SS$ contains a $vw$-shortcut $P_{vw}$.
+This implies $G^{\mathcal{P}}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
+This implies that $\mathcal{P}$ contains a $vw$-shortcut $P_{vw}$.
 Let $v'$ be the second-last vertex of $P_{vw}$ (so $v'w\in E(G)$).
@@
   \begin{compactenum}
-    \item There exists $w'\in V(G)$ such that $\SS$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
+    \item There exists $w'\in V(G)$ such that $\mathcal{P}$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
     \item $w\in Y_\delta$; in this case, define $w':=w$ and let $P_{ww'}$ be the 1-vertex path $w$.
@@
 
-At this point, the proof of \cref{ShortcutPartition} is almost immediate from \cref{general-width,general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\SS}$.
-In particular, $G^{\SS}$ may contain edges $vw$ with $v\in L_i$ and $w\in L_{i+j}$ for any $j\in\{0,\ldots,k\}$.
+At this point, the proof of \cref{ShortcutPartition} is almost immediate from \cref{general-width,general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.
+In particular, $G^{\mathcal{P}}$ may contain edges $vw$ with $v\in L_i$ and $w\in L_{i+j}$ for any $j\in\{0,\ldots,k\}$.
 To resolve this, we use a new layering $\mathcal{L}':=\langle L_0',\ldots,L_h'\rangle$ in which $L_i'=\bigcup_{j=ki}^{ki+k-1} L_i$.
@@
 An embedded graph $G$ is \defin{$k$-plane} if each edge of $G$ takes part in at most $k$ crossings.
-\note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}\note{PM}{No, under the current definition, $r$ edges crossing at a single point $p$ generate $\binom{r}{2}$ crossings, with each edge taking part in $r-1$ crossings, just like they would if we perturbed the edges so that each pair crosses at a distinct point.
-I guess we could mention this.
-(See the first paragraph in the proof of Theorem~2 below.)}  A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
+\note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}  A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
 Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
@@
 In this way, each edge $vw\in E(G)$ corresponds naturally to a path $P_{vw}$ of length at most $k+1$ in $G_0$.
-Let $\SS := \{P_{vw}: vw\in E(G)\}$.
-Observe that $\SS$ is a $(k+1,2)$-shortcut system for $G_0$ and that $G_0^{\SS}\supseteq G$.
-Specifically, $G_0^{\SS}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.
+Let $\PP := \{P_{vw}: vw\in E(G)\}$.
+Observe that $\PP$ is a $(k+1,2)$-shortcut system for $G_0$ and that $G_0^{\PP}\supseteq G$.
+Specifically, $G_0^{\PP}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.
 
 	Since $G_0$ is planar,  \cref{PlanarProduct}(b) and \cref{PartitionProduct} implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
-Applying \cref{ShortcutPartition} to $G_0$ and $\SS$ immediately implies that $G$ (an arbitrary $k$-planar graph) has an $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
+Applying \cref{ShortcutPartition} to $G_0$ and $\mathcal{P}$ immediately implies that $G$ (an arbitrary $k$-planar graph) has an $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
 
@@
 In this setting, the proof of \cref{general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $L_i\cap Y_x$.
-More precisely, each path $P\in\SS$ containing an internal (dummy) vertex $w\in Y_x\cap (L_{i-j}\cup L_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap L_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap L_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap L_j$ for $j > k+1$.
+More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (L_{i-j}\cup L_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap L_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap L_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap L_j$ for $j > k+1$.
 	Redoing the calculation at the end of the proof of \cref{general-width} then yields
@@
 
-\note{PM}{I agree that we should fix it and keep it.
-Another reason: For vertex ranking, the bound is $O(\log n/\log^{(t)} n)$ where $t:=\tw(H)$, so this is an example where $\tw(H)$ affects more than just the constant.}
-
 Let $G$ be an edge-maximal 1-plane multigraph with no two parallel edges on the boundary of a single face.
@@
 Right?
+This should be stated as a lemma.
 Does this viewpoint help?
@@
 		\end{compactenum}
-	\note{DW}{Should the vertices already grabbed by $P_1,P_2,P_3$ be mentioned in the setup?
-The new paths $Q_1,Q_2,Q_3$ need to avoid the vertices already grabbed by $P_1,P_2,P_3$.
+	\note{DW}{Should the vertices already grabbed by $P_1,P_2,P_3$ be mentioned in the setup (since the new paths $Q_1,Q_2,Q_3$ need to avoid the vertices already grabbed by $P_1,P_2,P_3$)?.
 Basically, I am concerned about what happens when distinct vertical paths in $T$ want to grab the same vertex.}
@@
 \label{gAddDummy}
-Every $(g,k)$-planar graph $G$ is a subgraph of $G_0^\SS$ for some graph $G_0$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\SS$ for $G_0$.
+Every $(g,k)$-planar graph $G$ is a subgraph of $G_0^\PP$ for some graph $G_0$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\PP$ for $G_0$.
 Moreover, $V(G) \subseteq V(G_0)$ and for every edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most $k+1$, such that every internal vertex in $P$ has degree at most $4$ in $G_0$.
@@
 \label{MapShortcut}
-Every $(g,d)$-map graph $G$ is a subgraph of $G_1^\SS$ for some graph $G_1$ with Euler genus at most $g$ and some $(2,\tfrac12 d(d-3) )$-shortcut system $\SS$ for $G_1$.
+Every $(g,d)$-map graph $G$ is a subgraph of $G_1^\PP$ for some graph $G_1$ with Euler genus at most $g$ and some $(2,\tfrac12 d(d-3) )$-shortcut system $\PP$ for $G_1$.
 \end{lem}
@@
 In the latter case, let $P_{vw}$ be the path $(v,x,w)$.
-Let $\SS$ be the set of all such paths $P_{vw}$.
-Each vertex $x\in V(G_0)$ is the middle vertex on at most $\tfrac12 d(d-3)$  paths in $\SS$.
-Thus $\SS$ is a $(2,\tfrac12 d(d-3))$-shortcut system for $G_1$, and by construction, $G \subseteq G_1^\SS$.
+Let $\PP$ be the set of all such paths $P_{vw}$.
+Each vertex $x\in V(G_0)$ is the middle vertex on at most $\tfrac12 d(d-3)$  paths in $\PP$.
+Thus $\PP$ is a $(2,\tfrac12 d(d-3))$-shortcut system for $G_1$, and by construction, $G \subseteq G_1^\PP$.
 \end{proof}
@@
 \label{StringShortcut}
-Every $(g,\delta)$-string graph $G$ is a subgraph of $G_0^\SS$ for some graph $G_0$ with Euler genus at most $g$ and some $(\delta+1,\delta+1 )$-shortcut system $\SS$ for $G_0$.
-%Moreover, every internal vertex on the paths in $\SS$ has degree 4 in $G_0$.
+Every $(g,\delta)$-string graph $G$ is a subgraph of $G_0^\PP$ for some graph $G_0$ with Euler genus at most $g$ and some $(\delta+1,\delta+1 )$-shortcut system $\PP$ for $G_0$.
+%Moreover, every internal vertex on the paths in $\PP$ has degree 4 in $G_0$.
 \end{lem}
@@
 For each edge $vw$ of $G$, there is a path $P_{vw}$ of length at most $2\ceil{\frac{\delta}{2}}\leq \delta+1$ in $G_0$ between $v$ and $w$.
-Let $\SS$ be the set of all such paths $P_{vw}$.
-Consider a vertex $x$ in $G_0$ that is an internal vertex on some path in $\SS$.
+Let $\PP$ be the set of all such paths $P_{vw}$.
+Consider a vertex $x$ in $G_0$ that is an internal vertex on some path in $\PP$.
 Then $x$ is at the intersection of $C_v$ and $C_w$ for some edge $vw\in E(G)$.
-If some path $P\in \SS$ passes through $x$, then $P=P_{vu}$ for some edge $vu$ incident to $v$, or $P=P_{wu}$ for some edge $wu$ incident to $w$.
-At most $\ceil{\frac{\delta}{2}}$ paths in $\SS$ corresponding to edges incident to $v$ pass through $x$, and similarly for edges incident to $w$.
-Thus at most $2\ceil{\frac{\delta}{2}}\leq\delta+1$ paths in $\SS$ use $x$ as an internal vertex.
-Thus $\SS$ is a $(\delta+1,\delta+1)$-shortcut system for $G_0$, and by construction, $G \subseteq G_0^\SS$.
+If some path $P\in \PP$ passes through $x$, then $P=P_{vu}$ for some edge $vu$ incident to $v$, or $P=P_{wu}$ for some edge $wu$ incident to $w$.
+At most $\ceil{\frac{\delta}{2}}$ paths in $\PP$ corresponding to edges incident to $v$ pass through $x$, and similarly for edges incident to $w$.
+Thus at most $2\ceil{\frac{\delta}{2}}\leq\delta+1$ paths in $\PP$ use $x$ as an internal vertex.
+Thus $\PP$ is a $(\delta+1,\delta+1)$-shortcut system for $G_0$, and by construction, $G \subseteq G_0^\PP$.
 \end{proof}
@@
 \label{PowerShortcut}
-If a graph $G$ has maximum degree $\Delta$, then $G^k = G^\SS$ for some $(k,2k \Delta^{k})$-shortcut system $\SS$.
+If a graph $G$ has maximum degree $\Delta$, then $G^k = G^\PP$ for some $(k,2k \Delta^{k})$-shortcut system $\PP$.
 \end{lem}
@@
 $\dist_G(x,y)$  in $G$.
-Let $\SS:=\{P_{xy}: \dist_G(x,y)\in\{1,\dots,k\} \}$.
+Let $\PP:=\{P_{xy}: \dist_G(x,y)\in\{1,\dots,k\} \}$.
 Say $P_{xy}$ uses some vertex $v$ as an internal vertex.
@@
 The number of vertices at distance $i$ from $v$ is at most $\Delta^i$.
-Thus the number of paths in $\SS$ that use $v$ as an internal vertex is at most
+Thus the number of paths in $\PP$ that use $v$ as an internal vertex is at most
 \[\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i\Delta^j
@@
 < 2k \Delta^k\enspace.\]
-Hence $\SS$ is a $(k, 2k \Delta^k)$-shortcut system.
+Hence $\PP$ is a $(k, 2k \Delta^k)$-shortcut system.
 \end{proof}

### 2022-03-07 k-planar.tex

--- 2022-03-07 k-planar.tex
@@
 We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.
-We say that a graph $G$ is \emph{contained in} a graph $X$ if $G$ is isomorphic to a subgraph of $X$.
+We say that a graph $G$ is \defin{contained in} a graph $X$ if $G$ is isomorphic to a subgraph of $X$.
 Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (improved by \citet{UWY}).
@@
 	\label{PlanarProduct}
-	Every planar graph is a subgraph of:
+	Every planar graph is contained in: 
+	%a subgraph of:
 	\begin{compactenum}[(a)]
@@
 
-% \note{DW}{I suggest we replace $8$ by $6$ here, and replace ``The starting point for this work is the following `product structure theorem' for planar graphs by \citet{DJMMUW20} (improved by \citet{UWY})''.
-Okay?}
+ \note{DW}{replace $8$ by $6$ and update elsewhere}
 
@@
 
+%Let $G$ be an edge-maximal 1-plane multigraph with no two parallel edges on the boundary of a single face.
+Here, edge-maximal should be taken to mean that, if any two vertices $v$ and $w$ appear on a common face\footnote{The \defin{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
+We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.
+\note{DW}{Each face of $G$ is one of two types, either bounded by three edges (none of which are crossed), or bounded by one edge that is not crossed plus portions of two crossing edges.
+Right?
+If this is true, we should say it here, which would help to clear up the confusion below.
+Add a figure.}
+%We assume that no two edges incident to a common vertex cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}
+%
+%\note{DW}{I suggest we say that it follows from Euler's formula that $G$ is finite.
+(To see this, say $G$ has $n$ vertices, $m$ non-crossed edges, and $k$ crossings.
+Let $G'$ be the plane graph obtained from $G$ by adding a dummy vertex at each crossing.
+So $G'$ is a plane multigraph in which every face is bounded by a 3-cycle.
+It follows from Euler's formula that $m+4k = |E(G')| = 3( |V(G')|-2) = 3 ( n+k-2)$, implying	$m+k = 3 (n-2)$ and $|E(G)| = m+2k \leq 6(n-2)$, which can also be shown by deleting one edge from each crossing point.
+) }
+%
+%\note{DW}{It seems to be that $G$ can be obtained from a plane multigraph with faces of size 3 or 4 such that no edge is in two 3-faces, by adding a pair of crossing edges across each 4-face.
+I think this characterises edge-maximal 1-planar multigraphs with no 2-faces.
+Right?
+This should be stated as a lemma.
+Does this viewpoint help?
+Is it known?}
+%
+%\note{DW}{Say $G$ is obtained from a plane multigraph by making each face a clique. Is $G\subseteq H \boxtimes P \boxtimes K_{f(\omega(G)}$? for some treewidth 3 graph $H$.}
+%
+%A \defin{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.
+It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
+\note{DW}{We need to define a kite to be the subgraph with vertex-set $\{v,w,x,y\}$ and edge-set $\{vx,xy,vx,vy,wy,wx\}$ where $vx,vy,wy,wx$ are the edges on the `outside' of the kite-faces.
+Then it really is a simple $K_4$.
+I think this is what we intended.
+We should add a figure showing a kite and kite-faces.
+}
+%The edges $vw$ and $xy$ are called \defin{spars} of $K$.
+The cycle $vxwy$ is called the \defin{sail} of $K$.
+It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$.
+%
+%\referee{2}{Page 13. Perhaps I am misunderstanding something, but I do not see why
+%'none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of
+%$G$.'  For example, see the attached PDF for a picture where $vw$, $xw$, $wy$,
+%and $yv$ are all crossed by other edges of $G$.\\
+%\includegraphics[width=60mm]{kite}}
+%\note{DW}{We need to explain more about how edge-maximality fixes this problem.
+In this case we can draw other $xv$, $vy$, $yw$ $wx$ edges close to the central crossing edges.} Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
+%
+%\referee{2}{Page 13. The attached PDF also shows that the definition of 'kite
+%face' may not be well-defined.  For example, I do not see why there
+%cannot be some edges of $G$ 'inside' a kite face.  Even if this is not
+%the case, I think it is more precise to say that a kite face has 'two
+%and a half edges' and two vertices of $G$ on its boundary' rather than
+%'three edges and two vertices of $G$ on its boundary'.}
+
+\note{DW}{This should be fixed if we define kite more precisely. A kite-face is bounded by one edge and two half-edges right?}
+
 \note{PM}{I agree that we should fix it and keep it.
@@
     \begin{tabular}{c@{\hspace{1cm}}c}
-      \includegraphics{figs/one_planar_example-1} &
-      \includegraphics{figs/one_planar_example-2} \\
+%      \includegraphics{figs/one_planar_example-1} &
+%      \includegraphics{figs/one_planar_example-2} \\
       (a) & (b) \\[1em]
-      \includegraphics{figs/one_planar_example-3} &
-      \includegraphics{figs/one_planar_example-4} \\
+%      \includegraphics{figs/one_planar_example-3} &
+%      \includegraphics{figs/one_planar_example-4} \\
       (c) & (d) \\[1em]

### 2022-03-07 k-planar.tex
Pfffft

--- 2022-03-07 k-planar.tex
@@
 Basically, I am concerned about what happens when distinct vertical paths in $T$ want to grab the same vertex.}
+  \note{PM}{No, at this point we're only promising an $H$-partition of $N$, the part of $G$ contained in the uncrossed cycle $F$.}
 		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
@@
 		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$
-		\note{DW}{Should this be $P_i\subseteq S_{x_i}$? (since the path grabs other vertices into its part)}; and
+		\note{DW}{Should this be $P_i\subseteq S_{x_i}$? (since the path grabs other vertices into its part)}\note{PM}{No, for the same reason as above.}; and
 		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
@@
 \begin{proof}
-	This proof is very similar to the proof of Lemma~14 by \citet{DJMMUW20}.
-Rather than duplicate every detail of that proof here, we focus on the differences and refer the reader to the original proof for the remaining details.
-
-	The proof is by induction on the number of vertices of $N$.
-	First note that $N'$ is a near-triangulation.
-If $k=3$, set $R_i := P_i$ for each $i\in\{1,2,3\}$.
+	This proof is very similar to the proof of Lemma~14 by \citet{DJMMUW20}, but is sufficiently complicated by the $1$-planar setting to warrant a full proof.
+   % 
+   % 
+   % 
+   % 
+   % Rather than duplicate every detail of that proof here, we focus on the differences and refer the reader to the original proof for the remaining details.
+  In particular, the reader should keep in mind that $G$ and $G'$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
+
+  The proof is by induction on the number of inner faces of $N$.
+If $N$ has one inner face, then we can simply take $\mathcal{P}:=\{P_1,\ldots,P_k\}$ and verify the preconditions of the lemma (the setup) ensure that $\mathcal{P}$ satisfies the requirements of the lemma.
+Therefore, we may assume that $N$ has at least two inner faces.
+
+	First note that every inner face of $N'$ is bounded by three edges of $G'$. 
+  Assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,2,3\}$ as follows: 
+  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.
+(This implies that $w=v$ for $v\in V(F)$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,2,3\}$ and we define $\alpha(v):=i$.
+
+  If $k<3$ then let $\tau:=v_1v_2v_3$ be any inner face of $N'$.
+Otherwise, $k=3$, in which case we apply Sperner's Lemma implies to obtain an inner face $\tau:=v_1v_2v_3$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+  For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to $V(F)$.
+   % 
+   % 
+   % If $k=3$, set $R_i := P_i$ for each $i\in\{1,2,3\}$.
 Otherwise, as in \citep{DJMMUW20}, split $P_1,\ldots,P_k$ to partition $V(F)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a non-empty path and each $R_i$ contains vertices from exactly one of $P_1,\ldots,P_k$.
-
-	Next, as in \citep{DJMMUW20}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
-
+  % 
+	% Next, as in \citep{DJMMUW20}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
 	Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.

### 2022-03-07 k-planar.tex
updates
Partial commit

@@
 
-  The proof is by induction on the number of inner faces of $N$.
-If $N$ has one inner face, then we can simply take $\mathcal{P}:=\{P_1,\ldots,P_k\}$ and verify the preconditions of the lemma (the setup) ensure that $\mathcal{P}$ satisfies the requirements of the lemma.
-Therefore, we may assume that $N$ has at least two inner faces.
+  The proof is by induction on the number of inner vertices of $N$, i.e., on $|V(N)\setminus V(F)|$.
+If $N$ has no inner vertices, then we can simply take $\mathcal{P}:=\{P_1,\ldots,P_k\}$ and verify the preconditions of the lemma (the setup) ensure that $\mathcal{P}$ satisfies the requirements of the lemma.
+Therefore, we may assume that $N$ has at least one inner vertex.
 

### 2022-03-07 k-planar.tex

--- 2022-03-07 k-planar.tex
@@
 }
+
 \note{PM}{I agree with this suggestion.  I've only added the other two references here, and haven't changed anything elsewhere yet.}
@@
 		\end{compactenum}
-	\note{DW}{Should the vertices already grabbed by $P_1,P_2,P_3$ be mentioned in the setup (since the new paths $Q_1,Q_2,Q_3$ need to avoid the vertices already grabbed by $P_1,P_2,P_3$)?.
+	\note{DW}{Should the vertices already grabbed by $P_1,P_2,P_3$ be mentioned in the setup?
+The new paths $Q_1,Q_2,Q_3$ need to avoid the vertices already grabbed by $P_1,P_2,P_3$.
 Basically, I am concerned about what happens when distinct vertical paths in $T$ want to grab the same vertex.}
+  \note{PM}{No, at this point we're only promising an $H$-partition of $N$, the part of $G$ contained in the uncrossed cycle $F$.}
 		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
@@
 \begin{proof}
-	This proof is very similar to the proof of Lemma~14 by \citet{DJMMUW20}.
-Rather than duplicate every detail of that proof here, we focus on the differences and refer the reader to the original proof for the remaining details.
-
-	The proof is by induction on the number of vertices of $N$.
-	First note that $N'$ is a near-triangulation.
-If $k=3$, set $R_i := P_i$ for each $i\in\{1,2,3\}$.
+	This proof is very similar to the proof of Lemma~14 by \citet{DJMMUW20}, but is sufficiently complicated by the $1$-planar setting to warrant a full proof.
+   % 
+   % 
+   % 
+   % 
+   % Rather than duplicate every detail of that proof here, we focus on the differences and refer the reader to the original proof for the remaining details.
+  In particular, the reader should keep in mind that $G$ and $G'$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
+
+  The proof is by induction on the number of inner vertices of $N$, i.e., on $|V(N)\setminus V(F)|$.
+If $N$ has no inner vertices, then we can simply take $\mathcal{P}:=\{P_1,\ldots,P_k\}$ and verify the preconditions of the lemma (the setup) ensure that $\mathcal{P}$ satisfies the requirements of the lemma.
+Therefore, we may assume that $N$ has at least one inner vertex.
+
+	First note that every inner face of $N'$ is bounded by three edges of $G'$. 
+  Assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,2,3\}$ as follows: 
+  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.
+(This implies that $w=v$ for $v\in V(F)$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,2,3\}$ and we define $\alpha(v):=i$.
+
+  If $k<3$ then let $\tau:=v_1v_2v_3$ be any inner face of $N'$.
+Otherwise, $k=3$, in which case we apply Sperner's Lemma implies to obtain an inner face $\tau:=v_1v_2v_3$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+  For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to $V(F)$.
+   % 
+   % 
+   % If $k=3$, set $R_i := P_i$ for each $i\in\{1,2,3\}$.
 Otherwise, as in \citep{DJMMUW20}, split $P_1,\ldots,P_k$ to partition $V(F)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a non-empty path and each $R_i$ contains vertices from exactly one of $P_1,\ldots,P_k$.
-
-	Next, as in \citep{DJMMUW20}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
-
+  % 
+	% Next, as in \citep{DJMMUW20}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
 	Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.

### 2022-03-07 k-planar.tex
updates
More PDFsssssss

@@
   The proof is by induction on the number of inner vertices of $N$, i.e., on $|V(N)\setminus V(F)|$.
-If $N$ has no inner vertices, then we can simply take $\mathcal{P}:=\{P_1,\ldots,P_k\}$ and verify the preconditions of the lemma (the setup) ensure that $\mathcal{P}$ satisfies the requirements of the lemma.
+If $N$ has no inner vertices, then we can simply take $\mathcal{P}:=\{P_1,\ldots,P_k\}$ and verify the preconditions of the lemma (the `setup') ensure that $\mathcal{P}$ satisfies the requirements of the lemma.
 Therefore, we may assume that $N$ has at least one inner vertex.
 
-	First note that every inner face of $N'$ is bounded by three edges of $G'$. 
   Assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,2,3\}$ as follows: 
@@
   If $k<3$ then let $\tau:=v_1v_2v_3$ be any inner face of $N'$.
-Otherwise, $k=3$, in which case we apply Sperner's Lemma implies to obtain an inner face $\tau:=v_1v_2v_3$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+Otherwise, $k=3$, in which case we apply Sperner's Lemma to obtain an inner face $\tau:=v_1v_2v_3$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
   For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to $V(F)$.
@@
 Let $\overline{Y}^+$ denote the subgraph of $N$ consisting of the vertices and edges of $\overline{Y}$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $\overline{Y}$.
+Let $Y:= V(\overline{Y}-F)$ and let $Y^+:=V(\overline{Y}^+-F)$.
 
@@
 	Let $M$ and $M^+$ denote the subgraph of $G$ containing the edges and vertices of $\overline{Y}$, respectively $\overline{Y}^+$, and the edges and vertices of $F$.
-The graph $M^+$ has some number of bounded faces, all contained in the interior of $F$.
-Some of the bounded faces of $M^+$ are kite faces.
-Let $F_1,\ldots,F_m$ be the non-kite bounded faces of $M^+$.
+By the choice of $\tau$, the vertices on the boundary of each inner face of $M$ are contained in $Y\cup P_a\cup P_b$ for some $a,b\in\{1,\ldots,k\}$.
+In particular if $k=3$ then, for each inner face $f$ of $M$ there exists some $c\in\{1,2,3\}$ such that no vertex of $P_c$ is on the boundary of $f$.
+
+  The graph $M^+$ has some number of inner faces, all contained in the interior of $F$. Some of the inner faces of $M^+$ are Type~(ii) faces. Let $F_1,\ldots,F_m$ be the Type~(i) inner faces of $M^+$.
 
@@
 Therefore $vw\in E(\overline{Y}^+)$ so $xy\in E(Y^+)$.
-But then the only faces of $M^+$ incident to $vw$ are kite faces.
-In particular $vw$ cannot be incident to the non-kite face $F_i$.
-
-	Observe that each of the faces $F_1,\ldots,F_m$ is contained in a single internal face of $M$.
+But then the only faces of $M^+$ incident to $vw$ are Type~(ii) faces.
+In particular $vw$ cannot be incident to the Type~(i) face $F_i$.
+
+	Observe that each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.
 Let $Y^+ := \overline{Y}^+-F$.

### 2022-03-07 k-planar.tex
updates
Hmmmm

@@
 	Let $M$ and $M^+$ denote the subgraph of $G$ containing the edges and vertices of $\overline{Y}$, respectively $\overline{Y}^+$, and the edges and vertices of $F$.
-By the choice of $\tau$, the vertices on the boundary of each inner face of $M$ are contained in $Y\cup P_a\cup P_b$ for some $a,b\in\{1,\ldots,k\}$.
-In particular if $k=3$ then, for each inner face $f$ of $M$ there exists some $c\in\{1,2,3\}$ such that no vertex of $P_c$ is on the boundary of $f$.
+(See \cref{boring_figure}.)  By the choice of $\tau$, the vertices on the boundary of each inner face of $M$ are contained in $Y\cup P_a\cup P_b$ for some $a,b\in\{1,\ldots,k\}$.
+In particular if $k=3$ then, for each inner face $F_i$ of $M$ there exists some $c\in\{1,2,3\}$ such that no vertex of $P_c$ is on the boundary of $F_i$.
 

### 2022-03-07 k-planar.tex
More figures and updates
Pfffft

@@
 
+  \begin{figure}
+    \begin{center}
+      \begin{tabular}{c@{\hspace{1cm}}c}
+        \includegraphics{figs/mmplus-1} & \includegraphics{figs/mmplus-2} \\
+        $M$ & $M^+$
+      \end{tabular}
+    \end{center}
+    \caption{The graphs $M$ and $M^+$.}
+    \label{boring_figure}
+  \end{figure}
+
   The graph $M^+$ has some number of inner faces, all contained in the interior of $F$. Some of the inner faces of $M^+$ are Type~(ii) faces. Let $F_1,\ldots,F_m$ be the Type~(i) inner faces of $M^+$.
@@
 	Observe that each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.
-Let $Y^+ := \overline{Y}^+-F$.
-Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.
-
-	Finally, the subgraph $N_i$ of $G$ consisting of the edges and vertices of $G$ contained in $F_i$ or its interior does not contain one of the three vertices of $\tau$.
-Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(C_i)$ to obtain the desired $H$-partition and tree decomposition of $N_i$.
+Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.\note{PM}{This latter requirement requires that $M^+$ be $2$-connected, we should mention this.}
+  Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(F_i)$ to obtain the desired $H$-partition $\mathcal{P}_i$ and tree decomposition of subgraph $N_i$ of $G$ contained in $F_i$.
 
 	The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
-	Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
-The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{DJMMUW20}, except that now each node $x$ has a child for each face $F_i$ of $M^+_x$ that contains a vertex of $G$ in its interior.
+  Elements in each of the $H$-partitions of $N_1,\ldots,N_m$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
+The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{DJMMUW20}.
+The root node contains the nodes of $H$ corresponding to $P_1,\ldots,P_k$.
+The root has one child that contains the at most $4$ nodes of $H$ corresponding to $P_1,\ldots,P_k$ and $Y^+$.
+This root has up to three children, each of which contains $Y^+$ and at most two of $P_1,\ldots,P_k$.
+For each $i\in\{1,\ldots,m\}$, the tree decomposition of $N_i/\mathcal{P}_i$ is then attached to one of these three children.
 

### 2022-03-08 response.tex
updates

@@
 
-\newcommand{\done}{\textcolor{blue}{done}}
-\newcommand{\tba}{\textcolor{red}{to be addresssed}}
+\newcommand{\done}{\textcolor{blue}{Done.}}
+\newcommand{\tba}{\textcolor{red}{To be addresssed.}}
 
@@
 
-\done
+\tba
+
+$P$ is used twice in the statement of Theorem 3.  It is probably
+safer (and necessary?) to use a different variable for the second
+occurrence.
+
+\textcolor{blue}{We now use $\mathcal{S}$ for shortcut systems.}
 
@@
 
-\tba
+\textcolor{blue}{The referee is right here: two edges may cross more than once.
+We do not want to `uncross' since this does not maintain the bound on the number of crossings per edge.
+We have revised the proof avoiding the claim that every pair of edges cross at most once.
+The lemma itself has not changed.
+}
 

### 2022-03-08 k-planar.tex
updates

@@
 
-\author{Vida Dujmovi\'c%
-        \thanks{School of Computer Science and Electrical Engineering,
-                University of Ottawa, Ottawa, Canada (\texttt{vida.dujmovic@uottawa.ca}).
+\author{%
+Vida Dujmovi\'c%
+\thanks{School of Computer Science and Electrical Engineering, University of Ottawa, Ottawa, Canada (\texttt{vida.dujmovic@uottawa.ca}).
 Research supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\,
 Pat Morin%
-        \thanks{School of Computer Science, Carleton University, Ottawa, Canada (\texttt{morin@scs.carleton.ca}).
-Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\, and
-        David R. Wood\thanks{School of Mathematics, Monash University, Melbourne, Australia (\texttt{david.wood@monash.edu}). Research supported by the Australian Research Council.}
+\thanks{School of Computer Science, Carleton University, Ottawa, Canada (\texttt{morin@scs.carleton.ca}). Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\, and
+David R. Wood%
+\thanks{School of Mathematics, Monash University, Melbourne, Australia (\texttt{david.wood@monash.edu}). Research supported by the Australian Research Council.}
 }
@@
 
+\note{DW}{I suggest we use consecutive numbering of theorems, lemmas, etc. Theorem 1, Lemma 2, Observation 3, etc.}
+
+\note{DW}{I suggest we delete ``We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$. '' This is absolutely standard.}
+
 The starting point for this work is the following `product structure theorem' for planar graphs\footnote{In this paper, all graphs are finite and undirected.
-Unless specifically mentioned otherwise, all graphs are also simple.
+Unless mentioned otherwise, all graphs are also simple.
 For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$), let $G[S]$  denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.
 We use $G-S$ as a shorthand for $G[V(G)\setminus S]$.
-We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$.
-We say that a graph $G$ is \defin{contained in} a graph $X$ if $G$ is isomorphic to a subgraph of $X$.
 Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (improved by \citet{UWY}).
+A graph $G$ is \defin{contained in} a graph $X$ if $G$ is isomorphic to a subgraph of $X$.
 
@@
 Every planar graph is contained in: 
-	%a subgraph of:
 \begin{compactenum}[(a)]
-		\item $H\boxtimes P$ for some graph $H$ of treewidth at most $8$ and for some path $P$,
+	\item $H\boxtimes P$ for some graph $H$ of treewidth at most $6$ and for some path $P$,
 	\item $H\boxtimes P \boxtimes K_3$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
@@
 
-% \note{DW}{In 	\cref{PlanarProduct} and elsewhere, we really should write, ``Every planar graph is isomorphic to a subgraph of''.
-One alternative is to define that $H$ is \defin{contained} in $G$ if $H$ is isomorphic to a subgraph of $G$.
-I suggest the latter.}
-
- \note{DW}{replace $8$ by $6$ and update elsewhere}
-
-Here $\boxtimes$ is the strong product,\!\footnote{The \emph{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
+Here $\boxtimes$ is the strong product,\!\footnote{The \defin{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
 	$v=w$ and $xy\in E(B)$, or
@@
 \item \citet{DFMS21} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs.
-\item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results starting with the work of \citet{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}).
+\item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results starting with the work of \citet{kannan.naor.ea:implicit}).
 \citet{DEJGMM21} have since used it to find asymptotically optimal adjacency labellings of planar graphs.
@@
 
-% \note{DW}{Delete `from 1992' and  `from 2007'? replace ``going back to 1988~\citep{kannan.naor.ea:implicit-stoc,kannan.naor.ea:implicit}'' by
-	% ``starting with the work of \citet{kannan.naor.ea:implicit}''}
-% \note{PM}{Was this a referee request, or just because we don't need to sell it anymore?  I'm fine with the change either way, but didn't implement it.}
-
-% \note{DW}{Add \citep{EJM} to the story.}
+\note{DW}{Mention other application such as vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, odd colourings~\citep{DMO}?
+We write a lot about adjacency labellings and universal induced subgraphs.
+Do we need so much detail here?
+We also mention this work in \cref{WrappingUp}.}
 
@@
 \label{kPlanarProduct}
-Every $k$-planar graph is a subgraph of $H\boxtimes P\boxtimes K_{18k^2+48k+30}$, for some graph $H$ of treewidth $\binom{k+4}{3}-1$ and for some path $P$.
+Every $k$-planar graph is contained in $H\boxtimes P\boxtimes K_{18k^2+48k+30}$, for some graph $H$ of treewidth $\binom{k+4}{3}-1$ and for some path $P$.
 \end{thm}
@@
 Justification: They're both short and they describe the exact bounds for bounded genus graphs, which you don't get from the one-paragraph description in Section~1.0.}
+
+\note{DW}{I suggest we write, ``This theorem has applications in diverse areas, which we explore in \cref{Applications}....''}
 
@@
 \label{AddDummy}
-Every $k$-planar graph is a subgraph of $G^\SS$ for some planar graph $G$ and some $(k+1,2)$-shortcut system $\SS$ for $G$.
+Every $k$-planar graph is contained in $G^\SS$ for some planar graph $G$ and some $(k+1,2)$-shortcut system $\SS$ for $G$.
 \end{obs}
@@
 \label{ShortcutProduct}
-Let $G$ be a subgraph of $H\boxtimes P \boxtimes K_\ell$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
-Let $\SS$ be a $(k,d)$-shortcut system for $G$. Then $G^\SS$ is a subgraph of $J\boxtimes P\boxtimes K_{d\ell(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
+Let $G$ be a graph contained in $H\boxtimes P \boxtimes K_\ell$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
+Let $\SS$ be a $(k,d)$-shortcut system for $G$.
+Then $G^\SS$ is contained in $J\boxtimes P\boxtimes K_{d\ell(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
 \end{thm}
@@
 \label{kPowerBasic}
-For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is a subgraph of $H\boxtimes P\boxtimes K_{6k^2(k^2+3)\Delta^{k}}$ for some graph $H$ of treewidth at most $\binom{k+3}{3}-1$ and some path $P$.
+For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is contained in $H\boxtimes P\boxtimes K_{6k^2(k^2+3)\Delta^{k}}$ for some graph $H$ of treewidth at most $\binom{k+3}{3}-1$ and some path $P$.
 \end{thm}
@@
 \label{GenusProduct}
-	Every graph of Euler genus $g$ is a subgraph of:
+Every graph of Euler genus $g$ is contained in:
 \begin{compactenum}[(a)]
@@
 
-\note{DW}{I suggest we replace $2g+8$ by $2g+6$ here and replace 4 by 3, and cite \citep{DHHW,UWY}.
-This will improve the bounds and (more importantly) simplify the presentation of various results (such as \cref{StringPartition,PowerGenus}) which currently distinguish the $g=0$ case.
-}
-
-\note{PM}{I agree with this suggestion.  I've only added the other two references here, and haven't changed anything elsewhere yet.}
-
 \citet{DJMMUW20} generalised \cref{GenusProduct} for apex-minor-free graphs as follows.
@@
 	\label{ApexMinorFree}
-	For every apex graph $X$, there exists $c\in\mathbb{N}$ such that every $X$-minor-free graph is a subgraph of $H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and some path $P$.
+	For every apex graph $X$, there exists $c\in\mathbb{N}$ such that every $X$-minor-free graph is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and some path $P$.
 \end{thm}
@@
 	\label{MinorFreeDegree}
-	For every graph $X$ there exists $c\in\mathbb{N}$ such that for every $\Delta\in\mathbb{N}$, every $X$-minor-free graph $G$ with maximum degree at most $\Delta$ is a subgraph of $H\boxtimes P$ for some graph $H$ with $\tw(H) \leq c\Delta$ and for some path $P$.
+	For every graph $X$ there exists $c\in\mathbb{N}$ such that for every $\Delta\in\mathbb{N}$, every $X$-minor-free graph $G$ with maximum degree at most $\Delta$ is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H) \leq c\Delta$ and for some path $P$.
 \end{thm}
@@
 %Let $\SS$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$.
-Then $G^\SS$ is a subgraph of $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
+Then $G^\SS$ is contained in $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
 %\end{thm}
@@
 A \defin{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle$ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
-For any partition $\mathcal{P}=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\mathcal{P}$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
-To highlight the importance of the quotient graph $H$, we call $\mathcal{P}$ an \defin{$H$-partition} and write this concisely as $\mathcal{P}=\{S_x : x\in V(H)\}$ so that each element of $\mathcal{P}$ is indexed by the vertex it creates in $H$.
-
-For any partition $\mathcal{P}$ of $V(G)$ and any layering $\mathcal{L}$ of $G$ we define the \defin{layered width} of $\mathcal{P}$ with respect to $\mathcal{L}$ as $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\mathcal{P}\}$.
-For any partition $\mathcal{P}$ of $V(G)$, we define the \defin{layered width} of $\mathcal{P}$ as the minimum, over all layerings $\mathcal{L}$ of $G$, of the layered width of $\mathcal{P}$ with respect to $\mathcal{L}$.
+For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
+To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the vertex it creates in $H$.
+
+For any partition $\PP$ of $V(G)$ and any layering $\mathcal{L}$ of $G$ we define the \defin{layered width} of $\PP$ with respect to $\mathcal{L}$ as $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\PP\}$.
+For any partition $\PP$ of $V(G)$, we define the \defin{layered width} of $\PP$ as the minimum, over all layerings $\mathcal{L}$ of $G$, of the layered width of $\PP$ with respect to $\mathcal{L}$.
 
@@
 \label{PartitionProduct}
-For every graph $H$, a graph $G$ has an $H$-partition of layered width at most $\ell$ if and only if $G$ is a subgraph of $H \boxtimes P \boxtimes K_\ell$ for some path $P$.
+For every graph $H$, a graph $G$ has an $H$-partition of layered width at most $\ell$ if and only if $G$ is contained in $H \boxtimes P \boxtimes K_\ell$ for some path $P$.
 \end{lem}
@@
 As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a),
-\citet{DJMMUW20} showed that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$.
+\citet{DJMMUW20} showed that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \cref{UWY}).
 The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.
@@
 \label{ps}
-Every planar triangulation $G$ has an $H$-partition $\mathcal{P}$
-such that $\tw(H)\leq 8$ and $G[P]$ is a shortest path in $G$ for each $P\in\mathcal{P}$.
+Every planar triangulation $G$ has an $H$-partition $\PP$
+such that $\tw(H)\leq 8$ and $G[P]$ is a shortest path in $G$ for each $P\in\PP$.
 \end{thm}
 
-Indeed, the above-mentioned result of \citet{DJMMUW20} is a slight strengthening of \cref{ps}, where for each $P\in\mathcal{P}$ no two vertices of $P$ have the same distance to some fixed root vertex $r$.
+Indeed, the above-mentioned result of \citet{DJMMUW20} is a slight strengthening of \cref{ps}, where for each $P\in\PP$ no two vertices of $P$ have the same distance to some fixed root vertex $r$.
 
@@
 
-\note{DW}{Use $\SS$ for the given shortcut system and $\SS$ for the partition?}
+\note{DW}{Use $\SS$ for the given shortcut system and $\PP$ for the partition?}
 \note{PM}{Yes, I've done this.  Again, I was carefuly, but need to reread.}
@@
 For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
-Observe that $\mathcal{P}:=(S_x : x\in V(T))$ is a partition of $V(G)$.\footnote{The sets in $\mathcal{P}$ are disjoint because $a:V(G)\to V(T)$ is a function.
-The sets in $\mathcal{P}$ cover $V(G)$ since, for each $v\in V(G)$, $v\in S_{a(v)}$.} Let $J:=G^\SS/\mathcal{P}$ denote the resulting quotient graph.
+Observe that $\PP:=(S_x : x\in V(T))$ is a partition of $V(G)$.\footnote{The sets in $\PP$ are disjoint because $a:V(G)\to V(T)$ is a function.
+The sets in $\PP$ cover $V(G)$ since, for each $v\in V(G)$, $v\in S_{a(v)}$.} Let $J:=G^\SS/\PP$ denote the resulting quotient graph.
 We consider $V(J)\subseteq V(T)$, where each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\SS}$.
@@
 
-% \referee{2}{Page 8.  Can you add a brief explanation why $\mathcal{P}$ is a
-% 	partition of $V(G)$?  It is clear that the sets in $\mathcal{P}$ are
+% \referee{2}{Page 8.  Can you add a brief explanation why $\PP$ is a
+% 	partition of $V(G)$?  It is clear that the sets in $\PP$ are
 % 	disjoint, but why do they cover $V(G)$?}
@@
 \begin{compactenum}[(i)]
-  \item $\mathcal{P}$ has small layered width with respect to the layering $\mathcal{L}$ of $G$ and that
+  \item $\PP$ has small layered width with respect to the layering $\mathcal{L}$ of $G$ and that
   \item $J$ has small treewidth.
@@
 
-First we need to understand the relationship between sets $S_x\in\mathcal{P}$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
+First we need to understand the relationship between sets $S_x\in\PP$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
 
@@
 
-Next we complete Step~(i) and show that $\mathcal{P}$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
+Next we complete Step~(i) and show that $\PP$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
 
@@
 
+  \note{DW}{replace Lemma~24 by Lemma~13, right?}
+
   At this point, the proof is complete, but let us summarize.
@@
 I guess we could mention this.
-(See the first paragraph in the proof of Theorem~2 below.)}  A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
+(See the first paragraph in the proof of Theorem~2 below.)} \note{DW}{The downside of this approach is that in the opening section we give an informal definition of $k$-planar, which needs the assumption of having no three edges cross at a point, since we use this assumption in the construction of a $(k+1,2)$-shortcut system for $k$-planar graphs give in the Introduction.
+I think we should not have different definitions in two parts of the paper.} A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
 Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
@@
 \label{1-planar}
-Every 1-planar graph is a subgraph of $H\boxtimes P\boxtimes K_{30}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+Every 1-planar graph is contained in $H\boxtimes P\boxtimes K_{30}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
 \end{thm}
@@
 %'three edges and two vertices of $G$ on its boundary'.}
-
-\note{DW}{This should be fixed if we define kite more precisely. A kite-face is bounded by one edge and two half-edges right?}
+%
+%\note{DW}{This should be fixed if we define kite more precisely. A kite-face is bounded by one edge and two half-edges right?}
 
@@
 %      \includegraphics{figs/one_planar_example-2} \\
-      (a) & (b) \\[1em]
-%      \includegraphics{figs/one_planar_example-3} &
-%      \includegraphics{figs/one_planar_example-4} \\
-      (c) & (d) \\[1em]
+%      (a) & (b) \\[1em]
+      \includegraphics{figs/one_planar_example-3} &
+      \includegraphics{figs/one_planar_example-4} \\
+      (a) & (b) 
    \end{tabular}
   \end{center}
-  \caption{Examples of (a)~a $1$-plane graph that is not edge maximal; (b)~an edge-maximal $1$-plane graph $G$; (c)~the kites in $G$; (d)~a plane multigraph $G'$ obtained by removing one spar from each kite of $G$.}
+%  \caption{Examples of (a)~a $1$-plane graph that is not edge maximal; (b)~an edge-maximal $1$-plane graph $G$; (c)~the kites in $G$; (d)~a plane multigraph $G'$ obtained by removing one spar from each kite of $G$.}
+  \caption{(a)~Kites in an edge-maximal $1$-plane multigraph $G$; (b)~Plane multigraph $G'$ obtained by removing one spar from each kite of $G$.} 
   \label{one_planar_example} 
-  \note{PM}{I suggest we only keep parts (c) and (d) of this figure.}
 \end{figure}
 
-To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \emph{dummy vertex} $p$.
+To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
 Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_r$ be the facial walk around $F$.
@@
 Each of these four faces of is of Type~(ii), so the boundary of $F_1\cup\cdots\cup F_4$ includes a $4$-cycle $C:=vxwy$ in $G$ whose edges are not involved in any crossings.
-We call the subgraph $K:=C\cup\{vw,xy\}$ a \defin{kite} of $G$.
-(See \cref{one_planar_example}(c).)  The edges $vw$ and $xy$ are called \defin{spars} of $K$.
+We call the subgraph $K:=C\cup\{vw,xy\}$ a \defin{kite} of $G$; see \cref{one_planar_example}(a).
+The edges $vw$ and $xy$ are called \defin{spars} of $K$.
 The cycle $vxwy$ is called the \defin{sail} of $K$.
@@
 
-\note{PM}{I've commented out a lot of red and blue here to make this readable. Those notes are still in the \LaTeX\ source.}
 % 
@@
 
-Refer to \cref{one_planar_example}(d).
-By removing one spar from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three edges.
+By removing one spar from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three edges; see \cref{one_planar_example}(b).
 Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
@@
 	\end{compactenum}
-	Then $N$ has an $H$-partition $\mathcal{P}=\{S_x : x\in V(H)\}$ such that:
+	Then $N$ has an $H$-partition $\PP=\{S_x : x\in V(H)\}$ such that:
 	\begin{compactenum}
@@
 \begin{proof}
-	This proof is very similar to the proof of Lemma~14 by \citet{DJMMUW20}, but is sufficiently complicated by the $1$-planar setting to warrant a full proof.
+	This proof is very similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is sufficiently complicated by the $1$-planar setting to warrant a full proof.
    % 
@@
   The proof is by induction on the number of inner vertices of $N$, i.e., on $|V(N)\setminus V(F)|$.
-If $N$ has no inner vertices, then we can simply take $\mathcal{P}:=\{P_1,\ldots,P_k\}$ and verify the preconditions of the lemma (the `setup') ensure that $\mathcal{P}$ satisfies the requirements of the lemma.
+If $N$ has no inner vertices, then we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements of the lemma.
 Therefore, we may assume that $N$ has at least one inner vertex.
@@
 Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.\note{PM}{This latter requirement requires that $M^+$ be $2$-connected, we should mention this.}
-  Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(F_i)$ to obtain the desired $H$-partition $\mathcal{P}_i$ and tree decomposition of subgraph $N_i$ of $G$ contained in $F_i$.
+  Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(F_i)$ to obtain the desired $H$-partition $\PP_i$ and tree decomposition of subgraph $N_i$ of $G$ contained in $F_i$.
 
@@
   Elements in each of the $H$-partitions of $N_1,\ldots,N_m$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
-The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{DJMMUW20}.
+
+  The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~13 in \cite{DJMMUW20}.
 The root node contains the nodes of $H$ corresponding to $P_1,\ldots,P_k$.
@@
 This root has up to three children, each of which contains $Y^+$ and at most two of $P_1,\ldots,P_k$.
-For each $i\in\{1,\ldots,m\}$, the tree decomposition of $N_i/\mathcal{P}_i$ is then attached to one of these three children.
-
-	\note{DW}{Check whether Lemma~14 is the right lemma}
+For each $i\in\{1,\ldots,m\}$, the tree decomposition of $N_i/\PP_i$ is then attached to one of these three children.
 
@@
 	\begin{compactenum}
-		\item $G/\mathcal{P}$ and $G^+/\mathcal{P^+}$ are isomorphic, where $G^+$ is the triangulation obtained by adding dummy vertices at each crossing in $G$ and $\mathcal{P}^+$ is the partition we obtain by adding a dummy vertex $z$ to $\overline{Y}^+$ if $\overline{Y}^+$ contains an edge $vw$ that contains $z$ in its interior.
+		\item $G/\PP$ and $G^+/\mathcal{P^+}$ are isomorphic, where $G^+$ is the triangulation obtained by adding dummy vertices at each crossing in $G$ and $\PP^+$ is the partition we obtain by adding a dummy vertex $z$ to $\overline{Y}^+$ if $\overline{Y}^+$ contains an edge $vw$ that contains $z$ in its interior.
 
@@
 	\end{compactenum}
-	Since $G^+$ is planar, the second point implies that $H=G^+/\mathcal{P}$ is planar.
+	Since $G^+$ is planar, the second point implies that $H=G^+/\PP$ is planar.
 	% \footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
-Repeatedly applying this fact implies that $\Delta/\mathcal{P}$ is planar provided that $\Delta[P]$ is connected for every $P\in\mathcal{P}$.} The first point then ensures that $G/\mathcal{P}$ is also planar.
-	\note{DW}{We need a sentence about why $H$ Has the claimed tree-decomposition.}
+Repeatedly applying this fact implies that $\Delta/\PP$ is planar provided that $\Delta[P]$ is connected for every $P\in\PP$.} The first point then ensures that $G/\PP$ is also planar.
 \end{proof}
@@
 \label{gAddDummy}
-Every $(g,k)$-planar graph $G$ is a subgraph of $G_0^\SS$ for some graph $G_0$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\SS$ for $G_0$.
+Every $(g,k)$-planar graph $G$ is contained in $G_0^\SS$ for some graph $G_0$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\SS$ for $G_0$.
 Moreover, $V(G) \subseteq V(G_0)$ and for every edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most $k+1$, such that every internal vertex in $P$ has degree at most $4$ in $G_0$.
@@
 \label{gkPlanarProduct}
-Every $(g,k)$-planar graph is a subgraph of $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with $\tw(H) \leq \binom{k+5}{4}-1$, where $\ell:=\max\{2g,3\}\cdot(6k^2+16k+10)$.
+Every $(g,k)$-planar graph is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with $\tw(H) \leq \binom{k+4}{3}-1$, where $\ell:=\max\{2g,3\}\cdot(6k^2+16k+10)$.
+\note{DW}{I have changed $\tw(H) \leq \binom{k+5}{4}-1$ to $\tw(H) \leq \binom{k+4}{3}-1$, okay?}
 \end{thm}
@@
 If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
-In particular, if $\mathcal{P}=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree-decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree-decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{DJMMUW20}.
+In particular, if $\PP=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree-decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree-decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{DJMMUW20}.
 Therefore, any property that holds for graphs of bounded layered treewidth also holds for $G$.
@@
 	\]
-	Draw each edge $vw$ of $G$ alongside $P_{vw}$ in $G_0$, so that every pair of edges cross at most once.
-	Every edge of $G$ that crosses $vw$ passes through a vertex on $P_{vw}$ (including $v$ and/or $w$ if they too have degree at most $\Delta$).
+	Fix a $(\Sigma,k)$-plane drawing of $G_0$ in a surface $\Sigma$ of Euler genus at most $g$. Choose $\epsilon>\epsilon'>0$. 
+	For each vertex $v$ of $G_0$, let $B_v$ be the \defin{ball} $\{x\in \mathbb{R}^2:\dist_(x,v)\leq\epsilon\}$ 
+	For each edge $vw$ of $G_0$, let $C_{vw}$ be the \defin{channel} $\{x\in \mathbb{R}^2:\dist(x,vw)\leq\epsilon'\}\setminus(B_v\cup B_w)$. 
+	We may choose $\epsilon>\epsilon'>0$ sufficiently small so that 
+	(i) $B_v\cap B_w=\emptyset$ for distinct $v,w\in V(G_0)$,
+	(ii) $C_{vw}\neq\emptyset$ for $vw\in E(G_0)$, 
+	(iii) $B_v\cap C_{xy}=\emptyset$ for $v\in V(G_0)$ and $xy\in E(G_0)$, and
+	(iv) $C_{vw}\cap C_{xy}=\emptyset$ for distinct $vw,xy\in E(G_0)$.
+	Draw each edge $vw$ of $G$ following the sequence of balls and channels defined by the vertices and edges in $P_{vw}$. 
+	This can be done so that whenever edges $vw$ and $xy$ of $G$ cross, the crossing point is in $B_z$ for some vertex $z$ of $G_0$ that is internal in $P_{vw}$ or in $P_{xy}$, and $vw$ and $xy$ cross at most once in each such $B_z$.
+
+Thus, for each edge $vw$ of $G$, every edge of $G$ that crosses $vw$ passes through a vertex on $P_{vw}$ (including $v$ and/or $w$ if they too have degree at most $\Delta$). 
 	Since $P_{vw}$ has at most $k+1$ vertices, and less than $2k\Delta^{k}$ edges of $G$ pass through each vertex on $P_{vw}$, the edge $vw$ is crossed by less than $2k(k+1)\Delta^{k}$ edges in $G$.
@@
 
-\referee{2}{Page 16.  At the end of Page 16, I do not see why 'every pair of edges
-cross at most once.'  Without further assumptions, it seems as if two
-paths in the path system could intersect several times.  Therefore, by
-drawing 'each edge vw of G alongside $P_vw$ in $G_0$', there may be two
-edges which cross several times.  It seems this can be fixed by
-choosing the path system carefully by 'uncrossing' paths which
-intersect at more than two internal vertices.}
-
-\note{DW}{But we do not want to uncross.
-Our claim that ``every pair of edges cross at most once'' is not true.
-I am convinced the result is still correct.
-Just need to think  of a better way to write it.
-Each crossing is in the vicinity of a vertex that has degree $\leq\Delta$.
-}
+\referee{2}{Page 16.
+At the end of Page 16, I do not see why 'every pair of edges cross at most once.'  Without further assumptions, it seems as if two paths in the path system could intersect several times.
+Therefore, by drawing 'each edge vw of G alongside $P_{vw}$ in $G_0$', there may be two edges which cross several times.
+It seems this can be fixed by choosing the path system carefully by 'uncrossing' paths which intersect at more than two internal vertices.}
+
+\note{DW}{I have revised the proof to address this error. I couldn't think of a way to do this without introducing balls and channels (and without vigorous hand-waving). Is it readable?}
 
@@
 
-\cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^4)}$.
-
+\cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$. \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
 Note that \citet{DJMMUW20} previously proved the bound of
 $O(g^{k+2})$ using \cref{GenusProduct} and an ad-hoc method.
-Our result provides a better bound when $g>2^{k^3}$.
+Our result provides a better bound when $g>2^{k^2}$ \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
 In the case of 1-planar graphs we can improve further.
@@
 \item	For every $1$-planar graph $G$, $\pi(G)\le 30\times 4^4=7680$.
-\item For every $k$-planar graph $G$,
-	$\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
+%\item For every $k$-planar graph $G$,
+%	$\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
 \item	For every $(g,k)$-planar graph $G$,
-	$  \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+5}{4}}.$
+	$  \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}.$
+\note{DW}{I have changed $4^{\binom{k+5}{4}}$ to $4^{\binom{k+4}{3}}$, okay?}
 \end{compactitem}
@@
 \note{DW}{For consistency with \cref{qn}, I suggest we combine \cref{p-centered,p-centered-treewidth}, and just write, ``For every graph $H$ of treewidth at most $t$ and for every path $P$, if $G\subseteq H\boxtimes P \boxtimes K_\ell$ then
-	\[\chi_p(G)\le \ell (p+1)\, \chi_p(H) \leq (p+1)\ell\, \tbinom{p+t}{t}.\]}
+	\[\chi_p(G)\le \ell (p+1)\, \chi_p(H) \leq \ell (p+1) \tbinom{p+t}{t}.\]}
 
 \begin{proof}
-	By \cref{PartitionProduct}, $G$ has an $H$-partition $(\mathcal{L}=\langle L_0,L_1,\ldots\rangle, \mathcal{P}=(B_x:x\in V(H))$ with layered width at most $\ell$.
+	By \cref{PartitionProduct}, $G$ has an $H$-partition $(\mathcal{L}=\langle L_0,L_1,\ldots\rangle, \PP=(B_x:x\in V(H))$ with layered width at most $\ell$.
 Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.
@@
 \item For every $1$-planar graph $G$,\; $ \chi_p(G)\le 5 (p+3)(p+2)(p+1)^2$.
-\item For every $k$-planar graph $G$,\; $\displaystyle \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}$.
+%\item For every $k$-planar graph $G$,\; $\displaystyle \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}$.
 \item For every $(g,k)$-planar graph $G$,\;
-$\displaystyle \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}$.
+$\displaystyle \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{p+\binom{k+4}{3}-1}{\binom{k+4}{3}-1}$.
+\note{DW}{I have changed $\binom{k+5}{4}$ to $\binom{k+4}{3}$, okay?}
  \end{compactitem}
@@
 \label{MapShortcut}
-Every $(g,d)$-map graph $G$ is a subgraph of $G_1^\SS$ for some graph $G_1$ with Euler genus at most $g$ and some $(2,\tfrac12 d(d-3) )$-shortcut system $\SS$ for $G_1$.
+Every $(g,d)$-map graph $G$ is contained in $G_1^\SS$ for some graph $G_1$ with Euler genus at most $g$ and some $(2,\tfrac12 d(d-3) )$-shortcut system $\SS$ for $G_1$.
 \end{lem}
@@
 
-Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,p-centered-treewidth,non-repetitive} imply the following results.
-
-\begin{thm}
-\label{PlaneMapPartition}
-Every $d$-map graph $G$:
-	\begin{compactitem}
-		\item is a subgraph of $H \boxtimes P \boxtimes K_{21d(d-3)}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
-		\item has queue-number $\qn(G) < 32225\, d(d-3)$.
-		\item has $p$-centered chromatic number $\chi_p(G) \leq 21d(d-3) (p+1)  \binom{p+9}{9}$,
-		\item has non-repetitive chromatic number $ \pi(G) \leq 21 \cdot 4^{10} d(d-3)$.
-	\end{compactitem}
-\end{thm}
+Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,p-centered-treewidth,non-repetitive} imply the following result.
+
+%\begin{thm}
+%\label{PlaneMapPartition}
+%\note{DW}{For every integer $d\geq 4$, }
+%Every $d$-map graph $G$:
+%	\begin{compactitem}
+%		\item is contained in $H \boxtimes P \boxtimes K_{21d(d-3)}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
+%		\item has queue-number $\qn(G) < 32225\, d(d-3)$.
+%		\item has $p$-centered chromatic number $\chi_p(G) \leq 21d(d-3) (p+1)  \binom{p+9}{9}$,
+%		\item has non-repetitive chromatic number $ \pi(G) \leq 21 \cdot 4^{10} d(d-3)$.
+%	\end{compactitem}
+%\end{thm}
 
@@
 \begin{compactitem}
-\item is a subgraph of $H \boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 14$,
-\item has queue-number $\qn(G) <  49151\, \ell $.
-\item has $p$-centered chromatic number $\chi_p(G) \leq \ell (p+1)\,  \binom{p+14}{14}$,
-\item has non-repetitive chromatic number $ \pi(G) \leq 4^{15}\,\ell $.
+\item is contained in $H \boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
+\item has queue-number $\qn(G) <  1535\, \ell $.
+\item has $p$-centered chromatic number $\chi_p(G) \leq \ell (p+1)\,  \binom{p+9}{9}$,
+\item has non-repetitive chromatic number $ \pi(G) \leq 4^{10}\,\ell $.
 \end{compactitem}
@@
 
-These results give
-the first constant upper bound on the non-repetitive chromatic number of map graphs,
-the first polynomial bounds on the $p$-centered chromatic number of map graphs, and
-the best known bounds on the queue-number of map graphs.
+\note{DW}{\cref{MapPartition} has been updated.}
+
+These results give the first constant upper bound on the non-repetitive chromatic number of map graphs, the first polynomial bounds on the $p$-centered chromatic number of map graphs, and the best known bounds on the queue-number of map graphs.
 
@@
 \label{StringShortcut}
-Every $(g,\delta)$-string graph $G$ is a subgraph of $G_0^\SS$ for some graph $G_0$ with Euler genus at most $g$ and some $(\delta+1,\delta+1 )$-shortcut system $\SS$ for $G_0$.
+Every $(g,\delta)$-string graph $G$ is contained in $G_0^\SS$ for some graph $G_0$ with Euler genus at most $g$ and some $(\delta+1,\delta+1 )$-shortcut system $\SS$ for $G_0$.
 %Moreover, every internal vertex on the paths in $\SS$ has degree 4 in $G_0$.
@@
 For integers $g\geq 0$ and $\delta\geq 2$, let $\ell:= \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$,
-and $t:= \binom{ \delta+5}{4}-1$ if $g\geq 1$ and $t:= \binom{ \delta+4}{3}-1$ if $g=0$.
+and $t:= \binom{ \delta+4}{3}-1$.
 Then  every $(g,\delta)$-string graph:
 \begin{compactitem}
-	\item is a subgraph of $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$,
+	\item is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$,
 	\item has queue-number $\qn(G^k) \leq 3 \ell \, 2^t - \ceil{\tfrac{3}{2}\ell}$.
@@
 \end{thm}
+
+\note{DW}{\cref{StringPartition} has been updated.}
 
@@
 \label{PowerProduct}
-Let $G$ be a subgraph of $H\boxtimes P\boxtimes K_\ell$ with maximum degree $\Delta$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
-Then for every integer $k\geq 1$, the $k$-th power $G^k$ is a subgraph of $J\boxtimes P\boxtimes K_{2k \ell \Delta^{k}(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
+Let $G$ be contained in $H\boxtimes P\boxtimes K_\ell$ with maximum degree $\Delta$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
+Then for every integer $k\geq 1$, the $k$-th power $G^k$ is contained in $J\boxtimes P\boxtimes K_{2k \ell \Delta^{k}(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
 \end{thm}
@@
 \label{PowerGenus}
-For integers $g\geq 0$ and $k,\Delta\geq 1$, let $\ell:= \max\{2g,3\} (2k^4+6k^2) \Delta^{k}$, and $t:= \binom{k+4}{4}-1$ if $g\geq 1$ and $t:= \binom{k+3}{3}-1$ if $g=0$.
-Then for every graph $G$ of Euler genus $g$ and maximum degree $\Delta$,
+For integers $g\geq 0$ and $k,\Delta\geq 1$, let $\ell:= \max\{2g,3\} (2k^4+6k^2) \Delta^{k}$ and $t:= \binom{k+3}{3}-1$. Then for every graph $G$ of Euler genus $g$ and maximum degree $\Delta$,
 \begin{compactitem}
-\item $G^k$ is a subgraph of $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$
+\item $G^k$ is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$
 \item $G^k$ is $(g, 2k(k+1)\Delta^{k} )$-planar,
@@
 
+\note{DW}{\cref{PowerGenus} has been updated}
+
 This result is the first constant upper bound on the queue-number of bounded powers of graphs with bounded degree and bounded Euler genus.
@@
 \begin{compactitem}
-	\item $G^k$ is a subgraph of $H\boxtimes P$ for some graph $H$ with treewidth $t$ and for some path $P$.
+	\item $G^k$ is contained in $H\boxtimes P$ for some graph $H$ with treewidth $t$ and for some path $P$.
 	\item $G^k$ has queue-number at most $3\cdot 2^t-2$,
@@
 \begin{compactitem}
-\item is a subgraph of $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $t$ and some path $P$,
+\item is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $t$ and some path $P$,
 \item has queue-number at most $2^{O(k^6)}$, and
@@
 \section{Wrapping Up}
+\label{WrappingUp}
 
@@
 We finish with an open problem.
-\cref{kPlanarProduct} shows that every $k$-planar graph is a subgraph of $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $O(k^3)$ where $\ell\leq O(k^2)$.
-Does there exist a function $\ell:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph is a subgraph of $H\boxtimes P \boxtimes K_{\ell(k)}$ for some graph $H$ with treewidth at most $C$?
+\cref{kPlanarProduct} shows that every $k$-planar graph is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $O(k^3)$ where $\ell\leq O(k^2)$.
+Does there exist a function $\ell:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph is contained in $H\boxtimes P \boxtimes K_{\ell(k)}$ for some graph $H$ with treewidth at most $C$?
 Perhaps $C=3$.

### 2022-03-08 k-planar.tex
Committting so that I can merge with David's changes

--- 2022-03-07 k-planar.tex
@@
 
-Let $G$ be an edge-maximal $1$-plane multigraph with no two parallel edges on the boundary of a single face.
-Here, edge-maximal should be taken to mean that, if any two vertices $v$ and $w$ appear on a common face\footnote{The \defin{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
-We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.
-(Refer to \cref{one_planar_example}(a,b).) We assume that no two edges incident to a common vertex of $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}
+In the following, we will assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$, where we treat each edge $vw$ of $G$ as the image of a simple curve $f_{vw}:[0,1]\to\R^2$.
+We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
+
+Let $G$ be a $1$-plane multigraph with no two parallel edges on the boundary of a single face.
+By Euler's Formula, $G$ has $O(|V(G)|)$ vertices.
+We further assume that $G$ is edge-maximal, in the following sense:  if two vertices $v$ and $w$ of $G$ appear on a common face $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.
+See  \cref{one_planar_example}(b) for an example.
+Any $1$-plane multigraph $G$ with at least three vertices can be made edge-maximal by the addition of at most $3|V(G)|-6$ edges since each edge added this way has endpoints in $V(G)$ and does not cross any other edge of $G$.
 
@@
     \begin{tabular}{c@{\hspace{1cm}}c}
-%      \includegraphics{figs/one_planar_example-1} &
-%      \includegraphics{figs/one_planar_example-2} \\
+     \includegraphics{figs/one_planar_example-1} &
+     \includegraphics{figs/one_planar_example-2} \\
       (a) & (b) \\[1em]
-%      \includegraphics{figs/one_planar_example-3} &
-%      \includegraphics{figs/one_planar_example-4} \\
+     \includegraphics{figs/one_planar_example-3} &
+     \includegraphics{figs/one_planar_example-4} \\
       (c) & (d) \\[1em]
@@
 To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \emph{dummy vertex} $p$.
-Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_r$ be the facial walk around $F$.
+Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$.
+In the following paragraphs, subscripts on the vertices in $W$ are implicitly taken modulo $r$.
 By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
 
-If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) is involved in at least two crossings.
+If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
 Therefore each dummy vertex $w_i$ of $W$ is surrounded by two vertices $w_{i-1},w_{i+1}\in V(G)$.
-Since $w_i$ has degree $4$ in $G_0$, $w_{i-1}\neq w_i$.
+Since $w_i$ has degree $4$ in $G_0$, $w_{i-1}\neq w_{i+1}$.
 Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.

### 2022-03-08 k-planar.tex

--- 2022-03-08 k-planar.tex
@@
     \begin{tabular}{c@{\hspace{1cm}}c}
-%      \includegraphics{figs/one_planar_example-1} &
-%      \includegraphics{figs/one_planar_example-2} \\
-%      (a) & (b) \\[1em]
         \includegraphics{figs/one_planar_example-3} &
@@
     \end{center}
-%  \caption{Examples of (a)~a $1$-plane graph that is not edge maximal; (b)~an edge-maximal $1$-plane graph $G$; (c)~the kites in $G$; (d)~a plane multigraph $G'$ obtained by removing one spar from each kite of $G$.}
     \caption{(a)~Kites in an edge-maximal $1$-plane multigraph $G$; (b)~Plane multigraph $G'$ obtained by removing one spar from each kite of $G$.}
@@
 
-To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
-Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_r$ be the facial walk around $F$.
+To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \emph{dummy vertex} $p$.
+Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$.
+In the following paragraphs, subscripts on the vertices in $W$ are implicitly taken modulo $r$.
 By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
 
-If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) is involved in at least two crossings.
+If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
 Therefore each dummy vertex $w_i$ of $W$ is surrounded by two vertices $w_{i-1},w_{i+1}\in V(G)$.
-Since $w_i$ has degree $4$ in $G_0$, $w_{i-1}\neq w_i$.
+Since $w_i$ has degree $4$ in $G_0$, $w_{i-1}\neq w_{i+1}$.
 Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.
@@
 
-Therefore, each face of $G$ is one of two types:
+Therefore each face of $G'$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
 \begin{inparaenum}[(i)]
-  \item bounded by three edges of $G$ (none of which are crossed) or 
+  \item bounded by three edges of $G$, none of which are crossed; or
   \item bounded by one edge of $G$ that is not crossed plus portions of two edges of $G$ that cross each other.
 \end{inparaenum}
+
 Consider some crossing $(p,vw,xy)$ in $G$.
@@
 Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
-It follows that $\dist_{G'}(v,w)\le 2$.
+Therefore $\dist_{G'}(v,w)\le 2$.
 
@@
 Prior to the current work, the strongest upper bound on the non-repetitive chromatic number of $n$-vertex  $k$-planar graphs was $O(k\log n)$ \cite{dujmovic.morin.ea:layered}.
+=======
+%      \includegraphics{figs/one_planar_example-1} &
+%      \includegraphics{figs/one_planar_example-2} \\
+%      (a) & (b) \\[1em]
+%  \caption{Examples of (a)~a $1$-plane graph that is not edge maximal; (b)~an edge-maximal $1$-plane graph $G$; (c)~the kites in $G$; (d)~a plane multigraph $G'$ obtained by removing one spar from each kite of $G$.}
+
+To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
+Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_r$ be the facial walk around $F$.
+By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
+>>>>>>> 432123323ba65fbac04503f0de5e50ec4ebd89ce
+
+If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) woul
 
@@
 The $k$-nearest-neighbour graph of a point set $P\subset\R^2$ is the geometric graph $G$ with vertex set $V(G)=P$, where the edge set is defined as follows.
-For each point $v\in P$, let $N_k(v)$ be the set of $k$ points in $P$ closest to $v$.
+For each point $v\in P$, let $N_k(v)$ be the set of $k$ points i
+=======
+%      \includegraphics{figs/one_planar_example-1} &
+%      \includegraphics{figs/one_planar_example-2} \\
+%      (a) & (b) \\[1em]
+%  \caption{Examples of (a)~a $1$-plane graph that is not edge maximal; (b)~an edge-maximal $1$-plane graph $G$; (c)~the kites in $G$; (d)~a plane multigraph $G'$ obtained by removing one spar from each kite of $G$.}
+
+To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
+Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_r$ be the facial walk around $F$.
+By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
+>>>>>>> 432123323ba65fbac04503f0de5e50ec4ebd89ce
+
+If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) wouln $P$ closest to $v$.
 Then $vw\in E(G)$ if and only if $w\in N_k(v)$ or $v\in N_k(w)$.

### 2022-03-08 k-planar.tex
Note to self

@@
   Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.
-(This implies that $w=v$ for $v\in V(F)$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,2,3\}$ and we define $\alpha(v):=i$.
+(This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,2,3\}$ and we define $\alpha(v):=i$.
 
   If $k<3$ then let $\tau:=v_1v_2v_3$ be any inner face of $N'$.
+\note{PM}{Not quite that easy.
+We want $M$ to be $2$-connected so have to make sure $\overline{Y}$ attaches to at least two different vertices of $F$.}
   Otherwise, $k=3$, in which case we apply Sperner's Lemma to obtain an inner face $\tau:=v_1v_2v_3$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.

### 2022-03-08 k-planar.tex
Some changes

--- 2022-03-08 k-planar.tex
@@
 
-  Assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,2,3\}$ as follows:
-  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.
-(This implies that $w=v$ for $v\in V(F)$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,2,3\}$ and we define $\alpha(v):=i$.
-
-  If $k<3$ then let $\tau:=v_1v_2v_3$ be any inner face of $N'$.
-Otherwise, $k=3$, in which case we apply Sperner's Lemma to obtain an inner face $\tau:=v_1v_2v_3$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+  If $k\in\{1,2\}$ then we let $\tau:=v_1v_2v_3$ be an inner face of $N$ that contains at least one edge of $F$ on its boundary.
+If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.
+(This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
+By Sperner's Lemma there exists an inner face $\tau:=v_1v_2v_3$ of $N$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+
+  % If $k=2$, then let $\tau:=v_1v_2$ be any edge of $N'$ whose interior is in the interior of $F$ and whose endpoints have different colours.
+The existence of such an edge is straightforward to establish by considering the auxilliary graph $A$ whose vertices are the faces of $N'$ and in which two faces are adjacent if and only if they share an edge whose endpoints have different colours.
+Every vertex in $A$ has degree $0$ or $2$ and the outer face $F_0$ of $N'$ has degree $2$ in $A$.
+Therefore $A$ contains a cycle $C$ that contains $F_0$.
+Two edges of $C$ correspond to bichromatic edges of $F$ and the the remaining edges correspond to bichromatic edges in the interior of $F$, so it only remains to verify that $C$ has length at least $3$.
+The only other possibility is that $C$ has length $2$, which
+
+  % See, for example, \cite[Lemma~6]{bose.morin.ea:optimal}.
+
   For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to $V(F)$.
@@
 	Let $M$ and $M^+$ denote the subgraph of $G$ containing the edges and vertices of $\overline{Y}$, respectively $\overline{Y}^+$, and the edges and vertices of $F$.
-(See \cref{boring_figure}.)  By the choice of $\tau$, the vertices on the boundary of each inner face of $M$ are contained in $Y\cup P_a\cup P_b$ for some $a,b\in\{1,\ldots,k\}$.
+(See \cref{boring_figure}.)  By our choice of $\tau$, the graph $M$ is $2$-connected.
+Furthermore, the vertices on the boundary of each inner face of $M$ are contained in $Y\cup P_a\cup P_b$ for some $a,b\in\{1,\ldots,k\}$.
 In particular if $k=3$ then, for each inner face $F_i$ of $M$ there exists some $c\in\{1,2,3\}$ such that no vertex of $P_c$ is on the boundary of $F_i$.
@@
 
-  The graph $M^+$ has some number of inner faces, all contained in the interior of $F$. Some of the inner faces of $M^+$ are Type~(ii) faces. Let $F_1,\ldots,F_m$ be the Type~(i) inner faces of $M^+$.
-
+  The graph $M^+$ has some number of inner faces, all contained in the interior of $F$.
+Some of the inner faces of $M^+$ are Type~(ii) faces of $G$.
+Let $F_1,\ldots,F_m$ be the inner faces of $M^+$ that are not Type~(ii) faces of $G$.
 We claim that, for each $i\in\{1,\ldots,m\}$, the boundary of $F_i$ is a cycle in $G'$ that contains no spars.
@@
 Then, $vw\not\in E(F)$ since no edge of $F$ is crossed by any edge of $G$.
-Therefore $vw\in E(\overline{Y}^+)$ so $xy\in E(Y^+)$.
-But then the only faces of $M^+$ incident to $vw$ are Type~(ii) faces.
-In particular $vw$ cannot be incident to the Type~(i) face $F_i$.
+Therefore $vw\in E(\overline{Y}^+)$ so $xy\in E(\overline{Y}^+)$.
+But then the only faces of $M^+$ with $vw$ on their boundary are Type~(ii) faces.
+In particular $vw$ cannot be on the boundary of $F_i$.
 
   Observe that each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.
-Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.\note{PM}{This latter requirement requires that $M^+$ be $2$-connected, we should mention this.}
+Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.
 Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(F_i)$ to obtain the desired $H$-partition $\PP_i$ and tree decomposition of subgraph $N_i$ of $G$ contained in $F_i$.
 
+  \note{PM}{Pick up here tomorrow afternoon. For now I have to teach and prepare a talk.}
   The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.

### 2022-03-08 k-planar.tex
changes

--- 2022-03-08 k-planar.tex
@@
 		\end{compactenum}
-	\note{DW}{Should the vertices already grabbed by $P_1,P_2,P_3$ be mentioned in the setup?
+% \note{DW}{Should the vertices already grabbed by $P_1,P_2,P_3$ be mentioned in the setup?
 The new paths $Q_1,Q_2,Q_3$ need to avoid the vertices already grabbed by $P_1,P_2,P_3$.
-Basically, I am concerned about what happens when distinct vertical paths in $T$ want to grab the same vertex.}
-  \note{PM}{No, at this point we're only promising an $H$-partition of $N$, the part of $G$ contained in the uncrossed cycle $F$.}
+Basically, I am concerned about what happens when distinct vertical paths in $T$ want to grab the same vertex.} \note{PM}{No, at this point we're only promising an $H$-partition of $N$, the part of $G$ contained in the uncrossed cycle $F$.}
 		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
@@
 		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 15$;
-		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$
-		\note{DW}{Should this be $P_i\subseteq S_{x_i}$? (since the path grabs other vertices into its part)}\note{PM}{No, for the same reason as above.}; and
+		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; 
+%		\note{DW}{Should this be $P_i\subseteq S_{x_i}$? (since the path grabs other vertices into its part)}\note{PM}{No, for the same reason as above.}
+ and
 		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
@@
 	Fix a $(\Sigma,k)$-plane drawing of $G_0$ in a surface $\Sigma$ of Euler genus at most $g$. Choose $\epsilon>\epsilon'>0$.
-	For each vertex $v$ of $G_0$, let $B_v$ be the \defin{ball} $\{x\in \mathbb{R}^2:\dist_(x,v)\leq\epsilon\}$
+	For each vertex $v$ of $G_0$, let $B_v$ be the \defin{ball} $\{x\in \mathbb{R}^2:\dist(x,v)\leq\epsilon\}$
 	For each edge $vw$ of $G_0$, let $C_{vw}$ be the \defin{channel} $\{x\in \mathbb{R}^2:\dist(x,vw)\leq\epsilon'\}\setminus(B_v\cup B_w)$.
@@
 Prior to the current work, the strongest upper bound on the non-repetitive chromatic number of $n$-vertex  $k$-planar graphs was $O(k\log n)$ \cite{dujmovic.morin.ea:layered}.
-=======
-%      \includegraphics{figs/one_planar_example-1} &
-%      \includegraphics{figs/one_planar_example-2} \\
-%      (a) & (b) \\[1em]
-%  \caption{Examples of (a)~a $1$-plane graph that is not edge maximal; (b)~an edge-maximal $1$-plane graph $G$; (c)~the kites in $G$; (d)~a plane multigraph $G'$ obtained by removing one spar from each kite of $G$.}
-
-To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
-Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_r$ be the facial walk around $F$.
-By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
->>>>>>> 432123323ba65fbac04503f0de5e50ec4ebd89ce
-
-If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) woul
 
@@
 \label{StringPartition}
-For integers $g\geq 0$ and $\delta\geq 2$, let $\ell:= \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$,
-and $t:= \binom{ \delta+4}{3}-1$.
+For integers $g\geq 0$ and $\delta\geq 2$, let $\ell:= \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$ and $t:= \binom{ \delta+4}{3}-1$.
 Then  every $(g,\delta)$-string graph:
@@
 The $k$-nearest-neighbour graph of a point set $P\subset\R^2$ is the geometric graph $G$ with vertex set $V(G)=P$, where the edge set is defined as follows.
-For each point $v\in P$, let $N_k(v)$ be the set of $k$ points i
-=======
-%      \includegraphics{figs/one_planar_example-1} &
-%      \includegraphics{figs/one_planar_example-2} \\
-%      (a) & (b) \\[1em]
-%  \caption{Examples of (a)~a $1$-plane graph that is not edge maximal; (b)~an edge-maximal $1$-plane graph $G$; (c)~the kites in $G$; (d)~a plane multigraph $G'$ obtained by removing one spar from each kite of $G$.}
-
-To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
-Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_r$ be the facial walk around $F$.
-By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
->>>>>>> 432123323ba65fbac04503f0de5e50ec4ebd89ce
-
-If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) wouln $P$ closest to $v$.
+For each point $v\in P$, let $N_k(v)$ be the set of $k$ points in $P$ closest to $v$.
 Then $vw\in E(G)$ if and only if $w\in N_k(v)$ or $v\in N_k(w)$.

### 2022-03-09 k-planar.tex
started appendix

@@
 
-\end{document}s
+\appendix
+\section{New Stuff}
+
+\note{DW}{Work in progress}
+
+For a plane multigraph $G$, let $\widehat{G}$ be the embedded graph obtained from $G$ as follows: for each face $f$ of $G$ and for each pair of non-consecutive vertices $v$ and $w$ on the boundary of $f$, add the edge $vw$ to $\widehat{G}$ drawn across $f$.
+
+\begin{thm}
+\label{CliqueFaceProductStructure}
+Let $G$ be a 2-connected plane multigraph, in which every facial cycle has length in $\{3,\dots,d\}$ for some fixed integer $d\geq 3$.
+Then $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d  \floor{d/2} }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
+\end{thm}
+
+Note that \cref{CliqueFaceProductStructure} with $d=3$ implies \cref{PlanarProduct}(b)  by \citet{DJMMUW20} (since every planar graph is a subgraph of a triangulation).
+Moreover, the proof of \cref{CliqueFaceProductStructure} is a direct generalisation of the proof of
+\cref{PlanarProduct}(b).
+
+For a cycle $C$, we write $C = [ P_1 , \dots , P_k ]$ if $P_1 , \dots, P_k$ are pairwise disjoint non-empty paths in $C$, and the endpoints of each path $P_i$ can be labelled $x_i$ and $y_i$ so that $y_ix_{i+1} \in E(C)$ for $i\in\{1,\dots,k\}$, where $x_{k+1}$ means
+$x_1$. This implies that $\{P_1,\dots,P_k\}$ is a vertex-partition of $C$.
+
+If $Q$ is a path in a graph, and $P_1, \dots,P_k$ are edge-disjoint non-empty subpaths of $Q$ ordered by $Q$ and partitioning $E(Q)$, then we write $Q=P_1,\dots,P_k$.
+
+We employ the following well-known variation of Sperner's Lemma (see~\citep{Proofs4}).
+
+\begin{lem}[Sperner's Lemma]
+\label{Sperner}
+Let $G$ be a near-triangulation whose vertices are coloured $1,2,3$, with the outer-face $F=[P_1,P_2,P_3]$ where each vertex in $P_i$ is coloured $i$.
+Then $G$ contains an internal face whose vertices are coloured $1,2,3$.
+\note{DW}{Does this allow parallel edges: check the proof in \citep{Proofs4}.}
+\end{lem}
+
+%If $C$ is a cycle in a graph, and $P_1, \dots,P_k$ are edge-disjoint non-empty subpaths of $C$ ordered around $C$ and partitioning $E(C)$, then we write $C=(P_1,\dots,P_k)$.
+
+Say $G$ is a 2-connected plane multigraph.
+Let $T$ be a spanning tree of $G$ rooted at a vertex $r$ on the outerface of $G$.
+A \defin{tripod} (with respect to $T$) is any subgraph of $G$ consisting of three vertical paths $P,Q,R$ in $T$ plus a facial cycle $F$ of $G$,
+where the lower endpoint of each of $P,Q,R$ is in $F$. A \defin{bipod} (with respect to $T$) is any (connected) subpath of a tripod (with respect to $T$).
+ %is a path $PXQ$ where $P$ and $Q$ are vertical paths in $T$ and $X$ is a facial path of $G$, where the lower endpoints of $P$ and $Q$ are the endpoints of $X$. 
+\note{DW}{Need to think about degenerate cases. We want every bipod to be a tripod}
+%, and every connected subgraph of a bipod to be a bipod.}.
+ These definitions  generalise definitions in \citep{DJMMUW20} where $G$ is assumed to be a plane triangulation, in which case $F$ is a 3-cycle and $X$ is an edge.
+\note{DW}{We actually managed to omit the definition of `bipod' in \citep{DJMMUW20}!}
+
+%In the original proof for planar triangulation, B_i is an edge.
+
+\begin{lem}
+Fix an integer $d \geq 3$.
+Say $G_0$ is a 2-connected plane multigraph, in which each facial cycle has length in $\{3,\dots,d\}$.
+\note{DW}{I expect $d$ will be irrlevant to this lemma.} Let $T$ be a spanning tree of $G_0$ rooted at some vertex $r$ on the boundary of the outerface of $G_0$.
+Let $k\in\{1,2,3\}$ and let $C = [A_1,\dots,A_k]$ be a cycle in $G_0$ with $r$ is in the exterior of $C$, where
+$A_1, \dots, A_k$ are bipods in $G_0$. Let $G$ be the plane graph consisting of all the edges and vertices of $G_0$ in $C$ and the interior of $C$.
+Then $G$ has a partition $\PP$ into tripods with resect to $T$ such that $P_1,\dots,P_k\in\PP$ and the graph $H : = G / \PP$ is planar and has a tree-decomposition in which every bag has size at most 4 and some bag contains all the vertices of $H$ corresponding to $P_1,\dots,P_k$.
+\note{DW}{Do we need to add that ``each face of $G_0$ is the centre of some tripod in $\PP$'' or something like that?
+Then each facial clique will be in some part.}
+\end{lem}
+
+\begin{proof}
+	\note{DW}{Is $C$ allowed to be a 2-cycle?}
+The proof is by induction on $|V(G)\setminus V(C)|$.
+If $V(G)\setminus V(C)=\emptyset$, then the desired partition is $\PP=\{P_1,\ldots,P_k\}$.
+The tree-decomposition of $H$ consists of a single bag that contains the $k\le 3$ vertices corresponding to $P_1,\ldots,P_k$.
+
+Now assume that $V(G)\setminus V(C)\neq\emptyset$. Partition the vertices of $C$ into paths $R_1$, $R_2$, and $R_3$ as follows:
+\begin{compactenum}
+	\item If $k=1$ then, since $C$ is a cycle, $P_1$ has at least three vertices, so $P_1=[v, P_1', w]$ for two distinct vertices $v$ and $w$.
+\note{DW}{need to be careful with parallel edges} We set $R_1:=v$, $R_2:=P_1'$ and $R_3:=w$.
+
+	\item If $k=2$ then we may assume without loss of generality that $P_1$ has at least two vertices so $P_1=[v,P_1']$ \note{DW}{need to be careful with parallel edges}.
+We set $R_1:=v$, $R_2:=P_1'$ and $R_3:=P_2$.
+
+	\item If $k=3$ then let $R_i:=P_i$ for each $i\in\{1,2,3\}$.
+\end{compactenum}
+
+Let $G'$ be an arbitrary triangulation of $G_0$.
+For $i\in\{1,2,3\}$, colour each vertex in $R_i$ by $i$.
+Now, for each remaining vertex $v$ in $G$, consider the path $P_v$ from $v$ to the root of $T$.
+Since $r$ is on the outer-face of $G_0$, $P_v$ contains at least one vertex of $C$.
+If the first vertex of $P_v$ that belongs to $F$ is in $R_i$ then assign the colour $i$ to $v$.
+We obtain a 3-colouring of the vertices of $G'$ that satisfies the conditions of Sperner's Lemma.
+By Sperner's Lemma there exists a triangular face $\tau=v_1v_2v_3$ of $G'$ whose vertices are coloured $1,2,3$ respectively.
+Let $F$ be the face of $G_0$ containing $\tau$.
+
+For each $i\in\{1,2,3\}$, let $Q_i$ be the path in $T$ from $v_i$ to the first ancestor $v_i'$ of $v_i$ in $T$ that is contained in $C$.
+Observe that $Q_1$, $Q_2$, and $Q_3$ are disjoint since $Q_i$ consists only of vertices coloured $i$.
+Note that $Q_i$ may consist of the single vertex $v_i=v_i'$.
+Let $Q_i'$ be $Q_i$ minus its final vertex $v_i'$.
+Imagine for a moment that the cycle $C$ is oriented clockwise, which defines an orientation of $R_1$, $R_2$ and $R_3$.
+Let $R_i^-$ be the subpath of $R_i$ that contains $v'_i$ and all vertices that precede it, and let $R_i^+$ be the subpath of $R_i$ that contains $v'_i$ and all vertices that succeed it.
+
+Say $F=v_1 S_1 v_2 S_2 v_3 S_3$, where $S_i$ is a facial path in $F$ clockwise from $v_i$ to $v_{i+1}$.
+
+, where we use the convention that $Q_4=Q_1$ and $R_4=R_1$. Note that $C_i$ may be \emph{degenerate} in the sense that $[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ may consist only of a single edge $v_iv_{i+1}$.
+
+Let $Y$ be the tripod consisting of $Q_1'\cup Q_2' \cup Q_3'$ plus the vertices and edges of $F$ between non-empty $Q_1',Q_2',Q_3'$.
+
+\note{DW}{Much of the following paragraph can be deleted} Consider any non-degenerate $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$.
+Note that these four paths  are pairwise disjoint, and thus $C_i$ is a cycle.
+If  $Q_i'$ and $Q_{i+1}'$ are non-empty, then each is a vertical path in $T$.
+Furthermore, each of $R_i^-$ and $R_{i+1}^+$ consists of at most two vertical paths in $T$.
+Thus, $C_i$ is the concatenation of at most six vertical paths in $T$ \note{DW}{re-do}.
+Let $G_i$ be the subgraph of $G_0$ consisting of all the edges and vertices of $G_0$ contained in $C_i$ and the interior of $C_i$.
+Observe that $G_i$ contains $v_i$ and $v_{i+1}$ but not the third vertex of $\tau$.
+
+Consider the graph consisting of the edges and vertices of $F$, the edges and vertices of $C$, and the edges and vertices of $Q_1,Q_2,Q_3$.
+This graph has up to three \note{DW}{can be more} internal faces $C_1,C_2,C_3$ where each $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ and $R_{i}^+$ and $R_{i}^-$ are the same portions of $R_i$ as defined in \cref{NearTriang} \note{DW}{move in here}    Observe that  $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$, where $R_i^+$ and $R_{i+1}^-$ are bipods, and $I_i$ is the bipod $Q_i' S_i Q_{i+1}'$.
+Let $G_i$ be the subgraph of $G$ whose vertices and edges are in $C_i$ or its interior.
+
+For $i\in\{1,2,3\}$, if $C_i$ is non-empty, then $G_i$ and $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$ satisfy the conditions of the lemma, and we may apply induction to $G_i$.
+(Note that one or two of $R_i^+$, $R_{i+1}^-$ and $I_i$ may be empty, in which case we apply the inductive hypothesis with $k=2$ or $k=1$, respectively.)\  This gives a partition $\PP_i$ of $G_i$ such that $H_i:= G_i / \PP_i$ satisfies the conclusions of the lemma.
+Let $(B^i_x:x\in V(J_i))$ be a tree-decomposition of $H_i$, in which every bag has size at most 4, and some bag $B^i_{u_i}$ contains the vertices of $H_i$ corresponding to  $R_i^+$, $R_{i+1}^-$ and $I_i$ (if they are non-empty).
+We do this for each non-degenerate $C_i$.
+
+We now construct the desired partition $\PP$ of $G$. 
+Initialise $\PP:= \{P_1,\ldots,P_k\}$. Then add each non-empty $Q_i'$ to $\PP$. 
+Now for each non-degenerate $F_i$, each path in  $\PP_i$ is either an \emph{external} path (that is, fully contained in $F_i$) or is an \emph{internal path} with none of its vertices in $F_i$. 
+Add all the internal paths of  $\PP_i$ to $\PP$. 
+By construction, $\PP$ partitions $V(G)$ into vertical paths in $T$ and $\PP$ contains $P_1,\ldots,P_k$.
+
+Let $H:=G/\PP$. Next we exhibit the desired tree-decomposition $(B_x:x\in V(J))$ of $H$. 
+Let $J$ be the tree obtained from the disjoint union of $J_i$, taken over the $i\in\{1,2,3\}$ such that $F_i$ is non-degenerate, by adding one new node $u$ adjacent to each $u_i$.
+(Recall that $u_i$ is the node of $J_i$ for which the bag $B^i_{u_i}$ contains the vertices of $H_i$ corresponding to the paths that form $F_i$.)\ 
+Let the bag $B_u$ contain all the vertices of $H$ corresponding to $P_1,\ldots,P_k,Q'_1,Q'_2,Q'_3$. 
+For each non-degenerate $F_i$, and for each node $x\in V(J_i)$, initialise $B_x:= B^i_x$. 
+Recall that vertices of $H_i$ correspond to contracted paths in  $\PP_i$.
+Each internal path in $\PP_i$ also lies  in $\PP$.
+Each external path $P$ in $\PP_i$ is a subpath of $P_j$ for some $j\in\{1,\dots,k \}$ or is one of the paths among $Q'_1, Q'_2, Q'_3$. 
+For each such path $P$, for every $x\in V(J)$, in bag $B_x$, 
+replace each instance of the vertex of $H_i$ corresponding to $P$ by the vertex of $H$ corresponding to the path among $P_1,\ldots, P_k, Q'_1,\ldots,Q'_3$ 
+that contains $P$.
+This completes the description of $(B_x : x\in V(J))$. 
+By construction, $|B_x|\leq 9$ for every $x\in V(J)$.
+
+First we show that for each vertex $a$ in $H$, the set $X:=\{x\in V(J) : a\in B_x\}$ forms a subtree of $J$.
+If $a$ corresponds to a path distinct from $P_1,\ldots,P_k,Q'_1,Q'_2,Q'_3$ then $X$ is fully contained in $J_i$ for some $i\in\{1,2,3\}$.
+Thus, by induction $X$ is non-empty and connected in $J_i$, so it is in $J$.
+If $a$ corresponds to $P$ which is one of the paths among $P_1,\ldots,P_k,Q'_1,Q'_2,Q'_3$ then 
+$u\in X$ and whenever $X$ contains a vertex of $J_i$ it is because some external path of $\PP_i$ was replaced by $P$.
+In particular, we would have $u_i \in X$ in that case. 
+Again by induction each $X\cap J_i$ is connected and since $uu_i\in E(T)$, we conclude that $X$ induces a (connected) subtree of $J$.
+
+Finally we show that, for every edge $ab$ of $H$, there is a bag $B_x$ that contains $a$ and $b$. 
+If $a$ and $b$ are both obtained by contracting any of $P_1,\ldots,P_k,Q'_1,Q'_2,Q'_3$, 
+then $a$ and $b$ both appear in $B_u$.  
+If $a$ and $b$ are both in $H_i$ for some $i\in\{1,2,3\}$, 
+then some bag $B^i_x$ contains both $a$ and $b$. 
+Finally, when $a$ is obtained by contracting a path $P_a$ in $G_i-V(F_i)$ and $b$ is obtained by contracting a path $P_b$ not in $G_i$, 
+then  the cycle $F_i$ separates $P_a$ from $P_b$ so the edge $ab$ is not present in $H$. 
+This concludes the proof that $(B_x:x\in V(J))$ is the desired tree-decomposition  of $H$.
+
+We construct $\PP$ as before.
+Initialise $\PP:=\{P_1,\ldots,P_k,Y\}$.
+Then, for $i\in \{1,2,3\}$, each tripod in  $\PP_i$ is either fully contained in $F_i$ or it is \emph{internal} with none of its vertices in $F_i$.
+Add all these internal tripods in  $\PP_i$ to $\PP$.
+By construction, $\PP$ partitions $V(G)$ into tripods.
+The graph $H:= G/ \PP$ is planar since $G$ is planar and each tripod in $\PP$ induces a connected subgraph of $G$.
+
+Next we produce the tree-decomposition $(B_x:x\in V(J))$ of $H$ that satisfies the requirements of the lemma.
+Let $J$ be the tree obtained from the disjoint union of $J_1$, $J_2$ and $J_3$ by adding one new node $u$ adjacent to $u_1$, $u_2$ and $u_3$.
+Let $B_u$ be the set of at most four vertices of $H$ corresponding to $Y,P_1,\dots,P_k$.
+For $i\in\{1,2,3\}$ and for each node $x\in V(J_i)$, initialise $B_x:= B^i_x$.
+
+As in  the proof of \cref{NearTriang},  the resulting structure,  $(B_x:x\in V(J))$, is not yet a tree-decomposition of $H$ since some bags may  contain vertices of $H_i$ that are not necessarily vertices of $H$.
+Note that unlike in  \cref{NearTriang} this does not only include elements of $\PP_i$  that are contained in $F$.
+In particular, $I_i$ is also not an element of $\PP$ and thus does not correspond to a vertex of $H$.   
+We remedy this  as follows. For $x\in V(J)$, in bag $B_x$, replace each instance of the vertex of $H_i$ corresponding to $I_i$ by the vertex of $H$ corresponding to $Y$. 
+Similarly, by construction, $R_i^+$ is a subgraph of $P_{\alpha_i}$ for some $\alpha_i\in \{1,\dots,k\}$. 
+For $x\in V(J)$, in bag $B_x$, replace each instance of the vertex of $H_i$ corresponding to $R_i^+$ 
+by the vertex of $H$ corresponding to $P_{\alpha_i}$. 
+Finally, $R_{i+1}^-$ is a subgraph of $P_{\beta_i}$ for some $\beta_i\in \{1,\dots,k\}$. 
+For $x\in V(J)$, in bag $B_x$, replace each instance of the vertex of $H_i$ corresponding to $R_{i+1}^-$  by the vertex of $H$ corresponding to $P_{\beta_i}$.
+
+This completes the description of $(B_x:x\in V(J))$. Clearly, every bag $B_x$ has size at most 4. We now prove that $(B_x:x\in V(J))$  is a tree-decomposition of $H$ 
+\end{proof}
+
+\begin{lem}
+	An embedded multigraph $G$ is edge-maximal 1-plane with no faces of length 2 if and only if there is a 2-connected plane graph $G_0$ such that: 
+	\begin{compactitem}
+		\item each face of $G_0$ is bounded by a 3-cycle or a 4-cycle,
+		\item no edge of $G_0$ is in the boundary of two facial 3-cycles,
+		\item $G=\widehat{G_0}$.
+	\end{compactitem}
+\end{lem}
+
+\begin{proof}
+\end{proof}
+
+\end{document}
+
+%%%%%%%%%%%%%%%%%%%%
+
+I claim that G_2 is contained in H * P * K_{ d * floor(d/2) } where tw(H) <= 3.
+
+With d=4, G_2 includes all edge-maximal 1-planar  multigraphs,
+and G_2 is contained in H * P * K_8 where tw(H) <= 3.
+
+Let T be a bfs spanning tree of G_0 rooted on the outerface.
+
+Run the original proof, finding three paths Q_1,Q_2,Q_3
+leading to a Sperner triangle t in G_1, which is a subset of a face F of G_0.
+The corresponding tripod in our partition is Q_1 \cup Q_2 \cup Q_3 \cup F.
+We can apply induction inside the regions bounded by C, Q_1, Q_2, Q_3 and F.
+
+The layering is in G_0. Each edge of G_2 has span at most floor(d/2)
+with respect to this layering.
+
+So G_2 is contained in H * P^{floor(d/2)} * K_d, which is contained in
+H * P * K_{d floor(d/2) }.
+
+This setup (if it works) also works for map graphs in which every face
+is a nation.
+
+I just thought we need to be careful with the vertex 3-colouring.
+If v is on the inside of C and the path in T from v to the root
+first hits C at some B_i, then what is the colour of v?
+This is not a problem there is a noose in G_0 including A_1..A_k, so
+the path in T from v to the root must intersect A_1...A_k.
+
+\end{document}
+
+	\begin{figure}[!b]
+	\begin{center}
+		\begin{tabular}{cc}
+			\includegraphics[scale=1.4]{induction-i-3} &
+			\includegraphics[scale=1.4]{induction-i-1} \\
+			(a) & (b) \\[1em]
+			\includegraphics[scale=1.4]{induction-i-2} &
+			\includegraphics[scale=1.4]{induction-iii} \\
+			(c) & (d) 
+		\end{tabular}
+	\end{center}
+	\caption{The inductive proof of \cref{NearTriang}: (a)~the spanning tree $T$ and the paths $P_1,\ldots,P_4$; 
+		(b) the paths $R_1$, $R_2$, $R_3$, and the Sperner triangle $\tau$; 
+		(c)~the paths $Q_1'$, $Q_2'$ and $Q_3'$; 
+		(d)~the near-triangulations $G_1$, $G_2$, and $G_3$, with the vertical paths of $T$ on $F_1$, $F_2$, and $F_3$.}
+	\label{NearTriangProof}
+\end{figure}

### 2022-03-09 k-planar.tex
The version with both spars removed is waaay cleaner

@@
 
-By removing one spar from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three edges; see \cref{one_planar_example}(b).
+By removing both spars from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three edges or $4$ edges; see \cref{one_planar_example}(b).
 Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
@@
 				\item $F[P_i]$ is a path; and
-				\item $|V(P_i)\cap L_j| \le 15$ for every integer $j\ge 0$.
+				\item $|V(P_i)\cap L_j| \le 3$ for every integer $j\ge 0$.
 			\end{compactenum}
@@
 	\end{compactenum}
-	Then $N$ has an $H$-partition $\PP=\{S_x : x\in V(H)\}$ such that:
+	Then $N$ has an $H$-partition $\PP=\{S_x\cup\{v_x\} : x\in V(H)\}$ such that:
 	\begin{compactenum}
 		\item $H$ is planar;
-		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 15$;
+		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 3$;
 		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; 
@@
 
-  If $k\in\{1,2\}$ then we let $\tau:=v_1v_2v_3$ be an inner face of $N$ that contains at least one edge of $F$ on its boundary.
+  If $k\in\{1,2\}$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge of $F$ on its boundary.
 If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
-By Sperner's Lemma there exists an inner face $\tau:=v_1v_2v_3$ of $N$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+Now add edges to the inner faces of $N'$ to obtain a supergraph $N''$ in which each inner face has three $3$ edges on its boundary.
+  By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.
+(Note that $p\in\{3,4\}$.)
 
@@
   Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
-Let $\overline{Y}^+$ denote the subgraph of $N$ consisting of the vertices and edges of $\overline{Y}$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $\overline{Y}$.
-Let $Y:= V(\overline{Y}-F)$ and let $Y^+:=V(\overline{Y}^+-F)$.
-
-	We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
+Let $S:=V(Q_1)\cup V(Q_2)\cup V(Q_3)$ and let $v:=v_p$.
+Observe that, for each $i\in\{1,2,3\}$ and each $j\ge 0$, $|V(Q_i)\cap L_j|\le 1$.
+Therefore $|S\cap L_j|\le 3$ for each integer $j\ge 0$.
+Therefore, taking $Y:=S\cup\{v\}$ as a part in the partition $\PP$ will satisfy Requirement~2 of the lemma.
+
+	% We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
 First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $L_i$, each incident on at most two edges of $\overline{Y}$.
@@
 
-	Let $M$ and $M^+$ denote the subgraph of $G$ containing the edges and vertices of $\overline{Y}$, respectively $\overline{Y}^+$, and the edges and vertices of $F$.
+	Let $M$ denote the subgraph of $G'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$.
 (See \cref{boring_figure}.)  By our choice of $\tau$, the graph $M$ is $2$-connected.
@@
     \end{center}
-    \caption{The graphs $M$ and $M^+$.}
+    \caption{The graphs $M$ and $M^+$. \note{PM}{This figure needs updating now.}}
     \label{boring_figure}
@@
 
-  The graph $M^+$ has some number of inner faces, all contained in the interior of $F$.
-Some of the inner faces of $M^+$ are Type~(ii) faces of $G$.
+  The graph $M$ has some number of inner faces $F_1,\ldots,F_m$, each of which is contained in the interior of $F$. 
+  % Some of the inner faces of $M^+$ are Type~(ii) faces of $G$.
 Let $F_1,\ldots,F_m$ be the inner faces of $M^+$ that are not Type~(ii) faces of $G$.
@@
 In particular $vw$ cannot be on the boundary of $F_i$.
-
-  Observe that each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.
+  Each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.
 Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.

### 2022-03-09 k-planar.tex
Don't add v_p if already in V(F)

@@
 Therefore $|S\cap L_j|\le 3$ for each integer $j\ge 0$.
-Therefore, taking $Y:=S\cup\{v\}$ as a part in the partition $\PP$ will satisfy Requirement~2 of the lemma.
+Therefore, taking $Y:=S\cup\{v\}\setminus V(F)$ as a part in the partition $\PP$ will satisfy Requirement~2 of the lemma.
 
@@
   Each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.
-Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.
+Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.
 Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(F_i)$ to obtain the desired $H$-partition $\PP_i$ and tree decomposition of subgraph $N_i$ of $G$ contained in $F_i$.

### 2022-03-09 k-planar.tex
Updated analysis of layered width

@@
 
-  \note{PM}{Pick up here tomorrow afternoon. For now I have to teach and prepare a talk.}
+  \note{PM}{Pick up here tomorrow, after teaching.}
   The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
@@
 
-	Let $G'$ be the plane graph obtained by removing one spar from each kite of $\overline{G}$ and let $T$ be a BFS tree of $G'$ rooted at $r$.
+	Let $G'$ be the plane graph obtained by removing both spars from each kite of $\overline{G}$ and let $T$ be a BFS tree of $G'$ rooted at $r$.
 Now apply \cref{induction} with $G=\overline{G}$, $G'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
-This gives an $H$-partition $\{S_x:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ is planar and has treewidth at most 3.
+This gives an $H$-partition $\{S_x\cup\{v_x\}:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ is planar and has treewidth at most 3.
 
@@
 That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le 2$ for every edge $vw\in E(G)$.
-Since $|L_j\cap S_x|\le 15$ for every integer $j\ge 0$, $|L_i'\cap S_x|\le 30$ for every integer $i\ge 0$ and every $x\in V(H)$.
+Since $|L_j\cap S_x|\le 3$ for every integer $j\ge 0$, $|L_i'\cap (S_x\cup\{v_x\})|\le 7$ for every integer $i\ge 0$ and every $x\in V(H)$.
 The result follows from \cref{PartitionProduct}.

### 2022-03-09 k-planar.tex
continuiing on appendix

--- 2022-03-09 k-planar.tex
@@
 
-By removing both spars from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three edges or $4$ edges; see \cref{one_planar_example}(b).
+By removing one spar from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three edges; see \cref{one_planar_example}(b).
 Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
@@
 				\item $F[P_i]$ is a path; and
-				\item $|V(P_i)\cap L_j| \le 3$ for every integer $j\ge 0$.
+				\item $|V(P_i)\cap L_j| \le 15$ for every integer $j\ge 0$.
 			\end{compactenum}
@@
 	\end{compactenum}
-	Then $N$ has an $H$-partition $\PP=\{S_x\cup\{v_x\} : x\in V(H)\}$ such that:
+	Then $N$ has an $H$-partition $\PP=\{S_x : x\in V(H)\}$ such that:
 	\begin{compactenum}
 		\item $H$ is planar;
-		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 3$;
+		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 15$;
 		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; 
@@
 
-  If $k\in\{1,2\}$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge of $F$ on its boundary.
+  If $k\in\{1,2\}$ then we let $\tau:=v_1v_2v_3$ be an inner face of $N$ that contains at least one edge of $F$ on its boundary.
 If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
-Now add edges to the inner faces of $N'$ to obtain a supergraph $N''$ in which each inner face has three $3$ edges on its boundary.
-  By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
-Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.
-(Note that $p\in\{3,4\}$.)
+By Sperner's Lemma there exists an inner face $\tau:=v_1v_2v_3$ of $N$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
 
@@
 	Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
-Let $S:=V(Q_1)\cup V(Q_2)\cup V(Q_3)$ and let $v:=v_p$.
-Observe that, for each $i\in\{1,2,3\}$ and each $j\ge 0$, $|V(Q_i)\cap L_j|\le 1$.
-Therefore $|S\cap L_j|\le 3$ for each integer $j\ge 0$.
-Therefore, taking $Y:=S\cup\{v\}$ as a part in the partition $\PP$ will satisfy Requirement~2 of the lemma.
-
-	% We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
+Let $\overline{Y}^+$ denote the subgraph of $N$ consisting of the vertices and edges of $\overline{Y}$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $\overline{Y}$.
+Let $Y:= V(\overline{Y}-F)$ and let $Y^+:=V(\overline{Y}^+-F)$.
+
+	We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
 First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $L_i$, each incident on at most two edges of $\overline{Y}$.
@@
 
-	Let $M$ denote the subgraph of $G'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$.
+	Let $M$ and $M^+$ denote the subgraph of $G$ containing the edges and vertices of $\overline{Y}$, respectively $\overline{Y}^+$, and the edges and vertices of $F$.
 (See \cref{boring_figure}.)  By our choice of $\tau$, the graph $M$ is $2$-connected.
@@
     \end{center}
-    \caption{The graphs $M$ and $M^+$. \note{PM}{This figure needs updating now.}}
+    \caption{The graphs $M$ and $M^+$.}
     \label{boring_figure}
@@
 
-  The graph $M$ has some number of inner faces $F_1,\ldots,F_m$, each of which is contained in the interior of $F$. 
-  % Some of the inner faces of $M^+$ are Type~(ii) faces of $G$.
+  The graph $M^+$ has some number of inner faces, all contained in the interior of $F$.
+Some of the inner faces of $M^+$ are Type~(ii) faces of $G$.
 Let $F_1,\ldots,F_m$ be the inner faces of $M^+$ that are not Type~(ii) faces of $G$.
@@
 In particular $vw$ cannot be on the boundary of $F_i$.
-  Each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.
+
+  Observe that each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.
 Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.
@@
 \begin{lem}
-Fix an integer $d \geq 3$.
-Say $G_0$ is a 2-connected plane multigraph, in which each facial cycle has length in $\{3,\dots,d\}$.
-\note{DW}{I expect $d$ will be irrlevant to this lemma.} Let $T$ be a spanning tree of $G_0$ rooted at some vertex $r$ on the boundary of the outerface of $G_0$.
-Let $k\in\{1,2,3\}$ and let $C = [A_1,\dots,A_k]$ be a cycle in $G_0$ with $r$ is in the exterior of $C$, where
-$A_1, \dots, A_k$ are bipods in $G_0$. Let $G$ be the plane graph consisting of all the edges and vertices of $G_0$ in $C$ and the interior of $C$.
-Then $G$ has a partition $\PP$ into tripods with resect to $T$ such that $P_1,\dots,P_k\in\PP$ and the graph $H : = G / \PP$ is planar and has a tree-decomposition in which every bag has size at most 4 and some bag contains all the vertices of $H$ corresponding to $P_1,\dots,P_k$.
-\note{DW}{Do we need to add that ``each face of $G_0$ is the centre of some tripod in $\PP$'' or something like that?
-Then each facial clique will be in some part.}
+Say $G_0$ is a 2-connected plane multigraph.
+Let $T$ be a spanning tree of $G_0$ rooted at some vertex $r$ on the boundary of the outerface of $G_0$.
+Let $k\in\{1,2,3\}$ and let $C = [A_1,\dots,A_k]$ be a cycle in $G_0$ with $r$ in the exterior of $C$, where $A_1, \dots, A_k$ are bipods in $G_0$.
+Let $G$ be the plane graph consisting of all the edges and vertices of $G_0$ in $C$ and the interior of $C$.
+Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that:
+\begin{compactitem}
+\item $P_1,\dots,P_k\in\PP$,
+\item the quotient $G / \PP$ is planar and has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
+\begin{compactitem}
+	\item some bag of $\mathcal{T}$ contains $P_1,\dots,P_k$, and 
+	\item for every face $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
+\end{compactitem}
+\end{compactitem}
 \end{lem}
@@
 \begin{proof}
+	\note{DW}{Can we always assume $k=3$ by allowing $A_i=\emptyset$? That would clean it up.}
+
 	\note{DW}{Is $C$ allowed to be a 2-cycle?}
 The proof is by induction on $|V(G)\setminus V(C)|$.
-If $V(G)\setminus V(C)=\emptyset$, then the desired partition is $\PP=\{P_1,\ldots,P_k\}$.
-The tree-decomposition of $H$ consists of a single bag that contains the $k\le 3$ vertices corresponding to $P_1,\ldots,P_k$.
+If $V(G)\setminus V(C)=\emptyset$, then $G$ has just one face bounded by $C$ and the desired partition is $\PP=\{P_1,\ldots,P_k\}$.
+The tree-decomposition of $G/\PP$ consists of a single bag that contains the $k\le 3$ vertices corresponding to $P_1,\ldots,P_k$.
 
@@
 
-Let $G'$ be an arbitrary triangulation of $G_0$.
+Let $G'$ be a near-triangulation obtained from $G$ by triangulating every internal face.
 For $i\in\{1,2,3\}$, colour each vertex in $R_i$ by $i$.
@@
 Observe that $Q_1$, $Q_2$, and $Q_3$ are disjoint since $Q_i$ consists only of vertices coloured $i$.
-Note that $Q_i$ may consist of the single vertex $v_i=v_i'$.
 Let $Q_i'$ be $Q_i$ minus its final vertex $v_i'$.
+Note that $Q_i$ may consist of the single vertex $v_i=v_i'$, in which case $Q_i'$ is empty.
 Imagine for a moment that the cycle $C$ is oriented clockwise, which defines an orientation of $R_1$, $R_2$ and $R_3$.
@@
 
+Let $Y$ be the tripod consisting of $Q_1'\cup Q_2' \cup Q_3'$ plus the vertices and edges of $F$.
+
 Say $F=v_1 S_1 v_2 S_2 v_3 S_3$, where $S_i$ is a facial path in $F$ clockwise from $v_i$ to $v_{i+1}$.
 
-, where we use the convention that $Q_4=Q_1$ and $R_4=R_1$. Note that $C_i$ may be \emph{degenerate} in the sense that $[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ may consist only of a single edge $v_iv_{i+1}$.
-
-Let $Y$ be the tripod consisting of $Q_1'\cup Q_2' \cup Q_3'$ plus the vertices and edges of $F$ between non-empty $Q_1',Q_2',Q_3'$.
-
-\note{DW}{Much of the following paragraph can be deleted} Consider any non-degenerate $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$.
+Let $L_i$ be the path $Q_i'S_iQ_{i+1}'$ for $i\in\{1,\dots,k\}$. So $L_i$ is a bipod.
+
+, where we use the convention that $Q_4=Q_1$ and $R_4=R_1$.
+
+%Note that $C_i$ may be \defin{degenerate} in the sense that $[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ may consist only of a single edge $v_iv_{i+1}$.
+
+% Consider any non-degenerate $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$.
 Note that these four paths  are pairwise disjoint, and thus $C_i$ is a cycle.
@@
 
-Consider the graph consisting of the edges and vertices of $F$, the edges and vertices of $C$, and the edges and vertices of $Q_1,Q_2,Q_3$.
-This graph has up to three \note{DW}{can be more} internal faces $C_1,C_2,C_3$ where each $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ and $R_{i}^+$ and $R_{i}^-$ are the same portions of $R_i$ as defined in \cref{NearTriang} \note{DW}{move in here}    Observe that  $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$, where $R_i^+$ and $R_{i+1}^-$ are bipods, and $I_i$ is the bipod $Q_i' S_i Q_{i+1}'$.
+Consider the plane graph $Z$ consisting of the edges and vertices of $F$, the edges and vertices of $C$, and the edges and vertices of $Q_1,Q_2,Q_3$.
+Then $Z$ is 2-connected, each face of $Z$ is bounded by a cycle, and the outerface of $Z$ is $C$.
+Let $Z_1,\dots,Z_m$ be the internal faces of $Z$.
+
+Consider $j\in\{1,\dots,m\}$.
+Let $G_j$ be the subgraph of $G$ consisting of all the edges and vertices of $G$ in $Z_j$ and the interior of $Z_j$.
+The outerface of $G_j$ is $Z_j$.
+Observe that for some $i\in\{1,\dots,k\}$, we have
+$Z_j=[R_{i,j}^+,R_{i+1,j}^-,S_{i,j}]$, where 
+$R_{i,j}^+$ is some subpath of $R_i^+$,
+$R_{i+1,j}^-$ is some subpath of $R_{i+1}^-$, and
+$S_{i,j}$ is some subpath of $S_i$.
+Since $R_i^+$, $R_{i+1}^-$ and $S_i$ are bipods, 
+$R_{i,j}^+$, $R_{i+1,j}^-$ and $S_{i,j}$ are bipods. 
+By construction, $v_{i-1}$ is not in $G_j$. 
+Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.  Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that:
+\begin{compactitem}
+	\item $R_{i,j}^+, R_{i+1,j}^-, S_{i,j}\in\PP_j$,
+	\item the quotient $G_j / \PP_j$ is planar and has a tree-decomposition $\mathcal{T}_j$ of width at most 3 such that: 
+	\begin{compactitem}
+		\item some bag $B_j$ of $\mathcal{T}_j$ contains $R_{i,j}^+, R_{i+1,j}^-, S_{i,j}$, and 
+		\item for every face $F$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
+	\end{compactitem}
+\end{compactitem}
+\end{lem}
+
+%This graph has up to three \note{DW}{can be more} internal faces $C_1,C_2,C_3$ where each $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ and $R_{i}^+$ and $R_{i}^-$ are the same portions of $R_i$ as defined in \cref{NearTriang} \note{DW}{move in here}    Observe that  $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$, where $R_i^+$ and $R_{i+1}^-$ are bipods, and $I_i$ is the bipod $Q_i' S_i Q_{i+1}'$.
 Let $G_i$ be the subgraph of $G$ whose vertices and edges are in $C_i$ or its interior.
 
-For $i\in\{1,2,3\}$, if $C_i$ is non-empty, then $G_i$ and $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$ satisfy the conditions of the lemma, and we may apply induction to $G_i$.
+%For $i\in\{1,2,3\}$, if $C_i$ is non-empty, then $G_i$ and $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$ satisfy the conditions of the lemma, and we may apply induction to $G_i$.
 (Note that one or two of $R_i^+$, $R_{i+1}^-$ and $I_i$ may be empty, in which case we apply the inductive hypothesis with $k=2$ or $k=1$, respectively.)\  This gives a partition $\PP_i$ of $G_i$ such that $H_i:= G_i / \PP_i$ satisfies the conclusions of the lemma.

### 2022-03-09 k-planar.tex

@@
   Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
-Let $\overline{Y}^+$ denote the subgraph of $N$ consisting of the vertices and edges of $\overline{Y}$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $\overline{Y}$.
-Let $Y:= V(\overline{Y}-F)$ and let $Y^+:=V(\overline{Y}^+-F)$.
-
-	We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
+Let $S:=V(Q_1)\cup V(Q_2)\cup V(Q_3)$ and let $v:=v_p$.
+Observe that, for each $i\in\{1,2,3\}$ and each $j\ge 0$, $|V(Q_i)\cap L_j|\le 1$.
+Therefore $|S\cap L_j|\le 3$ for each integer $j\ge 0$.
+Therefore, taking $Y:=S\cup\{v\}\setminus V(F)$ as a part in the partition $\PP$ will satisfy Requirement~2 of the lemma.
+
+	% We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
 First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $L_i$, each incident on at most two edges of $\overline{Y}$.
@@
 In particular $vw$ cannot be on the boundary of $F_i$.
-
-  Observe that each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.
-Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.
+  Each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.
+Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.
 Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(F_i)$ to obtain the desired $H$-partition $\PP_i$ and tree decomposition of subgraph $N_i$ of $G$ contained in $F_i$.
 
-  \note{PM}{Pick up here tomorrow afternoon. For now I have to teach and prepare a talk.}
+  \note{PM}{Pick up here tomorrow, after teaching.}
   The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.

### 2022-03-10 k-planar.tex
first draft of proof in appendix complete

@@
 \item $P_1,\dots,P_k\in\PP$,
-\item the quotient $G / \PP$ is planar and has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
+\item $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
 \begin{compactitem}
@@
 \begin{proof}
-	\note{DW}{Can we always assume $k=3$ by allowing $A_i=\emptyset$? That would clean it up.}
+	\note{DW}{Can we always assume $k=3$ by allowing $A_i=\emptyset$? That would clean up the proof.}
 
@@
 	\item $R_{i,j}^+, R_{i+1,j}^-, S_{i,j}\in\PP_j$,
-	\item the quotient $G_j / \PP_j$ is planar and has a tree-decomposition $\mathcal{T}_j$ of width at most 3 such that: 
+	\item the quotient $G_j / \PP_j$ has a tree-decomposition $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3 such that: 
 	\begin{compactitem}
-		\item some bag $B_j$ of $\mathcal{T}_j$ contains $R_{i,j}^+, R_{i+1,j}^-, S_{i,j}$, and 
+		\item some bag of $\mathcal{T}_j$ contains $R_{i,j}^+, R_{i+1,j}^-, S_{i,j}$, and 
 		\item for every face $F$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
@@
 \end{compactitem}
-\end{lem}
 
@@
 We now construct the desired partition $\PP$ of $G$. 
-Initialise $\PP:= \{P_1,\ldots,P_k\}$. Then add each non-empty $Q_i'$ to $\PP$. 
-Now for each non-degenerate $F_i$, each path in  $\PP_i$ is either an \emph{external} path (that is, fully contained in $F_i$) or is an \emph{internal path} with none of its vertices in $F_i$. 
-Add all the internal paths of  $\PP_i$ to $\PP$. 
-By construction, $\PP$ partitions $V(G)$ into vertical paths in $T$ and $\PP$ contains $P_1,\ldots,P_k$.
-
-Let $H:=G/\PP$. Next we exhibit the desired tree-decomposition $(B_x:x\in V(J))$ of $H$. 
-Let $J$ be the tree obtained from the disjoint union of $J_i$, taken over the $i\in\{1,2,3\}$ such that $F_i$ is non-degenerate, by adding one new node $u$ adjacent to each $u_i$.
-(Recall that $u_i$ is the node of $J_i$ for which the bag $B^i_{u_i}$ contains the vertices of $H_i$ corresponding to the paths that form $F_i$.)\ 
-Let the bag $B_u$ contain all the vertices of $H$ corresponding to $P_1,\ldots,P_k,Q'_1,Q'_2,Q'_3$. 
-For each non-degenerate $F_i$, and for each node $x\in V(J_i)$, initialise $B_x:= B^i_x$. 
-Recall that vertices of $H_i$ correspond to contracted paths in  $\PP_i$.
-Each internal path in $\PP_i$ also lies  in $\PP$.
-Each external path $P$ in $\PP_i$ is a subpath of $P_j$ for some $j\in\{1,\dots,k \}$ or is one of the paths among $Q'_1, Q'_2, Q'_3$. 
-For each such path $P$, for every $x\in V(J)$, in bag $B_x$, 
-replace each instance of the vertex of $H_i$ corresponding to $P$ by the vertex of $H$ corresponding to the path among $P_1,\ldots, P_k, Q'_1,\ldots,Q'_3$ 
-that contains $P$.
-This completes the description of $(B_x : x\in V(J))$. 
-By construction, $|B_x|\leq 9$ for every $x\in V(J)$.
-
-First we show that for each vertex $a$ in $H$, the set $X:=\{x\in V(J) : a\in B_x\}$ forms a subtree of $J$.
-If $a$ corresponds to a path distinct from $P_1,\ldots,P_k,Q'_1,Q'_2,Q'_3$ then $X$ is fully contained in $J_i$ for some $i\in\{1,2,3\}$.
-Thus, by induction $X$ is non-empty and connected in $J_i$, so it is in $J$.
-If $a$ corresponds to $P$ which is one of the paths among $P_1,\ldots,P_k,Q'_1,Q'_2,Q'_3$ then 
-$u\in X$ and whenever $X$ contains a vertex of $J_i$ it is because some external path of $\PP_i$ was replaced by $P$.
-In particular, we would have $u_i \in X$ in that case. 
-Again by induction each $X\cap J_i$ is connected and since $uu_i\in E(T)$, we conclude that $X$ induces a (connected) subtree of $J$.
-
-Finally we show that, for every edge $ab$ of $H$, there is a bag $B_x$ that contains $a$ and $b$. 
-If $a$ and $b$ are both obtained by contracting any of $P_1,\ldots,P_k,Q'_1,Q'_2,Q'_3$, 
-then $a$ and $b$ both appear in $B_u$.  
-If $a$ and $b$ are both in $H_i$ for some $i\in\{1,2,3\}$, 
-then some bag $B^i_x$ contains both $a$ and $b$. 
-Finally, when $a$ is obtained by contracting a path $P_a$ in $G_i-V(F_i)$ and $b$ is obtained by contracting a path $P_b$ not in $G_i$, 
-then  the cycle $F_i$ separates $P_a$ from $P_b$ so the edge $ab$ is not present in $H$. 
-This concludes the proof that $(B_x:x\in V(J))$ is the desired tree-decomposition  of $H$.
-
-We construct $\PP$ as before.
+Let $\PP$ be obtained from $\bigcup_{j\in\{1,\dots,m\}} \PP_j$ by replacing 
+$R_{i,j}^+\in\PP_j$ by $P_i$, 
+$R_{i+1,j}^-\in\PP_j$ by $P_i$, and
+$S_{i,j}\in\PP_j$ by $Y$.
+By construction, $\PP$ partitions $V(G)$ into tripods with respect to $T$ and $P_1,\ldots,P_k\in \PP$.
+
+We now construct the desired tree-decomposition of $G/\PP$.
+For each $j\in\{1,\dots,m\}$, apply the above replacements within the bags of $\mathcal{T}_j$, and let $u_j$ be the node of $D_j$ such that $R_{i,j}^+, R_{i+1,j}^-, S_{i,j} \in B_{u_j}$.
+Let $D$ be the tree obtained from the disjoint union of $D_1,\dots,D_m$ by adding one new node $u$ adjacent to $u_1,\dots,u_m$.
+Define $B_u:=\{P_1,P_2,P_3,Y\}$.
+We claim that $\{B_x:x\in V(D)\}$ is the desired tree-decomposition of $G/\PP$.
+By construction, every bag $B_x$ has size at most 4.
+
+First we show that for each part $X\in\PP$, the set $\{x\in V(D) : X\in B_x\}$ induces a subtree of $D$.
+If $X\not\in \{P_1,P_2,P_3,Y\}$ then $X$ is only in bags in some $D_j$, implying $\{x\in V(D) : X\in B_x\}$ induces a subtree of $D$.
+If $X\in \{P_1,P_2,P_3,Y\}$ then $X$ is in $B_u$, and if some part $X'$ of some $\PP_j$ was replaced by $X$ in the above construction, then $X'$ is in $B_{u_j}$ and $\{x\in V(D_j):X\in B_x\}$ induces a subtree of $D_j$, implying that  $\{x\in V(D) : X\in B_x\}$ induces a subtree of $D$ (since $uu_j\in E(D)$).
+
+We now show that, for every edge $XW$ of $G/\PP$, there is a node $x\in V(D)$ such that $X,W\in B_x$.
+If $X,W\in \{P_1,P_2,P_3,Y\}$ then $X,W\in B_u$ as desired.
+Otherwise, without loss of generality, $X\in \PP_j\setminus \{P_1,P_2,P_3,Y\}$ for some $j\in\{1,\dots,m\}$.
+Say $X'\in\PP'_j$ replaced by $X\in\PP_j$.
+Thus $X\subseteq V(G_j)\setminus V(Z_j)$.
+Since $Z_j$ separates $V(G_j)\setminus V(Z_j)$ and $V(G)\setminus V(G_j)$ and $XW\in E(G/\PP)$, it must be that $W\in \PP_j$ also, implying that $XW\in B_x$ for some node $x\in V(D_j)$.
+This concludes the proof that $(B_x:x\in V(D))$ is the desired tree-decomposition  of $G/\PP$.
+
+Every face $F$ of $G$ is a face of some $G_j$ or is $F$ \note{DW}{Avoid using $F$ for two different things}.
+In the latter case, if $B$ is the bag at $u$ then $V(F)\subseteq \bigcup_{X\in B} X$ since $V(F) \subseteq Y\in B$.
+In the first case, there is a bag $B$ of $\mathcal{T}_j$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
+This property is maintained by the above replacements since
+$R_{i,j}^+ \subseteq P_i$, 
+$R_{i+1,j}^- \subseteq P_i$, and
+$S_{i,j}\subseteq Y$.
+
+%For each non-degenerate $F_i$, and for each node $x\in V(J_i)$, initialise $B_x:= B^i_x$. 
+%Recall that vertices of $H_i$ correspond to contracted paths in  $\PP_i$.
+%Each internal path in $\PP_i$ also lies  in $\PP$.
+%Each external path $P$ in $\PP_i$ is a subpath of $P_j$ for some $j\in\{1,\dots,k \}$ or is one of the paths among $Q'_1, Q'_2, Q'_3$. 
+%For each such path $P$, for every $x\in V(J)$, in bag $B_x$, 
+%replace each instance of the vertex of $H_i$ corresponding to $P$ by the vertex of $H$ corresponding to the path among $P_1,\ldots, P_k, Q'_1,\ldots,Q'_3$ 
+%that contains $P$.
+%This completes the description of $(B_x : x\in V(J))$. 
+%By construction, $|B_x|\leq 9$ for every $x\in V(J)$.
+
+%We construct $\PP$ as before.
 Initialise $\PP:=\{P_1,\ldots,P_k,Y\}$.
@@
 
-Next we produce the tree-decomposition $(B_x:x\in V(J))$ of $H$ that satisfies the requirements of the lemma.
+%Next we produce the tree-decomposition $(B_x:x\in V(J))$ of $H$ that satisfies the requirements of the lemma.
 Let $J$ be the tree obtained from the disjoint union of $J_1$, $J_2$ and $J_3$ by adding one new node $u$ adjacent to $u_1$, $u_2$ and $u_3$.
@@
 
-As in  the proof of \cref{NearTriang},  the resulting structure,  $(B_x:x\in V(J))$, is not yet a tree-decomposition of $H$ since some bags may  contain vertices of $H_i$ that are not necessarily vertices of $H$.
+%As in  the proof of \cref{NearTriang},  the resulting structure,  $(B_x:x\in V(J))$, is not yet a tree-decomposition of $H$ since some bags may  contain vertices of $H_i$ that are not necessarily vertices of $H$.
 Note that unlike in  \cref{NearTriang} this does not only include elements of $\PP_i$  that are contained in $F$.
 In particular, $I_i$ is also not an element of $\PP$ and thus does not correspond to a vertex of $H$.
-We remedy this  as follows. For $x\in V(J)$, in bag $B_x$, replace each instance of the vertex of $H_i$ corresponding to $I_i$ by the vertex of $H$ corresponding to $Y$. 
+We remedy this  as follows.
+For $x\in V(J)$, in bag $B_x$, replace each instance of the vertex of $H_i$ corresponding to $I_i$ by the vertex of $H$ corresponding to $Y$.
 Similarly, by construction, $R_i^+$ is a subgraph of $P_{\alpha_i}$ for some $\alpha_i\in \{1,\dots,k\}$.
-For $x\in V(J)$, in bag $B_x$, replace each instance of the vertex of $H_i$ corresponding to $R_i^+$ 
-by the vertex of $H$ corresponding to $P_{\alpha_i}$. 
+For $x\in V(J)$, in bag $B_x$, replace each instance of the vertex of $H_i$ corresponding to $R_i^+$  by the vertex of $H$ corresponding to $P_{\alpha_i}$.
 Finally, $R_{i+1}^-$ is a subgraph of $P_{\beta_i}$ for some $\beta_i\in \{1,\dots,k\}$.
@@
 \begin{lem}
-	An embedded multigraph $G$ is edge-maximal 1-plane with no faces of length 2 if and only if there is a 2-connected plane graph $G_0$ such that: 
+	An embedded multigraph $G$ is edge-maximal 1-plane with no faces of length 2 if and only if there is a 2-connected plane multigraph $G_0$ such that: 
 	\begin{compactitem}

### 2022-03-10 k-planar.tex
cleaned up  proof in appendix, still need to think more about when k \neq 3

@@
 For a cycle $C$, we write $C = [ P_1 , \dots , P_k ]$ if $P_1 , \dots, P_k$ are pairwise disjoint non-empty paths in $C$, and the endpoints of each path $P_i$ can be labelled $x_i$ and $y_i$ so that $y_ix_{i+1} \in E(C)$ for $i\in\{1,\dots,k\}$, where $x_{k+1}$ means
-$x_1$. This implies that $\{P_1,\dots,P_k\}$ is a vertex-partition of $C$.
+$x_1$. This implies that $\{P_1,\dots,P_k\}$ is a vertex-partition of $C$. \note{DW}{Allow $P_i$ to be empty?}
 
 If $Q$ is a path in a graph, and $P_1, \dots,P_k$ are edge-disjoint non-empty subpaths of $Q$ ordered by $Q$ and partitioning $E(Q)$, then we write $Q=P_1,\dots,P_k$.
+
+\note{DW}{It would help to say that we consider a path to be an (unordered) subgraph. Would that be a problem anywhere else?}
 
@@
 Let $T$ be a spanning tree of $G$ rooted at a vertex $r$ on the outerface of $G$.
-A \defin{tripod} (with respect to $T$) is any subgraph of $G$ consisting of three vertical paths $P,Q,R$ in $T$ plus a facial cycle $F$ of $G$,
-where the lower endpoint of each of $P,Q,R$ is in $F$. A \defin{bipod} (with respect to $T$) is any (connected) subpath of a tripod (with respect to $T$).
- %is a path $PXQ$ where $P$ and $Q$ are vertical paths in $T$ and $X$ is a facial path of $G$, where the lower endpoints of $P$ and $Q$ are the endpoints of $X$. 
-\note{DW}{Need to think about degenerate cases. We want every bipod to be a tripod}
-%, and every connected subgraph of a bipod to be a bipod.}.
-These definitions generalise definitions in \citep{DJMMUW20} where $G$ is assumed to be a plane triangulation, in which case $F$ is a 3-cycle and $X$ is an edge.
+A \defin{bipod} in $G$ (with respect to $T$) is any path $P_1SP_2$, where $S$ is a facial path in $G$, and each $P_i$ is a vertical path in $T$ whose lower endpoint is in $S$.
+Here it is possible that $S$ or $P_i$ is empty.
+So every subpath of a bipod is a bipod.
+A \defin{tripod} in $G$ (with respect to $T$) is either a bipod in $G$ or a subgraph of $G$ consisting of a facial cycle $F$ of $G$, plus three distinct vertical paths $P_1,P_2,P_3$ in $T$ where the lower endpoint of each $P_i$ is in $F$.
+\note{DW}{Need to think about degenerate cases.}.
+These definitions generalise definitions in \citep{DJMMUW20} where $G$ is assumed to be a plane triangulation, in which case $F$ is a 3-cycle.
 \note{DW}{We actually managed to omit the definition of `bipod' in \citep{DJMMUW20}!}
@@
 \begin{lem}
-Say $G_0$ is a 2-connected plane multigraph.
+Let $G_0$ be a 2-connected plane multigraph.
 Let $T$ be a spanning tree of $G_0$ rooted at some vertex $r$ on the boundary of the outerface of $G_0$.
-Let $k\in\{1,2,3\}$ and let $C = [A_1,\dots,A_k]$ be a cycle in $G_0$ with $r$ in the exterior of $C$, where $A_1, \dots, A_k$ are bipods in $G_0$.
-Let $G$ be the plane graph consisting of all the edges and vertices of $G_0$ in $C$ and the interior of $C$.
-Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that:
+Let $k\in\{1,2,3\}$ and let $C = [P_1,\dots,P_k]$ be a cycle in $G_0$ with $r$ in the exterior of $C$, where $P_1, \dots, P_k$ are bipods in $G_0$.
+	\note{DW}{We need to assume that $|V(C)|\geq 3$ and ensure that when we apply induction $|V(Z_j)|\geq 3$. I think it is okay, but it needs an argument.}
+	Let $G$ be the plane graph consisting of all the edges and vertices of $G_0$ in $C$ and the interior of $C$. Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that:
 \begin{compactitem}
@@
 	\item some bag of $\mathcal{T}$ contains $P_1,\dots,P_k$, and 
-	\item for every face $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
+	\item for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
 \end{compactitem}
@@
 \begin{proof}
-	\note{DW}{Can we always assume $k=3$ by allowing $A_i=\emptyset$? That would clean up the proof.}
-
-	\note{DW}{Is $C$ allowed to be a 2-cycle?}
+\note{DW}{Can we assume $k=3$ by allowing $P_i=\emptyset$? That would clean up the proof. We could scrap $R_1,R_2,R_3$ and just use $P_1,P_2,P_3$.}
+
 The proof is by induction on $|V(G)\setminus V(C)|$.
 If $V(G)\setminus V(C)=\emptyset$, then $G$ has just one face bounded by $C$ and the desired partition is $\PP=\{P_1,\ldots,P_k\}$.
-The tree-decomposition of $G/\PP$ consists of a single bag that contains the $k\le 3$ vertices corresponding to $P_1,\ldots,P_k$.
+The desired tree-decomposition of $G/\PP$ consists of a single bag $\{P_1,\ldots,P_k\}$ indexed by a 1-node tree.
 
@@
 \begin{compactenum}
-	\item If $k=1$ then, since $C$ is a cycle, $P_1$ has at least three vertices, so $P_1=[v, P_1', w]$ for two distinct vertices $v$ and $w$.
-\note{DW}{need to be careful with parallel edges} We set $R_1:=v$, $R_2:=P_1'$ and $R_3:=w$.
-
-	\item If $k=2$ then we may assume without loss of generality that $P_1$ has at least two vertices so $P_1=[v,P_1']$ \note{DW}{need to be careful with parallel edges}.
-We set $R_1:=v$, $R_2:=P_1'$ and $R_3:=P_2$.
+	\item If $k=1$ then, since $C$ is a cycle, $P_1$ has at least three vertices, so $P_1=[v, P_1', w]$ for two distinct vertices $v$ and $w$. We set $R_1:=v$, $R_2:=P_1'$ and $R_3:=w$.
+
+	\item If $k=2$ then we may assume without loss of generality that $P_1$ has at least two vertices so $P_1=[v,P_1']$. We set $R_1:=v$, $R_2:=P_1'$ and $R_3:=P_2$.
 
@@
 Since $r$ is on the outer-face of $G_0$, $P_v$ contains at least one vertex of $C$.
-If the first vertex of $P_v$ that belongs to $F$ is in $R_i$ then assign the colour $i$ to $v$.
+If the first vertex of $P_v$ that belongs to $F$ is in $R_i$, then assign the colour $i$ to $v$.
 We obtain a 3-colouring of the vertices of $G'$ that satisfies the conditions of Sperner's Lemma.
 By Sperner's Lemma there exists a triangular face $\tau=v_1v_2v_3$ of $G'$ whose vertices are coloured $1,2,3$ respectively.
-Let $F$ be the face of $G_0$ containing $\tau$.
+Let $F$ be the cycle bounding the face of $G$ containing $\tau$.
 
@@
 
-Let $Y$ be the tripod consisting of $Q_1'\cup Q_2' \cup Q_3'$ plus the vertices and edges of $F$.
-
-Say $F=v_1 S_1 v_2 S_2 v_3 S_3$, where $S_i$ is a facial path in $F$ clockwise from $v_i$ to $v_{i+1}$.
-
-Let $L_i$ be the path $Q_i'S_iQ_{i+1}'$ for $i\in\{1,\dots,k\}$. So $L_i$ is a bipod.
-
-, where we use the convention that $Q_4=Q_1$ and $R_4=R_1$.
+Let $Y$ be the tripod $F\cup Q_1'\cup Q_2' \cup Q_3'$.
+Say $F=v_1 S_1 v_2 S_2 v_3 S_3$, where $S_i$ is a facial path in $F$ clockwise from $v_i$ to $v_{i+1}$ (where $v_4=v_1$).
+Let $L_i$ be the path $Q_i'S_iQ_{i+1}'$ for $i\in\{1,2,3\}$ (where $Q_4=Q_1$ and $R_4=R_1$).
+So $L_i$ is a bipod.
+\note{DW}{Note that $S_i\neq\emptyset$ since $v_i\neq v_{i+1}$.}
 
@@
 
-Consider the plane graph $Z$ consisting of the edges and vertices of $F$, the edges and vertices of $C$, and the edges and vertices of $Q_1,Q_2,Q_3$.
+Let $Z$ be the plane graph $F\cup C\cup Q_1\cup Q_2\cup Q_3$.
 Then $Z$ is 2-connected, each face of $Z$ is bounded by a cycle, and the outerface of $Z$ is $C$.
-Let $Z_1,\dots,Z_m$ be the internal faces of $Z$.
+Let $Z_1,\dots,Z_m$ be the internal facial cycles of $Z$ except $F$.
+
+\note{DW}{For a cycle $C$ in a plane graph $G$, define $G\langle C\rangle$ to be the subgraph of $G$ consisting of $C$ and all the edges and vertices of $G$ in the interior of $C$?}
 
 Consider $j\in\{1,\dots,m\}$.
-Let $G_j$ be the subgraph of $G$ consisting of all the edges and vertices of $G$ in $Z_j$ and the interior of $Z_j$.
+Let $G_j$ be the subgraph of $G$ consisting of $Z_j$ and all the edges and vertices of $G$ in the interior of $Z_j$.
 The outerface of $G_j$ is $Z_j$.
-Observe that for some $i\in\{1,\dots,k\}$, we have
+Observe that for some $i\in\{1,2,3\}$, we have
 $Z_j=[R_{i,j}^+,R_{i+1,j}^-,S_{i,j}]$, where 
@@
 $R_{i+1,j}^-$ is some subpath of $R_{i+1}^-$, and
-$S_{i,j}$ is some subpath of $S_i$.
-Since $R_i^+$, $R_{i+1}^-$ and $S_i$ are bipods, 
-$R_{i,j}^+$, $R_{i+1,j}^-$ and $S_{i,j}$ are bipods. 
-By construction, $v_{i-1}$ is not in $G_j$. 
+$L_{i,j}$ is the path $Z_j-V(C)$ (which is contained in $Y$). 
+Since $R_i^+$, $R_{i+1}^-$ and $L_i$ are bipods, 
+$R_{i,j}^+$, $R_{i+1,j}^-$ and $L_{i,j}$ are bipods. 
+By construction, $v_{i-1}$ is not in $G_j$ (where $v_0=v_3$). 
 Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.  Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that:
 \begin{compactitem}
-	\item $R_{i,j}^+, R_{i+1,j}^-, S_{i,j}\in\PP_j$,
+	\item $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$, 
 	\item the quotient $G_j / \PP_j$ has a tree-decomposition $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3 such that: 
 	\begin{compactitem}
-		\item some bag of $\mathcal{T}_j$ contains $R_{i,j}^+, R_{i+1,j}^-, S_{i,j}$, and 
-		\item for every face $F$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
+		\item some bag $B_{u_j}$ of $\mathcal{T}_j$ contains $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}$, and 
+		\item for every facial cycle $F'$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that $V(F')\subseteq \bigcup_{X\in B} X$.
 	\end{compactitem}
@@
 
-%This graph has up to three \note{DW}{can be more} internal faces $C_1,C_2,C_3$ where each $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ and $R_{i}^+$ and $R_{i}^-$ are the same portions of $R_i$ as defined in \cref{NearTriang} \note{DW}{move in here}    Observe that  $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$, where $R_i^+$ and $R_{i+1}^-$ are bipods, and $I_i$ is the bipod $Q_i' S_i Q_{i+1}'$.
+%This graph has up to three \note{DW}{can be more} internal faces $C_1,C_2,C_3$ where each $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ and $R_{i}^+$ and $R_{i}^-$ are the same portions of $R_i$ as defined in \cref{NearTriang} \note{DW}{move in here}    Observe that  $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$, where $R_i^+$ and $R_{i+1}^-$ are bipods, and $I_i$ is the bipod $Q_i' L_i Q_{i+1}'$.
 Let $G_i$ be the subgraph of $G$ whose vertices and edges are in $C_i$ or its interior.
@@
 We now construct the desired partition $\PP$ of $G$. 
-Let $\PP$ be obtained from $\bigcup_{j\in\{1,\dots,m\}} \PP_j$ by replacing 
-$R_{i,j}^+\in\PP_j$ by $P_i$, 
-$R_{i+1,j}^-\in\PP_j$ by $P_i$, and
-$S_{i,j}\in\PP_j$ by $Y$.
+Let $\PP$ be obtained from $\bigcup_{j\in\{1,\dots,m\}} \PP_j$ by 
+replacing $R_{i,j}^+\in\PP_j$ by $P_i$, 
+replacing $R_{i+1,j}^-\in\PP_j$ by $P_{i+1}$, and
+replacing $L_{i,j}\in\PP_j$ by $Y$. \note{DW}{What if $k\neq 3$? Then $P_2$ or $P_3$ might not be defined.}
 By construction, $\PP$ partitions $V(G)$ into tripods with respect to $T$ and $P_1,\ldots,P_k\in \PP$.
@@
 We now construct the desired tree-decomposition of $G/\PP$.
-For each $j\in\{1,\dots,m\}$, apply the above replacements within the bags of $\mathcal{T}_j$, and let $u_j$ be the node of $D_j$ such that $R_{i,j}^+, R_{i+1,j}^-, S_{i,j} \in B_{u_j}$.
 Let $D$ be the tree obtained from the disjoint union of $D_1,\dots,D_m$ by adding one new node $u$ adjacent to $u_1,\dots,u_m$.
-Define $B_u:=\{P_1,P_2,P_3,Y\}$.
-We claim that $\{B_x:x\in V(D)\}$ is the desired tree-decomposition of $G/\PP$.
-By construction, every bag $B_x$ has size at most 4.
+Define $B_u:=\{P_1,\dots,P_k,Y\}$.
+Apply the above replacements within the bags of $\mathcal{T}_1,\dots,\mathcal{T}_m$.
+We claim that $\mathcal{T}:=\{B_x:x\in V(D)\}$ is the desired tree-decomposition of $G/\PP$.
+By construction, every bag of $\mathcal{T}$ has size at most 4.
 
@@
 If $X\not\in \{P_1,P_2,P_3,Y\}$ then $X$ is only in bags in some $D_j$, implying $\{x\in V(D) : X\in B_x\}$ induces a subtree of $D$.
-If $X\in \{P_1,P_2,P_3,Y\}$ then $X$ is in $B_u$, and if some part $X'$ of some $\PP_j$ was replaced by $X$ in the above construction, then $X'$ is in $B_{u_j}$ and $\{x\in V(D_j):X\in B_x\}$ induces a subtree of $D_j$, implying that  $\{x\in V(D) : X\in B_x\}$ induces a subtree of $D$ (since $uu_j\in E(D)$).
+If $X\in \{P_1,P_2,P_3,Y\}$ then $X$ is in $B_u$, and if some part $X'$ of some $\PP_j$ was replaced by $X$ in the above construction, then $X'$ is in $B_{u_j}$ and $\{x\in V(D_j):X'\in B_x\}$ induces a subtree of $D_j$ (since $\mathcal{T}_j$ is a tree-decomposition of $G_j$).
+Thus $\{x\in V(D) : X\in B_x\}$ induces a subtree of $D$ (since $uu_j\in E(D)$).
 
@@
 Otherwise, without loss of generality, $X\in \PP_j\setminus \{P_1,P_2,P_3,Y\}$ for some $j\in\{1,\dots,m\}$.
-Say $X'\in\PP'_j$ replaced by $X\in\PP_j$.
 Thus $X\subseteq V(G_j)\setminus V(Z_j)$.
-Since $Z_j$ separates $V(G_j)\setminus V(Z_j)$ and $V(G)\setminus V(G_j)$ and $XW\in E(G/\PP)$, it must be that $W\in \PP_j$ also, implying that $XW\in B_x$ for some node $x\in V(D_j)$.
+Since $Z_j$ separates $V(G_j)\setminus V(Z_j)$ and $V(G)\setminus V(G_j)$ and $XW\in E(G/\PP)$, it must be that $W$ is also in $\PP_j$.
+Say $W$ replaced $W'$ in $\PP_j$ (where possibly $W=W'$).
+Since $\mathcal{T}_j$ is a tree-decomposition of $G_j$, $XW'\in B_x$ for some node $x\in V(D_j)$, implying $XW\in B_x$.
 This concludes the proof that $(B_x:x\in V(D))$ is the desired tree-decomposition of $G/\PP$.
 
-Every face $F$ of $G$ is a face of some $G_j$ or is $F$ \note{DW}{Avoid using $F$ for two different things}.
-In the latter case, if $B$ is the bag at $u$ then $V(F)\subseteq \bigcup_{X\in B} X$ since $V(F) \subseteq Y\in B$.
-In the first case, there is a bag $B$ of $\mathcal{T}_j$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
+Every facial cycle $F_0$ of $G$ is a face of some $G_j$ or is $F$.
+In the latter case, $V(F)\subseteq Y \subseteq \bigcup_{X\in B_u} X$.
+In the first case, there is a bag $B_x$ of $\mathcal{T}_j$ such that $V(F)\subseteq \bigcup_{X\in B_x} X$.
 This property is maintained by the above replacements since
 $R_{i,j}^+ \subseteq P_i$, 
-$R_{i+1,j}^- \subseteq P_i$, and
-$S_{i,j}\subseteq Y$.
+$R_{i+1,j}^- \subseteq P_{i+1}$, and
+$L_{i,j}\subseteq Y$. \note{DW}{What if $k\neq 3$?}
 
@@
 
-This completes the description of $(B_x:x\in V(J))$. Clearly, every bag $B_x$ has size at most 4. We now prove that $(B_x:x\in V(J))$  is a tree-decomposition of $H$ 
+This shows that $(B_x:x\in V(J))$ is the desired tree-decomposition of $G/\PP$. 
 \end{proof}

### 2022-03-10 k-planar.tex
Main theorem in appendix is now proved, subject to  a tweak to the main lemma

@@
 \DeclareMathOperator{\qn}{qn}
+
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
 \DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
+\DeclarePairedDelimiter{\blah}{\langle}{\rangle}
 
@@
 
+\newcommand{\TT}{\mathcal{T}}
 \newcommand{\PP}{\mathcal{P}}
@@
 
-A \defin{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle$ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
+A \defin{layering} of a graph $G$ is a sequence $\mathcal{L}=(L_0,L_1,\ldots)$ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
 For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
@@
 
-\note{DW}{Work in progress}
-
 For a plane multigraph $G$, let $\widehat{G}$ be the embedded graph obtained from $G$ as follows: for each face $f$ of $G$ and for each pair of non-consecutive vertices $v$ and $w$ on the boundary of $f$, add the edge $vw$ to $\widehat{G}$ drawn across $f$.
@@
 
-For a cycle $C$, we write $C = [ P_1 , \dots , P_k ]$ if $P_1 , \dots, P_k$ are pairwise disjoint non-empty paths in $C$, and the endpoints of each path $P_i$ can be labelled $x_i$ and $y_i$ so that $y_ix_{i+1} \in E(C)$ for $i\in\{1,\dots,k\}$, where $x_{k+1}$ means
-$x_1$. This implies that $\{P_1,\dots,P_k\}$ is a vertex-partition of $C$. \note{DW}{Allow $P_i$ to be empty?}
-
-If $Q$ is a path in a graph, and $P_1, \dots,P_k$ are edge-disjoint non-empty subpaths of $Q$ ordered by $Q$ and partitioning $E(Q)$, then we write $Q=P_1,\dots,P_k$.
-
-\note{DW}{It would help to say that we consider a path to be an (unordered) subgraph. Would that be a problem anywhere else?}
+For a cycle $C$, we write $C = [ P_1 , \dots , P_k ]$ if $P_1 , \dots, P_k$ are pairwise disjoint (possibly empty) path subgraphs in $C$ such that $V(C)=\bigcup_i V(P_i)$ and $P_1,\dots,P_k$ are ordered around $C$.
+
+%and the endpoints of each path $P_i$ can be labelled $x_i$ and $y_i$ so that $y_ix_{i+1} \in E(C)$ for $i\in\{1,\dots,k\}$, where $x_{k+1}$ means $x_1$.
+This implies that $\{P_1,\dots,P_k\}$ is a vertex-partition of $C$.
+\note{DW}{In the proof below, when we apply induction, we need to allow $P_i=\emptyset$, right?}
+
+If $Q$ is a path in a graph, and $P_1, \dots,P_k$ are edge-disjoint non-empty subpaths of $Q$ ordered by $Q$ and partitioning $E(Q)$, then we write $Q=P_1,\dots,P_k$. \note{DW}{Do we need this?}
 
@@
 Let $T$ be a spanning tree of $G$ rooted at a vertex $r$ on the outerface of $G$.
-A \defin{bipod} in $G$ (with respect to $T$) is any path $P_1SP_2$, where $S$ is a facial path in $G$, and each $P_i$ is a vertical path in $T$ whose lower endpoint is in $S$.
+A \defin{bipod} in $G$ (with respect to $T$) is any path $P_1SP_2$ (considered to be a subgraph), where $S$ is a facial path in $G$, and each $P_i$ is a vertical path in $T$ whose lower endpoint is in $S$.
 Here it is possible that $S$ or $P_i$ is empty.
@@
 A \defin{tripod} in $G$ (with respect to $T$) is either a bipod in $G$ or a subgraph of $G$ consisting of a facial cycle $F$ of $G$, plus three distinct vertical paths $P_1,P_2,P_3$ in $T$ where the lower endpoint of each $P_i$ is in $F$.
-\note{DW}{Need to think about degenerate cases.}.
 These definitions generalise definitions in \citep{DJMMUW20} where $G$ is assumed to be a plane triangulation, in which case $F$ is a 3-cycle.
@@
 
+For a cycle $C$ in a plane multigraph $G$, define \defin{$G\blah{C}$} to be the subgraph of $G$ consisting of $C$ and all the edges and vertices of $G$ in the interior of $C$.
+
 \begin{lem}
+	\label{NewLemma}
 Let $G_0$ be a 2-connected plane multigraph.
 Let $T$ be a spanning tree of $G_0$ rooted at some vertex $r$ on the boundary of the outerface of $G_0$.
-Let $k\in\{1,2,3\}$ and let $C = [P_1,\dots,P_k]$ be a cycle in $G_0$ with $r$ in the exterior of $C$, where $P_1, \dots, P_k$ are bipods in $G_0$.
-	\note{DW}{We need to assume that $|V(C)|\geq 3$ and ensure that when we apply induction $|V(Z_j)|\geq 3$. I think it is okay, but it needs an argument.}
-	Let $G$ be the plane graph consisting of all the edges and vertices of $G_0$ in $C$ and the interior of $C$. Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that:
+Let $C = [P_1,\dots,P_k]$ be a cycle in $G_0$ with $r$ in the exterior of $C$, where $k\in\{1,2,3\}$ and $P_1, \dots, P_k$ are bipods in $G_0$.
+Let $G:=G_0\blah{C}$.
+Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that:
 \begin{compactitem}
@@
 \item $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
-\begin{compactitem}
+\begin{compactenum}[(i)]
 	\item some bag of $\mathcal{T}$ contains $P_1,\dots,P_k$, and 
 	\item for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
-\end{compactitem}
+\end{compactenum}
+\note{DW}{We need to replace (ii) by ``for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$''.
+Curiously, this implies (i) (since $C$ is a facial cycle of $G$).}
 \end{compactitem}
@@
 \begin{proof}
-\note{DW}{Can we assume $k=3$ by allowing $P_i=\emptyset$? That would clean up the proof. We could scrap $R_1,R_2,R_3$ and just use $P_1,P_2,P_3$.}
-
-The proof is by induction on $|V(G)\setminus V(C)|$.
-If $V(G)\setminus V(C)=\emptyset$, then $G$ has just one face bounded by $C$ and the desired partition is $\PP=\{P_1,\ldots,P_k\}$.
-The desired tree-decomposition of $G/\PP$ consists of a single bag $\{P_1,\ldots,P_k\}$ indexed by a 1-node tree.
-
-Now assume that $V(G)\setminus V(C)\neq\emptyset$. Partition the vertices of $C$ into paths $R_1$, $R_2$, and $R_3$ as follows:
+We may assume that each of $P_1,\dots,P_k$ is non-empty (since if some $P_i$ is empty, then we just consider $C=[P_1,\dots,P_{i-1},P_{i+1},\dots,P_k]$).
+
+The proof is by induction on the number of faces of $G$.
+If $G$ has just one internal face (which is bounded by $C$), then the desired partition is $\PP=\{P_1,\ldots,P_k\}$ and the desired tree-decomposition of $G/\PP$ consists of a single bag $\{P_1,\ldots,P_k\}$ indexed by a 1-node tree.
+Now assume that $G$ has at least two internal faces.
+
+First suppose that $C$ is a 2-cycle.
+Thus $k\leq 2$.
+Let $e$ be one of the edges in $C$.
+Let $F$ be the internal facial cycle of $G$ that includes $e$.
+Since each facial cycle of $G_0$ has length at least 3, $F$ consists of a non-empty path $P$ plus the edge $e$.
+Let $C'$ be obtained by replacing $e$ in $C$ by $P$.
+Let $G':=G\langle{C'}\rangle$.
+Thus $G'$ has one fewer face than $G$.
+Note that $C'=[P_1,\dots,P_k,P]$ (\note{DW}{or it can be written that way?}).
+By induction, $G'$ has a partition $\PP$ into tripods with respect to $T$ such that:
+\begin{compactitem}
+	\item $P_1,\dots,P_k,P\in\PP$,
+	\item $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
+	\begin{compactenum}[(i)]
+	\item some bag of $\mathcal{T}$ contains $P_1,\dots,P_k,P$, and 
+	\item for every facial cycle $F'$ of $G\blah{C'}$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
+	\end{compactenum}
+\end{compactitem}
+The bag of $\mathcal{T}$ that contains $P_1,\dots,P_k,P$ satisfies property (ii) for $F$. 
+Thus, $\PP$ and $\TT$ satisfy the desired properties for $G$. Now assume that $|V(C)|\geq 3$.
+
+%\note{DW}{One option is to do induction on $|V(G)\setminus V(C)|-k$ and reduce to the $k=3$ case (similar to the $|V(C)|\leq 2$ case).
+When $k<3$ we can grab another face $F$ on the inside of $C$ and make the boundary of $F$ a new path.
+Then the main proof will only deal with the $k=3$ case.
+Is this any easier?}
+
+%Now suppose that $k=1$.
+Let $v$ and $w$ be the first and last vertex in $P_1$.
+So $C$ consists of $P_1$ plus the edge $vw$.
+Let $F$ be the internal facial cycle of $G$ containing $vw$..........
+Now suppose that $k=2$ .....
+
+Partition the vertices of $C$ into paths $R_1$, $R_2$, and $R_3$ as follows:
 \begin{compactenum}
-	\item If $k=1$ then, since $C$ is a cycle, $P_1$ has at least three vertices, so $P_1=[v, P_1', w]$ for two distinct vertices $v$ and $w$. We set $R_1:=v$, $R_2:=P_1'$ and $R_3:=w$.
+	\item If $k=1$ then, $P_1$ has at least three vertices (since $|V(C)|\geq 3$), so $P_1=[v, P_1', w]$ for two distinct vertices $v$ and $w$. We set $R_1:=v$, $R_2:=P_1'$ and $R_3:=w$.
 
@@
 \end{compactenum}
+
+Note that $R_i \subseteq P_{\alpha(i)}$ for some $\alpha(i)\in \{1,\dots,k\}$.
 
@@
 So $L_i$ is a bipod.
-\note{DW}{Note that $S_i\neq\emptyset$ since $v_i\neq v_{i+1}$.}
+\note{DW}{Note that $S_i\neq\emptyset$ since $v_i\neq v_{i+1}$.
+Is this useful.}
 
@@
 
-\note{DW}{For a cycle $C$ in a plane graph $G$, define $G\langle C\rangle$ to be the subgraph of $G$ consisting of $C$ and all the edges and vertices of $G$ in the interior of $C$?}
-
 Consider $j\in\{1,\dots,m\}$.
@@
 $R_{i+1,j}^-$ is some subpath of $R_{i+1}^-$, and
-$L_{i,j}$ is the path $Z_j-V(C)$ (which is contained in $Y$). 
+$L_{i,j}$ is the path $Z_j-V(C)$ (which is contained in $Y$). \note{DW}{Add a figure of a good case and a weird case.}
 Since $R_i^+$, $R_{i+1}^-$ and $L_i$ are bipods, 
@@
 Let $\PP$ be obtained from $\bigcup_{j\in\{1,\dots,m\}} \PP_j$ by 
-replacing $R_{i,j}^+\in\PP_j$ by $P_i$, 
-replacing $R_{i+1,j}^-\in\PP_j$ by $P_{i+1}$, and
-replacing $L_{i,j}\in\PP_j$ by $Y$. \note{DW}{What if $k\neq 3$? Then $P_2$ or $P_3$ might not be defined.}
+replacing $R_{i,j}^+\in\PP_j$ by $P_{\alpha(i)}$, 
+replacing $R_{i+1,j}^-\in\PP_j$ by $P_{\alpha(i+1)}$, and
+replacing $L_{i,j}\in\PP_j$ by $Y$. 
 By construction, $\PP$ partitions $V(G)$ into tripods with respect to $T$ and $P_1,\ldots,P_k\in \PP$.
@@
 This property is maintained by the above replacements since
-$R_{i,j}^+ \subseteq P_i$, 
-$R_{i+1,j}^- \subseteq P_{i+1}$, and
-$L_{i,j}\subseteq Y$. \note{DW}{What if $k\neq 3$?}
+$R_{i,j}^+ \subseteq P_{\alpha(i)}$ and 
+$R_{i+1,j}^- \subseteq P_{\alpha(i+1)}$ and
+$L_{i,j}\subseteq Y$.
 
@@
 
+\begin{proof}[Proof of \cref{CliqueFaceProductStructure}] 
+Let $G$ be a 2-connected plane multigraph, in which every facial cycle has length in $\{3,\dots,d\}$.
+Let $C$ be the outer-cycle of $G$.
+Let $G^+$ be obtained from $G$ by adding a new vertex $r$ adjacent to every vertex on the outerface of $G$.
+So $G^+$ is a 2-connected plane multigraph and $G=G^+\blah{C}$.
+Let $T$ be a BFS spanning tree of $G^+$ rooted at $r$.
+Let $\mathcal{L}$ be the corresponding BFS layering.
+By \cref{NewLemma} \note{DW}{with improved (ii)}, $G$ has a partition $\PP$ into tripods with respect to $T$ such that $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
+Since $G$ is a spanning subgraph of $\widehat{G}$, we may consider $\PP$ to be a partition of $\widehat{G}$, and $G/\PP \subseteq \widehat{G}/\PP$.
+Consider an edge $XY$ of $\widehat{G}/\PP$.
+By the definition of $\widehat{G}$, there is a facial cycle $F$ of $G$ such that both $X$ and $Y$ intersect $F$.
+By assumption, there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
+In particular, $X,Y\in B$.
+Hence $\TT$ is also a tree-decomposition of $\widehat{G}$.
+Each tripod in $\PP$ has at most $d$ vertices in each layer in $\mathcal{L}$.
+For each edge $vw$ of $\widehat{G}$, if $v\in L_i$ and $w\in L_j$ then $|i-j|\leq\floor{\frac{d}{2}}$.
+Grouping blocks of $\floor{\frac{d}{2}}$ consecutive layers gives a layering of $\widehat{H}$ with layered width $\floor{\frac{d}{2}}d$ with respect to $\PP$.
+By \cref{PartitionProduct}, $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d  \floor{d/2} }$ where $H:=\widehat{G}/\PP$, which has treewidth 3.
+\end{proof}
+
+We now give two graph classes for which \cref{CliqueFaceProductStructure} is applicable.
+
 \begin{lem}
@@
 
+\begin{lem}
+Every $d$-map graph is a subgraph of $\widehat{G}$ for some 2-connected plane graph $G$ in which every facial cycle has length in $\{3,\dots,d\}$.
+\end{lem}
+
+\begin{proof}
+\end{proof}
+
 \end{document}
@@
 %%%%%%%%%%%%%%%%%%%%
-
-I claim that G_2 is contained in H * P * K_{ d * floor(d/2) } where tw(H) <= 3.
-
-With d=4, G_2 includes all edge-maximal 1-planar  multigraphs,
-and G_2 is contained in H * P * K_8 where tw(H) <= 3.
-
-Let T be a bfs spanning tree of G_0 rooted on the outerface.
-
-Run the original proof, finding three paths Q_1,Q_2,Q_3
-leading to a Sperner triangle t in G_1, which is a subset of a face F of G_0.
-The corresponding tripod in our partition is Q_1 \cup Q_2 \cup Q_3 \cup F.
-We can apply induction inside the regions bounded by C, Q_1, Q_2, Q_3 and F.
 
@@
 
-This setup (if it works) also works for map graphs in which every face
-is a nation.
-
-I just thought we need to be careful with the vertex 3-colouring.
-If v is on the inside of C and the path in T from v to the root
-first hits C at some B_i, then what is the colour of v?
-This is not a problem there is a noose in G_0 including A_1..A_k, so
-the path in T from v to the root must intersect A_1...A_k.
-
 \end{document}
-
-	\begin{figure}[!b]
-	\begin{center}
-		\begin{tabular}{cc}
-			\includegraphics[scale=1.4]{induction-i-3} &
-			\includegraphics[scale=1.4]{induction-i-1} \\
-			(a) & (b) \\[1em]
-			\includegraphics[scale=1.4]{induction-i-2} &
-			\includegraphics[scale=1.4]{induction-iii} \\
-			(c) & (d) 
-		\end{tabular}
-	\end{center}
-	\caption{The inductive proof of \cref{NearTriang}: (a)~the spanning tree $T$ and the paths $P_1,\ldots,P_4$; 
-		(b) the paths $R_1$, $R_2$, $R_3$, and the Sperner triangle $\tau$; 
-		(c)~the paths $Q_1'$, $Q_2'$ and $Q_3'$; 
-		(d)~the near-triangulations $G_1$, $G_2$, and $G_3$, with the vertical paths of $T$ on $F_1$, $F_2$, and $F_3$.}
-	\label{NearTriangProof}
-\end{figure}

### 2022-03-10 k-planar.tex
Suggested response to referees about 1-planarity.

@@
 
+\note{PM}{Include in note to referee:  We choose to keep this section because
+  $1$-planar graphs are of special interest to the graph drawing community.
+For example, the annotated bibliography on $1$-planarity (\htmladdnormallink{https://arxiv.org/abs/1703.02261}{arXiv:1703.02261}) has 143 entries, and even this is now four years out of date.
+
+  An important aspect of our result for $1$-planar graphs is that graph $H$ in the product is planar and has treewidth $3$, i.e., it has simple treewidth $3$.
+In a more recent application of product structure \citet{BDJM} show that such graphs have $\ell$-vertex rankings using $O(\log n/\log\log\log n)$ colours, and this is tight.
+This result requires that $H$ have simple treewidth $3$, even treewidth $3$ is not sufficient.
+
+  Finally, while cleaning up this proof we have also improved the constant from $30$ to $7$.
+The proof is quite similar to the previous proof except that graph $G'$ is obtained by removing both crossing edges involved in a crossing (the spars of a kite), rather than just one.
+We thank the referee for forcing us to write this proof more carefully so that we could discover this improvement.
+}
+
 \begin{thm}
 \label{1-planar}
-Every 1-planar graph is contained in $H\boxtimes P\boxtimes K_{30}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+Every 1-planar graph is contained in $H\boxtimes P\boxtimes K_{7}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
 \end{thm}
@@
 
-Therefore each face of $G'$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
+Therefore each face of $G_0$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
 \begin{inparaenum}[(i)]
@@
 
-By removing both spars from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three edges or $4$ edges; see \cref{one_planar_example}(b).
+By removing both spars from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
 Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
@@
 
-\referee{2}{Page 13.  Regarding the proof of Lemma 4, I think it is better to
-include all the details from reference [14].  As far as I know, JCTB
-does not have a page limit, so I do not see an issue with just
-reproducing the entire proof and telling the reader that they can skip
-all the details if they wish.  This is just my personal opinion
-though, so the authors can ignore this request if they choose.}
-
-\note{DW}{I agree we need more details. I am having trouble verifying this proof. }
+% \referee{2}{Page 13.  Regarding the proof of Lemma 4, I think it is better to
+% include all the details from reference [14].  As far as I know, JCTB
+% does not have a page limit, so I do not see an issue with just
+% reproducing the entire proof and telling the reader that they can skip
+% all the details if they wish.  This is just my personal opinion
+% though, so the authors can ignore this request if they choose.}
+% 
+% \note{DW}{I agree we need more details. I am having trouble verifying this proof. }
 

### 2022-03-10 k-planar.tex
Small fries

@@
 
-\referee{1}{3.
+% \referee{1}{3.
 Theorem 10: Unfortunately I have to suggest removing it.
@@
 Is there another reason for including it?}
-
-\note{DW}{We can motivate \cref{1-planar} since the graph drawing community particularly cares about 1-planar graphs, since treewidth 3 is best possible, and since the planarity of $H$ is essential in the proof of \cref{1PlanarQueue}.
+% 
+% \note{DW}{We can motivate \cref{1-planar} since the graph drawing community particularly cares about 1-planar graphs, since treewidth 3 is best possible, and since the planarity of $H$ is essential in the proof of \cref{1PlanarQueue}.
 Futher motivation comes from the open problem at the end of the paper.
@@
 
-\note{PM}{I agree that we should fix it and keep it.
+% \note{PM}{I agree that we should fix it and keep it.
 Another reason: For vertex ranking, the bound is $O(\log n/\log^{(t)} n)$ where $t:=\tw(H)$, so this is an example where $\tw(H)$ affects more than just the constant.}

### 2022-03-10 k-planar.tex
Some stuff

@@
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+
+\usepackage{todonotes}
 
@@
 
-Let $G$ be a $1$-plane multigraph with no two parallel edges on the boundary of a single face.
-By Euler's Formula, $G$ has $O(|V(G)|)$ vertices.
-We further assume that $G$ is edge-maximal, in the following sense:  if two vertices $v$ and $w$ of $G$ appear on a common face $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.
-See  \cref{one_planar_example}(b) for an example.
-Any $1$-plane multigraph $G$ with at least three vertices can be made edge-maximal by the addition of at most $3|V(G)|-6$ edges since each edge added this way has endpoints in $V(G)$ and does not cross any other edge of $G$.
+Let $G$ be a $1$-plane graph.  
+% Then $G$ has $O(V(G)$ edges since the edges of $G$ can be partitioned into two planar graphs by placing one edge from each crossing pair into a second set.
+% \footnote{We may assume that $G$ has no parallel edges, since removing all but one copy from each set of parallel edges does not change the set of adjacent pairs of vertices in $G$, so any supergraph that contains}
+We make $G$ into an \emph{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add $vw$, embedded in the face $F$.
+This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.
+% This increases the number of edges of $G$, by $O(|V(G)|)$ since these newly added edges do not cross any other edges of $G$.
 
@@
 If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
-Therefore each dummy vertex $w_i$ of $W$ is surrounded by two vertices $w_{i-1},w_{i+1}\in V(G)$.
+Therefore each dummy vertex $w_i$ of $W$ is surrounded by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
 Since $w_i$ has degree $4$ in $G_0$, $w_{i-1}\neq w_{i+1}$.
@@
 The point $p$ at which this crossing occurs is a vertex of $G_0$ that has degree $4$.
-There are four faces $F_1,\ldots,F_4$ of $G_0$ that contain $p$.
+There are four faces $F_1,\ldots,F_4$ of $G_0$ each with $p$ on their boundary.
 Each of these four faces of is of Type~(ii), so the boundary of $F_1\cup\cdots\cup F_4$ includes a $4$-cycle $C:=vxwy$ in $G$ whose edges are not involved in any crossings.
@@
 The cycle $vxwy$ is called the \defin{sail} of $K$.
-Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
+Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.\todo{Drop the definitions of kite and spar, we hardly use them.}
 
@@
   The proof is by induction on the number of inner vertices of $N$, i.e., on $|V(N)\setminus V(F)|$.
-If $N$ has no inner vertices, then we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements of the lemma.
+If $N$ has no inner vertices, then we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements of the lemma.
 Therefore, we may assume that $N$ has at least one inner vertex.

### 2022-03-11 k-planar.tex
Pfffft
not man yupdates

@@
 	Let $G$ be a $k$-plane graph.
-We will assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$.
+We assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$.
 This assumption is justified since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
@@
 
-In the following, we will assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$, where we treat each edge $vw$ of $G$ as the image of a simple curve $f_{vw}:[0,1]\to\R^2$.
+In the following, we assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$, where we treat each edge $vw$ of $G$ as the image of a simple curve $f_{vw}:[0,1]\to\R^2$.
 We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
@@
 % \footnote{We may assume that $G$ has no parallel edges, since removing all but one copy from each set of parallel edges does not change the set of adjacent pairs of vertices in $G$, so any supergraph that contains}
-We make $G$ into an \emph{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add $vw$, embedded in the face $F$.
+We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add $vw$, embedded in the face $F$.
 This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.
@@
 
-To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \emph{dummy vertex} $p$.
+To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
 Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$.
@@
 
-If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
+If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
 Therefore each dummy vertex $w_i$ of $W$ is surrounded by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
@@
 		\begin{compactenum}
-			\item No edge of $F$ is crossed by any edge of $G$; and
+			\item No edge of $F$ is crossed by any edge of $G$; and \note{DW}{This can be deleted since $F$ is in $G'$, right?}
 			\item $V(F)$ can be partitioned into $P_1,\ldots,P_k$, for some $k\in\{1,2,3\}$ such that for each $i\in\{1,\ldots,k\}$,
@@
 Now add edges to the inner faces of $N'$ to obtain a supergraph $N''$ in which each inner face has three $3$ edges on its boundary.
-  By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+By Sperner's Lemma \note{DW}{Does this work for multigraph near-triangulations?} there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
 Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.
@@
 
-\referee{2}{Page 16.
+%\referee{2}{Page 16.
 At the end of Page 16, I do not see why 'every pair of edges cross at most once.'  Without further assumptions, it seems as if two paths in the path system could intersect several times.
@@
 It seems this can be fixed by choosing the path system carefully by 'uncrossing' paths which intersect at more than two internal vertices.}
-
-\note{DW}{I have revised the proof to address this error. I couldn't think of a way to do this without introducing balls and channels (and without vigorous hand-waving). Is it readable?}
+%
+%\note{DW}{I have revised the proof to address this error. I couldn't think of a way to do this without introducing balls and channels (and without vigorous hand-waving). Is it readable?}
 
@@
 \citet{PS21} and \citet{DFMS21} use \cref{ps} and \cref{PlanarProduct}(b), respectively, to show that $\chi_p(G)$ is polynomial in $p$ when $G$ is planar or of bounded Euler genus.  
-We use the following lemma due to \citet[Lemma~15]{PS21}.
-
-\begin{lem}[\citep{PS21}]
-\label{p-centered-treewidth}
-  Every graph $H$ of treewidth at most $t$ has $\chi_p(H)\leq \binom{p+t}{t}$.
-\end{lem}
-
-\note{DW}{In the journal version \citet{DFMS21} (under the insistence of the referee :-) ) proved that $\chi_p(H_1\boxtimes H_2) \leq \chi_p(H_1) \cdot \chi(H_2^p)$ where $H_2^p$ is the $p$-th power of $H_2$.
+Upper bounds on $\chi_p$ for graphs of given treewidth~\citep{PS21} and for graph products~\citep{DFMS21} imply the next lemma.
+
+%In the journal version, \citet{DFMS21} proved that $\chi_p(H_1\boxtimes H_2) \leq \chi_p(H_1) \cdot \chi(H_2^p)$ where $H_2^p$ is the $p$-th power of $H_2$.
 Apply this with $H_2:= P \boxtimes K_\ell$.
 Then $H_2^p= P^p\boxtimes K_\ell$ and $\chi(H_2^p)=(p+1)\ell$.
-So
-	$\chi_p(H\boxtimes P \boxtimes K_\ell) \leq \chi_p(H_1) \cdot (p+1)\ell$. So I suggest we replace the next paragraph by 	``\citet{DFMS21} proved that
-$\chi_p(H \boxtimes J) \leq \chi_p(H) \cdot \chi(J^p)$ for all graphs $H$ and $J$, which implies the following lemma'' and delete the proof. Okay?}
-
-The following lemma is implicitly in the work of \citet[Proof of Theorem~2.1]{DFMS21}. We include the proof for completeness.
-
-\begin{lem}[\citep{DFMS21}]
+So $\chi_p(H\boxtimes P \boxtimes K_\ell) \leq \chi_p(H_1) \cdot (p+1)\ell$.
+}
+
+\begin{lem}[\citep{DFMS21,PS21}]
 \label{p-centered}
-If $G\subseteq H\boxtimes P \boxtimes K_\ell$ then $\chi_p(G)\le \ell (p+1)\, \chi_p(H)$.
+For every graph $H$ of treewidth at most $t$ and for every path $P$, if $G\subseteq H\boxtimes P \boxtimes K_\ell$ then
+\[\chi_p(G)\le \ell (p+1)\, \chi_p(H) \leq \ell (p+1) \tbinom{p+t}{t}.\]
+%If $G\subseteq H\boxtimes P \boxtimes K_\ell$ then $\chi_p(G)\le \ell (p+1)\, \chi_p(H)$.
 \end{lem}
 
-\note{DW}{For consistency with \cref{qn}, I suggest we combine \cref{p-centered,p-centered-treewidth}, and just write, ``For every graph $H$ of treewidth at most $t$ and for every path $P$, if $G\subseteq H\boxtimes P \boxtimes K_\ell$ then
-	\[\chi_p(G)\le \ell (p+1)\, \chi_p(H) \leq \ell (p+1) \tbinom{p+t}{t}.\]}
-
-\begin{proof}
-	By \cref{PartitionProduct}, $G$ has an $H$-partition $(\mathcal{L}=\langle L_0,L_1,\ldots\rangle, \PP=(B_x:x\in V(H))$ with layered width at most $\ell$.
+%\note{DW}{For consistency with \cref{qn}, I suggest we combine \cref{p-centered,p-centered-treewidth}, and just write, ``}
+
+%\begin{proof}
+%	By \cref{PartitionProduct}, $G$ has an $H$-partition $(\mathcal{L}=\langle L_0,L_1,\ldots\rangle, \PP=(B_x:x\in V(H))$ with layered width at most $\ell$.
 Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.
 For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in L_i\cap B_x$ such that:
-   \begin{compactenum}
-     \item $\alpha(v)$ is unique among $\{\alpha(w): w\in L_i\cap B_x\}$, which is possible
-     since $|L_i\cap B_x|\le \ell$,
-     \item $\beta(v)= i\bmod (p+1)$, and
-     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
-   \end{compactenum}
- To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
-
- First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$.
+%   \begin{compactenum}
+%     \item $\alpha(v)$ is unique among $\{\alpha(w): w\in L_i\cap B_x\}$, which is possible
+%     since $|L_i\cap B_x|\le \ell$,
+%     \item $\beta(v)= i\bmod (p+1)$, and
+%     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
+%   \end{compactenum}
+% To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
+%
+% First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$.
 Since $X$ is connected, $X$ contains a path from $v$ to $w$.
@@
 Therefore, $|\{\beta(v'):v'\in V(X)\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
-
- Otherwise, $V(X)\subseteq L_{i}\cup\cdots\cup L_{i+s}$ for some $s<p$.
+%
+% Otherwise, $V(X)\subseteq L_{i}\cup\cdots\cup L_{i+s}$ for some $s<p$.
 Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.
@@
 Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
- For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
+% For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
 Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.
 Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
- \end{proof}
-
-\referee{2}{Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.}
-
-\cref{p-centered,p-centered-treewidth,kPlanarProduct,1-planar,gkPlanarProduct} immediately imply the following results, for every $p\ge 2$:
+% \end{proof}
+%
+%\referee{2}{Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.}
+
+\cref{p-centered,kPlanarProduct,1-planar,gkPlanarProduct} immediately imply the following results, for every $p\ge 2$:
 
@@
 
-Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,p-centered-treewidth,non-repetitive} imply the following result.
+Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,non-repetitive} imply the following result.
 
@@
 
-Theorems~\ref{PlanarProduct}(b), \ref{GenusProduct}(b) and \cref{ShortcutProduct} and  \cref{DrawG,qn,p-centered,p-centered-treewidth,non-repetitive,PowerShortcut} imply the following result, which with $g=0$ implies  \cref{kPowerBasic} in the introduction.
+Theorems~\ref{PlanarProduct}(b), \ref{GenusProduct}(b) and \cref{ShortcutProduct} and  \cref{DrawG,qn,p-centered,non-repetitive,PowerShortcut} imply the following result, which with $g=0$ implies  \cref{kPowerBasic} in the introduction.
 
@@
 
-\cref{qn,p-centered,p-centered-treewidth,non-repetitive,MinorFreeDegree,PowerProduct} imply the following analogous result for powers of graphs in any minor-closed class with bounded maximum degree.
+\cref{qn,p-centered,non-repetitive,MinorFreeDegree,PowerProduct} imply the following analogous result for powers of graphs in any minor-closed class with bounded maximum degree.
 
@@
 
-\cref{nearest-neighbour,kPlanarProduct,qn,p-centered,p-centered-treewidth} imply:
+\cref{nearest-neighbour,kPlanarProduct,qn,p-centered} imply:
 
@@
 Let $G:=G_0\blah{C}$.
-Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that:
-\begin{compactitem}
-\item $P_1,\dots,P_k\in\PP$,
-\item $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
-\begin{compactenum}[(i)]
-	\item some bag of $\mathcal{T}$ contains $P_1,\dots,P_k$, and 
-	\item for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
-\end{compactenum}
-\note{DW}{We need to replace (ii) by ``for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$''.
-Curiously, this implies (i) (since $C$ is a facial cycle of $G$).}
-\end{compactitem}
+Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that $P_1,\dots,P_k\in\PP$, and $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
 \end{lem}
@@
 Note that $C'=[P_1,\dots,P_k,P]$ (\note{DW}{or it can be written that way?}).
-By induction, $G'$ has a partition $\PP$ into tripods with respect to $T$ such that:
-\begin{compactitem}
-	\item $P_1,\dots,P_k,P\in\PP$,
-	\item $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
-	\begin{compactenum}[(i)]
-	\item some bag of $\mathcal{T}$ contains $P_1,\dots,P_k,P$, and 
-	\item for every facial cycle $F'$ of $G\blah{C'}$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
-	\end{compactenum}
-\end{compactitem}
-The bag of $\mathcal{T}$ that contains $P_1,\dots,P_k,P$ satisfies property (ii) for $F$. 
-Thus, $\PP$ and $\TT$ satisfy the desired properties for $G$. Now assume that $|V(C)|\geq 3$.
+By induction, $G'$ has a partition $\PP$ into tripods with respect to $T$ such that $P_1,\dots,P_k,P\in\PP$, and $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F'$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F'$ is in $B$.
+The bag of $\mathcal{T}$ that contains $P_1,\dots,P_k,P$ satisfies this property for $F$.
+Thus $\PP$ and $\TT$ satisfy the desired properties for $G$.
+Now assume that $|V(C)|\geq 3$.
 
@@
 By construction, $v_{i-1}$ is not in $G_j$ (where $v_0=v_3$). 
-Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.  Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that:
-\begin{compactitem}
-	\item $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$, 
-	\item the quotient $G_j / \PP_j$ has a tree-decomposition $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3 such that: 
-	\begin{compactitem}
-		\item some bag $B_{u_j}$ of $\mathcal{T}_j$ contains $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}$, and 
-		\item for every facial cycle $F'$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that $V(F')\subseteq \bigcup_{X\in B} X$.
-	\end{compactitem}
-\end{compactitem}
+Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.
+Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$, and 	$G_j / \PP_j$ has a tree-decomposition  $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3 such that for every facial cycle $F'$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that every part of $\PP_j$ that intersects $F'$ is in $B$.
 
@@
 
-Every facial cycle $F_0$ of $G$ is a face of some $G_j$ or is $F$.
-In the latter case, $V(F)\subseteq Y \subseteq \bigcup_{X\in B_u} X$.
-In the first case, there is a bag $B_x$ of $\mathcal{T}_j$ such that $V(F)\subseteq \bigcup_{X\in B_x} X$.
-This property is maintained by the above replacements since
+For every facial cycle $F_0$ of $G$, either $F_0=F$ or $F_0=C$ or $F_0$ is a facial cycle of some $G_j$.
+If $F=F_0$ then the only part of $\PP$ that intersects $F$ is $Y$, so $B_u$ is a bag of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
+If $F_0=C$ then $P_1,\dots,P_k$ are the only parts of $\PP$ that intersect $F_0$, so $B_u$ is a bag of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
+
+In the first case, there is a bag $B_x$ of $\mathcal{T}_j$ such that $V(F)\subseteq \bigcup_{X\in B_x} X$. This property is maintained by the above replacements since 
 $R_{i,j}^+ \subseteq P_{\alpha(i)}$ and 

### 2022-03-11 k-planar.tex
minor edits

@@
 This result requires that $H$ have simple treewidth $3$, even treewidth $3$ is not sufficient.
-
-  Finally, while cleaning up this proof we have also improved the constant from $30$ to $7$.
-The proof is quite similar to the previous proof except that graph $G'$ is obtained by removing both crossing edges involved in a crossing (the spars of a kite), rather than just one.
-We thank the referee for forcing us to write this proof more carefully so that we could discover this improvement.
 }
+
+\note{DW}{I will rewrite the intro to this subsection once we decide what our results are.}
 
@@
 
+\begin{cor}
+	\label{New1Planar}
+	Every 1-planar graph is contained in $H \boxtimes P \boxtimes K_8$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
+\end{cor}
+
 \begin{lem}
@@
 
+\begin{cor}
+	\label{NewMapGraph}
+	Every $d$-map graph is contained in $H \boxtimes P \boxtimes K_{ d  \floor{d/2} }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
+\end{cor}
+
 \end{document}

### 2022-03-11 k-planar.tex
Stuff

--- 2022-03-10 k-planar.tex
@@
 		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
-		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that
-		\begin{compactenum}
-			\item No edge of $F$ is crossed by any edge of $G$; and
-			\item $V(F)$ can be partitioned into $P_1,\ldots,P_k$, for some $k\in\{1,2,3\}$ such that for each $i\in\{1,\ldots,k\}$,
+		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that $V(F)$ can be partitioned into (possibly empty) sets $P_1,\ldots,P_3$ such that for each $i\in\{1,2,3\}$,
 		\begin{compactenum}
@@
 			\item $|V(P_i)\cap L_j| \le 3$ for every integer $j\ge 0$.
-			\end{compactenum}
 		\end{compactenum}
@@
 		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 3$;
-		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; 
+		\item for each $i\in\{1,2,3\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; 
 %		\note{DW}{Should this be $P_i\subseteq S_{x_i}$? (since the path grabs other vertices into its part)}\note{PM}{No, for the same reason as above.}
@@
 
-	Let $M$ denote the subgraph of $G'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$.
-(See \cref{boring_figure}.)  By our choice of $\tau$, the graph $M$ is $2$-connected.
-Furthermore, the vertices on the boundary of each inner face of $M$ are contained in $Y\cup P_a\cup P_b$ for some $a,b\in\{1,\ldots,k\}$.
-In particular if $k=3$ then, for each inner face $F_i$ of $M$ there exists some $c\in\{1,2,3\}$ such that no vertex of $P_c$ is on the boundary of $F_i$.
+  Refer to \cref{boring_figure}.
+	Let $M$ denote the subgraph of $G'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$ and let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
+By our choice of $\tau$, the graph $M$ is $2$-connected.
+  Furthermore, $V(F_i)$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset V(Y)$, $P_{i,2}\subseteq P_a$, $P_{i,3}\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_{i,j}]$ is a path, for each $j\in\{1,2,3\}$.
+Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.
+The remainder of the proof is now identical to that in \citet{DJMMUW20}.
+
+  % Furthermore, for each $i\in\{1,\ldots,m\}$, the vertices on the boundary $F_i$ are contained in $Y\cup P_a\cup P_b$ for some $a,b\in\{1,\ldots,k\}$.
+In particular if $k=3$ then there exists some $c\in\{1,2,3\}$ such that no vertex of $P_c$ is on the boundary of $F_i$.
 
@@
     \begin{center}
-      \begin{tabular}{c@{\hspace{1cm}}c}
-        \includegraphics{figs/mmplus-1} & \includegraphics{figs/mmplus-2} \\
-        $M$ & $M^+$
-      \end{tabular}
+      % \begin{tabular}{c@{\hspace{1cm}}c}
+        % \includegraphics{figs/mmplus-1} & 
+        \includegraphics{figs/mmplus-2}
+      % \end{tabular}
     \end{center}
-    \caption{The graphs $M$ and $M^+$. \note{PM}{This figure needs updating now.}}
+    \caption{The `Sperner' triangle $v_1v_2v_3$ in $N''$ and the Sperner face $\tau$ in $N'$.}
     \label{boring_figure}
@@
 
-  The graph $M$ has some number of inner faces $F_1,\ldots,F_m$, each of which is contained in the interior of $F$. 
+  % The graph $M$ has some number of inner faces $F_1,\ldots,F_m$, each of which is contained in the interior of $F$. 
   % Some of the inner faces of $M^+$ are Type~(ii) faces of $G$.
@@
 In particular $vw$ cannot be on the boundary of $F_i$.
-  Each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.
-Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.
-Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(F_i)$ to obtain the desired $H$-partition $\PP_i$ and tree decomposition of subgraph $N_i$ of $G$ contained in $F_i$.
-
-  \note{PM}{Pick up here tomorrow, after teaching.}
-  The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
-  Elements in each of the $H$-partitions of $N_1,\ldots,N_m$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
-
-  The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~13 in \cite{DJMMUW20}.
+
+  The result of the induction on $N_i$ is a partition $\mathcal{P}_i$ with each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$ in its own part. For $\mathcal{P}$ we use the partition
+  \[
+      \mathcal{P}:=\{Y, P_1,\ldots,P_k\} \cup \bigcup_{i=1}^m \{P\in\mathcal{P}_i: P\cap V(M)=\emptyset\} \enspace .
+  \]
+  All that remains is to verify that $\mathcal{P}$ satisfies Requirements~1--3.   
+  \begin{compactenum}
+    \item To see that $H:=G/\mathcal{P}$ is planar, observe that $G/\mathcal{P}=G'/\mathcal{P}$\todo{Explain why $G/\mathcal{P}=G'/\mathcal{P}$} and that each part $P\in\mathcal{P}$ induces a connected subgraph $G'[P]$ of $G'$.
+Since $G'$ is planar and planarity is preserved under edge contractions, this implies that $G'/\mathcal{P}$ is planar.
+
+    \item  The parts $P_1,P_2,P_3$ satisfy Requirement~2 by Assumption~4(b)ii,  we have already argued that the part $Y$ satisfies Requirement~2, and the inductive hypothesis ensures that each of the remaining parts satisfy Requirement~2.
+
+    \item Requirement~3 is satisfied since $\mathcal{P}$ contains parts $P_1,\ldots,P_k$.
+
+    \item Next we build a tree decomposition $(B_y:y\in V(T_0))$ of $H$ that satisifies Requirement~4.  
+    % We show that $\tw(H)\le 3$ by giving a tree-decomposition of $H$ of width at most $3$. 
+    Let $x_0$ be the vertex of $H$ obtained from contracting $Y$ and, for each $j\in\{1,\ldots,k\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
+
+    The tree $T_0$ has root node $y_0$ with $B_{y_0}:=\{x_0,\ldots,x_k\}$.    
+    For each $i\in\{1,\ldots,m\}$, let $N_i$ denote the subgraph of $N$ contained in $F_i$.
+The inductive hypothesis ensures that $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $(B_z:z\in V(T_i))$ whose largest bag has size at most $4$ whose root bag $B_{z_i}$ contains the vertices $a_i,b_i,c_i\in V(H_i)$ obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
+Recall that $P_{i,1}:=V(N_i)\cap Y$ and $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
+For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.
+Note that these replacements do not increase the size of any bag.
+Now add an edge that joins $y_0$ to $x_i$.
+  \end{compactenum}
+
+  % 
+  % 
+  % \note{PM}{Pick up here tomorrow, after teaching.}
+  % The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
+  % Elements in each of the $H$-partitions of $N_1,\ldots,N_m$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
+  % 
+  % The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~13 in \cite{DJMMUW20}.
 The root node contains the nodes of $H$ corresponding to $P_1,\ldots,P_k$.
@@
 For each $i\in\{1,\ldots,m\}$, the tree decomposition of $N_i/\PP_i$ is then attached to one of these three children.
-
-	The planarity of $H$ comes from two properties:
-	\begin{compactenum}
-		\item $G/\PP$ and $G^+/\mathcal{P^+}$ are isomorphic, where $G^+$ is the triangulation obtained by adding dummy vertices at each crossing in $G$ and $\PP^+$ is the partition we obtain by adding a dummy vertex $z$ to $\overline{Y}^+$ if $\overline{Y}^+$ contains an edge $vw$ that contains $z$ in its interior.
-
-		\item $G^+[\overline{Y}^+-F]$ is connected.
+  % 
+	% The planarity of $H$ comes from two properties:
+	% \begin{compactenum}
+	% 	\item $G/\PP$ and $G^+/\mathcal{P^+}$ are isomorphic, where $G^+$ is the triangulation obtained by adding dummy vertices at each crossing in $G$ and $\PP^+$ is the partition we obtain by adding a dummy vertex $z$ to $\overline{Y}^+$ if $\overline{Y}^+$ contains an edge $vw$ that contains $z$ in its interior.
+  % 
+	% 	\item $G^+[\overline{Y}^+-F]$ is connected.
 To see this, first observe that $\overline{Y}-F$ is connected, and then observe that every vertex of $\overline{Y}^+$ is either a vertex of $\overline{Y}$ or adjacent to a vertex of $\overline{Y}$.
-	\end{compactenum}
-	Since $G^+$ is planar, the second point implies that $H=G^+/\PP$ is planar.
-	% \footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
+	% \end{compactenum}
+	% Since $G^+$ is planar, the second point implies that $H=G^+/\PP$ is planar.
+	% % \footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
 Repeatedly applying this fact implies that $\Delta/\PP$ is planar provided that $\Delta[P]$ is connected for every $P\in\PP$.} The first point then ensures that $G/\PP$ is also planar.

### 2022-03-11 response.tex

@@
 
-\includegraphics[width=60mm]{kite}
+\textcolor{blue}{It appear that the referee forgot that we are dealing with an edge-maximal multigraph at this point. We have revised the proof so that this misunderstanding cannot happen.}
+
+%\includegraphics[width=60mm]{kite}
 
@@
 
-\tba
+\textcolor{blue}{Our definition of kite was not exactly what we meant. This section has been completely revised.}
 
@@
 
-\tba
+\textcolor{blue}{Upon reflection we agree with the referee here.
+Indeed, while cleaning up this proof we have also improved the constant from $30$ to $7$.
+The proof is quite similar to the previous proof except that graph $G'$ is obtained by removing both crossing edges involved in a crossing (the spars of a kite), rather than just one.
+We thank the referee for forcing us to write this proof more carefully so that we could discover this improvement.}
 

### 2022-03-11 k-planar.tex

--- 2022-03-11 k-planar.tex
@@
 		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
-		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that
-		\begin{compactenum}
-			\item No edge of $F$ is crossed by any edge of $G$; and \note{DW}{This can be deleted since $F$ is in $G'$, right?}
-			\item $V(F)$ can be partitioned into $P_1,\ldots,P_k$, for some $k\in\{1,2,3\}$ such that for each $i\in\{1,\ldots,k\}$,
+		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that $V(F)$ can be partitioned into (possibly empty) sets $P_1,\ldots,P_3$ such that for each $i\in\{1,2,3\}$,
 		\begin{compactenum}
@@
 		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 3$;
-		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; 
+		\item for each $i\in\{1,2,3\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$;
 %		\note{DW}{Should this be $P_i\subseteq S_{x_i}$? (since the path grabs other vertices into its part)}\note{PM}{No, for the same reason as above.}
@@
     \begin{center}
-      \begin{tabular}{c@{\hspace{1cm}}c}
-        \includegraphics{figs/mmplus-1} & \includegraphics{figs/mmplus-2} \\
-        $M$ & $M^+$
-      \end{tabular}
+      % \begin{tabular}{c@{\hspace{1cm}}c}
+        % \includegraphics{figs/mmplus-1} &
+        \includegraphics{figs/mmplus-2}
+      % \end{tabular}
     \end{center}
-    \caption{The graphs $M$ and $M^+$. \note{PM}{This figure needs updating now.}}
+    \caption{The `Sperner' triangle $v_1v_2v_3$ in $N''$ and the Sperner face $\tau$ in $N'$.}
     \label{boring_figure}
@@
 
-  The graph $M$ has some number of inner faces $F_1,\ldots,F_m$, each of which is contained in the interior of $F$. 
+  % The graph $M$ has some number of inner faces $F_1,\ldots,F_m$, each of which is contained in the interior of $F$.
   % Some of the inner faces of $M^+$ are Type~(ii) faces of $G$.
@@
 In particular $vw$ cannot be on the boundary of $F_i$.
-  Each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.
-Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.
-Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(F_i)$ to obtain the desired $H$-partition $\PP_i$ and tree decomposition of subgraph $N_i$ of $G$ contained in $F_i$.
-
-  \note{PM}{Pick up here tomorrow, after teaching.}
-  The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
-  Elements in each of the $H$-partitions of $N_1,\ldots,N_m$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
-
-  The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~13 in \cite{DJMMUW20}.
+
+  The result of the induction on $N_i$ is a partition $\mathcal{P}_i$ with each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$ in its own part. For $\mathcal{P}$ we use the partition
+  \[
+      \mathcal{P}:=\{Y, P_1,\ldots,P_k\} \cup \bigcup_{i=1}^m \{P\in\mathcal{P}_i: P\cap V(M)=\emptyset\} \enspace .
+  \]
+  All that remains is to verify that $\mathcal{P}$ satisfies Requirements~1--3.
+  \begin{compactenum}
+    \item To see that $H:=G/\mathcal{P}$ is planar, observe that $G/\mathcal{P}=G'/\mathcal{P}$\todo{Explain why $G/\mathcal{P}=G'/\mathcal{P}$} and that each part $P\in\mathcal{P}$ induces a connected subgraph $G'[P]$ of $G'$.
+Since $G'$ is planar and planarity is preserved under edge contractions, this implies that $G'/\mathcal{P}$ is planar.
+
+    \item  The parts $P_1,P_2,P_3$ satisfy Requirement~2 by Assumption~4(b)ii,  we have already argued that the part $Y$ satisfies Requirement~2, and the inductive hypothesis ensures that each of the remaining parts satisfy Requirement~2.
+
+    \item Requirement~3 is satisfied since $\mathcal{P}$ contains parts $P_1,\ldots,P_k$.
+
+    \item Next we build a tree decomposition $(B_y:y\in V(T_0))$ of $H$ that satisifies Requirement~4.
+    % We show that $\tw(H)\le 3$ by giving a tree-decomposition of $H$ of width at most $3$.
+    Let $x_0$ be the vertex of $H$ obtained from contracting $Y$ and, for each $j\in\{1,\ldots,k\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
+
+    The tree $T_0$ has root node $y_0$ with $B_{y_0}:=\{x_0,\ldots,x_k\}$.
+    For each $i\in\{1,\ldots,m\}$, let $N_i$ denote the subgraph of $N$ contained in $F_i$.
+The inductive hypothesis ensures that $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $(B_z:z\in V(T_i))$ whose largest bag has size at most $4$ whose root bag $B_{z_i}$ contains the vertices $a_i,b_i,c_i\in V(H_i)$ obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
+Recall that $P_{i,1}:=V(N_i)\cap Y$ and $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
+For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.
+Note that these replacements do not increase the size of any bag.
+Now add an edge that joins $y_0$ to $x_i$.
+  \end{compactenum}
+
+  %
+  %
+  % \note{PM}{Pick up here tomorrow, after teaching.}
+  % The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
+  % Elements in each of the $H$-partitions of $N_1,\ldots,N_m$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
+  %
+  % The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~13 in \cite{DJMMUW20}.
 The root node contains the nodes of $H$ corresponding to $P_1,\ldots,P_k$.
@@
 For each $i\in\{1,\ldots,m\}$, the tree decomposition of $N_i/\PP_i$ is then attached to one of these three children.
-
-	The planarity of $H$ comes from two properties:
-	\begin{compactenum}
-		\item $G/\PP$ and $G^+/\mathcal{P^+}$ are isomorphic, where $G^+$ is the triangulation obtained by adding dummy vertices at each crossing in $G$ and $\PP^+$ is the partition we obtain by adding a dummy vertex $z$ to $\overline{Y}^+$ if $\overline{Y}^+$ contains an edge $vw$ that contains $z$ in its interior.
-
-		\item $G^+[\overline{Y}^+-F]$ is connected.
+  %
+	% The planarity of $H$ comes from two properties:
+	% \begin{compactenum}
+	% 	\item $G/\PP$ and $G^+/\mathcal{P^+}$ are isomorphic, where $G^+$ is the triangulation obtained by adding dummy vertices at each crossing in $G$ and $\PP^+$ is the partition we obtain by adding a dummy vertex $z$ to $\overline{Y}^+$ if $\overline{Y}^+$ contains an edge $vw$ that contains $z$ in its interior.
+  %
+	% 	\item $G^+[\overline{Y}^+-F]$ is connected.
 To see this, first observe that $\overline{Y}-F$ is connected, and then observe that every vertex of $\overline{Y}^+$ is either a vertex of $\overline{Y}$ or adjacent to a vertex of $\overline{Y}$.
-	\end{compactenum}
-	Since $G^+$ is planar, the second point implies that $H=G^+/\PP$ is planar.
-	% \footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
+	% \end{compactenum}
+	% Since $G^+$ is planar, the second point implies that $H=G^+/\PP$ is planar.
+	% % \footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
 Repeatedly applying this fact implies that $\Delta/\PP$ is planar provided that $\Delta[P]$ is connected for every $P\in\PP$.} The first point then ensures that $G/\PP$ is also planar.

### 2022-03-11 k-planar.tex
Pffffft

@@
 Let $G$ be a $1$-plane graph.
-% Then $G$ has $O(V(G)$ edges since the edges of $G$ can be partitioned into two planar graphs by placing one edge from each crossing pair into a second set.
-% \footnote{We may assume that $G$ has no parallel edges, since removing all but one copy from each set of parallel edges does not change the set of adjacent pairs of vertices in $G$, so any supergraph that contains}
 We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add $vw$, embedded in the face $F$.
 This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.
-% This increases the number of edges of $G$, by $O(|V(G)|)$ since these newly added edges do not cross any other edges of $G$.
 
@@
 	Then $N$ has an $H$-partition $\PP=\{S_x\cup\{v_x\} : x\in V(H)\}$ such that:
-	\begin{compactenum}
+	\begin{compactenum}[(i)]
 		\item $H$ is planar;
@@
   This proof is very similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is sufficiently complicated by the $1$-planar setting to warrant a full proof.
-   %
-   %
-   %
-   %
-   % Rather than duplicate every detail of that proof here, we focus on the differences and refer the reader to the original proof for the remaining details.
 In particular, the reader should keep in mind that $G$ and $G'$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
 
-  The proof is by induction on the number of inner vertices of $N$, i.e., on $|V(N)\setminus V(F)|$.
-If $N$ has no inner vertices, then we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements of the lemma.
-Therefore, we may assume that $N$ has at least one inner vertex.
-
-  If $k\in\{1,2\}$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge of $F$ on its boundary.
-If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.
+  The proof is by induction on the number of inner faces of $N'$.
+If $N'$ has only one inner face then we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
+  \begin{inparaenum}[(i)]
+    \item $H:=N/\mathcal{P}$ is planar because any graph on $k\le 3$ vertices is planar.
+    \item Assumption 4(a) ensures that $|S_x\cap L_i|\le 3$ for each $x\in V(H)$.
+    \item By definition, each each of $P_1,\ldots,P_k$ is s part of $\mathcal{P}$.
+    \item The trivial tree decomposition of $H$ that places $x_1,\ldots,x_k$ in one bag satisfies this requirement.
+  \end{inparaenum}
+  Therefore, we may assume that $N$ has more than one inner face.
+
+  The $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a part $Y$, called a \emph{tripod}, that we now define.
+The tripod $Y$ is determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
+Depending on the value $k$, there are two cases to consider:
+  \begin{compactenum}
+    \item If $k\in\{1,2\}$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge of $F$ on its boundary.
+
+    \item If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
 Now add edges to the inner faces of $N'$ to obtain a supergraph $N''$ in which each inner face has three $3$ edges on its boundary.
-By Sperner's Lemma \note{DW}{Does this work for multigraph near-triangulations?} there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+By Sperner's Lemma \todo{DW:Does this work for multigraph near-triangulations?} there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
 Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.
 (Note that $p\in\{3,4\}$.)
-
-  % If $k=2$, then let $\tau:=v_1v_2$ be any edge of $N'$ whose interior is in the interior of $F$ and whose endpoints have different colours.
-The existence of such an edge is straightforward to establish by considering the auxilliary graph $A$ whose vertices are the faces of $N'$ and in which two faces are adjacent if and only if they share an edge whose endpoints have different colours.
-Every vertex in $A$ has degree $0$ or $2$ and the outer face $F_0$ of $N'$ has degree $2$ in $A$.
-Therefore $A$ contains a cycle $C$ that contains $F_0$.
-Two edges of $C$ correspond to bichromatic edges of $F$ and the the remaining edges correspond to bichromatic edges in the interior of $F$, so it only remains to verify that $C$ has length at least $3$.
-The only other possibility is that $C$ has length $2$, which
-
-  % See, for example, \cite[Lemma~6]{bose.morin.ea:optimal}.
-
+  \end{compactenum}
+
+  \todo[inline]{
+    Answering David's question: Yes, it only requires that the inner faces have three sides and that the outer face is bounded by a simple cycle.
+The proof goes through without change.
+Look at the subgraph $X$ of the dual that contains only the edges dual to red/blue edges.
+In our setting, the outer face has degree $1$ in $X$, which is odd.
+Any bichromatic inner face has degree $0$ or degree $2$ in $X$.
+There has to be another vertex of odd degree somewhere, which can only come from a trichromatic inner face.
+  }
+
+  We can now define the vertices in the tripod $Y$.
 For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to $V(F)$.
-   %
-   %
-   % If $k=3$, set $R_i := P_i$ for each $i\in\{1,2,3\}$.
-Otherwise, as in \citep{DJMMUW20}, split $P_1,\ldots,P_k$ to partition $V(F)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a non-empty path and each $R_i$ contains vertices from exactly one of $P_1,\ldots,P_k$.
-  %
-	% Next, as in \citep{DJMMUW20}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
 Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
@@
 Therefore $|S\cap L_j|\le 3$ for each integer $j\ge 0$.
-Therefore, taking $Y:=S\cup\{v\}\setminus V(F)$ as a part in the partition $\PP$ will satisfy Requirement~2 of the lemma.
+Therefore, taking $Y:=S\cup\{v\}\setminus V(F)$ as a part in the partition $\PP$ will not violate Condition~(ii).
 

### 2022-03-11 k-planar.tex
Pffft

@@
 
-% \referee{1}{3.
-Theorem 10: Unfortunately I have to suggest removing it.
-I don’t see the motivation for making $H$ planar or for improving the constants.
-Are the improved constants at all close to tight?
-Is there another reason for including it?}
-%
-% \note{DW}{We can motivate \cref{1-planar} since the graph drawing community particularly cares about 1-planar graphs, since treewidth 3 is best possible, and since the planarity of $H$ is essential in the proof of \cref{1PlanarQueue}.
-Futher motivation comes from the open problem at the end of the paper.
-We need to argue this in the response and in the paper.
-The alternative is to remove \cref{sec-1-planar}?
-We could keep the statement of \cref{1-planar} in the paper and refer to the arXiv paper for the proof.
-What do you think?
-All this depends on whether we can fix the errors below.
-I vote to fix the errors, add more motivation for \cref{1-planar}, keep the theorem and the proof.
-}
-
-%Let $G$ be an edge-maximal 1-plane multigraph with no two parallel edges on the boundary of a single face.
-Here, edge-maximal should be taken to mean that, if any two vertices $v$ and $w$ appear on a common face\footnote{The \defin{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
-We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.
-\note{DW}{Each face of $G$ is one of two types, either bounded by three edges (none of which are crossed), or bounded by one edge that is not crossed plus portions of two crossing edges.
-Right?
-If this is true, we should say it here, which would help to clear up the confusion below.
-Add a figure.}
-%We assume that no two edges incident to a common vertex cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}
-%
-%\note{DW}{I suggest we say that it follows from Euler's formula that $G$ is finite.
-(To see this, say $G$ has $n$ vertices, $m$ non-crossed edges, and $k$ crossings.
-Let $G'$ be the plane graph obtained from $G$ by adding a dummy vertex at each crossing.
-So $G'$ is a plane multigraph in which every face is bounded by a 3-cycle.
-It follows from Euler's formula that $m+4k = |E(G')| = 3( |V(G')|-2) = 3 ( n+k-2)$, implying	$m+k = 3 (n-2)$ and $|E(G)| = m+2k \leq 6(n-2)$, which can also be shown by deleting one edge from each crossing point.
-) }
-%
-%\note{DW}{It seems to be that $G$ can be obtained from a plane multigraph with faces of size 3 or 4 such that no edge is in two 3-faces, by adding a pair of crossing edges across each 4-face.
-I think this characterises edge-maximal 1-planar multigraphs with no 2-faces.
-Right?
-This should be stated as a lemma.
-Does this viewpoint help?
-Is it known?}
-%
-%\note{DW}{Say $G$ is obtained from a plane multigraph by making each face a clique. Is $G\subseteq H \boxtimes P \boxtimes K_{f(\omega(G)}$? for some treewidth 3 graph $H$.}
-%
-%A \defin{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.
-It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
-\note{DW}{We need to define a kite to be the subgraph with vertex-set $\{v,w,x,y\}$ and edge-set $\{vx,xy,vx,vy,wy,wx\}$ where $vx,vy,wy,wx$ are the edges on the `outside' of the kite-faces.
-Then it really is a simple $K_4$.
-I think this is what we intended.
-We should add a figure showing a kite and kite-faces.
-}
-%The edges $vw$ and $xy$ are called \defin{spars} of $K$.
-The cycle $vxwy$ is called the \defin{sail} of $K$.
-It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$.
-%
-%\referee{2}{Page 13. Perhaps I am misunderstanding something, but I do not see why
-%'none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of
-%$G$.'  For example, see the attached PDF for a picture where $vw$, $xw$, $wy$,
-%and $yv$ are all crossed by other edges of $G$.\\
-%\includegraphics[width=60mm]{kite}}
-%\note{DW}{We need to explain more about how edge-maximality fixes this problem.
-In this case we can draw other $xv$, $vy$, $yw$ $wx$ edges close to the central crossing edges.} Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
-%
-%\referee{2}{Page 13. The attached PDF also shows that the definition of 'kite
-%face' may not be well-defined.  For example, I do not see why there
-%cannot be some edges of $G$ 'inside' a kite face.  Even if this is not
-%the case, I think it is more precise to say that a kite face has 'two
-%and a half edges' and two vertices of $G$ on its boundary' rather than
-%'three edges and two vertices of $G$ on its boundary'.}
-%
-%\note{DW}{This should be fixed if we define kite more precisely. A kite-face is bounded by one edge and two half-edges right?}
-
-% \note{PM}{I agree that we should fix it and keep it.
-Another reason: For vertex ranking, the bound is $O(\log n/\log^{(t)} n)$ where $t:=\tw(H)$, so this is an example where $\tw(H)$ affects more than just the constant.}
-
 In the following, we assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$, where we treat each edge $vw$ of $G$ as the image of a simple curve $f_{vw}:[0,1]\to\R^2$.
@@
         \includegraphics{figs/one_planar_example-4} \\
-        (a) & (b)
+        $G$ & $G'$
      \end{tabular}
     \end{center}
-    \caption{(a)~Kites in an edge-maximal $1$-plane multigraph $G$; (b)~Plane multigraph $G'$ obtained by removing one spar from each kite of $G$.}
+    \caption{The graph $G'$ obtained by removing pairs of crossing edges from $G$ is a plane multigraph whose faces all have three or four sides.}
     \label{one_planar_example}
@@
 Each of these four faces of is of Type~(ii), so the boundary of $F_1\cup\cdots\cup F_4$ includes a $4$-cycle $C:=vxwy$ in $G$ whose edges are not involved in any crossings.
-We call the subgraph $K:=C\cup\{vw,xy\}$ a \defin{kite} of $G$; see \cref{one_planar_example}(a).
-The edges $vw$ and $xy$ are called \defin{spars} of $K$.
-The cycle $vxwy$ is called the \defin{sail} of $K$.
-Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.\todo{Drop the definitions of kite and spar, we hardly use them.}
-
-%
-% It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$.
-
-%
-% \note{DW}{Each face of $G$ is one of two types, either bounded by three edges (none of which are crossed), or bounded by one edge that is not crossed plus portions of two crossing edges.
-Right?
-If this is true, we should say it here, which would help to clear up the confusion below.
-Add a figure.}
-%
-% \note{DW}{I suggest we say that it follows from Euler's formula that $G$ is finite.
-(To see this, say $G$ has $n$ vertices, $m$ non-crossed edges, and $k$ crossings.
-Let $G'$ be the plane graph obtained from $G$ by adding a dummy vertex at each crossing.
-So $G'$ is a plane multigraph in which every face is bounded by a 3-cycle.
-It follows from Euler's formula that $m+4k = |E(G')| = 3( |V(G')|-2) = 3 ( n+k-2)$, implying	$m+k = 3 (n-2)$ and $|E(G)| = m+2k \leq 6(n-2)$, which can also be shown by deleting one edge from each crossing point.
-) }
-%
-% \note{DW}{It seems to be that $G$ can be obtained from a plane multigraph with faces of size 3 or 4 such that no edge is in two 3-faces, by adding a pair of crossing edges across each 4-face.
-I think this characterises edge-maximal 1-planar multigraphs with no 2-faces.
-Right?
-Does this viewpoint help?
-Is it known?}
-%
-% \note{DW}{Say $G$ is obtained from a plane multigraph by making each face a clique. Is $G\subseteq H \boxtimes P \boxtimes K_{f(\omega(G)}$? for some treewidth 3 graph $H$.}
-%
-% % A \defin{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.
-It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
-\note{DW}{We need to define a kite to be the subgraph with vertex-set $\{v,w,x,y\}$ and edge-set $\{vx,xy,vx,vy,wy,wx\}$ where $vx,vy,wy,wx$ are the edges on the `outside' of the kite-faces.
-Then it really is a simple $K_4$.
-I think this is what we intended.
-We should add a figure showing a kite and kite-faces.
-}
-%
-% \referee{2}{Page 13. Perhaps I am misunderstanding something, but I do not see why
-% 'none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of
-% $G$.'  For example, see the attached PDF for a picture where $vw$, $xw$, $wy$,
-% and $yv$ are all crossed by other edges of $G$.\\
-% \includegraphics[width=60mm]{kite}}
-% \note{DW}{We need to explain more about how edge-maximality fixes this problem.
-In this case we can draw other $xv$, $vy$, $yw$ $wx$ edges close to the central crossing edges.} \note{PM}{In the example above, $v$ and $y$ appear on a common (inner) face $F$, so by definition of edge-maximality there is an edge $xy$  on the boundary of $F$.
-}
-%
-% \referee{2}{Page 13. The attached PDF also shows that the definition of 'kite
-% face' may not be well-defined.  For example, I do not see why there
-% cannot be some edges of $G$ 'inside' a kite face.  Even if this is not
-% the case, I think it is more precise to say that a kite face has 'two
-% and a half edges' and two vertices of $G$ on its boundary' rather than
-% 'three edges and two vertices of $G$ on its boundary'.}
-% \note{DW}{This should be fixed if we define kite more precisely. A kite-face is bounded by one edge and two half-edges right?}
-%
-% \referee{2}{Page 13. Could it be that the subgraph induced by the vertices of a
-% 	kite is a $K_4$ with some parallel edges?  This seems relevant later in
-% 	the proof.}
-%
-% \note{DW}{If we define kite as above, this is not a problem}
-
-By removing both spars from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
-Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).
-Therefore $\dist_{G'}(v,w)\le 2$.
+
+By removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
+Since, for any pair of crossing edges $(vw,xy)$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.
+% It follows immediately that, for any pair of vertices $a,b\in V(G)$, $\dist_{G'}(a,b)\le 2\dist_G(a,b)$.
 
@@
 		\end{compactenum}
-% \note{DW}{Should the vertices already grabbed by $P_1,P_2,P_3$ be mentioned in the setup?
-The new paths $Q_1,Q_2,Q_3$ need to avoid the vertices already grabbed by $P_1,P_2,P_3$.
-Basically, I am concerned about what happens when distinct vertical paths in $T$ want to grab the same vertex.} \note{PM}{No, at this point we're only promising an $H$-partition of $N$, the part of $G$ contained in the uncrossed cycle $F$.}
 		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
@@
 		\item for each $i\in\{1,2,3\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$;
-%		\note{DW}{Should this be $P_i\subseteq S_{x_i}$? (since the path grabs other vertices into its part)}\note{PM}{No, for the same reason as above.}
- and
 		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
@@
 \end{lem}
-
-% \referee{2}{Page 13.  Regarding the proof of Lemma 4, I think it is better to
-% include all the details from reference [14].  As far as I know, JCTB
-% does not have a page limit, so I do not see an issue with just
-% reproducing the entire proof and telling the reader that they can skip
-% all the details if they wish.  This is just my personal opinion
-% though, so the authors can ignore this request if they choose.}
-%
-% \note{DW}{I agree we need more details. I am having trouble verifying this proof. }
 
@@
 The tripod $Y$ is determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
-Depending on the value $k$, there are two cases to consider:
+Depending on the value of $k$, there are two cases to consider:
   \begin{compactenum}
@@
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
-Now add edges to the inner faces of $N'$ to obtain a supergraph $N''$ in which each inner face has three $3$ edges on its boundary.
-By Sperner's Lemma \todo{DW:Does this work for multigraph near-triangulations?} there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+Now add one edge to each $4$-sided inner face of $N'$ to obtain a plane supergraph $N''$ of $N'$ in which each inner face has exactly three $3$ edges on its boundary.
+By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
 Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.
@@
 
-  \todo[inline]{
-    Answering David's question: Yes, it only requires that the inner faces have three sides and that the outer face is bounded by a simple cycle.
+  \todo[inline]{%
+    DW:Does this work for multigraph near-triangulations? \newline
+    PM: Yes, it only requires that the inner faces have three sides and that the outer face is bounded by a simple cycle.
 The proof goes through without change.
@@
   We can now define the vertices in the tripod $Y$.
-For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to $V(F)$.
+For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to a vertex in $V(F)$.
 Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
@@
   Refer to \cref{boring_figure}.
-	Let $M$ denote the subgraph of $G'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$ and let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
+Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$.
+Since $N'\subseteq G'$ contains only edges of $G$ that are not crossed, each component of $G-F-Y$ is contained in a single face of $M$.
+
+  Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
 By our choice of $\tau$, the graph $M$ is $2$-connected.
-  Furthermore, $V(F_i)$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset V(Y)$, $P_{i,2}\subseteq P_a$, $P_{i,3}\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_{i,j}]$ is a path, for each $j\in\{1,2,3\}$.
-Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.
-The remainder of the proof is now identical to that in \citet{DJMMUW20}.
-
-  % Furthermore, for each $i\in\{1,\ldots,m\}$, the vertices on the boundary $F_i$ are contained in $Y\cup P_a\cup P_b$ for some $a,b\in\{1,\ldots,k\}$.
-In particular if $k=3$ then there exists some $c\in\{1,2,3\}$ such that no vertex of $P_c$ is on the boundary of $F_i$.
+Furthermore, $V(F_i)$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset V(Y)$, $P_{i,2}\subseteq P_b$, $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
+In the following, in order to avoid introducing even more notation or constantly adding the modifier ``when it exists'', we will assume that $V(F_i)\cap Y\neq\emptyset$ and that there are exactly two distinct $b,c\in\{1,\ldots,k\}$ such that $V(F_i)\cap P_b\neq\emptyset$ and $V(F_i)\cap P_c\neq\emptyset$.
+When this assumption does not hold, exactly the same reasoning can be used, while omitting one or more of $P_{i,1}$, $P_{i,2}$ or $P_{i,3}$.
 
@@
 
-  % The graph $M$ has some number of inner faces $F_1,\ldots,F_m$, each of which is contained in the interior of $F$.
-  % Some of the inner faces of $M^+$ are Type~(ii) faces of $G$.
-Let $F_1,\ldots,F_m$ be the inner faces of $M^+$ that are not Type~(ii) faces of $G$.
-We claim that, for each $i\in\{1,\ldots,m\}$, the boundary of $F_i$ is a cycle in $G'$ that contains no spars.
-Otherwise, some edge $vw$ contributes to the boundary of $F_i$ but is crossed by an edge $xy\in E(G)$.
-Then, $vw\not\in E(F)$ since no edge of $F$ is crossed by any edge of $G$.
-Therefore $vw\in E(\overline{Y}^+)$ so $xy\in E(\overline{Y}^+)$.
-But then the only faces of $M^+$ with $vw$ on their boundary are Type~(ii) faces.
-In particular $vw$ cannot be on the boundary of $F_i$.
-
-  The result of the induction on $N_i$ is a partition $\mathcal{P}_i$ with each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$ in its own part. For $\mathcal{P}$ we use the partition
+  Since $M$ is $2$-connected and $\overline{Y}$ is connected, $F_i[P_{i,j}]$ is a path, for each $j\in\{1,2,3\}$.
+Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.
+The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of the subgraph $N_i$ of $N$ contained in $F_i$.
+The partition $\mathcal{P}_i$ contains each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$.
+For $\mathcal{P}$ we use the partition
   \[
@@
   \]
-  All that remains is to verify that $\mathcal{P}$ satisfies Requirements~1--3.
-  \begin{compactenum}
-    \item To see that $H:=G/\mathcal{P}$ is planar, observe that $G/\mathcal{P}=G'/\mathcal{P}$\todo{Explain why $G/\mathcal{P}=G'/\mathcal{P}$} and that each part $P\in\mathcal{P}$ induces a connected subgraph $G'[P]$ of $G'$.
+  It is straightforward to verify that $\mathcal{P}$ is indeed a partition of $V(N)$ and, by definition this is an $H$-partition of $N$ for the graph $H:=N/\mathcal{P}$.
+All that remains is to verify that $\mathcal{P}$ satisfies conditions~(i)--(iv).
+  \begin{compactenum}[(i)]
+    \item To see that $H$ is planar, observe that $G/\mathcal{P}=G'/\mathcal{P}$\todo{Explain why $G/\mathcal{P}=G'/\mathcal{P}$} and that each part $P\in\mathcal{P}$ induces a connected subgraph $G'[P]$ of $G'$.
 Since $G'$ is planar and planarity is preserved under edge contractions, this implies that $G'/\mathcal{P}$ is planar.
 
-    \item  The parts $P_1,P_2,P_3$ satisfy Requirement~2 by Assumption~4(b)ii,  we have already argued that the part $Y$ satisfies Requirement~2, and the inductive hypothesis ensures that each of the remaining parts satisfy Requirement~2.
-
-    \item Requirement~3 is satisfied since $\mathcal{P}$ contains parts $P_1,\ldots,P_k$.
-
-    \item Next we build a tree decomposition $(B_y:y\in V(T_0))$ of $H$ that satisifies Requirement~4.
+    \item  By Assumption~4(b) none of the parts $P_1,P_2,P_3$ violate condition~(ii).
+We have already argued that the part $Y$ does not violate condition~(ii).
+The inductive hypothesis ensures that each of the remaining parts satisfies condition~(ii).
+
+    \item Condition~(iii) is satisfied since $\mathcal{P}$ contains parts $P_1,\ldots,P_k$.
+
+    \item We now construct a tree decomposition $(B_y:y\in V(T_0))$ of $H$ that satisifies condition~(iv).
     % We show that $\tw(H)\le 3$ by giving a tree-decomposition of $H$ of width at most $3$.
@@
     The tree $T_0$ has root node $y_0$ with $B_{y_0}:=\{x_0,\ldots,x_k\}$.
-    For each $i\in\{1,\ldots,m\}$, let $N_i$ denote the subgraph of $N$ contained in $F_i$.
+    % For each $i\in\{1,\ldots,m\}$, let $N_i$ denote the subgraph of $N$ contained in $F_i$.
     The inductive hypothesis ensures that $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $(B_z:z\in V(T_i))$ whose largest bag has size at most $4$ whose root bag $B_{z_i}$ contains the vertices $a_i,b_i,c_i\in V(H_i)$ obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
-Recall that $P_{i,1}:=V(N_i)\cap Y$ and $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
+Recall that $P_{i,1}:=V(N_i)\cap Y$, $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
 For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.
-Note that these replacements do not increase the size of any bag.
+Note that these replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$.
 Now add an edge that joins $y_0$ to $x_i$.
+
+    It is now straightforward to verify that doing this for each $i\in\{1,\ldots,m\}$ yields a tree decomposition of $N$ that satisfies condition~(iv).
+(The key observation is that there is no edge of $N$  with one endpoint in $N_i-F_i$ and one endpoint in $N_j-F_j$ for any $i\neq j$.)
   \end{compactenum}
@@
 	Given a 1-plane graph $G$, add edges to make it edge-maximal so that it has an outer face $F=v_1v_2v_3$.
-Next, add a vertex $r$ adjacent to $v_1$, $v_2$, and $v_3$ to obtain an edge-maximal 1-plane graph $\overline{G}$ with one vertex $r$ of degree 3 on its outer face.
-
-	Let $G'$ be the plane graph obtained by removing both spars from each kite of $\overline{G}$ and let $T$ be a BFS tree of $G'$ rooted at $r$.
+Next, add a vertex $r$ adjacent to $v_1$, $v_2$, and $v_3$ to obtain an edge-maximal 1-plane multigraph $\overline{G}$ with one vertex $r$ of degree 3 on its outer face.
+
+	Let $G'$ be the plane graph obtained by removing each pair of crossing edges from $\overline{G}$ and let $T$ be a BFS tree of $G'$ rooted at $r$.
 Now apply \cref{induction} with $G=\overline{G}$, $G'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
-This gives an $H$-partition $\{S_x\cup\{v_x\}:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ is planar and has treewidth at most 3.
+This gives an $H$-partition $\mathcal{P}:=\{S_x\cup\{v_x\}:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ is planar and has treewidth at most 3.
 
 	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{2i}\cup L_{2i+1}$ for each integer $i\ge 0$.
-That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le 2$ for every edge $vw\in E(G)$.
-Since $|L_j\cap S_x|\le 3$ for every integer $j\ge 0$, $|L_i'\cap (S_x\cup\{v_x\})|\le 7$ for every integer $i\ge 0$ and every $x\in V(H)$.
+That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le 2$ for each edge $vw\in E(G)$.
+Since $|L_j\cap S_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap (S_x\cup\{v_x\})|\le 7$ for each integer $i\ge 0$ and each $x\in V(H)$.
 The result follows from \cref{PartitionProduct}.
 \end{proof}
+
+We remark that our proof of \cref{1-planar} generalizes to the class of $d$-framed graphs \cite{BDGGMR}.
+An embedded multigraph $G$ is a \defin{$d$-framed graph} if it contains a plane subgraph $G'$ (the \defin{frame}) whose edges are not involved in any crossings and such that each pair of crossing edges in $G$ are contained in a face of $G'$ that has at most $d$ vertices on its boundary.
+It is known that the class of $4$-framed multigraphs is exactly the class of $1$-plane multigraphs \cite{BDGGMR}.
+Indeed, the graph $G'$ in \cref{induction} is the frame of $G$.
+The proof of \cref{induction} can be repeated for any edge-maximal $d$-framed graph $G$ having frame $G'$.
+The proof in this case has only two differences:
+\begin{compactitem}
+   \item For any edge $vw$ of $G$ that is not in $G'$, $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$.
+   \item The face $\tau$ that contains the Sperner triangle $v_1,v_2,v_3$ may have up to $d$ vertices on its boundary, so up to $d-3$ of these vertices do not belong to $Q_i$ for any $i\in\{1,2,3\}$.
+\end{compactitem}
+The only effect this has on the final result is that the layered width becomes $3\lfloor d/2\rfloor + d-3$.  This yields the following result:
+
+\begin{thm}
+\label{1-planar}
+    Every $d$-framed graph is contained in $H\boxtimes P\boxtimes K_{3\lfloor d/2\rfloor + d -3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+\end{thm}
 
@@
 \label{1PlanarQueue}
-Every 1-planar graph has queue-number at most $3 \times 30 \times 5 + \floor{\tfrac{3}{2} \times 30} = 495$.
+Every 1-planar graph has queue-number at most $3 \times 7 \times 5 + \floor{\tfrac{3}{2} \times 7} = 115$.
 \end{cor}
@@
 \begin{compactitem}
-\item	For every $1$-planar graph $G$, $\pi(G)\le 30\times 4^4=7680$.
+\item	For every $1$-planar graph $G$, $\pi(G)\le 7\times 4^4=1792$.
 %\item For every $k$-planar graph $G$,

### 2022-03-11 k-planar.tex
Pfffft

@@
 
-\note{DW}{I suggest we use consecutive numbering of theorems, lemmas, etc. Theorem 1, Lemma 2, Observation 3, etc.}
-
-\note{DW}{I suggest we delete ``We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$. '' This is absolutely standard.}
+% \note{DW}{I suggest we use consecutive numbering of theorems, lemmas, etc. Theorem 1, Lemma 2, Observation 3, etc.}
+
+\note{PM}{Agreed. Done.}
+
+% \note{DW}{I suggest we delete ``We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$. '' This is absolutely standard.}
 
@@
 
-This theorem has applications in diverse areas, including queue layouts  \citep{DJMMUW20}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, $p$-centered colouring  \citep{DFMS21}, and adjacency labelling \citep{DEJGMM21}, which we explore in \cref{Applications}.
+This theorem has applications in diverse areas, which we explore in \cref{Applications}.
+
+% This theorem has applications in diverse areas, including queue layouts  \citep{DJMMUW20}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, $p$-centered colouring  \citep{DFMS21}, and adjacency labelling \citep{DEJGMM21}, which we explore in \cref{Applications}.
 For example, we prove that $k$-planar graphs have bounded non-repetitive chromatic number (for fixed $k$).
 Prior to the recent work of \citet{dujmovic.esperet.ea:planar}, it was even open whether planar graphs have bounded non-repetitive chromatic number.
-
-\referee{1}{3. pages 2 – 4: The paragraph following Theorem 2 and Section 1.3
-(except for the fact that the apex assumption is necessary) is repetition
-of things discussed earlier on in the introduction.}
-
-\note{PM}{Kind of, but I suggest we keep that paragraph and section anyway.
+%
+% \referee{1}{3. pages 2 – 4: The paragraph following Theorem 2 and Section 1.3
+% (except for the fact that the apex assumption is necessary) is repetition
+% of things discussed earlier on in the introduction.}
+%
+% \note{PM}{Kind of, but I suggest we keep that paragraph and section anyway.
 Justification: They're both short and they describe the exact bounds for bounded genus graphs, which you don't get from the one-paragraph description in Section~1.0.}
-
-\note{DW}{I suggest we write, ``This theorem has applications in diverse areas, which we explore in \cref{Applications}....''}
+%
+% \note{DW}{I suggest we write, ``This theorem has applications in diverse areas, which we explore in \cref{Applications}....''}
 
@@
 By removing repeated vertices this gives a path in the directed graph $\overrightarrow{H}^+$ obtained by directing each edge $xy\in E(H^+)$ from its $T$-descendant $x$ towards its $T$-ancestor $y$.
-  Finally, we are in a position to appeal to \cite[Lemma~24]{PS21} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
-
-  \note{DW}{replace Lemma~24 by Lemma~13, right?}
+  Finally, we are in a position to appeal to \cite[Lemma~13]{PS21} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
+
+  % \note{DW}{replace Lemma~24 by Lemma~13, right?}
+  % \note{PM}{Right. Changed.}
 

### 2022-03-11 k-planar.tex
Reworked opening paragraphs. We don't need to sell so hard.

@@
 
-\note{PM}{Agreed. Done.}
+% \note{PM}{Agreed. Done.}
 
@@
 We use $G-S$ as a shorthand for $G[V(G)\setminus S]$.
-Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (improved by \citet{UWY}).
+Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (with an improvement by \citet{UWY}).
 A graph $G$ is \defin{contained in} a graph $X$ if $G$ is isomorphic to a subgraph of $X$.
@@
 \item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
-\item \citet{DFMS21} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs.
-\item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results starting with the work of \citet{kannan.naor.ea:implicit}).
-\citet{DEJGMM21} have since used it to find asymptotically optimal adjacency labellings of planar graphs.
-\item The result of \citet{DEJGMM21} implies that, for every integer
-$n>0$, there is a `induced-universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
-subgraph of $U_n$.
+\item \citet{DEJGMM21} have since used it to find asymptotically optimal adjacency labellings of planar graphs (resolving a problem of \cite{kannan.naor.ea:implicit}).
+\item \citet{EJM} use it to show the existence of a 'universal graph' with $n^{1+o(1)}$ vertices and edges that contains every $n$-vertex planar graph as an induced subgraph (resolving a problem of \citet{babai.chung.ea:on}).
+\end{compactitem}
+In addition to these rather well-known problems, \cref{PlanarProduct} has been used to resolve or make substantial progress on a number of other problems on planar graphs, including $p$-centered colouring \cite{DFMS21}, $\ell$-vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, and odd colourings~\citep{DMO}.
+
+% \item \citet{DFMS21} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs.
+% \item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results starting with the work of \citet{kannan.naor.ea:implicit}).
+% \item \citet{BDJM} use it to show that every $n$-vertex planar graph has local vertex ranking using $O(\log n/\log\log\log n)$ colours.
+% % \item The result of \citet{DEJGMM21} implies that, for every integer
+% $n>0$, there is a `induced-universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
+% subgraph of $U_n$.
 This has recently been extended by \citet{EJM}, who show the existence of a `subgraph-universal' graph with $(1+\epsilon)n$ vertices that contains every $n$-vertex planar graph as a subgraph as well as the existence of an `induced universal' graph with $n^{1+o(1)}$ vertices \emph{and} edges.
@@
 % aking the first progress on this problem since the work of \citet{babai.xx} in 1984.
-\end{compactitem}
-
-\note{DW}{Mention other application such as vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, odd colourings~\citep{DMO}?
+% \end{compactitem}
+
+% \note{DW}{Mention other application such as vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, odd colourings~\citep{DMO}?
 We write a lot about adjacency labellings and universal induced subgraphs.

### 2022-03-11 k-planar.tex
Commenting out some conversation
Commenting out notes

@@
 
+% \note{PM}{Agreed. I cut this down to a list of famous open problem. We don't need to sell this so hard anymore.}
+
 All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this definition.} These applications motivate finding product structure theorems for other graph classes.
@@
 
-\referee{2}{Page 3. $P$ is used twice in the statement of Theorem 3.  It is probably
+% \referee{2}{Page 3. $P$ is used twice in the statement of Theorem 3.  It is probably
 safer (and necessary?) to use a different variable for the second
@@
 
-\note{DW}{Use $\SS$ for shortcut system, okay?}
-\note{PM}{Ok.  I've implemented this change.  I was careful, but will still reread.}
+% \note{DW}{Use $\SS$ for shortcut system, okay?}
+% \note{PM}{Ok.  I've implemented this change.  I was careful, but will still reread.}
 

### 2022-03-11 k-planar.tex
Always compile before committing!
Pfffft

@@
 \subsection{$k$-Planar Graphs}
+\label{k_planar_section}
 
@@
 % \referee{2}{Page 3. $P$ is used twice in the statement of Theorem 3.  It is probably
-safer (and necessary?) to use a different variable for the second
-occurrence.}
+% safer (and necessary?) to use a different variable for the second
+% occurrence.}
 
@@
 
-\note{DW}{Use $\SS$ for the given shortcut system and $\PP$ for the partition?}
-\note{PM}{Yes, I've done this.  Again, I was carefuly, but need to reread.}
+% \note{DW}{Use $\SS$ for the given shortcut system and $\PP$ for the partition?}
+% \note{PM}{Yes, I've done this.  Again, I was carefuly, but need to reread.}
 
@@
 When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
-A \defin{crossing} in an embedded graph $G$ is a triple $(p,vw,xy)$ with $p\in\R^2$, $vw,xy\in E(G)$ and such that $p\in (vw\cap xy)\setminus\{v,w,x,y\}$.
-An embedded graph $G$ is \defin{$k$-plane} if each edge of $G$ takes part in at most $k$ crossings.
-\note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}\note{PM}{No, under the current definition, $r$ edges crossing at a single point $p$ generate $\binom{r}{2}$ crossings, with each edge taking part in $r-1$ crossings, just like they would if we perturbed the edges so that each pair crosses at a distinct point.
+A \defin{crossing} in an embedded (multi)graph $G$ is a pair of distinct edges $vw,xy\in E(G)$ whose interior have a point in common.
+Without any loss of generality we may assume that the interiors of any two edges of $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.
+This assumption can be enforced by a local change that does not change the number of crossings each edge is involved in.
+If $p$ is a common point in the interior of two crossing edges $vw$ and $xy$ then we say that $vw$ and $xy$ \defin{cross at} $p$.
+
+% (This assumption is already present in the informal definition of $k$-planar graphs used in \cref{k_planar_section}.)
+
+%
+% \note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}\note{PM}{No, under the current definition, $r$ edges crossing at a single point $p$ generate $\binom{r}{2}$ crossings, with each edge taking part in $r-1$ crossings, just like they would if we perturbed the edges so that each pair crosses at a distinct point.
 I guess we could mention this.
@@
 	Let $G$ be a $k$-plane graph.
-We assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$.
+    % We assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$.
 This assumption is justified since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
-
 	As in the proof of \cref{AddDummy}, let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$.
@@
 
-\note{PM}{Include in note to referee:  We choose to keep this section because
-  $1$-planar graphs are of special interest to the graph drawing community.
+% \note{PM}{Include in note to referee:  We choose to keep this section because
+%   $1$-planar graphs are of special interest to the graph drawing community.
 For example, the annotated bibliography on $1$-planarity (\htmladdnormallink{https://arxiv.org/abs/1703.02261}{arXiv:1703.02261}) has 143 entries, and even this is now four years out of date.
-
-  An important aspect of our result for $1$-planar graphs is that graph $H$ in the product is planar and has treewidth $3$, i.e., it has simple treewidth $3$.
+%
+%   An important aspect of our result for $1$-planar graphs is that graph $H$ in the product is planar and has treewidth $3$, i.e., it has simple treewidth $3$.
 In a more recent application of product structure \citet{BDJM} show that such graphs have $\ell$-vertex rankings using $O(\log n/\log\log\log n)$ colours, and this is tight.
 This result requires that $H$ have simple treewidth $3$, even treewidth $3$ is not sufficient.
-}
-
-\note{DW}{I will rewrite the intro to this subsection once we decide what our results are.}
+% }
+%
+% \note{DW}{I will rewrite the intro to this subsection once we decide what our results are.}
 
@@
 Let $G$ be a $1$-plane graph.
-We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add $vw$, embedded in the face $F$.
+We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.
 This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.
@@
 
-To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
+To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ that cross at $p$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
 Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$.
@@
 
-Consider some crossing $(p,vw,xy)$ in $G$.
-The point $p$ at which this crossing occurs is a vertex of $G_0$ that has degree $4$.
-There are four faces $F_1,\ldots,F_4$ of $G_0$ each with $p$ on their boundary.
-Each of these four faces of is of Type~(ii), so the boundary of $F_1\cup\cdots\cup F_4$ includes a $4$-cycle $C:=vxwy$ in $G$ whose edges are not involved in any crossings.
+Consider some pair of edges $vw$ and $xy$ that cross at $p$.
+The point $p$ is a vertex of $G_0$ that has degree $4$.
+There are four faces $F_1,\ldots,F_4$ of $G$ each with $p$ on their boundary.
+Each of these four faces of $G$ is of Type~(ii), so the boundary of $F_1\cup\cdots\cup F_4$ includes a $4$-cycle $C:=vxwy$ in $G$ whose edges are not involved in any crossings.
 

### 2022-03-11 k-planar.tex
comments about framed graphs

@@
 \begin{thm}
-\label{1-planar}
+\label{framed}
     Every $d$-framed graph is contained in $H\boxtimes P\boxtimes K_{3\lfloor d/2\rfloor + d -3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
 \end{thm}
+
+\note{DW}{I think \cref{framed} should be proved explicitly and \cref{1-planar} should be a corollary, especially because \cref{NewMapGraph} is another corollary.
+This section should be about framed graphs, outside of \cref{sec-k-planar}.
+I suggest we make it a subsection of
+\cref{Structure} where we introduce framed graphs, say they can be described by $(2,\binom{d}{2})$-shortcut systems, so the main theorem is applicable, which gives a treewidth bound of $\binom{5}{3}-1=9$, but using  \cref{framed} we get a treewidth bound of 3, which is optional.
+Then in latter sections we give two applications, for 1-planar graphs and for $d$-map graphs.
+}
 
@@
 \appendix
-\section{New Stuff}
+\section{Framed Graphs}
+
+\note{DW}{Use the language of framed graphs}
 
@@
 Let $G$ be a 2-connected plane multigraph, in which every facial cycle has length in $\{3,\dots,d\}$ for some fixed integer $d\geq 3$.
-Then $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d  \floor{d/2} }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
+Then $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d +3( \floor{d/2}-1) }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
 \end{thm}
@@
 Let $T$ be a BFS spanning tree of $G^+$ rooted at $r$.
-Let $\mathcal{L}$ be the corresponding BFS layering.
-By \cref{NewLemma} \note{DW}{with improved (ii)}, $G$ has a partition $\PP$ into tripods with respect to $T$ such that $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
+Let $\mathcal{L}=(L_0,L_1,\dots)$ be the corresponding BFS layering.
+By \cref{NewLemma}, $G$ has a partition $\PP$ into tripods with respect to $T$ such that $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ for which every part of $\PP$ that intersects $F$ is in $B$.
 Since $G$ is a spanning subgraph of $\widehat{G}$, we may consider $\PP$ to be a partition of $\widehat{G}$, and $G/\PP \subseteq \widehat{G}/\PP$.
@@
 In particular, $X,Y\in B$.
-Hence $\TT$ is also a tree-decomposition of $\widehat{G}$.
-Each tripod in $\PP$ has at most $d$ vertices in each layer in $\mathcal{L}$.
+Hence $\TT$ is also a tree-decomposition of $\widehat{G}/\PP$.
 For each edge $vw$ of $\widehat{G}$, if $v\in L_i$ and $w\in L_j$ then $|i-j|\leq\floor{\frac{d}{2}}$.
-Grouping blocks of $\floor{\frac{d}{2}}$ consecutive layers gives a layering of $\widehat{H}$ with layered width $\floor{\frac{d}{2}}d$ with respect to $\PP$.
-By \cref{PartitionProduct}, $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d  \floor{d/2} }$ where $H:=\widehat{G}/\PP$, which has treewidth 3.
+Grouping blocks of $\floor{\frac{d}{2}}$ consecutive layers in $\mathcal{L}$ gives a layering $\mathcal{L}'$ of $\widehat{G}$.
+Observe that each tripod in $\PP$ has at most $d+3( \floor{\frac{d}{2}}-1)$ vertices in each set of $\floor{\frac{d}{2}}$ layers of $\mathcal{L}$.
+Hence $\mathcal{L}'$ has layered width $d+3( \floor{\frac{d}{2}}-1)$ with respect to $\PP$.
+By \cref{PartitionProduct}, $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d + 3( \floor{d/2} -1 )}$ where $H:=\widehat{G}/\PP$, which has treewidth 3.
 \end{proof}

### 2022-03-11 k-planar.tex
Last changes. Kids are here!

--- 2022-03-11 k-planar.tex
@@
 The points $f(0)$ and $f(1)$ are called the \defin{endpoints} of the curve.
-When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
-A \defin{crossing} in an embedded (multi)graph $G$ is a pair of distinct edges $vw,xy\in E(G)$ whose interior have a point in common.
+When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.
+The \emph{interior} of $f$ is the point set $\{f(t):0<t<1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
+A \defin{crossing} in an embedded (multi)graph $G$ is an unordered pair of distinct edges $\{vw,xy\}\subseteq E(G)$ whose interiors have a point in common.
 Without any loss of generality we may assume that the interiors of any two edges of $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.
-This assumption can be enforced by a local change that does not change the number of crossings each edge is involved in.
+This assumption can be enforced by local changes that do not change the number of crossings each edge is involved in.
 If $p$ is a common point in the interior of two crossing edges $vw$ and $xy$ then we say that $vw$ and $xy$ \defin{cross at} $p$.
@@
 Therefore each dummy vertex $w_i$ of $W$ is surrounded by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
-Since $w_i$ has degree $4$ in $G_0$, $w_{i-1}\neq w_{i+1}$.
+Since $w_i$ has degree $4$ in $G_0$ and $F$ does not have two parallel edges on its boundary, $w_{i-1}\neq w_{i+1}$.
 Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.
@@
 
-Therefore each face of $G_0$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
-\begin{inparaenum}[(i)]
-  \item bounded by three edges of $G$, none of which are crossed; or
-  \item bounded by one edge of $G$ that is not crossed plus portions of two edges of $G$ that cross each other.
-\end{inparaenum}
+% Therefore each face of $G_0$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
+% \begin{inparaenum}[(i)]
+%   \item bounded by three edges of $G$, none of which are crossed; or
+%   \item bounded by one edge of $G$ that is not crossed plus portions of two edges of $G$ that cross each other.
+% \end{inparaenum}
 
@@
 The point $p$ is a vertex of $G_0$ that has degree $4$.
-There are four faces $F_1,\ldots,F_4$ of $G$ each with $p$ on their boundary.
-Each of these four faces of $G$ is of Type~(ii), so the boundary of $F_1\cup\cdots\cup F_4$ includes a $4$-cycle $C:=vxwy$ in $G$ whose edges are not involved in any crossings.
-
-By removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
-Since, for any pair of crossing edges $(vw,xy)$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.
+There are four faces $F_1,\ldots,F_4$ of $G_0$ that contain $p$ and, from the preceding discussion each of $vx$, $xw$, $wy$, and $yv$ is an uncrossed edge of $G$.
+Therefore, by removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
+Since, for any pair of crossing edges $\{vw,xy\}$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.
 % It follows immediately that, for any pair of vertices $a,b\in V(G)$, $\dist_{G'}(a,b)\le 2\dist_G(a,b)$.
@@
 		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
-		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that $V(F)$ can be partitioned into (possibly empty) sets $P_1,\ldots,P_3$ such that for each $i\in\{1,2,3\}$,
+		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that $V(F)$ can be partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
 		\begin{compactenum}
@@
 		\end{compactenum}
-		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
+		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
 	\end{compactenum}
@@
 		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 3$;
-		\item for each $i\in\{1,2,3\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$;
+		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
 		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
@@
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
-Now add one edge to each $4$-sided inner face of $N'$ to obtain a plane supergraph $N''$ of $N'$ in which each inner face has exactly three $3$ edges on its boundary.
+Now add one edge to each $4$-sided inner face of $N'$ to obtain a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary.
 By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
@@
 Therefore $|S\cap L_j|\le 3$ for each integer $j\ge 0$.
-Therefore, taking $Y:=S\cup\{v\}\setminus V(F)$ as a part in the partition $\PP$ will not violate Condition~(ii).
+Therefore, taking $Y:=S\cup\{v\}\setminus V(F)$ as a part in the partition $\PP$ will not violate condition~(ii).
 
@@
 Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$.
-Since $N'\subseteq G'$ contains only edges of $G$ that are not crossed, each component of $G-F-Y$ is contained in a single face of $M$.
+Since $N'\subseteq G'$ contains only edges of $G$ that are not crossed, each component of $G-V(M)$ is contained in a single face of $M$.
+
+  \begin{figure}
+    \begin{center}
+      \begin{tabular}{c@{\hspace{1cm}}c}
+        \includegraphics{figs/mmplus-1} &
+        \includegraphics{figs/mmplus-2} \\
+        $k\le 2$ & $k=3$
+      \end{tabular}
+    \end{center}
+    \caption{Two cases when choosing the face $\tau$ of $N'$.}
+    \label{boring_figure}
+  \end{figure}
 
@@
 By our choice of $\tau$, the graph $M$ is $2$-connected.
-Furthermore, $V(F_i)$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset V(Y)$, $P_{i,2}\subseteq P_b$, $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
+Furthermore, $V(F_i)$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
 In the following, in order to avoid introducing even more notation or constantly adding the modifier ``when it exists'', we will assume that $V(F_i)\cap Y\neq\emptyset$ and that there are exactly two distinct $b,c\in\{1,\ldots,k\}$ such that $V(F_i)\cap P_b\neq\emptyset$ and $V(F_i)\cap P_c\neq\emptyset$.
@@
 
-  \begin{figure}
-    \begin{center}
-      % \begin{tabular}{c@{\hspace{1cm}}c}
-        % \includegraphics{figs/mmplus-1} &
-        \includegraphics{figs/mmplus-2}
-      % \end{tabular}
-    \end{center}
-    \caption{The `Sperner' triangle $v_1v_2v_3$ in $N''$ and the Sperner face $\tau$ in $N'$.}
-    \label{boring_figure}
-  \end{figure}
-
   Since $M$ is $2$-connected and $\overline{Y}$ is connected, $F_i[P_{i,j}]$ is a path, for each $j\in\{1,2,3\}$.
+Let $N_i$ and $N_i'$ be the subgraphs of $N$ and $N'$ contained in the closure of $F_i$.
+Every inner face of $N_i'$ is an inner face of $N'$ and $\tau$ is not an inner face of $N_i'$.
+Therefore $N_i'$ has fewer inner faces than $N'$.
 Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.
-The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of the subgraph $N_i$ of $N$ contained in $F_i$.
-The partition $\mathcal{P}_i$ contains each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$.
-For $\mathcal{P}$ we use the partition
+The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of the subgraph $N_i$ of $N$ contained in the closure of $F_i$.
+  % The partition $\mathcal{P}_i$ contains each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$.
+  We can now define our desired partition
   \[
@@
 
-    The tree $T_0$ has root node $y_0$ with $B_{y_0}:=\{x_0,\ldots,x_k\}$.
+    The tree $T_0$ has a node $y_0$ with $B_{y_0}:=\{x_0,\ldots,x_k\}$.
     % For each $i\in\{1,\ldots,m\}$, let $N_i$ denote the subgraph of $N$ contained in $F_i$.
-    The inductive hypothesis ensures that $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $(B_z:z\in V(T_i))$ whose largest bag has size at most $4$ whose root bag $B_{z_i}$ contains the vertices $a_i,b_i,c_i\in V(H_i)$ obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
+    The inductive hypothesis ensures that $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $(B_z:z\in V(T_i))$ whose largest bag has size at most $4$ and that has some bag $B_{z_i}$ that contains the vertices $a_i,b_i,c_i\in V(H_i)$ obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
 Recall that $P_{i,1}:=V(N_i)\cap Y$, $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
 For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.
-Note that these replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$.
-Now add an edge that joins $y_0$ to $x_i$.
-
-    It is now straightforward to verify that doing this for each $i\in\{1,\ldots,m\}$ yields a tree decomposition of $N$ that satisfies condition~(iv).
-(The key observation is that there is no edge of $N$  with one endpoint in $N_i-F_i$ and one endpoint in $N_j-F_j$ for any $i\neq j$.)
+These replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$.
+Now add an edge that joins $y_0$ to $z_i$.
+
+    Using the fact that each component of $G-V(M)$ is contained in a single face of $M$, it is now straightforward to verify that, by doing this for each $i\in\{1,\ldots,m\}$ we obtain a tree decomposition of $N$ that satisfies condition~(iv).
+\qedhere
   \end{compactenum}

### 2022-03-11 k-planar.tex
more comments about improving to tw 3
added comments

@@
 \usepackage[longnamesfirst,numbers,sort&compress]{natbib}
+
 \usepackage[noabbrev,capitalise]{cleveref}
+\crefname{lem}{Lemma}{Lemmas}
+\crefname{thm}{Theorem}{Theorems}
+\crefname{cor}{Corollary}{Corollaries}
+\crefname{prop}{Proposition}{Propositions}
+\crefname{conj}{Conjecture}{Conjectures}
+\crefname{open}{Open Problem}{Open Problems}
+\crefname{claim}{Claim}{Claims}
+\crefname{enumi}{Item}{Items}
+\crefname{obs}{Observation}{Observations}
+\crefname{clm}{Claim}{Claims}
+\crefname{openproblem}{Open Problem}{Open Problems}
+\crefformat{equation}{(#2#1#3)}
+\Crefformat{equation}{Equation #2(#1)#3}
 
@@
 \newcommand{\defin}[1]{\textcolor{Maroon}{\emph{#1}}}
-
-\crefname{lem}{Lemma}{Lemmas}
-\crefname{thm}{Theorem}{Theorems}
-\crefname{cor}{Corollary}{Corollaries}
-\crefname{conj}{Conjecture}{Conjectures}
-\crefname{obs}{Observation}{Observations}
-\crefname{prop}{Proposition}{Propositions}
-\crefname{clm}{Claim}{Claims}
-\crefname{openproblem}{Open Problem}{Open Problems}
-\crefformat{equation}{(#2#1#3)}
-\Crefformat{equation}{Equation #2(#1)#3}
 
@@
 \note{DW}{I think \cref{framed} should be proved explicitly and \cref{1-planar} should be a corollary, especially because \cref{NewMapGraph} is another corollary.
+The proof of \cref{framed} needs an extra clause; see the last sentence of \cref{NewLemma}.
 This section should be about framed graphs, outside of \cref{sec-k-planar}.
-I suggest we make it a subsection of
-\cref{Structure} where we introduce framed graphs, say they can be described by $(2,\binom{d}{2})$-shortcut systems, so the main theorem is applicable, which gives a treewidth bound of $\binom{5}{3}-1=9$, but using  \cref{framed} we get a treewidth bound of 3, which is optional.
+I suggest we make it a subsection of \cref{Structure} where we introduce framed graphs, say they can be described by $(2,\binom{d}{2})$-shortcut systems, so the main theorem is applicable, which gives a treewidth bound of $\binom{5}{3}-1=9$, but using  \cref{framed} we get a treewidth bound of 3, which is optional.
 Then in latter sections we give two applications, for 1-planar graphs and for $d$-map graphs.
@@
 \end{thm}
+
+\note{DW}{We should be able to get treewidth 3 for $(g,1)$-planar graphs, combining the method in \cref{1-planar} with \citep{DHHW}}
 
@@
 
+\note{DW}{We should be able to get treewidth 3 here using \cref{framed} and \citep{DHHW}.}
+
 These results give the first constant upper bound on the non-repetitive chromatic number of map graphs, the first polynomial bounds on the $p$-centered chromatic number of map graphs, and the best known bounds on the queue-number of map graphs.

### 2022-03-11 k-planar.tex
minor updates

@@
 
-\note{DW}{We should be able to get treewidth 3 for $(g,1)$-planar graphs, combining the method in \cref{1-planar} with \citep{DHHW}}
+\note{DW}{I wonder if one can get treewidth 3 for $(g,1)$-planar graphs, combining the method in \cref{1-planar} with \citep{DHHW}. Perhaps this should be in another paper?}
 
@@
 
-\note{DW}{We should be able to get treewidth 3 here using \cref{framed} and \citep{DHHW}.}
+\note{DW}{I wonder if one can get treewidth 3 here using \cref{framed} and \citep{DHHW}. 
+Perhaps this should be in another paper?}
 
@@
 \label{NewMapGraph}
-Every $d$-map graph is contained in $H \boxtimes P \boxtimes K_{ d  \floor{d/2} }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
+Every $d$-map graph is contained in $H \boxtimes P \boxtimes K_{ d + 3\floor{d/2} -3 }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
 \end{cor}
@@
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%
-
-The layering is in G_0. Each edge of G_2 has span at most floor(d/2)
-with respect to this layering.
-
-So G_2 is contained in H * P^{floor(d/2)} * K_d, which is contained in
-H * P * K_{d floor(d/2) }.
-
-\end{document}

### 2022-03-12 k-planar.tex
added comments

@@
 
-\note{DW}{I wonder if one can get treewidth 3 for $(g,1)$-planar graphs, combining the method in \cref{1-planar} with \citep{DHHW}. Perhaps this should be in another paper?}
+\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{framed} with the treewidth 3 result in \citep{DHHW}.
+I think the best place for this result is reference \citep{DHHW} (since the present paper pre-dates \citep{DHHW} and this addition will increase the significance of \citep{DHHW}).
+}
 
@@
 
-\note{DW}{I wonder if one can get treewidth 3 here using \cref{framed} and \citep{DHHW}. 
-Perhaps this should be in another paper?}
+\note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the treewidth 3 result in \citep{DHHW}.
+I think the best place for this result is reference \citep{DHHW} (since the present paper pre-dates \citep{DHHW} and this addition will increase the significance of \citep{DHHW}).}
 

### 2022-03-12 k-planar.tex
added comments

@@
 \note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{framed} with the treewidth 3 result in \citep{DHHW}.
-I think the best place for this result is reference \citep{DHHW} (since the present paper pre-dates \citep{DHHW} and this addition will increase the significance of \citep{DHHW}).
+I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
 }
@@
 \note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the treewidth 3 result in \citep{DHHW}.
-I think the best place for this result is reference \citep{DHHW} (since the present paper pre-dates \citep{DHHW} and this addition will increase the significance of \citep{DHHW}).}
+I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).}
 

### 2022-03-12 k-planar.tex
Pfffft

@@
   \begin{compactenum}[(i)]
-    \item To see that $H$ is planar, observe that $G/\mathcal{P}=G'/\mathcal{P}$\todo{Explain why $G/\mathcal{P}=G'/\mathcal{P}$} and that each part $P\in\mathcal{P}$ induces a connected subgraph $G'[P]$ of $G'$.
+    \item To see that $H$ is planar, observe that $N/\mathcal{P}=N'/\mathcal{P}$\todo{Explain why $N/\mathcal{P}=N'/\mathcal{P}$} and that each part $P\in\mathcal{P}$ induces a connected subgraph $G'[P]$ of $G'$.
 Since $G'$ is planar and planarity is preserved under edge contractions, this implies that $G'/\mathcal{P}$ is planar.
@@
 
-We remark that our proof of \cref{1-planar} generalizes to the class of $d$-framed graphs \cite{BDGGMR}.
-An embedded multigraph $G$ is a \defin{$d$-framed graph} if it contains a plane subgraph $G'$ (the \defin{frame}) whose edges are not involved in any crossings and such that each pair of crossing edges in $G$ are contained in a face of $G'$ that has at most $d$ vertices on its boundary.
-It is known that the class of $4$-framed multigraphs is exactly the class of $1$-plane multigraphs \cite{BDGGMR}.
-Indeed, the graph $G'$ in \cref{induction} is the frame of $G$.
-The proof of \cref{induction} can be repeated for any edge-maximal $d$-framed graph $G$ having frame $G'$.
-The proof in this case has only two differences:
+We remark that our proof of \cref{1-planar} generalizes to the class of $d$-framed graphs \cite{BDGGMR} and even $d$-framed multigraphs.
+An embedded multigraph $G$ is a \defin{$d$-framed multigraph} if it contains a plane subgraph $G'$ (the \defin{frame}) whose edges are not involved in any crossings and such that each pair of crossing edges in $G$ are contained in a face of $G'$ that has at most $d$ vertices on its boundary.
+It is known that every $3$-connected $1$-planar graph is contained in some $4$-framed graph.
+The discussion on the structure of $G$ and $G'$ that precedes \cref{induction} shows that every $1$-planar graph is contained in a $4$-framed multigraph.
+The proof of \cref{induction} can be repeated for any edge-maximal $d$-framed multigraphs $G$ having frame $G'$. The proof in this case has only two differences:
 \begin{compactitem}
@@
 \label{framed}
-   Every $d$-framed graph is contained in $H\boxtimes P\boxtimes K_{3\lfloor d/2\rfloor + d -3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+   Every $d$-framed multigraph is contained in $H\boxtimes P\boxtimes K_{3\lfloor d/2\rfloor + d -3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
 \end{thm}
@@
 }
+
+\note{PM}{I no longer agree about the restructuring. See my email essay for an explanation.}
+
+\note{PM}{Apropos the extra clause.
+It's related to the orange todo note (argue that $N'/\mathcal{P}=N/\mathcal{P}$).
+This doesn't need to be in the statement of the lemma since, unlike \cref{NewLemma} this one gives an $H$-partition of $N$, not $N'$, but something like it needs to be in the proof.}
 

### 2022-03-12 k-planar.tex
Tidying up

@@
 	\end{compactenum}
-	Then $N$ has an $H$-partition $\PP=\{S_x\cup\{v_x\} : x\in V(H)\}$ such that:
+	Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
 	\begin{compactenum}[(i)]
-		\item $H$ is planar;
-		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 3$;
 		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
-		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
+        \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le 1$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
+        \item $H$ has a plane embedding in which $x_1,\ldots,x_k$ appear on the outer face;
+		\item $\tw(H)\le 3$.
 	\end{compactenum}
@@
   The proof is by induction on the number of inner faces of $N'$.
-If $N'$ has only one inner face then we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
-  \begin{inparaenum}[(i)]
-    \item $H:=N/\mathcal{P}$ is planar because any graph on $k\le 3$ vertices is planar.
-    \item Assumption 4(a) ensures that $|S_x\cap L_i|\le 3$ for each $x\in V(H)$.
-    \item By definition, each each of $P_1,\ldots,P_k$ is s part of $\mathcal{P}$.
-    \item The trivial tree decomposition of $H$ that places $x_1,\ldots,x_k$ in one bag satisfies this requirement.
-  \end{inparaenum}
-  Therefore, we may assume that $N$ has more than one inner face.
-
-  The $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a part $Y$, called a \emph{tripod}, that we now define.
-The tripod $Y$ is determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
+A particular easy case occurs when $V(N')=V(F)$, for which we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
+  \begin{compactenum}[(i)]
+    \item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
+    \item For each $x\in V(H)$, let $Y_x:=S_x$ and $X_x=\emptyset$.  Then 4(a) ensures that $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
+    \item $H$ has $k\le 3$ vertices and any graph on $k\le 3$ vertices has a plane embedding with all its vertices on the outer face.
+    \item The trivial tree decomposition of $H$ that places $x_1,\ldots,x_k$ in one bag has width $k-1\le 2\le 3$.
+  \end{compactenum}
+  From this point on, we assume that $N$ has at least one inner vertex.
+
+  The $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
 Depending on the value of $k$, there are two cases to consider:
@@
 
-  We can now define the vertices in the tripod $Y$.
+  We can now define the vertices in $S$.
 For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to a vertex in $V(F)$.
-Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
-Let $S:=V(Q_1)\cup V(Q_2)\cup V(Q_3)$ and let $v:=v_p$.
+Let $\overline{S}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
+Let $S:=V(\overline{S})\setminus V(F)$, let $Y:=(V(Q_1)\cup V(Q_2)\cup V(Q_3))\cap S$, and let $X:=S\setminus Y$.
   Observe that, for each $i\in\{1,2,3\}$ and each $j\ge 0$, $|V(Q_i)\cap L_j|\le 1$.
-Therefore $|S\cap L_j|\le 3$ for each integer $j\ge 0$.
+Therefore $|Y\cap L_j|\le 3$ for each integer $j\ge 0$.
+Furthermore, $|X|\le p-3\le 1$, since $v_1,v_2,v_3\in Y\cup V(N)$.
 Therefore, taking $Y:=S\cup\{v\}\setminus V(F)$ as a part in the partition $\PP$ will not violate condition~(ii).
@@
 Furthermore, $V(F_i)$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
+
   In the following, in order to avoid introducing even more notation or constantly adding the modifier ``when it exists'', we will assume that $V(F_i)\cap Y\neq\emptyset$ and that there are exactly two distinct $b,c\in\{1,\ldots,k\}$ such that $V(F_i)\cap P_b\neq\emptyset$ and $V(F_i)\cap P_c\neq\emptyset$.
@@
   \begin{compactenum}[(i)]
-    \item To see that $H$ is planar, observe that $N/\mathcal{P}=N'/\mathcal{P}$\todo{Explain why $N/\mathcal{P}=N'/\mathcal{P}$} and that each part $P\in\mathcal{P}$ induces a connected subgraph $G'[P]$ of $G'$.
-Since $G'$ is planar and planarity is preserved under edge contractions, this implies that $G'/\mathcal{P}$ is planar.
+    \item By construction $\mathcal{P}$ contains parts $P_1,\ldots,P_k$.
 
     \item  By Assumption~4(b) none of the parts $P_1,P_2,P_3$ violate condition~(ii).
-We have already argued that the part $Y$ does not violate condition~(ii).
+We have already argued that the part $S$ does not violate condition~(ii).
 The inductive hypothesis ensures that each of the remaining parts satisfies condition~(ii).
 
-    \item Condition~(iii) is satisfied since $\mathcal{P}$ contains parts $P_1,\ldots,P_k$.
+    \item We only consider the (most difficult) case where $k=3$, the cases $\in\{1,2\}$ are handled easily.
+Let $x_0$ be the vertex of $H$ obtained from contracting $Y$ and, for each $j\in\{1,\ldots,3\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
+In this case the vertices $x_0,\ldots,x_3$ form a clique in $H$, which we embed with $x_1,\ldots,x_3$ on the outer face, as required.
+
+    We claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.
+Suppose to the contrary that there are two such faces $F_{i_1}$ and $F_{i_2}$.
+Create a graph $M^+$ from $M$ by placing a vertex $v_{i_b}$ inside $F_{i_b}$ and adjacent to each vertex in $V(F_{i_b})$ for each $b\in\{1,2\}$.
+Since $M$ is planar and $F_{i_1}$ and $F_{i_2}$ are distinct faces of $M$, $M^+$ is planar.
+However, contracting each of $S_{x_0},\ldots,S_{x_3}$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.
+
+    Therefore, by the inductive hypothesis, we can embed each such graph $H_i :=N_i/\mathcal{P}_i$ so that the three vertices on its outer face correspond to the appropriate inner face of the clique on $x_0,\ldots,x_3$.
+Any remaining graphs $H_i$ have outer faces with at most two vertices on their outer faces corresponding to at most two vertices of $x_0,\ldots,x_3$.
+These can easily be embedded, one a time, in some inner face that has the edge $x_ix_j$ on its boundary.
 
@@
     % For each $i\in\{1,\ldots,m\}$, let $N_i$ denote the subgraph of $N$ contained in $F_i$.
-    The inductive hypothesis ensures that $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $(B_z:z\in V(T_i))$ whose largest bag has size at most $4$ and that has some bag $B_{z_i}$ that contains the vertices $a_i,b_i,c_i\in V(H_i)$ obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
+    The inductive hypothesis ensures that $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $(B_z:z\in V(T_i))$ of width at most $3$.
+Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
+Since $a_i,b_i,c_i$ form a clique in $H_i$, some bag $B_{z_i}$ contains $a_i$, $b_i$, and $c_i$.
+
     Recall that $P_{i,1}:=V(N_i)\cap Y$, $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
@@
 
-    Using the fact that each component of $G-V(M)$ is contained in a single face of $M$, it is now straightforward to verify that, by doing this for each $i\in\{1,\ldots,m\}$ we obtain a tree decomposition of $N$ that satisfies condition~(iv).
+    Using the fact that each component of $G-V(M)$ is contained in a single face of $M$, it is now straightforward to verify that, by doing this for each $i\in\{1,\ldots,m\}$ we obtain a width-$3$ tree decomposition of $N$.
 \qedhere
@@
 
-	Let $G'$ be the plane graph obtained by removing each pair of crossing edges from $\overline{G}$ and let $T$ be a BFS tree of $G'$ rooted at $r$.
+	Let $G'$ be the plane graph obtained by removing each pair of crossing edges from $\overline{G}$ and let $T$ be a BFS tree of $G'$ rooted at $r$ and let $\langle L_0,L_1,\ldots,\rangle$ be the resulting BFS layering.
 Now apply \cref{induction} with $G=\overline{G}$, $G'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
@@
 That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le 2$ for each edge $vw\in E(G)$.
-Since $|L_j\cap S_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap (S_x\cup\{v_x\})|\le 7$ for each integer $i\ge 0$ and each $x\in V(H)$.
+Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 6$ and therefore $|L_i'\cap S_x|\le 7$ for each integer $i\ge 0$ and each $x\in V(H)$.
 The result follows from \cref{PartitionProduct}.

### 2022-03-13 k-planar.tex
A good solid proof for the 1-planar case.

@@
 In the following, we assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$, where we treat each edge $vw$ of $G$ as the image of a simple curve $f_{vw}:[0,1]\to\R^2$.
-We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
+% We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
 
@@
 
+A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $3$-element subset $\{u,v,w\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{u,v,w\}$.
 Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
@@
         \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le 1$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
-        \item $H$ has a plane embedding in which $x_1,\ldots,x_k$ appear on the outer face;
-		\item $\tw(H)\le 3$.
+		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$..
 	\end{compactenum}
@@
     \item For each $x\in V(H)$, let $Y_x:=S_x$ and $X_x=\emptyset$.  Then 4(a) ensures that $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
-    \item $H$ has $k\le 3$ vertices and any graph on $k\le 3$ vertices has a plane embedding with all its vertices on the outer face.
     \item The trivial tree decomposition of $H$ that places $x_1,\ldots,x_k$ in one bag has width $k-1\le 2\le 3$.
   \end{compactenum}
-  From this point on, we assume that $N$ has at least one inner vertex.
-
-  The $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
-Depending on the value of $k$, there are two cases to consider:
+
+  In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
+There is one case to consider for each possible value of $k$:
   \begin{compactenum}
-    \item If $k\in\{1,2\}$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge of $F$ on its boundary.
+    \item If $k=1$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
+
+    \item If $k= 2$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
 
@@
 Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.
-(Note that $p\in\{3,4\}$.)
+% (Note that $p\in\{3,4\}$.)
   \end{compactenum}
@@
 Let $\overline{S}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
-Let $S:=V(\overline{S})\setminus V(F)$, let $Y:=(V(Q_1)\cup V(Q_2)\cup V(Q_3))\cap S$, and let $X:=S\setminus Y$.
+Let $S:=V(\overline{S})\setminus V(F)$, let $Y:=(V(Q_1)\cup V(Q_2)\cup V(Q_3))\setminus V(F)$, and let $X:=S\setminus Y$.
   Observe that, for each $i\in\{1,2,3\}$ and each $j\ge 0$, $|V(Q_i)\cap L_j|\le 1$.
 Therefore $|Y\cap L_j|\le 3$ for each integer $j\ge 0$.
-Furthermore, $|X|\le p-3\le 1$, since $v_1,v_2,v_3\in Y\cup V(N)$.
-Therefore, taking $Y:=S\cup\{v\}\setminus V(F)$ as a part in the partition $\PP$ will not violate condition~(ii).
+Furthermore, $|X|\le p-3\le 1$, since $v_1,v_2,v_3\in Y\cup V(F)$.
+Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ will not violate condition~(ii).
 
@@
 
+  \begin{figure}
+    \begin{center}
+      \begin{tabular}{ccc}
+        \includegraphics{figs/mmplus-1} &
+        \includegraphics{figs/mmplus-2} &
+        \includegraphics{figs/mmplus-3} \\
+        $k=1$ & $k=2$ & $k=3$
+      \end{tabular}
+    \end{center}
+    \caption{Three cases when choosing the face $\tau$ of $N'$.}
+    \label{boring_figure}
+  \end{figure}
+
   Refer to \cref{boring_figure}.
 Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$.
+By our choice of $\tau$, the graph $M$ is $2$-connected.
 Since $N'\subseteq G'$ contains only edges of $G$ that are not crossed, each component of $G-V(M)$ is contained in a single face of $M$.
-
-  \begin{figure}
-    \begin{center}
-      \begin{tabular}{c@{\hspace{1cm}}c}
-        \includegraphics{figs/mmplus-1} &
-        \includegraphics{figs/mmplus-2} \\
-        $k\le 2$ & $k=3$
-      \end{tabular}
-    \end{center}
-    \caption{Two cases when choosing the face $\tau$ of $N'$.}
-    \label{boring_figure}
-  \end{figure}
-
   Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
-By our choice of $\tau$, the graph $M$ is $2$-connected.
-Furthermore, $V(F_i)$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
+By our choice of $v_1v_2v_3$, the vertices of each $F_i$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
 
@@
   It is straightforward to verify that $\mathcal{P}$ is indeed a partition of $V(N)$ and, by definition this is an $H$-partition of $N$ for the graph $H:=N/\mathcal{P}$.
-All that remains is to verify that $\mathcal{P}$ satisfies conditions~(i)--(iv).
-  \begin{compactenum}[(i)]
-    \item By construction $\mathcal{P}$ contains parts $P_1,\ldots,P_k$.
-
-    \item  By Assumption~4(b) none of the parts $P_1,P_2,P_3$ violate condition~(ii).
+All that remains is to verify that $\mathcal{P}$ satisfies (i)--(iii).
+
+  By construction $\mathcal{P}$ contains parts $P_1,\ldots,P_k$, so $\mathcal{P}$ satisifes (i).
+By Assumption~4(b) none of the parts $P_1,P_2,P_3$ violate condition~(ii).
 We have already argued that the part $S$ does not violate condition~(ii).
 The inductive hypothesis ensures that each of the remaining parts satisfies condition~(ii).
-
-    \item We only consider the (most difficult) case where $k=3$, the cases $\in\{1,2\}$ are handled easily.
+Therefore $\mathcal{P}$ satisfies (ii).
+
+    % \item We only consider the (most difficult) case where $k=3$, the cases $\in\{1,2\}$ are handled easily.
 Let $x_0$ be the vertex of $H$ obtained from contracting $Y$ and, for each $j\in\{1,\ldots,3\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
 In this case the vertices $x_0,\ldots,x_3$ form a clique in $H$, which we embed with $x_1,\ldots,x_3$ on the outer face, as required.
-
-    We claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.
+    %
+    % We claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.
 Suppose to the contrary that there are two such faces $F_{i_1}$ and $F_{i_2}$.
@@
 However, contracting each of $S_{x_0},\ldots,S_{x_3}$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.
-
-    Therefore, by the inductive hypothesis, we can embed each such graph $H_i :=N_i/\mathcal{P}_i$ so that the three vertices on its outer face correspond to the appropriate inner face of the clique on $x_0,\ldots,x_3$.
+    %
+    % Therefore, by the inductive hypothesis, we can embed each such graph $H_i :=N_i/\mathcal{P}_i$ so that the three vertices on its outer face correspond to the appropriate inner face of the clique on $x_0,\ldots,x_3$.
 Any remaining graphs $H_i$ have outer faces with at most two vertices on their outer faces corresponding to at most two vertices of $x_0,\ldots,x_3$.
@@
 
-    \item We now construct a tree decomposition $(B_y:y\in V(T_0))$ of $H$ that satisifies condition~(iv).
-    % We show that $\tw(H)\le 3$ by giving a tree-decomposition of $H$ of width at most $3$.
-    Let $x_0$ be the vertex of $H$ obtained from contracting $Y$ and, for each $j\in\{1,\ldots,k\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
-
-    The tree $T_0$ has a node $y_0$ with $B_{y_0}:=\{x_0,\ldots,x_k\}$.
-    % For each $i\in\{1,\ldots,m\}$, let $N_i$ denote the subgraph of $N$ contained in $F_i$.
-    The inductive hypothesis ensures that $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $(B_z:z\in V(T_i))$ of width at most $3$.
+  All that remains is to construct a tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$ of $H$ that satisfies (iii).
+Let $x_0$ be the vertex of $H$ obtained from contracting $S$ and, for each $j\in\{1,\ldots,k\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
+The tree $T_0$ has a node $y_0$ with bag $B_{y_0}:=\{x_0,\ldots,x_k\}$.
+The inductive hypothesis ensures that, for each $i\in\{1,\ldots,m\}$,  $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $(B_z:z\in V(T_i))$ that satisfies (iii).
 Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
-Since $a_i,b_i,c_i$ form a clique in $H_i$, some bag $B_{z_i}$ contains $a_i$, $b_i$, and $c_i$.
+Since $a_i$, $b_i$, $c_i$ form a clique in $H$, there exists a bag $B_{z_i}$ contains $\{a_i,b_i,c_i\}$.
 
@@
 These replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$.
-Now add an edge that joins $y_0$ to $z_i$.
-
-    Using the fact that each component of $G-V(M)$ is contained in a single face of $M$, it is now straightforward to verify that, by doing this for each $i\in\{1,\ldots,m\}$ we obtain a width-$3$ tree decomposition of $N$.
-\qedhere
-  \end{compactenum}
-
-  %
-  %
-  % \note{PM}{Pick up here tomorrow, after teaching.}
-  % The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
-  % Elements in each of the $H$-partitions of $N_1,\ldots,N_m$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
-  %
-  % The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~13 in \cite{DJMMUW20}.
-The root node contains the nodes of $H$ corresponding to $P_1,\ldots,P_k$.
-The root has one child that contains the at most $4$ nodes of $H$ corresponding to $P_1,\ldots,P_k$ and $Y^+$.
-This root has up to three children, each of which contains $Y^+$ and at most two of $P_1,\ldots,P_k$.
-For each $i\in\{1,\ldots,m\}$, the tree decomposition of $N_i/\PP_i$ is then attached to one of these three children.
-  %
-	% The planarity of $H$ comes from two properties:
-	% \begin{compactenum}
-	% 	\item $G/\PP$ and $G^+/\mathcal{P^+}$ are isomorphic, where $G^+$ is the triangulation obtained by adding dummy vertices at each crossing in $G$ and $\PP^+$ is the partition we obtain by adding a dummy vertex $z$ to $\overline{Y}^+$ if $\overline{Y}^+$ contains an edge $vw$ that contains $z$ in its interior.
-  %
-	% 	\item $G^+[\overline{Y}^+-F]$ is connected.
-To see this, first observe that $\overline{Y}-F$ is connected, and then observe that every vertex of $\overline{Y}^+$ is either a vertex of $\overline{Y}$ or adjacent to a vertex of $\overline{Y}$.
-	% \end{compactenum}
-	% Since $G^+$ is planar, the second point implies that $H=G^+/\PP$ is planar.
-	% % \footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
-Repeatedly applying this fact implies that $\Delta/\PP$ is planar provided that $\Delta[P]$ is connected for every $P\in\PP$.} The first point then ensures that $G/\PP$ is also planar.
+Now add an edge that joins $y_0$ to $z_i$ so that $T_i$ becomes a subgree of $T_0$ that is ajacent to $y_0$.
+We perform this operation for each $i\in\{1,\ldots,m\}$ to obtain our final tree decomposition $mathcal{T}:=(B_y:y\in V(T_0))$.
+
+  Using the fact that each component of $G-V(M)$ is contained in a single face of $M$, it is now straightforward to verify that $\mathcal{T}$ is a tree decomposition of $N$ in which each bag has size at most $4$.
+The bag $B_{y_0}$ contains $x_1,\ldots,x_k$.
+When $k=3$, each neighbour of $z_i$ of $y_0$ has a bag that contains at most two of $\{x_1,x_2,x_3\}$.
+Therefore, when $k=3$, exactly one bag of $\mathcal{T}$ contains $\{x_1,x_2,x_3\}$.
+
+  All that remains is to show that any triple of vertices of $H$ is contained in at most two bags of $\mathcal{T}$.
+By the inductive hypothesis, the only triples we need to be concerned about are the $3$-element subsets of  $\{x_0,x_1,\ldots,x_k\}$.
+The case $k=1$ is trivial since no triples can be formed by $\{x_0,x_1\}$.
+When $k=2$, $B_{y_0}=\{x_0,x_1,x_2\}$.
+In this case, there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_0$, and $P_1$ on its boundary.
+(Refer to \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only other bag in $\mathcal{T}$ that contains $\{x_0,x_1,x_2\}$.
+This establishes (iii) for $k\in\{1,2\}$ so, from this point on we assume that $k=3$.
+
+  We begin with the claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.
+Suppose to the contrary that there are two such faces $F_{i_1}$ and $F_{i_2}$.
+Create a graph $M^+$ from $M$ by placing a vertex $v_{i_b}$ inside $F_{i_b}$ and adjacent to each vertex in $V(F_{i_b})$ for each $b\in\{1,2\}$.
+Since $M$ is planar and $F_{i_1}$ and $F_{i_2}$ are distinct faces of $M$, $M^+$ is planar.
+However, contracting each of $S_{x_0},\ldots,S_{x_3}$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.
+(In this graph, the $x_0,x_a,x_b$ are on one side of the bipartition.)
+
+  We have already established that the only bag that contains $\{x_1,x_2,x_3\}$ is $B_{y_0}$, so we need only consider triples $\{x_0,x_a,x_b\}$ with $1\le a<b\le 3$.
+By the preceding claim, there is at most one face $F_i$ with vertices of $S$, $P_a$, and $P_b$ on its boundary.
+By the inductive hypothesis, the only other bag in $\mathcal{T}$ that contains $\{x_0,x_a,x_b\}$ is the bag $B_{z_i}$.
+This completes the proof.
 \end{proof}

### 2022-03-13 k-planar.tex
Added figure

@@
 
+\note{PM}{The trick I used (using the thm counter everywhere) to get consecutive Theorem, Lemma, etc numbering has broken cleverref.
+Everything is now references as a Theorem.
+Find a solution that works.}
 % \note{DW}{I suggest we use consecutive numbering of theorems, lemmas, etc. Theorem 1, Lemma 2, Observation 3, etc.}
@@
 
-\subsection{\boldmath  $1$-Planar Graphs}
+\subsection{\boldmath $d$-Framed and $1$-Planar Graphs}
 \label{sec-1-planar}
@@
 
-In the following, we assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$, where we treat each edge $vw$ of $G$ as the image of a simple curve $f_{vw}:[0,1]\to\R^2$.
-% We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
-
-Let $G$ be a $1$-plane graph.
-We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.
-This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c@{\hspace{1cm}}c}
-        \includegraphics{figs/one_planar_example-3} &
-        \includegraphics{figs/one_planar_example-4} \\
-        $G$ & $G'$
-     \end{tabular}
-    \end{center}
-    \caption{The graph $G'$ obtained by removing pairs of crossing edges from $G$ is a plane multigraph whose faces all have three or four sides.}
-    \label{one_planar_example}
-\end{figure}
-
-To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ that cross at $p$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
-Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$.
-In the following paragraphs, subscripts on the vertices in $W$ are implicitly taken modulo $r$.
-By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
-
-If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
-Therefore each dummy vertex $w_i$ of $W$ is surrounded by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
-Since $w_i$ has degree $4$ in $G_0$ and $F$ does not have two parallel edges on its boundary, $w_{i-1}\neq w_{i+1}$.
-Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.
-Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.
-
-% Therefore each face of $G_0$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
-% \begin{inparaenum}[(i)]
-%   \item bounded by three edges of $G$, none of which are crossed; or
-%   \item bounded by one edge of $G$ that is not crossed plus portions of two edges of $G$ that cross each other.
-% \end{inparaenum}
-
-Consider some pair of edges $vw$ and $xy$ that cross at $p$.
-The point $p$ is a vertex of $G_0$ that has degree $4$.
-There are four faces $F_1,\ldots,F_4$ of $G_0$ that contain $p$ and, from the preceding discussion each of $vx$, $xw$, $wy$, and $yv$ is an uncrossed edge of $G$.
-Therefore, by removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
-Since, for any pair of crossing edges $\{vw,xy\}$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.
-% It follows immediately that, for any pair of vertices $a,b\in V(G)$, $\dist_{G'}(a,b)\le 2\dist_G(a,b)$.
-
+We will actually prove this result in the more general setting of $d$-framed graphs \cite{BDGGMR}.
+For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing having a biconnected plane skeleton (multi)graph $G_0$, whose faces have size at most $d$, and whose crossing edges are in the interior of these faces.
+In \cref{1_planar_is_4_framed}, below, we show that every $1$-planar graph is a subgraph of a $4$-framed multigraph.
+This makes \cref{1-planar} an immediate consequence of the following theorem, whose proof is the subject of the rest of this section.
+
+\begin{thm}\label{d_framed_product_stucture}
+  Every $d$-framed graph is contained in $H\boxtimes P\boxtimes K_{\lfloor d/2\rfloor + d - 3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+\end{thm}
+
+\subsection{\boldmath $d$-Framed Graphs}
+
+To establish the planarity of $H$ in \cref{d_framed_product_stucture}, the following definition is helpful.
 A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $3$-element subset $\{u,v,w\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{u,v,w\}$.
@@
 	\begin{compactenum}
-		\item Let $G$ and $G'$ be defined as above.
+		\item Let $G$ be a $d$-framed graph and let $G'$ be the frame of $G$.
 		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
@@
 			\item $F[P_i]$ is a path; and
-			\item $|V(P_i)\cap L_j| \le 3$ for every integer $j\ge 0$.
+			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
 		\end{compactenum}
@@
 		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
-        \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le 1$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
-		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$..
+        \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
+		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
 	\end{compactenum}
@@
 \begin{proof}
-  This proof is very similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is sufficiently complicated by the $1$-planar setting to warrant a full proof.
+  This proof is very similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is complicated by several factors.
 In particular, the reader should keep in mind that $G$ and $G'$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
@@
     \item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
-    \item For each $x\in V(H)$, let $Y_x:=S_x$ and $X_x=\emptyset$.  Then 4(a) ensures that $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
-    \item The trivial tree decomposition of $H$ that places $x_1,\ldots,x_k$ in one bag has width $k-1\le 2\le 3$.
+    \item Assumption 4(a) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
+    \item The trivial tree decomposition of $H$ that places $x_1,\ldots,x_k$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
   \end{compactenum}
@@
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
-Now add one edge to each $4$-sided inner face of $N'$ to obtain a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary.
+Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N'$, add edges $d'-2$ edges inside of $F$ to split it into $3$-sided faces.
+This yields a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary.
 By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
@@
 Therefore $|Y\cap L_j|\le 3$ for each integer $j\ge 0$.
-Furthermore, $|X|\le p-3\le 1$, since $v_1,v_2,v_3\in Y\cup V(F)$.
+Furthermore, $|X|\le p-3\le d-3$, since $v_1,v_2,v_3\in Y\cup V(F)$.
 Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ will not violate condition~(ii).
@@
   \end{figure}
+  \todo[inline]{Update these figures with $d$-sided $\tau$.}
 
@@
 
-We remark that our proof of \cref{1-planar} generalizes to the class of $d$-framed graphs \cite{BDGGMR} and even $d$-framed multigraphs.
-An embedded multigraph $G$ is a \defin{$d$-framed multigraph} if it contains a plane subgraph $G'$ (the \defin{frame}) whose edges are not involved in any crossings and such that each pair of crossing edges in $G$ are contained in a face of $G'$ that has at most $d$ vertices on its boundary.
-It is known that every $3$-connected $1$-planar graph is contained in some $4$-framed graph.
-The discussion on the structure of $G$ and $G'$ that precedes \cref{induction} shows that every $1$-planar graph is contained in a $4$-framed multigraph.
-The proof of \cref{induction} can be repeated for any edge-maximal $d$-framed multigraphs $G$ having frame $G'$. The proof in this case has only two differences:
-\begin{compactitem}
-   \item For any edge $vw$ of $G$ that is not in $G'$, $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$.
-   \item The face $\tau$ that contains the Sperner triangle $v_1,v_2,v_3$ may have up to $d$ vertices on its boundary, so up to $d-3$ of these vertices do not belong to $Q_i$ for any $i\in\{1,2,3\}$.
-\end{compactitem}
-The only effect this has on the final result is that the layered width becomes $3\lfloor d/2\rfloor + d-3$.  This yields the following result:
-
-\begin{thm}
-\label{framed}
-   Every $d$-framed multigraph is contained in $H\boxtimes P\boxtimes K_{3\lfloor d/2\rfloor + d -3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
-\end{thm}
-
-\note{DW}{I think \cref{framed} should be proved explicitly and \cref{1-planar} should be a corollary, especially because \cref{NewMapGraph} is another corollary.
-The proof of \cref{framed} needs an extra clause; see the last sentence of \cref{NewLemma}.
-This section should be about framed graphs, outside of \cref{sec-k-planar}.
-I suggest we make it a subsection of \cref{Structure} where we introduce framed graphs, say they can be described by $(2,\binom{d}{2})$-shortcut systems, so the main theorem is applicable, which gives a treewidth bound of $\binom{5}{3}-1=9$, but using  \cref{framed} we get a treewidth bound of 3, which is optional.
-Then in latter sections we give two applications, for 1-planar graphs and for $d$-map graphs.
-}
-
-\note{PM}{I no longer agree about the restructuring. See my email essay for an explanation.}
-
-\note{PM}{Apropos the extra clause.
-It's related to the orange todo note (argue that $N'/\mathcal{P}=N/\mathcal{P}$).
-This doesn't need to be in the statement of the lemma since, unlike \cref{NewLemma} this one gives an $H$-partition of $N$, not $N'$, but something like it needs to be in the proof.}
+\subsection{\boldmath $1$-Planar Graphs}
+
+In the following, we assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$, where we treat each edge $vw$ of $G$ as the image of a simple curve $f_{vw}:[0,1]\to\R^2$.
+% We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
+
+\begin{lem}\label{1_planar_is_4_framed}
+  A graph $G$ is $1$-planar if and only if it is contained in a $4$-framed multigraph.
+\end{lem}
+
+\begin{proof}
+Let $G$ be a $1$-plane graph.
+We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.
+This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c@{\hspace{1cm}}c}
+        \includegraphics{figs/one_planar_example-3} &
+        \includegraphics{figs/one_planar_example-4} \\
+        $G$ & $G'$
+     \end{tabular}
+    \end{center}
+    \caption{The graph $G'$ obtained by removing pairs of crossing edges from $G$ is a plane multigraph whose faces all have three or four sides.}
+    \label{one_planar_example}
+\end{figure}
+
+To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ that cross at $p$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
+Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$.
+In the following paragraphs, subscripts on the vertices in $W$ are implicitly taken modulo $r$.
+By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
+
+If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
+Therefore each dummy vertex $w_i$ of $W$ is surrounded by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
+Since $w_i$ has degree $4$ in $G_0$ and $F$ does not have two parallel edges on its boundary, $w_{i-1}\neq w_{i+1}$.
+Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.
+Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.
+
+% Therefore each face of $G_0$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
+% \begin{inparaenum}[(i)]
+%   \item bounded by three edges of $G$, none of which are crossed; or
+%   \item bounded by one edge of $G$ that is not crossed plus portions of two edges of $G$ that cross each other.
+% \end{inparaenum}
+
+Consider some pair of edges $vw$ and $xy$ that cross at $p$.
+The point $p$ is a vertex of $G_0$ that has degree $4$.
+There are four faces $F_1,\ldots,F_4$ of $G_0$ that contain $p$ and, from the preceding discussion each of $vx$, $xw$, $wy$, and $yv$ is an uncrossed edge of $G$.
+Therefore, by removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
+Since, for any pair of crossing edges $\{vw,xy\}$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.
+% It follows immediately that, for any pair of vertices $a,b\in V(G)$, $\dist_{G'}(a,b)\le 2\dist_G(a,b)$.
+\end{proof}
 

### 2022-03-13 k-planar.tex
Switching to d-framed

@@
     \item Assumption 4(a) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
-    \item The trivial tree decomposition of $H$ that places $x_1,\ldots,x_k$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
+    \item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
   \end{compactenum}
@@
       \begin{tabular}{ccc}
-        \includegraphics{figs/mmplus-1} &
-        \includegraphics{figs/mmplus-2} &
-        \includegraphics{figs/mmplus-3} \\
+        \includegraphics{figs/zoomba-1} &
+        \includegraphics{figs/zoomba-2} &
+        \includegraphics{figs/zoomba-3} \\
         $k=1$ & $k=2$ & $k=3$
@@
     \caption{Three cases when choosing the face $\tau$ of $N'$.}
+    \todo[inline]{Update these figures with $d$-sided $\tau$.}
     \label{boring_figure}
   \end{figure}
-  \todo[inline]{Update these figures with $d$-sided $\tau$.}
 
@@
 
-    % \item We only consider the (most difficult) case where $k=3$, the cases $\in\{1,2\}$ are handled easily.
-Let $x_0$ be the vertex of $H$ obtained from contracting $Y$ and, for each $j\in\{1,\ldots,3\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
-In this case the vertices $x_0,\ldots,x_3$ form a clique in $H$, which we embed with $x_1,\ldots,x_3$ on the outer face, as required.
-    %
-    % We claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.
-Suppose to the contrary that there are two such faces $F_{i_1}$ and $F_{i_2}$.
-Create a graph $M^+$ from $M$ by placing a vertex $v_{i_b}$ inside $F_{i_b}$ and adjacent to each vertex in $V(F_{i_b})$ for each $b\in\{1,2\}$.
-Since $M$ is planar and $F_{i_1}$ and $F_{i_2}$ are distinct faces of $M$, $M^+$ is planar.
-However, contracting each of $S_{x_0},\ldots,S_{x_3}$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.
-    %
-    % Therefore, by the inductive hypothesis, we can embed each such graph $H_i :=N_i/\mathcal{P}_i$ so that the three vertices on its outer face correspond to the appropriate inner face of the clique on $x_0,\ldots,x_3$.
-Any remaining graphs $H_i$ have outer faces with at most two vertices on their outer faces corresponding to at most two vertices of $x_0,\ldots,x_3$.
-These can easily be embedded, one a time, in some inner face that has the edge $x_ix_j$ on its boundary.
-
   All that remains is to construct a tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$ of $H$ that satisfies (iii).
@@
 The tree $T_0$ has a node $y_0$ with bag $B_{y_0}:=\{x_0,\ldots,x_k\}$.
-The inductive hypothesis ensures that, for each $i\in\{1,\ldots,m\}$,  $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $(B_z:z\in V(T_i))$ that satisfies (iii).
+The inductive hypothesis ensures that, for each $i\in\{1,\ldots,m\}$,  $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $\mathcal{T}_i:=(B_z:z\in V(T_i))$ that satisfies (iii).
 Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
@@
 These replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$.
-Now add an edge that joins $y_0$ to $z_i$ so that $T_i$ becomes a subgree of $T_0$ that is ajacent to $y_0$.
+Now add an edge that joins $y_0$ to $z_i$ so that $T_i$ becomes a subtree of $T_0$ that is adjacent to $y_0$.
 We perform this operation for each $i\in\{1,\ldots,m\}$ to obtain our final tree decomposition $mathcal{T}:=(B_y:y\in V(T_0))$.
@@
 In this case, there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_0$, and $P_1$ on its boundary.
-(Refer to \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only other bag in $\mathcal{T}$ that contains $\{x_0,x_1,x_2\}$.
+(Refer to \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.
+Therefore there are at most two bags of $\mathcal{T}$ that contains $\{x_0,x_1,x_2\}$.
 This establishes (iii) for $k\in\{1,2\}$ so, from this point on we assume that $k=3$.
 
-  We begin with the claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.
+  We claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.
 Suppose to the contrary that there are two such faces $F_{i_1}$ and $F_{i_2}$.
@@
 
-Using \cref{induction}, the proof of \cref{1-planar} is now straightforward.
-
-\begin{proof}[Proof of \cref{1-planar}]
-	Given a 1-plane graph $G$, add edges to make it edge-maximal so that it has an outer face $F=v_1v_2v_3$.
-Next, add a vertex $r$ adjacent to $v_1$, $v_2$, and $v_3$ to obtain an edge-maximal 1-plane multigraph $\overline{G}$ with one vertex $r$ of degree 3 on its outer face.
-
-	Let $G'$ be the plane graph obtained by removing each pair of crossing edges from $\overline{G}$ and let $T$ be a BFS tree of $G'$ rooted at $r$ and let $\langle L_0,L_1,\ldots,\rangle$ be the resulting BFS layering.
-Now apply \cref{induction} with $G=\overline{G}$, $G'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
-This gives an $H$-partition $\mathcal{P}:=\{S_x\cup\{v_x\}:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ is planar and has treewidth at most 3.
-
-	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{2i}\cup L_{2i+1}$ for each integer $i\ge 0$.
-That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le 2$ for each edge $vw\in E(G)$.
-Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 6$ and therefore $|L_i'\cap S_x|\le 7$ for each integer $i\ge 0$ and each $x\in V(H)$.
+Using \cref{induction}, the proof of \cref{d_framed_product_stucture} is now straightforward.
+
+\begin{proof}[Proof of \cref{d_framed_product_stucture}]
+    Let $G$ be a $d$-framed multigraph with frame $G'$\todo{Deal with the case where $G'$ has no triangular face.} and assume that the outer face of $G'$ has exactly three vertices $v_1,v_2,v_3$ and that no edges of $G$ are in the outer face of $G'$.
+Add a vertex $r$ adjacent to $v_1$, $v_2$, and $v_3$ to obtain a $d$-framed multigraph $\overline{G}$ with frame $\overline{G}'$.
+
+	Let $T$ be a BFS tree of $\overline{G}'$ rooted at $r$ and let $\langle L_0,L_1,\ldots,\rangle$ be the resulting BFS layering.
+Now apply \cref{induction} with $G=\overline{G}$, $\overline{G}'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
+This gives an $H$-partition $\mathcal{P}:=\{S_x\cup\{v_x\}:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ has a $3$-simple tree decomposition.
+
+	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdot \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$.
+That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.
+Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x| |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
 The result follows from \cref{PartitionProduct}.
@@
 Therefore, by removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
-Since, for any pair of crossing edges $\{vw,xy\}$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.
+There $G$ is a $4$-framed graph with frame $G'$.
+% Since, for any pair of crossing edges $\{vw,xy\}$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.
 % It follows immediately that, for any pair of vertices $a,b\in V(G)$, $\dist_{G'}(a,b)\le 2\dist_G(a,b)$.

### 2022-03-13 k-planar.tex
small typo fix
Pffft

@@
 
-To establish the planarity of $H$ in \cref{d_framed_product_stucture}, the following definition is helpful.
-A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $3$-element subset $\{u,v,w\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{u,v,w\}$.
+The following definition is a convenient way to estalish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $3$-element subset $\{u,v,w\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{u,v,w\}$.
 Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
@@
     \caption{Three cases when choosing the face $\tau$ of $N'$.}
-    \todo[inline]{Update these figures with $d$-sided $\tau$.}
     \label{boring_figure}
@@
 Therefore, by removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
-There $G$ is a $4$-framed graph with frame $G'$.
+Therefore $G$ is a $4$-framed graph with frame $G'$.
 % Since, for any pair of crossing edges $\{vw,xy\}$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.

### 2022-03-13 k-planar.tex
Pfffft

@@
 
-The following definition is a convenient way to estalish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $3$-element subset $\{u,v,w\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{u,v,w\}$.
+The following definition is a convenient way to estalish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{v_1,\ldots,v_t\}$.
 Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
@@
 This gives an $H$-partition $\mathcal{P}:=\{S_x\cup\{v_x\}:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ has a $3$-simple tree decomposition.
+Since any graph having a $3$-simple tree decomposition is planar \cite{X}, this implies that $H$ is planar.
 

### 2022-03-13 response.tex
minor updates
added comment in proof of Thm 45

@@
 
-\tba
+\textcolor{blue}{We have shortened the paragraph, avoiding the repetition.}
 
@@
 
-\textcolor{blue}{It appear that the referee forgot that we are dealing with an edge-maximal multigraph at this point. We have revised the proof so that this misunderstanding cannot happen.}
+\textcolor{blue}{It appears that the referee forgot that we are dealing with an edge-maximal multigraph at this point. We have revised the proof so that this misunderstanding cannot happen.}
 
@@
 The proof is quite similar to the previous proof except that graph $G'$ is obtained by removing both crossing edges involved in a crossing (the spars of a kite), rather than just one.
-We thank the referee for forcing us to write this proof more carefully so that we could discover this improvement.}
+We use the language of framed multigraph, for consistency with the literature.
+Thanks to the referee for forcing us to write this proof more carefully so that we could discover this improvement.}
 

### 2022-03-13 k-planar.tex
added comment in proof of Thm 45

--- 2022-03-13 k-planar.tex
@@
 
-\note{PM}{The trick I used (using the thm counter everywhere) to get consecutive Theorem, Lemma, etc numbering has broken cleverref.
-Everything is now references as a Theorem.
-Find a solution that works.}
 % \note{DW}{I suggest we use consecutive numbering of theorems, lemmas, etc. Theorem 1, Lemma 2, Observation 3, etc.}
@@
 
-\subsection{\boldmath $d$-Framed and $1$-Planar Graphs}
+\subsection{\boldmath  $1$-Planar Graphs}
 \label{sec-1-planar}
@@
 
-We will actually prove this result in the more general setting of $d$-framed graphs \cite{BDGGMR}.
-For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing having a biconnected plane skeleton (multi)graph $G_0$, whose faces have size at most $d$, and whose crossing edges are in the interior of these faces.
-In \cref{1_planar_is_4_framed}, below, we show that every $1$-planar graph is a subgraph of a $4$-framed multigraph.
-This makes \cref{1-planar} an immediate consequence of the following theorem, whose proof is the subject of the rest of this section.
-
-\begin{thm}\label{d_framed_product_stucture}
-  Every $d$-framed graph is contained in $H\boxtimes P\boxtimes K_{\lfloor d/2\rfloor + d - 3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
-\end{thm}
-
-\subsection{\boldmath $d$-Framed Graphs}
-
-The following definition is a convenient way to estalish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $3$-element subset $\{u,v,w\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{u,v,w\}$.
+In the following, we assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$, where we treat each edge $vw$ of $G$ as the image of a simple curve $f_{vw}:[0,1]\to\R^2$.
+We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
+
+Let $G$ be a $1$-plane graph.
+We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.
+This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c@{\hspace{1cm}}c}
+        \includegraphics{figs/one_planar_example-3} &
+        \includegraphics{figs/one_planar_example-4} \\
+        $G$ & $G'$
+     \end{tabular}
+    \end{center}
+    \caption{The graph $G'$ obtained by removing pairs of crossing edges from $G$ is a plane multigraph whose faces all have three or four sides.}
+    \label{one_planar_example}
+\end{figure}
+
+To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ that cross at $p$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
+Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$.
+In the following paragraphs, subscripts on the vertices in $W$ are implicitly taken modulo $r$.
+By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
+
+If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
+Therefore each dummy vertex $w_i$ of $W$ is surrounded by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
+Since $w_i$ has degree $4$ in $G_0$ and $F$ does not have two parallel edges on its boundary, $w_{i-1}\neq w_{i+1}$.
+Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.
+Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.
+
+% Therefore each face of $G_0$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
+% \begin{inparaenum}[(i)]
+%   \item bounded by three edges of $G$, none of which are crossed; or
+%   \item bounded by one edge of $G$ that is not crossed plus portions of two edges of $G$ that cross each other.
+% \end{inparaenum}
+
+Consider some pair of edges $vw$ and $xy$ that cross at $p$.
+The point $p$ is a vertex of $G_0$ that has degree $4$.
+There are four faces $F_1,\ldots,F_4$ of $G_0$ that contain $p$ and, from the preceding discussion each of $vx$, $xw$, $wy$, and $yv$ is an uncrossed edge of $G$.
+Therefore, by removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
+Since, for any pair of crossing edges $\{vw,xy\}$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.
+% It follows immediately that, for any pair of vertices $a,b\in V(G)$, $\dist_{G'}(a,b)\le 2\dist_G(a,b)$.
+
 Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
@@
 	\begin{compactenum}
-		\item Let $G$ be a $d$-framed graph and let $G'$ be the frame of $G$.
+		\item Let $G$ and $G'$ be defined as above.
 		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
@@
 			\item $F[P_i]$ is a path; and
-			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
+			\item $|V(P_i)\cap L_j| \le 3$ for every integer $j\ge 0$.
 		\end{compactenum}
@@
 		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
-        \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
-		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
+        \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le 1$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
+        \item $H$ has a plane embedding in which $x_1,\ldots,x_k$ appear on the outer face;
+		\item $\tw(H)\le 3$.
 	\end{compactenum}
@@
 \begin{proof}
-  This proof is very similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is complicated by several factors.
+  This proof is very similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is sufficiently complicated by the $1$-planar setting to warrant a full proof.
 In particular, the reader should keep in mind that $G$ and $G'$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
@@
     \item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
-    \item Assumption 4(a) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
-    \item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
+    \item For each $x\in V(H)$, let $Y_x:=S_x$ and $X_x=\emptyset$.  Then 4(a) ensures that $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
+    \item $H$ has $k\le 3$ vertices and any graph on $k\le 3$ vertices has a plane embedding with all its vertices on the outer face.
+    \item The trivial tree decomposition of $H$ that places $x_1,\ldots,x_k$ in one bag has width $k-1\le 2\le 3$.
   \end{compactenum}
-
-  In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
-There is one case to consider for each possible value of $k$:
+  From this point on, we assume that $N$ has at least one inner vertex.
+
+  The $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
+Depending on the value of $k$, there are two cases to consider:
   \begin{compactenum}
-    \item If $k=1$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
-
-    \item If $k= 2$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
+    \item If $k\in\{1,2\}$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge of $F$ on its boundary.
 
@@
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
-Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N'$, add edges $d'-2$ edges inside of $F$ to split it into $3$-sided faces.
-This yields a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary.
+Now add one edge to each $4$-sided inner face of $N'$ to obtain a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary.
 By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
 Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.
-% (Note that $p\in\{3,4\}$.)
+(Note that $p\in\{3,4\}$.)
   \end{compactenum}
 
-  \todo[inline]{%
-    DW:Does this work for multigraph near-triangulations? \newline
-    PM: Yes, it only requires that the inner faces have three sides and that the outer face is bounded by a simple cycle.
+%  \todo[inline]{%
+%    DW:Does this work for multigraph near-triangulations? \newline
+%    PM: Yes, it only requires that the inner faces have three sides and that the outer face is bounded by a simple cycle.
 The proof goes through without change.
@@
 Any bichromatic inner face has degree $0$ or degree $2$ in $X$.
-There has to be another vertex of odd degree somewhere, which can only come from a trichromatic inner face.
-  }
+There has to be another vertex of odd degree somewhere, which can only come from a trichromatic inner face.\newline
+%    DW: great. This is basically the same as the proof of the Hex Lemma.
+%  }
 
@@
 Let $\overline{S}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
-Let $S:=V(\overline{S})\setminus V(F)$, let $Y:=(V(Q_1)\cup V(Q_2)\cup V(Q_3))\setminus V(F)$, and let $X:=S\setminus Y$.
+Let $S:=V(\overline{S})\setminus V(F)$, let $Y:=(V(Q_1)\cup V(Q_2)\cup V(Q_3))\cap S$, and let $X:=S\setminus Y$.
   Observe that, for each $i\in\{1,2,3\}$ and each $j\ge 0$, $|V(Q_i)\cap L_j|\le 1$.
 Therefore $|Y\cap L_j|\le 3$ for each integer $j\ge 0$.
-Furthermore, $|X|\le p-3\le d-3$, since $v_1,v_2,v_3\in Y\cup V(F)$.
-Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ will not violate condition~(ii).
+Furthermore, $|X|\le p-3\le 1$, since $v_1,v_2,v_3\in Y\cup V(N)$.
+Therefore, taking $Y:=S\cup\{v\}\setminus V(F)$ as a part in the partition $\PP$ will not violate condition~(ii).
 
@@
 
+  Refer to \cref{boring_figure}.
+Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$.
+Since $N'\subseteq G'$ contains only edges of $G$ that are not crossed, each component of $G-V(M)$ is contained in a single face of $M$.
+
   \begin{figure}
     \begin{center}
-      \begin{tabular}{ccc}
-        \includegraphics{figs/zoomba-1} &
-        \includegraphics{figs/zoomba-2} &
-        \includegraphics{figs/zoomba-3} \\
-        $k=1$ & $k=2$ & $k=3$
+      \begin{tabular}{c@{\hspace{1cm}}c}
+        \includegraphics{figs/mmplus-1} &
+        \includegraphics{figs/mmplus-2} \\
+        $k\le 2$ & $k=3$
       \end{tabular}
     \end{center}
-    \caption{Three cases when choosing the face $\tau$ of $N'$.}
+    \caption{Two cases when choosing the face $\tau$ of $N'$.}
     \label{boring_figure}
@@
 
-  Refer to \cref{boring_figure}.
-Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$.
+  Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
 By our choice of $\tau$, the graph $M$ is $2$-connected.
-Since $N'\subseteq G'$ contains only edges of $G$ that are not crossed, each component of $G-V(M)$ is contained in a single face of $M$.
-  Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
-By our choice of $v_1v_2v_3$, the vertices of each $F_i$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
+Furthermore, $V(F_i)$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
 
@@
   It is straightforward to verify that $\mathcal{P}$ is indeed a partition of $V(N)$ and, by definition this is an $H$-partition of $N$ for the graph $H:=N/\mathcal{P}$.
-All that remains is to verify that $\mathcal{P}$ satisfies (i)--(iii).
-
-  By construction $\mathcal{P}$ contains parts $P_1,\ldots,P_k$, so $\mathcal{P}$ satisifes (i).
-By Assumption~4(b) none of the parts $P_1,P_2,P_3$ violate condition~(ii).
+All that remains is to verify that $\mathcal{P}$ satisfies conditions~(i)--(iv).
+  \begin{compactenum}[(i)]
+    \item By construction $\mathcal{P}$ contains parts $P_1,\ldots,P_k$.
+
+    \item  By Assumption~4(b) none of the parts $P_1,P_2,P_3$ violate condition~(ii).
 We have already argued that the part $S$ does not violate condition~(ii).
 The inductive hypothesis ensures that each of the remaining parts satisfies condition~(ii).
-Therefore $\mathcal{P}$ satisfies (ii).
-
-  All that remains is to construct a tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$ of $H$ that satisfies (iii).
-Let $x_0$ be the vertex of $H$ obtained from contracting $S$ and, for each $j\in\{1,\ldots,k\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
-The tree $T_0$ has a node $y_0$ with bag $B_{y_0}:=\{x_0,\ldots,x_k\}$.
-The inductive hypothesis ensures that, for each $i\in\{1,\ldots,m\}$,  $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $\mathcal{T}_i:=(B_z:z\in V(T_i))$ that satisfies (iii).
-Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
-Since $a_i$, $b_i$, $c_i$ form a clique in $H$, there exists a bag $B_{z_i}$ contains $\{a_i,b_i,c_i\}$.
-
-  Recall that $P_{i,1}:=V(N_i)\cap Y$, $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
-For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.
-These replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$.
-Now add an edge that joins $y_0$ to $z_i$ so that $T_i$ becomes a subtree of $T_0$ that is adjacent to $y_0$.
-We perform this operation for each $i\in\{1,\ldots,m\}$ to obtain our final tree decomposition $mathcal{T}:=(B_y:y\in V(T_0))$.
-
-  Using the fact that each component of $G-V(M)$ is contained in a single face of $M$, it is now straightforward to verify that $\mathcal{T}$ is a tree decomposition of $N$ in which each bag has size at most $4$.
-The bag $B_{y_0}$ contains $x_1,\ldots,x_k$.
-When $k=3$, each neighbour of $z_i$ of $y_0$ has a bag that contains at most two of $\{x_1,x_2,x_3\}$.
-Therefore, when $k=3$, exactly one bag of $\mathcal{T}$ contains $\{x_1,x_2,x_3\}$.
-
-  All that remains is to show that any triple of vertices of $H$ is contained in at most two bags of $\mathcal{T}$.
-By the inductive hypothesis, the only triples we need to be concerned about are the $3$-element subsets of  $\{x_0,x_1,\ldots,x_k\}$.
-The case $k=1$ is trivial since no triples can be formed by $\{x_0,x_1\}$.
-When $k=2$, $B_{y_0}=\{x_0,x_1,x_2\}$.
-In this case, there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_0$, and $P_1$ on its boundary.
-(Refer to \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.
-Therefore there are at most two bags of $\mathcal{T}$ that contains $\{x_0,x_1,x_2\}$.
-This establishes (iii) for $k\in\{1,2\}$ so, from this point on we assume that $k=3$.
+
+    \item We only consider the (most difficult) case where $k=3$, the cases $\in\{1,2\}$ are handled easily.
+Let $x_0$ be the vertex of $H$ obtained from contracting $Y$ and, for each $j\in\{1,\ldots,3\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
+In this case the vertices $x_0,\ldots,x_3$ form a clique in $H$, which we embed with $x_1,\ldots,x_3$ on the outer face, as required.
 
@@
 However, contracting each of $S_{x_0},\ldots,S_{x_3}$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.
-(In this graph, the $x_0,x_a,x_b$ are on one side of the bipartition.)
-
-  We have already established that the only bag that contains $\{x_1,x_2,x_3\}$ is $B_{y_0}$, so we need only consider triples $\{x_0,x_a,x_b\}$ with $1\le a<b\le 3$.
-By the preceding claim, there is at most one face $F_i$ with vertices of $S$, $P_a$, and $P_b$ on its boundary.
-By the inductive hypothesis, the only other bag in $\mathcal{T}$ that contains $\{x_0,x_a,x_b\}$ is the bag $B_{z_i}$.
-This completes the proof.
+
+    Therefore, by the inductive hypothesis, we can embed each such graph $H_i :=N_i/\mathcal{P}_i$ so that the three vertices on its outer face correspond to the appropriate inner face of the clique on $x_0,\ldots,x_3$.
+Any remaining graphs $H_i$ have outer faces with at most two vertices on their outer faces corresponding to at most two vertices of $x_0,\ldots,x_3$.
+These can easily be embedded, one a time, in some inner face that has the edge $x_ix_j$ on its boundary.
+
+    \item We now construct a tree decomposition $(B_y:y\in V(T_0))$ of $H$ that satisifies condition~(iv).
+    % We show that $\tw(H)\le 3$ by giving a tree-decomposition of $H$ of width at most $3$.
+    Let $x_0$ be the vertex of $H$ obtained from contracting $Y$ and, for each $j\in\{1,\ldots,k\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
+
+    The tree $T_0$ has a node $y_0$ with $B_{y_0}:=\{x_0,\ldots,x_k\}$.
+    % For each $i\in\{1,\ldots,m\}$, let $N_i$ denote the subgraph of $N$ contained in $F_i$.
+    The inductive hypothesis ensures that $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $(B_z:z\in V(T_i))$ of width at most $3$.
+Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
+Since $a_i,b_i,c_i$ form a clique in $H_i$, some bag $B_{z_i}$ contains $a_i$, $b_i$, and $c_i$.
+
+    Recall that $P_{i,1}:=V(N_i)\cap Y$, $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
+For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.
+These replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$.
+Now add an edge that joins $y_0$ to $z_i$.
+
+    Using the fact that each component of $G-V(M)$ is contained in a single face of $M$, it is now straightforward to verify that, by doing this for each $i\in\{1,\ldots,m\}$ we obtain a width-$3$ tree decomposition of $N$.
+\qedhere
+  \end{compactenum}
+
+  %
+  %
+  % \note{PM}{Pick up here tomorrow, after teaching.}
+  % The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
+  % Elements in each of the $H$-partitions of $N_1,\ldots,N_m$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.
+  %
+  % The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~13 in \cite{DJMMUW20}.
+The root node contains the nodes of $H$ corresponding to $P_1,\ldots,P_k$.
+The root has one child that contains the at most $4$ nodes of $H$ corresponding to $P_1,\ldots,P_k$ and $Y^+$.
+This root has up to three children, each of which contains $Y^+$ and at most two of $P_1,\ldots,P_k$.
+For each $i\in\{1,\ldots,m\}$, the tree decomposition of $N_i/\PP_i$ is then attached to one of these three children.
+  %
+	% The planarity of $H$ comes from two properties:
+	% \begin{compactenum}
+	% 	\item $G/\PP$ and $G^+/\mathcal{P^+}$ are isomorphic, where $G^+$ is the triangulation obtained by adding dummy vertices at each crossing in $G$ and $\PP^+$ is the partition we obtain by adding a dummy vertex $z$ to $\overline{Y}^+$ if $\overline{Y}^+$ contains an edge $vw$ that contains $z$ in its interior.
+  %
+	% 	\item $G^+[\overline{Y}^+-F]$ is connected.
+To see this, first observe that $\overline{Y}-F$ is connected, and then observe that every vertex of $\overline{Y}^+$ is either a vertex of $\overline{Y}$ or adjacent to a vertex of $\overline{Y}$.
+	% \end{compactenum}
+	% Since $G^+$ is planar, the second point implies that $H=G^+/\PP$ is planar.
+	% % \footnote{It is well-known and easy to see that any graph $\Delta'$ obtained by contracting an edge in a planar graph $\Delta$ is also planar.
+Repeatedly applying this fact implies that $\Delta/\PP$ is planar provided that $\Delta[P]$ is connected for every $P\in\PP$.} The first point then ensures that $G/\PP$ is also planar.
 \end{proof}
 
-Using \cref{induction}, the proof of \cref{d_framed_product_stucture} is now straightforward.
-
-\begin{proof}[Proof of \cref{d_framed_product_stucture}]
-    Let $G$ be a $d$-framed multigraph with frame $G'$\todo{Deal with the case where $G'$ has no triangular face.} and assume that the outer face of $G'$ has exactly three vertices $v_1,v_2,v_3$ and that no edges of $G$ are in the outer face of $G'$.
-Add a vertex $r$ adjacent to $v_1$, $v_2$, and $v_3$ to obtain a $d$-framed multigraph $\overline{G}$ with frame $\overline{G}'$.
-
-	Let $T$ be a BFS tree of $\overline{G}'$ rooted at $r$ and let $\langle L_0,L_1,\ldots,\rangle$ be the resulting BFS layering.
-Now apply \cref{induction} with $G=\overline{G}$, $\overline{G}'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
-This gives an $H$-partition $\mathcal{P}:=\{S_x\cup\{v_x\}:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ has a $3$-simple tree decomposition.
-
-	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdot \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$.
-That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.
-Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x| |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
+Using \cref{induction}, the proof of \cref{1-planar} is now straightforward.
+
+\begin{proof}[Proof of \cref{1-planar}]
+	Given a 1-plane graph $G$, add edges to make it edge-maximal so that it has an outer face $F=v_1v_2v_3$.
+Next, add a vertex $r$ adjacent to $v_1$, $v_2$, and $v_3$ to obtain an edge-maximal 1-plane multigraph $\overline{G}$ with one vertex $r$ of degree 3 on its outer face.
+
+	Let $G'$ be the plane graph obtained by removing each pair of crossing edges from $\overline{G}$ and let $T$ be a BFS tree of $G'$ rooted at $r$ and let $\langle L_0,L_1,\ldots,\rangle$ be the resulting BFS layering.
+Now apply \cref{induction} with $G=\overline{G}$, $G'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
+This gives an $H$-partition $\mathcal{P}:=\{S_x\cup\{v_x\}:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ is planar and has treewidth at most 3.
+
+	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{2i}\cup L_{2i+1}$ for each integer $i\ge 0$.
+That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le 2$ for each edge $vw\in E(G)$.
+Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 6$ and therefore $|L_i'\cap S_x|\le 7$ for each integer $i\ge 0$ and each $x\in V(H)$.
 The result follows from \cref{PartitionProduct}.
@@
 
-\subsection{\boldmath $1$-Planar Graphs}
-
-In the following, we assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$, where we treat each edge $vw$ of $G$ as the image of a simple curve $f_{vw}:[0,1]\to\R^2$.
-% We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
-
-\begin{lem}\label{1_planar_is_4_framed}
-  A graph $G$ is $1$-planar if and only if it is contained in a $4$-framed multigraph.
-\end{lem}
-
-\begin{proof}
-Let $G$ be a $1$-plane graph.
-We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.
-This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c@{\hspace{1cm}}c}
-        \includegraphics{figs/one_planar_example-3} &
-        \includegraphics{figs/one_planar_example-4} \\
-        $G$ & $G'$
-     \end{tabular}
-    \end{center}
-    \caption{The graph $G'$ obtained by removing pairs of crossing edges from $G$ is a plane multigraph whose faces all have three or four sides.}
-    \label{one_planar_example}
-\end{figure}
-
-To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ that cross at $p$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
-Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$.
-In the following paragraphs, subscripts on the vertices in $W$ are implicitly taken modulo $r$.
-By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
-
-If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
-Therefore each dummy vertex $w_i$ of $W$ is surrounded by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
-Since $w_i$ has degree $4$ in $G_0$ and $F$ does not have two parallel edges on its boundary, $w_{i-1}\neq w_{i+1}$.
-Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.
-Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.
-
-% Therefore each face of $G_0$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
-% \begin{inparaenum}[(i)]
-%   \item bounded by three edges of $G$, none of which are crossed; or
-%   \item bounded by one edge of $G$ that is not crossed plus portions of two edges of $G$ that cross each other.
-% \end{inparaenum}
-
-Consider some pair of edges $vw$ and $xy$ that cross at $p$.
-The point $p$ is a vertex of $G_0$ that has degree $4$.
-There are four faces $F_1,\ldots,F_4$ of $G_0$ that contain $p$ and, from the preceding discussion each of $vx$, $xw$, $wy$, and $yv$ is an uncrossed edge of $G$.
-Therefore, by removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
-Therefore $G$ is a $4$-framed graph with frame $G'$.
-% Since, for any pair of crossing edges $\{vw,xy\}$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.
-% It follows immediately that, for any pair of vertices $a,b\in V(G)$, $\dist_{G'}(a,b)\le 2\dist_G(a,b)$.
-\end{proof}
+We remark that our proof of \cref{1-planar} generalizes to the class of $d$-framed graphs \cite{BDGGMR} and even $d$-framed multigraphs.
+An embedded multigraph $G$ is a \defin{$d$-framed multigraph} if it contains a plane subgraph $G'$ (the \defin{frame}) whose edges are not involved in any crossings and such that each pair of crossing edges in $G$ are contained in a face of $G'$ that has at most $d$ vertices on its boundary.
+It is known that every $3$-connected $1$-planar graph is contained in some $4$-framed graph.
+The discussion on the structure of $G$ and $G'$ that precedes \cref{induction} shows that every $1$-planar graph is contained in a $4$-framed multigraph.
+The proof of \cref{induction} can be repeated for any edge-maximal $d$-framed multigraphs $G$ having frame $G'$. The proof in this case has only two differences:
+\begin{compactitem}
+   \item For any edge $vw$ of $G$ that is not in $G'$, $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$.
+   \item The face $\tau$ that contains the Sperner triangle $v_1,v_2,v_3$ may have up to $d$ vertices on its boundary, so up to $d-3$ of these vertices do not belong to $Q_i$ for any $i\in\{1,2,3\}$.
+\end{compactitem}
+The only effect this has on the final result is that the layered width becomes $3\lfloor d/2\rfloor + d-3$.  This yields the following result:
+
+\begin{thm}
+\label{framed}
+   Every $d$-framed multigraph is contained in $H\boxtimes P\boxtimes K_{3\lfloor d/2\rfloor + d -3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+\end{thm}
+
+\note{DW}{I think \cref{framed} should be proved explicitly and \cref{1-planar} should be a corollary, especially because \cref{NewMapGraph} is another corollary.
+The proof of \cref{framed} needs an extra clause; see the last sentence of \cref{NewLemma}.
+This section should be about framed graphs, outside of \cref{sec-k-planar}.
+I suggest we make it a subsection of \cref{Structure} where we introduce framed graphs, say they can be described by $(2,\binom{d}{2})$-shortcut systems, so the main theorem is applicable, which gives a treewidth bound of $\binom{5}{3}-1=9$, but using  \cref{framed} we get a treewidth bound of 3, which is optional.
+Then in latter sections we give two applications, for 1-planar graphs and for $d$-map graphs.
+}
+
+\note{PM}{I no longer agree about the restructuring. See my email essay for an explanation.}
+
+\note{PM}{Apropos the extra clause.
+It's related to the orange todo note (argue that $N'/\mathcal{P}=N/\mathcal{P}$).
+This doesn't need to be in the statement of the lemma since, unlike \cref{NewLemma} this one gives an $H$-partition of $N$, not $N'$, but something like it needs to be in the proof.}
 
@@
 Let $G$ be a 2-connected plane multigraph, in which every facial cycle has length in $\{3,\dots,d\}$ for some fixed integer $d\geq 3$.
-Then $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d +3( \floor{d/2}-1) }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
+Then $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d +3 \floor{d/2}-3 }$ for some planar graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
 \end{thm}
@@
 Grouping blocks of $\floor{\frac{d}{2}}$ consecutive layers in $\mathcal{L}$ gives a layering $\mathcal{L}'$ of $\widehat{G}$.
-Observe that each tripod in $\PP$ has at most $d+3( \floor{\frac{d}{2}}-1)$ vertices in each set of $\floor{\frac{d}{2}}$ layers of $\mathcal{L}$.
+Observe that each tripod in $\PP$ has at most $d+3\floor{\frac{d}{2}}-3$ vertices in each set of $\floor{\frac{d}{2}}$ layers of $\mathcal{L}$.
 Hence $\mathcal{L}'$ has layered width $d+3( \floor{\frac{d}{2}}-1)$ with respect to $\PP$.
-By \cref{PartitionProduct}, $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d + 3( \floor{d/2} -1 )}$ where $H:=\widehat{G}/\PP$, which has treewidth 3.
+By \cref{PartitionProduct}, $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{d+3\floor{d/2}-3}$ where $H:=\widehat{G}/\PP$, which has treewidth 3.
+\note{DW}{This proof needs to be changed to show that $G/\PP = \widehat{G}/\PP$, so $H$ is a minor of $G$ and is thus planar.}
 \end{proof}
@@
 	\label{New1Planar}
-	Every 1-planar graph is contained in $H \boxtimes P \boxtimes K_8$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
+	Every 1-planar graph is contained in $H \boxtimes P \boxtimes K_7$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
 \end{cor}

### 2022-03-13 k-planar.tex

--- 2022-03-13 k-planar.tex
@@
 
-  \todo[inline]{%
-    DW:Does this work for multigraph near-triangulations? \newline
-    PM: Yes, it only requires that the inner faces have three sides and that the outer face is bounded by a simple cycle.
+%  \todo[inline]{%
+%    DW:Does this work for multigraph near-triangulations? \newline
+%    PM: Yes, it only requires that the inner faces have three sides and that the outer face is bounded by a simple cycle.
 The proof goes through without change.

### 2022-03-13 k-planar.tex
Removed pat.sty since it was messing  with cleveref

@@
 \usepackage{amsthm,mathtools}
-\usepackage{pat}
+%\usepackage{pat}
 \usepackage{paralist}
@@
 \usepackage[noabbrev,capitalise]{cleveref}
+\crefname{thm}{Theorem}{Theorems}
 \crefname{lem}{Lemma}{Lemmas}
-\crefname{thm}{Theorem}{Theorems}
 \crefname{cor}{Corollary}{Corollaries}
@@
 \Crefformat{equation}{Equation #2(#1)#3}
+
+\theoremstyle{plain}
+\newtheorem{thm}{Theorem}
+\newtheorem{lem}[thm]{Lemma}
+\newtheorem{cor}[thm]{Corollary}
+\newtheorem{obs}[thm]{Observation}
+\newtheorem{prop}[thm]{Proposition}
+\newtheorem{clm}{Claim}
+\theoremstyle{definition}
+\newtheorem{open}[thm]{Open Problem}
+\newtheorem{conj}[thm]{Conjecture}
+\newtheorem{ques}[thm]{Question}
 
@@
 
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\N}{\mathbb{N}}
+\newcommand{\Z}{\mathbb{Z}}
+
 \title{\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}}
@@
 
-\begin{lem}\lemlabel{nice-decomposition}
+\begin{lem}\label{nice-decomposition}
   For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a width-$t$ $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:
@@
 \begin{lem}\label{generalized-tripod}
-  Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \lemref{nice-decomposition}.
+  Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
 For each $x\in V(T)$, let $V_x := \bigcup_{y\in V(T_x)} Y_y$, $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x:=V_x\cup F_x$.
@@
 \subsection{Centered Colourings}
-\seclabel{centered-colourings}
+\label{centered-colourings}
 
@@
 Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$, and 	$G_j / \PP_j$ has a tree-decomposition  $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3 such that for every facial cycle $F'$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that every part of $\PP_j$ that intersects $F'$ is in $B$.
+Since $C$ is a facial cycle of $G_j$, there is a bag $B_{u_j}$ of $G_j$ such that $P_1,\dots,P_k\in B_{u_j}$.
 

### 2022-03-14 k-planar.tex
added subsection on framed graphs

@@
 %%%%%%%%%%%%%%%%%%%
+\subsection{Framed Graphs}
+
+For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane skeleton (multi)graph $G_0$, whose faces have size at most $d$, and whose crossing edges are in the interior of these faces.
+This definition (for simple graphs) was introduced by \citet{BDGGMR}.
+Every $d$-framed graph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar graph.
+Thus \cref{ShortcutProduct} is applicable.
+However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.
+
+\begin{thm}\label{d_framed_product_stucture}
+Every $d$-framed graph is contained in $H\boxtimes P\boxtimes K_{d+3\floor{d/2}-3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+\end{thm}
+
+This result is used below to obtain product structure theorems for 1-planar graphs and for map graphs. For 1-planar graphs it is essential that we allow skeletons with parallel edges.
+
+\cref{d_framed_product_stucture} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
+The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{v_1,\ldots,v_t\}$.
+
+\begin{lem}
+	\label{induction} The setup:
+	\begin{compactenum}
+		\item Let $G$ be a $d$-framed graph and let $G'$ be the frame of $G$.
+		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
+		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
+		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that $V(F)$ can be partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
+		\begin{compactenum}
+			\item $F[P_i]$ is a path; and
+			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
+		\end{compactenum}
+		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
+	\end{compactenum}
+	Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
+	\begin{compactenum}[(i)]
+		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
+		\item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
+		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
+	\end{compactenum}
+\end{lem}
+
+\begin{proof}
+	This proof is very similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is complicated by several factors.
+In particular, the reader should keep in mind that $G$ and $G'$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
+
+	The proof is by induction on the number of inner faces of $N'$.
+A particular easy case occurs when $V(N')=V(F)$, for which we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
+	\begin{compactenum}[(i)]
+		\item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
+		\item Assumption 4(a) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
+		\item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
+	\end{compactenum}
+
+	In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
+There is one case to consider for each possible value of $k$:
+	\begin{compactenum}
+		\item If $k=1$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
+
+		\item If $k= 2$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
+
+		\item If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.
+(This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
+Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N'$, add edges $d'-2$ edges inside of $F$ to split it into $3$-sided faces.
+This yields a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary.
+By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.
+% (Note that $p\in\{3,4\}$.)
+	\end{compactenum}
+
+	%  \todo[inline]{%
+	%    DW:Does this work for multigraph near-triangulations? \newline
+	%    PM: Yes, it only requires that the inner faces have three sides and that the outer face is bounded by a simple cycle.
+The proof goes through without change.
+Look at the subgraph $X$ of the dual that contains only the edges dual to red/blue edges.
+In our setting, the outer face has degree $1$ in $X$, which is odd.
+Any bichromatic inner face has degree $0$ or degree $2$ in $X$.
+There has to be another vertex of odd degree somewhere, which can only come from a trichromatic inner face.\newline
+	%    DW: great. This is basically the same as the proof of the Hex Lemma.
+	%  }
+
+	We can now define the vertices in $S$.
+For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to a vertex in $V(F)$.
+Let $\overline{S}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
+Let $S:=V(\overline{S})\setminus V(F)$, let $Y:=(V(Q_1)\cup V(Q_2)\cup V(Q_3))\setminus V(F)$, and let $X:=S\setminus Y$.
+	Observe that, for each $i\in\{1,2,3\}$ and each $j\ge 0$, $|V(Q_i)\cap L_j|\le 1$.
+Therefore $|Y\cap L_j|\le 3$ for each integer $j\ge 0$.
+Furthermore, $|X|\le p-3\le d-3$, since $v_1,v_2,v_3\in Y\cup V(F)$.
+Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ will not violate condition~(ii).
+
+	% We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
+First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $L_i$, each incident on at most two edges of $\overline{Y}$.
+Since $\dist_{G'}(v,w)\le 2$ for each $vw\in E(G)$, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $L_i$.
+These at most six edges have at most 12 endpoints.
+Therefore $|V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i|\le 6\times 2=12$, so $|V(\overline{Y}^+)\cap L_i|\le 12+3=15$.
+
+	\begin{figure}
+		\begin{center}
+			\begin{tabular}{ccc}
+				\includegraphics{figs/zoomba-1} &
+				\includegraphics{figs/zoomba-2} &
+				\includegraphics{figs/zoomba-3} \\
+				$k=1$ & $k=2$ & $k=3$
+			\end{tabular}
+		\end{center}
+		\caption{Three cases when choosing the face $\tau$ of $N'$.}
+		\label{boring_figure}
+	\end{figure}
+
+	Refer to \cref{boring_figure}.
+Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$.
+By our choice of $\tau$, the graph $M$ is $2$-connected.
+Since $N'\subseteq G'$ contains only edges of $G$ that are not crossed, each component of $G-V(M)$ is contained in a single face of $M$.
+	Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
+By our choice of $v_1v_2v_3$, the vertices of each $F_i$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
+
+	In the following, in order to avoid introducing even more notation or constantly adding the modifier ``when it exists'', we will assume that $V(F_i)\cap Y\neq\emptyset$ and that there are exactly two distinct $b,c\in\{1,\ldots,k\}$ such that $V(F_i)\cap P_b\neq\emptyset$ and $V(F_i)\cap P_c\neq\emptyset$.
+When this assumption does not hold, exactly the same reasoning can be used, while omitting one or more of $P_{i,1}$, $P_{i,2}$ or $P_{i,3}$.
+
+	Since $M$ is $2$-connected and $\overline{Y}$ is connected, $F_i[P_{i,j}]$ is a path, for each $j\in\{1,2,3\}$.
+Let $N_i$ and $N_i'$ be the subgraphs of $N$ and $N'$ contained in the closure of $F_i$.
+Every inner face of $N_i'$ is an inner face of $N'$ and $\tau$ is not an inner face of $N_i'$.
+Therefore $N_i'$ has fewer inner faces than $N'$.
+Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.
+The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of the subgraph $N_i$ of $N$ contained in the closure of $F_i$.
+	% The partition $\mathcal{P}_i$ contains each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$.
+	We can now define our desired partition
+	\[
+	\mathcal{P}:=\{Y, P_1,\ldots,P_k\} \cup \bigcup_{i=1}^m \{P\in\mathcal{P}_i: P\cap V(M)=\emptyset\} \enspace .
+	\]
+	It is straightforward to verify that $\mathcal{P}$ is indeed a partition of $V(N)$ and, by definition this is an $H$-partition of $N$ for the graph $H:=N/\mathcal{P}$.
+All that remains is to verify that $\mathcal{P}$ satisfies (i)--(iii).
+
+	By construction $\mathcal{P}$ contains parts $P_1,\ldots,P_k$, so $\mathcal{P}$ satisifes (i).
+By Assumption~4(b) none of the parts $P_1,P_2,P_3$ violate condition~(ii).
+We have already argued that the part $S$ does not violate condition~(ii).
+The inductive hypothesis ensures that each of the remaining parts satisfies condition~(ii).
+Therefore $\mathcal{P}$ satisfies (ii).
+
+	All that remains is to construct a tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$ of $H$ that satisfies (iii).
+Let $x_0$ be the vertex of $H$ obtained from contracting $S$ and, for each $j\in\{1,\ldots,k\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
+The tree $T_0$ has a node $y_0$ with bag $B_{y_0}:=\{x_0,\ldots,x_k\}$.
+The inductive hypothesis ensures that, for each $i\in\{1,\ldots,m\}$,  $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $\mathcal{T}_i:=(B_z:z\in V(T_i))$ that satisfies (iii).
+Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
+Since $a_i$, $b_i$, $c_i$ form a clique in $H$, there exists a bag $B_{z_i}$ contains $\{a_i,b_i,c_i\}$.
+
+	Recall that $P_{i,1}:=V(N_i)\cap Y$, $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
+For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.
+These replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$.
+Now add an edge that joins $y_0$ to $z_i$ so that $T_i$ becomes a subtree of $T_0$ that is adjacent to $y_0$.
+We perform this operation for each $i\in\{1,\ldots,m\}$ to obtain our final tree decomposition $mathcal{T}:=(B_y:y\in V(T_0))$.
+
+	Using the fact that each component of $G-V(M)$ is contained in a single face of $M$, it is now straightforward to verify that $\mathcal{T}$ is a tree decomposition of $N$ in which each bag has size at most $4$.
+The bag $B_{y_0}$ contains $x_1,\ldots,x_k$.
+When $k=3$, each neighbour of $z_i$ of $y_0$ has a bag that contains at most two of $\{x_1,x_2,x_3\}$.
+Therefore, when $k=3$, exactly one bag of $\mathcal{T}$ contains $\{x_1,x_2,x_3\}$.
+
+	All that remains is to show that any triple of vertices of $H$ is contained in at most two bags of $\mathcal{T}$.
+By the inductive hypothesis, the only triples we need to be concerned about are the $3$-element subsets of  $\{x_0,x_1,\ldots,x_k\}$.
+The case $k=1$ is trivial since no triples can be formed by $\{x_0,x_1\}$.
+When $k=2$, $B_{y_0}=\{x_0,x_1,x_2\}$.
+In this case, there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_0$, and $P_1$ on its boundary.
+(Refer to \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.
+Therefore there are at most two bags of $\mathcal{T}$ that contains $\{x_0,x_1,x_2\}$.
+This establishes (iii) for $k\in\{1,2\}$ so, from this point on we assume that $k=3$.
+
+	We claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.
+Suppose to the contrary that there are two such faces $F_{i_1}$ and $F_{i_2}$.
+Create a graph $M^+$ from $M$ by placing a vertex $v_{i_b}$ inside $F_{i_b}$ and adjacent to each vertex in $V(F_{i_b})$ for each $b\in\{1,2\}$.
+Since $M$ is planar and $F_{i_1}$ and $F_{i_2}$ are distinct faces of $M$, $M^+$ is planar.
+However, contracting each of $S_{x_0},\ldots,S_{x_3}$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.
+(In this graph, the $x_0,x_a,x_b$ are on one side of the bipartition.)
+
+	We have already established that the only bag that contains $\{x_1,x_2,x_3\}$ is $B_{y_0}$, so we need only consider triples $\{x_0,x_a,x_b\}$ with $1\le a<b\le 3$.
+By the preceding claim, there is at most one face $F_i$ with vertices of $S$, $P_a$, and $P_b$ on its boundary.
+By the inductive hypothesis, the only other bag in $\mathcal{T}$ that contains $\{x_0,x_a,x_b\}$ is the bag $B_{z_i}$.
+This completes the proof.
+\end{proof}
+
+Using \cref{induction}, the proof of \cref{d_framed_product_stucture} is now straightforward.
+
+\begin{proof}[Proof of \cref{d_framed_product_stucture}]
+	Let $G$ be a $d$-framed multigraph with frame $G'$\todo{Deal with the case where $G'$ has no triangular face.} and assume that the outer face of $G'$ has exactly three vertices $v_1,v_2,v_3$ and that no edges of $G$ are in the outer face of $G'$.
+Add a vertex $r$ adjacent to $v_1$, $v_2$, and $v_3$ to obtain a $d$-framed multigraph $\overline{G}$ with frame $\overline{G}'$.
+
+	Let $T$ be a BFS tree of $\overline{G}'$ rooted at $r$ and let $\langle L_0,L_1,\ldots,\rangle$ be the resulting BFS layering.
+Now apply \cref{induction} with $G=\overline{G}$, $\overline{G}'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
+This gives an $H$-partition $\mathcal{P}:=\{S_x\cup\{v_x\}:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ has a $3$-simple tree decomposition.
+Since any graph having a $3$-simple tree decomposition is planar \cite{X}, this implies that $H$ is planar.
+
+	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdot \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$.
+That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.
+Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x| |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
+The result follows from \cref{PartitionProduct}.
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%
 \section{Allowing Crossings}
@@
 When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.
-The \emph{interior} of $f$ is the point set $\{f(t):0<t<1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
+The \defin{interior} of $f$ is the point set $\{f(t):0<t<1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
+The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
 A \defin{crossing} in an embedded (multi)graph $G$ is an unordered pair of distinct edges $\{vw,xy\}\subseteq E(G)$ whose interiors have a point in common.
@@
 
-\subsection{\boldmath $d$-Framed and $1$-Planar Graphs}
+\subsection{\boldmath $1$-Planar Graphs}
 \label{sec-1-planar}
 
-In the important special case of 1-planar graphs we obtain better constants and an additional property (planarity) of $H$.
-
-% \note{PM}{Include in note to referee:  We choose to keep this section because
-%   $1$-planar graphs are of special interest to the graph drawing community.
-For example, the annotated bibliography on $1$-planarity (\htmladdnormallink{https://arxiv.org/abs/1703.02261}{arXiv:1703.02261}) has 143 entries, and even this is now four years out of date.
+$1$-planar graphs are of particular interest to the graph drawing community~\citep{kobourov.liotta.ea:annotated}.
+In this case, we obtain better constants and an additional property (planarity) of $H$.
+
+%(\htmladdnormallink{https://arxiv.org/abs/1703.02261}{arXiv:1703.02261}) has 143 entries, and even this is now four years out of date.
 %
@@
 % }
-%
-% \note{DW}{I will rewrite the intro to this subsection once we decide what our results are.}
 
@@
 
-We will actually prove this result in the more general setting of $d$-framed graphs \cite{BDGGMR}.
-For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing having a biconnected plane skeleton (multi)graph $G_0$, whose faces have size at most $d$, and whose crossing edges are in the interior of these faces.
-In \cref{1_planar_is_4_framed}, below, we show that every $1$-planar graph is a subgraph of a $4$-framed multigraph.
+%In \cref{1_planar_is_4_framed}, below, we show that every $1$-planar graph is a subgraph of a $4$-framed multigraph.
 This makes \cref{1-planar} an immediate consequence of the following theorem, whose proof is the subject of the rest of this section.
 
-\begin{thm}\label{d_framed_product_stucture}
-  Every $d$-framed graph is contained in $H\boxtimes P\boxtimes K_{\lfloor d/2\rfloor + d - 3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
-\end{thm}
-
-\subsection{\boldmath $d$-Framed Graphs}
-
-The following definition is a convenient way to estalish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{v_1,\ldots,v_t\}$.
-Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
-\begin{lem}
-	\label{induction} The setup:
-	\begin{compactenum}
-		\item Let $G$ be a $d$-framed graph and let $G'$ be the frame of $G$.
-		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
-		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
-		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that $V(F)$ can be partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
-		\begin{compactenum}
-			\item $F[P_i]$ is a path; and
-			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
-		\end{compactenum}
-		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
-	\end{compactenum}
-	Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
-	\begin{compactenum}[(i)]
-		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
-        \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
-		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
-	\end{compactenum}
-\end{lem}
-
-\begin{proof}
-  This proof is very similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is complicated by several factors.
-In particular, the reader should keep in mind that $G$ and $G'$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
-
-  The proof is by induction on the number of inner faces of $N'$.
-A particular easy case occurs when $V(N')=V(F)$, for which we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
-  \begin{compactenum}[(i)]
-    \item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
-    \item Assumption 4(a) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
-    \item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
-  \end{compactenum}
-
-  In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
-There is one case to consider for each possible value of $k$:
-  \begin{compactenum}
-    \item If $k=1$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
-
-    \item If $k= 2$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
-
-    \item If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.
-(This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
-Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N'$, add edges $d'-2$ edges inside of $F$ to split it into $3$-sided faces.
-This yields a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary.
-By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
-Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.
-% (Note that $p\in\{3,4\}$.)
-  \end{compactenum}
-
-%  \todo[inline]{%
-%    DW:Does this work for multigraph near-triangulations? \newline
-%    PM: Yes, it only requires that the inner faces have three sides and that the outer face is bounded by a simple cycle.
-The proof goes through without change.
-Look at the subgraph $X$ of the dual that contains only the edges dual to red/blue edges.
-In our setting, the outer face has degree $1$ in $X$, which is odd.
-Any bichromatic inner face has degree $0$ or degree $2$ in $X$.
-There has to be another vertex of odd degree somewhere, which can only come from a trichromatic inner face.\newline
-%    DW: great. This is basically the same as the proof of the Hex Lemma.
-%  }
-
-  We can now define the vertices in $S$.
-For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to a vertex in $V(F)$.
-Let $\overline{S}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
-Let $S:=V(\overline{S})\setminus V(F)$, let $Y:=(V(Q_1)\cup V(Q_2)\cup V(Q_3))\setminus V(F)$, and let $X:=S\setminus Y$.
-  Observe that, for each $i\in\{1,2,3\}$ and each $j\ge 0$, $|V(Q_i)\cap L_j|\le 1$.
-Therefore $|Y\cap L_j|\le 3$ for each integer $j\ge 0$.
-Furthermore, $|X|\le p-3\le d-3$, since $v_1,v_2,v_3\in Y\cup V(F)$.
-Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ will not violate condition~(ii).
-
-	% We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
-First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $L_i$, each incident on at most two edges of $\overline{Y}$.
-Since $\dist_{G'}(v,w)\le 2$ for each $vw\in E(G)$, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $L_i$.
-These at most six edges have at most 12 endpoints.
-Therefore $|V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i|\le 6\times 2=12$, so $|V(\overline{Y}^+)\cap L_i|\le 12+3=15$.
-
-  \begin{figure}
-    \begin{center}
-      \begin{tabular}{ccc}
-        \includegraphics{figs/zoomba-1} &
-        \includegraphics{figs/zoomba-2} &
-        \includegraphics{figs/zoomba-3} \\
-        $k=1$ & $k=2$ & $k=3$
-      \end{tabular}
-    \end{center}
-    \caption{Three cases when choosing the face $\tau$ of $N'$.}
-    \label{boring_figure}
-  \end{figure}
-
-  Refer to \cref{boring_figure}.
-Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$.
-By our choice of $\tau$, the graph $M$ is $2$-connected.
-Since $N'\subseteq G'$ contains only edges of $G$ that are not crossed, each component of $G-V(M)$ is contained in a single face of $M$.
-  Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
-By our choice of $v_1v_2v_3$, the vertices of each $F_i$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
-
-  In the following, in order to avoid introducing even more notation or constantly adding the modifier ``when it exists'', we will assume that $V(F_i)\cap Y\neq\emptyset$ and that there are exactly two distinct $b,c\in\{1,\ldots,k\}$ such that $V(F_i)\cap P_b\neq\emptyset$ and $V(F_i)\cap P_c\neq\emptyset$.
-When this assumption does not hold, exactly the same reasoning can be used, while omitting one or more of $P_{i,1}$, $P_{i,2}$ or $P_{i,3}$.
-
-  Since $M$ is $2$-connected and $\overline{Y}$ is connected, $F_i[P_{i,j}]$ is a path, for each $j\in\{1,2,3\}$.
-Let $N_i$ and $N_i'$ be the subgraphs of $N$ and $N'$ contained in the closure of $F_i$.
-Every inner face of $N_i'$ is an inner face of $N'$ and $\tau$ is not an inner face of $N_i'$.
-Therefore $N_i'$ has fewer inner faces than $N'$.
-Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.
-The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of the subgraph $N_i$ of $N$ contained in the closure of $F_i$.
-  % The partition $\mathcal{P}_i$ contains each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$.
-  We can now define our desired partition
-  \[
-      \mathcal{P}:=\{Y, P_1,\ldots,P_k\} \cup \bigcup_{i=1}^m \{P\in\mathcal{P}_i: P\cap V(M)=\emptyset\} \enspace .
-  \]
-  It is straightforward to verify that $\mathcal{P}$ is indeed a partition of $V(N)$ and, by definition this is an $H$-partition of $N$ for the graph $H:=N/\mathcal{P}$.
-All that remains is to verify that $\mathcal{P}$ satisfies (i)--(iii).
-
-  By construction $\mathcal{P}$ contains parts $P_1,\ldots,P_k$, so $\mathcal{P}$ satisifes (i).
-By Assumption~4(b) none of the parts $P_1,P_2,P_3$ violate condition~(ii).
-We have already argued that the part $S$ does not violate condition~(ii).
-The inductive hypothesis ensures that each of the remaining parts satisfies condition~(ii).
-Therefore $\mathcal{P}$ satisfies (ii).
-
-  All that remains is to construct a tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$ of $H$ that satisfies (iii).
-Let $x_0$ be the vertex of $H$ obtained from contracting $S$ and, for each $j\in\{1,\ldots,k\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
-The tree $T_0$ has a node $y_0$ with bag $B_{y_0}:=\{x_0,\ldots,x_k\}$.
-The inductive hypothesis ensures that, for each $i\in\{1,\ldots,m\}$,  $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $\mathcal{T}_i:=(B_z:z\in V(T_i))$ that satisfies (iii).
-Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
-Since $a_i$, $b_i$, $c_i$ form a clique in $H$, there exists a bag $B_{z_i}$ contains $\{a_i,b_i,c_i\}$.
-
-  Recall that $P_{i,1}:=V(N_i)\cap Y$, $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
-For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.
-These replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$.
-Now add an edge that joins $y_0$ to $z_i$ so that $T_i$ becomes a subtree of $T_0$ that is adjacent to $y_0$.
-We perform this operation for each $i\in\{1,\ldots,m\}$ to obtain our final tree decomposition $mathcal{T}:=(B_y:y\in V(T_0))$.
-
-  Using the fact that each component of $G-V(M)$ is contained in a single face of $M$, it is now straightforward to verify that $\mathcal{T}$ is a tree decomposition of $N$ in which each bag has size at most $4$.
-The bag $B_{y_0}$ contains $x_1,\ldots,x_k$.
-When $k=3$, each neighbour of $z_i$ of $y_0$ has a bag that contains at most two of $\{x_1,x_2,x_3\}$.
-Therefore, when $k=3$, exactly one bag of $\mathcal{T}$ contains $\{x_1,x_2,x_3\}$.
-
-  All that remains is to show that any triple of vertices of $H$ is contained in at most two bags of $\mathcal{T}$.
-By the inductive hypothesis, the only triples we need to be concerned about are the $3$-element subsets of  $\{x_0,x_1,\ldots,x_k\}$.
-The case $k=1$ is trivial since no triples can be formed by $\{x_0,x_1\}$.
-When $k=2$, $B_{y_0}=\{x_0,x_1,x_2\}$.
-In this case, there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_0$, and $P_1$ on its boundary.
-(Refer to \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.
-Therefore there are at most two bags of $\mathcal{T}$ that contains $\{x_0,x_1,x_2\}$.
-This establishes (iii) for $k\in\{1,2\}$ so, from this point on we assume that $k=3$.
-
-  We claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.
-Suppose to the contrary that there are two such faces $F_{i_1}$ and $F_{i_2}$.
-Create a graph $M^+$ from $M$ by placing a vertex $v_{i_b}$ inside $F_{i_b}$ and adjacent to each vertex in $V(F_{i_b})$ for each $b\in\{1,2\}$.
-Since $M$ is planar and $F_{i_1}$ and $F_{i_2}$ are distinct faces of $M$, $M^+$ is planar.
-However, contracting each of $S_{x_0},\ldots,S_{x_3}$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.
-(In this graph, the $x_0,x_a,x_b$ are on one side of the bipartition.)
-
-  We have already established that the only bag that contains $\{x_1,x_2,x_3\}$ is $B_{y_0}$, so we need only consider triples $\{x_0,x_a,x_b\}$ with $1\le a<b\le 3$.
-By the preceding claim, there is at most one face $F_i$ with vertices of $S$, $P_a$, and $P_b$ on its boundary.
-By the inductive hypothesis, the only other bag in $\mathcal{T}$ that contains $\{x_0,x_a,x_b\}$ is the bag $B_{z_i}$.
-This completes the proof.
-\end{proof}
-
-Using \cref{induction}, the proof of \cref{d_framed_product_stucture} is now straightforward.
-
-\begin{proof}[Proof of \cref{d_framed_product_stucture}]
-    Let $G$ be a $d$-framed multigraph with frame $G'$\todo{Deal with the case where $G'$ has no triangular face.} and assume that the outer face of $G'$ has exactly three vertices $v_1,v_2,v_3$ and that no edges of $G$ are in the outer face of $G'$.
-Add a vertex $r$ adjacent to $v_1$, $v_2$, and $v_3$ to obtain a $d$-framed multigraph $\overline{G}$ with frame $\overline{G}'$.
-
-	Let $T$ be a BFS tree of $\overline{G}'$ rooted at $r$ and let $\langle L_0,L_1,\ldots,\rangle$ be the resulting BFS layering.
-Now apply \cref{induction} with $G=\overline{G}$, $\overline{G}'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
-This gives an $H$-partition $\mathcal{P}:=\{S_x\cup\{v_x\}:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ has a $3$-simple tree decomposition.
-Since any graph having a $3$-simple tree decomposition is planar \cite{X}, this implies that $H$ is planar.
-
-	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdot \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$.
-That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.
-Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x| |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
-The result follows from \cref{PartitionProduct}.
-\end{proof}
-
-\subsection{\boldmath $1$-Planar Graphs}
-
-In the following, we assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$, where we treat each edge $vw$ of $G$ as the image of a simple curve $f_{vw}:[0,1]\to\R^2$.
-% We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
+\cref{1-planar} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma, which is of independent interest.
+We assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  % We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
+
+\note{DW}{Say something about how \cref{1_planar_is_4_framed} is better than similar results in the literature.}
 
@@
 I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
-}
+Okay?}
 
@@
 \note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the treewidth 3 result in \citep{DHHW}.
-I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).}
+I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
+Okay?}
 
@@
 
+\note{DW}{Work in progress}
+
 \note{DW}{Use the language of framed graphs}
@@
 Let $G:=G_0\blah{C}$.
-Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that $P_1,\dots,P_k\in\PP$, and $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
+Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that:
+\begin{compactenum}[(i)]
+	\item $P_1,\dots,P_k\in\PP$, 
+	\item $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3,
+	\item $G/\PP=\widehat{G}/\PP$.
+%	 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
+\end{compactenum}
 \end{lem}
@@
 
-The proof is by induction on the number of faces of $G$.
-If $G$ has just one internal face (which is bounded by $C$), then the desired partition is $\PP=\{P_1,\ldots,P_k\}$ and the desired tree-decomposition of $G/\PP$ consists of a single bag $\{P_1,\ldots,P_k\}$ indexed by a 1-node tree.
+The proof is by induction on the number of faces of $G$. Suppose $G$ has just one internal face (which is bounded by $C$). Let $\PP=\{P_1,\ldots,P_k\}$.
+Then $G/\PP=\widehat{G}/\PP = K_k$.
+Then desired tree-decomposition of $G/\PP$ consists of a single bag $\{P_1,\ldots,P_k\}$ indexed by a 1-node tree.
 Now assume that $G$ has at least two internal faces.
@@
 By construction, $v_{i-1}$ is not in $G_j$ (where $v_0=v_3$).
-Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.
-Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$, and 	$G_j / \PP_j$ has a tree-decomposition  $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3 such that for every facial cycle $F'$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that every part of $\PP_j$ that intersects $F'$ is in $B$.
-Since $C$ is a facial cycle of $G_j$, there is a bag $B_{u_j}$ of $G_j$ such that $P_1,\dots,P_k\in B_{u_j}$.
+Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.  Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that 
+\begin{compactenum}[(i)]
+	\item
+	$R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$, 
+	\item $G_j / \PP_j$ has a tree-decomposition $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3,
+	\item $G_j / \PP=\widehat{G_j}/\PP$.
+	%	 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
+\end{compactenum}
 
@@
 By construction, $\PP$ partitions $V(G)$ into tripods with respect to $T$ and $P_1,\ldots,P_k\in \PP$.
+
+%Since $P_1,
+%Since $C$ is a facial cycle of $G_j$, there is a bag $B_{u_j}$ of $G_j$ such that $P_1,\dots,P_k\in B_{u_j}$.
 

### 2022-03-14 k-planar.tex
added new  map graph lemma

@@
 \usepackage{amsthm,mathtools}
-%\usepackage{pat}
 \usepackage{paralist}
@@
 
-%\usepackage{doi} % To make plainnat handle doi's properly
-
-% \usepackage[mathlines]{lineno}
-% \setlength{\linenumbersep}{2em}
-% \linenumbers
-% \rightlinenumbers
-% \linenumbers
-% \newcommand*\patchAmsMathEnvironmentForLineno[1]{%
-%  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
-%  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
-%  \renewenvironment{#1}%
-%     {\linenomath\csname old#1\endcsname}%
-%     {\csname oldend#1\endcsname\endlinenomath}}%
-% \newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
-%  \patchAmsMathEnvironmentForLineno{#1}%
-%  \patchAmsMathEnvironmentForLineno{#1*}}%
-% \AtBeginDocument{%
-% \patchBothAmsMathEnvironmentsForLineno{equation}%
-% \patchBothAmsMathEnvironmentsForLineno{align}%
-% \patchBothAmsMathEnvironmentsForLineno{flalign}%
-% \patchBothAmsMathEnvironmentsForLineno{alignat}%
-% \patchBothAmsMathEnvironmentsForLineno{gather}%
-% \patchBothAmsMathEnvironmentsForLineno{multline}%
-% }
-%
-%
 % \allowdisplaybreaks
@@
 
-\note{PM}{The trick I used (using the thm counter everywhere) to get consecutive Theorem, Lemma, etc numbering has broken cleverref.
-Everything is now references as a Theorem.
-Find a solution that works.}
-% \note{DW}{I suggest we use consecutive numbering of theorems, lemmas, etc. Theorem 1, Lemma 2, Observation 3, etc.}
-
-% \note{PM}{Agreed. Done.}
-
 % \note{DW}{I suggest we delete ``We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$. '' This is absolutely standard.}
@@
 Every $k$-planar graph is contained in $H\boxtimes P\boxtimes K_{18k^2+48k+30}$, for some graph $H$ of treewidth $\binom{k+4}{3}-1$ and for some path $P$.
+\note{DW}{Do we want to mention ``every $1$-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ of treewidth $3$ and for some path $P$'' here?}
 \end{thm}
@@
 
-For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane skeleton (multi)graph $G_0$, whose faces have size at most $d$, and whose crossing edges are in the interior of these faces.
+For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and whose crossing edges are in the interior of these faces \note{DW}{I suggest we replace ``and whose crossing edges are in the interior of these faces'' by ``such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$.
+} The graph $G_0$ is called the \defin{frame} of $G_0$.
 This definition (for simple graphs) was introduced by \citet{BDGGMR}.
-Every $d$-framed graph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar graph.
+Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar graph.
 Thus \cref{ShortcutProduct} is applicable.
-However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.
+However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible\footnote{A multigraph $G$ is \defin{contained} in a graph $G'$ if the simple graph underlying $G$ is isomorphic to a subgraph of $G'$.}.
 
 \begin{thm}\label{d_framed_product_stucture}
-Every $d$-framed graph is contained in $H\boxtimes P\boxtimes K_{d+3\floor{d/2}-3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+Every $d$-framed multigraph is contained in $H\boxtimes P\boxtimes K_{d+3\floor{d/2}-3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
 \end{thm}
 
-This result is used below to obtain product structure theorems for 1-planar graphs and for map graphs. For 1-planar graphs it is essential that we allow skeletons with parallel edges.
+This result is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs. For simple 1-planar graphs it is essential that we allow frames with parallel edges.
 
@@
 \label{1-planar}
-Every 1-planar graph is contained in $H\boxtimes P\boxtimes K_{7}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+Every 1-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
 \end{thm}
@@
 \cref{1-planar} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma, which is of independent interest.
-We assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  % We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
 
@@
 Let $G$ be a $1$-plane graph.
+We may assume that no two edges incident to a common vertex of $G$ cross, since such a crossing can be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}.
+  % We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
 We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.
@@
 
-\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{framed} with the treewidth 3 result in \citep{DHHW}.
+\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{framed} with the method in \citep{DHHW}.
 I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
@@
 
+\note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph. Is the following lemma known (at least in the $g=0$ case)?}
+
+\begin{lem}
+\label{MewMapGraphLemma}
+for all integers $g\geq 0$ and $d\geq 3$, every $(g,d)$-map graph is a spanning subgraph of  $(g,d)$-framed graph.
+\end{lem}
+
+\begin{proof}
+Let $G_0$ be a graph embedded in a surface of Euler genus $g$, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations.
+If two lakes have a common edge in their boundaries, then we may delete the edges and merge the lakes.
+Now assume that no two lakes have a common edge in their boundaries.
+Let $G$ be the corresponding map graph. 
+Let $G_1$ be the graph with $V(G_1):=V(G)$, which is the set of nations. 
+Consider vertices $v$ and $w$ of $G_1$. 
+Let $A$ and $B$ be the nation faces of $G_0$ corresponding to $v$ and $w$.
+Then $vw\in E(G_1)$ if $A$ and $B$ have a common edge in their boundaries, 
+or there is a vertex $x$ of $G_0$ in the boundary of $A$ and $B$, and in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes. 
+Then $G_1$ has Euler genus at most $g$. 
+The faces of $G_1$ correspond to vertices of $G_0$ and lakes of $G_0$.
+Triangulate each face of $G_1$ corresponding to a lake of $G_0$. For each vertex 
+$x$ of $G_0$ the size of the face of $G_1$ corresponding to $x$ equals the number of nations incident to $x$, which is at most $d$. 
+So every face of $G_1$ has size at most $d$. 
+If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common, implying that  $v$ and $w$ are both on the face of $G_1$ corresponding to $x$, and $vw\in E(\widehat{G_1})$.
+Hence $G\subseteq \widehat{G_1}$.
+\end{proof}
+
+\begin{thm}
+\label{NewMapGraph}
+Every $d$-map graph is contained in $H \boxtimes P \boxtimes K_{ d + 3\floor{d/2} -3 }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
+\end{thm}
+
 Map graphs are defined as follows.
@@
 The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}. 
-So $(g,d)$-map graphs generalise graphs embedded in a surface, and we now assume that $d\geq 4$ for the remainder of this section.
+%So $(g,d)$-map graphs generalise graphs embedded in a surface, and we now assume that $d\geq 4$ for the remainder of this section.
 
@@
 
-\note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the treewidth 3 result in \citep{DHHW}.
+\note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the method in \citep{DHHW}.
 I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
@@
 
-\begin{lem}
-Every $d$-map graph is a subgraph of $\widehat{G}$ for some 2-connected plane graph $G$ in which every facial cycle has length in $\{3,\dots,d\}$.
-\end{lem}
-
-\begin{proof}
-\end{proof}
-
-\begin{cor}
-\label{NewMapGraph}
-Every $d$-map graph is contained in $H \boxtimes P \boxtimes K_{ d + 3\floor{d/2} -3 }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
-\end{cor}
-
 \end{document}

### 2022-03-14 k-planar.tex
Pfffft

@@
 Every $k$-planar graph is contained in $H\boxtimes P\boxtimes K_{18k^2+48k+30}$, for some graph $H$ of treewidth $\binom{k+4}{3}-1$ and for some path $P$.
+\end{thm}
+
 \note{DW}{Do we want to mention ``every $1$-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ of treewidth $3$ and for some path $P$'' here?}
+
+\note{PM}{Yes, I think so.  Two theorems, like this, or one?}
+
+For the important special case of $1$-planar graphs, we are able to obtain better constants and an additional property (planarity) on $H$.
+
+\begin{thm}\label{1_planar_product}
+  Every $1$-planar graph is contained in $H\boxtimes P\boxtimes K_{7}$, for some planar graph $H$ of treewidth $3$ and for some path $P$.
 \end{thm}
@@
 
-For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and whose crossing edges are in the interior of these faces \note{DW}{I suggest we replace ``and whose crossing edges are in the interior of these faces'' by ``such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$.
-} The graph $G_0$ is called the \defin{frame} of $G_0$.
+For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$.
+% \note{DW}{I suggest we replace ``and whose crossing edges are in the interior of these faces'' by ``such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$. }
+The graph $G_0$ is called the \defin{frame} of $G_0$.
 This definition (for simple graphs) was introduced by \citet{BDGGMR}.
@@
 	\begin{compactenum}
-		\item Let $G$ be a $d$-framed graph and let $G'$ be the frame of $G$.
+		\item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
 		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
@@
 
+% \todo[inline]{Alternative to the final three paragraphs in the preceding proof:
+%
+% To see that $H$ is planar, consider the plane multigraph $G''$ obtained by adding a vertex in the interior of each face of $G'$.
+% }
+
 Using \cref{induction}, the proof of \cref{d_framed_product_stucture} is now straightforward.
+
+\note{PM}{Annoying detail:  When we allow framed graphs to have crossings in the outer face, we can't guarantee that $G'$ has any face smaller than $d$.
+This makes it awkward to start the induction using \cref{induction}.
+Possible solutions:
+\begin{compactenum}
+  \item Disallow the addition of edges in the outer face of a frame.
+  \item Modify \cref{induction} so that $T$ is a BFS forest rooted at the set of vertices on the outer face of $G'$. Then we can start the induction with $k=1$.
+\end{compactenum}
+I prefer the second option.  Thoughts?
+}
 

### 2022-03-14 k-planar.tex
Other way of showing planarity

@@
 
-% \todo[inline]{Alternative to the final three paragraphs in the preceding proof:
-%
-% To see that $H$ is planar, consider the plane multigraph $G''$ obtained by adding a vertex in the interior of each face of $G'$.
-% }
+\todo[inline]{Alternative to the final three paragraphs in the preceding proof:
+
+To see that $H$ is planar, imagine adding a vertex consider a plane multigraph $G''$ obtained by adding a vertex $v_\tau$ in the interior of some of the faces of $G'$.
+When creating a new part $S:=V(\tau\cup Q_1\cup Q_2\cup Q_3-F)$ of $\mathcal{P}$, add a vertex $v_\tau$ to the interior of $\tau$ that is adjacent to each vertex of $\tau$.
+The graph $G''\supseteq G'$ obtained this way is clearly planar.
+Furthermore, for each part $P\in\mathcal{P}$, $G''[P]$ is connected.
+Since planarity is preserved under edge contraction, it follows that $G''/\mathcal{P}$ is planar.
+Now argue that $G/\mathcal{P}=G''/\mathcal{P}$.
+}
 

### 2022-03-14 k-planar.tex
Pfffff

@@
 
-To see that $H$ is planar, imagine adding a vertex consider a plane multigraph $G''$ obtained by adding a vertex $v_\tau$ in the interior of some of the faces of $G'$.
-When creating a new part $S:=V(\tau\cup Q_1\cup Q_2\cup Q_3-F)$ of $\mathcal{P}$, add a vertex $v_\tau$ to the interior of $\tau$ that is adjacent to each vertex of $\tau$.
+To see that $H$ is planar, consider a plane multigraph $G''$ obtained by adding stars in some of the faces of $G'$, as follows:  When creating the part $S$, add a vertex $v_\tau$ to the interior of $\tau$ that is adjacent to each vertex of $\tau$ and place $v_\tau$ in $S$.
 The graph $G''\supseteq G'$ obtained this way is clearly planar.
@@
 Since planarity is preserved under edge contraction, it follows that $G''/\mathcal{P}$ is planar.
-Now argue that $G/\mathcal{P}=G''/\mathcal{P}$.
+Now argue that $G/\mathcal{P}=G''/\mathcal{P}$\ldots
 }

### 2022-03-14 k-planar.tex
Using BFS forest

@@
 The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{v_1,\ldots,v_t\}$.
+Finally, a \emph{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that $\dist_G(v,V_0)=\dist_T(v,V_0)$ for each $v\in V(G)$.
 
@@
 		\item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
-		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
-		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
+		\item Let $F_0$ be the outer face of $G'$ and let $T$ be a BFS spanning forest of $G'$ rooted at $V(F_0)$.
+		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
 		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that $V(F)$ can be partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
 		\begin{compactenum}
-			\item $F[P_i]$ is a path; and
+			\item if $k\in\{2,3\}$ then $F[P_i]$ is a path; and
 			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
@@
 
-		\item If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.
+		\item If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the appropriate vertex of $F_0$.
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
@@
 
-\note{PM}{Annoying detail:  When we allow framed graphs to have crossings in the outer face, we can't guarantee that $G'$ has any face smaller than $d$.
+% \note{PM}{Annoying detail:  When we allow framed graphs to have crossings in the outer face, we can't guarantee that $G'$ has any face smaller than $d$.
 This makes it awkward to start the induction using \cref{induction}.
 Possible solutions:
-\begin{compactenum}
-  \item Disallow the addition of edges in the outer face of a frame.
-  \item Modify \cref{induction} so that $T$ is a BFS forest rooted at the set of vertices on the outer face of $G'$. Then we can start the induction with $k=1$.
-\end{compactenum}
-I prefer the second option.  Thoughts?
-}
+% \begin{compactenum}
+%   \item Disallow the addition of edges in the outer face of a frame.
+%   \item Modify \cref{induction} so that $T$ is a BFS forest rooted at the set of vertices on the outer face of $G'$. Then we can start the induction with $k=1$.
+% \end{compactenum}
+% I prefer the second option.  Thoughts?
+% }
 
 \begin{proof}[Proof of \cref{d_framed_product_stucture}]
-	Let $G$ be a $d$-framed multigraph with frame $G'$\todo{Deal with the case where $G'$ has no triangular face.} and assume that the outer face of $G'$ has exactly three vertices $v_1,v_2,v_3$ and that no edges of $G$ are in the outer face of $G'$.
-Add a vertex $r$ adjacent to $v_1$, $v_2$, and $v_3$ to obtain a $d$-framed multigraph $\overline{G}$ with frame $\overline{G}'$.
-
-	Let $T$ be a BFS tree of $\overline{G}'$ rooted at $r$ and let $\langle L_0,L_1,\ldots,\rangle$ be the resulting BFS layering.
-Now apply \cref{induction} with $G=\overline{G}$, $\overline{G}'$, $F$, and $P_i=\{v_i\}$ for each $i\in\{1,2,3\}$.
-This gives an $H$-partition $\mathcal{P}:=\{S_x\cup\{v_x\}:x\in V(H)\}$ of $\overline{G}-\{r\}\supseteq G$ in which $H$ has a $3$-simple tree decomposition.
+	Let $G$ be a $d$-framed multigraph with frame $G'$ having outer face $F_0$.	Let $T$ be a BFS forest of $G'$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.
+Let $P_1:= V(F_0)$, let $Y_1$ be any subset of $\min\{|P_1|,3\}$ vertices in $P_1$ and let $X_1:=P_1\setminus Y_1$.
+Now apply \cref{induction} on $G$, $G'$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$.
+    This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree decomposition.
 Since any graph having a $3$-simple tree decomposition is planar \cite{X}, this implies that $H$ is planar.
@@
 That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.
-Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x| |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
+Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
 The result follows from \cref{PartitionProduct}.

### 2022-03-14 k-planar.tex
Filled in some handwaving details

@@
 
-	Using the fact that each component of $G-V(M)$ is contained in a single face of $M$, it is now straightforward to verify that $\mathcal{T}$ is a tree decomposition of $N$ in which each bag has size at most $4$.
+  It is now straightforward to verify that $\mathcal{T}$ is a tree decomposition of $H':=N'/\mathcal{P}$ in which each bag has size at most $4$.
+To see that $\mathcal{T}$ is also a tree decomposition of $H:=N/\mathcal{P}$, consider any edge $vw\in E(N)\setminus E(N')$ with $v\in P_\alpha$ and $w\in P_\beta$.
+We must verify that some bag of $\mathcal{T}$ contains both $\alpha$ and $\beta$.
+If both $v$ and $w$ are in $V(M)$, then $\alpha=x_a$ and $\beta=x_b$ for some $a,b\in\{0,\ldots,3\}$, in which case $B_{y_0}=\{x_0,\ldots,x_k\}$ contains $\alpha$ and $\beta$.
+If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $G$ crosses an edge of $M$, $v$ or $w$ are both contained in the interior of the same face $F_i$ of $M$ so, by the inductive hypothesis there exists a bag of $\mathcal{T}_i$ that contains $\alpha$ and $\beta$ and this becomes a bag of $\mathcal{T}$ that contains $\alpha$ and $\beta$.
+Finally, if $v\in V(M)$ and $w\not\in V(M)$ then $v\in P_{i,j}$ for some $j\in\{1,2,3\}$ and $w$ is in the interior of some face $F_i$ of $M$.
+By induction, some bag of $\mathcal{T}_i$ contains $\beta$ and $\alpha'$, where $\alpha'$ is the vertex of $H_i$ obtained by contracting $P_{i,j}$.
+However, before $T_i$ is attached to $T_0$ each occurence of $\alpha'$ in $\mathcal{T}_i$ is replaced by $\alpha$, so some bag of $\mathcal{T}$ contains both $\alpha$ and $\beta$.
+is renamed to $\alpha$.
+Therefore $\mathcal{T}$ is a tree decomposition of $H$.
+
+  Next we verify that, when $k=3$ excactly one bag of $T$ contains $\{x_1,\ldots,x_3\}$.
 The bag $B_{y_0}$ contains $x_1,\ldots,x_k$.
@@
 When $k=2$, $B_{y_0}=\{x_0,x_1,x_2\}$.
-In this case, there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_0$, and $P_1$ on its boundary.
+In this case, the choice of $\tau$ ensures that there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_1$, and $P_2$ on its boundary.
 (Refer to \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.
@@
 However, contracting each of $S_{x_0},\ldots,S_{x_3}$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.
-(In this graph, the $x_0,x_a,x_b$ are on one side of the bipartition.)
+(In this contracted graph, $x_0,x_a,x_b$ are on one side of the bipartition and $v_{i_a}$, $v_{i_b}$, and the node in $\{x_1,x_2,x_3\}\setminus\{x_a,x_b\}$ is on the other side.)
 

### 2022-03-14 k-planar.tex
Touch ups

@@
 \newtheorem{ques}[thm]{Question}
+
+\definecolor{brew2}{rgb}{1.0, 1.0, 0.701960784314}
+\definecolor{brew8}{rgb}{0.988235294118, 0.803921568627, 0.898039215686}
 
@@
 
-For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$.
+For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
 % \note{DW}{I suggest we replace ``and whose crossing edges are in the interior of these faces'' by ``such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$. }
@@
 Thus \cref{ShortcutProduct} is applicable.
-However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible\footnote{A multigraph $G$ is \defin{contained} in a graph $G'$ if the simple graph underlying $G$ is isomorphic to a subgraph of $G'$.}.
+However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.\footnote{A multigraph $G$ is \defin{contained} in a graph $G'$ if the simple graph underlying $G$ is isomorphic to a subgraph of $G'$.}
 
@@
 
-This result is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs. For simple 1-planar graphs it is essential that we allow frames with parallel edges.
+This result is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs.
+\todo{Move this comment to the section on $1$-planar graphs?} For simple 1-planar graphs it is essential that we allow frames with parallel edges.
 
@@
 The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{v_1,\ldots,v_t\}$.
-Finally, a \emph{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that $\dist_G(v,V_0)=\dist_T(v,V_0)$ for each $v\in V(G)$.
+Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that $\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}$ for each $v\in V(G)$.
 
@@
 		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
-		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that $V(F)$ can be partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
+		\item Let $F$ be a cycle in $G'$ that $V(F)$ is partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
 		\begin{compactenum}
-			\item if $k\in\{2,3\}$ then $F[P_i]$ is a path; and
+			\item $F[P_i]$ is connected; and
 			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
@@
 		\item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
-		\item Assumption 4(a) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
+		\item Assumption 4(b) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
 		\item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
@@
 	In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
-There is one case to consider for each possible value of $k$:
+There is one case to consider for each possible value of $k$ (see \cref{boring_figure}):
 	\begin{compactenum}
@@
 		\begin{center}
-			\begin{tabular}{ccc}
+			\begin{tabular}{c@{}c@{}c}
 				\includegraphics{figs/zoomba-1} &
@@
 				\includegraphics{figs/zoomba-3} \\
+        \multicolumn{3}{c}{{\color{brew8}\raisebox{-3pt}{\rule{12pt}{12pt}}} $S$\qquad {\color{brew2}\raisebox{-3pt}{\rule{12pt}{12pt}}} $X$} \\
 				$k=1$ & $k=2$ & $k=3$
@@
 	Refer to \cref{boring_figure}.
-Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$.
+Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{S}$ and the edges and vertices of $F$.
 By our choice of $\tau$, the graph $M$ is $2$-connected.
-Since $N'\subseteq G'$ contains only edges of $G$ that are not crossed, each component of $G-V(M)$ is contained in a single face of $M$.
+  % Since $N'\subseteq G'$ contains only edges of $N\subseteq G$ that are not crossed, each component of $N-V(M)$ is contained in a single face of $M$.
 	Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
@@
 
-	In the following, in order to avoid introducing even more notation or constantly adding the modifier ``when it exists'', we will assume that $V(F_i)\cap Y\neq\emptyset$ and that there are exactly two distinct $b,c\in\{1,\ldots,k\}$ such that $V(F_i)\cap P_b\neq\emptyset$ and $V(F_i)\cap P_c\neq\emptyset$.
+	In the following, in order to avoid introducing even more notation or constantly adding the modifier ``when it exists'', we will assume that $V(F_i)\cap S\neq\emptyset$ and that there are exactly two distinct $b,c\in\{1,\ldots,k\}$ such that $V(F_i)\cap P_b\neq\emptyset$ and $V(F_i)\cap P_c\neq\emptyset$.
 When this assumption does not hold, exactly the same reasoning can be used, while omitting one or more of $P_{i,1}$, $P_{i,2}$ or $P_{i,3}$.
 
-	Since $M$ is $2$-connected and $\overline{Y}$ is connected, $F_i[P_{i,j}]$ is a path, for each $j\in\{1,2,3\}$.
+	Since $M$ is $2$-connected and $\overline{S}$ is connected, $F_i[P_{i,j}]$ is connected, for each $j\in\{1,2,3\}$.
 Let $N_i$ and $N_i'$ be the subgraphs of $N$ and $N'$ contained in the closure of $F_i$.
@@
 Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.
-The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of the subgraph $N_i$ of $N$ contained in the closure of $F_i$.
+The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of $N_i$.
 	% The partition $\mathcal{P}_i$ contains each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$.
-	We can now define our desired partition
+	We can now define our partition
 	\[
@@
 Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
-Since $a_i$, $b_i$, $c_i$ form a clique in $H$, there exists a bag $B_{z_i}$ contains $\{a_i,b_i,c_i\}$.
-
-	Recall that $P_{i,1}:=V(N_i)\cap Y$, $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
+Since $a_i$, $b_i$, $c_i$ form a clique in $H$, there exists a bag $B_{z_i}$ in $\mathcal{T}_i$ that contains $\{a_i,b_i,c_i\}$.
+
+	Recall that $P_{i,1}:=V(N_i)\cap S$, $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
 For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.
@@
 Now add an edge that joins $y_0$ to $z_i$ so that $T_i$ becomes a subtree of $T_0$ that is adjacent to $y_0$.
-We perform this operation for each $i\in\{1,\ldots,m\}$ to obtain our final tree decomposition $mathcal{T}:=(B_y:y\in V(T_0))$.
+We perform this operation for each $i\in\{1,\ldots,m\}$ to obtain our final tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$.
 
@@
 To see that $\mathcal{T}$ is also a tree decomposition of $H:=N/\mathcal{P}$, consider any edge $vw\in E(N)\setminus E(N')$ with $v\in P_\alpha$ and $w\in P_\beta$.
-We must verify that some bag of $\mathcal{T}$ contains both $\alpha$ and $\beta$.
+We must verify that some bag of $\mathcal{T}$ contains $\alpha$ and $\beta$.
 If both $v$ and $w$ are in $V(M)$, then $\alpha=x_a$ and $\beta=x_b$ for some $a,b\in\{0,\ldots,3\}$, in which case $B_{y_0}=\{x_0,\ldots,x_k\}$ contains $\alpha$ and $\beta$.
-If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $G$ crosses an edge of $M$, $v$ or $w$ are both contained in the interior of the same face $F_i$ of $M$ so, by the inductive hypothesis there exists a bag of $\mathcal{T}_i$ that contains $\alpha$ and $\beta$ and this becomes a bag of $\mathcal{T}$ that contains $\alpha$ and $\beta$.
+If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $N\subseteq G$ crosses an edge of $M\subseteq G'$, both $v$ and $w$ are contained in the interior of the same face $F_i$ of $M$ so, by the inductive hypothesis there exists a bag of $\mathcal{T}_i$ that contains $\alpha$ and $\beta$ and this becomes a bag of $\mathcal{T}$ that contains $\alpha$ and $\beta$.
 Finally, if $v\in V(M)$ and $w\not\in V(M)$ then $v\in P_{i,j}$ for some $j\in\{1,2,3\}$ and $w$ is in the interior of some face $F_i$ of $M$.
@@
 However, before $T_i$ is attached to $T_0$ each occurence of $\alpha'$ in $\mathcal{T}_i$ is replaced by $\alpha$, so some bag of $\mathcal{T}$ contains both $\alpha$ and $\beta$.
-is renamed to $\alpha$.
 Therefore $\mathcal{T}$ is a tree decomposition of $H$.
 
-  Next we verify that, when $k=3$ excactly one bag of $T$ contains $\{x_1,\ldots,x_3\}$.
+  Next we verify that, when $k=3$ exactly one bag of $\mathcal{T}$ contains $\{x_1,\ldots,x_3\}$.
 The bag $B_{y_0}$ contains $x_1,\ldots,x_k$.
 When $k=3$, each neighbour of $z_i$ of $y_0$ has a bag that contains at most two of $\{x_1,x_2,x_3\}$.
-Therefore, when $k=3$, exactly one bag of $\mathcal{T}$ contains $\{x_1,x_2,x_3\}$.
+Therefore, when $k=3$, $B_y$ is the unique bag in $\mathcal{T}$ that contains $\{x_1,x_2,x_3\}$.
 
@@
 (Refer to \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.
-Therefore there are at most two bags of $\mathcal{T}$ that contains $\{x_0,x_1,x_2\}$.
+Therefore there are at most two bags, $B_y$ and $B_{z_i}$ of $\mathcal{T}$ that contain $\{x_0,x_1,x_2\}$.
 This establishes (iii) for $k\in\{1,2\}$ so, from this point on we assume that $k=3$.
@@
 Since $M$ is planar and $F_{i_1}$ and $F_{i_2}$ are distinct faces of $M$, $M^+$ is planar.
-However, contracting each of $S_{x_0},\ldots,S_{x_3}$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.
+However, contracting each of $S_{x_0},\ldots,S_{x_3}$ in $M^+$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.
 (In this contracted graph, $x_0,x_a,x_b$ are on one side of the bipartition and $v_{i_a}$, $v_{i_b}$, and the node in $\{x_1,x_2,x_3\}\setminus\{x_a,x_b\}$ is on the other side.)
 
-	We have already established that the only bag that contains $\{x_1,x_2,x_3\}$ is $B_{y_0}$, so we need only consider triples $\{x_0,x_a,x_b\}$ with $1\le a<b\le 3$.
+	We have already established that, when $k=3$, the only bag that contains $\{x_1,x_2,x_3\}$ is $B_{y_0}$, so we need only consider triples $\{x_0,x_a,x_b\}$ for distinct $a,b\in\{1,2,3\}$.
 By the preceding claim, there is at most one face $F_i$ with vertices of $S$, $P_a$, and $P_b$ on its boundary.
@@
 
-	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdot \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$.
+	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdots \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$.
 That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.
 Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
-The result follows from \cref{PartitionProduct}.
+The result now follows from \cref{PartitionProduct}.
 \end{proof}

### 2022-03-15 k-planar.tex
Consistent notation

@@
 
-For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
+For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane spanning (multi) subgraph $G'$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G')$ is contained in the interior of some face of $G'$.
 % \note{DW}{I suggest we replace ``and whose crossing edges are in the interior of these faces'' by ``such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$. }
-The graph $G_0$ is called the \defin{frame} of $G_0$.
+The graph $G'$ is called the \defin{frame} of $G'$.
 This definition (for simple graphs) was introduced by \citet{BDGGMR}.

### 2022-03-15 response.tex
Started to write a response

@@
 \usepackage{stmaryrd}
+
+\usepackage{todonotes}
 
@@
 Major comments:
+
+The following comments are all related to the proof of Theorem~3.
+Much of the difficulty comes from the fact that there are many things to keep track of.
+We are a given a graph $G$, a shortcut system $\mathcal{S}$, a layering $\mathcal{L}$, a vertex-partition $\mathcal{P}$, and a tree decomposition $\mathcal{T}$ of the quotient graph $H:=G/\mathcal{P}$.
+Now we have to use these to show the existence of a new layering $\mathcal{L}'$, a new vertex partition $\mathcal{P'}$, a new tree decomposition $\mathcal{T}'$ of the new quotient graph $H':=G^{\mathcal{S}}$.
+
+During the proof, a number of related sets come up, which we have given names to. As a referee point out, these could remain unnamed and be described using an appropriate notation.
+
+\todo[inline]{Write something more here depending on how the experiment with new notation works out.}
+
+A secondary issue is that a number of the proofs in this section are essentially "definition checking".
+The relationships between the various vertex sets that arise in the proof really do just follow from their definitions (which were made with this purpose in mind).
+We have done this checking explicitly, though each of these proofs could also be replaced by a sentence of the form ``This [claim/lemma] follows from the definitions of\ldots''.
+It is not clear what the best choice is.
+
+The real action begins in the proof Theorem 9 (using numbering from the original submission) and especially Claim 3 (which involves some careful counting) and Claim 5 (which involves checking that we have everything needed to apply the Pillipczuk-Siebertz lemma on $k$-reachability in ``upward-directed'' graphs of treewidth $t$.)
+
+\todo[inline]{Write more here depending on how the notation experiment turns out.}
+
+ % and new vertex partition of a related graph $G^\mathcal{S}$ as we  A large part of the
 

### 2022-03-15 k-planar.tex
Pfffft

@@
 
+\note{PM}{I've added to the \texttt{response.tex} file explaining some of this.
+Each of (Y1), (Y3), (Y4), and (Y5) have a one-sentence justification.
+The justification for (Y2) is just definition unpacking and then using the observation that the referee recommends.}
+
+\note{PM}{I plan to do an experiment with the referee's suggested notation, in a separate git branch to see how it turns out.}
+
 Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:
@@
 Let $z_0,\ldots,z_r$ be the path in $T$ from $z_0:=x'$ to $z_r:=a$.
-For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $B_{z_i}$, since $x'a\in E(H)$.
-However, by \tref{subtree-root} $x'$ is not contained in $B_{x_i}$ for any $i\in\{1,\ldots,r\}$.
-Therefore $a\in B_{x_i}$ for each $i\in\{0,\ldots,r\}$.
+Since $x'a\in E(H)$, at least one of $a$ or $x'$ is in $B_{z_i}$, for each $i\in\{0,\ldots,r\}$ and there is at least one $i$ such that $B_{z_i}$ contains both $x'$ and $a$.
+However, by \tref{subtree-root} $x'$ is not contained in $B_{z_i}$ for any $i\in\{1,\ldots,r\}$.
+Therefore $a\in B_{z_i}$ for each $i\in\{0,\ldots,r\}$.
 In particular, $a$ is contained in $B_x$.
@@
 
-\referee{2}{Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
-(twice).  Moreover, the way the proof is written, it seems to suggest
-that we only get that $a \in B_{z_i}$ for each $i \in \{1, \dots, r\}$.
-The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
-true though.  In a normalized tree-decomposition, if $xy$ is an edge and
-$x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
-and $T_y$ intersect).}
+% \referee{2}{Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
+% (twice).  Moreover, the way the proof is written, it seems to suggest
+% that we only get that $a \in B_{z_i}$ for each $i \in \{1, \dots, r\}$.
+% The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
+% true though.  In a normalized tree-decomposition, if $xy$ is an edge and
+% $x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
+% and $T_y$ intersect).}
+
+%\note{PM}{I addressed the preceding referee's.}
 
@@
 notation as they are quite straightforward.}
+
+\note{PM}{Same comment: Yes, these claims are definition checking.  We'll see if the experimental notation change helps.}
 
@@
 
+\note{PM}{I also found the steps in this proof tedious.
+Unfortunately, when I tried to give a simple explanation for it in Banff, both Piotr and Torsten remained unconvinced until we opened the paper and walked through the steps of the proof together.
+I suspect the referee who can see through these proofs if one of the authors of \cite{PS21}.}
+
 \begin{proof}
@@
 
-\referee{2}{Page 18.  Say $H'$ is connected since $X$ is connected.}
+% \referee{2}{Page 18.  Say $H'$ is connected since $X$ is connected.}
+% \note{PM}{This is in the wrong place. It refers to the proof about centered colouring, which is now just a reference.}
 
@@
 \note{DW}{Say something about how \cref{1_planar_is_4_framed} is better than similar results in the literature.}
+\note{PM}{I don't know what to say without explaining that $4$-framed (simple) graphs only contain $3$-connected $1$-planar graphs.
+Vida's example shows that there is a $1$-plane embedding of a graph that can't be augmented into a $4$-framed graph, but I haven't checked if it can't be drawn some better way.
+And I also don't know if the Bekos claim that $8$-framed (smple) graphs contain all $1$-planar graphs is proven anywhere and, if so, is it tight?}
 
@@
 Every $(g,k)$-planar graph is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with $\tw(H) \leq \binom{k+4}{3}-1$, where $\ell:=\max\{2g,3\}\cdot(6k^2+16k+10)$.
-\note{DW}{I have changed $\tw(H) \leq \binom{k+5}{4}-1$ to $\tw(H) \leq \binom{k+4}{3}-1$, okay?}
+% \note{DW}{I have changed $\tw(H) \leq \binom{k+5}{4}-1$ to $\tw(H) \leq \binom{k+4}{3}-1$, okay?}
+%\note{PM}{Ok}
 \end{thm}
@@
 Okay?}
+\note{PM}{Okay to this and to all similar questions that come later (which I've now commented out.)}
 
@@
 \note{DW}{I don't want to do a complete restructuring. In our response, let's just say why we structure it the way we do.}
+\note{PM}{Sure, though I think they proposing a chop rather than a restructuring.}
 
@@
 
-\cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$. \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
+\cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$. 
+% \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
+% \note{OM}{Ok}
 Note that \citet{DJMMUW20} previously proved the bound of
-$O(g^{k+2})$ using \cref{GenusProduct} and an ad-hoc method.
-Our result provides a better bound when $g>2^{k^2}$ \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
+$O(g^{k+2})$ using \cref{GenusProduct} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$ 
+% \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}. 
+% \note{PM}{Ok}
 In the case of 1-planar graphs we can improve further.
@@
 	$  \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}.$
-\note{DW}{I have changed $4^{\binom{k+5}{4}}$ to $4^{\binom{k+4}{3}}$, okay?}
+% \note{DW}{I have changed $4^{\binom{k+5}{4}}$ to $4^{\binom{k+4}{3}}$, okay?}
+% \note{PM}{Ok}
 \end{compactitem}
@@
 $\displaystyle \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{p+\binom{k+4}{3}-1}{\binom{k+4}{3}-1}$.
-\note{DW}{I have changed $\binom{k+5}{4}$ to $\binom{k+4}{3}$, okay?}
+% \note{DW}{I have changed $\binom{k+5}{4}$ to $\binom{k+4}{3}$, okay?}
+% \note{PM}{Ok}
  \end{compactitem}
@@
 \note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph. Is the following lemma known (at least in the $g=0$ case)?}
-
+\note{PM}{I have no idea if this is known.  If so, it's probably not stated in terms of frames.}
 \begin{lem}
@@
 
-\note{DW}{\cref{MapPartition} has been updated.}
-
-\note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the method in \citep{DHHW}.
+% \note{DW}{\cref{MapPartition} has been updated.}
+
+% \note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the method in \citep{DHHW}.
 I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
@@
 
-\note{DW}{\cref{StringPartition} has been updated.}
+% \note{DW}{\cref{StringPartition} has been updated.}
 
@@
 
-\note{DW}{\cref{PowerGenus} has been updated}
+% \note{DW}{\cref{PowerGenus} has been updated}
 

### 2022-03-15 DavidsIncorrectWriteupFramedGraphs.tex
added a couple of comments

--- 2022-03-14 k-planar.tex
@@
-\documentclass{patmorin}
-\listfiles
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amsthm}
-\usepackage{graphicx}
-\usepackage{enumerate}
-\usepackage{amsfonts}
-\usepackage{amsthm,mathtools}
-\usepackage{paralist}
-\usepackage{stmaryrd}
-\usepackage{thm-restate}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-
-\usepackage{todonotes}
-
-\newcommand{\arXiv}[1]{arXiv:\,\href{https://arxiv.org/abs/#1}{#1}}
-%\newcommand{\doi}[1]{\href{https://doi.org/#1}{\tt https://doi.org/#1}}
-
-\usepackage[longnamesfirst,numbers,sort&compress]{natbib}
-
-\usepackage[noabbrev,capitalise]{cleveref}
-\crefname{thm}{Theorem}{Theorems}
-\crefname{lem}{Lemma}{Lemmas}
-\crefname{cor}{Corollary}{Corollaries}
-\crefname{prop}{Proposition}{Propositions}
-\crefname{conj}{Conjecture}{Conjectures}
-\crefname{open}{Open Problem}{Open Problems}
-\crefname{claim}{Claim}{Claims}
-\crefname{enumi}{Item}{Items}
-\crefname{obs}{Observation}{Observations}
-\crefname{clm}{Claim}{Claims}
-\crefname{openproblem}{Open Problem}{Open Problems}
-\crefformat{equation}{(#2#1#3)}
-\Crefformat{equation}{Equation #2(#1)#3}
-
-\theoremstyle{plain}
-\newtheorem{thm}{Theorem}
-\newtheorem{lem}[thm]{Lemma}
-\newtheorem{cor}[thm]{Corollary}
-\newtheorem{obs}[thm]{Observation}
-\newtheorem{prop}[thm]{Proposition}
-\newtheorem{clm}{Claim}
-\theoremstyle{definition}
-\newtheorem{open}[thm]{Open Problem}
-\newtheorem{conj}[thm]{Conjecture}
-\newtheorem{ques}[thm]{Question}
-
-\definecolor{brew2}{rgb}{1.0, 1.0, 0.701960784314}
-\definecolor{brew8}{rgb}{0.988235294118, 0.803921568627, 0.898039215686}
-
-% \allowdisplaybreaks
-%\sloppy
-%%%%%%%%%%
- \makeatletter
- \def\NAT@spacechar{~}
- \makeatother
-
-\newcommand{\snote}[1]{{\color{red}#1}}
-
-\newcommand{\defin}[1]{\textcolor{Maroon}{\emph{#1}}}
-
-\newcommand{\note}[2]{\noindent{\color{red}[#1:~#2]}}
-\newcommand{\notex}[2]{}
-\newcommand{\referee}[2]{\noindent\textcolor{blue}{\framebox{\begin{minipage}{\textwidth} Ref \#{#1}: #2\end{minipage}}}}
-
-\DeclareMathOperator{\dist}{dist}
-\DeclareMathOperator{\depth}{depth}
-\DeclareMathOperator{\ff}{f}
-\DeclareMathOperator{\tw}{tw}
-\DeclareMathOperator{\pw}{pw}
-\DeclareMathOperator{\qn}{qn}
-
-\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
-\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
-\DeclarePairedDelimiter{\blah}{\langle}{\rangle}
-
-\newcommand{\PRlabel}[1]{\label{PR:#1}}
-\newcommand{\PRref}[1]{(PR\ref{PR:#1})}
-
-\newcommand{\jlabel}[1]{\label{j:#1}}
-\newcommand{\jref}[1]{(J\ref{j:#1})}
-
-\newcommand{\tlabel}[1]{\label{t:#1}}
-\newcommand{\tref}[1]{(T\ref{t:#1})}
-\newcommand{\ylabel}[1]{\label{y:#1}}
-\newcommand{\yref}[1]{(Y\ref{y:#1})}
-
-\newcommand{\TT}{\mathcal{T}}
-\newcommand{\PP}{\mathcal{P}}
-\renewcommand{\SS}{\mathcal{S}}
-
-\renewcommand{\ge}{\geqslant}
-\renewcommand{\le}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\renewcommand{\leq}{\leqslant}
-
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\Z}{\mathbb{Z}}
-
-\title{\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}}
-
-\author{%
-Vida Dujmovi\'c%
-\thanks{School of Computer Science and Electrical Engineering, University of Ottawa, Ottawa, Canada (\texttt{vida.dujmovic@uottawa.ca}).
-Research supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\,
-Pat Morin%
-\thanks{School of Computer Science, Carleton University, Ottawa, Canada (\texttt{morin@scs.carleton.ca}). Research  supported by NSERC and the Ontario Ministry of Research and Innovation.},\,\, and
-David R. Wood%
-\thanks{School of Mathematics, Monash University, Melbourne, Australia (\texttt{david.wood@monash.edu}). Research supported by the Australian Research Council.}
-}
-
-\begin{document}
-\begin{titlepage}
-\maketitle
-
-\begin{abstract}
-Dujmovi\'c~et~al.~[\emph{J.~ACM}~'20] recently proved that every planar graph is isomorphic to a subgraph of the strong product of a graph of bounded treewidth and a path.
-Analogous results were obtained for graphs of bounded Euler genus or apex-minor-free graphs.
-These tools have been used to solve longstanding problems on queue layouts, non-repetitive colouring, $p$-centered colouring, and adjacency labelling.
-This paper proves analogous product structure theorems for various non-minor-closed classes.
-One noteable example is $k$-planar graphs (those with a drawing in the plane in which each edge is involved in at most $k$ crossings).
-We prove that every $k$-planar graph is isomorphic to a subgraph of the strong product of a graph of treewidth $O(k^5)$ and a path.
-This is the first result of this type for a non-minor-closed class of graphs.
-It implies, amongst other results, that $k$-planar graphs have non-repetitive chromatic number upper-bounded by a function of $k$.
-All these results generalise for drawings of graphs on arbitrary surfaces.
-In fact, we work in a much more general setting based on so-called shortcut systems that are of independent interest.
-This leads to analogous results for certain types of map graphs, string graphs, graph powers, and nearest neighbour graphs.
-\end{abstract}
-\end{titlepage}
-\pagenumbering{roman}
-\tableofcontents
-\newpage
-
-\pagenumbering{arabic}
-% \setcounter{\page}{1}
-\section{Introduction}
-\label{Introduction}
-
-% \note{DW}{I suggest we delete ``We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$. '' This is absolutely standard.}
-
-The starting point for this work is the following `product structure theorem' for planar graphs\footnote{In this paper, all graphs are finite and undirected.
-Unless mentioned otherwise, all graphs are also simple.
-For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$), let $G[S]$  denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.
-We use $G-S$ as a shorthand for $G[V(G)\setminus S]$.
-Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (with an improvement by \citet{UWY}).
-A graph $G$ is \defin{contained in} a graph $X$ if $G$ is isomorphic to a subgraph of $X$.
-
-\begin{thm}[\citep{DJMMUW20,UWY}]
-\label{PlanarProduct}
-Every planar graph is contained in:
-\begin{compactenum}[(a)]
-	\item $H\boxtimes P$ for some graph $H$ of treewidth at most $6$ and for some path $P$,
-	\item $H\boxtimes P \boxtimes K_3$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
-\end{compactenum}
-\end{thm}
-
-Here $\boxtimes$ is the strong product,\!\footnote{The \defin{strong product} of graphs $A$ and $B$, denoted by $A\boxtimes B$, is the graph with vertex set $V(A)\times V(B)$, where distinct vertices $(v,x),(w,y)\in V(A)\times V(B)$ are adjacent if
-	$v=w$ and $xy\in E(B)$, or
-	$x=y$ and $vw\in E(A)$, or
-	$vw\in E(A)$ and $xy\in E(B)$.}
-and treewidth\footnote{For a tree $T$, a \defin{$T$-decomposition} of a graph $G$ is a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that
-(i) for every $vw\in E(G)$, there exists some node $x\in V(T)$ with $v,w\in B_x$; and
-(ii) for every $v\in V(G)$, the induced subgraph $T[v] := T[\{x: v\in B_x\}]$ is connected.
-The \defin{width} of $\mathcal{T}$ is $\max\{|B_x|:x\in V(T)\}-1$.
-A \defin{tree-decomposition} is a $T$-decomposition for any tree $T$.
-The \defin{treewidth} $\tw(G)$ of a graph $G$ is the minimum width of a tree-decomposition of $G$.
-Treewidth is the standard measure of how similar a graph is to a tree.
-Indeed, a connected graph has treewidth 1 if and only if it is a tree.
-Treewidth is of fundamental importance in structural and algorithmic graph theory; see \citep{Reed03,HW17,Bodlaender-TCS98} for surveys.} is an invariant that measures how `tree-like' a given graph is; see \cref{ProductExample} for an example.
-Loosely speaking, \cref{PlanarProduct} says that every planar graph is contained in the product of a tree-like graph and a path.
-This enables combinatorial results for graphs of bounded treewidth to be generalised for planar graphs (with different constants).
-
-\begin{figure}[!h]
-\centering
-\includegraphics{ProductExample}
-\caption{Example of a strong product.
-\label{ProductExample}}
-\end{figure}
-
-\noindent\cref{PlanarProduct} has been the key tool in solving the following well-known open problems:
-\begin{compactitem}
-\item \citet{DJMMUW20} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92}).
-\item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
-\item \citet{DEJGMM21} have since used it to find asymptotically optimal adjacency labellings of planar graphs (resolving a problem of \cite{kannan.naor.ea:implicit}).
-\item \citet{EJM} use it to show the existence of a 'universal graph' with $n^{1+o(1)}$ vertices and edges that contains every $n$-vertex planar graph as an induced subgraph (resolving a problem of \citet{babai.chung.ea:on}).
-\end{compactitem}
-In addition to these rather well-known problems, \cref{PlanarProduct} has been used to resolve or make substantial progress on a number of other problems on planar graphs, including $p$-centered colouring \cite{DFMS21}, $\ell$-vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, and odd colourings~\citep{DMO}.
-
-% \item \citet{DFMS21} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs.
-% \item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results starting with the work of \citet{kannan.naor.ea:implicit}).
-% \item \citet{BDJM} use it to show that every $n$-vertex planar graph has local vertex ranking using $O(\log n/\log\log\log n)$ colours.
-% % \item The result of \citet{DEJGMM21} implies that, for every integer
-% $n>0$, there is a `induced-universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
-% subgraph of $U_n$.
-This has recently been extended by \citet{EJM}, who show the existence of a `subgraph-universal' graph with $(1+\epsilon)n$ vertices that contains every $n$-vertex planar graph as a subgraph as well as the existence of an `induced universal' graph with $n^{1+o(1)}$ vertices \emph{and} edges.
-The former result is the first progress on subgraph-universal graphs for planar graphs since the $O(n^{3/2})$-vertex construction of \citet{babai.chung.ea:on}.
-% aking the first progress on this problem since the work of \citet{babai.xx} in 1984.
-% \end{compactitem}
-
-% \note{DW}{Mention other application such as vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, odd colourings~\citep{DMO}?
-We write a lot about adjacency labellings and universal induced subgraphs.
-Do we need so much detail here?
-We also mention this work in \cref{WrappingUp}.}
-
-% \note{PM}{Agreed. I cut this down to a list of famous open problem. We don't need to sell this so hard anymore.}
-
-All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this definition.} These applications motivate finding product structure theorems for other graph classes.
-\citet{DJMMUW20} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
-The \textit{Euler genus} of the non-orientable surface with $c$ cross-caps is $c$.
-The \textit{Euler genus} of a graph $G$ is the minimum integer $g$ such that $G$ embeds in a surface of Euler genus $g$.
-Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.} and for apex-minor-free graphs\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
-A class $\mathcal{G}$ of graphs is \defin{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$.
-A minor-closed class is \defin{proper} if it is not the class of all graphs.
-For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class.
-A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar.
-A 1-apex graph is \defin{apex}.
-A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.}, and \citet{DEMWW22} do so for graphs in any minor-closed class and with bounded maximum degree.
-See \cref{Generalisations} for more precise statements and see \citep{DHJLW21} for a survey on this topic.
-
-The main purpose of this paper is to prove product structure theorems for several non-minor-closed classes of interest. Our results are the first of this type for non-minor-closed classes.
-
-\subsection{$k$-Planar Graphs}
-\label{k_planar_section}
-
-We start with the example of $k$-planar graphs.
-A graph is \defin{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings, where no three edges cross at a single point (see \cref{sec-k-planar} for a formal definition).
-Such graphs provide a natural generalisation of planar graphs, and are important in graph drawing research; see the recent bibliography on 1-planar graphs and the 140 references therein \citep{kobourov.liotta.ea:annotated}.
-It is well-known that the family of $k$-planar graphs is not minor-closed.
-Indeed, 1-planar graphs may contain arbitrarily large complete graph minors~\citep{dujmovic.eppstein.ea:structure}.
-Hence the above results are not applicable for  $k$-planar graphs.
-We extend \cref{PlanarProduct} as follows.
-
-\begin{thm}
-\label{kPlanarProduct}
-Every $k$-planar graph is contained in $H\boxtimes P\boxtimes K_{18k^2+48k+30}$, for some graph $H$ of treewidth $\binom{k+4}{3}-1$ and for some path $P$.
-\end{thm}
-
-\note{DW}{Do we want to mention ``every $1$-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ of treewidth $3$ and for some path $P$'' here?}
-
-\note{PM}{Yes, I think so.  Two theorems, like this, or one?}
-
-For the important special case of $1$-planar graphs, we are able to obtain better constants and an additional property (planarity) on $H$.
-
-\begin{thm}\label{1_planar_product}
-  Every $1$-planar graph is contained in $H\boxtimes P\boxtimes K_{7}$, for some planar graph $H$ of treewidth $3$ and for some path $P$.
-\end{thm}
-
-This theorem has applications in diverse areas, which we explore in \cref{Applications}.
-
-% This theorem has applications in diverse areas, including queue layouts  \citep{DJMMUW20}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, $p$-centered colouring  \citep{DFMS21}, and adjacency labelling \citep{DEJGMM21}, which we explore in \cref{Applications}.
-For example, we prove that $k$-planar graphs have bounded non-repetitive chromatic number (for fixed $k$).
-Prior to the recent work of \citet{dujmovic.esperet.ea:planar}, it was even open whether planar graphs have bounded non-repetitive chromatic number.
-%
-% \referee{1}{3. pages 2 – 4: The paragraph following Theorem 2 and Section 1.3
-% (except for the fact that the apex assumption is necessary) is repetition
-% of things discussed earlier on in the introduction.}
-%
-% \note{PM}{Kind of, but I suggest we keep that paragraph and section anyway.
-Justification: They're both short and they describe the exact bounds for bounded genus graphs, which you don't get from the one-paragraph description in Section~1.0.}
-%
-% \note{DW}{I suggest we write, ``This theorem has applications in diverse areas, which we explore in \cref{Applications}....''}
-
-\subsection{Shortcut Systems}
-
-Although $k$-planar graphs are the most high-profile target for a generalization of \cref{PlanarProduct}, we actually prove a substantially stronger result than \cref{kPlanarProduct} using the following definition.
-A non-empty set $\SS$ of non-trivial paths in a graph $G$ is a \defin{$(k,d)$-shortcut system} (for $G$) if:
-
-\begin{compactitem}
-\item every path in $\SS$ has length at most $k$,\footnote{A path of length $k$ consists of $k$ edges and $k+1$ vertices.
-A path is \defin{trivial} if it has length 0 and \defin{non-trivial} otherwise.} and
-\item for every $v\in V(G)$, the number of paths in $\SS$ that use $v$ as an internal vertex is at most $d$.
-\end{compactitem}
-Each path $P\in\SS$ is called a \defin{shortcut}; if $P$ has endpoints $v$ and $w$ then it is a \defin{$vw$-shortcut}.
-Given a graph $G$ and a $(k,d)$-shortcut system $\SS$ for $G$, let $G^{\SS}$ denote the supergraph of $G$ obtained by adding the edge $vw$ for each $vw$-shortcut in $\SS$.
-
-This definition is related to $k$-planarity because of the following observation:
-
-\begin{obs}
-\label{AddDummy}
-Every $k$-planar graph is contained in $G^\SS$ for some planar graph $G$ and some $(k+1,2)$-shortcut system $\SS$ for $G$.
-\end{obs}
-
-The proof of \cref{AddDummy} is trivial: Given a $k$-plane embedding of a graph $G'$, create a planar graph $G$ by adding a dummy vertex at each crossing point.
-For each edge $vw\in E(G')$ there is a path $P$ in $G$ between $v$ and $w$ of length at most $k+1$ in which every internal vertex is a dummy vertex.
-Let $\SS$ be the set of such paths $P$.
-For each vertex $v$ of $G$, at most two paths in $\SS$ use $v$ as an internal vertex (since no original vertex of $G'$ is an internal vertex of a path in $\SS$).
-Thus $\SS$ is a $(k+1,2)$-shortcut system for $G$, such that $G'\subseteq G^\SS$.
-This idea can be pushed further to obtain a rough characterisation of $k$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $k$-planar (see \cref{Characterisation}).
-
-The following theorem is the main contribution of the paper.
-It says that if a graph class $\mathcal{G}$ has a product structure theorem, then the class of graphs obtained by applying a shortcut system to graphs in $\mathcal{G}$ also has a product structure theorem.
-
-\begin{thm}
-\label{ShortcutProduct}
-Let $G$ be a graph contained in $H\boxtimes P \boxtimes K_\ell$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
-Let $\SS$ be a $(k,d)$-shortcut system for $G$.
-Then $G^\SS$ is contained in $J\boxtimes P\boxtimes K_{d\ell(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
-\end{thm}
-
-% \referee{2}{Page 3. $P$ is used twice in the statement of Theorem 3.  It is probably
-% safer (and necessary?) to use a different variable for the second
-% occurrence.}
-
-% \note{DW}{Use $\SS$ for shortcut system, okay?}
-% \note{PM}{Ok.  I've implemented this change.  I was careful, but will still reread.}
-
-Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$.
-Some further observations presented in \cref{sec-k-planar} lead to the improved result.
-
-\cref{ShortcutProduct} is applicable for many graph classes in addition to $k$-planar graphs.
-Here is one example.
-The \defin{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.\footnote{For a graph $G$ and two vertices $v,w\in V(G)$, $\dist_G(v,w)$  denotes the length of a shortest path, in $G$, with endpoints $v$ and $w$.
-We define $\dist_G(v,w):=\infty$ if $v$ and $w$ are in different connected components of $G$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\SS$ for some $(k,2k\Delta^{k})$-shortcut system $\SS$; see \cref{PowerShortcut}.
-Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} then imply:
-
-\begin{thm}
-\label{kPowerBasic}
-For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is contained in $H\boxtimes P\boxtimes K_{6k^2(k^2+3)\Delta^{k}}$ for some graph $H$ of treewidth at most $\binom{k+3}{3}-1$ and some path $P$.
-\end{thm}
-
-\cref{Examples} presents further examples of graph classes that can be constructed using shortcut systems, including certain types of map graphs, string graphs, and $k$-nearest neighbour graphs.
-\cref{ShortcutProduct} implies product structure theorems for each of these classes.
-All of the above-mentioned applications also hold for these examples.
-
-\subsection{Generalisations}
-\label{Generalisations}
-
-As mentioned above, product structure theorems have been established for several minor-closed classes in addition to planar graphs.
-The first generalises \cref{PlanarProduct} for graphs of bounded Euler genus.
-
-\begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
-\label{GenusProduct}
-Every graph of Euler genus $g$ is contained in:
-\begin{compactenum}[(a)]
-\item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
-\item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
-\end{compactenum}
-\end{thm}
-
-\citet{DJMMUW20} generalised \cref{GenusProduct} for apex-minor-free graphs as follows.
-
-\begin{thm}[\citep{DJMMUW20}]
-\label{ApexMinorFree}
-For every apex graph $X$, there exists $c\in\mathbb{N}$ such that every $X$-minor-free graph is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and some path $P$.
-\end{thm}
-
-The assumption that $X$ is apex is needed in \cref{ApexMinorFree}, since if the class of $X$-minor-free graphs has a product structure theorem analogous to \cref{PlanarProduct}, then $X$ is apex \citep{DJMMUW20}.
-On the other hand, \citet{DEMWW22} proved a product structure theorem for bounded degree graphs in any minor-closed class.
-
-\begin{thm}[\citep{DEMWW22}]
-	\label{MinorFreeDegree}
-	For every graph $X$ there exists $c\in\mathbb{N}$ such that for every $\Delta\in\mathbb{N}$, every $X$-minor-free graph $G$ with maximum degree at most $\Delta$ is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H) \leq c\Delta$ and for some path $P$.
-\end{thm}
-
-%The next three results are immediate corollaries of \cref{GenusProduct,ApexMinorFree,ShortcutProduct,MinorFreeDegree}.
-%
-%\begin{thm}
-%Let $\SS$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$.
-Then $G^\SS$ is contained in $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
-%\end{thm}
-%
-%\begin{thm}
-%For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\SS$ for $G$, $G^\SS\subseteq H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
-%\end{thm}
-%
-%\begin{thm}
-%For every graph $X$ and for all integers $k,\Delta\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ with maximum degree $\Delta$ and for every $(k,\Delta)$-shortcut system $\SS$ for $G$, $G^\SS \subseteq H \boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
-%\end{thm}
-%
-%\note{DW}{delete the above three theorems, they are never used}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Layered Partitions}
-\label{LayeredPartitions}
-
-While strong products enable concise statements of the theorems in \cref{Introduction}, to prove such results it is helpful to work with layerings and partitions, which we now introduce.
-
-A \defin{layering} of a graph $G$ is a sequence $\mathcal{L}=(L_0,L_1,\ldots)$ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
-For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
-To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the vertex it creates in $H$.
-
-For any partition $\PP$ of $V(G)$ and any layering $\mathcal{L}$ of $G$ we define the \defin{layered width} of $\PP$ with respect to $\mathcal{L}$ as $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\PP\}$.
-For any partition $\PP$ of $V(G)$, we define the \defin{layered width} of $\PP$ as the minimum, over all layerings $\mathcal{L}$ of $G$, of the layered width of $\PP$ with respect to $\mathcal{L}$.
-
-\citet{DJMMUW20} introduced the study of partitions with bounded layered width such that the quotient has some additional desirable property, like small treewidth.
-Dujmovi\'c~et~al.\ define a class $\mathcal{G}$ of graphs to \defin{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
-
-These definitions relate to strong products as follows.
-
-\begin{lem}[\citep{DJMMUW20}]
-\label{PartitionProduct}
-For every graph $H$, a graph $G$ has an $H$-partition of layered width at most $\ell$ if and only if $G$ is contained in $H \boxtimes P \boxtimes K_\ell$ for some path $P$.
-\end{lem}
-
-As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a),
-\citet{DJMMUW20} showed that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \cref{UWY}).
-The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.
-At the core of their work is the elegant proof by \citet{PS21} of the following result:
-
-\begin{thm}[\citep{PS21}]
-\label{ps}
-Every planar triangulation $G$ has an $H$-partition $\PP$
-such that $\tw(H)\leq 8$ and $G[P]$ is a shortest path in $G$ for each $P\in\PP$.
-\end{thm}
-
-Indeed, the above-mentioned result of \citet{DJMMUW20} is a slight strengthening of \cref{ps}, where for each $P\in\PP$ no two vertices of $P$ have the same distance to some fixed root vertex $r$.
-
-The following result is the main technical contribution of the paper.
-Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so too does $G^\SS$ for every shortcut system $\SS$ of $G$.
-
-\begin{restatable}{thm}{mmg}
-	\label{ShortcutPartition}
-	Let $G$ be a graph having an $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\SS$ be a $(k,d)$-shortcut system for $G$.
-Then $G^\SS$ has a $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$.
-\end{restatable}
-
-Note that \cref{ShortcutPartition} is equivalent to \cref{ShortcutProduct} by \cref{PartitionProduct}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Shortcut Systems}
-\label{Structure}
-
-The purpose of this section is to prove our main technical result, \cref{ShortcutPartition}.
-This theorem shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, a $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
-
-For convenience, it will be helpful to assume that $\SS$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.
-Since $G^\SS$ is defined to be a supergraph of $G$, this assumption has no effect on $G^{\SS}$ but eliminates special cases in some of our proofs.
-
-Consider a tree $T$ rooted at some node $x_0\in V(T)$.
-Each subtree $T'$ of $T$ is considered to be rooted at the node in $T'$ closest to $x_0$.
-A node $a\in V(T)$ is a \defin{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \defin{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
-Note that each node $x\in V(T)$ is a $T$-ancestor and $T$-descendant of itself.
-We say that a $T$-ancestor $a\in V(T)$ of $x\in V(T)$ is a \defin{strict} $T$-ancestor of $x$ if $a\neq x$.
-The \defin{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.  For each node $x\in V(T)$, define
-\[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
-to be the maximal subtree of $T$ rooted at $x$.
-
-We begin with a standard technique that allows us to work with a normalised tree-decomposition:
-
-\begin{lem}\label{nice-decomposition}
-  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a width-$t$ $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:
-  \begin{compactenum}[(T1)]
-    % \item\tlabel{rooted}\tlabel{first} $T$ is rooted at some node $x_0\in V(T)$ with $B_{x_0}=\emptyset$.
-    % \item\tlabel{node-set} $V(T)= V(H)$.
-    % \item\tlabel{diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
-    \item\tlabel{subtree-root} for each node $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$; and consequently
-    \item\tlabel{ancestor-edge}\tlabel{last} for each edge $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
-  \end{compactenum}
-\end{lem}
-
-\begin{proof}
-  That \tref{subtree-root} implies \tref{ancestor-edge} is a standard observation: If two subtrees intersect, then one contains the root of the other.
-Thus, it suffices to construct a width-$t$ tree-decomposition that satisfies \tref{subtree-root}.
-
-  Begin with any width-$t$ tree-decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some tree $T_0$.
-Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$.
-(The purpose of $x_0$ is to ensure that every node $x$ for which $B_x$ is non-empty has a parent.)  Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y\}]$.
-If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.
-Subdivide the edge between $a$ and its parent in $T_0$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
-Now $f(y)=a'$ and $f(x)=a$, so this modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
-
-  Next, remove the node $x_0$ from $T_0$, and consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.
-In this case, $B_{a}\subseteq B_{a'}$ where $a'$ is the parent of $a$ since any $x\in B_a\setminus B_{a'}$ would have $f(x)=a$.
-In this case, contract the edge $aa'$ in $T_0$, eliminating the node $a$.
-Repeating this operation will eventually produce a width-$t$ tree-decomposition of $(B_x:x\in V(T_0))$ where $f$ is a bijection between $V(H)$ and $V(T_0)$.
-Renaming each node $a\in V(T_0)$ as $f^{-1}(a)$ gives a tree-decomposition $(B_x:x\in V(T))$ with $V(T)=V(H)$.
-By the definition of $f$, the tree-decomposition $(B_x:x\in V(T))$ satisfies \tref{subtree-root}.
-  % To see that $(B_x:x\in V(T))$ satisifies \tref{ancestor-edge}, observe that,
-  % if $xy\in E(H)$, then at least one of $x$ or $y$ is contained in $B_z$ for every node $z$ on the path from $x$ to $y$ in $T$.
-If neither $x$ nor $y$ is an ancestor of the other, then some node $z$ on this path has $T$-depth less than that of $x$ and $y$.
-If $x\in B_z$ this contradicts the fact that $x$ is the root of $T[x]$.
-If $y\in B_z$ this contradicts the fact that $y$ is the root of $T[y]$.
-\end{proof}
-
-% \referee{1}{6. page 6, last paragraph of Lemma 2: This isn’t quite right since you have to be careful with $x_0$.}
-% \note{PM}{I addressed this by removing $x_0$ from $T_0$ at the beginning of the second paragraph.}
-
-%\subsection{Generalized Tripod Partitions}
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics{figs/tripoddo}
-  \end{center}
-  \caption{The sets $Y_x$, $F_x$, and $V_x$ associated with $x\in V(T)$
-  and the ancestors $a_1,\ldots,a_{t'}$ of $X$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.}
-  \label{fig:generalized-tripod}
-\end{figure}
-
-The following lemma shows how to interpret an $H$-partition of $G$ and a tree-decomposition of $H$ as a `hierarchical' decomposition of $G$; refer to \cref{fig:generalized-tripod}.
-
-\begin{lem}\label{generalized-tripod}
-  Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
-For each $x\in V(T)$, let $V_x := \bigcup_{y\in V(T_x)} Y_y$, $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x:=V_x\cup F_x$.
-Then,
-  \begin{compactenum}[(Y1)]
-    % \item\ylabel{thickness} $\mathcal{Y}=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
-    \item\ylabel{separator} For each $x\in V(T)$, there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$.
-    \item\ylabel{ancestor-edge} For each $x\in V(T)$, there is a set $\{a_1,\ldots,a_{t'}\}$ of $t'\le t$ strict $T$-ancestors of $x$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
-  \end{compactenum}
-\end{lem}
-
-\referee{1}{1.
-Lemma 3: This lemma is really about ``normalised'' tree-decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$.
-Call this property (T3) and remove (Y1) through (Y5), which should now be obvious.}
-
-Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:
-
-\begin{compactenum}[(Y1)]\setcounter{enumi}{2}
-  \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$.
-  \item\ylabel{containment-i} $V_x\subseteq V_a$ for every $T$-ancestor $a$ of $x$.
-  \item\ylabel{containment-ii}$N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
-\end{compactenum}
-
-Property~\yref{y-subsets} follows from the fact that $V_x$ is the union of several sets, one of which is $Y_x$.
-Property~\yref{containment-i} follows from the definition of $V_x$ and the fact that $V(T_x)\subseteq V(T_a)$.
-To show Property~\yref{containment-ii} first note that, by \yref{containment-i} it suffices to consider vertices $w\in F_x=N_x\setminus V_x$.
-By definition, every vertex $w\in F_x$ is adjacent, in $G$, to a vertex $v\in V_x$.
-By \yref{containment-i}, $v\in V_a$, so $w$ is either in $V_a$ or $w$ satisfies the condition $vw\in E(G)$, $v\in V_a$, and $w\not\in V_a$, so $w\in F_a$.
-In either case $w\in N_a=V_a\cup F_a$.
-Note that none of \yref{y-subsets}--\yref{containment-ii} depends on \yref{ancestor-edge} (which is important, since \yref{containment-i} is used to establish \yref{ancestor-edge} in the following proof).
-
-\begin{proof}[Proof of \cref{generalized-tripod}]
-  % Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
-  Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
-
-  To establish Property~\yref{ancestor-edge}, consider some vertex $w\in F_x$.
-Since $w\in F_x$, there exists an edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.
-Since $v\in V_x$, $v\in Y_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
-Since $\mathcal{Y}$ is a partition, $w\in Y_{a}$ for some $a\not\in V(T_x)$.
-Since $vw\in E(G)$, we have $x'a\in E(H)$.
-By \tref{ancestor-edge}, one of $a$ or $x'$ is a $T$-ancestor of the other.
-Since $w\in Y_a\subseteq V_a$ and $w\not\in V_x\supseteq V_{x'}$, \yref{containment-i} rules out the possibility that $x'$ is a $T$-ancestor of $a$.
-Therefore, $a$ is a $T$-ancestor of $x$ which is a $T$-ancestor of $x'$.
-Let $z_0,\ldots,z_r$ be the path in $T$ from $z_0:=x'$ to $z_r:=a$.
-For each $i\in\{0,\ldots,r\}$, at least one of $a$ or $x'$ is in $B_{z_i}$, since $x'a\in E(H)$.
-However, by \tref{subtree-root} $x'$ is not contained in $B_{x_i}$ for any $i\in\{1,\ldots,r\}$.
-Therefore $a\in B_{x_i}$ for each $i\in\{0,\ldots,r\}$.
-In particular, $a$ is contained in $B_x$.
-  Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $B_x$ contains $x$.
-\end{proof}
-
-\referee{2}{Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
-(twice).  Moreover, the way the proof is written, it seems to suggest
-that we only get that $a \in B_{z_i}$ for each $i \in \{1, \dots, r\}$.
-The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
-true though.  In a normalized tree-decomposition, if $xy$ is an edge and
-$x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
-and $T_y$ intersect).}
-
-We are now ready to prove our main result, which we restate here for convenience:
-
-\mmg*
-
-\referee{1}{2. Proof of Theorem 9: There is too much notation throughout. As
-	one example, it is hard to remember which of $T_x$, $Y_x$, $V_x$, $F_x$, $N_x$,
-	$S_x$, $X_v$, $B_x$, $C_x$ is which. If $\phi(S)$ is written for the subset of $V(G)$
-	corresponding to a set $S \subseteq V(H)$ and appropriate shorthand and
-	standard notation for neighbourhoods are used, then $Y_x$, $V_x$, $F_x$, and
-	$N_x$ can be replaced by $\phi(x)$, $\phi(T_x)$, $N(\phi(T_x))$, and $V(G) \setminus N[\phi(T_x)]$.
-	The set $X_v$ is only used in the statement of Claim~1, so it does not
-	need a name; say what these vertices are explicitly. It may be more
-	indicative to write $\phi'(x)$ for $S_x$ and $B'_x$ for $C_x$.}
-
-\begin{proof}
-Apply \cref{generalized-tripod} to $G$ and let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $Y_x$, $V_x$, $F_x$, and $N_x$ be defined as in \cref{generalized-tripod}, where the partition $\mathcal{Y}$ has width $\ell$ with respect to the layering $\mathcal{L}$.
-
-For a node $x\in V(T)$, we say that a shortcut $P\in\SS$ \defin{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
-We say that a vertex $v\in V(G)$ \defin{participates} in $x$ if $v\in Y_x$, or $\SS$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
-We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
-
-\begin{clm}\label{x-v-ancestor}
-  For any $v\in V(G)$ there exists a (unique) node $a(v)\in X_v$ such that
-  $a(v)$ is a $T$-ancestor of every node in $X_v$.
-\end{clm}
-
-\begin{proof}
-  Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\SS, v\in \{v_0,\ldots,v_r\}\}$.
-Then $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
-
-  We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
-If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\SS$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
-In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$, $(v_0,\ldots,v_r)\in\SS$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\SS$ that crosses $x$.
-
-  Let $X_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.  The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.
-  %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
-  Choose $a(v)\in X_H$ to be the member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.
-Transitivity of the $T$-ancestor relationship, \tref{ancestor-edge}, and connectivity of $H[X_H]$ implies that such an $a(v)$ exists and is a $T$-ancestor of every node $x\in X_H$, as required.
-\end{proof}
-
-% \note{DW}{Use $\SS$ for the given shortcut system and $\PP$ for the partition?}
-% \note{PM}{Yes, I've done this.  Again, I was carefuly, but need to reread.}
-
-For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
-Observe that $\PP:=(S_x : x\in V(T))$ is a partition of $V(G)$.\footnote{The sets in $\PP$ are disjoint because $a:V(G)\to V(T)$ is a function.
-The sets in $\PP$ cover $V(G)$ since, for each $v\in V(G)$, $v\in S_{a(v)}$.} Let $J:=G^\SS/\PP$ denote the resulting quotient graph.
-We consider $V(J)\subseteq V(T)$, where each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\SS}$.
-(Any node $x\in V(T)$ with $S_x=\emptyset$ does not contribute a vertex to $J$.)
-
-% \referee{2}{Page 8.  Can you add a brief explanation why $\PP$ is a
-% 	partition of $V(G)$?  It is clear that the sets in $\PP$ are
-% 	disjoint, but why do they cover $V(G)$?}
-% \note{PM}{I added an explanation in a footnote since this is obvious.  In particular, for any function $f$, $\{f^{-1}(x):x\in\operatorname{range}(f)\}$ is a partition of $\operatorname{domain}(f)$.}
-
-From this point onward, the plan is to show that:
-\begin{compactenum}[(i)]
-  \item $\PP$ has small layered width with respect to the layering $\mathcal{L}$ of $G$ and that
-  \item $J$ has small treewidth.
-\end{compactenum}
-Once we have established (i) and (ii), the result follows easily since a layering of $G^\SS$ is easily obtained from $\mathcal{L}$ by `compressing' groups of $k$ consecutive layers.
-% \referee{1}{7. page 8, item (i): ``$P$ has small layered width with respect to the layering $L$'' $\longrightarrow$ ``$P$ has small layered width with respect to the layering $L$ of $G$''.}
-% \referee{1}{Perhaps mention something about how it is easy to find the new layering.}
-
-First we need to understand the relationship between sets $S_x\in\PP$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
-
-\referee{1}{8. general comment: Claims 1, 2, and 4 are especially belaboured by
-notation as they are quite straightforward.}
-
-\begin{clm}
-	\label{s-subset}
-  For every $x\in V(T)$, $S_x\subseteq V_x$.
-\end{clm}
-
-\begin{proof}
-  For the sake of contradiction, assume otherwise, so there exists some $v\in S_x\setminus V_x$.
-By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.
-Therefore, $\SS$ contains a path $P$, with $v\in V(P)$, that crosses $x$.
-The path $P$ contains a proper subpath $v_0,v_1,\ldots,v_{r}$ such that $v=v_0$ and $v_r\in Y_x$.
-Since $v\not\in V_x$ and $v_r\in Y_x\subseteq V_x$, \yref{separator}, implies that $v_i\in F_x$ for some $i\in\{0,\ldots,r-1\}$.
-Now \yref{ancestor-edge} implies $v_i\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
-Therefore, either $v\in Y_a$ or $P$ crosses $a$.
-But this implies that $a(v)$ is a $T$-ancestor of $a$, which is a strict $T$-ancestor of $x$, contradicting the assumption that $v\in S_x$.
-\end{proof}
-
-Next we complete Step~(i) and show that $\PP$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
-
-\begin{clm}
-	\label{general-width}
-  For each $i\in\{1,\ldots,h\}$ and each $x\in V(J)$, $|S_x\cap L_i|\le d\ell(k^2+3)$.
-\end{clm}
-
-\begin{proof}
-  Recall that $S_x$ is defined by vertices that participate in $x$, and these are vertices that are either in $Y_x$ or in shortcuts that cross $Y_x$.
-We say that a vertex $w\in Y_x$ \defin{contributes} a vertex $v\in S_x$ if $v=w$ or if some path in $\SS$ that contains $v$ has $w$ as an internal vertex.
-  We upper bound the number of vertices in $S_x\cap L_i$ by upper-bounding the number of vertices contributed to $S_x\cap L_i$ by each $w\in Y_x$.
-
-  Refer to \cref{contribute}.  If $w\in Y_x\cap L_i$ and no path in $\SS$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
-  Otherwise, consider some path $P\in\SS$ that contains $w$ as an internal vertex.
-If $w\in L_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap L_i$.
-If $w\in L_{i-1}\cup L_{i+1}$, then $P$ contributes at most $k$ vertices to $S_x\cap L_i$.
-If $w\in L_{i-j}\cup L_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap L_i$.
-%
-% \referee{1}{9. page 8, line 10: ``We say that a vertex $w \in Y_x$ contributes a vertex
-%   $v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.}
-%
-% \note{DW}{What is wrong here?}
-
-  \begin{figure}[htbp]
-    \begin{center}
-      \includegraphics{figs/contribute}
-    \end{center}
-    \caption{A path $P$ containing an internal vertex $w\in Y_x\cap L_{i-j}$.}
-    \label{contribute}
-  \end{figure}
-
-  For any $j$, the number of vertices $w\in L_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\SS$. Therefore,
-  \[  |S_x\cap L_i|\le d\ell  \, \Big(k+1 + 2k + \sum_{j=2}^k 2(k-j)\Big) %= d(2k+1) + \sum_{i=1}^{k-2} i
-      = d\ell(k^2 +3) \enspace . \qedhere
-  \]
-\end{proof}
-
-We now proceed with Step~(ii), showing that $J$ has small treewidth.
-To accomplish this, we construct a small width tree-decomposition $\mathcal{C}:=(C_x:x\in V(T))$ of $J$ using the same tree $T$ used in $\mathcal{T}$.
-The following claim will be useful in showing that the resulting decomposition has small width.
-
-\begin{clm}\label{i-ancestor}
-  For each edge $xy\in E(J)$, one of $x$ or $y$ is a $T$-ancestor of the other.
-\end{clm}
-
-\begin{proof}
-  Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.
-Since $xy\in E(J)$, $G^\SS$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
-Since $vw\in E(G^{\SS})$,  $\SS$ contains a $vw$-shortcut $P$.\footnote{Recall that we have made the assumption that $\SS$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  By \cref{s-subset}, $v\in V_x$ and $w\in V_y$.
-By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
-By \yref{ancestor-edge}, $N_x$ and $V_y$ are also disjoint.
-By \yref{separator} $P$ contains an internal vertex $v'\in F_x$.
-By \yref{ancestor-edge}, $v'\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
-But this implies that $a(v)=a'$ so $v\in S_{a'}$ for some $T$-ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
-\end{proof}
-
-\begin{clm}
-\label{general-bag-size}
-The graph $J$ has a tree-decomposition in which every bag has size at most $\binom{k+t}{t}$.
-\end{clm}
-
-\referee{1}{10. page 10, Claim 5: Indices could mostly be removed by continuing
-to talk about $T$-ancestors instead. Move the definition of $H+$ and its
-directed version up, along with the explanation of what will be proven,
-so that it isn’t necessary to define the $s_i$.}
-
-\begin{proof}
-  For the tree-decomposition $(C_x:x\in V(T))$ of $J$ we use the same tree $T$ used in the tree-decomposition $(B_x:x\in V(T))$ of $H$.
-For each node $x$ of $T$, we define $C_x$ as follows: $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
-  \cref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, $a,x'\in C_{x'}$.
-The connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ follows from the fact that, for every node $x'\in T[a]$, every node $x$ on the path in $T$ from $x'$ to  $a$ is also a node of $T[a]$.
-Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $J$.
-It remains to bound the size of each bag $C_x$.
-
-  Consider an arbitrary node $x\in V(T)$ where $x_0,\ldots,x_r$ is the path from the root $x_0$ of $T$ to $x_r:=x$.
-To avoid triple-subscripts in what follows, we abuse notation slightly by using $V_i$, $F_i$, and $N_i$,  as shorthands for $V_{x_i}$, $F_{x_i}$ and $N_{x_i}$, respectively.
-
-  If $x_\delta\in C_x$, it is because $x_\delta x'\in E(J)$ for some $T$-descendant $x'$ of $x$.
-This implies $G^{\SS}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
-This implies that $\SS$ contains a $vw$-shortcut $P_{vw}$.
-Let $v'$ be the second-last vertex of $P_{vw}$ (so $v'w\in E(G)$).
-
-  Since $w\in S_{\delta}$, $a(w)=x_\delta$, so $w$ participates in $x_\delta$, so at least one of the following is true:
-  \begin{compactenum}
-    \item There exists $w'\in V(G)$ such that $\SS$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
-    \item $w\in Y_\delta$; in this case, define $w':=w$ and let $P_{ww'}$ be the 1-vertex path $w$.
-  \end{compactenum}
-  Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{\delta}$.
-
-  Let $w_0,w_1,\ldots,w_p$ be the path that begins $w_0:=v'$ and then follows the subpath of $P_{ww'}$ that begins at $w_1:=w$ and ends at $w_p:=w''$.
-For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{0,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.
-Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root, $x_0$, of $T$ is non-increasing.
-
-  We claim that $a_0=x_r$.
-Since $v\in S_{x'}$, $a(v)=x'$, so $V(P_{vw})\subseteq V_{x'}$, otherwise $v$ participates in $a$ for some node $a$ not in the subtree $T_{x'}$ rooted at $x'$, but this contradicts \cref{i-ancestor} since $a(v)=x'$ is a $T$-ancestor of every node in which $v$ participates.
-Since $v'=w_0\in V(P_{vw})$, $\{w_0\}\subseteq V_{x'}\subseteq V_{r}$, so $s_0=r$ and $a_0=x_r$.
-
-  We claim that $a_p=x_\delta$---that is, $s_p=\delta$.
-  To see this, first observe that, for each $i\in\{1,\ldots,p\}$, $w_i\in V_{\delta}$ since, otherwise, an internal vertex of $P_{ww'}$ belongs to $F_\delta$, which would imply (by \yref{ancestor-edge}) that $w\in S_{\delta'}$ for some $\delta' < \delta$, contradicting the assumption that $w\in S_\delta$.
-Therefore $s_p\ge\delta$.
-To see that $s_p<\delta+1$,
-  observe that either $w=w''\in Y_\delta$ or $P_{ww'}$ contains an internal vertex $w''$ in $Y_\delta$.  By the definition of $V_x$, $V_{\delta-1}$ does not contain $w''$, so $s_p<\delta+1$.
-
-  Let $H^+$ denote the supergraph of $H$ with vertex set $V(T)$ and in which $xy\in E(H^+)$ if and only there exists some $z\in V(T)$ such that $x,y\in B_z$.
-  % Note that, by \tref{ancestor-edge}, $xy\in E(H^+)$ if and only if $x\in B_y$ or $y\in B_x$.
-  We claim that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \defin{lazy walk} in a graph $H$ is a walk in the pseudograph $H'$ obtained by adding a loop at each vertex of $H$.} in $H^+$.
-Indeed, if $a_i\neq a_{i+1}$ for some $i\in\{0,\ldots,p-1\}$ then this is precisely because $w_i\in V_{a_i}$ but $w_{i+1}\not\in V_{a_i}$.
-By definition, $w_i\in Y_{a_i'}$ for some $T$-descendant $a_i'$ of $a_i$.
-  By \yref{separator}, $w_{i+1}\in F_{a_i}$ so by \yref{ancestor-edge} $w_{i+1}\in Y_{a_i''}$ for some strict $T$-ancestor $a_i''$ of $a_i$.
-Since $w_iw_{i+1}\in E(G)$, $a_i'a_i''\in E(H)$.
-By \tref{subtree-root}, $a_i''\in B_{a_i''}$ and $a_i''\in B_{a_i'}$.
-Since $a_i$ is on the path from $a_i'$ to $a_i''$ in $T$ this implies that $a_i''\in B_{a_i}$.
-Therefore $a_ia_i''\in E(H^+)$ as claimed.
-
-\referee{2}{Page 18.  Say $H'$ is connected since $X$ is connected.}
-
-  Thus, $a_0,\ldots,a_p$ is a lazy walk in $H^+$ of length $p\le k$ where the distance $s_i$ between $a_i$ and the root $x_0$ of $T$ is non-decreasing.
-By removing repeated vertices this gives a path in the directed graph $\overrightarrow{H}^+$ obtained by directing each edge $xy\in E(H^+)$ from its $T$-descendant $x$ towards its $T$-ancestor $y$.
-  Finally, we are in a position to appeal to \cite[Lemma~13]{PS21} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
-
-  % \note{DW}{replace Lemma~24 by Lemma~13, right?}
-  % \note{PM}{Right. Changed.}
-
-  At this point, the proof is complete, but let us summarize.
-For each node $x\in V(T)$, $C_x$ contains only $T$-ancestors of $x$ (including $x$ itself).
-For each ancestor $x_\delta$ of $x$ contained in $C_x$, there is path from $x$ to $x_\delta$ in $\overrightarrow{H}^+$ of length at most $k$.
-The number of ancestors of $x$ that can be reached by paths of length at most $k$ in $\overrightarrow{H}^+$ is at most $\binom{k+t}{t}$.
-Therefore $|C_x|\le \binom{k+t}{t}$, as required.
-\end{proof}
-
-At this point, the proof of \cref{ShortcutPartition} is almost immediate from \cref{general-width,general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\SS}$.
-In particular, $G^{\SS}$ may contain edges $vw$ with $v\in L_i$ and $w\in L_{i+j}$ for any $j\in\{0,\ldots,k\}$.
-To resolve this, we use a new layering $\mathcal{L}':=\langle L_0',\ldots,L_h'\rangle$ in which $L_i'=\bigcup_{j=ki}^{ki+k-1} L_i$.
-This increases the layered width given by \cref{general-width} from $d\ell(k^2+3)$ to $d\ell(k^3+3k)$.
-Therefore $G$ has an $H$-partition of layered width at most $d\ell(k^3+3k)$ in which $H$ has treewidth at most $\binom{k+t}{t}$, completing the proof of \cref{ShortcutPartition}.
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%
-\subsection{Framed Graphs}
-
-For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
-% \note{DW}{I suggest we replace ``and whose crossing edges are in the interior of these faces'' by ``such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$. }
-The graph $G_0$ is called the \defin{frame} of $G_0$.
-This definition (for simple graphs) was introduced by \citet{BDGGMR}.
-Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar graph.
-Thus \cref{ShortcutProduct} is applicable.
-However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.\footnote{A multigraph $G$ is \defin{contained} in a graph $G'$ if the simple graph underlying $G$ is isomorphic to a subgraph of $G'$.}
-
-\begin{thm}\label{d_framed_product_stucture}
-Every $d$-framed multigraph is contained in $H\boxtimes P\boxtimes K_{d+3\floor{d/2}-3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
-\end{thm}
-
-This result is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs.
-\todo{Move this comment to the section on $1$-planar graphs?} For simple 1-planar graphs it is essential that we allow frames with parallel edges.
-
-\cref{d_framed_product_stucture} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
-The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{v_1,\ldots,v_t\}$.
-Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that $\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}$ for each $v\in V(G)$.
-
-\begin{lem}
-	\label{induction} The setup:
-	\begin{compactenum}
-		\item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
-		\item Let $F_0$ be the outer face of $G'$ and let $T$ be a BFS spanning forest of $G'$ rooted at $V(F_0)$.
-		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
-		\item Let $F$ be a cycle in $G'$ that $V(F)$ is partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
-		\begin{compactenum}
-			\item $F[P_i]$ is connected; and
-			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
-		\end{compactenum}
-		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
-	\end{compactenum}
-	Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
-	\begin{compactenum}[(i)]
-		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
-		\item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
-		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
-	\end{compactenum}
-\end{lem}
-
-\begin{proof}
-	This proof is very similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is complicated by several factors.
-In particular, the reader should keep in mind that $G$ and $G'$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
-
-	The proof is by induction on the number of inner faces of $N'$.
-A particular easy case occurs when $V(N')=V(F)$, for which we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
-	\begin{compactenum}[(i)]
-		\item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
-		\item Assumption 4(b) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
-		\item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
-	\end{compactenum}
-
-	In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
-There is one case to consider for each possible value of $k$ (see \cref{boring_figure}):
-	\begin{compactenum}
-		\item If $k=1$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
-
-		\item If $k= 2$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
-
-		\item If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the appropriate vertex of $F_0$.
-(This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
-Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N'$, add edges $d'-2$ edges inside of $F$ to split it into $3$-sided faces.
-This yields a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary.
-By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
-Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.
-% (Note that $p\in\{3,4\}$.)
-	\end{compactenum}
-
-	%  \todo[inline]{%
-	%    DW:Does this work for multigraph near-triangulations? \newline
-	%    PM: Yes, it only requires that the inner faces have three sides and that the outer face is bounded by a simple cycle.
-The proof goes through without change.
-Look at the subgraph $X$ of the dual that contains only the edges dual to red/blue edges.
-In our setting, the outer face has degree $1$ in $X$, which is odd.
-Any bichromatic inner face has degree $0$ or degree $2$ in $X$.
-There has to be another vertex of odd degree somewhere, which can only come from a trichromatic inner face.\newline
-	%    DW: great. This is basically the same as the proof of the Hex Lemma.
-	%  }
-
-	We can now define the vertices in $S$.
-For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to a vertex in $V(F)$.
-Let $\overline{S}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
-Let $S:=V(\overline{S})\setminus V(F)$, let $Y:=(V(Q_1)\cup V(Q_2)\cup V(Q_3))\setminus V(F)$, and let $X:=S\setminus Y$.
-	Observe that, for each $i\in\{1,2,3\}$ and each $j\ge 0$, $|V(Q_i)\cap L_j|\le 1$.
-Therefore $|Y\cap L_j|\le 3$ for each integer $j\ge 0$.
-Furthermore, $|X|\le p-3\le d-3$, since $v_1,v_2,v_3\in Y\cup V(F)$.
-Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ will not violate condition~(ii).
-
-	% We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
-First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $L_i$, each incident on at most two edges of $\overline{Y}$.
-Since $\dist_{G'}(v,w)\le 2$ for each $vw\in E(G)$, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $L_i$.
-These at most six edges have at most 12 endpoints.
-Therefore $|V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i|\le 6\times 2=12$, so $|V(\overline{Y}^+)\cap L_i|\le 12+3=15$.
-
-	\begin{figure}
-		\begin{center}
-			\begin{tabular}{c@{}c@{}c}
-				\includegraphics{figs/zoomba-1} &
-				\includegraphics{figs/zoomba-2} &
-				\includegraphics{figs/zoomba-3} \\
-        \multicolumn{3}{c}{{\color{brew8}\raisebox{-3pt}{\rule{12pt}{12pt}}} $S$\qquad {\color{brew2}\raisebox{-3pt}{\rule{12pt}{12pt}}} $X$} \\
-				$k=1$ & $k=2$ & $k=3$
-			\end{tabular}
-		\end{center}
-		\caption{Three cases when choosing the face $\tau$ of $N'$.}
-		\label{boring_figure}
-	\end{figure}
-
-	Refer to \cref{boring_figure}.
-Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{S}$ and the edges and vertices of $F$.
-By our choice of $\tau$, the graph $M$ is $2$-connected.
-  % Since $N'\subseteq G'$ contains only edges of $N\subseteq G$ that are not crossed, each component of $N-V(M)$ is contained in a single face of $M$.
-	Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
-By our choice of $v_1v_2v_3$, the vertices of each $F_i$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
-
-	In the following, in order to avoid introducing even more notation or constantly adding the modifier ``when it exists'', we will assume that $V(F_i)\cap S\neq\emptyset$ and that there are exactly two distinct $b,c\in\{1,\ldots,k\}$ such that $V(F_i)\cap P_b\neq\emptyset$ and $V(F_i)\cap P_c\neq\emptyset$.
-When this assumption does not hold, exactly the same reasoning can be used, while omitting one or more of $P_{i,1}$, $P_{i,2}$ or $P_{i,3}$.
-
-	Since $M$ is $2$-connected and $\overline{S}$ is connected, $F_i[P_{i,j}]$ is connected, for each $j\in\{1,2,3\}$.
-Let $N_i$ and $N_i'$ be the subgraphs of $N$ and $N'$ contained in the closure of $F_i$.
-Every inner face of $N_i'$ is an inner face of $N'$ and $\tau$ is not an inner face of $N_i'$.
-Therefore $N_i'$ has fewer inner faces than $N'$.
-Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.
-The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of $N_i$.
-	% The partition $\mathcal{P}_i$ contains each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$.
-	We can now define our partition
-	\[
-	\mathcal{P}:=\{Y, P_1,\ldots,P_k\} \cup \bigcup_{i=1}^m \{P\in\mathcal{P}_i: P\cap V(M)=\emptyset\} \enspace .
-	\]
-	It is straightforward to verify that $\mathcal{P}$ is indeed a partition of $V(N)$ and, by definition this is an $H$-partition of $N$ for the graph $H:=N/\mathcal{P}$.
-All that remains is to verify that $\mathcal{P}$ satisfies (i)--(iii).
-
-	By construction $\mathcal{P}$ contains parts $P_1,\ldots,P_k$, so $\mathcal{P}$ satisifes (i).
-By Assumption~4(b) none of the parts $P_1,P_2,P_3$ violate condition~(ii).
-We have already argued that the part $S$ does not violate condition~(ii).
-The inductive hypothesis ensures that each of the remaining parts satisfies condition~(ii).
-Therefore $\mathcal{P}$ satisfies (ii).
-
-	All that remains is to construct a tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$ of $H$ that satisfies (iii).
-Let $x_0$ be the vertex of $H$ obtained from contracting $S$ and, for each $j\in\{1,\ldots,k\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
-The tree $T_0$ has a node $y_0$ with bag $B_{y_0}:=\{x_0,\ldots,x_k\}$.
-The inductive hypothesis ensures that, for each $i\in\{1,\ldots,m\}$,  $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $\mathcal{T}_i:=(B_z:z\in V(T_i))$ that satisfies (iii).
-Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
-Since $a_i$, $b_i$, $c_i$ form a clique in $H$, there exists a bag $B_{z_i}$ in $\mathcal{T}_i$ that contains $\{a_i,b_i,c_i\}$.
-
-	Recall that $P_{i,1}:=V(N_i)\cap S$, $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.
-For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.
-These replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$.
-Now add an edge that joins $y_0$ to $z_i$ so that $T_i$ becomes a subtree of $T_0$ that is adjacent to $y_0$.
-We perform this operation for each $i\in\{1,\ldots,m\}$ to obtain our final tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$.
-
-  It is now straightforward to verify that $\mathcal{T}$ is a tree decomposition of $H':=N'/\mathcal{P}$ in which each bag has size at most $4$.
-To see that $\mathcal{T}$ is also a tree decomposition of $H:=N/\mathcal{P}$, consider any edge $vw\in E(N)\setminus E(N')$ with $v\in P_\alpha$ and $w\in P_\beta$.
-We must verify that some bag of $\mathcal{T}$ contains $\alpha$ and $\beta$.
-If both $v$ and $w$ are in $V(M)$, then $\alpha=x_a$ and $\beta=x_b$ for some $a,b\in\{0,\ldots,3\}$, in which case $B_{y_0}=\{x_0,\ldots,x_k\}$ contains $\alpha$ and $\beta$.
-If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $N\subseteq G$ crosses an edge of $M\subseteq G'$, both $v$ and $w$ are contained in the interior of the same face $F_i$ of $M$ so, by the inductive hypothesis there exists a bag of $\mathcal{T}_i$ that contains $\alpha$ and $\beta$ and this becomes a bag of $\mathcal{T}$ that contains $\alpha$ and $\beta$.
-Finally, if $v\in V(M)$ and $w\not\in V(M)$ then $v\in P_{i,j}$ for some $j\in\{1,2,3\}$ and $w$ is in the interior of some face $F_i$ of $M$.
-By induction, some bag of $\mathcal{T}_i$ contains $\beta$ and $\alpha'$, where $\alpha'$ is the vertex of $H_i$ obtained by contracting $P_{i,j}$.
-However, before $T_i$ is attached to $T_0$ each occurence of $\alpha'$ in $\mathcal{T}_i$ is replaced by $\alpha$, so some bag of $\mathcal{T}$ contains both $\alpha$ and $\beta$.
-Therefore $\mathcal{T}$ is a tree decomposition of $H$.
-
-  Next we verify that, when $k=3$ exactly one bag of $\mathcal{T}$ contains $\{x_1,\ldots,x_3\}$.
-The bag $B_{y_0}$ contains $x_1,\ldots,x_k$.
-When $k=3$, each neighbour of $z_i$ of $y_0$ has a bag that contains at most two of $\{x_1,x_2,x_3\}$.
-Therefore, when $k=3$, $B_y$ is the unique bag in $\mathcal{T}$ that contains $\{x_1,x_2,x_3\}$.
-
-	All that remains is to show that any triple of vertices of $H$ is contained in at most two bags of $\mathcal{T}$.
-By the inductive hypothesis, the only triples we need to be concerned about are the $3$-element subsets of  $\{x_0,x_1,\ldots,x_k\}$.
-The case $k=1$ is trivial since no triples can be formed by $\{x_0,x_1\}$.
-When $k=2$, $B_{y_0}=\{x_0,x_1,x_2\}$.
-In this case, the choice of $\tau$ ensures that there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_1$, and $P_2$ on its boundary.
-(Refer to \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.
-Therefore there are at most two bags, $B_y$ and $B_{z_i}$ of $\mathcal{T}$ that contain $\{x_0,x_1,x_2\}$.
-This establishes (iii) for $k\in\{1,2\}$ so, from this point on we assume that $k=3$.
-
-	We claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.
-Suppose to the contrary that there are two such faces $F_{i_1}$ and $F_{i_2}$.
-Create a graph $M^+$ from $M$ by placing a vertex $v_{i_b}$ inside $F_{i_b}$ and adjacent to each vertex in $V(F_{i_b})$ for each $b\in\{1,2\}$.
-Since $M$ is planar and $F_{i_1}$ and $F_{i_2}$ are distinct faces of $M$, $M^+$ is planar.
-However, contracting each of $S_{x_0},\ldots,S_{x_3}$ in $M^+$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.
-(In this contracted graph, $x_0,x_a,x_b$ are on one side of the bipartition and $v_{i_a}$, $v_{i_b}$, and the node in $\{x_1,x_2,x_3\}\setminus\{x_a,x_b\}$ is on the other side.)
-
-	We have already established that, when $k=3$, the only bag that contains $\{x_1,x_2,x_3\}$ is $B_{y_0}$, so we need only consider triples $\{x_0,x_a,x_b\}$ for distinct $a,b\in\{1,2,3\}$.
-By the preceding claim, there is at most one face $F_i$ with vertices of $S$, $P_a$, and $P_b$ on its boundary.
-By the inductive hypothesis, the only other bag in $\mathcal{T}$ that contains $\{x_0,x_a,x_b\}$ is the bag $B_{z_i}$.
-This completes the proof.
-\end{proof}
-
-\todo[inline]{Alternative to the final three paragraphs in the preceding proof:
-
-To see that $H$ is planar, consider a plane multigraph $G''$ obtained by adding stars in some of the faces of $G'$, as follows:  When creating the part $S$, add a vertex $v_\tau$ to the interior of $\tau$ that is adjacent to each vertex of $\tau$ and place $v_\tau$ in $S$.
-The graph $G''\supseteq G'$ obtained this way is clearly planar.
-Furthermore, for each part $P\in\mathcal{P}$, $G''[P]$ is connected.
-Since planarity is preserved under edge contraction, it follows that $G''/\mathcal{P}$ is planar.
-Now argue that $G/\mathcal{P}=G''/\mathcal{P}$\ldots
-}
-
-Using \cref{induction}, the proof of \cref{d_framed_product_stucture} is now straightforward.
-
-% \note{PM}{Annoying detail:  When we allow framed graphs to have crossings in the outer face, we can't guarantee that $G'$ has any face smaller than $d$.
-This makes it awkward to start the induction using \cref{induction}.
-Possible solutions:
-% \begin{compactenum}
-%   \item Disallow the addition of edges in the outer face of a frame.
-%   \item Modify \cref{induction} so that $T$ is a BFS forest rooted at the set of vertices on the outer face of $G'$. Then we can start the induction with $k=1$.
-% \end{compactenum}
-% I prefer the second option.  Thoughts?
-% }
-
-\begin{proof}[Proof of \cref{d_framed_product_stucture}]
-	Let $G$ be a $d$-framed multigraph with frame $G'$ having outer face $F_0$.	Let $T$ be a BFS forest of $G'$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.
-Let $P_1:= V(F_0)$, let $Y_1$ be any subset of $\min\{|P_1|,3\}$ vertices in $P_1$ and let $X_1:=P_1\setminus Y_1$.
-Now apply \cref{induction} on $G$, $G'$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$.
-    This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree decomposition.
-Since any graph having a $3$-simple tree decomposition is planar \cite{X}, this implies that $H$ is planar.
-
-	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdots \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$.
-That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.
-Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
-The result now follows from \cref{PartitionProduct}.
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%
-\section{Allowing Crossings}
-\label{sec-k-planar}
-
-This section applies our main results for shortcut systems to prove a product structure theorem for graphs drawn with a bounded number of crossings per edge.
-Then we show how to improve the bounds in this case.
-
-\subsection{$k$-Planar Graphs}
-
-We first formally define $k$-planar graphs.
-An \defin{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a curve\footnote{A \defin{curve} in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
-The points $f(0)$ and $f(1)$ are called the \defin{endpoints} of the curve.
-When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.
-The \defin{interior} of $f$ is the point set $\{f(t):0<t<1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
-The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
-A \defin{crossing} in an embedded (multi)graph $G$ is an unordered pair of distinct edges $\{vw,xy\}\subseteq E(G)$ whose interiors have a point in common.
-Without any loss of generality we may assume that the interiors of any two edges of $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.
-This assumption can be enforced by local changes that do not change the number of crossings each edge is involved in.
-If $p$ is a common point in the interior of two crossing edges $vw$ and $xy$ then we say that $vw$ and $xy$ \defin{cross at} $p$.
-
-% (This assumption is already present in the informal definition of $k$-planar graphs used in \cref{k_planar_section}.)
-
-%
-% \note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}\note{PM}{No, under the current definition, $r$ edges crossing at a single point $p$ generate $\binom{r}{2}$ crossings, with each edge taking part in $r-1$ crossings, just like they would if we perturbed the edges so that each pair crosses at a distinct point.
-I guess we could mention this.
-(See the first paragraph in the proof of Theorem~2 below.)} \note{DW}{The downside of this approach is that in the opening section we give an informal definition of $k$-planar, which needs the assumption of having no three edges cross at a point, since we use this assumption in the construction of a $(k+1,2)$-shortcut system for $k$-planar graphs give in the Introduction.
-I think we should not have different definitions in two parts of the paper.} A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
-Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
-
-As mentioned in \cref{Introduction}, \cref{PlanarProduct,ShortcutProduct} imply a product structure theorem for $k$-planar graphs. We get improved bounds as follows.
-
-%\begin{thm}
-%\label{k-planar}
-%Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 48k+30$ in which $\tw(H)\leq \binom{k+4}{3}-1$.
-%\end{thm}
-
-\begin{proof}[Proof of \cref{kPlanarProduct}]
-	Let $G$ be a $k$-plane graph.
-    % We assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$.
-This assumption is justified since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
-	As in the proof of \cref{AddDummy}, let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$.
-In this way, each edge $vw\in E(G)$ corresponds naturally to a path $P_{vw}$ of length at most $k+1$ in $G_0$.
-Let $\SS := \{P_{vw}: vw\in E(G)\}$.
-Observe that $\SS$ is a $(k+1,2)$-shortcut system for $G_0$ and that $G_0^{\SS}\supseteq G$.
-Specifically, $G_0^{\SS}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.
-
-	Since $G_0$ is planar,  \cref{PlanarProduct}(b) and \cref{PartitionProduct} implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
-Applying \cref{ShortcutPartition} to $G_0$ and $\SS$ immediately implies that $G$ (an arbitrary $k$-planar graph) has an $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
-
-	We can reduce the layered width of the $H$-partition of $G$ from $O(k^3)$ to $O(k^2)$ by observing that the dummy vertices in $V(G_0)\setminus V(G)$ do not contribute to the layered width of this partition.
-In this setting, the proof of \cref{general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $L_i\cap Y_x$.
-More precisely, each path $P\in\SS$ containing an internal (dummy) vertex $w\in Y_x\cap (L_{i-j}\cup L_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap L_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap L_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap L_j$ for $j > k+1$.
-	Redoing the calculation at the end of the proof of \cref{general-width} then yields
-	\begin{align*}
-	|S_x\cap Y_i| \le d\ell\left(
-	2
-	+ 4\left\lfloor\tfrac{k+1}{2}\right\rfloor
-	+ 2\left\lceil\tfrac{k+1}{2}\right\rceil
-	\right)
-	 =
-	d\ell\left(
-	2 + 2(k+1) + 2\left\lfloor\tfrac{k+1}{2}\right\rfloor
-	\right)
-	 \le
-	d\ell(3k+5)
-	= 18k+30 \enspace .
-	\end{align*}
-	With this change, the layered width of the partition given by \cref{ShortcutPartition} becomes $(18k+30)(k+1)=18k^2+48k+30$.
-	The result follows from \cref{PartitionProduct}.
-	%This establishes \cref{k-planar}.
-	% Note to self-and-others. In our original SODA submission we had 18k^2+30k here because we forgot that the layer compression step has compress k+1 consecutive layers, not just k consecutive layers.
-\end{proof}
-
-\subsection{\boldmath $1$-Planar Graphs}
-\label{sec-1-planar}
-
-$1$-planar graphs are of particular interest to the graph drawing community~\citep{kobourov.liotta.ea:annotated}.
-In this case, we obtain better constants and an additional property (planarity) of $H$.
-
-%(\htmladdnormallink{https://arxiv.org/abs/1703.02261}{arXiv:1703.02261}) has 143 entries, and even this is now four years out of date.
-%
-%   An important aspect of our result for $1$-planar graphs is that graph $H$ in the product is planar and has treewidth $3$, i.e., it has simple treewidth $3$.
-In a more recent application of product structure \citet{BDJM} show that such graphs have $\ell$-vertex rankings using $O(\log n/\log\log\log n)$ colours, and this is tight.
-This result requires that $H$ have simple treewidth $3$, even treewidth $3$ is not sufficient.
-% }
-
-\begin{thm}
-\label{1-planar}
-Every 1-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
-\end{thm}
-
-%In \cref{1_planar_is_4_framed}, below, we show that every $1$-planar graph is a subgraph of a $4$-framed multigraph.
-This makes \cref{1-planar} an immediate consequence of the following theorem, whose proof is the subject of the rest of this section.
-
-\cref{1-planar} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma, which is of independent interest.
-
-\note{DW}{Say something about how \cref{1_planar_is_4_framed} is better than similar results in the literature.}
-
-\begin{lem}\label{1_planar_is_4_framed}
-  A graph $G$ is $1$-planar if and only if it is contained in a $4$-framed multigraph.
-\end{lem}
-
-\begin{proof}
-Let $G$ be a $1$-plane graph.
-We may assume that no two edges incident to a common vertex of $G$ cross, since such a crossing can be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}.
-  % We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
-We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.
-This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c@{\hspace{1cm}}c}
-        \includegraphics{figs/one_planar_example-3} &
-        \includegraphics{figs/one_planar_example-4} \\
-        $G$ & $G'$
-     \end{tabular}
-    \end{center}
-    \caption{The graph $G'$ obtained by removing pairs of crossing edges from $G$ is a plane multigraph whose faces all have three or four sides.}
-    \label{one_planar_example}
-\end{figure}
-
-To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ that cross at $p$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
-Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$.
-In the following paragraphs, subscripts on the vertices in $W$ are implicitly taken modulo $r$.
-By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
-
-If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
-Therefore each dummy vertex $w_i$ of $W$ is surrounded by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
-Since $w_i$ has degree $4$ in $G_0$ and $F$ does not have two parallel edges on its boundary, $w_{i-1}\neq w_{i+1}$.
-Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.
-Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.
-
-% Therefore each face of $G_0$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
-% \begin{inparaenum}[(i)]
-%   \item bounded by three edges of $G$, none of which are crossed; or
-%   \item bounded by one edge of $G$ that is not crossed plus portions of two edges of $G$ that cross each other.
-% \end{inparaenum}
-
-Consider some pair of edges $vw$ and $xy$ that cross at $p$.
-The point $p$ is a vertex of $G_0$ that has degree $4$.
-There are four faces $F_1,\ldots,F_4$ of $G_0$ that contain $p$ and, from the preceding discussion each of $vx$, $xw$, $wy$, and $yv$ is an uncrossed edge of $G$.
-Therefore, by removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
-Therefore $G$ is a $4$-framed graph with frame $G'$.
-% Since, for any pair of crossing edges $\{vw,xy\}$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.
-% It follows immediately that, for any pair of vertices $a,b\in V(G)$, $\dist_{G'}(a,b)\le 2\dist_G(a,b)$.
-\end{proof}
-
-\subsection{\boldmath $(g,k)$-Planar Graphs}
-
-The definition of $k$-planar graphs naturally generalises for other surfaces.
-A graph $G$ drawn on a surface $\Sigma$ is $(\Sigma,k)$-plane if every edge of $G$ is involved in at most $k$ crossings.
-A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
-\cref{AddDummy} immediately generalises as follows:
-
-\begin{obs}
-\label{gAddDummy}
-Every $(g,k)$-planar graph $G$ is contained in $G_0^\SS$ for some graph $G_0$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\SS$ for $G_0$.
-Moreover, $V(G) \subseteq V(G_0)$ and for every edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most $k+1$, such that every internal vertex in $P$ has degree at most $4$ in $G_0$.
-\end{obs}
-
-Theorems~\ref{ShortcutProduct}  and \ref{GenusProduct}(b) imply a product structure theorem for $(g,k)$-planar graphs.
-The obtained bounds are improved by the following result, which is proved using exactly the same approach used in the proof of \cref{kPlanarProduct} (applying \cref{GenusProduct}(b) instead of \cref{PlanarProduct}(b)).
-We omit repeating the details.
-
-\begin{thm}
-\label{gkPlanarProduct}
-Every $(g,k)$-planar graph is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with $\tw(H) \leq \binom{k+4}{3}-1$, where $\ell:=\max\{2g,3\}\cdot(6k^2+16k+10)$.
-\note{DW}{I have changed $\tw(H) \leq \binom{k+5}{4}-1$ to $\tw(H) \leq \binom{k+4}{3}-1$, okay?}
-\end{thm}
-
-\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{framed} with the method in \citep{DHHW}.
-I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
-Okay?}
-
-Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs (or any of the other classes presented in \cref{Examples}) was in terms of layered treewidth, which we now define.
-A \defin{layered tree-decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{T}$ of $G$.
-The layered width of $(\mathcal{L},\mathcal{T})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{T}\}$.
-The \defin{layered treewidth} of $G$ is the minimum layered width of any layered tree-decomposition of $G$.
-\citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph.
-\citet{dujmovic.eppstein.ea:structure} show that every $k$-planar graph has layered treewidth at most $6(k+1)$, and more generally that every $(g,k)$-planar graph has layered treewidth at most $(4g+6)(k+1)$.
-It follows from this result that $(g,k)$-planar graphs have treewidth $O(\sqrt{(g+1)(k+1)n})$ and thus have balanced separators of the same order, which can also be concluded from the work of \citet{FP08}.
-In related work, \citet{grigoriev.bodlaender:algorithms} used structural results to obtain approximation algorithms for $(g,k)$-planar graphs, and \citet{PachToth97} determined the maximum number of edges in a $k$-planar graph (up to a constant factor).
-
-If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
-In particular, if $\PP=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree-decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree-decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{DJMMUW20}.
-Therefore, any property that holds for graphs of bounded layered treewidth also holds for $G$.
-What sets layered partitions apart from layered treewidth is that they lead to constant upper bounds on the queue-number  and non-repetitive chromatic number, whereas for both these parameters, the best known upper bound obtainable via layered treewidth is $O(\log n)$; see \cref{Applications}.
-
-\subsection{Rough Characterisation}
-\label{Characterisation}
-
-\cref{gAddDummy} shows that $(g,k)$-planar graphs can be obtained by a shortcut system applied to a graph of Euler genus $g$, where internal vertices on the paths have bounded degree.
-This observation and the following converse result together provide a rough characterisation of $(g,k)$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $(g,k)$-planar.
-
-\begin{lem}
-	\label{DrawG}
-	Fix integers $g\geq 0$ and $k,\Delta\geq 2$.
-	Let $G_0$ be a graph of Euler genus at most $g$. Let $G$ be
-	a graph with $V(G) \subseteq V(G_0)$ such that for every edge $vw \in
-	E(G)$ there is a $vw$-path $P_{vw}$ in $G_0$ of length at most $k$, such
-	that every internal vertex on $P_{vw}$ has degree at most $\Delta$ in
-	$G_0$. Then $G$ is $(g, 2k(k+1)\Delta^{k} )$-planar.
-\end{lem}
-
-\begin{proof}
-	For a vertex $x$ of $G_0$ with degree at most $\Delta$, and for $i\in\{1,\dots,k-1\}$, say a vertex $v$ is \defin{$i$-close} to $x$ if there is a $vx$-path $P$ in $G_0$ of length at most $i$ such that every internal vertex in $P$  has degree at most $\Delta$ in $G_0$.
-	For each edge $vw$ of $G$, say that $vw$ \defin{passes through} each internal vertex on $P_{vw}$.	Say $vw$ passes through $x$.
-Then $v$ is $i$-close to $x$ and $w$ is $j$-close to $x$ for some $i,j\in\{1,\dots,k-1\}$ with $i+j\leq k$.
-At most $\Delta^{i}$ vertices are $i$-close to $x$.
-	Thus, the number of edges of $G$ that pass through $x$ is at most
-	\[
-	\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i \Delta^j
-	= \sum_{i=1}^{k-1} \Delta^i  \sum_{j=1}^{k-i} \Delta^j
-	< \sum_{i=1}^{k-1} \Delta^i  2 \Delta^{k-i}
-	= \sum_{i=1}^{k-1} 2\Delta^k
-	< 2k \Delta^k \enspace.
-	\]
-	Fix a $(\Sigma,k)$-plane drawing of $G_0$ in a surface $\Sigma$ of Euler genus at most $g$. Choose $\epsilon>\epsilon'>0$.
-	For each vertex $v$ of $G_0$, let $B_v$ be the \defin{ball} $\{x\in \mathbb{R}^2:\dist(x,v)\leq\epsilon\}$
-	For each edge $vw$ of $G_0$, let $C_{vw}$ be the \defin{channel} $\{x\in \mathbb{R}^2:\dist(x,vw)\leq\epsilon'\}\setminus(B_v\cup B_w)$.
-	We may choose $\epsilon>\epsilon'>0$ sufficiently small so that
-	(i) $B_v\cap B_w=\emptyset$ for distinct $v,w\in V(G_0)$,
-	(ii) $C_{vw}\neq\emptyset$ for $vw\in E(G_0)$,
-	(iii) $B_v\cap C_{xy}=\emptyset$ for $v\in V(G_0)$ and $xy\in E(G_0)$, and
-	(iv) $C_{vw}\cap C_{xy}=\emptyset$ for distinct $vw,xy\in E(G_0)$.
-	Draw each edge $vw$ of $G$ following the sequence of balls and channels defined by the vertices and edges in $P_{vw}$.
-	This can be done so that whenever edges $vw$ and $xy$ of $G$ cross, the crossing point is in $B_z$ for some vertex $z$ of $G_0$ that is internal in $P_{vw}$ or in $P_{xy}$, and $vw$ and $xy$ cross at most once in each such $B_z$.
-
-Thus, for each edge $vw$ of $G$, every edge of $G$ that crosses $vw$ passes through a vertex on $P_{vw}$ (including $v$ and/or $w$ if they too have degree at most $\Delta$).
-	Since $P_{vw}$ has at most $k+1$ vertices, and less than $2k\Delta^{k}$ edges of $G$ pass through each vertex on $P_{vw}$, the edge $vw$ is crossed by less than $2k(k+1)\Delta^{k}$ edges in $G$.
-Hence $G$ is $(g, 2k(k+1)\Delta^{k} )$-planar.
-\end{proof}
-
-%\referee{2}{Page 16.
-At the end of Page 16, I do not see why 'every pair of edges cross at most once.'  Without further assumptions, it seems as if two paths in the path system could intersect several times.
-Therefore, by drawing 'each edge vw of G alongside $P_{vw}$ in $G_0$', there may be two edges which cross several times.
-It seems this can be fixed by choosing the path system carefully by 'uncrossing' paths which intersect at more than two internal vertices.}
-%
-%\note{DW}{I have revised the proof to address this error. I couldn't think of a way to do this without introducing balls and channels (and without vigorous hand-waving). Is it readable?}
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Applications}
-\label{Applications}
-
-Here we discuss some of the consequences of the above theorems for $k$-planar and $(g,k)$-planar graphs.
-
-\referee{1}{4.
-Sections 4 and 5: I think as they are currently organized they distract from the purpose of the paper, as stated in the introduction, to ``prove product structure theorems for several non-minor-closed classes of interest.'' Which of Corollaries 1--4 and Theorems 12--17 are most important?
-I would remove any mention of the applications from the examples section, put that section first, and state in that section a single main theorem with all of the best bounds on the product structure.
-The applications section is mostly a survey and I think should be focused on the most exciting new corollaries.
-Theorems 18 and 19 should be in the applications section.}
-
-\note{DW}{I don't want to do a complete restructuring. In our response, let's just say why we structure it the way we do.}
-
-%%%%%%%%%%%%%%%%%%%
-\subsection{Queue Layouts}
-
-For an integer $k\geq 0$, a \textit{$k$-queue layout} of a graph $G$ consists of a linear ordering $\preceq$ of $V(G)$ and a partition $\{E_1,E_2,\dots,E_k\}$ of $E(G)$, such that for $i\in\{1,2,\dots,k\}$, no two edges in $E_i$ are nested with respect to $\preceq$.
-That is, it is not the case that $v\prec x \prec y \prec w$ for edges $vw,xy\in E_i$.
-The \textit{queue-number} of a graph $G$, denoted by $\qn(G)$, is the minimum integer $k$ such that $G$ has a $k$-queue layout.
-Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
-\citet{DJMMUW20} recently proved this conjecture using \cref{PlanarProduct} and the following lemma.
-Indeed, resolving this question was the motivation for the development of \cref{PlanarProduct}.
-
-\begin{lem}[\citep{DJMMUW20}]
-\label{qn}
-If $G\subseteq H \boxtimes P \boxtimes K_\ell$ then
-$\qn(G) \leq  3 \ell \, \qn(H) + \floor{\tfrac{3}{2}\ell}
-\leq 3 \ell \, 2^{\tw(H)}  - \ceil{\tfrac{3}{2}\ell}$.
-\end{lem}
-
-\cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$. \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
-Note that \citet{DJMMUW20} previously proved the bound of
-$O(g^{k+2})$ using \cref{GenusProduct} and an ad-hoc method.
-Our result provides a better bound when $g>2^{k^2}$ \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
-In the case of 1-planar graphs we can improve further.
-\citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
-Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$.
-\cref{qn,1-planar} then imply:
-
-\begin{cor}
-\label{1PlanarQueue}
-Every 1-planar graph has queue-number at most $3 \times 7 \times 5 + \floor{\tfrac{3}{2} \times 7} = 115$.
-\end{cor}
-
-%%%%%%%%%%%%%%%%%%%
-\subsection{Non-Repetitive Colouring}
-
-The next two applications are in the field of graph colouring. For our purposes, a \defin{$c$-colouring} of a graph $G$ is any function $\phi\colon V(G)\to C$, where $C$ is a set of size at most $c$.
-A $c$-colouring $\phi$ of $G$ is \defin{non-repetitive} if, for every path $v_1,\ldots,v_{2h}$ in $G$, there exists $i\in\{1,\ldots,h\}$ such that $\phi(v_i)\neq\phi(v_{i+h})$.
-The \defin{non-repetitive chromatic number} $\pi(G)$ of $G$ is the minimum integer $c$ such that $G$ has a non-repetitive $c$-colouring.
-This concept, introduced by \citet{AGHR-RSA02}, has since been widely studied; see \citep{dujmovic.esperet.ea:planar} for more than 40 references.
-Up until recently the main open problem in the field has been whether planar graphs have bounded non-repetitive chromatic number, first asked by \citet{AGHR-RSA02}.
-\citet{dujmovic.esperet.ea:planar} recently solved this question using \cref{PlanarProduct} and the following lemma.
-
-\begin{lem}[\citep{dujmovic.esperet.ea:planar}]
-\label{non-repetitive}
-If $G\subseteq H\boxtimes P \boxtimes K_\ell$ then $\pi(G)\le \ell\, 4^{\tw(H)+1}$.
-\end{lem}
-
-\cref{non-repetitive,kPlanarProduct,1-planar,gkPlanarProduct} imply the following result:
-
-\begin{cor}\quad
-\begin{compactitem}
-\item	For every $1$-planar graph $G$, $\pi(G)\le 7\times 4^4=1792$.
-%\item For every $k$-planar graph $G$,
-%	$\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
-\item	For every $(g,k)$-planar graph $G$,
-	$  \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}.$
-\note{DW}{I have changed $4^{\binom{k+5}{4}}$ to $4^{\binom{k+4}{3}}$, okay?}
-\end{compactitem}
-\end{cor}
-
-Prior to the current work, the strongest upper bound on the non-repetitive chromatic number of $n$-vertex  $k$-planar graphs was $O(k\log n)$ \cite{dujmovic.morin.ea:layered}.
-
-%%%%%%%%%%%%%%%%%%%
-\subsection{Centered Colourings}
-\label{centered-colourings}
-
-A $c$-colouring $\phi$ of $G$ is \defin{$p$-centered} if, for every connected subgraph $X\subseteq G$, $|\{\phi(v):v\in V(X)\}| > p$ or there exists some $v\in V(X)$ such that $\phi(v)\neq \phi(w)$ for every $w\in V(X)\setminus\{v\}$.
-In words, either $X$ receives more than $p$ colours or some vertex in $X$ receives a unique colour.
-Let $\chi_p(G)$ be the minimum integer $c$ such that $G$ has a $p$-centered $c$-colouring.
-Centered colourings are important since they characterise classes of bounded expansion, which is a key concept in the sparsity theory of \citet{Sparsity}.
-
-\citet{PS21} and \citet{DFMS21} use \cref{ps} and \cref{PlanarProduct}(b), respectively, to show that $\chi_p(G)$ is polynomial in $p$ when $G$ is planar or of bounded Euler genus.
-Upper bounds on $\chi_p$ for graphs of given treewidth~\citep{PS21} and for graph products~\citep{DFMS21} imply the next lemma.
-
-%In the journal version, \citet{DFMS21} proved that $\chi_p(H_1\boxtimes H_2) \leq \chi_p(H_1) \cdot \chi(H_2^p)$ where $H_2^p$ is the $p$-th power of $H_2$.
-Apply this with $H_2:= P \boxtimes K_\ell$.
-Then $H_2^p= P^p\boxtimes K_\ell$ and $\chi(H_2^p)=(p+1)\ell$.
-So $\chi_p(H\boxtimes P \boxtimes K_\ell) \leq \chi_p(H_1) \cdot (p+1)\ell$.
-}
-
-\begin{lem}[\citep{DFMS21,PS21}]
-\label{p-centered}
-For every graph $H$ of treewidth at most $t$ and for every path $P$, if $G\subseteq H\boxtimes P \boxtimes K_\ell$ then
-\[\chi_p(G)\le \ell (p+1)\, \chi_p(H) \leq \ell (p+1) \tbinom{p+t}{t}.\]
-%If $G\subseteq H\boxtimes P \boxtimes K_\ell$ then $\chi_p(G)\le \ell (p+1)\, \chi_p(H)$.
-\end{lem}
-
-%\note{DW}{For consistency with \cref{qn}, I suggest we combine \cref{p-centered,p-centered-treewidth}, and just write, ``}
-
-%\begin{proof}
-%	By \cref{PartitionProduct}, $G$ has an $H$-partition $(\mathcal{L}=\langle L_0,L_1,\ldots\rangle, \PP=(B_x:x\in V(H))$ with layered width at most $\ell$.
-Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.
-For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in L_i\cap B_x$ such that:
-%   \begin{compactenum}
-%     \item $\alpha(v)$ is unique among $\{\alpha(w): w\in L_i\cap B_x\}$, which is possible
-%     since $|L_i\cap B_x|\le \ell$,
-%     \item $\beta(v)= i\bmod (p+1)$, and
-%     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
-%   \end{compactenum}
-% To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
-%
-% First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$.
-Since $X$ is connected, $X$ contains a path from $v$ to $w$.
-By the definition of layering, this path contains at least one vertex from $L_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$.
-Therefore, $|\{\beta(v'):v'\in V(X)\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
-%
-% Otherwise, $V(X)\subseteq L_{i}\cup\cdots\cup L_{i+s}$ for some $s<p$.
-Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.
-If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in V(X)\}|> p$ so $|\{\phi(v):v\in V(X)\}|> p$ and we are done.
-Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
-% For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
-Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.
-Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
-% \end{proof}
-%
-%\referee{2}{Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.}
-
-\cref{p-centered,kPlanarProduct,1-planar,gkPlanarProduct} immediately imply the following results, for every $p\ge 2$:
-
-\begin{cor}\quad
-\begin{compactitem}
-\item For every $1$-planar graph $G$,\; $ \chi_p(G)\le 5 (p+3)(p+2)(p+1)^2$.
-%\item For every $k$-planar graph $G$,\; $\displaystyle \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}$.
-\item For every $(g,k)$-planar graph $G$,\;
-$\displaystyle \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{p+\binom{k+4}{3}-1}{\binom{k+4}{3}-1}$.
-\note{DW}{I have changed $\binom{k+5}{4}$ to $\binom{k+4}{3}$, okay?}
- \end{compactitem}
-\end{cor}
-
-Prior to the current work, the strongest known upper bounds on the $p$-centered chromatic number of $(g,k)$-planar graphs $G$ were doubly-exponential in $p$, as we now explain.
-\citet{dujmovic.eppstein.ea:structure} proved that $G$ has layered treewidth $(4g+6)(k+1)$.
-Van den Heuvel and Wood~\citep{vdHW17} showed that this implies that $G$  has $r$-strong colouring number at most $(4g + 6)(k + 1)(2r + 1)$.
-By a result of \citet{zhu:colouring}, $G$ has $r$-weak colouring number at most $( (4g + 6)(k + 1)(2r + 1) )^r$, which by another result of  \citet{zhu:colouring} implies that $G$ has  $p$-centered chromatic number at most $( (4g+6)(k+1)(2^{p-1} + 1) )^{2^{p-2}}$.
-The above results are substantial improvements, providing bounds on $\chi_p(G)$ that are polynomial in $p$ for fixed $g$ and $k$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Examples}
-\label{Examples}
-
-This section describes several examples of graph classes that can be obtained from a shortcut system typically applied to graphs of bounded Euler genus.
-
-\subsection{Map Graphs}
-
-\note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph. Is the following lemma known (at least in the $g=0$ case)?}
-
-\begin{lem}
-\label{MewMapGraphLemma}
-for all integers $g\geq 0$ and $d\geq 3$, every $(g,d)$-map graph is a spanning subgraph of  $(g,d)$-framed graph.
-\end{lem}
-
-\begin{proof}
-Let $G_0$ be a graph embedded in a surface of Euler genus $g$, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations.
-If two lakes have a common edge in their boundaries, then we may delete the edges and merge the lakes.
-Now assume that no two lakes have a common edge in their boundaries.
-Let $G$ be the corresponding map graph.
-Let $G_1$ be the graph with $V(G_1):=V(G)$, which is the set of nations.
-Consider vertices $v$ and $w$ of $G_1$.
-Let $A$ and $B$ be the nation faces of $G_0$ corresponding to $v$ and $w$.
-Then $vw\in E(G_1)$ if $A$ and $B$ have a common edge in their boundaries,
-or there is a vertex $x$ of $G_0$ in the boundary of $A$ and $B$, and in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes.
-Then $G_1$ has Euler genus at most $g$.
-The faces of $G_1$ correspond to vertices of $G_0$ and lakes of $G_0$.
-Triangulate each face of $G_1$ corresponding to a lake of $G_0$. For each vertex
-$x$ of $G_0$ the size of the face of $G_1$ corresponding to $x$ equals the number of nations incident to $x$, which is at most $d$.
-So every face of $G_1$ has size at most $d$.
-If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common, implying that  $v$ and $w$ are both on the face of $G_1$ corresponding to $x$, and $vw\in E(\widehat{G_1})$.
-Hence $G\subseteq \widehat{G_1}$.
-\end{proof}
-
-\begin{thm}
-\label{NewMapGraph}
-Every $d$-map graph is contained in $H \boxtimes P \boxtimes K_{ d + 3\floor{d/2} -3 }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
-\end{thm}
-
-Map graphs are defined as follows.
-Start with a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
-Let $G$ be the graph whose vertices are the nations of $G_0$, where two vertices are adjacent in $G$ if the corresponding faces in $G_0$ share a vertex.
-Then $G$ is called a \defin{$(g,d)$-map graph}.
-A $(0,d)$-map graph is called a (plane) \defin{$d$-map graph}; see \citep{FLS-SODA12,CGP02} for example.
-The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.
-%So $(g,d)$-map graphs generalise graphs embedded in a surface, and we now assume that $d\geq 4$ for the remainder of this section.
-
-There is a natural drawing of a map graph obtained by positioning each vertex of $G$ inside the corresponding nation and each edge of $G$ as a curve passing through the corresponding vertex of $G_0$.
-It is easily seen that each edge is in at most $\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}}$ crossings; see \citep{dujmovic.eppstein.ea:structure}.
-Thus $G$ is $(g,\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
-Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar.
-\cref{gkPlanarProduct} then establishes a product structure theorem for map graphs, but we get much better bounds by constructing a shortcut system directly.
-The following lemma is reminiscent of the characterisation of $(g,d)$-map graphs in terms of the half-square of bipartite graphs \citep{CGP02,dujmovic.eppstein.ea:structure}.
-
-\begin{lem}
-\label{MapShortcut}
-Every $(g,d)$-map graph $G$ is contained in $G_1^\SS$ for some graph $G_1$ with Euler genus at most $g$ and some $(2,\tfrac12 d(d-3) )$-shortcut system $\SS$ for $G_1$.
-\end{lem}
-
-\begin{proof}
-Let $G$ be a $(g,d)$-map graph.
-So there is a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
-Let $N$ be the set of nations.
-Then $V(G)=N$ where two vertices are adjacent in $G$ if the corresponding nation faces of $G_0$ share a vertex.
-Let $G_1$ be the graph with $V(G_1):=V(G_0) \cup N$, where distinct vertices $v,w\in N$ are adjacent in $G_1$ if the boundaries of the corresponding nations have an edge of $G_0$ in common, and $v\in V(G_0)$ and $w\in N$ are adjacent in $G_1$ if $v$ is on the boundary of the nation corresponding to $w$.
-Observe that $G_1$ embeds in the same surface as $G_0$ with no crossings, and that each vertex in $V(G_0)$ has degree at most $d$ in $G_1$.
-Consider an edge $vw\in E(G)$.
-If the nations corresponding to $v$ and $w$ share an edge of $G_0$, then $vw$ is an edge of $G_1$.
-Otherwise,  $v$ and $w$ have a common neighbour $x$ in $V(G_0)$.
-In the latter case, let $P_{vw}$ be the path $(v,x,w)$.
-Let $\SS$ be the set of all such paths $P_{vw}$.
-Each vertex $x\in V(G_0)$ is the middle vertex on at most $\tfrac12 d(d-3)$  paths in $\SS$.
-Thus $\SS$ is a $(2,\tfrac12 d(d-3))$-shortcut system for $G_1$, and by construction, $G \subseteq G_1^\SS$.
-\end{proof}
-
-Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,non-repetitive} imply the following result.
-
-%\begin{thm}
-%\label{PlaneMapPartition}
-%\note{DW}{For every integer $d\geq 4$, }
-%Every $d$-map graph $G$:
-%	\begin{compactitem}
-%		\item is contained in $H \boxtimes P \boxtimes K_{21d(d-3)}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
-%		\item has queue-number $\qn(G) < 32225\, d(d-3)$.
-%		\item has $p$-centered chromatic number $\chi_p(G) \leq 21d(d-3) (p+1)  \binom{p+9}{9}$,
-%		\item has non-repetitive chromatic number $ \pi(G) \leq 21 \cdot 4^{10} d(d-3)$.
-%	\end{compactitem}
-%\end{thm}
-
-\begin{thm}
-\label{MapPartition}
-For integers $g\geq 0$ and $d\geq 4$, if $\ell:=  7d(d-3)\, \max\{2g,3\}$ then every $(g,d)$-map graph $G$:
-\begin{compactitem}
-\item is contained in $H \boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
-\item has queue-number $\qn(G) <  1535\, \ell $.
-\item has $p$-centered chromatic number $\chi_p(G) \leq \ell (p+1)\,  \binom{p+9}{9}$,
-\item has non-repetitive chromatic number $ \pi(G) \leq 4^{10}\,\ell $.
-\end{compactitem}
-\end{thm}
-
-\note{DW}{\cref{MapPartition} has been updated.}
-
-\note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the method in \citep{DHHW}.
-I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
-Okay?}
-
-These results give the first constant upper bound on the non-repetitive chromatic number of map graphs, the first polynomial bounds on the $p$-centered chromatic number of map graphs, and the best known bounds on the queue-number of map graphs.
-
-%%%%%%%
-\subsection{String Graphs}
-
-A \defin{string graph} is the intersection graph of a set of curves in the plane with no three curves meeting at a single point; see  \cite{PachToth-DCG02,FP10,FP14} for example.
-For an integer $\delta\geq 2$, if each curve is in at most $\delta$ intersections with other curves, then the corresponding string graph is called a \defin{$\delta$-string graph}.
-A \defin{$(g,\delta)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
-
-\begin{lem}
-\label{StringShortcut}
-Every $(g,\delta)$-string graph $G$ is contained in $G_0^\SS$ for some graph $G_0$ with Euler genus at most $g$ and some $(\delta+1,\delta+1 )$-shortcut system $\SS$ for $G_0$.
-%Moreover, every internal vertex on the paths in $\SS$ has degree 4 in $G_0$.
-\end{lem}
-
-\begin{proof}
-Let $\mathcal{C}=\{C_v:v\in V(G)\}$ be a set of curves in a surface of Euler genus at most $g$ whose intersection graph is $G$.
-Let $G_0$ be the graph obtained by adding a vertex at the intersection point of every pair of curves in $\mathcal{C}$ that intersect,  where two such consecutive vertices on a curve $C_v$ are adjacent in $G_0$.
-For each vertex $v\in V(G)$, if $C_v$ intersects $k\leq\delta$ other curves, then introduce a new vertex called $v$ on $C_v$ between the
-$\floor{\frac{k}{2}}$-th vertex already on $C_v$ and the $\floor{\frac{k}{2}+1}$-th such vertex.
-For each edge $vw$ of $G$, there is a path $P_{vw}$ of length at most $2\ceil{\frac{\delta}{2}}\leq \delta+1$ in $G_0$ between $v$ and $w$.
-Let $\SS$ be the set of all such paths $P_{vw}$.
-Consider a vertex $x$ in $G_0$ that is an internal vertex on some path in $\SS$.
-Then $x$ is at the intersection of $C_v$ and $C_w$ for some edge $vw\in E(G)$.
-If some path $P\in \SS$ passes through $x$, then $P=P_{vu}$ for some edge $vu$ incident to $v$, or $P=P_{wu}$ for some edge $wu$ incident to $w$.
-At most $\ceil{\frac{\delta}{2}}$ paths in $\SS$ corresponding to edges incident to $v$ pass through $x$, and similarly for edges incident to $w$.
-Thus at most $2\ceil{\frac{\delta}{2}}\leq\delta+1$ paths in $\SS$ use $x$ as an internal vertex.
-Thus $\SS$ is a $(\delta+1,\delta+1)$-shortcut system for $G_0$, and by construction, $G \subseteq G_0^\SS$.
-\end{proof}
-
-A similar proof to that of \cref{StringShortcut} shows that every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.
-Theorems~\ref{PlanarProduct}(b), \ref{ShortcutProduct}  and \ref{GenusProduct}(b) and \cref{StringShortcut,qn} imply:
-
-\begin{thm}
-\label{StringPartition}
-For integers $g\geq 0$ and $\delta\geq 2$, let $\ell:= \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$ and $t:= \binom{ \delta+4}{3}-1$.
-Then  every $(g,\delta)$-string graph:
-\begin{compactitem}
-	\item is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$,
-	\item has queue-number $\qn(G^k) \leq 3 \ell \, 2^t - \ceil{\tfrac{3}{2}\ell}$.
-\end{compactitem}
-\end{thm}
-
-\note{DW}{\cref{StringPartition} has been updated.}
-
-Our results also give bounds on the non-repetitive chromatic number and the $p$-centered chromatic number of $(g,\delta)$-string graphs, but the bounds are weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ and $\chi_p(G)\le p(64\delta)^2$ by results of \citet{DJKW16} and \citet{DFMS21}, respectively.
-
-%%%%%%%%%%
-\subsection{Powers of Bounded Degree Graphs}
-\label{Powers}
-
-Recall that the \defin{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
-If $G$ is planar with maximum degree $\Delta$, then $G^k$ is $2k(k+1)\Delta^{k}$-planar by \cref{DrawG}.
-Thus we can immediately conclude that bounded powers of planar graphs of bounded degree admit bounded layered partitions.
-However, the bounds we obtain are improved by the following lemma that constructs a shortcut system directly.
-
-\begin{lem}
-\label{PowerShortcut}
-If a graph $G$ has maximum degree $\Delta$, then $G^k = G^\SS$ for some $(k,2k \Delta^{k})$-shortcut system $\SS$.
-\end{lem}
-
-\begin{proof}
-For each pair of vertices $x$ and $y$ in $G$ with $\dist_G(x,y)\in\{1,\dots,k\}$, fix an $xy$-path $P_{xy}$ of length
-$\dist_G(x,y)$  in $G$.
-Let $\SS:=\{P_{xy}: \dist_G(x,y)\in\{1,\dots,k\} \}$.
-Say $P_{xy}$ uses some vertex $v$ as an internal vertex.
-If $\dist_G(v,x)=i$ and $\dist_G(v,y)=j$, then $i,j\in\{1,\dots,k-1\}$ and $i+j\leq k$.
-The number of vertices at distance $i$ from $v$ is at most $\Delta^i$.
-Thus the number of paths in $\SS$ that use $v$ as an internal vertex is at most
-\[\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i\Delta^j
-= \sum_{i=1}^{k-1} \Delta^i \sum_{j=1}^{k-i} \Delta^j
-< \sum_{i=1}^{k-1} \Delta^i ( 2 \Delta^{k-i} )
-< 2k \Delta^k\enspace.\]
-Hence $\SS$ is a $(k, 2k \Delta^k)$-shortcut system.
-\end{proof}
-
-\cref{ShortcutProduct,PowerShortcut} imply:
-
-\begin{thm}
-\label{PowerProduct}
-Let $G$ be contained in $H\boxtimes P\boxtimes K_\ell$ with maximum degree $\Delta$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
-Then for every integer $k\geq 1$, the $k$-th power $G^k$ is contained in $J\boxtimes P\boxtimes K_{2k \ell \Delta^{k}(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
-\end{thm}
-
-Theorems~\ref{PlanarProduct}(b), \ref{GenusProduct}(b) and \cref{ShortcutProduct} and  \cref{DrawG,qn,p-centered,non-repetitive,PowerShortcut} imply the following result, which with $g=0$ implies  \cref{kPowerBasic} in the introduction.
-
-\begin{thm}
-\label{PowerGenus}
-For integers $g\geq 0$ and $k,\Delta\geq 1$, let $\ell:= \max\{2g,3\} (2k^4+6k^2) \Delta^{k}$ and $t:= \binom{k+3}{3}-1$. Then for every graph $G$ of Euler genus $g$ and maximum degree $\Delta$,
-\begin{compactitem}
-\item $G^k$ is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$
-\item $G^k$ is $(g, 2k(k+1)\Delta^{k} )$-planar,
-\item $G^k$ has queue-number $\qn(G^k) \leq 3 \ell \cdot 2^t - \ceil{\tfrac{3}{2}\ell}$.
-\item $G^k$ has  $p$-centered chromatic number $\chi_p(G^k) \leq \ell (p+1)\,  \binom{p+t}{t}$,
-\item $G^k$ has  non-repetitive chromatic number $ \pi(G^k) \leq \ell \, 4^{t+1}$.
-\end{compactitem}
-\end{thm}
-
-\note{DW}{\cref{PowerGenus} has been updated}
-
-This result is the first constant upper bound on the queue-number of bounded powers of graphs with bounded degree and bounded Euler genus.
-For every graph $G$, since $G^k$ has maximum degree at most $\Delta^k$, a result of \citet{DJKW16} implies that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$.
-\cref{PowerGenus} improves upon this bound when $k,g\ll\Delta$.
-Similarly, a result of \citet{DFMS21} implies that $\chi_p(G^k)\le 1024p\Delta^{2k}$ and \cref{PowerGenus} improves upon this bound when $p,k,g\ll\Delta$.
-
-\cref{qn,p-centered,non-repetitive,MinorFreeDegree,PowerProduct} imply the following analogous result for powers of graphs in any minor-closed class with bounded maximum degree.
-
-\begin{thm}
-\label{PowerMinor}
-For every graph $X$ there exists an integer $c$ such that for all integers $k,\Delta\geq 1$, if $t:= 2k\Delta^{k}(k^3+3k)\binom{k+c\Delta}{c\Delta}-1$ and $G$ is an $X$-minor-free graph with maximum degree $\Delta$, then:
-\begin{compactitem}
-	\item $G^k$ is contained in $H\boxtimes P$ for some graph $H$ with treewidth $t$ and for some path $P$.
-	\item $G^k$ has queue-number at most $3\cdot 2^t-2$,
-	\item $G^k$ has $p$-centered chromatic number at most $(p+1)\binom{p+t}{t}$.
-\end{compactitem}
-\end{thm}
-
-\subsection{$k$-Nearest-Neighbour Graphs}
-
-In this section, we show that $k$-nearest neighbour graphs of point sets in the plane are $O(k^2)$-planar.
-For two points $x,y\in\R^2$, let $d_2(x,y)$ denote the Euclidean distance between $x$ and $y$.
-The $k$-nearest-neighbour graph of a point set $P\subset\R^2$ is the geometric graph $G$ with vertex set $V(G)=P$, where the edge set is defined as follows.
-For each point $v\in P$, let $N_k(v)$ be the set of $k$ points in $P$ closest to $v$.
-Then $vw\in E(G)$ if and only if $w\in N_k(v)$ or $v\in N_k(w)$.
-(The edges of $G$ are straight-line segments joining their endpoints.) See \citep{ProximityGraphs} for a survey of results on $k$-nearest neighbour graphs and other related proximity graphs.
-
-The following result, which is immediate from \citet[Corollary~4.2.6]{abrego.munroy.ea:on} states that $k$-nearest-neighbour graphs have bounded maximum degree:
-\begin{lem}
-\label{k-nn-max-degree}
-The degree of every vertex in a $k$-nearest-neighbour graph is at most $6k$.
-\end{lem}
-
-We make use of the following well-known observation (see for example, \citet[Lemma~2]{bose.morin.ea:routing}):
-\begin{obs}
-\label{convex}
-If $v_0,\ldots,v_3$ are the vertices of a convex quadrilateral in counterclockwise order then there exists at least one $i\in\{0,\ldots,3\}$ such that $\max\{d_2(v_i,v_{i-1}), d_2(v_i,v_{i+1})\} < d_2(v_{i-1},v_{i+1})$, where subscripts are taken modulo 4.
-\end{obs}
-
-\begin{lem}
-\label{nearest-neighbour}
-  Every $k$-nearest-neighbour graph is $O(k^2)$-planar.
-\end{lem}
-
-\begin{proof}
-  Let $G$ be a $k$-nearest-neighbour graph and consider any edge $vw\in E(G)$.
-  Let $xy\in E(G)$ be an edge that crosses $vw$.  Note that $vxwy$ are the vertices of a convex quadrilateral in (without loss of generality) counter-clockwise order. Then we say that
-  \begin{compactenum}
-    \item $xy$ is of Type~$v$ if $\max\{d_2(v,x), d_2(v,y)\}< d_2(x,y)$;
-    \item $xy$ is of Type~$w$ if $\max\{d_2(w,x), d_2(w,y)\}< d_2(x,y)$; or
-    \item $xy$ is of Type~C otherwise.
-  \end{compactenum}
-  If $xy$ is of Type~C, then \cref{convex} implies that $\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)$ without loss of generality.
-   In this case, we call $x$ a Type~C vertex.
-We claim that $V(G)$ contains at most $k-1$ Type~C vertices.
-Indeed, more than $k-1$ Type~C vertices would contradict the fact that $vw\in E(G)$ since every Type~C vertex is closer to both $v$ and $w$ than $d_2(v,w)$.
-
-  Next observe that, if $xy$ is of Type~$v$, then at least one of $xv$ or $yv$ is in $E(G)$ in which case we call $x$ (respectively $y$) a Type~$v$ vertex.
-By \cref{k-nn-max-degree}, there are at most $6k$ Type~$v$ vertices.
-Similarly, there are at most $6k$ Type~$w$ vertices.
-
-  Thus, in total, there are at most $13k-1$ Type~$v$, Type~$w$, and Type~C vertices.
-By \cref{k-nn-max-degree}, each of these vertices is incident with at most $6k$ edges that cross $vw$.
-Therefore, there are at most $78k^2-6k$ edges of $G$ that cross $vw$.
-Since this is true for every edge $vw\in E(G)$, $G$ is $(78k^2-6k)$-planar.
-\end{proof}
-
-Note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n\ge k+1$ vertices has at least $kn/2$ edges and at most $kn$ edges.
-For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is $\Omega(k^2)$.
-
-\cref{nearest-neighbour,kPlanarProduct,qn,p-centered} imply:
-
-\begin{cor}
-\label{k-nn}
-For every integer $k\geq 1$ there exists integers $t\leq O(k^6)$ and $\ell\leq O(k^4)$ such that every $k$-nearest-neighbour graph:
-\begin{compactitem}
-\item is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $t$ and some path $P$,
-\item has queue-number at most $2^{O(k^6)}$, and
-\item has $p$-centered chromatic number at most $\ell (p+1)\binom{p+t}{t}$.
-\end{compactitem}
-\end{cor}
-
-\cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
-However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (36+o(1))k^2$ by a result of \citet{DJKW16}.
-
-\section{Wrapping Up}
-\label{WrappingUp}
-
-As mentioned in \cref{Introduction}, \citet{DEJGMM21} prove results about adjacency labelling schemes for planar graphs.
-Stated in terms of universal graphs, their main theorem is interpreted as follows:
-
-\begin{thm}[\citep{DEJGMM21}]
-	\label{Universal}
-	For every fixed integer $t$ and every integer $n>0$ there exists a
-	graph $U_n$ with $n^{1+o(1)}$ vertices such that for every graph $H$ of
-	treewidth at most $t$ and path $P$, every $n$-vertex subgraph of $H\boxtimes P$ is isomorphic to an induced subgraph of $U_n$.
-\end{thm}
-
-Combining this with \cref{gkPlanarProduct,k-nn,PowerMinor,StringPartition,MapPartition} yields the following:
-
-\begin{thm}
-	\label{UniversalUniversal}
-	For every fixed graph $X$ and all fixed integers $d,\delta,\Delta,g,k>0$, and every integer
-	$n>0$, there exists a graph $U_n$ with $n^{1+o(1)}$ vertices such that
-	$U_n$ contains all of the following graphs as induced subgraphs:
-	\begin{compactitem}
-		\item every $n$-vertex $(g,k)$-planar graph;
-		\item every $n$-vertex $(g,d)$-map graph;
-		\item every $n$-vertex $(g,\delta)$-string graph;
-		\item every $n$-vertex graph $G^k$ where $G$ is $X$-minor-free and has maximum degree at most $\Delta$;
-		\item every $k$-nearest neighbour graph of $n$ points in $\R^2$.
-	\end{compactitem}
-\end{thm}
-
-We finish with an open problem.
-\cref{kPlanarProduct} shows that every $k$-planar graph is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $O(k^3)$ where $\ell\leq O(k^2)$.
-Does there exist a function $\ell:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph is contained in $H\boxtimes P \boxtimes K_{\ell(k)}$ for some graph $H$ with treewidth at most $C$?
-Perhaps $C=3$.
-Note that $C\geq 3$ even for planar graphs \citep{DJMMUW20}.
-
-%\subsection*{Note Added in Proof} Subsequent to the initial release of this paper, the treewidth 8 bound in \cref{PlanarProduct} was improved to 6 by \citet{UWY}, which implies that the treewidth $2g+8$ bound in \cref{GenusProduct}(a) can be improved to $2g+6$.
-Simmilarly, the treewidth 4 bound in \cref{GenusProduct}(b) was improved to 3 by \citet{DHHW}.
-\note{DW}{Say more about the consequence for the bounds in the rest of the paper}
-
-% %%%  Squashing the bibliography
-   \let\oldthebibliography=\thebibliography
-   \let\endoldthebibliography=\endthebibliography
-   \renewenvironment{thebibliography}[1]{%
-     \begin{oldthebibliography}{#1}%
-       \setlength{\parskip}{0ex}%
-       \setlength{\itemsep}{0ex}%
-   }{\end{oldthebibliography}}
-
-\bibliographystyle{DavidNatbibStyle}
-\bibliography{k-planar}
-
 \appendix

### 2022-03-15 k-planar.tex
added a couple of comments

--- 2022-03-14 k-planar.tex
@@
 
-\note{DW}{Do we want to mention ``every $1$-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ of treewidth $3$ and for some path $P$'' here?}
-
-\note{PM}{Yes, I think so.  Two theorems, like this, or one?}
+%\note{DW}{Do we want to mention ``every $1$-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ of treewidth $3$ and for some path $P$'' here?}
+%\note{PM}{Yes, I think so.  Two theorems, like this, or one?}
+%\note{DW}{two theorems}
 
@@
 
-This theorem has applications in diverse areas, which we explore in \cref{Applications}.
+These theorems have applications in diverse areas, which we explore in \cref{Applications}.
 
@@
 
-For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
-% \note{DW}{I suggest we replace ``and whose crossing edges are in the interior of these faces'' by ``such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$. }
+For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane spanning (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
 The graph $G_0$ is called the \defin{frame} of $G_0$.
@@
 \todo{Move this comment to the section on $1$-planar graphs?} For simple 1-planar graphs it is essential that we allow frames with parallel edges.
+\note{DW}{I feel that at this point we need to justify the introduction of multigraphs.}
 
@@
 }
+
+\note{DW}{The proof via 3-simplicity is shorter, so just keep it as is.}
 
@@
 \bibliography{k-planar}
-
-\appendix
-\section{Framed Graphs}
-
-\note{DW}{Work in progress}
-
-\note{DW}{Use the language of framed graphs}
-
-For a plane multigraph $G$, let $\widehat{G}$ be the embedded graph obtained from $G$ as follows: for each face $f$ of $G$ and for each pair of non-consecutive vertices $v$ and $w$ on the boundary of $f$, add the edge $vw$ to $\widehat{G}$ drawn across $f$.
-
-\begin{thm}
-\label{CliqueFaceProductStructure}
-Let $G$ be a 2-connected plane multigraph, in which every facial cycle has length in $\{3,\dots,d\}$ for some fixed integer $d\geq 3$.
-Then $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d +3 \floor{d/2}-3 }$ for some planar graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
-\end{thm}
-
-Note that \cref{CliqueFaceProductStructure} with $d=3$ implies \cref{PlanarProduct}(b)  by \citet{DJMMUW20} (since every planar graph is a subgraph of a triangulation).
-Moreover, the proof of \cref{CliqueFaceProductStructure} is a direct generalisation of the proof of
-\cref{PlanarProduct}(b).
-
-For a cycle $C$, we write $C = [ P_1 , \dots , P_k ]$ if $P_1 , \dots, P_k$ are pairwise disjoint (possibly empty) path subgraphs in $C$ such that $V(C)=\bigcup_i V(P_i)$ and $P_1,\dots,P_k$ are ordered around $C$.
-
-%and the endpoints of each path $P_i$ can be labelled $x_i$ and $y_i$ so that $y_ix_{i+1} \in E(C)$ for $i\in\{1,\dots,k\}$, where $x_{k+1}$ means $x_1$.
-This implies that $\{P_1,\dots,P_k\}$ is a vertex-partition of $C$.
-\note{DW}{In the proof below, when we apply induction, we need to allow $P_i=\emptyset$, right?}
-
-If $Q$ is a path in a graph, and $P_1, \dots,P_k$ are edge-disjoint non-empty subpaths of $Q$ ordered by $Q$ and partitioning $E(Q)$, then we write $Q=P_1,\dots,P_k$. \note{DW}{Do we need this?}
-
-We employ the following well-known variation of Sperner's Lemma (see~\citep{Proofs4}).
-
-\begin{lem}[Sperner's Lemma]
-\label{Sperner}
-Let $G$ be a near-triangulation whose vertices are coloured $1,2,3$, with the outer-face $F=[P_1,P_2,P_3]$ where each vertex in $P_i$ is coloured $i$.
-Then $G$ contains an internal face whose vertices are coloured $1,2,3$.
-\note{DW}{Does this allow parallel edges: check the proof in \citep{Proofs4}.}
-\end{lem}
-
-%If $C$ is a cycle in a graph, and $P_1, \dots,P_k$ are edge-disjoint non-empty subpaths of $C$ ordered around $C$ and partitioning $E(C)$, then we write $C=(P_1,\dots,P_k)$.
-
-Say $G$ is a 2-connected plane multigraph.
-Let $T$ be a spanning tree of $G$ rooted at a vertex $r$ on the outerface of $G$.
-A \defin{bipod} in $G$ (with respect to $T$) is any path $P_1SP_2$ (considered to be a subgraph), where $S$ is a facial path in $G$, and each $P_i$ is a vertical path in $T$ whose lower endpoint is in $S$.
-Here it is possible that $S$ or $P_i$ is empty.
-So every subpath of a bipod is a bipod.
-A \defin{tripod} in $G$ (with respect to $T$) is either a bipod in $G$ or a subgraph of $G$ consisting of a facial cycle $F$ of $G$, plus three distinct vertical paths $P_1,P_2,P_3$ in $T$ where the lower endpoint of each $P_i$ is in $F$.
-These definitions generalise definitions in \citep{DJMMUW20} where $G$ is assumed to be a plane triangulation, in which case $F$ is a 3-cycle.
-\note{DW}{We actually managed to omit the definition of `bipod' in \citep{DJMMUW20}!}
-
-%In the original proof for planar triangulation, B_i is an edge.
-
-For a cycle $C$ in a plane multigraph $G$, define \defin{$G\blah{C}$} to be the subgraph of $G$ consisting of $C$ and all the edges and vertices of $G$ in the interior of $C$.
-
-\begin{lem}
-\label{NewLemma}
-Let $G_0$ be a 2-connected plane multigraph.
-Let $T$ be a spanning tree of $G_0$ rooted at some vertex $r$ on the boundary of the outerface of $G_0$.
-Let $C = [P_1,\dots,P_k]$ be a cycle in $G_0$ with $r$ in the exterior of $C$, where $k\in\{1,2,3\}$ and $P_1, \dots, P_k$ are bipods in $G_0$.
-Let $G:=G_0\blah{C}$.
-Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that:
-\begin{compactenum}[(i)]
-	\item $P_1,\dots,P_k\in\PP$,
-	\item $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3,
-	\item $G/\PP=\widehat{G}/\PP$.
-%	 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
-\end{compactenum}
-\end{lem}
-
-\begin{proof}
-We may assume that each of $P_1,\dots,P_k$ is non-empty (since if some $P_i$ is empty, then we just consider $C=[P_1,\dots,P_{i-1},P_{i+1},\dots,P_k]$).
-
-The proof is by induction on the number of faces of $G$. Suppose $G$ has just one internal face (which is bounded by $C$). Let $\PP=\{P_1,\ldots,P_k\}$.
-Then $G/\PP=\widehat{G}/\PP = K_k$.
-Then desired tree-decomposition of $G/\PP$ consists of a single bag $\{P_1,\ldots,P_k\}$ indexed by a 1-node tree.
-Now assume that $G$ has at least two internal faces.
-
-First suppose that $C$ is a 2-cycle.
-Thus $k\leq 2$.
-Let $e$ be one of the edges in $C$.
-Let $F$ be the internal facial cycle of $G$ that includes $e$.
-Since each facial cycle of $G_0$ has length at least 3, $F$ consists of a non-empty path $P$ plus the edge $e$.
-Let $C'$ be obtained by replacing $e$ in $C$ by $P$.
-Let $G':=G\langle{C'}\rangle$.
-Thus $G'$ has one fewer face than $G$.
-Note that $C'=[P_1,\dots,P_k,P]$ (\note{DW}{or it can be written that way?}).
-By induction, $G'$ has a partition $\PP$ into tripods with respect to $T$ such that $P_1,\dots,P_k,P\in\PP$, and $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F'$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F'$ is in $B$.
-The bag of $\mathcal{T}$ that contains $P_1,\dots,P_k,P$ satisfies this property for $F$.
-Thus $\PP$ and $\TT$ satisfy the desired properties for $G$.
-Now assume that $|V(C)|\geq 3$.
-
-%\note{DW}{One option is to do induction on $|V(G)\setminus V(C)|-k$ and reduce to the $k=3$ case (similar to the $|V(C)|\leq 2$ case).
-When $k<3$ we can grab another face $F$ on the inside of $C$ and make the boundary of $F$ a new path.
-Then the main proof will only deal with the $k=3$ case.
-Is this any easier?}
-
-%Now suppose that $k=1$.
-Let $v$ and $w$ be the first and last vertex in $P_1$.
-So $C$ consists of $P_1$ plus the edge $vw$.
-Let $F$ be the internal facial cycle of $G$ containing $vw$..........
-Now suppose that $k=2$ .....
-
-Partition the vertices of $C$ into paths $R_1$, $R_2$, and $R_3$ as follows:
-\begin{compactenum}
-	\item If $k=1$ then, $P_1$ has at least three vertices (since $|V(C)|\geq 3$), so $P_1=[v, P_1', w]$ for two distinct vertices $v$ and $w$. We set $R_1:=v$, $R_2:=P_1'$ and $R_3:=w$.
-
-	\item If $k=2$ then we may assume without loss of generality that $P_1$ has at least two vertices so $P_1=[v,P_1']$. We set $R_1:=v$, $R_2:=P_1'$ and $R_3:=P_2$.
-
-	\item If $k=3$ then let $R_i:=P_i$ for each $i\in\{1,2,3\}$.
-\end{compactenum}
-
-Note that $R_i \subseteq P_{\alpha(i)}$ for some $\alpha(i)\in \{1,\dots,k\}$.
-
-Let $G'$ be a near-triangulation obtained from $G$ by triangulating every internal face.
-For $i\in\{1,2,3\}$, colour each vertex in $R_i$ by $i$.
-Now, for each remaining vertex $v$ in $G$, consider the path $P_v$ from $v$ to the root of $T$.
-Since $r$ is on the outer-face of $G_0$, $P_v$ contains at least one vertex of $C$.
-If the first vertex of $P_v$ that belongs to $F$ is in $R_i$, then assign the colour $i$ to $v$.
-We obtain a 3-colouring of the vertices of $G'$ that satisfies the conditions of Sperner's Lemma.
-By Sperner's Lemma there exists a triangular face $\tau=v_1v_2v_3$ of $G'$ whose vertices are coloured $1,2,3$ respectively.
-Let $F$ be the cycle bounding the face of $G$ containing $\tau$.
-
-For each $i\in\{1,2,3\}$, let $Q_i$ be the path in $T$ from $v_i$ to the first ancestor $v_i'$ of $v_i$ in $T$ that is contained in $C$.
-Observe that $Q_1$, $Q_2$, and $Q_3$ are disjoint since $Q_i$ consists only of vertices coloured $i$.
-Let $Q_i'$ be $Q_i$ minus its final vertex $v_i'$.
-Note that $Q_i$ may consist of the single vertex $v_i=v_i'$, in which case $Q_i'$ is empty.
-Imagine for a moment that the cycle $C$ is oriented clockwise, which defines an orientation of $R_1$, $R_2$ and $R_3$.
-Let $R_i^-$ be the subpath of $R_i$ that contains $v'_i$ and all vertices that precede it, and let $R_i^+$ be the subpath of $R_i$ that contains $v'_i$ and all vertices that succeed it.
-
-Let $Y$ be the tripod $F\cup Q_1'\cup Q_2' \cup Q_3'$.
-Say $F=v_1 S_1 v_2 S_2 v_3 S_3$, where $S_i$ is a facial path in $F$ clockwise from $v_i$ to $v_{i+1}$ (where $v_4=v_1$).
-Let $L_i$ be the path $Q_i'S_iQ_{i+1}'$ for $i\in\{1,2,3\}$ (where $Q_4=Q_1$ and $R_4=R_1$).
-So $L_i$ is a bipod.
-\note{DW}{Note that $S_i\neq\emptyset$ since $v_i\neq v_{i+1}$.
-Is this useful.}
-
-%Note that $C_i$ may be \defin{degenerate} in the sense that $[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ may consist only of a single edge $v_iv_{i+1}$.
-
-% Consider any non-degenerate $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$.
-Note that these four paths  are pairwise disjoint, and thus $C_i$ is a cycle.
-If  $Q_i'$ and $Q_{i+1}'$ are non-empty, then each is a vertical path in $T$.
-Furthermore, each of $R_i^-$ and $R_{i+1}^+$ consists of at most two vertical paths in $T$.
-Thus, $C_i$ is the concatenation of at most six vertical paths in $T$ \note{DW}{re-do}.
-Let $G_i$ be the subgraph of $G_0$ consisting of all the edges and vertices of $G_0$ contained in $C_i$ and the interior of $C_i$.
-Observe that $G_i$ contains $v_i$ and $v_{i+1}$ but not the third vertex of $\tau$.
-
-Let $Z$ be the plane graph $F\cup C\cup Q_1\cup Q_2\cup Q_3$.
-Then $Z$ is 2-connected, each face of $Z$ is bounded by a cycle, and the outerface of $Z$ is $C$.
-Let $Z_1,\dots,Z_m$ be the internal facial cycles of $Z$ except $F$.
-
-Consider $j\in\{1,\dots,m\}$.
-Let $G_j$ be the subgraph of $G$ consisting of $Z_j$ and all the edges and vertices of $G$ in the interior of $Z_j$.
-The outerface of $G_j$ is $Z_j$.
-Observe that for some $i\in\{1,2,3\}$, we have
-$Z_j=[R_{i,j}^+,R_{i+1,j}^-,S_{i,j}]$, where
-$R_{i,j}^+$ is some subpath of $R_i^+$,
-$R_{i+1,j}^-$ is some subpath of $R_{i+1}^-$, and
-$L_{i,j}$ is the path $Z_j-V(C)$ (which is contained in $Y$). \note{DW}{Add a figure of a good case and a weird case.}
-Since $R_i^+$, $R_{i+1}^-$ and $L_i$ are bipods,
-$R_{i,j}^+$, $R_{i+1,j}^-$ and $L_{i,j}$ are bipods.
-By construction, $v_{i-1}$ is not in $G_j$ (where $v_0=v_3$).
-Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.  Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that
-\begin{compactenum}[(i)]
-	\item
-	$R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$,
-	\item $G_j / \PP_j$ has a tree-decomposition $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3,
-	\item $G_j / \PP=\widehat{G_j}/\PP$.
-	%	 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
-\end{compactenum}
-
-%This graph has up to three \note{DW}{can be more} internal faces $C_1,C_2,C_3$ where each $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ and $R_{i}^+$ and $R_{i}^-$ are the same portions of $R_i$ as defined in \cref{NearTriang} \note{DW}{move in here}    Observe that  $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$, where $R_i^+$ and $R_{i+1}^-$ are bipods, and $I_i$ is the bipod $Q_i' L_i Q_{i+1}'$.
-Let $G_i$ be the subgraph of $G$ whose vertices and edges are in $C_i$ or its interior.
-
-%For $i\in\{1,2,3\}$, if $C_i$ is non-empty, then $G_i$ and $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$ satisfy the conditions of the lemma, and we may apply induction to $G_i$.
-(Note that one or two of $R_i^+$, $R_{i+1}^-$ and $I_i$ may be empty, in which case we apply the inductive hypothesis with $k=2$ or $k=1$, respectively.)\  This gives a partition $\PP_i$ of $G_i$ such that $H_i:= G_i / \PP_i$ satisfies the conclusions of the lemma.
-Let $(B^i_x:x\in V(J_i))$ be a tree-decomposition of $H_i$, in which every bag has size at most 4, and some bag $B^i_{u_i}$ contains the vertices of $H_i$ corresponding to  $R_i^+$, $R_{i+1}^-$ and $I_i$ (if they are non-empty).
-We do this for each non-degenerate $C_i$.
-
-We now construct the desired partition $\PP$ of $G$.
-Let $\PP$ be obtained from $\bigcup_{j\in\{1,\dots,m\}} \PP_j$ by
-replacing $R_{i,j}^+\in\PP_j$ by $P_{\alpha(i)}$,
-replacing $R_{i+1,j}^-\in\PP_j$ by $P_{\alpha(i+1)}$, and
-replacing $L_{i,j}\in\PP_j$ by $Y$.
-By construction, $\PP$ partitions $V(G)$ into tripods with respect to $T$ and $P_1,\ldots,P_k\in \PP$.
-
-%Since $P_1,
-%Since $C$ is a facial cycle of $G_j$, there is a bag $B_{u_j}$ of $G_j$ such that $P_1,\dots,P_k\in B_{u_j}$.
-
-We now construct the desired tree-decomposition of $G/\PP$.
-Let $D$ be the tree obtained from the disjoint union of $D_1,\dots,D_m$ by adding one new node $u$ adjacent to $u_1,\dots,u_m$.
-Define $B_u:=\{P_1,\dots,P_k,Y\}$.
-Apply the above replacements within the bags of $\mathcal{T}_1,\dots,\mathcal{T}_m$.
-We claim that $\mathcal{T}:=\{B_x:x\in V(D)\}$ is the desired tree-decomposition of $G/\PP$.
-By construction, every bag of $\mathcal{T}$ has size at most 4.
-
-First we show that for each part $X\in\PP$, the set $\{x\in V(D) : X\in B_x\}$ induces a subtree of $D$.
-If $X\not\in \{P_1,P_2,P_3,Y\}$ then $X$ is only in bags in some $D_j$, implying $\{x\in V(D) : X\in B_x\}$ induces a subtree of $D$.
-If $X\in \{P_1,P_2,P_3,Y\}$ then $X$ is in $B_u$, and if some part $X'$ of some $\PP_j$ was replaced by $X$ in the above construction, then $X'$ is in $B_{u_j}$ and $\{x\in V(D_j):X'\in B_x\}$ induces a subtree of $D_j$ (since $\mathcal{T}_j$ is a tree-decomposition of $G_j$).
-Thus $\{x\in V(D) : X\in B_x\}$ induces a subtree of $D$ (since $uu_j\in E(D)$).
-
-We now show that, for every edge $XW$ of $G/\PP$, there is a node $x\in V(D)$ such that $X,W\in B_x$.
-If $X,W\in \{P_1,P_2,P_3,Y\}$ then $X,W\in B_u$ as desired.
-Otherwise, without loss of generality, $X\in \PP_j\setminus \{P_1,P_2,P_3,Y\}$ for some $j\in\{1,\dots,m\}$.
-Thus $X\subseteq V(G_j)\setminus V(Z_j)$.
-Since $Z_j$ separates $V(G_j)\setminus V(Z_j)$ and $V(G)\setminus V(G_j)$ and $XW\in E(G/\PP)$, it must be that $W$ is also in $\PP_j$.
-Say $W$ replaced $W'$ in $\PP_j$ (where possibly $W=W'$).
-Since $\mathcal{T}_j$ is a tree-decomposition of $G_j$, $XW'\in B_x$ for some node $x\in V(D_j)$, implying $XW\in B_x$.
-This concludes the proof that $(B_x:x\in V(D))$ is the desired tree-decomposition of $G/\PP$.
-
-For every facial cycle $F_0$ of $G$, either $F_0=F$ or $F_0=C$ or $F_0$ is a facial cycle of some $G_j$.
-If $F=F_0$ then the only part of $\PP$ that intersects $F$ is $Y$, so $B_u$ is a bag of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
-If $F_0=C$ then $P_1,\dots,P_k$ are the only parts of $\PP$ that intersect $F_0$, so $B_u$ is a bag of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
-
-In the first case, there is a bag $B_x$ of $\mathcal{T}_j$ such that $V(F)\subseteq \bigcup_{X\in B_x} X$. This property is maintained by the above replacements since
-$R_{i,j}^+ \subseteq P_{\alpha(i)}$ and
-$R_{i+1,j}^- \subseteq P_{\alpha(i+1)}$ and
-$L_{i,j}\subseteq Y$.
-
-%For each non-degenerate $F_i$, and for each node $x\in V(J_i)$, initialise $B_x:= B^i_x$.
-%Recall that vertices of $H_i$ correspond to contracted paths in  $\PP_i$.
-%Each internal path in $\PP_i$ also lies  in $\PP$.
-%Each external path $P$ in $\PP_i$ is a subpath of $P_j$ for some $j\in\{1,\dots,k \}$ or is one of the paths among $Q'_1, Q'_2, Q'_3$.
-%For each such path $P$, for every $x\in V(J)$, in bag $B_x$,
-%replace each instance of the vertex of $H_i$ corresponding to $P$ by the vertex of $H$ corresponding to the path among $P_1,\ldots, P_k, Q'_1,\ldots,Q'_3$
-%that contains $P$.
-%This completes the description of $(B_x : x\in V(J))$.
-%By construction, $|B_x|\leq 9$ for every $x\in V(J)$.
-
-%We construct $\PP$ as before.
-Initialise $\PP:=\{P_1,\ldots,P_k,Y\}$.
-Then, for $i\in \{1,2,3\}$, each tripod in  $\PP_i$ is either fully contained in $F_i$ or it is \emph{internal} with none of its vertices in $F_i$.
-Add all these internal tripods in  $\PP_i$ to $\PP$.
-By construction, $\PP$ partitions $V(G)$ into tripods.
-The graph $H:= G/ \PP$ is planar since $G$ is planar and each tripod in $\PP$ induces a connected subgraph of $G$.
-
-%Next we produce the tree-decomposition $(B_x:x\in V(J))$ of $H$ that satisfies the requirements of the lemma.
-Let $J$ be the tree obtained from the disjoint union of $J_1$, $J_2$ and $J_3$ by adding one new node $u$ adjacent to $u_1$, $u_2$ and $u_3$.
-Let $B_u$ be the set of at most four vertices of $H$ corresponding to $Y,P_1,\dots,P_k$.
-For $i\in\{1,2,3\}$ and for each node $x\in V(J_i)$, initialise $B_x:= B^i_x$.
-
-%As in  the proof of \cref{NearTriang},  the resulting structure,  $(B_x:x\in V(J))$, is not yet a tree-decomposition of $H$ since some bags may  contain vertices of $H_i$ that are not necessarily vertices of $H$.
-Note that unlike in  \cref{NearTriang} this does not only include elements of $\PP_i$  that are contained in $F$.
-In particular, $I_i$ is also not an element of $\PP$ and thus does not correspond to a vertex of $H$.
-We remedy this  as follows.
-For $x\in V(J)$, in bag $B_x$, replace each instance of the vertex of $H_i$ corresponding to $I_i$ by the vertex of $H$ corresponding to $Y$.
-Similarly, by construction, $R_i^+$ is a subgraph of $P_{\alpha_i}$ for some $\alpha_i\in \{1,\dots,k\}$.
-For $x\in V(J)$, in bag $B_x$, replace each instance of the vertex of $H_i$ corresponding to $R_i^+$  by the vertex of $H$ corresponding to $P_{\alpha_i}$.
-Finally, $R_{i+1}^-$ is a subgraph of $P_{\beta_i}$ for some $\beta_i\in \{1,\dots,k\}$.
-For $x\in V(J)$, in bag $B_x$, replace each instance of the vertex of $H_i$ corresponding to $R_{i+1}^-$  by the vertex of $H$ corresponding to $P_{\beta_i}$.
-
-This shows that $(B_x:x\in V(J))$ is the desired tree-decomposition of $G/\PP$.
-\end{proof}
-
-\begin{proof}[Proof of \cref{CliqueFaceProductStructure}]
-Let $G$ be a 2-connected plane multigraph, in which every facial cycle has length in $\{3,\dots,d\}$.
-Let $C$ be the outer-cycle of $G$.
-Let $G^+$ be obtained from $G$ by adding a new vertex $r$ adjacent to every vertex on the outerface of $G$.
-So $G^+$ is a 2-connected plane multigraph and $G=G^+\blah{C}$.
-Let $T$ be a BFS spanning tree of $G^+$ rooted at $r$.
-Let $\mathcal{L}=(L_0,L_1,\dots)$ be the corresponding BFS layering.
-By \cref{NewLemma}, $G$ has a partition $\PP$ into tripods with respect to $T$ such that $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ for which every part of $\PP$ that intersects $F$ is in $B$.
-Since $G$ is a spanning subgraph of $\widehat{G}$, we may consider $\PP$ to be a partition of $\widehat{G}$, and $G/\PP \subseteq \widehat{G}/\PP$.
-Consider an edge $XY$ of $\widehat{G}/\PP$.
-By the definition of $\widehat{G}$, there is a facial cycle $F$ of $G$ such that both $X$ and $Y$ intersect $F$.
-By assumption, there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
-In particular, $X,Y\in B$.
-Hence $\TT$ is also a tree-decomposition of $\widehat{G}/\PP$.
-For each edge $vw$ of $\widehat{G}$, if $v\in L_i$ and $w\in L_j$ then $|i-j|\leq\floor{\frac{d}{2}}$.
-Grouping blocks of $\floor{\frac{d}{2}}$ consecutive layers in $\mathcal{L}$ gives a layering $\mathcal{L}'$ of $\widehat{G}$.
-Observe that each tripod in $\PP$ has at most $d+3\floor{\frac{d}{2}}-3$ vertices in each set of $\floor{\frac{d}{2}}$ layers of $\mathcal{L}$.
-Hence $\mathcal{L}'$ has layered width $d+3( \floor{\frac{d}{2}}-1)$ with respect to $\PP$.
-By \cref{PartitionProduct}, $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{d+3\floor{d/2}-3}$ where $H:=\widehat{G}/\PP$, which has treewidth 3.
-\note{DW}{This proof needs to be changed to show that $G/\PP = \widehat{G}/\PP$, so $H$ is a minor of $G$ and is thus planar.}
-\end{proof}
-
-We now give two graph classes for which \cref{CliqueFaceProductStructure} is applicable.
-
-\begin{lem}
-	An embedded multigraph $G$ is edge-maximal 1-plane with no faces of length 2 if and only if there is a 2-connected plane multigraph $G_0$ such that:
-	\begin{compactitem}
-		\item each face of $G_0$ is bounded by a 3-cycle or a 4-cycle,
-		\item no edge of $G_0$ is in the boundary of two facial 3-cycles,
-		\item $G=\widehat{G_0}$.
-	\end{compactitem}
-\end{lem}
-
-\begin{proof}
-\end{proof}
-
-\begin{cor}
-	\label{New1Planar}
-	Every 1-planar graph is contained in $H \boxtimes P \boxtimes K_7$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
-\end{cor}
-
 \end{document}

### 2022-03-16 k-planar.tex
Added references, a comment about 1_planar_is_4_framed

@@
 This result is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs. 
-\todo{Move this comment to the section on $1$-planar graphs?} For simple 1-planar graphs it is essential that we allow frames with parallel edges.
-\note{DW}{I feel that at this point we need to justify the introduction of multigraphs.}
+% \todo{Move this comment to the section on $1$-planar graphs?} 
+For simple 1-planar graphs it is essential that we allow frames with parallel edges. 
+% \note{DW}{I feel that at this point we need to justify the introduction of multigraphs.}
 
@@
 
-\todo[inline]{Alternative to the final three paragraphs in the preceding proof:
-
-To see that $H$ is planar, consider a plane multigraph $G''$ obtained by adding stars in some of the faces of $G'$, as follows:  When creating the part $S$, add a vertex $v_\tau$ to the interior of $\tau$ that is adjacent to each vertex of $\tau$ and place $v_\tau$ in $S$.
+% \todo[inline]{Alternative to the final three paragraphs in the preceding proof:
+% 
+% To see that $H$ is planar, consider a plane multigraph $G''$ obtained by adding stars in some of the faces of $G'$, as follows:  When creating the part $S$, add a vertex $v_\tau$ to the interior of $\tau$ that is adjacent to each vertex of $\tau$ and place $v_\tau$ in $S$.
 The graph $G''\supseteq G'$ obtained this way is clearly planar.
@@
 Now argue that $G/\mathcal{P}=G''/\mathcal{P}$\ldots
-}
-
-\note{DW}{The proof via 3-simplicity is shorter, so just keep it as is.}
+% }
+% 
+% \note{DW}{The proof via 3-simplicity is shorter, so just keep it as is.}
 
@@
     This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree decomposition.
-Since any graph having a $3$-simple tree decomposition is planar \cite{X}, this implies that $H$ is planar.
+Since any graph having a $3$-simple tree decomposition is planar \cite{knauer.ueckerdt:simple,kratochvil.vaner:note,elmallah.colbourn:on}, this implies that $H$ is planar.
 
@@
 \note{DW}{Say something about how \cref{1_planar_is_4_framed} is better than similar results in the literature.}
-\note{PM}{I don't know what to say without explaining that $4$-framed (simple) graphs only contain $3$-connected $1$-planar graphs.
+
+\note{PM}{See my attempt below.}
+% \note{PM}{I don't know what to say without explaining that $4$-framed (simple) graphs only contain $3$-connected $1$-planar graphs.
 Vida's example shows that there is a $1$-plane embedding of a graph that can't be augmented into a $4$-framed graph, but I haven't checked if it can't be drawn some better way.
@@
 
-\begin{proof}
+Before proving \cref{1_planar_is_4_framed}, we remark that the lemma seems not to be true if we replace `multigraph' with `graph'.
+We are only aware of two previous results on the relationship between $1$-planar graphs and $d$-framed graphs \cite{BDGGMR}:
+\begin{inparaenum}[(i)]
+  \item Every $3$-connected $1$-planar graph is contained in some $4$-framed graph.
+  \item Every $1$-planar graph is contained in some $8$-framed graph.
+\end{inparaenum}
+Neither of these results allows a proof of \cref{1-planar} without sacrificing generality or accepting a larger constant.
+
+\begin{proof}[Proof of \cref{1_planar_is_4_framed}]
 Let $G$ be a $1$-plane graph.

### 2022-03-16 k-planar.tex
Fixed broken reference

@@
 As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a),
-\citet{DJMMUW20} showed that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \cref{UWY}).
+\citet{DJMMUW20} showed that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \cite{UWY}).
 The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.
@@
 When $k=2$, $B_{y_0}=\{x_0,x_1,x_2\}$.
-In this case, the choice of $\tau$ ensures that there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_1$, and $P_2$ on its boundary.
-(Refer to \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.
+In this case, the fact that $\tau$ contains an edge $v_1v_2$ of $F$ with $v_1\in P_{x_1}$ and $v_2\in P_{x_2}$ ensures that there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_1$, and $P_2$ on its boundary.
+(Refer to the middle of \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.
 Therefore there are at most two bags, $B_y$ and $B_{z_i}$ of $\mathcal{T}$ that contain $\{x_0,x_1,x_2\}$.

### 2022-03-16 response.tex
Pfft
Added a bit more to the response.

@@
 
-The following comments are all related to the proof of Theorem~3.
+The following comments (and several additional minor comments) are all related to the proof of Theorem~3 in the original submission.
 Much of the difficulty comes from the fact that there are many things to keep track of.
@@
 
-A secondary issue is that a number of the proofs in this section are essentially "definition checking".
-The relationships between the various vertex sets that arise in the proof really do just follow from their definitions (which were made with this purpose in mind).
-We have done this checking explicitly, though each of these proofs could also be replaced by a sentence of the form ``This [claim/lemma] follows from the definitions of\ldots''.
-It is not clear what the best choice is.
+One issue is that a number of the proofs in this section are essentially "definition checking".
+The relationships between the various vertex subsets that arise in the proof really do just follow (by design) from their definitions.
+In the paper, we do this checking explicitly, though each of these proofs could also be replaced by a sentence of the form ``This [claim/lemma] follows from the definitions of\ldots''.
+
+However, when discussing the proof of Theorem~3 with colleagues (very familiar with product structure) recently, we found them unconvinced by high-level explanations.
+It was only after carefully unpacking the definitions that they became convinced.
+These conversations usually end with someone saying ``Ah, of course!''
 

### 2022-03-16 k-planar.tex
Noticed a missing definition, to fix later
Much simpler.....

@@
 
-We begin with a standard technique that allows us to work with a normalised tree-decomposition:
+We begin with a standard technique that allows us to work with a normalised tree-decomposition whose tree has same vertex set as the graph it decomposes:
 
@@
 %\subsection{Generalized Tripod Partitions}
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics{figs/tripoddo}
-  \end{center}
-  \caption{The sets $Y_x$, $F_x$, and $V_x$ associated with $x\in V(T)$
-  and the ancestors $a_1,\ldots,a_{t'}$ of $X$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.}
-  \label{fig:generalized-tripod}
-\end{figure}
-
-The following lemma shows how to interpret an $H$-partition of $G$ and a tree-decomposition of $H$ as a `hierarchical' decomposition of $G$; refer to \cref{fig:generalized-tripod}.
-
-\begin{lem}\label{generalized-tripod}
-  Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
-For each $x\in V(T)$, let $V_x := \bigcup_{y\in V(T_x)} Y_y$, $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}$, and $N_x:=V_x\cup F_x$.
-Then,
-  \begin{compactenum}[(Y1)]
-    % \item\ylabel{thickness} $\mathcal{Y}=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
-    \item\ylabel{separator} For each $x\in V(T)$, there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$.
-    \item\ylabel{ancestor-edge} For each $x\in V(T)$, there is a set $\{a_1,\ldots,a_{t'}\}$ of $t'\le t$ strict $T$-ancestors of $x$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
-  \end{compactenum}
-\end{lem}
-
-\referee{1}{1.
+% 
+% 
+% \begin{figure}[htbp]
+%   \begin{center}
+%     \includegraphics{figs/tripoddo}
+%   \end{center}
+%   \caption{The sets $Y_x$, $F_x$, and $V_x$ associated with $x\in V(T)$
+%   and the ancestors $a_1,\ldots,a_{t'}$ of $X$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.}
+%   \label{fig:generalized-tripod}
+% \end{figure}
+
+% The following lemma shows how to interpret an $H$-partition of $G$ and a tree-decomposition of $H$ as a `hierarchical' decomposition of $G$; refer to \cref{fig:generalized-tripod}.
+
+% For a vertex $v$ of $G$, $N_G(v):=\{w\in V(G):vw\in E(G)\}$ and $N_G[v]:=N_G(v)\cup\{v\}$.  For a vertex set $S\subseteq V$, $N_G(S):=\bigcup_{v\in S} N_G(S)\setminus S$ and $N_G[S]:=N_G(S)\cup S$.
+
+% \note{PM}{
+% Guide to old notation:
+% \begin{compactitem}
+%   \item $V_x := \bigcup_{y\in V(T_x)} Y_y = G_x$
+%   \item $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}= N_G(G_x)$
+%   \item $N_x:=V_x\cup F_x=N_G[G_x]$
+% \end{compactitem}
+% }
+% 
+% \begin{lem}\label{generalized-tripod}
+%   Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
+%   For each $x\in V(T)$, let $H_x:=V(T_x)$ and let $G_x:=\bigcup_{y\in H_x}Y_y$.
+% 
+%   \begin{compactenum}[(Y1)]
+%     % \item\ylabel{thickness} $\mathcal{Y}=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
+%     % \item\ylabel{separator} For each $x\in V(T)$, $N_G(\phi(T_x))$ separates $\phi(T_x)$ from $N
+%     % there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$.
+% 
+%     \item\ylabel{ancestor-edge} For each $x\in V(T)$, there is a set $\{a_1,\ldots,a_{t'}\}$ of $t'\le t$ strict $T$-ancestors of $x$ such that $N_G(G_x) \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
+%   \end{compactenum}
+% \end{lem}
+% 
+% \begin{proof}
+%   Let $\{a_1,\ldots,a_{t'}\}:= B_x\setminus\{x\}$.
+Then, by the definition of a normalized tree decomposition $N_H(H_x))\subseteq\{a_1,\ldots,a_{t'}\}$.
+Therefore $N_G(G_x)\subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
+% \end{proof}
+
+% \referee{1}{1.
 Lemma 3: This lemma is really about ``normalised'' tree-decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$.
 Call this property (T3) and remove (Y1) through (Y5), which should now be obvious.}
-
-\note{PM}{I've added to the \texttt{response.tex} file explaining some of this.
+% 
+% \note{PM}{I've added to the \texttt{response.tex} file explaining some of this.
 Each of (Y1), (Y3), (Y4), and (Y5) have a one-sentence justification.
 The justification for (Y2) is just definition unpacking and then using the observation that the referee recommends.}
-
-\note{PM}{I plan to do an experiment with the referee's suggested notation, in a separate git branch to see how it turns out.}
-
-Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:
-
-\begin{compactenum}[(Y1)]\setcounter{enumi}{2}
-  \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$.
-  \item\ylabel{containment-i} $V_x\subseteq V_a$ for every $T$-ancestor $a$ of $x$.
-  \item\ylabel{containment-ii}$N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
-\end{compactenum}
-
-Property~\yref{y-subsets} follows from the fact that $V_x$ is the union of several sets, one of which is $Y_x$.
+% 
+% \note{PM}{I plan to do an experiment with the referee's suggested notation, in a separate git branch to see how it turns out.}
+% 
+% Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:
+% 
+% \begin{compactenum}[(Y1)]\setcounter{enumi}{2}
+%   \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$. \note{PM}{$Y_x$, for $x\in V(T)$ is not defined!}
+%   \item\ylabel{containment-i} $V_x\subseteq V_a$ for every $T$-ancestor $a$ of $x$.
+%   \item\ylabel{containment-ii}$N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
+% \end{compactenum}
+
+% Property~\yref{y-subsets} follows from the fact that $V_x$ is the union of several sets, one of which is $Y_x$.
 Property~\yref{containment-i} follows from the definition of $V_x$ and the fact that $V(T_x)\subseteq V(T_a)$.
@@
 Note that none of \yref{y-subsets}--\yref{containment-ii} depends on \yref{ancestor-edge} (which is important, since \yref{containment-i} is used to establish \yref{ancestor-edge} in the following proof).
-
-\begin{proof}[Proof of \cref{generalized-tripod}]
-  % Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
-  Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
-
-  To establish Property~\yref{ancestor-edge}, consider some vertex $w\in F_x$.
+% 
+% 
+% \begin{proof}[Proof of \cref{generalized-tripod}]
+%   % Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
+%   Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
+% 
+%   To establish Property~\yref{ancestor-edge}, consider some vertex $w\in F_x$.
 Since $w\in F_x$, there exists an edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.
@@
 In particular, $a$ is contained in $B_x$.
-  Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $B_x$ contains $x$.
-\end{proof}
+%   Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $B_x$ contains $x$.
+% \end{proof}
 
@@
 
-\referee{1}{2. Proof of Theorem 9: There is too much notation throughout. As
-	one example, it is hard to remember which of $T_x$, $Y_x$, $V_x$, $F_x$, $N_x$,
-	$S_x$, $X_v$, $B_x$, $C_x$ is which. If $\phi(S)$ is written for the subset of $V(G)$
-	corresponding to a set $S \subseteq V(H)$ and appropriate shorthand and
-	standard notation for neighbourhoods are used, then $Y_x$, $V_x$, $F_x$, and
-	$N_x$ can be replaced by $\phi(x)$, $\phi(T_x)$, $N(\phi(T_x))$, and $V(G) \setminus N[\phi(T_x)]$.
-	The set $X_v$ is only used in the statement of Claim~1, so it does not
-	need a name; say what these vertices are explicitly. It may be more
-	indicative to write $\phi'(x)$ for $S_x$ and $B'_x$ for $C_x$.}
+% \referee{1}{2. Proof of Theorem 9: There is too much notation throughout. As
+% 	one example, it is hard to remember which of $T_x$, $Y_x$, $V_x$, $F_x$, $N_x$,
+% 	$S_x$, $X_v$, $B_x$, $C_x$ is which. If $\phi(S)$ is written for the subset of $V(G)$
+% 	corresponding to a set $S \subseteq V(H)$ and appropriate shorthand and
+% 	standard notation for neighbourhoods are used, then $Y_x$, $V_x$, $F_x$, and
+% 	$N_x$ can be replaced by $\phi(x)$, $\phi(T_x)$, $N(\phi(T_x))$, and $V(G) \setminus N[\phi(T_x)]$.
+% 	The set $X_v$ is only used in the statement of Claim~1, so it does not
+% 	need a name; say what these vertices are explicitly. It may be more
+% 	indicative to write $\phi'(x)$ for $S_x$ and $B'_x$ for $C_x$.}
 
 \begin{proof}
-Apply \cref{generalized-tripod} to $G$ and let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $Y_x$, $V_x$, $F_x$, and $N_x$ be defined as in \cref{generalized-tripod}, where the partition $\mathcal{Y}$ has width $\ell$ with respect to the layering $\mathcal{L}$.
-
-For a node $x\in V(T)$, we say that a shortcut $P\in\SS$ \defin{crosses} $x$ if $Y_x$ contains an internal vertex of $P$, that is, $P=(v_0,\ldots,v_r)$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
+Let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
+% For each $x\in V(T)$, let $H_x:=V(T_x)$ and let $G_x:=\bigcup_{y\in H_x}Y_y$.
+
+% Apply \cref{generalized-tripod} to $G$ and let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $H_x$, and $G_x$ be defined as in \cref{generalized-tripod}, where the partition $\mathcal{Y}$ has width $\ell$ with respect to the layering $\mathcal{L}$.
+
+For a node $x\in V(T)$, we say that a shortcut $P\in\SS$ \defin{crosses} $x$ if $Y_x$ contains an internal vertex of $P$.
+In other words, if $P=v_0,\ldots,v_r$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
 We say that a vertex $v\in V(G)$ \defin{participates} in $x$ if $v\in Y_x$, or $\SS$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
-We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
-
-\begin{clm}\label{x-v-ancestor}
-  For any $v\in V(G)$ there exists a (unique) node $a(v)\in X_v$ such that
-  $a(v)$ is a $T$-ancestor of every node in $X_v$.
-\end{clm}
-
-\begin{proof}
-  Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\SS, v\in \{v_0,\ldots,v_r\}\}$.
+For each $v\in V(G)$, let $a(v)$ denote the lowest common $T$-ancestor of all the nodes $x\in V(T)$ in which $v$ participates.
+
+% \begin{clm}
+%   For each $v\in V(G)$, $v$ participates in $a(v)$.
+% \end{clm}
+% 
+% \begin{proof}
+%   Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\SS, v\in \{v_0,\ldots,v_r\}\}$.
 Then $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
-
-  We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
+% 
+%   We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
 If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\SS$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
 In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$, $(v_0,\ldots,v_r)\in\SS$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\SS$ that crosses $x$.
-
-  Let $X_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.  The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.
-  %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
-  Choose $a(v)\in X_H$ to be the member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.
+% 
+%   Let $X_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.
+Therefore $X_H$ contains every every node in which $v$ participates.
+The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.
+The fact that $\mathcal{T}$ is a normalized tree decomposition of $H$ implies that some node $a\in X_H$ is a $T$-ancestor of every node in $X_H$.
+Thus, $v$ participates in the node $a=a(v)$ that is the lowest common $T$-ancestor of every node in which $v$ participates, as required.
+% \end{proof}
+
+% \begin{proof}
+%   Let $a(v)$.
+If $v\in Y_{a(v)}$ then $v$ participates in $a(v)$, by definition, and we are done.
+Otherwise, there are two nodes $x,y\in V(T)$ in which $x$ participates and whose lowest common $T$-ancestor is $a(v)$.
+% 
+%   There are two cases to consider:  
+%   \begin{compactenum}
+%     \item If $v\in Y_x$ then $v\not\in Y_y$.
+Since $v$ participates in $Y_y$, there exist a shortcut $P:=v_0,\ldots,v_r$ in $\SS$ with $v=v_\alpha$ and $v_\beta\in Y_y$ for some $\alpha\in\{0,\ldots,r\}$ and $\beta\in\{1,\ldots,r-1\}$.
+Without loss of generality, $\alpha < \beta$.
+% 
+%     For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.
+Since $G[\{v_\alpha,\ldots,v_\beta\}]$ is connected, $H[\{x_\alpha,\ldots,x_\beta\}]$ connected.
+Since $\mathcal{T}$ is a normalized tree decomposition of $H$, $x_j$ is a $T$-ancestor of every node in $x_\alpha,\ldots,x_\beta$ for some $j\in\{\alpha,\ldots,\beta\}$.
+In particular $x_j$ is a $T$-ancestor of $x=x_\alpha$ and $y=x_\beta$, so $x_j$ is a $T$-ancestor of $a(v)$.
+Since $v_j\in V(P)$ and $v_j\in Y_{x_j}$ this implies that $v$ participates in $x_j$.
+Therefore $v$ participates in a $T$-ancestor of $a(v)$, which can only mean that $v$ participates in $a(v)$.
+% 
+%     \item If $v\not\in Y_x$ and $v\not\in Y_y$ then there are shortcuts $P_1,P_2\in\SS$ with $v\in V(P_1)$, $v\in V(P_2)$ where $P_1$ has an internal vertex $w_1\in Y_x$ and $P_2$ has an internal vertex $w_2\in Y_y$.
+% 
+% 
+%     .
+Since $v$ participates in $Y_y$, there exist a shortcut $v_0,\ldots,v_r$ in $\SS$ with $v=v_\alpha$ and $v_\beta\in Y_y$ for some $\alpha\in\{0,\ldots,r\}$ and $\beta\in\{1,\ldots,r-1\}$.
+Without loss of generality, $\alpha < \beta$.
+% 
+%     For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.
+Since $G[\{v_\alpha,\ldots,v_\beta\}]$ is connected, $H[\{x_\alpha,\ldots,x_\beta\}]$ connected.
+Since $\mathcal{T}$ is a normalized tree decomposition of $H$, $x_j$ is a $T$-ancestor of every node in $x_\alpha,\ldots,x_\beta$ for some $j\in\{\alpha,\ldots,\beta\}$.
+In particular $x_j$ is a $T$-ancestor of $x=x_\alpha$ and $y=x_\beta$, so $x_j$ is a $T$-ancestor of $a(v)$.
+Since $v_j\in Y_{x_j}$, and $v\not\in Y_{a(v)}$ this implies that $v$ participates in $Y_{x_j}$.
+Therefore $v$ participates in a $T$-ancestor of $a(v)$, which means that $v$ participates in $a(v)$.
+% 
+% 
+%   \end{compactenum}
+% \end{proof}
+
+% We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
+
+% \begin{clm}\label{x-v-ancestor}
+%   For any $v\in V(G)$ there exists a (unique) node $a(v)\in X_v$ such that
+%   $a(v)$ is a $T$-ancestor of every node in $X_v$.
+% \end{clm}
+% 
+% \begin{proof}
+%   Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\SS, v\in \{v_0,\ldots,v_r\}\}$.
+Then $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
+% 
+%   We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
+If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\SS$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
+In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$, $(v_0,\ldots,v_r)\in\SS$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\SS$ that crosses $x$.
+% 
+% 
+%   Let $X_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.  The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.  This implies that $T':=T[]
+% 
+% 
+% 
+%   %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
+%   Choose $a(v)\in X_H$ to be the member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.
 Transitivity of the $T$-ancestor relationship, \tref{ancestor-edge}, and connectivity of $H[X_H]$ implies that such an $a(v)$ exists and is a $T$-ancestor of every node $x\in X_H$, as required.
-\end{proof}
+% \end{proof}
 
@@
 For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
-Observe that $\PP:=(S_x : x\in V(T))$ is a partition of $V(G)$.\footnote{The sets in $\PP$ are disjoint because $a:V(G)\to V(T)$ is a function.
-The sets in $\PP$ cover $V(G)$ since, for each $v\in V(G)$, $v\in S_{a(v)}$.} Let $J:=G^\SS/\PP$ denote the resulting quotient graph.
+Observe that $\PP:=(S_x : x\in V(T))$ is a sequence of pairwise disjoint sets that cover $V(G)$.
+Let $J:=G^\SS/\PP$ denote the resulting quotient graph.
 We consider $V(J)\subseteq V(T)$, where each $x\in V(J)$ is the vertex obtained by contracting $S_x$ in $G^{\SS}$.
@@
 
-First we need to understand the relationship between sets $S_x\in\PP$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
-
-\referee{1}{8. general comment: Claims 1, 2, and 4 are especially belaboured by
-notation as they are quite straightforward.}
-
-\note{PM}{Same comment: Yes, these claims are definition checking.  We'll see if the experimental notation change helps.}
-
-\begin{clm}
-	\label{s-subset}
-  For every $x\in V(T)$, $S_x\subseteq V_x$.
-\end{clm}
-
-\begin{proof}
-  For the sake of contradiction, assume otherwise, so there exists some $v\in S_x\setminus V_x$.
-By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.
-Therefore, $\SS$ contains a path $P$, with $v\in V(P)$, that crosses $x$.
-The path $P$ contains a proper subpath $v_0,v_1,\ldots,v_{r}$ such that $v=v_0$ and $v_r\in Y_x$.
-Since $v\not\in V_x$ and $v_r\in Y_x\subseteq V_x$, \yref{separator}, implies that $v_i\in F_x$ for some $i\in\{0,\ldots,r-1\}$.
-Now \yref{ancestor-edge} implies $v_i\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
-Therefore, either $v\in Y_a$ or $P$ crosses $a$.
-But this implies that $a(v)$ is a $T$-ancestor of $a$, which is a strict $T$-ancestor of $x$, contradicting the assumption that $v\in S_x$.
-\end{proof}
-
-Next we complete Step~(i) and show that $\PP$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
+% First we need to understand the relationship between sets $S_x\in\PP$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
+
+% \referee{1}{8. general comment: Claims 1, 2, and 4 are especially belaboured by
+% notation as they are quite straightforward.}
+
+% \note{PM}{Same comment: Yes, these claims are definition checking.  We'll see if the experimental notation change helps.}
+
+% \begin{clm}
+% 	\label{s-subset}
+%   For every $x\in V(T)$, 
+%   % $S_x\subseteq V_x$.
+%   $S_x\subseteq \bigcup_{x'\in V(T_x)} Y_{x'}$.
+% \end{clm}
+% 
+% \begin{proof}
+%   Let $v$ be any vertex in $S_x$ (which implies that $a(v)=x$) and let $Y_y$ be the part of $\mathcal{Y}$ that contains $v$.
+By definition, $v$ participates in $y$ so by definition $a(v)$ is a $T$-ancestor of $y$, so $y\in V(T_{a(v)})$.
+Therefore, $v\in Y_y\subseteq \bigcup_{x'\in V(T_x)} Y_{x'}$, as required.
+% \end{proof}
+
+% Next we complete Step~(i) and show that $\PP$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
 
@@
 \begin{proof}
-  Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.
+  If $xy\in E(J)$ then there exists a shortcut $P:=v_0,\ldots,v_r\in\SS$ with endpoints $v_0\in S_x$ and $v_r\in S_y$.\footnote{Recall that we have made the assumption that $\SS$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part in $\mathcal{T}$ that contains $v_i$.
+Since $P$ is a connected subgraph of $G$, $H[\{x_0,\ldots,x_r\}]$ is a connected subgraph of $H$.
+Since $\mathcal{T}$ is a normalized tree decomposition of $H$, some node $x_i\in V(T)$ is a $T$-ancestor of all nodes in $x_0,\ldots,x_r$.
+
+  We claim that at least one of $v_0$ or $v_r$ participates in $x_i$.
+If $i=0$ then $v_0\in Y_{x_0}=Y_{x_i}$ so $v_0$ participates in $x_i$.
+Similarly, if $i=r$ then $v_r$ participates in $x_r$.
+Otherwise $i\in\{1,\ldots,r-1\}$ so $v_i$ is an internal vertex of the shortcut $P$, so $P$ crosses $x_i$.
+Since $v_0,v_r\in V(P)$, this implies that $v_0$ and $v_r$ both participate in $x_i$.
+
+  Suppose, without loss of generality, that $v_0$ participates in $x_i$.
+Then $a(v_0)=x$ is a $T$-ancestor of $x_i$ which a $T$-ancestor of $x_r$.
+Finally $a(v_r)=y$ is a $T$-ancestor of $x_r$.
+Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so one is a $T$-ancestor of the other.
+  % 
+  % 
+  % is a lazy walk in $H$.
+  % 
+  % Let $P:=v_0,\ldots,v_r$ (where $v_0=v$ and $v_r=w$)
+  % Suppose, for the sake of contradiction that neither $x$ nor $y$ is an ancestor of the other.  Then
+  % 
+  % 
+  % 
+  % 
+  % 
+  % 
+  % If $P$ has length $1$ (so $vw\in E(G)$) then $
+  % 
+  % 
+  % 
+  % By \cref{s-subset}, $v\in G_x$ and $w\in G_y$.
+  % 
+  % 
+  % 
+  % 
+  % Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.
 Since $xy\in E(J)$, $G^\SS$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
-Since $vw\in E(G^{\SS})$,  $\SS$ contains a $vw$-shortcut $P$.\footnote{Recall that we have made the assumption that $\SS$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  By \cref{s-subset}, $v\in V_x$ and $w\in V_y$.
+Since $vw\in E(G^{\SS})$,  $\SS$ contains a $vw$-shortcut $P$.
+By \cref{s-subset}, $v\in V_x$ and $w\in V_y$.
 By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
@@
 
-\referee{1}{10. page 10, Claim 5: Indices could mostly be removed by continuing
-to talk about $T$-ancestors instead. Move the definition of $H+$ and its
-directed version up, along with the explanation of what will be proven,
-so that it isn’t necessary to define the $s_i$.}
-
-\note{PM}{I also found the steps in this proof tedious.
+% \referee{1}{10. page 10, Claim 5: Indices could mostly be removed by continuing
+% to talk about $T$-ancestors instead. Move the definition of $H+$ and its
+% directed version up, along with the explanation of what will be proven,
+% so that it isn’t necessary to define the $s_i$.}
+% 
+% \note{PM}{I also found the steps in this proof tedious.
 Unfortunately, when I tried to give a simple explanation for it in Banff, both Piotr and Torsten remained unconvinced until we opened the paper and walked through the steps of the proof together.
@@
 Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $J$.
-It remains to bound the size of each bag $C_x$.
-
-  Consider an arbitrary node $x\in V(T)$ where $x_0,\ldots,x_r$ is the path from the root $x_0$ of $T$ to $x_r:=x$.
-To avoid triple-subscripts in what follows, we abuse notation slightly by using $V_i$, $F_i$, and $N_i$,  as shorthands for $V_{x_i}$, $F_{x_i}$ and $N_{x_i}$, respectively.
-
-  If $x_\delta\in C_x$, it is because $x_\delta x'\in E(J)$ for some $T$-descendant $x'$ of $x$.
-This implies $G^{\SS}$ contains an edge $vw$ with $v\in S_{x'}$ and $w\in S_{x_\delta}=S_\delta$.
-This implies that $\SS$ contains a $vw$-shortcut $P_{vw}$.
-Let $v'$ be the second-last vertex of $P_{vw}$ (so $v'w\in E(G)$).
-
-  Since $w\in S_{\delta}$, $a(w)=x_\delta$, so $w$ participates in $x_\delta$, so at least one of the following is true:
-  \begin{compactenum}
-    \item There exists $w'\in V(G)$ such that $\SS$ contains a $ww'$-shortcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$; or
-    \item $w\in Y_\delta$; in this case, define $w':=w$ and let $P_{ww'}$ be the 1-vertex path $w$.
-  \end{compactenum}
-  Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{\delta}$.
-
-  Let $w_0,w_1,\ldots,w_p$ be the path that begins $w_0:=v'$ and then follows the subpath of $P_{ww'}$ that begins at $w_1:=w$ and ends at $w_p:=w''$.
-For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{0,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.
-Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root, $x_0$, of $T$ is non-increasing.
-
-  We claim that $a_0=x_r$.
-Since $v\in S_{x'}$, $a(v)=x'$, so $V(P_{vw})\subseteq V_{x'}$, otherwise $v$ participates in $a$ for some node $a$ not in the subtree $T_{x'}$ rooted at $x'$, but this contradicts \cref{i-ancestor} since $a(v)=x'$ is a $T$-ancestor of every node in which $v$ participates.
-Since $v'=w_0\in V(P_{vw})$, $\{w_0\}\subseteq V_{x'}\subseteq V_{r}$, so $s_0=r$ and $a_0=x_r$.
-
-  We claim that $a_p=x_\delta$---that is, $s_p=\delta$.
-  To see this, first observe that, for each $i\in\{1,\ldots,p\}$, $w_i\in V_{\delta}$ since, otherwise, an internal vertex of $P_{ww'}$ belongs to $F_\delta$, which would imply (by \yref{ancestor-edge}) that $w\in S_{\delta'}$ for some $\delta' < \delta$, contradicting the assumption that $w\in S_\delta$.
-Therefore $s_p\ge\delta$.
-To see that $s_p<\delta+1$,
-  observe that either $w=w''\in Y_\delta$ or $P_{ww'}$ contains an internal vertex $w''$ in $Y_\delta$.  By the definition of $V_x$, $V_{\delta-1}$ does not contain $w''$, so $s_p<\delta+1$.
-
-  Let $H^+$ denote the supergraph of $H$ with vertex set $V(T)$ and in which $xy\in E(H^+)$ if and only there exists some $z\in V(T)$ such that $x,y\in B_z$.
-  % Note that, by \tref{ancestor-edge}, $xy\in E(H^+)$ if and only if $x\in B_y$ or $y\in B_x$.
-  We claim that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \defin{lazy walk} in a graph $H$ is a walk in the pseudograph $H'$ obtained by adding a loop at each vertex of $H$.} in $H^+$.
-Indeed, if $a_i\neq a_{i+1}$ for some $i\in\{0,\ldots,p-1\}$ then this is precisely because $w_i\in V_{a_i}$ but $w_{i+1}\not\in V_{a_i}$.
-By definition, $w_i\in Y_{a_i'}$ for some $T$-descendant $a_i'$ of $a_i$.
-  By \yref{separator}, $w_{i+1}\in F_{a_i}$ so by \yref{ancestor-edge} $w_{i+1}\in Y_{a_i''}$ for some strict $T$-ancestor $a_i''$ of $a_i$.
-Since $w_iw_{i+1}\in E(G)$, $a_i'a_i''\in E(H)$.
-By \tref{subtree-root}, $a_i''\in B_{a_i''}$ and $a_i''\in B_{a_i'}$.
-Since $a_i$ is on the path from $a_i'$ to $a_i''$ in $T$ this implies that $a_i''\in B_{a_i}$.
-Therefore $a_ia_i''\in E(H^+)$ as claimed.
-
-% \referee{2}{Page 18.  Say $H'$ is connected since $X$ is connected.}
-% \note{PM}{This is in the wrong place. It refers to the proof about centered colouring, which is now just a reference.}
-
-  Thus, $a_0,\ldots,a_p$ is a lazy walk in $H^+$ of length $p\le k$ where the distance $s_i$ between $a_i$ and the root $x_0$ of $T$ is non-decreasing.
-By removing repeated vertices this gives a path in the directed graph $\overrightarrow{H}^+$ obtained by directing each edge $xy\in E(H^+)$ from its $T$-descendant $x$ towards its $T$-ancestor $y$.
-  Finally, we are in a position to appeal to \cite[Lemma~13]{PS21} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
-
-  % \note{DW}{replace Lemma~24 by Lemma~13, right?}
-  % \note{PM}{Right. Changed.}
-
-  At this point, the proof is complete, but let us summarize.
-For each node $x\in V(T)$, $C_x$ contains only $T$-ancestors of $x$ (including $x$ itself).
-For each ancestor $x_\delta$ of $x$ contained in $C_x$, there is path from $x$ to $x_\delta$ in $\overrightarrow{H}^+$ of length at most $k$.
-The number of ancestors of $x$ that can be reached by paths of length at most $k$ in $\overrightarrow{H}^+$ is at most $\binom{k+t}{t}$.
-Therefore $|C_x|\le \binom{k+t}{t}$, as required.
+
+  It remains to show that each bag in our tree decomposition has size at most $\binom{k+t}{t}$.  We will do this by appealing to an elegant result of \citet{PS21}. 
+  % on the number of $k$-reachable vertices. in a directed graph obtained by directing the edges of a supergraph of $H$.  \
+  Specifically, let $H^+$ denote the supergraph of $H$ that contains the edge $xy$ if and only if some bag $B_z\in\mathcal{T}$ contains both $x$ and $y$ (i.e., $H^+$ is the chordal closure of $H$ with respect to $\mathcal{T}$).
+Let $\overrightarrow{H}^+$ be the directed graph obtained by directing each edge $xy$ of $H^+$ in the direction $\overrightarrow{xy}$ so that $y$ is a $T$-ancestor of $x$.
+\citet[Lemma~13]{PS21} show that, for any $x\in V(\overrightarrow{H}^+)$, the number of vertices of $\overrightarrow{H}^+$ that can be reached by a directed path that originates at $x$ and has length at most $k$ is at most $\binom{k+t}{t}$.
+
+  Consider an arbitrary node $x\in V(T)$ and suppose that some strict $T$-ancestor $a$ of $x$ is contained in $C_x$.
+We will show that $\overrightarrow{H}^+$ contains a directed path from $x$ to $a$ of length at most $k$.
+Since there $a$ a strict $T$-ancestor of $x$, $a\neq x$ so any path from $x$ to $a$ has length at least $1$.
+By \cite[Lemma~13]{PS21}, this implies that the number of strict $T$-ancestors of $x$ contained in $C_x$ is at most $\binom{k+t}{t}-1$ and therefore $|C_x|\le\binom{k+t}{t}$.
+
+  Since $a\in C_x$, there exists a $T$-descendant $x'$ of $x$ such that $ax'\in E(J)$.
+Therefore, there exists a shortcut $v_0,\ldots,v_r$ in $\SS$ with $v_0\in S_{x'}$ and $v_r\in S_a$.
+For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.
+Since $v_0,\ldots,v_r$ is a path in $G$, $x_0,\ldots,x_r$ is a lazy walk in $H$.\footnote{A \defin{lazy walk} is a graph $G$ is a walk in the pseudograph obtained by adding a self-loop at each vertex of $G$.}  Let $j$ be the minimum integer such that $x_j$ is a strict $T$-ancestor of $x$.
+Since $x'=x_0$ is a $T$-descendant of $x$ and $a=x_r$ is a strict $T$-ancestor of $x$, $j\in\{1,\ldots,r\}$ and $x_{j-1}$ is a $T$-descendant of $x$ (possibly $x_{j-1}=x$).
+Since $x_{j-1}x_j\in E(H)$, $x_j \in B_{x_{j-1}}$ and therefore $x_j\in B_x$.
+Therefore $xx_j\in E(H^+)$.
+
+  Therefore $W:=y_0,\ldots,y_s:=x,x_j,x_{j+1},\ldots,x_r$ is a lazy walk in $H^+$ of length $r\le k$.
+For each $j\in \{0,\ldots,s\}$, let $z_j$ be the lowest common $T$-ancestor of $y_0,\ldots,y_j$.
+By using the same reasoning used to conclude that $xx_j\in E(H^+)$ we conclude that $z_{i-1}=z_i$ or $z_{i-1}z_i\in E(H^+)$ for each $i\in\{1,\ldots,s\}$, so $z_0,\ldots,z_s$ is a directed lazy walk in $\overrightarrow{H}^+$.
+Finally, let $\overrightarrow{W}$ be the path in $\overrightarrow{H}^+$ obtained by removing removing duplicates from $z_0,\ldots,z_s$.
+Then $\overrightarrow{W}$ is a directed path in $\overrightarrow{H}^+$ from $x$ to $a$ of length $s\le r\le k$.
+This completes the proof.
 \end{proof}
@@
 
-\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{framed} with the method in \citep{DHHW}.
+\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{d_framed_product_stucture} with the method in \citep{DHHW}.
 I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).

### 2022-03-16 k-planar.tex
Added anote
Suggestion

@@
 
+\note{PM}{I have significantly simplified this section, which should make the referees happy. I will read it very carefully again tomorrow}
+
 We begin with a standard technique that allows us to work with a normalised tree-decomposition whose tree has same vertex set as the graph it decomposes:
@@
 Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that $\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}$ for each $v\in V(G)$.
+
+\note{PM}{Consider a change of notation:   Replace $F$ with $C$.  (I was using $F$ before to be consistent with the previous section, but that notation is gone now.)}
 

### 2022-03-17 k-planar.tex
Fixed wording

@@
 
+\note{PM}{I've changed my mind about this notation change.
+We end up applying the lemma inductively on the faces $F_1,\ldots,F_m$ of $M$ (which are cycles in $G$).
+It makes sense that the original cycle in $G$ is called $F$.}
+
 \begin{lem}
@@
     \item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
-		\item Let $F$ be a cycle in $G'$ that $V(F)$ is partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
+    \item Let $F$ be a cycle in $G'$ whose vertices are partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
     \begin{compactenum}

### 2022-03-17 k-planar.tex
Cleaned up and filled in an important missing detail

@@
 The purpose of this section is to prove our main technical result, \cref{ShortcutPartition}.
-This theorem shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, a $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
+This theorem shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, an $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
 
@@
 
-\note{PM}{I have significantly simplified this section, which should make the referees happy. I will read it very carefully again tomorrow}
-
-We begin with a standard technique that allows us to work with a normalised tree-decomposition whose tree has same vertex set as the graph it decomposes:
+% \note{PM}{I have significantly simplified this section, which should make the referees happy. I will read it very carefully again tomorrow}
+
+We begin with a standard technique that allows us to work with a normalised tree-decomposition whose tree has the same vertex set as the graph it decomposes:
 
@@
 By the definition of $f$, the tree-decomposition $(B_x:x\in V(T))$ satisfies \tref{subtree-root}.
-  % To see that $(B_x:x\in V(T))$ satisifies \tref{ancestor-edge}, observe that,
-  % if $xy\in E(H)$, then at least one of $x$ or $y$ is contained in $B_z$ for every node $z$ on the path from $x$ to $y$ in $T$.
-If neither $x$ nor $y$ is an ancestor of the other, then some node $z$ on this path has $T$-depth less than that of $x$ and $y$.
-If $x\in B_z$ this contradicts the fact that $x$ is the root of $T[x]$.
-If $y\in B_z$ this contradicts the fact that $y$ is the root of $T[y]$.
 \end{proof}
@@
 
-%\subsection{Generalized Tripod Partitions}
-%
-%
-% \begin{figure}[htbp]
-%   \begin{center}
-%     \includegraphics{figs/tripoddo}
-%   \end{center}
-%   \caption{The sets $Y_x$, $F_x$, and $V_x$ associated with $x\in V(T)$
-%   and the ancestors $a_1,\ldots,a_{t'}$ of $X$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.}
-%   \label{fig:generalized-tripod}
-% \end{figure}
-
-% The following lemma shows how to interpret an $H$-partition of $G$ and a tree-decomposition of $H$ as a `hierarchical' decomposition of $G$; refer to \cref{fig:generalized-tripod}.
-
-% For a vertex $v$ of $G$, $N_G(v):=\{w\in V(G):vw\in E(G)\}$ and $N_G[v]:=N_G(v)\cup\{v\}$.  For a vertex set $S\subseteq V$, $N_G(S):=\bigcup_{v\in S} N_G(S)\setminus S$ and $N_G[S]:=N_G(S)\cup S$.
-
-% \note{PM}{
-% Guide to old notation:
-% \begin{compactitem}
-%   \item $V_x := \bigcup_{y\in V(T_x)} Y_y = G_x$
-%   \item $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}= N_G(G_x)$
-%   \item $N_x:=V_x\cup F_x=N_G[G_x]$
-% \end{compactitem}
-% }
-%
-% \begin{lem}\label{generalized-tripod}
-%   Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
-%   For each $x\in V(T)$, let $H_x:=V(T_x)$ and let $G_x:=\bigcup_{y\in H_x}Y_y$.
-%
-%   \begin{compactenum}[(Y1)]
-%     % \item\ylabel{thickness} $\mathcal{Y}=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
-%     % \item\ylabel{separator} For each $x\in V(T)$, $N_G(\phi(T_x))$ separates $\phi(T_x)$ from $N
-%     % there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$.
-%
-%     \item\ylabel{ancestor-edge} For each $x\in V(T)$, there is a set $\{a_1,\ldots,a_{t'}\}$ of $t'\le t$ strict $T$-ancestors of $x$ such that $N_G(G_x) \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
-%   \end{compactenum}
-% \end{lem}
-%
-% \begin{proof}
-%   Let $\{a_1,\ldots,a_{t'}\}:= B_x\setminus\{x\}$.
-Then, by the definition of a normalized tree decomposition $N_H(H_x))\subseteq\{a_1,\ldots,a_{t'}\}$.
-Therefore $N_G(G_x)\subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
-% \end{proof}
-
-% \referee{1}{1.
-Lemma 3: This lemma is really about ``normalised'' tree-decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$.
-Call this property (T3) and remove (Y1) through (Y5), which should now be obvious.}
-%
-% \note{PM}{I've added to the \texttt{response.tex} file explaining some of this.
-Each of (Y1), (Y3), (Y4), and (Y5) have a one-sentence justification.
-The justification for (Y2) is just definition unpacking and then using the observation that the referee recommends.}
-%
-% \note{PM}{I plan to do an experiment with the referee's suggested notation, in a separate git branch to see how it turns out.}
-%
-% Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:
-%
-% \begin{compactenum}[(Y1)]\setcounter{enumi}{2}
-%   \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$. \note{PM}{$Y_x$, for $x\in V(T)$ is not defined!}
-%   \item\ylabel{containment-i} $V_x\subseteq V_a$ for every $T$-ancestor $a$ of $x$.
-%   \item\ylabel{containment-ii}$N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
-% \end{compactenum}
-
-% Property~\yref{y-subsets} follows from the fact that $V_x$ is the union of several sets, one of which is $Y_x$.
-Property~\yref{containment-i} follows from the definition of $V_x$ and the fact that $V(T_x)\subseteq V(T_a)$.
-To show Property~\yref{containment-ii} first note that, by \yref{containment-i} it suffices to consider vertices $w\in F_x=N_x\setminus V_x$.
-By definition, every vertex $w\in F_x$ is adjacent, in $G$, to a vertex $v\in V_x$.
-By \yref{containment-i}, $v\in V_a$, so $w$ is either in $V_a$ or $w$ satisfies the condition $vw\in E(G)$, $v\in V_a$, and $w\not\in V_a$, so $w\in F_a$.
-In either case $w\in N_a=V_a\cup F_a$.
-Note that none of \yref{y-subsets}--\yref{containment-ii} depends on \yref{ancestor-edge} (which is important, since \yref{containment-i} is used to establish \yref{ancestor-edge} in the following proof).
-%
-%
-% \begin{proof}[Proof of \cref{generalized-tripod}]
-%   % Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
-%   Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
-%
-%   To establish Property~\yref{ancestor-edge}, consider some vertex $w\in F_x$.
-Since $w\in F_x$, there exists an edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.
-Since $v\in V_x$, $v\in Y_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
-Since $\mathcal{Y}$ is a partition, $w\in Y_{a}$ for some $a\not\in V(T_x)$.
-Since $vw\in E(G)$, we have $x'a\in E(H)$.
-By \tref{ancestor-edge}, one of $a$ or $x'$ is a $T$-ancestor of the other.
-Since $w\in Y_a\subseteq V_a$ and $w\not\in V_x\supseteq V_{x'}$, \yref{containment-i} rules out the possibility that $x'$ is a $T$-ancestor of $a$.
-Therefore, $a$ is a $T$-ancestor of $x$ which is a $T$-ancestor of $x'$.
-Let $z_0,\ldots,z_r$ be the path in $T$ from $z_0:=x'$ to $z_r:=a$.
-Since $x'a\in E(H)$, at least one of $a$ or $x'$ is in $B_{z_i}$, for each $i\in\{0,\ldots,r\}$ and there is at least one $i$ such that $B_{z_i}$ contains both $x'$ and $a$.
-However, by \tref{subtree-root} $x'$ is not contained in $B_{z_i}$ for any $i\in\{1,\ldots,r\}$.
-Therefore $a\in B_{z_i}$ for each $i\in\{0,\ldots,r\}$.
-In particular, $a$ is contained in $B_x$.
-%   Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $B_x$ contains $x$.
-% \end{proof}
-
-% \referee{2}{Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
-% (twice).  Moreover, the way the proof is written, it seems to suggest
-% that we only get that $a \in B_{z_i}$ for each $i \in \{1, \dots, r\}$.
-% The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
-% true though.  In a normalized tree-decomposition, if $xy$ is an edge and
-% $x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
-% and $T_y$ intersect).}
-
-%\note{PM}{I addressed the preceding referee's.}
-
 We are now ready to prove our main result, which we restate here for convenience:
@@
 \mmg*
-
-% \referee{1}{2. Proof of Theorem 9: There is too much notation throughout. As
-% 	one example, it is hard to remember which of $T_x$, $Y_x$, $V_x$, $F_x$, $N_x$,
-% 	$S_x$, $X_v$, $B_x$, $C_x$ is which. If $\phi(S)$ is written for the subset of $V(G)$
-% 	corresponding to a set $S \subseteq V(H)$ and appropriate shorthand and
-% 	standard notation for neighbourhoods are used, then $Y_x$, $V_x$, $F_x$, and
-% 	$N_x$ can be replaced by $\phi(x)$, $\phi(T_x)$, $N(\phi(T_x))$, and $V(G) \setminus N[\phi(T_x)]$.
-% 	The set $X_v$ is only used in the statement of Claim~1, so it does not
-% 	need a name; say what these vertices are explicitly. It may be more
-% 	indicative to write $\phi'(x)$ for $S_x$ and $B'_x$ for $C_x$.}
 
@@
 Let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
-% For each $x\in V(T)$, let $H_x:=V(T_x)$ and let $G_x:=\bigcup_{y\in H_x}Y_y$.
-
-% Apply \cref{generalized-tripod} to $G$ and let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $H_x$, and $G_x$ be defined as in \cref{generalized-tripod}, where the partition $\mathcal{Y}$ has width $\ell$ with respect to the layering $\mathcal{L}$.
 
@@
 In other words, if $P=v_0,\ldots,v_r$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
-We say that a vertex $v\in V(G)$ \defin{participates} in $x$ if $v\in Y_x$, or $\SS$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
+We say that a vertex $v\in V(G)$ \defin{participates} in $x$ if $v\in Y_x$ or if $\SS$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
 For each $v\in V(G)$, let $a(v)$ denote the lowest common $T$-ancestor of all the nodes $x\in V(T)$ in which $v$ participates.
 
-% \begin{clm}
-%   For each $v\in V(G)$, $v$ participates in $a(v)$.
-% \end{clm}
-%
-% \begin{proof}
-%   Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\SS, v\in \{v_0,\ldots,v_r\}\}$.
-Then $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
-%
-%   We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
-If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\SS$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
-In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$, $(v_0,\ldots,v_r)\in\SS$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\SS$ that crosses $x$.
-%
-%   Let $X_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.
-Therefore $X_H$ contains every every node in which $v$ participates.
-The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.
-The fact that $\mathcal{T}$ is a normalized tree decomposition of $H$ implies that some node $a\in X_H$ is a $T$-ancestor of every node in $X_H$.
-Thus, $v$ participates in the node $a=a(v)$ that is the lowest common $T$-ancestor of every node in which $v$ participates, as required.
-% \end{proof}
-
-% \begin{proof}
-%   Let $a(v)$.
-If $v\in Y_{a(v)}$ then $v$ participates in $a(v)$, by definition, and we are done.
-Otherwise, there are two nodes $x,y\in V(T)$ in which $x$ participates and whose lowest common $T$-ancestor is $a(v)$.
-%
-%   There are two cases to consider:
-%   \begin{compactenum}
-%     \item If $v\in Y_x$ then $v\not\in Y_y$.
-Since $v$ participates in $Y_y$, there exist a shortcut $P:=v_0,\ldots,v_r$ in $\SS$ with $v=v_\alpha$ and $v_\beta\in Y_y$ for some $\alpha\in\{0,\ldots,r\}$ and $\beta\in\{1,\ldots,r-1\}$.
-Without loss of generality, $\alpha < \beta$.
-%
-%     For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.
-Since $G[\{v_\alpha,\ldots,v_\beta\}]$ is connected, $H[\{x_\alpha,\ldots,x_\beta\}]$ connected.
-Since $\mathcal{T}$ is a normalized tree decomposition of $H$, $x_j$ is a $T$-ancestor of every node in $x_\alpha,\ldots,x_\beta$ for some $j\in\{\alpha,\ldots,\beta\}$.
-In particular $x_j$ is a $T$-ancestor of $x=x_\alpha$ and $y=x_\beta$, so $x_j$ is a $T$-ancestor of $a(v)$.
-Since $v_j\in V(P)$ and $v_j\in Y_{x_j}$ this implies that $v$ participates in $x_j$.
-Therefore $v$ participates in a $T$-ancestor of $a(v)$, which can only mean that $v$ participates in $a(v)$.
-%
-%     \item If $v\not\in Y_x$ and $v\not\in Y_y$ then there are shortcuts $P_1,P_2\in\SS$ with $v\in V(P_1)$, $v\in V(P_2)$ where $P_1$ has an internal vertex $w_1\in Y_x$ and $P_2$ has an internal vertex $w_2\in Y_y$.
-%
-%
-%     .
-Since $v$ participates in $Y_y$, there exist a shortcut $v_0,\ldots,v_r$ in $\SS$ with $v=v_\alpha$ and $v_\beta\in Y_y$ for some $\alpha\in\{0,\ldots,r\}$ and $\beta\in\{1,\ldots,r-1\}$.
-Without loss of generality, $\alpha < \beta$.
-%
-%     For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.
-Since $G[\{v_\alpha,\ldots,v_\beta\}]$ is connected, $H[\{x_\alpha,\ldots,x_\beta\}]$ connected.
-Since $\mathcal{T}$ is a normalized tree decomposition of $H$, $x_j$ is a $T$-ancestor of every node in $x_\alpha,\ldots,x_\beta$ for some $j\in\{\alpha,\ldots,\beta\}$.
-In particular $x_j$ is a $T$-ancestor of $x=x_\alpha$ and $y=x_\beta$, so $x_j$ is a $T$-ancestor of $a(v)$.
-Since $v_j\in Y_{x_j}$, and $v\not\in Y_{a(v)}$ this implies that $v$ participates in $Y_{x_j}$.
-Therefore $v$ participates in a $T$-ancestor of $a(v)$, which means that $v$ participates in $a(v)$.
-%
-%
-%   \end{compactenum}
-% \end{proof}
-
-% We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
-
-% \begin{clm}\label{x-v-ancestor}
-%   For any $v\in V(G)$ there exists a (unique) node $a(v)\in X_v$ such that
-%   $a(v)$ is a $T$-ancestor of every node in $X_v$.
-% \end{clm}
-%
-% \begin{proof}
-%   Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\SS, v\in \{v_0,\ldots,v_r\}\}$.
-Then $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
-%
-%   We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
-If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\SS$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
-In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$, $(v_0,\ldots,v_r)\in\SS$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\SS$ that crosses $x$.
-%
-%
-%   Let $X_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.  The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.  This implies that $T':=T[]
-%
-%
-%
-%   %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
-%   Choose $a(v)\in X_H$ to be the member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.
-Transitivity of the $T$-ancestor relationship, \tref{ancestor-edge}, and connectivity of $H[X_H]$ implies that such an $a(v)$ exists and is a $T$-ancestor of every node $x\in X_H$, as required.
-% \end{proof}
-
-% \note{DW}{Use $\SS$ for the given shortcut system and $\PP$ for the partition?}
-% \note{PM}{Yes, I've done this.  Again, I was carefuly, but need to reread.}
-
 For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
-Observe that $\PP:=(S_x : x\in V(T))$ is a sequence of pairwise disjoint sets that cover $V(G)$.
+Since $a:V(G)\to V(T)$ is a function with domain $V(G)$ and range $V(T)$, $\PP:=(S_x : x\in V(T))$ is a sequence of pairwise disjoint sets that cover $V(G)$.
 Let $J:=G^\SS/\PP$ denote the resulting quotient graph.
@@
 (Any node $x\in V(T)$ with $S_x=\emptyset$ does not contribute a vertex to $J$.)
-
-% \referee{2}{Page 8.  Can you add a brief explanation why $\PP$ is a
-% 	partition of $V(G)$?  It is clear that the sets in $\PP$ are
-% 	disjoint, but why do they cover $V(G)$?}
-% \note{PM}{I added an explanation in a footnote since this is obvious.  In particular, for any function $f$, $\{f^{-1}(x):x\in\operatorname{range}(f)\}$ is a partition of $\operatorname{domain}(f)$.}
 
@@
 Once we have established (i) and (ii), the result follows easily since a layering of $G^\SS$ is easily obtained from $\mathcal{L}$ by `compressing' groups of $k$ consecutive layers.
-% \referee{1}{7. page 8, item (i): ``$P$ has small layered width with respect to the layering $L$'' $\longrightarrow$ ``$P$ has small layered width with respect to the layering $L$ of $G$''.}
-% \referee{1}{Perhaps mention something about how it is easy to find the new layering.}
-
-% First we need to understand the relationship between sets $S_x\in\PP$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
-
-% \referee{1}{8. general comment: Claims 1, 2, and 4 are especially belaboured by
-% notation as they are quite straightforward.}
-
-% \note{PM}{Same comment: Yes, these claims are definition checking.  We'll see if the experimental notation change helps.}
-
-% \begin{clm}
-% 	\label{s-subset}
-%   For every $x\in V(T)$,
-%   % $S_x\subseteq V_x$.
-%   $S_x\subseteq \bigcup_{x'\in V(T_x)} Y_{x'}$.
-% \end{clm}
-%
-% \begin{proof}
-%   Let $v$ be any vertex in $S_x$ (which implies that $a(v)=x$) and let $Y_y$ be the part of $\mathcal{Y}$ that contains $v$.
-By definition, $v$ participates in $y$ so by definition $a(v)$ is a $T$-ancestor of $y$, so $y\in V(T_{a(v)})$.
-Therefore, $v\in Y_y\subseteq \bigcup_{x'\in V(T_x)} Y_{x'}$, as required.
-% \end{proof}
-
-% Next we complete Step~(i) and show that $\PP$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
+We begin with Step~(i):
 
@@
 
-  Refer to \cref{contribute}.  If $w\in Y_x\cap L_i$ and no path in $\SS$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
+  Refer to \cref{contribute}.
+If $w\in Y_x\cap L_i$ and no path in $\SS$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
 Otherwise, consider some path $P\in\SS$ that contains $w$ as an internal vertex.
@@
 If $w\in L_{i-j}\cup L_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap L_i$.
-%
-% \referee{1}{9. page 8, line 10: ``We say that a vertex $w \in Y_x$ contributes a vertex
-%   $v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.}
-%
-% \note{DW}{What is wrong here?}
 
@@
   For any $j$, the number of vertices $w\in L_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\SS$. Therefore,
-  \[  |S_x\cap L_i|\le d\ell  \, \Big(k+1 + 2k + \sum_{j=2}^k 2(k-j)\Big) %= d(2k+1) + \sum_{i=1}^{k-2} i
+  \[  |S_x\cap L_i|\le d\ell  \, \Big(k+1 + 2k + \sum_{j=2}^k 2(k-j)\Big)
+  %= d(2k+1) + \sum_{i=1}^{k-2} i
       = d\ell(k^2 +3) \enspace . \qedhere
@@
 We now proceed with Step~(ii), showing that $J$ has small treewidth.
-To accomplish this, we construct a small width tree-decomposition $\mathcal{C}:=(C_x:x\in V(T))$ of $J$ using the same tree $T$ used in $\mathcal{T}$.
+To accomplish this, we construct a small width tree-decomposition $\mathcal{C}:=(C_x:x\in V(T))$ of $J$ using the same tree $T$ used in the tree decomposition $\mathcal{T}$ of $H$.
 The following claim will be useful in showing that the resulting decomposition has small width.
@@
 \begin{proof}
-  If $xy\in E(J)$ then there exists a shortcut $P:=v_0,\ldots,v_r\in\SS$ with endpoints $v_0\in S_x$ and $v_r\in S_y$.\footnote{Recall that we have made the assumption that $\SS$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part in $\mathcal{T}$ that contains $v_i$.
+  If $xy\in E(J)$ then there exists a shortcut $P:=v_0,\ldots,v_r\in\SS$ with endpoints $v_0\in S_x$ and $v_r\in S_y$.\footnote{Recall that we have made the assumption that $\SS$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part in $\mathcal{Y}$ that contains $v_i$.
 Since $P$ is a connected subgraph of $G$, $H[\{x_0,\ldots,x_r\}]$ is a connected subgraph of $H$.
-Since $\mathcal{T}$ is a normalized tree decomposition of $H$, some node $x_i\in V(T)$ is a $T$-ancestor of all nodes in $x_0,\ldots,x_r$.
+Since $\mathcal{T}$ is a normalized tree decomposition of $H$, some node $x_i$ is a $T$-ancestor of all nodes in $x_0,\ldots,x_r$.
 
@@
 Finally $a(v_r)=y$ is a $T$-ancestor of $x_r$.
-Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so one is a $T$-ancestor of the other.
-  %
-  %
-  % is a lazy walk in $H$.
-  %
-  % Let $P:=v_0,\ldots,v_r$ (where $v_0=v$ and $v_r=w$)
-  % Suppose, for the sake of contradiction that neither $x$ nor $y$ is an ancestor of the other.  Then
-  %
-  %
-  %
-  %
-  %
-  %
-  % If $P$ has length $1$ (so $vw\in E(G)$) then $
-  %
-  %
-  %
-  % By \cref{s-subset}, $v\in G_x$ and $w\in G_y$.
-  %
-  %
-  %
-  %
-  % Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.
-Since $xy\in E(J)$, $G^\SS$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
-Since $vw\in E(G^{\SS})$,  $\SS$ contains a $vw$-shortcut $P$.
-By \cref{s-subset}, $v\in V_x$ and $w\in V_y$.
-By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
-By \yref{ancestor-edge}, $N_x$ and $V_y$ are also disjoint.
-By \yref{separator} $P$ contains an internal vertex $v'\in F_x$.
-By \yref{ancestor-edge}, $v'\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
-But this implies that $a(v)=a'$ so $v\in S_{a'}$ for some $T$-ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
+Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so one of $x$ or $y$ is a $T$-ancestor of the other.
 \end{proof}
@@
 
-% \referee{1}{10. page 10, Claim 5: Indices could mostly be removed by continuing
-% to talk about $T$-ancestors instead. Move the definition of $H+$ and its
-% directed version up, along with the explanation of what will be proven,
-% so that it isn’t necessary to define the $s_i$.}
-%
-% \note{PM}{I also found the steps in this proof tedious.
-Unfortunately, when I tried to give a simple explanation for it in Banff, both Piotr and Torsten remained unconvinced until we opened the paper and walked through the steps of the proof together.
-I suspect the referee who can see through these proofs if one of the authors of \cite{PS21}.}
-
 \begin{proof}
@@
 For each node $x$ of $T$, we define $C_x$ as follows: $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
-  \cref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, $a,x'\in C_{x'}$.
-The connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ follows from the fact that, for every node $x'\in T[a]$, every node $x$ on the path in $T$ from $x'$ to  $a$ is also a node of $T[a]$.
+  \cref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, both $a$ and $x$ are contained in $C_{x'}$.
+The connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ follows from the fact that, for every node $x'\in V(T[a])$, every node $x$ on the path in $T$ from $a$ to  $x'$ is also a node of $T[a]$.
 Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $J$.
 
-  It remains to show that each bag in our tree decomposition has size at most $\binom{k+t}{t}$.  We will do this by appealing to an elegant result of \citet{PS21}.
-  % on the number of $k$-reachable vertices. in a directed graph obtained by directing the edges of a supergraph of $H$.  \
-  Specifically, let $H^+$ denote the supergraph of $H$ that contains the edge $xy$ if and only if some bag $B_z\in\mathcal{T}$ contains both $x$ and $y$ (i.e., $H^+$ is the chordal closure of $H$ with respect to $\mathcal{T}$).
+  It remains to show that each bag in our tree decomposition has size at most $\binom{k+t}{t}$.
+We will do this by appealing to an elegant result of \citet{PS21} on $k$-reachability in directed skeletons, which we now explain.
+
+  Let $H^+$ denote the supergraph of $H$ that contains the edge $xy$ if and only if some bag $B_z\in\mathcal{T}$ contains both $x$ and $y$ (i.e., $H^+$ is the chordal closure of $H$ with respect to $\mathcal{T}$).
 Let $\overrightarrow{H}^+$ be the directed graph obtained by directing each edge $xy$ of $H^+$ in the direction $\overrightarrow{xy}$ so that $y$ is a $T$-ancestor of $x$.
-\citet[Lemma~13]{PS21} show that, for any $x\in V(\overrightarrow{H}^+)$, the number of vertices of $\overrightarrow{H}^+$ that can be reached by a directed path that originates at $x$ and has length at most $k$ is at most $\binom{k+t}{t}$.
+Observe that directed path in $\overrightarrow{H}^+$ that begins at $x$ leads to a $T$-ancestor of $x$.
+\citet[Lemma~13]{PS21} show that, for any $x\in V(\overrightarrow{H}^+)$, the number of strict $T$-ancestors of $x$ that can be reached by a directed path of length at most $k$ that begins at $x$ is at most $\binom{k+t}{t}-1$.
 
@@
 We will show that $\overrightarrow{H}^+$ contains a directed path from $x$ to $a$ of length at most $k$.
-Since there $a$ a strict $T$-ancestor of $x$, $a\neq x$ so any path from $x$ to $a$ has length at least $1$.
 By \cite[Lemma~13]{PS21}, this implies that the number of strict $T$-ancestors of $x$ contained in $C_x$ is at most $\binom{k+t}{t}-1$ and therefore $|C_x|\le\binom{k+t}{t}$.
@@
   Since $a\in C_x$, there exists a $T$-descendant $x'$ of $x$ such that $ax'\in E(J)$.
-Therefore, there exists a shortcut $v_0,\ldots,v_r$ in $\SS$ with $v_0\in S_{x'}$ and $v_r\in S_a$.
+Therefore, there exists a shortcut $P:=v_0,\ldots,v_r$ in $\SS$ with $v_0\in S_{x'}$ and $v_r\in S_a$.
 For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.
+For each $i\in\{1,\ldots,r-1\}$, $v_i\in Y_{x_i}$ is an internal vertex of $P$, so $v_r$ participates in $x_i$.
+Therefore $a=a(v_r)$ is a $T$-ancestor of $x_i$ for each $i\in\{1,\ldots,r-1\}$.
+
   Since $v_0,\ldots,v_r$ is a path in $G$, $x_0,\ldots,x_r$ is a lazy walk in $H$.\footnote{A \defin{lazy walk} is a graph $G$ is a walk in the pseudograph obtained by adding a self-loop at each vertex of $G$.}  Let $j$ be the minimum integer such that $x_j$ is a strict $T$-ancestor of $x$.
@@
 For each $j\in \{0,\ldots,s\}$, let $z_j$ be the lowest common $T$-ancestor of $y_0,\ldots,y_j$.
-By using the same reasoning used to conclude that $xx_j\in E(H^+)$ we conclude that $z_{i-1}=z_i$ or $z_{i-1}z_i\in E(H^+)$ for each $i\in\{1,\ldots,s\}$, so $z_0,\ldots,z_s$ is a directed lazy walk in $\overrightarrow{H}^+$.
+The same reasoning used to conclude that $xx_j\in E(H^+)$ can be used to conclude that $z_{i-1}=z_i$ or $z_{i-1}z_i\in E(H^+)$ for each $i\in\{1,\ldots,s\}$.
+Therefore $z_0,\ldots,z_s$ is a directed lazy walk in $\overrightarrow{H}^+$.
 Finally, let $\overrightarrow{W}$ be the path in $\overrightarrow{H}^+$ obtained by removing removing duplicates from $z_0,\ldots,z_s$.
@@
 
-For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane spanning (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
-The graph $G_0$ is called the \defin{frame} of $G_0$.
+For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane spanning (multi) subgraph $G'$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G')$ is contained in the interior of some face of $G'$.
+The graph $G'$ is called the \defin{frame} of $G$.
 This definition (for simple graphs) was introduced by \citet{BDGGMR}.

### 2022-03-17 response.tex
Added osme justification

--- 2022-03-13 response.tex
@@
 
+% The following comments (and several additional minor comments) are all related to the proof of Theorem~3 in the original submission.
+Much of the difficulty comes from the fact that there are many things to keep track of.
+We are a given a graph $G$, a shortcut system $\mathcal{S}$, a layering $\mathcal{L}$, a vertex-partition $\mathcal{P}$, and a tree decomposition $\mathcal{T}$ of the quotient graph $H:=G/\mathcal{P}$.
+Now we have to use these to show the existence of a new layering $\mathcal{L}'$, a new vertex partition $\mathcal{P'}$, a new tree decomposition $\mathcal{T}'$ of the new quotient graph $H':=G^{\mathcal{S}}$.
+%
+% During the proof, a number of related sets come up, which we have given names to. As a referee point out, these could remain unnamed and be described using an appropriate notation.
+%
+% \todo[inline]{Write something more here depending on how the experiment with new notation works out.}
+%
+% One issue is that a number of the proofs in this section are essentially "definition checking".
+The relationships between the various vertex subsets that arise in the proof really do just follow (by design) from their definitions.
+In the paper, we do this checking explicitly, though each of these proofs could also be replaced by a sentence of the form ``This [claim/lemma] follows from the definitions of\ldots''.
+%
+% However, when discussing the proof of Theorem~3 with colleagues (very familiar with product structure) recently, we found them unconvinced by high-level explanations.
+It was only after carefully unpacking the definitions that they became convinced.
+These conversations usually end with someone saying ``Ah, of course!''
+%
+%
+% The real action begins in the proof Theorem 9 (using numbering from the original submission) and especially Claim 3 (which involves some careful counting) and Claim 5 (which involves checking that we have everything needed to apply the Pillipczuk-Siebertz lemma on $k$-reachability in ``upward-directed'' graphs of treewidth $t$.)
+%
+% \todo[inline]{Write more here depending on how the notation experiment turns out.}
+
+ % and new vertex partition of a related graph $G^\mathcal{S}$ as we  A large part of the
+
 1.
@@
 
-\tba
+This lemma has been eliminated completely.
 
@@
 
-\tba
+We agree that the notation was excessive (an unfortunate consequence of the general result on $(k,d)$-shortcut systems having evolved from a specific result about $k$-planar graphs).
+We have eliminated the notations $V_x$, $N_x$, $F_x$, $X_x$, and $T_x$.
+
+The notations $Y_x$ and $B_x$ are needed since we need names for the parts of the $H$-decomposition of $G$ and the bags in the tree-decomposition of $H$.
+The notation $C_x$ is limited to the proof of (newly-numbered) Claim~3, where it is used to denote the bags in the tree-decomposition that is the subject of Claim~3.
+
+We have avoided the introduction of any new notation.
+The new presentation is much shorter and clearer.
+We thank the referee for making us revisit this proof carefully.
+The final paper is better for it.
 
@@
 
-\tba
+Since the initial submission, there is now even more justification for including this special case.
+A result of Bose et al on vertex ranking relies critically on the simple treewidth of $H$.
+In particular, if $H$ is planar and has treewidth $3$, then it has an $\ell$-vertex ranking using $O(\log n/\log\log\log n)$ colours.
+The three logs in the denominator come from the fact
+
+\todo[inline]{Continue here...}
 

### 2022-03-17 k-planar.tex
Added osme justification
Specified that $T_0$ has no empty bags

@@
 We say that a $T$-ancestor $a\in V(T)$ of $x\in V(T)$ is a \defin{strict} $T$-ancestor of $x$ if $a\neq x$.
-The \defin{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.  For each node $x\in V(T)$, define
-\[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
-to be the maximal subtree of $T$ rooted at $x$.
+The \defin{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.
+% For each node $x\in V(T)$, define
+% \[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
+% to be the maximal subtree of $T$ rooted at $x$.
 
@@
 
-  Begin with any width-$t$ tree-decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some tree $T_0$.
+  Begin with any width-$t$ tree-decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some tree $T_0$ and has no empty bags.
 Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$.

### 2022-03-17 k-planar.tex
Change suggested by Vida

@@
 
-	In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
+	In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that will be determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
+We first explain how $v_1$, $v_2$ and $v_3$ are chosen and explain how these are used to define $S$.
 There is one case to consider for each possible value of $k$ (see \cref{boring_figure}):

### 2022-03-18 k-planar.tex
added comments about the framed proof

--- 2022-03-17 k-planar.tex
@@
 Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (with an improvement by \citet{UWY}).
-A graph $G$ is \defin{contained in} a graph $X$ if $G$ is isomorphic to a subgraph of $X$.
+A graph $G$ is \defin{contained} in a graph $X$ if $G$ is isomorphic to a subgraph of $X$.
 
@@
 \item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
-\item \citet{DEJGMM21} have since used it to find asymptotically optimal adjacency labellings of planar graphs (resolving a problem of \cite{kannan.naor.ea:implicit}).
+\item \citet{DEJGMM21} use it to find asymptotically optimal adjacency labellings of planar graphs (resolving a problem of \citet{kannan.naor.ea:implicit}).
 \item \citet{EJM} use it to show the existence of a 'universal graph' with $n^{1+o(1)}$ vertices and edges that contains every $n$-vertex planar graph as an induced subgraph (resolving a problem of \citet{babai.chung.ea:on}).
 \end{compactitem}
-In addition to these rather well-known problems, \cref{PlanarProduct} has been used to resolve or make substantial progress on a number of other problems on planar graphs, including $p$-centered colouring \cite{DFMS21}, $\ell$-vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, and odd colourings~\citep{DMO}.
+In addition, \cref{PlanarProduct} has been used to resolve or make substantial progress on a number of other problems on planar graphs, including $p$-centered colouring \cite{DFMS21}, $\ell$-vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, and odd colourings~\citep{DMO}.
 
@@
 
-All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this definition.} These applications motivate finding product structure theorems for other graph classes.
-\citet{DJMMUW20} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of the orientable surface with $h$ handles is $2h$.
-The \textit{Euler genus} of the non-orientable surface with $c$ cross-caps is $c$.
+All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this statement.} These applications motivate finding product structure theorems for other graph classes.
+\citet{DJMMUW20} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of a surface with $h$ handles and $c$ crosscaps is $2h+c$.
 The \textit{Euler genus} of a graph $G$ is the minimum integer $g$ such that $G$ embeds in a surface of Euler genus $g$.
@@
 
-This definition is related to $k$-planarity because of the following observation:
+This definition is related to $k$-planarity by the following observation:
 
@@
 
-A \defin{layering} of a graph $G$ is a sequence $\mathcal{L}=(L_0,L_1,\ldots)$ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
+A \defin{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ 
+\note{DW}{just write ``is an ordered partition $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that for every edge ...''}
+and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
 For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
-To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the vertex it creates in $H$.
+To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the vertex it creates in $H$ \note{DW}{``is indexed by the corresponding vertex in $H$''.}
 
@@
 The purpose of this section is to prove our main technical result, \cref{ShortcutPartition}.
-This theorem shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, an $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
+This theorem shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, a $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
 
@@
 Each subtree $T'$ of $T$ is considered to be rooted at the node in $T'$ closest to $x_0$.
-A node $a\in V(T)$ is a \defin{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \defin{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
+A node $a\in V(T)$ is a \defin{$T$-ancestor} of $x\in V(T)$ and $x$ is a \defin{$T$-descendant} of $a$ if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
 Note that each node $x\in V(T)$ is a $T$-ancestor and $T$-descendant of itself.
@@
 
-% \note{PM}{I have significantly simplified this section, which should make the referees happy. I will read it very carefully again tomorrow}
-
-We begin with a standard technique that allows us to work with a normalised tree-decomposition whose tree has the same vertex set as the graph it decomposes:
+\note{PM}{I have significantly simplified this section, which should make the referees happy. I will read it very carefully again tomorrow}
+
+We begin with a standard technique that allows us to work with a normalised tree-decomposition whose tree has same vertex set as the graph it decomposes:
 
@@
 By the definition of $f$, the tree-decomposition $(B_x:x\in V(T))$ satisfies \tref{subtree-root}.
+  % To see that $(B_x:x\in V(T))$ satisifies \tref{ancestor-edge}, observe that,
+  % if $xy\in E(H)$, then at least one of $x$ or $y$ is contained in $B_z$ for every node $z$ on the path from $x$ to $y$ in $T$.
+If neither $x$ nor $y$ is an ancestor of the other, then some node $z$ on this path has $T$-depth less than that of $x$ and $y$.
+If $x\in B_z$ this contradicts the fact that $x$ is the root of $T[x]$.
+If $y\in B_z$ this contradicts the fact that $y$ is the root of $T[y]$.
 \end{proof}
@@
 
+%\subsection{Generalized Tripod Partitions}
+% 
+% 
+% \begin{figure}[htbp]
+%   \begin{center}
+%     \includegraphics{figs/tripoddo}
+%   \end{center}
+%   \caption{The sets $Y_x$, $F_x$, and $V_x$ associated with $x\in V(T)$
+%   and the ancestors $a_1,\ldots,a_{t'}$ of $X$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.}
+%   \label{fig:generalized-tripod}
+% \end{figure}
+
+% The following lemma shows how to interpret an $H$-partition of $G$ and a tree-decomposition of $H$ as a `hierarchical' decomposition of $G$; refer to \cref{fig:generalized-tripod}.
+
+% For a vertex $v$ of $G$, $N_G(v):=\{w\in V(G):vw\in E(G)\}$ and $N_G[v]:=N_G(v)\cup\{v\}$.  For a vertex set $S\subseteq V$, $N_G(S):=\bigcup_{v\in S} N_G(S)\setminus S$ and $N_G[S]:=N_G(S)\cup S$.
+
+% \note{PM}{
+% Guide to old notation:
+% \begin{compactitem}
+%   \item $V_x := \bigcup_{y\in V(T_x)} Y_y = G_x$
+%   \item $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}= N_G(G_x)$
+%   \item $N_x:=V_x\cup F_x=N_G[G_x]$
+% \end{compactitem}
+% }
+% 
+% \begin{lem}\label{generalized-tripod}
+%   Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
+%   For each $x\in V(T)$, let $H_x:=V(T_x)$ and let $G_x:=\bigcup_{y\in H_x}Y_y$.
+% 
+%   \begin{compactenum}[(Y1)]
+%     % \item\ylabel{thickness} $\mathcal{Y}=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
+%     % \item\ylabel{separator} For each $x\in V(T)$, $N_G(\phi(T_x))$ separates $\phi(T_x)$ from $N
+%     % there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$.
+% 
+%     \item\ylabel{ancestor-edge} For each $x\in V(T)$, there is a set $\{a_1,\ldots,a_{t'}\}$ of $t'\le t$ strict $T$-ancestors of $x$ such that $N_G(G_x) \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
+%   \end{compactenum}
+% \end{lem}
+% 
+% \begin{proof}
+%   Let $\{a_1,\ldots,a_{t'}\}:= B_x\setminus\{x\}$.
+Then, by the definition of a normalized tree decomposition $N_H(H_x))\subseteq\{a_1,\ldots,a_{t'}\}$.
+Therefore $N_G(G_x)\subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
+% \end{proof}
+
+% \referee{1}{1.
+Lemma 3: This lemma is really about ``normalised'' tree-decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$.
+Call this property (T3) and remove (Y1) through (Y5), which should now be obvious.}
+% 
+% \note{PM}{I've added to the \texttt{response.tex} file explaining some of this.
+Each of (Y1), (Y3), (Y4), and (Y5) have a one-sentence justification.
+The justification for (Y2) is just definition unpacking and then using the observation that the referee recommends.}
+% 
+% \note{PM}{I plan to do an experiment with the referee's suggested notation, in a separate git branch to see how it turns out.}
+% 
+% Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:
+% 
+% \begin{compactenum}[(Y1)]\setcounter{enumi}{2}
+%   \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$. \note{PM}{$Y_x$, for $x\in V(T)$ is not defined!}
+%   \item\ylabel{containment-i} $V_x\subseteq V_a$ for every $T$-ancestor $a$ of $x$.
+%   \item\ylabel{containment-ii}$N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
+% \end{compactenum}
+
+% Property~\yref{y-subsets} follows from the fact that $V_x$ is the union of several sets, one of which is $Y_x$.
+Property~\yref{containment-i} follows from the definition of $V_x$ and the fact that $V(T_x)\subseteq V(T_a)$.
+To show Property~\yref{containment-ii} first note that, by \yref{containment-i} it suffices to consider vertices $w\in F_x=N_x\setminus V_x$.
+By definition, every vertex $w\in F_x$ is adjacent, in $G$, to a vertex $v\in V_x$.
+By \yref{containment-i}, $v\in V_a$, so $w$ is either in $V_a$ or $w$ satisfies the condition $vw\in E(G)$, $v\in V_a$, and $w\not\in V_a$, so $w\in F_a$.
+In either case $w\in N_a=V_a\cup F_a$.
+Note that none of \yref{y-subsets}--\yref{containment-ii} depends on \yref{ancestor-edge} (which is important, since \yref{containment-i} is used to establish \yref{ancestor-edge} in the following proof).
+% 
+% 
+% \begin{proof}[Proof of \cref{generalized-tripod}]
+%   % Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
+%   Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
+% 
+%   To establish Property~\yref{ancestor-edge}, consider some vertex $w\in F_x$.
+Since $w\in F_x$, there exists an edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.
+Since $v\in V_x$, $v\in Y_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
+Since $\mathcal{Y}$ is a partition, $w\in Y_{a}$ for some $a\not\in V(T_x)$.
+Since $vw\in E(G)$, we have $x'a\in E(H)$.
+By \tref{ancestor-edge}, one of $a$ or $x'$ is a $T$-ancestor of the other.
+Since $w\in Y_a\subseteq V_a$ and $w\not\in V_x\supseteq V_{x'}$, \yref{containment-i} rules out the possibility that $x'$ is a $T$-ancestor of $a$.
+Therefore, $a$ is a $T$-ancestor of $x$ which is a $T$-ancestor of $x'$.
+Let $z_0,\ldots,z_r$ be the path in $T$ from $z_0:=x'$ to $z_r:=a$.
+Since $x'a\in E(H)$, at least one of $a$ or $x'$ is in $B_{z_i}$, for each $i\in\{0,\ldots,r\}$ and there is at least one $i$ such that $B_{z_i}$ contains both $x'$ and $a$.
+However, by \tref{subtree-root} $x'$ is not contained in $B_{z_i}$ for any $i\in\{1,\ldots,r\}$.
+Therefore $a\in B_{z_i}$ for each $i\in\{0,\ldots,r\}$.
+In particular, $a$ is contained in $B_x$.
+%   Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $B_x$ contains $x$.
+% \end{proof}
+
+% \referee{2}{Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
+% (twice).  Moreover, the way the proof is written, it seems to suggest
+% that we only get that $a \in B_{z_i}$ for each $i \in \{1, \dots, r\}$.
+% The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
+% true though.  In a normalized tree-decomposition, if $xy$ is an edge and
+% $x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
+% and $T_y$ intersect).}
+
+%\note{PM}{I addressed the preceding referee's.}
+
 We are now ready to prove our main result, which we restate here for convenience:
@@
 \mmg*
+
+% \referee{1}{2. Proof of Theorem 9: There is too much notation throughout. As
+% 	one example, it is hard to remember which of $T_x$, $Y_x$, $V_x$, $F_x$, $N_x$,
+% 	$S_x$, $X_v$, $B_x$, $C_x$ is which. If $\phi(S)$ is written for the subset of $V(G)$
+% 	corresponding to a set $S \subseteq V(H)$ and appropriate shorthand and
+% 	standard notation for neighbourhoods are used, then $Y_x$, $V_x$, $F_x$, and
+% 	$N_x$ can be replaced by $\phi(x)$, $\phi(T_x)$, $N(\phi(T_x))$, and $V(G) \setminus N[\phi(T_x)]$.
+% 	The set $X_v$ is only used in the statement of Claim~1, so it does not
+% 	need a name; say what these vertices are explicitly. It may be more
+% 	indicative to write $\phi'(x)$ for $S_x$ and $B'_x$ for $C_x$.}
 
@@
 Let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
+% For each $x\in V(T)$, let $H_x:=V(T_x)$ and let $G_x:=\bigcup_{y\in H_x}Y_y$.
+
+% Apply \cref{generalized-tripod} to $G$ and let $\mathcal{L}$, $\mathcal{Y}$, $\mathcal{T}$, $T$, $H_x$, and $G_x$ be defined as in \cref{generalized-tripod}, where the partition $\mathcal{Y}$ has width $\ell$ with respect to the layering $\mathcal{L}$.
 
@@
 In other words, if $P=v_0,\ldots,v_r$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
-We say that a vertex $v\in V(G)$ \defin{participates} in $x$ if $v\in Y_x$ or if $\SS$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
+We say that a vertex $v\in V(G)$ \defin{participates} in $x$ if $v\in Y_x$, or $\SS$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
 For each $v\in V(G)$, let $a(v)$ denote the lowest common $T$-ancestor of all the nodes $x\in V(T)$ in which $v$ participates.
 
+% \begin{clm}
+%   For each $v\in V(G)$, $v$ participates in $a(v)$.
+% \end{clm}
+% 
+% \begin{proof}
+%   Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\SS, v\in \{v_0,\ldots,v_r\}\}$.
+Then $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
+% 
+%   We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
+If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\SS$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
+In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$, $(v_0,\ldots,v_r)\in\SS$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\SS$ that crosses $x$.
+% 
+%   Let $X_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.
+Therefore $X_H$ contains every every node in which $v$ participates.
+The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.
+The fact that $\mathcal{T}$ is a normalized tree decomposition of $H$ implies that some node $a\in X_H$ is a $T$-ancestor of every node in $X_H$.
+Thus, $v$ participates in the node $a=a(v)$ that is the lowest common $T$-ancestor of every node in which $v$ participates, as required.
+% \end{proof}
+
+% \begin{proof}
+%   Let $a(v)$.
+If $v\in Y_{a(v)}$ then $v$ participates in $a(v)$, by definition, and we are done.
+Otherwise, there are two nodes $x,y\in V(T)$ in which $x$ participates and whose lowest common $T$-ancestor is $a(v)$.
+% 
+%   There are two cases to consider:  
+%   \begin{compactenum}
+%     \item If $v\in Y_x$ then $v\not\in Y_y$.
+Since $v$ participates in $Y_y$, there exist a shortcut $P:=v_0,\ldots,v_r$ in $\SS$ with $v=v_\alpha$ and $v_\beta\in Y_y$ for some $\alpha\in\{0,\ldots,r\}$ and $\beta\in\{1,\ldots,r-1\}$.
+Without loss of generality, $\alpha < \beta$.
+% 
+%     For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.
+Since $G[\{v_\alpha,\ldots,v_\beta\}]$ is connected, $H[\{x_\alpha,\ldots,x_\beta\}]$ connected.
+Since $\mathcal{T}$ is a normalized tree decomposition of $H$, $x_j$ is a $T$-ancestor of every node in $x_\alpha,\ldots,x_\beta$ for some $j\in\{\alpha,\ldots,\beta\}$.
+In particular $x_j$ is a $T$-ancestor of $x=x_\alpha$ and $y=x_\beta$, so $x_j$ is a $T$-ancestor of $a(v)$.
+Since $v_j\in V(P)$ and $v_j\in Y_{x_j}$ this implies that $v$ participates in $x_j$.
+Therefore $v$ participates in a $T$-ancestor of $a(v)$, which can only mean that $v$ participates in $a(v)$.
+% 
+%     \item If $v\not\in Y_x$ and $v\not\in Y_y$ then there are shortcuts $P_1,P_2\in\SS$ with $v\in V(P_1)$, $v\in V(P_2)$ where $P_1$ has an internal vertex $w_1\in Y_x$ and $P_2$ has an internal vertex $w_2\in Y_y$.
+% 
+% 
+%     .
+Since $v$ participates in $Y_y$, there exist a shortcut $v_0,\ldots,v_r$ in $\SS$ with $v=v_\alpha$ and $v_\beta\in Y_y$ for some $\alpha\in\{0,\ldots,r\}$ and $\beta\in\{1,\ldots,r-1\}$.
+Without loss of generality, $\alpha < \beta$.
+% 
+%     For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.
+Since $G[\{v_\alpha,\ldots,v_\beta\}]$ is connected, $H[\{x_\alpha,\ldots,x_\beta\}]$ connected.
+Since $\mathcal{T}$ is a normalized tree decomposition of $H$, $x_j$ is a $T$-ancestor of every node in $x_\alpha,\ldots,x_\beta$ for some $j\in\{\alpha,\ldots,\beta\}$.
+In particular $x_j$ is a $T$-ancestor of $x=x_\alpha$ and $y=x_\beta$, so $x_j$ is a $T$-ancestor of $a(v)$.
+Since $v_j\in Y_{x_j}$, and $v\not\in Y_{a(v)}$ this implies that $v$ participates in $Y_{x_j}$.
+Therefore $v$ participates in a $T$-ancestor of $a(v)$, which means that $v$ participates in $a(v)$.
+% 
+% 
+%   \end{compactenum}
+% \end{proof}
+
+% We let $X_v$ denote the set of nodes $x\in V(T)$ such that $v$ participates in $x$.
+
+% \begin{clm}\label{x-v-ancestor}
+%   For any $v\in V(G)$ there exists a (unique) node $a(v)\in X_v$ such that
+%   $a(v)$ is a $T$-ancestor of every node in $X_v$.
+% \end{clm}
+% 
+% \begin{proof}
+%   Let $Z := \{v\} \cup \{\{v_1,\ldots,v_{r-1}\}:(v_0,\ldots,v_r)\in\SS, v\in \{v_0,\ldots,v_r\}\}$.
+Then $G[Z]$ is connected because $Z$ is the union of (vertex sets of) paths in $G$, each of which contains $v$.
+% 
+%   We claim that $v$ participates in a node $x\in V(T)$ if and only if $Z\cap Y_x\neq\emptyset$.
+If $v$ participates in $x$ then either $v\in Y_x$, so $Z\cap Y_x\supseteq\{v\}$; or $v\in \{v_0,\ldots,v_r\}$ for some shortcut $(v_0,\ldots,v_r)\in\SS$ that crosses $x$, so $Z\cap Y_x\supseteq \{v_i\}$ for some $i\in\{1,\ldots,r-1\}$.
+In the other direction, if $Z\cap Y_x\neq\emptyset$, then either $Z\cap Y_x\supseteq \{v\}$, so $v\in Y_x$; or $Z\cap Y_x\supseteq \{v_i\}$ where $i\in\{1,\ldots,r\}$, $(v_0,\ldots,v_r)\in\SS$ and $v\in\{v_0,\ldots,v_r\}$, so $v\in V(P)$ for a path $P=(v_0,\ldots,v_r)\in\SS$ that crosses $x$.
+% 
+% 
+%   Let $X_H:=\{x\in V(H): Z\cap Y_x\neq\emptyset\}$.  The connectivity of $G[Z]$ implies that $H[X_H]$ is connected.  This implies that $T':=T[]
+% 
+% 
+% 
+%   %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
+%   Choose $a(v)\in X_H$ to be the member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.
+Transitivity of the $T$-ancestor relationship, \tref{ancestor-edge}, and connectivity of $H[X_H]$ implies that such an $a(v)$ exists and is a $T$-ancestor of every node $x\in X_H$, as required.
+% \end{proof}
+
+% \note{DW}{Use $\SS$ for the given shortcut system and $\PP$ for the partition?}
+% \note{PM}{Yes, I've done this.  Again, I was carefuly, but need to reread.}
+
 For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
-Since $a:V(G)\to V(T)$ is a function with domain $V(G)$ and range $V(T)$, $\PP:=(S_x : x\in V(T))$ is a sequence of pairwise disjoint sets that cover $V(G)$.
+Observe that $\PP:=(S_x : x\in V(T))$ is a sequence of pairwise disjoint sets that cover $V(G)$.
 Let $J:=G^\SS/\PP$ denote the resulting quotient graph.
@@
 (Any node $x\in V(T)$ with $S_x=\emptyset$ does not contribute a vertex to $J$.)
+
+% \referee{2}{Page 8.  Can you add a brief explanation why $\PP$ is a
+% 	partition of $V(G)$?  It is clear that the sets in $\PP$ are
+% 	disjoint, but why do they cover $V(G)$?}
+% \note{PM}{I added an explanation in a footnote since this is obvious.  In particular, for any function $f$, $\{f^{-1}(x):x\in\operatorname{range}(f)\}$ is a partition of $\operatorname{domain}(f)$.}
 
@@
 Once we have established (i) and (ii), the result follows easily since a layering of $G^\SS$ is easily obtained from $\mathcal{L}$ by `compressing' groups of $k$ consecutive layers.
-We begin with Step~(i):
+% \referee{1}{7. page 8, item (i): ``$P$ has small layered width with respect to the layering $L$'' $\longrightarrow$ ``$P$ has small layered width with respect to the layering $L$ of $G$''.}
+% \referee{1}{Perhaps mention something about how it is easy to find the new layering.}
+
+% First we need to understand the relationship between sets $S_x\in\PP$ and the related sets $Y_x$, $V_x$, $F_x$, and $N_x$.
+
+% \referee{1}{8. general comment: Claims 1, 2, and 4 are especially belaboured by
+% notation as they are quite straightforward.}
+
+% \note{PM}{Same comment: Yes, these claims are definition checking.  We'll see if the experimental notation change helps.}
+
+% \begin{clm}
+% 	\label{s-subset}
+%   For every $x\in V(T)$, 
+%   % $S_x\subseteq V_x$.
+%   $S_x\subseteq \bigcup_{x'\in V(T_x)} Y_{x'}$.
+% \end{clm}
+% 
+% \begin{proof}
+%   Let $v$ be any vertex in $S_x$ (which implies that $a(v)=x$) and let $Y_y$ be the part of $\mathcal{Y}$ that contains $v$.
+By definition, $v$ participates in $y$ so by definition $a(v)$ is a $T$-ancestor of $y$, so $y\in V(T_{a(v)})$.
+Therefore, $v\in Y_y\subseteq \bigcup_{x'\in V(T_x)} Y_{x'}$, as required.
+% \end{proof}
+
+% Next we complete Step~(i) and show that $\PP$ has small layered width with respect to the layering $\mathcal{L}=\langle L_1,\ldots,L_h\rangle$:
 
@@
 
-  Refer to \cref{contribute}.
-If $w\in Y_x\cap L_i$ and no path in $\SS$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
+  Refer to \cref{contribute}.  If $w\in Y_x\cap L_i$ and no path in $\SS$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
   Otherwise, consider some path $P\in\SS$ that contains $w$ as an internal vertex.
@@
 If $w\in L_{i-j}\cup L_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap L_i$.
+%
+% \referee{1}{9. page 8, line 10: ``We say that a vertex $w \in Y_x$ contributes a vertex
+%   $v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.}
+%
+% \note{DW}{What is wrong here?}
 
@@
   For any $j$, the number of vertices $w\in L_{i+j}\cap Y_x$ is at most $\ell$. Each such vertex $w$ is an internal vertex of at most $d$ paths in $\SS$. Therefore,
-  \[  |S_x\cap L_i|\le d\ell  \, \Big(k+1 + 2k + \sum_{j=2}^k 2(k-j)\Big)
-  %= d(2k+1) + \sum_{i=1}^{k-2} i
+  \[  |S_x\cap L_i|\le d\ell  \, \Big(k+1 + 2k + \sum_{j=2}^k 2(k-j)\Big) %= d(2k+1) + \sum_{i=1}^{k-2} i
       = d\ell(k^2 +3) \enspace . \qedhere
@@
 We now proceed with Step~(ii), showing that $J$ has small treewidth.
-To accomplish this, we construct a small width tree-decomposition $\mathcal{C}:=(C_x:x\in V(T))$ of $J$ using the same tree $T$ used in the tree decomposition $\mathcal{T}$ of $H$.
+To accomplish this, we construct a small width tree-decomposition $\mathcal{C}:=(C_x:x\in V(T))$ of $J$ using the same tree $T$ used in $\mathcal{T}$.
 The following claim will be useful in showing that the resulting decomposition has small width.
@@
 \begin{proof}
-  If $xy\in E(J)$ then there exists a shortcut $P:=v_0,\ldots,v_r\in\SS$ with endpoints $v_0\in S_x$ and $v_r\in S_y$.\footnote{Recall that we have made the assumption that $\SS$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part in $\mathcal{Y}$ that contains $v_i$.
+  If $xy\in E(J)$ then there exists a shortcut $P:=v_0,\ldots,v_r\in\SS$ with endpoints $v_0\in S_x$ and $v_r\in S_y$.\footnote{Recall that we have made the assumption that $\SS$ contains a length-$1$ $vw$-shortcut for each edge $vw\in E(G)$.}  For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part in $\mathcal{T}$ that contains $v_i$.
 Since $P$ is a connected subgraph of $G$, $H[\{x_0,\ldots,x_r\}]$ is a connected subgraph of $H$.
-Since $\mathcal{T}$ is a normalized tree decomposition of $H$, some node $x_i$ is a $T$-ancestor of all nodes in $x_0,\ldots,x_r$.
+Since $\mathcal{T}$ is a normalized tree decomposition of $H$, some node $x_i\in V(T)$ is a $T$-ancestor of all nodes in $x_0,\ldots,x_r$.
 
@@
 Finally $a(v_r)=y$ is a $T$-ancestor of $x_r$.
-Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so one of $x$ or $y$ is a $T$-ancestor of the other.
+Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so one is a $T$-ancestor of the other.
+  % 
+  % 
+  % is a lazy walk in $H$.
+  % 
+  % Let $P:=v_0,\ldots,v_r$ (where $v_0=v$ and $v_r=w$)
+  % Suppose, for the sake of contradiction that neither $x$ nor $y$ is an ancestor of the other.  Then
+  % 
+  % 
+  % 
+  % 
+  % 
+  % 
+  % If $P$ has length $1$ (so $vw\in E(G)$) then $
+  % 
+  % 
+  % 
+  % By \cref{s-subset}, $v\in G_x$ and $w\in G_y$.
+  % 
+  % 
+  % 
+  % 
+  % Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.
+Since $xy\in E(J)$, $G^\SS$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.
+Since $vw\in E(G^{\SS})$,  $\SS$ contains a $vw$-shortcut $P$.
+By \cref{s-subset}, $v\in V_x$ and $w\in V_y$.
+By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.
+By \yref{ancestor-edge}, $N_x$ and $V_y$ are also disjoint.
+By \yref{separator} $P$ contains an internal vertex $v'\in F_x$.
+By \yref{ancestor-edge}, $v'\in Y_a$ for some strict $T$-ancestor $a$ of $x$.
+But this implies that $a(v)=a'$ so $v\in S_{a'}$ for some $T$-ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
 \end{proof}
@@
 
+% \referee{1}{10. page 10, Claim 5: Indices could mostly be removed by continuing
+% to talk about $T$-ancestors instead. Move the definition of $H+$ and its
+% directed version up, along with the explanation of what will be proven,
+% so that it isn’t necessary to define the $s_i$.}
+% 
+% \note{PM}{I also found the steps in this proof tedious.
+Unfortunately, when I tried to give a simple explanation for it in Banff, both Piotr and Torsten remained unconvinced until we opened the paper and walked through the steps of the proof together.
+I suspect the referee who can see through these proofs if one of the authors of \cite{PS21}.}
+
 \begin{proof}
@@
 For each node $x$ of $T$, we define $C_x$ as follows: $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
-  \cref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, both $a$ and $x$ are contained in $C_{x'}$.
-The connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ follows from the fact that, for every node $x'\in V(T[a])$, every node $x$ on the path in $T$ from $a$ to  $x'$ is also a node of $T[a]$.
+  \cref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, $a,x'\in C_{x'}$.
+The connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ follows from the fact that, for every node $x'\in T[a]$, every node $x$ on the path in $T$ from $x'$ to  $a$ is also a node of $T[a]$.
 Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $J$.
 
-  It remains to show that each bag in our tree decomposition has size at most $\binom{k+t}{t}$.
-We will do this by appealing to an elegant result of \citet{PS21} on $k$-reachability in directed skeletons, which we now explain.
-
-  Let $H^+$ denote the supergraph of $H$ that contains the edge $xy$ if and only if some bag $B_z\in\mathcal{T}$ contains both $x$ and $y$ (i.e., $H^+$ is the chordal closure of $H$ with respect to $\mathcal{T}$).
+  It remains to show that each bag in our tree decomposition has size at most $\binom{k+t}{t}$.  We will do this by appealing to an elegant result of \citet{PS21}. 
+  % on the number of $k$-reachable vertices. in a directed graph obtained by directing the edges of a supergraph of $H$.  \
+  Specifically, let $H^+$ denote the supergraph of $H$ that contains the edge $xy$ if and only if some bag $B_z\in\mathcal{T}$ contains both $x$ and $y$ (i.e., $H^+$ is the chordal closure of $H$ with respect to $\mathcal{T}$).
 Let $\overrightarrow{H}^+$ be the directed graph obtained by directing each edge $xy$ of $H^+$ in the direction $\overrightarrow{xy}$ so that $y$ is a $T$-ancestor of $x$.
-Observe that directed path in $\overrightarrow{H}^+$ that begins at $x$ leads to a $T$-ancestor of $x$.
-\citet[Lemma~13]{PS21} show that, for any $x\in V(\overrightarrow{H}^+)$, the number of strict $T$-ancestors of $x$ that can be reached by a directed path of length at most $k$ that begins at $x$ is at most $\binom{k+t}{t}-1$.
+\citet[Lemma~13]{PS21} show that, for any $x\in V(\overrightarrow{H}^+)$, the number of vertices of $\overrightarrow{H}^+$ that can be reached by a directed path that originates at $x$ and has length at most $k$ is at most $\binom{k+t}{t}$.
 
@@
 We will show that $\overrightarrow{H}^+$ contains a directed path from $x$ to $a$ of length at most $k$.
+Since there $a$ a strict $T$-ancestor of $x$, $a\neq x$ so any path from $x$ to $a$ has length at least $1$.
 By \cite[Lemma~13]{PS21}, this implies that the number of strict $T$-ancestors of $x$ contained in $C_x$ is at most $\binom{k+t}{t}-1$ and therefore $|C_x|\le\binom{k+t}{t}$.
@@
   Since $a\in C_x$, there exists a $T$-descendant $x'$ of $x$ such that $ax'\in E(J)$.
-Therefore, there exists a shortcut $P:=v_0,\ldots,v_r$ in $\SS$ with $v_0\in S_{x'}$ and $v_r\in S_a$.
+Therefore, there exists a shortcut $v_0,\ldots,v_r$ in $\SS$ with $v_0\in S_{x'}$ and $v_r\in S_a$.
 For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.
-For each $i\in\{1,\ldots,r-1\}$, $v_i\in Y_{x_i}$ is an internal vertex of $P$, so $v_r$ participates in $x_i$.
-Therefore $a=a(v_r)$ is a $T$-ancestor of $x_i$ for each $i\in\{1,\ldots,r-1\}$.
-
 Since $v_0,\ldots,v_r$ is a path in $G$, $x_0,\ldots,x_r$ is a lazy walk in $H$.\footnote{A \defin{lazy walk} is a graph $G$ is a walk in the pseudograph obtained by adding a self-loop at each vertex of $G$.}  Let $j$ be the minimum integer such that $x_j$ is a strict $T$-ancestor of $x$.
@@
 For each $j\in \{0,\ldots,s\}$, let $z_j$ be the lowest common $T$-ancestor of $y_0,\ldots,y_j$.
-The same reasoning used to conclude that $xx_j\in E(H^+)$ can be used to conclude that $z_{i-1}=z_i$ or $z_{i-1}z_i\in E(H^+)$ for each $i\in\{1,\ldots,s\}$.
-Therefore $z_0,\ldots,z_s$ is a directed lazy walk in $\overrightarrow{H}^+$.
+By using the same reasoning used to conclude that $xx_j\in E(H^+)$ we conclude that $z_{i-1}=z_i$ or $z_{i-1}z_i\in E(H^+)$ for each $i\in\{1,\ldots,s\}$, so $z_0,\ldots,z_s$ is a directed lazy walk in $\overrightarrow{H}^+$.
 Finally, let $\overrightarrow{W}$ be the path in $\overrightarrow{H}^+$ obtained by removing removing duplicates from $z_0,\ldots,z_s$.
@@
 
-For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane spanning (multi) subgraph $G'$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G')$ is contained in the interior of some face of $G'$.
-The graph $G'$ is called the \defin{frame} of $G$.
+For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane \note{DW}{Use the language of ``embedded graph''} having a biconnected plane spanning (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
+The graph $G_0$ is called the \defin{frame} of $G_0$.
 This definition (for simple graphs) was introduced by \citet{BDGGMR}.
-Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar graph.
+Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar multigraph (by adding a vertex inside each face $F$ adjacent to the vertices of $F$).
 Thus \cref{ShortcutProduct} is applicable.
@@
 \begin{thm}\label{d_framed_product_stucture}
-Every $d$-framed multigraph is contained in $H\boxtimes P\boxtimes K_{d+3\floor{d/2}-3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+For any integer $d\ge 3$, every $d$-framed multigraph is contained in $H\boxtimes P\boxtimes K_{d+3\floor{d/2}-3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
 \end{thm}
@@
 The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{v_1,\ldots,v_t\}$.
-Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that $\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}$ for each $v\in V(G)$.
-
-\note{PM}{Consider a change of notation:   Replace $F$ with $C$.  (I was using $F$ before to be consistent with the previous section, but that notation is gone now.)}
-
-\note{PM}{I've changed my mind about this notation change.
-We end up applying the lemma inductively on the faces $F_1,\ldots,F_m$ of $M$ (which are cycles in $G$).
-It makes sense that the original cycle in $G$ is called $F$.}
+Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that for each $v\in V(G)$,  \[\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.\]
+
+\note{PM}{Consider a change of notation:   Replace $F$ with $C$.  (I was using $F$ before to be consistent with the previous section, but that notation is gone now.)} \note{DW}{I agree.}
+
+\note{DW}{I suggest $G'$ be changed to $G_0$ throughout (since $G'$ feels like a graph derived from $G$, which it is not). Similarly, replace $N'$ by $N_0$.}
+
+\note{DW}{We need to define ``closure''}
 
@@
 		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
-    \item Let $F$ be a cycle in $G'$ whose vertices are partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
+		\item Let $F$ be a cycle in $G'$ that $V(F)$ is partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
 		\begin{compactenum}
@@
 			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
+			\note{DW}{suggested revision: ``$P_i$ has a partition $\{X_i,Y_i\}$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.''}
 		\end{compactenum}
@@
 		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
-    \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
+		\item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$ \note{DW}{suggested revision, ``$S_x$ has a partition $\{X_x,Y_x\}$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$''};
 		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
@@
 \begin{proof}
-	This proof is very similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is complicated by several factors.
+	This proof is similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is complicated by several factors.
 In particular, the reader should keep in mind that $G$ and $G'$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
 
-	The proof is by induction on the number of inner faces of $N'$.
-A particular easy case occurs when $V(N')=V(F)$, for which we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
+	The proof is by induction on the number of inner faces of $N'$.  A particular easy case occurs when $V(N')=V(F)$, for which we can simply take 
+	\note{DW}{Suggested revision, ``In the base case, $N'$ has only one inner face, and we can simply take ...} 	
+	$\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
 	\begin{compactenum}[(i)]
@@
 
+\note{DW}{Now assume that $N'$ has at least two internal faces.}
+
 	In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
@@
 	\begin{compactenum}
-		\item If $k=1$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
-
-		\item If $k= 2$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
+		\item If $k=1$ then let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
+
+		\item If $k= 2$ then let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
 
@@
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
-Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N'$, add edges $d'-2$ edges inside of $F$ to split it into $3$-sided faces.
+Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N'$, add edges $d'-2$
+		\note{DW}{should be $d'-3$, right?} edges inside of $F$ to split it into $3$-sided faces.
 This yields a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary.
 By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
-Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.
-% (Note that $p\in\{3,4\}$.)
+Let $\tau:=v_1,\ldots,v_p$ be the face \note{DW}{``be the facial cycle''} of $N'$ that contains $\tau''$.
 	\end{compactenum}
@@
 Furthermore, $|X|\le p-3\le d-3$, since $v_1,v_2,v_3\in Y\cup V(F)$.
-Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ will not violate condition~(ii).
+Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ will not violate condition~(ii) \note{DW}{Use ``satisfies'' instead of ``will not violate''}.
 
@@
 
+\note{DW}{These definitions should appear earlier, certainly before the definition of framed graph, which should be defined as an ``embedded graph''.
+One option is to put this whole paragraph as a footnote in the introduction where we first mention $k$-planar graphs.
+The experienced reader can skip this paragraph, which says to me it should be a footnote.}
 We first formally define $k$-planar graphs.
@@
 
-\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{d_framed_product_stucture} with the method in \citep{DHHW}.
+%\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{d_framed_product_stucture} with the method in \citep{DHHW}.
 I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
 Okay?}
-\note{PM}{Okay to this and to all similar questions that come later (which I've now commented out.)}
+%\note{PM}{Okay to this and to all similar questions that come later (which I've now commented out.)}
+%\note{DW}{Robert is working on it.}
 
@@
 \label{MewMapGraphLemma}
-for all integers $g\geq 0$ and $d\geq 3$, every $(g,d)$-map graph is a spanning subgraph of  $(g,d)$-framed graph.
+For all integers $g\geq 0$ and $d\geq 3$, every $(g,d)$-map graph is a spanning subgraph of a $(g,d)$-framed graph.
 \end{lem}

### 2022-03-18 response.tex

--- 2022-03-15 response.tex
@@
 \usepackage{stmaryrd}
-
+\usepackage{amssymb}
 \usepackage{todonotes}
@@
 \newcommand{\tba}{\textcolor{red}{To be addresssed.}}
+
+\newenvironment{response}{\color{blue}}{}
 
@@
 
-The following comments are all related to the proof of Theorem~3.
-Much of the difficulty comes from the fact that there are many things to keep track of.
-We are a given a graph $G$, a shortcut system $\mathcal{S}$, a layering $\mathcal{L}$, a vertex-partition $\mathcal{P}$, and a tree decomposition $\mathcal{T}$ of the quotient graph $H:=G/\mathcal{P}$.
-Now we have to use these to show the existence of a new layering $\mathcal{L}'$, a new vertex partition $\mathcal{P'}$, a new tree decomposition $\mathcal{T}'$ of the new quotient graph $H':=G^{\mathcal{S}}$.
-
-During the proof, a number of related sets come up, which we have given names to. As a referee point out, these could remain unnamed and be described using an appropriate notation.
-
-\todo[inline]{Write something more here depending on how the experiment with new notation works out.}
-
-A secondary issue is that a number of the proofs in this section are essentially "definition checking".
-The relationships between the various vertex sets that arise in the proof really do just follow from their definitions (which were made with this purpose in mind).
-We have done this checking explicitly, though each of these proofs could also be replaced by a sentence of the form ``This [claim/lemma] follows from the definitions of\ldots''.
-It is not clear what the best choice is.
-
-The real action begins in the proof Theorem 9 (using numbering from the original submission) and especially Claim 3 (which involves some careful counting) and Claim 5 (which involves checking that we have everything needed to apply the Pillipczuk-Siebertz lemma on $k$-reachability in ``upward-directed'' graphs of treewidth $t$.)
-
-\todo[inline]{Write more here depending on how the notation experiment turns out.}
-
- % and new vertex partition of a related graph $G^\mathcal{S}$ as we  A large part of the
-
 1.
@@
 
-\tba
+\textcolor{blue}{This lemma has been eliminated completely. (See the next response.)}
 
@@
 
-\tba
+\begin{response}
+  We agree that the notation was excessive (an unfortunate consequence of the general result on $(k,d)$-shortcut systems having evolved from a specific result about $k$-planar graphs).
+We have eliminated the notations $V_x$, $N_x$, $F_x$, $X_x$, and $T_x$.
+
+  The notations $Y_x$ and $B_x$ are needed since we need names for the parts of the $H$-decomposition of $G$ and the bags in the tree-decomposition of $H$.
+The notation $C_x$ is limited to the proof of (newly-numbered) Claim~3, where it is used to denote the bags in the tree-decomposition that is the subject of Claim~3.
+
+  We have avoided the introduction of any new notation.
+The new presentation is much shorter and clearer.
+We thank the referee for making us revisit this proof carefully.
+The final paper is better for it.
+\end{response}
 
@@
 
-\tba
+\begin{response}
+  Since the initial submission, there is now even more justification for including this special case.
+A result of Bose et al on vertex ranking relies critically on the simple treewidth of $H$.
+In particular, if $H$ is planar and has treewidth $3$, then $H\boxtimes P\boxtimes K_\ell$ an $r$-vertex ranking using $O(\log n/\log\log\log n)$ colours (for any fixed $r$ and $\ell$).
+The three logs in the denominator come from the fact that $H$ has simple treewidth $3$.
+Simple treewidth $4$ or even (non-simple) treewidth $3$ lead to more logs in the denominator.
+The number of logs in the denominator is exactly the simple treewidth of $H$, and this is tight.
+
+  Also, while adding details requested by Referee \#2 we discovered that a small change in the proof (removing both edges of a crossing pair instead of just one) results in a better constant and leads to a more general result that applies to $d$-framed graphs.
+This, in turn, leads to better results for $d$-map graphs.
+\end{response}
 
@@
 
-\tba
-
 6. page 6, last paragraph of Lemma 2: This isn’t quite right since you
@@
 
-\tba
+\textcolor{blue}{Both of these have been fixed by first insisting that we start with a tree decomposition having no empty bags and then adding the (empty) root $x_0$ which is used only in the first stage to ensure that every problematic node has a an edge to its parent that we can subdivide.
+Then the node $x_0$ is removed before the second stage where, by definition, every problematic/redundant node has a parent.}
 
@@
 
-\tba
+\textcolor{blue}{We have added ``Once we have established (i) and (ii), the result follows easily since a layering of $G^\mathcal{S}$ is easily obtained from $\mathcal{L}$ by `compressing' groups of $k$ consecutive layers.''}
 
@@
 
-\tba
+\textcolor{blue}{Claims 1 and 2 are now gone, leaving only Claims 3, 4, and 5 whose proofs have now been simplified.}
 
@@
 
-\tba
+\textcolor{blue}{Replaced with ``We say that a vertex $w\in Y_x$ \emph{contributes} a vertex $v\in S_x$ if $v=w$ or if some path in $\mathcal{S}$ that contains $v$ has $w$ as an internal vertex.''}
 
@@
 
-\tba
+\textcolor{blue}{This is greatly simplified in the new version.}
 
@@
 
-\textcolor{blue}{Rearranged the sentence so that it is more clear we are defining $w'=w$. }
+\textcolor{blue}{This has disappeared in the new version}
 
@@
 
-\textcolor{blue}{We have put `hierarchical' in quotes to indicate it is not a defined word.}
+\textcolor{blue}{The lemma that this sentence referred to is now gone, and so is the sentence.}
 
@@
 
-\tba
+\textcolor{blue}{We changed this sentence to ``Since $a:V(G)\to V(T)$ is a function with domain $V(G)$ and range $V(T)$, $\mathcal{P}:=(S_x : x\in V(T))$ is a sequence of pairwise disjoint sets that cover $V(G)$.''}
 
@@
 
-\done
+\textcolor{blue}{This is gone in the new version.}
 
@@
 
-\tba
+\textcolor{blue}{The referee is correct, the induced graph was the wrong thing to use here.
+Rather, it is the set of edges and vertices incident to the four faces that appear at a crossing.
+This issue no longer appears in the new version that works with $d$-framed graphs.
+The relationship between $4$-framed multigraphs and $1$-planar graphs is now stated as a lemma.}
 
@@
 The proof is quite similar to the previous proof except that graph $G'$ is obtained by removing both crossing edges involved in a crossing (the spars of a kite), rather than just one.
-We use the language of framed multigraph, for consistency with the literature.
+We now use the language of framed multigraph, for consistency with the literature.
 Thanks to the referee for forcing us to write this proof more carefully so that we could discover this improvement.}
@@
 
-\tba
-
 Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.
 
-\tba
+\textcolor{blue}{This proof is now gone since the authors of [10] helpfully include this result in the journal version of their paper.}
 
@@
 
-\textcolor{blue}{We have been told there is an error in their proof.}
+\textcolor{blue}{We have been told that there is an error in their proof.}
 

### 2022-03-18 k-planar.tex

--- 2022-03-17 k-planar.tex
@@
 \begin{thm}\label{d_framed_product_stucture}
-  Every $d$-framed multigraph is contained in $H\boxtimes P\boxtimes K_{d+3\floor{d/2}-3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+For any integer $d\ge 3$, every $d$-framed multigraph is contained in $H\boxtimes P\boxtimes K_{d+3\floor{d/2}-3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
 \end{thm}
@@
 The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{v_1,\ldots,v_t\}$.
-Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that $\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}$ for each $v\in V(G)$.
-
-\note{PM}{Consider a change of notation:   Replace $F$ with $C$.  (I was using $F$ before to be consistent with the previous section, but that notation is gone now.)}
-
-\note{PM}{I've changed my mind about this notation change.
-We end up applying the lemma inductively on the faces $F_1,\ldots,F_m$ of $M$ (which are cycles in $G$).
-It makes sense that the original cycle in $G$ is called $F$.}
+Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that for each $v\in V(G)$,  \[\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.\]
+
+%\note{PM}{Consider a change of notation:   Replace $F$ with $C$.  (I was using $F$ before to be consistent with the previous section, but that notation is gone now.)} \note{DW}{I agree.}
+
+\note{DW}{I suggest $G'$ be changed to $G_0$ throughout (since $G'$ feels like a graph derived from $G$, which it is not). Similarly, replace $N'$ by $N_0$.}
+
+\note{DW}{We need to define ``closure''}
 
@@
 		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
-    \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
+		\item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$ \note{DW}{suggested revision, ``$S_x$ has a partition $\{X_x,Y_x\}$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$''};
 		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
 	\end{compactenum}
+%\note{PM}{I've changed my mind about this notation change.
+We end up applying the lemma inductively on the faces $F_1,\ldots,F_m$ of $M$ (which are cycles in $G$).
+It makes sense that the original cycle in $G$ is called $F$.}
+%\note{DW}{okay}
+%\begin{lem}
+%  \label{induction} The setup:
+%  \begin{compactenum}
+%    \item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
+%    \item Let $F_0$ be the outer face of $G'$ and let $T$ be a BFS spanning forest of $G'$ rooted at $V(F_0)$.
+%    \item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
+%    \item Let $F$ be a cycle in $G'$ whose vertices are partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
+%    \begin{compactenum}
+%      \item $F[P_i]$ is connected; and
+%      \item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
+%    \end{compactenum}
+%    \item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
+%  \end{compactenum}
+%  Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
+%  \begin{compactenum}[(i)]
+%    \item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
+%    \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
+%    \item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
+%  \end{compactenum}
 \end{lem}
@@
 
+\note{DW}{Now assume that $N'$ has at least two internal faces.}
+
 In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that will be determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
@@
 There is one case to consider for each possible value of $k$ (see \cref{boring_figure}):
+
 \begin{compactenum}
-		\item If $k=1$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
-
-		\item If $k= 2$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
+		\item If $k=1$ then let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
+
+		\item If $k= 2$ then let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
 

### 2022-03-18 k-planar.tex
added comments

@@
 
-\note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph. Is the following lemma known (at least in the $g=0$ case)?}
-\note{PM}{I have no idea if this is known.  If so, it's probably not stated in terms of frames.}
+\note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph.
+Is the following lemma known (at least in the $g=0$ case)?} \note{PM}{I have no idea if this is known.
+If so, it's probably not stated in terms of frames.} \note{DW}{I will work this out.
+We should mention \citet{Brandenburg19,Brandenburg20}.
+\citet{Brandenburg20} also uses multigraphs, and this paper also mentions 1-planar graphs.
+So I suspect \cref{1_planar_is_4_framed} is known.
+I need to go now.}
+
 \begin{lem}

### 2022-03-18 k-planar.tex
added Brandenburg etc references and sentences

--- 2022-03-18 k-planar.tex
@@
 This result is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs. 
-% \todo{Move this comment to the section on $1$-planar graphs?}
 For simple 1-planar graphs it is essential that we allow frames with parallel edges.
-% \note{DW}{I feel that at this point we need to justify the introduction of multigraphs.}
 
@@
 
-\cref{1-planar} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma, which is of independent interest.
-
-\note{DW}{Say something about how \cref{1_planar_is_4_framed} is better than similar results in the literature.}
-
-\note{PM}{See my attempt below.}
-% \note{PM}{I don't know what to say without explaining that $4$-framed (simple) graphs only contain $3$-connected $1$-planar graphs.
-Vida's example shows that there is a $1$-plane embedding of a graph that can't be augmented into a $4$-framed graph, but I haven't checked if it can't be drawn some better way.
-And I also don't know if the Bekos claim that $8$-framed (smple) graphs contain all $1$-planar graphs is proven anywhere and, if so, is it tight?}
+\cref{1-planar} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma.
+Similar results appear in earlier works \cite{CGP06,BDGGMR,Brandenburg19,Brandenburg20} but we state the precise lemma and provide a proof here for the sake of completeness.
 
@@
 
-Before proving \cref{1_planar_is_4_framed}, we remark that the lemma seems not to be true if we replace `multigraph' with `graph'.
+\note{DW}{Write  ``A multigraph $G$ is $1$-planar if and only if ...''}
+
+%Before proving \cref{1_planar_is_4_framed}, we remark that the lemma seems not to be true if we replace `multigraph' with `graph'.
 We are only aware of two previous results on the relationship between $1$-planar graphs and $d$-framed graphs \cite{BDGGMR}:
-\begin{inparaenum}[(i)]
-  \item Every $3$-connected $1$-planar graph is contained in some $4$-framed graph.
-  \item Every $1$-planar graph is contained in some $8$-framed graph.
-\end{inparaenum}
-Neither of these results allows a proof of \cref{1-planar} without sacrificing generality or accepting a larger constant.
-
-\begin{proof}[Proof of \cref{1_planar_is_4_framed}]
+%\begin{inparaenum}[(i)]
+%  \item Every $3$-connected $1$-planar graph is contained in some $4$-framed graph.
+%  \item Every $1$-planar graph is contained in some $8$-framed graph.
+%\end{inparaenum}
+%Neither of these results allows a proof of \cref{1-planar} without sacrificing generality or accepting a larger constant.
+
+\begin{proof}
 Let $G$ be a $1$-plane graph.
@@
 \citet{dujmovic.eppstein.ea:structure} proved that $G$ has layered treewidth $(4g+6)(k+1)$.
-Van den Heuvel and Wood~\citep{vdHW17} showed that this implies that $G$  has $r$-strong colouring number at most $(4g + 6)(k + 1)(2r + 1)$.
+Van den Heuvel and Wood~\citep{vdHW18} showed that this implies that $G$  has $r$-strong colouring number at most $(4g + 6)(k + 1)(2r + 1)$.
 By a result of \citet{zhu:colouring}, $G$ has $r$-weak colouring number at most $( (4g + 6)(k + 1)(2r + 1) )^r$, which by another result of  \citet{zhu:colouring} implies that $G$ has  $p$-centered chromatic number at most $( (4g+6)(k+1)(2^{p-1} + 1) )^{2^{p-2}}$.
@@
 
-\note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph. Is the following lemma known (at least in the $g=0$ case)?}
-\note{PM}{I have no idea if this is known.  If so, it's probably not stated in terms of frames.}
+Map graphs are defined as follows.
+Start with a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
+Let $G$ be the graph whose vertices are the nations of $G_0$, where two vertices are adjacent in $G$ if the corresponding faces in $G_0$ share a vertex.
+Then $G$ is called a \defin{$(g,d)$-map graph}.
+A $(0,d)$-map graph is called a (plane) \defin{$d$-map graph}; see \citep{FLS-SODA12,CGP02} for example.
+The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.
+
+\note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph.}
+
+\begin{thm}
+\label{NewMapGraph}
+Every $d$-map graph is contained in $H \boxtimes P \boxtimes K_{ d + 3\floor{d/2} -3 }$ for some planar graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
+\end{thm}
+
+\cref{NewMapGraph} is an immediate consequence of \cref{d_framed_product_stucture}} and the next lemma.
+Similar results to \cref{MewMapGraphLemma} appear in earlier works \cite{CGP06,BDGGMR,Brandenburg19,Brandenburg20} but we state the precise lemma and provide a proof here for the sake of completeness.
+
 \begin{lem}
@@
 
-\begin{thm}
-\label{NewMapGraph}
-Every $d$-map graph is contained in $H \boxtimes P \boxtimes K_{ d + 3\floor{d/2} -3 }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
-\end{thm}
-
-Map graphs are defined as follows.
-Start with a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
-Let $G$ be the graph whose vertices are the nations of $G_0$, where two vertices are adjacent in $G$ if the corresponding faces in $G_0$ share a vertex.
-Then $G$ is called a \defin{$(g,d)$-map graph}.
-A $(0,d)$-map graph is called a (plane) \defin{$d$-map graph}; see \citep{FLS-SODA12,CGP02} for example.
-The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.
+\note{DW}{give a better version of \cref{MapPartition} with $g=0$. }
+
 %So $(g,d)$-map graphs generalise graphs embedded in a surface, and we now assume that $d\geq 4$ for the remainder of this section.
@@
 
-% \note{DW}{\cref{MapPartition} has been updated.}
+These results give the first constant upper bound on the non-repetitive chromatic number of map graphs, the first polynomial bounds on the $p$-centered chromatic number of map graphs, and the best known bounds on the queue-number of map graphs.
 
@@
 Okay?}
-
-These results give the first constant upper bound on the non-repetitive chromatic number of map graphs, the first polynomial bounds on the $p$-centered chromatic number of map graphs, and the best known bounds on the queue-number of map graphs.
 

### 2022-03-19 k-planar.tex
Fixed mismatched braces
Pfffft

@@
 The \textit{Euler genus} of a graph $G$ is the minimum integer $g$ such that $G$ embeds in a surface of Euler genus $g$.
-Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.} and for apex-minor-free graphs\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
+Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.} and for apex-minor-free graphs,\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges.
 A class $\mathcal{G}$ of graphs is \defin{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$.
@@
 A 1-apex graph is \defin{apex}.
-A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.}, and \citet{DEMWW22} do so for graphs in any minor-closed class and with bounded maximum degree.
+A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.} and \citet{DEMWW22} do so for graphs in any minor-closed class and with bounded maximum degree.
 See \cref{Generalisations} for more precise statements and see \citep{DHJLW21} for a survey on this topic.
@@
 
-A \defin{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ 
-\note{DW}{just write ``is an ordered partition $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that for every edge ...''}
-and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
+A \defin{layering} of a graph $G$
+% is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$
+% \note{DW}{just write ``
+is an ordered partition $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that for every edge
+ % ...''}
+$vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
 For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
-To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the vertex it creates in $H$ \note{DW}{``is indexed by the corresponding vertex in $H$''.}
+To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the corresponding vertex in $H$.
 
@@
 \citet{DJMMUW20} showed that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \cite{UWY}).
-The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.
-At the core of their work is the elegant proof by \citet{PS21} of the following result:
+The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.\footnote{\citet{bose.morin.ea:optimal} have recently given linear time algorithms for computing layered $H$-partitions of planar graphs.} At the core of their work is the elegant proof by \citet{PS21} of the following result:
 
@@
 \label{ps}
-Every planar triangulation $G$ has an $H$-partition $\PP$
-such that $\tw(H)\leq 8$ and $G[P]$ is a shortest path in $G$ for each $P\in\PP$.
+  Every planar triangulation $G$ has an $H$-partition $\PP$ such that $\tw(H)\leq 8$ and $G[P]$ is a shortest path in $G$ for each $P\in\PP$.
 \end{thm}
@@
 The \defin{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.
+The \defin{lowest common ancestor} of a non-empty set $S\subseteq V(T)$ is the maximum $T$-depth node of $T$ that is a $T$-ancestor of every node in $S$.
 % For each node $x\in V(T)$, define
@@
   Begin with any width-$t$ tree-decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some tree $T_0$ and has no empty bags.
-Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$.
+Select any node $x_1\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x_1$ and root $T_0$ at $x_0$.
 (The purpose of $x_0$ is to ensure that every node $x$ for which $B_x$ is non-empty has a parent.)  Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y\}]$.
@@
 Subdivide the edge between $a$ and its parent in $T_0$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
-Now $f(y)=a'$ and $f(x)=a$, so this modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
-
-  Next, remove the node $x_0$ from $T_0$, and consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.
+Now $f(y)=a'$ and $f(x)=a$, so this modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$.
+Repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.
+
+  Next, remove the node $x_0$ from $T_0$ (so that $x_1$ becomes the new root of $T_0$).
+Consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.
 In this case, $B_{a}\subseteq B_{a'}$ where $a'$ is the parent of $a$ since any $x\in B_a\setminus B_{a'}$ would have $f(x)=a$.
@@
 Finally $a(v_r)=y$ is a $T$-ancestor of $x_r$.
-Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so one of $x$ or $y$ is a $T$-ancestor of the other.
+Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so at least one of $x$ or $y$ is a $T$-ancestor of the other.
 \end{proof}
@@
 
-  It remains to show that each bag in our tree decomposition has size at most $\binom{k+t}{t}$.
+  It remains to show that each bag in this tree decomposition has size at most $\binom{k+t}{t}$.
 We will do this by appealing to an elegant result of \citet{PS21} on $k$-reachability in directed skeletons, which we now explain.
@@
 Let $\overrightarrow{H}^+$ be the directed graph obtained by directing each edge $xy$ of $H^+$ in the direction $\overrightarrow{xy}$ so that $y$ is a $T$-ancestor of $x$.
-Observe that directed path in $\overrightarrow{H}^+$ that begins at $x$ leads to a $T$-ancestor of $x$.
+Observe that any directed path in $\overrightarrow{H}^+$ that begins at $x$ leads to a $T$-ancestor of $x$.
 \citet[Lemma~13]{PS21} show that, for any $x\in V(\overrightarrow{H}^+)$, the number of strict $T$-ancestors of $x$ that can be reached by a directed path of length at most $k$ that begins at $x$ is at most $\binom{k+t}{t}-1$.
@@
 Finally, let $\overrightarrow{W}$ be the path in $\overrightarrow{H}^+$ obtained by removing removing duplicates from $z_0,\ldots,z_s$.
-Then $\overrightarrow{W}$ is a directed path in $\overrightarrow{H}^+$ from $x$ to $a$ of length $s\le r\le k$.
+Then $\overrightarrow{W}$ is a directed path in $\overrightarrow{H}^+$ from $x$ to $a$ of length at most $s\le r\le k$.
 This completes the proof.
@@
 This increases the layered width given by \cref{general-width} from $d\ell(k^2+3)$ to $d\ell(k^3+3k)$.
-Therefore $G$ has an $H$-partition of layered width at most $d\ell(k^3+3k)$ in which $H$ has treewidth at most $\binom{k+t}{t}$, completing the proof of \cref{ShortcutPartition}.
+Therefore $G$ has an $H$-partition of layered width at most $d\ell(k^3+3k)$ in which $H$ has treewidth at most $\binom{k+t}{t}-1$, completing the proof of \cref{ShortcutPartition}.
 \end{proof}
@@
 \subsection{Framed Graphs}
+
+\note{PM}{I don't like this structure anymore.
+The proof of \cref{kPlanarProduct} ($k$-planar) refers to the innards of the proof of \cref{ShortcutPartition} ($(k,d)$-shortcuts).
+The proofs should be close to each other.
+
+Let's consider something like this:
+\begin{compactenum}
+  \item Introduction
+  \item Shortcut systems and $k$-planar graphs
+  \item All graphs obtained from shortcut systems or $k$-planar graphs
+  \item $d$-framed graphs, $1$-planar graphs, and map graphs
+  \item All applications (colouring, labelling, etc).
+\end{compactenum}
+% (Possibly with the order of 2 and 3 reversed.)
+}
 
@@
 \note{DW}{I suggest $G'$ be changed to $G_0$ throughout (since $G'$ feels like a graph derived from $G$, which it is not). Similarly, replace $N'$ by $N_0$.}
+
+\note{PM}{I would also like to change $G'$ to $G_0$ and $N'$ to $N_0$.
+One issue is holding me back, though: We recursive on cycles $F_1,\ldots,F_m$ that define subgraphs $N_1,\ldots,N_m$ of $G$ and subgraphs $N_1',\ldots,N_m'$ of $G'$.
+Will these latter subgraphs becomes $N_{0,1},\ldots,N_{0,m}$?}
 
@@
 			\item $F[P_i]$ is connected; and
-			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
-			\note{DW}{suggested revision: ``$P_i$ has a partition $\{X_i,Y_i\}$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.''}
+			\item $P_i$ has a partition $\{X_i,Y_i\}$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
 		\end{compactenum}
-		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
+		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the interior or boundary of $F$.
 	\end{compactenum}
@@
 		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
-		\item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$ \note{DW}{suggested revision, ``$S_x$ has a partition $\{X_x,Y_x\}$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$''};
+		\item For each $x\in V(H)$, $S_x$ has a partition $\{X_x,Y_x\}$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
 		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
@@
 	The proof is by induction on the number of inner faces of $N'$.  A particular easy case occurs when $V(N')=V(F)$, for which we can simply take
-	\note{DW}{Suggested revision, ``In the base case, $N'$ has only one inner face, and we can simply take ...} 	
+	\note{DW}{Suggested revision, ``In the base case, $N'$ has only one inner face, and we can simply take ...} \note{PM}{I debated what to put here since, even when $N'$ has only one face, the rules below work for picking $\tau$.
+The \emph{real} base case occurs when $N'$ has zero faces, but that's not something that ever happens since we only ever apply induction on the inner faces of $M$ (except $\tau$).
+I settled on this ``base case'' since it's clear that, in this case, the partition $\mathcal{P}$ is already decided.}
 	$\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
@@
 In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that will be determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
-We first explain how $v_1$, $v_2$ and $v_3$ are chosen and explain how these are used to define $S$.
+We first explain how $v_1$, $v_2$ and $v_3$ are chosen and then explain how these are used to define $S$.
 There is one case to consider for each possible value of $k$ (see \cref{boring_figure}):
@@
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
-Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N'$, add edges $d'-2$
-		\note{DW}{should be $d'-3$, right?} edges inside of $F$ to split it into $3$-sided faces.
+Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N'$, add $d'-3$ edges inside of $F$ to split it into $3$-sided faces.
 This yields a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary.
 By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
-Let $\tau:=v_1,\ldots,v_p$ be the face \note{DW}{``be the facial cycle''} of $N'$ that contains $\tau''$.
+Let $\tau:=v_1,\ldots,v_p$ be  the facial cycle of $N'$ that contains $\tau''$.
 	\end{compactenum}
@@
 Therefore $|Y\cap L_j|\le 3$ for each integer $j\ge 0$.
-Furthermore, $|X|\le p-3\le d-3$, since $v_1,v_2,v_3\in Y\cup V(F)$.
-Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ will not violate condition~(ii) \note{DW}{Use ``satisfies'' instead of ``will not violate''}.
+Furthermore, $X\subseteq V(\tau)\setminus\{x_1,x_2,x_3\}$, so $|X|\le p-3\le d-3$.
+Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ satisfies condition~(ii).
 
@@
 	Since $M$ is $2$-connected and $\overline{S}$ is connected, $F_i[P_{i,j}]$ is connected, for each $j\in\{1,2,3\}$.
-Let $N_i$ and $N_i'$ be the subgraphs of $N$ and $N'$ contained in the closure of $F_i$.
+Let $N_i$ and $N_i'$ be the subgraphs of $N$ and $N'$ contained in the interior or boundary of $F_i$.
 Every inner face of $N_i'$ is an inner face of $N'$ and $\tau$ is not an inner face of $N_i'$.
@@
 	By construction $\mathcal{P}$ contains parts $P_1,\ldots,P_k$, so $\mathcal{P}$ satisifes (i).
-By Assumption~4(b) none of the parts $P_1,P_2,P_3$ violate condition~(ii).
-We have already argued that the part $S$ does not violate condition~(ii).
+By Assumption~4(b) each of the parts $P_1,P_2,P_3$ satisfies condition~(ii).
+We have already argued that the part $S$ satisfies condition~(ii).
 The inductive hypothesis ensures that each of the remaining parts satisfies condition~(ii).
@@
 Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
-Since $a_i$, $b_i$, $c_i$ form a clique in $H$, there exists a bag $B_{z_i}$ in $\mathcal{T}_i$ that contains $\{a_i,b_i,c_i\}$.
+Since $a_i$, $b_i$, $c_i$ form a clique in $H_i$, there exists a bag $B_{z_i}$ in $\mathcal{T}_i$ that contains $\{a_i,b_i,c_i\}$.
 
@@
 
-\cref{NewMapGraph} is an immediate consequence of \cref{d_framed_product_stucture}} and the next lemma.
+\cref{NewMapGraph} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma.
 Similar results to \cref{MewMapGraphLemma} appear in earlier works \cite{CGP06,BDGGMR,Brandenburg19,Brandenburg20} but we state the precise lemma and provide a proof here for the sake of completeness.

### 2022-03-19 slides/slides.tex
Pfffft

@@
 
-\title{Graph Product Structure for Non-Minor-Closed Classes}
-\author{Vida Dujmović \and Pat Morin \and David R. Wood}
+\title[$k$-planar product structure]{Graph Product Structure for Non-Minor-Closed Classes}
+\author[Dujmović, M, Wood]{Vida Dujmović \and Pat Morin \and David R. Wood}
 \titlegraphic{\includegraphics[height=1em]{by}}
@@
 
-\begin{frame}
-    \frametitle{Outline}
-
-    \begin{itemize}
-        \item $k$-planar graphs
-        \item $(k,d)$-shortcut systems
-        \item statement of main theorem
-        \item applications
-        \item proof of main theorem
-        \item open problem
-    \end{itemize}
-\end{frame}
+% \begin{frame}
+%     \frametitle{Outline}
+%
+%     \begin{itemize}
+%         \item $k$-planar graphs
+%         \item $(k,d)$-shortcut systems
+%         \item statement of main theorem
+%         \item applications
+%         \item proof of main theorem
+%         \item open problem
+%     \end{itemize}
+% \end{frame}
 
@@
     \begin{itemize}
+      \item $G'\subseteq G^{\mathcal{P}}$, with $G$ planar and $\mathcal{P}$ a $(k+1,2)$-shortcut system for $G$
       \item $t=3$, $\ell=3$, $d=2$, $k={\color{red}k}+1$
@@
 \begin{frame}
+  \frametitle{Progress so Far}
+
+  \begin{itemize}
+    \item $\mathcal{P'}$ is a $(k,dk\ell)$-shortcut system over $H\boxtimes P$
+    \item So far, $G^{\mathcal{P}}\subseteq (H\boxtimes P)^\mathcal{P'}\boxtimes K_{\ell k}$
+    \item Shortcuts in $\mathcal{P'}$ project onto edges of $P$
+    \item Shortcuts in $\mathcal{P'}$ project onto paths in $H$
+    \item Time to study $\tw(H')$ where $H':=H^{\mathcal{P}}$
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
   \frametitle{Nice Tree Decompositions}
@@
 
-  \textbf{Lemma (Pilipczuk--Siebertz):}\footnote{Yes, the same paper!} For any $z\in V(H)$ the number of $x\in V(H)$ such that $x\prec_T z\prec_T z'$ and $\mathop{dist}_H(x,z')\le k$ is at most $\binom{k+t}{t}$.
+  \begin{itemize}
+    \item $(B_x:x\in V(T))$ is a nice width-$t$ tree decomposition of $H$.
+
+    \item[]\textbf{Lemma (Pilipczuk--Siebertz):}\footnote{Yes, the same paper!} For any $z\in V(T)$ the number of $x\in V(T)$ such that $x\prec_T z\prec_T z'$ and $\mathop{dist}_H(x,z')\le k$ is at most $\binom{k+t}{t}$.
     \begin{center}
@@
     \end{center}
+  \end{itemize}
   \begin{itemize}
-      \item<2->Problem: $vw$-shortcut in $H\boxtimes P$ projects onto $xz'$ shortcut in $H$, but $x\not\preceq_T z'$ and $z'\not\preceq_T x$
+      \item<4->Problem: $vw$-shortcut in projects onto $xz'$ shortcut in $H$, but $x\not\preceq_T z'$ and $z'\not\preceq_T x$
   \end{itemize}
@@
     \item<2-> $\mathcal{S}:=(S_x:x\in V(T))$ is a partition of $V(G)$
-    \item<3->[$(\star)$] For any $vw$-shortcut in $H\boxtimes P$, $a(v)\preceq_T a(w)$ or $a(w)\preceq_T a(v)$
+    \item<3->[$(\star)$] For any $vw$-shortcut, $a(v)\preceq_T a(w)$ or $a(w)\preceq_T a(v)$
 \end{itemize}
@@
   \begin{itemize}
-      \item[$(\star)$] For any $vw$-shortcut in $H\boxtimes P$, $a(v)\preceq_T a(w)$ or $a(w)\preceq_T a(v)$
+      \item[$(\star)$] For any $vw$-shortcut, $a(v)\preceq_T a(w)$ or $a(w)\preceq_T a(v)$
   \end{itemize}

### 2022-03-19 k-planar.tex
new map graph apps theorem

@@
 \begin{thm}
-\label{NewMapGraph}
+\label{MapGraphProduct}
 Every $d$-map graph is contained in $H \boxtimes P \boxtimes K_{ d + 3\floor{d/2} -3 }$ for some planar graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
@@
 
-\cref{NewMapGraph} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma.
+\cref{MapGraphProduct} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma.
 Similar results to \cref{MewMapGraphLemma} appear in earlier works \cite{CGP06,BDGGMR,Brandenburg19,Brandenburg20} but we state the precise lemma and provide a proof here for the sake of completeness.
@@
 
-\note{DW}{give a better version of \cref{MapPartition} with $g=0$. }
+\cref{MapGraphProduct,qn,p-centered,non-repetitive} imply the following result.
+
+\begin{thm}
+\label{PlaneMapApps}
+Every $d$-map graph $G$:
+	\begin{compactitem}
+		\item has queue-number $\qn(G) \leq 24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$.
+		\item has $p$-centered chromatic number $\chi_p(G) \leq (d+3\floor{d/2}-3) (p+1)\,  \binom{p+3}{3}$,
+		\item has non-repetitive chromatic number $ \pi(G) \leq 256(d+3\floor{d/2}-3) $.
+	\end{compactitem}
+\end{thm}
 

### 2022-03-19 k-planar.tex
Trying a restructuring

--- 2022-03-19 k-planar.tex
@@
 
-The following theorem is the main contribution of the paper.
+The following theorem is one of the main contributions of the paper.
 It says that if a graph class $\mathcal{G}$ has a product structure theorem, then the class of graphs obtained by applying a shortcut system to graphs in $\mathcal{G}$ also has a product structure theorem.
@@
 Some further observations presented in \cref{sec-k-planar} lead to the improved result.
-
 \cref{ShortcutProduct} is applicable for many graph classes in addition to $k$-planar graphs.
-Here is one example.
+Some examples are explored in \cref{first_example,mid_example,last_example}.
+
+% Here is one example.
 The \defin{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.\footnote{For a graph $G$ and two vertices $v,w\in V(G)$, $\dist_G(v,w)$  denotes the length of a shortest path, in $G$, with endpoints $v$ and $w$.
@@
 
-\begin{thm}
-\label{kPowerBasic}
-For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is contained in $H\boxtimes P\boxtimes K_{6k^2(k^2+3)\Delta^{k}}$ for some graph $H$ of treewidth at most $\binom{k+3}{3}-1$ and some path $P$.
-\end{thm}
-
-\cref{Examples} presents further examples of graph classes that can be constructed using shortcut systems, including certain types of map graphs, string graphs, and $k$-nearest neighbour graphs.
+% \begin{thm}
+% \label{kPowerBasic}
+% For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is contained in $H\boxtimes P\boxtimes K_{6k^2(k^2+3)\Delta^{k}}$ for some graph $H$ of treewidth at most $\binom{k+3}{3}-1$ and some path $P$.
+% \end{thm}
+
+% \cref{Examples} presents further examples of graph classes that can be constructed using shortcut systems, including certain types of map graphs, string graphs, and $k$-nearest neighbour graphs.
 \cref{ShortcutProduct} implies product structure theorems for each of these classes.
 All of the above-mentioned applications also hold for these examples.
-
-\subsection{Generalisations}
-\label{Generalisations}
-
-As mentioned above, product structure theorems have been established for several minor-closed classes in addition to planar graphs.
-The first generalises \cref{PlanarProduct} for graphs of bounded Euler genus.
-
-\begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
-\label{GenusProduct}
-Every graph of Euler genus $g$ is contained in:
-\begin{compactenum}[(a)]
-\item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
-\item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
-\end{compactenum}
-\end{thm}
-
-\citet{DJMMUW20} generalised \cref{GenusProduct} for apex-minor-free graphs as follows.
-
-\begin{thm}[\citep{DJMMUW20}]
-\label{ApexMinorFree}
-For every apex graph $X$, there exists $c\in\mathbb{N}$ such that every $X$-minor-free graph is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and some path $P$.
-\end{thm}
-
-The assumption that $X$ is apex is needed in \cref{ApexMinorFree}, since if the class of $X$-minor-free graphs has a product structure theorem analogous to \cref{PlanarProduct}, then $X$ is apex \citep{DJMMUW20}.
-On the other hand, \citet{DEMWW22} proved a product structure theorem for bounded degree graphs in any minor-closed class.
-
-\begin{thm}[\citep{DEMWW22}]
-	\label{MinorFreeDegree}
-	For every graph $X$ there exists $c\in\mathbb{N}$ such that for every $\Delta\in\mathbb{N}$, every $X$-minor-free graph $G$ with maximum degree at most $\Delta$ is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H) \leq c\Delta$ and for some path $P$.
-\end{thm}
-
-%The next three results are immediate corollaries of \cref{GenusProduct,ApexMinorFree,ShortcutProduct,MinorFreeDegree}.
-%
-%\begin{thm}
-%Let $\SS$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$.
-Then $G^\SS$ is contained in $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
-%\end{thm}
-%
-%\begin{thm}
-%For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\SS$ for $G$, $G^\SS\subseteq H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
-%\end{thm}
-%
-%\begin{thm}
-%For every graph $X$ and for all integers $k,\Delta\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ with maximum degree $\Delta$ and for every $(k,\Delta)$-shortcut system $\SS$ for $G$, $G^\SS \subseteq H \boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
-%\end{thm}
-%
-%\note{DW}{delete the above three theorems, they are never used}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Shortcut Systems}
+\section{\boldmath Shortcut Systems and $k$-Planar Graphs}
 \label{Structure}
 
-The purpose of this section is to prove our main technical result, \cref{ShortcutPartition}.
-This theorem shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, an $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
+The purpose of this section is to prove our main technical result, \cref{ShortcutPartition}, and its specialization to $k$-planar graphs \cref{kPlanarProduct}.
+
+\subsection{Product Structure for Shortcut Systems}
+
+\cref{ShortcutPartition} shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, an $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
 
@@
 
-We are now ready to prove our main result, which we restate here for convenience:
-
-\mmg*
-
-\begin{proof}
+We are now ready to prove our main result.
+% , which we restate here for convenience:
+
+% \mmg*
+
+\begin{proof}[Proof of \cref{ShortcutPartition}]
 Let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:=(Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
@@
 
-%%%%%%%%%%%%%%%%%%%
-\subsection{Framed Graphs}
-
-\note{PM}{I don't like this structure anymore.
+\subsection{\boldmath Product Structure for $k$-Planar Graphs}
+
+% \note{DW}{These definitions should appear earlier, certainly before the definition of framed graph, which should be defined as an ``embedded graph''.
+One option is to put this whole paragraph as a footnote in the introduction where we first mention $k$-planar graphs.
+The experienced reader can skip this paragraph, which says to me it should be a footnote.}
+
+We first formally define $k$-planar graphs.
+An \defin{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a curve\footnote{A \defin{curve} in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
+The points $f(0)$ and $f(1)$ are called the \defin{endpoints} of the curve.
+When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.
+The \defin{interior} of $f$ is the point set $\{f(t):0<t<1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
+The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
+A \defin{crossing} in an embedded (multi)graph $G$ is an unordered pair of distinct edges $\{vw,xy\}\subseteq E(G)$ whose interiors have a point in common.
+Without any loss of generality we may assume that the interiors of any two edges of $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.
+This assumption can be enforced by local changes that do not change the number of crossings each edge is involved in.
+If $p$ is a common point in the interior of two crossing edges $vw$ and $xy$ then we say that $vw$ and $xy$ \defin{cross at} $p$.
+
+% (This assumption is already present in the informal definition of $k$-planar graphs used in \cref{k_planar_section}.)
+
+%
+% \note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}\note{PM}{No, under the current definition, $r$ edges crossing at a single point $p$ generate $\binom{r}{2}$ crossings, with each edge taking part in $r-1$ crossings, just like they would if we perturbed the edges so that each pair crosses at a distinct point.
+I guess we could mention this.
+(See the first paragraph in the proof of Theorem~2 below.)} \note{DW}{The downside of this approach is that in the opening section we give an informal definition of $k$-planar, which needs the assumption of having no three edges cross at a point, since we use this assumption in the construction of a $(k+1,2)$-shortcut system for $k$-planar graphs give in the Introduction.
+I think we should not have different definitions in two parts of the paper.} A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
+Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
+
+As mentioned in \cref{Introduction}, \cref{PlanarProduct,ShortcutProduct} imply a product structure theorem for $k$-planar graphs. We get improved bounds as follows.
+
+%\begin{thm}
+%\label{k-planar}
+%Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 48k+30$ in which $\tw(H)\leq \binom{k+4}{3}-1$.
+%\end{thm}
+
+\begin{proof}[Proof of \cref{kPlanarProduct}]
+	Let $G$ be a $k$-plane graph.
+    % We assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$.
+This assumption is justified since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
+	As in the proof of \cref{AddDummy}, let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$.
+In this way, each edge $vw\in E(G)$ corresponds naturally to a path $P_{vw}$ of length at most $k+1$ in $G_0$.
+Let $\SS := \{P_{vw}: vw\in E(G)\}$.
+Observe that $\SS$ is a $(k+1,2)$-shortcut system for $G_0$ and that $G_0^{\SS}\supseteq G$.
+Specifically, $G_0^{\SS}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.
+
+	Since $G_0$ is planar,  \cref{PlanarProduct}(b) and \cref{PartitionProduct} implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
+Applying \cref{ShortcutPartition} to $G_0$ and $\SS$ immediately implies that $G$ (an arbitrary $k$-planar graph) has an $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
+
+	We can reduce the layered width of the $H$-partition of $G$ from $O(k^3)$ to $O(k^2)$ by observing that the dummy vertices in $V(G_0)\setminus V(G)$ do not contribute to the layered width of this partition.
+In this setting, the proof of \cref{general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $L_i\cap Y_x$.
+More precisely, each path $P\in\SS$ containing an internal (dummy) vertex $w\in Y_x\cap (L_{i-j}\cup L_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap L_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap L_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap L_j$ for $j > k+1$.
+	Redoing the calculation at the end of the proof of \cref{general-width} then yields
+	\begin{align*}
+	|S_x\cap Y_i| \le d\ell\left(
+	2
+	+ 4\left\lfloor\tfrac{k+1}{2}\right\rfloor
+	+ 2\left\lceil\tfrac{k+1}{2}\right\rceil
+	\right)
+	 =
+	d\ell\left(
+	2 + 2(k+1) + 2\left\lfloor\tfrac{k+1}{2}\right\rfloor
+	\right)
+	 \le
+	d\ell(3k+5)
+	= 18k+30 \enspace .
+	\end{align*}
+	With this change, the layered width of the partition given by \cref{ShortcutPartition} becomes $(18k+30)(k+1)=18k^2+48k+30$.
+	The result follows from \cref{PartitionProduct}.
+	%This establishes \cref{k-planar}.
+	% Note to self-and-others. In our original SODA submission we had 18k^2+30k here because we forgot that the layer compression step has compress k+1 consecutive layers, not just k consecutive layers.
+\end{proof}
+
+% \subsection{Examples}
+
+%%%%%%%
+\subsection{String Graphs}
+\label{first_example}
+
+A \defin{string graph} is the intersection graph of a set of curves in the plane with no three curves meeting at a single point; see  \cite{PachToth-DCG02,FP10,FP14} for example.
+For an integer $\delta\geq 2$, if each curve is in at most $\delta$ intersections with other curves, then the corresponding string graph is called a \defin{$\delta$-string graph}.
+A \defin{$(g,\delta)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
+
+\begin{lem}
+\label{StringShortcut}
+Every $(g,\delta)$-string graph $G$ is contained in $G_0^\SS$ for some graph $G_0$ with Euler genus at most $g$ and some $(\delta+1,\delta+1 )$-shortcut system $\SS$ for $G_0$.
+%Moreover, every internal vertex on the paths in $\SS$ has degree 4 in $G_0$.
+\end{lem}
+
+\begin{proof}
+Let $\mathcal{C}=\{C_v:v\in V(G)\}$ be a set of curves in a surface of Euler genus at most $g$ whose intersection graph is $G$.
+Let $G_0$ be the graph obtained by adding a vertex at the intersection point of every pair of curves in $\mathcal{C}$ that intersect,  where two such consecutive vertices on a curve $C_v$ are adjacent in $G_0$.
+For each vertex $v\in V(G)$, if $C_v$ intersects $k\leq\delta$ other curves, then introduce a new vertex called $v$ on $C_v$ between the
+$\floor{\frac{k}{2}}$-th vertex already on $C_v$ and the $\floor{\frac{k}{2}+1}$-th such vertex.
+For each edge $vw$ of $G$, there is a path $P_{vw}$ of length at most $2\ceil{\frac{\delta}{2}}\leq \delta+1$ in $G_0$ between $v$ and $w$.
+Let $\SS$ be the set of all such paths $P_{vw}$.
+Consider a vertex $x$ in $G_0$ that is an internal vertex on some path in $\SS$.
+Then $x$ is at the intersection of $C_v$ and $C_w$ for some edge $vw\in E(G)$.
+If some path $P\in \SS$ passes through $x$, then $P=P_{vu}$ for some edge $vu$ incident to $v$, or $P=P_{wu}$ for some edge $wu$ incident to $w$.
+At most $\ceil{\frac{\delta}{2}}$ paths in $\SS$ corresponding to edges incident to $v$ pass through $x$, and similarly for edges incident to $w$.
+Thus at most $2\ceil{\frac{\delta}{2}}\leq\delta+1$ paths in $\SS$ use $x$ as an internal vertex.
+Thus $\SS$ is a $(\delta+1,\delta+1)$-shortcut system for $G_0$, and by construction, $G \subseteq G_0^\SS$.
+\end{proof}
+
+A similar proof to that of \cref{StringShortcut} shows that every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.
+Theorems~\ref{PlanarProduct}(b), \ref{ShortcutProduct}  and \ref{GenusProduct}(b) and \cref{StringShortcut,qn} imply:
+
+\begin{thm}\label{StringPartition}
+For integers $g\geq 0$ and $\delta\geq 2$, let $\ell:= \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$ and $t:= \binom{ \delta+4}{3}-1$.
+Then  every $(g,\delta)$-string graph is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$,
+\end{thm}
+
+% \begin{thm}
+% \label{StringPartition}
+% For integers $g\geq 0$ and $\delta\geq 2$, let $\ell:= \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$ and $t:= \binom{ \delta+4}{3}-1$.
+% Then  every $(g,\delta)$-string graph:
+% \begin{compactitem}
+% 	\item is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$,
+% 	\item has queue-number $\qn(G^k) \leq 3 \ell \, 2^t - \ceil{\tfrac{3}{2}\ell}$.
+% \end{compactitem}
+% \end{thm}
+
+% \note{DW}{\cref{StringPartition} has been updated.}
+
+% Our results also give bounds on the non-repetitive chromatic number and the $p$-centered chromatic number of $(g,\delta)$-string graphs, but the bounds are weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ and $\chi_p(G)\le p(64\delta)^2$ by results of \citet{DJKW16} and \citet{DFMS21}, respectively.
+
+%%%%%%%%%%
+\subsection{Powers of Bounded Degree Graphs}
+\label{Powers}
+\label{mid_example}
+
+Recall that the \defin{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
+If $G$ is planar with maximum degree $\Delta$, then $G^k$ is $2k(k+1)\Delta^{k}$-planar by \cref{DrawG}.
+Thus we can immediately conclude that bounded powers of planar graphs of bounded degree admit bounded layered partitions.
+However, the bounds we obtain are improved by the following lemma that constructs a shortcut system directly.
+
+\begin{lem}
+\label{PowerShortcut}
+If a graph $G$ has maximum degree $\Delta$, then $G^k = G^\SS$ for some $(k,2k \Delta^{k})$-shortcut system $\SS$.
+\end{lem}
+
+\begin{proof}
+For each pair of vertices $x$ and $y$ in $G$ with $\dist_G(x,y)\in\{1,\dots,k\}$, fix an $xy$-path $P_{xy}$ of length
+$\dist_G(x,y)$  in $G$.
+Let $\SS:=\{P_{xy}: \dist_G(x,y)\in\{1,\dots,k\} \}$.
+Say $P_{xy}$ uses some vertex $v$ as an internal vertex.
+If $\dist_G(v,x)=i$ and $\dist_G(v,y)=j$, then $i,j\in\{1,\dots,k-1\}$ and $i+j\leq k$.
+The number of vertices at distance $i$ from $v$ is at most $\Delta^i$.
+Thus the number of paths in $\SS$ that use $v$ as an internal vertex is at most
+\[\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i\Delta^j
+= \sum_{i=1}^{k-1} \Delta^i \sum_{j=1}^{k-i} \Delta^j
+< \sum_{i=1}^{k-1} \Delta^i ( 2 \Delta^{k-i} )
+< 2k \Delta^k\enspace.\]
+Hence $\SS$ is a $(k, 2k \Delta^k)$-shortcut system.
+\end{proof}
+
+\cref{ShortcutProduct,PowerShortcut} imply:
+
+\begin{thm}
+\label{PowerProduct}
+Let $G$ be contained in $H\boxtimes P\boxtimes K_\ell$ with maximum degree $\Delta$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
+Then for every integer $k\geq 1$, the $k$-th power $G^k$ is contained in $J\boxtimes P\boxtimes K_{2k \ell \Delta^{k}(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
+\end{thm}
+
+% Theorems~\ref{PlanarProduct}(b), \ref{GenusProduct}(b) and \cref{ShortcutProduct} and  \cref{DrawG,qn,p-centered,non-repetitive,PowerShortcut} imply the following result, which with $g=0$ implies  \cref{kPowerBasic} in the introduction.
+%
+% \begin{thm}
+% \label{PowerGenus}
+% For integers $g\geq 0$ and $k,\Delta\geq 1$, let $\ell:= \max\{2g,3\} (2k^4+6k^2) \Delta^{k}$ and $t:= \binom{k+3}{3}-1$. Then for every graph $G$ of Euler genus $g$ and maximum degree $\Delta$,
+% \begin{compactitem}
+% \item $G^k$ is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$
+% \item $G^k$ is $(g, 2k(k+1)\Delta^{k} )$-planar,
+% \item $G^k$ has queue-number $\qn(G^k) \leq 3 \ell \cdot 2^t - \ceil{\tfrac{3}{2}\ell}$.
+% \item $G^k$ has  $p$-centered chromatic number $\chi_p(G^k) \leq \ell (p+1)\,  \binom{p+t}{t}$,
+% \item $G^k$ has  non-repetitive chromatic number $ \pi(G^k) \leq \ell \, 4^{t+1}$.
+% \end{compactitem}
+% \end{thm}
+%
+% % \note{DW}{\cref{PowerGenus} has been updated}
+%
+% This result is the first constant upper bound on the queue-number of bounded powers of graphs with bounded degree and bounded Euler genus.
+For every graph $G$, since $G^k$ has maximum degree at most $\Delta^k$, a result of \citet{DJKW16} implies that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$.
+\cref{PowerGenus} improves upon this bound when $k,g\ll\Delta$.
+Similarly, a result of \citet{DFMS21} implies that $\chi_p(G^k)\le 1024p\Delta^{2k}$ and \cref{PowerGenus} improves upon this bound when $p,k,g\ll\Delta$.
+%
+% \cref{qn,p-centered,non-repetitive,MinorFreeDegree,PowerProduct} imply the following analogous result for powers of graphs in any minor-closed class with bounded maximum degree.
+%
+% \begin{thm}
+% \label{PowerMinor}
+% For every graph $X$ there exists an integer $c$ such that for all integers $k,\Delta\geq 1$, if $t:= 2k\Delta^{k}(k^3+3k)\binom{k+c\Delta}{c\Delta}-1$ and $G$ is an $X$-minor-free graph with maximum degree $\Delta$, then:
+% \begin{compactitem}
+% 	\item $G^k$ is contained in $H\boxtimes P$ for some graph $H$ with treewidth $t$ and for some path $P$.
+% 	\item $G^k$ has queue-number at most $3\cdot 2^t-2$,
+% 	\item $G^k$ has $p$-centered chromatic number at most $(p+1)\binom{p+t}{t}$.
+% \end{compactitem}
+% \end{thm}
+
+\subsection{$k$-Nearest-Neighbour Graphs}
+\label{last_example}
+
+In this section, we show that $k$-nearest neighbour graphs of point sets in the plane are $O(k^2)$-planar.
+For two points $x,y\in\R^2$, let $d_2(x,y)$ denote the Euclidean distance between $x$ and $y$.
+The $k$-nearest-neighbour graph of a point set $P\subset\R^2$ is the geometric graph $G$ with vertex set $V(G)=P$, where the edge set is defined as follows.
+For each point $v\in P$, let $N_k(v)$ be the set of $k$ points in $P$ closest to $v$.
+Then $vw\in E(G)$ if and only if $w\in N_k(v)$ or $v\in N_k(w)$.
+(The edges of $G$ are straight-line segments joining their endpoints.) See \citep{ProximityGraphs} for a survey of results on $k$-nearest neighbour graphs and other related proximity graphs.
+
+The following result, which is immediate from \citet[Corollary~4.2.6]{abrego.munroy.ea:on} states that $k$-nearest-neighbour graphs have bounded maximum degree:
+\begin{lem}
+\label{k-nn-max-degree}
+The degree of every vertex in a $k$-nearest-neighbour graph is at most $6k$.
+\end{lem}
+
+We make use of the following well-known observation (see for example, \citet[Lemma~2]{bose.morin.ea:routing}):
+\begin{obs}
+\label{convex}
+If $v_0,\ldots,v_3$ are the vertices of a convex quadrilateral in counterclockwise order then there exists at least one $i\in\{0,\ldots,3\}$ such that $\max\{d_2(v_i,v_{i-1}), d_2(v_i,v_{i+1})\} < d_2(v_{i-1},v_{i+1})$, where subscripts are taken modulo 4.
+\end{obs}
+
+\begin{lem}
+\label{nearest-neighbour}
+  Every $k$-nearest-neighbour graph is $O(k^2)$-planar.
+\end{lem}
+
+\begin{proof}
+  Let $G$ be a $k$-nearest-neighbour graph and consider any edge $vw\in E(G)$.
+  Let $xy\in E(G)$ be an edge that crosses $vw$.  Note that $vxwy$ are the vertices of a convex quadrilateral in (without loss of generality) counter-clockwise order. Then we say that
+  \begin{compactenum}
+    \item $xy$ is of Type~$v$ if $\max\{d_2(v,x), d_2(v,y)\}< d_2(x,y)$;
+    \item $xy$ is of Type~$w$ if $\max\{d_2(w,x), d_2(w,y)\}< d_2(x,y)$; or
+    \item $xy$ is of Type~C otherwise.
+  \end{compactenum}
+  If $xy$ is of Type~C, then \cref{convex} implies that $\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)$ without loss of generality.
+   In this case, we call $x$ a Type~C vertex.
+We claim that $V(G)$ contains at most $k-1$ Type~C vertices.
+Indeed, more than $k-1$ Type~C vertices would contradict the fact that $vw\in E(G)$ since every Type~C vertex is closer to both $v$ and $w$ than $d_2(v,w)$.
+
+  Next observe that, if $xy$ is of Type~$v$, then at least one of $xv$ or $yv$ is in $E(G)$ in which case we call $x$ (respectively $y$) a Type~$v$ vertex.
+By \cref{k-nn-max-degree}, there are at most $6k$ Type~$v$ vertices.
+Similarly, there are at most $6k$ Type~$w$ vertices.
+
+  Thus, in total, there are at most $13k-1$ Type~$v$, Type~$w$, and Type~C vertices.
+By \cref{k-nn-max-degree}, each of these vertices is incident with at most $6k$ edges that cross $vw$.
+Therefore, there are at most $78k^2-6k$ edges of $G$ that cross $vw$.
+Since this is true for every edge $vw\in E(G)$, $G$ is $(78k^2-6k)$-planar.
+\end{proof}
+
+Note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n\ge k+1$ vertices has at least $kn/2$ edges and at most $kn$ edges.
+For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is $\Omega(k^2)$.
+
+\cref{nearest-neighbour} and \cref{kPlanarProduct} immediately imply
+
+\begin{thm}
+  Every $k$-nearest neighbour graph is a subgraph of $H\boxtimes P\boxtimes K_{\ell}$ for some graph $H$ of treewidth $O(k^6)$ and some $\ell\in O(k^4)$.
+\end{thm}
+
+\section{Framed Graphs, $1$-Planar Graphs, and Map Graphs}
+
+% Next we consider
+%
+% %%%%%%%%%%%%%%%%%%%
+% \subsection{Framed Graphs}
+
+% \note{PM}{I don't like this structure anymore.
 The proof of \cref{kPlanarProduct} ($k$-planar) refers to the innards of the proof of \cref{ShortcutPartition} ($(k,d)$-shortcuts).
 The proofs should be close to each other.
-
-Let's consider something like this:
-\begin{compactenum}
-  \item Introduction
-  \item Shortcut systems and $k$-planar graphs
-  \item All graphs obtained from shortcut systems or $k$-planar graphs
-  \item $d$-framed graphs, $1$-planar graphs, and map graphs
-  \item All applications (colouring, labelling, etc).
-\end{compactenum}
-% (Possibly with the order of 2 and 3 reversed.)
-}
+%
+% Let's consider something like this:
+% \begin{compactenum}
+%   \item Introduction
+%   \item Shortcut systems and $k$-planar graphs
+%   \item All graphs obtained from shortcut systems or $k$-planar graphs
+%   \item $d$-framed graphs, $1$-planar graphs, and map graphs
+%   \item All applications (colouring, labelling, etc).
+% \end{compactenum}
+% % (Possibly with the order of 2 and 3 reversed.)
+% }
 
@@
 
-%%%%%%%%%%%%%%%%%%%
-\section{Allowing Crossings}
-\label{sec-k-planar}
-
-This section applies our main results for shortcut systems to prove a product structure theorem for graphs drawn with a bounded number of crossings per edge.
-Then we show how to improve the bounds in this case.
-
-\subsection{$k$-Planar Graphs}
-
-\note{DW}{These definitions should appear earlier, certainly before the definition of framed graph, which should be defined as an ``embedded graph''.
-One option is to put this whole paragraph as a footnote in the introduction where we first mention $k$-planar graphs.
-The experienced reader can skip this paragraph, which says to me it should be a footnote.}
-We first formally define $k$-planar graphs.
-An \defin{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a curve\footnote{A \defin{curve} in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
-The points $f(0)$ and $f(1)$ are called the \defin{endpoints} of the curve.
-When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.
-The \defin{interior} of $f$ is the point set $\{f(t):0<t<1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
-The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
-A \defin{crossing} in an embedded (multi)graph $G$ is an unordered pair of distinct edges $\{vw,xy\}\subseteq E(G)$ whose interiors have a point in common.
-Without any loss of generality we may assume that the interiors of any two edges of $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.
-This assumption can be enforced by local changes that do not change the number of crossings each edge is involved in.
-If $p$ is a common point in the interior of two crossing edges $vw$ and $xy$ then we say that $vw$ and $xy$ \defin{cross at} $p$.
-
-% (This assumption is already present in the informal definition of $k$-planar graphs used in \cref{k_planar_section}.)
-
-%
-% \note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}\note{PM}{No, under the current definition, $r$ edges crossing at a single point $p$ generate $\binom{r}{2}$ crossings, with each edge taking part in $r-1$ crossings, just like they would if we perturbed the edges so that each pair crosses at a distinct point.
-I guess we could mention this.
-(See the first paragraph in the proof of Theorem~2 below.)} \note{DW}{The downside of this approach is that in the opening section we give an informal definition of $k$-planar, which needs the assumption of having no three edges cross at a point, since we use this assumption in the construction of a $(k+1,2)$-shortcut system for $k$-planar graphs give in the Introduction.
-I think we should not have different definitions in two parts of the paper.} A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
-Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
-
-As mentioned in \cref{Introduction}, \cref{PlanarProduct,ShortcutProduct} imply a product structure theorem for $k$-planar graphs. We get improved bounds as follows.
-
-%\begin{thm}
-%\label{k-planar}
-%Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 48k+30$ in which $\tw(H)\leq \binom{k+4}{3}-1$.
-%\end{thm}
-
-\begin{proof}[Proof of \cref{kPlanarProduct}]
-	Let $G$ be a $k$-plane graph.
-    % We assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$.
-This assumption is justified since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
-	As in the proof of \cref{AddDummy}, let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$.
-In this way, each edge $vw\in E(G)$ corresponds naturally to a path $P_{vw}$ of length at most $k+1$ in $G_0$.
-Let $\SS := \{P_{vw}: vw\in E(G)\}$.
-Observe that $\SS$ is a $(k+1,2)$-shortcut system for $G_0$ and that $G_0^{\SS}\supseteq G$.
-Specifically, $G_0^{\SS}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.
-
-	Since $G_0$ is planar,  \cref{PlanarProduct}(b) and \cref{PartitionProduct} implies that $G_0$ has an $H$-partition of layered width 3 for some planar graph $H$ of treewidth at most 3.
-Applying \cref{ShortcutPartition} to $G_0$ and $\SS$ immediately implies that $G$ (an arbitrary $k$-planar graph) has an $H$-partition of layered width $6((k+1)^3+3(k+1))$ for some graph $H$ of treewidth at most $\binom{k+4}{3}-1$.
-
-	We can reduce the layered width of the $H$-partition of $G$ from $O(k^3)$ to $O(k^2)$ by observing that the dummy vertices in $V(G_0)\setminus V(G)$ do not contribute to the layered width of this partition.
-In this setting, the proof of \cref{general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $L_i\cap Y_x$.
-More precisely, each path $P\in\SS$ containing an internal (dummy) vertex $w\in Y_x\cap (L_{i-j}\cup L_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap L_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap L_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap L_j$ for $j > k+1$.
-	Redoing the calculation at the end of the proof of \cref{general-width} then yields
-	\begin{align*}
-	|S_x\cap Y_i| \le d\ell\left(
-	2
-	+ 4\left\lfloor\tfrac{k+1}{2}\right\rfloor
-	+ 2\left\lceil\tfrac{k+1}{2}\right\rceil
-	\right)
-	 =
-	d\ell\left(
-	2 + 2(k+1) + 2\left\lfloor\tfrac{k+1}{2}\right\rfloor
-	\right)
-	 \le
-	d\ell(3k+5)
-	= 18k+30 \enspace .
-	\end{align*}
-	With this change, the layered width of the partition given by \cref{ShortcutPartition} becomes $(18k+30)(k+1)=18k^2+48k+30$.
-	The result follows from \cref{PartitionProduct}.
-	%This establishes \cref{k-planar}.
-	% Note to self-and-others. In our original SODA submission we had 18k^2+30k here because we forgot that the layer compression step has compress k+1 consecutive layers, not just k consecutive layers.
-\end{proof}
-
 \subsection{\boldmath $1$-Planar Graphs}
@@
 
-\begin{thm}
-\label{1-planar}
-Every 1-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
-\end{thm}
+%
+% \note{PM}{Use restateable here...}
+%
+% \begin{thm}
+% \label{1-planar}
+% Every 1-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
+% \end{thm}
 
@@
 
-\cref{1-planar} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma.
+\cref{1_planar_product} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma.
 Similar results appear in earlier works \cite{CGP06,BDGGMR,Brandenburg19,Brandenburg20} but we state the precise lemma and provide a proof here for the sake of completeness.
@@
 \end{proof}
+
+\subsection{Map Graphs}
+
+Map graphs are defined as follows.
+Start with a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
+Let $G$ be the graph whose vertices are the nations of $G_0$, where two vertices are adjacent in $G$ if the corresponding faces in $G_0$ share a vertex.
+Then $G$ is called a \defin{$(g,d)$-map graph}.
+A $(0,d)$-map graph is called a (plane) \defin{$d$-map graph}; see \citep{FLS-SODA12,CGP02} for example.
+The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.
+
+\note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph.}
+
+\begin{thm}
+\label{NewMapGraph}
+Every $d$-map graph is contained in $H \boxtimes P \boxtimes K_{ d + 3\floor{d/2} -3 }$ for some planar graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
+\end{thm}
+
+\cref{NewMapGraph} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma.
+Similar results to \cref{MewMapGraphLemma} appear in earlier works \cite{CGP06,BDGGMR,Brandenburg19,Brandenburg20} but we state the precise lemma and provide a proof here for the sake of completeness.
+
+\begin{lem}
+\label{MewMapGraphLemma}
+For all integers $g\geq 0$ and $d\geq 3$, every $(g,d)$-map graph is a spanning subgraph of a $(g,d)$-framed graph.
+\end{lem}
+
+\begin{proof}
+Let $G_0$ be a graph embedded in a surface of Euler genus $g$, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations.
+If two lakes have a common edge in their boundaries, then we may delete the edges and merge the lakes.
+Now assume that no two lakes have a common edge in their boundaries.
+Let $G$ be the corresponding map graph.
+Let $G_1$ be the graph with $V(G_1):=V(G)$, which is the set of nations.
+Consider vertices $v$ and $w$ of $G_1$.
+Let $A$ and $B$ be the nation faces of $G_0$ corresponding to $v$ and $w$.
+Then $vw\in E(G_1)$ if $A$ and $B$ have a common edge in their boundaries,
+or there is a vertex $x$ of $G_0$ in the boundary of $A$ and $B$, and in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes.
+Then $G_1$ has Euler genus at most $g$.
+The faces of $G_1$ correspond to vertices of $G_0$ and lakes of $G_0$.
+Triangulate each face of $G_1$ corresponding to a lake of $G_0$. For each vertex
+$x$ of $G_0$ the size of the face of $G_1$ corresponding to $x$ equals the number of nations incident to $x$, which is at most $d$.
+So every face of $G_1$ has size at most $d$.
+If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common, implying that  $v$ and $w$ are both on the face of $G_1$ corresponding to $x$, and $vw\in E(\widehat{G_1})$.
+Hence $G\subseteq \widehat{G_1}$.
+\end{proof}
+
+\note{DW}{give a better version of \cref{MapPartition} with $g=0$. }
+
+%So $(g,d)$-map graphs generalise graphs embedded in a surface, and we now assume that $d\geq 4$ for the remainder of this section.
+
+There is a natural drawing of a map graph obtained by positioning each vertex of $G$ inside the corresponding nation and each edge of $G$ as a curve passing through the corresponding vertex of $G_0$.
+It is easily seen that each edge is in at most $\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}}$ crossings; see \citep{dujmovic.eppstein.ea:structure}.
+Thus $G$ is $(g,\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
+Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar.
+\cref{gkPlanarProduct} then establishes a product structure theorem for map graphs, but we get much better bounds by constructing a shortcut system directly.
+The following lemma is reminiscent of the characterisation of $(g,d)$-map graphs in terms of the half-square of bipartite graphs \citep{CGP02,dujmovic.eppstein.ea:structure}.
+
+\begin{lem}
+\label{MapShortcut}
+Every $(g,d)$-map graph $G$ is contained in $G_1^\SS$ for some graph $G_1$ with Euler genus at most $g$ and some $(2,\tfrac12 d(d-3) )$-shortcut system $\SS$ for $G_1$.
+\end{lem}
+
+\begin{proof}
+Let $G$ be a $(g,d)$-map graph.
+So there is a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
+Let $N$ be the set of nations.
+Then $V(G)=N$ where two vertices are adjacent in $G$ if the corresponding nation faces of $G_0$ share a vertex.
+Let $G_1$ be the graph with $V(G_1):=V(G_0) \cup N$, where distinct vertices $v,w\in N$ are adjacent in $G_1$ if the boundaries of the corresponding nations have an edge of $G_0$ in common, and $v\in V(G_0)$ and $w\in N$ are adjacent in $G_1$ if $v$ is on the boundary of the nation corresponding to $w$.
+Observe that $G_1$ embeds in the same surface as $G_0$ with no crossings, and that each vertex in $V(G_0)$ has degree at most $d$ in $G_1$.
+Consider an edge $vw\in E(G)$.
+If the nations corresponding to $v$ and $w$ share an edge of $G_0$, then $vw$ is an edge of $G_1$.
+Otherwise,  $v$ and $w$ have a common neighbour $x$ in $V(G_0)$.
+In the latter case, let $P_{vw}$ be the path $(v,x,w)$.
+Let $\SS$ be the set of all such paths $P_{vw}$.
+Each vertex $x\in V(G_0)$ is the middle vertex on at most $\tfrac12 d(d-3)$  paths in $\SS$.
+Thus $\SS$ is a $(2,\tfrac12 d(d-3))$-shortcut system for $G_1$, and by construction, $G \subseteq G_1^\SS$.
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%
+\section{Allowing Crossings}
+\label{sec-k-planar}
+
+This section applies our main results for shortcut systems to prove a product structure theorem for graphs drawn with a bounded number of crossings per edge.
+Then we show how to improve the bounds in this case.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Examples}
-\label{Examples}
-
-This section describes several examples of graph classes that can be obtained from a shortcut system typically applied to graphs of bounded Euler genus.
-
-\subsection{Map Graphs}
-
-Map graphs are defined as follows.
-Start with a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
-Let $G$ be the graph whose vertices are the nations of $G_0$, where two vertices are adjacent in $G$ if the corresponding faces in $G_0$ share a vertex.
-Then $G$ is called a \defin{$(g,d)$-map graph}.
-A $(0,d)$-map graph is called a (plane) \defin{$d$-map graph}; see \citep{FLS-SODA12,CGP02} for example.
-The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.
-
-\note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph.}
-
-\begin{thm}
-\label{NewMapGraph}
-Every $d$-map graph is contained in $H \boxtimes P \boxtimes K_{ d + 3\floor{d/2} -3 }$ for some planar graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
-\end{thm}
-
-\cref{NewMapGraph} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma.
-Similar results to \cref{MewMapGraphLemma} appear in earlier works \cite{CGP06,BDGGMR,Brandenburg19,Brandenburg20} but we state the precise lemma and provide a proof here for the sake of completeness.
-
-\begin{lem}
-\label{MewMapGraphLemma}
-For all integers $g\geq 0$ and $d\geq 3$, every $(g,d)$-map graph is a spanning subgraph of a $(g,d)$-framed graph.
-\end{lem}
-
-\begin{proof}
-Let $G_0$ be a graph embedded in a surface of Euler genus $g$, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations.
-If two lakes have a common edge in their boundaries, then we may delete the edges and merge the lakes.
-Now assume that no two lakes have a common edge in their boundaries.
-Let $G$ be the corresponding map graph.
-Let $G_1$ be the graph with $V(G_1):=V(G)$, which is the set of nations.
-Consider vertices $v$ and $w$ of $G_1$.
-Let $A$ and $B$ be the nation faces of $G_0$ corresponding to $v$ and $w$.
-Then $vw\in E(G_1)$ if $A$ and $B$ have a common edge in their boundaries,
-or there is a vertex $x$ of $G_0$ in the boundary of $A$ and $B$, and in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes.
-Then $G_1$ has Euler genus at most $g$.
-The faces of $G_1$ correspond to vertices of $G_0$ and lakes of $G_0$.
-Triangulate each face of $G_1$ corresponding to a lake of $G_0$. For each vertex
-$x$ of $G_0$ the size of the face of $G_1$ corresponding to $x$ equals the number of nations incident to $x$, which is at most $d$.
-So every face of $G_1$ has size at most $d$.
-If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common, implying that  $v$ and $w$ are both on the face of $G_1$ corresponding to $x$, and $vw\in E(\widehat{G_1})$.
-Hence $G\subseteq \widehat{G_1}$.
-\end{proof}
-
-\note{DW}{give a better version of \cref{MapPartition} with $g=0$. }
-
-%So $(g,d)$-map graphs generalise graphs embedded in a surface, and we now assume that $d\geq 4$ for the remainder of this section.
-
-There is a natural drawing of a map graph obtained by positioning each vertex of $G$ inside the corresponding nation and each edge of $G$ as a curve passing through the corresponding vertex of $G_0$.
-It is easily seen that each edge is in at most $\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}}$ crossings; see \citep{dujmovic.eppstein.ea:structure}.
-Thus $G$ is $(g,\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
-Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar.
-\cref{gkPlanarProduct} then establishes a product structure theorem for map graphs, but we get much better bounds by constructing a shortcut system directly.
-The following lemma is reminiscent of the characterisation of $(g,d)$-map graphs in terms of the half-square of bipartite graphs \citep{CGP02,dujmovic.eppstein.ea:structure}.
-
-\begin{lem}
-\label{MapShortcut}
-Every $(g,d)$-map graph $G$ is contained in $G_1^\SS$ for some graph $G_1$ with Euler genus at most $g$ and some $(2,\tfrac12 d(d-3) )$-shortcut system $\SS$ for $G_1$.
-\end{lem}
-
-\begin{proof}
-Let $G$ be a $(g,d)$-map graph.
-So there is a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
-Let $N$ be the set of nations.
-Then $V(G)=N$ where two vertices are adjacent in $G$ if the corresponding nation faces of $G_0$ share a vertex.
-Let $G_1$ be the graph with $V(G_1):=V(G_0) \cup N$, where distinct vertices $v,w\in N$ are adjacent in $G_1$ if the boundaries of the corresponding nations have an edge of $G_0$ in common, and $v\in V(G_0)$ and $w\in N$ are adjacent in $G_1$ if $v$ is on the boundary of the nation corresponding to $w$.
-Observe that $G_1$ embeds in the same surface as $G_0$ with no crossings, and that each vertex in $V(G_0)$ has degree at most $d$ in $G_1$.
-Consider an edge $vw\in E(G)$.
-If the nations corresponding to $v$ and $w$ share an edge of $G_0$, then $vw$ is an edge of $G_1$.
-Otherwise,  $v$ and $w$ have a common neighbour $x$ in $V(G_0)$.
-In the latter case, let $P_{vw}$ be the path $(v,x,w)$.
-Let $\SS$ be the set of all such paths $P_{vw}$.
-Each vertex $x\in V(G_0)$ is the middle vertex on at most $\tfrac12 d(d-3)$  paths in $\SS$.
-Thus $\SS$ is a $(2,\tfrac12 d(d-3))$-shortcut system for $G_1$, and by construction, $G \subseteq G_1^\SS$.
-\end{proof}
-
-Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,non-repetitive} imply the following result.
-
-%\begin{thm}
-%\label{PlaneMapPartition}
-%\note{DW}{For every integer $d\geq 4$, }
-%Every $d$-map graph $G$:
-%	\begin{compactitem}
-%		\item is contained in $H \boxtimes P \boxtimes K_{21d(d-3)}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
-%		\item has queue-number $\qn(G) < 32225\, d(d-3)$.
-%		\item has $p$-centered chromatic number $\chi_p(G) \leq 21d(d-3) (p+1)  \binom{p+9}{9}$,
-%		\item has non-repetitive chromatic number $ \pi(G) \leq 21 \cdot 4^{10} d(d-3)$.
-%	\end{compactitem}
-%\end{thm}
-
-\begin{thm}
-\label{MapPartition}
-For integers $g\geq 0$ and $d\geq 4$, if $\ell:=  7d(d-3)\, \max\{2g,3\}$ then every $(g,d)$-map graph $G$:
-\begin{compactitem}
-\item is contained in $H \boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
-\item has queue-number $\qn(G) <  1535\, \ell $.
-\item has $p$-centered chromatic number $\chi_p(G) \leq \ell (p+1)\,  \binom{p+9}{9}$,
-\item has non-repetitive chromatic number $ \pi(G) \leq 4^{10}\,\ell $.
-\end{compactitem}
-\end{thm}
-
-These results give the first constant upper bound on the non-repetitive chromatic number of map graphs, the first polynomial bounds on the $p$-centered chromatic number of map graphs, and the best known bounds on the queue-number of map graphs.
-
-% \note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the method in \citep{DHHW}.
-I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
-Okay?}
-
-%%%%%%%
-\subsection{String Graphs}
-
-A \defin{string graph} is the intersection graph of a set of curves in the plane with no three curves meeting at a single point; see  \cite{PachToth-DCG02,FP10,FP14} for example.
-For an integer $\delta\geq 2$, if each curve is in at most $\delta$ intersections with other curves, then the corresponding string graph is called a \defin{$\delta$-string graph}.
-A \defin{$(g,\delta)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
-
-\begin{lem}
-\label{StringShortcut}
-Every $(g,\delta)$-string graph $G$ is contained in $G_0^\SS$ for some graph $G_0$ with Euler genus at most $g$ and some $(\delta+1,\delta+1 )$-shortcut system $\SS$ for $G_0$.
-%Moreover, every internal vertex on the paths in $\SS$ has degree 4 in $G_0$.
-\end{lem}
-
-\begin{proof}
-Let $\mathcal{C}=\{C_v:v\in V(G)\}$ be a set of curves in a surface of Euler genus at most $g$ whose intersection graph is $G$.
-Let $G_0$ be the graph obtained by adding a vertex at the intersection point of every pair of curves in $\mathcal{C}$ that intersect,  where two such consecutive vertices on a curve $C_v$ are adjacent in $G_0$.
-For each vertex $v\in V(G)$, if $C_v$ intersects $k\leq\delta$ other curves, then introduce a new vertex called $v$ on $C_v$ between the
-$\floor{\frac{k}{2}}$-th vertex already on $C_v$ and the $\floor{\frac{k}{2}+1}$-th such vertex.
-For each edge $vw$ of $G$, there is a path $P_{vw}$ of length at most $2\ceil{\frac{\delta}{2}}\leq \delta+1$ in $G_0$ between $v$ and $w$.
-Let $\SS$ be the set of all such paths $P_{vw}$.
-Consider a vertex $x$ in $G_0$ that is an internal vertex on some path in $\SS$.
-Then $x$ is at the intersection of $C_v$ and $C_w$ for some edge $vw\in E(G)$.
-If some path $P\in \SS$ passes through $x$, then $P=P_{vu}$ for some edge $vu$ incident to $v$, or $P=P_{wu}$ for some edge $wu$ incident to $w$.
-At most $\ceil{\frac{\delta}{2}}$ paths in $\SS$ corresponding to edges incident to $v$ pass through $x$, and similarly for edges incident to $w$.
-Thus at most $2\ceil{\frac{\delta}{2}}\leq\delta+1$ paths in $\SS$ use $x$ as an internal vertex.
-Thus $\SS$ is a $(\delta+1,\delta+1)$-shortcut system for $G_0$, and by construction, $G \subseteq G_0^\SS$.
-\end{proof}
-
-A similar proof to that of \cref{StringShortcut} shows that every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.
-Theorems~\ref{PlanarProduct}(b), \ref{ShortcutProduct}  and \ref{GenusProduct}(b) and \cref{StringShortcut,qn} imply:
-
-\begin{thm}
-\label{StringPartition}
-For integers $g\geq 0$ and $\delta\geq 2$, let $\ell:= \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$ and $t:= \binom{ \delta+4}{3}-1$.
-Then  every $(g,\delta)$-string graph:
-\begin{compactitem}
-	\item is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$,
-	\item has queue-number $\qn(G^k) \leq 3 \ell \, 2^t - \ceil{\tfrac{3}{2}\ell}$.
-\end{compactitem}
-\end{thm}
-
-% \note{DW}{\cref{StringPartition} has been updated.}
-
-Our results also give bounds on the non-repetitive chromatic number and the $p$-centered chromatic number of $(g,\delta)$-string graphs, but the bounds are weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ and $\chi_p(G)\le p(64\delta)^2$ by results of \citet{DJKW16} and \citet{DFMS21}, respectively.
-
-%%%%%%%%%%
-\subsection{Powers of Bounded Degree Graphs}
-\label{Powers}
-
-Recall that the \defin{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.
-If $G$ is planar with maximum degree $\Delta$, then $G^k$ is $2k(k+1)\Delta^{k}$-planar by \cref{DrawG}.
-Thus we can immediately conclude that bounded powers of planar graphs of bounded degree admit bounded layered partitions.
-However, the bounds we obtain are improved by the following lemma that constructs a shortcut system directly.
-
-\begin{lem}
-\label{PowerShortcut}
-If a graph $G$ has maximum degree $\Delta$, then $G^k = G^\SS$ for some $(k,2k \Delta^{k})$-shortcut system $\SS$.
-\end{lem}
-
-\begin{proof}
-For each pair of vertices $x$ and $y$ in $G$ with $\dist_G(x,y)\in\{1,\dots,k\}$, fix an $xy$-path $P_{xy}$ of length
-$\dist_G(x,y)$  in $G$.
-Let $\SS:=\{P_{xy}: \dist_G(x,y)\in\{1,\dots,k\} \}$.
-Say $P_{xy}$ uses some vertex $v$ as an internal vertex.
-If $\dist_G(v,x)=i$ and $\dist_G(v,y)=j$, then $i,j\in\{1,\dots,k-1\}$ and $i+j\leq k$.
-The number of vertices at distance $i$ from $v$ is at most $\Delta^i$.
-Thus the number of paths in $\SS$ that use $v$ as an internal vertex is at most
-\[\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i\Delta^j
-= \sum_{i=1}^{k-1} \Delta^i \sum_{j=1}^{k-i} \Delta^j
-< \sum_{i=1}^{k-1} \Delta^i ( 2 \Delta^{k-i} )
-< 2k \Delta^k\enspace.\]
-Hence $\SS$ is a $(k, 2k \Delta^k)$-shortcut system.
-\end{proof}
-
-\cref{ShortcutProduct,PowerShortcut} imply:
-
-\begin{thm}
-\label{PowerProduct}
-Let $G$ be contained in $H\boxtimes P\boxtimes K_\ell$ with maximum degree $\Delta$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
-Then for every integer $k\geq 1$, the $k$-th power $G^k$ is contained in $J\boxtimes P\boxtimes K_{2k \ell \Delta^{k}(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
-\end{thm}
-
-Theorems~\ref{PlanarProduct}(b), \ref{GenusProduct}(b) and \cref{ShortcutProduct} and  \cref{DrawG,qn,p-centered,non-repetitive,PowerShortcut} imply the following result, which with $g=0$ implies  \cref{kPowerBasic} in the introduction.
-
-\begin{thm}
-\label{PowerGenus}
-For integers $g\geq 0$ and $k,\Delta\geq 1$, let $\ell:= \max\{2g,3\} (2k^4+6k^2) \Delta^{k}$ and $t:= \binom{k+3}{3}-1$. Then for every graph $G$ of Euler genus $g$ and maximum degree $\Delta$,
-\begin{compactitem}
-\item $G^k$ is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$
-\item $G^k$ is $(g, 2k(k+1)\Delta^{k} )$-planar,
-\item $G^k$ has queue-number $\qn(G^k) \leq 3 \ell \cdot 2^t - \ceil{\tfrac{3}{2}\ell}$.
-\item $G^k$ has  $p$-centered chromatic number $\chi_p(G^k) \leq \ell (p+1)\,  \binom{p+t}{t}$,
-\item $G^k$ has  non-repetitive chromatic number $ \pi(G^k) \leq \ell \, 4^{t+1}$.
-\end{compactitem}
-\end{thm}
-
-% \note{DW}{\cref{PowerGenus} has been updated}
-
-This result is the first constant upper bound on the queue-number of bounded powers of graphs with bounded degree and bounded Euler genus.
-For every graph $G$, since $G^k$ has maximum degree at most $\Delta^k$, a result of \citet{DJKW16} implies that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$.
-\cref{PowerGenus} improves upon this bound when $k,g\ll\Delta$.
-Similarly, a result of \citet{DFMS21} implies that $\chi_p(G^k)\le 1024p\Delta^{2k}$ and \cref{PowerGenus} improves upon this bound when $p,k,g\ll\Delta$.
-
-\cref{qn,p-centered,non-repetitive,MinorFreeDegree,PowerProduct} imply the following analogous result for powers of graphs in any minor-closed class with bounded maximum degree.
-
-\begin{thm}
-\label{PowerMinor}
-For every graph $X$ there exists an integer $c$ such that for all integers $k,\Delta\geq 1$, if $t:= 2k\Delta^{k}(k^3+3k)\binom{k+c\Delta}{c\Delta}-1$ and $G$ is an $X$-minor-free graph with maximum degree $\Delta$, then:
-\begin{compactitem}
-	\item $G^k$ is contained in $H\boxtimes P$ for some graph $H$ with treewidth $t$ and for some path $P$.
-	\item $G^k$ has queue-number at most $3\cdot 2^t-2$,
-	\item $G^k$ has $p$-centered chromatic number at most $(p+1)\binom{p+t}{t}$.
-\end{compactitem}
-\end{thm}
-
-\subsection{$k$-Nearest-Neighbour Graphs}
-
-In this section, we show that $k$-nearest neighbour graphs of point sets in the plane are $O(k^2)$-planar.
-For two points $x,y\in\R^2$, let $d_2(x,y)$ denote the Euclidean distance between $x$ and $y$.
-The $k$-nearest-neighbour graph of a point set $P\subset\R^2$ is the geometric graph $G$ with vertex set $V(G)=P$, where the edge set is defined as follows.
-For each point $v\in P$, let $N_k(v)$ be the set of $k$ points in $P$ closest to $v$.
-Then $vw\in E(G)$ if and only if $w\in N_k(v)$ or $v\in N_k(w)$.
-(The edges of $G$ are straight-line segments joining their endpoints.) See \citep{ProximityGraphs} for a survey of results on $k$-nearest neighbour graphs and other related proximity graphs.
-
-The following result, which is immediate from \citet[Corollary~4.2.6]{abrego.munroy.ea:on} states that $k$-nearest-neighbour graphs have bounded maximum degree:
-\begin{lem}
-\label{k-nn-max-degree}
-The degree of every vertex in a $k$-nearest-neighbour graph is at most $6k$.
-\end{lem}
-
-We make use of the following well-known observation (see for example, \citet[Lemma~2]{bose.morin.ea:routing}):
-\begin{obs}
-\label{convex}
-If $v_0,\ldots,v_3$ are the vertices of a convex quadrilateral in counterclockwise order then there exists at least one $i\in\{0,\ldots,3\}$ such that $\max\{d_2(v_i,v_{i-1}), d_2(v_i,v_{i+1})\} < d_2(v_{i-1},v_{i+1})$, where subscripts are taken modulo 4.
-\end{obs}
-
-\begin{lem}
-\label{nearest-neighbour}
-  Every $k$-nearest-neighbour graph is $O(k^2)$-planar.
-\end{lem}
-
-\begin{proof}
-  Let $G$ be a $k$-nearest-neighbour graph and consider any edge $vw\in E(G)$.
-  Let $xy\in E(G)$ be an edge that crosses $vw$.  Note that $vxwy$ are the vertices of a convex quadrilateral in (without loss of generality) counter-clockwise order. Then we say that
-  \begin{compactenum}
-    \item $xy$ is of Type~$v$ if $\max\{d_2(v,x), d_2(v,y)\}< d_2(x,y)$;
-    \item $xy$ is of Type~$w$ if $\max\{d_2(w,x), d_2(w,y)\}< d_2(x,y)$; or
-    \item $xy$ is of Type~C otherwise.
-  \end{compactenum}
-  If $xy$ is of Type~C, then \cref{convex} implies that $\max\{d_2(x,v),d_2(x,w)\} < d_2(v,w)$ without loss of generality.
-   In this case, we call $x$ a Type~C vertex.
-We claim that $V(G)$ contains at most $k-1$ Type~C vertices.
-Indeed, more than $k-1$ Type~C vertices would contradict the fact that $vw\in E(G)$ since every Type~C vertex is closer to both $v$ and $w$ than $d_2(v,w)$.
-
-  Next observe that, if $xy$ is of Type~$v$, then at least one of $xv$ or $yv$ is in $E(G)$ in which case we call $x$ (respectively $y$) a Type~$v$ vertex.
-By \cref{k-nn-max-degree}, there are at most $6k$ Type~$v$ vertices.
-Similarly, there are at most $6k$ Type~$w$ vertices.
-
-  Thus, in total, there are at most $13k-1$ Type~$v$, Type~$w$, and Type~C vertices.
-By \cref{k-nn-max-degree}, each of these vertices is incident with at most $6k$ edges that cross $vw$.
-Therefore, there are at most $78k^2-6k$ edges of $G$ that cross $vw$.
-Since this is true for every edge $vw\in E(G)$, $G$ is $(78k^2-6k)$-planar.
-\end{proof}
-
-Note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n\ge k+1$ vertices has at least $kn/2$ edges and at most $kn$ edges.
-For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is $\Omega(k^2)$.
-
-\cref{nearest-neighbour,kPlanarProduct,qn,p-centered} imply:
-
-\begin{cor}
-\label{k-nn}
-For every integer $k\geq 1$ there exists integers $t\leq O(k^6)$ and $\ell\leq O(k^4)$ such that every $k$-nearest-neighbour graph:
-\begin{compactitem}
-\item is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $t$ and some path $P$,
-\item has queue-number at most $2^{O(k^6)}$, and
-\item has $p$-centered chromatic number at most $\ell (p+1)\binom{p+t}{t}$.
-\end{compactitem}
-\end{cor}
-
-\cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
-However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (36+o(1))k^2$ by a result of \citet{DJKW16}.
-
-\section{Wrapping Up}
-\label{WrappingUp}
+\subsection{Universal Graphs}
 
@@
 
+\section{Wrapping Up}
+\label{WrappingUp}
+
 We finish with an open problem.
@@
 Note that $C\geq 3$ even for planar graphs \citep{DJMMUW20}.
+
+This section describes several examples of graph classes that can be obtained from a shortcut system typically applied to graphs of bounded Euler genus.
+
+\section{Orphaned Theorems}
+
+Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,non-repetitive} imply the following result.
+
+%\begin{thm}
+%\label{PlaneMapPartition}
+%\note{DW}{For every integer $d\geq 4$, }
+%Every $d$-map graph $G$:
+%	\begin{compactitem}
+%		\item is contained in $H \boxtimes P \boxtimes K_{21d(d-3)}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
+%		\item has queue-number $\qn(G) < 32225\, d(d-3)$.
+%		\item has $p$-centered chromatic number $\chi_p(G) \leq 21d(d-3) (p+1)  \binom{p+9}{9}$,
+%		\item has non-repetitive chromatic number $ \pi(G) \leq 21 \cdot 4^{10} d(d-3)$.
+%	\end{compactitem}
+%\end{thm}
+
+\begin{thm}
+\label{MapPartition}
+For integers $g\geq 0$ and $d\geq 4$, if $\ell:=  7d(d-3)\, \max\{2g,3\}$ then every $(g,d)$-map graph $G$:
+\begin{compactitem}
+\item is contained in $H \boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
+\item has queue-number $\qn(G) <  1535\, \ell $.
+\item has $p$-centered chromatic number $\chi_p(G) \leq \ell (p+1)\,  \binom{p+9}{9}$,
+\item has non-repetitive chromatic number $ \pi(G) \leq 4^{10}\,\ell $.
+\end{compactitem}
+\end{thm}
+
+These results give the first constant upper bound on the non-repetitive chromatic number of map graphs, the first polynomial bounds on the $p$-centered chromatic number of map graphs, and the best known bounds on the queue-number of map graphs.
+
+% \note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the method in \citep{DHHW}.
+I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
+Okay?}
+
+\section{Orphaned Section}
+
+\cref{nearest-neighbour,kPlanarProduct,qn,p-centered} imply:
+
+\begin{cor}
+\label{k-nn}
+For every integer $k\geq 1$ there exists integers $t\leq O(k^6)$ and $\ell\leq O(k^4)$ such that every $k$-nearest-neighbour graph:
+\begin{compactitem}
+\item is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $t$ and some path $P$,
+\item has queue-number at most $2^{O(k^6)}$, and
+\item has $p$-centered chromatic number at most $\ell (p+1)\binom{p+t}{t}$.
+\end{compactitem}
+\end{cor}
+
+\cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
+However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (36+o(1))k^2$ by a result of \citet{DJKW16}.
+
+\section{Orphaned Subsection}
+
+\subsection{Generalisations}
+\label{Generalisations}
+
+As mentioned above, product structure theorems have been established for several minor-closed classes in addition to planar graphs.
+The first generalises \cref{PlanarProduct} for graphs of bounded Euler genus.
+
+\begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
+\label{GenusProduct}
+Every graph of Euler genus $g$ is contained in:
+\begin{compactenum}[(a)]
+\item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
+\item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
+\end{compactenum}
+\end{thm}
+
+\citet{DJMMUW20} generalised \cref{GenusProduct} for apex-minor-free graphs as follows.
+
+\begin{thm}[\citep{DJMMUW20}]
+\label{ApexMinorFree}
+For every apex graph $X$, there exists $c\in\mathbb{N}$ such that every $X$-minor-free graph is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and some path $P$.
+\end{thm}
+
+The assumption that $X$ is apex is needed in \cref{ApexMinorFree}, since if the class of $X$-minor-free graphs has a product structure theorem analogous to \cref{PlanarProduct}, then $X$ is apex \citep{DJMMUW20}.
+On the other hand, \citet{DEMWW22} proved a product structure theorem for bounded degree graphs in any minor-closed class.
+
+\begin{thm}[\citep{DEMWW22}]
+	\label{MinorFreeDegree}
+	For every graph $X$ there exists $c\in\mathbb{N}$ such that for every $\Delta\in\mathbb{N}$, every $X$-minor-free graph $G$ with maximum degree at most $\Delta$ is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H) \leq c\Delta$ and for some path $P$.
+\end{thm}
+
+%The next three results are immediate corollaries of \cref{GenusProduct,ApexMinorFree,ShortcutProduct,MinorFreeDegree}.
+%
+%\begin{thm}
+%Let $\SS$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$.
+Then $G^\SS$ is contained in $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
+%\end{thm}
+%
+%\begin{thm}
+%For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\SS$ for $G$, $G^\SS\subseteq H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
+%\end{thm}
+%
+%\begin{thm}
+%For every graph $X$ and for all integers $k,\Delta\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ with maximum degree $\Delta$ and for every $(k,\Delta)$-shortcut system $\SS$ for $G$, $G^\SS \subseteq H \boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
+%\end{thm}
+%
+%\note{DW}{delete the above three theorems, they are never used}
 

### 2022-03-22 k-planar.tex
Pfffft
Lots of stuff
Working through......

@@
 
-These theorems have applications in diverse areas, which we explore in \cref{Applications}.
+% A planar graph of treewidth at most $3$ is said to have \emph{simple treewidth} $3$.
+
+For at least one application, the planarity of $H$ is critical in obtaining asymptotically tight bounds \cite{BDJM}.
+\cref{1_planar_product} is proven in the more general setting of $d$-framed multigraphs.
+This more general result implies a product structure theorem for $d$-map graphs in which $H$ is also planar.
+
+% 
+% 
+% These theorems have applications in diverse areas, which we explore in \cref{Applications}.
 
@@
 
-Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$.
+Theorem~\ref{PlanarProduct}(b), \cref{ShortcutProduct}, and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$.
 Some further observations presented in \cref{sec-k-planar} lead to the improved result.
@@
 
+\subsection{Outline}
+
+The remainder of this paper is organized as follows: 
+\begin{compactitem}
+  \item In \cref{Structure} we prove our main result for shortcut systems (\cref{ShortcutProduct}), and its optimisations for $k$-planar graphs and their genus-$g$ generalisation (\cref{kPlanarProduct,gkPlanarProduct}).
+  \item \Cref{examples} uses \cref{Structure} to derive product structure theorems for string graphs, powers of bounded-degree graphs, and $k$-nearest neighbour graphs.
+  \item \Cref{FramedSection} proves a product structure theorem for $d$-framed multigraphs in which the graph $H$ has treewidth $3$ and is planar.
+Using this result, we derive product structure theorems for $1$-planar graphs and map graphs, again with a planar graph $H$.
+  \item \Cref{Applications} quickly surveys applications of product structure and the consequences of the current work.  
+  \item \cref{WrappingUp} concludes with an open problem.
+\end{compactitem}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The purpose of this section is to prove our main technical result, \cref{ShortcutPartition}, and its specialization to $k$-planar graphs \cref{kPlanarProduct}.
-
-\subsection{Product Structure for Shortcut Systems}
+The purpose of this section is to prove \cref{ShortcutPartition} and its specialization to $k$-planar graphs \cref{kPlanarProduct}.
+
+\subsection{Shortcut Systems}
 
 \cref{ShortcutPartition} shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, an $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
-
 For convenience, it will be helpful to assume that $\SS$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.
@@
 
-We are now ready to prove our main result.
+% We are now ready to complete the proof.
 % , which we restate here for convenience:
@@
 
-\subsection{\boldmath Product Structure for $k$-Planar Graphs}
+\subsection{\boldmath $k$-Planar Graphs}
 
@@
 
+\subsection{\boldmath $(g,k)$-Planar Graphs}
+
+As mentioned in the introduction, product structure theorems have been established for several minor-closed classes in addition to planar graphs, including the following version of \cref{PlanarProduct} for graphs of bounded Euler genus.
+
+\begin{thm}[\citep{DJMMUW20,UWY,DHHW}]\label{GenusProduct}
+  Every graph of Euler genus $g$ is contained in:
+  \begin{compactenum}[(a)]
+  \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
+  \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
+  \end{compactenum}
+\end{thm}
+
+The definition of $k$-planar graphs naturally generalises to genus-$g$ surfaces.
+A graph $G$ embedded on a surface $\Sigma$ is $(\Sigma,k)$-plane if every edge of $G$ is involved in at most $k$ crossings.
+A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
+\cref{AddDummy} immediately generalises as follows:
+
+\begin{obs}
+\label{gAddDummy}
+Every $(g,k)$-planar graph $G$ is contained in $G_0^\SS$ for some graph $G_0$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\SS$ for $G_0$.
+Moreover, $V(G) \subseteq V(G_0)$ and for every edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most $k+1$, such that every internal vertex in $P$ has degree at most $4$ in $G_0$.
+\end{obs}
+
+Theorems~\ref{ShortcutProduct}  and \ref{GenusProduct}(b) imply a product structure theorem for $(g,k)$-planar graphs.
+The resulting bounds are improved by the following theorem, which is proved using exactly the same approach used in the proof of \cref{kPlanarProduct} (applying \cref{GenusProduct}(b) instead of \cref{PlanarProduct}(b)).
+We omit repeating the details.
+
+\begin{thm}
+\label{gkPlanarProduct}
+Every $(g,k)$-planar graph is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with $\tw(H) \leq \binom{k+4}{3}-1$, where $\ell:=\max\{2g,3\}\cdot(6k^2+16k+10)$.
+% \note{DW}{I have changed $\tw(H) \leq \binom{k+5}{4}-1$ to $\tw(H) \leq \binom{k+4}{3}-1$, okay?}
+%\note{PM}{Ok}
+\end{thm}
+
+\subsection{Rough Characterisation}
+\label{Characterisation}
+
+\cref{gAddDummy} shows that $(g,k)$-planar graphs can be obtained by a shortcut system applied to a graph of Euler genus $g$, where internal vertices on the paths have bounded degree.
+This observation and the following converse result together provide a rough characterisation of $(g,k)$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $(g,k)$-planar.
+
+\begin{lem}
+	\label{DrawG}
+	Fix integers $g\geq 0$ and $k,\Delta\geq 2$.
+	Let $G_0$ be a graph of Euler genus at most $g$. Let $G$ be
+	a graph with $V(G) \subseteq V(G_0)$ such that for every edge $vw \in
+	E(G)$ there is a $vw$-path $P_{vw}$ in $G_0$ of length at most $k$, such
+	that every internal vertex on $P_{vw}$ has degree at most $\Delta$ in
+	$G_0$. Then $G$ is $(g, 2k(k+1)\Delta^{k} )$-planar.
+\end{lem}
+
+\begin{proof}
+	For a vertex $x$ of $G_0$ with degree at most $\Delta$, and for $i\in\{1,\dots,k-1\}$, say a vertex $v$ is \defin{$i$-close} to $x$ if there is a $vx$-path $P$ in $G_0$ of length at most $i$ such that every internal vertex in $P$  has degree at most $\Delta$ in $G_0$.
+	For each edge $vw$ of $G$, say that $vw$ \defin{passes through} each internal vertex on $P_{vw}$.	Say $vw$ passes through $x$.
+Then $v$ is $i$-close to $x$ and $w$ is $j$-close to $x$ for some $i,j\in\{1,\dots,k-1\}$ with $i+j\leq k$.
+At most $\Delta^{i}$ vertices are $i$-close to $x$.
+	Thus, the number of edges of $G$ that pass through $x$ is at most
+	\[
+	\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i \Delta^j
+	= \sum_{i=1}^{k-1} \Delta^i  \sum_{j=1}^{k-i} \Delta^j
+	< \sum_{i=1}^{k-1} \Delta^i  2 \Delta^{k-i}
+	= \sum_{i=1}^{k-1} 2\Delta^k
+	< 2k \Delta^k \enspace.
+	\]
+	Fix a $(\Sigma,k)$-plane drawing of $G_0$ in a surface $\Sigma$ of Euler genus at most $g$. Choose $\epsilon>\epsilon'>0$.
+	For each vertex $v$ of $G_0$, let $B_v$ be the \defin{ball} $\{x\in \mathbb{R}^2:\dist(x,v)\leq\epsilon\}$
+	For each edge $vw$ of $G_0$, let $C_{vw}$ be the \defin{channel} $\{x\in \mathbb{R}^2:\dist(x,vw)\leq\epsilon'\}\setminus(B_v\cup B_w)$.
+	We may choose $\epsilon>\epsilon'>0$ sufficiently small so that
+	(i) $B_v\cap B_w=\emptyset$ for distinct $v,w\in V(G_0)$,
+	(ii) $C_{vw}\neq\emptyset$ for $vw\in E(G_0)$,
+	(iii) $B_v\cap C_{xy}=\emptyset$ for $v\in V(G_0)$ and $xy\in E(G_0)$, and
+	(iv) $C_{vw}\cap C_{xy}=\emptyset$ for distinct $vw,xy\in E(G_0)$.
+	Draw each edge $vw$ of $G$ following the sequence of balls and channels defined by the vertices and edges in $P_{vw}$.
+	This can be done so that whenever edges $vw$ and $xy$ of $G$ cross, the crossing point is in $B_z$ for some vertex $z$ of $G_0$ that is internal in $P_{vw}$ or in $P_{xy}$, and $vw$ and $xy$ cross at most once in each such $B_z$.
+
+  Thus, for each edge $vw$ of $G$, every edge of $G$ that crosses $vw$ passes through a vertex on $P_{vw}$ (including $v$ and/or $w$ if they too have degree at most $\Delta$).
+	Since $P_{vw}$ has at most $k+1$ vertices, and less than $2k\Delta^{k}$ edges of $G$ pass through each vertex on $P_{vw}$, the edge $vw$ is crossed by less than $2k(k+1)\Delta^{k}$ edges in $G$.
+Hence $G$ is $(g, 2k(k+1)\Delta^{k} )$-planar.
+\end{proof}
+
+%\referee{2}{Page 16.
+At the end of Page 16, I do not see why 'every pair of edges cross at most once.'  Without further assumptions, it seems as if two paths in the path system could intersect several times.
+Therefore, by drawing 'each edge vw of G alongside $P_{vw}$ in $G_0$', there may be two edges which cross several times.
+It seems this can be fixed by choosing the path system carefully by 'uncrossing' paths which intersect at more than two internal vertices.}
+%
+%\note{DW}{I have revised the proof to address this error. I couldn't think of a way to do this without introducing balls and channels (and without vigorous hand-waving). Is it readable?}
+
+\section{Graphs Derived from Shortcut Systems}
+\label{examples}
+
+In this section we give three examples of well-known graph families that can be expressed in terms of shortcut systems and which therefore have product structure theorems.
+
 %%%%%%%
@@
 A similar proof to that of \cref{StringShortcut} shows that every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.
-Theorems~\ref{PlanarProduct}(b), \ref{ShortcutProduct}  and \ref{GenusProduct}(b) and \cref{StringShortcut,qn} imply:
+Theorems~\ref{PlanarProduct}(b), \ref{ShortcutProduct}  and \ref{GenusProduct}(b):
+ % and \cref{StringShortcut,qn} imply:
 
@@
 \section{Framed Graphs, $1$-Planar Graphs, and Map Graphs}
+\label{FramedSection}
 
@@
 
-For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane \note{DW}{Use the language of ``embedded graph''} having a biconnected plane spanning (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
-The graph $G_0$ is called the \defin{frame} of $G_0$.
+For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it has an embedding in $\R^2$ having a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
+The graph $G_0$ is called the \defin{frame} of $G$.
 This definition (for simple graphs) was introduced by \citet{BDGGMR}.
-Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar multigraph (by adding a vertex inside each face $F$ adjacent to the vertices of $F$).
+
+Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar multigraph (by adding a vertex inside each face $F$ of $G_0$ adjacent to the vertices of $F$).
 Thus \cref{ShortcutProduct} is applicable.
@@
 
-This result is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs.
+\cref{d_framed_product_stucture} is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs.
 For simple 1-planar graphs it is essential that we allow frames with parallel edges.
@@
 \note{PM}{I would also like to change $G'$ to $G_0$ and $N'$ to $N_0$.
-One issue is holding me back, though: We recursive on cycles $F_1,\ldots,F_m$ that define subgraphs $N_1,\ldots,N_m$ of $G$ and subgraphs $N_1',\ldots,N_m'$ of $G'$.
+One issue is holding me back, though: We recurse on cycles $F_1,\ldots,F_m$ that define subgraphs $N_1,\ldots,N_m$ of $G$ and subgraphs $N_1',\ldots,N_m'$ of $G'$.
 Will these latter subgraphs becomes $N_{0,1},\ldots,N_{0,m}$?}
 
-\note{DW}{We need to define ``closure''}
+% \note{DW}{We need to define ``closure''}
 
@@
 
-$1$-planar graphs are of particular interest to the graph drawing community~\citep{kobourov.liotta.ea:annotated}.
+% $1$-planar graphs are of particular interest to the graph drawing community~\citep{kobourov.liotta.ea:annotated}.
 In this case, we obtain better constants and an additional property (planarity) of $H$.
@@
 \begin{lem}\label{1_planar_is_4_framed}
-A graph $G$ is $1$-planar if and only if it is contained in a $4$-framed multigraph.
+A multigraph $G$ is $1$-planar if and only if it is contained in a $4$-framed multigraph.
 \end{lem}
 
-\note{DW}{Write  ``A multigraph $G$ is $1$-planar if and only if ...''}
+% \note{DW}{Write  ``A multigraph $G$ is $1$-planar if and only if ...''}
 
@@
 \begin{proof}
-Let $G$ be a $1$-plane graph.
+Let $G$ be a $1$-plane multigraph.
 We may assume that no two edges incident to a common vertex of $G$ cross, since such a crossing can be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}.
@@
 \subsection{Map Graphs}
+
+\note{PM}{Check David's commit: a0e20aff83d12c74127331854ce9fb75e002efe3}
 
@@
 
-%%%%%%%%%%%%%%%%%%%
-\section{Allowing Crossings}
-\label{sec-k-planar}
-
-This section applies our main results for shortcut systems to prove a product structure theorem for graphs drawn with a bounded number of crossings per edge.
-Then we show how to improve the bounds in this case.
-
-\subsection{\boldmath $(g,k)$-Planar Graphs}
-
-The definition of $k$-planar graphs naturally generalises for other surfaces.
-A graph $G$ drawn on a surface $\Sigma$ is $(\Sigma,k)$-plane if every edge of $G$ is involved in at most $k$ crossings.
-A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
-\cref{AddDummy} immediately generalises as follows:
-
-\begin{obs}
-\label{gAddDummy}
-Every $(g,k)$-planar graph $G$ is contained in $G_0^\SS$ for some graph $G_0$ of Euler genus at most $g$ and some $(k+1,2)$-shortcut system $\SS$ for $G_0$.
-Moreover, $V(G) \subseteq V(G_0)$ and for every edge $vw \in E(G)$ there is a $vw$-path $P$ in $G_0$ of length at most $k+1$, such that every internal vertex in $P$ has degree at most $4$ in $G_0$.
-\end{obs}
-
-Theorems~\ref{ShortcutProduct}  and \ref{GenusProduct}(b) imply a product structure theorem for $(g,k)$-planar graphs.
-The obtained bounds are improved by the following result, which is proved using exactly the same approach used in the proof of \cref{kPlanarProduct} (applying \cref{GenusProduct}(b) instead of \cref{PlanarProduct}(b)).
-We omit repeating the details.
-
-\begin{thm}
-\label{gkPlanarProduct}
-Every $(g,k)$-planar graph is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with $\tw(H) \leq \binom{k+4}{3}-1$, where $\ell:=\max\{2g,3\}\cdot(6k^2+16k+10)$.
-% \note{DW}{I have changed $\tw(H) \leq \binom{k+5}{4}-1$ to $\tw(H) \leq \binom{k+4}{3}-1$, okay?}
-%\note{PM}{Ok}
-\end{thm}
-
-%\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{d_framed_product_stucture} with the method in \citep{DHHW}.
-I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
-Okay?}
-%\note{PM}{Okay to this and to all similar questions that come later (which I've now commented out.)}
-%\note{DW}{Robert is working on it.}
-
-Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs (or any of the other classes presented in \cref{Examples}) was in terms of layered treewidth, which we now define.
-A \defin{layered tree-decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{T}$ of $G$.
-The layered width of $(\mathcal{L},\mathcal{T})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{T}\}$.
-The \defin{layered treewidth} of $G$ is the minimum layered width of any layered tree-decomposition of $G$.
-\citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph.
-\citet{dujmovic.eppstein.ea:structure} show that every $k$-planar graph has layered treewidth at most $6(k+1)$, and more generally that every $(g,k)$-planar graph has layered treewidth at most $(4g+6)(k+1)$.
-It follows from this result that $(g,k)$-planar graphs have treewidth $O(\sqrt{(g+1)(k+1)n})$ and thus have balanced separators of the same order, which can also be concluded from the work of \citet{FP08}.
-In related work, \citet{grigoriev.bodlaender:algorithms} used structural results to obtain approximation algorithms for $(g,k)$-planar graphs, and \citet{PachToth97} determined the maximum number of edges in a $k$-planar graph (up to a constant factor).
-
-If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
-In particular, if $\PP=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree-decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree-decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{DJMMUW20}.
-Therefore, any property that holds for graphs of bounded layered treewidth also holds for $G$.
-What sets layered partitions apart from layered treewidth is that they lead to constant upper bounds on the queue-number  and non-repetitive chromatic number, whereas for both these parameters, the best known upper bound obtainable via layered treewidth is $O(\log n)$; see \cref{Applications}.
-
-\subsection{Rough Characterisation}
-\label{Characterisation}
-
-\cref{gAddDummy} shows that $(g,k)$-planar graphs can be obtained by a shortcut system applied to a graph of Euler genus $g$, where internal vertices on the paths have bounded degree.
-This observation and the following converse result together provide a rough characterisation of $(g,k)$-planar graphs, which is interesting in its own right, and is useful for showing that various classes of graphs are $(g,k)$-planar.
-
-\begin{lem}
-	\label{DrawG}
-	Fix integers $g\geq 0$ and $k,\Delta\geq 2$.
-	Let $G_0$ be a graph of Euler genus at most $g$. Let $G$ be
-	a graph with $V(G) \subseteq V(G_0)$ such that for every edge $vw \in
-	E(G)$ there is a $vw$-path $P_{vw}$ in $G_0$ of length at most $k$, such
-	that every internal vertex on $P_{vw}$ has degree at most $\Delta$ in
-	$G_0$. Then $G$ is $(g, 2k(k+1)\Delta^{k} )$-planar.
-\end{lem}
-
-\begin{proof}
-	For a vertex $x$ of $G_0$ with degree at most $\Delta$, and for $i\in\{1,\dots,k-1\}$, say a vertex $v$ is \defin{$i$-close} to $x$ if there is a $vx$-path $P$ in $G_0$ of length at most $i$ such that every internal vertex in $P$  has degree at most $\Delta$ in $G_0$.
-	For each edge $vw$ of $G$, say that $vw$ \defin{passes through} each internal vertex on $P_{vw}$.	Say $vw$ passes through $x$.
-Then $v$ is $i$-close to $x$ and $w$ is $j$-close to $x$ for some $i,j\in\{1,\dots,k-1\}$ with $i+j\leq k$.
-At most $\Delta^{i}$ vertices are $i$-close to $x$.
-	Thus, the number of edges of $G$ that pass through $x$ is at most
-	\[
-	\sum_{i=1}^{k-1} \sum_{j=1}^{k-i} \Delta^i \Delta^j
-	= \sum_{i=1}^{k-1} \Delta^i  \sum_{j=1}^{k-i} \Delta^j
-	< \sum_{i=1}^{k-1} \Delta^i  2 \Delta^{k-i}
-	= \sum_{i=1}^{k-1} 2\Delta^k
-	< 2k \Delta^k \enspace.
-	\]
-	Fix a $(\Sigma,k)$-plane drawing of $G_0$ in a surface $\Sigma$ of Euler genus at most $g$. Choose $\epsilon>\epsilon'>0$.
-	For each vertex $v$ of $G_0$, let $B_v$ be the \defin{ball} $\{x\in \mathbb{R}^2:\dist(x,v)\leq\epsilon\}$
-	For each edge $vw$ of $G_0$, let $C_{vw}$ be the \defin{channel} $\{x\in \mathbb{R}^2:\dist(x,vw)\leq\epsilon'\}\setminus(B_v\cup B_w)$.
-	We may choose $\epsilon>\epsilon'>0$ sufficiently small so that
-	(i) $B_v\cap B_w=\emptyset$ for distinct $v,w\in V(G_0)$,
-	(ii) $C_{vw}\neq\emptyset$ for $vw\in E(G_0)$,
-	(iii) $B_v\cap C_{xy}=\emptyset$ for $v\in V(G_0)$ and $xy\in E(G_0)$, and
-	(iv) $C_{vw}\cap C_{xy}=\emptyset$ for distinct $vw,xy\in E(G_0)$.
-	Draw each edge $vw$ of $G$ following the sequence of balls and channels defined by the vertices and edges in $P_{vw}$.
-	This can be done so that whenever edges $vw$ and $xy$ of $G$ cross, the crossing point is in $B_z$ for some vertex $z$ of $G_0$ that is internal in $P_{vw}$ or in $P_{xy}$, and $vw$ and $xy$ cross at most once in each such $B_z$.
-
-Thus, for each edge $vw$ of $G$, every edge of $G$ that crosses $vw$ passes through a vertex on $P_{vw}$ (including $v$ and/or $w$ if they too have degree at most $\Delta$).
-	Since $P_{vw}$ has at most $k+1$ vertices, and less than $2k\Delta^{k}$ edges of $G$ pass through each vertex on $P_{vw}$, the edge $vw$ is crossed by less than $2k(k+1)\Delta^{k}$ edges in $G$.
-Hence $G$ is $(g, 2k(k+1)\Delta^{k} )$-planar.
-\end{proof}
-
-%\referee{2}{Page 16.
-At the end of Page 16, I do not see why 'every pair of edges cross at most once.'  Without further assumptions, it seems as if two paths in the path system could intersect several times.
-Therefore, by drawing 'each edge vw of G alongside $P_{vw}$ in $G_0$', there may be two edges which cross several times.
-It seems this can be fixed by choosing the path system carefully by 'uncrossing' paths which intersect at more than two internal vertices.}
-%
-%\note{DW}{I have revised the proof to address this error. I couldn't think of a way to do this without introducing balls and channels (and without vigorous hand-waving). Is it readable?}
-
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Here we discuss some of the consequences of the above theorems for $k$-planar and $(g,k)$-planar graphs.
-
-\referee{1}{4.
+As discussed in the introduction, product structure has been used to resolve a number of problems on planar graphs.
+In most cases, these results hold for any graph with product structure.
+In this section, we survey some of the consequences of this for the graph classes considered in the previous two sections.
+
+% Here we discuss some of the consequences of the above theorems for $k$-planar and $(g,k)$-planar graphs.
+% 
+% \referee{1}{4.
 Sections 4 and 5: I think as they are currently organized they distract from the purpose of the paper, as stated in the introduction, to ``prove product structure theorems for several non-minor-closed classes of interest.'' Which of Corollaries 1--4 and Theorems 12--17 are most important?
@@
 Theorems 18 and 19 should be in the applications section.}
-
-\note{DW}{I don't want to do a complete restructuring. In our response, let's just say why we structure it the way we do.}
-\note{PM}{Sure, though I think they proposing a chop rather than a restructuring.}
+% 
+% \note{DW}{I don't want to do a complete restructuring. In our response, let's just say why we structure it the way we do.}
+% \note{PM}{Sure, though I think they proposing a chop rather than a restructuring.}
 
@@
 
-\cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$.
-% \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
-% \note{OM}{Ok}
-Note that \citet{DJMMUW20} previously proved the bound of
-$O(g^{k+2})$ using \cref{GenusProduct} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$
-% \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
-% \note{PM}{Ok}
-In the case of 1-planar graphs we can improve further.
+Since all the product structure theorems presented thus far upper bound the treewidth of $H$, 
+\cref{qn} immediately implies bounds on the queue number of all graphs in these classes.  An improvement can be made in the cases where $H$ is planar and has treewidth $3$, as follows.  
 \citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
-Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$.
-\cref{qn,1-planar} then imply:
+Thus the graph $H$ in \cref{1_planar_product} has queue-number at most $5$.
+\cref{qn,1_planar_product,NewMapGraph} then imply:
 
@@
 \end{cor}
+
+\begin{cor}
+\label{dMapQueue}
+Every $d$-map graph has queue-number at most $24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$.
+\end{cor}
+
+% \cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$.
+% \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
+% \note{OM}{Ok}
+% \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
+% \note{PM}{Ok}
+
+The following corollary summarizes the results for other graph classes.
+\begin{cor}\label{q_cor} 
+  The following bounds hold for the queue number of any graph from each of the following classes:
+  \begin{compactenum}
+    \item $k$-planar: $2^{O(k^3)}$
+    \item $(g,k)$-planar: $g2^{O(k^3)}$ 
+    \item $(g,\delta)$-string: $g2^{O(\delta^3)}$
+    \item $1$-planar: $115$
+    \item $d$-map: $24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$
+    \item $(g,d)$-map: $g2^{O(d)}$
+    \item $k$-NN: $2^{O(k^6)}$
+    % \item $(g,d)$-map graph
+    % \item 
+  \end{compactenum}
+\end{cor}
+
+Note that \citet{DJMMUW20} previously proved the bound of
+$O(g^{k+2})$ for $k$-planar graphs using \cref{GenusProduct} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$.
 
@@
 
-\cref{non-repetitive,kPlanarProduct,1-planar,gkPlanarProduct} imply the following result:
-
-\begin{cor}\quad
-\begin{compactitem}
-\item	For every $1$-planar graph $G$, $\pi(G)\le 7\times 4^4=1792$.
-%\item For every $k$-planar graph $G$,
-%	$\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
-\item	For every $(g,k)$-planar graph $G$,
-	$  \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}.$
-% \note{DW}{I have changed $4^{\binom{k+5}{4}}$ to $4^{\binom{k+4}{3}}$, okay?}
-% \note{PM}{Ok}
-\end{compactitem}
+% \cref{non-repetitive,kPlanarProduct,1-planar,gkPlanarProduct} imply the following result:
+% 
+% \begin{cor}\quad
+% \begin{compactitem}
+% \item	For every $1$-planar graph $G$, $\pi(G)\le 7\times 4^4=1792$.
+% %\item For every $k$-planar graph $G$,
+% %	$\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
+% \item	For every $(g,k)$-planar graph $G$,
+% 	$  \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}.$
+% % \note{DW}{I have changed $4^{\binom{k+5}{4}}$ to $4^{\binom{k+4}{3}}$, okay?}
+% % \note{PM}{Ok}
+% \end{compactitem}
+% \end{cor}
+
+The following corollary summarizes the results for other graph classes.
+\begin{cor}\label{q_cor} 
+  The following bounds hold for the nonrepetitive chromatic number of any graph from each of the following classes:
+  \begin{compactenum}
+    \item $k$-planar: $(18k^2+48k+30) 4^{\binom{k+4}{3}}$
+    \item $(g,k)$-planar: $\max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}$ 
+    \item $(g,\delta)$-string: $g4^{O(\delta^3)}$
+    \item $1$-planar: $1792$
+    \item $d$-map: $24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$
+    \item $(g,d)$-map: $g4^{O(d)}$
+    \item $k$-NN: $4^{O(k^6)}$
+    % \item $(g,d)$-map graph
+    % \item 
+  \end{compactenum}
 \end{cor}
@@
 
+\section{Orphaned Section on Structure}
+
+Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs (or any of the other classes presented in \cref{examples}) was in terms of layered treewidth, which we now define.
+A \defin{layered tree-decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{T}$ of $G$.
+The layered width of $(\mathcal{L},\mathcal{T})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{T}\}$.
+The \defin{layered treewidth} of $G$ is the minimum layered width of any layered tree-decomposition of $G$.
+\citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph.
+\citet{dujmovic.eppstein.ea:structure} show that every $k$-planar graph has layered treewidth at most $6(k+1)$, and more generally that every $(g,k)$-planar graph has layered treewidth at most $(4g+6)(k+1)$.
+It follows from this result that $(g,k)$-planar graphs have treewidth $O(\sqrt{(g+1)(k+1)n})$ and thus have balanced separators of the same order, which can also be concluded from the work of \citet{FP08}.
+In related work, \citet{grigoriev.bodlaender:algorithms} used structural results to obtain approximation algorithms for $(g,k)$-planar graphs, and \citet{PachToth97} determined the maximum number of edges in a $k$-planar graph (up to a constant factor).
+
+If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
+In particular, if $\PP=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree-decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree-decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{DJMMUW20}.
+Therefore, any property that holds for graphs of bounded layered treewidth also holds for $G$.
+What sets layered partitions apart from layered treewidth is that they lead to constant upper bounds on the queue-number  and non-repetitive chromatic number, whereas for both these parameters, the best known upper bound obtainable via layered treewidth is $O(\log n)$; see \cref{Applications}.
+
+%%%%%%%%%%%%%%%%%%%
+\label{sec-k-planar}
+
+%\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{d_framed_product_stucture} with the method in \citep{DHHW}.
+I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
+Okay?}
+%\note{PM}{Okay to this and to all similar questions that come later (which I've now commented out.)}
+%\note{DW}{Robert is working on it.}
+
 This section describes several examples of graph classes that can be obtained from a shortcut system typically applied to graphs of bounded Euler genus.
@@
 
-As mentioned above, product structure theorems have been established for several minor-closed classes in addition to planar graphs.
+% As mentioned above, product structure theorems have been established for several minor-closed classes in addition to planar graphs.
 The first generalises \cref{PlanarProduct} for graphs of bounded Euler genus.
-
-\begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
-\label{GenusProduct}
-Every graph of Euler genus $g$ is contained in:
-\begin{compactenum}[(a)]
-\item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
-\item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
-\end{compactenum}
-\end{thm}
+% 
+% \begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
+% \label{GenusProduct}
+% Every graph of Euler genus $g$ is contained in:
+% \begin{compactenum}[(a)]
+% \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
+% \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
+% \end{compactenum}
+% \end{thm}
 

### 2022-03-23 k-planar.tex
Made a table

@@
 
+\note{PM}{Move this section out of the introduction, into a ``preliminaries'' section?}
+
 While strong products enable concise statements of the theorems in \cref{Introduction}, to prove such results it is helpful to work with layerings and partitions, which we now introduce.
@@
 
-\subsection{Outline}
-
+\subsection{Overview and Outline}
+
+\cref{summary_table} summarizes existing results on product structure theorems for minor-closed graph classes and new results for non-minor-closed graph classes.
 The remainder of this paper is organized as follows:
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{lccl}
+      \textbf{Graph class} & \boldmath $\tw(H)$ & \boldmath $\ell$ & \textbf{Ref.} \\ \hline\hline
+      Planar & $3^*$ & $3$ & \cite{DJMMUW20} \\
+      Planar & $6^*$ & $1$ & \cite{UWY} \\
+      Genus-$g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW} \\
+      Genus-$g$ & $2g+6$ & 1 & \cite{UWY} \\
+      Apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\[1ex]
+      % B-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
+      $k$-planar & $\binom{k+4}{3}-1$ & $18k^2 + 48k + 30$ & \cref{kPlanarProduct} \\
+      $(g,k)$-planar & $\binom{k+4}{3}-1$ & $\max\{2g,3\}(6k^2+16k+10)$ & \cref{gkPlanarProduct} \\
+      $(g,\delta)$-string & $\binom{\delta+4}{3}-1$ & $\max\{2g,3\}(\delta^4 + 4\delta^3 + 9\delta^2 + 10\delta+4)$ & \cref{StringPartition} \\
+      $k$-NN & $O(k^6)$ & $O(k^4)$ & \cref{nn_product_structure} \\
+      $d$-framed & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{d_framed_product_stucture} \\
+      $1$-planar & $3^*$ & $7$ & \cref{1_planar_product} \\
+      $d$-map & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{NewMapGraph} \\
+      $(g,d)$-map & $\textcolor{red}{?}$ & $\textcolor{red}{?}$ & \textcolor{red}{\cref{NewMapGraph}} \\ \hline
+      \\[.2ex]
+      \multicolumn{4}{l}{$^*$ the bound also applies to the simple treewidth of $H$.}
+    \end{tabular}
+  \end{center}
+  \caption{Summary of known product structure theorems.}
+  \label{summary_table}
+\end{table}
+
 \begin{compactitem}
@@
 
-\begin{thm}
+\begin{thm}\label{nn_product_structure}
   Every $k$-nearest neighbour graph is a subgraph of $H\boxtimes P\boxtimes K_{\ell}$ for some graph $H$ of treewidth $O(k^6)$ and some $\ell\in O(k^4)$.

### 2022-03-23 k-planar.tex
Pfffft

@@
 
-\citet{DJMMUW20} introduced the study of partitions with bounded layered width such that the quotient has some additional desirable property, like small treewidth.
+% \citet{DJMMUW20} introduced the study of partitions with bounded layered width such that the quotient has some additional desirable property, like small treewidth.
 Dujmovi\'c~et~al.\ define a class $\mathcal{G}$ of graphs to \defin{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
@@
 The following result is the main technical contribution of the paper.
-Loosely speaking, it shows that if a graph $G$ admits bounded layered partitions, then so too does $G^\SS$ for every shortcut system $\SS$ of $G$.
+Loosely speaking, it shows that if a graph $G$ has product structure, then so too does $G^\SS$ for every shortcut system $\SS$ of $G$.
 
@@
 If $G$ is planar with maximum degree $\Delta$, then $G^k$ is $2k(k+1)\Delta^{k}$-planar by \cref{DrawG}.
-Thus we can immediately conclude that bounded powers of planar graphs of bounded degree admit bounded layered partitions.
+Thus we can immediately conclude that bounded powers of planar graphs of bounded degree have product structure.
 However, the bounds we obtain are improved by the following lemma that constructs a shortcut system directly.
@@
 
+\subsection{Map Graphs}
+
+% \note{PM}{Check David's commit: a0e20aff83d12c74127331854ce9fb75e002efe3}
+
+Map graphs are defined as follows.
+Start with a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
+Let $G$ be the graph whose vertices are the nations of $G_0$, where two vertices are adjacent in $G$ if the corresponding faces in $G_0$ share a vertex.
+Then $G$ is called a \defin{$(g,d)$-map graph}.
+A $(0,d)$-map graph is called a (plane) \defin{$d$-map graph}; see \citep{FLS-SODA12,CGP02} for example.
+The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.
+
+% \note{DW}{give a better version of \cref{MapPartition} with $g=0$. }
+
+%So $(g,d)$-map graphs generalise graphs embedded in a surface, and we now assume that $d\geq 4$ for the remainder of this section.
+
+There is a natural drawing of a map graph obtained by positioning each vertex of $G$ inside the corresponding nation and each edge of $G$ as a curve passing through the corresponding vertex of $G_0$.
+It is easily seen that each edge is in at most $\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}}$ crossings; see \citep{dujmovic.eppstein.ea:structure}.
+Thus $G$ is $(g,\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
+Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar.
+\cref{gkPlanarProduct} then establishes a product structure theorem for map graphs, but we get much better bounds by constructing a shortcut system directly.
+The following lemma is reminiscent of the characterisation of $(g,d)$-map graphs in terms of the half-square of bipartite graphs \citep{CGP02,dujmovic.eppstein.ea:structure}.
+
+\begin{lem}
+\label{MapShortcut}
+Every $(g,d)$-map graph $G$ is contained in $G_1^\SS$ for some graph $G_1$ with Euler genus at most $g$ and some $(2,\tfrac12 d(d-3) )$-shortcut system $\SS$ for $G_1$.
+\end{lem}
+
+\begin{proof}
+Let $G$ be a $(g,d)$-map graph.
+So there is a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
+Let $N$ be the set of nations.
+Then $V(G)=N$ where two vertices are adjacent in $G$ if the corresponding nation faces of $G_0$ share a vertex.
+Let $G_1$ be the graph with $V(G_1):=V(G_0) \cup N$, where distinct vertices $v,w\in N$ are adjacent in $G_1$ if the boundaries of the corresponding nations have an edge of $G_0$ in common, and $v\in V(G_0)$ and $w\in N$ are adjacent in $G_1$ if $v$ is on the boundary of the nation corresponding to $w$.
+Observe that $G_1$ embeds in the same surface as $G_0$ with no crossings, and that each vertex in $V(G_0)$ has degree at most $d$ in $G_1$.
+Consider an edge $vw\in E(G)$.
+If the nations corresponding to $v$ and $w$ share an edge of $G_0$, then $vw$ is an edge of $G_1$.
+Otherwise,  $v$ and $w$ have a common neighbour $x$ in $V(G_0)$.
+In the latter case, let $P_{vw}$ be the path $(v,x,w)$.
+Let $\SS$ be the set of all such paths $P_{vw}$.
+Each vertex $x\in V(G_0)$ is the middle vertex on at most $\tfrac12 d(d-3)$  paths in $\SS$.
+Thus $\SS$ is a $(2,\tfrac12 d(d-3))$-shortcut system for $G_1$, and by construction, $G \subseteq G_1^\SS$.
+\end{proof}
+
+\cref{ShortcutProduct,MapShortcut,GenusProduct} imply:
+
+\begin{thm}
+\label{gdMapProduct}
+For integers $g\geq 0$ and $d\geq 3$, let $\ell:= 14\max\{2g,3\}\,\binom{d}{2}$.
+Then every $(g,d)$-string graph is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $9$,
+\end{thm}
+
 \subsection{$k$-Nearest-Neighbour Graphs}
@@
 
-\section{Framed Graphs, $1$-Planar Graphs, and Map Graphs}
+\section{Framed Graphs, $1$-Planar Graphs, and (Plane) Map Graphs}
 \label{FramedSection}
@@
 
-\subsection{Map Graphs}
-
-\note{PM}{Check David's commit: a0e20aff83d12c74127331854ce9fb75e002efe3}
-
-Map graphs are defined as follows.
+\subsection{Plane Map Graphs}
+
+% \note{PM}{Check David's commit: a0e20aff83d12c74127331854ce9fb75e002efe3}
+
+% Map graphs are defined as follows.
 Start with a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
@@
 The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.
-
-\note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph.}
+% 
+% \note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph.}
+
+In addition to $1$-planar graphs, framed graphs can be used to obtain the following improved product structure theorem for (plane) $d$-map graphs.
 
 \begin{thm}
-\label{NewMapGraph}
+\label{dMapProduct}
 Every $d$-map graph is contained in $H \boxtimes P \boxtimes K_{ d + 3\floor{d/2} -3 }$ for some planar graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
@@
 
-\cref{NewMapGraph} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma.
+\cref{dMapProduct} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma.
 Similar results to \cref{MewMapGraphLemma} appear in earlier works \cite{CGP06,BDGGMR,Brandenburg19,Brandenburg20} but we state the precise lemma and provide a proof here for the sake of completeness.
@@
 \label{MewMapGraphLemma}
-For all integers $g\geq 0$ and $d\geq 3$, every $(g,d)$-map graph is a spanning subgraph of a $(g,d)$-framed graph.
+For all integers $g\geq 0$ and $d\geq 3$, every $d$-map graph is a spanning subgraph of a $d$-framed graph.\todo{All I did here was remove references to $g$}
 \end{lem}
@@
 \begin{proof}
-Let $G_0$ be a graph embedded in a surface of Euler genus $g$, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations.
+\note{PM}{I fixed some issues here that come from the fact that some vertices of $G_0$ do not contribute faces to $G_1$.
+This happens when two nations and two lakes meet at a vertex $x$.
+It's not an issue but it means we can't write ``the face of $G_1$ that corresponds to $x$.''}
+Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations.
 If two lakes have a common edge in their boundaries, then we may delete the edges and merge the lakes.
@@
 Let $G$ be the corresponding map graph.
-Let $G_1$ be the graph with $V(G_1):=V(G)$, which is the set of nations.
-Consider vertices $v$ and $w$ of $G_1$.
+Let $G_1$ be the graph with $V(G_1):=V(G)$, which is the set of nations, and whose edge set we now define.  Consider vertices $v$ and $w$ of $G_1$.
 Let $A$ and $B$ be the nation faces of $G_0$ corresponding to $v$ and $w$.
@@
 or there is a vertex $x$ of $G_0$ in the boundary of $A$ and $B$, and in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes.
-Then $G_1$ has Euler genus at most $g$.
-The faces of $G_1$ correspond to vertices of $G_0$ and lakes of $G_0$.
-Triangulate each face of $G_1$ corresponding to a lake of $G_0$. For each vertex
-$x$ of $G_0$ the size of the face of $G_1$ corresponding to $x$ equals the number of nations incident to $x$, which is at most $d$.
-So every face of $G_1$ has size at most $d$.
-If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common, implying that  $v$ and $w$ are both on the face of $G_1$ corresponding to $x$, and $vw\in E(\widehat{G_1})$.
-Hence $G\subseteq \widehat{G_1}$.
+% Then $G_1$ has Euler genus at most $g$.
+Each face of $G_1$ corresponds to a vertex of $G_0$ or a lake of $G_0$.
+Triangulate each face of $G_1$ corresponding to a lake of $G_0$.
+
+For each face $F$ of $G_1$ that corresponds to a vertex $x$ of $G_0$, the size of $F$ equals the number of nations incident to $x$, which is at most $d$.
+So every face of $G_1$ has size at most $d$.  Let $\hat{G}_1$ be the $d$-framed graph whose frame is $G_1$.
+
+To complete the proof, we now show that $\hat{G}_1$ contains $G$.
+If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common.
+If some face of $G_1$ corresponds to $x$ then $v$ and $w$ are on a common face of $G_1$ so $vw\in E(\hat{G}_1)$.
+If no face of $G_1$ corresponds to $x$ then the only two nation faces of $G_0$ with $x$ on their boundary those corresponding to $v$ and $w$, in which case $vw\in E(G_1)\subseteq E(\hat{G}_1)$.
+Therefore $G\subseteq\hat{G}_1$.
 \end{proof}
 
-\note{DW}{give a better version of \cref{MapPartition} with $g=0$. }
+% \note{DW}{give a better version of \cref{MapPartition} with $g=0$. }
 
 %So $(g,d)$-map graphs generalise graphs embedded in a surface, and we now assume that $d\geq 4$ for the remainder of this section.
-
-There is a natural drawing of a map graph obtained by positioning each vertex of $G$ inside the corresponding nation and each edge of $G$ as a curve passing through the corresponding vertex of $G_0$.
-It is easily seen that each edge is in at most $\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}}$ crossings; see \citep{dujmovic.eppstein.ea:structure}.
-Thus $G$ is $(g,\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.
-Also note that \cref{DrawG} with $k=2$ implies that $G$ is $(g, O(d^{2}) )$-planar.
-\cref{gkPlanarProduct} then establishes a product structure theorem for map graphs, but we get much better bounds by constructing a shortcut system directly.
-The following lemma is reminiscent of the characterisation of $(g,d)$-map graphs in terms of the half-square of bipartite graphs \citep{CGP02,dujmovic.eppstein.ea:structure}.
-
-\begin{lem}
-\label{MapShortcut}
-Every $(g,d)$-map graph $G$ is contained in $G_1^\SS$ for some graph $G_1$ with Euler genus at most $g$ and some $(2,\tfrac12 d(d-3) )$-shortcut system $\SS$ for $G_1$.
-\end{lem}
-
-\begin{proof}
-Let $G$ be a $(g,d)$-map graph.
-So there is a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
-Let $N$ be the set of nations.
-Then $V(G)=N$ where two vertices are adjacent in $G$ if the corresponding nation faces of $G_0$ share a vertex.
-Let $G_1$ be the graph with $V(G_1):=V(G_0) \cup N$, where distinct vertices $v,w\in N$ are adjacent in $G_1$ if the boundaries of the corresponding nations have an edge of $G_0$ in common, and $v\in V(G_0)$ and $w\in N$ are adjacent in $G_1$ if $v$ is on the boundary of the nation corresponding to $w$.
-Observe that $G_1$ embeds in the same surface as $G_0$ with no crossings, and that each vertex in $V(G_0)$ has degree at most $d$ in $G_1$.
-Consider an edge $vw\in E(G)$.
-If the nations corresponding to $v$ and $w$ share an edge of $G_0$, then $vw$ is an edge of $G_1$.
-Otherwise,  $v$ and $w$ have a common neighbour $x$ in $V(G_0)$.
-In the latter case, let $P_{vw}$ be the path $(v,x,w)$.
-Let $\SS$ be the set of all such paths $P_{vw}$.
-Each vertex $x\in V(G_0)$ is the middle vertex on at most $\tfrac12 d(d-3)$  paths in $\SS$.
-Thus $\SS$ is a $(2,\tfrac12 d(d-3))$-shortcut system for $G_1$, and by construction, $G \subseteq G_1^\SS$.
-\end{proof}
 
@@
     \item $k$-planar: $2^{O(k^3)}$
-    \item $(g,k)$-planar: $g2^{O(k^3)}$
-    \item $(g,\delta)$-string: $g2^{O(\delta^3)}$
+    \item $(g,k)$-planar: $\max\{1,g\}\cdot 2^{O(k^3)}$
+    \item $(g,\delta)$-string: $\max\{1,g\}\cdot 2^{O(\delta^3)}$
     \item $1$-planar: $115$
     \item $d$-map: $24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$
-    \item $(g,d)$-map: $g2^{O(d)}$
+    \item $(g,d)$-map: $\max\{1,g\}\cdot 2^{O(d)}$
     \item $k$-NN: $2^{O(k^6)}$
@@
 
-The following corollary summarizes the results for other graph classes.
-\begin{cor}\label{q_cor}
+Combining \cref{non-repetitive} with our results on product structure, we obtain the following corollary:
+\begin{cor}\label{nonrepetitive_cor}
   The following bounds hold for the nonrepetitive chromatic number of any graph from each of the following classes:
   \begin{compactenum}
-    \item $k$-planar: $(18k^2+48k+30) 4^{\binom{k+4}{3}}$
+    \item $k$-planar: $(18k^2+48k+30)\cdot 4^{\binom{k+4}{3}}$
     \item $(g,k)$-planar: $\max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}$
-    \item $(g,\delta)$-string: $g4^{O(\delta^3)}$
+    % \item $(g,\delta)$-string: $g4^{O(\delta^3)}$
     \item $1$-planar: $1792$
-    \item $d$-map: $24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$
-    \item $(g,d)$-map: $g4^{O(d)}$
-    \item $k$-NN: $4^{O(k^6)}$
-    % \item $(g,d)$-map graph
+    \item $d$-map: $256(d+3\floor{d/2}-3)$
+    % \item $(g,d)$-map: $g4^{O(d)}$
+    % \item $k$-NN: $4^{O(k^6)}$
+    \item $(g,d)$-map graph: \textcolor{red}{??}
     % \item
@@
 Prior to the current work, the strongest upper bound on the non-repetitive chromatic number of $n$-vertex  $k$-planar graphs was $O(k\log n)$ \cite{dujmovic.morin.ea:layered}.
+Our results also give bounds on the nonrepetitive chromatic number of $(g,\delta)$-string graphs and $k$-nearest-neighbour graphs, but these bounds are weaker than existing results based on maximum degree.
+The bounds obtained from product structure are exponential in $\delta$ and $k^2$, respectively.
+However, each of these graph classes has degree bounded by $O(\delta)$ and $O(k^2)$, respectively, and therefore have nonrepetitive chromatic number $O(\delta^2)$ and $O(k^{4})$, respectively \cite{DJKW16}.
 
@@
 
-\cref{p-centered,kPlanarProduct,1-planar,gkPlanarProduct} immediately imply the following results, for every $p\ge 2$:
-
-\begin{cor}\quad
-\begin{compactitem}
-\item For every $1$-planar graph $G$,\; $ \chi_p(G)\le 5 (p+3)(p+2)(p+1)^2$.
-%\item For every $k$-planar graph $G$,\; $\displaystyle \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}$.
-\item For every $(g,k)$-planar graph $G$,\;
-$\displaystyle \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{p+\binom{k+4}{3}-1}{\binom{k+4}{3}-1}$.
-% \note{DW}{I have changed $\binom{k+5}{4}$ to $\binom{k+4}{3}$, okay?}
-% \note{PM}{Ok}
- \end{compactitem}
+% \cref{p-centered,kPlanarProduct,1-planar,gkPlanarProduct} immediately imply the following results, for every $p\ge 2$:
+% 
+% \begin{cor}\quad
+% \begin{compactitem}
+% \item For every $1$-planar graph $G$,\; $ \chi_p(G)\le 5 (p+3)(p+2)(p+1)^2$.
+% %\item For every $k$-planar graph $G$,\; $\displaystyle \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}$.
+% \item For every $(g,k)$-planar graph $G$,\;
+% $\displaystyle \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{p+\binom{k+4}{3}-1}{\binom{k+4}{3}-1}$.
+% % \note{DW}{I have changed $\binom{k+5}{4}$ to $\binom{k+4}{3}$, okay?}
+% % \note{PM}{Ok}
+%  \end{compactitem}
+% \end{cor}
+
+For planar graphs of treewidth at most $3$ an $O(p^2\log p)$ upper bound is known:
+
+\begin{lem}\cite{DFMS21}\label{simple_3tree_centered_colouring}
+  For every planar graph $H$ of treewidth at most $3$, 
+  \[ 
+    \chi_p(H)\le (p+1)(p\ceil{\log(p+1)}+2p+1) \enspace .
+  \]
+\end{lem}
+
+Combining \cref{p-centered,simple_3tree_centered_colouring} with our results on product structure, we obtain the following corollary:
+
+\begin{cor}\label{p_centered_cor}
+  The following bounds hold for the $p$-centered chromatic number of any graph from each of the following classes:
+  \begin{compactenum}
+    \item $k$-planar: $(18k^2+48k+30)\cdot 4^{\binom{k+4}{3}}$
+    \item $(g,k)$-planar: $\max\{2g,3\}\cdot(6k^2+16k+10)(p+1)\binom{p+\binom{k+4}{3}-1}{\binom{k+4}{3}-1}$
+    % \item $(g,\delta)$-string: $g4^{O(\delta^3)}$
+    \item $1$-planar: $7(p+1)^2(p\ceil{\log(p+1)}+2p+1)$
+    \item $d$-map: $(d+3\floor{d/2}-3)\cdot (p+1)^2(p\ceil{\log(p+1)}+2p+1)$
+    % \item $(g,d)$-map: $g4^{O(d)}$
+    % \item $k$-NN: $4^{O(k^6)}$
+    \item $(g,d)$-map: $\max\{2g,3\}\cdot 7d(d-3)(p+1)\cdot \binom{p+9}{9}$
+    % \item
+  \end{compactenum}
 \end{cor}
@@
 
+As with nonrepetitive chromatic number, our results also imply bounds on the $p$-centered chromatic numer of $(g,\delta)$-string graphs and $k$-nearest-neighbour graphs, but these bounds are weaker than the bounds implied by the fact that these graphs have bounded maximum degree \cite{DFMS21}.
+
 \subsection{Universal Graphs}
@@
 
-\begin{thm}[\citep{DEJGMM21}]
+\begin{thm}[\citep{DEJGMM21,EJM}]
 	\label{Universal}
 	For every fixed integer $t$ and every integer $n>0$ there exists a
-	graph $U_n$ with $n^{1+o(1)}$ vertices such that for every graph $H$ of
-	treewidth at most $t$ and path $P$, every $n$-vertex subgraph of $H\boxtimes P$ is isomorphic to an induced subgraph of $U_n$.
+	graph $U_n$ with $n^{1+o(1)}$ vertices and edges such that for every graph $H$ of treewidth at most $t$ and path $P$, every $n$-vertex subgraph of $H\boxtimes P$ is isomorphic to an induced subgraph of $U_n$.
 \end{thm}
 
-Combining this with \cref{gkPlanarProduct,k-nn,PowerMinor,StringPartition,MapPartition} yields the following:
-
-\begin{thm}
+Combining \cref{Universal} with our results on product structure yields the following:
+% this with \cref{gkPlanarProduct,k-nn,PowerMinor,StringPartition,MapPartition} yields the following:
+
+\begin{cor}\label{universal_cor}
 	\label{UniversalUniversal}
-	For every fixed graph $X$ and all fixed integers $d,\delta,\Delta,g,k>0$, and every integer
-	$n>0$, there exists a graph $U_n$ with $n^{1+o(1)}$ vertices such that
-	$U_n$ contains all of the following graphs as induced subgraphs:
+	For every fixed graph $X$ and all fixed integers $d,\delta,\Delta,g,k>0$, and every integer $n>0$, there exists a graph $U_n$ with $n^{1+o(1)}$ vertices and edges such that $U_n$ contains all of the following graphs as induced subgraphs:
 	\begin{compactitem}
@@
 	\end{compactitem}
-\end{thm}
+\end{cor}
 
@@
 
-\section{Orphaned Section on Structure}
+\section{Orphaned Text}
+
+\subsection{Text following \cref{gkPlanarProduct}}
 
@@
 
-\section{Orphaned Section}
-
-\cref{nearest-neighbour,kPlanarProduct,qn,p-centered} imply:
-
-\begin{cor}
-\label{k-nn}
-For every integer $k\geq 1$ there exists integers $t\leq O(k^6)$ and $\ell\leq O(k^4)$ such that every $k$-nearest-neighbour graph:
-\begin{compactitem}
-\item is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $t$ and some path $P$,
-\item has queue-number at most $2^{O(k^6)}$, and
-\item has $p$-centered chromatic number at most $\ell (p+1)\binom{p+t}{t}$.
-\end{compactitem}
-\end{cor}
-
-\cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
+% \subsection{Orphaned Section}
+% 
+% \cref{nearest-neighbour,kPlanarProduct,qn,p-centered} imply:
+% 
+% \begin{cor}
+% \label{k-nn}
+% For every integer $k\geq 1$ there exists integers $t\leq O(k^6)$ and $\ell\leq O(k^4)$ such that every $k$-nearest-neighbour graph:
+% \begin{compactitem}
+% \item is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $t$ and some path $P$,
+% \item has queue-number at most $2^{O(k^6)}$, and
+% \item has $p$-centered chromatic number at most $\ell (p+1)\binom{p+t}{t}$.
+% \end{compactitem}
+% \end{cor}
+% 
+% \cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
 However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (36+o(1))k^2$ by a result of \citet{DJKW16}.
-
-\section{Orphaned Subsection}
 

### 2022-03-23 k-planar.tex

@@
       $1$-planar & $3^*$ & $7$ & \cref{1_planar_product} \\
-      $d$-map & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{NewMapGraph} \\
-      $(g,d)$-map & $\textcolor{red}{?}$ & $\textcolor{red}{?}$ & \textcolor{red}{\cref{NewMapGraph}} \\ \hline
+      $d$-map & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{dMapProduct} \\
+      $(g,d)$-map & $\textcolor{red}{?}$ & $\textcolor{red}{?}$ & \textcolor{red}{\cref{dMapProduct}} \\ \hline
       \\[.2ex]
@@
 Thus the graph $H$ in \cref{1_planar_product} has queue-number at most $5$.
-\cref{qn,1_planar_product,NewMapGraph} then imply:
+\cref{qn,1_planar_product,dMapProduct} then imply:
 

### 2022-03-23 k-planar.tex
Big reorganization done

@@
 \item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
+\item \citet{DFMS21} use it to prove that planar graphs have $p$-centered chromatic number $O(p^2\log p)$ and give a matching lower bound.
 \item \citet{DEJGMM21} use it to find asymptotically optimal adjacency labellings of planar graphs (resolving a problem of \citet{kannan.naor.ea:implicit}).
@@
 \end{compactitem}
-In addition, \cref{PlanarProduct} has been used to resolve or make substantial progress on a number of other problems on planar graphs, including $p$-centered colouring \cite{DFMS21}, $\ell$-vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, and odd colourings~\citep{DMO}.
+In addition, \cref{PlanarProduct} has been used to resolve or make substantial progress on a number of other problems on planar graphs, including $\ell$-vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, and odd colourings~\citep{DMO}.
 
@@
 A 1-apex graph is \defin{apex}.
-A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.} and \citet{DEMWW22} do so for graphs in any minor-closed class and with bounded maximum degree.
-See \cref{Generalisations} for more precise statements and see \citep{DHJLW21} for a survey on this topic.
+A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.} and \citet{DEMWW22} do so for bounded-degree graphs in any minor-closed class.
+See
+% \cref{Generalisations} for more precise statements and see 
+\citep{DHJLW21} for a survey on this topic.
 
@@
 \subsection{$k$-Planar Graphs}
-\label{k_planar_section}
+% \label{k_planar_section}
 
 We start with the example of $k$-planar graphs.
-A graph is \defin{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings, where no three edges cross at a single point (see \cref{sec-k-planar} for a formal definition).
+A graph is \defin{$k$-planar} if it has a drawing in the plane in which each edge is involved in at most $k$ crossings, where no three edges cross at a single point (see \cref{k_planar_section} for a formal definition).
 Such graphs provide a natural generalisation of planar graphs, and are important in graph drawing research; see the recent bibliography on 1-planar graphs and the 140 references therein \citep{kobourov.liotta.ea:annotated}.
@@
 
-For the important special case of $1$-planar graphs, we are able to obtain better constants and an additional property (planarity) on $H$.
+For the important special case of $1$-planar graphs, we are able to obtain better constants and an additional property (planarity) of $H$.
 
@@
 
-For at least one application, the planarity of $H$ is critical in obtaining asymptotically tight bounds \cite{BDJM}.
+For at least two applications, the planarity of $H$ is critical in obtaining asymptotically tight bounds \cite{BDJM,DFMS21}.
 \cref{1_planar_product} is proven in the more general setting of $d$-framed multigraphs.
@@
 
-Although $k$-planar graphs are the most high-profile target for a generalization of \cref{PlanarProduct}, we actually prove a substantially stronger result than \cref{kPlanarProduct} using the following definition.
+Although $k$-planar graphs are the most high-profile target for a generalization of \cref{PlanarProduct}, we actually prove a more general result than \cref{kPlanarProduct} using the following definition.
 A non-empty set $\SS$ of non-trivial paths in a graph $G$ is a \defin{$(k,d)$-shortcut system} (for $G$) if:
@@
 The following theorem is one of the main contributions of the paper.
-It says that if a graph class $\mathcal{G}$ has a product structure theorem, then the class of graphs obtained by applying a shortcut system to graphs in $\mathcal{G}$ also has a product structure theorem.
+It says that if a graph class $\mathcal{G}$ has a product structure theorem like \cref{PlanarProduct}(b), then so too does the class of graphs obtained by applying shortcut systems to graphs in $\mathcal{G}$.
 
@@
 Theorem~\ref{PlanarProduct}(b), \cref{ShortcutProduct}, and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$.
-Some further observations presented in \cref{sec-k-planar} lead to the improved result.
+Some further observations presented in \cref{k_planar_section} lead to the improved result.
 \cref{ShortcutProduct} is applicable for many graph classes in addition to $k$-planar graphs.
@@
 All of the above-mentioned applications also hold for these examples.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Layered Partitions}
-\label{LayeredPartitions}
-
-\note{PM}{Move this section out of the introduction, into a ``preliminaries'' section?}
-
-While strong products enable concise statements of the theorems in \cref{Introduction}, to prove such results it is helpful to work with layerings and partitions, which we now introduce.
-
-A \defin{layering} of a graph $G$
-% is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$
-% \note{DW}{just write ``
-is an ordered partition $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that for every edge
- % ...''}
-$vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
-For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
-To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the corresponding vertex in $H$.
-
-For any partition $\PP$ of $V(G)$ and any layering $\mathcal{L}$ of $G$ we define the \defin{layered width} of $\PP$ with respect to $\mathcal{L}$ as $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\PP\}$.
-For any partition $\PP$ of $V(G)$, we define the \defin{layered width} of $\PP$ as the minimum, over all layerings $\mathcal{L}$ of $G$, of the layered width of $\PP$ with respect to $\mathcal{L}$.
-
-% \citet{DJMMUW20} introduced the study of partitions with bounded layered width such that the quotient has some additional desirable property, like small treewidth.
-Dujmovi\'c~et~al.\ define a class $\mathcal{G}$ of graphs to \defin{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
-
-These definitions relate to strong products as follows.
-
-\begin{lem}[\citep{DJMMUW20}]
-\label{PartitionProduct}
-For every graph $H$, a graph $G$ has an $H$-partition of layered width at most $\ell$ if and only if $G$ is contained in $H \boxtimes P \boxtimes K_\ell$ for some path $P$.
-\end{lem}
-
-As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a),
-\citet{DJMMUW20} showed that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \cite{UWY}).
-The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.\footnote{\citet{bose.morin.ea:optimal} have recently given linear time algorithms for computing layered $H$-partitions of planar graphs.} At the core of their work is the elegant proof by \citet{PS21} of the following result:
-
-\begin{thm}[\citep{PS21}]
-\label{ps}
-  Every planar triangulation $G$ has an $H$-partition $\PP$ such that $\tw(H)\leq 8$ and $G[P]$ is a shortest path in $G$ for each $P\in\PP$.
-\end{thm}
-
-Indeed, the above-mentioned result of \citet{DJMMUW20} is a slight strengthening of \cref{ps}, where for each $P\in\PP$ no two vertices of $P$ have the same distance to some fixed root vertex $r$.
-
-The following result is the main technical contribution of the paper.
-Loosely speaking, it shows that if a graph $G$ has product structure, then so too does $G^\SS$ for every shortcut system $\SS$ of $G$.
-
-\begin{restatable}{thm}{mmg}
-	\label{ShortcutPartition}
-	Let $G$ be a graph having an $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\SS$ be a $(k,d)$-shortcut system for $G$.
-Then $G^\SS$ has a $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$.
-\end{restatable}
-
-Note that \cref{ShortcutPartition} is equivalent to \cref{ShortcutProduct} by \cref{PartitionProduct}.
 
@@
       $(g,k)$-planar & $\binom{k+4}{3}-1$ & $\max\{2g,3\}(6k^2+16k+10)$ & \cref{gkPlanarProduct} \\
-      $(g,\delta)$-string & $\binom{\delta+4}{3}-1$ & $\max\{2g,3\}(\delta^4 + 4\delta^3 + 9\delta^2 + 10\delta+4)$ & \cref{StringPartition} \\
+      $(g,\delta)$-string & $\binom{\delta+4}{3}-1$ & $\max\{2g,3\}(\delta^4 + 4\delta^3 + 9\delta^2 + 10\delta+4)$ & \cref{StringProduct} \\
       $k$-NN & $O(k^6)$ & $O(k^4)$ & \cref{nn_product_structure} \\
@@
       $d$-map & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{dMapProduct} \\
-      $(g,d)$-map & $\textcolor{red}{?}$ & $\textcolor{red}{?}$ & \textcolor{red}{\cref{dMapProduct}} \\ \hline
+      $(g,d)$-map & $9$ & $\max\{2g,3\}\,(7d^2 -21d)$ & \cref{gdMapProduct} \\ \hline
       \\[.2ex]
@@
   \end{center}
-  \caption{Summary of known product structure theorems.}
+  \caption{Known product structure theorems (of the form $G\subseteq H\boxtimes P\boxtimes K_\ell$).}
   \label{summary_table}
@@
   \item In \cref{Structure} we prove our main result for shortcut systems (\cref{ShortcutProduct}), and its optimisations for $k$-planar graphs and their genus-$g$ generalisation (\cref{kPlanarProduct,gkPlanarProduct}).
-  \item \Cref{examples} uses \cref{Structure} to derive product structure theorems for string graphs, powers of bounded-degree graphs, and $k$-nearest neighbour graphs.
+  \item \Cref{examples} uses \cref{ShortcutProduct,kPlanarProduct,gkPlanarProduct} to derive product structure theorems for string graphs, powers of bounded-degree graphs, map graphs, and $k$-nearest neighbour graphs.
   \item \Cref{FramedSection} proves a product structure theorem for $d$-framed multigraphs in which the graph $H$ has treewidth $3$ and is planar.
-Using this result, we derive product structure theorems for $1$-planar graphs and map graphs, again with a planar graph $H$.
+Using this result, we derive improved product structure theorems for $1$-planar graphs and planar map graphs, again with a planar graph $H$.
   \item \Cref{Applications} quickly surveys applications of product structure and the consequences of the current work.
@@
 
-The purpose of this section is to prove \cref{ShortcutPartition} and its specialization to $k$-planar graphs \cref{kPlanarProduct}.
+The purpose of this section is to prove \cref{ShortcutProduct} and its specialization to $k$-planar graphs, \cref{kPlanarProduct}.
+While strong products enable concise statements of the theorems in \cref{Introduction}, to prove such results it is helpful to work with layerings and partitions, which we now introduce.
+
+\subsection{Layered Partitions}
+
+A \defin{layering} of a graph $G$
+% is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$
+% \note{DW}{just write ``
+is an ordered partition $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that for every edge
+ % ...''}
+$vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
+For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
+To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the corresponding vertex in $H$.
+
+For any partition $\PP$ of $V(G)$ and any layering $\mathcal{L}$ of $G$ we define the \defin{layered width} of $\PP$ with respect to $\mathcal{L}$ as $\max\{|L\cap P|: L\in\mathcal{L},\, P\in\PP\}$.
+For any partition $\PP$ of $V(G)$, we define the \defin{layered width} of $\PP$ as the minimum, over all layerings $\mathcal{L}$ of $G$, of the layered width of $\PP$ with respect to $\mathcal{L}$.
+
+% \citet{DJMMUW20} introduced the study of partitions with bounded layered width such that the quotient has some additional desirable property, like small treewidth.
+Dujmovi\'c~et~al.\ define a class $\mathcal{G}$ of graphs to \defin{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
+
+These definitions relate to strong products as follows.
+
+\begin{lem}[\citep{DJMMUW20}]
+\label{PartitionProduct}
+For every graph $H$, a graph $G$ has an $H$-partition of layered width at most $\ell$ if and only if $G$ is contained in $H \boxtimes P \boxtimes K_\ell$ for some path $P$.
+\end{lem}
+
+As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a),
+\citet{DJMMUW20} build on an earlier result of \citet{PS21} to show that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \cite{UWY}).
+The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.\footnote{\citet{bose.morin.ea:optimal} have recently given linear time algorithms for computing layered $H$-partitions of planar graphs.}
+
+% \note{PM}{This feels out of place now that this section has been moved out of the introduction.  Delete?
+% 
+% At the core of their work is the elegant proof by \citet{PS21} of the following result:
+% 
+% \begin{thm}[\citep{PS21}]
+% \label{ps}
+%   Every planar triangulation $G$ has an $H$-partition $\PP$ such that $\tw(H)\leq 8$ and $G[P]$ is a shortest path in $G$ for each $P\in\PP$.
+% \end{thm}
+% 
+% Indeed, the above-mentioned result of \citet{DJMMUW20} is a slight strengthening of \cref{ps}, where for each $P\in\PP$ no two vertices of $P$ have the same distance to some fixed root vertex $r$.
+% }
+
+By \cref{PartitionProduct}, \cref{ShortcutProduct} is equivalent to the following result, whose proof is the subject of the next section.
+
+\begin{thm}
+	\label{ShortcutPartition}
+	Let $G$ be a graph having an $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$ and let $\SS$ be a $(k,d)$-shortcut system for $G$.
+Then $G^\SS$ has a $J$-partition of layered width at most $d\ell(k^3+3k)$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$.
+\end{thm}
 
@@
 
-\cref{ShortcutPartition} shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, an $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.
+We now prove \cref{ShortcutPartition}.
 For convenience, it will be helpful to assume that $\SS$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.
@@
 
-Consider a tree $T$ rooted at some node $x_0\in V(T)$.
-Each subtree $T'$ of $T$ is considered to be rooted at the node in $T'$ closest to $x_0$.
+Let $T$ be a tree rooted at some node $x_0\in V(T)$.
 A node $a\in V(T)$ is a \defin{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \defin{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.
@@
 The \defin{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.
-The \defin{lowest common ancestor} of a non-empty set $S\subseteq V(T)$ is the maximum $T$-depth node of $T$ that is a $T$-ancestor of every node in $S$.
+The \defin{lowest common $T$-ancestor} of a non-empty set $S\subseteq V(T)$ is the maximum $T$-depth node of $T$ that is a $T$-ancestor of every node in $S$.
+In the following, we treat any subtree $T'$ of $T$ as a rooted tree whose root is the lowest common $T$-ancestor of $V(T')$.
+
 % For each node $x\in V(T)$, define
@@
 \subsection{\boldmath $k$-Planar Graphs}
+\label{k_planar_section}
 
@@
 \subsection{\boldmath $(g,k)$-Planar Graphs}
+\label{gk_planar_section}
 
@@
   \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
-  \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
+  \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some planar graph $H$ of treewidth at most $3$ and for some path $P$.
   \end{compactenum}
@@
 
-In this section we give three examples of well-known graph families that can be expressed in terms of shortcut systems and which therefore have product structure theorems.
+In this section we give four examples of well-known graph families that can be expressed in terms of shortcut systems and which therefore have product structure theorems.
 
@@
 
-A similar proof to that of \cref{StringShortcut} shows that every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.
-Theorems~\ref{PlanarProduct}(b), \ref{ShortcutProduct}  and \ref{GenusProduct}(b):
+% A similar proof to that of \cref{StringShortcut} shows that every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.
+
+\cref{StringShortcut,ShortcutProduct,GenusProduct} now imply:
  % and \cref{StringShortcut,qn} imply:
 
-\begin{thm}\label{StringPartition}
+\begin{thm}\label{StringProduct}
 For integers $g\geq 0$ and $\delta\geq 2$, let $\ell:= \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$ and $t:= \binom{ \delta+4}{3}-1$.
@@
 \label{gdMapProduct}
-For integers $g\geq 0$ and $d\geq 3$, let $\ell:= 14\max\{2g,3\}\,\binom{d}{2}$.
-Then every $(g,d)$-string graph is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $9$,
+For integers $g\geq 0$ and $d\geq 3$, let $\ell:= \max\{2g,3\}(7d^2-21d)$.
+Then every $(g,d)$-map graph is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $9$,
 \end{thm}
@@
 
+\note{PM}{Got rid of genus-$g$ in the following. We use shortcuts for $(g,d)$-map graphs and now we don't have to define $(g,d)$-frame.}
+
 \begin{lem}
 \label{MewMapGraphLemma}
-For all integers $g\geq 0$ and $d\geq 3$, every $d$-map graph is a spanning subgraph of a $d$-framed graph.\todo{All I did here was remove references to $g$}
+For all integers $g\geq 0$ and $d\geq 3$, every $d$-map graph is a spanning subgraph of a $d$-framed graph.
 \end{lem}
@@
 For each face $F$ of $G_1$ that corresponds to a vertex $x$ of $G_0$, the size of $F$ equals the number of nations incident to $x$, which is at most $d$.
-So every face of $G_1$ has size at most $d$.  Let $\hat{G}_1$ be the $d$-framed graph whose frame is $G_1$.
-
-To complete the proof, we now show that $\hat{G}_1$ contains $G$.
+So every face of $G_1$ has size at most $d$.  Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.
+
+To complete the proof, we now show that $\widehat{G}_1$ contains every edge of $G$.
 If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common.
-If some face of $G_1$ corresponds to $x$ then $v$ and $w$ are on a common face of $G_1$ so $vw\in E(\hat{G}_1)$.
-If no face of $G_1$ corresponds to $x$ then the only two nation faces of $G_0$ with $x$ on their boundary those corresponding to $v$ and $w$, in which case $vw\in E(G_1)\subseteq E(\hat{G}_1)$.
-Therefore $G\subseteq\hat{G}_1$.
+If some face of $G_1$ corresponds to $x$ then $v$ and $w$ are on a common face of $G_1$ so $vw\in E(\widehat{G}_1)$.
+If no face of $G_1$ corresponds to $x$\footnote{This can occur when $x$ is on the boundary of exactly two nation faces and two lake faces.} then the only two nation faces of $G_0$ with $x$ on their boundary are those corresponding to $v$ and $w$, in which case $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.
+Therefore $G\subseteq\widehat{G}_1$.
 \end{proof}
@@
     % \item $k$-NN: $4^{O(k^6)}$
-    \item $(g,d)$-map graph: \textcolor{red}{??}
+    \item $(g,d)$-map graph: $(7d^2-21d)\cdot 4^{10}$
     % \item
@@
 
-\citet{PS21} and \citet{DFMS21} use \cref{ps} and \cref{PlanarProduct}(b), respectively, to show that $\chi_p(G)$ is polynomial in $p$ when $G$ is planar or of bounded Euler genus.
+% \citet{PS21} and 
+\citet{DFMS21} use \cref{PlanarProduct} and \cref{PlanarProduct}(b), respectively, to show that $\chi_p(G)\in O(p^2\log p)$ when $G$ is planar or of bounded Euler genus.
 Upper bounds on $\chi_p$ for graphs of given treewidth~\citep{PS21} and for graph products~\citep{DFMS21} imply the next lemma.

### 2022-03-23 k-planar.tex
Section to subsection
better  qn d-map bound

@@
 \citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
-Thus the graph $H$ in \cref{1_planar_product} has queue-number at most $5$.
-\cref{qn,1_planar_product,dMapProduct} then imply:
+Thus the graph $H$ in \cref{1_planar_product,dMapProduct} has queue-number at most $5$, which with \cref{qn} implies:
 
@@
 \label{1PlanarQueue}
-Every 1-planar graph has queue-number at most $3 \times 7 \times 5 + \floor{\tfrac{3}{2} \times 7} = 115$.
+Every 1-planar graph has queue-number at most 
+$3 \times 7 \times 5 + \floor{\tfrac{3}{2} \times 7} = 115$.
 \end{cor}
@@
 \label{dMapQueue}
-Every $d$-map graph has queue-number at most $24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$.
+Every $d$-map graph has queue-number at most 
+$15 (d+3\floor{d/2}-3)  + \floor{\tfrac{3}{2}(d+3\floor{d/2}-3)}$.
+\note{DW}{I have updated this bound using the better qn bound for planar 3-trees. Note that with $d=3$ we get the original bound of 49.}
 \end{cor}
@@
     \item $(g,\delta)$-string: $\max\{1,g\}\cdot 2^{O(\delta^3)}$
-    \item $1$-planar: $115$
-    \item $d$-map: $24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$
+    \item $1$-planar: $115$ \note{DW}{delete?}
+    \item $d$-map: $24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$ \note{DW}{delete?}
     \item $(g,d)$-map: $\max\{1,g\}\cdot 2^{O(d)}$
@@
 
-\section{Orphaned Theorems}
+\subsection{Orphaned Theorems}
 

### 2022-03-24 k-planar.tex
minor update
minor edits and comments

@@
 \item \citet{DEJGMM21} use it to find asymptotically optimal adjacency labellings of planar graphs (resolving a problem of \citet{kannan.naor.ea:implicit}).
-\item \citet{EJM} use it to show the existence of a 'universal graph' with $n^{1+o(1)}$ vertices and edges that contains every $n$-vertex planar graph as an induced subgraph (resolving a problem of \citet{babai.chung.ea:on}).
+\item \citet{EJM} use it to show the existence of a `universal graph' with $n^{1+o(1)}$ vertices and edges that contains every $n$-vertex planar graph as an induced subgraph (resolving a problem of \citet{babai.chung.ea:on}).
 \end{compactitem}
@@
 
-For at least two applications, the planarity of $H$ is critical in obtaining asymptotically tight bounds \cite{BDJM,DFMS21}.
+For at least two applications, the planarity of $H$ in \cref{1_planar_product} is critical in obtaining asymptotically tight bounds \cite{BDJM,DFMS21}.
 \cref{1_planar_product} is proven in the more general setting of $d$-framed multigraphs.
-This more general result implies a product structure theorem for $d$-map graphs in which $H$ is also planar.
+This more general result also implies a product structure theorem for $d$-map graphs in which $H$ is a planar.
 
@@
 \cref{summary_table} summarizes existing results on product structure theorems for minor-closed graph classes and new results for non-minor-closed graph classes.
-The remainder of this paper is organized as follows:
-
-\begin{table}
+
+\begin{table}[h]
   \begin{center}
     \begin{tabular}{lccl}
-      \textbf{Graph class} & \boldmath $\tw(H)$ & \boldmath $\ell$ & \textbf{Ref.} \\ \hline\hline
-      Planar & $3^*$ & $3$ & \cite{DJMMUW20} \\
-      Planar & $6^*$ & $1$ & \cite{UWY} \\
-      Genus-$g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW} \\
-      Genus-$g$ & $2g+6$ & 1 & \cite{UWY} \\
-      Apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\[1ex]
+      \textbf{Graph class} & \boldmath $\tw(H)$ & \boldmath $\ell$ & \textbf{Reference} \\ \hline
+      planar & $3^*$ & $3$ & \cite{DJMMUW20} \\
+      planar & $6^*$ & $1$ & \cite{UWY} \\
+      genus $g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW} \\
+%      genus $g$ & $2g+6$ & 1 & \cite{UWY} \\
+      apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\[1ex]
       % B-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
@@
 
+The remainder of this paper is organized as follows:
 \begin{compactitem}
@@
 The definition of $k$-planar graphs naturally generalises to genus-$g$ surfaces.
-A graph $G$ embedded on a surface $\Sigma$ is $(\Sigma,k)$-plane if every edge of $G$ is involved in at most $k$ crossings.
-A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
+A graph $G$ embedded on a surface $\Sigma$ is \defin{$(\Sigma,k)$-plane} if every edge of $G$ is involved in at most $k$ crossings.
+A graph $G$ is \defin{$(g,k)$-planar} if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
 \cref{AddDummy} immediately generalises as follows:
@@
 
-\cref{d_framed_product_stucture} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
+\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite{DJMMUW20}.
 The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{v_1,\ldots,v_t\}$.
@@
 Every $d$-map graph has queue-number at most 
-$15 (d+3\floor{d/2}-3)  + \floor{\tfrac{3}{2}(d+3\floor{d/2}-3)}$.
+$\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$.
 \note{DW}{I have updated this bound using the better qn bound for planar 3-trees. Note that with $d=3$ we get the original bound of 49.}
@@
     \item $(g,d)$-map graph: $(7d^2-21d)\cdot 4^{10}$
-    % \item
   \end{compactenum}
+\note{DW}{I prefer bullets here}
 \end{cor}
@@
 For planar graphs of treewidth at most $3$ an $O(p^2\log p)$ upper bound is known:
+\note{DW}{I prefer ``a $O(p^2\log p)$'', pronounced ``a big-O...''}
 
@@
     \item $(g,d)$-map: $\max\{2g,3\}\cdot 7d(d-3)(p+1)\cdot \binom{p+9}{9}$
-    % \item
   \end{compactenum}
+\note{DW}{I prefer bullets here}
 \end{cor}
@@
 
+\note{DW}{This problem does not deserve a section of its own. Move to end of \cref{gk_planar_section}?}
+
 We finish with an open problem.

### 2022-03-24 response.tex
updated the response

@@
 \newcommand{\done}{\textcolor{blue}{Done.}}
+\newcommand{\changed}{\textcolor{blue}{Revised accordingly.}}
 \newcommand{\tba}{\textcolor{red}{To be addresssed.}}
@@
 
+Both referees strongly recommend acceptance. We have revised the paper according to the referee reports as detailed below.
+
 \subsection*{Referee \#1}
 
-This paper proves that if a class of graphs F admits a product structure
-theorem, then the class of all graphs which have a 1-subdivision in F admits
-a product structure theorem as well. The main theorem, Theorem 3, while
-equivalent to the above statement using standard techniques, is stated in
-a different way which is considerably more useful for applications and for
-obtaining good bounds. As applications of Theorem 3, the paper shows that
-the following classes admit a product structure theorem:
-
-\begin{itemize}
-\item graphs drawn on a fixed surface with each edge in $\leq k$ crossings,
-\item map graphs of degree $\leq k$ graphs embedded in a fixed surface,
-\item string graphs with each curve in $\leq k$ intersections,
-\item $k$-nearest neighbour graphs in the plane, and
-\item powers of classes of graphs of degree $\leq k$ which admit a product structure theorem.
-\end{itemize}
-
-All but the last bullet point are shown to follow from the first.
-The main results in the paper are correct, strong, and of high interest,
-and the paper should definitely be accepted to JCTB after some
-(fairly significant) revisions. In particular, the proof of Theorem 3 is
-difficult to read due to the notation, and the paper could be substantially
-streamlined without losing the above results. Detailed suggestions follow.
-
-Major comments:
+%This paper proves that if a class of graphs F admits a product structure
+%theorem, then the class of all graphs which have a 1-subdivision in F admits
+%a product structure theorem as well. The main theorem, Theorem 3, while
+%equivalent to the above statement using standard techniques, is stated in
+%a different way which is considerably more useful for applications and for
+%obtaining good bounds. As applications of Theorem 3, the paper shows that
+%the following classes admit a product structure theorem:
+%
+%\begin{itemize}
+%\item graphs drawn on a fixed surface with each edge in $\leq k$ crossings,
+%\item map graphs of degree $\leq k$ graphs embedded in a fixed surface,
+%\item string graphs with each curve in $\leq k$ intersections,
+%\item $k$-nearest neighbour graphs in the plane, and
+%\item powers of classes of graphs of degree $\leq k$ which admit a product structure theorem.
+%\end{itemize}
+%
+%All but the last bullet point are shown to follow from the first.
+%The main results in the paper are correct, strong, and of high interest,
+%and the paper should definitely be accepted to JCTB after some
+%(fairly significant) revisions. In particular, the proof of Theorem 3 is
+%difficult to read due to the notation, and the paper could be substantially
+%streamlined without losing the above results. Detailed suggestions follow.
+%
+%Major comments:
 
@@
 
-\tba
+\textcolor{blue}{We have reorganised the paper, addressing this comment. }
 
@@
 5. page 6, line 16: ``Select any node $x \in V (T_0)$'' $\longrightarrow$ ``Select any nonempty node $x \in V (T_0)$''
+
+\tba
 
@@
 
-The Graph Product Structure Theorem of Dujmović, Joret,
-Micek, Morin, Ueckerdt, and Wood (FOCS '19) asserts that every planar
-graph is a subgraph of the strong product of a graph of treewidth at
-most 8 and a path.  This has been the key tool in resolving several
-longstanding open problems on planar graphs (and graphs in
-minor-closed classes), including problems on queue layouts,
-non-repetitive colouring, $p$-centered colouring, and universal graphs.
-This paper extends the Graph Product Structure Theorem beyond
-minor-closed classes.  The authors prove product structure theorems
-for k-planar graphs, map graphs, string graphs, graph powers, and
-nearest neighbour graphs.   The authors also present several
-applications of their results.  These are essentially the same
-applications as for the original Graph Product Structure Theorem; the
-point being that the applications typically follow quite easily once a
-product structure theorem is established.
-
-This paper is a significant step in answering the (ambitious) question
-of which graph classes admit product structure theorems. It is
-extremely well-written (with very few typos) and was a pleasure to
-read.  It is certainly well within the scope of JCTB and I strongly
-recommend it for acceptance.
-
-Below are some typos and suggestions for the authors.
+%The Graph Product Structure Theorem of Dujmović, Joret,
+%Micek, Morin, Ueckerdt, and Wood (FOCS '19) asserts that every planar
+%graph is a subgraph of the strong product of a graph of treewidth at
+%most 8 and a path.  This has been the key tool in resolving several
+%longstanding open problems on planar graphs (and graphs in
+%minor-closed classes), including problems on queue layouts,
+%non-repetitive colouring, $p$-centered colouring, and universal graphs.
+%This paper extends the Graph Product Structure Theorem beyond
+%minor-closed classes.  The authors prove product structure theorems
+%for k-planar graphs, map graphs, string graphs, graph powers, and
+%nearest neighbour graphs.   The authors also present several
+%applications of their results.  These are essentially the same
+%applications as for the original Graph Product Structure Theorem; the
+%point being that the applications typically follow quite easily once a
+%product structure theorem is established.
+%
+%This paper is a significant step in answering the (ambitious) question
+%of which graph classes admit product structure theorems. It is
+%extremely well-written (with very few typos) and was a pleasure to
+%read.  It is certainly well within the scope of JCTB and I strongly
+%recommend it for acceptance.
+%
+%Below are some typos and suggestions for the authors.
 
@@
 
-\done
+\changed
 
@@
 
-\textcolor{blue}{We have written the entire paper starting with simpler cases before moving onto the more general.
+\textcolor{blue}{We have written the paper starting with simpler cases before moving onto the more general.
 This approach may use a few more words, but we think the paper becomes more accessible for more readers this way.}
@@
 We have revised the proof avoiding the claim that every pair of edges cross at most once.
-The lemma itself has not changed.
-}
+The lemma itself has not changed.}
 
@@
 
-\done
+\changed
 
 Page 18.  Say $H'$ is connected since $X$ is connected.
+
+\tba
 

### 2022-03-24 k-planar.tex
minor edits and comments

@@
 
+\note{DW}{I suggest we use the language of ``embedded graph'' here}
 For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it has an embedding in $\R^2$ having a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
@@
 
-%\note{PM}{Consider a change of notation:   Replace $F$ with $C$.  (I was using $F$ before to be consistent with the previous section, but that notation is gone now.)} \note{DW}{I agree.}
-
 \note{DW}{I suggest $G'$ be changed to $G_0$ throughout (since $G'$ feels like a graph derived from $G$, which it is not). Similarly, replace $N'$ by $N_0$.}
@@
 
-% \note{DW}{We need to define ``closure''}
+\note{DW}{Some suggestions: Replace $G'$ by $F$ (the frame), replace $F_0$ by $C$ (the outer-cycle), replace $F_1,\dots,F_m$ by $C_1,\dots,C_m$.
+Use $F_1,\dots,F_m$ for the subgraph of $F$ inside $C_i$.
+Use $G_i$ for the subgraph of $G$ inside $C_i$.}
 
@@
 		\item For each $x\in V(H)$, $S_x$ has a partition $\{X_x,Y_x\}$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
-		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
+		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$ then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
 	\end{compactenum}
@@
 We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.
-This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.
-
-\begin{figure}
+This may introduce parallel edges into $G$, but no two such edges  appear on the boundary of a common face.
+
+\begin{figure}[h]
   \begin{center}
@@
 
-Consider some pair of edges $vw$ and $xy$ that cross at $p$.
-The point $p$ is a vertex of $G_0$ that has degree $4$.
+Consider some pair of edges $vw$ and $xy$ that cross at point $p$, which a degree-4 vertex of $G_0$.
 There are four faces $F_1,\ldots,F_4$ of $G_0$ that contain $p$ and, from the preceding discussion each of $vx$, $xw$, $wy$, and $yv$ is an uncrossed edge of $G$.
@@
 
-\note{PM}{Got rid of genus-$g$ in the following. We use shortcuts for $(g,d)$-map graphs and now we don't have to define $(g,d)$-frame.}
+%\note{PM}{Got rid of genus-$g$ in the following. We use shortcuts for $(g,d)$-map graphs and now we don't have to define $(g,d)$-frame.}
 
@@
 \label{MewMapGraphLemma}
-For all integers $g\geq 0$ and $d\geq 3$, every $d$-map graph is a spanning subgraph of a $d$-framed graph.
+For every %all 
+integer %s $g\geq 0$ and
+ $d\geq 3$, every $d$-map graph is a spanning subgraph of a $d$-framed graph.
 \end{lem}

### 2022-03-24 k-planar.tex
Numbers to bullets , got rid of final section

@@
 % \subsection{Examples}
+
+\cref{kPlanarProduct} shows that every $k$-planar graph is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $O(k^3)$ where $\ell\leq O(k^2)$.
+In some applications, the treewidth of $H$ is much more important than the value of $\ell$, which leads to the following question:
+
+\noindent\textbf{Open Problem:}
+  Does there exist a function $\ell:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph is contained in $H\boxtimes P \boxtimes K_{\ell(k)}$ for some graph $H$ with treewidth at most $C$?
+Perhaps $C=3$, which is the case for planar graphs (\cref{PlanarProduct}(b)) and for $1$-planar graphs (\cref{1_planar_product}).
+Note that $C\geq 3$ even for planar graphs \citep{DJMMUW20}.
 
@@
   The following bounds hold for the nonrepetitive chromatic number of any graph from each of the following classes:
-  \begin{compactenum}
+  \begin{compactitem}
     \item $k$-planar: $(18k^2+48k+30)\cdot 4^{\binom{k+4}{3}}$
@@
     \item $(g,d)$-map graph: $(7d^2-21d)\cdot 4^{10}$
-  \end{compactenum}
-\note{DW}{I prefer bullets here}
+  \end{compactitem}
 \end{cor}
@@
   The following bounds hold for the $p$-centered chromatic number of any graph from each of the following classes:
-  \begin{compactenum}
+  \begin{compactitem}
     \item $k$-planar: $(18k^2+48k+30)\cdot 4^{\binom{k+4}{3}}$
@@
     \item $(g,d)$-map: $\max\{2g,3\}\cdot 7d(d-3)(p+1)\cdot \binom{p+9}{9}$
-  \end{compactenum}
-\note{DW}{I prefer bullets here}
+  \end{compactitem}
 \end{cor}
@@
 
-\section{Wrapping Up}
-\label{WrappingUp}
-
-\note{DW}{This problem does not deserve a section of its own. Move to end of \cref{gk_planar_section}?}
-
-We finish with an open problem.
-\cref{kPlanarProduct} shows that every $k$-planar graph is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $O(k^3)$ where $\ell\leq O(k^2)$.
-Does there exist a function $\ell:\N\to\N$ and a universal constant $C$ such that every $k$-planar graph is contained in $H\boxtimes P \boxtimes K_{\ell(k)}$ for some graph $H$ with treewidth at most $C$?
-Perhaps $C=3$.
-Note that $C\geq 3$ even for planar graphs \citep{DJMMUW20}.
-
 \section{Orphaned Text}

### 2022-03-24 k-planar.tex
Got rid of reference to last section
Formatting
Pffft

@@
 \begin{table}[h]
-  \begin{center}
+  \centering{
     \begin{tabular}{lccl}
@@
       $(g,d)$-map & $9$ & $\max\{2g,3\}\,(7d^2 -21d)$ & \cref{gdMapProduct} \\ \hline 
-      \\[.2ex]
-      \multicolumn{4}{l}{$^*$ the bound also applies to the simple treewidth of $H$.}
+      \multicolumn{4}{l}{$^*$ these bounds also applies to the simple treewidth of $H$.}
     \end{tabular}
-  \end{center}
+  }
   \caption{Known product structure theorems (of the form $G\subseteq H\boxtimes P\boxtimes K_\ell$).}
@@
   \item \Cref{Applications} quickly surveys applications of product structure and the consequences of the current work.
-  \item \cref{WrappingUp} concludes with an open problem.
 \end{compactitem}
@@
 I settled on this ``base case'' since it's clear that, in this case, the partition $\mathcal{P}$ is already decided.}
-	$\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
+  \note{PM}{The longwinded version is that induction is on the pair $(x,y)$ where $x=|V(N-F)|$ and $y$ is the number of inner faces of $N'$.  The base cases are $(0,y)$ for any $y$.}
+	$\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iii):
 	\begin{compactenum}[(i)]
@@
 Since all the product structure theorems presented thus far upper bound the treewidth of $H$,
-\cref{qn} immediately implies bounds on the queue number of all graphs in these classes.  An improvement can be made in the cases where $H$ is planar and has treewidth $3$, as follows.
-\citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
-Thus the graph $H$ in \cref{1_planar_product,dMapProduct} has queue-number at most $5$, which with \cref{qn} implies:
-
-\begin{cor}
-\label{1PlanarQueue}
-Every 1-planar graph has queue-number at most 
-$3 \times 7 \times 5 + \floor{\tfrac{3}{2} \times 7} = 115$.
-\end{cor}
-
-\begin{cor}
-\label{dMapQueue}
-Every $d$-map graph has queue-number at most 
-$\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$.
-\note{DW}{I have updated this bound using the better qn bound for planar 3-trees. Note that with $d=3$ we get the original bound of 49.}
-\end{cor}
-
+\cref{qn} immediately implies bounds on the queue number of all graphs in these classes.
+An improvement can be made for $1$-planar and $d$-map graphs since, in the cases $H$ is planar and has treewidth $3$.
+\citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$. Thus the graph $H$ in \cref{1_planar_product,dMapProduct} has queue-number at most $5$.
+% , which with \cref{qn} implies:
+% 
+% \begin{cor}
+% \label{1PlanarQueue}
+% Every 1-planar graph has queue-number at most 
+% $3 \times 7 \times 5 + \floor{\tfrac{3}{2} \times 7} = 115$.
+% \end{cor}
+% 
+% \begin{cor}
+% \label{dMapQueue}
+% Every $d$-map graph has queue-number at most 
+% $\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$.
+% \note{DW}{I have updated this bound using the better qn bound for planar 3-trees. Note that with $d=3$ we get the original bound of 49.}
+% \end{cor}
+% 
+% 
+% 
 % \cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$.
@@
 % \note{PM}{Ok}
-
-The following corollary summarizes the results for other graph classes.
+The following corollary summarizes all of these results.
 \begin{cor}\label{q_cor}
@@
     \item $(g,\delta)$-string: $\max\{1,g\}\cdot 2^{O(\delta^3)}$
-    \item $1$-planar: $115$ \note{DW}{delete?}
-    \item $d$-map: $24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$ \note{DW}{delete?}
+    \item $1$-planar: $115$ 
+    \item $d$-map: $\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$ 
     \item $(g,d)$-map: $\max\{1,g\}\cdot 2^{O(d)}$
@@
 \end{cor}
+\note{PM}{For uniformity, I got rid of the two corollaries on $d$-map and $1$-planar graphs and kept them in \cref{q_cor}.
+The text is still commented out in the source in case someone feels strongly about keeping them.}
 
@@
 
-For planar graphs of treewidth at most $3$ an $O(p^2\log p)$ upper bound is known:
-\note{DW}{I prefer ``a $O(p^2\log p)$'', pronounced ``a big-O...''}
+For planar graphs of treewidth at most $3$ a $O(p^2\log p)$ upper bound is known:
+% \note{DW}{I prefer ``a $O(p^2\log p)$'', pronounced ``a big-O...''}
 
@@
 
-As with nonrepetitive chromatic number, our results also imply bounds on the $p$-centered chromatic numer of $(g,\delta)$-string graphs and $k$-nearest-neighbour graphs, but these bounds are weaker than the bounds implied by the fact that these graphs have bounded maximum degree \cite{DFMS21}.
+As with nonrepetitive chromatic number, our results also imply bounds on the $p$-centered chromatic numer of $(g,\delta)$-string graphs, powers of bounded-degree graphs, and $k$-nearest-neighbour graphs, but these bounds are weaker than the bounds implied by the fact that these graphs have bounded maximum degree \cite{DFMS21}.
 

### 2022-03-24 k-planar.tex
Small fixes/clarifications in the proof of 1_planar_is_4_framed

@@
 
-\note{DW}{I suggest we use the language of ``embedded graph'' here}
-For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it has an embedding in $\R^2$ having a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
-The graph $G_0$ is called the \defin{frame} of $G$.
+% \note{DW}{I suggest we use the language of ``embedded graph'' here}
+For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
+The embedded (multi)graph $G_0$ is called the \defin{frame} of $G$.
 This definition (for simple graphs) was introduced by \citet{BDGGMR}.
 
-Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar multigraph (by adding a vertex inside each face $F$ of $G_0$ adjacent to the vertices of $F$).
+Every $d$-framed multigraph can be described by a $(2,d(d-3)/2)$-shortcut system applied to a plane multigraph by adding a vertex inside each face $F$ of $G_0$ adjacent to the vertices of $F$ and creating a shortcut between each non-consecutive pair of vertices of $F$.
 Thus \cref{ShortcutProduct} is applicable.
@@
 
-	The proof is by induction on the number of inner faces of $N'$.  A particular easy case occurs when $V(N')=V(F)$, for which we can simply take
-	\note{DW}{Suggested revision, ``In the base case, $N'$ has only one inner face, and we can simply take ...} \note{PM}{I debated what to put here since, even when $N'$ has only one face, the rules below work for picking $\tau$.
+	The proof is by induction on the number of inner faces of $N'$.  A particular easy case occurs $N'$ has no inner vertices, i.e., $V(N')=V(F)$. In this case we can simply take
+	\note{DW}{Suggested revision, ``In the base case, $N'$ has only one inner face, and we can simply take ...} 
+  % \note{PM}{I debated what to put here since, even when $N'$ has only one face, the rules below work for picking $\tau$.
 The \emph{real} base case occurs when $N'$ has zero faces, but that's not something that ever happens since we only ever apply induction on the inner faces of $M$ (except $\tau$).
 I settled on this ``base case'' since it's clear that, in this case, the partition $\mathcal{P}$ is already decided.}
-  \note{PM}{The longwinded version is that induction is on the pair $(x,y)$ where $x=|V(N-F)|$ and $y$ is the number of inner faces of $N'$.  The base cases are $(0,y)$ for any $y$.}
+  \note{PM}{The longwinded version is that induction is on $(x,y)$ where $x$ is the number of inner vertices of $N'$ and $y$ is the number of inner faces of $N'$.
+The base cases are $(0,y)$ for any $y$.
+Unfortunately, we can't just induct on $x$ because it doesn't always decrease.}
 	$\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iii):
@@
 
-\note{DW}{Now assume that $N'$ has at least two internal faces.}
-
-In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that will be determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
+% \note{DW}{Now assume that $N'$ has at least two internal faces.}
+
+We now move onto the case in which $N'$ contains at least one inner vertex.
+Our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that will be determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
 We first explain how $v_1$, $v_2$ and $v_3$ are chosen and then explain how these are used to define $S$.
@@
 
+Refer to \cref{one_planar_example}.
 To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ that cross at $p$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
@@
 If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
-Therefore each dummy vertex $w_i$ of $W$ is surrounded by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
-Since $w_i$ has degree $4$ in $G_0$ and $F$ does not have two parallel edges on its boundary, $w_{i-1}\neq w_{i+1}$.
-Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.
+Therefore each dummy vertex $w_i$ of $W$ is flanked by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
+Since $w_i$ has degree $4>1$ in $G_0$ and no two edges incident to a common vertex cross each other, $w_{i-1}\neq w_{i+1}$.
+Since $G$ is edge-maximal the edge $w_{i-1}w_{i+1}$ is an edge of $G$ on the boundary of $F$.
 Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.
@@
 
-Consider some pair of edges $vw$ and $xy$ that cross at point $p$, which a degree-4 vertex of $G_0$.
-There are four faces $F_1,\ldots,F_4$ of $G_0$ that contain $p$ and, from the preceding discussion each of $vx$, $xw$, $wy$, and $yv$ is an uncrossed edge of $G$.
-Therefore, by removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).
+Consider some pair of edges $vw$ and $xy$ that cross at point $p$, which is a degree-4 vertex of $G_0$.
+Since no pair of edges incident to a common vertex cross each other, $v$, $w$, $x$, and $y$ are all distinct.
+There are four faces $F_1,\ldots,F_4$ of $G_0$ with $p$ on their boundary and, from the preceding discussion each of $vx$, $xw$, $wy$, and $yv$ is an uncrossed edge of $G$.
+Therefore, by removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges.
 Therefore $G$ is a $4$-framed graph with frame $G'$.
@@
 
-As mentioned in \cref{Introduction}, \citet{DEJGMM21} prove results about adjacency labelling schemes for planar graphs.
+As mentioned in \cref{Introduction}, \citet{DEJGMM21,EJM} prove results about adjacency labelling schemes for planar graphs.
 Stated in terms of universal graphs, their main theorem is interpreted as follows:
@@
 
-\section{Orphaned Text}
-
-\subsection{Text following \cref{gkPlanarProduct}}
-
-Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs (or any of the other classes presented in \cref{examples}) was in terms of layered treewidth, which we now define.
+% \section{Orphaned Text}
+% 
+% \subsection{Text following \cref{gkPlanarProduct}}
+% 
+% Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs (or any of the other classes presented in \cref{examples}) was in terms of layered treewidth, which we now define.
 A \defin{layered tree-decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{T}$ of $G$.
@@
 In related work, \citet{grigoriev.bodlaender:algorithms} used structural results to obtain approximation algorithms for $(g,k)$-planar graphs, and \citet{PachToth97} determined the maximum number of edges in a $k$-planar graph (up to a constant factor).
-
-If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
+% 
+% If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
 In particular, if $\PP=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree-decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree-decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{DJMMUW20}.
@@
 What sets layered partitions apart from layered treewidth is that they lead to constant upper bounds on the queue-number  and non-repetitive chromatic number, whereas for both these parameters, the best known upper bound obtainable via layered treewidth is $O(\log n)$; see \cref{Applications}.
-
-%%%%%%%%%%%%%%%%%%%
-\label{sec-k-planar}
-
-%\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{d_framed_product_stucture} with the method in \citep{DHHW}.
+% 
+% 
+% 
+% %%%%%%%%%%%%%%%%%%%
+% \label{sec-k-planar}
+% 
+% 
+% %\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{d_framed_product_stucture} with the method in \citep{DHHW}.
 I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
 Okay?}
-%\note{PM}{Okay to this and to all similar questions that come later (which I've now commented out.)}
-%\note{DW}{Robert is working on it.}
-
-This section describes several examples of graph classes that can be obtained from a shortcut system typically applied to graphs of bounded Euler genus.
-
-\subsection{Orphaned Theorems}
-
-Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,non-repetitive} imply the following result.
-
-%\begin{thm}
-%\label{PlaneMapPartition}
-%\note{DW}{For every integer $d\geq 4$, }
-%Every $d$-map graph $G$:
+% %\note{PM}{Okay to this and to all similar questions that come later (which I've now commented out.)}
+% %\note{DW}{Robert is working on it.}
+% 
+% 
+% 
+% 
+% 
+% This section describes several examples of graph classes that can be obtained from a shortcut system typically applied to graphs of bounded Euler genus.
+% 
+% \subsection{Orphaned Theorems}
+% 
+% Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,non-repetitive} imply the following result.
+% 
+% %\begin{thm}
+% %\label{PlaneMapPartition}
+% %\note{DW}{For every integer $d\geq 4$, }
+% %Every $d$-map graph $G$:
+% %	\begin{compactitem}
+% %		\item is contained in $H \boxtimes P \boxtimes K_{21d(d-3)}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
+% %		\item has queue-number $\qn(G) < 32225\, d(d-3)$.
+% %		\item has $p$-centered chromatic number $\chi_p(G) \leq 21d(d-3) (p+1)  \binom{p+9}{9}$,
+% %		\item has non-repetitive chromatic number $ \pi(G) \leq 21 \cdot 4^{10} d(d-3)$.
+% %	\end{compactitem}
+% %\end{thm}
+% 
+% \begin{thm}
+% \label{MapPartition}
+% For integers $g\geq 0$ and $d\geq 4$, if $\ell:=  7d(d-3)\, \max\{2g,3\}$ then every $(g,d)$-map graph $G$:
 % \begin{compactitem}
-%		\item is contained in $H \boxtimes P \boxtimes K_{21d(d-3)}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
-%		\item has queue-number $\qn(G) < 32225\, d(d-3)$.
-%		\item has $p$-centered chromatic number $\chi_p(G) \leq 21d(d-3) (p+1)  \binom{p+9}{9}$,
-%		\item has non-repetitive chromatic number $ \pi(G) \leq 21 \cdot 4^{10} d(d-3)$.
+% \item is contained in $H \boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
+% \item has queue-number $\qn(G) <  1535\, \ell $.
+% \item has $p$-centered chromatic number $\chi_p(G) \leq \ell (p+1)\,  \binom{p+9}{9}$,
+% \item has non-repetitive chromatic number $ \pi(G) \leq 4^{10}\,\ell $.
 % \end{compactitem}
-%\end{thm}
-
-\begin{thm}
-\label{MapPartition}
-For integers $g\geq 0$ and $d\geq 4$, if $\ell:=  7d(d-3)\, \max\{2g,3\}$ then every $(g,d)$-map graph $G$:
-\begin{compactitem}
-\item is contained in $H \boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
-\item has queue-number $\qn(G) <  1535\, \ell $.
-\item has $p$-centered chromatic number $\chi_p(G) \leq \ell (p+1)\,  \binom{p+9}{9}$,
-\item has non-repetitive chromatic number $ \pi(G) \leq 4^{10}\,\ell $.
-\end{compactitem}
-\end{thm}
-
-These results give the first constant upper bound on the non-repetitive chromatic number of map graphs, the first polynomial bounds on the $p$-centered chromatic number of map graphs, and the best known bounds on the queue-number of map graphs.
-
-% \note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the method in \citep{DHHW}.
+% \end{thm}
+% 
+% These results give the first constant upper bound on the non-repetitive chromatic number of map graphs, the first polynomial bounds on the $p$-centered chromatic number of map graphs, and the best known bounds on the queue-number of map graphs.
+% 
+% % \note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the method in \citep{DHHW}.
 I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
 Okay?}
-
-% \subsection{Orphaned Section}
-% 
-% \cref{nearest-neighbour,kPlanarProduct,qn,p-centered} imply:
-% 
-% \begin{cor}
-% \label{k-nn}
-% For every integer $k\geq 1$ there exists integers $t\leq O(k^6)$ and $\ell\leq O(k^4)$ such that every $k$-nearest-neighbour graph:
-% \begin{compactitem}
-% \item is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $t$ and some path $P$,
-% \item has queue-number at most $2^{O(k^6)}$, and
-% \item has $p$-centered chromatic number at most $\ell (p+1)\binom{p+t}{t}$.
-% \end{compactitem}
-% \end{cor}
-% 
-% \cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
+% 
+% 
+% % \subsection{Orphaned Section}
+% % 
+% % \cref{nearest-neighbour,kPlanarProduct,qn,p-centered} imply:
+% % 
+% % \begin{cor}
+% % \label{k-nn}
+% % For every integer $k\geq 1$ there exists integers $t\leq O(k^6)$ and $\ell\leq O(k^4)$ such that every $k$-nearest-neighbour graph:
+% % \begin{compactitem}
+% % \item is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $t$ and some path $P$,
+% % \item has queue-number at most $2^{O(k^6)}$, and
+% % \item has $p$-centered chromatic number at most $\ell (p+1)\binom{p+t}{t}$.
+% % \end{compactitem}
+% % \end{cor}
+% % 
+% % \cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
 However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (36+o(1))k^2$ by a result of \citet{DJKW16}.
-
-\subsection{Generalisations}
-\label{Generalisations}
-
-% As mentioned above, product structure theorems have been established for several minor-closed classes in addition to planar graphs.
+% 
+% 
+% 
+% 
+% \subsection{Generalisations}
+% \label{Generalisations}
+% 
+% % As mentioned above, product structure theorems have been established for several minor-closed classes in addition to planar graphs.
 The first generalises \cref{PlanarProduct} for graphs of bounded Euler genus.
-%
-% \begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
-% \label{GenusProduct}
-% Every graph of Euler genus $g$ is contained in:
-% \begin{compactenum}[(a)]
-% \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
-% \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
-% \end{compactenum}
+% %
+% % \begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
+% % \label{GenusProduct}
+% % Every graph of Euler genus $g$ is contained in:
+% % \begin{compactenum}[(a)]
+% % \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
+% % \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
+% % \end{compactenum}
+% % \end{thm}
+% 
+% \citet{DJMMUW20} generalised \cref{GenusProduct} for apex-minor-free graphs as follows.
+% 
+% \begin{thm}[\citep{DJMMUW20}]
+% \label{ApexMinorFree}
+% For every apex graph $X$, there exists $c\in\mathbb{N}$ such that every $X$-minor-free graph is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and some path $P$.
 % \end{thm}
-
-\citet{DJMMUW20} generalised \cref{GenusProduct} for apex-minor-free graphs as follows.
-
-\begin{thm}[\citep{DJMMUW20}]
-\label{ApexMinorFree}
-For every apex graph $X$, there exists $c\in\mathbb{N}$ such that every $X$-minor-free graph is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and some path $P$.
-\end{thm}
-
-The assumption that $X$ is apex is needed in \cref{ApexMinorFree}, since if the class of $X$-minor-free graphs has a product structure theorem analogous to \cref{PlanarProduct}, then $X$ is apex \citep{DJMMUW20}.
+% 
+% The assumption that $X$ is apex is needed in \cref{ApexMinorFree}, since if the class of $X$-minor-free graphs has a product structure theorem analogous to \cref{PlanarProduct}, then $X$ is apex \citep{DJMMUW20}.
 On the other hand, \citet{DEMWW22} proved a product structure theorem for bounded degree graphs in any minor-closed class.
-
-\begin{thm}[\citep{DEMWW22}]
-	\label{MinorFreeDegree}
-	For every graph $X$ there exists $c\in\mathbb{N}$ such that for every $\Delta\in\mathbb{N}$, every $X$-minor-free graph $G$ with maximum degree at most $\Delta$ is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H) \leq c\Delta$ and for some path $P$.
-\end{thm}
-
-%The next three results are immediate corollaries of \cref{GenusProduct,ApexMinorFree,ShortcutProduct,MinorFreeDegree}.
-%
-%\begin{thm}
-%Let $\SS$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$.
+% 
+% \begin{thm}[\citep{DEMWW22}]
+% 	\label{MinorFreeDegree}
+% 	For every graph $X$ there exists $c\in\mathbb{N}$ such that for every $\Delta\in\mathbb{N}$, every $X$-minor-free graph $G$ with maximum degree at most $\Delta$ is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H) \leq c\Delta$ and for some path $P$.
+% \end{thm}
+% 
+% %The next three results are immediate corollaries of \cref{GenusProduct,ApexMinorFree,ShortcutProduct,MinorFreeDegree}.
+% %
+% %\begin{thm}
+% %Let $\SS$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$.
 Then $G^\SS$ is contained in $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
-%\end{thm}
-%
-%\begin{thm}
-%For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\SS$ for $G$, $G^\SS\subseteq H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
-%\end{thm}
-%
-%\begin{thm}
-%For every graph $X$ and for all integers $k,\Delta\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ with maximum degree $\Delta$ and for every $(k,\Delta)$-shortcut system $\SS$ for $G$, $G^\SS \subseteq H \boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
-%\end{thm}
-%
-%\note{DW}{delete the above three theorems, they are never used}
-
-%\subsection*{Note Added in Proof} Subsequent to the initial release of this paper, the treewidth 8 bound in \cref{PlanarProduct} was improved to 6 by \citet{UWY}, which implies that the treewidth $2g+8$ bound in \cref{GenusProduct}(a) can be improved to $2g+6$.
+% %\end{thm}
+% %
+% %\begin{thm}
+% %For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\SS$ for $G$, $G^\SS\subseteq H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
+% %\end{thm}
+% %
+% %\begin{thm}
+% %For every graph $X$ and for all integers $k,\Delta\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ with maximum degree $\Delta$ and for every $(k,\Delta)$-shortcut system $\SS$ for $G$, $G^\SS \subseteq H \boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
+% %\end{thm}
+% %
+% %\note{DW}{delete the above three theorems, they are never used}
+% 
+% 
+% 
+% 
+% %\subsection*{Note Added in Proof} Subsequent to the initial release of this paper, the treewidth 8 bound in \cref{PlanarProduct} was improved to 6 by \citet{UWY}, which implies that the treewidth $2g+8$ bound in \cref{GenusProduct}(a) can be improved to $2g+6$.
 Simmilarly, the treewidth 4 bound in \cref{GenusProduct}(b) was improved to 3 by \citet{DHHW}.

### 2022-03-24 k-planar.tex
Deleted note

@@
 \begin{proof}
-\note{PM}{I fixed some issues here that come from the fact that some vertices of $G_0$ do not contribute faces to $G_1$.
+% \note{PM}{I fixed some issues here that come from the fact that some vertices of $G_0$ do not contribute faces to $G_1$.
 This happens when two nations and two lakes meet at a vertex $x$.

### 2022-03-24 k-planar.tex
G_0 and N_0

@@
 
-\note{DW}{I suggest $G'$ be changed to $G_0$ throughout (since $G'$ feels like a graph derived from $G$, which it is not). Similarly, replace $N'$ by $N_0$.}
-
-\note{PM}{I would also like to change $G'$ to $G_0$ and $N'$ to $N_0$.
+% \note{DW}{I suggest $G'$ be changed to $G_0$ throughout (since $G'$ feels like a graph derived from $G$, which it is not). Similarly, replace $N'$ by $N_0$.}
+% 
+% \note{PM}{I would also like to change $G'$ to $G_0$ and $N'$ to $N_0$.
 One issue is holding me back, though: We recurse on cycles $F_1,\ldots,F_m$ that define subgraphs $N_1,\ldots,N_m$ of $G$ and subgraphs $N_1',\ldots,N_m'$ of $G'$.
 Will these latter subgraphs becomes $N_{0,1},\ldots,N_{0,m}$?}
-
-\note{DW}{Some suggestions: Replace $G'$ by $F$ (the frame), replace $F_0$ by $C$ (the outer-cycle), replace $F_1,\dots,F_m$ by $C_1,\dots,C_m$.
+% 
+% \note{DW}{Some suggestions: Replace $G'$ by $F$ (the frame), replace $F_0$ by $C$ (the outer-cycle), replace $F_1,\dots,F_m$ by $C_1,\dots,C_m$.
 Use $F_1,\dots,F_m$ for the subgraph of $F$ inside $C_i$.
@@
 	\begin{compactenum}
-		\item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
-		\item Let $F_0$ be the outer face of $G'$ and let $T$ be a BFS spanning forest of $G'$ rooted at $V(F_0)$.
+		\item Let $G$ be a $d$-framed multigraph and let $G_0$ be a frame of $G$.
+		\item Let $F_0$ be the outer face of $G_0$ and let $T$ be a BFS spanning forest of $G_0$ rooted at $V(F_0)$.
 		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
@@
 		\end{compactenum}
-		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the interior or boundary of $F$.
+		\item Let $N$ and $N_0$ be the subgraphs of $G$ and $G_0$ consisting only of those edges and vertices contained in the interior or boundary of $F$.
 	\end{compactenum}
@@
 	This proof is similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is complicated by several factors.
-In particular, the reader should keep in mind that $G$ and $G'$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
-
-	The proof is by induction on the number of inner faces of $N'$.  A particular easy case occurs $N'$ has no inner vertices, i.e., $V(N')=V(F)$. In this case we can simply take
-	\note{DW}{Suggested revision, ``In the base case, $N'$ has only one inner face, and we can simply take ...} 
+In particular, the reader should keep in mind that $G$ and $G_0$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
+
+	The proof is by induction on the number of inner faces of $N_0$.  A particular easy case occurs when $N_0$ has no inner vertices, i.e., $V(N_0)=V(F)$. In this case we can simply take
+	\note{DW}{Suggested revision, ``In the base case, $N_0$ has only one inner face, and we can simply take ...} 
   % \note{PM}{I debated what to put here since, even when $N'$ has only one face, the rules below work for picking $\tau$.
@@
 I settled on this ``base case'' since it's clear that, in this case, the partition $\mathcal{P}$ is already decided.}
-  \note{PM}{The longwinded version is that induction is on $(x,y)$ where $x$ is the number of inner vertices of $N'$ and $y$ is the number of inner faces of $N'$.
+  \note{PM}{The longwinded version is that induction is on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$.
 The base cases are $(0,y)$ for any $y$.
@@
 
-We now move onto the case in which $N'$ contains at least one inner vertex.
-Our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that will be determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.
+We now move onto the case in which $N_0$ contains at least one inner vertex.
+Our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that will be determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N_0$.
 We first explain how $v_1$, $v_2$ and $v_3$ are chosen and then explain how these are used to define $S$.
@@
 \begin{compactenum}
-		\item If $k=1$ then let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
-
-		\item If $k= 2$ then let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
-
-		\item If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the appropriate vertex of $F_0$.
+		\item If $k=1$ then let $\tau:=v_1,\ldots,v_p$ be an inner face of $N_0$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
+
+		\item If $k= 2$ then let $\tau:=v_1,\ldots,v_p$ be an inner face of $N_0$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
+
+		\item If $k=3$ then assign each vertex $v$ of $N_0$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the appropriate vertex of $F_0$.
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
-Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N'$, add $d'-3$ edges inside of $F$ to split it into $3$-sided faces.
-This yields a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary.
-By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
-Let $\tau:=v_1,\ldots,v_p$ be  the facial cycle of $N'$ that contains $\tau''$.
+Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N_0$, add $d'-3$ edges inside of $F$ to split it into $3$-sided faces.
+This yields a plane supergraph $N_0'$ of $N_0$ in which each inner face has exactly three edges on its boundary.
+By Sperner's Lemma there exists an inner face $\tau':=v_1,v_2,v_3$ of $N_0'$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+Let $\tau:=v_1,\ldots,v_p$ be  the facial cycle of $N_0$ that contains $\tau'$.
 	\end{compactenum}
@@
 For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to a vertex in $V(F)$.
-Let $\overline{S}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
+Let $\overline{S}$ denote the subgraph of $N_0$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
 	Let $S:=V(\overline{S})\setminus V(F)$, let $Y:=(V(Q_1)\cup V(Q_2)\cup V(Q_3))\setminus V(F)$, and let $X:=S\setminus Y$.
@@
 		\end{center}
-		\caption{Three cases when choosing the face $\tau$ of $N'$.}
+		\caption{Three cases when choosing the face $\tau$ of $N_0$.}
 		\label{boring_figure}
@@
 	Refer to \cref{boring_figure}.
-Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{S}$ and the edges and vertices of $F$.
+Let $M$ denote the subgraph of $N_0$ containing the edges and vertices of $\overline{S}$ and the edges and vertices of $F$.
 By our choice of $\tau$, the graph $M$ is $2$-connected.
@@
 	Since $M$ is $2$-connected and $\overline{S}$ is connected, $F_i[P_{i,j}]$ is connected, for each $j\in\{1,2,3\}$.
-Let $N_i$ and $N_i'$ be the subgraphs of $N$ and $N'$ contained in the interior or boundary of $F_i$.
-Every inner face of $N_i'$ is an inner face of $N'$ and $\tau$ is not an inner face of $N_i'$.
-Therefore $N_i'$ has fewer inner faces than $N'$.
+Let $N_i$ and $N_{0,i}$ be the subgraphs of $N$ and $N_0$ contained in the interior or boundary of $F_i$.
+Every inner face of $N_{0,i}$ is an inner face of $N_0$ and $\tau$ is not an inner face of $N_{0,i}$.
+Therefore $N_{0,i}$ has fewer inner faces than $N_{0}$.
 Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.
@@
 
-  It is now straightforward to verify that $\mathcal{T}$ is a tree decomposition of $H':=N'/\mathcal{P}$ in which each bag has size at most $4$.
-To see that $\mathcal{T}$ is also a tree decomposition of $H:=N/\mathcal{P}$, consider any edge $vw\in E(N)\setminus E(N')$ with $v\in P_\alpha$ and $w\in P_\beta$.
+  It is now straightforward to verify that $\mathcal{T}$ is a tree decomposition of $H_0:=N_0/\mathcal{P}$ in which each bag has size at most $4$.
+To see that $\mathcal{T}$ is also a tree decomposition of $H:=N/\mathcal{P}$, consider any edge $vw\in E(N)\setminus E(N_0)$ with $v\in P_\alpha$ and $w\in P_\beta$.
 We must verify that some bag of $\mathcal{T}$ contains $\alpha$ and $\beta$.
 If both $v$ and $w$ are in $V(M)$, then $\alpha=x_a$ and $\beta=x_b$ for some $a,b\in\{0,\ldots,3\}$, in which case $B_{y_0}=\{x_0,\ldots,x_k\}$ contains $\alpha$ and $\beta$.
-If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $N\subseteq G$ crosses an edge of $M\subseteq G'$, both $v$ and $w$ are contained in the interior of the same face $F_i$ of $M$ so, by the inductive hypothesis there exists a bag of $\mathcal{T}_i$ that contains $\alpha$ and $\beta$ and this becomes a bag of $\mathcal{T}$ that contains $\alpha$ and $\beta$.
+If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $N\subseteq G$ crosses an edge of $M\subseteq G_0$, both $v$ and $w$ are contained in the interior of the same face $F_i$ of $M$ so, by the inductive hypothesis there exists a bag of $\mathcal{T}_i$ that contains $\alpha$ and $\beta$ and this becomes a bag of $\mathcal{T}$ that contains $\alpha$ and $\beta$.
 Finally, if $v\in V(M)$ and $w\not\in V(M)$ then $v\in P_{i,j}$ for some $j\in\{1,2,3\}$ and $w$ is in the interior of some face $F_i$ of $M$.
@@
 \begin{proof}[Proof of \cref{d_framed_product_stucture}]
-	Let $G$ be a $d$-framed multigraph with frame $G'$ having outer face $F_0$.	Let $T$ be a BFS forest of $G'$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.
+	Let $G$ be a $d$-framed multigraph with frame $G_0$ having outer face $F_0$.	Let $T$ be a BFS forest of $G_0$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.
 Let $P_1:= V(F_0)$, let $Y_1$ be any subset of $\min\{|P_1|,3\}$ vertices in $P_1$ and let $X_1:=P_1\setminus Y_1$.
-Now apply \cref{induction} on $G$, $G'$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$.
+Now apply \cref{induction} on $G$, $G_0$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$.
     This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree decomposition.
@@
 	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdots \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$.
-That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.
+That this is a layering of $G$ follows from the fact that $\dist_{G_0}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.
 Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
@@
         \includegraphics{figs/one_planar_example-4} \\
-        $G$ & $G'$
+        $G$ & $G_0$
      \end{tabular}
     \end{center}
-    \caption{The graph $G'$ obtained by removing pairs of crossing edges from $G$ is a plane multigraph whose faces all have three or four sides.}
+    \caption{The graph $G_0$ obtained by removing pairs of crossing edges from $G$ is a plane multigraph whose faces all have three or four sides.}
     \label{one_planar_example}
@@
 Refer to \cref{one_planar_example}.
-To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ that cross at $p$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
+To understand the structure of $G$, it is helpful to consider the plane graph $G'$ obtained by replacing each pair of edges $vw$ and $xy$ that cross at $p$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
 Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$.
@@
 Therefore each dummy vertex $w_i$ of $W$ is flanked by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
-Since $w_i$ has degree $4>1$ in $G_0$ and no two edges incident to a common vertex cross each other, $w_{i-1}\neq w_{i+1}$.
+Since $w_i$ has degree $4>1$ in $G'$ and no two edges incident to a common vertex cross each other, $w_{i-1}\neq w_{i+1}$.
 Since $G$ is edge-maximal the edge $w_{i-1}w_{i+1}$ is an edge of $G$ on the boundary of $F$.
-Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.
+Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G'$, each of which is contained in a different edge of $G$.
 
@@
 
-Consider some pair of edges $vw$ and $xy$ that cross at point $p$, which is a degree-4 vertex of $G_0$.
+Consider some pair of edges $vw$ and $xy$ that cross at point $p$, which is a degree-4 vertex of $G'$.
 Since no pair of edges incident to a common vertex cross each other, $v$, $w$, $x$, and $y$ are all distinct.
-There are four faces $F_1,\ldots,F_4$ of $G_0$ with $p$ on their boundary and, from the preceding discussion each of $vx$, $xw$, $wy$, and $yv$ is an uncrossed edge of $G$.
-Therefore, by removing each pair of crossing edges from $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges.
-Therefore $G$ is a $4$-framed graph with frame $G'$.
+There are four faces $F_1,\ldots,F_4$ of $G'$ with $p$ on their boundary and, from the preceding discussion each of $vx$, $xw$, $wy$, and $yv$ is an uncrossed edge of $G$.
+Therefore, by removing each pair of crossing edges from $G$ we obtain a plane multigraph $G_0$ each of whose faces is bounded by three or four edges.
+Therefore $G$ is a $4$-framed graph with frame $G_0$.
 % Since, for any pair of crossing edges $\{vw,xy\}$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.

### 2022-03-24 response.tex
Updated some responses

@@
 
-  We have avoided the introduction of any new notation.
+  We have avoided the introduction of any new notation, including $\phi$.
 The new presentation is much shorter and clearer.
@@
   Since the initial submission, there is now even more justification for including this special case.
-A result of Bose et al on vertex ranking relies critically on the simple treewidth of $H$.
-In particular, if $H$ is planar and has treewidth $3$, then $H\boxtimes P\boxtimes K_\ell$ an $r$-vertex ranking using $O(\log n/\log\log\log n)$ colours (for any fixed $r$ and $\ell$).
-The three logs in the denominator come from the fact that $H$ has simple treewidth $3$.
-Simple treewidth $4$ or even (non-simple) treewidth $3$ lead to more logs in the denominator.
-The number of logs in the denominator is exactly the simple treewidth of $H$, and this is tight.
+A result of Bose et al on vertex ranking relies critically on the simple treewidth of $H$, as does the result of D\,ebski et al on $p$-centered colouring.
+In both case, if $H$ does not have simple treewidth $3$ then one does not get an asymptotically tight bound.
 
@@
 This, in turn, leads to better results for $d$-map graphs.
+
+  The former Theorem~10 (on $1$-planar graphs) is now a consequence of  Theorem~23 on ($d$-framed graphs).
 \end{response}
@@
 
-\done
+\textcolor{blue}{Fixed, though in a different way.
+Since we have to define lowest common $T$-ancestors anyway, we define the subtree $T'$ of $T$ to be rooted at at the lowest common $T$-ancestor of $V(T')$.}
 
@@
 
-\done
+\textcolor{blue}{This sentence is now gone.}
 

### 2022-03-24 response.tex
comments and revisisions

@@
 \begin{response}
-  Since the initial submission, there is now even more justification for including this special case.
-A result of Bose et al on vertex ranking relies critically on the simple treewidth of $H$, as does the result of D\,ebski et al on $p$-centered colouring.
+There is a lot of motivation for proving this result with $H$ of treewidth 3 and planar.
+First, the treewidth bound is best possible (even for planar graphs).
+Having $H$ planar means that $H$ is in fact a planar 3-tree (also called stack polytope or Appolonian etwork).
+It follows that we get better bounds for queue-number for example.
+
+Since the initial submission, there is now even more justification for including this material.
+A result of Bose et al on vertex ranking relies critically on the simple treewidth of $H$, as does the result of Debski et al on $p$-centered colouring.
 In both case, if $H$ does not have simple treewidth $3$ then one does not get an asymptotically tight bound.

### 2022-03-24 k-planar.tex
comments and revisisions

@@
 %      genus $g$ & $2g+6$ & 1 & \cite{UWY} \\
-      apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\[1ex]
+      apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
       % B-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
@@
       $(g,\delta)$-string & $\binom{\delta+4}{3}-1$ & $\max\{2g,3\}(\delta^4 + 4\delta^3 + 9\delta^2 + 10\delta+4)$ & \cref{StringProduct} \\
-      $k$-NN & $O(k^6)$ & $O(k^4)$ & \cref{nn_product_structure} \\
+      $k$-nearest-neighbour & $O(k^6)$ & $O(k^4)$ & \cref{nn_product_structure} \\
       $d$-framed & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{d_framed_product_stucture} \\
@@
       $(g,d)$-map & $9$ & $\max\{2g,3\}\,(7d^2 -21d)$ & \cref{gdMapProduct} \\ \hline 
-      \multicolumn{4}{l}{$^*$ these bounds also applies to the simple treewidth of $H$.}
+      \multicolumn{4}{l}{$^*$ these bounds also apply to the simple treewidth of $H$.}
     \end{tabular}
@@
 \end{table}
+
+\note{DW}{why is the $2g+6$ commented out from the table?}
 
@@
 
+\note{DW}{Suggested section name ``Graphs Derived From Framed Graphs'' or just 
+``Framed Graphs''}
+
 % Next we consider
@@
 Thus \cref{ShortcutProduct} is applicable.
-However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.\footnote{A multigraph $G$ is \defin{contained} in a graph $G'$ if the simple graph underlying $G$ is isomorphic to a subgraph of $G'$.}
+However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.\footnote{A multigraph $G$ is \defin{contained} in a graph $X$ if the simple graph underlying $G$ is contained in $X$.}
 
@@
 The base cases are $(0,y)$ for any $y$.
-Unfortunately, we can't just induct on $x$ because it doesn't always decrease.}
+Unfortunately, we can't just induct on $x$ because it doesn't always decrease.} \note{DW}{We need to make it clear what we are doing induction on.
+I suggest we write, ``we proceed by  
+induction on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$. ''}
 	$\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iii):
@@
 This happens when two nations and two lakes meet at a vertex $x$.
-It's not an issue but it means we can't write ``the face of $G_1$ that corresponds to $x$.''}
+It's not an issue but it means we can't write ``the face of $G_1$ that corresponds to $x$.''} \note{DW}{Good}
 Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations.
-If two lakes have a common edge in their boundaries, then we may delete the edges and merge the lakes.
+If two lakes have a common edge in their boundaries, then we may delete the edges and merge the lakes \note{DW}{Delete the previous sentence?
+This sentence is incompatible with ``in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes'' below.}.
 Now assume that no two lakes have a common edge in their boundaries.
@@
 
-For each face $F$ of $G_1$ that corresponds to a vertex $x$ of $G_0$, the size of $F$ equals the number of nations incident to $x$, which is at most $d$.
-So every face of $G_1$ has size at most $d$.  Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.
+For each face $F$ of $G_1$ that corresponds to a vertex $x$ of $G_0$, the size of $F$ \note{DW}{Is ``size'' defined?
+Do we mean length of the facial cycle?
+} equals the number of nations incident to $x$, which is at most $d$.
+So every face of $G_1$ has size at most $d$.
+Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.
+\note{DW}{I like this notation.
+Should we use it throughout?}
 
@@
 If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common.
-If some face of $G_1$ corresponds to $x$ then $v$ and $w$ are on a common face of $G_1$ so $vw\in E(\widehat{G}_1)$.
-If no face of $G_1$ corresponds to $x$\footnote{This can occur when $x$ is on the boundary of exactly two nation faces and two lake faces.} then the only two nation faces of $G_0$ with $x$ on their boundary are those corresponding to $v$ and $w$, in which case $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.
+If some face of $G_1$ corresponds to $x$, then $v$ and $w$ are on a common face of $G_1$ so $vw\in E(\widehat{G}_1)$.
+If no face of $G_1$ corresponds to $x$\footnote{This can occur when $x$ is on the boundary of exactly two nation faces and two lake faces.}, then the only two nation faces of $G_0$ with $x$ on their boundary are those corresponding to $v$ and $w$, in which case $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.
 Therefore $G\subseteq\widehat{G}_1$.
@@
 As discussed in the introduction, product structure has been used to resolve a number of problems on planar graphs.
-In most cases, these results hold for any graph with product structure.
+In most cases, these results hold for any graph class with product structure.
 In this section, we survey some of the consequences of this for the graph classes considered in the previous two sections.
@@
 
-For an integer $k\geq 0$, a \textit{$k$-queue layout} of a graph $G$ consists of a linear ordering $\preceq$ of $V(G)$ and a partition $\{E_1,E_2,\dots,E_k\}$ of $E(G)$, such that for $i\in\{1,2,\dots,k\}$, no two edges in $E_i$ are nested with respect to $\preceq$.
+For an integer $k\geq 0$, a \defin{$k$-queue layout} of a graph $G$ consists of a linear ordering $\preceq$ of $V(G)$ and a partition $\{E_1,E_2,\dots,E_k\}$ of $E(G)$, such that for $i\in\{1,2,\dots,k\}$, no two edges in $E_i$ are nested with respect to $\preceq$.
 That is, it is not the case that $v\prec x \prec y \prec w$ for edges $vw,xy\in E_i$.
-The \textit{queue-number} of a graph $G$, denoted by $\qn(G)$, is the minimum integer $k$ such that $G$ has a $k$-queue layout.
+The \defin{queue-number} of a graph $G$, denoted by \defin{$\qn(G)$}, is the minimum integer $k$ such that $G$ has a $k$-queue layout.
 Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
@@
 
-Since all the product structure theorems presented thus far upper bound the treewidth of $H$,
-\cref{qn} immediately implies bounds on the queue number of all graphs in these classes.
-An improvement can be made for $1$-planar and $d$-map graphs since, in the cases $H$ is planar and has treewidth $3$.
-\citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$. Thus the graph $H$ in \cref{1_planar_product,dMapProduct} has queue-number at most $5$.
+Since all the product structure theorems presented thus far upper bound the treewidth of $H$, \cref{qn} immediately implies bounds on the queue number of all graphs in these classes.
+An improvement can be made for $1$-planar and $d$-map graphs since then $H$ is planar with treewidth at most $3$.
+\citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$.
+Thus the graph $H$ in \cref{1_planar_product,dMapProduct} has queue-number at most $5$.
 % , which with \cref{qn} implies:
@@
   The following bounds hold for the queue number of any graph from each of the following classes:
-  \begin{compactenum}
+  \begin{compactitem}
     \item $k$-planar: $2^{O(k^3)}$
@@
     \item $(g,d)$-map: $\max\{1,g\}\cdot 2^{O(d)}$
-    \item $k$-NN: $2^{O(k^6)}$
-    % \item $(g,d)$-map graph
-    % \item
-  \end{compactenum}
+    \item $k$-nearest-neighbour: $2^{O(k^6)}$
+  \end{compactitem}
 \end{cor}
-\note{PM}{For uniformity, I got rid of the two corollaries on $d$-map and $1$-planar graphs and kept them in \cref{q_cor}.
+%\note{PM}{For uniformity, I got rid of the two corollaries on $d$-map and $1$-planar graphs and kept them in \cref{q_cor}.
 The text is still commented out in the source in case someone feels strongly about keeping them.}
@@
     % \item $(g,d)$-map: $g4^{O(d)}$
-    % \item $k$-NN: $4^{O(k^6)}$
+    % \item $k$-nearest-neighbour: $4^{O(k^6)}$
     \item $(g,d)$-map graph: $(7d^2-21d)\cdot 4^{10}$
@@
     % \item $(g,d)$-map: $g4^{O(d)}$
-    % \item $k$-NN: $4^{O(k^6)}$
+    % \item $k$-nearest-neighbour: $4^{O(k^6)}$
     \item $(g,d)$-map: $\max\{2g,3\}\cdot 7d(d-3)(p+1)\cdot \binom{p+9}{9}$
@@
 
-As mentioned in \cref{Introduction}, \citet{DEJGMM21,EJM} prove results about adjacency labelling schemes for planar graphs.
+As mentioned in \cref{Introduction}, \citet{DEJGMM21} and \citet{EJM} prove results about adjacency labelling schemes for planar graphs.
 Stated in terms of universal graphs, their main theorem is interpreted as follows:
@@
 	\label{UniversalUniversal}
-	For every fixed graph $X$ and all fixed integers $d,\delta,\Delta,g,k>0$, and every integer $n>0$, there exists a graph $U_n$ with $n^{1+o(1)}$ vertices and edges such that $U_n$ contains all of the following graphs as induced subgraphs:
+	For every fixed graph $X$ and all fixed integers $d,\delta,\Delta,g,k>0$, and every integer $n>0$, there exists a graph $U_n$ with $n^{1+o(1)}$ vertices and edges such that $U_n$ contains the following graphs as induced subgraphs:
 	\begin{compactitem}

### 2022-03-25 k-planar.tex
refs clean-up

@@
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-
 \usepackage{todonotes}
-
-\newcommand{\arXiv}[1]{arXiv:\,\href{https://arxiv.org/abs/#1}{#1}}
-%\newcommand{\doi}[1]{\href{https://doi.org/#1}{\tt https://doi.org/#1}}
 
@@
 \begin{proof}[Proof of \cref{d_framed_product_stucture}]
-	Let $G$ be a $d$-framed multigraph with frame $G_0$ having outer face $F_0$.	Let $T$ be a BFS forest of $G_0$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.
+	Let $G$ be a $d$-framed multigraph with frame $G_0$ having outer face $F_0$
+	\note{DW}{Replace ``outer face'' by ``outer-cycle'' since a face is a defined to be a portion of the plane.
+}.	Let $T$ be a BFS forest of $G_0$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.
 Let $P_1:= V(F_0)$, let $Y_1$ be any subset of $\min\{|P_1|,3\}$ vertices in $P_1$ and let $X_1:=P_1\setminus Y_1$.
@@
     This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree decomposition.
-Since any graph having a $3$-simple tree decomposition is planar \cite{knauer.ueckerdt:simple,kratochvil.vaner:note,elmallah.colbourn:on}, this implies that $H$ is planar.
-
-	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdots \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$.
-That this is a layering of $G$ follows from the fact that $\dist_{G_0}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.
-Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
+Since any graph having a $3$-simple tree decomposition is planar \cite{knauer.ueckerdt:simple,kratochvil.vaner:note,elmallah.colbourn:on}, $H$ is planar.
+
+	Define $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdots \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$.
+This is a layering of $G$ since $\dist_{G_0}(v,w)\le \floor{ \frac{d}{2} }$ for each edge $vw\in E(G)$.
+Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\floor{\frac{d}{2}}$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \floor{ \frac{d}{2}} + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
 The result now follows from \cref{PartitionProduct}.
@@
     \item $1$-planar: $115$ 
-    \item $d$-map: $\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$ 
+    \item $d$-map: $\floor{ \frac{33}{2} (d+3\floor{\frac{d}{2}}-3) }$ 
     \item $(g,d)$-map: $\max\{1,g\}\cdot 2^{O(d)}$
@@
     \item $1$-planar: $1792$
-    \item $d$-map: $256(d+3\floor{d/2}-3)$
+    \item $d$-map: $256(d+3\floor{\frac{d}{2}}-3)$
     % \item $(g,d)$-map: $g4^{O(d)}$

### 2022-03-25 k-planar.tex
Boarding a flight now....

@@
       planar & $3^*$ & $3$ & \cite{DJMMUW20} \\
+      planar & $4^*$ & $2$ & \cite{bose.morin.ea:optimal} \\
       planar & $6^*$ & $1$ & \cite{UWY} \\
       genus $g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW} \\
-%      genus $g$ & $2g+6$ & 1 & \cite{UWY} \\
+      genus $g$ & $2g+6$ & 1 & \cite{UWY} \\
       apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
@@
 
-\note{DW}{why is the $2g+6$ commented out from the table?}
+% \note{DW}{why is the $2g+6$ commented out from the table?}
 
@@
 
-\section{Framed Graphs, $1$-Planar Graphs, and (Plane) Map Graphs}
+\section{Framed Graphs}
 \label{FramedSection}
 
-\note{DW}{Suggested section name ``Graphs Derived From Framed Graphs'' or just 
-``Framed Graphs''}
+% \note{DW}{Suggested section name ``Graphs Derived From Framed Graphs'' or just``Framed Graphs''}
 
@@
 
-	The proof is by induction on the number of inner faces of $N_0$.  A particular easy case occurs when $N_0$ has no inner vertices, i.e., $V(N_0)=V(F)$. In this case we can simply take
-	\note{DW}{Suggested revision, ``In the base case, $N_0$ has only one inner face, and we can simply take ...} 
+  We proceed by induction on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$.
+The induction base case occurs when $N_0$ has no inner vertices, i.e., $V(N_0)=V(F)$.
+In this case we can simply take
+	% \note{DW}{Suggested revision, ``In the base case, $N_0$ has only one inner face, and we can simply take ...}
   % \note{PM}{I debated what to put here since, even when $N'$ has only one face, the rules below work for picking $\tau$.
@@
 I settled on this ``base case'' since it's clear that, in this case, the partition $\mathcal{P}$ is already decided.}
-  \note{PM}{The longwinded version is that induction is on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$.
+  % \note{PM}{The longwinded version is that induction is on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$.
 The base cases are $(0,y)$ for any $y$.
 Unfortunately, we can't just induct on $x$ because it doesn't always decrease.} \note{DW}{We need to make it clear what we are doing induction on.
-I suggest we write, ``we proceed by  
-induction on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$. ''}
+% I suggest we write, ``we proceed by induction on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$. ''}
 	$\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iii):

### 2022-03-25 k-planar.tex
Suggesting new proof of d-dmap to d-frame

@@
 
+\note{PM}{Suggested new Proof}
+\begin{proof}
+  Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations and let $G$ be the corresponding map graph.
+Let $G_0^*$ be the dual graph of $G_0$, so that each vertex of $G_0^*$ corresponds to face of $G_0$ and vice-versa.
+Thus, each vertex of $G_0^*$ is a nation vertex or a lake vertex.
+
+  Let $x$ be a vertex of $G_0$, let $F_x$ be the corresponding face of $G_0^*$, and let $v_0,\ldots,v_r$ be the facial cycle of $F_x$.
+Let $C:=w_0,\ldots,w_s$ be the subsequence of $v_0,\ldots,v_r$ consisting of only the nation vertices.
+Since $x$ is incident to at most $d$ nations, $r\le d$.
+We call $C_x$ the \emph{nation cycle} of $F_x$.
+
+  Let $G_1$ be a plane supergraph of $G_0^*$ obtained by adding each nation cycle of each face $F$ of $G^0$ and then triangulating each face with more than $d$ vertices on its boundary.
+Each nation cycle is now a face of $G_1$ and no face of $G_1$ has more than $d$ vertices on its boundary.
+Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.
+
+  Since $\widehat{G}_1$ is a $d$-framed graph, it suffices to show that these definitions imply that $E(G)\subseteq E(\widehat{G}_1)$.
+Indeed, if $vw\in E(G)$ then the nation faces corresponding to $v$ and $w$ have a common vertex $x$ on their boundary.
+The vertex $x$ corresponds to a face $F_x$ in $G_0^*$ and the facial cycle of $F_x$ contains $v$ and $w$.
+Therefore, the nation cycle $C_x$ of $F_x$ contains $v$ and $w$.
+Since $C_x$ a face in $G_1$ the vertices $vw\in E(G_1)$.
+\end{proof}
+
+\note{PM}{Old proof}
+
 \begin{proof}

### 2022-03-25 k-planar.tex
Some definitions around embedded graphs

@@
 We first formally define $k$-planar graphs.
-An \defin{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a curve\footnote{A \defin{curve} in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
+For a surface $\Sigma$,  an \defin{$\Sigma$-embedded graph} $G$ is a graph with $V(G)\subset\Sigma$ in which each edge $vw\in E(G)$ is a curve\footnote{A \defin{curve} in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
 The points $f(0)$ and $f(1)$ are called the \defin{endpoints} of the curve.
 When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.
-The \defin{interior} of $f$ is the point set $\{f(t):0<t<1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
+The \defin{interior} of $f$ is the point set $\{f(t):0<t<1\}$.} in $\Sigma$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
 The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
-A \defin{crossing} in an embedded (multi)graph $G$ is an unordered pair of distinct edges $\{vw,xy\}\subseteq E(G)$ whose interiors have a point in common.
-Without any loss of generality we may assume that the interiors of any two edges of $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.
+A \defin{crossing} in a $\Sigma$-embedded (multi)graph $G$ is an unordered pair of distinct edges $\{vw,xy\}\subseteq E(G)$ whose interiors have a point in common.
+
+If an embedded graph $G$  has no crossings then it is \defin{non-crossing} graph.
+When $G$ is non-crossing and $F$ is a face of $G$ then $V(F)$ denotes the set of vertices in $G$ on the boundary of $F$.
+When $F$ is homeomorphic to a disk, the \defin{facial walk} of $F$ is the cycle consisting of the edges and vertices of $G$ on the boundary of $F$.
+When $\Sigma=\R^2$ is the Euclidean plane and each edge of $G$ is involved in at most $k$ crossings then $G$ is a $k$-plane graph.
+A graph $G$ is \defin{$k$-planar} if is isomorphic to some $k$-plane graph.
+A $0$-plane graph is a \defin{plane} graph and a $0$-planar graph is a \defin{planar} graph.
+
+Without any loss of generality we may assume that the interiors of any two edges of a $\sigma$-embedded graph $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.
 This assumption can be enforced by local changes that do not change the number of crossings each edge is involved in.
@@
 The definition of $k$-planar graphs naturally generalises to genus-$g$ surfaces.
-A graph $G$ embedded on a surface $\Sigma$ is \defin{$(\Sigma,k)$-plane} if every edge of $G$ is involved in at most $k$ crossings.
+A graph $\Sigma$-embedded graph $G$ is \defin{$(\Sigma,k)$-plane} if every edge of $G$ is involved in at most $k$ crossings.
 A graph $G$ is \defin{$(g,k)$-planar} if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
@@
 	Let $G$ be a $d$-framed multigraph with frame $G_0$ having outer face $F_0$
-	\note{DW}{Replace ``outer face'' by ``outer-cycle'' since a face is a defined to be a portion of the plane.
-}.	Let $T$ be a BFS forest of $G_0$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.
+	% \note{DW}{Replace ``outer face'' by ``outer-cycle'' since a face is a defined to be a portion of the plane. }.
+  %\note{PM}{I added some conventions around the definition of $\Sigma$-embedded graph so that $V(F)$ is defined for a face $F$ in a non-crossing embedded graph.}
+  Let $T$ be a BFS forest of $G_0$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.
 Let $P_1:= V(F_0)$, let $Y_1$ be any subset of $\min\{|P_1|,3\}$ vertices in $P_1$ and let $X_1:=P_1\setminus Y_1$.

### 2022-03-25 k-planar.tex
I'm happy with this....

@@
 \begin{abstract}
-Dujmovi\'c~et~al.~[\emph{J.~ACM}~'20] recently proved that every planar graph is isomorphic to a subgraph of the strong product of a graph of bounded treewidth and a path.
+Dujmovi\'c~et~al.~[\emph{J.~ACM}~'20] recently proved that every planar graph is isomorphic to a subgraph of the strong product of a bounded treewidth graph and a path.
 Analogous results were obtained for graphs of bounded Euler genus or apex-minor-free graphs.
@@
 A path is \defin{trivial} if it has length 0 and \defin{non-trivial} otherwise.} and
-\item for every $v\in V(G)$, the number of paths in $\SS$ that use $v$ as an internal vertex is at most $d$.
+\item for every $v\in V(G)$, the number of paths in $\SS$ that include $v$ as an internal vertex is at most $d$.
 \end{compactitem}
@@
 Using this result, we derive improved product structure theorems for $1$-planar graphs and planar map graphs, again with a planar graph $H$.
-  \item \Cref{Applications} quickly surveys applications of product structure and the consequences of the current work.
+  \item \Cref{Applications} quickly surveys applications of product structure theorems and the consequences of the current work.
 \end{compactitem}
@@
 Since $P$ is a connected subgraph of $G$, $H[\{x_0,\ldots,x_r\}]$ is a connected subgraph of $H$.
-Since $\mathcal{T}$ is a normalized tree decomposition of $H$, some node $x_i$ is a $T$-ancestor of all nodes in $x_0,\ldots,x_r$.
+Since $\mathcal{T}$ is a normalized tree decomposition of $H$, \tref{ancestor-edge} implies that some node $x_i$ is a $T$-ancestor of all nodes in $x_0,\ldots,x_r$.
 
@@
 Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so at least one of $x$ or $y$ is a $T$-ancestor of the other.
+This completes the proof of \cref{i-ancestor}
 \end{proof}
@@
 Then $\overrightarrow{W}$ is a directed path in $\overrightarrow{H}^+$ from $x$ to $a$ of length at most $s\le r\le k$.
-This completes the proof.
+This completes the proof of \cref{general-bag-size}.
 \end{proof}
@@
 \cref{d_framed_product_stucture} is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs.
-For simple 1-planar graphs it is essential that we allow frames with parallel edges.
+Even for simple 1-planar graphs it is essential that we allow frames with parallel edges.
 
@@
 		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
-		\item Let $F$ be a cycle in $G'$ that $V(F)$ is partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
+		\item Let $F$ be a cycle in $G_0$ whose vertices are partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
 		\begin{compactenum}
@@
 Let $N_i$ and $N_{0,i}$ be the subgraphs of $N$ and $N_0$ contained in the interior or boundary of $F_i$.
+Every inner vertex of $N_{0,i}$ is an inner vertex of $N_{0}$.
 Every inner face of $N_{0,i}$ is an inner face of $N_0$ and $\tau$ is not an inner face of $N_{0,i}$.
-Therefore $N_{0,i}$ has fewer inner faces than $N_{0}$.
+Therefore $N_{0,i}$ has no more inner vertices than $N_0$ and has fewer inner faces than $N_{0}$.
 Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.
@@
 \note{DW}{I like this notation.
-Should we use it throughout?}
+Should we use it throughout?} \note{PM}{No.
+For most things I prefer $G$ and $G_0$ rather than $\widehat{G}_0$ and $G_0$.}
 
@@
 Combining \cref{non-repetitive} with our results on product structure, we obtain the following corollary:
-\begin{cor}\label{nonrepetitive_cor}
-  The following bounds hold for the nonrepetitive chromatic number of any graph from each of the following classes:
+\begin{cor}\label{non-repetitive_cor}
+  The following bounds hold for the non-repetitive chromatic number of any graph from each of the following classes:
   \begin{compactitem}
@@
 Prior to the current work, the strongest upper bound on the non-repetitive chromatic number of $n$-vertex  $k$-planar graphs was $O(k\log n)$ \cite{dujmovic.morin.ea:layered}.
-Our results also give bounds on the nonrepetitive chromatic number of $(g,\delta)$-string graphs and $k$-nearest-neighbour graphs, but these bounds are weaker than existing results based on maximum degree.
+Our results also give bounds on the non-repetitive chromatic number of $(g,\delta)$-string graphs and $k$-nearest-neighbour graphs, but these bounds are weaker than existing results based on maximum degree.
 The bounds obtained from product structure are exponential in $\delta$ and $k^2$, respectively.
-However, each of these graph classes has degree bounded by $O(\delta)$ and $O(k^2)$, respectively, and therefore have nonrepetitive chromatic number $O(\delta^2)$ and $O(k^{4})$, respectively \cite{DJKW16}.
+However, each of these graph classes has degree bounded by $O(\delta)$ and $O(k^2)$, respectively, and therefore have non-repetitive chromatic number $O(\delta^2)$ and $O(k^{4})$, respectively \cite{DJKW16}.
 
@@
 
-As with nonrepetitive chromatic number, our results also imply bounds on the $p$-centered chromatic numer of $(g,\delta)$-string graphs, powers of bounded-degree graphs, and $k$-nearest-neighbour graphs, but these bounds are weaker than the bounds implied by the fact that these graphs have bounded maximum degree \cite{DFMS21}.
+As with non-repetitive chromatic number, our results also imply bounds on the $p$-centered chromatic numer of $(g,\delta)$-string graphs, powers of bounded-degree graphs, and $k$-nearest-neighbour graphs, but these bounds are weaker than the bounds implied by the fact that these graphs have bounded maximum degree \cite{DFMS21}.
 

### 2022-03-29 k-planar.tex
minor updates
comments on map graph proof

@@
 All these results generalise for drawings of graphs on arbitrary surfaces.
-In fact, we work in a much more general setting based on so-called shortcut systems that are of independent interest.
+In fact, we work in a more general setting based on so-called shortcut systems, which are of independent interest.
 This leads to analogous results for certain types of map graphs, string graphs, graph powers, and nearest neighbour graphs.
@@
   }
-  \caption{Known product structure theorems (of the form $G\subseteq H\boxtimes P\boxtimes K_\ell$).}
+  \caption{Product structure theorems (of the form $G\subseteq H\boxtimes P\boxtimes K_\ell$).}
   \label{summary_table}
@@
  $d\geq 3$, every $d$-map graph is a spanning subgraph of a $d$-framed graph.
+\note{DW}{Does this need to be ``spanning subgraph of a $d$-framed multigraph''?}
 \end{lem}
@@
 \note{PM}{Suggested new Proof}
+\note{DW}{I am happy with the new proof, once we address the issues.}
+
 \begin{proof}
   Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations and let $G$ be the corresponding map graph.
-Let $G_0^*$ be the dual graph of $G_0$, so that each vertex of $G_0^*$ corresponds to face of $G_0$ and vice-versa.
-Thus, each vertex of $G_0^*$ is a nation vertex or a lake vertex.
-
-  Let $x$ be a vertex of $G_0$, let $F_x$ be the corresponding face of $G_0^*$, and let $v_0,\ldots,v_r$ be the facial cycle of $F_x$.
-Let $C:=w_0,\ldots,w_s$ be the subsequence of $v_0,\ldots,v_r$ consisting of only the nation vertices.
-Since $x$ is incident to at most $d$ nations, $r\le d$.
-We call $C_x$ the \emph{nation cycle} of $F_x$.
-
-  Let $G_1$ be a plane supergraph of $G_0^*$ obtained by adding each nation cycle of each face $F$ of $G^0$ and then triangulating each face with more than $d$ vertices on its boundary.
+Let $G_0^*$ be the dual graph of $G_0$.
+So the vertices of $G_0^*$ correspond to faces of $G_0$, and each vertex of $G_0^*$ is a nation vertex or a lake vertex.
+
+  Let $x$ be a vertex of $G_0$, let $F_x$ be the corresponding face of $G_0^*$, and let $v_0,\ldots,v_s$ be the facial cycle of $F_x$.
+Let $C:=w_0,\ldots,w_r$ be the subsequence of $v_0,\ldots,v_s$ consisting of only the nation vertices.
+Since $x$ is incident to at most $d$ nations, $r\le d$ \note{DW}{Why not $r+1\leq d$?
+Do we mean that $v_0=v_r$?
+I suggest we write $(v_1,\dots,v_s)$ and $(w_1,\dots,w_r)$.
+}.
+Call $C_x$ the \defin{nation cycle} of $F_x$.
+\note{DW}{What if there are 1 or 2 nations incident to $x$?
+Is $C_x$ a cycle?
+Do we need parallel edges here?}
+
+  Let $G_1$ be a plane supergraph of $G_0^*$ obtained by adding each nation cycle of each face $F$ of $G^0$ \note{DW}{We mean `` face $F$ of $G^*_0$'', right?} and then triangulating each face with more than $d$ vertices on its boundary.
 Each nation cycle is now a face of $G_1$ and no face of $G_1$ has more than $d$ vertices on its boundary.
@@
 
-  Since $\widehat{G}_1$ is a $d$-framed graph, it suffices to show that these definitions imply that $E(G)\subseteq E(\widehat{G}_1)$.
+By definition, $V(G)=V(\widehat{G}_1)$.
+To prove the claim, it suffices to show that $E(G)\subseteq E(\widehat{G}_1)$.
 Indeed, if $vw\in E(G)$ then the nation faces corresponding to $v$ and $w$ have a common vertex $x$ on their boundary.
@@
 Therefore, the nation cycle $C_x$ of $F_x$ contains $v$ and $w$.
-Since $C_x$ a face in $G_1$ the vertices $vw\in E(G_1)$.
+Since $C_x$ bounds a face in $G_1$, $vw\in E(\widehat{G}_1)$.
 \end{proof}

### 2022-03-31 k-planar.tex
read the whole paper, lots of  minor touch-ups, added a  few comments

@@
 
-For the important special case of $1$-planar graphs, we are able to obtain better constants and an additional property (planarity) of $H$.
+For the important special case of $1$-planar graphs, we obtain the following result with $H$ planar and a best possible bound on the treewidth of $H$.
 
@@
 For at least two applications, the planarity of $H$ in \cref{1_planar_product} is critical in obtaining asymptotically tight bounds \cite{BDJM,DFMS21}.
-\cref{1_planar_product} is proven in the more general setting of $d$-framed multigraphs.
-This more general result also implies a product structure theorem for $d$-map graphs in which $H$ is a planar.
-
-%
-%
+\cref{1_planar_product} is proven in the more general setting of $d$-framed multigraphs, which also implies a product structure theorem for $d$-map graphs in which $H$ is a planar with treewidth 3.
+
 % These theorems have applications in diverse areas, which we explore in \cref{Applications}.
@@
 \subsection{Shortcut Systems}
+
+\note{DW}{I'm not comfortable about saying that $k$-planar graphs are our most high-profile target. Why are they more important than map graphs, say?}
 
@@
 Some examples are explored in \cref{first_example,mid_example,last_example}.
+\note{DW}{I suggest we replace ``\cref{first_example,mid_example,last_example}''
+by ``\cref{examples}''.}
 
@@
 
+\note{DW}{I suggest the table headers are not in bold}
+
 \begin{table}[h]
@@
     \begin{tabular}{lccl}
+    	\hline
       \textbf{Graph class} & \boldmath $\tw(H)$ & \boldmath $\ell$ & \textbf{Reference} \\ \hline
@@
 
-The purpose of this section is to prove \cref{ShortcutProduct} and its specialization to $k$-planar graphs, \cref{kPlanarProduct}.
+This section proves \cref{ShortcutProduct} and its specialization to $k$-planar graphs, \cref{kPlanarProduct}.
 While strong products enable concise statements of the theorems in \cref{Introduction}, to prove such results it is helpful to work with layerings and partitions, which we now introduce.
@@
 The \defin{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.
-The \defin{lowest common $T$-ancestor} of a non-empty set $S\subseteq V(T)$ is the maximum $T$-depth node of $T$ that is a $T$-ancestor of every node in $S$.
+The \defin{lowest common $T$-ancestor} of a non-empty set $S\subseteq V(T)$ is the $T$-ancestor of every node in $S$ with maximum $T$-depth.
 In the following, we treat any subtree $T'$ of $T$ as a rooted tree whose root is the lowest common $T$-ancestor of $V(T')$.
@@
 \begin{compactenum}[(i)]
-  \item $\PP$ has small layered width with respect to the layering $\mathcal{L}$ of $G$ and that
+  \item $\PP$ has small layered width with respect to the layering $\mathcal{L}$ of $G$, and 
   \item $J$ has small treewidth.
@@
 We now proceed with Step~(ii), showing that $J$ has small treewidth.
-To accomplish this, we construct a small width tree-decomposition $\mathcal{C}:=(C_x:x\in V(T))$ of $J$ using the same tree $T$ used in the tree decomposition $\mathcal{T}$ of $H$.
+To accomplish this, we construct a small width tree-decomposition $\mathcal{C}:=(C_x:x\in V(T))$ of $J$ using the same tree $T$ used in the tree-decomposition $\mathcal{T}$ of $H$.
 The following claim will be useful in showing that the resulting decomposition has small width.
@@
 Since $P$ is a connected subgraph of $G$, $H[\{x_0,\ldots,x_r\}]$ is a connected subgraph of $H$.
-Since $\mathcal{T}$ is a normalized tree decomposition of $H$, \tref{ancestor-edge} implies that some node $x_i$ is a $T$-ancestor of all nodes in $x_0,\ldots,x_r$.
+Since $\mathcal{T}$ is a normalized tree-decomposition of $H$, \tref{ancestor-edge} implies that some node $x_i$ is a $T$-ancestor of all nodes in $x_0,\ldots,x_r$.
 
@@
 
-  It remains to show that each bag in this tree decomposition has size at most $\binom{k+t}{t}$.
+  It remains to show that each bag in this tree-decomposition has size at most $\binom{k+t}{t}$.
 We will do this by appealing to an elegant result of \citet{PS21} on $k$-reachability in directed skeletons, which we now explain.
 
-  Let $H^+$ denote the supergraph of $H$ that contains the edge $xy$ if and only if some bag $B_z\in\mathcal{T}$ contains both $x$ and $y$ (i.e., $H^+$ is the chordal closure of $H$ with respect to $\mathcal{T}$).
+  Let $H^+$ denote the supergraph of $H$ that contains the edge $xy$ if and only if some bag $B_z\in\mathcal{T}$ contains both $x$ and $y$ (that is, $H^+$ is the chordal closure of $H$ with respect to $\mathcal{T}$).
 Let $\overrightarrow{H}^+$ be the directed graph obtained by directing each edge $xy$ of $H^+$ in the direction $\overrightarrow{xy}$ so that $y$ is a $T$-ancestor of $x$.
@@
 We first formally define $k$-planar graphs.
-For a surface $\Sigma$,  an \defin{$\Sigma$-embedded graph} $G$ is a graph with $V(G)\subset\Sigma$ in which each edge $vw\in E(G)$ is a curve\footnote{A \defin{curve} in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
+For a surface $\Sigma$,  a \defin{$\Sigma$-embedded graph} $G$ is a graph with $V(G)\subset\Sigma$ in which each edge $vw\in E(G)$ is a curve\footnote{A \defin{curve} in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$.
 The points $f(0)$ and $f(1)$ are called the \defin{endpoints} of the curve.
@@
 When $F$ is homeomorphic to a disk, the \defin{facial walk} of $F$ is the cycle consisting of the edges and vertices of $G$ on the boundary of $F$.
-When $\Sigma=\R^2$ is the Euclidean plane and each edge of $G$ is involved in at most $k$ crossings then $G$ is a $k$-plane graph.
+If  $\Sigma=\R^2$ is the Euclidean plane and each edge of $G$ is involved in at most $k$ crossings, then $G$ is a $k$-plane graph.
 A graph $G$ is \defin{$k$-planar} if is isomorphic to some $k$-plane graph.
@@
 
-Without any loss of generality we may assume that the interiors of any two edges of a $\sigma$-embedded graph $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.
+Without any loss of generality we may assume that the interiors of any two edges of a $\Sigma$-embedded graph $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.
 This assumption can be enforced by local changes that do not change the number of crossings each edge is involved in.
-If $p$ is a common point in the interior of two crossing edges $vw$ and $xy$ then we say that $vw$ and $xy$ \defin{cross at} $p$.
+If $p$ is a common point in the interior of two crossing edges $vw$ and $xy$, then we say that $vw$ and $xy$ \defin{cross at} $p$.
 
@@
 \cref{kPlanarProduct} shows that every $k$-planar graph is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $O(k^3)$ where $\ell\leq O(k^2)$.
-In some applications, the treewidth of $H$ is much more important than the value of $\ell$, which leads to the following question:
+In some applications, the treewidth of $H$ is much more significant than the value of $\ell$, which leads to the following question:
 
@@
 The definition of $k$-planar graphs naturally generalises to genus-$g$ surfaces.
-A graph $\Sigma$-embedded graph $G$ is \defin{$(\Sigma,k)$-plane} if every edge of $G$ is involved in at most $k$ crossings.
+A $\Sigma$-embedded graph $G$ is \defin{$(\Sigma,k)$-plane} if every edge of $G$ is involved in at most $k$ crossings.
 A graph $G$ is \defin{$(g,k)$-planar} if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$.
@@
 	Fix a $(\Sigma,k)$-plane drawing of $G_0$ in a surface $\Sigma$ of Euler genus at most $g$. Choose $\epsilon>\epsilon'>0$.
-	For each vertex $v$ of $G_0$, let $B_v$ be the \defin{ball} $\{x\in \mathbb{R}^2:\dist(x,v)\leq\epsilon\}$
+	For each vertex $v$ of $G_0$, let $B_v$ be the \defin{ball} $\{x\in \mathbb{R}^2:\dist(x,v)\leq\epsilon\}$. 
 	For each edge $vw$ of $G_0$, let $C_{vw}$ be the \defin{channel} $\{x\in \mathbb{R}^2:\dist(x,vw)\leq\epsilon'\}\setminus(B_v\cup B_w)$.
@@
 \begin{thm}\label{nn_product_structure}
-  Every $k$-nearest neighbour graph is a subgraph of $H\boxtimes P\boxtimes K_{\ell}$ for some graph $H$ of treewidth $O(k^6)$ and some $\ell\in O(k^4)$.
+  Every $k$-nearest neighbour graph is a subgraph of $H\boxtimes P\boxtimes K_{\ell}$ for some graph $H$ of treewidth $O(k^6)$ and for some $\ell\in O(k^4)$.
 \end{thm}
@@
 
-Every $d$-framed multigraph can be described by a $(2,d(d-3)/2)$-shortcut system applied to a plane multigraph by adding a vertex inside each face $F$ of $G_0$ adjacent to the vertices of $F$ and creating a shortcut between each non-consecutive pair of vertices of $F$.
+Every $d$-framed multigraph can be described by a $(2,d(d-3)/2)$-shortcut system applied to a plane multigraph by adding a vertex inside each face $F$ of $G_0$ adjacent to the vertices of $F$ and creating a shortcut between each non-adjacent pair of vertices in $V(F)$.
 Thus \cref{ShortcutProduct} is applicable.
@@
 \cref{d_framed_product_stucture} is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs.
-Even for simple 1-planar graphs it is essential that we allow frames with parallel edges.
+Note that even for simple 1-planar graphs it is essential that we allow frames with parallel edges.
 
 \cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite{DJMMUW20}.
-The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{v_1,\ldots,v_t\}$.
+The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A $T$-decomposition of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags contain $v_1,\ldots,v_t$.
 Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that for each $v\in V(G)$,  \[\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.\]
@@
 		\item For each $x\in V(H)$, $S_x$ has a partition $\{X_x,Y_x\}$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
-		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$ then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
+		\item $H$ has a $3$-simple tree-decomposition $\mathcal{T}$, and if $k=3$ then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
 	\end{compactenum}
@@
 %    \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
-%    \item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
+%    \item $H$ has a $3$-simple tree-decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
 %  \end{compactenum}
@@
   We proceed by induction on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$.
-The induction base case occurs when $N_0$ has no inner vertices, i.e., $V(N_0)=V(F)$.
+The induction base case occurs when $N_0$ has no inner vertices (that is, $V(N_0)=V(F)$).
 In this case we can simply take
@@
 		\item Assumption 4(b) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
-		\item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
+		\item The trivial tree-decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
 	\end{compactenum}
@@
 
-		\item If $k=3$ then assign each vertex $v$ of $N_0$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the appropriate vertex of $F_0$.
+		\item If $k=3$ then assign each vertex $v$ of $N_0$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $V(F)$ encountered on the path in $T$ from $v$ to the appropriate vertex of $F_0$.
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
@@
 This yields a plane supergraph $N_0'$ of $N_0$ in which each inner face has exactly three edges on its boundary.
-By Sperner's Lemma there exists an inner face $\tau':=v_1,v_2,v_3$ of $N_0'$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
+By Sperner's Lemma (see \citep{Proofs4}) there exists an inner face $\tau':=v_1,v_2,v_3$ of $N_0'$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
 Let $\tau:=v_1,\ldots,v_p$ be  the facial cycle of $N_0$ that contains $\tau'$.
@@
 
-	We can now define the vertices in $S$.
-For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to a vertex in $V(F)$.
-Let $\overline{S}$ denote the subgraph of $N_0$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
-	Let $S:=V(\overline{S})\setminus V(F)$, let $Y:=(V(Q_1)\cup V(Q_2)\cup V(Q_3))\setminus V(F)$, and let $X:=S\setminus Y$.
-Observe that, for each $i\in\{1,2,3\}$ and each $j\ge 0$, $|V(Q_i)\cap L_j|\le 1$.
-Therefore $|Y\cap L_j|\le 3$ for each integer $j\ge 0$.
-Furthermore, $X\subseteq V(\tau)\setminus\{x_1,x_2,x_3\}$, so $|X|\le p-3\le d-3$.
-Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ satisfies condition~(ii).
-
-	% We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
-First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $L_i$, each incident on at most two edges of $\overline{Y}$.
-Since $\dist_{G'}(v,w)\le 2$ for each $vw\in E(G)$, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $L_i$.
-These at most six edges have at most 12 endpoints.
-Therefore $|V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i|\le 6\times 2=12$, so $|V(\overline{Y}^+)\cap L_i|\le 12+3=15$.
-
-	\begin{figure}
+\begin{figure}[!b]
 	\begin{center}
@@
 
-	Refer to \cref{boring_figure}.
-Let $M$ denote the subgraph of $N_0$ containing the edges and vertices of $\overline{S}$ and the edges and vertices of $F$.
-By our choice of $\tau$, the graph $M$ is $2$-connected.
+	We now define the vertices in $S$.
+For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to a vertex in $V(F)$.
+Let $\overline{S}$ denote the subgraph of $N_0$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.
+	Let $S:=V(\overline{S})\setminus V(F)$, let $Y:=(V(Q_1)\cup V(Q_2)\cup V(Q_3))\setminus V(F)$, and let $X:=S\setminus Y$.
+Observe that $|V(Q_i)\cap L_j|\le 1$ 	for each $i\in\{1,2,3\}$ and each $j\ge 0$.
+Therefore $|Y\cap L_j|\le 3$ for each $j\ge 0$.
+Furthermore, $X\subseteq V(\tau)\setminus\{x_1,x_2,x_3\}$, so $|X|\le p-3\le d-3$.
+Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ satisfies condition~(ii).
+
+	% We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
+First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $L_i$, each incident on at most two edges of $\overline{Y}$.
+Since $\dist_{G'}(v,w)\le 2$ for each $vw\in E(G)$, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $L_i$.
+These at most six edges have at most 12 endpoints.
+Therefore $|V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i|\le 6\times 2=12$, so $|V(\overline{Y}^+)\cap L_i|\le 12+3=15$.
+
+Let $M$ denote the subgraph of $N_0$ containing the edges and vertices of $\overline{S}$ and the edges and vertices of $F$. By our choice of $\tau$, the graph $M$ is $2$-connected.
   % Since $N'\subseteq G'$ contains only edges of $N\subseteq G$ that are not crossed, each component of $N-V(M)$ is contained in a single face of $M$.
@@
 Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.
-The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of $N_i$.
-	% The partition $\mathcal{P}_i$ contains each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$.
-	We can now define our partition
+The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of $N_i$ satisfying (i)--(iii) above.
+We now define our partition
 	\[
@@
 	It is straightforward to verify that $\mathcal{P}$ is indeed a partition of $V(N)$ and, by definition this is an $H$-partition of $N$ for the graph $H:=N/\mathcal{P}$.
-All that remains is to verify that $\mathcal{P}$ satisfies (i)--(iii).
-
-	By construction $\mathcal{P}$ contains parts $P_1,\ldots,P_k$, so $\mathcal{P}$ satisifes (i).
+It remains to verify that $\mathcal{P}$ satisfies (i)--(iii).
+
+	By construction $\mathcal{P}$ contains parts $P_1,\ldots,P_k$, so $\mathcal{P}$ satisfies (i).
 By Assumption~4(b) each of the parts $P_1,P_2,P_3$ satisfies condition~(ii).
@@
 
-	All that remains is to construct a tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$ of $H$ that satisfies (iii).
+	It remains to construct a tree-decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$ of $H$ that satisfies (iii).
 Let $x_0$ be the vertex of $H$ obtained from contracting $S$ and, for each $j\in\{1,\ldots,k\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.
 The tree $T_0$ has a node $y_0$ with bag $B_{y_0}:=\{x_0,\ldots,x_k\}$.
-The inductive hypothesis ensures that, for each $i\in\{1,\ldots,m\}$,  $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $\mathcal{T}_i:=(B_z:z\in V(T_i))$ that satisfies (iii).
+The inductive hypothesis ensures that, for each $i\in\{1,\ldots,m\}$,  $H_i:=N_i/\mathcal{P}_i$ has a tree-decomposition $\mathcal{T}_i:=(B_z:z\in V(T_i))$ that satisfies (iii).
 Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively.
@@
 For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.
-These replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$.
+These replacements do not increase the size of any bag and they result in a tree-decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$.
 Now add an edge that joins $y_0$ to $z_i$ so that $T_i$ becomes a subtree of $T_0$ that is adjacent to $y_0$.
-We perform this operation for each $i\in\{1,\ldots,m\}$ to obtain our final tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$.
-
-  It is now straightforward to verify that $\mathcal{T}$ is a tree decomposition of $H_0:=N_0/\mathcal{P}$ in which each bag has size at most $4$.
-To see that $\mathcal{T}$ is also a tree decomposition of $H:=N/\mathcal{P}$, consider any edge $vw\in E(N)\setminus E(N_0)$ with $v\in P_\alpha$ and $w\in P_\beta$.
+We perform this operation for each $i\in\{1,\ldots,m\}$ to obtain our final tree-decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$.
+
+  It is now straightforward to verify that $\mathcal{T}$ is a tree-decomposition of $H_0:=N_0/\mathcal{P}$ in which each bag has size at most $4$.
+To see that $\mathcal{T}$ is also a tree-decomposition of $H:=N/\mathcal{P}$, consider any edge $vw\in E(N)\setminus E(N_0)$ with $v\in P_\alpha$ and $w\in P_\beta$.
 We must verify that some bag of $\mathcal{T}$ contains $\alpha$ and $\beta$.
 If both $v$ and $w$ are in $V(M)$, then $\alpha=x_a$ and $\beta=x_b$ for some $a,b\in\{0,\ldots,3\}$, in which case $B_{y_0}=\{x_0,\ldots,x_k\}$ contains $\alpha$ and $\beta$.
-If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $N\subseteq G$ crosses an edge of $M\subseteq G_0$, both $v$ and $w$ are contained in the interior of the same face $F_i$ of $M$ so, by the inductive hypothesis there exists a bag of $\mathcal{T}_i$ that contains $\alpha$ and $\beta$ and this becomes a bag of $\mathcal{T}$ that contains $\alpha$ and $\beta$.
-Finally, if $v\in V(M)$ and $w\not\in V(M)$ then $v\in P_{i,j}$ for some $j\in\{1,2,3\}$ and $w$ is in the interior of some face $F_i$ of $M$.
+If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $N\subseteq G$ crosses an edge of $M\subseteq G_0$, both $v$ and $w$ are contained in the interior of the same face $F_i$ of $M$.
+So, by the inductive hypothesis there exists a bag of $\mathcal{T}_i$ that contains $\alpha$ and $\beta$ and this becomes a bag of $\mathcal{T}$ that contains $\alpha$ and $\beta$.
+Finally, if $v\in V(M)$ and $w\not\in V(M)$, then $v\in P_{i,j}$ for some $j\in\{1,2,3\}$ and $w$ is in the interior of some face $F_i$ of $M$.
 By induction, some bag of $\mathcal{T}_i$ contains $\beta$ and $\alpha'$, where $\alpha'$ is the vertex of $H_i$ obtained by contracting $P_{i,j}$.
 However, before $T_i$ is attached to $T_0$ each occurence of $\alpha'$ in $\mathcal{T}_i$ is replaced by $\alpha$, so some bag of $\mathcal{T}$ contains both $\alpha$ and $\beta$.
-Therefore $\mathcal{T}$ is a tree decomposition of $H$.
+Therefore $\mathcal{T}$ is a tree-decomposition of $H$.
 
@@
 
-	All that remains is to show that any triple of vertices of $H$ is contained in at most two bags of $\mathcal{T}$.
+It remains to show that any triple of vertices of $H$ is contained in at most two bags of $\mathcal{T}$.
 By the inductive hypothesis, the only triples we need to be concerned about are the $3$-element subsets of  $\{x_0,x_1,\ldots,x_k\}$.
@@
 When $k=2$, $B_{y_0}=\{x_0,x_1,x_2\}$.
-In this case, the fact that $\tau$ contains an edge $v_1v_2$ of $F$ with $v_1\in P_{x_1}$ and $v_2\in P_{x_2}$ ensures that there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_1$, and $P_2$ on its boundary.
-(Refer to the middle of \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.
+In this case, the fact that $\tau$ contains an edge $v_1v_2$ of $F$ with $v_1\in P_{x_1}$ and $v_2\in P_{x_2}$ ensures that there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_1$, and $P_2$ on its boundary  (see the middle part of \cref{boring_figure}).
+By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.
 Therefore there are at most two bags, $B_y$ and $B_{z_i}$ of $\mathcal{T}$ that contain $\{x_0,x_1,x_2\}$.
@@
 
-	We claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.
+	We claim that, for distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.
 Suppose to the contrary that there are two such faces $F_{i_1}$ and $F_{i_2}$.
@@
 Now apply \cref{induction} on $G$, $G_0$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$.
-    This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree decomposition.
-Since any graph having a $3$-simple tree decomposition is planar \cite{knauer.ueckerdt:simple,kratochvil.vaner:note,elmallah.colbourn:on}, $H$ is planar.
+    This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree-decomposition.
+Since any graph having a $3$-simple tree-decomposition is planar \cite{knauer.ueckerdt:simple,kratochvil.vaner:note,elmallah.colbourn:on}, $H$ is planar.
 
@@
 In the following paragraphs, subscripts on the vertices in $W$ are implicitly taken modulo $r$.
-By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
+By edge-maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
 
@@
 
-\note{PM}{Old proof}
-
-\begin{proof}
-% \note{PM}{I fixed some issues here that come from the fact that some vertices of $G_0$ do not contribute faces to $G_1$.
+%\note{PM}{Old proof}
+%
+%\begin{proof}
+%% \note{PM}{I fixed some issues here that come from the fact that some vertices of $G_0$ do not contribute faces to $G_1$.
 This happens when two nations and two lakes meet at a vertex $x$.
 It's not an issue but it means we can't write ``the face of $G_1$ that corresponds to $x$.''} \note{DW}{Good}
-Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations.
+%Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations.
 If two lakes have a common edge in their boundaries, then we may delete the edges and merge the lakes \note{DW}{Delete the previous sentence?
@@
 Now assume that no two lakes have a common edge in their boundaries.
-Let $G$ be the corresponding map graph.
-Let $G_1$ be the graph with $V(G_1):=V(G)$, which is the set of nations, and whose edge set we now define.  Consider vertices $v$ and $w$ of $G_1$.
-Let $A$ and $B$ be the nation faces of $G_0$ corresponding to $v$ and $w$.
-Then $vw\in E(G_1)$ if $A$ and $B$ have a common edge in their boundaries,
-or there is a vertex $x$ of $G_0$ in the boundary of $A$ and $B$, and in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes.
-% Then $G_1$ has Euler genus at most $g$.
-Each face of $G_1$ corresponds to a vertex of $G_0$ or a lake of $G_0$.
-Triangulate each face of $G_1$ corresponding to a lake of $G_0$.
-
-For each face $F$ of $G_1$ that corresponds to a vertex $x$ of $G_0$, the size of $F$ \note{DW}{Is ``size'' defined?
+%Let $G$ be the corresponding map graph.
+%Let $G_1$ be the graph with $V(G_1):=V(G)$, which is the set of nations, and whose edge set we now define.  Consider vertices $v$ and $w$ of $G_1$.
+%Let $A$ and $B$ be the nation faces of $G_0$ corresponding to $v$ and $w$.
+%Then $vw\in E(G_1)$ if $A$ and $B$ have a common edge in their boundaries,
+%or there is a vertex $x$ of $G_0$ in the boundary of $A$ and $B$, and in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes.
+%% Then $G_1$ has Euler genus at most $g$.
+%Each face of $G_1$ corresponds to a vertex of $G_0$ or a lake of $G_0$.
+%Triangulate each face of $G_1$ corresponding to a lake of $G_0$.
+%
+%For each face $F$ of $G_1$ that corresponds to a vertex $x$ of $G_0$, the size of $F$ \note{DW}{Is ``size'' defined?
 Do we mean length of the facial cycle?
@@
 For most things I prefer $G$ and $G_0$ rather than $\widehat{G}_0$ and $G_0$.}
-
-To complete the proof, we now show that $\widehat{G}_1$ contains every edge of $G$.
-If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common.
+%
+%To complete the proof, we now show that $\widehat{G}_1$ contains every edge of $G$.
+%If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common.
 If some face of $G_1$ corresponds to $x$, then $v$ and $w$ are on a common face of $G_1$ so $vw\in E(\widehat{G}_1)$.
@@
 Therefore $G\subseteq\widehat{G}_1$.
-\end{proof}
-
-% \note{DW}{give a better version of \cref{MapPartition} with $g=0$. }
-
-%So $(g,d)$-map graphs generalise graphs embedded in a surface, and we now assume that $d\geq 4$ for the remainder of this section.
+%\end{proof}
 
@@
 \citet{dujmovic.eppstein.ea:structure} proved that $G$ has layered treewidth $(4g+6)(k+1)$.
-Van den Heuvel and Wood~\citep{vdHW18} showed that this implies that $G$  has $r$-strong colouring number at most $(4g + 6)(k + 1)(2r + 1)$.
+Van den Heuvel and Wood~\citep{vdHW18} showed this implies that $G$  has $r$-strong colouring number at most $(4g + 6)(k + 1)(2r + 1)$.
 By a result of \citet{zhu:colouring}, $G$ has $r$-weak colouring number at most $( (4g + 6)(k + 1)(2r + 1) )^r$, which by another result of  \citet{zhu:colouring} implies that $G$ has  $p$-centered chromatic number at most $( (4g+6)(k+1)(2^{p-1} + 1) )^{2^{p-2}}$.

### 2022-04-05 response.tex
added some suggestions, updated response

@@
 First, the treewidth bound is best possible (even for planar graphs).
-Having $H$ planar means that $H$ is in fact a planar 3-tree (also called stack polytope or Appolonian etwork).
+Having $H$ planar means that $H$ is in fact a planar 3-tree (also called stack polytope or Appolonian network).
 It follows that we get better bounds for queue-number for example.
@@
 
-\tba
+\textcolor{blue}{All the bags at this point are non-empty. We have revised to make this clear. }
 
@@
 Indeed, while cleaning up this proof we have also improved the constant from $30$ to $7$.
-The proof is quite similar to the previous proof except that graph $G'$ is obtained by removing both crossing edges involved in a crossing (the spars of a kite), rather than just one.
+The proof is similar to the previous proof except that graph $G'$ is obtained by removing both crossing edges involved in a crossing (the spars of a kite), rather than just one.
 We now use the language of framed multigraph, for consistency with the literature.
@@
 
-\tba
+\textcolor{blue}{This proof is now removed since the authors of [10] included this result in the journal version of their paper.}
 
@@
 
-\textcolor{blue}{This proof is now gone since the authors of [10] helpfully include this result in the journal version of their paper.}
+\textcolor{blue}{This proof is now removed since the authors of [10] included this result in the journal version of their paper.}
 

### 2022-04-05 k-planar.tex
added some suggestions, updated response

@@
 
-\note{DW}{I'm not comfortable about saying that $k$-planar graphs are our most high-profile target. Why are they more important than map graphs, say?}
+\note{DW}{I'm uncomfortable about saying that $k$-planar graphs are our most high-profile target. Why are they more important than map graphs or nearest neighbour graphs? I suggest ``
+The above result for $k$-planar graphs (\cref{kPlanarProduct}) is in fact a special case of a more general result that relies on the following definition. A non-empty set $\SS$  ....''}
 

### 2022-04-06 k-planar.tex
Last little bits

--- 2022-03-31 k-planar.tex
@@
 
-\note{DW}{I'm not comfortable about saying that $k$-planar graphs are our most high-profile target. Why are they more important than map graphs, say?}
-
-Although $k$-planar graphs are the most high-profile target for a generalization of \cref{PlanarProduct}, we actually prove a more general result than \cref{kPlanarProduct} using the following definition.
+% \note{DW}{I'm uncomfortable about saying that $k$-planar graphs are our most high-profile target. Why are they more important than map graphs or nearest neighbour graphs? I suggest ``
+% The above result for $k$-planar graphs (\cref{kPlanarProduct}) is in fact a special case of a more general result that relies on the following definition. A non-empty set $\SS$  ....''}
+
+The above result for $k$-planar graphs (\cref{kPlanarProduct}) is in fact a special case of a more general result that relies on the following definition.
 A non-empty set $\SS$ of non-trivial paths in a graph $G$ is a \defin{$(k,d)$-shortcut system} (for $G$) if:
@@
 \cref{ShortcutProduct} is applicable for many graph classes in addition to $k$-planar graphs.
-Some examples are explored in \cref{first_example,mid_example,last_example}.
-\note{DW}{I suggest we replace ``\cref{first_example,mid_example,last_example}''
-by ``\cref{examples}''.}
+Some examples are explored in \cref{examples}.
+
+% \note{DW}{I suggest we replace ``\cref{first_example,mid_example,last_example}''by ``\cref{examples}''.}
 
@@
 
-\note{DW}{I suggest the table headers are not in bold}
+% \note{DW}{I suggest the table headers are not in bold}
 
@@
     	\hline
-      \textbf{Graph class} & \boldmath $\tw(H)$ & \boldmath $\ell$ & \textbf{Reference} \\ \hline
+      {Graph class} & $\tw(H)$ & $\ell$ & {Reference} \\ \hline
       planar & $3^*$ & $3$ & \cite{DJMMUW20} \\
@@
 integer %s $g\geq 0$ and
- $d\geq 3$, every $d$-map graph is a spanning subgraph of a $d$-framed graph.
-\note{DW}{Does this need to be ``spanning subgraph of a $d$-framed multigraph''?}
+ $d\geq 3$, every $d$-map graph is a spanning subgraph of a $d$-framed multigraph.
 \end{lem}
 
-\note{PM}{Suggested new Proof}
-\note{DW}{I am happy with the new proof, once we address the issues.}
+% \note{PM}{Suggested new Proof}
+% \note{DW}{I am happy with the new proof, once we address the issues.}
 
@@
 
-  Let $x$ be a vertex of $G_0$, let $F_x$ be the corresponding face of $G_0^*$, and let $v_0,\ldots,v_s$ be the facial cycle of $F_x$.
-Let $C:=w_0,\ldots,w_r$ be the subsequence of $v_0,\ldots,v_s$ consisting of only the nation vertices.
-Since $x$ is incident to at most $d$ nations, $r\le d$ \note{DW}{Why not $r+1\leq d$?
-Do we mean that $v_0=v_r$?
-I suggest we write $(v_1,\dots,v_s)$ and $(w_1,\dots,w_r)$.
-}.
+  Let $x$ be a vertex of $G_0$, let $F_x$ be the corresponding face of $G_0^*$, and let $v_1,\ldots,v_s$ be the facial cycle of $F_x$.
+Let $C:=w_1,\ldots,w_r$ be the subsequence of $v_1,\ldots,v_s$ consisting of only the nation vertices.
+Since $x$ is incident to at most $d$ nations, $r\le d$.
   Call $C_x$ the \defin{nation cycle} of $F_x$.
-\note{DW}{What if there are 1 or 2 nations incident to $x$?
-Is $C_x$ a cycle?
-Do we need parallel edges here?}
-
-  Let $G_1$ be a plane supergraph of $G_0^*$ obtained by adding each nation cycle of each face $F$ of $G^0$ \note{DW}{We mean `` face $F$ of $G^*_0$'', right?} and then triangulating each face with more than $d$ vertices on its boundary.
+
+  \note{DW}{What if there are 1 or 2 nations incident to $x$? Is $C_x$ a cycle? Do we need parallel edges here?}
+
+  \note{PM}{If $r=1$ then the ``nation cycle'' has no edges, so this doesn't contribute any edges to $G_1$, which is the right thing to do.}
+
+  \note{PM}{If $r=2$ then the ``nation cycle'' has one edge, so this contributes one edge to $G_1$ (which is needed, because the corresponding nations are adjacent, but no other edges are needed.)}
+
+  Let $G_1$ be a plane supergraph of $G_0^*$ obtained by adding each nation cycle of each face $F$ of $G_0$ and then triangulating each face with more than $d$ vertices on its boundary.
 Each nation cycle is now a face of $G_1$ and no face of $G_1$ has more than $d$ vertices on its boundary.

### 2022-04-07 k-planar.tex
addedc omment re map graphs

@@
 
-% \note{DW}{I suggest we delete ``We use $G'\subseteq G$ to denote subgraph containment; that is, $V(G')\subseteq V(G)$ and $E(G')\subseteq E(G)$. '' This is absolutely standard.}
-
 The starting point for this work is the following `product structure theorem' for planar graphs\footnote{In this paper, all graphs are finite and undirected.
@@
 
-% \note{DW}{Mention other application such as vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, odd colourings~\citep{DMO}?
-We write a lot about adjacency labellings and universal induced subgraphs.
-Do we need so much detail here?
-We also mention this work in \cref{WrappingUp}.}
-
-% \note{PM}{Agreed. I cut this down to a list of famous open problem. We don't need to sell this so hard anymore.}
-
 All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this statement.} These applications motivate finding product structure theorems for other graph classes.
@@
 
-%\note{DW}{Do we want to mention ``every $1$-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ of treewidth $3$ and for some path $P$'' here?}
-%\note{PM}{Yes, I think so.  Two theorems, like this, or one?}
-%\note{DW}{two theorems}
-
 For the important special case of $1$-planar graphs, we obtain the following result with $H$ planar and a best possible bound on the treewidth of $H$.
@@
 
-% A planar graph of treewidth at most $3$ is said to have \emph{simple treewidth} $3$.
-
 For at least two applications, the planarity of $H$ in \cref{1_planar_product} is critical in obtaining asymptotically tight bounds \cite{BDJM,DFMS21}.
@@
 \subsection{Shortcut Systems}
-
-% \note{DW}{I'm uncomfortable about saying that $k$-planar graphs are our most high-profile target. Why are they more important than map graphs or nearest neighbour graphs? I suggest ``
-% The above result for $k$-planar graphs (\cref{kPlanarProduct}) is in fact a special case of a more general result that relies on the following definition. A non-empty set $\SS$  ....''}
 
@@
 
-% \note{PM}{Suggested new Proof}
-% \note{DW}{I am happy with the new proof, once we address the issues.}
-
 \begin{proof}
@@
 Call $C_x$ the \defin{nation cycle} of $F_x$.
-
-  \note{DW}{What if there are 1 or 2 nations incident to $x$? Is $C_x$ a cycle? Do we need parallel edges here?}
-
-  \note{PM}{If $r=1$ then the ``nation cycle'' has no edges, so this doesn't contribute any edges to $G_1$, which is the right thing to do.}
-
-  \note{PM}{If $r=2$ then the ``nation cycle'' has one edge, so this contributes one edge to $G_1$ (which is needed, because the corresponding nations are adjacent, but no other edges are needed.)}
-
-  Let $G_1$ be a plane supergraph of $G_0^*$ obtained by adding each nation cycle of each face $F$ of $G_0$ and then triangulating each face with more than $d$ vertices on its boundary.
-Each nation cycle is now a face of $G_1$ and no face of $G_1$ has more than $d$ vertices on its boundary.
+Note that if $r=1$ then the ``nation cycle'' has no edges, and if $r=2$ then the ``nation cycle'' has one edge.
+\note{DW}{Perhaps we shouldn't call it the ``nation cycle'' since it is not necessarily a cycle.}
+
+  Let $G_1$ be a plane supergraph of $G_0^*$ obtained by adding the nation cycle of each face $F$ of $G_0$, and then triangulating each face with more than $d$ vertices on its boundary.
+Each nation cycle \note{DW}{add `` of length at least 3''} is now a face of $G_1$, and no face of $G_1$ has more than $d$ vertices on its boundary.
 Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.

### 2022-04-07 k-planar.tex
Done(?!)

@@
 Note that if $r=1$ then the ``nation cycle'' has no edges, and if $r=2$ then the ``nation cycle'' has one edge.
-\note{DW}{Perhaps we shouldn't call it the ``nation cycle'' since it is not necessarily a cycle.}
+  % \note{DW}{Perhaps we shouldn't call it the ``nation cycle'' since it is not necessarily a cycle.}
+  % \note{PM}{I'm happy to call it a nation cycle and then use sarcastic quotes for the two oddball cases.}
 
   Let $G_1$ be a plane supergraph of $G_0^*$ obtained by adding the nation cycle of each face $F$ of $G_0$, and then triangulating each face with more than $d$ vertices on its boundary.
-Each nation cycle \note{DW}{add `` of length at least 3''} is now a face of $G_1$, and no face of $G_1$ has more than $d$ vertices on its boundary.
+Each nation cycle of length at least $3$ is now a face of $G_1$, and no face of $G_1$ has more than $d$ vertices on its boundary.
 Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.
@@
 Therefore, the nation cycle $C_x$ of $F_x$ contains $v$ and $w$.
-Since $C_x$ bounds a face in $G_1$, $vw\in E(\widehat{G}_1)$.
+\note{PM}{I distinguished the length $2$ case here.} If $C_x$ has length $2$ then $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.
+If $C_x$ has length at least $3$ then $C_x$ bounds a face in $G_1$, so $vw\in E(\widehat{G}_1)$.
 \end{proof}

### 2022-04-07 k-planar.tex
minor
removed all the \notes, minor revisions

--- 2022-04-07 k-planar.tex
@@
 
-% \item \citet{DFMS21} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs.
-% \item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results starting with the work of \citet{kannan.naor.ea:implicit}).
-% \item \citet{BDJM} use it to show that every $n$-vertex planar graph has local vertex ranking using $O(\log n/\log\log\log n)$ colours.
-% % \item The result of \citet{DEJGMM21} implies that, for every integer
-% $n>0$, there is a `induced-universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
-% subgraph of $U_n$.
-This has recently been extended by \citet{EJM}, who show the existence of a `subgraph-universal' graph with $(1+\epsilon)n$ vertices that contains every $n$-vertex planar graph as a subgraph as well as the existence of an `induced universal' graph with $n^{1+o(1)}$ vertices \emph{and} edges.
-The former result is the first progress on subgraph-universal graphs for planar graphs since the $O(n^{3/2})$-vertex construction of \citet{babai.chung.ea:on}.
-% aking the first progress on this problem since the work of \citet{babai.xx} in 1984.
-% \end{compactitem}
-
 All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this statement.} These applications motivate finding product structure theorems for other graph classes.
@@
 A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.} and \citet{DEMWW22} do so for bounded-degree graphs in any minor-closed class.
-See
-% \cref{Generalisations} for more precise statements and see
-\citep{DHJLW21} for a survey on this topic.
+See \citep{DHJLW21} for a survey on this topic.
 
@@
 \subsection{$k$-Planar Graphs}
-% \label{k_planar_section}
 
@@
 
-% These theorems have applications in diverse areas, which we explore in \cref{Applications}.
-
-% This theorem has applications in diverse areas, including queue layouts  \citep{DJMMUW20}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, $p$-centered colouring  \citep{DFMS21}, and adjacency labelling \citep{DEJGMM21}, which we explore in \cref{Applications}.
-For example, we prove that $k$-planar graphs have bounded non-repetitive chromatic number (for fixed $k$).
-Prior to the recent work of \citet{dujmovic.esperet.ea:planar}, it was even open whether planar graphs have bounded non-repetitive chromatic number.
-%
-% \referee{1}{3. pages 2 – 4: The paragraph following Theorem 2 and Section 1.3
-% (except for the fact that the apex assumption is necessary) is repetition
-% of things discussed earlier on in the introduction.}
-%
-% \note{PM}{Kind of, but I suggest we keep that paragraph and section anyway.
-Justification: They're both short and they describe the exact bounds for bounded genus graphs, which you don't get from the one-paragraph description in Section~1.0.}
-%
-% \note{DW}{I suggest we write, ``This theorem has applications in diverse areas, which we explore in \cref{Applications}....''}
-
 \subsection{Shortcut Systems}
@@
 
-% \referee{2}{Page 3. $P$ is used twice in the statement of Theorem 3.  It is probably
-% safer (and necessary?) to use a different variable for the second
-% occurrence.}
-
-% \note{DW}{Use $\SS$ for shortcut system, okay?}
-% \note{PM}{Ok.  I've implemented this change.  I was careful, but will still reread.}
-
-Theorem~\ref{PlanarProduct}(b), \cref{ShortcutProduct}, and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$.
+\cref{PlanarProduct}(b), \cref{ShortcutProduct}, and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$.
 Some further observations presented in \cref{k_planar_section} lead to the improved result.
@@
 
-% \note{DW}{I suggest we replace ``\cref{first_example,mid_example,last_example}''by ``\cref{examples}''.}
-
-% Here is one example.
-The \defin{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.\footnote{For a graph $G$ and two vertices $v,w\in V(G)$, $\dist_G(v,w)$  denotes the length of a shortest path, in $G$, with endpoints $v$ and $w$.
-We define $\dist_G(v,w):=\infty$ if $v$ and $w$ are in different connected components of $G$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\SS$ for some $(k,2k\Delta^{k})$-shortcut system $\SS$; see \cref{PowerShortcut}.
-Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} then imply:
-
-% \begin{thm}
-% \label{kPowerBasic}
-% For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is contained in $H\boxtimes P\boxtimes K_{6k^2(k^2+3)\Delta^{k}}$ for some graph $H$ of treewidth at most $\binom{k+3}{3}-1$ and some path $P$.
-% \end{thm}
-
-% \cref{Examples} presents further examples of graph classes that can be constructed using shortcut systems, including certain types of map graphs, string graphs, and $k$-nearest neighbour graphs.
-\cref{ShortcutProduct} implies product structure theorems for each of these classes.
-All of the above-mentioned applications also hold for these examples.
-
 \subsection{Overview and Outline}
@@
 \cref{summary_table} summarizes existing results on product structure theorems for minor-closed graph classes and new results for non-minor-closed graph classes.
-
-% \note{DW}{I suggest the table headers are not in bold}
 
@@
 
-% \note{DW}{why is the $2g+6$ commented out from the table?}
-
 The remainder of this paper is organized as follows:
@@
 
-A \defin{layering} of a graph $G$
-% is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$
-% \note{DW}{just write ``
-is an ordered partition $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that for every edge
- % ...''}
-$vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
+A \defin{layering} of a graph $G$ is an ordered partition $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.
 For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$.
@@
 For any partition $\PP$ of $V(G)$, we define the \defin{layered width} of $\PP$ as the minimum, over all layerings $\mathcal{L}$ of $G$, of the layered width of $\PP$ with respect to $\mathcal{L}$.
-
-% \citet{DJMMUW20} introduced the study of partitions with bounded layered width such that the quotient has some additional desirable property, like small treewidth.
-Dujmovi\'c~et~al.\ define a class $\mathcal{G}$ of graphs to \defin{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.
 
@@
 The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.\footnote{\citet{bose.morin.ea:optimal} have recently given linear time algorithms for computing layered $H$-partitions of planar graphs.}
-
-% \note{PM}{This feels out of place now that this section has been moved out of the introduction.  Delete?
-%
-% At the core of their work is the elegant proof by \citet{PS21} of the following result:
-%
-% \begin{thm}[\citep{PS21}]
-% \label{ps}
-%   Every planar triangulation $G$ has an $H$-partition $\PP$ such that $\tw(H)\leq 8$ and $G[P]$ is a shortest path in $G$ for each $P\in\PP$.
-% \end{thm}
-%
-% Indeed, the above-mentioned result of \citet{DJMMUW20} is a slight strengthening of \cref{ps}, where for each $P\in\PP$ no two vertices of $P$ have the same distance to some fixed root vertex $r$.
-% }
 
@@
 
-% For each node $x\in V(T)$, define
-% \[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
-% to be the maximal subtree of $T$ rooted at $x$.
-
-% \note{PM}{I have significantly simplified this section, which should make the referees happy. I will read it very carefully again tomorrow}
-
 We begin with a standard technique that allows us to work with a normalised tree-decomposition whose tree has the same vertex set as the graph it decomposes:
@@
   \begin{compactenum}[(T1)]
-    % \item\tlabel{rooted}\tlabel{first} $T$ is rooted at some node $x_0\in V(T)$ with $B_{x_0}=\emptyset$.
-    % \item\tlabel{node-set} $V(T)= V(H)$.
-    % \item\tlabel{diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
     \item\tlabel{subtree-root} for each node $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$; and consequently
@@
 
-% \referee{1}{6. page 6, last paragraph of Lemma 2: This isn’t quite right since you have to be careful with $x_0$.}
-% \note{PM}{I addressed this by removing $x_0$ from $T_0$ at the beginning of the second paragraph.}
-
-%\subsection{Generalized Tripod Partitions}
-%
-%
-% \begin{figure}[htbp]
-%   \begin{center}
-%     \includegraphics{figs/tripoddo}
-%   \end{center}
-%   \caption{The sets $Y_x$, $F_x$, and $V_x$ associated with $x\in V(T)$
-%   and the ancestors $a_1,\ldots,a_{t'}$ of $X$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.}
-%   \label{fig:generalized-tripod}
-% \end{figure}
-
-% The following lemma shows how to interpret an $H$-partition of $G$ and a tree-decomposition of $H$ as a `hierarchical' decomposition of $G$; refer to \cref{fig:generalized-tripod}.
-
-% For a vertex $v$ of $G$, $N_G(v):=\{w\in V(G):vw\in E(G)\}$ and $N_G[v]:=N_G(v)\cup\{v\}$.  For a vertex set $S\subseteq V$, $N_G(S):=\bigcup_{v\in S} N_G(S)\setminus S$ and $N_G[S]:=N_G(S)\cup S$.
-
-% \note{PM}{
-% Guide to old notation:
-% \begin{compactitem}
-%   \item $V_x := \bigcup_{y\in V(T_x)} Y_y = G_x$
-%   \item $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}= N_G(G_x)$
-%   \item $N_x:=V_x\cup F_x=N_G[G_x]$
-% \end{compactitem}
-% }
-%
-% \begin{lem}\label{generalized-tripod}
-%   Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
-%   For each $x\in V(T)$, let $H_x:=V(T_x)$ and let $G_x:=\bigcup_{y\in H_x}Y_y$.
-%
-%   \begin{compactenum}[(Y1)]
-%     % \item\ylabel{thickness} $\mathcal{Y}=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
-%     % \item\ylabel{separator} For each $x\in V(T)$, $N_G(\phi(T_x))$ separates $\phi(T_x)$ from $N
-%     % there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$.
-%
-%     \item\ylabel{ancestor-edge} For each $x\in V(T)$, there is a set $\{a_1,\ldots,a_{t'}\}$ of $t'\le t$ strict $T$-ancestors of $x$ such that $N_G(G_x) \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
-%   \end{compactenum}
-% \end{lem}
-%
-% \begin{proof}
-%   Let $\{a_1,\ldots,a_{t'}\}:= B_x\setminus\{x\}$.
-Then, by the definition of a normalized tree decomposition $N_H(H_x))\subseteq\{a_1,\ldots,a_{t'}\}$.
-Therefore $N_G(G_x)\subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
-% \end{proof}
-
-% \referee{1}{1.
-Lemma 3: This lemma is really about ``normalised'' tree-decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$.
-Call this property (T3) and remove (Y1) through (Y5), which should now be obvious.}
-%
-% \note{PM}{I've added to the \texttt{response.tex} file explaining some of this.
-Each of (Y1), (Y3), (Y4), and (Y5) have a one-sentence justification.
-The justification for (Y2) is just definition unpacking and then using the observation that the referee recommends.}
-%
-% \note{PM}{I plan to do an experiment with the referee's suggested notation, in a separate git branch to see how it turns out.}
-%
-% Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:
-%
-% \begin{compactenum}[(Y1)]\setcounter{enumi}{2}
-%   \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$. \note{PM}{$Y_x$, for $x\in V(T)$ is not defined!}
-%   \item\ylabel{containment-i} $V_x\subseteq V_a$ for every $T$-ancestor $a$ of $x$.
-%   \item\ylabel{containment-ii}$N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
-% \end{compactenum}
-
-% Property~\yref{y-subsets} follows from the fact that $V_x$ is the union of several sets, one of which is $Y_x$.
-Property~\yref{containment-i} follows from the definition of $V_x$ and the fact that $V(T_x)\subseteq V(T_a)$.
-To show Property~\yref{containment-ii} first note that, by \yref{containment-i} it suffices to consider vertices $w\in F_x=N_x\setminus V_x$.
-By definition, every vertex $w\in F_x$ is adjacent, in $G$, to a vertex $v\in V_x$.
-By \yref{containment-i}, $v\in V_a$, so $w$ is either in $V_a$ or $w$ satisfies the condition $vw\in E(G)$, $v\in V_a$, and $w\not\in V_a$, so $w\in F_a$.
-In either case $w\in N_a=V_a\cup F_a$.
-Note that none of \yref{y-subsets}--\yref{containment-ii} depends on \yref{ancestor-edge} (which is important, since \yref{containment-i} is used to establish \yref{ancestor-edge} in the following proof).
-%
-%
-% \begin{proof}[Proof of \cref{generalized-tripod}]
-%   % Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
-%   Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
-%
-%   To establish Property~\yref{ancestor-edge}, consider some vertex $w\in F_x$.
-Since $w\in F_x$, there exists an edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.
-Since $v\in V_x$, $v\in Y_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$).
-Since $\mathcal{Y}$ is a partition, $w\in Y_{a}$ for some $a\not\in V(T_x)$.
-Since $vw\in E(G)$, we have $x'a\in E(H)$.
-By \tref{ancestor-edge}, one of $a$ or $x'$ is a $T$-ancestor of the other.
-Since $w\in Y_a\subseteq V_a$ and $w\not\in V_x\supseteq V_{x'}$, \yref{containment-i} rules out the possibility that $x'$ is a $T$-ancestor of $a$.
-Therefore, $a$ is a $T$-ancestor of $x$ which is a $T$-ancestor of $x'$.
-Let $z_0,\ldots,z_r$ be the path in $T$ from $z_0:=x'$ to $z_r:=a$.
-Since $x'a\in E(H)$, at least one of $a$ or $x'$ is in $B_{z_i}$, for each $i\in\{0,\ldots,r\}$ and there is at least one $i$ such that $B_{z_i}$ contains both $x'$ and $a$.
-However, by \tref{subtree-root} $x'$ is not contained in $B_{z_i}$ for any $i\in\{1,\ldots,r\}$.
-Therefore $a\in B_{z_i}$ for each $i\in\{0,\ldots,r\}$.
-In particular, $a$ is contained in $B_x$.
-%   Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $B_x$ contains $x$.
-% \end{proof}
-
-% \referee{2}{Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
-% (twice).  Moreover, the way the proof is written, it seems to suggest
-% that we only get that $a \in B_{z_i}$ for each $i \in \{1, \dots, r\}$.
-% The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
-% true though.  In a normalized tree-decomposition, if $xy$ is an edge and
-% $x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
-% and $T_y$ intersect).}
-
-%\note{PM}{I addressed the preceding referee's.}
-
-% We are now ready to complete the proof.
-% , which we restate here for convenience:
-
-% \mmg*
-
 \begin{proof}[Proof of \cref{ShortcutPartition}]
@@
 
-% \note{DW}{These definitions should appear earlier, certainly before the definition of framed graph, which should be defined as an ``embedded graph''.
-One option is to put this whole paragraph as a footnote in the introduction where we first mention $k$-planar graphs.
-The experienced reader can skip this paragraph, which says to me it should be a footnote.}
-
 We first formally define $k$-planar graphs.
@@
 If an embedded graph $G$  has no crossings then it is \defin{non-crossing} graph.
-When $G$ is non-crossing and $F$ is a face of $G$ then $V(F)$ denotes the set of vertices in $G$ on the boundary of $F$.
+When $G$ is non-crossing and $F$ is a face of $G$, then let \defin{$V(F)$} be the set of vertices in $G$ on the boundary of $F$.
 When $F$ is homeomorphic to a disk, the \defin{facial walk} of $F$ is the cycle consisting of the edges and vertices of $G$ on the boundary of $F$.
@@
 
-% (This assumption is already present in the informal definition of $k$-planar graphs used in \cref{k_planar_section}.)
-
-%
-% \note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}\note{PM}{No, under the current definition, $r$ edges crossing at a single point $p$ generate $\binom{r}{2}$ crossings, with each edge taking part in $r-1$ crossings, just like they would if we perturbed the edges so that each pair crosses at a distinct point.
-I guess we could mention this.
-(See the first paragraph in the proof of Theorem~2 below.)} \note{DW}{The downside of this approach is that in the opening section we give an informal definition of $k$-planar, which needs the assumption of having no three edges cross at a point, since we use this assumption in the construction of a $(k+1,2)$-shortcut system for $k$-planar graphs give in the Introduction.
-I think we should not have different definitions in two parts of the paper.} A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$.
-Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.
-
 As mentioned in \cref{Introduction}, \cref{PlanarProduct,ShortcutProduct} imply a product structure theorem for $k$-planar graphs. We get improved bounds as follows.
 
-%\begin{thm}
-%\label{k-planar}
-%Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 48k+30$ in which $\tw(H)\leq \binom{k+4}{3}-1$.
-%\end{thm}
-
 \begin{proof}[Proof of \cref{kPlanarProduct}]
-	Let $G$ be a $k$-plane graph.
-    % We assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$.
-This assumption is justified since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
-	As in the proof of \cref{AddDummy}, let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$.
+	Let $G$ be a $k$-plane graph.	As in the proof of \cref{AddDummy}, let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$.
 In this way, each edge $vw\in E(G)$ corresponds naturally to a path $P_{vw}$ of length at most $k+1$ in $G_0$.
@@
 	The result follows from \cref{PartitionProduct}.
-	%This establishes \cref{k-planar}.
-	% Note to self-and-others. In our original SODA submission we had 18k^2+30k here because we forgot that the layer compression step has compress k+1 consecutive layers, not just k consecutive layers.
-\end{proof}
-
-% \subsection{Examples}
+\end{proof}
 
@@
 
-%\referee{2}{Page 16.
-At the end of Page 16, I do not see why 'every pair of edges cross at most once.'  Without further assumptions, it seems as if two paths in the path system could intersect several times.
-Therefore, by drawing 'each edge vw of G alongside $P_{vw}$ in $G_0$', there may be two edges which cross several times.
-It seems this can be fixed by choosing the path system carefully by 'uncrossing' paths which intersect at more than two internal vertices.}
-%
-%\note{DW}{I have revised the proof to address this error. I couldn't think of a way to do this without introducing balls and channels (and without vigorous hand-waving). Is it readable?}
-
 \section{Graphs Derived from Shortcut Systems}
@@
 
-% A similar proof to that of \cref{StringShortcut} shows that every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.
-
 \cref{StringShortcut,ShortcutProduct,GenusProduct} now imply:
- % and \cref{StringShortcut,qn} imply:
 
@@
 \end{thm}
-
-% \begin{thm}
-% \label{StringPartition}
-% For integers $g\geq 0$ and $\delta\geq 2$, let $\ell:= \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$ and $t:= \binom{ \delta+4}{3}-1$.
-% Then  every $(g,\delta)$-string graph:
-% \begin{compactitem}
-% 	\item is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$,
-% 	\item has queue-number $\qn(G^k) \leq 3 \ell \, 2^t - \ceil{\tfrac{3}{2}\ell}$.
-% \end{compactitem}
-% \end{thm}
-
-% \note{DW}{\cref{StringPartition} has been updated.}
-
-% Our results also give bounds on the non-repetitive chromatic number and the $p$-centered chromatic number of $(g,\delta)$-string graphs, but the bounds are weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ and $\chi_p(G)\le p(64\delta)^2$ by results of \citet{DJKW16} and \citet{DFMS21}, respectively.
 
@@
 
-% Theorems~\ref{PlanarProduct}(b), \ref{GenusProduct}(b) and \cref{ShortcutProduct} and  \cref{DrawG,qn,p-centered,non-repetitive,PowerShortcut} imply the following result, which with $g=0$ implies  \cref{kPowerBasic} in the introduction.
-%
-% \begin{thm}
-% \label{PowerGenus}
-% For integers $g\geq 0$ and $k,\Delta\geq 1$, let $\ell:= \max\{2g,3\} (2k^4+6k^2) \Delta^{k}$ and $t:= \binom{k+3}{3}-1$. Then for every graph $G$ of Euler genus $g$ and maximum degree $\Delta$,
-% \begin{compactitem}
-% \item $G^k$ is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$
-% \item $G^k$ is $(g, 2k(k+1)\Delta^{k} )$-planar,
-% \item $G^k$ has queue-number $\qn(G^k) \leq 3 \ell \cdot 2^t - \ceil{\tfrac{3}{2}\ell}$.
-% \item $G^k$ has  $p$-centered chromatic number $\chi_p(G^k) \leq \ell (p+1)\,  \binom{p+t}{t}$,
-% \item $G^k$ has  non-repetitive chromatic number $ \pi(G^k) \leq \ell \, 4^{t+1}$.
-% \end{compactitem}
-% \end{thm}
-%
-% % \note{DW}{\cref{PowerGenus} has been updated}
-%
-% This result is the first constant upper bound on the queue-number of bounded powers of graphs with bounded degree and bounded Euler genus.
-For every graph $G$, since $G^k$ has maximum degree at most $\Delta^k$, a result of \citet{DJKW16} implies that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$.
-\cref{PowerGenus} improves upon this bound when $k,g\ll\Delta$.
-Similarly, a result of \citet{DFMS21} implies that $\chi_p(G^k)\le 1024p\Delta^{2k}$ and \cref{PowerGenus} improves upon this bound when $p,k,g\ll\Delta$.
-%
-% \cref{qn,p-centered,non-repetitive,MinorFreeDegree,PowerProduct} imply the following analogous result for powers of graphs in any minor-closed class with bounded maximum degree.
-%
-% \begin{thm}
-% \label{PowerMinor}
-% For every graph $X$ there exists an integer $c$ such that for all integers $k,\Delta\geq 1$, if $t:= 2k\Delta^{k}(k^3+3k)\binom{k+c\Delta}{c\Delta}-1$ and $G$ is an $X$-minor-free graph with maximum degree $\Delta$, then:
-% \begin{compactitem}
-% 	\item $G^k$ is contained in $H\boxtimes P$ for some graph $H$ with treewidth $t$ and for some path $P$.
-% 	\item $G^k$ has queue-number at most $3\cdot 2^t-2$,
-% 	\item $G^k$ has $p$-centered chromatic number at most $(p+1)\binom{p+t}{t}$.
-% \end{compactitem}
-% \end{thm}
-
 \subsection{Map Graphs}
-
-% \note{PM}{Check David's commit: a0e20aff83d12c74127331854ce9fb75e002efe3}
 
@@
 The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.
-
-% \note{DW}{give a better version of \cref{MapPartition} with $g=0$. }
-
-%So $(g,d)$-map graphs generalise graphs embedded in a surface, and we now assume that $d\geq 4$ for the remainder of this section.
 
@@
 
-% \note{DW}{Suggested section name ``Graphs Derived From Framed Graphs'' or just``Framed Graphs''}
-
-% Next we consider
-%
-% %%%%%%%%%%%%%%%%%%%
-% \subsection{Framed Graphs}
-
-% \note{PM}{I don't like this structure anymore.
-The proof of \cref{kPlanarProduct} ($k$-planar) refers to the innards of the proof of \cref{ShortcutPartition} ($(k,d)$-shortcuts).
-The proofs should be close to each other.
-%
-% Let's consider something like this:
-% \begin{compactenum}
-%   \item Introduction
-%   \item Shortcut systems and $k$-planar graphs
-%   \item All graphs obtained from shortcut systems or $k$-planar graphs
-%   \item $d$-framed graphs, $1$-planar graphs, and map graphs
-%   \item All applications (colouring, labelling, etc).
-% \end{compactenum}
-% % (Possibly with the order of 2 and 3 reversed.)
-% }
-
-% \note{DW}{I suggest we use the language of ``embedded graph'' here}
 For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
@@
 Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that for each $v\in V(G)$,  \[\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.\]
-
-% \note{DW}{I suggest $G'$ be changed to $G_0$ throughout (since $G'$ feels like a graph derived from $G$, which it is not). Similarly, replace $N'$ by $N_0$.}
-%
-% \note{PM}{I would also like to change $G'$ to $G_0$ and $N'$ to $N_0$.
-One issue is holding me back, though: We recurse on cycles $F_1,\ldots,F_m$ that define subgraphs $N_1,\ldots,N_m$ of $G$ and subgraphs $N_1',\ldots,N_m'$ of $G'$.
-Will these latter subgraphs becomes $N_{0,1},\ldots,N_{0,m}$?}
-%
-% \note{DW}{Some suggestions: Replace $G'$ by $F$ (the frame), replace $F_0$ by $C$ (the outer-cycle), replace $F_1,\dots,F_m$ by $C_1,\dots,C_m$.
-Use $F_1,\dots,F_m$ for the subgraph of $F$ inside $C_i$.
-Use $G_i$ for the subgraph of $G$ inside $C_i$.}
 
@@
 	\end{compactenum}
-%\note{PM}{I've changed my mind about this notation change.
-We end up applying the lemma inductively on the faces $F_1,\ldots,F_m$ of $M$ (which are cycles in $G$).
-It makes sense that the original cycle in $G$ is called $F$.}
-%\note{DW}{okay}
-%\begin{lem}
-%  \label{induction} The setup:
-%  \begin{compactenum}
-%    \item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
-%    \item Let $F_0$ be the outer face of $G'$ and let $T$ be a BFS spanning forest of $G'$ rooted at $V(F_0)$.
-%    \item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
-%    \item Let $F$ be a cycle in $G'$ whose vertices are partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
-%    \begin{compactenum}
-%      \item $F[P_i]$ is connected; and
-%      \item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
-%    \end{compactenum}
-%    \item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
-%  \end{compactenum}
-%  Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
-%  \begin{compactenum}[(i)]
-%    \item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
-%    \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
-%    \item $H$ has a $3$-simple tree-decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
-%  \end{compactenum}
 \end{lem}
@@
 The induction base case occurs when $N_0$ has no inner vertices (that is, $V(N_0)=V(F)$).
-In this case we can simply take
-	% \note{DW}{Suggested revision, ``In the base case, $N_0$ has only one inner face, and we can simply take ...}
-  % \note{PM}{I debated what to put here since, even when $N'$ has only one face, the rules below work for picking $\tau$.
-The \emph{real} base case occurs when $N'$ has zero faces, but that's not something that ever happens since we only ever apply induction on the inner faces of $M$ (except $\tau$).
-I settled on this ``base case'' since it's clear that, in this case, the partition $\mathcal{P}$ is already decided.}
-  % \note{PM}{The longwinded version is that induction is on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$.
-The base cases are $(0,y)$ for any $y$.
-Unfortunately, we can't just induct on $x$ because it doesn't always decrease.} \note{DW}{We need to make it clear what we are doing induction on.
-% I suggest we write, ``we proceed by induction on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$. ''}
-	$\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iii):
+In this case, simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iii):
 	\begin{compactenum}[(i)]
@@
 	\end{compactenum}
-
-% \note{DW}{Now assume that $N'$ has at least two internal faces.}
 
@@
 	\end{compactenum}
-
-	%  \todo[inline]{%
-	%    DW:Does this work for multigraph near-triangulations? \newline
-	%    PM: Yes, it only requires that the inner faces have three sides and that the outer face is bounded by a simple cycle.
-The proof goes through without change.
-Look at the subgraph $X$ of the dual that contains only the edges dual to red/blue edges.
-In our setting, the outer face has degree $1$ in $X$, which is odd.
-Any bichromatic inner face has degree $0$ or degree $2$ in $X$.
-There has to be another vertex of odd degree somewhere, which can only come from a trichromatic inner face.\newline
-	%    DW: great. This is basically the same as the proof of the Hex Lemma.
-	%  }
 
@@
 
-	% We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.
-First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $L_i$, each incident on at most two edges of $\overline{Y}$.
-Since $\dist_{G'}(v,w)\le 2$ for each $vw\in E(G)$, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $L_i$.
-These at most six edges have at most 12 endpoints.
-Therefore $|V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i|\le 6\times 2=12$, so $|V(\overline{Y}^+)\cap L_i|\le 12+3=15$.
-
-Let $M$ denote the subgraph of $N_0$ containing the edges and vertices of $\overline{S}$ and the edges and vertices of $F$. By our choice of $\tau$, the graph $M$ is $2$-connected.
-  % Since $N'\subseteq G'$ contains only edges of $N\subseteq G$ that are not crossed, each component of $N-V(M)$ is contained in a single face of $M$.
+Let $M$ denote the subgraph of $N_0$ containing the edges and vertices of $\overline{S}$ and the edges and vertices of $F$.
+By our choice of $\tau$, the graph $M$ is $2$-connected.
 Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
@@
 
-	In the following, in order to avoid introducing even more notation or constantly adding the modifier ``when it exists'', we will assume that $V(F_i)\cap S\neq\emptyset$ and that there are exactly two distinct $b,c\in\{1,\ldots,k\}$ such that $V(F_i)\cap P_b\neq\emptyset$ and $V(F_i)\cap P_c\neq\emptyset$.
+	In the following, in order to avoid introducing even more notation or constantly adding the modifier ``when it exists'', we assume that $V(F_i)\cap S\neq\emptyset$ and that there are exactly two distinct $b,c\in\{1,\ldots,k\}$ such that $V(F_i)\cap P_b\neq\emptyset$ and $V(F_i)\cap P_c\neq\emptyset$.
 When this assumption does not hold, exactly the same reasoning can be used, while omitting one or more of $P_{i,1}$, $P_{i,2}$ or $P_{i,3}$.
@@
 
-% \todo[inline]{Alternative to the final three paragraphs in the preceding proof:
-%
-% To see that $H$ is planar, consider a plane multigraph $G''$ obtained by adding stars in some of the faces of $G'$, as follows:  When creating the part $S$, add a vertex $v_\tau$ to the interior of $\tau$ that is adjacent to each vertex of $\tau$ and place $v_\tau$ in $S$.
-The graph $G''\supseteq G'$ obtained this way is clearly planar.
-Furthermore, for each part $P\in\mathcal{P}$, $G''[P]$ is connected.
-Since planarity is preserved under edge contraction, it follows that $G''/\mathcal{P}$ is planar.
-Now argue that $G/\mathcal{P}=G''/\mathcal{P}$\ldots
-% }
-%
-% \note{DW}{The proof via 3-simplicity is shorter, so just keep it as is.}
-
 Using \cref{induction}, the proof of \cref{d_framed_product_stucture} is now straightforward.
-
-% \note{PM}{Annoying detail:  When we allow framed graphs to have crossings in the outer face, we can't guarantee that $G'$ has any face smaller than $d$.
-This makes it awkward to start the induction using \cref{induction}.
-Possible solutions:
-% \begin{compactenum}
-%   \item Disallow the addition of edges in the outer face of a frame.
-%   \item Modify \cref{induction} so that $T$ is a BFS forest rooted at the set of vertices on the outer face of $G'$. Then we can start the induction with $k=1$.
-% \end{compactenum}
-% I prefer the second option.  Thoughts?
-% }
 
@@
 Let $G$ be a $d$-framed multigraph with frame $G_0$ having outer face $F_0$
-	% \note{DW}{Replace ``outer face'' by ``outer-cycle'' since a face is a defined to be a portion of the plane. }.
-  %\note{PM}{I added some conventions around the definition of $\Sigma$-embedded graph so that $V(F)$ is defined for a face $F$ in a non-crossing embedded graph.}
 Let $T$ be a BFS forest of $G_0$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.
@@
 
-% $1$-planar graphs are of particular interest to the graph drawing community~\citep{kobourov.liotta.ea:annotated}.
-In this case, we obtain better constants and an additional property (planarity) of $H$.
-
-%(\htmladdnormallink{https://arxiv.org/abs/1703.02261}{arXiv:1703.02261}) has 143 entries, and even this is now four years out of date.
-%
-%   An important aspect of our result for $1$-planar graphs is that graph $H$ in the product is planar and has treewidth $3$, i.e., it has simple treewidth $3$.
-In a more recent application of product structure \citet{BDJM} show that such graphs have $\ell$-vertex rankings using $O(\log n/\log\log\log n)$ colours, and this is tight.
-This result requires that $H$ have simple treewidth $3$, even treewidth $3$ is not sufficient.
-% }
-
-%
-% \note{PM}{Use restateable here...}
-%
-% \begin{thm}
-% \label{1-planar}
-% Every 1-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
-% \end{thm}
-
-%In \cref{1_planar_is_4_framed}, below, we show that every $1$-planar graph is a subgraph of a $4$-framed multigraph.
-This makes \cref{1-planar} an immediate consequence of the following theorem, whose proof is the subject of the rest of this section.
-
 \cref{1_planar_product} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma.
@@
 \end{lem}
-
-% \note{DW}{Write  ``A multigraph $G$ is $1$-planar if and only if ...''}
-
-%Before proving \cref{1_planar_is_4_framed}, we remark that the lemma seems not to be true if we replace `multigraph' with `graph'.
-We are only aware of two previous results on the relationship between $1$-planar graphs and $d$-framed graphs \cite{BDGGMR}:
-%\begin{inparaenum}[(i)]
-%  \item Every $3$-connected $1$-planar graph is contained in some $4$-framed graph.
-%  \item Every $1$-planar graph is contained in some $8$-framed graph.
-%\end{inparaenum}
-%Neither of these results allows a proof of \cref{1-planar} without sacrificing generality or accepting a larger constant.
 
@@
 We may assume that no two edges incident to a common vertex of $G$ cross, since such a crossing can be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}.
-  % We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
 We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.
@@
 
-% Therefore each face of $G_0$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
-% \begin{inparaenum}[(i)]
-%   \item bounded by three edges of $G$, none of which are crossed; or
-%   \item bounded by one edge of $G$ that is not crossed plus portions of two edges of $G$ that cross each other.
-% \end{inparaenum}
-
 Consider some pair of edges $vw$ and $xy$ that cross at point $p$, which is a degree-4 vertex of $G'$.
@@
 Therefore $G$ is a $4$-framed graph with frame $G_0$.
-% Since, for any pair of crossing edges $\{vw,xy\}$ that we remove this way, $vxwy$ is a cycle in $G'$, $\dist_{G'}(v,w)\le 2$ and $\dist_{G'}(x,y)\le 2$.
-% It follows immediately that, for any pair of vertices $a,b\in V(G)$, $\dist_{G'}(a,b)\le 2\dist_G(a,b)$.
 \end{proof}
@@
 \subsection{Plane Map Graphs}
-
-% \note{PM}{Check David's commit: a0e20aff83d12c74127331854ce9fb75e002efe3}
-
-% Map graphs are defined as follows.
-Start with a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations.
-Let $G$ be the graph whose vertices are the nations of $G_0$, where two vertices are adjacent in $G$ if the corresponding faces in $G_0$ share a vertex.
-Then $G$ is called a \defin{$(g,d)$-map graph}.
-A $(0,d)$-map graph is called a (plane) \defin{$d$-map graph}; see \citep{FLS-SODA12,CGP02} for example.
-The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.
-%
-% \note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph.}
 
@@
 
-%\note{PM}{Got rid of genus-$g$ in the following. We use shortcuts for $(g,d)$-map graphs and now we don't have to define $(g,d)$-frame.}
-
 \begin{lem}
 \label{MewMapGraphLemma}
-For every %all
-integer %s $g\geq 0$ and
- $d\geq 3$, every $d$-map graph is a spanning subgraph of a $d$-framed multigraph.
+For every integer $d\geq 3$, every $d$-map graph is a spanning subgraph of a $d$-framed multigraph.
 \end{lem}
@@
 Note that if $r=1$ then the ``nation cycle'' has no edges, and if $r=2$ then the ``nation cycle'' has one edge.
-\note{DW}{Perhaps we shouldn't call it the ``nation cycle'' since it is not necessarily a cycle.}
 
 Let $G_1$ be a plane supergraph of $G_0^*$ obtained by adding the nation cycle of each face $F$ of $G_0$, and then triangulating each face with more than $d$ vertices on its boundary.
-Each nation cycle \note{DW}{add `` of length at least 3''} is now a face of $G_1$, and no face of $G_1$ has more than $d$ vertices on its boundary.
+Each nation cycle of length at least $3$ is now a facial cycle of $G_1$, and no face of $G_1$ has more than $d$ vertices on its boundary.
 Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.
@@
 Therefore, the nation cycle $C_x$ of $F_x$ contains $v$ and $w$.
-Since $C_x$ bounds a face in $G_1$, $vw\in E(\widehat{G}_1)$.
-\end{proof}
-
-%\note{PM}{Old proof}
-%
-%\begin{proof}
-%% \note{PM}{I fixed some issues here that come from the fact that some vertices of $G_0$ do not contribute faces to $G_1$.
-This happens when two nations and two lakes meet at a vertex $x$.
-It's not an issue but it means we can't write ``the face of $G_1$ that corresponds to $x$.''} \note{DW}{Good}
-%Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations.
-If two lakes have a common edge in their boundaries, then we may delete the edges and merge the lakes \note{DW}{Delete the previous sentence?
-This sentence is incompatible with ``in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes'' below.}.
-Now assume that no two lakes have a common edge in their boundaries.
-%Let $G$ be the corresponding map graph.
-%Let $G_1$ be the graph with $V(G_1):=V(G)$, which is the set of nations, and whose edge set we now define.  Consider vertices $v$ and $w$ of $G_1$.
-%Let $A$ and $B$ be the nation faces of $G_0$ corresponding to $v$ and $w$.
-%Then $vw\in E(G_1)$ if $A$ and $B$ have a common edge in their boundaries,
-%or there is a vertex $x$ of $G_0$ in the boundary of $A$ and $B$, and in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes.
-%% Then $G_1$ has Euler genus at most $g$.
-%Each face of $G_1$ corresponds to a vertex of $G_0$ or a lake of $G_0$.
-%Triangulate each face of $G_1$ corresponding to a lake of $G_0$.
-%
-%For each face $F$ of $G_1$ that corresponds to a vertex $x$ of $G_0$, the size of $F$ \note{DW}{Is ``size'' defined?
-Do we mean length of the facial cycle?
-} equals the number of nations incident to $x$, which is at most $d$.
-So every face of $G_1$ has size at most $d$.
-Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.
-\note{DW}{I like this notation.
-Should we use it throughout?} \note{PM}{No.
-For most things I prefer $G$ and $G_0$ rather than $\widehat{G}_0$ and $G_0$.}
-%
-%To complete the proof, we now show that $\widehat{G}_1$ contains every edge of $G$.
-%If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common.
-If some face of $G_1$ corresponds to $x$, then $v$ and $w$ are on a common face of $G_1$ so $vw\in E(\widehat{G}_1)$.
-If no face of $G_1$ corresponds to $x$\footnote{This can occur when $x$ is on the boundary of exactly two nation faces and two lake faces.}, then the only two nation faces of $G_0$ with $x$ on their boundary are those corresponding to $v$ and $w$, in which case $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.
-Therefore $G\subseteq\widehat{G}_1$.
-%\end{proof}
+If $C_x$ has length $2$ then $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.
+If $C_x$ has length at least $3$ then $C_x$ bounds a face in $G_1$, so $vw\in E(\widehat{G}_1)$.
+\end{proof}
 
@@
 In this section, we survey some of the consequences of this for the graph classes considered in the previous two sections.
-
-% Here we discuss some of the consequences of the above theorems for $k$-planar and $(g,k)$-planar graphs.
-%
-% \referee{1}{4.
-Sections 4 and 5: I think as they are currently organized they distract from the purpose of the paper, as stated in the introduction, to ``prove product structure theorems for several non-minor-closed classes of interest.'' Which of Corollaries 1--4 and Theorems 12--17 are most important?
-I would remove any mention of the applications from the examples section, put that section first, and state in that section a single main theorem with all of the best bounds on the product structure.
-The applications section is mostly a survey and I think should be focused on the most exciting new corollaries.
-Theorems 18 and 19 should be in the applications section.}
-%
-% \note{DW}{I don't want to do a complete restructuring. In our response, let's just say why we structure it the way we do.}
-% \note{PM}{Sure, though I think they proposing a chop rather than a restructuring.}
 
@@
 Thus the graph $H$ in \cref{1_planar_product,dMapProduct} has queue-number at most $5$.
-% , which with \cref{qn} implies:
-%
-% \begin{cor}
-% \label{1PlanarQueue}
-% Every 1-planar graph has queue-number at most
-% $3 \times 7 \times 5 + \floor{\tfrac{3}{2} \times 7} = 115$.
-% \end{cor}
-%
-% \begin{cor}
-% \label{dMapQueue}
-% Every $d$-map graph has queue-number at most
-% $\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$.
-% \note{DW}{I have updated this bound using the better qn bound for planar 3-trees. Note that with $d=3$ we get the original bound of 49.}
-% \end{cor}
-%
-%
-%
-% \cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$.
-% \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
-% \note{OM}{Ok}
-% \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
-% \note{PM}{Ok}
+
 The following corollary summarizes all of these results.
@@
 \end{cor}
-%\note{PM}{For uniformity, I got rid of the two corollaries on $d$-map and $1$-planar graphs and kept them in \cref{q_cor}.
-The text is still commented out in the source in case someone feels strongly about keeping them.}
 
@@
 \end{lem}
-
-% \cref{non-repetitive,kPlanarProduct,1-planar,gkPlanarProduct} imply the following result:
-%
-% \begin{cor}\quad
-% \begin{compactitem}
-% \item	For every $1$-planar graph $G$, $\pi(G)\le 7\times 4^4=1792$.
-% %\item For every $k$-planar graph $G$,
-% %	$\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
-% \item	For every $(g,k)$-planar graph $G$,
-% 	$  \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}.$
-% % \note{DW}{I have changed $4^{\binom{k+5}{4}}$ to $4^{\binom{k+4}{3}}$, okay?}
-% % \note{PM}{Ok}
-% \end{compactitem}
-% \end{cor}
 
@@
 
-%In the journal version, \citet{DFMS21} proved that $\chi_p(H_1\boxtimes H_2) \leq \chi_p(H_1) \cdot \chi(H_2^p)$ where $H_2^p$ is the $p$-th power of $H_2$.
-Apply this with $H_2:= P \boxtimes K_\ell$.
-Then $H_2^p= P^p\boxtimes K_\ell$ and $\chi(H_2^p)=(p+1)\ell$.
-So $\chi_p(H\boxtimes P \boxtimes K_\ell) \leq \chi_p(H_1) \cdot (p+1)\ell$.
-}
-
 \begin{lem}[\citep{DFMS21,PS21}]
@@
 \[\chi_p(G)\le \ell (p+1)\, \chi_p(H) \leq \ell (p+1) \tbinom{p+t}{t}.\]
-%If $G\subseteq H\boxtimes P \boxtimes K_\ell$ then $\chi_p(G)\le \ell (p+1)\, \chi_p(H)$.
-\end{lem}
-
-%\note{DW}{For consistency with \cref{qn}, I suggest we combine \cref{p-centered,p-centered-treewidth}, and just write, ``}
-
-%\begin{proof}
-%	By \cref{PartitionProduct}, $G$ has an $H$-partition $(\mathcal{L}=\langle L_0,L_1,\ldots\rangle, \PP=(B_x:x\in V(H))$ with layered width at most $\ell$.
-Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.
-For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in L_i\cap B_x$ such that:
-%   \begin{compactenum}
-%     \item $\alpha(v)$ is unique among $\{\alpha(w): w\in L_i\cap B_x\}$, which is possible
-%     since $|L_i\cap B_x|\le \ell$,
-%     \item $\beta(v)= i\bmod (p+1)$, and
-%     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
-%   \end{compactenum}
-% To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
-%
-% First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$.
-Since $X$ is connected, $X$ contains a path from $v$ to $w$.
-By the definition of layering, this path contains at least one vertex from $L_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$.
-Therefore, $|\{\beta(v'):v'\in V(X)\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
-%
-% Otherwise, $V(X)\subseteq L_{i}\cup\cdots\cup L_{i+s}$ for some $s<p$.
-Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.
-If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in V(X)\}|> p$ so $|\{\phi(v):v\in V(X)\}|> p$ and we are done.
-Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
-% For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$.
-Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$.
-Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
-% \end{proof}
-%
-%\referee{2}{Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.}
-
-% \cref{p-centered,kPlanarProduct,1-planar,gkPlanarProduct} immediately imply the following results, for every $p\ge 2$:
-%
-% \begin{cor}\quad
-% \begin{compactitem}
-% \item For every $1$-planar graph $G$,\; $ \chi_p(G)\le 5 (p+3)(p+2)(p+1)^2$.
-% %\item For every $k$-planar graph $G$,\; $\displaystyle \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}$.
-% \item For every $(g,k)$-planar graph $G$,\;
-% $\displaystyle \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{p+\binom{k+4}{3}-1}{\binom{k+4}{3}-1}$.
-% % \note{DW}{I have changed $\binom{k+5}{4}$ to $\binom{k+4}{3}$, okay?}
-% % \note{PM}{Ok}
-%  \end{compactitem}
-% \end{cor}
+\end{lem}
 
@@
 Combining \cref{Universal} with our results on product structure yields the following:
+
 % this with \cref{gkPlanarProduct,k-nn,PowerMinor,StringPartition,MapPartition} yields the following:
@@
 
-% \section{Orphaned Text}
-%
-% \subsection{Text following \cref{gkPlanarProduct}}
-%
-% Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs (or any of the other classes presented in \cref{examples}) was in terms of layered treewidth, which we now define.
-A \defin{layered tree-decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{T}$ of $G$.
-The layered width of $(\mathcal{L},\mathcal{T})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{T}\}$.
-The \defin{layered treewidth} of $G$ is the minimum layered width of any layered tree-decomposition of $G$.
-\citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph.
-\citet{dujmovic.eppstein.ea:structure} show that every $k$-planar graph has layered treewidth at most $6(k+1)$, and more generally that every $(g,k)$-planar graph has layered treewidth at most $(4g+6)(k+1)$.
-It follows from this result that $(g,k)$-planar graphs have treewidth $O(\sqrt{(g+1)(k+1)n})$ and thus have balanced separators of the same order, which can also be concluded from the work of \citet{FP08}.
-In related work, \citet{grigoriev.bodlaender:algorithms} used structural results to obtain approximation algorithms for $(g,k)$-planar graphs, and \citet{PachToth97} determined the maximum number of edges in a $k$-planar graph (up to a constant factor).
-%
-% If a graph class admits bounded layered partitions, then it also has bounded layered treewidth.
-In particular, if $\PP=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree-decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree-decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{DJMMUW20}.
-Therefore, any property that holds for graphs of bounded layered treewidth also holds for $G$.
-What sets layered partitions apart from layered treewidth is that they lead to constant upper bounds on the queue-number  and non-repetitive chromatic number, whereas for both these parameters, the best known upper bound obtainable via layered treewidth is $O(\log n)$; see \cref{Applications}.
-%
-%
-%
-% %%%%%%%%%%%%%%%%%%%
-% \label{sec-k-planar}
-%
-%
-% %\note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{d_framed_product_stucture} with the method in \citep{DHHW}.
-I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
-Okay?}
-% %\note{PM}{Okay to this and to all similar questions that come later (which I've now commented out.)}
-% %\note{DW}{Robert is working on it.}
-%
-%
-%
-%
-%
-% This section describes several examples of graph classes that can be obtained from a shortcut system typically applied to graphs of bounded Euler genus.
-%
-% \subsection{Orphaned Theorems}
-%
-% Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,non-repetitive} imply the following result.
-%
-% %\begin{thm}
-% %\label{PlaneMapPartition}
-% %\note{DW}{For every integer $d\geq 4$, }
-% %Every $d$-map graph $G$:
-% %	\begin{compactitem}
-% %		\item is contained in $H \boxtimes P \boxtimes K_{21d(d-3)}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
-% %		\item has queue-number $\qn(G) < 32225\, d(d-3)$.
-% %		\item has $p$-centered chromatic number $\chi_p(G) \leq 21d(d-3) (p+1)  \binom{p+9}{9}$,
-% %		\item has non-repetitive chromatic number $ \pi(G) \leq 21 \cdot 4^{10} d(d-3)$.
-% %	\end{compactitem}
-% %\end{thm}
-%
-% \begin{thm}
-% \label{MapPartition}
-% For integers $g\geq 0$ and $d\geq 4$, if $\ell:=  7d(d-3)\, \max\{2g,3\}$ then every $(g,d)$-map graph $G$:
-% \begin{compactitem}
-% \item is contained in $H \boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
-% \item has queue-number $\qn(G) <  1535\, \ell $.
-% \item has $p$-centered chromatic number $\chi_p(G) \leq \ell (p+1)\,  \binom{p+9}{9}$,
-% \item has non-repetitive chromatic number $ \pi(G) \leq 4^{10}\,\ell $.
-% \end{compactitem}
-% \end{thm}
-%
-% These results give the first constant upper bound on the non-repetitive chromatic number of map graphs, the first polynomial bounds on the $p$-centered chromatic number of map graphs, and the best known bounds on the queue-number of map graphs.
-%
-% % \note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the method in \citep{DHHW}.
-I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}).
-Okay?}
-%
-%
-% % \subsection{Orphaned Section}
-% %
-% % \cref{nearest-neighbour,kPlanarProduct,qn,p-centered} imply:
-% %
-% % \begin{cor}
-% % \label{k-nn}
-% % For every integer $k\geq 1$ there exists integers $t\leq O(k^6)$ and $\ell\leq O(k^4)$ such that every $k$-nearest-neighbour graph:
-% % \begin{compactitem}
-% % \item is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $t$ and some path $P$,
-% % \item has queue-number at most $2^{O(k^6)}$, and
-% % \item has $p$-centered chromatic number at most $\ell (p+1)\binom{p+t}{t}$.
-% % \end{compactitem}
-% % \end{cor}
-% %
-% % \cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$.
-However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (36+o(1))k^2$ by a result of \citet{DJKW16}.
-%
-%
-%
-%
-% \subsection{Generalisations}
-% \label{Generalisations}
-%
-% % As mentioned above, product structure theorems have been established for several minor-closed classes in addition to planar graphs.
-The first generalises \cref{PlanarProduct} for graphs of bounded Euler genus.
-% %
-% % \begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
-% % \label{GenusProduct}
-% % Every graph of Euler genus $g$ is contained in:
-% % \begin{compactenum}[(a)]
-% % \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
-% % \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
-% % \end{compactenum}
-% % \end{thm}
-%
-% \citet{DJMMUW20} generalised \cref{GenusProduct} for apex-minor-free graphs as follows.
-%
-% \begin{thm}[\citep{DJMMUW20}]
-% \label{ApexMinorFree}
-% For every apex graph $X$, there exists $c\in\mathbb{N}$ such that every $X$-minor-free graph is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and some path $P$.
-% \end{thm}
-%
-% The assumption that $X$ is apex is needed in \cref{ApexMinorFree}, since if the class of $X$-minor-free graphs has a product structure theorem analogous to \cref{PlanarProduct}, then $X$ is apex \citep{DJMMUW20}.
-On the other hand, \citet{DEMWW22} proved a product structure theorem for bounded degree graphs in any minor-closed class.
-%
-% \begin{thm}[\citep{DEMWW22}]
-% 	\label{MinorFreeDegree}
-% 	For every graph $X$ there exists $c\in\mathbb{N}$ such that for every $\Delta\in\mathbb{N}$, every $X$-minor-free graph $G$ with maximum degree at most $\Delta$ is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H) \leq c\Delta$ and for some path $P$.
-% \end{thm}
-%
-% %The next three results are immediate corollaries of \cref{GenusProduct,ApexMinorFree,ShortcutProduct,MinorFreeDegree}.
-% %
-% %\begin{thm}
-% %Let $\SS$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$.
-Then $G^\SS$ is contained in $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
-% %\end{thm}
-% %
-% %\begin{thm}
-% %For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\SS$ for $G$, $G^\SS\subseteq H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
-% %\end{thm}
-% %
-% %\begin{thm}
-% %For every graph $X$ and for all integers $k,\Delta\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ with maximum degree $\Delta$ and for every $(k,\Delta)$-shortcut system $\SS$ for $G$, $G^\SS \subseteq H \boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
-% %\end{thm}
-% %
-% %\note{DW}{delete the above three theorems, they are never used}
-%
-%
-%
-%
-% %\subsection*{Note Added in Proof} Subsequent to the initial release of this paper, the treewidth 8 bound in \cref{PlanarProduct} was improved to 6 by \citet{UWY}, which implies that the treewidth $2g+8$ bound in \cref{GenusProduct}(a) can be improved to $2g+6$.
-Simmilarly, the treewidth 4 bound in \cref{GenusProduct}(b) was improved to 3 by \citet{DHHW}.
-\note{DW}{Say more about the consequence for the bounds in the rest of the paper}
-
 % %%%  Squashing the bibliography

### 2022-04-07 orphaned.tex
removed all the \notes, minor revisions

 \section{Orphaned Text}

 \item \citet{DFMS21} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs.
 \item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results starting with the work of \citet{kannan.naor.ea:implicit}).
 \item \citet{BDJM} use it to show that every $n$-vertex planar graph has local vertex ranking using $O(\log n/\log\log\log n)$ colours.
  \item The result of \citet{DEJGMM21} implies that, for every integer
 $n>0$, there is a `induced-universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
 subgraph of $U_n$. This has recently been extended by \citet{EJM}, who show the existence of a `subgraph-universal' graph with $(1+\epsilon)n$ vertices that contains every $n$-vertex planar graph as a subgraph as well as the existence of an `induced universal' graph with $n^{1+o(1)}$ vertices \emph{and} edges.  The former result is the first progress on subgraph-universal graphs for planar graphs since the $O(n^{3/2})$-vertex construction of \citet{babai.chung.ea:on}.
 aking the first progress on this problem since the work of \citet{babai.xx} in 1984.
 \end{compactitem}

 These theorems have applications in diverse areas, which we explore in \cref{Applications}.

 This theorem has applications in diverse areas, including queue layouts  \citep{DJMMUW20}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, $p$-centered colouring  \citep{DFMS21}, and adjacency labelling \citep{DEJGMM21}, which we explore in \cref{Applications}. For example, we prove that $k$-planar graphs have bounded non-repetitive chromatic number (for fixed $k$). Prior to the recent work of \citet{dujmovic.esperet.ea:planar}, it was even open whether planar graphs have bounded non-repetitive chromatic number.

 \referee{1}{3. pages 2 – 4: The paragraph following Theorem 2 and Section 1.3
 (except for the fact that the apex assumption is necessary) is repetition
 of things discussed earlier on in the introduction.}

 \note{PM}{Kind of, but I suggest we keep that paragraph and section anyway. Justification: They're both short and they describe the exact bounds for bounded genus graphs, which you don't get from the one-paragraph description in Section~1.0.}

 \note{DW}{I suggest we write, ``This theorem has applications in diverse areas, which we explore in \cref{Applications}....''}

 \note{DW}{I suggest we replace ``\cref{first_example,mid_example,last_example}''by ``\cref{examples}''.}

 Here is one example. The \defin{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.\footnote{For a graph $G$ and two vertices $v,w\in V(G)$, $\dist_G(v,w)$  denotes the length of a shortest path, in $G$, with endpoints $v$ and $w$.  We define $\dist_G(v,w):=\infty$ if $v$ and $w$ are in different connected components of $G$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\SS$ for some $(k,2k\Delta^{k})$-shortcut system $\SS$; see \cref{PowerShortcut}. Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} then imply:

 \begin{thm}
 \label{kPowerBasic}
 For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is contained in $H\boxtimes P\boxtimes K_{6k^2(k^2+3)\Delta^{k}}$ for some graph $H$ of treewidth at most $\binom{k+3}{3}-1$ and some path $P$.
 \end{thm}

 \cref{Examples} presents further examples of graph classes that can be constructed using shortcut systems, including certain types of map graphs, string graphs, and $k$-nearest neighbour graphs. \cref{ShortcutProduct} implies product structure theorems for each of these classes. All of the above-mentioned applications also hold for these examples.

 \citet{DJMMUW20} introduced the study of partitions with bounded layered width such that the quotient has some additional desirable property, like small treewidth. Dujmovi\'c~et~al.\ define a class $\mathcal{G}$ of graphs to \defin{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.

 \note{PM}{This feels out of place now that this section has been moved out of the introduction.  Delete?

 At the core of their work is the elegant proof by \citet{PS21} of the following result:

 \begin{thm}[\citep{PS21}]
 \label{ps}
   Every planar triangulation $G$ has an $H$-partition $\PP$ such that $\tw(H)\leq 8$ and $G[P]$ is a shortest path in $G$ for each $P\in\PP$.
 \end{thm}

 Indeed, the above-mentioned result of \citet{DJMMUW20} is a slight strengthening of \cref{ps}, where for each $P\in\PP$ no two vertices of $P$ have the same distance to some fixed root vertex $r$.
 }

 \referee{1}{6. page 6, last paragraph of Lemma 2: This isn’t quite right since you have to be careful with $x_0$.}
 \note{PM}{I addressed this by removing $x_0$ from $T_0$ at the beginning of the second paragraph.}

\subsection{Generalized Tripod Partitions}

 \begin{figure}[htbp]
   \begin{center}
     \includegraphics{figs/tripoddo}
   \end{center}
   \caption{The sets $Y_x$, $F_x$, and $V_x$ associated with $x\in V(T)$
   and the ancestors $a_1,\ldots,a_{t'}$ of $X$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.}
   \label{fig:generalized-tripod}
 \end{figure}

 The following lemma shows how to interpret an $H$-partition of $G$ and a tree-decomposition of $H$ as a `hierarchical' decomposition of $G$; refer to \cref{fig:generalized-tripod}.

 For a vertex $v$ of $G$, $N_G(v):=\{w\in V(G):vw\in E(G)\}$ and $N_G[v]:=N_G(v)\cup\{v\}$.  For a vertex set $S\subseteq V$, $N_G(S):=\bigcup_{v\in S} N_G(S)\setminus S$ and $N_G[S]:=N_G(S)\cup S$.

 \note{PM}{
 Guide to old notation:
 \begin{compactitem}
   \item $V_x := \bigcup_{y\in V(T_x)} Y_y = G_x$
   \item $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}= N_G(G_x)$
   \item $N_x:=V_x\cup F_x=N_G[G_x]$
 \end{compactitem}
 }

 \begin{lem}\label{generalized-tripod}
   Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
   For each $x\in V(T)$, let $H_x:=V(T_x)$ and let $G_x:=\bigcup_{y\in H_x}Y_y$.

   \begin{compactenum}[(Y1)]
      \item\ylabel{thickness} $\mathcal{Y}=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
      \item\ylabel{separator} For each $x\in V(T)$, $N_G(\phi(T_x))$ separates $\phi(T_x)$ from $N
      there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$.

     \item\ylabel{ancestor-edge} For each $x\in V(T)$, there is a set $\{a_1,\ldots,a_{t'}\}$ of $t'\le t$ strict $T$-ancestors of $x$ such that $N_G(G_x) \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
   \end{compactenum}
 \end{lem}

 \begin{proof}
   Let $\{a_1,\ldots,a_{t'}\}:= B_x\setminus\{x\}$.  Then, by the definition of a normalized tree decomposition $N_H(H_x))\subseteq\{a_1,\ldots,a_{t'}\}$.  Therefore $N_G(G_x)\subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
 \end{proof}

 \referee{1}{1. Lemma 3: This lemma is really about ``normalised'' tree-decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$. Call this property (T3) and remove (Y1) through (Y5), which should now be obvious.}

 \note{PM}{I've added to the \texttt{response.tex} file explaining some of this.  Each of (Y1), (Y3), (Y4), and (Y5) have a one-sentence justification. The justification for (Y2) is just definition unpacking and then using the observation that the referee recommends.}

 \note{PM}{I plan to do an experiment with the referee's suggested notation, in a separate git branch to see how it turns out.}

 Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:

 \begin{compactenum}[(Y1)]\setcounter{enumi}{2}
   \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$. \note{PM}{$Y_x$, for $x\in V(T)$ is not defined!}
   \item\ylabel{containment-i} $V_x\subseteq V_a$ for every $T$-ancestor $a$ of $x$.
   \item\ylabel{containment-ii}$N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
 \end{compactenum}

 Property~\yref{y-subsets} follows from the fact that $V_x$ is the union of several sets, one of which is $Y_x$.  Property~\yref{containment-i} follows from the definition of $V_x$ and the fact that $V(T_x)\subseteq V(T_a)$. To show Property~\yref{containment-ii} first note that, by \yref{containment-i} it suffices to consider vertices $w\in F_x=N_x\setminus V_x$. By definition, every vertex $w\in F_x$ is adjacent, in $G$, to a vertex $v\in V_x$.  By \yref{containment-i}, $v\in V_a$, so $w$ is either in $V_a$ or $w$ satisfies the condition $vw\in E(G)$, $v\in V_a$, and $w\not\in V_a$, so $w\in F_a$.  In either case $w\in N_a=V_a\cup F_a$.  Note that none of \yref{y-subsets}--\yref{containment-ii} depends on \yref{ancestor-edge} (which is important, since \yref{containment-i} is used to establish \yref{ancestor-edge} in the following proof).

 \begin{proof}[Proof of \cref{generalized-tripod}]
    Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
   Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.

   To establish Property~\yref{ancestor-edge}, consider some vertex $w\in F_x$.  Since $w\in F_x$, there exists an edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.  Since $v\in V_x$, $v\in Y_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$). Since $\mathcal{Y}$ is a partition, $w\in Y_{a}$ for some $a\not\in V(T_x)$.  Since $vw\in E(G)$, we have $x'a\in E(H)$.  By \tref{ancestor-edge}, one of $a$ or $x'$ is a $T$-ancestor of the other. Since $w\in Y_a\subseteq V_a$ and $w\not\in V_x\supseteq V_{x'}$, \yref{containment-i} rules out the possibility that $x'$ is a $T$-ancestor of $a$. Therefore, $a$ is a $T$-ancestor of $x$ which is a $T$-ancestor of $x'$.  Let $z_0,\ldots,z_r$ be the path in $T$ from $z_0:=x'$ to $z_r:=a$.  Since $x'a\in E(H)$, at least one of $a$ or $x'$ is in $B_{z_i}$, for each $i\in\{0,\ldots,r\}$ and there is at least one $i$ such that $B_{z_i}$ contains both $x'$ and $a$.  However, by \tref{subtree-root} $x'$ is not contained in $B_{z_i}$ for any $i\in\{1,\ldots,r\}$.  Therefore $a\in B_{z_i}$ for each $i\in\{0,\ldots,r\}$.  In particular, $a$ is contained in $B_x$.
   Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $B_x$ contains $x$.
 \end{proof}

 \referee{2}{Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
 (twice).  Moreover, the way the proof is written, it seems to suggest
 that we only get that $a \in B_{z_i}$ for each $i \in \{1, \dots, r\}$.
 The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
 true though.  In a normalized tree-decomposition, if $xy$ is an edge and
 $x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
 and $T_y$ intersect).}

\note{PM}{I addressed the preceding referee's.}

 We are now ready to complete the proof.
 , which we restate here for convenience:

 \mmg*

 (This assumption is already present in the informal definition of $k$-planar graphs used in \cref{k_planar_section}.)

 \note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}\note{PM}{No, under the current definition, $r$ edges crossing at a single point $p$ generate $\binom{r}{2}$ crossings, with each edge taking part in $r-1$ crossings, just like they would if we perturbed the edges so that each pair crosses at a distinct point.  I guess we could mention this.  (See the first paragraph in the proof of Theorem~2 below.)} \note{DW}{The downside of this approach is that in the opening section we give an informal definition of $k$-planar, which needs the assumption of having no three edges cross at a point, since we use this assumption in the construction of a $(k+1,2)$-shortcut system for $k$-planar graphs give in the Introduction. I think we should not have different definitions in two parts of the paper.} A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$. Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.

\begin{thm}
\label{k-planar}
Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 48k+30$ in which $\tw(H)\leq \binom{k+4}{3}-1$.
\end{thm}

 We assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$. This assumption is justified since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.

 \begin{thm}
 \label{StringPartition}
 For integers $g\geq 0$ and $\delta\geq 2$, let $\ell:= \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$ and $t:= \binom{ \delta+4}{3}-1$.
 Then  every $(g,\delta)$-string graph:
 \begin{compactitem}
 	\item is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$,
 	\item has queue-number $\qn(G^k) \leq 3 \ell \, 2^t - \ceil{\tfrac{3}{2}\ell}$.
 \end{compactitem}
 \end{thm}

 \note{DW}{\cref{StringPartition} has been updated.}

 Our results also give bounds on the non-repetitive chromatic number and the $p$-centered chromatic number of $(g,\delta)$-string graphs, but the bounds are weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ and $\chi_p(G)\le p(64\delta)^2$ by results of \citet{DJKW16} and \citet{DFMS21}, respectively.

 Theorems~\ref{PlanarProduct}(b), \ref{GenusProduct}(b) and \cref{ShortcutProduct} and  \cref{DrawG,qn,p-centered,non-repetitive,PowerShortcut} imply the following result, which with $g=0$ implies  \cref{kPowerBasic} in the introduction.

 \begin{thm}
 \label{PowerGenus}
 For integers $g\geq 0$ and $k,\Delta\geq 1$, let $\ell:= \max\{2g,3\} (2k^4+6k^2) \Delta^{k}$ and $t:= \binom{k+3}{3}-1$. Then for every graph $G$ of Euler genus $g$ and maximum degree $\Delta$,
 \begin{compactitem}
 \item $G^k$ is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$
 \item $G^k$ is $(g, 2k(k+1)\Delta^{k} )$-planar,
 \item $G^k$ has queue-number $\qn(G^k) \leq 3 \ell \cdot 2^t - \ceil{\tfrac{3}{2}\ell}$.
 \item $G^k$ has  $p$-centered chromatic number $\chi_p(G^k) \leq \ell (p+1)\,  \binom{p+t}{t}$,
 \item $G^k$ has  non-repetitive chromatic number $ \pi(G^k) \leq \ell \, 4^{t+1}$.
 \end{compactitem}
 \end{thm}

  \note{DW}{\cref{PowerGenus} has been updated}

 This result is the first constant upper bound on the queue-number of bounded powers of graphs with bounded degree and bounded Euler genus.  For every graph $G$, since $G^k$ has maximum degree at most $\Delta^k$, a result of \citet{DJKW16} implies that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$. \cref{PowerGenus} improves upon this bound when $k,g\ll\Delta$.  Similarly, a result of \citet{DFMS21} implies that $\chi_p(G^k)\le 1024p\Delta^{2k}$ and \cref{PowerGenus} improves upon this bound when $p,k,g\ll\Delta$.

 \cref{qn,p-centered,non-repetitive,MinorFreeDegree,PowerProduct} imply the following analogous result for powers of graphs in any minor-closed class with bounded maximum degree.

 \begin{thm}
 \label{PowerMinor}
 For every graph $X$ there exists an integer $c$ such that for all integers $k,\Delta\geq 1$, if $t:= 2k\Delta^{k}(k^3+3k)\binom{k+c\Delta}{c\Delta}-1$ and $G$ is an $X$-minor-free graph with maximum degree $\Delta$, then:
 \begin{compactitem}
 	\item $G^k$ is contained in $H\boxtimes P$ for some graph $H$ with treewidth $t$ and for some path $P$.
 	\item $G^k$ has queue-number at most $3\cdot 2^t-2$,
 	\item $G^k$ has $p$-centered chromatic number at most $(p+1)\binom{p+t}{t}$.
 \end{compactitem}
 \end{thm}

 $1$-planar graphs are of particular interest to the graph drawing community~\citep{kobourov.liotta.ea:annotated}. In this case, we obtain better constants and an additional property (planarity) of $H$.

(\htmladdnormallink{https://arxiv.org/abs/1703.02261}{arXiv:1703.02261}) has 143 entries, and even this is now four years out of date.

   An important aspect of our result for $1$-planar graphs is that graph $H$ in the product is planar and has treewidth $3$, i.e., it has simple treewidth $3$.  In a more recent application of product structure \citet{BDJM} show that such graphs have $\ell$-vertex rankings using $O(\log n/\log\log\log n)$ colours, and this is tight. This result requires that $H$ have simple treewidth $3$, even treewidth $3$ is not sufficient.
 }

 \note{PM}{Use restateable here...}

 \begin{thm}
 \label{1-planar}
 Every 1-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
 \end{thm}

In \cref{1_planar_is_4_framed}, below, we show that every $1$-planar graph is a subgraph of a $4$-framed multigraph. This makes \cref{1-planar} an immediate consequence of the following theorem, whose proof is the subject of the rest of this section.

 \note{DW}{Write  ``A multigraph $G$ is $1$-planar if and only if ...''}

Before proving \cref{1_planar_is_4_framed}, we remark that the lemma seems not to be true if we replace `multigraph' with `graph'.  We are only aware of two previous results on the relationship between $1$-planar graphs and $d$-framed graphs \cite{BDGGMR}:
\begin{inparaenum}[(i)]
  \item Every $3$-connected $1$-planar graph is contained in some $4$-framed graph.
  \item Every $1$-planar graph is contained in some $8$-framed graph.
\end{inparaenum}
Neither of these results allows a proof of \cref{1-planar} without sacrificing generality or accepting a larger constant.

 \note{PM}{Check David's commit: a0e20aff83d12c74127331854ce9fb75e002efe3}

 Map graphs are defined as follows. Start with a graph $G_0$ embedded in a surface of Euler genus $g$, with each face labelled a `nation' or a `lake', where each vertex of $G_0$ is incident with at most $d$ nations. Let $G$ be the graph whose vertices are the nations of $G_0$, where two vertices are adjacent in $G$ if the corresponding faces in $G_0$ share a vertex. Then $G$ is called a \defin{$(g,d)$-map graph}.  A $(0,d)$-map graph is called a (plane) \defin{$d$-map graph}; see \citep{FLS-SODA12,CGP02} for example. The $(g,3)$-map graphs are precisely the graphs of Euler genus at most $g$; see \citep{dujmovic.eppstein.ea:structure}.

 \note{DW}{This subsection needs a clean-up; define $(g,d)$-framed graph.}

\note{PM}{Got rid of genus-$g$ in the following. We use shortcuts for $(g,d)$-map graphs and now we don't have to define $(g,d)$-frame.}

\note{PM}{Old proof}

\begin{proof}
 \note{PM}{I fixed some issues here that come from the fact that some vertices of $G_0$ do not contribute faces to $G_1$.  This happens when two nations and two lakes meet at a vertex $x$.  It's not an issue but it means we can't write ``the face of $G_1$ that corresponds to $x$.''} \note{DW}{Good}
Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations. If two lakes have a common edge in their boundaries, then we may delete the edges and merge the lakes \note{DW}{Delete the previous sentence? This sentence is incompatible with ``in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes'' below.}. Now assume that no two lakes have a common edge in their boundaries.
Let $G$ be the corresponding map graph.
Let $G_1$ be the graph with $V(G_1):=V(G)$, which is the set of nations, and whose edge set we now define.  Consider vertices $v$ and $w$ of $G_1$.
Let $A$ and $B$ be the nation faces of $G_0$ corresponding to $v$ and $w$.
Then $vw\in E(G_1)$ if $A$ and $B$ have a common edge in their boundaries,
or there is a vertex $x$ of $G_0$ in the boundary of $A$ and $B$, and in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes.
 Then $G_1$ has Euler genus at most $g$.
Each face of $G_1$ corresponds to a vertex of $G_0$ or a lake of $G_0$.
Triangulate each face of $G_1$ corresponding to a lake of $G_0$.

For each face $F$ of $G_1$ that corresponds to a vertex $x$ of $G_0$, the size of $F$ \note{DW}{Is ``size'' defined? Do we mean length of the facial cycle? } equals the number of nations incident to $x$, which is at most $d$. So every face of $G_1$ has size at most $d$.  Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.  \note{DW}{I like this notation. Should we use it throughout?} \note{PM}{No. For most things I prefer $G$ and $G_0$ rather than $\widehat{G}_0$ and $G_0$.}

To complete the proof, we now show that $\widehat{G}_1$ contains every edge of $G$.
If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common.   If some face of $G_1$ corresponds to $x$, then $v$ and $w$ are on a common face of $G_1$ so $vw\in E(\widehat{G}_1)$.  If no face of $G_1$ corresponds to $x$\footnote{This can occur when $x$ is on the boundary of exactly two nation faces and two lake faces.}, then the only two nation faces of $G_0$ with $x$ on their boundary are those corresponding to $v$ and $w$, in which case $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.  Therefore $G\subseteq\widehat{G}_1$.
\end{proof}

 Here we discuss some of the consequences of the above theorems for $k$-planar and $(g,k)$-planar graphs.

 \referee{1}{4. Sections 4 and 5: I think as they are currently organized they distract from the purpose of the paper, as stated in the introduction, to ``prove product structure theorems for several non-minor-closed classes of interest.'' Which of Corollaries 1--4 and Theorems 12--17 are most important? I would remove any mention of the applications from the examples section, put that section first, and state in that section a single main theorem with all of the best bounds on the product structure. The applications section is mostly a survey and I think should be focused on the most exciting new corollaries. Theorems 18 and 19 should be in the applications section.}

 \note{DW}{I don't want to do a complete restructuring. In our response, let's just say why we structure it the way we do.}
 \note{PM}{Sure, though I think they proposing a chop rather than a restructuring.}

 , which with \cref{qn} implies:

 \begin{cor}
 \label{1PlanarQueue}
 Every 1-planar graph has queue-number at most
 $3 \times 7 \times 5 + \floor{\tfrac{3}{2} \times 7} = 115$.
 \end{cor}

 \begin{cor}
 \label{dMapQueue}
 Every $d$-map graph has queue-number at most
 $\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$.
 \note{DW}{I have updated this bound using the better qn bound for planar 3-trees. Note that with $d=3$ we get the original bound of 49.}
 \end{cor}

 \cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$.
 \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
 \note{OM}{Ok}
 \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
 \note{PM}{Ok}

 \cref{non-repetitive,kPlanarProduct,1-planar,gkPlanarProduct} imply the following result:

 \begin{cor}\quad
 \begin{compactitem}
 \item	For every $1$-planar graph $G$, $\pi(G)\le 7\times 4^4=1792$.
 \item For every $k$-planar graph $G$,
 	$\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
 \item	For every $(g,k)$-planar graph $G$,
 	$  \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}.$
  \note{DW}{I have changed $4^{\binom{k+5}{4}}$ to $4^{\binom{k+4}{3}}$, okay?}
  \note{PM}{Ok}
 \end{compactitem}
 \end{cor}

In the journal version, \citet{DFMS21} proved that $\chi_p(H_1\boxtimes H_2) \leq \chi_p(H_1) \cdot \chi(H_2^p)$ where $H_2^p$ is the $p$-th power of $H_2$. Apply this with $H_2:= P \boxtimes K_\ell$. Then $H_2^p= P^p\boxtimes K_\ell$ and $\chi(H_2^p)=(p+1)\ell$. So $\chi_p(H\boxtimes P \boxtimes K_\ell) \leq \chi_p(H_1) \cdot (p+1)\ell$. }

\note{DW}{For consistency with \cref{qn}, I suggest we combine \cref{p-centered,p-centered-treewidth}, and just write, ``}

\begin{proof}
	By \cref{PartitionProduct}, $G$ has an $H$-partition $(\mathcal{L}=\langle L_0,L_1,\ldots\rangle, \PP=(B_x:x\in V(H))$ with layered width at most $\ell$. Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.  For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in L_i\cap B_x$ such that:
   \begin{compactenum}
     \item $\alpha(v)$ is unique among $\{\alpha(w): w\in L_i\cap B_x\}$, which is possible
     since $|L_i\cap B_x|\le \ell$,
     \item $\beta(v)= i\bmod (p+1)$, and
     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
   \end{compactenum}
 To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.

 First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$. Since $X$ is connected, $X$ contains a path from $v$ to $w$.  By the definition of layering, this path contains at least one vertex from $L_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$. Therefore, $|\{\beta(v'):v'\in V(X)\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.

 Otherwise, $V(X)\subseteq L_{i}\cup\cdots\cup L_{i+s}$ for some $s<p$.  Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.  If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in V(X)\}|> p$ so $|\{\phi(v):v\in V(X)\}|> p$ and we are done.  Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
 For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$. Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$. Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
 \end{proof}

\referee{2}{Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.}

 \cref{p-centered,kPlanarProduct,1-planar,gkPlanarProduct} immediately imply the following results, for every $p\ge 2$:

 \begin{cor}\quad
 \begin{compactitem}
 \item For every $1$-planar graph $G$,\; $ \chi_p(G)\le 5 (p+3)(p+2)(p+1)^2$.
 \item For every $k$-planar graph $G$,\; $\displaystyle \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}$.
 \item For every $(g,k)$-planar graph $G$,\;
 $\displaystyle \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{p+\binom{k+4}{3}-1}{\binom{k+4}{3}-1}$.
  \note{DW}{I have changed $\binom{k+5}{4}$ to $\binom{k+4}{3}$, okay?}
  \note{PM}{Ok}
  \end{compactitem}
 \end{cor}

 \subsection{Text following \cref{gkPlanarProduct}}

 Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs (or any of the other classes presented in \cref{examples}) was in terms of layered treewidth, which we now define.  A \defin{layered tree-decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{T}$ of $G$. The layered width of $(\mathcal{L},\mathcal{T})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{T}\}$.  The \defin{layered treewidth} of $G$ is the minimum layered width of any layered tree-decomposition of $G$. \citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph. \citet{dujmovic.eppstein.ea:structure} show that every $k$-planar graph has layered treewidth at most $6(k+1)$, and more generally that every $(g,k)$-planar graph has layered treewidth at most $(4g+6)(k+1)$. It follows from this result that $(g,k)$-planar graphs have treewidth $O(\sqrt{(g+1)(k+1)n})$ and thus have balanced separators of the same order, which can also be concluded from the work of \citet{FP08}. In related work, \citet{grigoriev.bodlaender:algorithms} used structural results to obtain approximation algorithms for $(g,k)$-planar graphs, and \citet{PachToth97} determined the maximum number of edges in a $k$-planar graph (up to a constant factor).

 If a graph class admits bounded layered partitions, then it also has bounded layered treewidth. In particular, if $\PP=(P_x:x\in V(H))$ is an $H$-partition of $G$ of layered width $\ell$ with respect to some layering $\mathcal{L}$ of $G$ and $(B_x:x\in V(T))$ is a width-$t$ tree-decomposition of $H$, then setting $C_x = \bigcup_{y\in B_x} P_y$ for each $x\in V(T)$ gives a tree-decomposition $(C_x:x\in V(T))$ of $G$ that has layered treewidth $(t+1)\ell$ \cite{DJMMUW20}. Therefore, any property that holds for graphs of bounded layered treewidth also holds for $G$. What sets layered partitions apart from layered treewidth is that they lead to constant upper bounds on the queue-number  and non-repetitive chromatic number, whereas for both these parameters, the best known upper bound obtainable via layered treewidth is $O(\log n)$; see \cref{Applications}.

 %%%%%%%%%%%%%%%%
 \label{sec-k-planar}

 \note{DW}{I am pretty sure one can get treewidth 3 for $(g,1)$-planar graphs, combining \cref{d_framed_product_stucture} with the method in \citep{DHHW}. I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}). Okay?}
 \note{PM}{Okay to this and to all similar questions that come later (which I've now commented out.)}
 \note{DW}{Robert is working on it.}

 This section describes several examples of graph classes that can be obtained from a shortcut system typically applied to graphs of bounded Euler genus.

 \subsection{Orphaned Theorems}

 Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,non-repetitive} imply the following result.

 \begin{thm}
 \label{PlaneMapPartition}
 \note{DW}{For every integer $d\geq 4$, }
 Every $d$-map graph $G$:
 	\begin{compactitem}
 		\item is contained in $H \boxtimes P \boxtimes K_{21d(d-3)}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
 		\item has queue-number $\qn(G) < 32225\, d(d-3)$.
 		\item has $p$-centered chromatic number $\chi_p(G) \leq 21d(d-3) (p+1)  \binom{p+9}{9}$,
 		\item has non-repetitive chromatic number $ \pi(G) \leq 21 \cdot 4^{10} d(d-3)$.
 	\end{compactitem}
 \end{thm}

 \begin{thm}
 \label{MapPartition}
 For integers $g\geq 0$ and $d\geq 4$, if $\ell:=  7d(d-3)\, \max\{2g,3\}$ then every $(g,d)$-map graph $G$:
 \begin{compactitem}
 \item is contained in $H \boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with $\tw(H)\leq 9$,
 \item has queue-number $\qn(G) <  1535\, \ell $.
 \item has $p$-centered chromatic number $\chi_p(G) \leq \ell (p+1)\,  \binom{p+9}{9}$,
 \item has non-repetitive chromatic number $ \pi(G) \leq 4^{10}\,\ell $.
 \end{compactitem}
 \end{thm}

 These results give the first constant upper bound on the non-repetitive chromatic number of map graphs, the first polynomial bounds on the $p$-centered chromatic number of map graphs, and the best known bounds on the queue-number of map graphs.

  \note{DW}{I am pretty sure one can get treewidth 3 here using \cref{framed} and the method in \citep{DHHW}. I think the best place for this result is reference \citep{DHHW} since the present paper pre-dates \citep{DHHW} (and this addition will increase the significance of \citep{DHHW}). Okay?}

  \subsection{Orphaned Section}

  \cref{nearest-neighbour,kPlanarProduct,qn,p-centered} imply:

  \begin{cor}
  \label{k-nn}
  For every integer $k\geq 1$ there exists integers $t\leq O(k^6)$ and $\ell\leq O(k^4)$ such that every $k$-nearest-neighbour graph:
  \begin{compactitem}
  \item is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with treewidth $t$ and some path $P$,
  \item has queue-number at most $2^{O(k^6)}$, and
  \item has $p$-centered chromatic number at most $\ell (p+1)\binom{p+t}{t}$.
  \end{compactitem}
  \end{cor}

  \cref{non-repetitive,k-nn} also give bounds on the non-repetitive chromatic number of a $k$-nearest neighbour graph $G$. However, the bound is weak, since $G$ has maximum degree at most $6k$, implying that $\pi(G) \leq (36+o(1))k^2$ by a result of \citet{DJKW16}.

 \subsection{Generalisations}
 \label{Generalisations}

  As mentioned above, product structure theorems have been established for several minor-closed classes in addition to planar graphs. The first generalises \cref{PlanarProduct} for graphs of bounded Euler genus.

  \begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
  \label{GenusProduct}
  Every graph of Euler genus $g$ is contained in:
  \begin{compactenum}[(a)]
  \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
  \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
  \end{compactenum}
  \end{thm}

 \citet{DJMMUW20} generalised \cref{GenusProduct} for apex-minor-free graphs as follows.

 \begin{thm}[\citep{DJMMUW20}]
 \label{ApexMinorFree}
 For every apex graph $X$, there exists $c\in\mathbb{N}$ such that every $X$-minor-free graph is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and some path $P$.
 \end{thm}

 The assumption that $X$ is apex is needed in \cref{ApexMinorFree}, since if the class of $X$-minor-free graphs has a product structure theorem analogous to \cref{PlanarProduct}, then $X$ is apex \citep{DJMMUW20}. On the other hand, \citet{DEMWW22} proved a product structure theorem for bounded degree graphs in any minor-closed class.

 \begin{thm}[\citep{DEMWW22}]
 	\label{MinorFreeDegree}
 	For every graph $X$ there exists $c\in\mathbb{N}$ such that for every $\Delta\in\mathbb{N}$, every $X$-minor-free graph $G$ with maximum degree at most $\Delta$ is contained in $H\boxtimes P$ for some graph $H$ with $\tw(H) \leq c\Delta$ and for some path $P$.
 \end{thm}

 The next three results are immediate corollaries of \cref{GenusProduct,ApexMinorFree,ShortcutProduct,MinorFreeDegree}.

 \begin{thm}
 Let $\SS$ be a $(k,d)$-shortcut system for a graph $G$ of Euler genus $g$. Then $G^\SS$ is contained in $H\boxtimes P$ for some graph $H$ of treewidth at most $d(k^3+3k)\binom{k+2g+8}{2g+8}-1$ and for some path $P$.
 \end{thm}

 \begin{thm}
 For every apex graph $X$ and for all integers $k,d\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ and for every $(k,d)$-shortcut system $\SS$ for $G$, $G^\SS\subseteq H\boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
 \end{thm}

 \begin{thm}
 For every graph $X$ and for all integers $k,\Delta\geq 1$, there is an integer $c$ such that for every $X$-minor-free graph $G$ with maximum degree $\Delta$ and for every $(k,\Delta)$-shortcut system $\SS$ for $G$, $G^\SS \subseteq H \boxtimes P$ for some graph $H$ with $\tw(H)\leq c$ and for some path $P$.
 \end{thm}

 \note{DW}{delete the above three theorems, they are never used}

 \subsection*{Note Added in Proof} Subsequent to the initial release of this paper, the treewidth 8 bound in \cref{PlanarProduct} was improved to 6 by \citet{UWY}, which implies that the treewidth $2g+8$ bound in \cref{GenusProduct}(a) can be improved to $2g+6$. Simmilarly, the treewidth 4 bound in \cref{GenusProduct}(b) was improved to 3 by \citet{DHHW}. \note{DW}{Say more about the consequence for the bounds in the rest of the paper}

 \todo[inline]{Alternative to the final three paragraphs in the preceding proof:

 To see that $H$ is planar, consider a plane multigraph $G''$ obtained by adding stars in some of the faces of $G'$, as follows:  When creating the part $S$, add a vertex $v_\tau$ to the interior of $\tau$ that is adjacent to each vertex of $\tau$ and place $v_\tau$ in $S$.  The graph $G''\supseteq G'$ obtained this way is clearly planar.  Furthermore, for each part $P\in\mathcal{P}$, $G''[P]$ is connected.  Since planarity is preserved under edge contraction, it follows that $G''/\mathcal{P}$ is planar.  Now argue that $G/\mathcal{P}=G''/\mathcal{P}$\ldots
 }

### 2022-04-07 k-planar.tex
correctly sorted the references

@@
 
-\bibliographystyle{DavidNatbibStyle}
-\bibliography{k-planar}
+%\bibliographystyle{DavidNatbibStyle}
+%\bibliography{k-planar}
+
+\begin{thebibliography}{49}
+	\providecommand{\natexlab}[1]{#1}
+	\providecommand{\msn}[1]{MR:\,\href{http://www.ams.org/mathscinet-getitem?mr=MR{#1}}{#1}}
+	\providecommand{\ZBL}[1]{Zbl:\,\href{https://www.zentralblatt-math.org/zmath/en/search/?q=an:#1}{#1}}
+	\providecommand{\url}[1]{\texttt{#1}}
+	\providecommand{\urlprefix}{}
+	\expandafter\ifx\csname urlstyle\endcsname\relax
+	\providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
+	\providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
+		\urlstyle{rm}\Url}\fi
+
+	\bibitem[{{\'{A}}brego et~al.(2011){\'{A}}brego, Monroy,
+		Fern{\'{a}}ndez{-}Merchant, Flores{-}Pe{\~{n}}aloza, Hurtado,
+		Sacrist{\'{a}}n, and Saumell}]{abrego.munroy.ea:on}
+	\textsc{Bernardo~M. {\'{A}}brego, Ruy~Fabila Monroy, Silvia
+		Fern{\'{a}}ndez{-}Merchant, David Flores{-}Pe{\~{n}}aloza, Ferran Hurtado,
+		Vera Sacrist{\'{a}}n, and Maria Saumell}.
+	\newblock \href{https://doi.org/10.1016/j.comgeo.2010.11.003}{On crossing
+		numbers of geometric proximity graphs}.
+	\newblock \emph{Comput. Geom.}, 44(4):216--233, 2011.
+
+	\bibitem[{Aigner and Ziegler(2010)}]{Proofs4}
+	\textsc{Martin Aigner and G{\"u}nter~M. Ziegler}.
+	\newblock Proofs from {T}he {B}ook.
+	\newblock Springer, 4th edn., 2010.
+
+	\bibitem[{Ajtai et~al.(1982)Ajtai, Chv{\'a}tal, Newborn, and
+		Szemer{\'e}di}]{ajtai.chvatal.ea:crossing-free}
+	\textsc{Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, and Endre
+		Szemer{\'e}di}.
+	\newblock \href{https://doi.org/10.1016/S0304-0208(08)73484-4}{Crossing-free
+		subgraphs}.
+	\newblock \emph{Ann. Discrete Math.}, 12:9--12, 1982.
+
+	\bibitem[{Alam et~al.(2020)Alam, Bekos, Gronemann, Kaufmann, and
+		Pupyrev}]{ABGKP20}
+	\textsc{Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael
+		Kaufmann, and Sergey Pupyrev}.
+	\newblock \href{https://doi.org/10.1007/s00453-020-00697-4}{Queue layouts of
+		planar 3-trees}.
+	\newblock \emph{Algorithmica}, 82:2564--2585, 2020.
+
+	\bibitem[{Alon et~al.(2002)Alon, Grytczuk, Ha{\l}uszczak, and
+		Riordan}]{AGHR-RSA02}
+	\textsc{Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, and Oliver
+		Riordan}.
+	\newblock \href{https://doi.org/10.1002/rsa.10057}{Nonrepetitive colorings of
+		graphs}.
+	\newblock \emph{Random Structures Algorithms}, 21(3-4):336--346, 2002.
+
+	\bibitem[{Babai et~al.(1982)Babai, Chung, Erd\H{o}s, Graham, and
+		Spencer}]{babai.chung.ea:on}
+	\textsc{Laszlo Babai, F.~R.~K. Chung, Paul Erd\H{o}s, Ronald~L. Graham, and
+		J.~H. Spencer}.
+	\newblock On graphs which contain all sparse graphs.
+	\newblock In \emph{Theory and practice of combinatorics. A collection of
+		articles honoring Anton Kotzig on the occasion of his sixtieth birthday}, pp.
+	21--26. Elsevier, 1982.
+
+	\bibitem[{Bekos et~al.(2020)Bekos, Lozzo, Griesbach, Gronemann, Montecchiani,
+		and Raftopoulou}]{BDGGMR}
+	\textsc{Michael~A. Bekos, Giordano~Da Lozzo, Svenja Griesbach, Martin
+		Gronemann, Fabrizio Montecchiani, and Chrysanthi~N. Raftopoulou}.
+	\newblock \href{https://doi.org/10.4230/LIPIcs.SoCG.2020.16}{Book embeddings of
+		nonplanar graphs with small faces in few pages}.
+	\newblock In \textsc{Sergio Cabello and Danny~Z. Chen}, eds., \emph{36th
+		International Symposium on Computational Geometry, SoCG 2020, June 23-26,
+		2020, Z{\"{u}}rich, Switzerland}, vol. 164 of \emph{LIPIcs}, pp. 16:1--16:17.
+	Schloss Dagstuhl - Leibniz-Zentrum f{\"{u}}r Informatik, 2020.
+	\newblock arXiv:2003.07655.
+
+	\bibitem[{Bodlaender(1998)}]{Bodlaender-TCS98}
+	\textsc{Hans~L. Bodlaender}.
+	\newblock \href{https://doi.org/10.1016/S0304-3975(97)00228-4}{A partial
+		$k$-arboretum of graphs with bounded treewidth}.
+	\newblock \emph{Theoret. Comput. Sci.}, 209(1-2):1--45, 1998.
+
+	\bibitem[{Bonnet et~al.(2022)Bonnet, Kwon, and Wood}]{BKW}
+	\textsc{\'Edouard Bonnet, {O-joung} Kwon, and David~R. Wood}.
+	\newblock \href{http://arxiv.org/abs/2202.11858}{Reduced bandwidth: a
+		qualitative strengthening of twin-width in minor-closed classes (and
+		beyond)}.
+	\newblock 2022, arXiv:2202.11858.
+
+	\bibitem[{Bose et~al.(2012)Bose, Dujmovic, Hurtado, Iacono, Langerman, Meijer,
+		Adinolfi, Saumell, and Wood}]{ProximityGraphs}
+	\textsc{Prosenjit Bose, Vida Dujmovic, Ferran Hurtado, John Iacono, Stefan
+		Langerman, Henk Meijer, Vera~Sacrist{\'{a}}n Adinolfi, Maria Saumell, and
+		David~R. Wood}.
+	\newblock \href{https://doi.org/10.1142/S0218195912500112}{Proximity graphs: E,
+		{\(\delta\)}, {\(\Delta\)}, {\(\chi\)} and {\(\omega\)}}.
+	\newblock \emph{Int. J. Comput. Geometry Appl.}, 22(5):439--470, 2012.
+
+	\bibitem[{Bose et~al.(2020)Bose, Dujmovi\'c, Javarsineh, and Morin}]{BDJM}
+	\textsc{Prosenjit Bose, Vida Dujmovi\'c, Mehrnoosh Javarsineh, and Pat Morin}.
+	\newblock \href{http://arxiv.org/abs/2007.06455}{Asymptotically optimal vertex
+		ranking of planar graphs}.
+	\newblock 2020, arXiv:2007.06455.
+
+	\bibitem[{Bose et~al.(2022)Bose, Morin, and Odak}]{bose.morin.ea:optimal}
+	\textsc{Prosenjit Bose, Pat Morin, and Saeed Odak}.
+	\newblock \href{http://arxiv.org/abs/2202.08870}{An optimal algorithm for
+		product structure in planar graphs}.
+	\newblock 2022, arXiv:2202.08870.
+
+	\bibitem[{Bose et~al.(2001)Bose, Morin, Stojmenović, and
+		Urrutia}]{bose.morin.ea:routing}
+	\textsc{Prosenjit Bose, Pat Morin, Ivan Stojmenović, and Jorge Urrutia}.
+	\newblock \href{https://doi.org/10.1023/A:1012319418150}{Routing with
+		guaranteed delivery in ad hoc wireless networks}.
+	\newblock \emph{Wireless Networks}, 7(6):609--616, 2001.
+
+	\bibitem[{Brandenburg(2019)}]{Brandenburg19}
+	\textsc{Franz~J. Brandenburg}.
+	\newblock \href{https://doi.org/10.1007/s00453-018-0510-x}{Characterizing and
+		recognizing 4-map graphs}.
+	\newblock \emph{Algorithmica}, 81(5):1818--1843, 2019.
+
+	\bibitem[{Brandenburg(2020)}]{Brandenburg20}
+	\textsc{Franz~J. Brandenburg}.
+	\newblock \href{http://arxiv.org/abs/2012.06874}{Book embeddings of $k$-map
+		graphs}.
+	\newblock 2020, arXiv:2012.06874.
+
+	\bibitem[{Chen et~al.(2002)Chen, Grigni, and Papadimitriou}]{CGP02}
+	\textsc{Zhi-Zhong Chen, Michelangelo Grigni, and Christos~H. Papadimitriou}.
+	\newblock \href{https://doi.org/10.1145/506147.506148}{Map graphs}.
+	\newblock \emph{J. ACM}, 49(2):127--138, 2002.
+
+	\bibitem[{Chen et~al.(2006)Chen, Grigni, and Papadimitriou}]{CGP06}
+	\textsc{Zhi-Zhong Chen, Michelangelo Grigni, and Christos~H. Papadimitriou}.
+	\newblock \href{https://doi.org/10.1007/s00453-005-1184-8}{Recognizing
+		hole-free 4-map graphs in cubic time}.
+	\newblock \emph{Algorithmica}, 45(2):227--262, 2006.
+
+	\bibitem[{Diestel(2018)}]{Diestel5}
+	\textsc{Reinhard Diestel}.
+	\newblock \href{http://diestel-graph-theory.com/}{Graph theory}, vol. 173 of
+	\emph{Graduate Texts in Mathematics}.
+	\newblock Springer, 5th edn., 2018.
+
+	\bibitem[{Distel et~al.(2021)Distel, Hickingbotham, Huynh, and Wood}]{DHHW}
+	\textsc{Marc Distel, Robert Hickingbotham, Tony Huynh, and David~R. Wood}.
+	\newblock \href{http://arxiv.org/abs/2112.10025}{Improved product structure for
+		graphs on surfaces}.
+	\newblock 2021, arXiv:2112.10025.
+
+	\bibitem[{D\k{e}bski et~al.(2021)D\k{e}bski, Felsner, Micek, and
+		Schr\"{o}der}]{DFMS21}
+	\textsc{Micha{\l} D\k{e}bski, Stefan Felsner, Piotr Micek, and Felix
+		Schr\"{o}der}.
+	\newblock \href{https://doi.org/10.19086/aic.27351}{Improved bounds for
+		centered colorings}.
+	\newblock \emph{Adv. Comb.}, \#8, 2021.
+
+	\bibitem[{Dujmovi\'c et~al.(2017{\natexlab{a}})Dujmovi\'c, Eppstein, and
+		Wood}]{dujmovic.eppstein.ea:structure}
+	\textsc{Vida Dujmovi\'c, David Eppstein, and David~R. Wood}.
+	\newblock \href{https://doi.org/10.1137/16M1062879}{Structure of graphs with
+		locally restricted crossings}.
+	\newblock \emph{{SIAM} J. Discrete Math.}, 31(2):805--824, 2017{\natexlab{a}}.
+
+	\bibitem[{Dujmovi\'c et~al.(2021)Dujmovi\'c, Esperet, Gavoille, Joret, Micek,
+		and Morin}]{DEJGMM21}
+	\textsc{Vida Dujmovi\'c, Louis Esperet, Cyril Gavoille, Gwena\"el Joret, Piotr
+		Micek, and Pat Morin}.
+	\newblock \href{https://doi.org/10.1145/3477542}{Adjacency labelling for planar
+		graphs (and beyond)}.
+	\newblock \emph{J. ACM}, 68(6):42, 2021.
+
+	\bibitem[{Dujmovi\'c et~al.(2020)Dujmovi\'c, Esperet, Joret, Walczak, and
+		Wood}]{dujmovic.esperet.ea:planar}
+	\textsc{Vida Dujmovi\'c, Louis Esperet, Gwena{\"{e}}l Joret, Bartosz Walczak,
+		and David~R. Wood}.
+	\newblock \href{https://doi.org/10.19086/aic.12100}{Planar graphs have bounded
+		nonrepetitive chromatic number}.
+	\newblock \emph{Adv. Comb.}, \#5, 2020.
+
+	\bibitem[{Dujmovi{\'c} et~al.(2022)Dujmovi{\'c}, Esperet, Morin, Walczak, and
+		Wood}]{DEMWW22}
+	\textsc{Vida Dujmovi{\'c}, Louis Esperet, Pat Morin, Bartosz Walczak, and
+		David~R. Wood}.
+	\newblock \href{https://doi.org/10.1017/S0963548321000213}{Clustered
+		3-colouring graphs of bounded degree}.
+	\newblock \emph{Combin. Probab. Comput.}, 31(1):123--135, 2022.
+
+	\bibitem[{Dujmovi\'{c} et~al.(2016)Dujmovi\'{c}, Joret, Kozik, and
+		Wood}]{DJKW16}
+	\textsc{Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, and David~R. Wood}.
+	\newblock \href{https://doi.org/10.1007/s00493-015-3070-6}{Nonrepetitive
+		colouring via entropy compression}.
+	\newblock \emph{Combinatorica}, 36(6):661--686, 2016.
+
+	\bibitem[{Dujmovi{\'c} et~al.(2020)Dujmovi{\'c}, Joret, Micek, Morin, Ueckerdt,
+		and Wood}]{DJMMUW20}
+	\textsc{Vida Dujmovi{\'c}, Gwena\"{e}l Joret, Piotr Micek, Pat Morin, Torsten
+		Ueckerdt, and David~R. Wood}.
+	\newblock \href{https://doi.org/10.1145/3385731}{Planar graphs have bounded
+		queue-number}.
+	\newblock \emph{J. ACM}, 67(4):\#22, 2020.
+
+	\bibitem[{Dujmovi\'c et~al.(2022)Dujmovi\'c, Morin, and Odak}]{DMO}
+	\textsc{Vida Dujmovi\'c, Pat Morin, and Saeed Odak}.
+	\newblock \href{http://arxiv.org/abs/2202.12882}{Odd colourings of graph
+		products}.
+	\newblock 2022, arXiv:2202.12882.
+
+	\bibitem[{Dujmovi\'c et~al.(2017{\natexlab{b}})Dujmovi\'c, Morin, and
+		Wood}]{dujmovic.morin.ea:layered}
+	\textsc{Vida Dujmovi\'c, Pat Morin, and David~R. Wood}.
+	\newblock \href{https://doi.org/10.1016/j.jctb.2017.05.006}{Layered separators
+		in minor-closed graph classes with applications}.
+	\newblock \emph{J. Comb. Theory, Ser. {B}}, 127:111--147, 2017{\natexlab{b}}.
+
+	\bibitem[{Dvo{\v{r}}{\'a}k et~al.(2021)Dvo{\v{r}}{\'a}k, Huynh, Joret, Liu, and
+		Wood}]{DHJLW21}
+	\textsc{Zden{\v{e}}k Dvo{\v{r}}{\'a}k, Tony Huynh, Gwena\"el Joret, Chun-Hung
+		Liu, and David~R. Wood}.
+	\newblock \href{https://doi.org/10.1007/978-3-030-62497-2_32}{Notes on graph
+		product structure theory}.
+	\newblock In \textsc{David~R. Wood, Jan de~Gier, Cheryl~E. Praeger, and Terence
+		Tao}, eds., \emph{2019-20 MATRIX Annals}, pp. 513--533. Springer, 2021.
+	\newblock arXiv:2001.08860.
+
+	\bibitem[{Elmallah and Colbourn(1990)}]{elmallah.colbourn:on}
+	\textsc{Ehab~S. Elmallah and Charles~J. Colbourn}.
+	\newblock \href{https://doi.org/10.1016/0012-365X(90)90293-Q}{On two dual
+		classes of planar graphs}.
+	\newblock \emph{Discrete. Math.}, 80(1):21--40, 1990.
+
+	\bibitem[{Esperet et~al.(2020)Esperet, Joret, and Morin}]{EJM}
+	\textsc{Louis Esperet, Gwena\"{e}l Joret, and Pat Morin}.
+	\newblock \href{http://arxiv.org/abs/2010.05779}{Sparse universal graphs for
+		planarity}.
+	\newblock 2020, arXiv:2010.05779.
+
+	\bibitem[{Fomin et~al.(2012)Fomin, Lokshtanov, and Saurabh}]{FLS-SODA12}
+	\textsc{Fedor~V. Fomin, Daniel Lokshtanov, and Saket Saurabh}.
+	\newblock \href{https://doi.org/10.1137/1.9781611973099.124}{Bidimensionality
+		and geometric graphs}.
+	\newblock In \emph{Proc. 23rd {A}nnual {ACM}-{SIAM} {S}ymposium on {D}iscrete
+		{A}lgorithms}, pp. 1563--1575. 2012.
+
+	\bibitem[{Fox and Pach(2010)}]{FP10}
+	\textsc{Jacob Fox and J{\'a}nos Pach}.
+	\newblock \href{https://doi.org/10.1017/S0963548309990459}{A separator theorem
+		for string graphs and its applications}.
+	\newblock \emph{Combin. Probab. Comput.}, 19(3):371--390, 2010.
+
+	\bibitem[{Fox and Pach(2014)}]{FP14}
+	\textsc{Jacob Fox and J{\'a}nos Pach}.
+	\newblock \href{https://doi.org/10.1017/S0963548313000412}{Applications of a
+		new separator theorem for string graphs}.
+	\newblock \emph{Combin. Probab. Comput.}, 23(1):66--74, 2014.
+
+	\bibitem[{Harvey and Wood(2017)}]{HW17}
+	\textsc{Daniel~J. Harvey and David~R. Wood}.
+	\newblock \href{https://doi.org/10.1002/jgt.22030}{Parameters tied to
+		treewidth}.
+	\newblock \emph{J. Graph Theory}, 84(4):364--385, 2017.
+
+	\bibitem[{Heath et~al.(1992)Heath, Leighton, and Rosenberg}]{HLR92}
+	\textsc{Lenwood~S. Heath, F.~Thomson Leighton, and Arnold~L. Rosenberg}.
+	\newblock \href{https://doi.org/10.1137/0405031}{Comparing queues and stacks as
+		mechanisms for laying out graphs}.
+	\newblock \emph{SIAM J. Discrete Math.}, 5(3):398--412, 1992.
+
+\bibitem[{van~den Heuvel and Wood(2018)}]{vdHW18}
+\textsc{Jan van~den Heuvel and David~R. Wood}.
+\newblock \href{https://doi.org/10.1112/jlms.12127}{Improper colourings
+	inspired by {H}adwiger's conjecture}.
+\newblock \emph{J. London Math. Soc.}, 98:129--148, 2018.
+\newblock arXiv:1704.06536.
+
+	\bibitem[{Kannan et~al.(1992)Kannan, Naor, and
+		Rudich}]{kannan.naor.ea:implicit}
+	\textsc{Sampath Kannan, Moni Naor, and Steven Rudich}.
+	\newblock \href{https://doi.org/10.1137/0405049}{Implicit representation of
+		graphs}.
+	\newblock \emph{{SIAM} J. Discrete Math.}, 5(4):596--603, 1992.
+
+	\bibitem[{Knauer and Ueckerdt(2012)}]{knauer.ueckerdt:simple}
+	\textsc{Kolja Knauer and Torsten Ueckerdt}.
+	\newblock
+	\href{https://kam.mff.cuni.cz/workshops/mcw/work18/mcw2012booklet.pdf}{Simple
+		treewidth}.
+	\newblock In \textsc{Pavel Rytír}, ed., \emph{Midsummer Combinatorial Workshop
+		Prague}. 2012.
+
+	\bibitem[{Kobourov et~al.(2017)Kobourov, Liotta, and
+		Montecchiani}]{kobourov.liotta.ea:annotated}
+	\textsc{Stephen~G. Kobourov, Giuseppe Liotta, and Fabrizio Montecchiani}.
+	\newblock \href{https://doi.org/10.1016/j.cosrev.2017.06.002}{An annotated
+		bibliography on 1-planarity}.
+	\newblock \emph{Comput. Sci. Rev.}, 25:49--67, 2017.
+
+	\bibitem[{Kratochv{\'{\i}}l and Vaner(2012)}]{kratochvil.vaner:note}
+	\textsc{Jan Kratochv{\'{\i}}l and Michal Vaner}.
+	\newblock \href{http://arxiv.org/abs/1210.8113}{A note on planar partial
+		3-trees}.
+	\newblock 2012, arXiv:1210.8113.
+
+	\bibitem[{Leighton(1983)}]{leighton:complexity}
+	\textsc{F.~Thomson Leighton}.
+	\newblock Complexity issues in {VLSI}.
+	\newblock MIT Press, Cambridge, MA, 1983.
+
+	\bibitem[{Mohar and Thomassen(2001)}]{mohar.thomassen:graphs}
+	\textsc{Bojan Mohar and Carsten Thomassen}.
+	\newblock Graphs on surfaces.
+	\newblock Johns Hopkins University Press, 2001.
+
+	\bibitem[{Ne{\v{s}}et{\v{r}}il and Ossona~de Mendez(2012)}]{Sparsity}
+	\textsc{Jaroslav Ne{\v{s}}et{\v{r}}il and Patrice Ossona~de Mendez}.
+	\newblock \href{https://doi.org/10.1007/978-3-642-27875-4}{Sparsity}, vol.~28
+	of \emph{Algorithms and Combinatorics}.
+	\newblock Springer, 2012.
+
+	\bibitem[{Pach and T{\'o}th(2002)}]{PachToth-DCG02}
+	\textsc{J{\'a}nos Pach and G{\'e}za T{\'o}th}.
+	\newblock \href{https://doi.org/10.1007/s00454-002-2891-4}{Recognizing string
+		graphs is decidable}.
+	\newblock \emph{Discrete Comput. Geom.}, 28(4):593--606, 2002.
+
+	\bibitem[{Pilipczuk and Siebertz(2021)}]{PS21}
+	\textsc{Micha{\l} Pilipczuk and Sebastian Siebertz}.
+	\newblock \href{https://doi.org/10.1016/j.jctb.2021.06.002}{Polynomial bounds
+		for centered colorings on proper minor-closed graph classes}.
+	\newblock \emph{J. Comb. Theory, Series B}, 151:111--147, 2021.
+
+	\bibitem[{Reed(2003)}]{Reed03}
+	\textsc{Bruce~A. Reed}.
+	\newblock \href{https://doi.org/10.1007/0-387-22444-0_4}{Algorithmic aspects of
+		tree width}.
+	\newblock In \emph{Recent advances in algorithms and combinatorics}, vol.~11,
+	pp. 85--107. Springer, 2003.
+
+	\bibitem[{Ueckerdt et~al.(2021)Ueckerdt, Wood, and Yi}]{UWY}
+	\textsc{Torsten Ueckerdt, David~R. Wood, and Wendy Yi}.
+	\newblock \href{http://arxiv.org/abs/2108.00198}{An improved planar graph
+		product structure theorem}.
+	\newblock 2021, arXiv:2108.00198.
+
+	\bibitem[{Zhu(2009)}]{zhu:colouring}
+	\textsc{Xuding Zhu}.
+	\newblock \href{https://doi.org/10.1016/j.disc.2008.03.024}{Colouring graphs
+		with bounded generalized colouring number}.
+	\newblock \emph{Discrete Math.}, 309(18):5562--5568, 2009.
+
+\end{thebibliography}
+
 \end{document}

### 2022-05-17 k-planar.tex
Lemma 27 update

@@
 \label{MewMapGraphLemma}
-For every integer $d\geq 3$, every $d$-map graph is a spanning subgraph of a $d$-framed multigraph.
+For every integer $d\geq 3$, every $d$-map graph is a spanning \note{DW}{delete ``spanning''?} 
+subgraph of a $d$-framed multigraph.
 \end{lem}
@@
 \begin{proof}	
-Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations and let $G$ be the corresponding map graph.
+Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a lake, and where each vertex of $G_0$ is incident with at most $d$ nations and let $G$ be the corresponding map graph.
+
+We may assume that $G_0$ is an edge-maximal multigraph with no faces of length 2 (and with each face labelled a nation or a lake), such that $G$ is the corresponding map graph.
+Suppose that some face $f$ of $G_0$ has a disconnected boundary (implying $G_0$ is disconnected).
+Let $v$ and $w$ be vertices in distinct components of the boundary of $f$.
+Add the edge $vw$ to $G_0$ across $f$.
+The corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$.
+Thus no face of $G_0$ is disconnected.
+Suppose that some face $f$ of $G_0$ has a repeated vertex in the boundary walk of $f$.
+Let $u,v,w$ be consecutive vertices on the boundary of $f$.
+So $u,v,w$ are distinct and $vw$ is not an edge in the boundary of $f$.
+Add the edge $uw$ across $f$ and consider the resulting face $uvw$ to be a lake.
+Since $v$ appears elsewhere in the boundary of $f$, the corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$.
+Thus no facial walk of $G_0$ has a repeated vertex.
+Since each facial walk is connected, every face of $G_0$ is bounded by a cycle, and $G_0$ is biconnected.
+
 Let $G_0^*$ be the dual graph of $G_0$.
 So the vertices of $G_0^*$ correspond to faces of $G_0$, and each vertex of $G_0^*$ is a nation vertex or a lake vertex.
+Since $G_0$ is biconnected, so is $G^*_0$.
 
@@
 Let $G_1$ be a plane supergraph of $G_0^*$ obtained by adding the nation cycle of each face $F$ of $G_0$, and then triangulating each face with more than $d$ vertices on its boundary.
+$G_1$ is biconnected since $G^*_0$ is a biconnected spanning subgraph of $G_1$.
 Each nation cycle of length at least $3$ is now a facial cycle of $G_1$, and no face of $G_1$ has more than $d$ vertices on its boundary.
@@
 
-By definition, $V(G)=V(\widehat{G}_1)$.
+By definition, $V(G)=V(\widehat{G}_1)$ \note{DW}{$V(\widehat{G}_1)=V(G_1)=V(G^*_0)$ which includes vertices corresponding to lakes of $G_0$.
+So I think this claim is wrong (or am I missing something?).
+I suggest we write $V(G) \subseteq V(\widehat{G}_1)$ and drop `spanning' from the lemma}.
 To prove the claim, it suffices to show that $E(G)\subseteq E(\widehat{G}_1)$.

### 2022-05-21 k-planar.tex
fixed typo
added a little more to the map graph proof, and removed `spanning'

@@
 \label{MewMapGraphLemma}
-For every integer $d\geq 3$, every $d$-map graph is a spanning \note{DW}{delete ``spanning''?} 
-subgraph of a $d$-framed multigraph.
+For every integer $d\geq 3$, every $d$-map graph is a subgraph of a $d$-framed multigraph.
 \end{lem}
@@
 
-We may assume that $G_0$ is an edge-maximal multigraph with no faces of length 2 (and with each face labelled a nation or a lake), such that $G$ is the corresponding map graph.
+If $G_0$ has a face $f$ of length 2, then add a new vertex inside $f$ adjacent to both vertices on the boundary of $f$, which creates two new triangular faces $f_1$ and $f_2$.
+If $f$ is a lake, then make $f_1$ and $f_2$ lakes.
+If $F$ is a nation, then make $f_1$ a nation and make $f_2$ a lake.
+The resulting map graph is still $M$.
+So we may assume that $G_0$ is an edge-maximal multigraph with no faces of length 2 (and with each face labelled a nation or a lake), such that $G$ is the corresponding map graph.
 Suppose that some face $f$ of $G_0$ has a disconnected boundary (implying $G_0$ is disconnected).
@@
 Let $u,v,w$ be consecutive vertices on the boundary of $f$.
-So $u,v,w$ are distinct and $vw$ is not an edge in the boundary of $f$.
+So $u,v,w$ are distinct and $uw$ is not an edge in the boundary of $f$.
 Add the edge $uw$ across $f$ and consider the resulting face $uvw$ to be a lake.
@@
 
-By definition, $V(G)=V(\widehat{G}_1)$ \note{DW}{$V(\widehat{G}_1)=V(G_1)=V(G^*_0)$ which includes vertices corresponding to lakes of $G_0$.
-So I think this claim is wrong (or am I missing something?).
-I suggest we write $V(G) \subseteq V(\widehat{G}_1)$ and drop `spanning' from the lemma}.
+By definition, $V(G) \subseteq V(\widehat{G}_1)$.
 To prove the claim, it suffices to show that $E(G)\subseteq E(\widehat{G}_1)$.

### 2022-10-18 k-planar.tex
added a  sentence to the mao graph proof
fixed a bib entry

@@
 If $F$ is a nation, then make $f_1$ a nation and make $f_2$ a lake.
-The resulting map graph is still $M$.
+The resulting map graph is still $G$.
 So we may assume that $G_0$ is an edge-maximal multigraph with no faces of length 2 (and with each face labelled a nation or a lake), such that $G$ is the corresponding map graph.
+This is well-defined since the assumption of having no face of length 2 implies that $|E(G_0|\leq 3|V(G)|-6$.
+
 Suppose that some face $f$ of $G_0$ has a disconnected boundary (implying $G_0$ is disconnected).
@@
 	\textsc{Torsten Ueckerdt, David~R. Wood, and Wendy Yi}.
-	\newblock \href{http://arxiv.org/abs/2108.00198}{An improved planar graph
-		product structure theorem}.
-	\newblock 2021, arXiv:2108.00198.
+	\newblock \href{https://doi.org/10.37236/10614}{An improved planar graph product structure theorem}.
+	\newblock Electronic J. Combinatorics 29:P2.51, 2022.
 

### 2022-10-18 k-planar.tex
created response.* and started doing the referee's revisions

@@
 For at least two applications, the planarity of $H$ in \cref{1_planar_product} is critical in obtaining asymptotically tight bounds \cite{BDJM,DFMS21}.
-\cref{1_planar_product} is proven in the more general setting of $d$-framed multigraphs, which also implies a product structure theorem for $d$-map graphs in which $H$ is a planar with treewidth 3.
+\cref{1_planar_product} is proven in the more general setting of $d$-framed multigraphs, which also implies a product structure theorem for $d$-map graphs in which $H$ is a planar graph with treewidth 3.
 

### 2022-10-18 k-planar.tex
continued doing referee revisions

@@
 The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A $T$-decomposition of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags contain $v_1,\ldots,v_t$.
+To show the planarity of $H$ in \cref{d_framed_product_stucture} we use the fact that a graph has a 3-simple tree-decomposition if and only if it has treewidth at most 3 and is planar~\citep{KV12}.
 Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that for each $v\in V(G)$,  \[\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.\]
@@
 		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
-		\item Let $F$ be a cycle in $G_0$ whose vertices are partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
+		\item Let $F$ be a cycle in $G_0$ whose vertices are partitioned into $k\le 3$ non-empty sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
 		\begin{compactenum}
@@
 We now move onto the case in which $N_0$ contains at least one inner vertex.
-Our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that will be determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N_0$.
+Our $H$-partition $\mathcal{P}$ will include $P_1,\dots,P_k$ and a (possibly empty) part $S$ that will be determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N_0$.
 We first explain how $v_1$, $v_2$ and $v_3$ are chosen and then explain how these are used to define $S$.

### 2022-10-18 response2.tex
continued doing referee revisions

\documentclass[12pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=20mm]{geometry}
\usepackage{amsmath,color,amsthm,enumerate,amsfonts,mathtools}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{amsfonts}
\usepackage{paralist}
\usepackage{stmaryrd}
\usepackage{amssymb}
\usepackage{todonotes}

\newcommand{\done}{\textcolor{blue}{Done.}}
\newcommand{\changed}{\textcolor{blue}{Revised accordingly.}}
\newcommand{\tba}{\textcolor{red}{To be addresssed.}}

\newenvironment{response}{\color{blue}}{}

\setlength{\parindent}{0ex}
\setlength{\parskip}{1ex}

\begin{document}

\subsection*{Graph Product Structure for Non-Minor-Closed Classes\\
Vida Dujmovi\'c, Pat Morin, David R. Wood\\
Journal of Combinatorial Theory Series B (JCTB9758)\\
Response to Referee Report received 4th October 2022}

Thanks to the referee for thoroughly checking the paper, and for highlighting several places for improvement. Here we respond to the referee's comments.

\hrulefill

1. In the proof of Theorem 7, I see no reason why we should have $a = x_r$
(or $x'=x_0$, for that matter). I think you need to use the fact that $v_r$ is
in $S_a$. (Also, if we did have that $a = x_r$, then $j$ would be $r$, so maybe
there is a different mistake, or the intention is different.) It might be
helpful to define $H^+$ first and then have a separate claim which shows
how to use paths in $H$ to get paths in $H^+$.

\tba

2. Are the nations incident to a vertex counted with multiplicity in dmap graphs? The definition does not say so, but the proof of Lemma 27 (as currently stated) assumes not only this, but that the embedded graph under consideration is 2-connected. It also seems that frames are loopless, so slightly more care is called for when they are constructed.
It may be worth noting explicitly that you allow d-framed graphs to
have edges outside the outer face of their frame - I was unsure about
this given the phrase “interior of the face”.

\tba

3. Sperner’s Lemma is usually stated for simple graphs (as is the case
in the source given) so there should be a reduction argument. It is
more than just removing parallel edges because there may be separating
cycles of length two.

\tba

Other comments:

1. There are some proofreading things to do - some lines overflow, there
are citations on a new line (instead of using tilde), and there are typos
like missing words (page 3, line 7, should read “H is a planar graph”).

\begin{response}
	We have fixed ``$H$ is a planar graph'', but do not see the point in fixing minor line overflows, since the paper will be reformatted in the journal style anyway.
\end{response}

2. Page 4, second bullet point. It says that you derive product structure
theorems for string graphs, which is clearly not possible. Please use
more precise language. This comment was in the previous review as
well. Some but not all instances have been corrected.

\begin{response}
We do not understand this comment. Why can't there be a product structure
theorem for string graphs?
\end{response}

3. page 6, line -6: The elements of P are not ordered so it is not a sequence.

\tba

4. Page 7, line 6: It is confusing to say that Sx “is defined by vertices
that participate in x” because it suggests that Sx is the set of vertices
which participate in x. Moreover, for the proven upper bound to be
enough, you need to show that the set of all nodes which participate in
x induces a subtree of T (and thus that a(v) is in this set).

\tba

5. The paper regularly says things like “let P be an H-partition” without
saying what H is. Typically this is understandable, but it could be
written more clearly. For instance, on page 8, line 2, it says “let Yxi
be the part in Y...” without saying what xi
is (and similarly in other
places). This shouldn’t be a serious hang-up for anyone, but still, it
would be better to acknowledge technical inaccuracies.

\tba

6. page 8, Claim 3. The definition of Cx, a subset of V (J), includes things
which may not be vertices of J. This, of course, is not a real problem,
but the proof treats the fact that V (J) may be a proper subset of
V (T) sloppily. Also, the notation T[x] is unfortunately ambiguous - it
is strange to need to define something which has already been defined.

\tba

7. page 16, line 6: Say why that is a convenient way to establish the
planarity of H. (It does not actually say here that any such graph G
is planar.)

\begin{response}
	Added the desired sentence.
\end{response}

8. page 16, line 8. I would not call these “BFS spanning forests rooted
at V0” since they need not be BFS spanning forests (unless the same
terminology has already appeared elsewhere?).

\tba

9. page 16, Lemma 24. The lemma is rather long to state and I would
imagine that anyone reading it has already read the proof of the product
structure theorem for planar graphs. So I would add a sentence or two
beforehand about how the set-up relates to that proof/tripods.

\tba

10. page 16, Lemma 24: Add “non-empty” before “sets $P_1,\dots,P_k$”.

\done

11. page 16, line -2. “$P_1, P_2, P_3$” should be $P_1,\dots,P_k$.

\done

12. page 17, top. In the first two cases there may not be a vertex v3 defined
as we could have p = 2.

\tba

13. page 17, right after 3. The end of Qi should be the same as the vertex w
chosen in case 3 because there may be multiple “appropriate” vertices.
It should also really say what “appropriate” means.

\tba

14. page 17, line -9. “$\{x_1, x_2, x_3\}$” should be “$\{v_1, v_2, v_3\}$”

\tba

15. page 19, second to last paragraph. Add a sentence about why such
a forest exists. The definition of a BFS layering is missing. A reader
may expect that you will just use the assumptions about H and P
that are stated, so it would be better to say something like “This gives
an H-partition... and conditions (i), (ii), and (iii) of Lemma 24 are
satisfied”.

\tba

16. page 20, first paragraph of Lemma 25. Well, parallel edges can appear
on the boundary of a common face if they were there before. Add a
sentence about why the process ends.

\tba

17. page 20, second to last paragraph. Figure 4 is mentioned several paragraphs before the graph which it depicts is defined. The graph $G_0$ in
this paragraph should be $G'$, and check for this throughout.

\tba

18. page 21, Lemma 27. Should be “labelled a nation or a lake” in the
first paragraph. In the third paragraph, the graph $\hat{G}_1$ should be “edge-maximal” or similar.

\tba
\end{document}

### 2022-10-18 k-planar.tex
recompiled the bibliography

@@
 We use $G-S$ as a shorthand for $G[V(G)\setminus S]$.
-Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (with an improvement by \citet{UWY}).
+Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (with an improvement by \citet{UWY22}).
 A graph $G$ is \defin{contained} in a graph $X$ if $G$ is isomorphic to a subgraph of $X$.
 
-\begin{thm}[\citep{DJMMUW20,UWY}]
+\begin{thm}[\citep{DJMMUW20,UWY22}]
 \label{PlanarProduct}
@@
       planar & $4^*$ & $2$ & \cite{bose.morin.ea:optimal} \\
-      planar & $6^*$ & $1$ & \cite{UWY} \\
+      planar & $6^*$ & $1$ & \cite{UWY22} \\
       genus $g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW} \\
-      genus $g$ & $2g+6$ & 1 & \cite{UWY} \\
+      genus $g$ & $2g+6$ & 1 & \cite{UWY22} \\
       apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
@@
 As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a),
-\citet{DJMMUW20} build on an earlier result of \citet{PS21} to show that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \cite{UWY}).
+\citet{DJMMUW20} build on an earlier result of \citet{PS21} to show that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \citep{UWY22}).
 The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.\footnote{\citet{bose.morin.ea:optimal} have recently given linear time algorithms for computing layered $H$-partitions of planar graphs.}
@@
 
-\begin{thm}[\citep{DJMMUW20,UWY,DHHW}]\label{GenusProduct}
+\begin{thm}[\citep{DJMMUW20,UWY22,DHHW}]\label{GenusProduct}
   Every graph of Euler genus $g$ is contained in:
@@
 
-%\bibliographystyle{DavidNatbibStyle}
-%\bibliography{k-planar}
-
-\begin{thebibliography}{49}
-	\providecommand{\natexlab}[1]{#1}
-	\providecommand{\msn}[1]{MR:\,\href{http://www.ams.org/mathscinet-getitem?mr=MR{#1}}{#1}}
-	\providecommand{\ZBL}[1]{Zbl:\,\href{https://www.zentralblatt-math.org/zmath/en/search/?q=an:#1}{#1}}
-	\providecommand{\url}[1]{\texttt{#1}}
-	\providecommand{\urlprefix}{}
-	\expandafter\ifx\csname urlstyle\endcsname\relax
-	\providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
-	\providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
-		\urlstyle{rm}\Url}\fi
-
-	\bibitem[{{\'{A}}brego et~al.(2011){\'{A}}brego, Monroy,
-		Fern{\'{a}}ndez{-}Merchant, Flores{-}Pe{\~{n}}aloza, Hurtado,
-		Sacrist{\'{a}}n, and Saumell}]{abrego.munroy.ea:on}
-	\textsc{Bernardo~M. {\'{A}}brego, Ruy~Fabila Monroy, Silvia
-		Fern{\'{a}}ndez{-}Merchant, David Flores{-}Pe{\~{n}}aloza, Ferran Hurtado,
-		Vera Sacrist{\'{a}}n, and Maria Saumell}.
-	\newblock \href{https://doi.org/10.1016/j.comgeo.2010.11.003}{On crossing
-		numbers of geometric proximity graphs}.
-	\newblock \emph{Comput. Geom.}, 44(4):216--233, 2011.
-
-	\bibitem[{Aigner and Ziegler(2010)}]{Proofs4}
-	\textsc{Martin Aigner and G{\"u}nter~M. Ziegler}.
-	\newblock Proofs from {T}he {B}ook.
-	\newblock Springer, 4th edn., 2010.
-
-	\bibitem[{Ajtai et~al.(1982)Ajtai, Chv{\'a}tal, Newborn, and
-		Szemer{\'e}di}]{ajtai.chvatal.ea:crossing-free}
-	\textsc{Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, and Endre
-		Szemer{\'e}di}.
-	\newblock \href{https://doi.org/10.1016/S0304-0208(08)73484-4}{Crossing-free
-		subgraphs}.
-	\newblock \emph{Ann. Discrete Math.}, 12:9--12, 1982.
-
-	\bibitem[{Alam et~al.(2020)Alam, Bekos, Gronemann, Kaufmann, and
-		Pupyrev}]{ABGKP20}
-	\textsc{Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael
-		Kaufmann, and Sergey Pupyrev}.
-	\newblock \href{https://doi.org/10.1007/s00453-020-00697-4}{Queue layouts of
-		planar 3-trees}.
-	\newblock \emph{Algorithmica}, 82:2564--2585, 2020.
-
-	\bibitem[{Alon et~al.(2002)Alon, Grytczuk, Ha{\l}uszczak, and
-		Riordan}]{AGHR-RSA02}
-	\textsc{Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, and Oliver
-		Riordan}.
-	\newblock \href{https://doi.org/10.1002/rsa.10057}{Nonrepetitive colorings of
-		graphs}.
-	\newblock \emph{Random Structures Algorithms}, 21(3-4):336--346, 2002.
-
-	\bibitem[{Babai et~al.(1982)Babai, Chung, Erd\H{o}s, Graham, and
-		Spencer}]{babai.chung.ea:on}
-	\textsc{Laszlo Babai, F.~R.~K. Chung, Paul Erd\H{o}s, Ronald~L. Graham, and
-		J.~H. Spencer}.
-	\newblock On graphs which contain all sparse graphs.
-	\newblock In \emph{Theory and practice of combinatorics. A collection of
-		articles honoring Anton Kotzig on the occasion of his sixtieth birthday}, pp.
-	21--26. Elsevier, 1982.
-
-	\bibitem[{Bekos et~al.(2020)Bekos, Lozzo, Griesbach, Gronemann, Montecchiani,
-		and Raftopoulou}]{BDGGMR}
-	\textsc{Michael~A. Bekos, Giordano~Da Lozzo, Svenja Griesbach, Martin
-		Gronemann, Fabrizio Montecchiani, and Chrysanthi~N. Raftopoulou}.
-	\newblock \href{https://doi.org/10.4230/LIPIcs.SoCG.2020.16}{Book embeddings of
-		nonplanar graphs with small faces in few pages}.
-	\newblock In \textsc{Sergio Cabello and Danny~Z. Chen}, eds., \emph{36th
-		International Symposium on Computational Geometry, SoCG 2020, June 23-26,
-		2020, Z{\"{u}}rich, Switzerland}, vol. 164 of \emph{LIPIcs}, pp. 16:1--16:17.
-	Schloss Dagstuhl - Leibniz-Zentrum f{\"{u}}r Informatik, 2020.
-	\newblock arXiv:2003.07655.
-
-	\bibitem[{Bodlaender(1998)}]{Bodlaender-TCS98}
-	\textsc{Hans~L. Bodlaender}.
-	\newblock \href{https://doi.org/10.1016/S0304-3975(97)00228-4}{A partial
-		$k$-arboretum of graphs with bounded treewidth}.
-	\newblock \emph{Theoret. Comput. Sci.}, 209(1-2):1--45, 1998.
-
-	\bibitem[{Bonnet et~al.(2022)Bonnet, Kwon, and Wood}]{BKW}
-	\textsc{\'Edouard Bonnet, {O-joung} Kwon, and David~R. Wood}.
-	\newblock \href{http://arxiv.org/abs/2202.11858}{Reduced bandwidth: a
-		qualitative strengthening of twin-width in minor-closed classes (and
-		beyond)}.
-	\newblock 2022, arXiv:2202.11858.
-
-	\bibitem[{Bose et~al.(2012)Bose, Dujmovic, Hurtado, Iacono, Langerman, Meijer,
-		Adinolfi, Saumell, and Wood}]{ProximityGraphs}
-	\textsc{Prosenjit Bose, Vida Dujmovic, Ferran Hurtado, John Iacono, Stefan
-		Langerman, Henk Meijer, Vera~Sacrist{\'{a}}n Adinolfi, Maria Saumell, and
-		David~R. Wood}.
-	\newblock \href{https://doi.org/10.1142/S0218195912500112}{Proximity graphs: E,
-		{\(\delta\)}, {\(\Delta\)}, {\(\chi\)} and {\(\omega\)}}.
-	\newblock \emph{Int. J. Comput. Geometry Appl.}, 22(5):439--470, 2012.
-
-	\bibitem[{Bose et~al.(2020)Bose, Dujmovi\'c, Javarsineh, and Morin}]{BDJM}
-	\textsc{Prosenjit Bose, Vida Dujmovi\'c, Mehrnoosh Javarsineh, and Pat Morin}.
-	\newblock \href{http://arxiv.org/abs/2007.06455}{Asymptotically optimal vertex
-		ranking of planar graphs}.
-	\newblock 2020, arXiv:2007.06455.
-
-	\bibitem[{Bose et~al.(2022)Bose, Morin, and Odak}]{bose.morin.ea:optimal}
-	\textsc{Prosenjit Bose, Pat Morin, and Saeed Odak}.
-	\newblock \href{http://arxiv.org/abs/2202.08870}{An optimal algorithm for
-		product structure in planar graphs}.
-	\newblock 2022, arXiv:2202.08870.
-
-	\bibitem[{Bose et~al.(2001)Bose, Morin, Stojmenović, and
-		Urrutia}]{bose.morin.ea:routing}
-	\textsc{Prosenjit Bose, Pat Morin, Ivan Stojmenović, and Jorge Urrutia}.
-	\newblock \href{https://doi.org/10.1023/A:1012319418150}{Routing with
-		guaranteed delivery in ad hoc wireless networks}.
-	\newblock \emph{Wireless Networks}, 7(6):609--616, 2001.
-
-	\bibitem[{Brandenburg(2019)}]{Brandenburg19}
-	\textsc{Franz~J. Brandenburg}.
-	\newblock \href{https://doi.org/10.1007/s00453-018-0510-x}{Characterizing and
-		recognizing 4-map graphs}.
-	\newblock \emph{Algorithmica}, 81(5):1818--1843, 2019.
-
-	\bibitem[{Brandenburg(2020)}]{Brandenburg20}
-	\textsc{Franz~J. Brandenburg}.
-	\newblock \href{http://arxiv.org/abs/2012.06874}{Book embeddings of $k$-map
-		graphs}.
-	\newblock 2020, arXiv:2012.06874.
-
-	\bibitem[{Chen et~al.(2002)Chen, Grigni, and Papadimitriou}]{CGP02}
-	\textsc{Zhi-Zhong Chen, Michelangelo Grigni, and Christos~H. Papadimitriou}.
-	\newblock \href{https://doi.org/10.1145/506147.506148}{Map graphs}.
-	\newblock \emph{J. ACM}, 49(2):127--138, 2002.
-
-	\bibitem[{Chen et~al.(2006)Chen, Grigni, and Papadimitriou}]{CGP06}
-	\textsc{Zhi-Zhong Chen, Michelangelo Grigni, and Christos~H. Papadimitriou}.
-	\newblock \href{https://doi.org/10.1007/s00453-005-1184-8}{Recognizing
-		hole-free 4-map graphs in cubic time}.
-	\newblock \emph{Algorithmica}, 45(2):227--262, 2006.
-
-	\bibitem[{Diestel(2018)}]{Diestel5}
-	\textsc{Reinhard Diestel}.
-	\newblock \href{http://diestel-graph-theory.com/}{Graph theory}, vol. 173 of
-	\emph{Graduate Texts in Mathematics}.
-	\newblock Springer, 5th edn., 2018.
-
-	\bibitem[{Distel et~al.(2021)Distel, Hickingbotham, Huynh, and Wood}]{DHHW}
-	\textsc{Marc Distel, Robert Hickingbotham, Tony Huynh, and David~R. Wood}.
-	\newblock \href{http://arxiv.org/abs/2112.10025}{Improved product structure for
-		graphs on surfaces}.
-	\newblock 2021, arXiv:2112.10025.
-
-	\bibitem[{D\k{e}bski et~al.(2021)D\k{e}bski, Felsner, Micek, and
-		Schr\"{o}der}]{DFMS21}
-	\textsc{Micha{\l} D\k{e}bski, Stefan Felsner, Piotr Micek, and Felix
-		Schr\"{o}der}.
-	\newblock \href{https://doi.org/10.19086/aic.27351}{Improved bounds for
-		centered colorings}.
-	\newblock \emph{Adv. Comb.}, \#8, 2021.
-
-	\bibitem[{Dujmovi\'c et~al.(2017{\natexlab{a}})Dujmovi\'c, Eppstein, and
-		Wood}]{dujmovic.eppstein.ea:structure}
-	\textsc{Vida Dujmovi\'c, David Eppstein, and David~R. Wood}.
-	\newblock \href{https://doi.org/10.1137/16M1062879}{Structure of graphs with
-		locally restricted crossings}.
-	\newblock \emph{{SIAM} J. Discrete Math.}, 31(2):805--824, 2017{\natexlab{a}}.
-
-	\bibitem[{Dujmovi\'c et~al.(2021)Dujmovi\'c, Esperet, Gavoille, Joret, Micek,
-		and Morin}]{DEJGMM21}
-	\textsc{Vida Dujmovi\'c, Louis Esperet, Cyril Gavoille, Gwena\"el Joret, Piotr
-		Micek, and Pat Morin}.
-	\newblock \href{https://doi.org/10.1145/3477542}{Adjacency labelling for planar
-		graphs (and beyond)}.
-	\newblock \emph{J. ACM}, 68(6):42, 2021.
-
-	\bibitem[{Dujmovi\'c et~al.(2020)Dujmovi\'c, Esperet, Joret, Walczak, and
-		Wood}]{dujmovic.esperet.ea:planar}
-	\textsc{Vida Dujmovi\'c, Louis Esperet, Gwena{\"{e}}l Joret, Bartosz Walczak,
-		and David~R. Wood}.
-	\newblock \href{https://doi.org/10.19086/aic.12100}{Planar graphs have bounded
-		nonrepetitive chromatic number}.
-	\newblock \emph{Adv. Comb.}, \#5, 2020.
-
-	\bibitem[{Dujmovi{\'c} et~al.(2022)Dujmovi{\'c}, Esperet, Morin, Walczak, and
-		Wood}]{DEMWW22}
-	\textsc{Vida Dujmovi{\'c}, Louis Esperet, Pat Morin, Bartosz Walczak, and
-		David~R. Wood}.
-	\newblock \href{https://doi.org/10.1017/S0963548321000213}{Clustered
-		3-colouring graphs of bounded degree}.
-	\newblock \emph{Combin. Probab. Comput.}, 31(1):123--135, 2022.
-
-	\bibitem[{Dujmovi\'{c} et~al.(2016)Dujmovi\'{c}, Joret, Kozik, and
-		Wood}]{DJKW16}
-	\textsc{Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, and David~R. Wood}.
-	\newblock \href{https://doi.org/10.1007/s00493-015-3070-6}{Nonrepetitive
-		colouring via entropy compression}.
-	\newblock \emph{Combinatorica}, 36(6):661--686, 2016.
-
-	\bibitem[{Dujmovi{\'c} et~al.(2020)Dujmovi{\'c}, Joret, Micek, Morin, Ueckerdt,
-		and Wood}]{DJMMUW20}
-	\textsc{Vida Dujmovi{\'c}, Gwena\"{e}l Joret, Piotr Micek, Pat Morin, Torsten
-		Ueckerdt, and David~R. Wood}.
-	\newblock \href{https://doi.org/10.1145/3385731}{Planar graphs have bounded
-		queue-number}.
-	\newblock \emph{J. ACM}, 67(4):\#22, 2020.
-
-	\bibitem[{Dujmovi\'c et~al.(2022)Dujmovi\'c, Morin, and Odak}]{DMO}
-	\textsc{Vida Dujmovi\'c, Pat Morin, and Saeed Odak}.
-	\newblock \href{http://arxiv.org/abs/2202.12882}{Odd colourings of graph
-		products}.
-	\newblock 2022, arXiv:2202.12882.
-
-	\bibitem[{Dujmovi\'c et~al.(2017{\natexlab{b}})Dujmovi\'c, Morin, and
-		Wood}]{dujmovic.morin.ea:layered}
-	\textsc{Vida Dujmovi\'c, Pat Morin, and David~R. Wood}.
-	\newblock \href{https://doi.org/10.1016/j.jctb.2017.05.006}{Layered separators
-		in minor-closed graph classes with applications}.
-	\newblock \emph{J. Comb. Theory, Ser. {B}}, 127:111--147, 2017{\natexlab{b}}.
-
-	\bibitem[{Dvo{\v{r}}{\'a}k et~al.(2021)Dvo{\v{r}}{\'a}k, Huynh, Joret, Liu, and
-		Wood}]{DHJLW21}
-	\textsc{Zden{\v{e}}k Dvo{\v{r}}{\'a}k, Tony Huynh, Gwena\"el Joret, Chun-Hung
-		Liu, and David~R. Wood}.
-	\newblock \href{https://doi.org/10.1007/978-3-030-62497-2_32}{Notes on graph
-		product structure theory}.
-	\newblock In \textsc{David~R. Wood, Jan de~Gier, Cheryl~E. Praeger, and Terence
-		Tao}, eds., \emph{2019-20 MATRIX Annals}, pp. 513--533. Springer, 2021.
-	\newblock arXiv:2001.08860.
-
-	\bibitem[{Elmallah and Colbourn(1990)}]{elmallah.colbourn:on}
-	\textsc{Ehab~S. Elmallah and Charles~J. Colbourn}.
-	\newblock \href{https://doi.org/10.1016/0012-365X(90)90293-Q}{On two dual
-		classes of planar graphs}.
-	\newblock \emph{Discrete. Math.}, 80(1):21--40, 1990.
-
-	\bibitem[{Esperet et~al.(2020)Esperet, Joret, and Morin}]{EJM}
-	\textsc{Louis Esperet, Gwena\"{e}l Joret, and Pat Morin}.
-	\newblock \href{http://arxiv.org/abs/2010.05779}{Sparse universal graphs for
-		planarity}.
-	\newblock 2020, arXiv:2010.05779.
-
-	\bibitem[{Fomin et~al.(2012)Fomin, Lokshtanov, and Saurabh}]{FLS-SODA12}
-	\textsc{Fedor~V. Fomin, Daniel Lokshtanov, and Saket Saurabh}.
-	\newblock \href{https://doi.org/10.1137/1.9781611973099.124}{Bidimensionality
-		and geometric graphs}.
-	\newblock In \emph{Proc. 23rd {A}nnual {ACM}-{SIAM} {S}ymposium on {D}iscrete
-		{A}lgorithms}, pp. 1563--1575. 2012.
-
-	\bibitem[{Fox and Pach(2010)}]{FP10}
-	\textsc{Jacob Fox and J{\'a}nos Pach}.
-	\newblock \href{https://doi.org/10.1017/S0963548309990459}{A separator theorem
-		for string graphs and its applications}.
-	\newblock \emph{Combin. Probab. Comput.}, 19(3):371--390, 2010.
-
-	\bibitem[{Fox and Pach(2014)}]{FP14}
-	\textsc{Jacob Fox and J{\'a}nos Pach}.
-	\newblock \href{https://doi.org/10.1017/S0963548313000412}{Applications of a
-		new separator theorem for string graphs}.
-	\newblock \emph{Combin. Probab. Comput.}, 23(1):66--74, 2014.
-
-	\bibitem[{Harvey and Wood(2017)}]{HW17}
-	\textsc{Daniel~J. Harvey and David~R. Wood}.
-	\newblock \href{https://doi.org/10.1002/jgt.22030}{Parameters tied to
-		treewidth}.
-	\newblock \emph{J. Graph Theory}, 84(4):364--385, 2017.
-
-	\bibitem[{Heath et~al.(1992)Heath, Leighton, and Rosenberg}]{HLR92}
-	\textsc{Lenwood~S. Heath, F.~Thomson Leighton, and Arnold~L. Rosenberg}.
-	\newblock \href{https://doi.org/10.1137/0405031}{Comparing queues and stacks as
-		mechanisms for laying out graphs}.
-	\newblock \emph{SIAM J. Discrete Math.}, 5(3):398--412, 1992.
-
-\bibitem[{van~den Heuvel and Wood(2018)}]{vdHW18}
-\textsc{Jan van~den Heuvel and David~R. Wood}.
-\newblock \href{https://doi.org/10.1112/jlms.12127}{Improper colourings
-	inspired by {H}adwiger's conjecture}.
-\newblock \emph{J. London Math. Soc.}, 98:129--148, 2018.
-\newblock arXiv:1704.06536.
-
-	\bibitem[{Kannan et~al.(1992)Kannan, Naor, and
-		Rudich}]{kannan.naor.ea:implicit}
-	\textsc{Sampath Kannan, Moni Naor, and Steven Rudich}.
-	\newblock \href{https://doi.org/10.1137/0405049}{Implicit representation of
-		graphs}.
-	\newblock \emph{{SIAM} J. Discrete Math.}, 5(4):596--603, 1992.
-
-	\bibitem[{Knauer and Ueckerdt(2012)}]{knauer.ueckerdt:simple}
-	\textsc{Kolja Knauer and Torsten Ueckerdt}.
-	\newblock
-	\href{https://kam.mff.cuni.cz/workshops/mcw/work18/mcw2012booklet.pdf}{Simple
-		treewidth}.
-	\newblock In \textsc{Pavel Rytír}, ed., \emph{Midsummer Combinatorial Workshop
-		Prague}. 2012.
-
-	\bibitem[{Kobourov et~al.(2017)Kobourov, Liotta, and
-		Montecchiani}]{kobourov.liotta.ea:annotated}
-	\textsc{Stephen~G. Kobourov, Giuseppe Liotta, and Fabrizio Montecchiani}.
-	\newblock \href{https://doi.org/10.1016/j.cosrev.2017.06.002}{An annotated
-		bibliography on 1-planarity}.
-	\newblock \emph{Comput. Sci. Rev.}, 25:49--67, 2017.
-
-	\bibitem[{Kratochv{\'{\i}}l and Vaner(2012)}]{kratochvil.vaner:note}
-	\textsc{Jan Kratochv{\'{\i}}l and Michal Vaner}.
-	\newblock \href{http://arxiv.org/abs/1210.8113}{A note on planar partial
-		3-trees}.
-	\newblock 2012, arXiv:1210.8113.
-
-	\bibitem[{Leighton(1983)}]{leighton:complexity}
-	\textsc{F.~Thomson Leighton}.
-	\newblock Complexity issues in {VLSI}.
-	\newblock MIT Press, Cambridge, MA, 1983.
-
-	\bibitem[{Mohar and Thomassen(2001)}]{mohar.thomassen:graphs}
-	\textsc{Bojan Mohar and Carsten Thomassen}.
-	\newblock Graphs on surfaces.
-	\newblock Johns Hopkins University Press, 2001.
-
-	\bibitem[{Ne{\v{s}}et{\v{r}}il and Ossona~de Mendez(2012)}]{Sparsity}
-	\textsc{Jaroslav Ne{\v{s}}et{\v{r}}il and Patrice Ossona~de Mendez}.
-	\newblock \href{https://doi.org/10.1007/978-3-642-27875-4}{Sparsity}, vol.~28
-	of \emph{Algorithms and Combinatorics}.
-	\newblock Springer, 2012.
-
-	\bibitem[{Pach and T{\'o}th(2002)}]{PachToth-DCG02}
-	\textsc{J{\'a}nos Pach and G{\'e}za T{\'o}th}.
-	\newblock \href{https://doi.org/10.1007/s00454-002-2891-4}{Recognizing string
-		graphs is decidable}.
-	\newblock \emph{Discrete Comput. Geom.}, 28(4):593--606, 2002.
-
-	\bibitem[{Pilipczuk and Siebertz(2021)}]{PS21}
-	\textsc{Micha{\l} Pilipczuk and Sebastian Siebertz}.
-	\newblock \href{https://doi.org/10.1016/j.jctb.2021.06.002}{Polynomial bounds
-		for centered colorings on proper minor-closed graph classes}.
-	\newblock \emph{J. Comb. Theory, Series B}, 151:111--147, 2021.
-
-	\bibitem[{Reed(2003)}]{Reed03}
-	\textsc{Bruce~A. Reed}.
-	\newblock \href{https://doi.org/10.1007/0-387-22444-0_4}{Algorithmic aspects of
-		tree width}.
-	\newblock In \emph{Recent advances in algorithms and combinatorics}, vol.~11,
-	pp. 85--107. Springer, 2003.
-
-	\bibitem[{Ueckerdt et~al.(2021)Ueckerdt, Wood, and Yi}]{UWY}
-	\textsc{Torsten Ueckerdt, David~R. Wood, and Wendy Yi}.
-	\newblock \href{https://doi.org/10.37236/10614}{An improved planar graph product structure theorem}.
-	\newblock Electronic J. Combinatorics 29:P2.51, 2022.
-
-	\bibitem[{Zhu(2009)}]{zhu:colouring}
-	\textsc{Xuding Zhu}.
-	\newblock \href{https://doi.org/10.1016/j.disc.2008.03.024}{Colouring graphs
-		with bounded generalized colouring number}.
-	\newblock \emph{Discrete Math.}, 309(18):5562--5568, 2009.
-
-\end{thebibliography}
-
+\bibliographystyle{DavidNatbibStyle}
+\bibliography{k-planar}
 \end{document}

### 2022-11-03 k-planar.tex
One small fix

@@
 For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
-Since $a:V(G)\to V(T)$ is a function with domain $V(G)$ and range $V(T)$, $\PP:=(S_x : x\in V(T))$ is a sequence of pairwise disjoint sets that cover $V(G)$.
+Since $a:V(G)\to V(T)$ is a function with domain $V(G)$ and range $V(T)$, $\PP:=\{S_x : x\in V(T)\}$ is a partition of $V(G)$.
 Let $J:=G^\SS/\PP$ denote the resulting quotient graph.
@@
 For each node $x$ of $T$, we define $C_x$ as follows: $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
-  \cref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, both $a$ and $x$ are contained in $C_{x'}$.
+  \cref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, both $a$ and $x$ are contained in $C_{x'}$\todo{Do we mean $C_x$ here?}.
 The connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ follows from the fact that, for every node $x'\in V(T[a])$, every node $x$ on the path in $T$ from $a$ to  $x'$ is also a node of $T[a]$.

### 2022-11-03 k-planar.tex
Import typo fix, x -> x'
Typo fix

@@
   Suppose, without loss of generality, that $v_0$ participates in $x_i$.
-Then $a(v_0)=x$ is a $T$-ancestor of $x_i$ which a $T$-ancestor of $x_r$.
+Then $a(v_0)=x$ is a $T$-ancestor of $x_i$ which is a $T$-ancestor of $x_r$.
 Finally $a(v_r)=y$ is a $T$-ancestor of $x_r$.
@@
 For each node $x$ of $T$, we define $C_x$ as follows: $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
-  \cref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, both $a$ and $x$ are contained in $C_{x'}$\todo{Do we mean $C_x$ here?}.
+  \cref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, both $a$ and $x'$ are contained in $C_{x'}$.
 The connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ follows from the fact that, for every node $x'\in V(T[a])$, every node $x$ on the path in $T$ from $a$ to  $x'$ is also a node of $T[a]$.

### 2022-11-04 k-planar.tex
Another typo fix
Fixed big hole in proof of Theorem 7

@@
 
+\begin{clm}\label{short_path_in_h}
+  For each $v\in V(G)$, $H$ contains a path $x_0,\ldots,x_r$ of length $r\le k-1$ with $v\in Y_{x_0}$ and $x_r:=a(v)$ such that $v$ participates in each of $x_0,\ldots,x_r$.
+\end{clm}
+
+\begin{proof}
+  Let $x_0\in V(T)$ be the unique node with $v\in Y_{x_0}$ and let $x_r:=a(v)$.
+If $x_0=x_r$ then the length-$0$ path with the single vertex $x_0=x_r$ satisfies the requirements of the lemma.
+Assume now that $x_0\neq x_r$.
+
+  Consider the subgraph $H_v:=H[x\in V(T):\text{$v$ participates in $x$}\}]$.
+Since $v\in Y_{x_0}$, $v$ participates in $x_0$, so $x_0\in V(H_v)$.
+We claim that, for each $x\in V(H_v)\setminus\{x_0\}$, $H_v$ contains a path $y_0,\ldots,y_s$ of length $s\le k-1$ from $y_0:=x_0$ to $y_s:=x$.
+Indeed, since $v$ participates in $x\neq x_0$, $\mathcal{S}$ contains a shortcut $P$ with $v\in V(P)$ and that crosses $x$.
+Therefore $P$ contains a path $v_0,\ldots,v_s$ from $v_0:=v$ to some $v_s\in Y_x$.
+Furthermore $v_s$ is an internal vertex of $P$ and $P$ has length at most $k$, so $s\le k-1$.
+For each $i\in\{0,\ldots,s\}$, let $Y_{y_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.
+For each $i\in\{1,\ldots,s\}$, the existence of $v_{i-1}v_{i}\in E(G)$ implies that $y_{i-1}=y_i$ or $y_{i-1}y_i\in E(H)$.
+  Therefore $y_0,\ldots,y_s$ is a lazy walk\footnote{A \defin{lazy walk} in a graph $G$ is a walk in the pseudograph obtained by adding a self-loop at each vertex of $G$.} in $H_v$ of length at most $s\le k-1$.
+This lazy walk contains a path from $y_0$ to $y_s$ of length at most $s\le k-1$, as promised.
+
+  Since $H_v$ contains a path from $x_0$ to $x$ for each $x\in V(H_v)\setminus\{x_0\}$,  $H_v$ is connected.
+Since $H_v$ is connected, $x_r$---which is the lowest common $T$-ancestor of $V(H_v)$---is itself a node of $V(H_v)$.
+Therefore $H_v$ contains a path $x_0,\ldots,x_r$ of length most $k-1$.
+By definition, $v$ participates in each of $x_0,\ldots,x_r$.
+\end{proof}
+
 For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
@@
 Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so at least one of $x$ or $y$ is a $T$-ancestor of the other.
-This completes the proof of \cref{i-ancestor}
+This completes the proof of \cref{i-ancestor}.
 \end{proof}
@@
 For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.
-For each $i\in\{1,\ldots,r-1\}$, $v_i\in Y_{x_i}$ is an internal vertex of $P$, so $v_r$ participates in $x_i$.
-Therefore $a=a(v_r)$ is a $T$-ancestor of $x_i$ for each $i\in\{1,\ldots,r-1\}$.
-
-  Since $v_0,\ldots,v_r$ is a path in $G$, $x_0,\ldots,x_r$ is a lazy walk in $H$.\footnote{A \defin{lazy walk} is a graph $G$ is a walk in the pseudograph obtained by adding a self-loop at each vertex of $G$.}  Let $j$ be the minimum integer such that $x_j$ is a strict $T$-ancestor of $x$.
-Since $x'=x_0$ is a $T$-descendant of $x$ and $a=x_r$ is a strict $T$-ancestor of $x$, $j\in\{1,\ldots,r\}$ and $x_{j-1}$ is a $T$-descendant of $x$ (possibly $x_{j-1}=x$).
-Since $x_{j-1}x_j\in E(H)$, $x_j \in B_{x_{j-1}}$ and therefore $x_j\in B_x$.
-Therefore $xx_j\in E(H^+)$.
-
-  Therefore $W:=y_0,\ldots,y_s:=x,x_j,x_{j+1},\ldots,x_r$ is a lazy walk in $H^+$ of length $r\le k$.
-For each $j\in \{0,\ldots,s\}$, let $z_j$ be the lowest common $T$-ancestor of $y_0,\ldots,y_j$.
-The same reasoning used to conclude that $xx_j\in E(H^+)$ can be used to conclude that $z_{i-1}=z_i$ or $z_{i-1}z_i\in E(H^+)$ for each $i\in\{1,\ldots,s\}$.
-Therefore $z_0,\ldots,z_s$ is a directed lazy walk in $\overrightarrow{H}^+$.
-Finally, let $\overrightarrow{W}$ be the path in $\overrightarrow{H}^+$ obtained by removing removing duplicates from $z_0,\ldots,z_s$.
-Then $\overrightarrow{W}$ is a directed path in $\overrightarrow{H}^+$ from $x$ to $a$ of length at most $s\le r\le k$.
+Since $x_r\in S_{a}$, $a(v_r)=a$.
+By \cref{short_path_in_h}, $H$ contains a path $y_0,\ldots,y_s$ from $y_0:=x_r$ to $y_s:=a$ of length at most $k-1$ such that $v_r$ participates in each of $y_0,\ldots,y_s$.
+We will use this to show that $H$ contains a path $z_0,\ldots,z_{s'}$ from $z_0:=x$ to $z_{s'}:=a$ such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.
+There are three cases to consider, depending on the location of $x_r$ relative to $x'$.
+
+  \begin{enumerate}
+    \item $x_r$ is a strict $T$-ancestor of $x$.
+The shortcut $P=x_0,\ldots,x_r$ implies that $x_0$ participates in $x_{r-1}$.
+Therefore $x'=a(x_0)$ is a $T$-ancestor of $x_{r-1}$, so $x$ is also a $T$-ancestor of $x_{r-1}$.
+Since $x_{r-1}$ is a $T$-descendant of $x$ and $x_{r-1}x_r\in E(H)$, $x_r=y_0$ is contained in $B_x$.
+Therefore $xy_0$ is an edge of $H^+$, so the path $x,y_0,\ldots,y_s$ is a path in $H^+$ from $x$ to $a$ of length at most $s+1\le k$ and $v_r$ participates in each of $y_0,\ldots,y_s$.
+
+    \item $x_r=x$.  In this case, $y_0,\ldots,y_s$ is a path from $x$ to $a$ of length at most $k-1$ and $v_r$ participates in each of $y_0,\ldots,y_s$.
+
+    \item $x_r$ is a strict $T$-descendant of $x$.
+      Since  $y_0$ is a strict $T$-descendant of $x$ and $y_s$ is a $T$-ancestor of $x$, at least one of $y_0,\ldots,y_s$, say $y_i$, is contained in $B_x$.
+Therefore $xy_i$ is an edge of $H^{+}$.
+Therefore $x,y_i,\ldots,y_s$ is a path in $H^+$ from $x$ to $a$ of length at most $s+1\le k$ and $v_r$ participates in each of $y_i,\ldots,y_s$.
+
+  \end{enumerate}
+  In each case, we obtain a path $z_0,\ldots,z_{s'}$ of length at most $k$ from $z_0:=x$ to $z_{s'}:=a$ such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.
+By definition, $a$ is a $T$-ancestor of $x$.
+Since $v_r$ participates in each of $z_1,\ldots,z_{s'}$, $a=a(v_r)$ is also a $T$-ancestor of each of $z_1,\ldots,z_{s'}$.
+Let $W$ be the sequence obtained from $z_0,\ldots,z_{s'}$ by replacing $z_i$ with the lowest common $T$-ancestor of $z_0,\ldots,z_i$, for each $i\in\{0,\ldots,s'\}$.
+Then $W$ is a lazy walk in $H^+$ of length at most $k$ that begins at $z_0=x$ and ends at $z_{s'}=a$.
+By removing duplicates from $W$ we obtain a path in $H^+$ that is also a directed path in $\overrightarrow{H}^+$.
 This completes the proof of \cref{general-bag-size}.

### 2022-11-04 response2.tex
One small fix
Fixed big hole in proof of Theorem 7

@@
 
-\tba
+\begin{response}
+  Thanks to the referee for catching this embarrassing mistake.
+In our previous round of simplifications, we oversimplified here.
+The existence of the directed path in $\overrightarrow{H}^+$ does not come from the shortcut $v_0,\ldots,v_r$.
+Rather, it comes from some shortcut that contains $v_r$ and some vertex in $Y_a$ (which must exist since $a=a(v_r)$ so $v_r\in Y_a$ or some shortcut that contains $v_r$ crosses $a$).
+We have rewritten this part accordingly.
+\end{response}
 
@@
 
-\tba
+\begin{response}
+  We have replaced \\
+  ``$P := (S_x : x \in V (T))$ is a sequence of pairwise disjoint sets that cover $V(G)$'' \\
+  with \\
+  ``$\mathcal{P}:=\{S_x : x\in V(T)\}$ is a partition of $V(G)$.''
+\end{response}
 

### 2022-11-04 k-planar.tex
Some little cleanups in the proof of Theorem 7

@@
 If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.
-Subdivide the edge between $a$ and its parent in $T_0$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
+Subdivide the edge between $a$ and its $T$-parent by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.
 Now $f(y)=a'$ and $f(x)=a$, so this modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$.
@@
 In other words, if $P=v_0,\ldots,v_r$ and $\{v_1,\ldots,v_{r-1}\}\cap Y_x\neq\emptyset$.
-We say that a vertex $v\in V(G)$ \defin{participates} in $x$ if $v\in Y_x$ or if $\SS$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
+We say that a vertex $v\in V(G)$ \defin{participates} in $x\in V(T)$ if $v\in Y_x$ or if $\SS$ contains a shortcut $P$ with $v\in V(P)$ and $P$ crosses $x$.
 For each $v\in V(G)$, let $a(v)$ denote the lowest common $T$-ancestor of all the nodes $x\in V(T)$ in which $v$ participates.
@@
 
-  Consider the subgraph $H_v:=H[x\in V(T):\text{$v$ participates in $x$}\}]$.
+  Consider the subgraph $H_v:=H[\{x\in V(T):\text{$v$ participates in $x$}\}]$.
 Since $v\in Y_{x_0}$, $v$ participates in $x_0$, so $x_0\in V(H_v)$.
@@
 Therefore $H_v$ contains a path $x_0,\ldots,x_r$ of length most $k-1$.
-By definition, $v$ participates in each of $x_0,\ldots,x_r$.
+By the definition of $H_v$, $v$ participates in each of $x_0,\ldots,x_r$.
 \end{proof}
@@
 For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
-Since $a:V(G)\to V(T)$ is a function with domain $V(G)$ and range $V(T)$, $\PP:=\{S_x : x\in V(T)\}$ is a partition of $V(G)$.
+Since $a:V(G)\to V(T)$ is a function with domain $V(G)$ and range $V(T)$, $\PP:=(S_x : x\in V(T))$ is a collection of disjoint sets that cover $V(G)$.
 Let $J:=G^\SS/\PP$ denote the resulting quotient graph.
@@
     \item $x_r$ is a strict $T$-ancestor of $x$.
-The shortcut $P=x_0,\ldots,x_r$ implies that $x_0$ participates in $x_{r-1}$.
-Therefore $x'=a(x_0)$ is a $T$-ancestor of $x_{r-1}$, so $x$ is also a $T$-ancestor of $x_{r-1}$.
+The shortcut $P=v_0,\ldots,v_r$ implies that $v_0$ participates in $x_{r-1}$.
+Therefore $x'=a(v_0)$ is a $T$-ancestor of $x_{r-1}$, so $x$ is also a $T$-ancestor of $x_{r-1}$.
 Since $x_{r-1}$ is a $T$-descendant of $x$ and $x_{r-1}x_r\in E(H)$, $x_r=y_0$ is contained in $B_x$.
@@
 
-    \item $x_r=x$.  In this case, $y_0,\ldots,y_s$ is a path from $x$ to $a$ of length at most $k-1$ and $v_r$ participates in each of $y_0,\ldots,y_s$.
+    \item $x_r=x$.  In this case, $y_0,\ldots,y_s$ is a path from $x$ to $a$ of length at most $k-1\le k$ and $v_r$ participates in each of $y_0,\ldots,y_s$.
 
     \item $x_r$ is a strict $T$-descendant of $x$.
-      Since  $y_0$ is a strict $T$-descendant of $x$ and $y_s$ is a $T$-ancestor of $x$, at least one of $y_0,\ldots,y_s$, say $y_i$, is contained in $B_x$.
+      Since  $y_0=x_r$ is a strict $T$-descendant of $x$ and $y_s$ is a $T$-ancestor of $x$, at least one of $y_0,\ldots,y_s$, say $y_i$, is contained in $B_x$.
 Therefore $xy_i$ is an edge of $H^{+}$.
@@
   \end{enumerate}
-  In each case, we obtain a path $z_0,\ldots,z_{s'}$ of length at most $k$ from $z_0:=x$ to $z_{s'}:=a$ such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.
+  In each case, we obtain a path $z_0,\ldots,z_{s'}$ in $H$ of length at most $k$ from $z_0:=x$ to $z_{s'}:=a$ such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.
 By definition, $a$ is a $T$-ancestor of $x$.

### 2022-11-04 k-planar.tex
updated refs

@@
       planar & $6^*$ & $1$ & \cite{UWY22} \\
-      genus $g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW} \\
+      genus $g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW22} \\
       genus $g$ & $2g+6$ & 1 & \cite{UWY22} \\
@@
 
-\begin{thm}[\citep{DJMMUW20,UWY22,DHHW}]\label{GenusProduct}
+\begin{thm}[\citep{DJMMUW20,UWY22,DHHW22}]\label{GenusProduct}
   Every graph of Euler genus $g$ is contained in:
@@
 The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A $T$-decomposition of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags contain $v_1,\ldots,v_t$.
-To show the planarity of $H$ in \cref{d_framed_product_stucture} we use the fact that a graph has a 3-simple tree-decomposition if and only if it has treewidth at most 3 and is planar~\citep{KV12}.
+To show the planarity of $H$ in \cref{d_framed_product_stucture} we use the fact that a graph has a 3-simple tree-decomposition if and only if it has treewidth at most 3 and is planar~\citep{kratochvil.vaner:note}.
 Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that for each $v\in V(G)$,  \[\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.\]

### 2022-11-07 k-planar.tex
Minor touch-ups

@@
 For each $i\in\{0,\ldots,s\}$, let $Y_{y_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.
-For each $i\in\{1,\ldots,s\}$, the existence of $v_{i-1}v_{i}\in E(G)$ implies that $y_{i-1}=y_i$ or $y_{i-1}y_i\in E(H)$.
+For each $i\in\{1,\ldots,s\}$, the existence of $v_{i-1}v_{i}\in E(G)$ implies that $y_{i-1}=y_i$ or that $y_{i-1}y_i\in E(H)$.
 Therefore $y_0,\ldots,y_s$ is a lazy walk\footnote{A \defin{lazy walk} in a graph $G$ is a walk in the pseudograph obtained by adding a self-loop at each vertex of $G$.} in $H_v$ of length at most $s\le k-1$.
@@
 By \cref{short_path_in_h}, $H$ contains a path $y_0,\ldots,y_s$ from $y_0:=x_r$ to $y_s:=a$ of length at most $k-1$ such that $v_r$ participates in each of $y_0,\ldots,y_s$.
-We will use this to show that $H$ contains a path $z_0,\ldots,z_{s'}$ from $z_0:=x$ to $z_{s'}:=a$ such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.
+We will use this to show that $H$ contains a path $z_0,\ldots,z_{s'}$ of length $s'\le s$, from $z_0:=x$ to $z_{s'}:=a$, and such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.
 There are three cases to consider, depending on the location of $x_r$ relative to $x'$.
+
+  % \begin{figure}
+  %   \begin{center}
+  %     \begin{tabular}{ccc}
+  %       % (1) & (2) & (3) \\
+  %       \includegraphics[page=2]{figs/width} &
+  %       \includegraphics[page=3]{figs/width} &
+  %       \includegraphics[page=4]{figs/width}
+  %     \end{tabular}
+  %   \end{center}
+  %   \caption{Three cases in the proof of \cref{general-bag-size}}
+  %   \label{width_figure}
+  % \end{figure}
 
@@
     \item $x_r$ is a strict $T$-descendant of $x$.
-      Since  $y_0=x_r$ is a strict $T$-descendant of $x$ and $y_s$ is a $T$-ancestor of $x$, at least one of $y_0,\ldots,y_s$, say $y_i$, is contained in $B_x$.
+      Since  $y_0=x_r$ is a strict $T$-descendant of $x$ and $y_s$ is a $T$-ancestor of $x$, the path in $T$ from $y_s$ to $y_0$ includes $x$.
+Since $H[\{y_0,\ldots,y_s\}]$ is connected, at least one of $y_0,\ldots,y_s$, say $y_i$, is contained in $B_x$.
 Therefore $xy_i$ is an edge of $H^{+}$.
 Therefore $x,y_i,\ldots,y_s$ is a path in $H^+$ from $x$ to $a$ of length at most $s+1\le k$ and $v_r$ participates in each of $y_i,\ldots,y_s$.
-
   \end{enumerate}

### 2022-11-07 k-planar.tex
A bunch of changes

@@
   \item In \cref{Structure} we prove our main result for shortcut systems (\cref{ShortcutProduct}), and its optimisations for $k$-planar graphs and their genus-$g$ generalisation (\cref{kPlanarProduct,gkPlanarProduct}).
-  \item \Cref{examples} uses \cref{ShortcutProduct,kPlanarProduct,gkPlanarProduct} to derive product structure theorems for string graphs, powers of bounded-degree graphs, map graphs, and $k$-nearest neighbour graphs.
+  \item \Cref{examples} uses \cref{ShortcutProduct,kPlanarProduct,gkPlanarProduct} to derive product structure theorems for bounded-degree string graphs, powers of bounded-degree graphs, map graphs, and $k$-nearest neighbour graphs.
   \item \Cref{FramedSection} proves a product structure theorem for $d$-framed multigraphs in which the graph $H$ has treewidth $3$ and is planar.
@@
 Since $G^\SS$ is defined to be a supergraph of $G$, this assumption has no effect on $G^{\SS}$ but eliminates special cases in some of our proofs.
+For a given $H$-partition $\mathcal{H}:=\{B_x:x\in V(H)\}$ of a graph $G$, we are frequently interested in the unique vertex $x\in V(H)$ such that $B_x$ contains a particular vertex $v\in V(G)$.
+We will write this concisely by writing that ``$P_x$ is the part in $\mathcal{H}$ that contains $v$.''
 
@@
 \begin{proof}
-  Recall that $S_x$ is defined by vertices that participate in $x$, and these are vertices that are either in $Y_x$ or in shortcuts that cross $Y_x$.
+  % Recall that $S_x$ is defined by vertices that participate in $x$, and these are vertices that are either in $Y_x$ or in shortcuts that cross $Y_x$.
+  It follows from \cref{short_path_in_h} that, for each $v\in V(H)$, $v$ participates in $x:=a(v)$.
+Therefore, the contents of $S_x$ are limited to elements that are in $Y_x$ or vertices in shortcuts that cross $x$.
 We say that a vertex $w\in Y_x$ \defin{contributes} a vertex $v\in S_x$ if $v=w$ or if some path in $\SS$ that contains $v$ has $w$ as an internal vertex.
@@
   For the tree-decomposition $(C_x:x\in V(T))$ of $J$ we use the same tree $T$ used in the tree-decomposition $(B_x:x\in V(T))$ of $H$.
-For each node $x$ of $T$, we define $C_x$ as follows: $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
+For each node $x$ of $T$, we define $C_x$ as follows: $C_x$ contains every $T$-ancestor $a$ of $x$ such that $J$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
+Additionally, $C_x$ includes $x$ if $x\in V(J)$.
   \cref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, both $a$ and $x'$ are contained in $C_{x'}$.
@@
 \begin{proof}[Proof of \cref{kPlanarProduct}]
-	Let $G$ be a $k$-plane graph.	As in the proof of \cref{AddDummy}, let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$.
+  Let $G$ be a $k$-plane graph.
+As in the proof of \cref{AddDummy}, let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$.
 In this way, each edge $vw\in E(G)$ corresponds naturally to a path $P_{vw}$ of length at most $k+1$ in $G_0$.
@@
   For a vertex $x$ of $G_0$ with degree at most $\Delta$, and for $i\in\{1,\dots,k-1\}$, say a vertex $v$ is \defin{$i$-close} to $x$ if there is a $vx$-path $P$ in $G_0$ of length at most $i$ such that every internal vertex in $P$  has degree at most $\Delta$ in $G_0$.
-	For each edge $vw$ of $G$, say that $vw$ \defin{passes through} each internal vertex on $P_{vw}$.	Say $vw$ passes through $x$.
+  For each edge $vw$ of $G$, say that $vw$ \defin{passes through} each internal vertex on $P_{vw}$.
+Say $vw$ passes through $x$.
 Then $v$ is $i$-close to $x$ and $w$ is $j$-close to $x$ for some $i,j\in\{1,\dots,k-1\}$ with $i+j\leq k$.
@@
 
-\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite{DJMMUW20}.
+\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite[Lemma~17]{DJMMUW20}.
 The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A $T$-decomposition of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags contain $v_1,\ldots,v_t$.
 To show the planarity of $H$ in \cref{d_framed_product_stucture} we use the fact that a graph has a 3-simple tree-decomposition if and only if it has treewidth at most 3 and is planar~\citep{kratochvil.vaner:note}.
-Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that for each $v\in V(G)$,  \[\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.\]
+Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ that contains a tree rooted at each vertex of $V_0$ with the property that for each $v\in V(G)$,
+\[  \min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.  \]
+
+% The following technical lemma is analogous to Lemma~17 in \cite{dujmovic.joret.ea:planar}, which performs the inductive step in the construction of a so-called \emph{tripod decomposition}.
+%
+%
+%
+%
+%
+%  cycle $F$ in a near-triangulation $G$ and a partition of $V(F)$ into at most three bipods,
+%
+%
+% to how a near-triangulation bounded by a cycle $F$ in in a triangulation $G$ whose vertices have already been assigned
+%
+% we offer some comparison.  \cite[Lemma~17]{dujmovic.joret.ea:planar} shows that, given a partial partition of $V(G)$ into tripods, and a cycle $F$ in $G$ whose vert
+
+For readers already familiar with the proof in \cite{dujmovic.joret.ea:planar}, the main difference here is that a typical tripod in the proof of \cite[Lemma~17]{dujmovic.joret.ea:planar} consists of three upward paths in a BFS tree/forest whose lower endpoints belong to a single triangular face $\tau$.
+In the current setting, a typical tripod consists of three upward paths belonging to a common face $\tau$ of $G_0$ plus all of the (at most $d$) vertices of $\tau$.
 
@@
 \begin{proof}
-This proof is similar to the proof of Lemma~13 by \citet{DJMMUW20}, but is complicated by several factors.
+This proof is similar to the proofs of Lemmas~13 and 17 in \cite{DJMMUW20}, but is complicated by several factors.
 In particular, the reader should keep in mind that $G$ and $G_0$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
@@
 We now move onto the case in which $N_0$ contains at least one inner vertex.
+If $N_0$ contains a face bounded by two parallel edge $e$ and $e'$ with $e\in E(F)$, then we can apply the inductive hypothesis on the cycle $F'$ obtained from $F$ by replacing $e$ with $e'$.
+(This is a valid application of induction since $F'$ has fewer faces of $G_0$ in its interior than $F$ does.)  In this way, we may assume that each edge of $F$ bounds a face of $N_0$ that contains at least three vertices.
+
 Our $H$-partition $\mathcal{P}$ will include $P_1,\dots,P_k$ and a (possibly empty) part $S$ that will be determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N_0$.
@@
 
-		\item If $k=3$ then assign each vertex $v$ of $N_0$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $V(F)$ encountered on the path in $T$ from $v$ to the appropriate vertex of $F_0$.
+    \item If $k=3$ then assign each vertex $v$ of $N_0$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $V(F)$ encountered on the path in $T$ from $v$ to the root of the tree that contains $v$.
 (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.
@@
 This yields a plane supergraph $N_0'$ of $N_0$ in which each inner face has exactly three edges on its boundary.
-By Sperner's Lemma (see \citep{Proofs4}) there exists an inner face $\tau':=v_1,v_2,v_3$ of $N_0'$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.
-Let $\tau:=v_1,\ldots,v_p$ be  the facial cycle of $N_0$ that contains $\tau'$.
+By Sperner's Lemma (see \citep{Proofs4}) there exists an inner face $\tau':=v_1,v_2,v_3$ of $N_0'$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.\footnote{Although Sperner's Lemma is usually stated for simple graphs, the proof considers a certain graph defined over the faces of $N_0'$.
+The proof relies only on the fact that the number of odd-degree vertices in this graph is even (known as the \defin{Handshake Lemma}), which is true even when this graph is non-simple.}  Let $\tau:=v_1,\ldots,v_p$ be  the facial cycle of $N_0$ that contains $\tau'$.
   \end{compactenum}
@@
 Therefore $|Y\cap L_j|\le 3$ for each $j\ge 0$.
-Furthermore, $X\subseteq V(\tau)\setminus\{x_1,x_2,x_3\}$, so $|X|\le p-3\le d-3$.
+Furthermore, $X\subseteq V(\tau)\setminus\{v_1,v_2,v_3\}$, so $|X|\le p-3\le d-3$.
 Therefore, using $S:=Y\cup X$ as a part in the partition $\PP$ satisfies condition~(ii).
@@
 Let $G$ be a $d$-framed multigraph with frame $G_0$ having outer face $F_0$
-Let $T$ be a BFS forest of $G_0$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.
+Let $T$ be a BFS forest of $G_0$ rooted at $V(F_0)$.
+For each integer $j\ge 0$, let $L_j:=\{w\in V(G_0):\dist_G(w,V_0)=j\}$, so that $\langle L_0,L_1,\ldots\rangle$ is a layering of $G_0$.
 Let $P_1:= V(F_0)$, let $Y_1$ be any subset of $\min\{|P_1|,3\}$ vertices in $P_1$ and let $X_1:=P_1\setminus Y_1$.
-Now apply \cref{induction} on $G$, $G_0$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$.
-This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree-decomposition.
-Since any graph having a $3$-simple tree-decomposition is planar \cite{knauer.ueckerdt:simple,kratochvil.vaner:note,elmallah.colbourn:on}, $H$ is planar.
+Then $G$, $G_0$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$ satisfy the conditions of \cref{induction}.
+Apply \cref{induction} to obtain an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ that satisfies Conditions~(i), (ii), and (iii).
+Since any graph having a $3$-simple tree-decomposition is planar \cite{knauer.ueckerdt:simple,kratochvil.vaner:note,elmallah.colbourn:on}, Condition~(iii) implies that $H$ is planar.
 
@@
 This is a layering of $G$ since $\dist_{G_0}(v,w)\le \floor{ \frac{d}{2} }$ for each edge $vw\in E(G)$.
-Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\floor{\frac{d}{2}}$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \floor{ \frac{d}{2}} + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
+By Condition~(ii), $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, so $|L_i'\cap Y_x|\le 3\floor{\frac{d}{2}}$.
+Therefore, $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \floor{ \frac{d}{2}} + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
 The result now follows from \cref{PartitionProduct}.

### 2022-11-07 response2.tex
Changed response. Remembered that some referee didn't like partition
Pfft
A bunch of changes

@@
 
-\tba
+\begin{response}
+  We have added a footnote here explaining that the proof of Sperner's Lemma is based on a certain graph defined on the faces of the embedding (triangles plus outer face).
+It relies only on the fact that the number of odd-degree vertices in this graph is even, which is true even if this graph is not simple.
+\end{response}
 
@@
 \begin{response}
-We do not understand this comment. Why can't there be a product structure
-theorem for string graphs?
+  We have added the qualifier "bounded-degree" here.
 \end{response}
@@
 \begin{response}
-  We have replaced \\
-  ``$P := (S_x : x \in V (T))$ is a sequence of pairwise disjoint sets that cover $V(G)$'' \\
-  with \\
-  ``$\mathcal{P}:=\{S_x : x\in V(T)\}$ is a partition of $V(G)$.''
+  We have replaced ``sequence'' with ``collection''.
 \end{response}
@@
 
-\tba
+\begin{response}
+  As part of addressing the referee's primary comment, we have introduced an extra claim that which shows that, for any $v\in V(G)$, there is a path in $H$ from $v$ to $a(v)$ of length at most $k-1$ in which $v$ participates in every vertex on this path, including $a(v)$.
+We have rephrased this confusing sentence and replaced it with a reference to this claim.
+\end{response}
 
@@
 
-\tba
+\begin{response}
+  We have added a note at the beginning of Section 2.2 stating that, rather write ``$x$ is the unique vertex in $V(H)$ such that $v\in B_x$'' we write
+  ``$P_x$ is the part in $\mathcal{H}$ that contains $v$.''  (Where the $H$-partition $\mathcal{H}:=\{B_x:\in V(H)\}$ has already been defined.)
+\end{response}
 
@@
 
-\tba
+\begin{response}
+  We have adjusted the definition of $C_x$ so that it only includes $x$ if $x\in V(J)$.
+We agree that the overloading of the notation $T[\cdot]$ is slightly unfortunate, but can live with it since the two uses can be distinguished from the type of the argument.
+For a single node $a$, $T[a]$ denote the subtree induced by all nodes whose bags contain $a$.
+For a set $S$, $T[S]$ is the subgraph of $T$ induced by $S$.
+\end{response}
 
@@
 
-\tba
+\begin{response}
+  We now specify that this forest contains one tree rooted at each vertex of $V_0$.
+\end{response}
 
@@
 
-\tba
+\begin{response}
+  We have added a couple of sentences here pointing to the exact lemma (Lemma~17) in the JACM paper and discussing the differences between the tripods here and there.
+\end{response}
 
@@
 
-\tba
+\begin{response}
+  We have added an explanation, prior to this, of how to eliminate the possibility that an edge of $F$ bounds a face with with only two vertices and two parallel edges.
+Simply apply induction on the cycle $F'$ that uses the second of the parallel edges rather than the first.
+This is valid since $F'$ has fewer faces in its interior.
+\end{response}
 
@@
 
-\tba
+\begin{response}
+  We have replaced "appropriate" with the "root of tree in $T$ that contains $v$".
+\end{response}
 
@@
 
-\tba
+\done
 
 15. page 19, second to last paragraph. Add a sentence about why such
-a forest exists. The definition of a BFS layering is missing. A reader
-may expect that you will just use the assumptions about H and P
-that are stated, so it would be better to say something like “This gives
-an H-partition... and conditions (i), (ii), and (iii) of Lemma 24 are
-satisfied”.
+a forest exists. The definition of a BFS layering is missing.
 
-\tba
+\begin{response}
+  The more precise definition of ``BFS forest rooted $V_0$'' takes care of this.
+\end{response}
+
+A reader may expect that you will just use the assumptions about H and P that are stated, so it would be better to say something like “This gives an H-partition...
+and conditions (i), (ii), and (iii) of Lemma 24 are satisfied”.
+
+\done
 

### 2022-11-08 k-planar.tex
last three referee comments

@@
 This may introduce parallel edges into $G$, but no two such edges  appear on the boundary of a common face.
-
-\begin{figure}[h]
+\note{DW}{Add ``So each face is bounded by a triangle, and this process terminates by Euler's formula.''}
+
+\note{DW}{I think $G_0$ should be replaced by $G'$ throughout this proof.}
+
+\begin{figure}[!h]
 	\begin{center}
@@
 Since $w_i$ has degree $4>1$ in $G'$ and no two edges incident to a common vertex cross each other, $w_{i-1}\neq w_{i+1}$.
-Since $G$ is edge-maximal the edge $w_{i-1}w_{i+1}$ is an edge of $G$ on the boundary of $F$.
+Since $G$ is edge-maximal, the edge $w_{i-1}w_{i+1}$ is an edge of $G$ on the boundary of $F$.
 Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G'$, each of which is contained in a different edge of $G$.
+
+\note{DW}{Put \cref{one_planar_example} here?}
 
@@
 \begin{proof}
-Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a lake, and where each vertex of $G_0$ is incident with at most $d$ nations and let $G$ be the corresponding map graph.
+Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a lake, and where each vertex of $G_0$ is incident with at most $d$ nations. Let $G$ be the corresponding map graph.
 
@@
 Each nation cycle of length at least $3$ is now a facial cycle of $G_1$, and no face of $G_1$ has more than $d$ vertices on its boundary.
-Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.
+Let $\widehat{G}_1$ be the \note{DW}{Add ``edge-maximal''.
+Do we need to write this anywhere else?} $d$-framed graph whose frame is $G_1$.
 

### 2022-11-08 response2.tex
last three referee comments

@@
 \begin{response}
-  We have added a footnote here explaining that the proof of Sperner's Lemma is based on a certain graph defined on the faces of the embedding (triangles plus outer face).
+  We have added a footnote here explaining that the proof of Sperner's Lemma is based on a certain graph defined on the faces of the embedding (triangles plus outer-face).
 It relies only on the fact that the number of odd-degree vertices in this graph is even, which is true even if this graph is not simple.
@@
 \begin{response}
-  As part of addressing the referee's primary comment, we have introduced an extra claim that which shows that, for any $v\in V(G)$, there is a path in $H$ from $v$ to $a(v)$ of length at most $k-1$ in which $v$ participates in every vertex on this path, including $a(v)$.
+  As part of addressing the referee's primary comment, we have introduced an extra claim which shows that, for any $v\in V(G)$, there is a path in $H$ from $v$ to $a(v)$ of length at most $k-1$ in which $v$ participates in every vertex on this path, including $a(v)$.
 We have rephrased this confusing sentence and replaced it with a reference to this claim.
@@
 8. page 16, line 8. I would not call these “BFS spanning forests rooted
-at V0” since they need not be BFS spanning forests (unless the same
+at $V_0$” since they need not be BFS spanning forests (unless the same
 terminology has already appeared elsewhere?).
@@
 
-12. page 17, top. In the first two cases there may not be a vertex v3 defined
-as we could have p = 2.
+12. page 17, top. In the first two cases there may not be a vertex $v_3$ defined
+as we could have $p=2$.
 
@@
 
+\textcolor{orange}{DW: I have suggested a sentence to add. }
+
 \tba
@@
 this paragraph should be $G'$, and check for this throughout.
+
+\textcolor{orange}{DW: I think $G_0$ should be replaced by $G'$ throughout this proof.}
 
@@
 
+\textcolor{orange}{DW: First point fixed. I have suggested words to add. }
+
 \tba

### 2022-11-08 k-planar.tex
Almost done......

@@
 \begin{proof}
-Let $G$ be a $1$-plane multigraph.
+Let $G$ be a (connected) $1$-plane multigraph.
 We may assume that no two edges incident to a common vertex of $G$ cross, since such a crossing can be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}.
-We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.
-This may introduce parallel edges into $G$, but no two such edges  appear on the boundary of a common face.
-\note{DW}{Add ``So each face is bounded by a triangle, and this process terminates by Euler's formula.''}
-
-\note{DW}{I think $G_0$ should be replaced by $G'$ throughout this proof.}
+We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If distinct vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.
+This may introduce parallel edges into $G$, but no newly-introduced edge $vw$ is on the boundary of a face with an edge parallel to $vw$.
+Since $G$ is connected, each edge added during this process splits a face $F$ with a facial walk of length at least $4$ into two faces each having a shorter facial walk than that of $F$.
+Therefore this process eventually reaches a stage where each face is either bounded by a $3$-cycle or (a pre-existing) $2$-cycle and the process ends.
+
+% \note{DW}{I think $G_0$ should be replaced by $G'$ throughout this proof.}
+% \note{PM}{Agreed. Done.}
+
+To understand the structure of $G$, it is helpful to consider the plane graph $G'$  obtained by replacing each pair of edges $vw$ and $xy$ that cross at $p$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
+Let $F$ be a face of $G'$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$.
+In the following paragraphs, subscripts on the vertices in $W$ are implicitly taken modulo $r$.
+By edge-maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
+
+If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
+Therefore each dummy vertex $w_i$ of $W$ is flanked by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
+Since $w_i$ has degree $4>1$ in $G'$ and no two edges incident to a common vertex cross each other, $w_{i-1}\neq w_{i+1}$.
+Since $G$ is edge-maximal, the edge $w_{i-1}w_{i+1}$ is an edge of $G$ on the boundary of $F$.
+Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G'$, each of which is contained in a different edge of $G$.
 
@@
 Refer to \cref{one_planar_example}.
-To understand the structure of $G$, it is helpful to consider the plane graph $G'$  obtained by replacing each pair of edges $vw$ and $xy$ that cross at $p$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$.
-Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$.
-In the following paragraphs, subscripts on the vertices in $W$ are implicitly taken modulo $r$.
-By edge-maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
-
-If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
-Therefore each dummy vertex $w_i$ of $W$ is flanked by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.
-Since $w_i$ has degree $4>1$ in $G'$ and no two edges incident to a common vertex cross each other, $w_{i-1}\neq w_{i+1}$.
-Since $G$ is edge-maximal, the edge $w_{i-1}w_{i+1}$ is an edge of $G$ on the boundary of $F$.
-Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G'$, each of which is contained in a different edge of $G$.
-
-\note{DW}{Put \cref{one_planar_example} here?}
-
 Consider some pair of edges $vw$ and $xy$ that cross at point $p$, which is a degree-4 vertex of $G'$.
@@
 
+\note{PM}{Here we're using $f$ to denote a face, earlier we were using $F$.}
 If $G_0$ has a face $f$ of length 2, then add a new vertex inside $f$ adjacent to both vertices on the boundary of $f$, which creates two new triangular faces $f_1$ and $f_2$.
@@
 So we may assume that $G_0$ is an edge-maximal multigraph with no faces of length 2 (and with each face labelled a nation or a lake), such that $G$ is the corresponding map graph.
-This is well-defined since the assumption of having no face of length 2 implies that $|E(G_0|\leq 3|V(G)|-6$.
-
-Suppose that some face $f$ of $G_0$ has a disconnected boundary (implying $G_0$ is disconnected).
+This is well-defined since the assumption of having no face of length 2 implies that $|E(G_0|\leq 3|V(G_0)|-6$.
+
+Suppose that some face $f$ of $G_0$ has a disconnected boundary (implying that $G_0$ is disconnected).
 Let $v$ and $w$ be vertices in distinct components of the boundary of $f$.
@@
 Thus no face of $G_0$ is disconnected.
-Suppose that some face $f$ of $G_0$ has a repeated vertex in the boundary walk of $f$.
-Let $u,v,w$ be consecutive vertices on the boundary of $f$.
-So $u,v,w$ are distinct and $uw$ is not an edge in the boundary of $f$.
+
+Suppose that some face $f$ of $G_0$ has a repeated vertex $v$ in the boundary walk of $f$ and let $u,v,w$ be consecutive vertices in this boundary walk.
+Since $v$ appears more than once in the boundary walk of $f$, the degree of $v$ is greater than $1$.
+Since the degree of $v$ is greater than $1$ and the length of $f$ is greater than $2$, $u\neq w$.
+Since $v$ is repeated, $uw$ is not an edge in the boundary of $f$.
 Add the edge $uw$ across $f$ and consider the resulting face $uvw$ to be a lake.
@@
 Thus no facial walk of $G_0$ has a repeated vertex.
-Since each facial walk is connected, every face of $G_0$ is bounded by a cycle, and $G_0$ is biconnected.
+Since each facial walk is connected, each face of $G_0$ is bounded by a cycle, and $G_0$ is biconnected.
 

### 2022-11-08 response2.tex
Almost done......

@@
 It also seems that frames are loopless, so slightly more care is called for when they are constructed.
+
+\begin{response}
+  We have addressed this by adding the assumption that the underlying planar graph $G_0$ is edge-maximal.
+More precisely, it is not possible to add an edge to $G_0$ (and relabel faces as nation or lake) to obtain the same map graph $G$.
+We then argue that this implies that $G_0$ is biconnected.
+\end{response}
+
 It may be worth noting explicitly that you allow d-framed graphs to
@@
 this given the phrase “interior of the face”.
-
-\tba
 
@@
 16. page 20, first paragraph of Lemma 25. Well, parallel edges can appear
-on the boundary of a common face if they were there before. Add a
-sentence about why the process ends.
+on the boundary of a common face if they were there before.
 
-\textcolor{orange}{DW: I have suggested a sentence to add. }
+\begin{response}
+  We rephrased this to clarify that it only applies to newly-introduced edges.
+\end{response}
 
-\tba
+Add a sentence about why the process ends.
+
+\done
 
@@
 
-\textcolor{orange}{DW: I think $G_0$ should be replaced by $G'$ throughout this proof.}
-
-\tba
+\begin{response}
+  We have moved Figure 4 a bit later and, indeed, several occurrences of $G_0$ should have been $G'$.  These are corrected now.
+\end{response}
 

### 2022-11-08 k-planar.tex
Just noticed issue with definitions

@@
 This definition (for simple graphs) was introduced by \citet{BDGGMR}.
-
-Every $d$-framed multigraph can be described by a $(2,d(d-3)/2)$-shortcut system applied to a plane multigraph by adding a vertex inside each face $F$ of $G_0$ adjacent to the vertices of $F$ and creating a shortcut between each non-adjacent pair of vertices in $V(F)$.
-Thus \cref{ShortcutProduct} is applicable.
-However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.\footnote{A multigraph $G$ is \defin{contained} in a graph $X$ if the simple graph underlying $G$ is contained in $X$.}
+A $d$-framed graph $G$ is \defin{edge-maximal} with respect to its frame $G_0$ if, for each face $F$ of $G_0$ and each pair of distinct vertices $v,w\in V(F)$, $vw\in E(F)$ or $G$ contains an edge $vw$ through $F$.
+
+Every edge-maximal $d$-framed multigraph can be described by a $(2,d(d-3)/2)$-shortcut system applied to a plane multigraph by adding a vertex inside each face $F$ of $G_0$ adjacent to the vertices of $F$ and creating a length-$2$ shortcut between each non-adjacent pair of vertices in $V(F)$.
+Thus \cref{ShortcutProduct,PlanarProduct} imply that any $d$-framed multigraph $G$ is a subgraph of $H\boxtimes P\boxtimes K_{O(d^2)}$ where the treewidth of $H$ is at most $10$.
+However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible and dependence on $d$ is only linear.\footnote{A multigraph $G$ is \defin{contained} in a graph $X$ if the simple graph underlying $G$ is contained in $X$.}
 
@@
 For readers already familiar with the proof in \cite{dujmovic.joret.ea:planar}, the main difference here is that a typical tripod in the proof of \cite[Lemma~17]{dujmovic.joret.ea:planar} consists of three upward paths in a BFS tree/forest whose lower endpoints belong to a single triangular face $\tau$.
-In the current setting, a typical tripod consists of three upward paths belonging to a common face $\tau$ of $G_0$ plus all of the (at most $d$) vertices of $\tau$.
+In the current setting, a typical tripod consists of three upward paths whose lower endpoints belong to a common face $\tau$ of $G_0$ plus all of the (at most $d$) vertices of $\tau$ that are not already included in tripods.
 
@@
 
-\note{PM}{Here we're using $f$ to denote a face, earlier we were using $F$.}
-If $G_0$ has a face $f$ of length 2, then add a new vertex inside $f$ adjacent to both vertices on the boundary of $f$, which creates two new triangular faces $f_1$ and $f_2$.
-If $f$ is a lake, then make $f_1$ and $f_2$ lakes.
-If $F$ is a nation, then make $f_1$ a nation and make $f_2$ a lake.
+If $G_0$ has a face $F$ of length 2, then add a new vertex inside $F$ adjacent to both vertices on the boundary of $F$, which creates two new triangular faces $F_1$ and $F_2$.
+If $F$ is a lake, then make $F_1$ and $F_2$ lakes.
+If $F$ is a nation, then make $F_1$ a nation and make $F_2$ a lake.
 The resulting map graph is still $G$.
-So we may assume that $G_0$ is an edge-maximal multigraph with no faces of length 2 (and with each face labelled a nation or a lake), such that $G$ is the corresponding map graph.
-This is well-defined since the assumption of having no face of length 2 implies that $|E(G_0|\leq 3|V(G_0)|-6$.
-
-Suppose that some face $f$ of $G_0$ has a disconnected boundary (implying that $G_0$ is disconnected).
-Let $v$ and $w$ be vertices in distinct components of the boundary of $f$.
-Add the edge $vw$ to $G_0$ across $f$.
+We may therefore assume that $G_0$ has no face of length $2$.
+We may also assume that $G_0$ is edge-maximal in the following sense: Adding any edge to $G_0$ will introduce a crossing or make it impossible to label the faces of $G_0$ as nations or lakes to obtain $G$ as the resulting map graph.
+This edge-maximal graph is well-defined since the assumption of having no face of length $2$ implies that $|E(G_0|\leq 3|V(G_0)|-6$.
+
+We now argue that $G_0$ is biconnected.
+Suppose that some face $F$ of $G_0$ has a disconnected boundary (implying that $G_0$ is disconnected).
+Let $v$ and $w$ be vertices in distinct components of the boundary of $F$.
+Add the edge $vw$ to $G_0$ across $F$.
 The corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$.
-Thus no face of $G_0$ is disconnected.
-
-Suppose that some face $f$ of $G_0$ has a repeated vertex $v$ in the boundary walk of $f$ and let $u,v,w$ be consecutive vertices in this boundary walk.
-Since $v$ appears more than once in the boundary walk of $f$, the degree of $v$ is greater than $1$.
-Since the degree of $v$ is greater than $1$ and the length of $f$ is greater than $2$, $u\neq w$.
-Since $v$ is repeated, $uw$ is not an edge in the boundary of $f$.
-Add the edge $uw$ across $f$ and consider the resulting face $uvw$ to be a lake.
-Since $v$ appears elsewhere in the boundary of $f$, the corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$.
+Thus no face of $G_0$ is disconnected and therefore $G_0$ is connected.
+
+Suppose that some face $F$ of $G_0$ has a repeated vertex $v$ in the boundary walk of $F$ and let $u,v,w$ be consecutive vertices in this boundary walk.
+Since $v$ appears more than once in the boundary walk of $F$, the degree of $v$ is greater than $1$.
+Since the degree of $v$ is greater than $1$ and the length of $F$ is greater than $2$, $u\neq w$.
+Since $v$ is repeated in the boundary walk of $F$, $uw$ is not an edge in the boundary of $f$.
+Add the edge $uw$ across $F$ and consider the resulting face $uvw$ to be a lake.
+Since $v$ appears elsewhere in the boundary of $F$, the corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$.
 Thus no facial walk of $G_0$ has a repeated vertex.
@@
 Each nation cycle of length at least $3$ is now a facial cycle of $G_1$, and no face of $G_1$ has more than $d$ vertices on its boundary.
-Let $\widehat{G}_1$ be the \note{DW}{Add ``edge-maximal''.
-Do we need to write this anywhere else?} $d$-framed graph whose frame is $G_1$.
+Let $\widehat{G}_1$ be the edge-maximal $d$-framed graph whose frame is $G_1$.
 

### 2022-11-08 response2.tex
Just noticed issue with definitions
minor comments

@@
 
-It may be worth noting explicitly that you allow d-framed graphs to
+It may be worth noting explicitly that you allow $d$-framed graphs to
 have edges outside the outer face of their frame - I was unsure about
 this given the phrase “interior of the face”.
+
+\tba
 

### 2022-11-08 k-planar.tex
added acknolwedgement

--- 2022-11-08 k-planar.tex
@@
 
-For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
+For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$ \note{DW}{The word ``interior'' may be misinterpreted when refering to the outerface (see referee's comment).
+Shall we just write ``contained in some face''?
+Do we need to change this elsewhere?}.
 The embedded (multi)graph $G_0$ is called the \defin{frame} of $G$.
@@
 
+\note{DW}{I suggest we do not start a new paragraph here}
 If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
@@
 
+\subsection*{Acknowledgements} Thanks to the referees for many helpful comments.
+
 % %%%  Squashing the bibliography

### 2022-11-09 k-planar.tex

--- 2022-11-08 k-planar.tex
@@
 
-For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
+For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in some face of $G_0$.
 The embedded (multi)graph $G_0$ is called the \defin{frame} of $G$.
@@
 A $d$-framed graph $G$ is \defin{edge-maximal} with respect to its frame $G_0$ if, for each face $F$ of $G_0$ and each pair of distinct vertices $v,w\in V(F)$, $vw\in E(F)$ or $G$ contains an edge $vw$ through $F$.
+
+\note{DW}{The word ``interior'' may be misinterpreted when refering to the outerface (see referee's comment). Shall we just write ``contained in some face''? Do we need to change this elsewhere?}
+\note{PM}{Done.  Changing occurrences elsewhere also.}
 
@@
   Since $M$ is $2$-connected and $\overline{S}$ is connected, $F_i[P_{i,j}]$ is connected, for each $j\in\{1,2,3\}$.
-Let $N_i$ and $N_{0,i}$ be the subgraphs of $N$ and $N_0$ contained in the interior or boundary of $F_i$.
+Let $N_i$ and $N_{0,i}$ be the subgraphs of $N$ and $N_0$ contained in the closure of $F_i$.
 Every inner vertex of $N_{0,i}$ is an inner vertex of $N_{0}$.
@@
 If both $v$ and $w$ are in $V(M)$, then $\alpha=x_a$ and $\beta=x_b$ for some $a,b\in\{0,\ldots,3\}$, in which case $B_{y_0}=\{x_0,\ldots,x_k\}$ contains $\alpha$ and $\beta$.
-If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $N\subseteq G$ crosses an edge of $M\subseteq G_0$, both $v$ and $w$ are contained in the interior of the same face $F_i$ of $M$.
+If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $N\subseteq G$ crosses an edge of $M\subseteq G_0$, both $v$ and $w$ are contained in the same face $F_i$ of $M$.
 So, by the inductive hypothesis there exists a bag of $\mathcal{T}_i$ that contains $\alpha$ and $\beta$ and this becomes a bag of $\mathcal{T}$ that contains $\alpha$ and $\beta$.
-Finally, if $v\in V(M)$ and $w\not\in V(M)$, then $v\in P_{i,j}$ for some $j\in\{1,2,3\}$ and $w$ is in the interior of some face $F_i$ of $M$.
+Finally, if $v\in V(M)$ and $w\not\in V(M)$, then $v\in P_{i,j}$ for some $j\in\{1,2,3\}$ and $w$ is in some face $F_i$ of $M$.
 By induction, some bag of $\mathcal{T}_i$ contains $\beta$ and $\alpha'$, where $\alpha'$ is the vertex of $H_i$ obtained by contracting $P_{i,j}$.

### 2022-11-09 k-planar.tex
Notes
Went back to old definition of crossing and removed incorrect statement about two edges crossing at most once.

@@
 The \defin{interior} of $f$ is the point set $\{f(t):0<t<1\}$.} in $\Sigma$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
-The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
-A \defin{crossing} in a $\Sigma$-embedded (multi)graph $G$ is an unordered pair of distinct edges $\{vw,xy\}\subseteq E(G)$ whose interiors have a point in common.
-
-If an embedded graph $G$  has no crossings then it is \defin{non-crossing} graph.
+The \defin{faces} of $G$ are the connected components of $\Sigma\setminus \bigcup_{vw\in E(G)} vw$.
+If some point $p\in\Sigma$ is a common point in the interior of two distinct edges $vw$ and $xy$, then we say that $vw$ and $xy$ \defin{cross at} $p$ and we call the pair $(\{vw,xy\},p)$ a \defin{crossing} in $G$.
+Without any loss of generality we may assume that no point in $\Sigma$ is contained in the interior of three or more edges of $G$.
+This assumption can be enforced by local changes that do not change the number of crossings each edge is involved in.
+
+% A \defin{crossing} in a $\Sigma$-embedded (multi)graph $G$ is a pair $(\{vw,xy\},p)$ where $vw$ and $xy$ are distinct edges of $G$ and $p$ is a point that is in the interior of both $vw$ and $xy$.
+
+If an embedded graph $G$  has no crossings then it is \defin{non-crossing}.
 When $G$ is non-crossing and $F$ is a face of $G$, then let \defin{$V(F)$} be the set of vertices in $G$ on the boundary of $F$.
-When $F$ is homeomorphic to a disk, the \defin{facial walk} of $F$ is the cycle consisting of the edges and vertices of $G$ on the boundary of $F$.
+When $F$ has a single boundary component, the \defin{facial walk} of $F$ is the walk consisting of the edges and vertices of $G$ on the boundary of $F$ in the cyclic order they are encountered while traversing the boundary of $F$.
 If  $\Sigma=\R^2$ is the Euclidean plane and each edge of $G$ is involved in at most $k$ crossings, then $G$ is a $k$-plane graph.
-A graph $G$ is \defin{$k$-planar} if is isomorphic to some $k$-plane graph.
+A graph $G$ is \defin{$k$-planar} if it is isomorphic to some $k$-plane graph.
 A $0$-plane graph is a \defin{plane} graph and a $0$-planar graph is a \defin{planar} graph.
-
-Without any loss of generality we may assume that the interiors of any two edges of a $\Sigma$-embedded graph $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.
-This assumption can be enforced by local changes that do not change the number of crossings each edge is involved in.
-If $p$ is a common point in the interior of two crossing edges $vw$ and $xy$, then we say that $vw$ and $xy$ \defin{cross at} $p$.
 
@@
 \note{DW}{The word ``interior'' may be misinterpreted when refering to the outerface (see referee's comment). Shall we just write ``contained in some face''? Do we need to change this elsewhere?}
-\note{PM}{Done.  Changing occurrences elsewhere also.}
+\note{PM}{Done.
+I've changed occurrences elsewhere also.
+Under our definitions (in Section 2.3) the faces of $G$ are open sets so in $F$ already means in the interior of $F$.
+In one instance, I changed ``"``in the interior or boundary of $F$'' to ``in the closure of $F$.''}
 

### 2022-11-09 k-planar.tex
Small change

@@
 This definition (for simple graphs) was introduced by \citet{BDGGMR}.
-A $d$-framed graph $G$ is \defin{edge-maximal} with respect to its frame $G_0$ if, for each face $F$ of $G_0$ and each pair of distinct vertices $v,w\in V(F)$, $vw\in E(F)$ or $G$ contains an edge $vw$ through $F$.
+A $d$-framed graph $G$ is \defin{edge-maximal} with respect to its frame $G_0$ if, for each face $F$ of $G_0$ and each pair of distinct vertices $v,w\in V(F)$, $vw\in E(F)$ or $G$ contains an edge $vw$ whose interior is contained in $F$.
 
@@
 Under our definitions (in Section 2.3) the faces of $G$ are open sets so in $F$ already means in the interior of $F$.
-In one instance, I changed ``"``in the interior or boundary of $F$'' to ``in the closure of $F$.''}
+In one instance, I changed ``in the interior or boundary of $F$'' to ``in the closure of $F$.''}
 

### 2022-11-09 k-planar.tex
Touched up figure

@@
 
-\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite[Lemma~17]{DJMMUW20}.
 The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A $T$-decomposition of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags contain $v_1,\ldots,v_t$.
 To show the planarity of $H$ in \cref{d_framed_product_stucture} we use the fact that a graph has a 3-simple tree-decomposition if and only if it has treewidth at most 3 and is planar~\citep{kratochvil.vaner:note}.
-Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ that contains a tree rooted at each vertex of $V_0$ with the property that for each $v\in V(G)$,
-\[  \min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.  \]
 
@@
 
+A \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ that contains a tree rooted at each vertex of $V_0$ with the property that for each $v\in V(G)$,
+\[  \min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.  \]
+\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite[Lemma~17]{DJMMUW20}.
 For readers already familiar with the proof in \cite{dujmovic.joret.ea:planar}, the main difference here is that a typical tripod in the proof of \cite[Lemma~17]{dujmovic.joret.ea:planar} consists of three upward paths in a BFS tree/forest whose lower endpoints belong to a single triangular face $\tau$.

### 2022-11-09 k-planar.tex
A bunch of small changes noticed when proofreading

@@
 If an embedded graph $G$  has no crossings then it is \defin{non-crossing}.
-When $G$ is non-crossing and $F$ is a face of $G$, then let \defin{$V(F)$} be the set of vertices in $G$ on the boundary of $F$.
+When $G$ is non-crossing and $F$ is a face of $G$, we use \defin{$V(F)$} to denote the set of vertices in $G$ on the boundary of $F$.
 When $F$ has a single boundary component, the \defin{facial walk} of $F$ is the walk consisting of the edges and vertices of $G$ on the boundary of $F$ in the cyclic order they are encountered while traversing the boundary of $F$.
@@
 Thus \cref{ShortcutProduct,PlanarProduct} imply that any $d$-framed multigraph $G$ is a subgraph of $H\boxtimes P\boxtimes K_{O(d^2)}$ where the treewidth of $H$ is at most $10$.
-However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible and dependence on $d$ is only linear.\footnote{A multigraph $G$ is \defin{contained} in a graph $X$ if the simple graph underlying $G$ is contained in $X$.}
+However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible and the dependence on $d$ is only linear.\footnote{A multigraph $G$ is \defin{contained} in a graph $X$ if the simple graph underlying $G$ is contained in $X$.}
 
@@
     \end{compactenum}
-    \item Let $N$ and $N_0$ be the subgraphs of $G$ and $G_0$ consisting only of those edges and vertices contained in the interior or boundary of $F$.
+    \item Let $N$ and $N_0$ be the subgraphs of $G$ and $G_0$ consisting only of those edges and vertices contained in the closure of the interior of the Jordan curve defined by $F$.
   \end{compactenum}
@@
     \item For each $x\in V(H)$, $S_x$ has a partition $\{X_x,Y_x\}$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
-    \item $H$ has a $3$-simple tree-decomposition $\mathcal{T}$, and if $k=3$ then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
+    \item $H$ has a $3$-simple tree-decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
   \end{compactenum}
@@
 \begin{compactenum}
-    \item If $k=1$ then let $\tau:=v_1,\ldots,v_p$ be an inner face of $N_0$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
-
-    \item If $k= 2$ then let $\tau:=v_1,\ldots,v_p$ be an inner face of $N_0$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
+    \item If $k=1$ then let $\tau:=v_1,\ldots,v_p$ be (the facial cycle of) an inner face of $N_0$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
+
+    \item If $k= 2$ then let $\tau:=v_1,\ldots,v_p$ be (the facial cycle of) an inner face of $N_0$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
 
@@
 By Sperner's Lemma (see \citep{Proofs4}) there exists an inner face $\tau':=v_1,v_2,v_3$ of $N_0'$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.\footnote{Although Sperner's Lemma is usually stated for simple graphs, the proof considers a certain graph defined over the faces of $N_0'$.
-The proof relies only on the fact that the number of odd-degree vertices in this graph is even (known as the \defin{Handshake Lemma}), which is true even when this graph is non-simple.}  Let $\tau:=v_1,\ldots,v_p$ be  the facial cycle of $N_0$ that contains $\tau'$.
+The proof relies only on the fact that the number of odd-degree vertices in this graph is even (known as the \defin{Handshake Lemma}), which is true even when this graph is non-simple.}  Let $\tau:=v_1,\ldots,v_p$ be (the facial cycle of) the face in $N_0$ that contains $\tau'$.
   \end{compactenum}
@@
 Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
-By our choice of $v_1v_2v_3$, the vertices of each $F_i$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
+For each $i\in\{1,\ldots,m\}$, the choice of $v_1v_2v_3$ ensures that $V(F_i)$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subseteq S$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
 
@@
   Since $M$ is $2$-connected and $\overline{S}$ is connected, $F_i[P_{i,j}]$ is connected, for each $j\in\{1,2,3\}$.
-Let $N_i$ and $N_{0,i}$ be the subgraphs of $N$ and $N_0$ contained in the closure of $F_i$.
+Let $N_i$ and $N_{0,i}$ be the subgraphs of $N$ and $N_0$ contained in the closure of the interior of $F_i$.
 Every inner vertex of $N_{0,i}$ is an inner vertex of $N_{0}$.
@@
 This is a layering of $G$ since $\dist_{G_0}(v,w)\le \floor{ \frac{d}{2} }$ for each edge $vw\in E(G)$.
-By Condition~(ii), $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, so $|L_i'\cap Y_x|\le 3\floor{\frac{d}{2}}$.
-Therefore, $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \floor{ \frac{d}{2}} + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
+By Condition~(ii), $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, so $|L_i'\cap Y_x|\le 3\floor{\tfrac{d}{2}}$.
+Therefore, $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le 3\floor{ \tfrac{d}{2}} + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
 The result now follows from \cref{PartitionProduct}.
@@
 Since $G$ is connected, each edge added during this process splits a face $F$ with a facial walk of length at least $4$ into two faces each having a shorter facial walk than that of $F$.
-Therefore this process eventually reaches a stage where each face is either bounded by a $3$-cycle or (a pre-existing) $2$-cycle and the process ends.
+Therefore this process eventually reaches a stage where each face is either bounded by a $3$-cycle or a (pre-existing) $2$-cycle and the process ends.
 
@@
 By edge-maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
-
-\note{DW}{I suggest we do not start a new paragraph here}
 If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
@@
 Let $v$ and $w$ be vertices in distinct components of the boundary of $F$.
-Add the edge $vw$ to $G_0$ across $F$.
+Add an edge $vw$ to $G_0$ inside of $F$.
 The corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$.
-Thus no face of $G_0$ is disconnected and therefore $G_0$ is connected.
+Thus no face of $G_0$ has a disconnected boundary and therefore $G_0$ is connected.
 
@@
 Since $v$ is repeated in the boundary walk of $F$, $uw$ is not an edge in the boundary of $f$.
-Add the edge $uw$ across $F$ and consider the resulting face $uvw$ to be a lake.
+Add the edge $uw$ inside of $F$ and consider the resulting face $uvw$ to be a lake.
 Since $v$ appears elsewhere in the boundary of $F$, the corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$.
@@
 By definition, $V(G) \subseteq V(\widehat{G}_1)$.
-To prove the claim, it suffices to show that $E(G)\subseteq E(\widehat{G}_1)$.
+To finish the proof, it suffices to show that $E(G)\subseteq E(\widehat{G}_1)$.
 Indeed, if $vw\in E(G)$ then the nation faces corresponding to $v$ and $w$ have a common vertex $x$ on their boundary.
@@
 Note that \citet{DJMMUW20} previously proved the bound of
-$O(g^{k+2})$ for $k$-planar graphs using \cref{GenusProduct} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$.
+$O(g^{k+2})$ for $(g,k)$-planar graphs using \cref{GenusProduct} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$.
 

### 2022-11-09 response2.tex
Small updates to responses

@@
 \begin{response}
-  As part of addressing the referee's primary comment, we have introduced an extra claim which shows that, for any $v\in V(G)$, there is a path in $H$ from $v$ to $a(v)$ of length at most $k-1$ in which $v$ participates in every vertex on this path, including $a(v)$.
-We have rephrased this confusing sentence and replaced it with a reference to this claim.
+  As part of addressing the referee's primary comment, we have introduced an extra claim which shows that, for any $x\in V(H)$ and any $v\in Y_x$, there is a path in $H$ from the $x$ to $a(v)$ of length at most $k-1$ in which $v$ participates in every vertex on this path, including $a(v)$.
+We have rephrased the confusing sentence and replaced it with a reference to this claim.
 \end{response}
@@
 \begin{response}
-  We have added an explanation, prior to this, of how to eliminate the possibility that an edge of $F$ bounds a face with with only two vertices and two parallel edges.
+  We have added an explanation, prior to this, of how to eliminate the possibility that an edge of $F$ bounds an inner face with with only two vertices and two parallel edges.
 Simply apply induction on the cycle $F'$ that uses the second of the parallel edges rather than the first.
@@
 \begin{response}
-  We have replaced "appropriate" with the "root of tree in $T$ that contains $v$".
+  We have replaced ``appropriate vertex'' with ``the root of tree in $T$ that contains $v$.''
 \end{response}
@@
 18. page 21, Lemma 27. Should be “labelled a nation or a lake” in the
-first paragraph. In the third paragraph, the graph $\hat{G}_1$ should be “edge-maximal” or similar.
+first paragraph.
 
-\textcolor{orange}{DW: First point fixed. I have suggested words to add. }
+\done
 
-\tba
+In the third paragraph, the graph $\hat{G}_1$ should be “edge-maximal” or similar.
+
+\begin{response}
+  We have added the edge-maximal qualifier and (earlier) defined what it means for a framed graph to be edge-maximal (with respect to its frame).
+\end{response}
+
 \end{document}

### 2022-11-09 k-planar.tex
Pfft

@@
 \end{compactitem}
-In addition, \cref{PlanarProduct} has been used to resolve or make substantial progress on a number of other problems on planar graphs, including $\ell$-vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, and odd colourings~\citep{DMO}.
+In addition, \cref{PlanarProduct} has been used to resolve or make substantial progress on a number of other problems on planar graphs, including $\ell$-vertex ranking~\citep{BDJM} and twin-width~\citep{BKW}.
 
@@
 Indeed, 1-planar graphs may contain arbitrarily large complete graph minors~\citep{dujmovic.eppstein.ea:structure}.
-Hence the above results are not applicable for  $k$-planar graphs.
+Hence the preceding results are not applicable for  $k$-planar graphs.
 We extend \cref{PlanarProduct} as follows.
@@
 
-The above result for $k$-planar graphs (\cref{kPlanarProduct}) is in fact a special case of a more general result that relies on the following definition.
+The preceding result for $k$-planar graphs (\cref{kPlanarProduct}) is in fact a special case of a more general result that relies on the following definition.
 A non-empty set $\SS$ of non-trivial paths in a graph $G$ is a \defin{$(k,d)$-shortcut system} (for $G$) if:
@@
       $(g,d)$-map & $9$ & $\max\{2g,3\}\,(7d^2 -21d)$ & \cref{gdMapProduct} \\ \hline
-      \multicolumn{4}{l}{$^*$ these bounds also apply to the simple treewidth of $H$.}
+      \multicolumn{4}{l}{$^*$ these bounds also apply to the simple treewidth \cite{knauer.ueckerdt:simple} of $H$.}
     \end{tabular}
@@
   \item In \cref{Structure} we prove our main result for shortcut systems (\cref{ShortcutProduct}), and its optimisations for $k$-planar graphs and their genus-$g$ generalisation (\cref{kPlanarProduct,gkPlanarProduct}).
-  \item \Cref{examples} uses \cref{ShortcutProduct,kPlanarProduct,gkPlanarProduct} to derive product structure theorems for bounded-degree string graphs, powers of bounded-degree graphs, map graphs, and $k$-nearest neighbour graphs.
+  \item \Cref{examples} uses \cref{ShortcutProduct,kPlanarProduct,gkPlanarProduct} to derive product structure theorems for bounded-degree string graphs, powers of bounded-degree graphs, map graphs of bounded clique number, and $k$-nearest neighbour graphs.
   \item \Cref{FramedSection} proves a product structure theorem for $d$-framed multigraphs in which the graph $H$ has treewidth $3$ and is planar.
@@
 For a given $H$-partition $\mathcal{H}:=\{B_x:x\in V(H)\}$ of a graph $G$, we are frequently interested in the unique vertex $x\in V(H)$ such that $B_x$ contains a particular vertex $v\in V(G)$.
-We will write this concisely by writing that ``$P_x$ is the part in $\mathcal{H}$ that contains $v$.''
+We express this concisely by writing that ``$P_x$ is the part in $\mathcal{H}$ that contains $v$.''
 
@@
 The \defin{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.
-The \defin{lowest common $T$-ancestor} of a non-empty set $S\subseteq V(T)$ is the $T$-ancestor of every node in $S$ with maximum $T$-depth.
+The \defin{lowest common $T$-ancestor} of a non-empty set $S\subseteq V(T)$ is the maximum $T$-depth node $a\in V(T)$ that is a $T$-ancestor of every node in $S$.
 In the following, we treat any subtree $T'$ of $T$ as a rooted tree whose root is the lowest common $T$-ancestor of $V(T')$.
@@
 Select any node $x_1\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x_1$ and root $T_0$ at $x_0$.
-(The purpose of $x_0$ is to ensure that every node $x$ for which $B_x$ is non-empty has a parent.)  Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y\}]$.
+(The sole purpose of $x_0$ is to ensure that every node $x$ for which $B_x$ is non-empty has a parent.)  Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y\}]$.
 If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.
@@
 Since $H_v$ is connected, $x_r$---which is the lowest common $T$-ancestor of $V(H_v)$---is itself a node of $V(H_v)$.
-Therefore $H_v$ contains a path $x_0,\ldots,x_r$ of length most $k-1$.
+Therefore $H_v$ contains a path $x_0,\ldots,x_r$ of length at most $k-1$.
 By the definition of $H_v$, $v$ participates in each of $x_0,\ldots,x_r$.
@@
   It follows from \cref{short_path_in_h} that, for each $v\in V(H)$, $v$ participates in $x:=a(v)$.
-Therefore, the contents of $S_x$ are limited to elements that are in $Y_x$ or vertices in shortcuts that cross $x$.
+Therefore, for each $x\in V(J)$, the contents of $S_x$ are limited to vertices that are in $Y_x$ or in shortcuts that cross $x$.
 We say that a vertex $w\in Y_x$ \defin{contributes} a vertex $v\in S_x$ if $v=w$ or if some path in $\SS$ that contains $v$ has $w$ as an internal vertex.
@@
 Additionally, $C_x$ includes $x$ if $x\in V(J)$.
-  \cref{i-ancestor} ensures that, for every edge $ax'\in E(J)$, both $a$ and $x'$ are contained in $C_{x'}$.
+  \cref{i-ancestor} ensures that, for each edge $ax'\in E(J)$, both $a$ and $x'$ are contained in $C_{x'}$.
 The connectivity of $T[a]:=T[\{x\in V(T):a\in C_x\}]$ follows from the fact that, for every node $x'\in V(T[a])$, every node $x$ on the path in $T$ from $a$ to  $x'$ is also a node of $T[a]$.
@@
 Observe that any directed path in $\overrightarrow{H}^+$ that begins at $x$ leads to a $T$-ancestor of $x$.
-\citet[Lemma~13]{PS21} show that, for any $x\in V(\overrightarrow{H}^+)$, the number of strict $T$-ancestors of $x$ that can be reached by a directed path of length at most $k$ that begins at $x$ is at most $\binom{k+t}{t}-1$.
+\citet[Lemma~13]{PS21} show that, for any $x\in V(\overrightarrow{H}^+)$, the number of strict $T$-ancestors of $x$ that can be reached by a directed path in $\overrightarrow{H}^+$ of length at most $k$ that begins at $x$ is at most $\binom{k+t}{t}-1$.
 
@@
     \item $x_r$ is a strict $T$-descendant of $x$.
-      Since  $y_0=x_r$ is a strict $T$-descendant of $x$ and $y_s$ is a $T$-ancestor of $x$, the path in $T$ from $y_s$ to $y_0$ includes $x$.
+      Since  $y_0=x_r$ is a strict $T$-descendant of $x$ and $y_s=a$ is a $T$-ancestor of $x$, the path in $T$ from $y_0$ to $y_s$ includes $x$.
 Since $H[\{y_0,\ldots,y_s\}]$ is connected, at least one of $y_0,\ldots,y_s$, say $y_i$, is contained in $B_x$.
@@
   \end{enumerate}
-  In each case, we obtain a path $z_0,\ldots,z_{s'}$ in $H$ of length at most $k$ from $z_0:=x$ to $z_{s'}:=a$ such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.
+  In each case, we obtain a path $z_0,\ldots,z_{s'}$ in $H$ of length $s'\le k$ from $z_0:=x$ to $z_{s'}:=a$ such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.
 By definition, $a$ is a $T$-ancestor of $x$.
@@
 %%%%%%%
-\subsection{String Graphs}
+\subsection{Bounded-Degree String Graphs}
 \label{first_example}
@@
 
-\subsection{Map Graphs}
+\subsection{Map Graphs of Bounded Clique Number}
+
+\note{PM}{To make the referee happy I've added the ``bounded-degree'' qualifier to string graphs and the ``bounded clique number'' qualifier to map graphs.
+Do we want to explcitly state that the clique number of a $(g,d)$-map graph is bounded by a function of $g$ and $d$?
+Proof (for the $g=0$ case): A $d$-map representation gives a drawing with at most $\floor{\tfrac{d-2}{2}}\ceil{\tfrac{d-2}{2}}\in O(d^2)$ crossings per edge, but any drawing of $K_{\omega}$ on a surface of genus-$g$ has an edge that is crossed $\Omega(\omega^2)$ times.
+So if $G$ is a $d$-map graph then $\omega(G) \in O(d)$.}
 
@@
 The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A $T$-decomposition of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags contain $v_1,\ldots,v_t$.
-To show the planarity of $H$ in \cref{d_framed_product_stucture} we use the fact that a graph has a 3-simple tree-decomposition if and only if it has treewidth at most 3 and is planar~\citep{kratochvil.vaner:note}.
+To show the planarity of $H$ in \cref{d_framed_product_stucture} we use the fact that a graph has a 3-simple tree-decomposition if and only if it has treewidth at most 3 and is planar~\citep{knauer.ueckerdt:simple,kratochvil.vaner:note}.
 

### 2022-11-09 k-planar.tex
Pffft

@@
 \label{PowerProduct}
-Let $G$ be contained in $H\boxtimes P\boxtimes K_\ell$ with maximum degree $\Delta$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
+Let $G$ be a graph of maximum degree $\Delta$ that is contained in $H\boxtimes P\boxtimes K_\ell$, for some graph $H$ of treewidth at most $t$ and for some path $P$.
 Then for every integer $k\geq 1$, the $k$-th power $G^k$ is contained in $J\boxtimes P\boxtimes K_{2k \ell \Delta^{k}(k^3+3k)}$ for some graph $J$ of treewidth at most $\binom{k+t}{t}-1$ and some path $P$.
@@
 There is a natural drawing of a map graph obtained by positioning each vertex of $G$ inside the corresponding nation and each edge of $G$ as a curve passing through the corresponding vertex of $G_0$.
-It is easily seen that each edge is in at most $\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}}$ crossings; see \citep{dujmovic.eppstein.ea:structure}.
+It is easily seen that each edge of this drawing is involved in at most $\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}}$ crossings; see \citep{dujmovic.eppstein.ea:structure}.
 Thus $G$ is $(g,\floor{\frac{d-2}{2}}\ceil{\frac{d-2}{2}})$-planar.

### 2022-11-09 response2.tex
Pffft

@@
 
-\tba
+\begin{response}
+  We no longer talk about ``interiors of faces''.
+Faces are already defined as the connected components one gets after cutting a surface along all the edges of the (embedded) graph, so faces are already open sets.
+In this particular instance we now write ``the interior of each edge in $E(G)\setminus E(G_0)$ is contained in some face (possibly the outer face) of $G_0$.''
+\end{response}
 

### 2022-11-09 k-planar.tex
nothing changed

--- 2022-11-09 k-planar.tex
@@
 The \defin{interior} of $f$ is the point set $\{f(t):0<t<1\}$.} in $\Sigma$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior.
-The \defin{faces} of $G$ are the connected components of $\Sigma\setminus \bigcup_{vw\in E(G)} vw$.
-If some point $p\in\Sigma$ is a common point in the interior of two distinct edges $vw$ and $xy$, then we say that $vw$ and $xy$ \defin{cross at} $p$ and we call the pair $(\{vw,xy\},p)$ a \defin{crossing} in $G$.
-Without any loss of generality we may assume that no point in $\Sigma$ is contained in the interior of three or more edges of $G$.
+The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.
+A \defin{crossing} in a $\Sigma$-embedded (multi)graph $G$ is an unordered pair of distinct edges $\{vw,xy\}\subseteq E(G)$ whose interiors have a point in common.
+
+If an embedded graph $G$  has no crossings then it is \defin{non-crossing} graph.
+When $G$ is non-crossing and $F$ is a face of $G$, then let \defin{$V(F)$} be the set of vertices in $G$ on the boundary of $F$.
+When $F$ is homeomorphic to a disk, the \defin{facial walk} of $F$ is the cycle consisting of the edges and vertices of $G$ on the boundary of $F$.
+If  $\Sigma=\R^2$ is the Euclidean plane and each edge of $G$ is involved in at most $k$ crossings, then $G$ is a $k$-plane graph.
+A graph $G$ is \defin{$k$-planar} if is isomorphic to some $k$-plane graph.
+A $0$-plane graph is a \defin{plane} graph and a $0$-planar graph is a \defin{planar} graph.
+
+Without any loss of generality we may assume that the interiors of any two edges of a $\Sigma$-embedded graph $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.
 This assumption can be enforced by local changes that do not change the number of crossings each edge is involved in.
-
-% A \defin{crossing} in a $\Sigma$-embedded (multi)graph $G$ is a pair $(\{vw,xy\},p)$ where $vw$ and $xy$ are distinct edges of $G$ and $p$ is a point that is in the interior of both $vw$ and $xy$.
-
-If an embedded graph $G$  has no crossings then it is \defin{non-crossing}.
-When $G$ is non-crossing and $F$ is a face of $G$, we use \defin{$V(F)$} to denote the set of vertices in $G$ on the boundary of $F$.
-When $F$ has a single boundary component, the \defin{facial walk} of $F$ is the walk consisting of the edges and vertices of $G$ on the boundary of $F$ in the cyclic order they are encountered while traversing the boundary of $F$.
-If  $\Sigma=\R^2$ is the Euclidean plane and each edge of $G$ is involved in at most $k$ crossings, then $G$ is a $k$-plane graph.
-A graph $G$ is \defin{$k$-planar} if it is isomorphic to some $k$-plane graph.
-A $0$-plane graph is a \defin{plane} graph and a $0$-planar graph is a \defin{planar} graph.
+If $p$ is a common point in the interior of two crossing edges $vw$ and $xy$, then we say that $vw$ and $xy$ \defin{cross at} $p$.
 
@@
 
-For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in some face of $G_0$.
+For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$ \note{DW}{The word ``interior'' may be misinterpreted when refering to the outerface (see referee's comment).
+Shall we just write ``contained in some face''?
+Do we need to change this elsewhere?}.
 The embedded (multi)graph $G_0$ is called the \defin{frame} of $G$.
 This definition (for simple graphs) was introduced by \citet{BDGGMR}.
-A $d$-framed graph $G$ is \defin{edge-maximal} with respect to its frame $G_0$ if, for each face $F$ of $G_0$ and each pair of distinct vertices $v,w\in V(F)$, $vw\in E(F)$ or $G$ contains an edge $vw$ whose interior is contained in $F$.
-
-\note{DW}{The word ``interior'' may be misinterpreted when refering to the outerface (see referee's comment). Shall we just write ``contained in some face''? Do we need to change this elsewhere?}
-\note{PM}{Done.
-I've changed occurrences elsewhere also.
-Under our definitions (in Section 2.3) the faces of $G$ are open sets so in $F$ already means in the interior of $F$.
-In one instance, I changed ``in the interior or boundary of $F$'' to ``in the closure of $F$.''}
-
-Every edge-maximal $d$-framed multigraph can be described by a $(2,d(d-3)/2)$-shortcut system applied to a plane multigraph by adding a vertex inside each face $F$ of $G_0$ adjacent to the vertices of $F$ and creating a length-$2$ shortcut between each non-adjacent pair of vertices in $V(F)$.
-Thus \cref{ShortcutProduct,PlanarProduct} imply that any $d$-framed multigraph $G$ is a subgraph of $H\boxtimes P\boxtimes K_{O(d^2)}$ where the treewidth of $H$ is at most $10$.
-However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible and the dependence on $d$ is only linear.\footnote{A multigraph $G$ is \defin{contained} in a graph $X$ if the simple graph underlying $G$ is contained in $X$.}
+
+Every $d$-framed multigraph can be described by a $(2,d(d-3)/2)$-shortcut system applied to a plane multigraph by adding a vertex inside each face $F$ of $G_0$ adjacent to the vertices of $F$ and creating a shortcut between each non-adjacent pair of vertices in $V(F)$.
+Thus \cref{ShortcutProduct} is applicable.
+However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.\footnote{A multigraph $G$ is \defin{contained} in a graph $X$ if the simple graph underlying $G$ is contained in $X$.}
 
@@
 
+\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite[Lemma~17]{DJMMUW20}.
 The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A $T$-decomposition of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags contain $v_1,\ldots,v_t$.
 To show the planarity of $H$ in \cref{d_framed_product_stucture} we use the fact that a graph has a 3-simple tree-decomposition if and only if it has treewidth at most 3 and is planar~\citep{kratochvil.vaner:note}.
+Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ that contains a tree rooted at each vertex of $V_0$ with the property that for each $v\in V(G)$,
+\[  \min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.  \]
 
@@
 
-A \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ that contains a tree rooted at each vertex of $V_0$ with the property that for each $v\in V(G)$,
-\[  \min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.  \]
-\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite[Lemma~17]{DJMMUW20}.
 For readers already familiar with the proof in \cite{dujmovic.joret.ea:planar}, the main difference here is that a typical tripod in the proof of \cite[Lemma~17]{dujmovic.joret.ea:planar} consists of three upward paths in a BFS tree/forest whose lower endpoints belong to a single triangular face $\tau$.
-In the current setting, a typical tripod consists of three upward paths whose lower endpoints belong to a common face $\tau$ of $G_0$ plus all of the (at most $d$) vertices of $\tau$ that are not already included in tripods.
+In the current setting, a typical tripod consists of three upward paths belonging to a common face $\tau$ of $G_0$ plus all of the (at most $d$) vertices of $\tau$.
 
@@
     \end{compactenum}
-    \item Let $N$ and $N_0$ be the subgraphs of $G$ and $G_0$ consisting only of those edges and vertices contained in the closure of the interior of the Jordan curve defined by $F$.
+    \item Let $N$ and $N_0$ be the subgraphs of $G$ and $G_0$ consisting only of those edges and vertices contained in the interior or boundary of $F$.
   \end{compactenum}
@@
     \item For each $x\in V(H)$, $S_x$ has a partition $\{X_x,Y_x\}$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
-    \item $H$ has a $3$-simple tree-decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
+    \item $H$ has a $3$-simple tree-decomposition $\mathcal{T}$, and if $k=3$ then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
   \end{compactenum}
@@
 \begin{compactenum}
-    \item If $k=1$ then let $\tau:=v_1,\ldots,v_p$ be (the facial cycle of) an inner face of $N_0$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
-
-    \item If $k= 2$ then let $\tau:=v_1,\ldots,v_p$ be (the facial cycle of) an inner face of $N_0$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
+    \item If $k=1$ then let $\tau:=v_1,\ldots,v_p$ be an inner face of $N_0$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.
+
+    \item If $k= 2$ then let $\tau:=v_1,\ldots,v_p$ be an inner face of $N_0$ that contains at least one edge $v_1v_2$ of $F$ on its boundary and such that $v_1\in P_1$ and $v_2\in P_2$.
 
@@
 By Sperner's Lemma (see \citep{Proofs4}) there exists an inner face $\tau':=v_1,v_2,v_3$ of $N_0'$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.\footnote{Although Sperner's Lemma is usually stated for simple graphs, the proof considers a certain graph defined over the faces of $N_0'$.
-The proof relies only on the fact that the number of odd-degree vertices in this graph is even (known as the \defin{Handshake Lemma}), which is true even when this graph is non-simple.}  Let $\tau:=v_1,\ldots,v_p$ be (the facial cycle of) the face in $N_0$ that contains $\tau'$.
+The proof relies only on the fact that the number of odd-degree vertices in this graph is even (known as the \defin{Handshake Lemma}), which is true even when this graph is non-simple.}  Let $\tau:=v_1,\ldots,v_p$ be  the facial cycle of $N_0$ that contains $\tau'$.
   \end{compactenum}
@@
 Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$.
-For each $i\in\{1,\ldots,m\}$, the choice of $v_1v_2v_3$ ensures that $V(F_i)$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subseteq S$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
+By our choice of $v_1v_2v_3$, the vertices of each $F_i$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.
 
@@
   Since $M$ is $2$-connected and $\overline{S}$ is connected, $F_i[P_{i,j}]$ is connected, for each $j\in\{1,2,3\}$.
-Let $N_i$ and $N_{0,i}$ be the subgraphs of $N$ and $N_0$ contained in the closure of the interior of $F_i$.
+Let $N_i$ and $N_{0,i}$ be the subgraphs of $N$ and $N_0$ contained in the interior or boundary of $F_i$.
 Every inner vertex of $N_{0,i}$ is an inner vertex of $N_{0}$.
@@
 If both $v$ and $w$ are in $V(M)$, then $\alpha=x_a$ and $\beta=x_b$ for some $a,b\in\{0,\ldots,3\}$, in which case $B_{y_0}=\{x_0,\ldots,x_k\}$ contains $\alpha$ and $\beta$.
-If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $N\subseteq G$ crosses an edge of $M\subseteq G_0$, both $v$ and $w$ are contained in the same face $F_i$ of $M$.
+If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $N\subseteq G$ crosses an edge of $M\subseteq G_0$, both $v$ and $w$ are contained in the interior of the same face $F_i$ of $M$.
 So, by the inductive hypothesis there exists a bag of $\mathcal{T}_i$ that contains $\alpha$ and $\beta$ and this becomes a bag of $\mathcal{T}$ that contains $\alpha$ and $\beta$.
-Finally, if $v\in V(M)$ and $w\not\in V(M)$, then $v\in P_{i,j}$ for some $j\in\{1,2,3\}$ and $w$ is in some face $F_i$ of $M$.
+Finally, if $v\in V(M)$ and $w\not\in V(M)$, then $v\in P_{i,j}$ for some $j\in\{1,2,3\}$ and $w$ is in the interior of some face $F_i$ of $M$.
 By induction, some bag of $\mathcal{T}_i$ contains $\beta$ and $\alpha'$, where $\alpha'$ is the vertex of $H_i$ obtained by contracting $P_{i,j}$.
@@
 This is a layering of $G$ since $\dist_{G_0}(v,w)\le \floor{ \frac{d}{2} }$ for each edge $vw\in E(G)$.
-By Condition~(ii), $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, so $|L_i'\cap Y_x|\le 3\floor{\tfrac{d}{2}}$.
-Therefore, $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le 3\floor{ \tfrac{d}{2}} + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
+By Condition~(ii), $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, so $|L_i'\cap Y_x|\le 3\floor{\frac{d}{2}}$.
+Therefore, $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \floor{ \frac{d}{2}} + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$.
 The result now follows from \cref{PartitionProduct}.
@@
 Since $G$ is connected, each edge added during this process splits a face $F$ with a facial walk of length at least $4$ into two faces each having a shorter facial walk than that of $F$.
-Therefore this process eventually reaches a stage where each face is either bounded by a $3$-cycle or a (pre-existing) $2$-cycle and the process ends.
+Therefore this process eventually reaches a stage where each face is either bounded by a $3$-cycle or (a pre-existing) $2$-cycle and the process ends.
 
@@
 By edge-maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.
+
+\note{DW}{I suggest we do not start a new paragraph here}
 If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.
@@
 
-If $G_0$ has a face $F$ of length 2, then add a new vertex inside $F$ adjacent to both vertices on the boundary of $F$, which creates two new triangular faces $F_1$ and $F_2$.
-If $F$ is a lake, then make $F_1$ and $F_2$ lakes.
-If $F$ is a nation, then make $F_1$ a nation and make $F_2$ a lake.
+\note{PM}{Here we're using $f$ to denote a face, earlier we were using $F$.}
+If $G_0$ has a face $f$ of length 2, then add a new vertex inside $f$ adjacent to both vertices on the boundary of $f$, which creates two new triangular faces $f_1$ and $f_2$.
+If $f$ is a lake, then make $f_1$ and $f_2$ lakes.
+If $F$ is a nation, then make $f_1$ a nation and make $f_2$ a lake.
 The resulting map graph is still $G$.
-We may therefore assume that $G_0$ has no face of length $2$.
-We may also assume that $G_0$ is edge-maximal in the following sense: Adding any edge to $G_0$ will introduce a crossing or make it impossible to label the faces of $G_0$ as nations or lakes to obtain $G$ as the resulting map graph.
-This edge-maximal graph is well-defined since the assumption of having no face of length $2$ implies that $|E(G_0|\leq 3|V(G_0)|-6$.
-
-We now argue that $G_0$ is biconnected.
-Suppose that some face $F$ of $G_0$ has a disconnected boundary (implying that $G_0$ is disconnected).
-Let $v$ and $w$ be vertices in distinct components of the boundary of $F$.
-Add an edge $vw$ to $G_0$ inside of $F$.
+So we may assume that $G_0$ is an edge-maximal multigraph with no faces of length 2 (and with each face labelled a nation or a lake), such that $G$ is the corresponding map graph.
+This is well-defined since the assumption of having no face of length 2 implies that $|E(G_0|\leq 3|V(G_0)|-6$.
+
+Suppose that some face $f$ of $G_0$ has a disconnected boundary (implying that $G_0$ is disconnected).
+Let $v$ and $w$ be vertices in distinct components of the boundary of $f$.
+Add the edge $vw$ to $G_0$ across $f$.
 The corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$.
-Thus no face of $G_0$ has a disconnected boundary and therefore $G_0$ is connected.
-
-Suppose that some face $F$ of $G_0$ has a repeated vertex $v$ in the boundary walk of $F$ and let $u,v,w$ be consecutive vertices in this boundary walk.
-Since $v$ appears more than once in the boundary walk of $F$, the degree of $v$ is greater than $1$.
-Since the degree of $v$ is greater than $1$ and the length of $F$ is greater than $2$, $u\neq w$.
-Since $v$ is repeated in the boundary walk of $F$, $uw$ is not an edge in the boundary of $f$.
-Add the edge $uw$ inside of $F$ and consider the resulting face $uvw$ to be a lake.
-Since $v$ appears elsewhere in the boundary of $F$, the corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$.
+Thus no face of $G_0$ is disconnected.
+
+Suppose that some face $f$ of $G_0$ has a repeated vertex $v$ in the boundary walk of $f$ and let $u,v,w$ be consecutive vertices in this boundary walk.
+Since $v$ appears more than once in the boundary walk of $f$, the degree of $v$ is greater than $1$.
+Since the degree of $v$ is greater than $1$ and the length of $f$ is greater than $2$, $u\neq w$.
+Since $v$ is repeated, $uw$ is not an edge in the boundary of $f$.
+Add the edge $uw$ across $f$ and consider the resulting face $uvw$ to be a lake.
+Since $v$ appears elsewhere in the boundary of $f$, the corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$.
 Thus no facial walk of $G_0$ has a repeated vertex.
@@
 Each nation cycle of length at least $3$ is now a facial cycle of $G_1$, and no face of $G_1$ has more than $d$ vertices on its boundary.
-Let $\widehat{G}_1$ be the edge-maximal $d$-framed graph whose frame is $G_1$.
+Let $\widehat{G}_1$ be the \note{DW}{Add ``edge-maximal''.
+Do we need to write this anywhere else?} $d$-framed graph whose frame is $G_1$.
 
 By definition, $V(G) \subseteq V(\widehat{G}_1)$.
-To finish the proof, it suffices to show that $E(G)\subseteq E(\widehat{G}_1)$.
+To prove the claim, it suffices to show that $E(G)\subseteq E(\widehat{G}_1)$.
 Indeed, if $vw\in E(G)$ then the nation faces corresponding to $v$ and $w$ have a common vertex $x$ on their boundary.
@@
 Note that \citet{DJMMUW20} previously proved the bound of
-$O(g^{k+2})$ for $(g,k)$-planar graphs using \cref{GenusProduct} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$.
+$O(g^{k+2})$ for $k$-planar graphs using \cref{GenusProduct} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$.
 

### 2022-11-09 k-planar.tex

--- 2022-11-09 k-planar.tex
@@
 
-For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in some face of $G_0$.
+For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in some face (possibly the outer face) of $G_0$.
 The embedded (multi)graph $G_0$ is called the \defin{frame} of $G$.
@@
 \note{DW}{The word ``interior'' may be misinterpreted when refering to the outerface (see referee's comment). Shall we just write ``contained in some face''? Do we need to change this elsewhere?}
-\note{PM}{Done.
-I've changed occurrences elsewhere also.
-Under our definitions (in Section 2.3) the faces of $G$ are open sets so in $F$ already means in the interior of $F$.
-In one instance, I changed ``in the interior or boundary of $F$'' to ``in the closure of $F$.''}
+\note{PM}{Done.  I've changed occurrences elsewhere also.  Under our definitions (in Section 2.3) the faces of $G$ are open sets so ``in $F$'' already means ``in the interior of $F$.''}
 

### 2022-11-09 k-planar.tex
answered comments

@@
   \item In \cref{Structure} we prove our main result for shortcut systems (\cref{ShortcutProduct}), and its optimisations for $k$-planar graphs and their genus-$g$ generalisation (\cref{kPlanarProduct,gkPlanarProduct}).
-  \item \Cref{examples} uses \cref{ShortcutProduct,kPlanarProduct,gkPlanarProduct} to derive product structure theorems for bounded-degree string graphs, powers of bounded-degree graphs, map graphs of bounded clique number, and $k$-nearest neighbour graphs.
+  \item \Cref{examples} uses \cref{ShortcutProduct,kPlanarProduct,gkPlanarProduct} to derive product structure theorems for bounded-degree \note{DW}{We actually make a stronger assumption than bounded degree; we assume each string is crossed at most $k$ times.
+I am happy to leave our write-up as is, but be aware of the difference.
+} string graphs, powers of bounded-degree graphs, map graphs of bounded clique number, and $k$-nearest neighbour graphs.
   \item \Cref{FramedSection} proves a product structure theorem for $d$-framed multigraphs in which the graph $H$ has treewidth $3$ and is planar.
@@
 Proof (for the $g=0$ case): A $d$-map representation gives a drawing with at most $\floor{\tfrac{d-2}{2}}\ceil{\tfrac{d-2}{2}}\in O(d^2)$ crossings per edge, but any drawing of $K_{\omega}$ on a surface of genus-$g$ has an edge that is crossed $\Omega(\omega^2)$ times.
-So if $G$ is a $d$-map graph then $\omega(G) \in O(d)$.}
+So if $G$ is a $d$-map graph then $\omega(G) \in O(d)$.} \note{DW}{I see no need to state that the clique number of a $(g,d)$-map graph is bounded by a function of $g$ and $d$.
+This is implied by \cref{gdMapProduct} and is not used in its proof.
+Also, surely there is a more combinatorial proof (with better bounds) than what you get by counting crossings.}
 
@@
 
-\note{DW}{The word ``interior'' may be misinterpreted when refering to the outerface (see referee's comment). Shall we just write ``contained in some face''? Do we need to change this elsewhere?}
-\note{PM}{Done.  I've changed occurrences elsewhere also.  Under our definitions (in Section 2.3) the faces of $G$ are open sets so ``in $F$'' already means ``in the interior of $F$.''}
+%\note{DW}{The word ``interior'' may be misinterpreted when refering to the outerface (see referee's comment).
+Shall we just write ``contained in some face''?
+Do we need to change this elsewhere?} \note{PM}{Done.
+I've changed occurrences elsewhere also.
+Under our definitions (in Section 2.3) the faces of $G$ are open sets so ``in $F$'' already means ``in the interior of $F$.''} \note{DW}{great}
 

### 2022-11-12 k-planar.tex
added minor comments

@@
 \begin{abstract}
-Dujmovi\'c~et~al.~[\emph{J.~ACM}~'20] recently proved that every planar graph is isomorphic to a subgraph of the strong product of a bounded treewidth graph and a path.
+Dujmovi\'c~et~al.~[\emph{J.~ACM}~'20] recently \note{DW}{delete ``recently''?} proved that every planar graph is isomorphic to a subgraph of the strong product of a bounded treewidth graph and a path.
 Analogous results were obtained for graphs of bounded Euler genus or apex-minor-free graphs.
@@
 
-For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in some face (possibly the outer face) of $G_0$.
+For any integer $d\ge 3$, an \note{DW}{add ``$\R^2$-''}embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in some face (possibly the outer face) of $G_0$.
 The embedded (multi)graph $G_0$ is called the \defin{frame} of $G$.
@@
 Since $G$ is connected, each edge added during this process splits a face $F$ with a facial walk of length at least $4$ into two faces each having a shorter facial walk than that of $F$.
-Therefore this process eventually reaches a stage where each face is either bounded by a $3$-cycle or a (pre-existing) $2$-cycle and the process ends.
+Therefore this process eventually reaches a stage where each face is either bounded by a $3$-cycle or a (pre-existing) $2$-cycle, and the process ends.
 
@@
 Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
-\citet{DJMMUW20} recently proved this conjecture using \cref{PlanarProduct} and the following lemma.
+\citet{DJMMUW20} recently \note{DW}{delete ``recently''?} proved this conjecture using \cref{PlanarProduct} and the following lemma.
 Indeed, resolving this question was the motivation for the development of \cref{PlanarProduct}.
@@
 The \defin{non-repetitive chromatic number} $\pi(G)$ of $G$ is the minimum integer $c$ such that $G$ has a non-repetitive $c$-colouring.
-This concept, introduced by \citet{AGHR-RSA02}, has since been widely studied; see \citep{dujmovic.esperet.ea:planar} for more than 40 references.
-Up until recently the main open problem in the field has been whether planar graphs have bounded non-repetitive chromatic number, first asked by \citet{AGHR-RSA02}.
-\citet{dujmovic.esperet.ea:planar} recently solved this question using \cref{PlanarProduct} and the following lemma.
+This concept, introduced by \citet{AGHR-RSA02}, has since been widely studied; see \citep{Wood21} for an extensive survey.
+Up until recently the main open problem in the field has been \note{DW}{replace ``has been'' by ``was''?} whether planar graphs have bounded non-repetitive chromatic number, first asked by \citet{AGHR-RSA02}.
+\citet{dujmovic.esperet.ea:planar} recently \note{DW}{delete ``recently''?}  solved this question using \cref{PlanarProduct} and the following lemma.
 

### 2022-11-15 k-planar.tex
nada
minor changes implemented

--- 2022-11-09 k-planar.tex
@@
 \begin{abstract}
-Dujmovi\'c~et~al.~[\emph{J.~ACM}~'20] recently proved that every planar graph is isomorphic to a subgraph of the strong product of a bounded treewidth graph and a path.
+Dujmovi\'c~et~al.~[\emph{J.~ACM}~'20] proved that every planar graph is isomorphic to a subgraph of the strong product of a bounded treewidth graph and a path.
 Analogous results were obtained for graphs of bounded Euler genus or apex-minor-free graphs.
@@
 We use $G-S$ as a shorthand for $G[V(G)\setminus S]$.
-Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (with an improvement by \citet{UWY22}).
+Undefined terms are in \citep{Diestel5}.} by \citet{dujmovic.joret.ea:planar} (with an improvement by \citet{UWY22}).
 A graph $G$ is \defin{contained} in a graph $X$ if $G$ is isomorphic to a subgraph of $X$.
 
-\begin{thm}[\citep{DJMMUW20,UWY22}]
+\begin{thm}[\citep{dujmovic.joret.ea:planar,UWY22}]
 \label{PlanarProduct}
@@
 \begin{compactitem}
-\item \citet{DJMMUW20} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92}).
+\item \citet{dujmovic.joret.ea:planar} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92}).
 \item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
@@
 All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this statement.} These applications motivate finding product structure theorems for other graph classes.
-\citet{DJMMUW20} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of a surface with $h$ handles and $c$ crosscaps is $2h+c$.
+\citet{dujmovic.joret.ea:planar} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of a surface with $h$ handles and $c$ crosscaps is $2h+c$.
 The \textit{Euler genus} of a graph $G$ is the minimum integer $g$ such that $G$ embeds in a surface of Euler genus $g$.
@@
       {Graph class} & $\tw(H)$ & $\ell$ & {Reference} \\ \hline
-      planar & $3^*$ & $3$ & \cite{DJMMUW20} \\
+      planar & $3^*$ & $3$ & \cite{dujmovic.joret.ea:planar} \\
       planar & $4^*$ & $2$ & \cite{bose.morin.ea:optimal} \\
@@
       genus $g$ & $2g+6$ & 1 & \cite{UWY22} \\
-      apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
-      % B-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
+      apex-minor-free & $O(1)$ & $O(1)$ & \cite{dujmovic.joret.ea:planar} \\
+      % B-minor-free & $O(1)$ & $O(1)$ & \cite{dujmovic.joret.ea:planar} \\
       $k$-planar & $\binom{k+4}{3}-1$ & $18k^2 + 48k + 30$ & \cref{kPlanarProduct} \\
@@
   \item In \cref{Structure} we prove our main result for shortcut systems (\cref{ShortcutProduct}), and its optimisations for $k$-planar graphs and their genus-$g$ generalisation (\cref{kPlanarProduct,gkPlanarProduct}).
-  \item \Cref{examples} uses \cref{ShortcutProduct,kPlanarProduct,gkPlanarProduct} to derive product structure theorems for bounded-degree string graphs, powers of bounded-degree graphs, map graphs of bounded clique number, and $k$-nearest neighbour graphs.
+  \item \Cref{examples} uses \cref{ShortcutProduct,kPlanarProduct,gkPlanarProduct} to derive product structure theorems for bounded-degree 
+  %\note{DW}{We actually make a stronger assumption than bounded degree; we assume each string is crossed at most $k$ times. I am happy to leave our write-up as is, but be aware of the difference. } 
+  string graphs, powers of bounded-degree graphs, map graphs of bounded clique number, and $k$-nearest neighbour graphs.
   \item \Cref{FramedSection} proves a product structure theorem for $d$-framed multigraphs in which the graph $H$ has treewidth $3$ and is planar.
@@
 
-\begin{lem}[\citep{DJMMUW20}]
+\begin{lem}[\citep{dujmovic.joret.ea:planar}]
 \label{PartitionProduct}
@@
 As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a),
-\citet{DJMMUW20} build on an earlier result of \citet{PS21} to show that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \citep{UWY22}).
+\citet{dujmovic.joret.ea:planar} build on an earlier result of \citet{PS21} to show that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \citep{UWY22}).
 The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.\footnote{\citet{bose.morin.ea:optimal} have recently given linear time algorithms for computing layered $H$-partitions of planar graphs.}
@@
 Perhaps $C=3$, which is the case for planar graphs (\cref{PlanarProduct}(b)) and for $1$-planar graphs (\cref{1_planar_product}).
-Note that $C\geq 3$ even for planar graphs \citep{DJMMUW20}.
+Note that $C\geq 3$ even for planar graphs \citep{dujmovic.joret.ea:planar}.
 
@@
 
-\begin{thm}[\citep{DJMMUW20,UWY22,DHHW22}]\label{GenusProduct}
+\begin{thm}[\citep{dujmovic.joret.ea:planar,UWY22,DHHW22}]\label{GenusProduct}
   Every graph of Euler genus $g$ is contained in:
@@
 Every $(g,k)$-planar graph is contained in $H\boxtimes P \boxtimes K_\ell$ for some graph $H$ with $\tw(H) \leq \binom{k+4}{3}-1$, where $\ell:=\max\{2g,3\}\cdot(6k^2+16k+10)$.
-% \note{DW}{I have changed $\tw(H) \leq \binom{k+5}{4}-1$ to $\tw(H) \leq \binom{k+4}{3}-1$, okay?}
-%\note{PM}{Ok}
 \end{thm}
@@
 
-\note{PM}{To make the referee happy I've added the ``bounded-degree'' qualifier to string graphs and the ``bounded clique number'' qualifier to map graphs.
+%\note{PM}{To make the referee happy I've added the ``bounded-degree'' qualifier to string graphs and the ``bounded clique number'' qualifier to map graphs.
 Do we want to explcitly state that the clique number of a $(g,d)$-map graph is bounded by a function of $g$ and $d$?
 Proof (for the $g=0$ case): A $d$-map representation gives a drawing with at most $\floor{\tfrac{d-2}{2}}\ceil{\tfrac{d-2}{2}}\in O(d^2)$ crossings per edge, but any drawing of $K_{\omega}$ on a surface of genus-$g$ has an edge that is crossed $\Omega(\omega^2)$ times.
-So if $G$ is a $d$-map graph then $\omega(G) \in O(d)$.}
+So if $G$ is a $d$-map graph then $\omega(G) \in O(d)$.} \note{DW}{I see no need to state that the clique number of a $(g,d)$-map graph is bounded by a function of $g$ and $d$.
+This is implied by \cref{gdMapProduct} and is not used in its proof.
+Also, surely there is a more combinatorial proof (with better bounds) than what you get by counting crossings.}
 
@@
 
-For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in some face (possibly the outer face) of $G_0$.
+For any integer $d\ge 3$, an $\R^2$-embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in some face (possibly the outer face) of $G_0$.
 The embedded (multi)graph $G_0$ is called the \defin{frame} of $G$.
@@
 A $d$-framed graph $G$ is \defin{edge-maximal} with respect to its frame $G_0$ if, for each face $F$ of $G_0$ and each pair of distinct vertices $v,w\in V(F)$, $vw\in E(F)$ or $G$ contains an edge $vw$ whose interior is contained in $F$.
-
-\note{DW}{The word ``interior'' may be misinterpreted when refering to the outerface (see referee's comment). Shall we just write ``contained in some face''? Do we need to change this elsewhere?}
-\note{PM}{Done.  I've changed occurrences elsewhere also.  Under our definitions (in Section 2.3) the faces of $G$ are open sets so ``in $F$'' already means ``in the interior of $F$.''}
 
@@
 \[  \min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.  \]
-\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite[Lemma~17]{DJMMUW20}.
+\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite[Lemma~17]{dujmovic.joret.ea:planar}.
 For readers already familiar with the proof in \cite{dujmovic.joret.ea:planar}, the main difference here is that a typical tripod in the proof of \cite[Lemma~17]{dujmovic.joret.ea:planar} consists of three upward paths in a BFS tree/forest whose lower endpoints belong to a single triangular face $\tau$.
@@
 \begin{proof}
-This proof is similar to the proofs of Lemmas~13 and 17 in \cite{DJMMUW20}, but is complicated by several factors.
+This proof is similar to the proofs of Lemmas~13 and 17 in \cite{dujmovic.joret.ea:planar}, but is complicated by several factors.
 In particular, the reader should keep in mind that $G$ and $G_0$ are multigraphs and that the cycle $F$ may consist of two vertices and two (parallel) edges.
@@
 
-  \begin{figure}[!b]
+  \begin{figure}[!h]
     \begin{center}
@@
 Since $G$ is connected, each edge added during this process splits a face $F$ with a facial walk of length at least $4$ into two faces each having a shorter facial walk than that of $F$.
-Therefore this process eventually reaches a stage where each face is either bounded by a $3$-cycle or a (pre-existing) $2$-cycle and the process ends.
-
-% \note{DW}{I think $G_0$ should be replaced by $G'$ throughout this proof.}
-% \note{PM}{Agreed. Done.}
+Therefore this process eventually reaches a stage where each face is either bounded by a $3$-cycle or a (pre-existing) $2$-cycle, and the process ends.
 
@@
 Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number.
-\citet{DJMMUW20} recently proved this conjecture using \cref{PlanarProduct} and the following lemma.
+\citet{dujmovic.joret.ea:planar} proved this conjecture using \cref{PlanarProduct} and the following lemma.
 Indeed, resolving this question was the motivation for the development of \cref{PlanarProduct}.
 
-\begin{lem}[\citep{DJMMUW20}]
+\begin{lem}[\citep{dujmovic.joret.ea:planar}]
 \label{qn}
@@
 
-Note that \citet{DJMMUW20} previously proved the bound of
+Note that \citet{dujmovic.joret.ea:planar} previously proved the bound of
 $O(g^{k+2})$ for $(g,k)$-planar graphs using \cref{GenusProduct} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$.
@@
 The \defin{non-repetitive chromatic number} $\pi(G)$ of $G$ is the minimum integer $c$ such that $G$ has a non-repetitive $c$-colouring.
-This concept, introduced by \citet{AGHR-RSA02}, has since been widely studied; see \citep{dujmovic.esperet.ea:planar} for more than 40 references.
-Up until recently the main open problem in the field has been whether planar graphs have bounded non-repetitive chromatic number, first asked by \citet{AGHR-RSA02}.
-\citet{dujmovic.esperet.ea:planar} recently solved this question using \cref{PlanarProduct} and the following lemma.
+This concept, introduced by \citet{AGHR-RSA02}, has since been widely studied; see \citep{Wood21} for an extensive survey.
+Up until recently the main open problem in the field was whether planar graphs have bounded non-repetitive chromatic number, first asked by \citet{AGHR-RSA02}.
+\citet{dujmovic.esperet.ea:planar} solved this question using \cref{PlanarProduct} and the following lemma.
 
@@
 For planar graphs of treewidth at most $3$ a $O(p^2\log p)$ upper bound is known:
-% \note{DW}{I prefer ``a $O(p^2\log p)$'', pronounced ``a big-O...''}
 

### 2022-11-15 response2.tex
answered comments
Added missing "the"
Small rephrasing

@@
 \usepackage[utf8]{inputenc}
-\usepackage[margin=20mm]{geometry}
+\usepackage[margin=25mm]{geometry}
 \usepackage{amsmath,color,amsthm,enumerate,amsfonts,mathtools}
@@
 \begin{response}
-  We have replaced ``appropriate vertex'' with ``the root of tree in $T$ that contains $v$.''
+  We have replaced ``appropriate vertex'' with ``the root of the tree in $T$ that contains $v$.''
 \end{response}
@@
 \begin{response}
-  We have added the edge-maximal qualifier and (earlier) defined what it means for a framed graph to be edge-maximal (with respect to its frame).
+  We have added the edge-maximal qualifier and (earlier) defined what it means for a framed graph to be edge-maximal (with respect to a particular frame).
 \end{response}

### 2022-11-15 k-planar.tex
Distinction between δ-string and max-degree-δ string graph

@@
 
-A \defin{string graph} is the intersection graph of a set of curves in the plane with no three curves meeting at a single point; see  \cite{PachToth-DCG02,FP10,FP14} for example.
-For an integer $\delta\geq 2$, if each curve is in at most $\delta$ intersections with other curves, then the corresponding string graph is called a \defin{$\delta$-string graph}.
-A \defin{$(g,\delta)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
+A \defin{string graph} is the intersection graph of a set of curves (\defin{strings}) in the plane with no three curves meeting at a single point; see  \cite{PachToth-DCG02,FP10,FP14} for example.
+For an integer $\delta\geq 2$, if each curve is in at most $\delta$ intersections with other curves, then the corresponding string graph is called a \defin{$\delta$-string graph}.\footnote{Every $\delta$-string graph has maximum degree at most $\delta$.
+However, it is not necessarily the case that every string graph of maximum degree $\delta$ is a $\delta$-string graph.
+If a pair of strings intersect in $c\ge 1$ points, then this counts as $c$ crossings for each string, but only contributes one to the degree of each.} A \defin{$(g,\delta)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
 

### 2022-11-15 response2.tex
Distinction between δ-string and max-degree-δ string graph

@@
 \begin{response}
-  We have added the qualifier "bounded-degree" here.
+  We have added the qualifier "bounded-degree" here, even though the $\delta$-string condition is actually stronger.
+(A pair of strings may cross more than once, in which case each of these crossings counts against $\delta$.)   At this point in the introduction (not yet having defined a string graph), this is hopefully a good enough approximation.
+The first paragraph of Section 3.1 gives the definition of $\delta$-string graph and makes note of the distinction between $\delta$-string graph and string graph of maximum degree at most $\delta$.
+Hopefully anyone that wants to make use of Theorem~14 will read the definition of $\delta$-string graph.
 \end{response}

### 2022-11-15 k-planar.tex
toned down  response

@@
 However, it is not necessarily the case that every string graph of maximum degree $\delta$ is a $\delta$-string graph.
-If a pair of strings intersect in $c\ge 1$ points, then this counts as $c$ crossings for each string, but only contributes one to the degree of each.} A \defin{$(g,\delta)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
+If a pair of strings intersect in $c\ge 1$ points, then this counts as $c$ crossings for each string, but only contributes one to their degree.} A \defin{$(g,\delta)$-string} graph is defined analogously for curves on a surface of Euler genus at most $g$.
 

### 2022-11-15 response2.tex
toned down  response

@@
 \begin{response}
-  We have added the qualifier "bounded-degree" here, even though the $\delta$-string condition is actually stronger.
+  We have added the qualifier "bounded-degree" here, although the $\delta$-string condition is actually stronger.
 (A pair of strings may cross more than once, in which case each of these crossings counts against $\delta$.)   At this point in the introduction (not yet having defined a string graph), this is hopefully a good enough approximation.
 The first paragraph of Section 3.1 gives the definition of $\delta$-string graph and makes note of the distinction between $\delta$-string graph and string graph of maximum degree at most $\delta$.
-Hopefully anyone that wants to make use of Theorem~14 will read the definition of $\delta$-string graph.
+  %Hopefully anyone that wants to make use of Theorem~14 will read the definition of $\delta$-string graph. DW: This could be deemed an attack on the referee. 
 \end{response}

### 2022-11-17 k-planar.tex
Trying out the term "indexed family"

@@
   $vw\in E(A)$ and $xy\in E(B)$.}
-and treewidth\footnote{For a tree $T$, a \defin{$T$-decomposition} of a graph $G$ is a collection $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that
+and treewidth\footnote{For a tree $T$, a \defin{$T$-decomposition} of a graph $G$ is an indexed family $\mathcal{T}=(B_x:x\in V(T))$ of subsets of $V(G)$ indexed by the nodes of $T$ such that
 (i) for every $vw\in E(G)$, there exists some node $x\in V(T)$ with $v,w\in B_x$; and
@@
 For each $x\in V(T)$, define $S_x := \{v\in V(G): a(v)= x\}$.
-Since $a:V(G)\to V(T)$ is a function with domain $V(G)$ and range $V(T)$, $\PP:=(S_x : x\in V(T))$ is a collection of disjoint sets that cover $V(G)$.
+Since $a:V(G)\to V(T)$ is a function with domain $V(G)$ and range $V(T)$, $\PP:=(S_x : x\in V(T))$ is an indexed family of disjoint sets that cover $V(G)$.
 Let $J:=G^\SS/\PP$ denote the resulting quotient graph.

### 2022-11-17 k-planar.tex
removed \notes

@@
   \item In \cref{Structure} we prove our main result for shortcut systems (\cref{ShortcutProduct}), and its optimisations for $k$-planar graphs and their genus-$g$ generalisation (\cref{kPlanarProduct,gkPlanarProduct}).
-  \item \Cref{examples} uses \cref{ShortcutProduct,kPlanarProduct,gkPlanarProduct} to derive product structure theorems for bounded-degree
-  %\note{DW}{We actually make a stronger assumption than bounded degree; we assume each string is crossed at most $k$ times. I am happy to leave our write-up as is, but be aware of the difference. }
-  string graphs, powers of bounded-degree graphs, map graphs of bounded clique number, and $k$-nearest neighbour graphs.
+  \item \Cref{examples} uses \cref{ShortcutProduct,kPlanarProduct,gkPlanarProduct} to derive product structure theorems for bounded-degree string graphs, powers of bounded-degree graphs, map graphs of bounded clique number, and $k$-nearest neighbour graphs.
   \item \Cref{FramedSection} proves a product structure theorem for $d$-framed multigraphs in which the graph $H$ has treewidth $3$ and is planar.
@@
 
-%\note{PM}{To make the referee happy I've added the ``bounded-degree'' qualifier to string graphs and the ``bounded clique number'' qualifier to map graphs.
-Do we want to explcitly state that the clique number of a $(g,d)$-map graph is bounded by a function of $g$ and $d$?
-Proof (for the $g=0$ case): A $d$-map representation gives a drawing with at most $\floor{\tfrac{d-2}{2}}\ceil{\tfrac{d-2}{2}}\in O(d^2)$ crossings per edge, but any drawing of $K_{\omega}$ on a surface of genus-$g$ has an edge that is crossed $\Omega(\omega^2)$ times.
-So if $G$ is a $d$-map graph then $\omega(G) \in O(d)$.} \note{DW}{I see no need to state that the clique number of a $(g,d)$-map graph is bounded by a function of $g$ and $d$.
-This is implied by \cref{gdMapProduct} and is not used in its proof.
-Also, surely there is a more combinatorial proof (with better bounds) than what you get by counting crossings.}
-
 Map graphs are defined as follows.
